### 2020-07-10 abstract/IQFA 11_19/GDR-IQFA_Template.tex
tout d'un coup

\documentclass[aps,twocolumn,showpacs]{revtex4-1}

\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{theorem}
\usepackage{amsmath}
\usepackage{times}
\usepackage{color}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\ifpdf
\usepackage{epstopdf}   
\usepackage{url}
\fi

\begin{document}
\title{Sub-GHz linewidths ensembles of SiV centers in a diamond nano-pyramid revealed by charge state conversion}

\author{Cl{\'e}ment Pellet-Mary$^1$, Louis Nicolas$^1$, Tom Delord $^1$, Paul Huillery $^1$, and Gabriel H{\'e}tet$^{1}$}
\email{clement.pellet-mary@phys.ens.fr}
\affiliation{$^1$ Laboratoire de Physique de l'Ecole normale sup{\'e}rieure, ENS, Universit{\'e} PSL, CNRS,Sorbonne Universit{\'e}, Universit{\'e} Paris-Diderot, Sorbonne Paris Cit{\'e}, Paris, France
}

\begin{abstract}
\normalsize
\noindent
Efficient interfaces between two-level emitters and photons are key components of quantum networks and sensing. %Diamond negatively charged silicon vacancy centers (SiV$^-$) can be used for such applications as they are photostable and reveal a large emission in the zero phonon line. Nanostructures are required in order to avoid reflections to ensure a strong coupling between the emitters and the light field. 
SiV$^-$ centers embedded into a diamond nanostructure are promising for such purpose. We have studied photoluminescence properties at cryogenic temperatures of diamond AFM probes [1,2] and observed photochromism of a very dense ensemble of SiV$^-$ centers consisting in trapping into a dark state under resonant excitation when nitrogen impurities are present. We suppose that this effect is a charge state switching between negatively charged and neutral SiV charge states.  This effect is used to perform persistent hole burning which reveals very low homogeneous broading, only twice the lifetime limit. It is promising  for quantum optics experiments and paves the way to sub-wavelength microscopy technics. %such as ground state depletion microscopy

We will also present our latest results on the Zeeman splitting and possible lifetime measurement of the SiV$^-$ $1/2$ spin.

\includegraphics[width=.7\columnwidth]{fig_hyp}\\
\noindent\textbf{Figure 1.} a) Averaged PLE measured at 6K by scanning the resonant laser at 737 nm around an electronic transition transition in the presence of green laser light between scans. b) PLE spectra after three different exposure times at 6K : in black without hole burning, in blue t$_{exp}$ = 300 ms and in orange, t$_{exp}$ = 1500 ms. The dots represent raw data and the continuous lines are fitted data. Here P$_{737}$= 0.877 $\mu$W. The inset shows the width of the holes $\Gamma_h$ inferred from the fit and normalized to the homogeneous linewidth $\Gamma_e$(W= $\Gamma_h/\Gamma_e$).

\end{abstract}

\maketitle

\begin{thebibliography}{}

\bibitem{Ref1} AR. Nelz et al., Applied Physics Letters, 109, 19, 193105, (2016).

\bibitem{Ref2} L. Nicolas et al., AIP Advances, 8, 6, 065102, (2018).

\end{thebibliography}
\end{document}

### 2020-07-10 abstract/IQFA 11_19/abstract_CPM.tex
tout d'un coup

--- 2020-07-10 abstract/IQFA 11_19/GDR-IQFA_Template.tex
@@
-\documentclass[aps,twocolumn,showpacs]{revtex4-1}
-
-\usepackage{epsfig}
+\documentclass[a4paper,11pt]{article}
+\linespread{1.1}
+\usepackage[top=3cm, bottom=2.54cm, left=2.54cm, right=2.54cm]{geometry}
+\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
-\usepackage{mathrsfs}
-\usepackage{theorem}
-\usepackage{amsmath}
-\usepackage{times}
-\usepackage{color}
-\usepackage[french]{babel}
-\usepackage[T1]{fontenc}
-\usepackage{ifpdf}
-\ifpdf
-\usepackage{epstopdf}   
-\usepackage{url}
-\fi
+\usepackage{graphicx}
 
 \begin{document}
-\title{Sub-GHz linewidths ensembles of SiV centers in a diamond nano-pyramid revealed by charge state conversion}
+\pagenumbering{gobble}
 
-\author{Cl{\'e}ment Pellet-Mary$^1$, Louis Nicolas$^1$, Tom Delord $^1$, Paul Huillery $^1$, and Gabriel H{\'e}tet$^{1}$}
-\email{clement.pellet-mary@phys.ens.fr}
-\affiliation{$^1$ Laboratoire de Physique de l'Ecole normale sup{\'e}rieure, ENS, Universit{\'e} PSL, CNRS,Sorbonne Universit{\'e}, Universit{\'e} Paris-Diderot, Sorbonne Paris Cit{\'e}, Paris, France
-}
+\Large
+ \begin{center}
+Sub-GHz linewidths ensembles of SiV centers in a diamond nano-pyramid revealed by charge state conversion
 
-\begin{abstract}
+\hspace{10pt}
+
 \normalsize
+% Author names and affiliations
+Cl{\'e}ment Pellet-Mary $^1$, Louis Nicolas $^1$, Tom Delord $^1$, Paul Huillery $^1$, Gabriel H{\'e}tet$^1$\\
+
+\hspace{10pt}
+
+\small  
+$^1$ Laboratoire de Physique de l'Ecole normale sup{\'e}rieure, ENS, Université PSL, CNRS,Sorbonne Universit{\'e}, Universit{\'e} Paris-Diderot, Sorbonne Paris Cit{\'e}, Paris, France.\\
+
+\end{center}
+
+\hspace{10pt}
+
+\normalsize
+
 \noindent
@@
 
+%In our search for a suitable sample, we have studied in detail and at cryogenic temperature sharp single crystal diamond AFM probes from Artech Carbon.
+Due to the growth conditions, a lot of SiV centers are present at the apex of such a tip whose radius of curvature is as low as 10 nm [3].
+They exhibit remarkable optical properties at low temperature.
+%We have performed measurements with a scanning Fabry-P{\'e}rot interferometer [4].
+This way, we have shown that the inhomogeneous linewidth of the ensemble of emitters is 15 GHz at 6 K.
+That means that the structure has a very low strain which is probably due to the slow and controlled CVD growth along the $\langle001\rangle$ axis [5].
+
+%Moreover, when nitrogen impurities are also present, SiV centers are trapped into a dark state under resonant excitation.
+Its lifetime is very long (more than hours).
+Illumination with a green laser is needed to recover the photoluminescence.
+We suppose that this effect is a charge state switching between negatively charged and neutral SiV charge states [6].
+%The charge transfer dynamics enables persistent spectral hole burning which unveils the homogeneous width as shown in Figure 1.
+Thanks to a rate model, we estimate the homogeneous linewidth $\Gamma_e$ to be 390 MHz at 6 K [7].
+
+%Figure
+\hspace{10pt}
+\begin{figure}[h!]\centering
 \includegraphics[width=.7\columnwidth]{fig_hyp}\\
@@
 The inset shows the width of the holes $\Gamma_h$ inferred from the fit and normalized to the homogeneous linewidth $\Gamma_e$(W= $\Gamma_h/\Gamma_e$).
+\end{figure}
 
-\end{abstract}
+%This effect reveals that the linewidth is close to be lifetime limited as required for quantum optics experiments and paves the way to sub-wavelength microscopy technics such as ground state depletion microscopy [8].
+%
+%We will also present our latest results on the modification of lifetime of solid state emitters in a half-cavity set-up.
 
-\maketitle
+%References
+\hspace{10pt}
 
-\begin{thebibliography}{}
-
-\bibitem{Ref1} AR. Nelz et al., Applied Physics Letters, 109, 19, 193105, (2016).
-
-\bibitem{Ref2} L. Nicolas et al., AIP Advances, 8, 6, 065102, (2018).
-
-\end{thebibliography}
+\noindent{[}1{]} R. Nelz et al., Applied Physics Letters, 109, 19, 193105, (2016).\\
+\noindent{[}2{]} L. Nicolas et al., AIP Advances, 8, 6, 065102, (2018).\\
+\noindent{[}2{]} L. Nicolas et al., AIP Advances, 8, 6, 065102, (2018).\\
 \end{document}

### 2020-07-10 abstract/IQFA 11_19/abstract_LN.tex
tout d'un coup

--- 2020-07-10 abstract/IQFA 11_19/abstract_CPM.tex
@@
 % Author names and affiliations
-Cl{\'e}ment Pellet-Mary $^1$, Louis Nicolas $^1$, Tom Delord $^1$, Paul Huillery $^1$, Gabriel H{\'e}tet$^1$\\
+Louis Nicolas $^1$, Cl{\'e}ment Pellet-Mary $^1$, Tom Delord $^1$, Paul Huillery $^1$, Gabriel H{\'e}tet$^1$\\
 
@@
 %such as ground state depletion microscopy
-
-We will also present our latest results on the Zeeman splitting and possible lifetime measurement of the SiV$^-$ $1/2$ spin.
+We will also present our latest results on the modification of the lifetime of solid state emitters in a half-cavity set-up.
 

### 2020-07-10 articles/fiche de lecture.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Fiche de lecture des articles}
  \section{Nano-pyramides Louis 2019}
  \subsection{Description des échantillons}
  Les échantillons c'est des pointes d'AFM de $\approx 15 \mu$m de long et $\leq 10$ nm de large au niveau de la pointe et qui, à cause de leur synthèse sur un substrat de silicium, ont des centres SiV, particulièrement au niveau de la pointe.

  Du coup, on peut focus sur une pointe unique, et dans ces conditions tu as très peu d'élargissement inhomogène ($\approx 10$ GHZ), ce qui est inférieur à la structure fine du SiV qui est de 50 et 250 GHZ. Donc tu vois les 4 pics au spectro.

  \end{document}

### 2020-07-10 dico/voc.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Vocabulaire}
\section{Spectro}
\begin{itemize}
\item \textbf{Fluorescence} : émission spontanée
\item \textbf{ZPL} : Zero Phonon Line : la plage de fréquence de l'émissions/absorption dans un solide qui correspond à une recombinaison radiative sans émission de phonon, pour une transition donnée.
\item \textbf{PSB} : Phonon Side Band : au contraire, la plage qui correspond à l'émission d'un photon et d'au moins un phonon (vibronic transition). Les énergies de la PSB sont nécessairement plus faibles que celles de la ZPL pour l'émission, et plus hautes pour l'absorption.
\item \textbf{Frank-Codon} : (conséquence) Le diagramme d'absorption et de fluorescence sont en gros symétriques par rapport à la longueur d'onde centrale (c'est pas une science exacte)
\item \textbf{PLE} : Photoluminescence excitation : le but c'est d’exciter à résonance avec un laser accordable qui va balayer la zone qui t’intéresse (typiquement la ZPL) tout en observant la photoluminescence d'une autre région (typiquement la PSB) que tu espères proportionnelle à l'absorption, elle même proportionnelle à l'émission. Ca te permet donc de faire de la spectro avec une précision limitée par la largeur de ton laser. Pour faire tout ça, tu auras besoin d'un filtre pour retirer la zone à résonance, et d'un moyen de mesurer la fréquence de ton laser (par exemple un Michelson, au moins pour les valeures relatives)
\item \textbf{Effet Stark} : Décalage énergétique du à un champ électrique, s'applique donc aux dipôles électriques qui sont sensibles à un effet $D\cdot E$
\item \textbf{Effet Zeeman} : Décalage énergétique du à un champ magnétique, s'applique donc aux dipôles électriques qui sont sensibles à un effet $\mu \cdot B$
\item \textbf{EPR / ESR} (en spectro, c'est aussi lié aux inégalités de Bell) Electron Paramgnetic / Spin Resonnance : c'est en gros de la MNR sur des électrons : on split les niveaux d'énergie avec un champ mag permanent important ($\approx$ 0.3 T) par effet Zeeman, et on mesure l'absorbance d'une micro-onde scannée autour de la transition. (tu peux aussi scanner le champ mag permanent, comme de la MNR)

Dans le cas des NV, tu n'as même pas besoin d'un champ mag vu que tu as naturellement un splitting entre $\ket{0}$ et $\ket{\pm 1}$. La différence avec l'ODMR c'est que c'est directement l'absorption des photons mirco-onde qu'on regarde.

\item \textbf{ODMR} : Optically Detected Magnetic Resonnance : Assez clair, c'est de l'ESR mais ou tu mesures un signal optique plutôt que l'absorption de la micro-onde. Dans le cas du NV, c'est parce que le niveau $\ket{0}$ est plus brillant que l'état $\ket{\pm 1}$, et que le laser de pompe polarise dans l'état $\ket{0}$.

\item \textbf{Jahn-Teller Effect} : c'es un effet de brisure spontanée de symétrie (à suivre...)
\end{itemize}
  \section{Matériel}
  \begin{itemize}
  \item \textbf{FEL0700} (c'est un zéro): Filtre longpass, c'est à dire passe haut en longueur d'onde, donc passe-bas en énergie, jusqu'au $\lambda$ indiqué (en nm). C'est des filtres interférentiels, donc il faut faire un peu gaffe à l'angle d'incidence. La petite flèche indique le sens de transmission.
  \item \textbf{FES1200} : Pareil mais shortpass.
  \end{itemize}

  \end{document}

### 2020-07-10 echantillons/échantillons.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Échantillons}

  \section{SiV}

  \subsection{template}
  \subsubsection{Mesures réalisées}
  \subsubsection{Observations}
  \subsubsection{Rangé}

  \subsection{"échantillon d'alex"} 
  C'est un diamant bulk (procédé de fabrication ?) avec une fine couche de SiV près de la surface.
  \subsubsection{Mesures réalisées}
  \begin{itemize}
  \item spectres à chaud et à froid : 2019/09/02 $\to$ 2019/09/12
  \item tentative de PLE : 2019/09/09 $\to$ 2019/09/12
  \end{itemize}
  \subsubsection{Observations}
  Spectre de SiV assez compact (~3 nm de ZPL), la PLE n'a pas réussie à isoler de centre unique, soit parce qu’il y en a trop, soit parce que je me suis loupé.

  PL avec le laser vert : max 8E4, PLE avec le laser résonnant : max 4E4.
  \subsubsection{Rangé}
  Dans une boite d'Alex marqué ...Si...Recuit 1100 (pas sur que ça corresponde)

 \subsection{Nano-pyramides "Elke Neu"}
  \subsubsection{Mesures réalisées}
  \begin{itemize}
  \item Des spectres à chaud(2019/09/19)
  \item Cf manip pour plus de précisions
  \item 22/10/2019 : du Hole-burning pour des pyramides posées directement sur un aimant néodyme.
  \end{itemize}

  \subsubsection{Observations}
  Ca crache bien, 6E6 de PL au laser vert et à l'objectif x10 sans forcer, et à travers la plaque de quartz. Par contre le NV est beaucoup plus visible que le SiV.

  (15/10/19) En faite les SiV crachent pas mal, 1.3E6 avec les filtre pour la ZPL, 5E5 avec le filtre pour la PSB (PLE)
  \subsubsection{Rangé}
  \begin{itemize}
  \item Solution mère : Dans une boite "gel" marquée "nano-diamant pyramide CVD" (ou un truc du genre) dans le tiroir des échantillons.
  \item Échantillon sur lame de quartz épaisse: sur son porte-échantillon t dans le cryostat.
  \end{itemize}

 \section{Rubis}
 \subsection{L'échantillon du stage}
 Le joli, même si je crois que techniquement c'est pas celui des photos au MEB.
 \subsubsection{Mesures réalisées}
 cf le stage, des temps de vie, de la saturation, des spectres ... 
 \subsubsection{Observations}
 Bah pas de modifications de temps de vie pour l'instant
 \subsubsection{Rangé}
 \begin{itemize}
 \item Solution mère dans un tube marqué "rubis" dans un pot avec d'autres tubes, dans le tiroir échantillon (j'espère)
 \item Échantillon qui va bien : sur une fine lame de quartz, sur le porte-échantillon, dans le cryostat.
 \end{itemize}

 \section{Plaques de quartz}
 Plaques de quartz de 3x3 mm nettoyées à l'acétone + IPA et propres au microscope (certaines en tout cas)
 \subsubsection{Rangé}
 Dans une boite "gel" marqué "propre" dans le tiroir des échantillons.
  \end{document}

### 2020-07-10 ensignement/TD_math/CC1/CC1_2PY013-P2.tex
tout d'un coup

\documentclass[10pt,a4paper,notitlepage]{article}
%Mise en page
\usepackage[left=2cm, right=2cm, lines=45, top=0.8in, bottom=0.7in]{geometry}
\usepackage{fancyhdr}
\usepackage{fancybox}
\usepackage{pdfpages} 
\renewcommand{\headrulewidth}{1.5pt}
\renewcommand{\footrulewidth}{1.5pt}
\pagestyle{fancy}
\newcommand\Loadedframemethod{TikZ}
\usepackage[framemethod=\Loadedframemethod]{mdframed}
\usepackage{tikz}
\usetikzlibrary{calc,through,backgrounds}
\usetikzlibrary{matrix,positioning}
%Desssins geometriques
\usetikzlibrary{arrows}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{datavisualization}
\usetikzlibrary{automata} % LATEX and plain TEX
\usetikzlibrary{shapes.multipart}
\usetikzlibrary{decorations.pathmorphing} 
\usepackage{pgfplots}
\usepackage{physics}
\usepackage{titletoc}
\usepackage{mathpazo} 
\usepackage{algpseudocode}
\usepackage{algorithmicx} 
\usepackage{bohr} 
\usepackage{xlop} 
\usepackage{bbding} 
%\usepackage{minibox} 
%Ecriture arabe
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage{mathdesign}
\usepackage{bbding} 
\usepackage{romande}

\lhead{
\textbf{LU2PY013-PeiP }
}
\rhead{\textbf{03 mars 2020}
}
\chead{\textbf{
CC1
 }}

\lfoot{}
\cfoot{\textbf{$\mathsf{2019-2020}$}}
%\rfoot{\textit{Pr. $\mathcal{A}$.Kaal}}
%=====================Algo setup
\algblock{If}{EndIf}
\algcblock[If]{If}{ElsIf}{EndIf}
\algcblock{If}{Else}{EndIf}
\algrenewtext{If}{\textbf{si}}
\algrenewtext{Else}{\textbf{sinon}}
\algrenewtext{EndIf}{\textbf{finsi}}
\algrenewtext{Then}{\textbf{alors}}
\algrenewtext{While}{\textbf{tant que}}
\algrenewtext{EndWhile}{\textbf{fin tant que}}
\algrenewtext{Repeat}{\textbf{r\'ep\'eter}}
\algrenewtext{Until}{\textbf{jusqu'\`a}}
\algcblockdefx[Strange]{If}{Eeee}{Oooo}
[1]{\textbf{Eeee} "#1"}
{\textbf{Wuuuups\dots}}

\algrenewcommand\algorithmicwhile{\textbf{TantQue}}
\algrenewcommand\algorithmicdo{\textbf{Faire}}
\algrenewcommand\algorithmicend{\textbf{Fin}}
\algrenewcommand\algorithmicrequire{\textbf{Variables}}
\algrenewcommand\algorithmicensure{\textbf{Constante}}% replace ensure by constante
\algblock[block]{Begin}{End}
\newcommand\algo[1]{\textbf{algorithme} #1;}
\newcommand\vars{\textbf{variables } }
\newcommand\consts{\textbf{constantes}}
\algrenewtext{Begin}{\textbf{debut}}
\algrenewtext{End}{\textbf{fin}}
%================================
%================================

\setlength{\parskip}{1cm}
\setlength{\parindent}{1cm}
\tikzstyle{titregris} =
[draw=gray,fill=white, shading = exersicetitle, %
text=gray, rectangle, rounded corners, right,minimum height=.3cm]
\pgfdeclarehorizontalshading{exersicebackground}{100bp}
{color(0bp)=(green!40); color(100bp)=(black!5)}
\pgfdeclarehorizontalshading{exersicetitle}{100bp}
{color(0bp)=(red!40);color(100bp)=(black!5)}
\newcounter{exercise}
\renewcommand*\theexercise{exercice \textbf{Exercice}~n\arabic{exercise}}
\makeatletter
\def\mdf@@exercisepoints{}%new mdframed key:
\define@key{mdf}{exercisepoints}{%
\def\mdf@@exercisepoints{#1}
}
\mdfdefinestyle{exercisestyle}{%
outerlinewidth=1em,outerlinecolor=white,%
leftmargin=-1em,rightmargin=-1em,%
middlelinewidth=0.5pt,roundcorner=3pt,linecolor=black,
apptotikzsetting={\tikzset{mdfbackground/.append style ={%
shading = exersicebackground}}},
innertopmargin=0.1\baselineskip,
skipabove={\dimexpr0.1\baselineskip+0\topskip\relax},
skipbelow={-0.1em},
needspace=0.5\baselineskip,
frametitlefont=\sffamily\bfseries,
settings={\global\stepcounter{exercise}},
singleextra={%
\node[titregris,xshift=0.5cm] at (P-|O) %
{~\mdf@frametitlefont{\theexercise}~};
\ifdefempty{\mdf@@exercisepoints}%
{}%
{\node[titregris,left,xshift=-1cm] at (P)%
{~\mdf@frametitlefont{\mdf@@exercisepoints points}~};}%
},
firstextra={%
\node[titregris,xshift=1cm] at (P-|O) %
{~\mdf@frametitlefont{\theexercise}~};
\ifdefempty{\mdf@@exercisepoints}%
{}%
{\node[titregris,left,xshift=-1cm] at (P)%
{~\mdf@frametitlefont{\mdf@@exercisepoints points}~};}%
},
}
\makeatother

%%%%%%%%%

%%%%%%%%%%%%%%%
\mdfdefinestyle{theoremstyle}{%
outerlinewidth=0.01em,linecolor=black,middlelinewidth=0.5pt,%
frametitlerule=true,roundcorner=2pt,%
apptotikzsetting={\tikzset{mfframetitlebackground/.append style={%
shade,left color=white, right color=blue!20}}},
frametitlerulecolor=black,innertopmargin=1\baselineskip,%green!60,
innerbottommargin=0.5\baselineskip,
frametitlerulewidth=0.1pt,
innertopmargin=0.7\topskip,skipabove={\dimexpr0.2\baselineskip+0.1\topskip\relax},
frametitleaboveskip=1pt,
frametitlebelowskip=1pt
}
\setlength{\parskip}{0mm}
\setlength{\parindent}{10mm}
\mdtheorem[style=theoremstyle]{definition}{\textbf{Exercice}}
%================Liste definition--numList-and alphList=============
\newcounter{alphListCounter}
\newenvironment
{alphList}
{\begin{list}
{\alph{alphListCounter})}
{\usecounter{alphListCounter}
\setlength{\rightmargin}{0cm}
\setlength{\leftmargin}{0.5cm}
\setlength{\itemsep}{0.2cm}
\setlength{\partopsep}{0cm}
\setlength{\parsep}{0cm}}
}
{\end{list}}
\newcounter{numListCounter}
\newenvironment
{numList}
{\begin{list}
{\arabic{numListCounter})}
{\usecounter{numListCounter}
\setlength{\rightmargin}{0cm}
\setlength{\leftmargin}{0.5cm}
\setlength{\itemsep}{0cm}
\setlength{\partopsep}{0cm}
\setlength{\parsep}{0cm}}
}
{\end{list}}
%===========================================================
\begin{document}

\begin{center}
\large{\textbf{Corrigé}}
\end{center}
\begin{definition}[]
\hspace{0ex}\textbf{(30')} - Calculer les intégrales suivantes et préciser en une phrase si elles sont généralisées ou non :
\begin{enumerate}
   \item ${\displaystyle \int_0^1 x^2 \mathrm{ln}(x) dx}$

   \begin{center}
   En intégrant par partie,
   \begin{align*}
   \int_0^1 x^2 \mathrm{ln}(x) dx = \left[\dfrac{x^3}{3} \mathrm{ln}(x)\right]_0^1 - \int_0^1 \dfrac{x^2}{3}dx   =-\left[\dfrac{x^3}{9}\right]_0^1 =-\dfrac{1}{9}
   \end{align*}
   L'intégrale n'est pas généralisée car $\lim_{x \to 0} \left(x^2 \mathrm{ln}(x)\right)= 0$ par croissance comparée, la fonction est bien bornée sur le domaine d'intégration.
   \end{center}
   \item ${\displaystyle \int_0^{\pi/2} \dfrac{\cos(x)}{\sqrt{1+3\sin(x)}}dx}$
   \begin{center}
   En posant le changement de variable $y^2=1+3\sin(x)$, qui est bien bijectif sur l'intervalle considéré, alors $2y\mathrm{d}y=3\cos(x)dx$ et les nouvelles bornes vont de $y=1$ à $y=2$.
   \begin{align*}
   \int_0^{\pi/2} \dfrac{\cos(x)}{\sqrt{1+3\sin(x)}}dx = \int_1^2 \dfrac{\dfrac{2}{3}y\mathrm{d}y}{\sqrt{y^2}}=\int_1^2 \dfrac{2}{3}dy=\dfrac{2}{3}
   \end{align*}
   L'intégrale n'est pas généralisée.
   \end{center}
   \item  ${\displaystyle \int_{-\pi/2}^{\pi/2} \dfrac{\cos(x)-1}{\sin^2(x)}dx}$
   \begin{center}
   En remarquant que $$\dfrac{\mathrm{d}(1/\sin(x))}{dx}=-\dfrac{\cos(x)}{\sin^2(x)}$$
   Et que $$ \dfrac{1}{\sin^2(x)}=\dfrac{\sin^2(x)+\cos^2(x)}{\sin^2(x)}=1+\mathrm{cotan}^2(x)=-\mathrm{cotan}'(x)$$
   Il vient que $$\int_{-\pi/2}^{\pi/2} \dfrac{\cos(x)-1}{\sin^2(x)}dx = \left[-\dfrac{1}{\sin(x)}+\mathrm{cotan}(x)\right]_{-\pi/2}^{\pi/2} =\left[\dfrac{\cos(x)-1}{\sin(x)}\right]_{-\pi/2}^{\pi/2}=-2$$
   L'intégrale n'est pas généralisée car $\lim_{x \to 0}\left(\dfrac{\cos(x)-1}{\sin^2(x)}\right)=-\dfrac{1}{2}$ par les développement limités de $\sin^2(x)$ et $\cos(x) -1$ en $0$. La fonction est bien bornée sur son domaine d'intégration.
   \end{center}
\end{enumerate}
\end{definition}
%----------------
\begin{definition}
\hspace{0ex} \textbf{(15')} - Calcul de volumes en utilisant des intégrales simples :

\subsection*{Le cylindre}
\includegraphics[scale=.15]{cylindre}

Le cylindre est un objet formidable (observez la figure ci dessus si vous ne nous croyez pas), surtout lorsqu'il a un rayon R et une hauteur h. La valeur de son volume va vous étonner ! Calculez-la en utilisant une intégrale simple. (Pour ce faire, vous vous rappelerez, avec nostalgie, l'équation d'un cercle de rayon R : $x^2+y^2=R^2$).

\subsection*{La toupie}
\includegraphics[scale=.15]{toupie}

Les toupies ne sont pas mal non plus. Celle-ci possède le petit désavantage d'être infinie. Mais son équation cartésienne est plutôt lisible :

\begin{equation*}
\left\{ 
        \begin{array}{rl}
    x^2+y^2&=\frac{1}{\sqrt{\abs{z}}}\\
    \abs{z}&\leq 5
    \end{array}
    \right.
\end{equation*}
Trouvez la valeur du volume de cette toupie (que vous pouvez admirer sur la figure ci dessus) en utilisant une intégrale simple.

\end{definition}
%----------------
\begin{definition}[]
%-- begin{minipage}{0.7\textwidth}
\hspace{0ex} \textbf{(10')} - Soit
%--\newline
\begin{equation}
y'(x)+\sin (x) y = \dfrac{e^{\cos(x)}}{\cos^2 (x)}
\end{equation}
\begin{enumerate}
\item Est-ce une équation différentielle linéaire ? (justifiez). De quel ordre ? (justifiez)
\item Résoudre cette équation sur un intervalle que l'on précisera (le plus grand possible).
\end{enumerate}

%--\end{minipage}
\end{definition}
%----------------
\begin{definition}[]
\textbf{(20')} - On cherche les fonctions y, solutions de :
\begin{equation}
\label{Eq2.2}
\dfrac{dy}{dx}e^{-y^2}(1-2y)=2xe^{-y}
\end{equation}
et vérifiant $y(1)=1$.
\begin{enumerate}
\item Est-ce que (\ref{Eq2.2}) est une équation différentielle linéaire ? (justifiez). De quel ordre ? (justifiez)
\item Calculer la dérivée de $e^{y(1-y)}$
\item Donner une EDO équivalente à (\ref{Eq2.2}) en utilisant la dérivée de $e^{y(1-y)}$ calculée juste avant. L'intégrer sur un intervalle que l'on précisera (le plus grand possible).
\end{enumerate}

%-\end{minipage}

\end{definition}
%----------------
\begin{definition}[]
\subsection*{(15') - Une balade en voiture en guise de dessert  }
\subsubsection*{- À l'arrêt }
La suspension d'une voiture de masse $m$ = 800 kg est équivalente à un ressort vertical de constante de raideur $k$ = 1,36 10$^4$ N/m, qui est associé à un frottement visqueux de constante $f$ = 1,6 10$^2$ kg/s.

L'équation du déplacement vertival $y(t)$ de la voiture s'écrit :

\begin {equation}
\quad  \quad \quad \quad \quad \quad \quad \quad \quad m\frac{d^{2}y}{dt^{2}}+f \frac{dy}{dt} +ky = 0
\end{equation}

\begin{enumerate}
   \item Cette équation différentielle est-elle linéaire ? Si oui, de quel ordre est-elle ? 
   \item Résoudre cette équation différentielle
\end{enumerate}
\subsubsection*{- Sur une route "ondulée"}
 On suppose maintenant que la voiture roule sur piste "ondulée" qui force une oscillation verticale de pulsation $\sqrt{17}$ et transforme l'équation différentielle initiale en :
\begin {equation}
\quad  \quad \quad \quad \quad \quad \quad m\frac{d^{2}y}{dt^{2}}+f \frac{dy}{dt} +ky = F\text{cos} \sqrt{17}t
\end{equation}
avec $F$= 800 N
\begin{enumerate}
   \item Simplifier cette équation différentielle en effectuant les applications numériques.
   \item Donner la solution générale en précisant l'intervalle de définition de $t$.
   \item Déterminer la solution particulière alors associée aux conditions initiales $y(t=0)=1$ et $y'(t=0) = -1$.

\end{enumerate}
\end{definition}

%% Macros for ``successive divisions'' 
%%
%\def\Division#1#2#3{ % Dividend, divisor, remainder
% \matrix (D) [matrix of nodes,
%              below=0pt of D-1-2.south east,
%              row sep=1pt, column sep=1pt,
%              every node/.append style={minimum width=12mm}] {
%   #1 \pgfmatrixnextcell #2 \\
%   |[marcar] (R#1)| #3      \\
% };
% \draw[shorten >=2pt, shorten <=2pt]
%   (D-1-2.north west) |- (D-1-2.south east);
%}
%\def\FinDivision#1{
%\node[marcar, below=2pt of D-1-2.south] (C)(C)  {#1};
%}
%\tikzset{marcar/.style={circle,draw,inner sep=2pt,minimum width=0pt,
%fill=yellow!10}}
%
%
%\begin{tikzpicture}
%  \coordinate (D-1-2) at (0,0) {}; % We must start with this command.
%  \Division{25}{2}{1} % First dividend, divisor, remainder
%  \Division{12}{2}{0} % Dividend (previous quotient), divisor, remainder
%  \Division{6}{2}{0}  
%  \Division{3}{2}{1}  
%  \FinDivision{1}     % Last remainder.
%
%% We can draw an arrow jumping from one remainder 
%% to the next one. Every reminder is a node called
%% Rdividend. Last remainder is node C.
%  \draw[shorten <=1mm, ->, dashed] (C) to[out=-150,in=-65] (R3);
%  \draw[shorten <=1mm, ->, dashed] (R3) to[out=-150,in=-65] (R6);
%  \draw[shorten <=1mm, ->, dashed] (R6) to[out=-150,in=-65] (R12);
%  \draw[shorten <=1mm, ->, dashed] (R12) to[out=-150,in=-65] (R25);
%
%% Some more information:
%  \node (MSB) at ([yshift=-1.3cm]R6.south) {Most significant bit (MSB)};    
%  \node (LSB) at ([yshift=-2mm]MSB.south) {Less significant bit (LSB)}; 
%\draw[ ->] (MSB.east) to[out=30,in=-55] (C);
%\draw[ ->] (LSB.west) to[out=150,in=-95] (R25);
%\end{tikzpicture}
%\begin{center}\SnowflakeChevronBold \SnowflakeChevronBold \SnowflakeChevronBold \end{center}
%%----------------------------------------------------------------------------------------------------------------------------------------
%
%%\includepdf[doublepages=true]{serie55}

\end{document}

### 2020-07-10 ensignement/TD_math/CC1/CC1_corrigé.tex
tout d'un coup

\documentclass[fleqn]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{mathexam}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{mathtools, physics}

\ExamClass{LUPY013 PEIP}
\ExamName{Controle Continu 1}
\ExamHead{3 Mars 2020}

\let\ds\displaystyle

\begin{document}
\ExamInstrBox{
\begin{center}
\textbf{\textbf{Corrigé}}
\end{center}
}

\section{Exercice 1 : Intégrales }

\begin{enumerate}
   \item \textbf{(3 pts)} En intégrant par partie,
   \begin{align*}
   \int_0^1 x^2 \mathrm{ln}(x) dx = \left[\dfrac{x^3}{3} \mathrm{ln}(x)\right]_0^1 - \int_0^1 \dfrac{x^2}{3}dx   =-\left[\dfrac{x^3}{9}\right]_0^1 =-\dfrac{1}{9}
   \end{align*}
   L'intégrale n'est pas généralisée car $\lim_{x \to 0} \left(x^2 \mathrm{ln}(x)\right)= 0$ par croissance comparée, la fonction est bien bornée sur son domaine d'intégration.
   \item \textbf{(3 pts)} En posant le changement de variable $y^2=1+3\sin(x)$, qui est bien bijectif sur l'intervalle considéré, alors $2y\mathrm{d}y=3\cos(x)dx$ et les nouvelles bornes vont de $y=1$ à $y=2$.
   \begin{align*}
   \int_0^{\pi/2} \dfrac{\cos(x)}{\sqrt{1+3\sin(x)}}dx = \int_1^2 \dfrac{\dfrac{2}{3}y\mathrm{d}y}{\sqrt{y^2}}=\int_1^2 \dfrac{2}{3}dy=\dfrac{2}{3}
   \end{align*}
   L'intégrale n'est pas généralisée.
   \item \textbf{(4 pts)} En remarquant que $$\dfrac{\mathrm{d}(1/\sin(x))}{dx}=-\dfrac{\cos(x)}{\sin^2(x)}$$
   Et que $$ \dfrac{1}{\sin^2(x)}=\dfrac{\sin^2(x)+\cos^2(x)}{\sin^2(x)}=1+\mathrm{cotan}^2(x)=-\mathrm{cotan}'(x)$$
   Il vient que $$\int_{-\pi/2}^{\pi/2} \dfrac{\cos(x)-1}{\sin^2(x)}dx = \left[-\dfrac{1}{\sin(x)}+\mathrm{cotan}(x)\right]_{-\pi/2}^{\pi/2} =\left[\dfrac{\cos(x)-1}{\sin(x)}\right]_{-\pi/2}^{\pi/2}=-2$$
   L'intégrale n'est pas généralisée car $\lim_{x \to 0}\left(\dfrac{\cos(x)-1}{\sin^2(x)}\right)=-\dfrac{1}{2}$ par les développement limités de $\sin^2(x)$ et $\cos(x) -1$ en $0$. La fonction est bien bornée sur son domaine d'intégration.
\end{enumerate}
\section{Exercice 2 : calcul de volumes}
\begin{enumerate}
\item \textbf{(3 pts)} On décompose notre cylindre en petit cylindre de hauteur $\mathrm{d} z$ et additionne les volumes de tous ces cylindres pour trouver le volume du cylindre. Le volume de chacun de ces cylindres est $\pi r(z)^2 \mathrm{d} z$. Il ne reste plus qu'à trouver $r(z)$.\newline
On connaît les équations du cylindre $x^2 + y^2 = R^2$ et $0\leq z \leq h$. Pour chacun de ces cylindres, on est dans un plan parallèle à $xOy$, c'est à dire qu'on est à $z$ fixé. On connaît l'équation d'un cercle dans un tel plan $x^2+y^2 = r(z)^2$, on en déduit que $x^2 + y^2 = R^2 = r(z)^2$. Donc, le volume du cylindre pour un $z$ donné est $\pi R^2 \mathrm{d} z $, il ne reste plus qu'à sommer :
\begin{align*}
    V_{\mathrm{cyl}} = \displaystyle \int_{0}^{h} \pi R^2 \, \mathrm{d} z = \pi R^2 h
\end{align*}
\item \textbf{(2 pts)} On décompose la toupie en petit cylindre de hauteur $\mathrm{d} z$ et on  additionne les volumes de tous ces cylindres pour trouver le volume de la toupie. Le volume de chacun des cylindres est $\pi r(z)^2 \mathrm{d} y$.\newline
A $z$ fixé, on connaît l'équation d'un cercle dans le plan $xOy$: $x^2+y^2 = \frac{1}{\sqrt{\abs{z}}}$, on en déduit que $x^2 + y^2 =  \frac{1}{\sqrt{\abs{z}}} = r(z)^2$. Donc, le volume du cylindre pour un $z$ donné est $\pi  \frac{1}{\sqrt{\abs{z}}} \mathrm{d} z $, il ne reste plus qu'à sommer :
\begin{equation}
    V_{\mathrm{toupie}} = \displaystyle \int_{-5}^{5} \pi  \frac{1}{\sqrt{\abs{z}}} \, \mathrm{d} z = 2\displaystyle \int_{0}^{5} \pi  \frac{1}{\sqrt{z}} \, \mathrm{d} z
\end{equation}
Qui est une intégrale généralisée. On trouve finalement
\begin{equation}
    V_{\mathrm{toupie}} = 2\pi\lim \limits_{\epsilon \rightarrow 0}\left[2\sqrt{z} \right]_\epsilon^5 = 4\pi\sqrt{5}
\end{equation}
\end{enumerate}
\section{Exercice 3 \textbf{(4 pts)}}

{1. \textbf{(1 pts)} C'est une équation différentielle linéaire (en
y) ordinaire du premier ordre (seule la première dérivée de y intervient)}\\
{}\\

{2. \textbf{(3 pts)} Méthode de la variation de la constante.}

{On commence par résoudre l'EDO sans second membre
:
\[
\frac{dy_{0}}{dx}+\textrm{sin}\left(x\right)y_{0}=0
\]
c'est-à-dire
\[
\frac{y_{0}'}{y_{0}}=-\textrm{sin}\left(x\right)
\]
ou
\[
\frac{d}{dx}\left(\textrm{ln}\left(y_{0}\right)\right)=\frac{d}{dx}\left(\textrm{cos}\left(x\right)+cte\right)
\]
dont la solution générale s'écrit
\[
y_{0}\left(x\right)=\textrm{e}^{\textrm{cos}\left(x\right)+cte}=A\textrm{e}^{\textrm{cos}\left(x\right)}
\]
avec $A$ une constante réelle que l'on suppose être une fonction
$f\left(x\right)$ dans un second temps. On remplace donc $y\left(x\right)$
par $f\left(x\right)\textrm{e}^{\textrm{cos}\left(x\right)}$ dans
l'EDO et l'on obtient
\[
f'\left(x\right)=\frac{1}{\textrm{cos}^{2}\left(x\right)}
\]
en divisant par $\textrm{e}^{\textrm{cos}\left(x\right)}$ de chaque
côté.}

{Cette équation s'intègre alors facilement en reconnaissant
la dérivée de la fonction tangente à droite :
\[
f\left(x\right)=\textrm{tan}\left(x\right)+C
\]
avec $C$ une constante d'intégration réelle.}

{La solution générale recherchée est donc égale à
\[
y_{sol}\left(x\right)=\left(\textrm{tan}\left(x\right)+C\right)\textrm{e}^{\textrm{cos}\left(x\right)}
\]
et son support de définition est celui de la fonction tangente, c'est-à-dire
l'ensemble des réels privé des nombres $\pi/2$ modulo $\pi$ (relatif).}

\section{Exercice 4 \textbf{(6 pts)}}

{1.\textbf{(1 pts)} C'est une équation différentielle non-linéaire
(en y) du premier ordre.}\\
{}\\

{2.\textbf{(1 pts)} La dérivée de la fonction $\textrm{e}^{y\left(1-y\right)}$
s'écrit :
\[
\left(\textrm{e}^{y\left(1-y\right)}\right)'=y'\left(1-2y\right)\textrm{e}^{y\left(1-y\right)}
\]
}\\
{}\\

{3.\textbf{(3 pts)} Cette dérivée conduit à réexprimer l'EDO comme
\[
\frac{d}{dx}\left(\textrm{e}^{y\left(1-y\right)}\right)=2x
\]
d'où
\[
y\left(1-y\right)=\textrm{ln}\left(x^{2}+cte\right)
\]
La condition $y\left(1\right)=1$ élimine la constante : $cte=0$
puisque $\textrm{ln}\left(1+cte\right)=0$.}

{Le support de définition de $y\left(x\right)$ est
donc au maximum la droite réelle privée du point $0$.}

{Il reste à exprimer $y\left(x\right)$ de façon explicite.
L'équation $y\left(1-y\right)=\textrm{ln}\left(x^{2}\right)$ s'écrit
aussi
\[
y^{2}-y+\textrm{ln}\left(x^{2}\right)=0
\]
qui donne deux solutions a priori possibles
\[
\frac{1\pm\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
\]
dont en fait seule celle avec le signe $+$ correspond à la solution
recherchée car devant vérifier la condition $y\left(1\right)=1$ :
\[
y_{sol}\left(x\right)=\frac{1+\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
\]
On remarque que le support de définition de $y_{sol}\left(x\right)$
est plus restreint que prévu ci-dessus : il doit correspondre à la
condition de réalité de la racine carrée : $1-4\textrm{ln}\left(x^{2}\right)\geq0$,
c'est-à-dire 
\[
\left|x\right|\leq\textrm{e}^{1/8}
\]
Finalement, l'ensemble de définition de $y_{sol}\left(x\right)$ est
\[
\left[-\textrm{e}^{1/8},0\right[U\left]0,\textrm{e}^{1/8}\right]
\]
}
\section{Exercice 5 \textit{\normalsize{Une ballade en voiture en guise de dessert}{[}5 points{]}}}
\subsubsection*{- \`a l'arr\^et }

\begin{enumerate}
\item \textbf{(0.5 pts)}Cette \'equation diff\'erentielle est lin\'eaire, \`a coefficients constants et elle est du second ordre.
\item \textbf{(1.5 pts)}On \'ecrit le polyn\^ome associ\'e et on cherche ses solutions.
\begin{align*}
mp^{2}+fp+k=0
\end{align*}
avec $\Delta= f^2-4mk$.

Si $\Delta >0$  (c'est \`a dire si $f^{2} > 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent :
\begin{align*}
p_{1,2}=\frac{- f \pm \sqrt{\Delta}}{2m}
\end{align*}
et les solutions de l'\'equation diff\'erentielle s'expriment :
\begin{align*}
y= Ae^{p_{1}t}+Be^{p_{2}t}
\end{align*}
o\`u $A$ et $B$ sont des constantes r\'eelles.

Si $\Delta =0$ (c'est \`a dire si $f^{2} = 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
\begin{align*}
p_{1} = p_{2} = \frac{-f }{2m}
\end{align*}
et les solutions de l'\'equation diff\'erentielle s'expriment :
\begin{align*}
y= (At+B)e^{p_{1}t}
\end{align*}
o\`u $A$ et $B$ sont des constantes r\'eelles.

Si $\Delta <0$ (c'est \`a dire si $f^{2} < 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
\begin{align*}
p_{1,2} = \alpha \pm i \omega   
\end{align*}
avec $\alpha= \frac{- f}{2m}$ et $\omega =\frac{\sqrt{-\Delta}}{2m}$

et les solutions de l'\'equation diff\'erentielle s'expriment :
\begin{align*}
y= e^{\alpha t} [A\  \text{sin}(\omega t) + B\  \text{cos} (\omega t)]
\end{align*}
o\`u $A$ et $B$ sont des constantes r\'eelles.
\end{enumerate}

\subsection{- Sur une route "ondul\'ee"}

L'\'equation diff\'erentielle a maintenant un second membre. Elle s'\'ecrit 
\begin{align*}
m\frac{d^{2}y}{dt^{2}}+f \frac{dy}{dt} +ky = F\text{cos} \sqrt{17}t
\end{align*}

\begin{enumerate}
\item \textbf{(0.5 pts)}En effectuant les applications num\'eriques, on obtient :
\begin{align*}
\frac{d^{2}y}{dt^{2}}+2 \frac{dy}{dt} +17y = \text{cos} \sqrt{17}t
\end{align*}
\item \textbf{(1.5 pts)}L'intervalle de d\'efinition de $t$ est ici l'ensemble des r\'eels positifs, la variable $t$ \'etant ici le temps.  Pour obtenir la solution g\'en\'erale de cette \'equation diff\'erentielle, on cherche d'abord la solution g\'en\'erale de l'\'equation sans second membre. On se reporte aux r\'esultats pr\'ec\'edents. Ici $\Delta$ est n\'egatif et on a $\sqrt{-\Delta} = 8$, ce qui conduit \`a $\alpha =-1$ et $\omega=4$. La solution g\'en\'erale de l'\'equation sans second membre s'\'ecrit donc :
\begin{center}

{$y= e^{- t}[A\  \text{sin}(4 t) + B\  \text{cos} (4 t)]$ , A et B \'etant deux r\'eels.}
\end{center}

{On cherche maintenant une solution particuli\`ere de l'\'equation avec second membre, de la forme:}
\begin{center}

{$y= C\ \text{sin}(\sqrt{17} t) + D\  \text{cos} (\sqrt{17} t))$ , C et D \'etant deux r\'eels.}
\end{center}

{Les d\'eriv\'ees $\frac{dy}{dt}$ et $\frac{d^{2}y}{dt^{2}}$ s'\'ecrivent alors:}
\begin{center}

{ $ \frac{dy}{dt}\ = \sqrt{17} C\ \text{cos}(\sqrt{17} t) - \sqrt{17\ } D\ \text{sin} (\sqrt{17} t))$}

{$\frac{d^{2}y}{dt^{2}}= - \ 17 C\ \text{sin}(\sqrt{17} t) \ \  -17\ D\ \text{cos} (\sqrt{17} t))$}
\end{center}

{En rempla\c cant $\frac{d^{2}y}{dt^{2}}$, $\frac{dy}{dt}$ et $y$ dans l'\'equation diff\'erentielle compl\`ete, on obtient en r\'eunissant les termes en sinus d'un c\^ot\'e et ceux en cosinus de l'autre:}

\begin{center}
{$[- 17C -2 \sqrt{17} D+17C]\ \text{sin}(\sqrt{17} t) + [-17 D+2\sqrt{17} C +17D]\ \text{cos}(\sqrt{17} t) = \text{cos}( \sqrt{17} t)$}
\end{center}

{Par identification des termes en $\text{sin}(\sqrt{17} t)$ et ceux en $\text{cos}(\sqrt{17} t)$ de part et d'autre de l'\'egalit\'e, on obtient :}

\begin{center}
{$D=0$ et $C=\frac{1}{2\sqrt{17}}$.}
\end{center}

{Soit une solution particuli\`ere de l'\'equation avec second membre \'egale \`a:}

\begin{center}
{$y= \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$,}
\end{center}

{et une solution g\'en\'erale de l'\'equation compl\`ete avec second membre :}

\begin{center}
{$y= e^{- t} [A\  \text{sin}(4 t) + B\  \text{cos}(4 t)] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$, A et B \'etant deux r\'eels}
\end{center}
\item \textbf{(1 pts)} Pour trouver la solution particuli\`ere de l'\'equation compl\`ete avec second membre, associ\'ee aux conditions aux limites $y(t=0)=1$ et $y'(t=0)=-1$ on calcule $y(t=0)$ et $y'(t=0)$:

\begin{center}
{$y(t=0)= B+ \frac{1}{2\sqrt{17}}*0=1$, }

{$y'(t=0)= [4A - B + \frac{1}{2}]=-1$.}
\end{center}

{On obtient donc $B=1$ et $A=-\frac{1}{8}$ et la solution s''ecrit :}

\begin{center}
{$y= e^{- t} [-\frac{1}{8} \  \text{sin}(4 t) +  \text{cos}(4 t) ] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$.}
\end{center}
\end{enumerate}

\end{document}

### 2020-07-10 ensignement/TD_math/CC1_projet.tex
tout d'un coup

--- 2020-07-10 ensignement/TD_math/CC1/CC1_corrigé.tex
@@
 \usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{color}
-\usepackage{mathtools, physics}
 
@@
 \begin{center}
-\textbf{\textbf{Corrigé}}
+\textbf{Polycopié de cours autorisé. \\ Calculatrice et TD interdits}
 \end{center}
@@
 
-\section{Exercice 1 : Intégrales }
-
+\section{Exercice 1 : intégrales (45')}
+\subsection{Calcul d'intégrales (3x10')}
+Calculer les intégrales suivantes et préciser en une phrase si elles sont généralisées ou non :
 \begin{enumerate}
-   \item \textbf{(3 pts)} En intégrant par partie,
-   \begin{align*}
-   \int_0^1 x^2 \mathrm{ln}(x) dx = \left[\dfrac{x^3}{3} \mathrm{ln}(x)\right]_0^1 - \int_0^1 \dfrac{x^2}{3}dx   =-\left[\dfrac{x^3}{9}\right]_0^1 =-\dfrac{1}{9}
-   \end{align*}
-   L'intégrale n'est pas généralisée car $\lim_{x \to 0} \left(x^2 \mathrm{ln}(x)\right)= 0$ par croissance comparée, la fonction est bien bornée sur son domaine d'intégration.
-   \item \textbf{(3 pts)} En posant le changement de variable $y^2=1+3\sin(x)$, qui est bien bijectif sur l'intervalle considéré, alors $2y\mathrm{d}y=3\cos(x)dx$ et les nouvelles bornes vont de $y=1$ à $y=2$.
-   \begin{align*}
-   \int_0^{\pi/2} \dfrac{\cos(x)}{\sqrt{1+3\sin(x)}}dx = \int_1^2 \dfrac{\dfrac{2}{3}y\mathrm{d}y}{\sqrt{y^2}}=\int_1^2 \dfrac{2}{3}dy=\dfrac{2}{3}
-   \end{align*}
-   L'intégrale n'est pas généralisée.
-   \item \textbf{(4 pts)} En remarquant que $$\dfrac{\mathrm{d}(1/\sin(x))}{dx}=-\dfrac{\cos(x)}{\sin^2(x)}$$
-   Et que $$ \dfrac{1}{\sin^2(x)}=\dfrac{\sin^2(x)+\cos^2(x)}{\sin^2(x)}=1+\mathrm{cotan}^2(x)=-\mathrm{cotan}'(x)$$
-   Il vient que $$\int_{-\pi/2}^{\pi/2} \dfrac{\cos(x)-1}{\sin^2(x)}dx = \left[-\dfrac{1}{\sin(x)}+\mathrm{cotan}(x)\right]_{-\pi/2}^{\pi/2} =\left[\dfrac{\cos(x)-1}{\sin(x)}\right]_{-\pi/2}^{\pi/2}=-2$$
-   L'intégrale n'est pas généralisée car $\lim_{x \to 0}\left(\dfrac{\cos(x)-1}{\sin^2(x)}\right)=-\dfrac{1}{2}$ par les développement limités de $\sin^2(x)$ et $\cos(x) -1$ en $0$.
-La fonction est bien bornée sur son domaine d'intégration.
+   \item ${\displaystyle \int_0^1 x^2 \mathrm{ln}(x) dx}$ 
+   \item ${\displaystyle \int_0^{\pi/2} \dfrac{\cos(x)}{\sqrt{1+3\sin(x)}}dx}$
+   \item ${\displaystyle \int_{-\pi/2}^{\pi/2} \dfrac{\cos(x)-1}{\sin^2(x)}dx}$
 \end{enumerate}
-\section{Exercice 2 : calcul de volumes}
-\begin{enumerate}
-\item \textbf{(3 pts)} On décompose notre cylindre en petit cylindre de hauteur $\mathrm{d} z$ et additionne les volumes de tous ces cylindres pour trouver le volume du cylindre.
-Le volume de chacun de ces cylindres est $\pi r(z)^2 \mathrm{d} z$.
-Il ne reste plus qu'à trouver $r(z)$.\newline
-On connaît les équations du cylindre $x^2 + y^2 = R^2$ et $0\leq z \leq h$.
-Pour chacun de ces cylindres, on est dans un plan parallèle à $xOy$, c'est à dire qu'on est à $z$ fixé.
-On connaît l'équation d'un cercle dans un tel plan $x^2+y^2 = r(z)^2$, on en déduit que $x^2 + y^2 = R^2 = r(z)^2$.
-Donc, le volume du cylindre pour un $z$ donné est $\pi R^2 \mathrm{d} z $, il ne reste plus qu'à sommer :
-\begin{align*}
-    V_{\mathrm{cyl}} = \displaystyle \int_{0}^{h} \pi R^2 \, \mathrm{d} z = \pi R^2 h
-\end{align*}
-\item \textbf{(2 pts)} On décompose la toupie en petit cylindre de hauteur $\mathrm{d} z$ et on  additionne les volumes de tous ces cylindres pour trouver le volume de la toupie.
-Le volume de chacun des cylindres est $\pi r(z)^2 \mathrm{d} y$.\newline
-A $z$ fixé, on connaît l'équation d'un cercle dans le plan $xOy$: $x^2+y^2 = \frac{1}{\sqrt{\abs{z}}}$, on en déduit que $x^2 + y^2 =  \frac{1}{\sqrt{\abs{z}}} = r(z)^2$.
-Donc, le volume du cylindre pour un $z$ donné est $\pi  \frac{1}{\sqrt{\abs{z}}} \mathrm{d} z $, il ne reste plus qu'à sommer :
-\begin{equation}
-    V_{\mathrm{toupie}} = \displaystyle \int_{-5}^{5} \pi  \frac{1}{\sqrt{\abs{z}}} \, \mathrm{d} z = 2\displaystyle \int_{0}^{5} \pi  \frac{1}{\sqrt{z}} \, \mathrm{d} z
-\end{equation}
-Qui est une intégrale généralisée. On trouve finalement
-\begin{equation}
-    V_{\mathrm{toupie}} = 2\pi\lim \limits_{\epsilon \rightarrow 0}\left[2\sqrt{z} \right]_\epsilon^5 = 4\pi\sqrt{5}
-\end{equation}
-\end{enumerate}
-\section{Exercice 3 \textbf{(4 pts)}}
-
-{1. \textbf{(1 pts)} C'est une équation différentielle linéaire (en
-y) ordinaire du premier ordre (seule la première dérivée de y intervient)}\\
-{}\\
-
-{2. \textbf{(3 pts)} Méthode de la variation de la constante.}
-
-{On commence par résoudre l'EDO sans second membre
-:
-\[
-\frac{dy_{0}}{dx}+\textrm{sin}\left(x\right)y_{0}=0
-\]
-c'est-à-dire
-\[
-\frac{y_{0}'}{y_{0}}=-\textrm{sin}\left(x\right)
-\]
-ou
-\[
-\frac{d}{dx}\left(\textrm{ln}\left(y_{0}\right)\right)=\frac{d}{dx}\left(\textrm{cos}\left(x\right)+cte\right)
-\]
-dont la solution générale s'écrit
-\[
-y_{0}\left(x\right)=\textrm{e}^{\textrm{cos}\left(x\right)+cte}=A\textrm{e}^{\textrm{cos}\left(x\right)}
-\]
-avec $A$ une constante réelle que l'on suppose être une fonction
-$f\left(x\right)$ dans un second temps. On remplace donc $y\left(x\right)$
-par $f\left(x\right)\textrm{e}^{\textrm{cos}\left(x\right)}$ dans
-l'EDO et l'on obtient
-\[
-f'\left(x\right)=\frac{1}{\textrm{cos}^{2}\left(x\right)}
-\]
-en divisant par $\textrm{e}^{\textrm{cos}\left(x\right)}$ de chaque
-côté.}
-
-{Cette équation s'intègre alors facilement en reconnaissant
-la dérivée de la fonction tangente à droite :
-\[
-f\left(x\right)=\textrm{tan}\left(x\right)+C
-\]
-avec $C$ une constante d'intégration réelle.}
-
-{La solution générale recherchée est donc égale à
-\[
-y_{sol}\left(x\right)=\left(\textrm{tan}\left(x\right)+C\right)\textrm{e}^{\textrm{cos}\left(x\right)}
-\]
-et son support de définition est celui de la fonction tangente, c'est-à-dire
-l'ensemble des réels privé des nombres $\pi/2$ modulo $\pi$ (relatif).}
-
-\section{Exercice 4 \textbf{(6 pts)}}
-
-{1.\textbf{(1 pts)} C'est une équation différentielle non-linéaire
-(en y) du premier ordre.}\\
-{}\\
-
-{2.\textbf{(1 pts)} La dérivée de la fonction $\textrm{e}^{y\left(1-y\right)}$
-s'écrit :
-\[
-\left(\textrm{e}^{y\left(1-y\right)}\right)'=y'\left(1-2y\right)\textrm{e}^{y\left(1-y\right)}
-\]
-}\\
-{}\\
-
-{3.\textbf{(3 pts)} Cette dérivée conduit à réexprimer l'EDO comme
-\[
-\frac{d}{dx}\left(\textrm{e}^{y\left(1-y\right)}\right)=2x
-\]
-d'où
-\[
-y\left(1-y\right)=\textrm{ln}\left(x^{2}+cte\right)
-\]
-La condition $y\left(1\right)=1$ élimine la constante : $cte=0$
-puisque $\textrm{ln}\left(1+cte\right)=0$.}
-
-{Le support de définition de $y\left(x\right)$ est
-donc au maximum la droite réelle privée du point $0$.}
-
-{Il reste à exprimer $y\left(x\right)$ de façon explicite.
-L'équation $y\left(1-y\right)=\textrm{ln}\left(x^{2}\right)$ s'écrit
-aussi
-\[
-y^{2}-y+\textrm{ln}\left(x^{2}\right)=0
-\]
-qui donne deux solutions a priori possibles
-\[
-\frac{1\pm\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
-\]
-dont en fait seule celle avec le signe $+$ correspond à la solution
-recherchée car devant vérifier la condition $y\left(1\right)=1$ :
-\[
-y_{sol}\left(x\right)=\frac{1+\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
-\]
-On remarque que le support de définition de $y_{sol}\left(x\right)$
-est plus restreint que prévu ci-dessus : il doit correspondre à la
-condition de réalité de la racine carrée : $1-4\textrm{ln}\left(x^{2}\right)\geq0$,
-c'est-à-dire 
-\[
-\left|x\right|\leq\textrm{e}^{1/8}
-\]
-Finalement, l'ensemble de définition de $y_{sol}\left(x\right)$ est
-\[
-\left[-\textrm{e}^{1/8},0\right[U\left]0,\textrm{e}^{1/8}\right]
-\]
-}
-\section{Exercice 5 \textit{\normalsize{Une ballade en voiture en guise de dessert}{[}5 points{]}}}
-\subsubsection*{- \`a l'arr\^et }
-
-\begin{enumerate}
-\item \textbf{(0.5 pts)}Cette \'equation diff\'erentielle est lin\'eaire, \`a coefficients constants et elle est du second ordre.
-\item \textbf{(1.5 pts)}On \'ecrit le polyn\^ome associ\'e et on cherche ses solutions.
-\begin{align*}
-mp^{2}+fp+k=0
-\end{align*}
-avec $\Delta= f^2-4mk$.
-
-Si $\Delta >0$  (c'est \`a dire si $f^{2} > 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent :
-\begin{align*}
-p_{1,2}=\frac{- f \pm \sqrt{\Delta}}{2m}
-\end{align*}
-et les solutions de l'\'equation diff\'erentielle s'expriment :
-\begin{align*}
-y= Ae^{p_{1}t}+Be^{p_{2}t}
-\end{align*}
-o\`u $A$ et $B$ sont des constantes r\'eelles.
-
-Si $\Delta =0$ (c'est \`a dire si $f^{2} = 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
-\begin{align*}
-p_{1} = p_{2} = \frac{-f }{2m}
-\end{align*}
-et les solutions de l'\'equation diff\'erentielle s'expriment :
-\begin{align*}
-y= (At+B)e^{p_{1}t}
-\end{align*}
-o\`u $A$ et $B$ sont des constantes r\'eelles.
-
-Si $\Delta <0$ (c'est \`a dire si $f^{2} < 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
-\begin{align*}
-p_{1,2} = \alpha \pm i \omega   
-\end{align*}
-avec $\alpha= \frac{- f}{2m}$ et $\omega =\frac{\sqrt{-\Delta}}{2m}$
-
-et les solutions de l'\'equation diff\'erentielle s'expriment :
-\begin{align*}
-y= e^{\alpha t} [A\  \text{sin}(\omega t) + B\  \text{cos} (\omega t)]
-\end{align*}
-o\`u $A$ et $B$ sont des constantes r\'eelles.
-\end{enumerate}
-
-\subsection{- Sur une route "ondul\'ee"}
-
-L'\'equation diff\'erentielle a maintenant un second membre. Elle s'\'ecrit 
-\begin{align*}
-m\frac{d^{2}y}{dt^{2}}+f \frac{dy}{dt} +ky = F\text{cos} \sqrt{17}t
-\end{align*}
-
-\begin{enumerate}
-\item \textbf{(0.5 pts)}En effectuant les applications num\'eriques, on obtient :
-\begin{align*}
-\frac{d^{2}y}{dt^{2}}+2 \frac{dy}{dt} +17y = \text{cos} \sqrt{17}t
-\end{align*}
-\item \textbf{(1.5 pts)}L'intervalle de d\'efinition de $t$ est ici l'ensemble des r\'eels positifs, la variable $t$ \'etant ici le temps.
-Pour obtenir la solution g\'en\'erale de cette \'equation diff\'erentielle, on cherche d'abord la solution g\'en\'erale de l'\'equation sans second membre.
-On se reporte aux r\'esultats pr\'ec\'edents.
-Ici $\Delta$ est n\'egatif et on a $\sqrt{-\Delta} = 8$, ce qui conduit \`a $\alpha =-1$ et $\omega=4$.
-La solution g\'en\'erale de l'\'equation sans second membre s'\'ecrit donc :
-\begin{center}
-
-{$y= e^{- t}[A\  \text{sin}(4 t) + B\  \text{cos} (4 t)]$ , A et B \'etant deux r\'eels.}
-\end{center}
-
-{On cherche maintenant une solution particuli\`ere de l'\'equation avec second membre, de la forme:}
-\begin{center}
-
-{$y= C\ \text{sin}(\sqrt{17} t) + D\  \text{cos} (\sqrt{17} t))$ , C et D \'etant deux r\'eels.}
-\end{center}
-
-{Les d\'eriv\'ees $\frac{dy}{dt}$ et $\frac{d^{2}y}{dt^{2}}$ s'\'ecrivent alors:}
-\begin{center}
-
-{ $ \frac{dy}{dt}\ = \sqrt{17} C\ \text{cos}(\sqrt{17} t) - \sqrt{17\ } D\ \text{sin} (\sqrt{17} t))$}
-
-{$\frac{d^{2}y}{dt^{2}}= - \ 17 C\ \text{sin}(\sqrt{17} t) \ \  -17\ D\ \text{cos} (\sqrt{17} t))$}
-\end{center}
-
-{En rempla\c cant $\frac{d^{2}y}{dt^{2}}$, $\frac{dy}{dt}$ et $y$ dans l'\'equation diff\'erentielle compl\`ete, on obtient en r\'eunissant les termes en sinus d'un c\^ot\'e et ceux en cosinus de l'autre:}
-
-\begin{center}
-{$[- 17C -2 \sqrt{17} D+17C]\ \text{sin}(\sqrt{17} t) + [-17 D+2\sqrt{17} C +17D]\ \text{cos}(\sqrt{17} t) = \text{cos}( \sqrt{17} t)$}
-\end{center}
-
-{Par identification des termes en $\text{sin}(\sqrt{17} t)$ et ceux en $\text{cos}(\sqrt{17} t)$ de part et d'autre de l'\'egalit\'e, on obtient :}
-
-\begin{center}
-{$D=0$ et $C=\frac{1}{2\sqrt{17}}$.}
-\end{center}
-
-{Soit une solution particuli\`ere de l'\'equation avec second membre \'egale \`a:}
-
-\begin{center}
-{$y= \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$,}
-\end{center}
-
-{et une solution g\'en\'erale de l'\'equation compl\`ete avec second membre :}
-
-\begin{center}
-{$y= e^{- t} [A\  \text{sin}(4 t) + B\  \text{cos}(4 t)] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$, A et B \'etant deux r\'eels}
-\end{center}
-\item \textbf{(1 pts)} Pour trouver la solution particuli\`ere de l'\'equation compl\`ete avec second membre, associ\'ee aux conditions aux limites $y(t=0)=1$ et $y'(t=0)=-1$ on calcule $y(t=0)$ et $y'(t=0)$:
-
-\begin{center}
-{$y(t=0)= B+ \frac{1}{2\sqrt{17}}*0=1$, }
-
-{$y'(t=0)= [4A - B + \frac{1}{2}]=-1$.}
-\end{center}
-
-{On obtient donc $B=1$ et $A=-\frac{1}{8}$ et la solution s''ecrit :}
-
-\begin{center}
-{$y= e^{- t} [-\frac{1}{8} \  \text{sin}(4 t) +  \text{cos}(4 t) ] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$.}
-\end{center}
-\end{enumerate}
-
 \end{document}

### 2020-07-10 ensignement/TD_math/correction CC1/CC1-2020-Exos-3-et-4-CORRIGE.tex
tout d'un coup

%% LyX 2.3.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,french]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=20mm,rmargin=20mm}
\setcounter{tocdepth}{2}
\usepackage{color}
\usepackage{babel}
\makeatletter
\addto\extrasfrench{%
   \providecommand{\og}{\leavevmode\flqq~}%
   \providecommand{\fg}{\ifdim\lastskip>\z@\unskip\fi~\frqq}%
}

\makeatother
\usepackage{amstext}
\usepackage[unicode=true,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=1,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=true]
 {hyperref}
\hypersetup{pdftitle={Le Manuel d'apprentissage de LYX},
 pdfauthor={LyX Team},
 pdfsubject={LyX-documentation Tutorial},
 pdfkeywords={LyX, documentation},
 linkcolor=black, citecolor=black, urlcolor=blue, filecolor=blue, pdfpagelayout=OneColumn, pdfnewwindow=true, pdfstartview=XYZ, plainpages=false}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
% DO NOT ALTER THIS PREAMBLE!!!
%
%This preamble is designed to ensure that the document prints
% out as advertised. If you mess with this preamble,
% parts of the document may not print out as expected.  If you
% have problems LaTeXing this file, please contact 
% the documentation team
% email: lyx-docs@lists.lyx.org

\usepackage{ifpdf} % part of the hyperref bundle
\ifpdf % if pdflatex is used

 % set fonts for nicer pdf view
 \IfFileExists{lmodern.sty}{\usepackage{lmodern}}{}

 \fi % end if pdflatex is used

% the pages of the TOC is numbered roman
% and a pdf-bookmark for the TOC is added
\let\myTOC\tableofcontents
\renewcommand\tableofcontents{%
  \frontmatter
  \pdfbookmark[1]{\contentsname}{}
  \myTOC
  \mainmatter }

\makeatother

\begin{document}
\begin{center}
{\small{}Sorbonne Universités} 
\par\end{center}

LU2PY013-PeiP\hfill{}03 mars 2020

\begin{center}
\medskip{}
\textbf{\textcolor{blue}{\Large{}CC1}}\medskip{}
\par\end{center}

\begin{center}
\textbf{Polycopié de cours autorisé - Calculatrice et TD interdits}\medskip{}
\par\end{center}

\section{Exercice 3 \textit{\normalsize{}{[}4 points{]}}}

Soit
\begin{equation}
y'\left(x\right)+\textrm{sin}\left(x\right)y=\frac{\textrm{e}^{\textrm{cos}\left(x\right)}}{\textrm{cos}^{2}\left(x\right)}
\end{equation}

\begin{enumerate}
\item Est-ce une équation différentielle \textbf{linéaire} ? (justifiez).
De quel \textbf{ordre} ? (justifiez)\textbf{ }\textbf{\textit{{[}0,5
+ 0,5 point{]}}}
\item \textbf{Résoudre} cette équation sur un \textbf{intervalle} que l\textquoteright on
précisera (le plus grand possible).\textbf{ }\textbf{\textit{{[}2
+ 1 points{]}}}\bigskip{}
\end{enumerate}
\textcolor{blue}{CORRECTION :}

\textcolor{blue}{1. C'est une équation différentielle linéaire (en
y) ordinaire du premier ordre (seule la première dérivée de y intervient)}\\
\textcolor{blue}{}\\

\textcolor{blue}{2. Méthode de la variation de la constante.}

\textcolor{blue}{On commence par résoudre l'EDO sans second membre
:
\[
\frac{dy_{0}}{dx}+\textrm{sin}\left(x\right)y_{0}=0
\]
c'est-à-dire
\[
\frac{y_{0}'}{y_{0}}=-\textrm{sin}\left(x\right)
\]
ou
\[
\frac{d}{dx}\left(\textrm{ln}\left(y_{0}\right)\right)=\frac{d}{dx}\left(\textrm{cos}\left(x\right)+cte\right)
\]
dont la solution générale s'écrit
\[
y_{0}\left(x\right)=\textrm{e}^{\textrm{cos}\left(x\right)+cte}=A\textrm{e}^{\textrm{cos}\left(x\right)}
\]
avec $A$ une constante réelle que l'on suppose être une fonction
$f\left(x\right)$ dans un second temps. On remplace donc $y\left(x\right)$
par $f\left(x\right)\textrm{e}^{\textrm{cos}\left(x\right)}$ dans
l'EDO et l'on obtient
\[
f'\left(x\right)=\frac{1}{\textrm{cos}^{2}\left(x\right)}
\]
en divisant par $\textrm{e}^{\textrm{cos}\left(x\right)}$ de chaque
côté.}

\textcolor{blue}{Cette équation s'intègre alors facilement en reconnaissant
la dérivée de la fonction tangente à droite :
\[
f\left(x\right)=\textrm{tan}\left(x\right)+C
\]
avec $C$ une constante d'intégration réelle.}

\textcolor{blue}{La solution générale recherchée est donc égale à
\[
y_{sol}\left(x\right)=\left(\textrm{tan}\left(x\right)+C\right)\textrm{e}^{\textrm{cos}\left(x\right)}
\]
et son support de définition est celui de la fonction tangente, c'est-à-dire
l'ensemble des réels privé des nombres $\pi/2$ modulo $\pi$ (relatif).}

\section{Exercice 4 \textit{\normalsize{}{[}6 points{]}}}

On cherche les fonctions $y$, solutions de : 
\begin{equation}
\frac{dy}{dx}\textrm{e}^{-y^{2}}\left(1-2y\right)=2x\textrm{e}^{-y}
\end{equation}
et vérifiant $y\left(1\right)=1$.
\begin{enumerate}
\item Est-ce que (2) est une équation différentielle \textbf{linéaire} ?
(justifiez). De quel \textbf{ordre} ? (justifiez) \textbf{\textit{{[}0,5
+ 0,5 point{]}}}
\item \textbf{Calculer} la dérivée de $\textrm{e}^{y\left(1-y\right)}$
\textbf{\textit{{[}1 point{]}}}
\item Donner une \textbf{EDO équivalente} à (2) en utilisant la dérivée
de $\textrm{e}^{y\left(1-y\right)}$ calculée juste avant. L\textquoteright \textbf{intégrer}
sur un \textbf{intervalle} que l\textquoteright on précisera (le plus
grand possible). \textbf{\textit{{[}0,5 + 1,5 + 2 points{]}}}\bigskip{}
\end{enumerate}
\textcolor{blue}{CORRECTION :}

\textcolor{blue}{1. C'est une équation différentielle non-linéaire
(en y) du premier ordre.}\\
\textcolor{blue}{}\\

\textcolor{blue}{2. La dérivée de la fonction $\textrm{e}^{y\left(1-y\right)}$
s'écrit :
\[
\left(\textrm{e}^{y\left(1-y\right)}\right)'=y'\left(1-2y\right)\textrm{e}^{y\left(1-y\right)}
\]
}\\
\textcolor{blue}{}\\

\textcolor{blue}{3. Cette dérivée conduit à réexprimer l'EDO comme
\[
\frac{d}{dx}\left(\textrm{e}^{y\left(1-y\right)}\right)=2x
\]
d'où
\[
y\left(1-y\right)=\textrm{ln}\left(x^{2}+cte\right)
\]
La condition $y\left(1\right)=1$ élimine la constante : $cte=0$
puisque $\textrm{ln}\left(1+cte\right)=0$.}

\textcolor{blue}{Le support de définition de $y\left(x\right)$ est
donc au maximum la droite réelle privée du point $0$.}

\textcolor{blue}{Il reste à exprimer $y\left(x\right)$ de façon explicite.
L'équation $y\left(1-y\right)=\textrm{ln}\left(x^{2}\right)$ s'écrit
aussi
\[
y^{2}-y+\textrm{ln}\left(x^{2}\right)=0
\]
qui donne deux solutions a priori possibles
\[
\frac{1\pm\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
\]
dont en fait seule celle avec le signe $+$ correspond à la solution
recherchée car devant vérifier la condition $y\left(1\right)=1$ :
\[
y_{sol}\left(x\right)=\frac{1+\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
\]
On remarque que le support de définition de $y_{sol}\left(x\right)$
est plus restreint que prévu ci-dessus : il doit correspondre à la
condition de réalité de la racine carrée : $1-4\textrm{ln}\left(x^{2}\right)\geq0$,
c'est-à-dire 
\[
\left|x\right|\leq\textrm{e}^{1/8}
\]
Finalement, l'ensemble de définition de $y_{sol}\left(x\right)$ est
\[
\left[-\textrm{e}^{1/8},0\right[U\left]0,\textrm{e}^{1/8}\right]
\]
}
\end{document}

### 2020-07-10 ensignement/TD_math/correction CC1/VF 20200310/CC1_corrigé.tex
tout d'un coup

--- 2020-07-10 ensignement/TD_math/CC1/CC1_corrigé.tex
@@
 \item \textbf{(0.5 pts)}Cette \'equation diff\'erentielle est lin\'eaire, \`a coefficients constants et elle est du second ordre.
-\item \textbf{(1.5 pts)}On \'ecrit le polyn\^ome associ\'e et on cherche ses solutions.
-\begin{align*}
-mp^{2}+fp+k=0
-\end{align*}
-avec $\Delta= f^2-4mk$.
+\item \textbf{(1.5 pts)}{On \'ecrit le polyn\^ome associ\'e et on cherche ses solutions.}
+\begin{center}
+{$mp^{2}+fp+k=0$ }
 
-Si $\Delta >0$  (c'est \`a dire si $f^{2} > 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent :
-\begin{align*}
-p_{1,2}=\frac{- f \pm \sqrt{\Delta}}{2m}
-\end{align*}
-et les solutions de l'\'equation diff\'erentielle s'expriment :
-\begin{align*}
-y= Ae^{p_{1}t}+Be^{p_{2}t}
-\end{align*}
-o\`u $A$ et $B$ sont des constantes r\'eelles.
+{avec $\Delta= f^2-4mk$.}
+\end{center}
 
-Si $\Delta =0$ (c'est \`a dire si $f^{2} = 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
-\begin{align*}
-p_{1} = p_{2} = \frac{-f }{2m}
-\end{align*}
-et les solutions de l'\'equation diff\'erentielle s'expriment :
-\begin{align*}
-y= (At+B)e^{p_{1}t}
-\end{align*}
-o\`u $A$ et $B$ sont des constantes r\'eelles.
+{On remarque qu'ici (compte tenu des valeurs num\'eriques de $f$, $m$ et $k$), on a $f^{2} << 4mk$ et donc $\Delta \sim -4mk$ < 0. Les solutions du polyn\^ome associ\'e s'\'ecrivent alors:}
 
-Si $\Delta <0$ (c'est \`a dire si $f^{2} < 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
-\begin{align*}
-p_{1,2} = \alpha \pm i \omega   
-\end{align*}
-avec $\alpha= \frac{- f}{2m}$ et $\omega =\frac{\sqrt{-\Delta}}{2m}$
+\begin{center}
+{ $p_{1,2} = \alpha \pm i \omega$   avec $\alpha= \frac{- f}{2m}$ et $\omega =\frac{\sqrt{-\Delta}}{2m}$}
 
-et les solutions de l'\'equation diff\'erentielle s'expriment :
-\begin{align*}
-y= e^{\alpha t} [A\  \text{sin}(\omega t) + B\  \text{cos} (\omega t)]
-\end{align*}
-o\`u $A$ et $B$ sont des constantes r\'eelles.
+\
+
+{et les solutions de l'\'equation diff\'erentielle s'expriment :}
+
+\
+
+{$y= e^{\alpha t} [A\  \text{sin}(\omega t) + B\  \text{cos} (\omega t)]$}
+
+{o\`u $A$ et $B$ sont des constantes r\'eelles. } 
+\end{center}
 \end{enumerate}
@@
 \begin{enumerate}
-\item \textbf{(0.5 pts)}En effectuant les applications num\'eriques, on obtient :
-\begin{align*}
-\frac{d^{2}y}{dt^{2}}+2 \frac{dy}{dt} +17y = \text{cos} \sqrt{17}t
-\end{align*}
-\item \textbf{(1.5 pts)}L'intervalle de d\'efinition de $t$ est ici l'ensemble des r\'eels positifs, la variable $t$ \'etant ici le temps.
-Pour obtenir la solution g\'en\'erale de cette \'equation diff\'erentielle, on cherche d'abord la solution g\'en\'erale de l'\'equation sans second membre.
-On se reporte aux r\'esultats pr\'ec\'edents.
-Ici $\Delta$ est n\'egatif et on a $\sqrt{-\Delta} = 8$, ce qui conduit \`a $\alpha =-1$ et $\omega=4$.
-La solution g\'en\'erale de l'\'equation sans second membre s'\'ecrit donc :
+\item \textbf{(0.5 pts)}{En effectuant les applications num\'eriques, on obtient :}
+
 \begin{center}
 
-{$y= e^{- t}[A\  \text{sin}(4 t) + B\  \text{cos} (4 t)]$ , A et B \'etant deux r\'eels.}
+{$\frac{d^{2}y}{dt^{2}}+\frac{1}{5} \frac{dy}{dt} +17y = \text{cos} \sqrt{17}t$}
+\end{center}
+
+\item \textbf{(2 pts)}{L'intervalle de d\'efinition de $t$ est ici l'ensemble des r\'eels positifs, la variable $t$ \'etant ici le temps.
+Pour obtenir la solution g\'en\'erale de cette \'equation diff\'erentielle, on cherche d'abord la solution g\'en\'erale de l'\'equation sans second membre.
+On se reporte aux r\'esultats pr\'ec\'edents en prenant les valeurs num\'eriques de l'equation simplifi\'ee.
+Ici $\Delta$ est n\'egatif et on a $\sqrt{-\Delta} \sim 2 \sqrt{17}$, ce qui conduit \`a $\alpha =-0.1$ et $\omega \sim \sqrt{17}$.
+La solution g\'en\'erale de l'\'equation sans second membre s'\'ecrit donc :}
+\begin{center}
+
+{$y= e^{- 0.1t}[A\  \text{sin}(\sqrt{17} t) + B\  \text{cos} (\sqrt{17} t)]$ , A et B \'etant deux r\'eels.}
 \end{center}
@@
 \begin{center}
-{$[- 17C -2 \sqrt{17} D+17C]\ \text{sin}(\sqrt{17} t) + [-17 D+2\sqrt{17} C +17D]\ \text{cos}(\sqrt{17} t) = \text{cos}( \sqrt{17} t)$}
+{$[- 17C -\frac{1}{5} \sqrt{17} D+17C]\ \text{sin}(\sqrt{17} t) + [-17 D+\frac{1}{5}\sqrt{17} C +17D]\ \text{cos}(\sqrt{17} t) = \text{cos}( \sqrt{17} t)$}
 \end{center}
@@
 \begin{center}
-{$D=0$ et $C=\frac{1}{2\sqrt{17}}$.}
+{$D=0$ et $C=\frac{5}{\sqrt{17}}$.}
 \end{center}
@@
 \begin{center}
-{$y= \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$,}
+{$y= \frac{5}{\sqrt{17}}\ \text{sin}(\sqrt{17} t)$,}
 \end{center}
@@
 \begin{center}
-{$y= e^{- t} [A\  \text{sin}(4 t) + B\  \text{cos}(4 t)] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$, A et B \'etant deux r\'eels}
+{$y= e^{- 0.1t} [A\  \text{sin}(\sqrt{17} t) + B\  \text{cos}(\sqrt{17} t)] + \frac{5}{\sqrt{17}}\ \text{sin}(\sqrt{17} t)$, A et B \'etant deux r\'eels}
 \end{center}
-\item \textbf{(1 pts)} Pour trouver la solution particuli\`ere de l'\'equation compl\`ete avec second membre, associ\'ee aux conditions aux limites $y(t=0)=1$ et $y'(t=0)=-1$ on calcule $y(t=0)$ et $y'(t=0)$:
+
+\item \textbf{(0.5 pts)} {Pour trouver la solution particuli\`ere de l'\'equation compl\`ete avec second membre, associ\'ee aux conditions aux limites $y(t=0)=1$ et $y'(t=0)=-1$ on calcule $y(t=0)$ et $y'(t=0)$:}
 
 \begin{center}
-{$y(t=0)= B+ \frac{1}{2\sqrt{17}}*0=1$, }
+{$y(t=0)= B=1$, }
 
-{$y'(t=0)= [4A - B + \frac{1}{2}]=-1$.}
+{$y'(t=0)= [\sqrt{17}A - 0.1 B + 5]=-1$.}
 \end{center}
 
-{On obtient donc $B=1$ et $A=-\frac{1}{8}$ et la solution s''ecrit :}
+{On obtient donc $B=1$ et $A=-\frac{5,9}{\sqrt{17}}$ et la solution s'\'ecrit :}
 
 \begin{center}
-{$y= e^{- t} [-\frac{1}{8} \  \text{sin}(4 t) +  \text{cos}(4 t) ] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$.}
+{$y= e^{- 0.1 t} [-\frac{5,9}{\sqrt{17}} \  \text{sin}(\sqrt{17} t) +  \text{cos}(\sqrt{17} t) ] + \frac{5}{\sqrt{17}}\ \text{sin}(\sqrt{17} t)$.}
 \end{center}

### 2020-07-10 ensignement/TD_math/correction CC1/main.tex
tout d'un coup

\documentclass[a4paper, 11pt]{article}

%% Language and font encodings
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}

\usepackage[sfdefault,lf]{carlito}
%% The 'lf' option for lining figures
%% The 'sfdefault' option to make the base font sans serif
\usepackage[parfill]{parskip}
\usepackage{framed}
%\renewcommand*\oldstylenums[1]{\carlitoOsF #1}
\usepackage{fancyhdr}
\usepackage{authblk}
\setlength{\headheight}{41pt}
\usepackage[a4paper,top=3cm,bottom=2cm,left=2cm,right=2.5cm,marginparwidth=1.75cm]{geometry}
\usepackage[unicode=true,pdfusetitle,bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,breaklinks=false,pdfborder={0 0 0},backref=false,colorlinks=true,allcolors=magenta]{hyperref}
%% Sets page size and margins

%\usepackage[left=1.5cm, right=2.5cm, marginparwidth=3.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}
\usepackage{mathtools, physics}
\usepackage{graphicx}
\setcounter{secnumdepth}{3}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[textsize=small]{todonotes}
\presetkeys{todonotes}{color=yellow!20}{}
\newcommand{\CZ}{\mathit{CZ}}
\newcommand{\Ket}[1]{\ket{#1}} %adapt syntax of the braket package to the one of the physics package
\newcommand{\ksq}{\Ket{{}\mathop{\square}{}}}
\newcommand{\sq}{{}\mathop{\square}{}}%spaces to be like ×
%next needed because empty still takes room for missing indices
\NewDocumentCommand{\decosq}{O{} O{} O{} O{}}%
%        {\Ket{{\scriptstyle{#4}}\overset{#1}{\underset{#3}{\square}}{\scriptstyle{#2}}}}
        {{\scriptstyle{#4}}%
         \smashoperator{\mathop\square\limits^{#1}_{#3}}%
        {\scriptstyle{#2}}}
\NewDocumentCommand{\decoksq}{O{} O{} O{} O{}}%
         {\Ket{\decosq[#1][#2][#3][#4]}}
\newcommand{\fred}[2][]{\todo[color=red!30, #1]{\textbf{FG:} #2}}
%\newcommand{\kX}{\Ket{\times}}
\NewDocumentCommand{\kX}{O{} O{}}{\Ket{\prescript{#1}{#2}\times}}
\NewDocumentCommand{\mX}{O{} O{}}{\prescript{#1}{#2}\times}
\usepackage{colortbl}
\newcommand{\cm}[1]{\textcolor{blue}{#1}}
\newcommand{\intg}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}

%% Useful packages

%\renewcommand{\headrulewidth}{0pt}
\pagestyle{plain}
\title{Exercice 2}
\author{}

\begin{document}
\maketitle
\section{La première [3pt; 1pt pour la décomposition en cylindre, 2pt pour le résultat]}
Le cylindre est un objet formidable (observez la figure \ref{fig:cyl} si vous ne me croyez pas), surtout lorsqu'il a un rayon R et une hauteur h. La valeur de son volume va vous étonner ! Calculez-la en utilisant une intégrale simple. (Pour ce faire, vous vous rappelerez, avec nostalgie, l'équation d'un cercle de rayon R : $x^2+y^2=R^2$).

\begin{centering}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{cylindre.png}
    \caption{Un cylindre, dans toute sa simplicité}
    \label{fig:cyl}
\end{figure}

\end{centering}

\subsection{Soluce}
On décompose notre cylindre en petit cylindre de hauteur $\intg z$ et additionne les volumes de tous ces cylindres pour trouver le volume du cylindre. Le volume de chacun de ces cylindres est $\pi r(z)^2 \intg y$. Il ne reste plus qu'à trouver $r(z)$.\newline
On connaît les équations du cylindre $x^2 + y^2 = R^2$ et $0\leq z \leq h$. Pour chacun de ces cylindres, on est dans un plan parallèle à $xOy$, c'est à dire qu'on est à $z$ fixé. On connaît l'équation d'un cercle dans un tel plan $x^2+y^2 = r(z)^2$, on en déduit que $x^2 + y^2 = R^2 = r(z)^2$. Donc, le volume du cylindre pour un $z$ donné est $\pi R^2 \intg z $, il ne reste plus qu'à sommer :
\begin{equation}
    V_{\mathrm{cyl}} = \displaystyle \int_{0}^{h} \pi R^2 \, \intg z = \pi R^2 h
\end{equation}

\section{La deuxième [2pt; 0.5pt pour la décompo en cylindre, 1pt pour avoir posé la bonne intégrale, 0.5 pour le résultat]}
Les toupies ne sont pas mal non plus. Celle-ci possède le petit désavantage d'être infinie. Mais son équation cartésienne est plutôt lisible :
\begin{equation*}
\left\{ 
        \begin{array}{rl}
    x^2+y^2&=\frac{1}{\sqrt{\abs{z}}}\\
    \abs{z}&\leq 5
    \end{array}
    \right.
\end{equation*}
Trouvez la valeur du volume de cette toupie (que vous pouvez admirer sur la figure \ref{fig:toup}) en utilisant une intégrale.

\subsection{Soluce}
On décompose la toupie en petit cylindre de hauteur $\intg z$ et on  additionne les volumes de tous ces cylindres pour trouver le volume de la toupie. Le volume de chacun des cylindres est $\pi r(z)^2 \intg y$.\newline
A $z$ fixé, on connaît l'équation d'un cercle dans le plan $xOy$: $x^2+y^2 = \frac{1}{\sqrt{\abs{z}}}$, on en déduit que $x^2 + y^2 =  \frac{1}{\sqrt{\abs{z}}} = r(z)^2$. Donc, le volume du cylindre pour un $z$ donné est $\pi  \frac{1}{\sqrt{\abs{z}}} \intg z $, il ne reste plus qu'à sommer :
\begin{equation}
    V_{\mathrm{toupie}} = \displaystyle \int_{-5}^{5} \pi  \frac{1}{\sqrt{\abs{z}}} \, \intg z = 2\displaystyle \int_{0}^{5} \pi  \frac{1}{\sqrt{z}} \, \intg z
\end{equation}
Qui est une intégrale généralisée. On trouve finalement
\begin{equation}
    V_{\mathrm{toupie}} = 2\pi\lim \limits_{\epsilon \rightarrow 0}\left[2\sqrt{z} \right]_\epsilon^5 = 4\pi\sqrt{5}
\end{equation}

\begin{centering}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{toupie}
    \caption{Une toupie, dans toute son infinité}
    \label{fig:toup}
\end{figure}

\end{centering}
\end{document}

### 2020-07-10 experience/fiche matos/fiche.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Le matos}
  \section{Attocube}
  \subsection{Axes}
  Susceptibles de changer si tu rebranches mal derrière
  \begin{itemize}
  \item Gauche : axe y, haut=haut, bas=bas
  \item Milieu : axe z, haut=avant , bas=arrière
  \item Droite : axe x, ???
  \end{itemize}
  \subsection{Positioners}
  ref : https://www.attocube.com/en/products/nanopositioners/low-temperature-nanopositioners/anpz51ltuhv-linear-z-nanopositioner

  et https://www.attocube.com/en/products/nanopositioners/low-temperature-nanopositioners/anpx51ltuhv-linear-x-nanopositioner (à vérifier)

  \subsubsection{Masse limite}
  La structure c'est que t'as le z-positionner en bas (pour y) qui peut porter 50g et les deux x-positionners dessus (pour x et z) qui pèsent 7 g chacun et peuvent porter 25g.

  Donc a priori on ne peut pas porter qqchose de plus de 18g.
  \section{Cryo}
  \subsection{Références}
  http://www.mycryofirm.com/nos-produits/page-accueil/
  \subsection{Allumage}
  \begin{itemize}
  \item Pomper jusqu'à $10^{-2}$ au moins
  \item Allumer le cooler (gros truc à l'entrée, t'as le bouton d'arrêt d'urgence dessus et le bouton on sur le tableau de bord à droite sur le mur).
  \item Allumer le cryo dans la salle des serveurs (gros bouton vert).
  \item Éteindre la pompe (et fermer la vanne) avant d'atteindre la température de liquéfaction de l'air.
  \end{itemize}
  \subsection{Sonde thermique}
  Tu as un doigt froid qui vient de l'enceinte vers la chambre à vide et qui arrive en dessous de la pile atto-cube. Tu as un filament doré qui fait le contact thermique avec une petite plaque au sommet de la pile atto-cube, sur laquelle tu pose ton porte-échantillon. C'est au niveau de cette plaque qu'est la sonde de platine.
 \section{Thermomètre (?)}
 A compléter, mais en gros tu atteins la température minimum aux alentours de $1000 \; \Omega$.

 Utilisation de la fonction chauffage : appuyer sur "loop1", fixer la températture visée avec la valeur de la résistance, appuyer sur commande pour lancer le truc (c'est une boucle de rétro-action) et appuyer sur "stop" pour arrêter. (et "home" pour revenir aux températures.
 \section{spectro}
 \subsection{Résolution}

  \end{document}

### 2020-07-10 idées/To_do_list.tex
tout d'un coup

--- 2020-07-10 experience/fiche matos/fiche.tex
@@
 \documentclass[a4paper]{report}
-\usepackage[french]{babel}
 \usepackage[utf8]{inputenc}
@@
 \begin{document}
-\chapter{Le matos}
-  \section{Attocube}
-  \subsection{Axes}
-  Susceptibles de changer si tu rebranches mal derrière
+\chapter{La liste que je regarderai jamais}
+  \section{Théorie}
   \begin{itemize}
-  \item Gauche : axe y, haut=haut, bas=bas
-  \item Milieu : axe z, haut=avant , bas=arrière
-  \item Droite : axe x, ???
+  \item Changement de référentiel, représentation de Heisenberg etc...
+  \item Spins : additions \& co
+  \item Règles de sélection
+  \item ESR, CPT et co
+  \item origine du dipole de transition, lien avec la polarisibilité etc.
+  \item un peu d'optics f2f
   \end{itemize}
-  \subsection{Positioners}
-  ref : https://www.attocube.com/en/products/nanopositioners/low-temperature-nanopositioners/anpz51ltuhv-linear-z-nanopositioner
-
-  et https://www.attocube.com/en/products/nanopositioners/low-temperature-nanopositioners/anpx51ltuhv-linear-x-nanopositioner (à vérifier)
-
-  \subsubsection{Masse limite}
-  La structure c'est que t'as le z-positionner en bas (pour y) qui peut porter 50g et les deux x-positionners dessus (pour x et z) qui pèsent 7 g chacun et peuvent porter 25g.
-
-  Donc a priori on ne peut pas porter qqchose de plus de 18g.
-  \section{Cryo}
-  \subsection{Références}
-  http://www.mycryofirm.com/nos-produits/page-accueil/
-  \subsection{Allumage}
-  \begin{itemize}
-  \item Pomper jusqu'à $10^{-2}$ au moins
-  \item Allumer le cooler (gros truc à l'entrée, t'as le bouton d'arrêt d'urgence dessus et le bouton on sur le tableau de bord à droite sur le mur).
-  \item Allumer le cryo dans la salle des serveurs (gros bouton vert).
-  \item Éteindre la pompe (et fermer la vanne) avant d'atteindre la température de liquéfaction de l'air.
-  \end{itemize}
-  \subsection{Sonde thermique}
-  Tu as un doigt froid qui vient de l'enceinte vers la chambre à vide et qui arrive en dessous de la pile atto-cube.
-Tu as un filament doré qui fait le contact thermique avec une petite plaque au sommet de la pile atto-cube, sur laquelle tu pose ton porte-échantillon.
-C'est au niveau de cette plaque qu'est la sonde de platine.
- \section{Thermomètre (?)}
- A compléter, mais en gros tu atteins la température minimum aux alentours de $1000 \; \Omega$.
-
- Utilisation de la fonction chauffage : appuyer sur "loop1", fixer la températture visée avec la valeur de la résistance, appuyer sur commande pour lancer le truc (c'est une boucle de rétro-action) et appuyer sur "stop" pour arrêter.
-(et "home" pour revenir aux températures.
- \section{spectro}
- \subsection{Résolution}
 

### 2020-07-10 manips/Hole_Burning_SiV_NanoP/manip.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Hole burning de SiV dans des nano-pyramides CVD}
\subsubsection{Date}
2019/09/16 $\to$ 2019/10
  \section{But}
  On étudie les pointes d'AFM que Louis a déja étudié dans l'article d'ACS. Le but dans un premier temps c'est d'observer un splitting zeeman sur une des 4 raies en faisant un trou/anti-trou, comme dans le PRL 2014 de Lukin.

  Edit 15/10 : Le but a un peu changé, en faite on a réalisé qu'on pourrait observer Zeeman sans avoir à bouger l'aimant : si tu creuses un trou avec un champ mag déja présent, étant donné le T1 de spin des SiV (~40 ns), il va se creuser instantanément des petits pics à coté qui correspondent au spin $\pm 1/2$ complémentaire des spins que tu creuses au milieu.
  \section{Éléments de la manip}
  \subsection{L'échantillon}
  Les pyramides sont sur un wafer de silicone. Ici pour pouvoir approcher un aimant le plus près possible on a décidé de mettre des pyramides sur une plaque de quartz (cf après). J'ai nettoyé des vieilles plaques de quartz (acéone + IPA) et pour faire l'échantillon j'ai simplement tapoté la plaque sur le wafer, et vérifié la présence de pyramides au microscope (objectif x20 vert).
  \subsection{Setup dans le cryostat}
  (Inclure photo)

  Le setup est le suivant : l'échantillon est placé sur un porte-échantillon en t, de l'autre coté de l'objectif (coté aimant donc). L'objectif utilisé est le jaune x10 pour être placé en dehors de l'écran thermique (en vrai c'est limite, l'objectif touche quasiment le hublot et l'attocube est quasiment en bout de course vers l'avant.

  L'aimant pour l'instant est un aimant fin de Nd qui es a priori orienté dans le mauvais sens (le champ sort de la "tranche", et pas du "bout"), et monté sur la platine motorisé (ci-après "le tchou-tchou"), il arrive a quelques mm de l'échantillon et peut s'écarter de 1.5cm.

  \subsubsection{Améliorations}
  Utiliser les aimants ronds, vissés, plutôt que l'aimant fin. Vérifier les champs avec un tesla-mètre pour voir si on peut pas retourner le porte-échantillon.

  \subsection{Influence de la température}
  Déjà, en fonction des pointes, tu as les 4 pics qui peuvent se distinguer ou non. Mais en plus t'as des effets assez fourbes avec la température, notamment au niveau du chauffage du laser : Pour l'instant (25/09) j'utilise une densité 0.15 (de la roue car les grosses densité dévie le faisceau), et faut aussi optimiser la ou tu tapes. Je m'explique :

  Je pense qu'en fonction de si tu excites beaucoup de NV ou pas, ton échantillon chauffe plus ou moins (pk ?). Du coup faut réussir à exciter un max de SiV sans trop de NV. Pour ça ma technique c'est de passer en laser pleine puissance (pour avoir du signal au spectro en mode caméra), et de bouger un peu jusqu'à voir les raies se séparer légèrement, sans perdre trop de signal (regarder le moment ou les raies sont grandes mais la base est petite).

  \subsubsection{Note}
  Normalement, avec un hublot d'ouvert j'arrive à une température de $\approx 1020 \omega $, ce qui est suffisant, même avec le laser à fond, pour voir la séparation de 250 GHz (niveaux excités) au spectro (réseau vert ou rouge). Par contre il faut passer au réseau rouge pour voir la séparation à 50GHz du fondamental.

  \subsubsection{Amélioration}
  Mettre du papier d'alu sur le hublot de l'aimant autour de l'aimant (p-e moins utile avec l'aimant rond, à vérifier).

  Edit 15/10 : Effectivement l'aimant rond fait un plutot bon écran thermique. D'après Louis il est sensé être à 300K et donc par rayonnement il devrait réchauffer l'aimant, mais c'est pas ce que j'observe : en raprochant l'aimant, la température diminue (donc je pense que la convection est plus importante que le rayonnement). Si il faut, on peut relier thermiquement l'aimant à l'écran thermique, mais ce sera plus dur de le déplacer.

  \subsection{Programmes Labview (outdated) (je m'avance un peu)}
  \subsubsection{Hole burning}
  \subsubsection{Principe}
  Le programme va burn pendant un certain temps (c'est à dire fermer le laser vert)
  \subsubsection{Branchements}
  Avec les post-its actuels, le signal de commande es en "signal 1", l'APD dans "In APD", le shutter rouge dans "out shutter 1", le shutter vert dans "out shutter 2".
  \subsubsection{Instructions}
  Le programme que j'utilise pour l'instant (25/09) c'est HoleBurning V2. Les paramètres sont les suivants : \begin{itemize}
  \item tps intégration 1 \& 2 : mettre la même valeur que dt (valeur en s, typiquement 1e-5)
  \item plage totale en V : amplitude du scan (valeur en V, typiquement 1)
  \item pas en V : le pas, typiquement 1e-4
  \item offset en V : bah l'offset, 0 par défaut
  \item dt : le temps que tu vas passer sur chaque point en s. Le nombre de point étant défini comme plage en V/pas en V, le temps total d'un cycle vaut plage/pas*dt (en sachant qu'il y a un break de 0.1s entre chaque cycle, donc ça sert à rien de vouloir descendre en dessous. Aussi le programme ne marche pas pour d=1e-6)
  \item nb burning : ???
  \item nb itérations : le nombre de cycle. Mettre 999999 et arrêter quand t'en as marre
  \item rate : mettre une valeur plus grand que dt, genre 999 (pk ??)
  \item 
  \end{itemize}

  \end{document}

### 2020-07-10 manips/Influence Inter dipolaire pyramide SiV/manip.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Non influence de l'interaction dipolaire sur les nano-pyramides de SiV}
\subsubsection{Date}
2019/12 pendant une petite semaine
  \section{But}
  Le but était d'observer la modification des largeurs (homogènes et inhomogènes) dues à l'interaction dipolaire entre les différents centres SiV. Pour ça l'idée est de réduire progressivement la quantité de centres SiV en les brulant de façon homogène (avec une ou plusieurs rampes) et de voir comment changent : la bosse totale (dominée par la largeur inhomogène) et les trous creusés (largeur homogène élargie par puissance).
  \section{Méthodes expérimentales}
  Globalement similaire au hole-burning, sauf que la séquence de burning inclue des rampes, en plus de l'éventuel trou à creuser.

  Deux points importants : 
  \begin{itemize}
  \item Le laser est loin d'être monomode : ça se voit très bien sur cette manip ou tu peux creuser plusieurs trous en même temps. Pour éviter ça : Michelson déja, et ensuite jouer sur le courant et la température, l'essentiel étant d'avoir des franges d'interférences bien contrastées sur l'ensemble du scan. Les sauts de mode c'est gênant mais moins important. (et je suis pas sur qu'on puisse faire 3V de scans sans saut de modes)
  \item Pas oublier la puissance du laser qui peut complètement changer la tronche du scan. En particulier si tes scans commencent toujours par une descente de PL, c'est sans doute que le laser est trop fort.
  \end{itemize}
  \section{Résultats}
 Les résultats sont pour l'instant dans le dossier python la haut, il faudra penser à les ramener.
 Concrètement on a vu aucune modification, ni sur la largeur homogène, ni sur la largeur inhomogène (plus logique).
  \end{document}

### 2020-07-10 manips/PL en fonction de l'angle de B/Double quantums/fiche.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Double quantum}
  \section{Principe}
  L'idée c'est que dans le hamiltonien d’interaction dipolaire tu as des termes en $|0-><+0|$ qui sont souvent négligés (en tout cas dans les travaux de Lukin) car non résonnants d'un point de vue énergie, mais si tu travailles à champ quasi nul et que le strain est limité, ils commencent à jouer et à modifier le T1. (jusqu'à un facteur 5 par rapport à une théorie sans double-quantum d'après Gabriel).
  Le principal soucis pour les observer, c'est le strain qui empêche la résonance (il faudra aussi un peu clarifier l'intercation dipolaire dans la base mixée par le strain, ainsi que le moyennage dans l'espace, mais c'est du fine-tuning je pense).

  Un des interêts pour moi des double quantum, ce serait l'idée que si on arrivait par ingénérie à limiter le strain suffisament par rapport au T2, on aurait un détecteur de bruit magnétique en champ nul, un peu comme ce qu'on a déja avec les adamas (la PL monte quand on approche un aimant) mais en bcp plus sensible : deja l'effet est plus violent (facteur 5 etc), mais surtout les raies s'éloignent les unes des autres plus vite que dans le cas des simples quantum (elles vont dans des directions différentes, plutot qu'à des vitesses différentes dans la même direction).

  \section{Manips}
  Les manips ont globalement commencées à la reprise après le confinement (le premier lol), le 18/05/2020.
  \subsection{Scan du champ mag selon la 100}
  Première manip réalisée, l'idée est assez simple : mesurer un dip de PL autour de 0G en alignant l'Electro-aiamant (EM) selon la direction 100.

  Pas de résultats sur le rose il me semble, en revanche un petit résultat sur le sumitumo 4. Dans les jours qui viennent ce sera sans doute le point à pousser avec un sétup similaire mais sur des adamas à faible strain.
  \subsubsection{astuces}
  Pour aligner tout ça j'ai refait un sétup à bas de monture de lambda/2 (j'ai utilisé le socle rotatif pour la rotation principale) ou j'ai plaqué l'EM avec de la patafix sur la monture de lambda/2.

  L'aspect un peu sioux c'est qu'on voyait pas grand chose à cause de l'hystéresis de l'EM, du coup pour bien passer par 0G je met le courant à fond dans l'autres sens avant chaque scan. Sauf que le générateur de courant peut pas prendre du courant négatif donc je suis obligé de débrancher et rebrancher pour chaque scan avec un programme qui fait bip-bip. Ce que j'ai réalisé avec les manips de bruit c'est qu'un GBF peut quand même cracher un peu de patate (10 VPP dans du 50hms donc 2W, plus que ce que j'utilise avec le générateur de courant), donc je pense que je me suis fait chier pour rien mais c'était marrant.

 Pour ce qui est de l'alignement j'ai l'impression que c'est assez robuste, parceque quand j'étais un peu désaligné je voyais des remontées au moment ou les raies se splittent. Ce sera à confirmer par des calculs/simu je pense.

 \subsection{Variation de T1 pour des Adamas de différents strains}
 La l'idée c'est de comparer les T1 dans un configuration 4x0 et en champ nul, pour des adamas avec un fort strain et des adamas avec faible strain, dans l'idée que dans le premier cas on observe pas de différence, alors qu'on en observe dans le second cas.

 Le gros soucis ici bah c'est la mesure de T1. Je l'ai fait avec la méthode de la soustraction avec un pulse pi des polaks, mais avec la pulse pi à la fin du tau, comme Lukin. En gros dans la config des polaks t'as un problème si ton temps de vie dans l'état |-1> est pas le meme que celui dans l'état |0>, ce qui semble être le cas ici. Et l'autre gros problème c'est que si tu fais ta pulse au début et que tu laisses l'AOM en continue et pas en pulsé, bah tu mesures quand meme un temps de vie, mais un temps de vie avec le laser allumé (temps de repolarisation). Et ça peut prendre 3 jours de s'en rendre compte...

 Mais même avec la soustraction il y l'air d'avoir des blagues : Lukin laisse un temps de retour à l'équilibre des charges (j'ai essayé mais ça fait rien chez moi), L'influence du laser a pas l'air énorme non plu, par contre j'ai essayé de mettre une micro-onde plus faible et la je mesure des tdv bcp plus courts. J'ai aussi vu des fois des petites remontés au début ou des courbes qui ont pas la même tronche quand on zoom...

 Et surtout le gros problème que j'ai actuellement c'est le fit exp Vs stretch. Sans que je comprenne trop la logique, certaines courbes se fittent mieux avec l'un ou l'autre, et le temps de vie obtenus sont pas comparabales (deja que les temps de vie stretch sont difficilement comparables... Tu peux prendre un facteur 3 juste en zoomant ou en dézoomant. Il faudrait un programme avec des points rapprochés au début mais quand même une grande queue)

 \subsection{Elargissement des raies par bruit magnétique}
 Bon la c'est un échec mais j'ai vu des trucs rigolo.

 Le but c'était d'élargir (homogènement) les raies en champ nul pour qu'il y ait un recouvrement plus important des double quantums. En gros le soucis c'est qu'on a un croisement évité des états +/- avec un ZFS de qq10MHz, et d'après Gabriel, et je veux bien le croire, pour battre ce croisement il faut un bruit avec une fréquence > ces qq10MHz, ce qui élimine toutes les bobines et EM (fréquence de coupure que j'ai mesuré entre 300Hz pour le gros EM et 15KHz pour une bobine avec dans les deux cas un R de 50Ohms en série).

 J'ai quand même fait l'expérience avec un bruit à 100K et l'EM en pointe (branché direct à l'ampli qui descend à 100KHz et avec un signal GBF) et j'observe bien un élargissement sur les bords, mais pas au centre (cf 11/06/2020) L'allure du truc est rigolo quand meme, mais j'ai pas trop poussé parce que l'EM chauffait vite. En faite je pense que les paliers viennent du fait que les différentes classes de NV n'oscillent pas avec la même amplitude (ça dépend de leur orientation avec le champ mag)

 Du coup l'autre idée c'est d'utiliser une mirco-onde avec un fréquence assez élevée (mais pas résonnante) comme source de bruit magnétique. J'ai l'impression d'avoir rien vu mais faudrait quand meme mettre les chifres dessus sachant que j'ai des pulses pi de 50ns avec 0dB de uW (sur le R\&S avec le switch à -11dB et l'atténuateur à -3dB donc en théorie y'a encore un peu de marge).

 Et ma dernière idée c'était d'utiliser le couplage dipolaire aux autres centres NV, en mettant une uW à résonnance avec une autre classe de NV. En faite c'est assez similaire aux manips de DEER qu'à fait Lukin par exemple dans le PRL de 2018, et comme l'a fait remarquer Gabriel c'est des modifications à l'échelle du T2écho, qui est très différent du T2réél en faite. Donc a priori c'est pas réaliste.

 Par contre en faisant mumuse avec les deux uW j'ai pu observer les side-band du C13 first-shell autour de 2.8 et 2.95Ghz (en champ nul en mettant la uW à résonnance à 2.87GHz), cf 12/06/2020. C'est un peu rigolo en vrai.
  \end{document}

### 2020-07-10 manips/PL en fonction de l'angle de B/Intercation dipolaire/interaction_dipolaire.tex
tout d'un coup

--- 2020-07-10 manips/PL en fonction de l'angle de B/Double quantums/fiche.tex
@@
 \begin{document}
-\chapter{Double quantum}
-  \section{Principe}
-  L'idée c'est que dans le hamiltonien d’interaction dipolaire tu as des termes en $|0-><+0|$ qui sont souvent négligés (en tout cas dans les travaux de Lukin) car non résonnants d'un point de vue énergie, mais si tu travailles à champ quasi nul et que le strain est limité, ils commencent à jouer et à modifier le T1.
-(jusqu'à un facteur 5 par rapport à une théorie sans double-quantum d'après Gabriel).
-  Le principal soucis pour les observer, c'est le strain qui empêche la résonance (il faudra aussi un peu clarifier l'intercation dipolaire dans la base mixée par le strain, ainsi que le moyennage dans l'espace, mais c'est du fine-tuning je pense).
+\chapter{Interaction dipolaire}
+  \section{Hamiltonien général}
+  Le hamiltonien général est :
+  \begin{equation}
+  \hat{\mathcal{H}}_{dd}= \dfrac{\mu_0}{4\pi}\gamma_1\gamma_2\hbar^2\dfrac{1}{r^3}(\hat{\mathbf{S_1}} \cdot \hat{\mathbf{S_2}} - 3(\hat{\mathbf{S_1}} \cdot \mathbf{r})(\hat{\mathbf{S_2}} \cdot \mathbf{r}))
+  \end{equation}
+  \begin{center}
+  Avec $gamma_i$ le rapport gyromagnétique du spin i et $J_0=\dfrac{\mu_0}{4\pi}\gamma_1\gamma_2\hbar^2=52\, \rm MHz / nm^3$ pour l'interaction $\rm NV-NV$
+\end{center}
 
-  Un des interêts pour moi des double quantum, ce serait l'idée que si on arrivait par ingénérie à limiter le strain suffisament par rapport au T2, on aurait un détecteur de bruit magnétique en champ nul, un peu comme ce qu'on a déja avec les adamas (la PL monte quand on approche un aimant) mais en bcp plus sensible : deja l'effet est plus violent (facteur 5 etc), mais surtout les raies s'éloignent les unes des autres plus vite que dans le cas des simples quantum (elles vont dans des directions différentes, plutot qu'à des vitesses différentes dans la même direction).
+  \section{Réécriture pour deux spins 1}
+  La tout de suite j'ai la flemme de me taper des heures de latex, en gros il faut se baser sur le supplementary de l'article "critical thermalization" de Lukin, il y a trois sortes de termes :
+  \begin{
 
-  \section{Manips}
-  Les manips ont globalement commencées à la reprise après le confinement (le premier lol), le 18/05/2020.
-  \subsection{Scan du champ mag selon la 100}
-  Première manip réalisée, l'idée est assez simple : mesurer un dip de PL autour de 0G en alignant l'Electro-aiamant (EM) selon la direction 100.
-
-  Pas de résultats sur le rose il me semble, en revanche un petit résultat sur le sumitumo 4.
-Dans les jours qui viennent ce sera sans doute le point à pousser avec un sétup similaire mais sur des adamas à faible strain.
-  \subsubsection{astuces}
-  Pour aligner tout ça j'ai refait un sétup à bas de monture de lambda/2 (j'ai utilisé le socle rotatif pour la rotation principale) ou j'ai plaqué l'EM avec de la patafix sur la monture de lambda/2.
-
-  L'aspect un peu sioux c'est qu'on voyait pas grand chose à cause de l'hystéresis de l'EM, du coup pour bien passer par 0G je met le courant à fond dans l'autres sens avant chaque scan.
-Sauf que le générateur de courant peut pas prendre du courant négatif donc je suis obligé de débrancher et rebrancher pour chaque scan avec un programme qui fait bip-bip.
-Ce que j'ai réalisé avec les manips de bruit c'est qu'un GBF peut quand même cracher un peu de patate (10 VPP dans du 50hms donc 2W, plus que ce que j'utilise avec le générateur de courant), donc je pense que je me suis fait chier pour rien mais c'était marrant.
-
- Pour ce qui est de l'alignement j'ai l'impression que c'est assez robuste, parceque quand j'étais un peu désaligné je voyais des remontées au moment ou les raies se splittent.
-Ce sera à confirmer par des calculs/simu je pense.
-
- \subsection{Variation de T1 pour des Adamas de différents strains}
- La l'idée c'est de comparer les T1 dans un configuration 4x0 et en champ nul, pour des adamas avec un fort strain et des adamas avec faible strain, dans l'idée que dans le premier cas on observe pas de différence, alors qu'on en observe dans le second cas.
-
- Le gros soucis ici bah c'est la mesure de T1.
-Je l'ai fait avec la méthode de la soustraction avec un pulse pi des polaks, mais avec la pulse pi à la fin du tau, comme Lukin.
-En gros dans la config des polaks t'as un problème si ton temps de vie dans l'état |-1> est pas le meme que celui dans l'état |0>, ce qui semble être le cas ici.
-Et l'autre gros problème c'est que si tu fais ta pulse au début et que tu laisses l'AOM en continue et pas en pulsé, bah tu mesures quand meme un temps de vie, mais un temps de vie avec le laser allumé (temps de repolarisation).
-Et ça peut prendre 3 jours de s'en rendre compte...
-
- Mais même avec la soustraction il y l'air d'avoir des blagues : Lukin laisse un temps de retour à l'équilibre des charges (j'ai essayé mais ça fait rien chez moi), L'influence du laser a pas l'air énorme non plu, par contre j'ai essayé de mettre une micro-onde plus faible et la je mesure des tdv bcp plus courts.
-J'ai aussi vu des fois des petites remontés au début ou des courbes qui ont pas la même tronche quand on zoom...
-
- Et surtout le gros problème que j'ai actuellement c'est le fit exp Vs stretch.
-Sans que je comprenne trop la logique, certaines courbes se fittent mieux avec l'un ou l'autre, et le temps de vie obtenus sont pas comparabales (deja que les temps de vie stretch sont difficilement comparables...
-Tu peux prendre un facteur 3 juste en zoomant ou en dézoomant.
-Il faudrait un programme avec des points rapprochés au début mais quand même une grande queue)
-
- \subsection{Elargissement des raies par bruit magnétique}
- Bon la c'est un échec mais j'ai vu des trucs rigolo.
-
- Le but c'était d'élargir (homogènement) les raies en champ nul pour qu'il y ait un recouvrement plus important des double quantums.
-En gros le soucis c'est qu'on a un croisement évité des états +/- avec un ZFS de qq10MHz, et d'après Gabriel, et je veux bien le croire, pour battre ce croisement il faut un bruit avec une fréquence > ces qq10MHz, ce qui élimine toutes les bobines et EM (fréquence de coupure que j'ai mesuré entre 300Hz pour le gros EM et 15KHz pour une bobine avec dans les deux cas un R de 50Ohms en série).
-
- J'ai quand même fait l'expérience avec un bruit à 100K et l'EM en pointe (branché direct à l'ampli qui descend à 100KHz et avec un signal GBF) et j'observe bien un élargissement sur les bords, mais pas au centre (cf 11/06/2020) L'allure du truc est rigolo quand meme, mais j'ai pas trop poussé parce que l'EM chauffait vite.
-En faite je pense que les paliers viennent du fait que les différentes classes de NV n'oscillent pas avec la même amplitude (ça dépend de leur orientation avec le champ mag)
-
- Du coup l'autre idée c'est d'utiliser une mirco-onde avec un fréquence assez élevée (mais pas résonnante) comme source de bruit magnétique.
-J'ai l'impression d'avoir rien vu mais faudrait quand meme mettre les chifres dessus sachant que j'ai des pulses pi de 50ns avec 0dB de uW (sur le R\&S avec le switch à -11dB et l'atténuateur à -3dB donc en théorie y'a encore un peu de marge).
-
- Et ma dernière idée c'était d'utiliser le couplage dipolaire aux autres centres NV, en mettant une uW à résonnance avec une autre classe de NV.
-En faite c'est assez similaire aux manips de DEER qu'à fait Lukin par exemple dans le PRL de 2018, et comme l'a fait remarquer Gabriel c'est des modifications à l'échelle du T2écho, qui est très différent du T2réél en faite.
-Donc a priori c'est pas réaliste.
-
- Par contre en faisant mumuse avec les deux uW j'ai pu observer les side-band du C13 first-shell autour de 2.8 et 2.95Ghz (en champ nul en mettant la uW à résonnance à 2.87GHz), cf 12/06/2020.
-C'est un peu rigolo en vrai.
   \end{document}

### 2020-07-10 manips/PL en fonction de l'angle de B/manip.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Dépendance angulaire du champ magnétique sur un ensemble de NV-}

\section{Théorie}
La théorie pour l'instant c'est le PRL 2017 de Lukin. Je sais pas encore trop quoi en penser. Sans rentrer dans les détails, l'essentiel de ce qui m’intéresse est dans les équations (S36), (S38), (S40), (S41) (tout dans le suplementary). Faut juste prendre en compte que "slightly lower" pour Lukin c'est une diminution de 40\%.
Je trouve pour les différentes configurations de dégénérescences :
\begin{itemize}
\item \textbf{1x1x1x1} : $T_1=\tau$
\item \textbf{1x2x1} : $T_1\approx \dfrac{\tau}{4.1}$
\item \textbf{2x2} : $T_1\approx \dfrac{\tau}{7.3}$
\item \textbf{3x1} : $T_1\approx \dfrac{\tau}{14.7}$
\item \textbf{4x} : $T_1\approx \dfrac{\tau}{37}$
\end{itemize}

La dépendance en NRJ ensuite, c'est en gros une Lorentzienne tant que ton $\gamma_f$ (c'est à dire la largeur spectrale de ton bruit, qui correspond au taux de dépolarisation des fluctuators) est grand devant ton élargissement inhomogène de chaque classe ($\delta \omega$ chez Lukin), ce qui est le cas pour lui et pour nous avec le diamant rose d'Alex et tous les échantillons assez concentrés qu'on a.

Plus précisement, la formule générale c'est celle-la :
\begin{equation}
\dfrac{T_1}{T_{1\mathrm{ref}}} = \dfrac{1}{4} \sum_{i=1}^4 \left(1+\sum_{i \neq j} 1.69 \sqrt{\dfrac{\delta \omega ^2 + 4 \gamma_f^2}{(\delta \omega + \delta_{ij})^2 + 4\gamma_f^2}} \right)^2
\end{equation}

Où $T_{1\mathrm{ref}}$ est le temps de vie obtenu dans la configuration 1x1x1x1 (avec uniquement de l'interaction intra-classe), et où on somme sur les 4 orientations possibles. Le 1.69 vient de $\dfrac{0.6507}{2/3\sqrt{3}}$, c'est ce slightly qui vaut 1 chez Lukin.

Pour la photoluminescence maintenant, on fait simplement une petite équation de tau et on trouve (sans approximation): 
\begin{equation}
\rho_0=\dfrac{\Gamma_{las} + 1/t_1}{\Gamma_{las} + 3/t_1}
\end{equation}
Avec $\Gamma_{las}$ le taux de pompage optique dans l'état $\ket{0}$ et 
\begin{equation}
\dfrac{1}{t_1}=\dfrac{1}{t_1^{phonons}}+\dfrac{1}{t_1^*}(\delta)
\end{equation}

\subsection{Géométrie}
on se place dans le référentiel (x,y,z) ou x,y et z ont les normales aux faces du cube de la maille cristalline. Alors les coordonnées des différentes orientations dans cette base sont :
\begin{equation}
\begin{pmatrix}
1 \\ 1 \\ 1
\end{pmatrix} ,
\begin{pmatrix}
1 \\ -1 \\ -1
\end{pmatrix} ,
\begin{pmatrix}
-1 \\ 1 \\ -1
\end{pmatrix} ,
\begin{pmatrix}
-1 \\ -1 \\ 1
\end{pmatrix} 
\end{equation}
Qu'on peut également noter $(111), (1\bar{1}\bar{1}),(\bar{1}1\bar{1}),(\bar{1}\bar{1}1)$

Alors :
\begin{itemize}
\item Les dégénérescences 4x0 correspondent aux directions $(100), (010), (001)$ ou simplement x,y et z.
\item Les dégénérescences 3x1 correspondent aux directions $(111), (1\bar{1}\bar{1}),(\bar{1}1\bar{1}),(\bar{1}\bar{1}1)$, soit la direction des différents centres.
\item Les dégénérescences 2x2 correspondent aux plans normaux à $(100), (010), (001)$, c'est à dire les plans (xy), (yz) et (xz).
\item Les dégénérescences 1x2x1 correspondent aux plans normaux à :

$(110), (1\bar{1}0),(101), (10\bar{1}),(011), (01\bar{1})$ chacune des ces directions correspondant à une dégénérescence 2x2 particulière. 
\end{itemize}
  \section{Map 2d à la main}
  \subsection{Principes physiques}
  La référence c'est principalement "Depolarization Dynamics in a Strongly Interacting Solid-State Spin Ensemble" PRL 2017 de lukin, où il explique avec un modèle de spin-fluctuators pourquoi le $T_1$ de spin dépend de la dégénérescence des niveaux d'énergies de spin, et donc de l'angle du champ mag.

  L'idée c'est que : Champ aligné avec plusieurs orientations de NV- $\to$ états dégénérés $\to$ $T_1$ plus faible $\to$ Polarisation plus faible dans le fondamental $\to$ Moins de PL

  \subsection{Principe de la manip}
  La manip "à la main" repose sur l'utilisation de montures de $\lambda/2$ pour faire tourner un aimant autour de l'échantillon avec un mouvement sphérique (pour avoir une amplitude de champ constante). Du coup l'échantillon est sur une pointe dure (pour pas bouger, c'est important que la PL soit stable dans le temps), au milieu d'une monture qui peut elle même pivoter, ça te fait un angle de nutation et un angle de précession.
  (Include picture, c'est sur mon portable).

  \subsubsection{Subtilités de la manip}
  Y'en a pas mal, je note comme ça me vient :
  \begin{itemize}
  \item Ce qui m'a le plus bloqué, c'est que pour voir un effet, il faut que le temps de polarisation des spins (par le laser vert, proportionnel à la puissance verte) soit du même ordre de grandeur que le $T_1$ des spins, tu peux faire le calcul mais en gros c'est la ou la variation de $T_1$ modifie le plus la PL.

  Y'a 2 façons de regarder ça : soit à la barbare en mesurant les 2, et en ajustant la puissance du laser avec AOM/densité, soit en regardant la différence de contraste quand tu approches un aimant : avec un échantillon fortement dopé, tu devrais voir une hausse de la PL qui peut monter à 3/4\% je pense, en approchant l'aimant dans une direction quelconque avec la bonne puissance incidente.

  Pour ma carte j'avais une densité 0.1 et 0.3 sur le laser, et une E-3 sur la collection

  \item Stabiliser la photolum : Tu regardes des variations de qq \% au mieux, assez éloignées dans le temps et sans moyennage : il te faut une photolum stable. Le principal truc à retenir c'est d'utiliser une fibre multi-mode. L'autre truc que j'ai pas fait mais qui est à creuser c'est de ne pas focaliser le laser complètement, pour avoir une tache plus grande que l'échantillon. Faut quand même garder une puissance suffisante, cf avant.

  \item Faire de la place : faut de la place pour tourner l'aimant, à la fin j'avais 30 degrés de précession (je suis monté à 40) et genre 270 degrés de nutation. Et pour faire ça il a fallu ruser :

 Déposer l'échantillon sur un truc long et fin (une pointe en métal ici) et approcher l'antenne micro-onde selon le même axe pour pas que ça gêne trop la rotation (l'échantillon est pas posé sur l'antenne car elle chauffe et se déplace, même si en vrai tu pourrais faire tous les ESR à la fin)

 Utiliser une lentille type lentille d'AOM au lieu d'un objectif pour l'éclairage/la collection. Ca limite aussi pas mal la puissance max.

  \item Centrer l'échantillon : la pas trop de techniques, bien penser à regarder sous différents angles. Le truc le plus astucieux que j'ai trouvé c'est de mettre une tige au milieu de la monture fixée à la table (précession), ça rend les différents éléments (échantillon puis monture 2) bcp plus simples à centrer. Et sinon je me suis chié sur la hauteur pour ma carte et ç se voit.

  \item Trouver un diamant unique : c'est la partie ou je suis le moins sur de moi. Ce que j'ai trouvé globalement, c'est soit de foutre des diamants partout, et d'essuyer sur le dessus et le dessous de la tige (puisque c'est les diamant sur les coté que tu peux voir directement), soit d'utiliser un petit fil chargé de diamants et le frotter sur la tige pour pas trop en déposer. A la fin cherche en un qui crache bien, en vrai je suis pas sur que ce soit si dur.
  \end{itemize}

  \subsection{Résultats expérimentaux}
  \begin{itemize}
  \item 2019/11/04 : 2 map et des ESR correspondants. La 2eme map m'a quand même pris plus d'une heure de mesure (900 points, intégrés 1s chacun), et j'ai du essayer de piffer des variations de 1 degré avec des graduations de 2 degrés. En plus le diamant était pas au centre de la monture "nutation", donc j'ai un gradient de PL que j'ai normalisé un peu à la zob. On voit un grosse tache au centre avec 6 ou 7 branches qui partent. Une ptite simu pour comprendre ce que sont ces branches ferait pas de mal.
  \end{itemize}

  \section{Map automatisée}
  date : fin novembre 2019 à Janvier 2020 au moins
  \subsection{Matos}
  La grosse nouveauté qui m'aura bien bouffé un mois c'est le gonio motorisé Thorlabs. Actuellement, si j'ai le temps d'écrire, c'est parce que je l'ai un peu flingué (et aussi que les rer du soir ont enfin repris), mais ça se trouve je l'ai réparé. Dans tous les cas à l'avenir il vaudrait mieux éviter de construire des tours Eiffel dessus, pendant un moment je vais sans doute me contenter d'un aimant posé sur la base.

  \subsubsection{Utilisation}
  C'est assez plug and play en mode "joystick", faut juste brancher les deux controlleurs aux deux moteurs. A noter qu'on a qu'une seule bonne alim, du coup l'autre j'ai bricolé un truc un peu crado avec 3 étages d'adaptateurs.
  \subsubsection{Interfacage}
  Après avoir bien galéré avec les triggers je me suis rendu compte que c'était infiniment plus simple en mode USB avec une bilbi thorlabs-apt en python. A retenir pour la suite, les bibli propriétaire ca fait gagner du temps. Pour l'utilisation regarder les programmes map new gen ou line new gen, c'est du très basique à base de move-to.

  Globalement j'utilise toujours l'option "fast" des cartes ou je prend la PL à la volée, mais forcément c'est pas hyper précis angulairement (pour la direction qui est balayée) vu que le moteur est pas parfaitement régulier (accélération/décélération entre autre). Donc éventuellement faut check que tu obtiens un truc similaire dans les 2 directions. J'ai aussi rajouté un petit "over wait" dans la dernière version qui te dit le temps que tu attends après ton scan de PL pour que le moteur arrive. Dans l'idéal ce truc est autour de 0, mais comme je peux pas voir les valeurs négatives, on va dire qu'il faut que ça tourne autour de 5 (unité $\approx 0.01$ s).

  Le dernier truc qui est peut-être le plus important : la bibli chie grave sa race, et très souvent python arrive même pas à la charger. C'est sans doute un truc autour du clean-up, et peut être du fait que j'utilise une interface graphique. J'ai bricolé un truc en modifiant les fichiers de la bibli mais c'est pas très sérieux. Bref, le truc à retenir c'est : \textbf{SI CA PLANTE FAUT LANCER KINESIS}. Kinesis c'est le programme Thorlabs plus récent mais qu'on peut pas interfacer à ma connaissance. Dans Kinesis faut s'assurer que les 2 moteurs sont chargé, et éventuellement les charger soi même. Si tu les vois pas, débranche et rebranche les USB.
  \subsubsection{Entretien (lol)}
  Je suis un peu obligé de faire cette partie, ne serait-ce que pour dire que c'est pas la peine d'essayer de dévisser les moteurs. La seule chose que ta vas réussir à dévisser c'est la colle qui maintien l'enveloppe plastique du moteur.

  Pour accéder à la vis du moteur, en faite il faut enlever le stage 1 du stage 2 et le stage 2 de la petite plaque grace à 4 vis noires verticales dans les deux cas (ou 3 dans un cas vu que j'en ai paumé une) et retourner le stage à l'envers. Ca continue de couiner même après l'avoir graissé mais bon il a quand même l'air de marcher.
  \subsection{Résultats}
  Déja, des cartes qui pètent la classe (et que je pourrais peut-être pas refaire en fonction d'à quel point le gonio est abîmé). Et plein d'autres trucs, mais ce sera au prochain épisode, je commence à avoir faim.

 \section{Électro-aimant}
 date : début : février 2020
 \subsection{Matos}
 Le matos utilisé c'est l'électro aimant fabriqué par Paul avec le gros entre-fer et beaucoup d'enroulement (du coup il prend max 1A, et sous vide je sais pas mais moins que ça), et l'alimentation Rhode \& Shwartz avec trois sorties, qui est pilotée en USB via Python.
 \subsection{Interfaçage}
 Il y a plusieurs programmes comme scan\_electroaimant.py qui montrent comment faire, les infos viennent toutes du site de R\&S. En gros ça passe par du VISA, il faut trouver l'ID de l'appareil (dans mon cas j'ai fait ça avec un utilitaire de R\&S) et ensuite les commandes sont assez intuitives, elle se séparent en QUERY quand tu veux interroger l'appareil, et WRITE pour envoyer des instructions. Il faut bien penser à mettre un WAIT entre chaque instruction, sinon des fois les instruction suivantes viennent override.
 Aussi, le wait te dis juste que l'instruction a été reçue, pas qu'elle est appliquée. Du style une consigne en tension va revenir immédiatement mais ensuite il faut compter au moins 0.5s pour que la tension soit effectivement appliquée (c'est un générateur DC, il est pas vraiment fait pour faire de la dynamique)
 \subsection{Manips}
 Pour l'instant des scans à chaud sur les sumitomo (des diamants HPHT avec 4 concentrations de NV différentes filés par Alex) en tension (donc en courant donc en champ mag) ou je regarde la PL et en ce moment le temps de polarisation (à voir si ça mène quelque part). Les programmes sont un peu chelou : j'utilise l'option EasyRamp (tm) du générateur qui permet de faire des rampes de tension, mais il y a quelques subtils problèmes :
 \begin{itemize}
 \item Le plus important c'est que le générateur est pas triggé. Y'a peut être moyen de le faire mais en attendant je mesure en même temps (plus ou moins) la tension et la PL et je re-plot ça comme je peux ensuite pour avoir la PL en fonction de la tension.
 \item Ensuite les rampes ne se font qu'à l'allumage de la tension (quand tu passe en output ON). Donc tu pars forcément de 0 (au passage le fait de couper à chaque fois l'output joue sur le cycle d’hystérésis je pense : j'ai l'impression qu'on passe par des tension négatives du coup au moment ou il s'éteint)
 \item Et le dernier truc c'est que ce crétin ne connait pas les nombres à virgule, donc je peux faire des rampes que de 1s ou genre 10-2 (le minimum).
 \end{itemize}

  L'étape suivante c'est de choper une résistance (je voulais un rhéostat mais Hadrien m'a gentiment filé un vieux truc qui trainait en R18) parce que le générateur ne peux pas faire des pas de moins de 0.01V (d'ailleurs la rampe est en escalier, on peut le voir) et que la zone qui m'intersse pour l'instant est entre 0.4 et 0.5V. Donc une résistance en série ça me permettrait d'augmenter les tensions tout en maintenant le même courant.

  Et la vraie étape suivante c'est de mesurer les largeurs sur les différents sumitumo (et éventuellement le diamant rose) pour voir un peu comment ça varie avec la concentration. A la base je voulais faire ça avec le gonio mais il boude en ce moment, et de toutes façon c'est moins reproductible les mesures au gonio, et ça donne pas non plus l'énergie.

  Edit coronavirus (1er Avril 2020) : J'ai bien ajouté une resistance ce qui a pas mal augmenté la résolution, et j'ai fait des mesures de largeurs sur les sumi 2,4 et le rose. J'ai aussi essayé l'electro-aimant à froid mais sans écran thermique on descend à peine à 250K, c'est vraiment minable. Pour les mesures de largeur, il y a quand même pas mal de soucis :
  \begin{itemize}
  \item Deja il faut que je mette au clair cette mesure HWHM qui a l'air complètement fausse sur les sumi
  \item Ensuite y'a le probleme de la soustraction du background, qui joue sans doute beaucoup dans le fait que le fit soit ensuite plutot gaussien que lorentzien, même si a priori en ajoutant une constante libre je pense que ça amoindrit le pb. Pour l'instant je soustrait simplement une droite entre les deux points qui limitent la bosse (interpolation linéaire). Je pourrai faire une interpolation quadratique ou plus mais je sais pas si j'y gagnerai beaucoup.
  \item Enfin y'a le problème du rose : Deja sons contraste est vraiment flingué, c'est à se demander comment j'ai réussi à faire une carte, et la seule transition 1x2x1 que j'ai réussie à isoler est très proche de 2.88 GHz. Je sais pas si c'est lié mais ma raie est aussi très asymétrique.
  \end{itemize}
  \subsubsection{Mesures}
  Les principales mesures grace à l'électro-aimant
  \end{document}

### 2020-07-10 théorie/fiches/Signal et bruit/Signal et bruit.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\tableofcontents
\chapter{Signal et bruit}
  \section{Conventions, définitions}
  \subsection{Transformée de Fourier}
  On prend la définition "du physicien". Soit $s(t)$ un signal, alors ; 
  \begin{align}
  \tilde{s}_1(\omega)&=\int_{-\infty}^{+\infty} s(t) e^{-i\omega t} dt \\
  s(t)&=\int_{-\infty}^{+\infty}\tilde{s}_1(\omega) e^{i\omega t} \dfrac{d\omega}{2\pi}
  \end{align}
  On peut également choisir la convention :
  \begin{align}
  \tilde{s}_2(\nu)&=\int_{-\infty}^{+\infty} s(t) e^{-i2\pi \nu t} dt \\
  s(t)&=\int_{-\infty}^{+\infty}\tilde{s}_2(\nu) e^{i2\pi \nu t} d\nu
  \end{align}
  Et dans ce cas là tu as simplement $\tilde{s}_1(\omega)=\tilde{s}_2(2\pi \nu)$. Par contre les $\sqrt{2\pi}$ on oublie.

  Dans tous les cas, ton spectre à pour dimension $[\tilde{s}]=[s]\cdot[Hz]^{-1}$
  \subsection{Propriétés de la TF}
  \begin{itemize}
  \item Translation : \begin{equation}
  \tilde{f}(t-\tau)[\omega]=e^{-i\omega \tau} \tilde{f}(t)[\omega]
  \end{equation}
  \item Scaling : \begin{equation}
  \tilde{f}(a\cdot t)[\omega]= \dfrac{1}{|a|} \tilde{f}(t)[\dfrac{\omega}{a}]
  \end{equation}
  En particulier, $\tilde{f}(-t)[\omega]=\tilde{f}(t)[-\omega]$. C'est la "time reversibility"
  \item $f$ réelle $\implies$ $\tilde f$ Hermitienne, soit $\tilde f^*(\omega)= \tilde f(-\omega)$
  \item Inversement, si $f$ est hermitienne (donc paire pour une fonction réelle), $\tilde{f}$ est réelle.
  \item Convolution : 
  \begin{equation}
  TF[(f*g)(\tau](\omega)=\tilde{f}(\omega)\tilde{g}(\omega)
  \end{equation}
  \item Auto-corrélation : \begin{align*}
  \tilde S_{ff}(\omega)&=TF[(f(t)*f^*(-t))] \\
  &=TF[f(t)] \cdot TF[f^*(-t)] \\
  &= TF[f(t)] \cdot TF[f(t)]^* \\
  &= |\tilde{f}(\omega)|^2
  \end{align*}

  \end{itemize}
  \subsection{Transformée de Fourier tronquée}
  L'infini c'est long, surtout dans le négatif. Donc expérimentalement on va plutôt utiliser la TF tronquée : 
  \begin{equation}
  \hat{s}(\omega)=\dfrac{1}{\sqrt{T}}\int_{0}^{T} s(t) e^{-i\omega t} dt
  \end{equation}
  \subsection{Transformée de Fourier discrète}
  \subsection{Produit de convolution}
  \begin{equation}
  (f*g)(\tau)=\int_{-\infty}^{+\infty} f(\tau-t) g(t) dt = \int_{-\infty}^{+\infty} f(t) g(\tau-t) dt
  \end{equation}
  \subsection{Fonction d'auto-corrélation}
  La fonction d'auto-corrélation d'un signal est définie comme :
  \begin{equation}
  R_{ss}(t_1,t_2)=\langle s(t_1)s^*(t_2)\rangle
  \end{equation}
  où $\langle ...\rangle$ représente a priori une moyenne d'ensemble.
  Mais nous on aime bien les processus Markovien, donc $R_{ss}$ n'est plus fonction que de $\tau=|t_1-t_2|$. En plus de ça, on fait une petite hypothèse ergodique est la moyenne d'ensemble devient une moyenne temporelle, et on oublie au passage le facteur de normalisation Alors :
  \begin{equation}
  R_{ss}(\tau)=\int_{-\infty}^{+\infty} s(t) s^*(t-\tau)= (s(t)*s^*(-t))(\tau)
  \end{equation}
  Si jamais ton signal est réel, la fonction d'auto-corrélation est paire, donc $R_{ss}(\tau)=R_{ss}(-\tau)$. Si ton signal est complexe c'est un peu plus chiant.

  La \textbf{Fonction d'auto-covariance} c'est simplement la fonction d'auto-corrélation à la quelle tu soustrais $\langle s(t)\rangle ^2$.
  \subsection{Densité spectrale de puissance}
  On peut définir proprement la DSP à partir de la TF tronquée : 
  \begin{equation}
  S_{ss}(\omega)=\lim_{T \to \infty} \langle |\hat{s}(\omega)|^2 \rangle
  \end{equation}
 Or, \begin{align*}
 \langle |\hat{s}(\omega)|^2 \rangle &=\langle \hat{s}(\omega)\hat{s^*}(\omega) \rangle \\
 &=\langle \dfrac{1}{T} \int_{0}^{T} s(t) e^{-i\omega t} dt \int_{0}^{T} s^*(t') e^{i\omega t'} dt' \rangle\\
 &=\dfrac{1}{T} \int_{0}^{T} \int_{0}^{T}\langle s(t)s^*(t') \rangle e^{i\omega(t-t')} dt\; dt'
 \end{align*}
 Et donc tu vois que, en posant $\tau=t-t'$ et en bidouillant un peu les bornes d'intégrations, tu te retrouves avec :
 \begin{equation}
 S_{ss}(\omega)=\int_{-\infty}^\infty R_{ss}(\tau) e^{-i\omega\tau}\,d \tau=\tilde R_{ss}(\omega)=|\tilde s (\omega)|^2
 \end{equation}
 La densité spectrale de puissance est la TF de la fonction d'auto-corrélation (pour un processus Markovien). C'est le théorème de \textbf{Wiener–Khinchin}.

 La version plus sale c'est d'écrire la DSP comme :
 \begin{equation}
 \langle \tilde{s}(\omega) \tilde{s}^*(\omega') \rangle = S_{ss}(\omega)2\pi  \delta(\omega-\omega')
 \end{equation}

 \section{Résolution d'un Michelson}
 Michelson mais ça marche pour toutes les interférences à deux ondes (Ramsay entre autre).

 \subsection{Réponse spectrale}
 On va commencer par la réponse spectrale d'un Micheslon, c'est à dire le rapport des densités spectrales de puissance en entrée ($I_{in}(\omega)$) et en sortie ($I_{out}(\omega)$) de l'interféromètre. On note $A$ les amplitudes correspondantes, et $\tau$ le retard dans l'interféromètre. Alors :
 \begin{equation}
 I_{in}(\omega)=\tilde{A}_{in}(\omega)\tilde{A}_{in}^*(\omega)=|\tilde{A}_{in}(\omega)|^2=|TF[A_{in}(t)](\omega)|^2
 \end{equation}
 et 
 \begin{align*}
I_{out}(\omega)&= |TF[A_{out}(t)](\omega)|^2 \\
 &= |TF[\dfrac{1}{2} (A_{in}(t)+A_{in}(t-\tau)](\omega)|^2 \\
 &= \left|TF[(A_{in}(t)](\omega)\left(\dfrac{1+e^{i\omega \tau}}{2}\right)\right|^2 \\
 &= I_{in}(\omega) \left(\dfrac{1+e^{i\omega \tau}}{2}\right)^2 \\
 &= I_{in}(\omega) \left(\dfrac{1+\cos \omega tau}{2}\right)
 \end{align*}

 Remarque : le 1/2 dans le $A_{out}(t)$ vient du fait que tu as deux séparatrices, qui ajoutent $1/\sqrt{2}$ chacune.

 On trouve donc l'intervalle spectral libre : \begin{equation}
 \Delta \nu = \dfrac{1}{\tau} = \dfrac{(2)l}{c}
 \end{equation}

 En résonnant maintenant en nombre de particules (photon ou autre), en supposant que le flux incident comporte $N$ particules et le flux sortant (flux transmis, ou nombre d'atomes excités dans le cas de Ramsay) est $N_e$, on a tout simplement \begin{equation}
 N_e=N\left(\dfrac{1+\cos \Phi}{2}\right)
 \end{equation}
Avec $\Phi$ le déphasage entre les deux voies.
\subsection{Pouvoir de résolution}
On va maintenant considérer que $N_e$ est une variable aléatoire, chaque particule ayant une chance $p_e=\dfrac{1+\cos \Phi}{2}$ de passer. On a donc :
\begin{align}
\langle N_e\rangle &= N\left(\dfrac{1+\cos \Phi}{2}\right) \\
\Delta^2 N_e &= N(p_e)(1-p_e) = N\dfrac{\sin ^2 \Phi}{4} \\
\Delta N_e &=\sqrt{N} \dfrac{|\sin \Phi|}{2}
\end{align}

Maintenant pour s'intéresser au pouvoir de résolution, on va considérer un résultat $N_{e1}$ de $N$ particules à $\omega$ (où $\Delta$ pour un Ramsay) et un autre résultat $N_{e2}$ de $N$ particules à $\omega + \delta \omega$. Pour pouvoir distinguer le résultat entre les deux pulsations, il faut que la différence des valeurs moyennes soit plus grande que ($\sqrt{2}$ fois) l'écart type du résultat. ($\sqrt{2}$ car il y a deux variables aléatoires dont tu fais la différence, donc tu sommes les variances, un peu comme une mesure à la règle. Mais en vrai c'est du chipotage). D'où : \begin{equation}
\langle N_e (\omega+ \delta \omega)\rangle - \langle N_e (\omega)\rangle \approx -N\tau\dfrac{\sin \omega \tau}{2} \geq \sqrt{2N} \dfrac{\sin \omega \tau}{2} 
\end{equation}

D'où finalement \begin{equation}
\delta \omega \geq \dfrac{\sqrt{2}}{\sqrt{N}\tau}
\end{equation}

On constate que le pouvoir de résolution d'un Michelson ne dépend pas de la phase : Le signal et le bruit sont tous les deux plus importants quand ta phase vaut $\pi/2$

On remarque également qu'on a une limite en $1/\tau$, qui vient de la transformée de Fourier / Heisenberg temps/énergie, et en $1/\sqrt{N}$, qui est la limite "classique" pour tout processus stochastique à N particules. En utilisant des états quantiques à N particules (intriquées donc) on peut idéalement monter à des résolutions en $1/N$.
  \end{document}

### 2020-07-10 théorie/fiches/Spectro/Spectroscopie.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{bm} % pour faire des vecteurs gras
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{L'atome}
  \section{Le moment cinétique}
  \subsection{Définition, relations de commutation}
  \textbf{Convention} : Dans tout ce chapitre, les vecteurs seront notés en \textbf{gras} et les opérateurs avec des chapeaux.

  En partant de la définition classique $\bm{L}=\bm{r} \times \bm{p}$ pour une particule, on obtient la version quantique avec la magie des chapeaux : \begin{equation}
  \hat{\bm{L}}=\hat{\bm{r}} \times \hat{ \bm p}
  \end{equation}
 Alors avec les relations de commutation sur $\hat{\bm{r}}$ et $\hat{ \bm p}$, on trouve : \begin{equation}
 [\hat L_x, \hat L_y]=i\hbar \hat L_z \quad , \quad [\hat L_y, \hat L_z]=i\hbar \hat L_x \quad , \quad [\hat L_z, \hat L_x]=i\hbar \hat L_y
 \end{equation}
 Qu'on peut également synthétiser en : \begin{equation}
 \hat{\bm L} \times \hat{\bm L} = i\hbar \hat{\bm L}
 \end{equation}
 C'est cette relation de commutation qu'on prendra comme définition d'un opérateur de moment cinétique, y compris quand $\bm{L}=\bm{r} \times \bm{p}$ n'existent pas (coucou le spin). Pour généraliser, on écrira $\hat{\bm J}$ pour un moment cinétique quelconque
 \subsection{Valeurs propres}
 On définit $\hat J^2= \hat J_x^2 + \hat J_y^2 + \hat J_z^2$. $ \hat J^2$ commute avec chacune des composantes de $\hat{\bm J}$ : \begin{equation}
 [\hat J^2, \hat{\bm J}]=0
\end{equation}  
(Composantes qui ne commutent pas entre elles, cf avant).

On va alors former un ECOC à partir de $\hat{J^2}$ et d'une des trois composante qui sera par convention $\hat{J_z}$, c'est à dire que l'état du moment cinétique sera entièrement défini par la velur des deux nombres quantiques $j$ et $m$ sans dimension tels que :\begin{align}
\hat{J^2}\ket{j,m}&=j(j+1) \hbar^2 \ket{j,m} \\
\hat J_z\ket{j,m}&= m\hbar \ket{j,m}
\end{align}
$j$ est forcément positif car $\hat{J^2}$ est défini positif, et tu trouves que tu ne peux avoir que des valeurs demi-entières et que $-j \leq m \leq j$ par l'application répétée de $\hat L_\pm$ défini par :\begin{equation}
\hat L_\pm= \hat L_x \pm i \hat L_y
\end{equation}
Et qui vérifie :\begin{equation}
\hat L_\pm \ket{j,m} \propto \ket{j,m \pm 1}
\end{equation}
\subsection{Moment cinétique orbital et harmoniques sphériques}
On va maintenant s'intéresser à une particule ponctuelle qui a un moment cinétique classique. On peut donc bien écrire ici $\hat{\bm{L}}=\hat{\bm{r}} \times \hat{ \bm p}$

On se placera dans les coordonnées polaires habituelles ($r$,$\theta$,$\phi$) avec l'axe z comme axe polaire, $0 \leq \theta \leq \pi$ la colatitude en bon français et $0 \leq \phi \leq 2\pi$ l'angle azimutal.

Avec ces notations, en se rappelant que $\hat p_x = -i\hbar\dfrac{d}{dx}$ lorsqu'on raisonne en fonction d'onde, on trouve en coordonnées polaires : \begin{equation}
\hat L_z= -i\hbar \dfrac{d}{d\phi}
\end{equation}

Ce qui veut dire que les orbitales, c'est à dire les fonctions d'onde des états propres de $\hat L^2$ et $\hat L_z$, vérifient : \begin{equation}
\hat L_z \psi_{lm}(r,\theta, \phi)= -i\hbar \dfrac{d}{d\phi} \psi_{lm}(r,\theta, \phi) = \hbar m \psi_{lm}(r,\theta, \phi)
\end{equation}
D'où : \begin{equation}
\psi_{lm}(r,\theta, \phi)= \varphi(r,\theta)e^{i m \phi}
\end{equation}

Puisqu'une rotation de $\phi$ de $2\pi$ doit donner la même fonction d'onde (? pourquoi c'est pas vrai sur la sphère de Bloch, parce que ce n'est pas l'espace physique ?), on en déduit que \textbf{m et don l sont nécessairement entiers}

C'est également vrai pour $\hat L^2$ mais sa formule est plus dégueu : \begin{equation}
\hat L^2= -\hbar ^2 \left( \dfrac{1}{\sin \theta}\dfrac{d}{d\theta}\sin \theta\dfrac{d}{d\theta}+ \dfrac{1}{\sin ^2 \theta} \dfrac{d}{d\phi} \right)
\end{equation}

En faite c'est la partie angulaire du laplacien sphérique : \begin{equation}
\Delta = \dfrac{1}{r}\dfrac{d^2}{dr^2}-\dfrac{1}{r^2\hbar^2}\hat L^2
\end{equation}

On constate également que $\hat L^2$ et $\hat L_z$ n'agissent pas sur la partie radiale de l'orbitale, on va donc décomposer l'orbitale en sa partie radiale et sphérique : \begin{equation}
\psi_{lm}(r,\theta,\phi)=R_{l}(r)Y_{lm}(\theta,\phi)
\end{equation}
Où les $Y_{lm}$ sont les \textbf{Harmoniques sphériques}, c'est à dire les fonction propres de $\hat L^2$ et $\hat L_z$ normalisées à 1.

On remarque que $L_z$ n'agit pas sur la partie radiale de $\psi_lm$, d'où le fait que $R_{l}$ ne dépendent pas de $m$

\subsection{Lien avec le moment magnétique}
Toute particule chargée avec un moment cinétique (orbital ou non) dispose également d'un moment magnétique.

Pour rappel, l'interaction d'un moment magnétique avec un champ s'écrit : \begin{align}
W &= - \bm \mu \cdot \bm B \\
\bm \Gamma &= \bm \mu \times \bm B
\end{align}
\subsubsection{Pour un moment cinétique orbital}
Analogie classique : tu as une spire de courant formée par ta particule et son mouvement (circulaire ici mais je crois que c'est général) :
\begin{align}
\bm L &= \bm r \times \bm p = m_q r v \bm u \\
\bm \mu &= I S \bm u = \dfrac{qv}{2\pi r}\pi r^2 \bm u = \dfrac{q}{2} v r \bm u
\end{align}
D'où : \begin{equation}
\bm \mu = \gamma_0 \bm L \quad \mathrm{avec} \quad \gamma_0=\dfrac{q}{2 m_q}
\end{equation}
\emph{note :} $q$ est algébrique, dans le cas de l'électron $q=-e$

$\gamma_0$ est appelé le \textbf{rapport gyromagnétique} de ta particule. \\

On définit également le \textbf{magnéton de Bohr} comme le quanta de moment magnétique (du au mouvement orbital) pour l'électron : 
\begin{equation}
\mu_B=\gamma_0 \hbar= -\dfrac{q \hbar}{2 m_e}
\end{equation}
\subsubsection{Pour un moment cinétique intrinsèque}
De la même façon, on postule (je crois pas qu'il y ait de preuve, si ce n'est expérimentale) que les moments cinétiques intrinsèques sont proportionnels à un moment magnétique.

On doit quand même introduire un facteur sans dimension supplémentaire par rapport au moment cinétique orbital : le \textbf{facteur de Landée $g$} : \begin{equation}
\bm \mu_s= g\dfrac{q}{2m_q} \bm J
\end{equation}

Les valeurs numériques de g sont : \begin{align*}
\mathrm{electron} &\approx -2.00 \\
\mathrm{proton} &\approx +5.59 \\
\mathrm{neutron} &\approx -3.83
\end{align*}

\subsection{Composition de moments cinétiques}
\begin{figure}[h]
\centering
\includegraphics[width=.5\linewidth]{Clebs_Jordan.jpg}
\caption{Clebs-Jordan, le cousin de Clebsch-Gordan}
\end{figure}
Ça peut concerner deux particules en interaction ou un moment orbital et un moment intrinsèque. On va donc considérer deux moments : $\hat{\bm J_1}$ et $\hat{\bm J_1}$

Dans la base découplée, ton système est donc décrit par les 4 nombres quantiques naturels : $\ket{j_1,m_j,j_2,m_2}$.

Cependant on constate que $\hat{\bm J}=\hat{\bm J_1}+\hat{\bm J_2}$ est également un moment cinétique, on peut alors former un nouvel ECOC avec les 4 observables : $\hat J^2, \hat J_z, \hat J_1^2$ et $\hat J_2^2$. Les 4 nouveaux nombres quantiques de la base couplée seront donc : $\ket{j,m,j_1,j_2}$.

Il y a deux avantages à passer dans la base couplée : 
\begin{itemize}
\item D'une part dans le cas de moment cinétiques orbitaux, $J$ représente bien le moment cinétique total du système
\item D'autre part dans le cas d'une interaction, tu vas très souvent te retrouver (au premier ordre) avec un terme dans le hamiltonien en $\hat{\bm J_1} \cdot \hat{\bm J_2}$. Or $\hat{\bm J_1} \cdot \hat{\bm J_2}= \dfrac{1}{2}(\hat J^2- \hat J_1^2 - \hat J_2^2$ (source : mon cours de troisième). Donc les états de la base couplée sont états propres du Hamiltonien, et ça c'est toujours sympathique.
\end{itemize}

\subsubsection{Passage d'une base à l'autre}

Dans la théorie tu utilises $J^-=J_1^-+J_2^-$ en partant de l'état $m=j$ qui est présent dans les deux bases, dans la pratique tu regardes les coefs de Clebsch-Gordon :

\begin{center}
\includegraphics[scale=.35]{Clebsch_Gordan}
\end{center}

(je sais c'est écrit petit, mais c'est ces vicieux de la physique des particules).

Dans tous les cas tu constates que $\hat J_1^2$ et $\hat J_2^2$ sont présents dans les deux bases donc tu n'as que 2 nombres quantiques à changer.

\subsection{Lois de conservations}

Pour la projection c'est très clair, vu que $m=m_1+m_2$. Le moment total c'est plus compliqué vu qu'il y a la conservation de l'énergie qui s'en mêle.

\section{Nombres quantiques d'un état atomique}
Dans toute cette partie on s’intéresse à l'atome d'hydrogène. Ça se généralise assez bien aux alcalins mais tout le reste c'est plus compliqué.
\subsection{Hamiltonien}
Dans l'absolu c'est le problème à deux corps, que tu sépare en une partie avec les coordonnées du centre de masse, et une partie avec les coordonnées relatives. Dans la pratique on va considérer l'électron dans un potentiel central généré par le proton immobile dans le référentiel du labo, parce qu'il y a un rapport 2000 entre les masses. De toute façon les calculs sont identiques.

\begin{equation}
H=\dfrac{p^2}{2m}+V(r)
\end{equation}

On s’intéresse aux solutions stables, donc ESIT : \begin{equation}
\Delta \psi (\bm r) + V(r) \psi (\bm r) = E \psi (\bm r)
\end{equation}

Avec la définition du laplacien sphérique en fonction de $\hat L^2$ et la décomposition en harmonique sphérique $\psi_{lm}(r,\theta,\phi)=R_{l}(r)Y_{lm}(\theta,\phi)$, on trouve finalement une EDO sur $R_l$ : \begin{equation}
\left(-\dfrac{\hbar^2}{2m}\dfrac{1}{r}\dfrac{d^2}{dr^2}r + \dfrac{l(l+1)\hbar^2}{2mr^2} + V(r) \right) R_l(r) = E R_l(r)
\end{equation} \\

\textbf{Remarque} : C'est parce que $\hat H$, c'est à dire $\hat V$, commute avec $L_z$ et $L^2$ que tu peux décomposer les solutions comme ça. Et c'est vrai ici parce que V est invariant selon $\theta$ et $\phi$.
\subsection{Solutions de l'ESIT}

Bon en vrai c'est un peu chaud les solutions de l'EDO, ce qu'on peut retenir c'est que : \begin{itemize}
\item Comme pour tout système lié, il y a discrétisation des $E<0$ possibles
\item Pour un $l$ donné, il n'existe qu'une solution pour chaque valeur propre $E_{n'}$ donnée. On peut donc définir sans équivoque $\ket{l,n'}$
\item La partie radiale de la fonction d'onde $R_{n'l}(r)$ a exactement $n'$ zéros.
\item L'énergie $E_{n'l}$ correspondante est \begin{equation}
E_{n'l}=-\dfrac{E_0}{(n'+l+1)^2}
\end{equation}
Avec $E_0$ l'énergie du fondamental, c'est à dire l'énergie d'ionisation de l'atome d'hydrogène, c'est à dire un Ryldberg (techniquement les Ry c'est en $\mathrm{cm}^{-1}$ je crois) : \begin{equation}
E_0=\dfrac{m_e e^4}{2\hbar^2}\approx13.6 \mathrm{eV}
\end{equation}
\item A cause de l'expression de l'énergie, on utilise comme nombre quantique \begin{equation}
n=n'+l+1
\end{equation}  Donc maintenant l'énergie peut ne s'exprimer qu'en fonction de $n$, mais elle dépend quand même de $l$. Bref en vrai ça complique la vie plus qu'autre chose.
\end{itemize}
\subsection{Notations spectro}
Bon je t'épargne l'histoire (c'est à cause des séries de raies de l'atome d'hydrogène : sharp, principal, diffuse...) mais 3d $\Rightarrow n=3\; \mathrm{et} \;l=2$
\subsection{Solutions particulières}
\subsubsection{Orbitale 1s}
$l=0$ et $n'=0$, $Y_{l,m}(\theta, \phi)=1/\sqrt{4\pi}$, ta fonction est : \begin{equation}
\psi_{1,0,0}(\bm r) = \dfrac{e^{-r/a_0}}{\sqrt{\pi a_0^3}}
\end{equation}
\textbf{note :} pour tout $n'$ tu es dominé aux grands $r$ par un $e^{-r}$. \\

Si on s'intéresse à la densité de présence radiale, il faut intégrer sur une sphère de rayon $r$, d'où :
\begin{equation}
\rho(r)=|\psi_{1,0,0}(\bm r)|^2 4 \pi r^2= e^{-2r/a_0} \dfrac{r^2}{a_0^3}
\end{equation}

Et on constate que $<r> = a_0$, la distance moyenne de l'électron est bien le rayon de Bohr

\subsubsection{Orbitale 2s}
Toujours une symétrie sphérique car $l=0$, mais cette fois ci ta fonction s'annule et change de signe en $r=a_0/2$. Tu as donc une région interdite pour ton électron. \begin{equation}
\psi_{2,0,0}(\bm r) \propto \left(1-\dfrac{r}{2 a_0}\right) e^{-r/a0}
\end{equation}

\subsubsection{Orbitale 2p}
Ici on commence les embrouilles car $l=1$ et donc $m$ commence à prendre des valeurs non nulles. En plus $p_x$ et $p_y$ c'est des c.l des états propres (ils ne sont donc pas état propre de $L_z$, ce qui est plutôt logique vu que c'est les états propres de $L_x$ et $L_y$).

\begin{align}
\psi_{pz}(\bm r) &\propto Y_{1,0}(\theta,\phi) e^{-r/a_0} \propto \cos \theta  e^{-r/a_0} \\
\psi_{px}(\bm r) &\propto (Y_{1,1}+Y_{1,-1})(\theta,\phi) e^{-r/a_0} \propto \sin \theta \sin \phi e^{-r/a_0} \\
\psi_{py}(\bm r) &\propto (Y_{1,1}-Y_{1,-1})(\theta,\phi) e^{-r/a_0} \propto \sin \theta \cos \phi e^{-r/a_0}
\end{align}

\subsection{Et le spin dans tout ça}
Ton système c'est un électron, donc $s=1/2$ on ne le note pas. Sa projection $m_s=\pm 1/2$ n'influe ni sur les niveaux d'énergie (sans champ magnétique), ni sur les orbitales. En revanche ça joue sur la population des états.

\section{Modification des niveaux d'énergie}
\subsection{Couplage spin-orbite : structure fine}
\subsection{Effet Zeeman}
\subsection{Effet Stark}
\subsection{Couplage spin-spin : structure hyper-fine}
\subsection{Quantification du champ : Lamb shift}

  \end{document}

### 2020-07-10 théorie/fiches/Système à 2 niveaux/fiche.tex
tout d'un coup

\documentclass[a4paper]{report}
%\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{empheq} % Tout ca pour encadrer des équations
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Système à deux niveaux}
  \section{Représentation, sphère de Bloch}
  \begin{figure}[h]
  \centering
  \includegraphics[width=.8\textwidth]{BlochSphere}
  \caption{$\ket{\psi}=\cos (\theta/2)\ket{0}+\sin (\theta/2)e^{i \phi} \ket{1}$}
  \end{figure}

  On va prendre le formalisme des spins (qui ne correspond pas au dessin mais bon).

  Ta base de base va être $\{\ket{-},\ket{+}\}$, les vecteurs propres de $S_z$, et
  \begin{itemize}
  \item $\ket{+}_x=\dfrac{\ket{+}+\ket{-}}{\sqrt(2)} \quad \ket{-}_x=\dfrac{\ket{+}-\ket{-}}{\sqrt(2)}$
  \item $\ket{+}_y=\dfrac{\ket{+}+i\ket{-}}{\sqrt(2)} \quad \ket{-}_y=\dfrac{\ket{+}-i\ket{-}}{\sqrt(2)}$
  \end{itemize}

  Avec $S_x= \begin{pmatrix} 0 & 1 \\ 1 & 0  \end{pmatrix}$ et $S_y= \begin{pmatrix} 0 & -i \\ i & 0  \end{pmatrix}$

  Pour un vecteur quelconque sur la sphère de Bloch (état pur), \begin{equation}
  \ket{\psi}=\cos (\theta/2)\ket{+}+\sin (\theta/2)e^{i \phi} \ket{-}
  \end{equation}

  Le truc à retenir c'est que $\ket{\psi}$ est le vecteur $\ket{+}_\textbf{u}$où $\textbf{u}$ est la direction dans la sphère de Bloch.

  \section{Évolution d'un système pur à 2 niveaux}
  Ton Hamiltonien va être typiquement de la forme $H=H_0 +H_1$ avec
  \begin{itemize}
  \item $H_0=\hbar \begin{pmatrix} -\omega_0/2 & 0 \\ 0 & \omega_0/2  \end{pmatrix}$ (l'ordre de la base c'est $\{\ket{-},\ket{+}\}$). Pour un spin, ca va venir du champ magnétique selon z : $-M.B_0$, donc $\omega_0=\gamma B_0$. Dans le cas d'un atome c'est simplement le niveau d'énergie de l'électron.
  \item $H_1$ est le terme de couplage, dont on ne retiendra que les termes non-diagonaux. On traitera une perturbation statique, et une perturbation (quasi)résonnante à $\omega_0$ (Rabi).
  \end{itemize}
   \subsection{Perturbation statique : anti-croisement}
   \includegraphics[width=\textwidth]{anti_croisement}

   Hamiltonien indépendant du temps, on diagonalise. $E_1$ et $E_2$ sont les énergies propres de $H_0$, $E_\pm$ celles de  $H= \begin{pmatrix} E_1 & W_{12} \\ W_{21} & E_2  \end{pmatrix}$

   On trouve : \begin{align}  
   E_\pm &= \dfrac{E_1-E_2}{2}\pm\dfrac{1}{2}\sqrt{(E_1-E_2)^2+4|W_{12}|^2} \\   
   \ket{\psi_+} &= \cos \dfrac{\theta}{2} e^{-i\phi/2} \ket{\psi_1}+\sin \dfrac{\theta}{2} e^{i\phi/2} \ket{\psi_2} \\   
   \ket{\psi_-}&= -\sin \dfrac{\theta}{2} e^{-i\phi/2} \ket{\psi_1}+\cos \dfrac{\theta}{2} e^{i\phi/2} \ket{\psi_2} 
  \end{align}

   Avec $\tan \theta = 2 \dfrac{|W_{12}|}{E_1-E_2}$ et $W_{12}=e^{i\phi}|W_{12}|$

   En particulier, on voit que pour 2 niveaux dégénérés, le fait qu'il existe un couplage entre les deux va lever la dégénérescence, et créer un état de plus basse énergie, donc plus stable (d'où la stabilité des cycles aromatiques, genre benzène).

   \subsection{Perturbation sinusoïdale : Rabi}
   La perturbation peut être un champ magnétique transverse oscillant à $\omega$ pour un spin, une onde électromagnétique pour un système atomique, ou n'importe quel couplage sinusoïdal ne faisant intervenir que des termes hors-diagonaux.

   Remarque pour le couplage lumière-matière, c'est parce que l'opérateur dipôle $\textbf{D}=q\textbf{r}\begin{pmatrix}
   0 & d \\ d^* & 0
   \end{pmatrix}$ n'a pas de termes diagonaux, car les états propres du système ont une parité définie (pas de sens privilégié, symétrie centrale).

   On a alors \begin{equation}
    H=\dfrac{\hbar}{2} \begin{pmatrix} \omega_0 & \omega_1 e^{-i\omega t} \\ \omega_1 e^{i\omega t} & -\omega_0  \end{pmatrix}
\end{equation}    où $\omega_1$ correspond à l'intensité du couplage, et $\omega$ la pulsation de l'oscillation.

   On va passer dans le référentiel tournant par la transformation \begin{equation}
   U=\begin{pmatrix} e^{i\omega t/2} & 0 \\ 0 & e^{-i\omega t/2}  \end{pmatrix}
   \end{equation}On fait également l'approximation de l'onde tournante : Rotating wave / secular approx, pour virer le terme en $2\omega$. Le résultat n'est donc valable que proche de la résonance

   $H$ devient alors \begin{equation}
   \tilde{H}=\dfrac{\hbar}{2} \begin{pmatrix} -\Delta\omega & \omega_1  \\ \omega_1 & \Delta\omega  \end{pmatrix}
   \end{equation} (A revoir) qui est indépendante du temps, donc on résout, on repasse dans le référentiel classique et on trouve que, pour un état initialement en $\ket{+}$, la probabilité de passage dans l'état $\ket{-}$ vaut : \begin{equation}
   \mathcal{P}_{+-}(t)=\dfrac{\omega_1^2}{\omega_1^2+(\Delta\omega)^2}\sin^2(\sqrt{\omega_1^2+(\Delta\omega)^2}\dfrac{t}{2})
   \end{equation}
   On retrouve l'évolution temporelle, sinusoïdale à la fréquence de Rabi $\Omega=\sqrt{\omega_1^2+(\Delta\omega)^2}$ (le /2 vient du fais que tu as un sinus carré, qui oscille donc deux fois plus vite qu'un sinus normal. Tu fais donc une période en $2\pi/\Omega$.)

   On retrouve également la Lorentzienne pour l'amplitude maximale en fonction du désaccord.

   \subsection{Franges de Ramsey}
   Le but ici ça va être de mesurer la fréquence d'une transition. Jusque là assez classique en spectro. Une méthode pour ça, c'est de se fixer un temps $\tau$, et de choisir $\Omega$ pour avoir une pulse $\pi$, soit $\Omega=\pi/\tau$. Tu vas ensuite scanner en fréquence ton laser résonnant : tu vas mesurer une lorentzienne de largeur $\Omega \approx 1/\tau$. Donc pour avoir une mesure précise, il te faut $\tau$ le plus grand possible. Ca c'est Fourier/Heinsnberg temps/énergie. Sauf que pour un temps long, bah t'as l'émission spontanée qui vient t'emmerder. D'où le lifetime-limited. Et bah Ramsay c'est exactement les mêmes limitations, mais c'est plus pratique dans certains cas visiblement.

   Ramsey, lui, c'est la fameuse méthode des 2 pulses $\pi/2$ : 2 pulses identiques (avec la même phase !) très courte (durée $t$) et séparée d'un temps long ($T >> t$).

   Le mieux c'est de se représenter ça graphiquement : ton laser est légèrement detuné, à la fréquence $\omega$. Tu vas imaginer ta sphère de Bloch dans le référentiel tournant à $\omega$. A cause du detuning, elle tourne (lentement). Comme les pulses sont très rapides, tu vas négliger la rotation pendant la durée des pulses (ce qui revient à dire que les pulses ont un spectre très large, donc forcément résonnant). La population de l'état excité (proportionnelle à la fluorescence) après la deuxième pulse est donc \begin{equation}
   p_e=\dfrac{1}{2}(1+\cos\Delta T)
   \end{equation}
   Tu as donc un résultat aussi précis que si tu avais réalisé une pulse $\pi$ de durée $T$.

   Pour l'appellation "Interférométre de Ramsay" ou "Franges de Ramsay", c'est parce que tu as un processus d'interférence type Mach-Zender, où les séparatrices sont tes pulses, et les deux voies possibles pour passer du fondamental à l'état excité viennent du fait que tu peut être excité à la première ou à la deuxième pulse, et ces deux chemins peuvent être constructifs ou destructifs.

   Pour la résolution, c'est celle d'un interferomètre à deux voies : \begin{equation}
   \delta_{min}=\dfrac{\sqrt{2}}{T\sqrt{N}}
   \end{equation}
   où $N$ est le nombre d'atomes sondés.

   \section{Système impur : Optical Bloch Equations}

   \subsection{Origine de l'impureté et approximations du modèle}
   On considère qu'un système est impur à partir du moment où il est couplé à un autre système qu'on ne veut/peut pas décrire (à savoir "l'environnement"). Dans ce cas on fait une réduction du système total en "traçant" sur l'environnement.

   Dans le cas qui nous intéresse ici, on va considérer notre système à 2 niveaux couplé à un reversoir (à savoir le vide et tout ses modes) qui vérifie ces deux hypothèses :\begin{itemize}
   \item \textbf{Born approximation} : On ne garde le terme d'interaction entre le système et l'environnement qu'à l'ordre le plus faible. D'après J-M Raymond ça signifie que toute perturbation du système envoyée au réservoir est "perdue".
   \item \textbf{Markov approximation} : Le réservoir n'a pas de mémoire, ou bien encore, dans la représentation de Heisenberg, que les observables du réservoir un temps de corrélation extrêmement court (de l'ordre de $\tau_c \approx 10^{-15} s$, c'est à dire la fréquence caractéristique des modes qui nous intéressent).
   \end{itemize}
   \subsection{Dérivation des OBE}
   On va considérer 3 phénomènes : \begin{itemize}
   \item L’interaction cohérente (donc évolution Hamiltonienne) avec le champ quasi-résonnant. On se place dans la représentation en interaction, et H s'écrit \begin{equation}
   H=-\dfrac{\hbar \Delta }{2} \sigma_z + \dfrac{\hbar \Omega}{2} \left( \sigma_+ e^{i\phi} + \sigma_- e^{-i\phi} \right)
   \end{equation}
   Avec $\Delta = \omega - \omega_{eg}$, $\Omega = -\dfrac{d\cdot E_0}{\hbar}$ et l'amplitude complexe du champ s'écrit  \begin{equation}
   E_1=E_0 e^{i\phi}
   \end{equation}
   \item La relaxation due à l'émission spontanée (jump operator $\sigma^-$ et temps caractéristique $T_1$)
   \item Le phase damping/pure dephasing du aux fluctuations locales du champ électrique/magnétique qui vont faire varier les niveaux d'énergie (par effet Stark / Zeeman) et provoquer un déphasage entre les deux états, mais sans modifier les populations (à creuser un peu).

   Le déphasage maximum éant un déphasage de $\pi$ entre les deux états, le jump operator qu'on va considérer est tout simplement $\sigma_z$, et le temps caractéristique associé est $T_2$
\end{itemize}

Dans ces conditions, on met tout ça dans le Lindbaldien et on résout. On trouve finalement les OBE : 
\begin{empheq}[box=\fbox]{align}
\dfrac{d \, \rho_{ee}}{dt} &= \dfrac{d}{\hbar} \mathrm{Im} (\rho_{eg} E_1^*) - \Gamma \rho_{ee} \\ 
\dfrac{d \, \rho_{eg}}{dt} &= i\Delta \rho_{eg} -i\dfrac{d}{2\hbar} E_1(\rho_{ee}-\rho_{gg}) - \gamma ' \rho_{eg}
\end{empheq}

Avec \begin{equation}
\gamma ' = \gamma + \dfrac{\Gamma}{2}
\end{equation}

Ou encore \begin{equation}
\dfrac{1}{T_2^*} = \dfrac{1}{T_2} + \dfrac{1}{2T_1}
\end{equation}

Remarque : le $\Gamma/2$ pour les cohérences il vient directement des termes en $-\dfrac{1}{2} L^\dagger_\mu L_\mu \rho$ et $-\dfrac{1}{2} \rho L_\mu^\dagger L_\mu $ dans le Lindbladien.
   \subsection{Autre formulations}
   \subsubsection{Vecteur de la sphère de Bloch}
   Soit la matrice densité : 
   \begin{equation}
   \rho=\begin{pmatrix}
   \rho_{ee} & \rho_{eg} \\ \rho_{ge} & \rho_{gg}
   \end{pmatrix}
   \end{equation}
   Avec $\rho_{ee}+\rho_{gg}=1$ et $\rho_{eg}=\rho_{ge}*$ donc 3 paramètres libres, on va donc réécrire  \begin{equation}
   \rho=\dfrac{1}{2}\begin{pmatrix}
   1+Z & X-iY \\ X+iY & 1-Z
   \end{pmatrix}
   \end{equation}
   Il se trouve que $X,Y$ et $Z$ sont les coordonnées de ton vecteur dans la sphère de Bloch. Plutôt pratique, et ça te donne une contrainte graphique sur tes cohérences.

  \end{document}

### 2020-07-10 théorie/fiches/Système à 2 niveaux/système à 2 niveaux.tex
tout d'un coup

--- 2020-07-10 théorie/fiches/Système à 2 niveaux/fiche.tex
@@
 \usepackage{empheq} % Tout ca pour encadrer des équations
+\usepackage{tcolorbox} % toujours plus de boites
+\usepackage{amsthm}% Pour faire des remarques (et de théorèmes et tout)
+
 \usetikzlibrary{optics}
@@
 \date\today
+\newtheorem*{remark}{Remarque}
 
@@
 
-   Remarque pour le couplage lumière-matière, c'est parce que l'opérateur dipôle $\textbf{D}=q\textbf{r}\begin{pmatrix}
+   \begin{tcolorbox}
+   \textbf{Remarque} : Pour le couplage lumière-matière, on considère que l'opérateur dipôle $\textbf{D}=q\textbf{r}\begin{pmatrix}
    0 & d \\ d^* & 0
-   \end{pmatrix}$ n'a pas de termes diagonaux, car les états propres du système ont une parité définie (pas de sens privilégié, symétrie centrale).
+   \end{pmatrix}$ n'a pas de termes diagonaux, car les états propres du système ont une parité définie par rapport à $\textbf{r} \to -\textbf{r}$.
+Dans ce cas, la distribution de charge (qui ne dépend pas du signe de la fonction d'onde) à une symétrie centrale, donc il n'y a pas de dipôle permanent (c'est à dire de terme diagonal).
+C'est le cas pour le SiV par exemple, mais pas pour le NV.
+
+   Les termes hors-diagonaux sont des dipôles de transition, et viennent du fait qu'il y a un déplacement de charge dû au changement d'état, d'où le fait que les transitions (dipolaires) entre états de même symétrie soit interdites.
+   \end{tcolorbox}
 
@@
 Remarque : le $\Gamma/2$ pour les cohérences il vient directement des termes en $-\dfrac{1}{2} L^\dagger_\mu L_\mu \rho$ et $-\dfrac{1}{2} \rho L_\mu^\dagger L_\mu $ dans le Lindbladien.
-   \subsection{Autre formulations}
-   \subsubsection{Vecteur de la sphère de Bloch}
+
+\begin{tcolorbox}
+\textbf{Attention} : $\Gamma, \gamma, T_1, T_2$ et les temps de vie en général sont homogènes à des pulsations (penser $e^{-\Gamma t} \simeq e^{i \omega t}$).
+
+La largeur naturelle en fréquence est donc de $\dfrac{\Gamma}{2 \pi}$ [Hz]
+\end{tcolorbox}
+   \subsubsection{Autre formulation : X,Y,Z}
    Soit la matrice densité : 
@@
    \end{equation}
-   Il se trouve que $X,Y$ et $Z$ sont les coordonnées de ton vecteur dans la sphère de Bloch. Plutôt pratique, et ça te donne une contrainte graphique sur tes cohérences.
+   Il se trouve que $X,Y$ et $Z$ sont les coordonnées de ton vecteur dans la sphère de Bloch. Plutôt pratique, et ça te donne une contrainte graphique sur tes cohérences. \\
+
+   Avec ce formalisme et en notant $E_1=E_1'+iE_1"$, les OBEs s'écrivent : 
+   \begin{align}
+   \dfrac{dZ}{dt}&=-\dfrac{d}{\hbar}(XE_1"+YE_1')-\Gamma (1+Z) \\
+   \dfrac{dX}{dt}&= +\Delta X +-\dfrac{d}{\hbar}ZE_1"-\gamma 'X \\
+   \dfrac{dY}{dt}&= -\Delta Y +-\dfrac{d}{\hbar}ZE_1'-\gamma 'Y
+   \end{align}
+
+   \subsection{Résolution dans le cas résonnant : Rabi amorti}
+   On a besoin d'un certain nombre d'hypothèses pour faire le calcul de Rabi proprement : \begin{itemize}
+   \item L'état est initialement dans $\ket{g}$, soit $z=-1$
+   \item Le champ extérieur est résonnant : $\Delta = 0$
+   \item le champ est purement réel : la rotation se fait autour de l'axe X (pas contrraignant si tu peux choisir X et Y)
+   \end{itemize}
+   Dans ces conditions, tu gardes toujours $X=0$ et tu te retrouves avec 2 équations du premier ordre en Z et Y que tu peux combiner en une équation du second ordre : 
+   \begin{equation}
+   \dfrac{d^2Z}{dt^2}+(\Gamma + \gamma ')\dfrac{dZ}{dt}+(\Omega^2 + \gamma ' \Gamma)Z = -\gamma ' \Gamma
+   \end{equation}
+
+   Tu trouves alors un steady state : \begin{align}
+   Z_s&=-\dfrac{\gamma' \Gamma}{\Omega ^2 + \gamma ' \Gamma} \\
+   Y_s &= \dfrac{\Omega \Gamma}{\Omega ^2 + \Gamma \gamma '}
+   \end{align}
+
+   On remarque donc d'une part que $Z_s$ est forcément négatif, contrairement à Rabi sans damping ou tu peux finir à Z=+1.
+Tu retrouves d'une certaine façon le résultat d'Einstein où tu ne peux pas avoir d'inversion de population.
+
+   D'autre part que l'état final à une certaine cohérence en Y (obligé pour avoir un effet de pompage) \\
+
+   Par ailleurs, en rajoutant comme hypothèse que :\begin{itemize}
+   \item La décohérence est lifetime limited : $\gamma ' = \Gamma /2$
+   \item Le damping est faible devant le couplage (couplage fort ?) : $\Omega \gg \Gamma$
+   \end{itemize}
+
+   Dans ces conditions, la solution temporelle devient simplement une oscillateur amorti : \begin{equation}
+   Z(t)=-\cos (\Omega t) e^{-3/2\Gamma t}
+   \end{equation}
+
+   On voit que d'une certaine façon, le damping en Y et en Z s'additionnent dans l'exponentielle.
+   \subsection{Steady state et saturation}
+   La suite au prochain épisode
 

### 2020-07-10 théorie/fiches/Système à 3 niveaux/lambda_system.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Lambda system}
  \section{Présentation}
  \includegraphics[width=\textwidth]{lambdasystem}
  On va considérer un couplage (sinusoïdal) uniquement entre $\ket{e}$ et $\ket{g_1}$ et entre $\ket{e}$ et $\ket{g_2}$. Le Hamiltonien du système s'écrit donc :
  \begin{equation}
  H=\hbar \begin{pmatrix}
  \omega_e & -\dfrac{\Omega_{R1}}{2}e^{-i\phi_1}e^{-i\omega_1t} & -\dfrac{\Omega_{R2}}{2}e^{-i\phi_2}e^{-i\omega_2t} \\
  -\dfrac{\Omega_{R1}}{2}e^{i\phi_1}e^{i\omega_1t} & \omega_{g1} & 0 \\
  -\dfrac{\Omega_{R2}}{2}e^{i\phi_2}e^{i\omega_2t} & 0 & \omega_{g2}
  \end{pmatrix}
  \end{equation}
  où $\phi_1$ et $\phi_2$ sont les composantes complexes des pulsations de Rabi $\Omega_{R1}$ et $\Omega_{R2}$.

  Dans ces conditions, on écrira un état quelconque du système à 3 niveaux sous la forme : 
  \begin{equation}
  \ket{\psi}=c_e e^{-i\omega_et} \ket{e} + c_{g1} e^{-i\omega_{g1}t} \ket{g1} + c_{g2} e^{-i\omega_{g2}t} \ket{g2}
  \end{equation}
  \subsection{Résolution dans le cas résonnant}
  En supposant que $\omega_1 = \omega_e - \omega_{g1}$ et $\omega_2 = \omega_e - \omega_{g2}$, on déduit de Schrödinger :

  \begin{align}
  \label{schro_resonnant}
  \dot{c_e}&=\dfrac{i}{2}(\Omega_{R1}e^{-i\phi_1}c_{g1}+\Omega_{R2}e^{-i\phi_2}c_{g2})\\
  \dot{c_{g1}}&=\dfrac{i}{2}\Omega_{R1}e^{i\phi_1}c_e \\
  \dot{c_{g2}}&=\dfrac{i}{2}\Omega_{R2}e^{i\phi_2}c_e 
  \end{align}

\section{Dark State}
On va se placer dans le cas résonnant, et on va considérer un état initialement dans une superposition des états fondamentaux : 
\begin{equation}
\label{psi0}
\ket{\psi}(t=0)=\cos\dfrac{\theta}{2} \ket{g1} + \sin\dfrac{\theta}{2} e^{i \psi}\ket{g2}
\end{equation}

Alors, en résolvant les équations \ref{schro_resonnant}, on trouve que :
\begin{equation}
c_e(t)=\dfrac{i \sin (\Omega t /2)}{\Omega}\left[\Omega_{R1}e^{-i\phi_1} \cos\dfrac{\theta}{2} + \Omega_{R2}e^{-i(\phi_2+\psi)} \sin\dfrac{\theta}{2} \right]
\end{equation}
où $\Omega = \sqrt{\Omega_{R1}^2+\Omega_{R2}^2}$. $c_{g1}$ et $c_{g2}$ peuvent se calculer mais c'est chiant.

On voit que pour certaines valeurs de $\theta$ et certains paramètres de couplage, tu peux annuler $c_e$, et donc créer un dark state, un état fondamental qui n'est pas excité par des champs résonnants. La condition c'est que :
\begin{equation}
\dfrac{\sin (\theta/2)}{\cos (\theta/2)} = \tan (\theta/2) = - \dfrac{\Omega_{R1}}{\Omega_{R2}} e^{-i(\phi_1-\phi_2-\psi)}
\end{equation}

ce qui te permet d'écrire, en remplaçant dans \ref{psi0} que le dark state généralisé s'écrit comme : 
\begin{equation}
\ket{\psi(t)}=\dfrac{\Omega_{R2}(t)e^{-i\phi_2}\ket{g1}+\Omega_{R1}(t)e^{-i\phi_1}\ket{g2}}{\sqrt{\Omega_{R1}^2+\Omega_{R2}^2}}
\end{equation}

Dans le cas où tu peux négliger les composantes complexes, tu vois qu'une façon de créer un dark state c'est de partir de $\ket{g1}$, avec $\Omega_{R2}$ allumé, mais $\Omega_{R1}$ éteint, puis d'allumer progressivement (adiabatiquement) $\Omega_{R1}$. (D'ailleurs ça marche même si tu ne négliges pas les composantes complexes).

Ca s'appelle du STImulated Raman Adiabatic Passage (STIRAP), et a priori, "lentement" ici c'est par rapport aux fréquences de Rabi, puisqu'il faut que les populations ait le temps de s'équilibrer. Bon c'est un peu bizarre vu que c'est justement la pulsation de Rabi que tu modifies, mais je le comprend comme : si tu veux atteindre un couplage $\Omega_R$, tu dois le faire dans un temps grand devant $1/\Omega_R$. Ca peut permettre de passer de $\ket{g1}$ à $\ket{g2}$ de façon plus certaine que par une pulse Rabi $\pi$

\section{Electromagnetically Induced Transparency (EIT)}

La suite au prochain épisode

  \end{document}

### 2020-07-10 théorie/fiches/optique pas quantique/optique.tex
tout d'un coup

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Optique pas quantique}
  \section{Ordres de grandeurs}
  \subsection{Force électrique Vs Force magnétique}
  La raison pour laquelle on s’intéresse par défaut plus au champ électrique que magnétique, c'est parce que le rapport des deux termes dans la force de Lorentz vaut :\begin{equation}
  \dfrac{F_E}{F_B}=\dfrac{q\textbf{E}}{q\textbf{v}\times \textbf{B}}=\dfrac{c}{v}
  \end{equation}
  Il se trouve qu'on est pas si loin des vitesse relativistes pour les électrons d'un atome, en effet, en partant du quantum de moment cinétique \begin{equation}
  \hbar=mvr=m_e\,v\,a_0
  \end{equation}
  Avec $m_e$ la masse de l'électron et $a_0$ le rayon de Bohr, que l'on détermine en posant le PFD : $\dfrac{m v^2}{r}=\dfrac{q^2}{4\pi \epsilon_0r^2}$, d'où :\begin{equation}
  a_0=\dfrac{\hbar^2 4\pi\epsilon_0}{m_e q_e^2}\approx 0.53 \mathrm{\AA}
  \end{equation}

  Et donc finalement : \begin{equation}
  \dfrac{F_E}{F_B}=\dfrac{c\hbar}{m_e a_0}=\dfrac{1}{\alpha}\approx 137
  \end{equation}

  C'est également le ratio des transitions Dipolaire électrique Vs Dipolaire magnétique par défaut.

  \end{document}

### 2020-07-10 théorie/fiches/these.tex
tout d'un coup

--- 2020-07-10 théorie/fiches/Spectro/Spectroscopie.tex
@@
 \usepackage{tikz}
-\usepackage{bm} % pour faire des vecteurs gras
 \usepackage{subcaption} % package pour faire des subfigures
@@
 \begin{document}
-\chapter{L'atome}
-  \section{Le moment cinétique}
-  \subsection{Définition, relations de commutation}
-  \textbf{Convention} : Dans tout ce chapitre, les vecteurs seront notés en \textbf{gras} et les opérateurs avec des chapeaux.
+\chapter{Introduction}
+  \section{The diamond}
+  The Diamond is a crystal comprised solely of carbon atoms, each one being bonded to four other atoms through covalent bonds.
+Because the covalent bonds are the strongest kind of chemical bonds, even more so for light atoms such as carbon, the diamond shows a a great
 
-  En partant de la définition classique $\bm{L}=\bm{r} \times \bm{p}$ pour une particule, on obtient la version quantique avec la magie des chapeaux : \begin{equation}
-  \hat{\bm{L}}=\hat{\bm{r}} \times \hat{ \bm p}
+  \section{The NV Center}
+  A compléter à un moment,
+
+  Hamiltonien de spin simplifié : 
+  \begin{equation}
+  \hat{\mathcal{H}}_s=D S_z^2 + \gamma_e \textbf{B}\cdot\hat{\textbf{S}}
   \end{equation}
- Alors avec les relations de commutation sur $\hat{\bm{r}}$ et $\hat{ \bm p}$, on trouve : \begin{equation}
- [\hat L_x, \hat L_y]=i\hbar \hat L_z \quad , \quad [\hat L_y, \hat L_z]=i\hbar \hat L_x \quad , \quad [\hat L_z, \hat L_x]=i\hbar \hat L_y
- \end{equation}
- Qu'on peut également synthétiser en : \begin{equation}
- \hat{\bm L} \times \hat{\bm L} = i\hbar \hat{\bm L}
- \end{equation}
- C'est cette relation de commutation qu'on prendra comme définition d'un opérateur de moment cinétique, y compris quand $\bm{L}=\bm{r} \times \bm{p}$ n'existent pas (coucou le spin).
-Pour généraliser, on écrira $\hat{\bm J}$ pour un moment cinétique quelconque
- \subsection{Valeurs propres}
- On définit $\hat J^2= \hat J_x^2 + \hat J_y^2 + \hat J_z^2$. $ \hat J^2$ commute avec chacune des composantes de $\hat{\bm J}$ : \begin{equation}
- [\hat J^2, \hat{\bm J}]=0
-\end{equation}  
-(Composantes qui ne commutent pas entre elles, cf avant).
-
-On va alors former un ECOC à partir de $\hat{J^2}$ et d'une des trois composante qui sera par convention $\hat{J_z}$, c'est à dire que l'état du moment cinétique sera entièrement défini par la velur des deux nombres quantiques $j$ et $m$ sans dimension tels que :\begin{align}
-\hat{J^2}\ket{j,m}&=j(j+1) \hbar^2 \ket{j,m} \\
-\hat J_z\ket{j,m}&= m\hbar \ket{j,m}
-\end{align}
-$j$ est forcément positif car $\hat{J^2}$ est défini positif, et tu trouves que tu ne peux avoir que des valeurs demi-entières et que $-j \leq m \leq j$ par l'application répétée de $\hat L_\pm$ défini par :\begin{equation}
-\hat L_\pm= \hat L_x \pm i \hat L_y
-\end{equation}
-Et qui vérifie :\begin{equation}
-\hat L_\pm \ket{j,m} \propto \ket{j,m \pm 1}
-\end{equation}
-\subsection{Moment cinétique orbital et harmoniques sphériques}
-On va maintenant s'intéresser à une particule ponctuelle qui a un moment cinétique classique. On peut donc bien écrire ici $\hat{\bm{L}}=\hat{\bm{r}} \times \hat{ \bm p}$
-
-On se placera dans les coordonnées polaires habituelles ($r$,$\theta$,$\phi$) avec l'axe z comme axe polaire, $0 \leq \theta \leq \pi$ la colatitude en bon français et $0 \leq \phi \leq 2\pi$ l'angle azimutal.
-
-Avec ces notations, en se rappelant que $\hat p_x = -i\hbar\dfrac{d}{dx}$ lorsqu'on raisonne en fonction d'onde, on trouve en coordonnées polaires : \begin{equation}
-\hat L_z= -i\hbar \dfrac{d}{d\phi}
-\end{equation}
-
-Ce qui veut dire que les orbitales, c'est à dire les fonctions d'onde des états propres de $\hat L^2$ et $\hat L_z$, vérifient : \begin{equation}
-\hat L_z \psi_{lm}(r,\theta, \phi)= -i\hbar \dfrac{d}{d\phi} \psi_{lm}(r,\theta, \phi) = \hbar m \psi_{lm}(r,\theta, \phi)
-\end{equation}
-D'où : \begin{equation}
-\psi_{lm}(r,\theta, \phi)= \varphi(r,\theta)e^{i m \phi}
-\end{equation}
-
-Puisqu'une rotation de $\phi$ de $2\pi$ doit donner la même fonction d'onde (?
-pourquoi c'est pas vrai sur la sphère de Bloch, parce que ce n'est pas l'espace physique ?), on en déduit que \textbf{m et don l sont nécessairement entiers}
-
-C'est également vrai pour $\hat L^2$ mais sa formule est plus dégueu : \begin{equation}
-\hat L^2= -\hbar ^2 \left( \dfrac{1}{\sin \theta}\dfrac{d}{d\theta}\sin \theta\dfrac{d}{d\theta}+ \dfrac{1}{\sin ^2 \theta} \dfrac{d}{d\phi} \right)
-\end{equation}
-
-En faite c'est la partie angulaire du laplacien sphérique : \begin{equation}
-\Delta = \dfrac{1}{r}\dfrac{d^2}{dr^2}-\dfrac{1}{r^2\hbar^2}\hat L^2
-\end{equation}
-
-On constate également que $\hat L^2$ et $\hat L_z$ n'agissent pas sur la partie radiale de l'orbitale, on va donc décomposer l'orbitale en sa partie radiale et sphérique : \begin{equation}
-\psi_{lm}(r,\theta,\phi)=R_{l}(r)Y_{lm}(\theta,\phi)
-\end{equation}
-Où les $Y_{lm}$ sont les \textbf{Harmoniques sphériques}, c'est à dire les fonction propres de $\hat L^2$ et $\hat L_z$ normalisées à 1.
-
-On remarque que $L_z$ n'agit pas sur la partie radiale de $\psi_lm$, d'où le fait que $R_{l}$ ne dépendent pas de $m$
-
-\subsection{Lien avec le moment magnétique}
-Toute particule chargée avec un moment cinétique (orbital ou non) dispose également d'un moment magnétique.
-
-Pour rappel, l'interaction d'un moment magnétique avec un champ s'écrit : \begin{align}
-W &= - \bm \mu \cdot \bm B \\
-\bm \Gamma &= \bm \mu \times \bm B
-\end{align}
-\subsubsection{Pour un moment cinétique orbital}
-Analogie classique : tu as une spire de courant formée par ta particule et son mouvement (circulaire ici mais je crois que c'est général) :
-\begin{align}
-\bm L &= \bm r \times \bm p = m_q r v \bm u \\
-\bm \mu &= I S \bm u = \dfrac{qv}{2\pi r}\pi r^2 \bm u = \dfrac{q}{2} v r \bm u
-\end{align}
-D'où : \begin{equation}
-\bm \mu = \gamma_0 \bm L \quad \mathrm{avec} \quad \gamma_0=\dfrac{q}{2 m_q}
-\end{equation}
-\emph{note :} $q$ est algébrique, dans le cas de l'électron $q=-e$
-
-$\gamma_0$ est appelé le \textbf{rapport gyromagnétique} de ta particule. \\
-
-On définit également le \textbf{magnéton de Bohr} comme le quanta de moment magnétique (du au mouvement orbital) pour l'électron : 
-\begin{equation}
-\mu_B=\gamma_0 \hbar= -\dfrac{q \hbar}{2 m_e}
-\end{equation}
-\subsubsection{Pour un moment cinétique intrinsèque}
-De la même façon, on postule (je crois pas qu'il y ait de preuve, si ce n'est expérimentale) que les moments cinétiques intrinsèques sont proportionnels à un moment magnétique.
-
-On doit quand même introduire un facteur sans dimension supplémentaire par rapport au moment cinétique orbital : le \textbf{facteur de Landée $g$} : \begin{equation}
-\bm \mu_s= g\dfrac{q}{2m_q} \bm J
-\end{equation}
-
-Les valeurs numériques de g sont : \begin{align*}
-\mathrm{electron} &\approx -2.00 \\
-\mathrm{proton} &\approx +5.59 \\
-\mathrm{neutron} &\approx -3.83
-\end{align*}
-
-\subsection{Composition de moments cinétiques}
-\begin{figure}[h]
-\centering
-\includegraphics[width=.5\linewidth]{Clebs_Jordan.jpg}
-\caption{Clebs-Jordan, le cousin de Clebsch-Gordan}
-\end{figure}
-Ça peut concerner deux particules en interaction ou un moment orbital et un moment intrinsèque. On va donc considérer deux moments : $\hat{\bm J_1}$ et $\hat{\bm J_1}$
-
-Dans la base découplée, ton système est donc décrit par les 4 nombres quantiques naturels : $\ket{j_1,m_j,j_2,m_2}$.
-
-Cependant on constate que $\hat{\bm J}=\hat{\bm J_1}+\hat{\bm J_2}$ est également un moment cinétique, on peut alors former un nouvel ECOC avec les 4 observables : $\hat J^2, \hat J_z, \hat J_1^2$ et $\hat J_2^2$.
-Les 4 nouveaux nombres quantiques de la base couplée seront donc : $\ket{j,m,j_1,j_2}$.
-
-Il y a deux avantages à passer dans la base couplée : 
-\begin{itemize}
-\item D'une part dans le cas de moment cinétiques orbitaux, $J$ représente bien le moment cinétique total du système
-\item D'autre part dans le cas d'une interaction, tu vas très souvent te retrouver (au premier ordre) avec un terme dans le hamiltonien en $\hat{\bm J_1} \cdot \hat{\bm J_2}$.
-Or $\hat{\bm J_1} \cdot \hat{\bm J_2}= \dfrac{1}{2}(\hat J^2- \hat J_1^2 - \hat J_2^2$ (source : mon cours de troisième).
-Donc les états de la base couplée sont états propres du Hamiltonien, et ça c'est toujours sympathique.
-\end{itemize}
-
-\subsubsection{Passage d'une base à l'autre}
-
-Dans la théorie tu utilises $J^-=J_1^-+J_2^-$ en partant de l'état $m=j$ qui est présent dans les deux bases, dans la pratique tu regardes les coefs de Clebsch-Gordon :
-
   \begin{center}
-\includegraphics[scale=.35]{Clebsch_Gordan}
+  $D=2.87 \,$GHz et $\gamma_e=2.8\,\mathrm{MHz}/\mathrm{G}$
 \end{center}
 
-(je sais c'est écrit petit, mais c'est ces vicieux de la physique des particules).
-
-Dans tous les cas tu constates que $\hat J_1^2$ et $\hat J_2^2$ sont présents dans les deux bases donc tu n'as que 2 nombres quantiques à changer.
-
-\subsection{Lois de conservations}
-
-Pour la projection c'est très clair, vu que $m=m_1+m_2$. Le moment total c'est plus compliqué vu qu'il y a la conservation de l'énergie qui s'en mêle.
-
-\section{Nombres quantiques d'un état atomique}
-Dans toute cette partie on s’intéresse à l'atome d'hydrogène. Ça se généralise assez bien aux alcalins mais tout le reste c'est plus compliqué.
-\subsection{Hamiltonien}
-Dans l'absolu c'est le problème à deux corps, que tu sépare en une partie avec les coordonnées du centre de masse, et une partie avec les coordonnées relatives.
-Dans la pratique on va considérer l'électron dans un potentiel central généré par le proton immobile dans le référentiel du labo, parce qu'il y a un rapport 2000 entre les masses.
-De toute façon les calculs sont identiques.
-
-\begin{equation}
-H=\dfrac{p^2}{2m}+V(r)
-\end{equation}
-
-On s’intéresse aux solutions stables, donc ESIT : \begin{equation}
-\Delta \psi (\bm r) + V(r) \psi (\bm r) = E \psi (\bm r)
-\end{equation}
-
-Avec la définition du laplacien sphérique en fonction de $\hat L^2$ et la décomposition en harmonique sphérique $\psi_{lm}(r,\theta,\phi)=R_{l}(r)Y_{lm}(\theta,\phi)$, on trouve finalement une EDO sur $R_l$ : \begin{equation}
-\left(-\dfrac{\hbar^2}{2m}\dfrac{1}{r}\dfrac{d^2}{dr^2}r + \dfrac{l(l+1)\hbar^2}{2mr^2} + V(r) \right) R_l(r) = E R_l(r)
-\end{equation} \\
-
-\textbf{Remarque} : C'est parce que $\hat H$, c'est à dire $\hat V$, commute avec $L_z$ et $L^2$ que tu peux décomposer les solutions comme ça.
-Et c'est vrai ici parce que V est invariant selon $\theta$ et $\phi$.
-\subsection{Solutions de l'ESIT}
-
-Bon en vrai c'est un peu chaud les solutions de l'EDO, ce qu'on peut retenir c'est que : \begin{itemize}
-\item Comme pour tout système lié, il y a discrétisation des $E<0$ possibles
-\item Pour un $l$ donné, il n'existe qu'une solution pour chaque valeur propre $E_{n'}$ donnée. On peut donc définir sans équivoque $\ket{l,n'}$
-\item La partie radiale de la fonction d'onde $R_{n'l}(r)$ a exactement $n'$ zéros.
-\item L'énergie $E_{n'l}$ correspondante est \begin{equation}
-E_{n'l}=-\dfrac{E_0}{(n'+l+1)^2}
-\end{equation}
-Avec $E_0$ l'énergie du fondamental, c'est à dire l'énergie d'ionisation de l'atome d'hydrogène, c'est à dire un Ryldberg (techniquement les Ry c'est en $\mathrm{cm}^{-1}$ je crois) : \begin{equation}
-E_0=\dfrac{m_e e^4}{2\hbar^2}\approx13.6 \mathrm{eV}
-\end{equation}
-\item A cause de l'expression de l'énergie, on utilise comme nombre quantique \begin{equation}
-n=n'+l+1
-\end{equation}  Donc maintenant l'énergie peut ne s'exprimer qu'en fonction de $n$, mais elle dépend quand même de $l$. Bref en vrai ça complique la vie plus qu'autre chose.
-\end{itemize}
-\subsection{Notations spectro}
-Bon je t'épargne l'histoire (c'est à cause des séries de raies de l'atome d'hydrogène : sharp, principal, diffuse...) mais 3d $\Rightarrow n=3\; \mathrm{et} \;l=2$
-\subsection{Solutions particulières}
-\subsubsection{Orbitale 1s}
-$l=0$ et $n'=0$, $Y_{l,m}(\theta, \phi)=1/\sqrt{4\pi}$, ta fonction est : \begin{equation}
-\psi_{1,0,0}(\bm r) = \dfrac{e^{-r/a_0}}{\sqrt{\pi a_0^3}}
-\end{equation}
-\textbf{note :} pour tout $n'$ tu es dominé aux grands $r$ par un $e^{-r}$. \\
-
-Si on s'intéresse à la densité de présence radiale, il faut intégrer sur une sphère de rayon $r$, d'où :
-\begin{equation}
-\rho(r)=|\psi_{1,0,0}(\bm r)|^2 4 \pi r^2= e^{-2r/a_0} \dfrac{r^2}{a_0^3}
-\end{equation}
-
-Et on constate que $<r> = a_0$, la distance moyenne de l'électron est bien le rayon de Bohr
-
-\subsubsection{Orbitale 2s}
-Toujours une symétrie sphérique car $l=0$, mais cette fois ci ta fonction s'annule et change de signe en $r=a_0/2$. Tu as donc une région interdite pour ton électron. \begin{equation}
-\psi_{2,0,0}(\bm r) \propto \left(1-\dfrac{r}{2 a_0}\right) e^{-r/a0}
-\end{equation}
-
-\subsubsection{Orbitale 2p}
-Ici on commence les embrouilles car $l=1$ et donc $m$ commence à prendre des valeurs non nulles.
-En plus $p_x$ et $p_y$ c'est des c.l des états propres (ils ne sont donc pas état propre de $L_z$, ce qui est plutôt logique vu que c'est les états propres de $L_x$ et $L_y$).
-
-\begin{align}
-\psi_{pz}(\bm r) &\propto Y_{1,0}(\theta,\phi) e^{-r/a_0} \propto \cos \theta  e^{-r/a_0} \\
-\psi_{px}(\bm r) &\propto (Y_{1,1}+Y_{1,-1})(\theta,\phi) e^{-r/a_0} \propto \sin \theta \sin \phi e^{-r/a_0} \\
-\psi_{py}(\bm r) &\propto (Y_{1,1}-Y_{1,-1})(\theta,\phi) e^{-r/a_0} \propto \sin \theta \cos \phi e^{-r/a_0}
-\end{align}
-
-\subsection{Et le spin dans tout ça}
-Ton système c'est un électron, donc $s=1/2$ on ne le note pas.
-Sa projection $m_s=\pm 1/2$ n'influe ni sur les niveaux d'énergie (sans champ magnétique), ni sur les orbitales.
-En revanche ça joue sur la population des états.
-
-\section{Modification des niveaux d'énergie}
-\subsection{Couplage spin-orbite : structure fine}
-\subsection{Effet Zeeman}
-\subsection{Effet Stark}
-\subsection{Couplage spin-spin : structure hyper-fine}
-\subsection{Quantification du champ : Lamb shift}
+  toto
 

### 2020-07-10 data/20200213/scans_t_pola_1/test_tf.py

@@ -1,0 +1,28 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+#Bien essayé mais c'est non la tf
+N=197
+file='13.txt'
+
+for file in ['13.txt','23.txt','33.txt','43.txt'] :
+	x=np.zeros(N)
+	y=np.zeros(N)
+	with open(file,'r') as f:
+		f.readline()
+		f.readline()
+		for i in range(N):
+			line=f.readline()
+			line=line.split()
+			x[i]=float(line[0])
+			y[i]=float(line[1])
+
+	y_avg=sum(y[100:])/len(y[100:])
+	y_plt=np.log(-y[:75]+y_avg)
+	x_plt=x[:75]
+
+
+	plt.plot(x_plt,y_plt)
+
+	
+plt.show()

### 2020-07-10 data/20200225/scans_t1/plot_t1.py

@@ -1,0 +1,62 @@
+import sys
+import time
+import random
+import os
+import matplotlib.pyplot as plt
+import numpy as np
+
+from scipy.optimize import curve_fit
+
+voltage=np.arange(1.30,2.3,0.05)
+
+
+fnames=['%i.txt'%(v*100) for v in voltage]
+taux=[]
+error=[]
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def stretch_exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-np.sqrt(x/tau))+ss
+	p0=[Amp,ss,tau]
+	popt, pcov = curve_fit(f, x, y, p0,sigma=np.ones(len(y))*2E3, absolute_sigma=True)
+	return(popt,pcov,f(x,popt[0],popt[1],popt[2]))
+
+for fname in fnames :
+	x,y=extract_data(fname)
+	popt,pcov,yfit=stretch_exp_fit(x,y)
+	taux+=[popt[2]]
+	perr = np.sqrt(np.diag(pcov))
+	error+=[perr[2]]
+
+taux=np.array(taux)
+
+print(taux,error)
+x,y=extract_data('scan_sumi2_correspond_ESR.txt')
+x=x[88:155]
+y=y[88:155]
+y=y-min(y)
+taux=taux-min(taux)
+plt.plot(x,y/max(y))
+plt.plot(voltage,taux/max(taux))
+#plt.errorbar(voltage,taux,yerr=error)
+
+plt.show()

### 2020-07-10 data/20200226/scans_t1/plot_t1.py

@@ -1,0 +1,75 @@
+import sys
+import time
+import random
+import os
+import matplotlib.pyplot as plt
+import numpy as np
+
+from scipy.optimize import curve_fit
+
+voltage=np.arange(1.0,2.5,0.02)
+
+
+fnames=['%i.txt'%(v*100) for v in voltage]
+taux=[]
+error=[]
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def stretch_exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-np.sqrt(x/tau))+ss
+	p0=[Amp,ss,tau]
+	popt, pcov = curve_fit(f, x, y, p0,sigma=np.ones(len(y))*2E3, absolute_sigma=True)
+	return(popt,pcov,f(x,popt[0],popt[1],popt[2]))
+
+def exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-x/tau)+ss
+	p0=[Amp,ss,tau]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,pcov,f(x,popt[0],popt[1],popt[2]))
+
+for fname in fnames :
+	x,y=extract_data(fname)
+	popt,pcov,yfit=exp_fit(x,y)
+	taux+=[popt[2]]
+	perr = np.sqrt(np.diag(pcov))
+	error+=[perr[2]]
+
+taux=np.array(taux)
+
+print(taux,error)
+x,y=extract_data('scan_sumi2_correspond_ESR.txt')
+x=x[88:155]
+y=y[88:155]
+y=y-min(y)
[…]

### 2020-07-10 data/20200304/scans_t1/traitement.py

@@ -1,0 +1,102 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,x0,sigma) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def lor_fit(x,y,AAmp=None,x0=None,sigma=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,x0,sigma) :
+		return Amp*1/(1+((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-x/tau)+ss
[…]

### 2020-07-10 data/20200305/traitement.py

@@ -1,0 +1,102 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,x0,sigma) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def lor_fit(x,y,AAmp=None,x0=None,sigma=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,x0,sigma) :
+		return Amp*1/(1+((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-x/tau)+ss
[…]

### 2020-07-10 data/20200527/traitement_tdv.py

@@ -1,0 +1,99 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,x0,sigma) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def lor_fit(x,y,Amp=None,x0=None,sigma=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,x0,sigma) :
+		return Amp*1/(1+((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-x/tau)+ss
[…]

### 2020-07-10 data/20200617/traitement_scan_gbf.py

@@ -1,0 +1,136 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None,ss=0) :
+	if not ss :
+		ss=y[0]
+	if not Amp :
+		if max(y)-ss > ss-min(y) :
+			Amp=max(y)-ss
+		else :
+			Amp=min(y)-ss
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/5)]-x[0]
+	def f(x,Amp,x0,sigma,ss) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)+ss
+	p0=[Amp,x0,sigma,ss]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2],popt[3]))
+
+def lor_fit(x,y,Amp=None,x0=None,sigma=None) :
+	if not ss :
+		ss=y[0]
+	if not Amp :
+		if max(y)-ss > ss-min(y) :
+			Amp=max(y)-ss
+		else :
+			Amp=min(y)-ss
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/5)]-x[0]
+	def f(x,Amp,x0,sigma,ss) :
+		return ss+Amp*1/(1+((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma,ss]
+	popt, pcov = curve_fit(f, x, y, p0)
[…]

### 2020-07-10 data/20200619/traitement_scan_gbf.py

@@ -1,0 +1,133 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None,ss=0) :
+	if not ss :
+		ss=y[0]
+	if not Amp :
+		if max(y)-ss > ss-min(y) :
+			Amp=max(y)-ss
+		else :
+			Amp=min(y)-ss
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/5)]-x[0]
+	def f(x,Amp,x0,sigma,ss) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)+ss
+	p0=[Amp,x0,sigma,ss]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2],popt[3]))
+
+def lor_fit(x,y,Amp=None,x0=None,sigma=None,ss=None) :
+	if not ss :
+		ss=y[0]
+	if not Amp :
+		if max(y)-ss > ss-min(y) :
+			Amp=max(y)-ss
+		else :
+			Amp=min(y)-ss
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/5)]-x[0]
+	def f(x,Amp,x0,sigma,ss) :
+		return ss+Amp*1/(1+((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma,ss]
+	popt, pcov = curve_fit(f, x, y, p0)
[…]

### 2020-07-10 manips/PL en fonction de l'angle de B/largeur diamants/data_sumi4/traitement.py

@@ -1,0 +1,60 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b)
+
+def gauss_fit(x,y,Amp,x0,sigma) :
+	def f(x,Amp,x0,sigma) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(x for x in popt)
+
+def lor_fit(x,y,Amp,x0,sigma) :
+	def f(x,Amp,x0,sigma) :
+		return Amp*1/(1+((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(x for x in popt)
+
+def ask_name():
+	qapp = QApplication(sys.argv)
+	fname,filters=QFileDialog.getOpenFileName()	
+	return fname
+
+
+fname='ESR_sumi4_5-00V_80mA.txt'
+
+x,y=extract_data(fname)
+y=y/max(y)
+
+plt.plot(x*1000,y,label='data')
+x,y=extract_data('ESR_5V_simu.txt')
+plt.plot(x,y,label='simu')
+
+
+plt.legend()
+
+plt.show()

### 2020-07-10 manips/PL en fonction de l'angle de B/simu python/anim/data.py

@@ -1,0 +1,260 @@
+import numpy as np
+from numpy import cos,sin,sqrt,pi
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+
+
+
+Sz=np.array([[1,0,0],[0,0,0],[0,0,-1]])
+Sy=np.array([[0,-1j,0],[1j,0,-1j],[0,1j,0]])
+Sx=np.array([[0,1,0],[1,0,1],[0,1,0]])
+Sz2=np.array([[1,0,0],[0,0,0],[0,0,1]]) # Pour éviter une multilplcation matricielle
+
+c0=np.array([0,0,0])
+c1=np.array([-1,1.,-1])/sqrt(3)
+c2=np.array([1,1,1])/sqrt(3)
+c3=np.array([-1,-1,1])/sqrt(3)
+c4=np.array([1,-1,-1])/sqrt(3)
+
+carb=[c0,c1,c2,c3,c4]
+
+sigma=5E6 #Hz
+gamma=2.8E6 #Hz.G-1
+N=500
+h=6.63E-34 #SI
+D=2.88E9
+
+
+
+def ChampMag(theta,phi,amp): #amp en gauss, angles en degrés
+	theta=theta*pi/180
+	phi=phi*pi/180
+	return amp*np.array([sin(theta)*cos(phi),cos(theta),sin(theta)*sin(phi)])
+
+
+
+def show(B) :
+	fig = plt.figure()
+	ax = fig.add_subplot(111, projection='3d')
+	xs=[]
+	ys=[]
+	zs=[]
+	for c in carb :
+		xs+=[c[0]]
+		ys+=[c[1]]
+		zs+=[c[2]]
+		ax.scatter(xs, ys, zs, marker='o')
+	B=B/np.linalg.norm(B)
+	ax.scatter(B[0],B[1],B[2], marker='o')
+	for c in carb[1:] :
+		ax.plot3D([c0[0],c[0]],[c0[1],c[1]],[c0[2],c[2]],'blue')
+
+	ax.plot3D([c0[0],B[0]],[c0[1],B[1]],[c0[2],B[2]],'red')
+	B=-B
+	ax.scatter(B[0],B[1],B[2], marker='o')
+	ax.plot3D([c0[0],B[0]],[c0[1],B[1]],[c0[2],B[2]],'red')
+	ax.set_xlabel('Axe x')
+	ax.set_ylabel('Axe y')
[…]

### 2020-07-10 manips/PL en fonction de l'angle de B/simu python/contraste.py

@@ -1,0 +1,17 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+gamma_min=1/200
+gamma_max=1/5
+gamma_phonon=1/3
+
+def contraste(x):
+    return 1-((x+gamma_max+gamma_phonon)*(x+3*(gamma_min+gamma_phonon)))/((x+gamma_min+gamma_phonon)*(x+3*(gamma_max+gamma_phonon)))
+
+
+absc=np.linspace(0,2,500)
+
+ord=[contraste(x) for x in absc]
+
+plt.plot(absc,ord)
+plt.show()

### 2020-07-10 manips/PL en fonction de l'angle de B/simu python/data.py

@@ -1,0 +1,394 @@
+import numpy as np
+from numpy import cos,sin,sqrt,pi
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+
+
+
+Sz=np.array([[1,0,0],[0,0,0],[0,0,-1]])
+Sy=np.array([[0,-1j,0],[1j,0,-1j],[0,1j,0]])
+Sx=np.array([[0,1,0],[1,0,1],[0,1,0]])
+Sz2=np.array([[1,0,0],[0,0,0],[0,0,1]]) # Pour éviter une multilplcation matricielle
+
+c0=np.array([0,0,0])
+c1=np.array([-1,1.,-1])/sqrt(3)
+c2=np.array([1,1,1])/sqrt(3)
+c3=np.array([-1,-1,1])/sqrt(3)
+c4=np.array([1,-1,-1])/sqrt(3)
+
+carb=[c0,c1,c2,c3,c4]
+
+sigma=5E6 #Hz
+gamma=2.8E6 #Hz.G-1
+N=500
+h=6.63E-34 #SI
+D=2.88E9 #Hz
+
+inhomogenous_broad=0 #Hz
+gamma_f=4E6 #Hz
+
+tau_las=1e-4 #s
+t1_phonons=3e-3 #s
+t1_non_deg=200e-3 #s
+
+mag_field = 100 #G
+
+
+
+def ChampMag(theta,phi,amp): #amp en gauss, angles en degrés
+    theta=theta*pi/180
+    phi=phi*pi/180
+    return amp*np.array([sin(theta)*cos(phi),cos(theta),sin(theta)*sin(phi)])
+
+
+
+def show(B) :
+    def link(ax,c1,c2,linewidth=1,color='blue',linestyle='-'):
+        ax.plot3D([c1[0],c2[0]],[c1[1],c2[1]],[c1[2],c2[2]],color=color,linewidth=linewidth,linestyle=linestyle)
+
+    def sphere(ax,c,linewidth=1,color='black'):
+        ax.scatter(c[0],c[1],c[2], marker='o',linewidth=linewidth,color=color)
+
+
+    fig = plt.figure()
+    ax = fig.add_subplot(111, projection='3d')
+    
+    B=B/np.linalg.norm(B)/np.sqrt(3)
+
+    for c in carb :
+        link(ax,c0,c,2)
+        sphere(ax,c,linewidth=10)
+
[…]

### 2020-07-10 manips/PL en fonction de l'angle de B/simu t1 sqrt/simu_t1_stretch.py

@@ -1,0 +1,56 @@
+import numpy as np
+from numpy import exp,sqrt
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from numpy.random import random,seed
+
+
+
+def exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-x/tau)+ss
+	p0=[Amp,ss,tau]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def stretch_exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-np.sqrt(x/tau))+ss
+	p0=[Amp,ss,tau]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+seed(1234)
+
+taux=random(1000)
+
+x=np.linspace(0,5,200)
+y=np.zeros(200)
+
+for tau in taux :
+	y+=exp(-x/tau)
+
+y=y/len(taux)
+plt.plot(x,y,'x')
+
+popt,yfit=stretch_exp_fit(x,y)
+plt.plot(x,yfit,label='stretch')
+
+popt,yfit=exp_fit(x,y)
+plt.plot(x,yfit,label='exp')
+
+plt.legend()
+plt.show()
+

### 2020-07-10 manips/PL en fonction de l'angle de B/tdv_diamant_rose/traitement.py

@@ -1,0 +1,121 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	def f(x,Amp,x0,sigma) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)
+	p0=[Amp,x0,sigma]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2]))
+
+def lor_fit(x,y,Amp=None,x0=None,sigma=None,origin=None) :
+	if not Amp :
+		Amp=max(y)
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/10)]-x[0]
+	if not origin :
+		origin=y[0]
+	def f(x,Amp,x0,sigma,origin) :
+		return Amp*1/(1+((x-x0)/(sigma))**2)+origin
+	p0=[Amp,x0,sigma,origin]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2],popt[3]))
+
+def exp_fit(x,y,Amp=None,ss=None,tau=None) :
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
[…]

### 2020-07-10 python/data.py

@@ -1,0 +1,49 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+def lecture (fichier,sep=None, colonne=0): #colonne=0 => rend le tableau complet
+	data=[]
+	with open(fichier,"r") as f:	
+		for line in f :
+			line=line.split(sep)
+			try :
+				float(line[0])
+			except :
+				continue
+			ligne=[]
+			for elem in line :
+				ligne+=[float(elem)]
+			data+=[ligne]
+	npdata=np.array(data)
+	if colonne==0 :
+		return npdata
+	else :
+		return npdata[:,colonne-1]
+		
+def cs(x,n=3): # retourne x avec n chiffres significatifs
+	y=round(x,-int(floor(log10(abs(x)))+1-n))
+	if y-int(y)==0 :
+		y=int(y)
+	return str(y)
+		
+def linfit(x,y): #retourne a,b,label avec f(x)=ax+b 
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b=np.linalg.lstsq(A,y,rcond=None)[0]
+	lab='fit {0:1.2e}*X{1:+1.2e}'.format(a,b)
+	return a,b,lab
+
+
+		
+def s1():
+	data=lecture("s1")
+	plt.plot(data[:,0],data[:,1],'b-x',label='label',ms=7, mew=3)
+	plt.xlabel('Axe x')
+	plt.ylabel('Axe y')
+	
+s1()
+plt.legend()
+plt.show()
+				
+        
+
+

### 2020-07-10 python/poubelle.py

@@ -1,0 +1,57 @@
+import numpy as np
+from numpy import exp,cos,sin
+import matplotlib.pyplot as plt
+
+sigma=1
+gamma_las=0.1 #puissance laser
+ratio_t1=0.5 #ratio de r1 venant de l'effet
+
+def f1(x,ratio_t1):
+	#return (1-ratio_t1)+ratio_t1*exp(-x**2/(2*sigma)**2)
+	return (1-ratio_t1)+ratio_t1*1/(1+(x/sigma)**2)
+
+def f2(x,gamma_las):
+	return((x+gamma_las)/(2*x+gamma_las))
+
+def width(x,y):
+	hm=(max(y)+min(y))/2
+	i=0
+	while (y[i]-hm)*(y[i+1]-hm) > 0 :
+		i+=1
+	return x[i]
+
+x=np.linspace(-10,10,100)
+y=f1(x,ratio_t1)
+z=f2(y,gamma_las)
+
+def simple() :
+	print(width(x,y),width(x,z))
+	plt.plot(x,y/max(y),label='R1')
+	plt.plot(x,z/max(z),label='PL')
+	plt.legend()
+
+def widths():
+	n_x=1000
+	n_gamma=1000
+	ratios=[0.2,0.5,1]
+	fig,ax=plt.subplots(2)
+	x=np.linspace(-10,10,n_x)	
+	for ratio_t1 in ratios :
+		delta_width=[]	
+		contraste=[]	
+		gamma_range=np.linspace(0,5*ratio_t1,n_gamma)
+		y=f1(x,ratio_t1)
+		for gamma_las in gamma_range :		
+			z=f2(y,gamma_las)
+			delta_width+=[width(x,y)-width(x,z)]
+			contraste+=[-(f2(f1(0,ratio_t1),gamma_las)-f2(f1(10,ratio_t1),gamma_las))/f2(f1(10,ratio_t1),gamma_las)]
+		ax[0].plot(gamma_range/ratio_t1,delta_width,label='ratio=%f'%ratio_t1)
+		ax[1].plot(gamma_range/ratio_t1,contraste,label='ratio=%f'%ratio_t1)
+		ax[0].legend()
+		ax[1].legend()
+		ax[1].set_xlabel('Gamma_las')
+		ax[0].set_ylabel('Width(PL)-Width(R1)')
+		ax[1].set_ylabel('Contrast')
+
+widths()
+plt.show()

### 2020-07-10 python/python_clément/compteur de photon 2.py

@@ -1,0 +1,101 @@
+import sys
+import time
+import nidaqmx
+import nidaqmx.stream_writers
+import nidaqmx.task
+import nidaqmx.system
+
+import numpy as np
+
+from PyQt5.QtCore import (Qt, QTimer)
+from PyQt5.QtWidgets import (QWidget, QPushButton, 
+    QHBoxLayout, QVBoxLayout, QApplication, QDesktopWidget, QMainWindow)
+#from matplotlib.backends.qt_compat import QtCore, QtWidgets, is_pyqt5
+from matplotlib.backends.backend_qt5agg import (FigureCanvas, NavigationToolbar2QT as NavigationToolbar)
+from matplotlib.figure import Figure
+
+
+
+
+class ApplicationWindow(QMainWindow):
+    def __init__(self):
+        super().__init__()
+        
+        self._main = QWidget()
+        self.setCentralWidget(self._main)
+        Vbox = QVBoxLayout()
+        x=QDesktopWidget.width(self._main)*0.7
+        y=QDesktopWidget.height(self._main)*0.7
+        #self.setFixedSize(x,y)
+        layout= QHBoxLayout()
+        layout.addLayout(Vbox)
+        stop=QPushButton('Stop')
+        start=QPushButton('Start')
+        Vbox2=QVBoxLayout()
+        Vbox2.addWidget(start)
+        Vbox2.addWidget(stop)
+        
+        layout.addLayout(Vbox2)
+        self._main.setLayout(layout)
+        #static_canvas = FigureCanvas(Figure(figsize=(5, 3)))
+        #layout.addWidget(static_canvas)
+        #self.addToolBar(NavigationToolbar(static_canvas, self))
+
+        dynamic_canvas = FigureCanvas(Figure(figsize=(10, 5)))
+        Vbox.addStretch(1)
+        Vbox.addWidget(dynamic_canvas)
+        self.addToolBar(Qt.BottomToolBarArea,
+                        NavigationToolbar(dynamic_canvas, self))
+
+        #self._static_ax = static_canvas.figure.subplots()
+        self.t = np.linspace(0, 10, 501)
+        #self._static_ax.plot(self.t, np.tan(self.t), ".")
+
+        self._dynamic_ax = dynamic_canvas.figure.subplots()
+        self.y=np.zeros(len(self.t))
+        self.line2,=self._dynamic_ax.plot(self.t, self.y)
+        
+        self.dt=0.03
[…]

### 2020-07-10 python/python_clément/print_map.py

@@ -1,0 +1,24 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+with open("map_2.txt",'r') as f:
+	xy=[]
+	line=f.readline()
+	while line != "" :
+		line=line.split()
+		phi=[]
+		for v in line :
+			phi+=[float(v)]
+		xy+=[phi]
+		line=f.readline()
+
+
+thetaphi=np.array(xy)
+
+thetaAxis=np.linspace(50,80,30)
+phiAxis=np.linspace(200,170,32)
+
+plt.plot(thetaphi[:,0])
+#plt.pcolor(phiAxis, thetaAxis, thetaphi)
+plt.show()

### 2020-07-10 python_clément/qudi/qudi-master/hardware/national_instruments_x_series.py

@@ -1,0 +1,2215 @@
+# -*- coding: utf-8 -*-
+
+"""
+This file contains the Qudi Hardware module NICard class.
+
+Qudi is free software: you can redistribute it and/or modify
+it under the terms of the GNU General Public License as published by
+the Free Software Foundation, either version 3 of the License, or
+(at your option) any later version.
+
+Qudi is distributed in the hope that it will be useful,
+but WITHOUT ANY WARRANTY; without even the implied warranty of
+MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License
+along with Qudi. If not, see <http://www.gnu.org/licenses/>.
+
+Copyright (c) the Qudi Developers. See the COPYRIGHT.txt file at the
+top-level directory of this distribution and at <https://github.com/Ulm-IQO/qudi/>
+"""
+
+import numpy as np
+import re
+
+import PyDAQmx as daq
+
+from core.module import Base
+from core.configoption import ConfigOption
+from interface.slow_counter_interface import SlowCounterInterface
+from interface.slow_counter_interface import SlowCounterConstraints
+from interface.slow_counter_interface import CountingMode
+from interface.odmr_counter_interface import ODMRCounterInterface
+from interface.confocal_scanner_interface import ConfocalScannerInterface
+
+
+class NationalInstrumentsXSeries(Base, SlowCounterInterface, ConfocalScannerInterface, ODMRCounterInterface):
+    """ A National Instruments device that can count and control microvave generators.
+
+    !!!!!! NI USB 63XX, NI PCIe 63XX and NI PXIe 63XX DEVICES ONLY !!!!!!
+
+    See [National Instruments X Series Documentation](@ref nidaq-x-series) for details.
+
+    stable: Kay Jahnke, Alexander Stark
+
+    Example config for copy-paste:
+
+    nicard_6343:
+        module.Class: 'national_instruments_x_series.NationalInstrumentsXSeries'
+        photon_sources:
+            - '/Dev1/PFI8'
+        #    - '/Dev1/PFI9'
+        clock_channel: '/Dev1/Ctr0'
+        default_clock_frequency: 100 # optional, in Hz
+        counter_channels:
+            - '/Dev1/Ctr1'
+        counter_ai_channels:
+            - '/Dev1/AI0'
[…]

### 2020-07-10 python_clément/scan_electro_aimant_V2.py

@@ -1,0 +1,334 @@
+import sys
+import os
+import time
+import random
+import nidaqmx
+import nidaqmx.stream_writers
+import nidaqmx.stream_readers
+import nidaqmx.task
+import nidaqmx.system
+import nidaqmx.constants
+
+
+import visa
+import numpy as np
+
+
+from PyQt5.QtGui import QFont
+from PyQt5.QtCore import (Qt, QTimer)
+from PyQt5.QtWidgets import (QWidget, QPushButton, 
+    QHBoxLayout, QVBoxLayout, QApplication, QDesktopWidget, QMainWindow, QLineEdit, QLabel, QCheckBox, QFileDialog, QComboBox)
+
+from matplotlib.backends.backend_qt5agg import (FigureCanvas, NavigationToolbar2QT as NavigationToolbar)
+from matplotlib.figure import Figure
+
+
+
+class Photon_Counter(QMainWindow):
+    def __init__(self):
+        super().__init__()
+        
+        #Timing Parameter ##
+        #The program aquires the total number of photons at a rate defined by real_sampling_rate, but will only display an average of it every dt
+
+        self.N_scan=201 #nb of points per scan
+        self.Vmin=0 #V
+        self.Vmax=1 #V
+
+        self.Imax=1 #A
+
+        self.time_scan=11 #s
+        self.time_rampe=10 #s, doit être un multiple de 1s pour une raison mysterieuse
+
+        self.sampling_rate=self.N_scan/self.time_scan
+
+
+
+        
+
+        ##Creation of the graphical interface##
+
+        self.setWindowTitle("Line Scan")
+
+        self.main = QWidget()
+        self.setCentralWidget(self.main)
+
+        layout= QHBoxLayout()
+        Vbox_gauche=QVBoxLayout()
+        Vbox = QVBoxLayout()
+        Vbox_droite=QVBoxLayout()
+
+
+        layout.addLayout(Vbox_gauche)
+        layout.addLayout(Vbox)
+        layout.addLayout(Vbox_droite)
+        self.main.setLayout(layout)
+
+
+        #Fields on the left
+        self.labelVmin=QLabel("V min (V)")
+        self.labelVmax=QLabel("V max (V)")
+        self.labelNscan=QLabel("Number of points")
+
+        self.lectVmin=QLineEdit(str(self.Vmin))
+        self.lectVmin.setEnabled(False)
+        self.lectVmax=QLineEdit(str(self.Vmax))
+        self.lectNscan=QLineEdit(str(self.N_scan))
+
[…]

### 2020-07-10 python_clément/switch_AOM.py

@@ -1,0 +1,62 @@
+import sys
+import nidaqmx
+import nidaqmx.task
+from PyQt5.QtGui import QFont
+from PyQt5.QtCore import (Qt, QTimer)
+from PyQt5.QtWidgets import (QWidget, QPushButton, 
+	QHBoxLayout, QVBoxLayout, QApplication, QDesktopWidget, QMainWindow, QLineEdit, QLabel, QCheckBox, QFileDialog, QComboBox)
+
+class Photon_Counter(QMainWindow):
+	def __init__(self):
+		super().__init__()
+		layout= QHBoxLayout()
+		self.setWindowTitle("Switch AOM")
+		self.main = QWidget()
+		self.setCentralWidget(self.main)
+		Vbox = QVBoxLayout()
+		Vbox_right = QVBoxLayout()
+		layout.addLayout(Vbox)
+		layout.addLayout(Vbox_right)
+		self.main.setLayout(layout)
+
+		self.stop=QPushButton('AOM closed')
+		self.start=QPushButton('AOM opened')
+		Vbox.addWidget(self.start)
+		Vbox.addWidget(self.stop)
+
+		self.stop2=QPushButton('uW closed')
+		self.start2=QPushButton('uW opened')
+		Vbox_right.addWidget(self.start2)
+		Vbox_right.addWidget(self.stop2)
+
+		self.start.clicked.connect(self.open_AOM)
+		self.stop.clicked.connect(self.close_AOM)
+		self.start2.clicked.connect(self.open_uW)
+		self.stop2.clicked.connect(self.close_uW)
+
+
+	def open_AOM(self):
+		with nidaqmx.Task() as write :
+			write.do_channels.add_do_chan('Dev1/port0/line2')
+			write.write([True])
+			
+	def close_AOM(self):
+		with nidaqmx.Task() as write :
+			write.do_channels.add_do_chan('Dev1/port0/line2')
+			write.write([False])
+
+	def open_uW(self):
+		with nidaqmx.Task() as write :
+			write.do_channels.add_do_chan('Dev1/port0/line3')
+			write.write([False])
+			
+	def close_uW(self):
+		with nidaqmx.Task() as write :
+			write.do_channels.add_do_chan('Dev1/port0/line3')
+			write.write([True])
+
+qapp = QApplication(sys.argv)
+app = Photon_Counter()
+app.show()
+qapp.exec_()
+		

### 2020-07-10 python_clément/tdv_soustraction.py

@@ -1,0 +1,413 @@
+import sys
+import os
+import time
+import random
+import nidaqmx
+import nidaqmx.stream_writers
+import nidaqmx.stream_readers
+import nidaqmx.task
+import nidaqmx.system
+import nidaqmx.constants
+
+
+import visa
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+from PyQt5.QtGui import QFont
+from PyQt5.QtCore import (Qt, QTimer)
+from PyQt5.QtWidgets import (QWidget, QPushButton, 
+	QHBoxLayout, QVBoxLayout, QApplication, QDesktopWidget, QMainWindow, QLineEdit, QLabel, QCheckBox, QFileDialog, QComboBox)
+
+from matplotlib.backends.backend_qt5agg import (FigureCanvas, NavigationToolbar2QT as NavigationToolbar)
+from matplotlib.figure import Figure
+
+
+
+class Photon_Counter(QMainWindow):
+	def __init__(self):
+		super().__init__()
+		
+
+
+		self.n_points=100
+		self.f_uW=2.88   
+		self.level=0 #dBm
+		self.time_acq=1E-3 #in s
+
+		self.time_polarization=2E-4 #in s
+		self.time_read=1E-4 #in s
+		self.time_pulse_uW=1E-6 #in s. The pulse time might not pe precise
+
+		self.refresh_rate=0.1 #Screen max refresh rate (in s), below 0.1 is probably too fast
+
+
+		self.sampling_rate=self.n_points/self.time_acq
+
+		self.sampling_rate_switch_uW=((1/self.time_pulse_uW)//self.sampling_rate)*self.sampling_rate
+
+		self.ratio_sample_clocks=int(self.sampling_rate_switch_uW//self.sampling_rate)
+
+		self.n_pola=int(self.sampling_rate*self.time_polarization)
+		self.n_read=int(self.sampling_rate*self.time_read)
+
+		
+
+
+		
+
+		##Creation of the graphical interface##
+
+		self.setWindowTitle("TdV soustraction")
+
+		self.main = QWidget()
+		self.setCentralWidget(self.main)
+
+		layout= QHBoxLayout()
+		Vbox = QVBoxLayout()
+		Vbox_droite=QVBoxLayout()
+		Vbox_gauche=QVBoxLayout()
+
+		layout.addLayout(Vbox_gauche)
+		layout.addLayout(Vbox)
[…]

### 2020-08-19 Mes papiers/Double quantum/PAPER/dble_quantum.tex
ordibureau2020-08-19

\documentclass[preprintnumbers,amsmath,amssymb,twocolumn]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}

\def\sgn{\mathop{\rm sgn}}

\begin{document}

\title{Optical detection of hydrogen paramagnetic defects in diamonds}

\author{C. Pellet-Mary$^{1}$}
\author{P. Huillery $^{1}$}
\author{A. Tallaire $^{1}$...}
\author{M. Perdriat $^{1}$}
\author{G. H\'etet$^{1}$}

\affiliation{$^1$Laboratoire de Physique de l'Ecole normale sup\'erieure, ENS, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e Paris-Diderot, Sorbonne Paris Cit\'e, Paris, France.}

\begin{abstract}
We observe cross-relaxation amongst closely packed negatively charged nitrogen-vacancy centers (NV$^-$) that involve dipolar interaction between NV centers and paramagnetic defects. 
Specifically, we observe dipolar coupling between pairs of NV$^-$ centers and between single  NV$^-$ centers and NV$^- -^{13}$C pairs as well as novel optical cross-relaxation with hydrogen-vacancy complexes.  These processes are detected using magnetic fields scans along the $\langle 100 \rangle$ cristalline direction of diamond.
Contrary the more commonly chosen $\langle 111 \rangle$ direction, $\langle 100 \rangle$ scans select dipolar couplings involving the exchange of two quanta of spin-projection.
The technique can be used to detect the degree of polarisation of $^{13}$C atoms at room temperature.
 %Scanning a magnetic field from zero to a few hundred Gauss along the $\langle 100 \rangle$ cristalline direction of diamond, we identify double-quantum cross-relaxation processes amongst NV centers. 
%In particular, the technique is shown to enable efficient micro-wave free detection of the $^{13}$C atoms on the first shell around the NV centers.centers. 
\end{abstract}
\maketitle

The precise control of the spin of negatively charged nitrogen-vacancy (NV$^-$) centers in diamond has given rise to a wealth of applications in nanoscale sensing, quantum information science etc.. 
One major reason for the interest in NV centers is that they can be optically polarized at room temperature.  
This property can be employed to manipulate and to detect other dark spins inside or outside the diamond.

Cross relaxation.
Most of the measurements are done at the ESLAC/GLSAC. OORT, BUDKER, HOLLENBERG, BAJAJ, MANSON ?

Polsarisation of 13C at low fields : PINES, MERILES. 
Theory GALI.
Detection at low fields was also demonstrated \cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}. It was suggested that spin relaxation beyond the phonon-relaxation timescales comes from electron tunneling between nearby impurities. 
Dipolar coupling amongst closely packed NV was used further to demonstrate many body effects such as critcal thermalisation, or Dicke time crystals or beyond Heisenberg limited sensing (Lukin).

To the best of our knowledge, the process was however not employed to detect other impurities.

D. SUTER.
However, 13C was not detected. The main reason is the predominance of single-quantum processes.
13C detected using B field scans ? nop. 
%Recently, this high degree of control was employed to study strongly interacting, dense NV ensembles to explore quantum many-body dynamics.
Here, we characterize the cross-relaxation of high-density NV ensembles with a focus on double quantum resonances.

\begin{figure}[!ht]
  \centering \scalebox{0.5}{\includegraphics{Cristallo.eps}}
  \caption{a) Schematics showing the tetrahedron in one of the interstitial sites of the diamond structure. The NV centers can only be in of the four depicted orientations. 
   }\label{Cristallo}
\end{figure}

\begin{figure}[!ht]
  \centering {\includegraphics[width=0.4\textwidth]{map_full.eps}}
  \caption{a) 
Graph (theory) showing the locus of the degeneracies of the 4 NV centers electronic spins resonances in the $(\theta,\phi)$ parameter space. See text for details.  
 }
\end{figure}

\begin{figure}[!ht]
  \centering{\includegraphics[width=0.4\textwidth]{map_zoom.eps}}
  \caption{a) 
Zoom on the $\langle 100 \rangle$ direction. 
 }
\end{figure}

Fig. \ref{Map} shows the NV Photoluminescence as a function of magnetic field direction, referenced by azimuthal and polar angles $(\phi, \theta)$. The PL is seen to drop for particular values of the angle. The symmetric pattern suggests that the diamond cristalline main directions play a role. Because of the anisotropy of the NV center, see \ref{Cristallo}, the NV PL drop can indeed be a signature of the cristal direction.
The drop of the PL along the cristal axis was already observed by many groups. LUKIN, HEMMER, BUDKER. JARMOLA. OORT
\cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}

Charge transfer processes :

IONISATION
2 photon Photoionisation from NV- to NV0.
OR 
Tunneling (Manson/Degen) : NV- excited state ionizes via tunneling of electrons to Ns+.

RECHARGING  : 
NV0+N0-> NV-+N+ (in the dark). 
It can occur via tunneling of electrons among closely packed NVs.

Each of the lines is very well approximated by a Gaussian curve.

\begin{figure}[!ht]
  \centering{\includegraphics[width=0.40\textwidth]{Carte.eps}}
  \caption{a) 
NV Photoluminescence as a function of magnetic field direction, referenced by azimuthal and polar angles $(\phi, \theta)$.  }\label{Map}
\end{figure}

Fig.~\ref{T1} shows measurements of the $T_1$/ Tunneling is strong in the presence of nitrogen impurities. 
To remove this effect, we apply a microwave....

\begin{figure}[!ht]
  \centering{\includegraphics[width=0.40\textwidth]{T_1.eps}}
  \caption{a) 
Lifetime saturation theory
 }\label{T1}
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.5}{\includegraphics{R1_vs_J.pdf}}
  \caption{a) 
Lifetime saturation theory
 }
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.4}{\includegraphics{13C_transitions.eps}}
  \caption{Transitions of NV-13C and NV.
 }
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.5}{\includegraphics{rose_10V_13C.eps}}
  \caption{NV Photoluminescence as a function of a magnetic field scanned along the 100 direction. 
 }
\end{figure}

Two peaks cannot be explained.

Hypotheses that did not work.

$NV^0, P1, 13C$ other shells, NVN (H3) center (diamagnetic in ground state), paramagnetic only in excited state but requires UV light. 
No other species were observed in the optical spectrum or EPR.

\section*{acknowledgements}
We would like to thank Neil Manson and ... for fruitful discussions. 
GH acknowledges funding by the French National Research Agency (ANR) through the T-ERC project QUOVADIS.

\bibliography{Biblio_dbl_quantum}

\end{document}

### 2020-08-19 manips/PL en fonction de l'angle de B/manip.tex
ordibureau2020-08-19

@@
   Les principales mesures grace à l'électro-aimant
+
+  \section{Double quantums et VH-}
+  Manips réalisées de mi Mai à Aout (et peut être un peu après, faut voir).
+
+  J'ai déja écrit une partie dans une autre fiche qui est dans "double quantums", je vais décrire le setup final.
+  \subsection{Matos}
+
+  En gros le meilleur setup c'était, toujours avec l'EM de Paul et ses 5000 enroulages, de fixer l'EM avec de la patafix sur une monture de lambda/2 et de monter cette lambda/2 sur une platine tournante, et de brancher cet EM directement sur le GBF (!).
+
+  Setup du GBF (un HP, on doit pas en avoir des milliers) : Je me suis mis en mode triangle, à une fréquence de 1HZ (au dela le signal se distort à cause de l'inductance énorme) et je suis passé en mode burst (bouton burst...) en envoyant des pulses avec le programme d'acquisition pour générer une rampe toute les 1.1s (pour être safe, le GBF ne réagit pas si t'envoie une pulse alors qu'il a pas fini son burst).
+Le piège c'est d'appuyer sur "Single", sinon ce con de GBF est en mode auto et renvoie un burst des qu'il a fini le précédent.
+Tu peux aussi modifier la phase du burst dans les options, je te laisse chercher dans le manuel parce que c'est chiant (menu, flèche du bas, flèche de droite *5 de mémoire)
+
+  Le programme c'est scan\_em\_gbf\_2.py, et normalement il ne regarde que la partie du scan qui correspond à une rampe.
+
+  Aussi faut que je signale qu'on a vu des grosses asymétries dans les scans, et qu'on sait toujours pas bien ce que c'est :
+ \begin{itemize}
+ \item C'est pas un processus dynamique parce que ça ne dépend pas de la vitesse de scan
+ \item C'est pas du à l'hysteresis de l'aimant parce que les asymétries ne changent pas de sens quand on scan dans l'autre sens
+ \item c'est peut-etre du aux gradient de champs parce que l'effet a l'air moins fort au milieu de l'EM
+ \end{itemize}
+ Du coup j'ai fini par conclure que c'était des gradients de champ qui sont pas les même en fonction du signe du champ.
+Je ne vois pas comment c'est possible mais ça me suffira pour l'instant.
+Et la vraie solution c'est de couper les scans pour en voir qu'une moitié.
+Merci Gabriel pour les cours d'intégrité scientifique.
+
+ \subsection{Manips}
+ J'ai analysé 4 échantillons avec ce setup : le diamant rose d'Alex (3.5 PPM de NV), le sumitumo \#4, HPHT irradié par électrons (à vue de nez 5 PPM de NV, sans doute un peu plus), un Adamas sur une antenne uW (5 PPM de NV aussi), et un diamant de Ludo HPHT, irradié aux protons avec 20 PPM de NV.
+
+ J'ai essentiellement fait des scans selon la (100), avec aussi pour certains des temps de vie (le Adamas en particulier) et des scans (111).
+
+ TdV : le 06/07/2020
+
+ scans 100 : de fin Juin à fin Juillet
+
+ \subsection{Resultats}
+ cf article, je completerai peut-etre ce qu'on a pas mis dedans.
   \end{document}

### 2020-08-19 manips/PL en fonction de l'angle de B/degenerescences analytiques/plot.py

@@ -83,12 +83,16 @@ def map_full() :
 
 
 
-	plt.xlabel(r'$\phi$(°)')
-	plt.ylabel(r'$\theta$(°)')
+	plt.xlabel(r'$\phi$(°)',fontsize=25)
+	plt.ylabel(r'$\theta$(°)',fontsize=25)
+	plt.xticks(fontsize=15)
+	plt.yticks(fontsize=15)
 
 
-	plt.legend()
+	# plt.legend()
 	plt.show()
+
+
 
 def map_zoom():
 	fig,ax=plt.subplots()
@@ -145,13 +149,15 @@ def map_zoom():
 	for line in ax.get_lines() :
 		line.set_data(line.get_xdata()-180,line.get_ydata()-90)
 
-	plt.xlabel(r'$\phi$(°)')
-	plt.ylabel(r'$\theta$(°)')
+	plt.xlabel(r'$\phi$(°)',fontsize=25)
+	plt.ylabel(r'$\theta$(°)',fontsize=25)
+	plt.xticks(fontsize=15)
+	plt.yticks(fontsize=15)
 	plt.xlim([-9,9])
 	plt.ylim([-6,6])
 
 
-	plt.legend()
+	plt.legend(loc='upper right', fontsize=15)
 	plt.show()
 
 map_zoom()

### 2020-08-19 python/simu Hamiltonien NV/template.py

@@ -269,6 +269,7 @@ def NV_simple(orientation='111'):
 
 
 
+
 def egv_C13(): #Les 4 croisements sont à : 17.96 G, 19.70 G, 22.11 G, 24.31 G 
 	amps=np.linspace(0,100,100)
 	transi=[]
@@ -938,7 +939,7 @@ def spin_NV_propre_base(B,gamma_las=1E-3,gamma_phonon=2E-4): #Attention : gamma 
 	dm=steadystate(H,make_collapse_list(gamma_las,gamma_phonon))
 	rho=np.array(dm)
 	s=np.array([np.trace(rho.dot(Sx)),np.trace(rho.dot(Sy)),np.trace(rho.dot(Sz))])
-	return(s)
+	return(s,rho[1,1])
 
 
 
@@ -983,9 +984,9 @@ def torque_3nvx_1classe(amp):
 	# plt.show()
 	plt.savefig('torque une classe/simu torque 3 nvx 1 classe %iG.png'%amp)
 
-for amp in np.arange(50,2050,50) :
-	torque_3nvx_1classe(amp)
-	print(amp)
+# for amp in np.arange(50,2050,50) :
+# 	torque_3nvx_1classe(amp)
+# 	print(amp)
 
 def torque_1classe_amplitude(theta):
 	amps=np.linspace(0,5000,200)
@@ -1029,11 +1030,13 @@ def spin_3vx_4classes(B):
 	Rretour=[Rzmoins.dot(Ryplus),Rzmoins.dot(Rymoins),Rzmoins.dot(Rxmoins),Rzmoins.dot(Rxplus)]
 
 	spin=np.zeros(3,dtype='complex128')
+	rho_0=0
 	for k in range(4) :
 		B_base=Raller[k].dot(B)
-		spin_base=spin_NV_propre_base(B_base)
+		spin_base,rho_0_base=spin_NV_propre_base(B_base)
+		rho_0+=rho_0_base
 		spin+=Rretour[k].dot(spin_base)
-	return(spin.real)	
+	return(spin.real,rho_0)	
 
 def cart_to_spher(r):
 	x=r[0]
@@ -1111,21 +1114,26 @@ def NRJ_map(amp):
 	thetas=np.linspace(pi*0.1,pi*0.9,nthetas)
 	phis=np.linspace(-pi*0.9,pi*0.9,nphis)
 	NRJs=np.zeros((nthetas,nphis))
[…]

### 2020-09-02 manips/PL en fonction de l'angle de B/manip.tex
bureau

@@
  \subsection{Manips}
- J'ai analysé 4 échantillons avec ce setup : le diamant rose d'Alex (3.5 PPM de NV), le sumitumo \#4, HPHT irradié par électrons (à vue de nez 5 PPM de NV, sans doute un peu plus), un Adamas sur une antenne uW (5 PPM de NV aussi), et un diamant de Ludo HPHT, irradié aux protons avec 20 PPM de NV.
+ J'ai analysé 4 échantillons avec ce setup : le diamant rose d'Alex (3.5 PPM de NV), le sumitumo \#4, HPHT irradié par électrons (à vue de nez 5 PPM de NV, sans doute un peu plus), un Adamas sur une antenne uW (5 PPM de NV aussi), et un diamant de Ludo HPHT, irradié aux protons avec 20 PPM de NV (Update : il a été irradié à l'Hélium, du coup c'est chelou).
 

### 2020-09-02 manips/PL en fonction de l'angle de B/VH/traitement.py

@@ -128,15 +128,6 @@ def ask_name():
 
 x_transi=[17.953273372377286, 19.705551221857355, 24.31251459360322, 22.117301530072833]
 
-# fname='scan_rose_100_7V_'
-# x,y=extract_data(fname+'.txt')
-# y=y/max(y)
-# xmin=258
-# x=x-x[xmin]
-# x=x[:xmin]
-# y=y[:xmin]
-# x=x*64
-# plt.plot(x,y,label='Photoluminescence')
 
 fname='scan_100_rose_10V'
 x,y=extract_data(fname+'.txt')
@@ -146,47 +137,38 @@ x=x-x[xmin]
 x=x[xmin:]
 y=y[xmin:]
 x=x*62
-plt.plot(-x,y,label='Photoluminescence')
+plt.plot(-x,y,lw=2,label='Photoluminescence')
 
-
-# fname='scan_100_sample_ludo_aligne_vitedeuf'
-# x,y=extract_data(fname+'.txt')
-# y=y/max(y)
-# x=x*62
-# xmin=251
-# x=x-x[xmin]
-# x=x[:xmin]
-# y=y[:xmin]
-# plt.plot(x,y,label='Photoluminescence')
-
-# fname='scan_10V_uW_2796'
-# x,y=extract_data(fname+'.txt')
-# y=y/max(y)
-# x=x-x[253]
-# x=x*62
-# plt.plot(-x,y,label='Photoluminescence')
 
 
 ax=plt.gca()
 ylim=ax.get_ylim()
+ymin=ylim[0]
+ymax=ylim[1]
+
 color = next(ax._get_lines.prop_cycler)['color']
-# x=x_transi[0]
-# plt.plot([x,x],[0,2],'--',color=color,label='P1 cross relaxation')
-# # plt.plot([-x,-x],[0,2],'--',color=color)
-# for x in x_transi[1:] :
-# 	plt.plot([x,x],[0,2],'--',color=color)
-# 	# plt.plot([-x,-x],[0,2],'--',color=color)
+x=0
+plt.plot([x,x],[ymin,ymax],'--',color=color,label='Double quantums')
+
 # color = next(ax._get_lines.prop_cycler)['color']
-x=50.5276
-plt.plot([x,x],[0,2],'--',color=color,label='VH cross relaxation')
-color = next(ax._get_lines.prop_cycler)['color']
-x=x_transi[0]
[…]

### 2020-09-02 python_clément/write_pb.py

@@ -1,0 +1,40 @@
+import numpy as np
+
+#unit = s,ms,us,ns
+
+
+def rabi(t_scan,n_scan,t_lect,t_ecl):
+	with open('PB_instructions.txt','w') as f:
+		for s in t_scan.split() :
+			try :
+				val=float(s)
+			except :
+				unit=s
+		taux=np.linspace(0,val,n_scan)
+		for i in range(1,len(taux)) :
+			tau=str(taux[i])+unit
+			if i == 1:
+				f.write('label : ')
+			else : 
+				f.write('\t')
+
+			f.write('0b 0100, '+t_ecl+'\n')
+			f.write('\t0b 1000, '+tau+'\n')
+			f.write('\t0b 0110, '+t_lect) #Est-ce qu'on peut compter sur les deux fronts ?
+			if i==len(taux)-1:
+				f.write(', branch, label')
+			else :
+				f.write('\n')
+
+
+
+# rabi('1000 us',101,'500 us','500 us')
+
+
+def AOM_on():
+	with open('PB_instructions.txt','w') as f:
+		f.write('label: 0b 0100, 100 ms, branch, label')
+
+AOM_on()
+
+

### 2020-09-11 matos/Cryo + microscope/fiche.tex
malaubide

--- 2020-07-10 experience/fiche matos/fiche.tex
@@
   \item Milieu : axe z, haut=avant , bas=arrière
-  \item Droite : axe x, ???
+  \item Droite : axe x, bas=droite, haut=gauche
   \end{itemize}
@@
 
+ \section{microscope}
+ Le microscope est un peu gitan, j'ai pas réussi à trouver d'objectif qui se visse mais tu peut simplement en poser un.
+Ensuite pour regarder un échantillon, le mieux c'est de le poser sur de la patafix que tu as mis sur une tige.
+
+ Faut aussi bien orienter la lampe (la tu regardes en réflexion donc faut la faire passer par en dessous). C'est con mais ça marche.
+
+ Tu as les molettes à droite de l'oculaire qui te permettent de translater rapidement en x/y, mais j'ai pas encore trouvé comment translater vite en z.
   \end{document}

### 2020-09-11 data/20200831/traitement_brut.py

@@ -1,7 +1,8 @@
 import sys
 import numpy as np
 import matplotlib.pyplot as plt
-from scipy.optimize import curve_fit
+import glob
+from scipy.optimize import curve_fit,root_scalar
 from PyQt5.QtWidgets import QApplication,QFileDialog
 
 
@@ -21,9 +22,18 @@ def extract_data(filename,xcol=0,ycol=1):
 	return(np.array(x),np.array(y))
 
 def lin_fit(x,y) :
+	x=np.array(x)
+	y=np.array(y)
 	A=np.vstack([x,np.ones(len(x))]).T
 	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
 	return(a,b,a*x+b)
+
+def quad_fit(x,y) :
+	x=np.array(x)
+	y=np.array(y)
+	A=np.vstack([x**2,x,np.ones(len(x))]).T
+	q,a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([q,a,b],q*x**2+a*x+b)
 
 def gauss_fit(x,y,Amp=None,x0=None,sigma=None,ss=0) :
 	if not ss :
@@ -34,7 +44,10 @@ def gauss_fit(x,y,Amp=None,x0=None,sigma=None,ss=0) :
 		else :
 			Amp=min(y)-ss
 	if not x0 :
-		x0=x[int(len(x)/2)]
+		if max(y)-ss > ss-min(y) :
+			x0=x[np.argmax(y)]
+		else :
+			x0=x[np.argmin(y)]
 	if not sigma :
 		sigma=x[int(len(x)/5)]-x[0]
 	def f(x,Amp,x0,sigma,ss) :
@@ -125,53 +138,64 @@ def ask_name():
 	fname,filters=QFileDialog.getOpenFileName()	
 	return fname
 
-
-
-
-
-fname='scan_100_rose_10V'
-x,y=extract_data(fname+'.txt')
+fname='blanc.txt'
+x,y=extract_data(fname)
+bmin=243
+bmax=707
+x=x[bmin:bmax]
+y=y[bmin:bmax]
+y=y-min(y)
 y=y/max(y)
-xmin=507
-x=x-x[xmin]
-x=x[xmin:]
-y=y[xmin:]
-x=x*66
-plt.plot(-x,y,lw=2,label='Photoluminescence')
+plt.plot(x,y)
+
[…]

### 2020-09-21 matos/photodiode/photodiode.tex

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Photodiodes}
  \section{Principe}
  Bon le principe d'une photodiode je l'ai su mais j'ai oublié. En gros ça transforme des photons en électrons mobiles et donc ça fait du courant. Dans le visible on a essentiellement interet à utiliser du Si, ça va jusqu'à 1100 nm je crois.

  Deux trucs importants à retenir :
  \begin{itemize}
  \item De base ça produit du courant une photodiode (si t'es polarisé en inverse, blabla. On va pas partir d'un bloc de silicium non plus), donc y'a un moment ou il faut le convertir en tension. T'as deux modèles de photodiodes qu'on a : avec et sans gain. Sans gain (celle sur pile), t'as directement le courant, donc en fonction de la résistance que tu branche dessus, t'auras plus ou moins de contraste : si tu la branche direct sur l'oscillo ($R\approx M\Omega$) t'auras un super contraste mais un temps de réponse moisi (cf après). Avec gain (celle sur secteur), tu dois avoir un petit circuit avec AO qui fait qu'elle délivre simplement une tension, toujours la même et toujours avec le même temps de réponse (en fonction du gain bien sur).
  \item Le temps de réponse : ça dépend de ta capacité de jonction (taille du capteur) et de la résistance de charge ($\tau=RC$). Avec un temps de réponse lent c'est chiant pour tes mesures, mais ça peut être un avantage pour le bruit (c'est compliqué le bruit, mais plus grosse bande passante = plus de bruit)
  \end{itemize}
  \section{Performances}
  Les perfs décrites sur le site thorlabs me semblent assez justes, mais je vais noter vite fait ce que j'ai vu.

  \subsection{La photodiode sur pile}
  D'après Gabriel, c'est mieux d'être sur pile pour éviter le 50 Hz. Moi de ce que j'ai vu elle est pas ouf, mais j'ai pas vraiment accès au bruit de la photodiode directement.
  La puissance optique que j'avais à chaque fois était de genre $P=700 nW$, puissance de PL raisonnable sur des adamas avec l'objectif vert (x40, NA=0.4).

  Branché direct à l'oscillo, j'avais un contraste de l'ordre de 0.5V, mais un temps de réponse à la ms.

  Branché sur la carte NI j'avais rien du tout (saturation, elle doit avoir une impédance d'entrée trop grande)

  Branché avec un bouchon de $20 k\Omega$ sur un T(fait maison STP, et j'ai galéré. Il est rangé dans le tiroir en haut à gauche du foutoir électronique), j'avais un temps de réponse de $50 \mu s$ mais un contraste de 10 mV.

 \subsection{La photodiode sur secteur}
 En gros il faut l'utiliser sur les gains 60 ou 70 dB pour de la PL, mais les chiffres sont pas dégueux : contratse de 0.5V et 1.5V, et temps de réponse de $20 \mu s$ et $50 \mu s$. Même à $20 \mu s$ j'ai pas vu de différence entre la monté et la descente de la PL (la montée devrait être plus lente à cause de la polarisation des spins), mais c'est parce que le laser vert est à fond pour avoir du signal (à peu près le même ordre de grandeur $t_{pol}$ mesuré à l'APD).
 Du coup c'est peut être jouable. Par contre j'ai vu un bruit chelou à 130 kHz.

 Par ailleurs j'y avais pas pensé, mais je sais pas du tout comment dire à Labview de moyenner la tension entre deux points. En faite je sais pas sur combien de temps il lit une tension.

  \end{document}

### 2020-10-01 Mes papiers/Méca CR/Meca_cross.tex
covid

\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-1}
%\documentclass[preprintnumbers,amsmath,amssymb]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}

%\newcommand{|}{Y$_2$SiO$_5$}
\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{Mechanical Detection of Dipole-Dipole Interactions \\between Electronic Spins in a Solid}

\author{C. Pellet-Mary, P. Huillery, M. Perdriat, G. H\'etet}

\affiliation{$^1$Laboratoire Pierre Aigrain, Ecole normale sup\'erieure, PSL Research University, CNRS, Universit\'e Pierre et Marie Curie, Sorbonne Universit\'es, Universit\'e Paris Diderot, Sorbonne Paris-Cit\'e, 24 rue Lhomond, 75231 Paris Cedex 05, France.}

\begin{abstract}
Magnetic-dipole interactions between the spins of defects in crystals play a key role in a vast range of applications. In typical experiments on nuclear or electronic spin resonance, the interactions are identified in the electromagnetic emission or absorption spectrum of the spins.
Here, we demonstrate mechanical detection of dipolar interactions. 
Using a levitating diamond containing nitrogen-vacancy (NV) centers, we employ cross-relaxation to alter 
the spin-torque when pairs of NV centers' orientations become resonant.
Our approach opens a path towards the use of mechanical oscillators to detect paramagnetic defects that lack optical transitions. 
\end{abstract}

\maketitle

Defects in solid state materials with long-lived electronic or nuclear spins play a key role in the fields of quantum computing, communication and sensing. 
Such defects can be hosted by various materials, the most studied examples being nuclear impurities in quantum dots \cite{Qdots}, dopants in silicon \cite{Zwanenburg}, or negatively charged nitrogen vacancy (NV$^-$) centers in diamond \cite{Doherty}.
The detection of their spins is generically performed {\it via} emission or absorption of electromagnetic radiation, both in the radio/micro-wave or optical domain, or purely electrically \cite{Zwanenburg, Hopper}.
Electron-Paramagnetic Resonance (EPR) and Optically Detected Magnetic Resonance (ODMR) are two such detection methods, both routinely operating under ambient conditions and with large signal to noise ratios.
EPR can detect most paramagnetic species with great accuracy, while the latter reaches single spin resolution \cite{Wrachtrup1}, albeit only with optically active defects.
Notably, both techniques can also efficiently detect spins through magnetic dipolar interactions. 
Dedicated Pulsed-EPR \cite{Mims} and ODMR protocols such as the Double-Electron-Electron-Resonance (DEER), Electron-Nuclear-Double-Resonance (ENDOR) or Cross-Relaxation (CR) at the Hartmann Hahn resonance have all indeed been proven essential for characterizing dipolar interactions, offering ways to detect distant spins \cite{Mamin}.

Another more recent route for detecting spins is to couple them to mechanical oscillators.
Electronic spins have first been detected this way in \cite{Alzetta} and nuclear spins can be sensed using state of the art magnetic resonance force microscope (MRFM) \cite{rugar, MaminH}. MRFM is a very efficient probe for paramagnetic impurities with nanometer resolution.
Conversely, torque sensors lends themselves naturally to the detection of solid state defects, which thanks to anisotropic terms in the spin hamiltonian can exert torques in a homogeneous magnetic field. 
Torque sensing was employed recently to detect the long-lived spins of NV centers in diamonds and to demonstrate spin-cooling of a mechanical oscillator \cite{DelordNat}.
In general, spin-mechanical experiments with long-lived spins also offer prospects for performing quantum mechanical experiments with large objects \cite{yin, Wan, Scala, Lee_2017}. However, while dipolar interactions can be detected using EPR or ODMR, mechanical detection of dipolar interactions between long-lived spins has not been observed. Compared to these two established techniques, it could offer the advantage of sensing angular momentum conservation during dipolar relaxation \cite{Zangara} as well as providing an alternative path for controlling the quantum motion of mechanical oscillators.

Here, we report on the detection of spin-interactions using a mechanical torque sensor.
Specifically, we employ a levitating diamond in a Paul trap with embedded NV centers and use the diamond orientation as a probe of the NV spin interactions.
%We use closely packed nitrogen-vacancy centers to ensure both large spin-torques on the mechanical oscillator together with large dipolar-interactions amongst the spins.

%When this occurs, the spin-torque is modified so that subsequently, the diamond rotates.

\begin{figure}[!ht]
  \centering \scalebox{0.3}{\includegraphics{scan_adamas_121.eps}}
  \caption{a) PL versus magnetic field. b) 4 ESRs showing the degeneracies}
  \label{scans1}
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.4}{\includegraphics{T_1.eps}}
  \caption{Longitudinal relaxation from NV centers for different degeneracies. 
  }\label{CR_deposited}
\end{figure}

%We first present the results and conclude by discussing the potential applications behind our observations.
The key idea of our mechanical detection of dipolar interactions lies in the cross-relaxation (CR) that takes place when two spins with different orientations and polarization become resonant. This can be achieved by tuning a magnetic field when using spin-1 defects such as the NV center.
The NV center has two electrons in the ground state, and a zero-field splitting $D=2.87$ GHz.
The ground states are labelled $| m_s={0}, {\pm 1}\rangle$.
Due to an intersystem crossing in the excited state of the NV centers \cite{Doherty}, the state $m_s=\ket{0}$ is brighter than the state $m_s=\ket{\pm 1}$ upon green laser optical illumination \cite{Hopper}. This provides a means to read out the Zeeman splitting by scanning a microwave tone around the resonance, carrying out ODMR.

The diamonds that we use are in the form of a powder of particles that have a diameter of 15 $\mu$m. They are supplied by the company Adamas, which produces diamonds with a concentration of NV centers in the 3-4 ppm range. 
We operate with a Paul trap that is similar to the one used in \cite{delordPRL} with particles are stably trapped at the bottleneck region, where both the electric field gradient and anisotropy are stronger.
The microwave is applied directly on the trapping electrode, which provides an efficient means to excite the spins.
The photoluminescence is detected using standard confocal microscopy. We use about 100 $\mu$W of laser light at 532nm to polarise the NV centers. The laser is focussed via a lens inside the vacuum chamber which has a numerical aperture of 0.5 and a working distance of 8mm. The focal point of the laser is kept few tens of micrometer away from the micro-diamond to mitigate the effect of radiation pressure and to enable laser excitation of the whole diamond \cite{delordPRL,delord2016}.

Let us first show that dipolar interaction in our diamonds is effective by studying cross-relaxation with diamonds that are simply attached to the trap.
In diamonds that contain more than 1 ppm of nitrogen-vacancy centers cross relaxation effects have been observed by many groups \cite{van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, mrozek_longitudinal_2015, jarmola_temperature-_2012, choi_depolarization_2017}.
It manifests itself as a change in the longitudinal relaxation rate of NV centers, which can then translate as a reduction of the overall photo-luminescence.

This effect has been routinely observed in bulk materials, the origin of which was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}. The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{Manson}) have a reduced longitudinal spin lifetime $T_1$. When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx ms$), the latter can drop 
to a few hundreds of micro-seconds \cite{Jarmola} when the NV concentration is large enough (typically in the ppm range)
This process has recently revived interest recently because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}. 
%In our spin-mechanics experiment, we use micro-diamond particules subjected to a green laser field and a magnetic field.
%To observe the dipolar effects in a levitating platform, we employ heavily doped micro-diamonds.
In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias}$ of a few 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet. 
Figure \ref{CR_deposited}-a) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
For these measurement we use GREEN POWER ? REPOLARISATION RATE ? 
Three dips in the fluorescence can be observed. 
We attribute these to cross-relaxation that takes place when two orientation meet. To be sure, we recorded Electronic spin resonance spectra at the values ...
There is indeed a strong correspondance between the PL dips and the degeneracy conditions. 
(See SI for details on planes that cross etc.). 
We tentatively attrtibute those to CR. 
%It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
%where, to our knowledge, such cross-relaxations have not been observed. 
%Indeed, surface effects may start to play a role and change quantitatively. 
To confirm our interpretation, and to extract the spin-torque dynamics in the mechanical experiment, we measure the longitudinal relaxation time ($T_1$ time) of the NV spins under different degeneracy conditions.

The measurement consists in applying a green laser to polarise the NV centers and to measure the photoluminescence at a later time. 
Such a measurement turns out to be hidden by recharging of NV centers in the dark, as has been observed and analyzed by several groups \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use a sequence presented in the SI, where a microwave pulse is applied or not prior to relaxation. Subtracting the two resulting signals eliminates charge dynamics from the spin decay.
The results of the measurements are shown in Fig. \label{CR_deposited}b) for different degeneracies. 
It can be seen that $T_1$ is greatly shortened, from X to Y when approaching degeneracy, confirming this interpretation.

\begin{figure}[!ht]
  \centering \scalebox{0.13}{\includegraphics{dilutednv.eps}}
  \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond cristalline structure. Here, the magnetic field is tuned so that no cross-relaxation can occur.
The resulting non-resonant spin-torque $\tau_{NR}$ applied to the levitating diamond is sketched below. 
ii) Here, the magnetic field is tuned so that cross-relaxation can occur between two classes of NVs.
The resulting resonant spin-torque $\tau_{R}$ applied to the levitating diamond (below) is thus modified. 
  }\label{principle}
\end{figure}

%
%\begin{figure}[!ht]
%  \centering \scalebox{0.6}{\includegraphics{cross_121_200G.png}}
%  \caption{Numerics : PL change in the presence of cross-relaxation (top) and spin-mechanical detection (below). 
%  }\label{CR_deposited}
%\end{figure}

Let us now study these heavily doped diamonds in the trap, with the goal to mechanically detect the dipolar interactions.
Specifically, we will detect the diamond angle and seek for changes in the diamond orientation at CR conditions. The detection is optimized by collecting the back reflected green light from the diamond interface, separated from the excitation light using a polarizing beam splitter. The best sensitivity is achieved by using a speckle pattern produced by the rough surface of the micro-diamond under coherent illumination. When the particle is stably levitating, at the particle image plane, a speckle feature is formed. To detect the diamond motion, we then focus a small area of this image onto an optical fibre and detect the photons transmitted through the fibre with a single-photon avalanche photodiode. The detected signal is then highly sensitive to the particle position and orientation.

%For a given levitating particle, we can optimize, in real time, the signal coming from the angular displacement of the particle by selecting the most favorable region of the particle image. To do this, we monitor our optical signal while switching a microwave field tuned to one of the ESR transitions, at a frequency of 1 Hz.  Alignment is realized by maximizing the variation of the coupled light as the diamond rotates between two angular positions.

One major ingredient of the present study compared to \cite{DelordNat} is the spin-torque coming from the NV centers in the absence of the microwave tone.
For this, let us consider first the torque coming from a single spin due to the transverse component of the magnetic field when the NV is in the ground state.
The hamiltonian for one NV orientation with quantization axis $z$ reads 
\begin{equation}\hat{H}_{\rm NV}=\hbar D \hat{S}_z^2+ \hbar \gamma_e \bf B  \cdot \bf\hat S.
\end{equation}
Under the condition $\gamma B \ll D$, $H_{B}= \gamma_e B  ( S_x \cos\theta + S_z \sin\theta)$ can be treated as a perturbation to the anisotropic part of the hamiltonian. 
The perturbed energy $\epsilon_g$ of $\ket{0}$ due to the transverse B field is then
\begin{equation} \epsilon_g=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B)^2 }{D} \cos^2\theta.
\end{equation}
One can deduce the total torque when N spins of a given orientation are polarised in the ground state to be 
\begin{equation}\tau_g=-N \frac{\partial \epsilon_0}{\partial \theta}=\hbar N  \frac{(\gamma_e B)^2}{D} \sin 2\theta.\end{equation}

At an angle $\theta=\pi/4$, where the torque is maximized, using $10^9$ spins polarized in the ground state and a B field of 100 G, the static torque is about $10^{-19}$ N.m.
We checked numerically that at these magnetic field values and angle, the NV are still polarized mostly in the ground state. 
Note that the torque is zero at $\theta=mod (\pi/2)$ where the eigenstates of $S_z$ are not mixed. The hamiltonian ground state is the eigenstate $| S=1,m_s=0\rangle$, which is is non magnetic. 
%Note also that the torque tends to zero if  tends to zero at any angle $\theta$. This is the case in the presence of strong spin-depolarisation, i.e. when $\hat\rho\rightarrow\frac{1}{3}\hat{I}$ (with $\hat{I}$ the identity operator).

Although the physics of the torque differs from the torque observed in \cite{DelordNat}, which came from the microwave excitation to magnetic states, this torque is comparable in magnitude. With our angular resolution of 100 $\mu$rad/$\sqrt{\rm Hz}$, it should thus enable a measurable shift in the angular position to be detected and possibly modified when cross-relaxation occurs. 
One major difference however, compared to \cite{DelordNat} is that the microwave selectively excited one class only. 
Here, we are interested in the modification of the total torque in the NV ground state, where all the NV orientation play a role. We thus also need to evaluate their contributions.
When estimating the total torque including the four classes, we found (see SI) that is is one order of magnitude weaker than for the above single class estimate.

\begin{figure*}[!ht]
  \centering \scalebox{0.45}{\includegraphics{CRmeca}}
  \caption{a) Angular detection of the diamond as a function of microwave frequency for three different magnetic field values. 
b) PL detection as a function of Bem across a co-resonance. trace i) exp, trace ii) theory. 
c) Angular detection... 
  }
  \label{data}
\end{figure*}

Let us now explain why CR modifies the NV spin torque.
Fig. \ref{principle}-i) is a depiction of a levitating diamond containing NV centers that have four possible directions along the $<111>$ axes. 
In the presence of a magnetic field, each NV centers has a different magnetic field projection, which means a different torque. 
Below, we show the sum of the ground state spin-torques $\tau_{NR}$, resulting from the four classes. 
Fig. \ref{principle}-ii) shows the same situation, with a magnetic field that is tuned so that two NV classes are co-resonant. 
When the re-polarisation rate due to the green laser is smaller than the cross-relaxation rate, the magnetization coming from these two classes (indicated by a circle) tends to cancel. The population in the three NV eigenstates will indeed equilibrate, giving a smaller magnetization.
The resulting torque is thus changed to $\tau_{R}$ and the diamond then turns about another angle.
This is the essence of the proposed dipolar detection.

%shows the total torque when 1-2-3-4 NV transitions are in the ground state. It is very small because the torque from 
%2-4 and 1-3 cancel each other. Looking at each pair 2-4 and 1-3 however, one can see that the torque is large (here 20 MHz for one spin).

To observe this effect we use similar parameters and magnetic field arrangement than when the diamonds where outside the trap.
Let us first monitor the spin-mechanical resonances. 
Fig. \ref{data}-a) shows the signal coming from the laser reflected off the diamond surface as a function of the microwave signal driving the NV centers' spins, with a zoom on the $m_s=-1$ states frequencies.
Three ESR have been taken for three different $\bf B_{\rm em}$ amplitudes. 
At 0.4 and 0.6 V, one can observe 4 pics in the spectrum, that are demonstrate spin induced torque on the diamond from the 4 classes of NVs.
At 0.5 V however, two classes merge at 2.75 GHz. This is where we expect CR.

Analysis (SI) suggest that we cross a single degeneracy that is X.
 Fig. \ref{data}-b) shows the photoluminescence as a function of $\bf B_{\rm em}$ both experimentally (trace i) and numerically (trace ii)
As expected, the PL decreases as the magnetic sweeps across the degeneracies at the same point. 
 Fig. \ref{data}-c), trace i) is a measurement of the angle in the same conditions (same acquisition time in particular). 
A pronounced variation of the reflected signal is also observed. Subtracting the varying background (see SI) and approximating this curve by a Gaussian, we obtain a width that is in the same ballpark as the PL (TO BE DONE).
There is a close correspondence between degeneracy and diamond rotation, letting us conclude that we are observing a change in the spin-torque as two spin-classes cross-relax. 
In SI, we show similar SM curve using different bias magnetic fields for different degeneracy conditions. 
Trace ii) is a corresponding calculation, which give the correct order of magnitude.  
Note that the present effect cannot be attributed to a change in the angular momentum, as in the Einstein de Haas effect. 
The latter would manifest itself for much smaller particles sizes. Further, the rotation would occur dynamically and not manifest itself as a static torque as is the case in the present experiment.

Such tunable mechanically-detected dipolar interactions can be used to control the temperature and stiffness of mechanical oscillators.
The starting point for this effect would be to obtain a dynamical back action to the mechanical oscillator \cite{aspelmeyer}. 
Note that it is the essence of the detection in MRFM : change in the stiffness of MRFM oscillators. Here, one could also change the imaginary part of the susceptibility.
Under small enough laser, at a magnetic field value on the side of the cross-relaxation peak, the fluctuator will partly depolarize the spins and let the two other transitions apply the torque until the spins re-polarize back. This produces a cooling cycle.  
In essence this forms the analogue of the spin-cooling shown in \cite{DelordNat}, but without microwave and using dipolar-coupled spins.

Mechanical detection of dipolar interactions can also find applications for detecting spins that cannot be polarized optically. 
To give an example, the nitrogen defect itself (also called the $P_1$ center) is a paramagnetic defect that cannot be read-out directly. 
Using the above cross-relaxation effect, one could use a magnetic field that is oriented along the 111 direction etc..

One last implication is the Einstein de Haas effect.
ODMR and EPR detect energies shifts but not angular momenta. 
It would be beneficial to detect the dipolar interaction using the motion of the body itself in order to gain knowledge about angular momentum exchange.
as proposed in \cite{Zangara}.
%The dipolar interaction can readily be detected indirectly using the motion of the object itself or via the Zeeman shift or motion of close paramagnetic or ferro-magnetic objects in the presence of a magnetic field.

To conclude, we observe tunable dipolar-interactions amongst long lived spins using a mechanical detection scheme.
We enlarge the toolbox of spin-mechanical platforms and ....
One could consider this work to be a kind of bottom-up approach for magnetism, where both the interaction and the relaxation can be tuned in order to reproduce the behaviors of a magnet. 
The detailed microscopic origin of magnetism depends a lot on the material and the relaxation mechanism is not controlled and is very short, making measurements of the  Gilbert-damping a complicated task.
The microscopic origin of the angular momentum exchange is also not well-understood. 
Our demonstration is thus great because it will solve all these problems and because we are very good.

\section*{acknowledgements}
GH acknowledges funding by the French National Research Agency (ANR) through the project SMEQUI and by the T-ERC program through the project QUOVADIS.

\bibliography{trap}

\end{document}

### 2020-10-01 Mes papiers/Méca CR/Meca_cross_SI.tex
covid

\documentclass[preprintnumbers,amsmath,amssymb,onecolumn,12pt]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage[caption=false]{subfig}

\def\sgn{\mathop{\rm sgn}}

\begin{document}

\vspace{0.2in}
{\Large \hspace{1.6in}\textsc{Supplementary Material} }\\
\
\title{Mechanical Detection of Dipole-Dipole Interactions \\between Electronic Spins in a Solid}

\author{C. Pellet-Mary, P. Huillery, M. Perdriat, G. H\'etet}

\affiliation{$^1$
Laboratoire de Physique de l'Ecole normale sup\'erieure, ENS, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e Paris-Diderot, Sorbonne Paris Cit\'e, Paris, France.
}

\begin{abstract}
\end{abstract}

\maketitle

\tableofcontents

\newpage

\section*{Experimental details}

\subsection{Experimental setup}

As illustrated in Fig.\ref{Optics}, the diamond sample is typically illuminated with 1mW of 532 nm laser light, focused by a NA = 0.5 objective. An acousto-optic modulator (AOM) is used to switch on and off the 532nm laser and to finely tuned its power. The photo-luminescence (PL) is collected by the objective, separated form the excitation light using a dichroic mirror (DM) and a 532nm notch filter (NF), and detected using a multimode-fibered single-photon avalanche photo-detector (APD) (SPCM-AQRH-15 from Perkin Elmer). Typically, we detect PL photons at a rate of 1MHz.

\begin{figure}[!ht]
  \centering \scalebox{0.7}{\includegraphics{map_200G_juste_CR}}
  \caption{?}
	\label{Optics}
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.7}{\includegraphics{map_cross_22}}
  \caption{?}
	\label{Optics}
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.7}{\includegraphics{map_cross_121}}
  \caption{?}
	\label{Optics}
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.5}{\includegraphics{cross_22_200G}}
  \caption{?}
	\label{Optics}
\end{figure}

\subsection{Magnetic field calibration}

A permanent magnet is placed a few cm away from the diamond sample in order to apply a uniform magnetic field to the NV centers together with an electro-magnet.

To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonace (ODMR) spectra....

\subsection{Spin-mechanical detection}

The microwave detuning is scanned in 2~MHz steps with a duration of 10 ms per points. During those 10ms, the diamond orientation has enough time to reach its equilibrium position and the spin torque effect can be observed. The average count-rate is about 1 MCounts/s.

\subsection{PL detection}

In these measurements, compared to \cite{DelordNat}, to detect the PL we use another detection channel that is more resilient to motion. We therefore do not need to detect prior to ring down.

\bibliography{Bib-with-NV-Nuclear-spins}

\end{document}

### 2020-10-01 data/20200806/1-2-1/simu.py

@@ -128,16 +128,23 @@ def spher_to_cart(theta,phi):
 	return(np.array([sin(theta)*cos(phi),sin(theta)*sin(phi),cos(theta)]))
 
 
+h=6.626*1E-34
+gamma_e=2.8*1E6 #en Hz/Gauss
+n_centers=1E9/4 #nb de spins par classe
 
-fig,ax=plt.subplots(2)
+xmin=80
+
+fig,ax=plt.subplots()
 
 
 theta_111=arctan(sqrt(2))
 gamma_las=5E-3
 
-B_in=np.array([72.65829175, 30.55410235, 85.39414061])
-B_out=np.array([ 27.41941162,  50.08273284, 114.89942708])
-n=200
+B_in=np.array([77.05651621, 28.6554855 , 82.52557109])
+B_out=np.array([ 26.79109384,  50.35396382, 115.30922273])
+n=201
+xs=np.linspace(0,np.linalg.norm(B_out-B_in),n)
+xs=xs-xs[xmin]
 Bxs=np.linspace(B_in[0],B_out[0],n)
 Bys=np.linspace(B_in[1],B_out[1],n)
 Bzs=np.linspace(B_in[2],B_out[2],n)
@@ -148,7 +155,7 @@ PLs_sans=[]
 for i in range(n):
 	B=[Bxs[i],Bys[i],Bzs[i]]
 	H=Hamiltonian_NV_propre_base(B)
-	gamma_t1=make_t1_list(B)
+	gamma_t1=make_t1_list(B,width=6)
 	s,rho_0=spin_3vx_4classes(B,gamma_las,gamma_t1)
 	PLs+=[rho_0]
 	torques+=[np.cross(s,B)]
@@ -157,21 +164,68 @@ for i in range(n):
 	PLs_sans+=[rho_0]
 	torques_sans+=[np.cross(s,B)]
 
-torques=np.array(torques)
-torques_sans=np.array(torques_sans)
-color = next(ax[0]._get_lines.prop_cycler)['color']
-ax[0].plot(PLs,label='PL',color=color)
-ax[0].plot(PLs_sans,'--',color=color)
-color = next(ax[1]._get_lines.prop_cycler)['color']
-ax[1].plot(torques[:,0],label='Torque x',color=color)
-ax[1].plot(torques_sans[:,0],'--',color=color)
-color = next(ax[1]._get_lines.prop_cycler)['color']
-ax[1].plot(torques[:,1],label='Torque y',color=color)
[…]

### 2020-10-01 data/20200922/traitement.py

@@ -1,0 +1,169 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def quad_fit(x,y) :
+	A=np.vstack([x**2,x,np.ones(len(x))]).T
+	a,b,c = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([a,b,c],a*x**2+b*x+c)
+
+def fit_ordre_4(x,y) :
+	A=np.vstack([x**4,x**3,x**2,x,np.ones(len(x))]).T
+	a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([a,b,c,d,e],a*x**4+b*x**3+c*x**2+d*x+e)
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None,ss=0) :
+	if not ss :
+		ss=y[0]
+	if not Amp :
+		if max(y)-ss > ss-min(y) :
+			Amp=max(y)-ss
+		else :
+			Amp=min(y)-ss
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/5)]-x[0]
+	def f(x,Amp,x0,sigma,ss) :
+		return Amp*np.exp(-((x-x0)/(2*sigma))**2)+ss
+	p0=[Amp,x0,sigma,ss]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,popt[0],popt[1],popt[2],popt[3]))
+
+def lor_fit(x,y,Amp=None,x0=None,sigma=None,ss=None) :
+	if not ss :
+		ss=y[0]
[…]

### 2020-10-06 Mes papiers/Méca CR/Meca_cross_SI.tex
toto

@@
 
+\subsection{$T_1$ Measurement}
+%Include graphics...
+The protocol to measure the spin lifetime of the NV centers is described in Fig.
+...
+.
+The spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolove in the dark for a variable dark time $\tau$.
+The spin state is finally read out with a 10 $\mu$s laser pulse, shorter than the polarization time of the spins.
+
+The sequence is repeated a second time, adding a resonnant microwave $\pi$ pulse on a transition of one of the four classes of NV$^-$, exchanging the population of the $m_s=\ket{0}$ state with one of the $m_s=\ket{\pm 1}$ state.
+
+The two sequences are then substracted, allowing us to measure the spin state population of a single class, and removing unwanted constributions such as charge state transfer in the dark.
+
 \subsection{Magnetic field calibration}

### 2020-10-06 data/20200922/traitement.py

@@ -35,6 +35,10 @@ def fit_ordre_4(x,y) :
 	a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
 	return([a,b,c,d,e],a*x**4+b*x**3+c*x**2+d*x+e)
 
+def fit_ordre_6(x,y) :
+	A=np.vstack([x**6,x**5,x**4,x**3,x**2,x,np.ones(len(x))]).T
+	w,z,a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([w,z,a,b,c,d,e],w*x**6+z*x**5+a*x**4+b*x**3+c*x**2+d*x+e)
 def gauss_fit(x,y,Amp=None,x0=None,sigma=None,ss=0) :
 	if not ss :
 		ss=y[0]
@@ -130,21 +134,66 @@ def stretch_et_phonon(x,y,Amp=None,tau=None) :
 	popt, pcov = curve_fit(f, x, y, p0)
 	return(popt,f(x,popt[0],popt[1]))
 
+def lissage(y):
+	yl=np.zeros(len(y))
+	yl[0]=y[0]
+	yl[-1]=y[-1]
+	for i in range(1,len(y)-1):
+		yl[i]=(y[i]/2+y[i-1]/4+y[i+1]/4)
+	return(yl)
+
+
 def ask_name():
 	qapp = QApplication(sys.argv)
 	fname,filters=QFileDialog.getOpenFileName()	
 	return fname
 
-# xmin=0
-# xmax=-1
-indices=[294,313,327,339,352,368,383,405,456,474]
-xmin=295
-xmax=474#295+113#474
+
+
+
+def fit_droite():
+	# xmin=0
+	# xmax=-1
+	indices=[294,313,327,339,352,368,383,405,456,474]
+	xmin=295
+	xmax=474#295+113#474
+
+	fname='scan_rose_111'
+	x,y=extract_data(fname+'.txt',ycol=1)
+	x=x*56.528132258549384-9.624229803755462
+	y=y/max(y)
+	xs=[]
+	ys=[]
+	for i in indices :
+		xs+=[x[i]]
+		ys+=[y[i]]
+	xs=np.array(xs)
+	ys=np.array(ys)
+	plt.plot(x,y)
+
+
+	# fname='scan_uW_2783'
[…]

### 2020-10-07 Mes papiers/Méca CR/Meca_cross.tex
tutu

@@
 We attribute these to cross-relaxation that takes place when two orientation meet. To be sure, we recorded Electronic spin resonance spectra at the values ...
-There is indeed a strong correspondance between the PL dips and the degeneracy conditions. 
+There is indeed a strong correspondence between the PL dips and the degeneracy conditions. 
 (See SI for details on planes that cross etc.). 
-We tentatively attrtibute those to CR. 
+We tentatively attribute those to CR. 
 %It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
@@
 
-The measurement consists in applying a green laser to polarise the NV centers and to measure the photoluminescence at a later time. 
+The measurement consists in applying a green laser to polarize the NV centers and to measure the photoluminescence at a later time. 
 Such a measurement turns out to be hidden by recharging of NV centers in the dark, as has been observed and analyzed by several groups \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
@@
   \centering \scalebox{0.13}{\includegraphics{dilutednv.eps}}
-  \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond cristalline structure.
+  \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond crystalline structure.
 Here, the magnetic field is tuned so that no cross-relaxation can occur.

### 2020-10-16 Mes papiers/Méca CR/Meca_cross.tex
weekend

@@
 \begin{figure}[!ht]
-  \centering \scalebox{0.3}{\includegraphics{scan_adamas_121.eps}}
-  \caption{a) PL versus magnetic field. b) 4 ESRs showing the degeneracies}
-  \label{scans1}
+  \centering \scalebox{0.45}{\includegraphics{Adamas_fixe.eps}}
+  \caption{a) Photoluminescence signal in function of the scanning magnetic field.
+b) Predicted transition frequencies for the four classes of NV centers (dashed).
+ESR spectra for particular value of the magnetic field (plain, vertical)}
+  \label{CR_deposited}
 \end{figure}
@@
 \begin{figure}[!ht]
-  \centering \scalebox{0.4}{\includegraphics{T_1.eps}}
-  \caption{Longitudinal relaxation from NV centers for different degeneracies. 
-  }\label{CR_deposited}
+  \centering \scalebox{0.5}{\includegraphics{T1.eps}}
+  \caption{Longitudinal relaxation from a single class of NV centers with exponential fit.
+i) The class is not at resonance with any other classes : T$_1$=1.02 ms.
+ii) The class is at resonance with another class : T$_1$=0.28 ms.
+iii) The class is at resonance with the three other classes : T$_1$=0.15 ms.
+  }\label{T_1}
 \end{figure}
@@
 We use about 100 $\mu$W of laser light at 532nm to polarise the NV centers.
-The laser is focussed via a lens inside the vacuum chamber which has a numerical aperture of 0.5 and a working distance of 8mm.
+The laser is focused via a lens inside the vacuum chamber which has a numerical aperture of 0.5 and a working distance of 8mm.
 The focal point of the laser is kept few tens of micrometer away from the micro-diamond to mitigate the effect of radiation pressure and to enable laser excitation of the whole diamond \cite{delordPRL,delord2016}.
@@
 The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{Manson}) have a reduced longitudinal spin lifetime $T_1$.
-When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx ms$), the latter can drop
+When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx 5 ms$), the latter can drop
 to a few hundreds of micro-seconds \cite{Jarmola} when the NV concentration is large enough (typically in the ppm range)
@@
 %To observe the dipolar effects in a levitating platform, we employ heavily doped micro-diamonds.
-In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias}$ of a few 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet.
+In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias} \approx$ 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet.
 Figure \ref{CR_deposited}-a) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
-For these measurement we use GREEN POWER ? REPOLARISATION RATE ? 
+For these measurement we use a 1 mW green laser excitation which yields a repolarisation rate of about 100 $\mu$s. 
 Three dips in the fluorescence can be observed. 
-We attribute these to cross-relaxation that takes place when two orientation meet. To be sure, we recorded Electronic spin resonance spectra at the values ...
+We attribute these to cross-relaxation that takes place when two orientation meet.
+To be sure, we recorded Electronic spin resonance spectra for different values which allows us to measure the total magnetic field felt by the spins, and therefore to extrapolate the transition frequencies of the four classes of NV for any applied $\bf B_{\rm em}$.
 There is indeed a strong correspondence between the PL dips and the degeneracy conditions. 
@@
 Subtracting the two resulting signals eliminates charge dynamics from the spin decay.
-The results of the measurements are shown in Fig. \label{CR_deposited}b) for different degeneracies. 
+The results of the measurements are shown in Fig. \ref{T_1} for different degeneracies. 
 It can be seen that $T_1$ is greatly shortened, from X to Y when approaching degeneracy, confirming this interpretation.
@@
 \begin{figure}[!ht]
-  \centering \scalebox{0.13}{\includegraphics{dilutednv.eps}}
+  \centering \scalebox{0.13}{\includegraphics{dilutednv.pdf}}
   \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond crystalline structure.
@@
   \caption{a) Angular detection of the diamond as a function of microwave frequency for three different magnetic field values. 
-b) PL detection as a function of Bem across a co-resonance. trace i) exp, trace ii) theory. 
+b) PL detection as a function of B$_{em}$ across a co-resonance. trace i) exp, trace ii) theory. 
 c) Angular detection... 
@@
 Three ESR have been taken for three different $\bf B_{\rm em}$ amplitudes. 
-At 0.4 and 0.6 V, one can observe 4 pics in the spectrum, that are demonstrate spin induced torque on the diamond from the 4 classes of NVs.
-At 0.5 V however, two classes merge at 2.75 GHz. This is where we expect CR.
+At 10 and 25 G, one can observe 4 pics in the spectrum, that are demonstrate spin induced torque on the diamond from the 4 classes of NVs.
+At 17 G however, two classes merge at 2.75 GHz. This is where we expect CR.
 
-Analysis (SI) suggest that we cross a single degeneracy that is X.
+Analysis (SI) suggest that we cross a single degeneracy that is the magnetic field crosses a $(110)_\perp$ plane.
  Fig. \ref{data}-b) shows the photoluminescence as a function of $\bf B_{\rm em}$ both experimentally (trace i) and numerically (trace ii)
@@
  Fig. \ref{data}-c), trace i) is a measurement of the angle in the same conditions (same acquisition time in particular). 
-A pronounced variation of the reflected signal is also observed.
-Subtracting the varying background (see SI) and approximating this curve by a Gaussian, we obtain a width that is in the same ballpark as the PL (TO BE DONE).
+A pronounced variation of the reflected signal is also observed. Approximating this curve by a Gaussian, we obtain a width that is in the same ballpark as the PL (1.3G and 1.7G respectively).
 There is a close correspondence between degeneracy and diamond rotation, letting us conclude that we are observing a change in the spin-torque as two spin-classes cross-relax. 
@@
 
-To conclude, we observe tunable dipolar-interactions amongst long lived spins using a mechanical detection scheme.
+To conclude, we observe tunable dipolar-interactions among long lived spins using a mechanical detection scheme.
 We enlarge the toolbox of spin-mechanical platforms and ....

### 2020-10-16 data/20200917/find_B.py

@@ -1,0 +1,198 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import minimize
+
+#unité naturelle : Gauss et MHz
+
+D=2870 #Mhz
+gamma_e=2.8 #Mhz/gauss
+
+Sz=np.array([[1,0,0],[0,0,0],[0,0,-1]])
+Sy=np.array([[0,-1j,0],[1j,0,-1j],[0,1j,0]])*1/np.sqrt(2)
+Sx=np.array([[0,1,0],[1,0,1],[0,1,0]])*1/np.sqrt(2)
+Sz2=np.array([[1,0,0],[0,0,0],[0,0,1]]) # Pour éviter une multilplcation matricielle
+
+c1=np.array([-1,1.,-1])/np.sqrt(3)
+c2=np.array([1,1,1])/np.sqrt(3)
+c3=np.array([-1,-1,1])/np.sqrt(3)
+c4=np.array([1,-1,-1])/np.sqrt(3)
+
+carb=[c1,c2,c3,c4]
+
+angle_111_theta=np.arccos(1/np.sqrt(3))
+angle_111_phi=np.arccos(1/np.sqrt(3)/np.sin(angle_111_theta))
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def find_nearest(value,array): #suppose the array is sorted
+	i=0
+	while i<len(array)-1 and array[i]<value :
+		diff=value-array[i]
+		i+=1
+	if i==len(array)-1:
+		return i
+	elif i==0:
+		return i
+	elif diff < array[i]-value :
+		return i-1
+	else :
+		return i
+
+
+
+
+class mag_field() : #l'amplitude est en Gauss
+	def __init__(self,amp,theta,phi,radian=True) :
+		self.amp=amp
+		if not radian :
+			theta=theta*np.pi/180
+			phi=phi*np.pi/180
+		self.theta=theta
+		self.phi=phi
+		self.z=amp*np.cos(theta)
[…]

### 2020-10-19 Mes papiers/Méca CR/Meca_cross.tex
toto

@@
   \centering \scalebox{0.45}{\includegraphics{CRmeca}}
-  \caption{a) Angular detection of the diamond as a function of microwave frequency for three different magnetic field values. 
-b) PL detection as a function of B$_{em}$ across a co-resonance. trace i) exp, trace ii) theory. 
-c) Angular detection... 
+  \caption{a) Angular detection of the diamond as a function of microwave frequency for three different magnetic field values and representation of the co-resonance crossing.
+b) PL detection as a function of B$_{em}$ across a co-resonance.
+i) experimental data with gaussian fit, ii) simulation of the population in $\ket{m_s=0}$ state, taking into account the CR (plain) or not (dashed).
+c) Angular detection as a function of B$_{em}$ across a co-resonance.
+i) experimental data with gaussian fit, ii) simulation of the magnetic torque applied to the diamond, taking into account the CR (plain) or not (dashed).
   }

### 2020-10-19 Mes papiers/Méca CR/Meca_cross_SI.tex
weekend
toto

--- 2020-10-01 Mes papiers/Méca CR/Meca_cross_SI.tex
@@
 \usepackage{natbib}
+\usepackage{physics}
 \usepackage[caption=false]{subfig}
@@
 
+\section*{NV$^-$ center Theory}
+\subsection{NV spin Hamiltonian}
+The NV spin Hamiltonian in its fundamental level can be written as :
+\begin{equation*}
+  \hat{\mathcal{H}}_s=D S_z^2 + \gamma_e \textbf{B}\cdot\hat{\textbf{S}},
+  \end{equation*} 
+Where $D = (2\pi) 2.87$ GHz is the crystal field splitting and $\gamma_e = 28 $GHz/T is the electron gyromagnetic ratio.
+We neglected contributions of the strain and local electric field since we are working with magnetic fields of the order of 100G, as well as the hyper-fine interaction with $^{14}$N since we are working with ensembles with a typical inhomogeneous broadening $\frac{1}{T_2^*} = (2\pi) 5$ MHz.
+
+The \textbf{z} axis of the $S_z$ operator here is the axis formed by the nitrogen atom and the vacancy, hence why we do have different energy transitions for different NV orientations, depending on the projection of the magnetic field on the NV axis.
+
+\subsection{Diamond crystalline axes and degeneracy conditions}
+\begin{figure}[!ht]
+  \centering \scalebox{0.45}{\includegraphics{SI_Cristallo}}
+  \caption{Representation of the magnetic field in the diamond crystalline basis.}
+	\label{cristallo}
+\end{figure}
+There are four possible crystalline axis for the NV centers (so-called "classes" of NV) which are depicted in Fig.
+\ref{cristallo} b) and correspond to the crystalline directions [$111$], [$1\bar 1 \bar 1$], [$\bar 1 1 \bar 1$] and [$\bar 1 \bar 1 1$].
+
+The magnetic field is represented in the diamond basis in Fig.
+\ref{cristallo} a) where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z} ([$001$]) direction.
+
+For some orientations of the magnetic field, the projection of the magnetic field on two or more NV axes will be identical, and therefore the energy level of the corresponding classes will be the same.
+These degeneracies are represented on fig \ref{cristallo} c), where the dashed lines are the locus of the crystalline planes orthogonal to the [110] directions (6 planes in total).
+When the magnetic field belongs to these plane, we observe a degeneracy between two classes of NV, as can be seen in the Fig.
+4 of the main paper.
+
+The plain lines are the locus of the planes orthogonal to the [100] directions (3 planes total).
+When the magnetic field is in these planes, every class is at resonance with another class, as can be seen in Fig.
+X.
+
+The red circles correspond to the [100] directions, for which the four classes of NV or at degeneracy, and the green squares to the [111] direction where one class is aligned with the magnetic field, and the three other are at degeneracy.
+
+Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 4.
+of the main text, where we can see a degeneracy plane of the [110]$_\perp$ family being crossed.
+
 \section*{Experimental details}
@@
 
-As illustrated in Fig.\ref{Optics}, the diamond sample is typically illuminated with 1mW of 532 nm laser light, focused by a NA = 0.5 objective.
+\begin{figure}[!ht]
+  \centering \scalebox{0.3}{\includegraphics{setup}}
+  \caption{Illustration of the experimental setup}
+	\label{Optics}
+\end{figure}
+
+The experimental setup illustrated in Fig.\ref{Optics} is very similar to the one used in \citep{DelordPRL} with the addition of a permanent magnet and an electromagnetic (EM) coil in order to perform magnetic filed scans.
+The diamond sample is typically illuminated with 1mW of 532 nm laser light, focused by a NA = 0.5 objective.
 An acousto-optic modulator (AOM) is used to switch on and off the 532nm laser and to finely tuned its power.
@@
 
+The Paul trap is a pseudo-ring as can be seen in \citep{DelordPhD} which acts both as trap through the high voltage (HV) and as a microwave (MW) antenna.
+
+The magnetic field generated by the (homemade) EM coil is controlled by a frequency generator (HP 33120A) with a triangle signal in order to perform magnetic field ramps.
+For scans of less than 100 G, no amplification of the signal is required.
+
+While the levitating setup is located in a vacuum chamber, all the experiments presented in this article are performed at atmospheric pressure.
+
+\subsection{$T_1$ Measurement}
 \begin{figure}[!ht]
-  \centering \scalebox{0.7}{\includegraphics{map_200G_juste_CR}}
-  \caption{?}
-	\label{Optics}
+  \centering \scalebox{1}{\includegraphics{T1_soustraction_shema}}
+  \caption{T$_1$ measurement protocol for a single class in an ensemble of NV centers}
+	\label{T1_protocol}
 \end{figure}
+The protocol to measure the spin lifetime of the NV centers, as presented in Fig 2.
+of the main text, is described in Fig.
+\ref{T1_protocol} .
+The spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$.
+The spin state is finally read out with a 10 $\mu$s laser pulse, shorter than the polarization time of the spins.
 
-\begin{figure}[!ht]
-  \centering \scalebox{0.7}{\includegraphics{map_cross_22}}
-  \caption{?}
-	\label{Optics}
-\end{figure}
+The sequence is repeated a second time, adding a resonant microwave $\pi$ pulse on a transition of one of the four classes of NV$^-$, exchanging the population of the $m_s=\ket{0}$ state with one of the $m_s=\ket{\pm 1}$ state.
 
-\begin{figure}[!ht]
-  \centering \scalebox{0.7}{\includegraphics{map_cross_121}}
-  \caption{?}
-	\label{Optics}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering \scalebox{0.5}{\includegraphics{cross_22_200G}}
-  \caption{?}
-	\label{Optics}
-\end{figure}
+The two sequences are then subtracted, allowing us to measure the spin state population of a single class, and removing unwanted contributions to the photoluminescence, such as charge state transfer in the dark.
 
@@
 
-\bibliography{Bib-with-NV-Nuclear-spins}
+\section*{Cross-relaxation detection for another type of degeneracy}
+
+\section*{Simulation details}
+In this part we will discuss the method used to simulate the average torque as well as the $\ket{m_s=0}$ population of the stationary state.
+The numerical solving of the master equations was performed thanks to Quantum Toolbox in Python (QuTiP) \citep{qutip1} \citep{qutip2}.
+
+In order to describe the dynamics of our spin ensemble, besides the spin Hamiltonian described in the first part, we introduce the incoherent optical pumping through the jump operators $\mathcal{L}_+ = \Gamma_l \ket{0}\bra{+1} $ and $\mathcal{L}_- = \Gamma_l \ket{0}\bra{-1} $, where $\Gamma_l \approx (2\pi) 10$kHz is the polarizing rate due to the laser.
+
+We also introduce the usual T1 jump operators...
+
+In order to describe the T1 modification induced by the cross-relaxations, we use a phenomenological model ...
+
+According to previous measurements \citep{choi_depolarization_2017}, only the $\ket{0}\bra{\pm1}$ (corresponding to a single quantum exchange in the dipole-dipole interaction) operators are modified by the cross-relaxations.
+
+\bibliography{trap}
 

### 2020-10-28 Mes papiers/Méca CR/Meca_cross.tex
confinement?

@@
 
-\author{C. Pellet-Mary, P. Huillery, M. Perdriat, G. H\'etet}
+\author{C. Pellet-Mary$^1$, P. Huillery$^1$, M. Perdriat$^1$, G. H\'etet}
 
-\affiliation{$^1$Laboratoire Pierre Aigrain, Ecole normale sup\'erieure, PSL Research University, CNRS, Universit\'e Pierre et Marie Curie, Sorbonne Universit\'es, Universit\'e Paris Diderot, Sorbonne Paris-Cit\'e, 24 rue Lhomond, 75231 Paris Cedex 05, France.}
+\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e de Paris , 24 rue Lhomond, 75231 Paris Cedex 05, France.}
 
@@
 
-\begin{figure}[!ht]
-  \centering \scalebox{0.45}{\includegraphics{Adamas_fixe.eps}}
-  \caption{a) Photoluminescence signal in function of the scanning magnetic field.
-b) Predicted transition frequencies for the four classes of NV centers (dashed).
-ESR spectra for particular value of the magnetic field (plain, vertical)}
-  \label{CR_deposited}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering \scalebox{0.5}{\includegraphics{T1.eps}}
-  \caption{Longitudinal relaxation from a single class of NV centers with exponential fit.
+\begin{figure*}[!ht]
+  \centering \scalebox{0.33}{\includegraphics{Adamas_fixe.eps}}
+  \caption{Depolarization of NV ensembles induced by co-resonances.
+a) Representation of the NV-NV depolarization process.
+Green arrows represent the optical pumping in the brighter $\ket{m_s=0}$ state, curvy blue arrows represent the longitudinal relaxation where NV$_2$ has a considerably shorter T$_1$ than NV$_1$.
+Red circles represent the population in each state and red dashed arrows represent the resonant dipole-dipole interaction between the two NV centers.
+b) Photoluminescence signal in function of the scanning magnetic field.
+c) Predicted transition frequencies for the four classes of NV centers (dashed).
+ESR spectra for particular value of the magnetic field (plain, vertical).
+d) Longitudinal relaxation from a single class of NV centers with exponential fit.
 i) The class is not at resonance with any other classes : T$_1$=1.02 ms.
 ii) The class is at resonance with another class : T$_1$=0.28 ms.
-iii) The class is at resonance with the three other classes : T$_1$=0.15 ms.
-  }\label{T_1}
-\end{figure}
+iii) The class is at resonance with the three other classes : T$_1$=0.15 ms.}
+  \label{CR_deposited}
+\end{figure*}
 
@@
 They are supplied by the company Adamas, which produces diamonds with a concentration of NV centers in the 3-4 ppm range.
-We operate with a Paul trap that is similar to the one used in \cite{delordPRL} with particles are stably trapped at the bottleneck region, where both the electric field gradient and anisotropy are stronger.
+We operate with a Paul trap that is similar to the one used in \cite{delordPRL} with particles stably trapped at the bottleneck region, where both the electric field gradient and anisotropy are stronger.
 The microwave is applied directly on the trapping electrode, which provides an efficient means to excite the spins.
@@
 Let us first show that dipolar interaction in our diamonds is effective by studying cross-relaxation with diamonds that are simply attached to the trap.
-In diamonds that contain more than 1 ppm of nitrogen-vacancy centers cross relaxation effects have been observed by many groups \cite{van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, mrozek_longitudinal_2015, jarmola_temperature-_2012, choi_depolarization_2017}.
-It manifests itself as a change in the longitudinal relaxation rate of NV centers, which can then translate as a reduction of the overall photo-luminescence.
+Fig.
+\ref{CR_deposited} a) shows the interaction between two NV centers, one of which (NV$_2$) has a considerably reduced relaxation time T$_1$, almost negating the optical polarization by the laser.
+When considering the resonant dipole-dipole interaction with another center NV$_1$ (whose T$_1$ is phonon-limited), the centers will exchange excitation quanta through flip-flop process, resulting in a depolarization of NV$_1$.
+Generalizing, this means that a few fast-decaying NV can depolarize an ensemble of "normal" NV centers through dipolar interaction, reducing the average T$_1$ of the ensemble to a few to a few hundreds of micro-seconds \cite{Jarmola} and lowering the total photoluminescence.
+This effect has been observed by many groups\cite{van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, mrozek_longitudinal_2015, jarmola_temperature-_2012, choi_depolarization_2017}.
 
-This effect has been routinely observed in bulk materials, the origin of which was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}.
-The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{Manson}) have a reduced longitudinal spin lifetime $T_1$.
-When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx 5 ms$), the latter can drop
-to a few hundreds of micro-seconds \cite{Jarmola} when the NV concentration is large enough (typically in the ppm range)
-This process has recently revived interest recently because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}. 
+This effect has been routinely observed in bulk materials with large enough  NV concentration, typically in the ppm range.
+The origin of it (the fast-decaying NV centers)  was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}.
+The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{manson_nv_2018}) have a largely reduced longitudinal spin lifetime $T_1$.
+This process has recently revived interest because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}.
+
 %In our spin-mechanics experiment, we use micro-diamond particules subjected to a green laser field and a magnetic field.
@@
 In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias} \approx$ 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet.
-Figure \ref{CR_deposited}-a) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
+Figure \ref{CR_deposited}-b) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
 For these measurement we use a 1 mW green laser excitation which yields a repolarisation rate of about 100 $\mu$s. 
@@
 We attribute these to cross-relaxation that takes place when two orientation meet.
-To be sure, we recorded Electronic spin resonance spectra for different values which allows us to measure the total magnetic field felt by the spins, and therefore to extrapolate the transition frequencies of the four classes of NV for any applied $\bf B_{\rm em}$.
+To be sure, we recorded Electronic spin resonance spectra for different field values in Fig.
+\ref{CR_deposited}-c) which allows us to measure the total magnetic field felt by the spins, and therefore to extrapolate the transition frequencies of the four classes of NV for any applied $\bf B_{\rm em}$.
 There is indeed a strong correspondence between the PL dips and the degeneracy conditions. 
-(See SI for details on planes that cross etc.). 
-We tentatively attribute those to CR. 
+This is expected for NV ensembles presenting a NV-NV depolarization process, since when two orientations or more are at resonance, the effective density of NV being able to exchange spin quanta increases, resulting in a stronger depolarization process.
 %It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
@@
 %Indeed, surface effects may start to play a role and change quantitatively.
+
 To confirm our interpretation, and to extract the spin-torque dynamics in the mechanical experiment, we measure the longitudinal relaxation time ($T_1$ time) of the NV spins under different degeneracy conditions.
-
-The measurement consists in applying a green laser to polarize the NV centers and to measure the photoluminescence at a later time. 
+%Clément : je comprend pas la partie "to extract the spin-torque dynamics"
+The measurement presented in Fig \ref{CR_deposited}-d) consists in applying a green laser to polarize the NV centers and to measure the photoluminescence at a later time. 
 Such a measurement turns out to be hidden by recharging of NV centers in the dark, as has been observed and analyzed by several groups \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
@@
 Subtracting the two resulting signals eliminates charge dynamics from the spin decay.
-The results of the measurements are shown in Fig. \ref{T_1} for different degeneracies. 
-It can be seen that $T_1$ is greatly shortened, from X to Y when approaching degeneracy, confirming this interpretation.
+The results are shown for different degeneracy configuration, starting with a case without degeneracy where we extract a $T_1=1.08$ ms, already shorter than the expected phonon limited lifetime, and this lifetime is even further reduced to 0.28 ms (resp.
+0.15 ms) when two (resp.
+four) orientations are brought to resonance.
 
@@
 For this, let us consider first the torque coming from a single spin due to the transverse component of the magnetic field when the NV is in the ground state.
-The hamiltonian for one NV orientation with quantization axis $z$ reads 
+The Hamiltonian for one NV orientation with quantization axis $z$ reads 
 \begin{equation}\hat{H}_{\rm NV}=\hbar D \hat{S}_z^2+ \hbar \gamma_e \bf B  \cdot \bf\hat S.
 \end{equation}
-Under the condition $\gamma B \ll D$, $H_{B}= \gamma_e B  ( S_x \cos\theta + S_z \sin\theta)$ can be treated as a perturbation to the anisotropic part of the hamiltonian. 
+Under the condition $\gamma B \ll D$, $H_{B}= \gamma_e B  ( S_x \cos\theta + S_z \sin\theta)$ can be treated as a perturbation to the anisotropic part of the Hamiltonian. 
 The perturbed energy $\epsilon_g$ of $\ket{0}$ due to the transverse B field is then
@@
 \end{equation}
-One can deduce the total torque when N spins of a given orientation are polarised in the ground state to be 
+One can deduce the total torque when N spins of a given orientation are polarized in the ground state to be 
 \begin{equation}\tau_g=-N \frac{\partial \epsilon_0}{\partial \theta}=\hbar N  \frac{(\gamma_e B)^2}{D} \sin 2\theta.\end{equation}
@@
 We checked numerically that at these magnetic field values and angle, the NV are still polarized mostly in the ground state. 
-Note that the torque is zero at $\theta=mod (\pi/2)$ where the eigenstates of $S_z$ are not mixed. The hamiltonian ground state is the eigenstate $| S=1,m_s=0\rangle$, which is is non magnetic. 
+Note that the torque is zero at $\theta=mod (\pi/2)$ where the eigenstates of $S_z$ are not mixed. The Hamiltonian ground state is the eigenstate $| S=1,m_s=0\rangle$, which is is non magnetic. 
 %Note also that the torque tends to zero if  tends to zero at any angle $\theta$.
@@
 Fig. \ref{principle}-ii) shows the same situation, with a magnetic field that is tuned so that two NV classes are co-resonant. 
-When the re-polarisation rate due to the green laser is smaller than the cross-relaxation rate, the magnetization coming from these two classes (indicated by a circle) tends to cancel.
+When the re-polarization rate due to the green laser is smaller than the cross-relaxation rate, the magnetization coming from these two classes (indicated by a circle) tends to cancel.
 The population in the three NV eigenstates will indeed equilibrate, giving a smaller magnetization.
@@
 Three ESR have been taken for three different $\bf B_{\rm em}$ amplitudes. 
-At 10 and 25 G, one can observe 4 pics in the spectrum, that are demonstrate spin induced torque on the diamond from the 4 classes of NVs.
+At 10 and 25 G, one can observe 4 peaks in the spectrum that demonstrate spin induced torque on the diamond from the 4 classes of NVs.
 At 17 G however, two classes merge at 2.75 GHz. This is where we expect CR.

### 2020-10-28 Mes papiers/Méca CR/Meca_cross_SI.tex
confinement?

@@
 
-\section*{NV$^-$ center Theory}
+\section{NV$^-$ center Theory}
 \subsection{NV spin Hamiltonian}
@@
   \end{equation*} 
-Where $D = (2\pi) 2.87$ GHz is the crystal field splitting and $\gamma_e = 28 $GHz/T is the electron gyromagnetic ratio.
-We neglected contributions of the strain and local electric field since we are working with magnetic fields of the order of 100G, as well as the hyper-fine interaction with $^{14}$N since we are working with ensembles with a typical inhomogeneous broadening $\frac{1}{T_2^*} = (2\pi) 5$ MHz.
+Where $D = (2\pi) 2.87$ GHz is the crystal field splitting originating from a spin-spin interaction, and $\gamma_e = 28 $GHz/T is the electron gyromagnetic ratio.
+We neglected contributions of the strain and local electric field since we are working with magnetic fields of the order of 100G, as well as the hyper-fine interaction with $^{14}$N since we are working with ensembles with a typical inhomogeneous broadening $\frac{1}{T_2^*} \approx (2\pi) 5$ MHz.
 
@@
   \centering \scalebox{0.45}{\includegraphics{SI_Cristallo}}
-  \caption{Representation of the magnetic field in the diamond crystalline basis.}
+  \caption{a) Representation of the four crystalline axes of the diamond b) Representation of the magnetic field in the diamond crystalline basis c) Representation in the $(\theta , \phi)$ basis of the crystalline planes ([110]$_\perp$ family of planes in dashed lines, [100]$_\perp$ in plain lines), [100] direction as a red circles, [111] direction as a green squares and the path of the magnetic field in the experiment of Fig.
+3 of the main text in blue dots.}
 	\label{cristallo}
@@
 
-\section*{Experimental details}
+\section{Experimental details}
 
@@
 
-The experimental setup illustrated in Fig.\ref{Optics} is very similar to the one used in \citep{DelordPRL} with the addition of a permanent magnet and an electromagnetic (EM) coil in order to perform magnetic filed scans.
+The experimental setup illustrated in Fig.\ref{Optics} is very similar to the one used in \citep{DelordPRL} with the addition of a permanent magnet and an electromagnetic (EM) coil in order to perform magnetic field scans.
 The diamond sample is typically illuminated with 1mW of 532 nm laser light, focused by a NA = 0.5 objective.
@@
 The photo-luminescence (PL) is collected by the objective, separated form the excitation light using a dichroic mirror (DM) and a 532nm notch filter (NF), and detected using a multimode-fibered single-photon avalanche photo-detector (APD) (SPCM-AQRH-15 from Perkin Elmer).
-Typically, we detect PL photons at a rate of 1MHz.
+Typically, we detect PL photons at a rate of 1 MHz.
 
-The Paul trap is a pseudo-ring as can be seen in \citep{DelordPhD} which acts both as trap through the high voltage (HV) and as a microwave (MW) antenna.
+The Paul trap is a pseudo-ring with a diameter of approximately 200 $\mu$m, as can be seen in \citep{DelordPhD}. It acts both as trap through the high voltage (HV) and as a microwave (MW) antenna.
 
-The magnetic field generated by the (homemade) EM coil is controlled by a frequency generator (HP 33120A) with a triangle signal in order to perform magnetic field ramps.
-For scans of less than 100 G, no amplification of the signal is required.
+The magnetic field generated by the (homemade) EM coil is controlled by a programmable power supply (Rhode \& Shwartz NGE 103) performing current ramps.
 
@@
 \begin{figure}[!ht]
-  \centering \scalebox{1}{\includegraphics{T1_soustraction_shema}}
-  \caption{T$_1$ measurement protocol for a single class in an ensemble of NV centers}
+  \centering \scalebox{1.2}{\includegraphics{T1_soustraction_shema}}
+  \caption{T$_1$ measurement protocol for a single class in an ensemble of NV centers. Green bars represent laser excitation, red bar represent microwave $\pi$ pulse for the selected class}
 	\label{T1_protocol}
@@
 \ref{T1_protocol} .
-The spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$.
+In the first sequence ($S_1$), the spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$.
 The spin state is finally read out with a 10 $\mu$s laser pulse, shorter than the polarization time of the spins.
 
-The sequence is repeated a second time, adding a resonant microwave $\pi$ pulse on a transition of one of the four classes of NV$^-$, exchanging the population of the $m_s=\ket{0}$ state with one of the $m_s=\ket{\pm 1}$ state.
+The second sequence ($S_2$) uses the same parameters (polarization time, dark time and readout time) than the first sequence but adds a resonant microwave $\pi$ pulse on a transition of one of the four classes of NV$^-$ right before the readout pulse, exchanging the population of the $m_s=\ket{0}$ state with one of the $m_s=\ket{\pm 1}$ state.
 
-The two sequences are then subtracted, allowing us to measure the spin state population of a single class, and removing unwanted contributions to the photoluminescence, such as charge state transfer in the dark.
+By looking a the subtraction of both signal $S_1-S_2$, we can then measure the spin state population of a single class, and remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark.
 
@@
 
-To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonace (ODMR) spectra....
+To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonance (ODMR) spectra....
 
@@
 
-\section*{Cross-relaxation detection for another type of degeneracy}
+\section{Cross-relaxation detection for another type of degeneracy}
 
-\section*{Simulation details}
+\begin{figure}[!ht]
+  \centering \scalebox{0.32}{\includegraphics{CR_meca_22}}
+  \caption{Mechanical detection of a degeneracy in a [100]$_\perp$ plane.
+a) Theoretical frequencies for the four $\ket{0} \to \ket{-1}$ transitions in dashed lines and mechanical ESR spectra vertically.
+b) Photoluminescence of the NV centers as a function of the scanning magnetic field (blue crosses) and gaussian fit (orange line).
+c) Simulated population in $\ket{m_s=0}$ for the stationary state with (plain) or without (dashed) taking into account the decrease of the T$_1$ induced by the cross-relaxations.
+  d) Reflected signal off the diamond as a function of the scanning magnetic field (blue crosses) and double gaussian fit (orange line) e) Simulated torque applied by the spins on the diamond, with (plain) or without(dashed) taking into account the cross-relaxations.}
+	\label{CR_22}
+\end{figure}
+
+Similarly to Fig.
+3 of the main text, we managed to mechanically detect another type of NV-NV resonance.
+Fig.
+\ref{CR_22}a) shows the theoretical frequencies of the $\ket{0} \to \ket{-1}$ transitions for all four classes of NV and ESR spectra measured through the reflected laser light off the diamond for various magnetic field values.
+Unlike The experiment of the main text, this time all four classes of NV are at resonance with another class at B=23 G, telling us that we are crossing a [100]$_\perp$ plane instead of a [110]$_\perp$ one.
+
+Fig.
+\ref{CR_22}b) shows the recorded photoluminescence of the NV centers during the magnetic field scan.
+As previously, we see a drop in the luminosity when the resonance occurs, that we fitted with a gaussian.
+The drop is slightly more pronounced in this case since all classes are depolarized, instead of only two in the previous experiment.
+This behavior is predicted by the simulation of the $\ket{m_s=0}$ population in the stationary state in Fig.
+\ref{CR_22}c).
+
+Fig.
+\ref{CR_22}d) shows the signal of the reflected laser off the diamond, allowing us to measure angular displacement.
+This time we see a clear difference with the experiment of the main text, instead of a single drop centered on the resonance, we see two bumps on both sides of the resonance, fitted with two gaussians.
+This is in accordance with the simulated torque in Fig.
+\ref{CR_22}e) where we can see a dispersive profile with an almost zero torque at the resonance.
+The reason we do not observe a change of sign in the experiment (with two positive bumps instead of a positive and a negative one) might be because of the strong non-linearity of our detection : if the signal initially correspond to a dark spot of the speckle, then a change in the motion of the diamond can only result in an increased signal.
+We also note a drop both in the predicted torque amplitude, and in the signal-to-noise ratio  of the angular measurement compared to the single degeneracy case, implying a weaker effect.
+
+Our physical interpretation of the angular signal we observe, as well as the dispersive profile of the torque is that in this case, for symmetry reason, the magnetic torque generated by the four classes of NV is not modified when we are exactly at resonance, since all four classes are affected identically.
+What happens when you are near, but not at, resonance is that the classes will not be identically depolarized : looking at Fig.
+\ref{CR_22}a), we can see that the two classes of higher frequency are always slightly closer to each other than the two classes of lower frequency (we can see the slope of the frequencies being a bit smaller for the two upper classes), resulting in a bit more depolarization for these two classes, except when they are exactly at resonance.
+This interpretation accounts both for the overall shape of the responses, and the comparatively weakness of the phenomenon.
+
+\section{Simulation details}
 In this part we will discuss the method used to simulate the average torque as well as the $\ket{m_s=0}$ population of the stationary state.
-The numerical solving of the master equations was performed thanks to Quantum Toolbox in Python (QuTiP) \citep{qutip1} \citep{qutip2}.
+The numerical solving of the master equations was performed thanks to the Quantum Toolbox in Python (QuTiP) \citep{qutip1} \citep{qutip2}.
 
@@
 
-We also introduce the usual T1 jump operators...
+We also introduce the usual T$_1$ jump operators $\mathcal{L}_i^j= \frac{1}{T_1}\ket{i}\bra{j}$ where $\ket{i,j}$=$\ket{0, \pm 1}$
 
-In order to describe the T1 modification induced by the cross-relaxations, we use a phenomenological model ...
+In order to describe the T$_1$ modification induced by the cross-relaxations, we use a phenomenological model where each class as its own T$_1^i$ (i from 1 to 4) that depends on the energy levels of the other classes with the formula :
+\begin{equation*}
+\frac{1}{T_1^i}=\frac{1}{T_1^0}+\sum_{j \neq i} \frac{1}{T_1^{dd}}e^{-\frac{(\nu_i-\nu_j)^2}{2(\sigma^{dd})^2}},
+\end{equation*}
+where $\nu_i$ and $\nu_j$ are the transition frequencies of the classes i and j (we are arbitrarily considering the $\ket{0} \to \ket{-1}$ transition here, since the resonance condition is the same for both transitions at the magnetic fields we are working at.
+This would not always be true for magnetic fields greater than 592 G\citep{van_oort_cross-relaxation_1989}).
 
-According to previous measurements \citep{choi_depolarization_2017}, only the $\ket{0}\bra{\pm1}$ (corresponding to a single quantum exchange in the dipole-dipole interaction) operators are modified by the cross-relaxations.
+$\sigma^{dd} = 6$ MHz is the the dipolar interaction width, which we measured to be similar to the inhomogeneous broadening.
 
+$T_1^0=1.03$ ms and $T_1^{dd}=0.38$ ms were chosen to match the T$_1$ measured in Fig.
+1 of the main text.
+We only focus on the T$_1$ without degeneracy and the one with a single degeneracy since our experiments will not have more than two classes at resonance at once.
+Our model is probably not suited to deal with triple our quadruple resonances.
+
+Finally, according to previous measurements \citep{choi_depolarization_2017}, only the $\ket{0}\bra{\pm1}$ and $\ket{\pm1}\bra{0}$ (corresponding to a single quantum exchange in the dipole-dipole interaction) operators are modified by the cross-relaxations.
+
+With this model in consideration, we can numerically solve the master equation and get the density matrix in the stationary state $\rho_s$.
+With $\rho_s$ we can directly obtain the $\ket{m_s=0}$ population, corresponding experimentally to the photoluminescence.
+
+As for the torque, we use a semi-classical formula :
+\begin{equation*}
+\mathbf \Gamma = N_0 \gamma_e <\hat{\mathbf S}> \times \mathbf B,
+\end{equation*}
+where $N_0 \approx 10^9$ is an estimate of the number of spins in our sample based on the average size and NV density of our diamonds, $\gamma_e$ is the gyromagnetic ratio of the electron and $<\hat{\mathbf S}> =\mathrm{Tr}(\rho_s \mathbf{\hat S})$ is the averaged spin vector in the stationary sate, averaged again on the four possible orientations of NV.
+This formula assumes that the spin dynamics is faster than the dynamics of the motion of the diamond, which is the case in our experiments.
+
+In our plots in Fig. X and Fig. 3 of the main ext, we only represent one spatial component (e.g. $\Gamma_x$) of the torque, but the three components behave similarly.
 \bibliography{trap}

### 2020-10-28 abstract/IQFA 2020/Abstract_Pellet-Mary.tex
confinement?

\documentclass[aps,twocolumn,showpacs]{revtex4-1}

\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{theorem}
\usepackage{amsmath}
\usepackage{times}
\usepackage{color}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\ifpdf
\usepackage{epstopdf}   
\usepackage{url}
\fi

\begin{document}
\title{Novel detection of spin interactions in diamond}

\author{Clément Pellet-Mary$^1$, Maxime Perdriat$^1$, Paul Huillery$^1$, and Gabriel Hétet$^{1}$}
\email{gabriel.hetet@phys.ens.fr}
\affiliation{$^1$Laboratoire de physique de l'ENS, ENS Paris, France\\
}

\begin{abstract}
\normalsize
NV$^-$ centers in diamond are a widely used quantum system, both for applications and fundamental research, thanks to their good coherence properties, long lifetimes and most importantly their ability to be optically polarized.
We have recently found new ways to use these properties of NV centers in order to study the dipolar interactions between ensemble of spins in diamond, including NV centers as well as other spin defects.

On one part, we have observed cross-relaxations (CR) between NV centers and other defects (see Fig. 1), namely VH- \cite{VH}, War1 (first defect found by EPR at the university of Warwick, still chemically unknown) and $^{13}$C First-shell modified NV. None of these defects had been observed through CR before. This is a first step toward hyper-polarization of new spin defects in diamond.

On another part, we have used the mechanical detection of a levitating diamond to measure the torque applied by the NV centers on the diamond when two classes of NV with different orientation are put to resonance(see Fig. 2). We have linked this effect to the dipolar-mediated modification of the T$_1$ of NV centers happening with dense ensembles.\cite{Lukin}. This paves the way toward the mechanical detection of other spin impurities in diamond, as well as the observation of the Einstein-de Haas effect with paramagnetic systems. \cite{Meriles}

\includegraphics[scale=.25]{CR_total}
Figure 1 : i) Photoluminescence change while scanning the magnetic field along the crystalline [100] direction. ii) Energy levels for various species : Plain : NV centers, dotted : NV centers with hyper-fine coupling of a first-shell $^{13}$C, dashed : VH$^-$, dash-dotted : War1. The predicted cross relaxation with NV centers are at 18-24 G for first-shell $^{13}$C, 54 G for VH$^-$ and 123 G for War1

\includegraphics[scale=.5]{CRmeca}
Figure 2 : i) ESR spectra while scanning B$_{em}$ with some bias field applied. ii)Angular position of the diamond while scanning B$_{em}$ without a microwave. The rotation of the diamond at 17 G corresponds to the resonance between two classes of NV
\end{abstract}

\maketitle

\begin{thebibliography}{}

\bibitem{VH} Glover, Claire, et al. "Hydrogen incorporation in diamond: The nitrogen-vacancy-hydrogen complex." Physical review letters 90.18 (2003): 185507.

\bibitem{Lukin} Choi, Joonhee, et al. "Depolarization dynamics in a strongly interacting solid-state spin ensemble." Physical review letters 118.9 (2017): 093601.

\bibitem{Meriles} Zangara, Pablo R., et al. "Mechanical rotation via optical pumping of paramagnetic impurities." Physical Review B 100.23 (2019): 235410.

\end{thebibliography}
\end{document}

### 2020-10-28 data/20200806/2-2/traitement.py

@@ -125,6 +125,39 @@ def ask_name():
 	fname,filters=QFileDialog.getOpenFileName()	
 	return fname
 
+def ESR_n_pics(x,y,cs,width=8,ss=None,amp=None) :
+	if not ss :
+		ss=y[0]
+	if not amp :
+		if max(y)-ss > ss-min(y) :
+			amp=max(y)-ss
+		else :
+			amp=min(y)-ss
+	n=len(cs)
+	widths=np.ones(n)*width
+	amps=np.ones(n)*amp
+	p0=[ss]
+	for c in cs:
+		p0+=[c]
+	for w in widths:
+		p0+=[w]
+	for a in amps:
+		p0+=[a]
+	def f(x,*params):
+		ss=params[0]
+		n=(len(params)-1)//3
+		y=ss
+		for i in range(n):
+			c=params[1+i]
+			width=params[1+n+i]
+			amp=params[1+2*n+i]
+			y+=amp*np.exp(-((x-c)**2/(2*width**2)))
+		return(y)
+	popt, pcov = curve_fit(f, x, y, p0)
+	variables=(x,)
+	for p in popt :
+		variables+=(p,)
+	return(variables[1:],f(*variables))
 
 xmin=30
 xmax=200
@@ -137,22 +170,30 @@ y_tot=np.zeros(xmax-xmin)
 # 	y_tot+=y[xmin:xmax]
 
 
-# pool=['0-0.5V_PLdown','0-0.5V_PLdown-c','0-0.5V_PLdown-e-14uW','0-0.5V_PLdown-f-18.5uW']
-# for fname in pool :
-# 	x,y=extract_data(fname+'.txt',ycol=1)
-# 	y=y/max(y)
-# 	y_tot+=y[xmin:xmax]
+pool=['0-0.5V_PLdown','0-0.5V_PLdown-c','0-0.5V_PLdown-e-14uW','0-0.5V_PLdown-f-18.5uW','0-0.5V_PLdown-d-14uW_was9']
+for fname in pool :
+	x,y=extract_data(fname+'.txt',ycol=3)
+	y=y/max(y)
+	y_tot+=y[xmin:xmax]
+	x=x[xmin:xmax]
 
-fname='0-0.5V_PLdown'
-x,y=extract_data(fname+'.txt',ycol=1)
+y_tot=y_tot/max(y_tot)
[…]

### 2020-10-28 python/TF filtering/traitement.py

@@ -1,0 +1,230 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit
+from scipy import fftpack
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def quad_fit(x,y) :
+	A=np.vstack([x**2,x,np.ones(len(x))]).T
+	a,b,c = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([a,b,c],a*x**2+b*x+c)
+
+def fit_ordre_4(x,y) :
+	A=np.vstack([x**4,x**3,x**2,x,np.ones(len(x))]).T
+	a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([a,b,c,d,e],a*x**4+b*x**3+c*x**2+d*x+e)
+
+def fit_ordre_6(x,y) :
+	A=np.vstack([x**6,x**5,x**4,x**3,x**2,x,np.ones(len(x))]).T
+	w,z,a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([w,z,a,b,c,d,e],w*x**6+z*x**5+a*x**4+b*x**3+c*x**2+d*x+e)
+
+
+def gauss_fit(x,y,Amp=None,x0=None,sigma=None,ss=0) :
+	if not ss :
+		ss=y[0]
+	if not Amp :
+		if max(y)-ss > ss-min(y) :
+			Amp=max(y)-ss
+		else :
+			Amp=min(y)-ss
+	if not x0 :
+		x0=x[int(len(x)/2)]
+	if not sigma :
+		sigma=x[int(len(x)/5)]-x[0]
+	def f(x,Amp,x0,sigma,ss) :
[…]

### 2020-10-28 python/simu Hamiltonien NV/template.py

@@ -177,14 +177,16 @@ def transitions(val,vec,Sx,rho_0,show=True):
 
 def NV_simple(orientation='111'):
 
+	plt.figure(num=1,figsize=(11,7),dpi=80)
 	# B=[5000,0,0]
 	# H=Hamiltonian_0(B,classe=1,E=3)
 	# val,vec=egvect(H)
 	# show_vpropres(val,vec,bnamez)
 	# transitions(val,vec,Sx,rho_0)
-	D_VH=2466
+	D_VH=2694 #+/-5 , Cruddace a 2706 +/- 30
+	D_War1=2466 #la meme valeur que Cruddace, lui il a +/- 60
 	if orientation=='111' :
-		amps=np.linspace(0,200,201)
+		amps=np.linspace(0,400,201)
 		transi_NV=[]
 		for amp in amps :
 			# B=[amp,0,0]
@@ -258,7 +260,7 @@ def NV_simple(orientation='111'):
 		ax.set_ylim(ylim)
 		plt.show()
 	if orientation=='100' :
-		amps=np.linspace(0,200,100)
+		amps=np.linspace(0,145.8369,100)
 		transi_NV=[]
 		for amp in amps :
 			B=[amp,0,0]
@@ -268,8 +270,8 @@ def NV_simple(orientation='111'):
 		transi_NV=np.array(transi_NV)
 		ax=plt.gca()
 		color = next(ax._get_lines.prop_cycler)['color']
-		plt.plot(amps,transi_NV[:,0],color=color,label='NV-')
-		plt.plot(amps,transi_NV[:,1],color=color)
+		plt.plot(amps,transi_NV[:,0],color=color,label='NV-',lw=5)
+		plt.plot(amps,transi_NV[:,1],color=color,lw=5)
 		transi_NV=[]
 		for amp in amps :
 			B=[amp,0,0]
@@ -277,21 +279,40 @@ def NV_simple(orientation='111'):
 			val,vec=egvect(H)
 			transi_NV+=[[val[2]-val[0],val[1]-val[0]]]
 		transi_NV=np.array(transi_NV)
-		ax=plt.gca()
 		color = next(ax._get_lines.prop_cycler)['color']
-		plt.plot(amps,transi_NV[:,0],color=color,label='VH-')
-		plt.plot(amps,transi_NV[:,1],color=color)
-		def zero():
[…]

### 2020-10-29 abstract/IQFA 2020/Abstract_Pellet-Mary.tex
tata

@@
 \begin{document}
-\title{Novel detection of spin interactions in diamond}
+\title{Mechanical and relaxation-based detection of dipolar-interactions between spins in diamond}
 
-\author{Clément Pellet-Mary$^1$, Maxime Perdriat$^1$, Paul Huillery$^1$, and Gabriel Hétet$^{1}$}
-\email{gabriel.hetet@phys.ens.fr}
-\affiliation{$^1$Laboratoire de physique de l'ENS, ENS Paris, France\\
+\author{Clément Pellet-Mary$^1$, Maxime Perdriat$^1$, Paul Huillery$^1$, and Gabriel Hétet$^{1*}$}
+
+\affiliation{$^1$Laboratoire de Physique de l’Ecole Normale Supérieure, ENS, Université PSL, CNRS, Sorbonne Université, Université de Paris, F-75005 Paris, France.\\$^*$ gabriel.hetet@phys.ens.fr
 }
@@
 \normalsize
-NV$^-$ centers in diamond are a widely used quantum system, both for applications and fundamental research, thanks to their good coherence properties, long lifetimes and most importantly their ability to be optically polarized.
+The negatively charged nitrogen vacancy center (NV$^-$) is a well established quantum system, employed both for applications and fundamental research, thanks to its good coherence properties, long lifetimes and most importantly their ability to be optically polarized.
 We have recently found new ways to use these properties of NV centers in order to study the dipolar interactions between ensemble of spins in diamond, including NV centers as well as other spin defects.
 
-On one part, we have observed cross-relaxations (CR) between NV centers and other defects (see Fig.
-1), namely VH- \cite{VH}, War1 (first defect found by EPR at the university of Warwick, still chemically unknown) and $^{13}$C First-shell modified NV.
+On the one hand, we have observed cross-relaxations (CR) between NV centers and other defects (see Fig.
+1), namely VH$^-$ \cite{VH}, War1 (first defect found by EPR at the university of Warwick, still chemically unknown) and $^{13}$C First-shell modified NV.
 None of these defects had been observed through CR before.
@@
 
-On another part, we have used the mechanical detection of a levitating diamond to measure the torque applied by the NV centers on the diamond when two classes of NV with different orientation are put to resonance(see Fig.
+On the other hand, we have used the mechanical detection of a levitating diamond \cite{nature} to measure the torque applied by the NV centers on the diamond when two classes of NV with different orientation are brought into resonance(see Fig.
 2).
-We have linked this effect to the dipolar-mediated modification of the T$_1$ of NV centers happening with dense ensembles.\cite{Lukin}.
+This effect is due to the dipolar-mediated modification of the T$_1$ of NV centers happening with dense ensembles.\cite{Lukin}.
 This paves the way toward the mechanical detection of other spin impurities in diamond, as well as the observation of the Einstein-de Haas effect with paramagnetic systems.
@@
 
+\begin{center}
 \includegraphics[scale=.25]{CR_total}
-Figure 1 : i) Photoluminescence change while scanning the magnetic field along the crystalline [100] direction.
-ii) Energy levels for various species : Plain : NV centers, dotted : NV centers with hyper-fine coupling of a first-shell $^{13}$C, dashed : VH$^-$, dash-dotted : War1.
-The predicted cross relaxation with NV centers are at 18-24 G for first-shell $^{13}$C, 54 G for VH$^-$ and 123 G for War1
+\end{center}
 
-\includegraphics[scale=.5]{CRmeca}
-Figure 2 : i) ESR spectra while scanning B$_{em}$ with some bias field applied.
-ii)Angular position of the diamond while scanning B$_{em}$ without a microwave.
-The rotation of the diamond at 17 G corresponds to the resonance between two classes of NV
+Figure 1 : Photoluminescence change while scanning the magnetic field along the crystalline [100] direction. Three dips are observed for the CR with $^{13}$C-NV, VH$^-$ and War1.
+
+\begin{center}
+\includegraphics[scale=.25]{CRmeca}
+\end{center}
+
+Figure 2 : Angular position of the diamond while scanning B$_{\mathrm{Scan}}$ without a microwave. The rotation of the diamond at 17 G corresponds to the resonance between two classes of NV
 \end{abstract}
@@
 
-\bibitem{VH} Glover, Claire, et al. "Hydrogen incorporation in diamond: The nitrogen-vacancy-hydrogen complex." Physical review letters 90.18 (2003): 185507.
+\bibitem{VH} Glover, C., et al. Physical review letters 90.18 (2003): 185507.
 
-\bibitem{Lukin} Choi, Joonhee, et al. "Depolarization dynamics in a strongly interacting solid-state spin ensemble." Physical review letters 118.9 (2017): 093601.
+\bibitem{nature} Delord, T., et al. Nature 580.7801 (2020): 56-59.
 
-\bibitem{Meriles} Zangara, Pablo R., et al. "Mechanical rotation via optical pumping of paramagnetic impurities." Physical Review B 100.23 (2019): 235410.
+\bibitem{Lukin} Choi, J., et al. Physical review letters 118.9 (2017): 093601.
+
+\bibitem{Meriles} Zangara, P.R., et al. Physical Review B 100.23 (2019): 235410.
 

### 2020-11-01 Mes papiers/Méca CR/Meca_cross_SI.tex
tata
demenagement

@@
 
+\section{Principle of the mechanical detection}
+
+\subsection{Torque sensing with a levitating diamond}
+
+The way we can experimentally measure torques applied on a levitating diamond is by looking at the displacement of equilibrium caused by them. Here is a short model detailing it :
+
+We model the trap as a pure harmonical potential, both for the center of mass and for the rotational degrees of freedom of the diamond (REF qq part), with trapping frequencies $\omega_T \approx 1$ kHz that we can experimentally measure.
+If we focus on a single rotational degree, we can write the torque generated by the trap as $\Gamma_T^\theta=-K(\theta-\theta_{eq})$, where $K=I \omega_t^2$ is the stiffness of the trap, $I$ being the moment of inertia of the diamond.
+
+The application of an external torque $\Gamma_{\mathrm{ext}}$ on the diamond will therefore shift the angular position at equilibrium in such a way that : 
+\begin{align*}
+-K(\theta-\theta_{\mathrm{eq}}) + \Gamma_{\mathrm{ext}} &= -K(\theta-\theta_{\mathrm{eq}}') \\
+\delta\theta = \theta_{\mathrm{eq}}'-\theta_{\mathrm{eq}} &= \frac{\Gamma_{\mathrm{ext}}}{K}=\frac{\Gamma_{\mathrm{ext}}}{I \omega_t^2}
+\end{align*}
+
+In our case, $\Gamma_{\mathrm{ext}}$ is the magnetic torque exerted by the NV$^-$ spins directly on the diamond, such that we can write $\Gamma_{\mathrm{ext}} = N_{NV} \langle \Gamma_{\mathrm{1 spin}} \rangle$  where $\langle \Gamma_{\mathrm{1 spin}} \rangle = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B \approx 10^{-28}$ Nm is the classical magnetic torque applied by one spin (see Section  \ref{Simu}).
+
+By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle [NV^-]}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-3} rad$$ where $[NV^-] \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 1$ kHz is the typical value of the trap frequency.
+
+It should be noted that the main uncertainty comes here from the diamond size, which along the other uncertainties can change the expected result by an order of magnitude.
+
+\subsection{Change of torque at the cross-relaxation}
+
 \section{Cross-relaxation detection for another type of degeneracy}
 
+%Faire 2 lignes avec sur la première ligne l'ESR et le trajet sur la carte du champ mag
 \begin{figure}[!ht]
@@
 Our physical interpretation of the angular signal we observe, as well as the dispersive profile of the torque is that in this case, for symmetry reason, the magnetic torque generated by the four classes of NV is not modified when we are exactly at resonance, since all four classes are affected identically.
-What happens when you are near, but not at, resonance is that the classes will not be identically depolarized : looking at Fig.
+What happens when being near, but not at, resonance is that the classes will not be identically depolarized : looking at Fig.
 \ref{CR_22}a), we can see that the two classes of higher frequency are always slightly closer to each other than the two classes of lower frequency (we can see the slope of the frequencies being a bit smaller for the two upper classes), resulting in a bit more depolarization for these two classes, except when they are exactly at resonance.
@@
 \section{Simulation details}
+\label{Simu}
 In this part we will discuss the method used to simulate the average torque as well as the $\ket{m_s=0}$ population of the stationary state.
@@
 \begin{equation*}
-\mathbf \Gamma = N_0 \gamma_e <\hat{\mathbf S}> \times \mathbf B,
+\mathbf \Gamma = N_0 \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B,
 \end{equation*}
-where $N_0 \approx 10^9$ is an estimate of the number of spins in our sample based on the average size and NV density of our diamonds, $\gamma_e$ is the gyromagnetic ratio of the electron and $<\hat{\mathbf S}> =\mathrm{Tr}(\rho_s \mathbf{\hat S})$ is the averaged spin vector in the stationary sate, averaged again on the four possible orientations of NV.
+where $N_0 \approx 10^9$ is an estimate of the number of spins in our sample based on the average size and NV density of our diamonds, $\gamma_e$ is the gyromagnetic ratio of the electron and $\langle\hat{\mathbf S}\rangle =\mathrm{Tr}(\rho_s \mathbf{\hat S})$ is the averaged spin vector in the stationary sate, averaged again on the four possible orientations of NV.
 This formula assumes that the spin dynamics is faster than the dynamics of the motion of the diamond, which is the case in our experiments.
 
-In our plots in Fig. X and Fig. 3 of the main ext, we only represent one spatial component (e.g. $\Gamma_x$) of the torque, but the three components behave similarly.
+In our plots in Fig. \ref{CR_22} and Fig. 3 of the main ext, we only represent one spatial component (e.g. $\Gamma_x$) of the torque, but the three components behave similarly.
 \bibliography{trap}

### 2020-11-01 théorie/fiches de lecture/Ashcroft-Mermin/fiche.tex
demenagement

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\chapter{Théorie de Drude}
  \section{Historique}
  Découverte des électrons par Thomson en 1897, Modèle de Drude à partir de 1900, découverte du noyau par Rutherford en 1909. (Avant y'avait toute cette histoire de flan au pruneau)
  \section{le modèle}
  C'est un modèle pour décrire le comportement des conducteurs (donc métaux) : Les électrons de valence sont libres et forment un gaz, les noyaux et les électrons de coeurs (les "ions") sont fixes. En ordre de grandeur, ça fait $\approx 10^{22} e^-/cm^3$ soit environ 10 x plus que la densité d'un gaz à 300K et $P_0$. En plus les électrons sont chargés, donc fortement en interaction, mais Drude il s'en ballec il dit que c'est un gaz dillué.

  En gros il a un modèle purement mécanico-collisionel, ou les électrons ne sont sensibles qu'au champ exterieur entre deux collisions. Il ne considère que les collisions $e^-$-ions, ce qui est assez justifié parce que la section efficace des ions est bcp plus grosse, et après chaque collision, l'$e^-$ est thermalisé avec la matrice cristalline : sa vitesse suit la loi de ditribution de Boltzmann et a une direction aléatoire. La fréquence des collisions est donnée par $1/\tau$ ou $\tau$ est une constante dans le modèle de base.
  \section{Résultats du modèle}
  \begin{enumerate}

  \item \textbf{La loi d'Ohm (locale)} : $\mathbf{j}=\sigma\mathbf{E}$ avec la conductivité $\sigma = \frac{ne^2\tau}{m}$ (La formule est pas vraiment prédictive vu qu'on connait pas $\tau$). Je refais pas la démo, en gros tu trouve que le modèle soumis à une Force constante revient à ajouter une force de frottement.

  \item \textbf{L'effet Hall (classique)} L'effet Hall c'est toujours la merde en vrai. L'idée est toute simple, tu as un champ mag transverse, donc tes $e^-$ tournent et se mettent en excès sur un bord, ce qui te créé un champ transverse (car surplus de charge d'un coté), donc un courant transverse. Sauf que en régime stattionaire tu peux pas avoir de courant transverse vu que tes électrons ont nulle part ou aller. Donc NTM il y un champ E transverse mais pas de courant (en gros c'est juste que la Force E et B se compensent pour la composante transverse). A la fin, ta resitivité est la même vu qu'il y a pas de courant transverse (magnétoresistance = 0, ce qui est pas toujours vrai en quantique), et tu as une constante qui apparait : $R_H=\frac{E_\perp}{j_\parallel B}=\frac{1}{ne}$ la constante de Hall, qui ne dépend a priori que de la densité électronique.

  \begin{tcolorbox}
  \textbf{Attention :} Le Ashcroft est un petit coquin, il utilise $H$ au lieu de $B$ en disant que c'est pareil pour un materiau pas magnétique (pourquoi pas). Sauf que ce trou du cul normalise $H=B/c$ plutot que $H=B/\mu_0$ (H est homogène à un champ électrique, ce qui il faut le reconnaitre est pratique pour les OEM) Donc méfiance, je ferai pas forcément gaffe partout.
 \end{tcolorbox}

 La constante de Hall, c'est un truc qui marche en gros en statique, à froid et à gros champ mag. Donc c'est quand meme un peu limité.
 Point plus intéressant, l'angle $\phi$ entre le courant et le champ Elec (angle de Hall) s'écrit $\tan \phi = \omega_c \tau$ ou $\omega_c=\frac{eB}{m}$ est la fréquence cyclotron (1/le temps pour qu'un  $e^-$ fasse un tour autour du champ mag). Donc $\phi$ te dit en gros combien de tours tu fais entre chaque collsion.

 \item \textbf{Courant alternatif} En courant alternatif il faut passer en Fourier : $\mathbf{j}(\omega)=\sigma(\omega)\mathbf{E}(\omega)$, et la conductivité prend une composante complexe $\sigma=\frac{\sigma_0}{1-i\omega\tau}$. Attention quand même, ce n'est vrai que dans la limite ou l'$e^-$ voit un champ constant entre deux collsions, cad la longueur d'onde de l'OEM est grande devant le libre parcours moyen (lpm) des $e^-$. En appliquant Maxwell comme des bourrins, on trouve l'équation d'onde :
 \begin{equation}
 \Delta E + \frac{\omega^2}{c^2} \epsilon(\omega) E =0
 \end{equation}
 $\epsilon(\omega)$ est la constante diélectrique dans le voc du livre et elle vaut $$ \epsilon(\omega)= \frac{4\pi i \sigma}{\omega} \approx 1-\frac{\omega_p^2}{\omega^2}$$ ce qui fait sortir la fréquence plasma $\omega_p=\frac{4\pi n e^2}{m}$. Pour $\omega < \omega_p$, les solutions sont exponentielles (donc décroissante parce qu'on est en physique), le métal absorbe les OEM avec une épaisseur de peau $l=c/\omega_p$. Pour $\omega > \omega_p$, les solutions sont des oscillations : le métal devient transparent aux OEM.
 Bon en vrai ça marche un peu pour les alcalins et c'est tout, et encore c'est de la chatte.

 Une conséquence tout de même c'est l'apparition des \textbf{plasmons} dans le modèle. En gros tu trouve que tu as des oscillations de densité locale de charge exactement à la fréquence $\omega_p$. De ce que j'en comprend c'est la fréquence de résonnance des $e^-$ elastiquement liés au cristal ($\omega_p$ est simplement lié à la constante de raideur de la liaison).

 \item \textbf{Conductivité thermique}

   \end{enumerate}
  \end{document}

### 2020-11-17 Mes papiers/Double quantum/template legacy/OSA-journal-template.tex
okou

\documentclass[9pt,twocolumn,twoside]{osajnl}
%% Please use 11pt if submitting to AOP
% \documentclass[11pt,twocolumn,twoside]{osajnl}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}

\journal{ol} % Choose journal (ao, aop, josaa, josab, ol, optica, pr)

% See template introduction for guidance on setting shortarticle option
\setboolean{shortarticle}{true}
% true = letter / tutorial
% false = research / review article
% (depending on journal).

\title{Legacy \LaTeX\ template for preparing an article for submission to OSA journals \emph{Applied Optics}, \emph{Advances in Optics and Photonics}, JOSA A, JOSA B, \emph{Optics Letters}, \emph{Optica}, and \emph{Photonics Research}}

\author[1,2,3]{Author One}
\author[2,*]{Author Two}
\author[1]{Author Three}

\affil[1]{Publications Department, The Optical Society (OSA), 2010 Massachusetts Avenue NW, Washington D.C., 20036}
\affil[2]{School of Science, University of Technology, 2000 J St. NW, Washington DC, 20036}
\affil[3]{School of Optics, University of Technology, 2000 J St. NW, Washington DC, 20036}

\affil[*]{Corresponding author: email@my-email.com}

%% To be edited by editor
% \dates{Compiled \today}

%\ociscodes{(140.3490) Lasers, distributed feedback; (060.2420) Fibers, polarization-maintaining;(060.3735) Fiber Bragg gratings.}

%% To be edited by editor
% \doi{\url{http://dx.doi.org/10.1364/XX.XX.XXXXXX}}

\begin{abstract}
This legacy template can be used to prepare a research article for submission to OSA’s journals \emph{Applied Optics}, \emph{Advances in Optics and Photonics}, JOSA A, JOSA B, \emph{Optics Letters}, and \emph{Optica}. \emph{Optics Letters} authors and authors of \emph{Optica} letters and memoranda may use this template for a precise length estimate. Use the shortarticle/true option for \emph{Optics Letters} and short \emph{Optica} articles. 
\end{abstract}

\setboolean{displaycopyright}{true}

\begin{document}

\maketitle

\section{Introduction}
This legacy template is designed to assist with creating an article to submit to \emph{Applied Optics}, \emph{Advances in Optics and Photonics}, JOSA A, JOSA B, \emph{Optics Letters} or \emph{Optica}. See the OSA's \href{http://www.opticsinfobase.org/submit/style/}{Style Guide} and \href{http://www.opticsinfobase.org/submit/templates/}{Manuscript Templates} pages for more details. Please select the appropriate journal abbreviation (ao, aop, josaa, josab, ol, optica) in the document preamble.

Use the shortarticle/false option for \emph{Applied Optics}, JOSA A, and JOSA B. Use the shortarticle/true option for \emph{Optics Letters}. For \emph{Advances in Optics and Photonics}, use the shortarticle/false option for Review Articles, and the shortarticle/true option for Tutorials.

If you have a question while using this template on {Overleaf}, please use the help menu (``?'') on the top bar to search for help or ask us a question using our \href{https://www.overleaf.com/contact}{contact form}.

\section{Examples of Article Components}
\label{sec:examples}

The sections below show examples of different article components.

\section{Figures and Tables}

It is not necessary to place figures and tables at the back of the manuscript. Figures and tables should be sized as they are to appear in the final article. Do not include a separate list of figure captions and table titles.

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

\begin{figure}[htbp]
\centering
\fbox{\includegraphics[width=\linewidth]{Carte_annotation}}
\caption{NV$^-$ photoluminescence as a function of a scanning magnetic field around the [100] crystalline directions with a fixed amplitude $|B|=115G$. The planes orthogonal to the [010],[001],[011], and [01$\bar 1$] directions are indicated in dashed lines.}
\label{map}
\end{figure}

\begin{figure}[htbp]
\centering
\fbox{\includegraphics[width=\linewidth]{Scans_fig4}}
\caption{Optical detection of cross-relaxations. \textbf{Top} NV$^-$ photoluminescence while scanning the magnetic field along the [100] crystalline direction (plain line) and polynomial fit to the fourth order (dashed line). \textbf{Bottom} Subtraction of the previous signal by the polynomial fit (circles), simulated CR magnetic field amplitude (dashed,vertical) and gaussian fits for the second and third dips (plain line)}
\label{scan}
\end{figure}

\begin{figure}[htbp]
\centering
\fbox{\includegraphics[width=\linewidth]{Theorie_CR}}
\caption{Simulated energy levels as a function of a magnetic field aligned along the [100] axis for the various spins considered. NV centers are in plain lines, VH$^-$ are in dashed line, WAR1 in dashdotted line and $^{13}$C-NV pair in dotted line. The amplitudes of the magnetic field where the energy level of the NV center crosses the one of another species are represented by vertical dashed lines}
\label{energy-levels}
\end{figure}

\subsection{Sample Table}

Table \ref{tab:shape-functions} shows an example table.

\begin{table}[htbp]
\centering
\caption{\bf Zero-field splitting parameter $D_z$ for the different spin-1 species}
\begin{tabular}{ccc}
\hline
$D_z$ estimation (MHz) & Cruddace's work\citep{cruddace2007magnetic} & Our work \\
\hline
NV$^-$ & 2872(7) & * \\
VH$^-$ & 2706(30) & 2694(5)  \\
WAR1 & 2466(60) & 2470(10) \\
\hline
\end{tabular}
  \label{tab:shape-functions}
\end{table}

\section{Sample Equation}

Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
\begin{equation}
S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
      = \frac{1}{n}\sum_{i}^{n} X_i
\label{eq:refname1}
\end{equation}
denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.

\subsection{Supplementary materials in OSA journals}
\citep{glover_hydrogen_2003} OSA journals allow authors to include supplementary materials as integral parts of a manuscript. Such materials are subject to peer-review procedures along with the rest of the paper and should be uploaded and described using OSA's Prism manuscript system. Please refer to the \href{https://www.opticsinfobase.org/submit/style/supplementary_materials.cfm}{Author Guidelines for Supplementary Materials in OSA Journals} for more detailed instructions on labeling supplementary materials and your manuscript. Visualizations, Data Files, Datasets, and Code must be associated with a figure, table, or equation, OR be referenced in the results section of the manuscript.

\textbf{Authors may also include Supplemental Documents} (PDF documents with expanded descriptions or methods) with the primary manuscript. At this time, supplemental PDF files are not accepted for partner titles, JOCN and Photonics Research. To reference the supplementary document, the statement ``See Supplement 1 for supporting content.'' should appear at the bottom of the manuscript (above the References heading). Please note that to create text color for supplementary materials links, use of the command \\
\verb|\textcolor{urlblue}{Visualization 1}| is preferred to using the command\\
\verb|\url{Visualization 1}|.

\subsection{Sample Dataset Citation}

1. M. Partridge, "Spectra evolution during coating," figshare (2014) [retrieved 13 May 2015], http://dx.doi.org/10.6084/m9.figshare.1004612.

\subsection{Sample Code Citation}

2. C. Rivers, "Epipy: Python tools for epidemiology" (Figshare, 2014) [retrieved 13 May 2015], http://dx.doi.org/10.6084/m9.figshare.1005064.

\section{Funding}
Content in the funding section will be generated entirely from details submitted to Prism. Authors may add placeholder text in the manuscript to assess length, but any text added to this section in the manuscript will be replaced during production and will display official funder names along with any grant numbers provided. If additional details about a funder are required, they may be added to the Acknowledgments, even if this duplicates information in the funding section. See the example below in Acknowledgements.

\section{Acknowledgments}
Acknowledgments should be included at the end of the document. The section title should not follow the numbering scheme of the body of the paper. Additional information crediting individuals who contributed to the work being reported, clarifying who received funding from a particular source, or other information that does not fit the criteria for the funding block may also be included; for example, ``K. Flockhart thanks the National Science Foundation for help identifying collaborators for this work.''

\section{Disclosures}

Disclosures should be listed in a separate section at the end of the manuscript. List the Disclosures codes identified on OSA's \href{http://www.osapublishing.org/submit/review/conflicts-interest-policy.cfm}{Conflict of Interest policy page}. If there are no disclosures, then list ``The authors declare no conflicts of interest.''

Here are examples of disclosures:

\medskip

\noindent\textbf{Disclosures.} ABC: 123 Corporation (I,E,P), DEF: 456 Corporation (R,S). GHI: 789 Corporation (C).

\medskip

\noindent\textbf{Disclosures.} The authors declare no conflicts of interest.

\section{References}

Note that \emph{Optics Letters} and \emph{Optica} short articles use an abbreviated reference style. Citations to journal articles should omit the article title and final page number; this abbreviated reference style is produced automatically when the \emph{Optics Letters} journal option is selected in the template, if you are using a .bib file for your references.

However, full references (to aid the editor and reviewers) must be included as well on a fifth informational page that will not count against page length; again this will be produced automatically if you are using a .bib file.

\bigskip
\noindent Add citations manually or use BibTeX. See \cite{Zhang:14,OSA,FORSTER2007,testthesis,manga_rao_single_2007}.

% Bibliography
\bibliography{Biblio_CVD-CR}

% Full bibliography added automatically for Optics Letters submissions; the following line will simply be ignored if submitting to other journals.
% Note that this extra page will not count against page length
\bibliographyfullrefs{sample}

\end{document}

### 2020-11-17 Mes papiers/Méca CR/Meca_cross_SI.tex
okou

@@
 When the magnetic field is in these planes, every class is at resonance with another class, as can be seen in Fig.
-X.
+\ref{CR_22}.
 
@@
 
-Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 4.
+Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 3.
 of the main text, where we can see a degeneracy plane of the [110]$_\perp$ family being crossed.
@@
 \ref{T1_protocol} .
-In the first sequence ($S_1$), the spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$.
+In the first sequence the spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$.
 The spin state is finally read out with a 10 $\mu$s laser pulse, shorter than the polarization time of the spins.
 
-The second sequence ($S_2$) uses the same parameters (polarization time, dark time and readout time) than the first sequence but adds a resonant microwave $\pi$ pulse on a transition of one of the four classes of NV$^-$ right before the readout pulse, exchanging the population of the $m_s=\ket{0}$ state with one of the $m_s=\ket{\pm 1}$ state.
+The second sequence uses the same parameters (polarization time, dark time and readout time) than the first sequence but adds a resonant microwave $\pi$ pulse on a transition of one of the four classes of NV$^-$ right before the readout pulse, exchanging the population of the $m_s=\ket{0}$ state with one of the $m_s=\ket{\pm 1}$ state.
 
-By looking a the subtraction of both signal $S_1-S_2$, we can then measure the spin state population of a single class, and remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark.
+Noting $S_1$ the signal given by the first sequence and $S_2$ the one given by the second sequence, then by looking at $S_1-S_2$ we can measure the spin state population of a single class, and remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark (which would be common to the two signals and therefore subtracted).
+In order to avoid low frequency noises or drift, we alternate both sequence when performing the measurement.
 
@@
 To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonance (ODMR) spectra....
+
+\subsection{Signal drift for levitating particles}
+
+Measurements on levitating diamonds have to be relatively short (few minutes at most) because of a slow drift on the particle orientation.
+The most likely origin of this drift is the loss of charges of the diamond due to photoionization by the laser, which changes the trapping conditions over time.
+
+This drift is the reason why the photoluminescence scan on Fig.3 of the main text has a lower signal to noise ratio than the scan in Fig.1 of the main text, which was done on a static diamond for a few hours.
 
@@
 
+\subsection{Origin of the magnetic torque and link to cross-relaxation}
+
+\begin{figure}[ht]
+\centering{ \scalebox{0.4} {\includegraphics{map_torque_1_classe}}}
+  \caption{\textbf{a)} Skecth of an NV center aligned in the \textbf{z} direction in the presence of an eternal magnetic field \textbf{B} forming an angle $\theta$ with the NV axis, and resulting magnetization $\gamma_e \langle \mathbf{ \hat S} \rangle$ of the NV center
+   \textbf{b)} Plotting of the longitudinal ($S_\parallel$ in dashed line) and transverse ($S_\perp$ in plain line) components of the spin operator average value, in $\hbar$ unit, as a function of $\theta$ and for a magnetic field amplitude $|\mathbf{B}|=100$ G.
+   \textbf{c)} Amplitude of the magnetic torque $\mathbf \Gamma = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B$ acting on a single spin as a function of $\theta$ for $|\mathbf{B}|=100$ G.
+   \textbf{d)} Amplitude of the same magnetic torque in the crystalline basis with $\theta$ and $\phi$ being the polar and azimuthal angle with respect to the [100] direction.}
+  		\label{Torque1classe}
+\end{figure}
+
+The magnetic torque responsible for the motion of the diamond fundamentally comes from the crystalline anisotropy of the NV centers and from the transverse field $B_\perp$ responsible for mixing the eigenstates in the stationary state.
+We will start by focusing on a single NV center.
+
+As can be seen in Fig \ref{Torque1classe} b), when under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, at the low magnetic fields ($<$ 200 G) we are working at, will be oriented at a 90$^\circ$ angle from the NV axis : $\langle \hat S_z \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$ This magnetization vanish when the magnetic field is aligned with the NV center since there is no longer a transverse field responsible for the mixing of the eigenstates.
+
+The magnetization of the NV center is therefore not aligned with the magnetic field, except when the field is also at a 90$^\circ$ angle from the NV axis, which means that the magnetic torque $\mathbf \Gamma = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B $ will be non-zero everywhere except when the field is aligned with the center, or in the plane normal to the direction of the center.
+We can describe each NV center as a paramgnetic defect with the anisotropic magnetic suceptibility $\chi = \begin{pmatrix}-\chi_\perp & 0 & 0\\ 0 & -\chi_\perp & 0 \\ 0 & 0 & 0  \end{pmatrix}$ in the $(\mathbf{x},\mathbf{y},\mathbf{z})$ basis where $\mathbf{z}$ is the orientation of the NV center.
+
+The amplitude of the torque with respect to the magnetic field orientation (for an amplitude of 100 G) is represented in 1D in Fig \ref{Torque1classe} c) where we can see a behavior very close to $|sin(2\theta)|$, which is the formula (3) of the main text that we obtained through a perturbative approach.
+The same torque amplitude is represented in 2D in Fig \ref{Torque1classe} d) where the two dots correspond to the [111] direction, when the magnetic field is aligned with the centers, and the curvy line is the plane $[111]_\perp$.
+Important to note is the maximum torque value $1\cdot 10^{-26}$ Nm for a single spin.
+
+\begin{figure}[ht]
+\centering{ \scalebox{0.4} {\includegraphics{map_torque_4_classe}}}
+  \caption{Magnetic torque amplitude for 4 spins, one in each possible orientation, as a function of $\theta$ and $\phi$, the polar and azimuthal angle with respect to the [100] direction.
+\textbf{a)} Not taking into account the CR between NV centers.
+\textbf{b)} Taking into account the CR between NV centers.}
+  		\label{Torque4classes}
+\end{figure}
+
+Fig \ref{Torque4classes} a) represents the same map but for four NV centers, one in each possible orientation.
+We can see that the maximum torque actually decreased to $3\cdot 10^{-28}$ Nm even though we increased the number of NV centers by four.
+This is due to the directional averaging of the torque generated by the four centers.
+The torque per NV center is decreased by more than two order of magnitude when taking the directional averaging in consideration.
+
+Finally, Fig \ref{Torque4classes} b) shows the same map, this time taking into account the modification of the spins lifetime due to cross-relaxations. There are two things to note here : 
+\begin{enumerate}
+\item The maximum torque has increased by an order of magnitude compared to the previous case, up to $3\cdot 10^{-27}$ Nm for four spins, or about $10^{-27}$ Nm per spin.
+Qualitatively, this is because cross-relaxation will lower the torque contribution of specific classes (the ones that get depolarized), meaning the the end result is closer to the single spin case (there is less directional averaging)
+\item The change in magnetic torque is resonant, it happens only when different classes are brought to resonance (we can see in Fig \ref{Torque4classes} b) the [110]$_\perp$ planes that were drawn in Fig.
+\ref{cristallo}), meaning that the change in signal when scanning a magnetic field will be much sharper.
+\end{enumerate}
+
 \subsection{Torque sensing with a levitating diamond}
@@
 
-We model the trap as a pure harmonical potential, both for the center of mass and for the rotational degrees of freedom of the diamond (REF qq part), with trapping frequencies $\omega_T \approx 1$ kHz that we can experimentally measure.
-If we focus on a single rotational degree, we can write the torque generated by the trap as $\Gamma_T^\theta=-K(\theta-\theta_{eq})$, where $K=I \omega_t^2$ is the stiffness of the trap, $I$ being the moment of inertia of the diamond.
+We model the trap as a pure harmonical potential, both for the center of mass and for the rotational degrees of freedom of the diamond (REF ?), with trapping frequencies $\omega_T \approx (2\pi) \cdot 1$ kHz that we can experimentally measure.
+If we focus on a single rotational degree, we can write the torque generated by the trap as $\Gamma_T=-K(\theta-\theta_{eq})$, where $K=I \omega_t^2$ is the stiffness of the trap, $I$ being the moment of inertia of the diamond.
 
@@
 
-In our case, $\Gamma_{\mathrm{ext}}$ is the magnetic torque exerted by the NV$^-$ spins directly on the diamond, such that we can write $\Gamma_{\mathrm{ext}} = N_{NV} \langle \Gamma_{\mathrm{1 spin}} \rangle$  where $\langle \Gamma_{\mathrm{1 spin}} \rangle = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B \approx 10^{-28}$ Nm is the classical magnetic torque applied by one spin (see Section  \ref{Simu}).
+In our case, $\Gamma_{\mathrm{ext}}$ is the magnetic torque exerted by the NV$^-$ spins on the diamond, such that we can write $\Gamma_{\mathrm{ext}} = N_{NV} \langle \Gamma_{\mathrm{1 spin}} \rangle$  where $\langle \Gamma_{\mathrm{1 spin}} \rangle = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B \approx 10^{-27}$ Nm is the expected magnetic torque applied by one spin (see Fig \ref{Torque4classes}).
 
-By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle [NV^-]}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-3} rad$$ where $[NV^-] \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 1$ kHz is the typical value of the trap frequency.
+By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle n(NV^-)}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-4} rad$$ where $n(NV^-) \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 6.3\cdot 10^3$ rad/s is the typical value of the trap angular frequency.
 
 It should be noted that the main uncertainty comes here from the diamond size, which along the other uncertainties can change the expected result by an order of magnitude.
-
-\subsection{Change of torque at the cross-relaxation}
 
@@
 \begin{figure}[!ht]
-  \centering \scalebox{0.32}{\includegraphics{CR_meca_22}}
-  \caption{Mechanical detection of a degeneracy in a [100]$_\perp$ plane.
-a) Theoretical frequencies for the four $\ket{0} \to \ket{-1}$ transitions in dashed lines and mechanical ESR spectra vertically.
-b) Photoluminescence of the NV centers as a function of the scanning magnetic field (blue crosses) and gaussian fit (orange line).
-c) Simulated population in $\ket{m_s=0}$ for the stationary state with (plain) or without (dashed) taking into account the decrease of the T$_1$ induced by the cross-relaxations.
-  d) Reflected signal off the diamond as a function of the scanning magnetic field (blue crosses) and double gaussian fit (orange line) e) Simulated torque applied by the spins on the diamond, with (plain) or without(dashed) taking into account the cross-relaxations.}
+  \centering \scalebox{0.4}{\includegraphics{22-total}}
+  \caption{Mechanical detection of a degeneracy in a [100]$_\perp$ plane.\textbf{a)} Path of the magnetic field along the scan (red dots) in the ($\theta , \phi$) basis with respect to the [100] direction.
+[100]$_\perp$ family of planes in plain blue lines, [110]$_\perp$ family of planes in orange dashed lines.\textbf{b)} Theoretical frequencies for the four $\ket{0} \to \ket{-1}$ transitions in dashed lines and experimental ESR spectra vertically.
+\textbf{c)} Photoluminescence of the NV centers as a function of the scanning magnetic field (blue crosses) and gaussian fit (orange line).
+\textbf{d)} Simulated population in $\ket{m_s=0}$ for the stationary state with (plain) or without (dashed) taking into account the decrease of the T$_1$ induced by the cross-relaxations.
+  \textbf{e)} Reflected signal off the diamond as a function of the scanning magnetic field (blue crosses) and double gaussian fit (orange line) \textbf{f)} Simulated torque applied by the spins on the diamond, with (plain) or without(dashed) taking into account the cross-relaxations.}
+
 	\label{CR_22}
@@
 Fig.
-\ref{CR_22}a) shows the theoretical frequencies of the $\ket{0} \to \ket{-1}$ transitions for all four classes of NV and ESR spectra measured through the reflected laser light off the diamond for various magnetic field values.
-Unlike The experiment of the main text, this time all four classes of NV are at resonance with another class at B=23 G, telling us that we are crossing a [100]$_\perp$ plane instead of a [110]$_\perp$ one.
+\ref{CR_22}b) shows the theoretical frequencies of the $\ket{0} \to \ket{-1}$ transitions for all four classes of NV and ESR spectra measured through the reflected laser light off the diamond for various magnetic field values.
+Unlike The experiment of the main text, this time all four classes of NV are at resonance with another class at B=23 G, telling us that we are crossing a [100]$_\perp$ plane instead of a [110]$_\perp$ one, as can be seen on fig \ref{CR_22}a)
 
 Fig.
-\ref{CR_22}b) shows the recorded photoluminescence of the NV centers during the magnetic field scan.
+\ref{CR_22}c) shows the recorded photoluminescence of the NV centers during the magnetic field scan.
 As previously, we see a drop in the luminosity when the resonance occurs, that we fitted with a gaussian.
@@
 This behavior is predicted by the simulation of the $\ket{m_s=0}$ population in the stationary state in Fig.
-\ref{CR_22}c).
+\ref{CR_22}d).
 
 Fig.
-\ref{CR_22}d) shows the signal of the reflected laser off the diamond, allowing us to measure angular displacement.
+\ref{CR_22}e) shows the signal of the reflected laser off the diamond, allowing us to measure angular displacement.
 This time we see a clear difference with the experiment of the main text, instead of a single drop centered on the resonance, we see two bumps on both sides of the resonance, fitted with two gaussians.
 This is in accordance with the simulated torque in Fig.
-\ref{CR_22}e) where we can see a dispersive profile with an almost zero torque at the resonance.
+\ref{CR_22}f) where we can see a dispersive profile with an almost zero torque at the resonance.
 The reason we do not observe a change of sign in the experiment (with two positive bumps instead of a positive and a negative one) might be because of the strong non-linearity of our detection : if the signal initially correspond to a dark spot of the speckle, then a change in the motion of the diamond can only result in an increased signal.
@@
 What happens when being near, but not at, resonance is that the classes will not be identically depolarized : looking at Fig.
-\ref{CR_22}a), we can see that the two classes of higher frequency are always slightly closer to each other than the two classes of lower frequency (we can see the slope of the frequencies being a bit smaller for the two upper classes), resulting in a bit more depolarization for these two classes, except when they are exactly at resonance.
+\ref{CR_22}b), we can see that the two classes of higher frequency are always slightly closer to each other than the two classes of lower frequency (we can see the slope of the frequencies being a bit smaller for the two upper classes), resulting in a bit more depolarization for these two classes, except when they are exactly at resonance.
 This interpretation accounts both for the overall shape of the responses, and the comparatively weakness of the phenomenon.

### 2020-11-17 Talks/IQFA 2020/Script.tex
okou

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\section{intro}
Thanks to the organizer for inviting me. Today I'm going to present to you two of our recent results regarding dipolar spin interaction in diamond. The first one is the optical observation of interaction between Natrogen-Vacancy (or NV) spins and new spin species in diamond, and the second is the mechanical detection of dipolar interaction between NV centers.

But first, I will have to briefly introduce the physics of NV centers.

\section{NV center}
%Slide avec un NV dans sa maille cristalline et les niveaux d'énergie (juste excité
% et fondamental) avec un zoom sur le niveau d'énergie fondamentale.
A Nitrogen-vacancy center, which consists of nitrogen and a vacancy in the diamond crystal lattice, is a colored center, because it emits red light when optically excited (in our case with a green laser). However, the main reasons NV centers are widely used in quantum technology are not its optical properties, which are rather poor, but its spin properties.

An NV center is a Spin-1 specie with no spin-orbit coupling and relatively good spin properties : a lifetime of a few milliseconds at room temperature, and a coherence time that can reach ??? for single spins. In our case since we are working with relatively dense ensembles, we do have an inhomogenous broadening of a few MHz.

But the most important property of the NV spin is that it can be optically polarized : At zero magnetic field, because of (origin of D???), the ms=0 state has an energy level about 3 GHz lower than that of the magnetic states ms=+/-1. At room temperature, this means that the ms=0 state is more populated by a whole .0000001\%. But when you are optically exciting (Reformuler ca) the NV center,because the optical cycle is different for the ms=0 and the magnetic states, you start pumping population in the 0 state up to about 80\%, which would be equivalent to a Temperature of 0.6 K. Another important property, which is related, is that the ms=0 state is brighter than the magnetic state, (quantum ratios ???).

So to sum up, exciting an NV center with a green laser allows us to polarize the NV spin, and looking at the photoluminescence emitted by the NV allows us to measure that polarization.

\section{Magnetometry with NV centers}
The most common use of NV centers is probably magnetometry. The reason is rather simple : thanks to the optical pumping and the ability to read the quantum state, you can quite easily find the energy levels of your spin states, and since these levels depend on the magnetic field though the Zeeman effect, you can deduce the magnetic field. 
The simplest protocol is called Optically Detected Magnetic Resonance (ODMR). This protocol consist in continuously shining a green laser on your diamond and looking at the photoluminescence (or PL) while scanning a microwave frequency. Whenever the microwave is resonant with a 0 to -1 or 0 to +1 transition, the spin is partially depolarized in a magnetic state and the PL drops.% Plot avec un ODMR de single spin.

This would be what you observe with a single spin, but we are dealing with ensemble. And an important aspect I did not mention so far is that the ms=0,+1 or -1 states are implicitly the eigenstates of a Sz operator, with a particular direction z. And since we have a zero-field splitting term, the direction of z is not imposed by the magnetic field. To the fist order, as long as the zeeman splitting is lower than than the zero field splitting, we can consider that the z axis is given by the zero field splitting, and the Zeeman splitting will simply be the projection of the magnetic field along the z axis. 
The physical z axis would then simply be the one formed by the nitrogen and the vacancy. And since there are four possible crystalline orientation in a diamond, one for each valence bond of the carbon, there are four possible z axis for our NV centers, and therefor 4 possible projection of the magnetic field. %Un peu de géométrie

Going back to the ODMR, when dealing with ensemble, this is what you would observe : 4 transitions of 0 to -1, one for each possible orientation, or class, of NV centers, and 4 0 to +1 transitions.

\section{Polarization transfer}
Ref : Kamp, E. J., Carvajal, B., \& Samarth, N. (2018). Continuous wave protocol for simultaneous polarization and optical detection of P1-center electron spin resonance. Physical Review B, 97(4), 045204.

le science 2016 de Lukin, le PRX de Wallsworth

Another thing you can do when you have a polarized spin is to transfer its polarization, through dipolar interaction,  to another spin species. This is an especially active area for NV application with Dynamic Nuclear Polarization where you can transfer some of the NV polarization to nuclear spins either in the diamond (papier 13C et 14N), or outside the diamond (papier Lukin et wallsworth).
But you can also transfer its polarization to a (dark) electron spin. So far, as far as y know, there has only been transfer to electric spin inside the diamond, and only to one particular spin which is the substitutional Nitrogen N minus defect, often called P1 center.

There are several ways to transfer polarization from one spin to another, but the simplest one is cross-relaxation : by tuning the magnetic field you can sometimes match the energy transition of diferent spins species and they will start to exchange spin quanta through RESONANT dipolar interaction, which tend to balance the spin polarization of the two spins.

This leads me to our first result :
\section{Cross-Relaxation with NV centers}
%1ere slide : Scan + en dessous 100 et puis soustraction par dessus.
What we see here is the change in photoluminescence of our NV centers when we scan a magnetic field along the 100 direction of the diamond. I'll start detailing what we observed.

The first thing to discuss is the 100 direction. It is a defined by the crystaline planes of the diamond, and it correspond to a direction where the magnetic field projection is the same on all four classes of the diamond, so in a way we are back to the single spin case which will make analysis much easier.

The second thing to discuss is the overall decrease of the photoluminescence with the magnetic field. This is because every class is misaligned with the magnetic field, which means that by increasing the field amplitude, you will increase both the longitudinal field, so the zeeman splitting Bz*Sz in the Hamiltonian, but you also increase the transverse magnetic field component, so Bx*Sx which will mix the bright ms=0 state with the dark ms=+/-1 states and decrease the overall luminosity.

So we expect a decreasing enveloppe (slide), and when we subtract the enveloppe of the signal, what remains is this :(new slide avec la soustraction) three clear features, none of which at been observed before and that we attributed to cross-relaxation of the NV center with three different species. The first

  \end{document}

### 2020-11-17 abstract/IQFA 2020/Abstract_Pellet-Mary.tex
demenagement
okou

@@
 \normalsize
-The negatively charged nitrogen vacancy center (NV$^-$) is a well established quantum system, employed both for applications and fundamental research, thanks to its good coherence properties, long lifetimes and most importantly their ability to be optically polarized.
+The electronic spin of the negatively charged nitrogen vacancy center (NV$^-$) is employed both for applications and fundamental research, thanks to its good coherence properties, long lifetimes and most importantly their ability to be optically polarized.
 We have recently found new ways to use these properties of NV centers in order to study the dipolar interactions between ensemble of spins in diamond, including NV centers as well as other spin defects.
@@
 On the one hand, we have observed cross-relaxations (CR) between NV centers and other defects (see Fig.
-1), namely VH$^-$ \cite{VH}, War1 (first defect found by EPR at the university of Warwick, still chemically unknown) and $^{13}$C First-shell modified NV.
+1), namely VH$^-$ \cite{VH}, War1 (first defect found by EPR at the university of Warwick, still chemically unknown) and $^{13}$C-NV pair.
 None of these defects had been observed through CR before.
-This is a first step toward hyper-polarization of new spin defects in diamond.
+This observation is a first step toward hyper-polarization of new spin defects in diamond.
 
-On the other hand, we have used the mechanical detection of a levitating diamond \cite{nature} to measure the torque applied by the NV centers on the diamond when two classes of NV with different orientation are brought into resonance(see Fig.
-2).
-This effect is due to the dipolar-mediated modification of the T$_1$ of NV centers happening with dense ensembles.\cite{Lukin}.
-This paves the way toward the mechanical detection of other spin impurities in diamond, as well as the observation of the Einstein-de Haas effect with paramagnetic systems.
-\cite{Meriles}
+On the other hand, we have used a levitating diamond \cite{nature} to measure the torque applied by the NV centers on the diamond when two classes of NV with different orientation are brought into resonance (see Fig.
+2) because of the dipolar-mediated modification of the T$_1$ of NV in dense ensembles\cite{Lukin}.
+This demonstration paves the way toward the mechanical detection of other spin impurities in diamond, as well as the observation of the Einstein-de Haas effect with paramagnetic systems\cite{Meriles}.
 
@@
 
-Figure 1 : Photoluminescence change while scanning the magnetic field along the crystalline [100] direction. Three dips are observed for the CR with $^{13}$C-NV, VH$^-$ and War1.
+Figure 1 : \textit{Relaxation-based detection of dipolar interaction.} \textbf{Left} : Sketch of the various spins interacting.
+\textbf{Right} : Photoluminescence change while scanning the magnetic field along the crystalline [100] direction.
+Three dips are observed for the CR with $^{13}$C-NV, VH$^-$ and War1.
 
@@
 
-Figure 2 : Angular position of the diamond while scanning B$_{\mathrm{Scan}}$ without a microwave. The rotation of the diamond at 17 G corresponds to the resonance between two classes of NV
+Figure 2 : \textit{Mechanical detection of dipolar interactions.} \textbf{Left} : Sketch of the experimental setup.
+\textbf{Right :} Angular position of the diamond while scanning B$_{\mathrm{Scan}}$ without a microwave.
+The rotation of the diamond at 17 G corresponds to the resonance between two classes of NV.
 \end{abstract}

### 2020-11-17 théorie/fiches de lecture/Ashcroft-Mermin/fiche.tex
okou

@@
 En ordre de grandeur, ça fait $\approx 10^{22} e^-/cm^3$ soit environ 10 x plus que la densité d'un gaz à 300K et $P_0$.
-En plus les électrons sont chargés, donc fortement en interaction, mais Drude il s'en ballec il dit que c'est un gaz dillué.
+En plus les électrons sont chargés, donc fortement en interaction, mais Drude il s'en ballec il dit que c'est un gaz parfait.
 
@@
  $\epsilon(\omega)$ est la constante diélectrique dans le voc du livre et elle vaut $$ \epsilon(\omega)= \frac{4\pi i \sigma}{\omega} \approx 1-\frac{\omega_p^2}{\omega^2}$$ ce qui fait sortir la fréquence plasma $\omega_p=\frac{4\pi n e^2}{m}$.
-Pour $\omega < \omega_p$, les solutions sont exponentielles (donc décroissante parce qu'on est en physique), le métal absorbe les OEM avec une épaisseur de peau $l=c/\omega_p$.
+Pour $\omega < \omega_p$, les solutions sont exponentielles (donc décroissante parce qu'on est en physique), le métal absorbe/réfléchit les OEM avec une épaisseur de peau $l=c/\omega_p$.
 Pour $\omega > \omega_p$, les solutions sont des oscillations : le métal devient transparent aux OEM.
@@
 
- \item \textbf{Conductivité thermique}
+ \item \textbf{Conductivité thermique} : Le modèle propose une explication à la loi empirique de Wiedemann-Franz qui dit que $$ \frac{\kappa}{\sigma} \propto \frac{1}{T}$$ Ou  $\kappa$ est la conductivité thermique.
+Et Drude a réussi à retrouver ça avec son modèle (en disant que la conduction thermique était purement électronique, et que la propagation de la chaleur vient du fait que après une collision dans une zone chaude, la vitesse de sortie des $e^-$ est plus grande que dans une zone froide.
+Mais surtout il a réussi à calculer la constante de proportionalité avec des constantes fondamentales, en suposant que la chaleur spécifique d'un électron vaut $\frac{3}{2} k_B$ et son énergie cinétique $\frac{1}{2}mv^2=\frac{3}{2}k_BT$.
+
+ Sauf que ce con, de un il s'est gourré d'un facteur 2 (je la connais celle la, petit filou), et surtout son raisonement est complètement faux.
+Le problème dans son modèle c'est que du coup c'est les électrons qui contiennent l'essentiel de l'énergie thermique, alors qu'en vrai ils en contiennent 100x moins que ce qu'il prédit, mais ont une vitesse 100x plus grande.
+Parce que en vrai c'est pas un gaz parfait les $e^-$ dans un métal.
+
+ Une conséquence par contre du modèle, c'est que cette histoire de collisions avec plus ou moins de vitesse, bah en faite ca occasionne un mouvement d'ensemble des $e^-$ (vers les basses Temp.).
+Sauf que t'as pas de courant qui peut circuler dans ton métal si t'en en circuit ouvert, du coup comme pour l'effet Hall, t'as un champ électrique dans le sens opposé qui s'installe.
+Ca s'apelle \textbf{l'effet Seebeck}, ou le \textbf{champ thermoélectrique}.
+Et la par contre la prédiction de Drude est 100x trop grande et ça se compense pas.
 
  \end{enumerate}
+
+ \chapter{Théorie de Sommerfeld des métaux}
+ Fini de rire, on passe en quantique
+ \section{Statistique quantique}
+ Si on s'intéresse à la vitesse, la formule de Maxwell-Boltzmann pour un gaz parfait c'est : $$ f_B(\mathbf{v}) = n_0 \left(\frac{m}{2 \pi k_B T}\right)^{3/2} e^{-mv^2/2 k_B T} $$
+
+ ou $f_B(\mathbf{v})$ est la densité volumique de particules avec une vitesse \textbf{v} et $n_0$ la densité totale de particule ($\int f_B(\mathbf{v}) d\mathbf{v} = n_0$).
+Maintenant Si on passe en quantique et qu'on regarde la distribution de Fermi-Dirac plutot, on obtient : $$ f(\mathbf{v})=\frac{(m/\hbar)^3}{4\pi^3}\frac{1}{1+exp[(mv^2/2-k_b T_0)/k_B T]}$$
+
+ Le $T_0$ vient de la normalisation de f à $n_0$, j'avoue que j'aimerai en savoir un peu plus  mais ce sera pour plus tard.
+En attendant la conclusion c'est que pour $mv^2/2 < k_b T_0$ , et $T_0 \approx 10^3 K$, le profil de distribution des vitesse est plat vu qu'il ne dépend (presque) pas de v, alors que $f_B$ est une exponentielle doublement décroissante.
+
+ \section{Modèle quantique du gaz d'électrons libres}
+ En faite dans un premier temps on peut continuer de négliger les interactions $e^-$- $e^-$, on garde des électrons sans interactions qui ne voient qu'un potentiel exterieur, mais qui suivent la stat de Fermi-Dirac, ce qui revient à dire qu'on prend en compte le principe d'exclusion de Pauli.
+
+ Ensuite on va chercher les états propres, et en particulier l'état fondamental, de ce gaz d'électrons.
+Comme les $e^-$ sont indépendant, il suffit de chercher les états propres pour 1 électron, et donc on resoud l'eq de Shrodinger pour un électron sans potentiel (j'ai la flemme de la taper).
+On choisit que l'électron est confiné dans un cube $V=L^3$ avec des CL périodiques : lui son argument c'est que c'est purement mathématique, mais que comme on s'interesse aux propriétés de bulk, les conditions surfaciques ne doivent rien changer.
+Bon on va le croire.
+
+ Bref les solutions pour 1 $e^-$ c'est des OPPM, c'est à dire les vecteurs propres de $\mathbf{\hat p}$ que l'on peut noter $\ket{\mathbf{k}}$ d'énergie propre $\mathcal{E}_k=\frac{(\hbar k)^2}{2m}$.
+On peut donc leur associer une vitesse $v=\hbar k /m$ est une longueur d'onde (de de Broglie) $\lambda = 2\pi /k$
+
+ Et qui dit $\ket{\mathbf{k}}$ dit espace réciproque, et discrétisation à cause du volume (réel) fini.
+Donc dans l'espace des k, chaque état correspond à un volume élementaire $(2\pi/L)^3$, et dans la limite ou on a un grand nombre de k, la densité volumique d'état (toujours dans l'espace des k) vaut $$ \rho(k)= (L/2\pi)^3 = \frac{V}{8 \pi ^3} $$.
+
+ \section{Le niveau fondamental du gaz d'électrons : La grand-mère à Fermi}
+
+ En oubliant pas la dégénérescence du spin, on peut donc construire l'état fondamental en remplissant les niveaux d'énergies en partant du bas (quis de l'anti-symmétrisation ?), ce qui dans l'espace des k revient à remplir une boule vu que l'énergie c'est la distance à l'origine dans l'espace des k : La sphère de Fermi, de rayon $k_f$ (pour bien se rappeler que c'est dans l'espace réciproque).
+Et la relation avec la densité volumique (dans l'espace réel) d'état $n$ est : $$n=\frac{k_f^3}{3\pi^2}$$ Et $n$ on le connait, c'est globalement le même d'un métal à l'autre (en fonction du paramètre de maille et du nombre d'électrons de Valence mais toujours le meme ordre de grandeur).
+Donc $k_f$ a aussi toujours le meme ordre de grandeur, ce qui nous donne une vitesse de Fermi (vitesse des électrons de plus haute énergie, qui sont aussi les plus nombreux) : $$ v_f = \frac{\hbar}{m} k_f \approx 4 \cdot 10^6 \mathrm{m/s} $$
+ Et ce alors qu'on est dans l'état fondamental, donc à T=0.
+A T=300K, la vitesse thermique d'une particule de la masse d'un électron est toujours 10x plus petite que cette vitesse de Fermi.
+Le calcul numérique des énergies de Fermi donne lui des valeurs $\mathcal{E}_f \approx 1.5 \to 15 $ eV, comparables donc à des liaisons chimiques.
+Et la température de Fermi $T_f=\mathcal{E}_f/k_B \approx 10^4 \to 10^5$ K.
+
+ L'énergie (par électron) totale du fondamental est $E/N= \frac{3}{5} k_B T_f$, sachant que l'énergie par électron classique dans un gaz parfait est $E/N= \frac{3}{2} k_B T$ , ça te donne idée de quand est-ce que "l'agitation quantique" l'emporte sur l'agitation thermique.
+
+ Un dernier point c'est la pression exercée par le gaz d'électron : Puisque l'énergie (pour un nombre d'électrons N donné) varie avec le volume, ça te donne une pression électronique$$P=-\left(\frac{\partial E}{\partial V}\right)_N=\frac{2E}{3V}$$ De la tu peux en déduire le module de compression électronique $B=-V\frac{\partial P}{\partial V}$ qui en l'occurence est comparable au coef de compression des métaux (je sais pas dans quelle proportion ça joue par rapport a la matrice cristalline, mais visiblement c'est pas complètement négligeable).
+
+ \section{Température non nulle : distribution de Fermi-Dirac}
+
+ \subsection{Obtention de la statistique}
+ La démo de la distribution de FD à partir du formulisme micro-cannonique et du principe d'exclusion de Pauli est très élégante. Je vais juste noter les définitions thermo du début histoire de.
+
+ La probabitilité d'avoir un état à N électrons d'énergie E dans un système à l'équilibre thermique à la température T est : $$P_N(E) = \frac{e^{-E/k_BT}}{\sum e^{-E_\alpha/k_BT}}$$
+
+ Ou les états $\alpha$ sont TOUS les états stationaires possibles (que tu peux voir en terme de nombre d'occupation des états à 1 $e^-$ en mode seconde quantification), donc de toutes les énergies possibles.
+
+ L'énergie libre $F=U-TS$ du système à N $e^-$ peut être reliée à la fonction de partition par $\sum e^{-E_\alpha/k_BT} = e^{-F_N/k_BT}$.
+F est fonction de N et de T, elle ne dépend pas d'une configuration particulière de ton état (je suis pas forcément hyper au point mais ok).
+
+ Ensuite tu t'interesse à la probabilité que l'état à 1 électron $i$ soit occupé par un état à N électron $f_i^N$, et avec un raisonement récursif astucieux tu trouves que : $$f_i^N=\frac{1}{e^{(\mathcal{E}_i-\mu)/k_BT}+1}$$
+
+ Ou $\mathcal{E}_i$ est l'énergie (à 1 électron, donc $(\hbar k)^2/2m$) de l'état i, et $\mu$ est le \textbf{potentiel chimique} défini ici par $\mu(N_0,T_0)=F_{N+1}-F_{N}=\frac{\partial F}{\partial N}_{N_0,T_0}$
+
+   Finalement, tu peux relier le potentiel chimique, la température et le nombre de particules (ou la densité) par $$ N=\sum_i f_i =\sum_i \frac{1}{e^{(\mathcal{E}_i-\mu)/k_BT}+1} $$
+
+   \subsection{Applications}
+
+   \subsubsection{Potentiel chimique et énergie de Fermi}
+   Pour un métal à 0K, on sait que la distribution des $f_i$ doit être telle que $f_i=1$ si $\mathcal{E}_i < \mathcal{E}_f$ et $f_i=0$ si $\mathcal{E}_i > \mathcal{E}_f$.
+Vu la forme des $f_i$, ca veut dire qu'il faut $\mathcal{E}_i-\mu < 0$ si $\mathcal{E}_i < \mathcal{E}_f$, et inversement.
+Bref $$\mathcal{E}_f = \lim_{T \to 0} \mu(T)$$
+
+   Ce qui pour un métal (et uniquement pour un métal) reste valable à Température ambiante.
+
+  \subsubsection{Chaleur spécifique des électrons}
+  La chaleur spécifique c'est défini par $$c_v=\frac{T}{V}\left(\frac{\partial S}{\partial T}\right)_V = \left(\frac{\partial u}{\partial T}\right)_V$$
+  ou $u=U/V$ (et $n=N/V$).
+Donc le premier enjeu ca va être de calculer U, ce qui est faisable car on a suposé qu'il n'y a pas d'interactions et que donc U est la somme des énergies des états à 1 $e^-$ : $U=2\sum_k \mathcal{E}(k) f(\mathcal{E}(k))$ ou $f$ est la fonction de Fermi $$f_T(\mathcal{E})=\frac{1}{e^{(\mathcal{E}-\mu)/k_BT}+1}$$ qui ne dépend effectivement que de l'énergie d'un état à 1 particule (et le 2 viens de la dégénrescence de spin, comme d'hab).
+
+  Ensuite pour simplifer le calcul, on va donc compter le nombre d'état d'énergie $\mathcal{E}$ et introduire une densité (volumique, attention!) d'état d'énergie entre $\mathcal{E}$ et $\mathcal{E}+d\mathcal{E}$ : $$g(\mathcal{E})=\frac{3}{2}\frac{n}{\mathcal{E}_f}\left(\frac{\mathcal{E}}{\mathcal{E}_f}\right)^{1/2}$$
+
+  Tu peux aussi l'écrire avec des constantes fondamentales mais c'est un peu plus joli comme ça.
+Pour être rigoureux il faut aussi préciser que $g(\mathcal{E} < 0) =0$, mais bon tu peux t'en douter.
+La dégénrésecence de spin est aussi incluse dans le g, comme ça c'est bon.
+
+  Avec ça on a une nouvelle formulation de u et n : $$u=\int_{-\infty}^{+\infty} \mathcal{E} g(\mathcal{E}) f_T(\mathcal{E}) d\mathcal{E} $$ $$n=\int_{-\infty}^{+\infty} g(\mathcal{E}) f_T(\mathcal{E}) d\mathcal{E} $$
+
+  \subsubsection{Calcul de grandeurs à T$\neq 0$ : Développement de Sommerfeld}
+
+  Les intégrales du type $H(T)=\int_{-\infty}^{+\infty} h (\mathcal{E}) g(\mathcal{E}) f_T(\mathcal{E}) d\mathcal{E}$ sont dans l'absolu impossible à calculer pour T$\neq 0$.
+Par contre c'est beaucoup plus simple à T=0 parce que ça devient simplement $\int_{-\infty}^{\mathcal{E}_f} h(\mathcal{E}) g(\mathcal{E})d\mathcal{E}$.
+
+  L'idée c'est que aux températures qui nous intéressent, $f_T(\mathcal{E})$ change assez peu, donc on peut faire un dl de $H$ autour de $H(0)$.
+En plus, le seul endroit ou $f_T(\mathcal{E})$ change avec T c'est autour de $\mathcal{E}=\mathcal{E}_f$ (pour $\mathcal{E}<<\mathcal{E}_f$, $f_T(\mathcal{E})=1$ et pour $\mathcal{E}>>\mathcal{E}_f$, $f_T(\mathcal{E})=0$, quelque soit T.) Donc il suffit de s'intéresser au comportement de $h$ autour de $\mathcal{E}_f$.
+Le développement en série de Sommerfeld s'écrit au premier ordre : $$H(T)=H(0)+\frac{\pi^2}{6}(k_B T)^2 \frac{\partial h}{\partial \mathcal{E}}(\mathcal{E}_f)$$
+
+  Avec ça on peut facilement en conclure que $$u(T) = u(0) + \frac{\pi^2}{6}(k_B T)^2 \frac{\partial (\mathcal{E}g(\mathcal{E}))}{\partial \mathcal{E}}(\mathcal{E}_f) = u(0) + \frac{3 \pi^2}{8} (k_B T)^2 \frac{n}{\mathcal{E}_f}$$
+
+  Puis $$c_V(T)=\frac{\partial u}{\partial T}= \frac{3 \pi^2}{4} \frac{k_BT}{\mathcal{E}_f} nk_B$$
+
+  Bon il se trouve que je suis peut-etre allé un peu vite en besogne, et que le dl de Sommerfeld doit se faire autour de $\mu$ plutot que de $\mathcal{E}_f$, ce qui change légèrement le facteur numérique.
+Mais bon le scaling est bon, et par rapport à un gaz classique $c_V(T)=\frac{3}{2}nk_B$, on voit qu'on a pris un facteur $\frac{k_BT}{\mathcal{E}_f}$ dans les detns, d'ou le fait que la chaleur spécifique des $e^-$ est 100x moins importante que pour un gaz parfait.
+Par contre ça fait un scaling linéaire avec T, alors que les solides à basse température c'est en $T^3$, donc à basse température (qq K) les métaux ont une chaleur spécifique essentiellement due aux $e^-$ de conductions.
+
+  \subsubsection{Vitesse des électrons}
+  Vu qu'on est en quantique, maintenant la première question c'est de savoir si on peut réelement parler des vitesse, et de position pour les électrons.
+Dans notre modèle les électrons sont complètement délocalisés vu qu'ils sont défini par $\mathbf{k}=\mathbf{v}/\hbar$.
+Si on veut un modèle un peu plus réaliste, il faut inclure des fluctuations autour de \textbf{k} faibles devant $k_f$, donc des fluctuation en positions grandes devant $1/k_f$, qui est plus ou moins la distance inter-atomique.
+Donc bref les électrons ne peuvent pas être localisés à l'atome près, par contre sur une centaine d'\AA c'est jouable.
+Typiquement du point de vue de la lumière visible, on peut considérer des $e^-$ ponctuels, par contre pour les rayons X il faut regarder le comportement ondulatoire.
+
+  Pour des phénomènes qui ont lieu sur des distances assez grande, on peut a nouveau considérer les $e^-$ comme des bouboules avec une certaine vitesse, mais cette fois en utilisant la ditribution de FD des vitesse : en gros une distribution uniforme des vitesses (dans l'espace, pas en norme) jusqu'a $v_f$ vu qu'on a une distrib uniforme des \textbf{k}, et une vitesse moyenne de l'ordre de $v_f$ qui est environ 10x plus rapide que la vitesse thermique à température ambiante.
+
+  Pour le reste on revient à un modèle collisionel, dont le temps carac $\tau$ est toujours défini par la résitivité/conductivité (et qui ne dépendait pas de la vitesse).
+Idem pour l'effet Hall et ou la (non)magnéto-résistance qui ne dépendent pas de la vitesse.
+Ce qui va changer en revanche c'est le lpm ($\approx$ 10 x plus court, entre 10 et 100 \AA , ce qui est problèmatique pour les petits lpm vu qu'on s'approche de $\Delta x$), la conductivité thermique ne change quasi pas parce qu'elle est en $c_v v^2$ et que les erreurs numériques se compensent, et l'effet thermoélectrique qui est $\propto c_v$ 0est donc 100x plus petit que dans le modèle de Drude.
+
+  \chapter{Défaut du modèle des électrons libres}
+
+  Bon même avec une statistique quantique, y'a tout un tas de phénomènes qui sont soient inexpliqués, soit numériquement faux.
+Le plus important c'est qu'on a toujours aucunue idée de quels sont ces électrons libres, et pourquoi certains solide sont des métaux et d'autres non, meme avec une chimie équivalente (genre Aluminium vs Bore).
+
+  Si on reprend les hypothèses jusque la, on avait :
+  \begin{enumerate}
+  \item \textbf{Hypothèse des électrons libres.} On supose qu'entre 2 collisions, les $e^-$ n'interagissent pas avec le cristal. On supose aussi le cristal comme parfaitement immobile.
+  \item \textbf{Hypothèse des électrons indépendants.} On supose que les électrons n'interagissent pas entre eux.
+  \item \textbf{Hypothèse du temps de relaxation} On supose que le résultat d'une collision ne dépend pas de la configuration des $e^-$.
+  \end{enumerate}
+
+  En faite le plus gros soucis vient de la première hypothèse. Etonnement (pour moi) considérer les électrons comme indépendants ne pose pas trop de soucis. Bref on va devoir parler de cristaux.
+
+  \chapter{Réseaux cristallins}
+  Je vais sans doute aller un peu plus vite dans ces chapitres vu que, bon, la cristallo c'est pas vraiment de la physique.
+  \section{Réseau de Bravais (réseau direct)}
+  C'est le réseau cristallin en tant que tel, c'est à dire c'est la collection de points (en 3D) qui peuvent s'écrire $$\mathbf{R}=n_1 \mathbf{a_1}+n_2 \mathbf{a_2}+n_3 \mathbf{a_3}$$ ou $(\mathbf{a_1},\mathbf{a_2},\mathbf{a_3})$Sont les \textbf{vecteurs primitifs}qui engendrent le réseau, et qui ne sont pas forcément orthogonaux.
+
+  Les noeuds du réseaux peuvent être des atomes, mais aussi des structures plus complexes, l'essentiel c'est que \textbf{une translation dans le réseau de bravais laisse la structure inchangée}.
+En particulier, si tu as un système hexagonal en nid d'abeil (2D), tu ne peux pas considérer les atomes comme des noeuds du réseau, parce que tu as 3 directions possibles de translation par atome.
+Si je dis pas de bétise il te faut au moins 2 atomes par maille.
+  \subsubsection{Remarque sur le CFC}
+  Pour le cubique face centré, tu peux choisr un réseau cubique dont la maille contiendra 4 atomes, mais tu peux aussi considérer le réseau qui relie un sommet aux 3 centres de faces les plus proches (réseau non-orthogonal) et avoir un réseau mono-atomique.
+  \section{Réseau fini}
+  Au coeur de la théorie des réseau, tu as le fait qu'ils sont infinis (pour avoir l'invariance par translation).
+On serait pas contre considérer des cristaux infinis, sauf qu'on a besoin qu'ils sooient finis pour discrétiser les états quantiques.
+Du coup on va considérer que le cristal occupe une partie finie d'un réseau infini, et on notera ses noeuds  avec les indices $(n_1,n_2,n_3)$ tels que $$0\leq n_i \leq N_i$$ avec $N=N_1N_2N_3$ le nombre total de sites du cristal.
+  \section{Maille}
+  Une fois que tu as les noeuds de ton reseau de Bravais, tu peux remplir l'espace comme tu veux autour de ces noeuds (tant que tu laisse pas de trous).
+Le choix par défaut c'est le parallélépipède formé par les 3 vecteurs primitifs.
+Si ta maille ne contient qu'un seul noeud, elle est \textbf{primitive}.
+Par exemple pour le cfc (ou le cc), la maille primitive n'est pas cubique vu que la maille cubique contient 4 (resp.
+2) noeuds.
+Dans ce cas les mailles cubiques sont nommées \textbf{conventionelles}.
+La maille de \textbf{Wigner-Seitz} c'est une maille primitive qui vérifie les propriétés de symmétries de ton réseau.
+En 2D ce sera toujours un hexagone, en 3D ça peut être plus funky.
+  \section{structure cristalline}
+  La structure cristalline c'est tout simplement le contenu physique dans chaque maille (x atomes/molécules, positionnées de telle ou telle façon)
+  \section{Diamant}
+  Une façon de voir le diamant c'est 2 réseaux cfc dont l'un est décalé d'un quart de longeur dans les 3 directions.
+Du coup le diamant c'est un CFC dont la maille primitive contient 2 atomes de carbones, positionnés en $\mathbf{(0,0,0)}$, et $\mathbf{(\frac{a_0}{4},\frac{a_0}{4},\frac{a_0}{4})}$.
+Note que $\mathbf{a_0}=3.57$ \AA , ici c'est un coté du cube, les vecteurs primitifs (qui relient un coin au centre des trois faces voisines) sont de longueur $a_0/\sqrt{2}$.
+  \chapter{Le réseau réciproque}
+  Pour l'instant pas de mécanqiue quantique, on définit un réseau réciproque uniquement à partir.
+  \section{Définition}
+  Le réseau réciproque c'est l'ensemble des \textbf{k} tels que l'onde plane $e^{i\mathbf{k}\cdot\mathbf{r}}$ soit identique sur l'ensemble des points \textbf{R} du réseau direct, c'est à dire que $e^{i\mathbf{k}\cdot\mathbf{(r+R)}}=e^{i\mathbf{k}\cdot\mathbf{r}}$ ou bien encore $e^{i\mathbf{k}\cdot\mathbf{R}}=1$.
+
+  Tu t'en doute vu le titre, mais l'ensemble de ces \textbf{k} forment également un réseau de bravais (c'est à dire que c'est un groupe pour l'addition tout simplement), et l'expression des 3 vecteurs de base sont :
+  \begin{align*}
+  b_1&=2\pi \frac{a_2 \times a_3}{a_1 \cdot (a_2 \times a_3)} \\
+  b_2&=2\pi \frac{a_3 \times a_1}{a_2 \cdot (a_3 \times a_1)} \\
+  b_3&=2\pi \frac{a_1 \times a_2}{a_3 \cdot (a_1 \times a_2)}
+  \end{align*}
+  Qui doivent vérifier $b_i \cdot a_j = 2\pi \delta_{ij}$.
+Bref pour un cube c'est un cube de coté $\frac{2\pi}{a_0}$ (et les ondes sont les différentes harmoniques de chaque "corde vibrante" élementaire dans les 3 directions)
+
+  Donc le réseau réciproque c'est les \textbf{k}=$k_1\mathbf{b_1}+k_2\mathbf{b_2}+k_3\mathbf{b_3}$ (on confond vecteur en noeud mais bon tu vois l'idée)
     \end{document}

### 2020-11-17 data/20200722/traitement2.py

@@ -1,0 +1,398 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit,root_scalar
+from scipy import fftpack
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1):
+	x=[]
+	y=[]
+	with open(filename,'r',encoding = "ISO-8859-1") as f:
+		for line in f :
+			line=line.split()
+			try :
+				x+=[float(line[xcol])]
+				y+=[float(line[ycol])]
+			except :
+				pass
+	return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def quad_fit(x,y) :
+	A=np.vstack([x**2,x,np.ones(len(x))]).T
+	a,b,c = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([a,b,c],a*x**2+b*x+c)
+
+def fit_ordre_4(x,y) :
+	A=np.vstack([x**4,x**3,x**2,x,np.ones(len(x))]).T
+	a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([a,b,c,d,e],a*x**4+b*x**3+c*x**2+d*x+e)
+
+def fit_ordre_5(x,y) :
+	A=np.vstack([x**5,x**4,x**3,x**2,x,np.ones(len(x))]).T
+	z,a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([z,a,b,c,d,e],z*x**5+a*x**4+b*x**3+c*x**2+d*x+e)
+
+def fit_ordre_6(x,y) :
+	A=np.vstack([x**6,x**5,x**4,x**3,x**2,x,np.ones(len(x))]).T
+	w,z,a,b,c,d,e = np.linalg.lstsq(A, y, rcond=None)[0]
+	return([w,z,a,b,c,d,e],w*x**6+z*x**5+a*x**4+b*x**3+c*x**2+d*x+e)
+
+def fit_ordre_6_CI(x,y,a,b,c,d,e) :
+	p0=[0,0,a,b,c,d,e]
+	def f(x,w,z,a,b,c,d,e) :
[…]

### 2020-11-17 data/20201026/traitement.py

@@ -187,6 +187,8 @@ def ask_name():
 fname='scan_rose_EM_week_end_100'
 x,y=extract_data(fname+'.txt')
 
+y=y/max(y)
+
 V_min=0
 V_max=10
 n_V=len(x)
@@ -235,6 +237,22 @@ sig_fft[np.abs(sample_freq) > 5]=0
 
 filtered_sig = fftpack.ifft(sig_fft)
 plt.plot(x_plot,filtered_sig,label='filtre passe bas')
+ax=plt.gca()
+ylim=ax.get_ylim()
 
+c2 = next(ax._get_lines.prop_cycler)['color']
+
+x_v=54.5 #+/- 1
+plt.plot([x_v,x_v],[1,-1],'--',color=c2)
+
+x_v=121 #+/- 2
+plt.plot([x_v,x_v],[1,-1],'--',color=c2)
+
+
+x_C13=[17.953273372377286, 19.705551221857355, 24.31251459360322, 22.117301530072833]
+for x_v in x_C13 :
+	plt.plot([x_v,x_v],[1,-1],'--',color=c2)
+
+ax.set_ylim(ylim)
 plt.legend()
 plt.show()

### 2020-11-17 python/simu Hamiltonien NV/template.py

@@ -184,7 +184,7 @@ def NV_simple(orientation='111'):
 	# show_vpropres(val,vec,bnamez)
 	# transitions(val,vec,Sx,rho_0)
 	D_VH=2694 #+/-5 , Cruddace a 2706 +/- 30
-	D_War1=2466 #la meme valeur que Cruddace, lui il a +/- 60
+	D_War1=2468 #+/- 10 , Cruddace il a 2466 +/- 60
 	if orientation=='111' :
 		amps=np.linspace(0,400,201)
 		transi_NV=[]

### 2020-11-23 Talks/IQFA 2020/presentation.tex
toto

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage{graphicx}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage[backend=biber]{biblatex}

\addbibresource{Biblio_dbl_quantum.bib}

%\bibliographystyle{stylename}
%\bibliography{Biblio_dbl_quantum}

\title{Mechanical and relaxation-based detection of dipolar-interactions between spins in diamond}
\author{Clément Pellet-Mary\\ Laboratoire de physique de l'ENS\\ \textit{ENS, Paris}}
\date\today

\mode<presentation> {\usetheme{Rochester}}

\begin{document}
\begin{frame}
\maketitle
\end{frame}
\begin{frame}{Two experiments around cross-relaxations in diamond}
\includegraphics[scale=.3]{Shémas_intro}
\end{frame}
\begin{frame}{Outline}
\tableofcontents
\end{frame}
\section{Nitrogen Vacancy center presentation}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{Minimalist approach to NV center}
\includegraphics[scale=.28]{NV_prez_generale}
\end{frame}
\begin{frame}{Triplet spin levels at rest}
\includegraphics[scale=.4]{3_niveaux_1}
\end{frame}
\begin{frame}{Triplet spin levels with optical pumping}
\includegraphics[scale=.4]{3_niveaux_2}
\end{frame}
\begin{frame}{Magnetometry with NV centers : ODMR with a single spin}%"Efficient population transfer"
\includegraphics[scale=.65]{ESR_1spin_simu}
\end{frame}
\begin{frame}{Spin Hamiltonian and orientation of the centers}
\includegraphics[scale=.25]{projection_4_classes}
\end{frame}
\begin{frame}{ODMR with an ensemble of NV centers}
\includegraphics[scale=.65]{ESR_4spin_simu}
\end{frame}
\section{Cross-relaxations with new spin species}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{Photoluminescence change with magnetic field amplitude}
\includegraphics[scale=.62]{Scan_100_0}
\end{frame}
\begin{frame}{PL change with magnetic field : [100] direction}% "Cubic cell"
\includegraphics[scale=.62]{Scan_100_1}
\end{frame}
\begin{frame}{PL change with magnetic field : transverse field}
\includegraphics[scale=.62]{Scan_100_2}
\end{frame}
\begin{frame}{Subtracting the transverse field envelope}
\includegraphics[scale=.48]{soustraction_0}
\end{frame}
\begin{frame}{Cross-relaxation between NV$^-$ and VH$^-$}
\includegraphics[scale=.48]{soustraction_1}
\end{frame}
\begin{frame}{Cross-relaxation between NV$^-$ and WAR1}
\includegraphics[scale=.48]{soustraction_2}
\end{frame}
\begin{frame}{EPR spectroscopy}
Electron Paramagnetic Resonance : A spectroscopy technique using absorption of a microwave at a given frequency (typically 9.5 GHz) as a function of magnetic field to detect paramagnetic defects.
\bigbreak
VH$^-$\footfullcite{glover_hydrogen_2004} and WAR1 \footfullcite{cruddace2007magnetic} have been observed by EPR spectrocopy in CVD diamonds by Mark Newton's team at the University of Warwick.
\end{frame}
\begin{frame}{Comparison between CR and EPR}
\begin{itemize}
\item CR experiments are much simpler to setup (low $B$, no microwave)
\item NV centers produce a calibration for $B$ $\to$ Better precision on the ZFS measurement
\item Potential for polarization transfer
\bigbreak
\item Requires a high NV concentration and quickly becomes unreadable on non-[100] directions.
\end{itemize}
\end{frame}

\begin{frame}{Cross-relaxation between NV$^-$ and $^{13}$C$-$NV$^-$}
\includegraphics[scale=.48]{soustraction_3}
\end{frame}
\section{Cross-relaxation between NV centers}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{Experimental proofs of NV-NV CR}
\includegraphics[scale=.28]{NV-NV_0}
\end{frame}
\begin{frame}{Microscopic explanation of the NV-NV CR}
\centering
\includegraphics[scale=.35]{NVNV-micro}\footfullcite{choi_depolarization_2017}
\end{frame}
%ajouter la slide avec la carte du coup ?
\section{Torque induced by cross-relaxation on a levitating diamond}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{Torque measurement with a levitating diamond}
\centering
\includegraphics[scale=.3]{Trap_0}%"This trap not only locks the center of mass of the diamond, but it also locks its orientations, because of the anistropy of the trap and of the diamond" "The trap can be roughly modeled by an harmonic potential centered around a particular angle at equilibrium"
\end{frame}
\begin{frame}{Torque potential energy}%"The equivalent of a torque in term of potential energy is a slope"
\centering
\includegraphics[scale=.3]{Trap_1}
\end{frame}
\begin{frame}{Displacement of equilibrium}
\centering
\includegraphics[scale=.3]{Trap_2}
\end{frame}
\begin{frame}{Back-scattered laser detection}
\centering
\includegraphics[scale=.3]{Trap_3}%And how do we measure this rotation ? well, since we have to send a laser on the diamond to polarise the NV centers..." "We do not observe a clear reflection of the laser, instead we observe a speckle pattern, which is produced by the many interferences on the rough surface of the diamond" "Whenever the diamond moves, the speckle pattern will either shift or change its shape..."
\end{frame}
\begin{frame}{Torque caused by cross-relaxations}
\centering
\includegraphics[scale=.4]{CRmeca_121}
\end{frame}
\begin{frame}{Origin of the magnetic torque}
\centering
\includegraphics[scale=.3]{Explication_torque}
\end{frame}
\begin{frame}{Torque caused by CR : other configuration}
\centering
\includegraphics[scale=.28]{CRmeca_22}
\end{frame}
\begin{frame}{Conclusion}
\begin{itemize}
\setlength\itemsep{1em}
\item{Detection of new spin defects in diamond through Cross-Relaxations
\begin{itemize}
\item Potential for hyperpolariaztion of new dark electronic spins
\end{itemize}}
\item{Observations of NV$-$NV Cross-Relaxations}
\item{Mechanical detection of NV$-$NV Cross relaxations :
\begin{itemize}
\item Potential for mechanical detection of new spin species
\item Potential for Resonant Einstein-De-Haas effect 
\end{itemize} }
\end{itemize}
\end{frame}
\end{document}

### 2020-11-23 théorie/fiches de lecture/Ashcroft-Mermin/fiche.tex
toto

@@
   Donc le réseau réciproque c'est les \textbf{k}=$k_1\mathbf{b_1}+k_2\mathbf{b_2}+k_3\mathbf{b_3}$ (on confond vecteur en noeud mais bon tu vois l'idée)
+  \section{Première zone de Brillouin}
+  La définition est simple : \textbf{La première zone de Brillouin (PZB) est la maille de Wigner-Seitz du réseau réciproque}
+
+  Un truc important à comprendre c'est que la PZB est énorme dans l'espace des k en faite : vu les conditions $b_i \cdot a_j = 2\pi \delta_{ij}$, ca veut dire que si $a_j$ est de l'ordre de quelques \AA , alors $b_i$ est de l'ordre de quelques $10^{10}$ m$^{-1}$.
+Les points du réseau réciproque sont très très espacés (avec des longueurs d'onde associées égales aux distance atomiques, donc dans les rayons X).
+
+  De façon générale, le volume de la PZB est $$V_k = \frac{(2\pi)^3}{v}$$ ou $v$ est \textbf{le volume de la maille élémentaire (en espace réel)}.
+J'insiste parce que naturellement j'ai envie de considérer le volume total du solide, mais ça n'a pas de sens ici ou on parle uniquement de réseaux (infinis).
+La taille finie du cristal elle va jouer sur la discrétisation des phonons dans la PZB (spoils).
+  \section{Plans réticulaires}
+  Les plans réticulaires sont simplement définis dans le réseau de Bravais par un plan contenant 3 points du réseau non alignés.
+On peut alors toujours crééer une famille de ces plans réticulaires, équidistants d'une distance $d$, qui contiennent l'ensemble des points du réseau de Bravais.
+les vecteurs normaux à ces plans sont alors toujours alignés avec des points du réseau réciproque, et le plus court vecteur du réseau réciproque (deux points consécutifs en gros) est de longueur $2\pi/d$.
+
+  Inversement, pour toute droite dans la réseau de Bravais, dont deux points consécutifs sont séparés d'une distance $2\pi/d$, il existe une famille de plans réticulaires normaux à ce vecteurs, et séparés d'une distance $d$.
+  \subsection{Indices de Miller des plans réticulaires}
+  Vu qu'il y a une équivalence entre plans réticulaires et plus court vecteur du réseau réciproque, tu peux référencer une maille de plans réticulaires par 3 indices entiers (h,k,l) tels que $$ \mathbf{k} = h \mathbf{b1} + k \mathbf{b2} + l \mathbf{b3} $$.
+
+  Au passage tu vois que pour que ce soit le plus petit vecteur possible, il suffit que le PGCD de h,k, et l soit 1.
+
+  \textbf{Remarque :} Attention quand même (je me posais la questions) dans les réseaux CC et CFC, on considère en général la maille conventionelle (donc cubique simple, et son RR cubique simple également) pour parler des indices de Miller, plutot que la maille primitive qui donnerait des trucs chelous.
+Donc c'est bien la définition intuitive les plans (100) et (111) dans le diamant.
+Par contre pour l'hexagonal compact par exemple t'es obligé de te faire le vrai truc.
+
+  \subsection{Notations (si c'est encore d'actualité)}
+  \begin{itemize}
+  \item (111) la famille de plans d'indice de Miller h=1, k=1, l=1. Pour de nombres négatifs on utilisera plutot ($1\bar 1 1$). Fait référence au réseau réciproque.
+  \item $[111]$ : La direction dans le réseau direct définie par le vecteur $\mathbf{r} = 1 \mathbf{a1} + 1 \mathbf{a2} + 1 \mathbf{a3} $
+  \item $\{ 100 \}$ : Les familles de plans équivalentes (100), (010) et (001).
+  \item $\langle 100 \rangle$ : Les directions équivalentes [100], [010], [001], [$\bar 1 00$], [$0\bar 1 0$], [$00\bar 1$]
+  \end{itemize}
     \end{document}

### 2020-11-23 data/20200806/1-2-1/traitement.py

@@ -127,7 +127,7 @@ def ask_name():
 plt.figure(num=1,figsize=(11,7),dpi=80)
 ax=plt.gca()
 fname='0-0.8V_PLdown-Db'
-x,y=extract_data(fname+('.txt'),ycol=1)
+x,y=extract_data(fname+('.txt'),ycol=3)
 y=y/max(y)
 xmin=100
 xmax=165
@@ -136,7 +136,7 @@ xplotmin=80
 x=(x-x[xplotmin])*31
 
 
-ax.plot(x[xplotmin:],y[xplotmin:],'x',ms=10,mew=3)
+ax.plot(x[xplotmin:],y[xplotmin:],'o',markerfacecolor="None",ms=10,mew=3)
 
 popt,yfit=gauss_fit(x[xmin:xmax],y[xmin:xmax])
 ax.plot(x[xmin:xmax],yfit,linewidth=5)

### 2020-11-27 Mes papiers/Double quantum/pre_SI.tex
manger

\documentclass{article}
%% Please use 11pt if submitting to AOP
% \documentclass[11pt,twocolumn,twoside]{osajnl}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}

\begin{document}

\section{NV-$^{13}$C Hamiltonian}
The full spin Hamiltonian of the NV-$^{13}$C complex can be written as follow : 
\begin{equation*}
\mathcal{H}=\mathcal{H}_{NV}+\mathcal{H}_{^{13}C}+\mathcal{H}_{HF}
\end{equation*}
Where $\mathcal{H}_{NV}$ is the NV$^-$ spin Hamiltonian described in XX, $\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian : $\mathcal{H}_{HF}= \hat{\mathbf{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\mathbf{I}}_C$.

In the case of first shell $^{13}$C, the hyper-fine tensor $\mathcal{A}$ can be written as \citep{simanovskaia_sidebands_2013} : $$ \mathcal{A} = \begin{pmatrix}
\mathcal{A}_{xx} & 0 & \mathcal{A}_{xz} \\ 0 & \mathcal{A}_{yy} & 0 \\ \mathcal{A}_{zx} & 0 & \mathcal{A}_{zz}
\end{pmatrix} $$

Where $\mathcal{A}_{xx}=190.2(2)$ MHz, $\mathcal{A}_{yy}=120.3(2)$ MHz, $\mathcal{A}_{zz}=129.1(2)$ MHz, and  $\mathcal{A}_{xz}=\mathcal{A}_{zx}=-25.0(1)$ MHz.

Diagonalizing the total Hamiltonian, we notice that the quantization axis of the nuclear spin (in the limit $\gamma_{n} B << A_{ij}$) is not the same in the manifold of the $\ket{m_s=0}$ state and in the $\ket{m_s=\pm 1}$ states \citep{alvarez_local_2015}, meaning that the nuclear spin is not preserved by the electron spin flip, and giving rise to a splitting of the $\ket{m_s=0} \to \ket{m_s=\pm 1}$ transition in 4 distinct lines when the magnetic field is not aligned with the NV center. \citep{jiang2018estimation}

\section{NV-P1 mutual flip transitions}
Transitions corresponding to a simultaneous flip of the NV$^-$ and P1 spin states, mediated by the dipolar interaction between NV$^-$ and P1 electronic spins, are commonly observed on ODMR spectra \citep{simanovskaia_sidebands_2013} \citep{kamp2018continuous} \citep{alfasi2019detection} \citep{lazda2020cross}. We investigated whether the cross-relaxations processes we observed could be due a three body interaction between two NV centers and a P1 center.

In order for this process to be energy conservative, the P1 transition frequency has to match the energy difference between the $\ket{m_s=0} \to \ket{m_s=-1}$ and $\ket{m_s=0} \to \ket{m_s=+1}$ transitions, giving us the equation \begin{equation}
\label{eq_P1}
\nu^i_{P1}(B)=\nu^{0 \to -1}_{NV}(B)+\nu^{0 \to +1}_{NV}(B)
\end{equation}
Where $\nu^i_{P1}$ is a transition between any of the P1 spin Hamiltonian eigenstates. Note that since we are scanning the magnetic field in the crystalline [100] direction, we do not have to take into account the fours classes of NV centers and P1.

The P1 spin states is defined by its  electron 1/2-spin and its nuclear 1-spin, giving a manifold of 6 spin states. When considering any possible transitions between the 6 eignestates of the P1 Hamiltonian, we have therefore a total of 15 possible transitions, including the nuclear-like transitions which have been observed through mutual flip with NV centers \citep{alfasi2019detection}.

Solving eq. \ref{eq_P1} for all possible P1 transitions (see fig. \ref{fig_P1}) therefore gives us all the possible magnetic field amplitude in the [100] direction where we could observe NV-P1 mutual flip transitions. The predicted magnetic field values are 0, 3.89, 5.96, 6.58, 17.90, 28.93, 35.87, 49.44, 81.33, 83.28, 137.52, 154.20 and 246.34 G. None of these values correspond to one of the feature our scans. We then concluded that the Features we observed were not due to NV-P1 mutual flips. 
\begin{figure}
\label{fig_P1}
\includegraphics[scale=.6]{Transis_P1}
\caption{Cross-relaxation condition for mutual flip between NV and P1 centers. The orange plain lines correspond to the NV center $\ket{m_s=0} \to \ket{m_s=-1}$ and $\ket{m_s=0} \to \ket{m_s=+1}$ transition frequencies as function of a magnetic field along the [100] crystalline axis, the blue dashed lines correspond to the frequencies of the 15 theoretical transitions of the P1 centers, added to the frequency of the NV $\ket{m_s=0} \to \ket{m_s=-1}$ transition. The green circles correspond to the particular magnetic fields where eq. \ref{eq_P1} is verified, meaning the energy of the P1 transition matches the difference of energy between the two NV transitions.}
\end{figure}
% Bibliography
\bibliographystyle{plain}

\section{Scan in the [111] direction}
\begin{figure}
\label{scan_PL}
\includegraphics[scale=.48]{scan_111_PL}
\caption{\textbf{a)} Plain blue line : Photoluminescence change as a function of the magnetic field amplitude in the [111] crystalline direction. Dashed orange line : order-4 polynomial fit of the decreasing envelope. \textbf{b)} Subtraction of the previous signal by the envelope fit. The vertical dotted orange lines correspond to the expected cross-relaxations with VH$^-$ and the vertical dashed green lines to the expected cross-relaxations with WAR1.}
\end{figure}
\begin{figure}
\label{transis_VHWAR}
\includegraphics[scale=.48]{Transis_111_VHWAR}
\caption{Predicted transitions frequencies for NV centers (Plain lines), VH$^-$ (dashed lines) and WAR1 (dashdotted lines) when the magnetic field is scanned in the [111] direction. The magnetic field where a transition of the NV center crosses one of VH$^-$ or WAR1 are reported on fig \ref{scan_PL}}
\end{figure}

\begin{figure}
\label{transis_13C}
\includegraphics[scale=.55]{Transis_111_13C}
\caption{Predicted transitions frequencies for NV centers (Plain lines) and NV-$^{13}$C complex when the magnetic field is scanned in the [111] direction. Vertical lines represent the various degeneracy conditions.}
\end{figure}

We have performed similar magnetic field scan as presented in Fig.3 of the main text, but this time scanning the magnetic field in the crystalline [111] direction instead of the [100], meaning that this time one class of NV centers is aligned with the magnetic field while the three others are disaligned and degenerated. The results in term of cross-relaxations generally means 4 times as many degeneracy conditions as the [100] case, since the two spins involved in the cross-relaxation process now have two possible orientations, and therefore two separate transitions.

It is then significantly harder to find cross-relaxations signatures, since they are both more numerous and less contrasted than in the [100] case. The experimental results are shown in fig. \ref{scan_PL}. Similarly to fig. 3 of the main text, we have subtracted the envelope associated with the transverse field in order to better see the fine features of the scan. The predicted cross-relaxation conditions between NV centers and VH$^-$ or WAR1 are presented in fig. \ref{transis_VHWAR} and reported on fig. \ref{scan_PL}. While the predicted values do not exactly match the experimental results, it seems like every predicted transition falls in the vicinity of an experimental line. Importantly, it should be noted that, unlike Fig.3 of the main text, the magnetic field was not calibrated with NV ODMR at every field value in this case. Instead we have simply applied an homotethy on the current intensity in the electromagnet to match the field values at the maxima of the scan. This means that the magnetic field of the points of low field value in particular are more susceptible to the non-linearities due to the electromagnet hysteresis cycle.

Finally we have plotted on fig.\ref{transis_13C} the expected cross-relaxations between NV and NV-$^{13}$C complex. It seems that these transitions are too diluted to appear in our experimental scan. The peak at $\sim$37 G in fig.\ref{scan_PL} which in the only peak unpaired with any of the NV-VH$^-$ or NV-WAR1 resonance could come from the resonance between NV and NV-$^{13}$C since they seem particularly dense at this particular field values.
\bibliography{Biblio_CVD-CR}

% Full bibliography added automatically for Optics Letters submissions; the following line will simply be ignored if submitting to other journals.
% Note that this extra page will not count against page length

\end{document}

### 2020-12-02 Talks/Défaut 2020/Script.tex
toto

--- 2020-11-17 Talks/IQFA 2020/Script.tex
@@
 \section{NV center}
+$\mathcal{H}_{dd} \approx - \frac{\mu_0}{4\pi r^3} \vec{\mu_1}\cdot \vec{\mu_2}$
 %Slide avec un NV dans sa maille cristalline et les niveaux d'énergie (juste excité
@@
 %1ere slide : Scan + en dessous 100 et puis soustraction par dessus.
-What we see here is the change in photoluminescence of our NV centers when we scan a magnetic field along the 100 direction of the diamond. I'll start detailing what we observed.
+What we see here is the change in photoluminescence of our NV centers when we scan a magnetic field along the 100 direction of a CVD grown irradiated diamond with about 5ppm of NV centers.
+I'll start detailing what we observed.
 
@@
 So we expect a decreasing enveloppe (slide), and when we subtract the enveloppe of the signal, what remains is this :(new slide avec la soustraction) three clear features, none of which at been observed before and that we attributed to cross-relaxation of the NV center with three different species.
-The first
+The peak at 55 G is due to another spin defects : VH- (niveaux d'énergie sur la slide d'en dessous).
+VH- is a spin 1 defect with a spin structure close to that of NV- but is a dark spin.
+It was observed previously in EPR measurement for other CVD grown diamond, but never optically.
+(article ?).
+We can see that its +1 to 0 transition crosses the -1 to 0 transition at 55 G.
+We can also see its signature in scan along 111 although the signal is not as clear.
+The scan at 123 G is yet another spin defect, which was probably observed before and named WAR1 , because it was the first defect detected at the university of Warwick, again with EPR technique.
+We do not know its chemical nature, but it is another spin 1 defect with a slightly lower zero field splitting.
+Finally the peak at 20G is not exactly a new defect because this are NV centers, but with one of its neighboring carbon being a carbon 13 nucleus.
+Carbon 12, which has a 99\% nutural abundance, does not have a nuclear spin, but Carbon 13 which is about 1\% abundant does have a one half nuclear spin.
+Most of the carbon-13 spins, and other spin defects in the diamond, will contibute to the inhomogenous broadening of the NV centers.
+However when a carbon-13 is a direct neighbor, we say there is a first shell carbon-13, the hyper-fine coupling to the electronic spin of the NV center is strong enough that we can observe resolved/distinct sidebands (ESR with distinct sidebands ?), separating the carbon-13 first shell NV centers to the other NV centers.
+Based on ???, there are four possible values for the hyper fine coupling energy splitting, which leads to four possible cross-relaxation conditions that are somewhat confounded in the 20G peak.
+However, this peak is different from the last ones, because in the previous cases we were looking at cross-relaxation between a bright polarized spin and dark unpolarized spins, so we were expecting a drop in the photolmuniescence.
+However, here, we are looking at the cross-relaxation between two bright polarized spins, so why do we still observe a drop in the PL ?
+This brings me to our second results.
 
+\section{NV-NV cross relaxation}
+Before discussing our levitation experiment, I will briefly discuss what we observed for the first shell carbon-13 cross-relaxation.
+The fact that NV centers will spontaneously depolarize when brought to resonance with other NV centers is not particularly linked to the carbon-13.
+This is a phenomenon that has been observed and studied by multiple groups (Budker, les polaks, Lukin), and withouth going into to much details, this probably due to some NV centers having a much shorter spin lifetime than the average NV centers, most likely because their electron tunnels in and out of the NV site, which causes a depolarization of the ensemble of spins through dipolar interaction.
+This phenomoenon is only clearly observed on dense ensemble of more than 1 ppm of NV centers.
+
+Here are two experimental proofs that we recorded on another sample, a micro-diamond similar to the one we use in our levitation experiment.
+On the left side you can look at the change in fluorescence when we scan the magnetic field in a random direction with a magnetic field offset.
+You can see in the bottom panel the energy level of the four different classes of NV, that we recorded through ODMR, and you can notice that the drops in photoluminescence of the NV centers correspond to the part were several classes are at degeneracy.
+In our explanation, this because when two or more classes are at degeneracy, the number of NV centers being able to resonantly exchange spin quanta increases, therefore the depolarization by the flucuators increases and we observe cross-relaxations.
+
+The experiment on the right side is a lifetime measurement for various resonance conditions that confirm what we observed in the cross-relaxation on the left.
+The top curve correspond to the case where all four classes are clearly separated, and the measured lifetime of about 1 ms is already shorter than the phonon-limited lifetime of 5ms that we observe in dilute ensemble.
+The second curve correspond to the case where two classes are degenerated and we observe a clear decrease to .25 ms.
+Finally the bottom curve is the 100 alignment where all four classes are at degeneracy and we observe an increased reduction of the lifetime to .15 ms.
+
+\section{Torque measurement with a levitating diamond}
+The second ingredient to explain our levitating experiment is how we can measure a torque thanks to our levitating diamond.
+
+In order to trap the center of mass of our diamonds, we use a Paul trap which is an electro-static based trap, mostly used to trap ions, but since there are always a few patch charges on the surface of objects, we can trap any micrometric objects, including in this case 15 microns diameter diamonds.
+
+A side effect to this trap is that because of the anisotropy of shape of the diamond, and potentially of the charges on its surface, and because the anisotropy of the trap, the angular orientation of the diamond is also confined.
+
+If we look at the potential angular energy, the action of the trap will look something like this : a harmonic potential centered around the equilibrium angular position.
+Now if we apply a static torque on the diamond, which can be represented as a pure slope in the potential energy, we see that the bottom of the energy well is displaced, meaning that the diamond will slightly rotate to a new angular position.
+
+In order to detect this rotation, we simply shine a laser on the diamond and, which also serve to excite the NV centers, and look at its reflection off the diamond.
+Since the diamond's facettes are of the order of 1 microns, we do not observe a clear refelction of the laser beam, however we observe a speckle pattern, and whenever the diamond moves, the pattern will change.
+So by collecting the light on a particular bright or dark spot of the speckle, we will observe either an increase or a decrease in the signal whenever the diamond rotates.
+
+%Slide avec du speckle et un carré rouge sur la collection
+
+\section{Magnetic torque caused by the cross-relaxation}
+With all this being discussed, I can show you the result of our experiment.
+What we did was again to scan a magnetic field with an offset in order to observe NV-NV cross-relaxation, but this time in levitation.
+And indeed we observe again a drop in the photoluminescence of the NV centers when two classes are brought to resonance.
+But when looking at the diamond orientation, we can also see a movement around the resonance, indicating a torque (a microwave-free torque) at the cross-relaxation.
+
+The origin of the torque is in fact due to the natural magnetization of the spins, and by natural I mean the magnetization you get when polarizing your spin with a green laser but without any microwave.
+This magnetization depends mostly on the transverse field, and will be different for every class of NVs.
+Then by looking at the expected torque gamma times the average S value cross the magnetic field, here is what we get for the four classes of NV for an arbitrary direction $\Gamma_x$.
+When we take the sum of all four contributions, we get a total torque which is almost zero, due to directional averaging by the four classes.
+However, when we take into consideration the depolarization due to the cross-relaxation between NVs, we can reduce the contribution of two of the four classes for a specific magnetic field, and therefore we observe a non-zero torque specifically at the cross-relaxation transition.
+
+\section{Conclusion and outlook}
+-Observation of two news electronic spin defects in diamond through cross-relaxation : new candidates for electronic hyperpolarization.
+
+-Observation of NV-NV cross-relaxation : implications for magnetometry with ensembles
+
+-Mechanical detection of cross-relaxations : - first observation of the "natural" magnetic torque of diamonds, -potential for mechanical detection of new spin defects - and Einstein de Haas effect with paramagnetic systems.
   \end{document}

### 2020-12-02 Talks/Défaut 2020/presentation.tex
toto

--- 2020-11-23 Talks/IQFA 2020/presentation.tex
@@
 \end{frame}
-\section{Nitrogen Vacancy center presentation}
+\section{NV center and Cross-Relaxations}
 \begin{frame}{Outline}
 \tableofcontents[currentsection]
-\end{frame}
-\begin{frame}{Minimalist approach to NV center}
-\includegraphics[scale=.28]{NV_prez_generale}
-\end{frame}
-\begin{frame}{Triplet spin levels at rest}
-\includegraphics[scale=.4]{3_niveaux_1}
-\end{frame}
-\begin{frame}{Triplet spin levels with optical pumping}
-\includegraphics[scale=.4]{3_niveaux_2}
-\end{frame}
-\begin{frame}{Magnetometry with NV centers : ODMR with a single spin}%"Efficient population transfer"
-\includegraphics[scale=.65]{ESR_1spin_simu}
 \end{frame}
@@
 \end{frame}
+\begin{frame}{Non-Resonant dipolar interaction}
+\centering
+\includegraphics[scale=.35]{CR_non_resonnant}
+\end{frame}
+\begin{frame}{Resonant dipolar interaction : Cross-Relaxation}
+\centering
+\includegraphics[scale=.35]{CR_resonnant}
+\end{frame}
+\begin{frame}{Exemple of C.R. : scan along the [111]}
+\includegraphics[scale=.30]{CR_111}\footfullcite{armstrong_nvnv_2010}
+\end{frame}
 \section{Cross-relaxations with new spin species}
@@
 \end{frame}
-\begin{frame}{Photoluminescence change with magnetic field amplitude}
-\includegraphics[scale=.62]{Scan_100_0}
+\begin{frame}{Motivations of the study}
+\includegraphics[scale=.3]{Motivations}\footfullcite{tallaire2020high}
 \end{frame}
-\begin{frame}{PL change with magnetic field : [100] direction}% "Cubic cell"
-\includegraphics[scale=.62]{Scan_100_1}
-\end{frame}
-\begin{frame}{PL change with magnetic field : transverse field}
-\includegraphics[scale=.62]{Scan_100_2}
-\end{frame}
-\begin{frame}{Subtracting the transverse field envelope}
-\includegraphics[scale=.48]{soustraction_0}
+\begin{frame}{Photoluminescence change with $\vec B$ in the [100] direction}
+\centering
+\includegraphics[scale=.45]{Scan_100_brut}
 \end{frame}
@@
 \begin{frame}{EPR spectroscopy}
-Electron Paramagnetic Resonance : A spectroscopy technique using absorption of a microwave at a given frequency (typically 9.5 GHz) as a function of magnetic field to detect paramagnetic defects.
+Electron Paramagnetic Resonance : A spectroscopy technique using absorption of a microwave at a given frequency (9.5 GHz for example) as a function of magnetic field to detect paramagnetic defects.
 \bigbreak
@@
 \item CR experiments are much simpler to setup (low $B$, no microwave)
+\item Doable on micro and potentially nano-diamonds
 \item NV centers produce a calibration for $B$ $\to$ Better precision on the ZFS measurement
@@
 \bigbreak
+\bigbreak
 \item Requires a high NV concentration and quickly becomes unreadable on non-[100] directions.
 \end{itemize}
-\end{frame}
-
-\begin{frame}{Cross-relaxation between NV$^-$ and $^{13}$C$-$NV$^-$}
-\includegraphics[scale=.48]{soustraction_3}
 \end{frame}
@@
 \end{frame}
+\begin{frame}{Cross-relaxation between NV$^-$ and $^{13}$C$-$NV$^-$}
+\includegraphics[scale=.45]{soustraction_3}
+\end{frame}
 %ajouter la slide avec la carte du coup ?
@@
 \end{frame}
-\begin{frame}{Torque potential energy}%"The equivalent of a torque in term of potential energy is a slope"
-\centering
-\includegraphics[scale=.3]{Trap_1}
-\end{frame}
-\begin{frame}{Displacement of equilibrium}
-\centering
-\includegraphics[scale=.3]{Trap_2}
-\end{frame}
 \begin{frame}{Back-scattered laser detection}
@@
 \end{frame}
-\begin{frame}{Torque caused by cross-relaxations}
+\begin{frame}{Measurement of the torque caused by C.R.}
 \centering
-\includegraphics[scale=.4]{CRmeca_121}
+\includegraphics[scale=.08]{CRmeca_121_mieux}
 \end{frame}
 \begin{frame}{Origin of the magnetic torque}
+\centering
+\includegraphics[scale=.23]{Explication_torque_alalouche}
+\end{frame}
+\begin{frame}{Simulation of the C.R.-caused torque}
 \centering

### 2020-12-04 Talks/Défaut 2020/presentation.tex
totto

@@
 
-\title{Mechanical and relaxation-based detection of dipolar-interactions between spins in diamond}
+\title{Mechanical and luminescence-based detection of dipolar-interactions between spins in diamond}
 \author{Clément Pellet-Mary\\ Laboratoire de physique de l'ENS\\ \textit{ENS, Paris}}
@@
 \end{frame}
-\begin{frame}{Two experiments around cross-relaxations in diamond}
+\begin{frame}
 \includegraphics[scale=.3]{Shémas_intro}
-\end{frame}
-\begin{frame}{Outline}
-\tableofcontents
 \end{frame}
@@
 \begin{frame}{Spin Hamiltonian and orientation of the centers}
-\includegraphics[scale=.25]{projection_4_classes}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\includegraphics[scale=.25]{projection_4_classes_0}
+    \onslide<2>\includegraphics[scale=.25]{projection_4_classes_1}
+    \end{overprint}
+\end{figure}
 \end{frame}
@@
 \end{frame}
-\begin{frame}{Exemple of C.R. : scan along the [111]}
-\includegraphics[scale=.30]{CR_111}\footfullcite{armstrong_nvnv_2010}
+\begin{frame}{Exemple of Cross-Relaxation NV-P1}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\includegraphics[scale=.30]{CR_111_0}\footfullcite{armstrong_nvnv_2010}
+    \onslide<2>\includegraphics[scale=.30]{CR_111_05}
+    \onslide<3>\includegraphics[scale=.30]{CR_111_1}
+    \onslide<4>\includegraphics[scale=.30]{CR_111_2}
+    \end{overprint}
+\end{figure}
 \end{frame}
@@
 \begin{frame}{Motivations of the study}
-\includegraphics[scale=.3]{Motivations}\footfullcite{tallaire2020high}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\includegraphics[scale=.30]{Motivations_0}
+    \onslide<2>\includegraphics[scale=.30]{Motivations_1}\footfullcite{tallaire2020high}
+    \onslide<3>\includegraphics[scale=.30]{Motivations_2}
+    \onslide<4>\includegraphics[scale=.30]{Motivations_3}
+    \end{overprint}
+\end{figure}
 \end{frame}
 \begin{frame}{Photoluminescence change with $\vec B$ in the [100] direction}
-\centering
-\includegraphics[scale=.45]{Scan_100_brut}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\centering\includegraphics[scale=.45]{Scan_100_brut_0}
+    \onslide<2>\centering\includegraphics[scale=.45]{Scan_100_brut_1}
+    \onslide<3>\centering\includegraphics[scale=.45]{Scan_100_brut_2}
+    \end{overprint}
+\end{figure}
 \end{frame}
@@
 \end{frame}
-\begin{frame}{Comparison between CR and EPR}
+\begin{frame}{Comparison between cross-relaxation and EPR}
+\pause
+Upsides :
 \begin{itemize}
-\item CR experiments are much simpler to setup (low $B$, no microwave)
+\item CR experiments are much simpler to setup (low $\vec B$, no microwave)
+\pause
 \item Doable on micro and potentially nano-diamonds
-\item NV centers produce a calibration for $B$ $\to$ Better precision on the ZFS measurement
-\item Potential for polarization transfer
-\bigbreak
-\bigbreak
-\item Requires a high NV concentration and quickly becomes unreadable on non-[100] directions.
+\pause
+\item NV centers produce a calibration for $\vec B$ $\to$ Better precision on the ZFS measurement
+\pause
+\item Potential for hyperpolarization of new spins
+\pause
+\end{itemize}
+Downsides :
+\begin{itemize}
+\item Requires a high NV concentration
+\pause
+\item Loss of contrast in non-[100] directions
 \end{itemize}
@@
 \end{frame}
-\begin{frame}{Experimental proofs of NV-NV CR}
-\includegraphics[scale=.28]{NV-NV_0}
+\begin{frame}{Experimental proofs of NV-NV cross-relaxations}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\includegraphics[scale=.28]{NV-NV_0}
+    \onslide<2>\includegraphics[scale=.28]{NV-NV_05}
+    \onslide<3>\includegraphics[scale=.28]{NV-NV_1}
+    \onslide<4>\includegraphics[scale=.28]{NV-NV_2}
+    \onslide<5>\includegraphics[scale=.28]{NV-NV_3}
+    \end{overprint}
+\end{figure}
 \end{frame}
@@
 \centering
-\includegraphics[scale=.3]{Trap_3}%And how do we measure this rotation ?
+\includegraphics[scale=.085]{Trap_3}%And how do we measure this rotation ?
 well, since we have to send a laser on the diamond to polarise the NV centers..." "We do not observe a clear reflection of the laser, instead we observe a speckle pattern, which is produced by the many interferences on the rough surface of the diamond" "Whenever the diamond moves, the speckle pattern will either shift or change its shape..."
@@
 \begin{frame}{Measurement of the torque caused by C.R.}
-\centering
-\includegraphics[scale=.08]{CRmeca_121_mieux}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\includegraphics[scale=.08]{CRmeca_121_mieux_0}
+    \onslide<2>\includegraphics[scale=.08]{CRmeca_121_mieux_1}
+    \onslide<3>\includegraphics[scale=.08]{CRmeca_121_mieux_2}
+    \end{overprint}
+\end{figure}
 \end{frame}
 \begin{frame}{Origin of the magnetic torque}
-\centering
-\includegraphics[scale=.23]{Explication_torque_alalouche}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\includegraphics[scale=.23]{Explication_torque_alalouche_0}
+    \onslide<2>\includegraphics[scale=.23]{Explication_torque_alalouche_1}
+    \onslide<3>\includegraphics[scale=.23]{Explication_torque_alalouche_2}
+    \onslide<4>\includegraphics[scale=.23]{Explication_torque_alalouche_3}
+    \onslide<5>\includegraphics[scale=.23]{Explication_torque_alalouche_4}
+    \onslide<6>\includegraphics[scale=.23]{Explication_torque_alalouche_5}
+    \end{overprint}
+\end{figure}
 \end{frame}

### 2020-12-09 Talks/Défaut 2020/presentation.tex
papa

@@
     \begin{overprint}
-    \onslide<1>\includegraphics[scale=.25]{projection_4_classes_0}
-    \onslide<2>\includegraphics[scale=.25]{projection_4_classes_1}
+    \onslide<1>\centering\includegraphics[scale=.25]{projection_4_classes_0}
+    \onslide<2>\centering\includegraphics[scale=.25]{projection_4_classes_1}
     \end{overprint}
 \end{figure}
-\end{frame}
-\begin{frame}{ODMR with an ensemble of NV centers}
-\includegraphics[scale=.65]{ESR_4spin_simu}
 \end{frame}
@@
     \begin{overprint}
-    \onslide<1>\includegraphics[scale=.28]{NV-NV_0}
-    \onslide<2>\includegraphics[scale=.28]{NV-NV_05}
-    \onslide<3>\includegraphics[scale=.28]{NV-NV_1}
-    \onslide<4>\includegraphics[scale=.28]{NV-NV_2}
-    \onslide<5>\includegraphics[scale=.28]{NV-NV_3}
+    \onslide<1>\includegraphics[scale=.24]{NV-NV_1}
+    \onslide<2>\includegraphics[scale=.24]{NV-NV_2}
     \end{overprint}
@@
 \end{frame}
-\begin{frame}{Microscopic explanation of the NV-NV CR}
+\begin{frame}{Explanation of the NV-NV cross-relaxations}
+\fullcite{jarmola_temperature-_2012}
+
+\fullcite{mrozek_longitudinal_2015}
+
+\fullcite{choi_depolarization_2017}
+
+\fullcite{giri_coupled_2018}
+
+\fullcite{akhmedzhanov_magnetometry_2019}
+
+\pause
 \centering
-\includegraphics[scale=.35]{NVNV-micro}\footfullcite{choi_depolarization_2017}
-\end{frame}
-\begin{frame}{Cross-relaxation between NV$^-$ and $^{13}$C$-$NV$^-$}
-\includegraphics[scale=.45]{soustraction_3}
+\includegraphics[scale=.35]{NVNV-micro_2}
 \end{frame}
 %ajouter la slide avec la carte du coup ?
-\section{Torque induced by cross-relaxation on a levitating diamond}
+\section{Mechanical detection of cross-relaxations with a levitating diamond}
 \begin{frame}{Outline}
 \tableofcontents[currentsection]
+\end{frame}
+\begin{frame}{Objectives of the experiment}
+\begin{figure}
+    \begin{overprint}
+    \onslide<1>\includegraphics[scale=.23]{Explication_torque_alalouche_0}
+    \onslide<2>\includegraphics[scale=.23]{Explication_torque_alalouche_1}
+    \onslide<3>\includegraphics[scale=.23]{Explication_torque_alalouche_2}
+    \onslide<4>\includegraphics[scale=.23]{Explication_torque_alalouche_3}
+    \onslide<5>\includegraphics[scale=.23]{Explication_torque_alalouche_4}
+    \onslide<6>\includegraphics[scale=.23]{Explication_torque_alalouche_5}
+    \end{overprint}
+\end{figure}
 \end{frame}
@@
 \end{frame}
-\begin{frame}{Origin of the magnetic torque}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\includegraphics[scale=.23]{Explication_torque_alalouche_0}
-    \onslide<2>\includegraphics[scale=.23]{Explication_torque_alalouche_1}
-    \onslide<3>\includegraphics[scale=.23]{Explication_torque_alalouche_2}
-    \onslide<4>\includegraphics[scale=.23]{Explication_torque_alalouche_3}
-    \onslide<5>\includegraphics[scale=.23]{Explication_torque_alalouche_4}
-    \onslide<6>\includegraphics[scale=.23]{Explication_torque_alalouche_5}
-    \end{overprint}
-\end{figure}
+\begin{frame}{Conclusion}
+\begin{tcolorbox}[title=Summary,colbacktitle=red!80!white]
+\begin{itemize}
+\item{Optical detection of new spin defects in CVD diamond through cross-relaxations with NV centers}
+\item{Mechanical detection of NV magnetization enhanced by cross-relaxations}
+\end{itemize}
+\end{tcolorbox}
+\pause
+\begin{tcolorbox}[title=Prospects,colbacktitle=blue!80!white]
+Observation of cross-relaxation between NV and another spin in a levitating (CVD ?) diamond.
+\begin{itemize}
+\item Mechanical detection of a non-NV spin thanks to hyperpolarization
+\item Angular momentum conservation with non-spin preserving dipolar coupling (Einstein-de Haas effect)
+\end{itemize}
+\end{tcolorbox}
+\end{frame}
+\begin{frame}{Cross-relaxation between NV$^-$ and $^{13}$C$-$NV$^-$}
+\includegraphics[scale=.45]{soustraction_3}
 \end{frame}
@@
 \end{frame}
-\begin{frame}{Conclusion}
-\begin{itemize}
-\setlength\itemsep{1em}
-\item{Detection of new spin defects in diamond through Cross-Relaxations
-\begin{itemize}
-\item Potential for hyperpolariaztion of new dark electronic spins
-\end{itemize}}
-\item{Observations of NV$-$NV Cross-Relaxations}
-\item{Mechanical detection of NV$-$NV Cross relaxations :
-\begin{itemize}
-\item Potential for mechanical detection of new spin species
-\item Potential for Resonant Einstein-De-Haas effect 
-\end{itemize} }
-\end{itemize}
-\end{frame}
 \end{document}

### 2021-01-05 Mes papiers/CR-CVD/CR-CVD.tex
tutu

\documentclass{osa-article}

%% Select the journal you're submitting to
%% oe, boe, ome, osac, osajournal
\journal{osac}
% Key:
% Express journals must have the correct journal selected:
% {oe} Optics Express
% {boe} Biomedical Optics Express
% {ome} Optical Material Express
% {osac} OSAC Continuum
% Other OSA journals may use:
% {osajournal} Applied Optics, Advances in Optics and Photonics, Journal of the Optical Society of America A/B, Optics Letters, Optica, Photonics Research

% Uncomment if submitting to Photonics Research.
% ONLY APPLICABLE FOR \journal{osajournal}
% \setprjcopyright

% Set the article type
\articletype{Research Article}
% Note that article type is not required for Express journals (OE, BOE, OME and OSAC)

\begin{document}

\title{Universal manuscript template for OSA journals}

\author{Author One,\authormark{1} Author Two,\authormark{2,*} and Author Three\authormark{2,3}}

\address{\authormark{1}Peer Review, Publications Department, The Optical Society (OSA), 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{2}Publications Department, The Optical Society (OSA), 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}Currently with the Department of Electronic Journals, The Optical Society (OSA), 2010 Massachusetts Avenue NW, Washington, DC 20036, USA}

\email{\authormark{*}opex@osa.org} %% email address is required

% \homepage{http:...} %% author's URL, if desired

%%%%%%%%%%%%%%%%%%% abstract %%%%%%%%%%%%%%%%
%% [use \begin{abstract*}...\end{abstract*} if exempt from copyright]

\begin{abstract}
\LaTeX{} manuscripts submitted to OSA journals may use these instructions and this universal template format. The template simplifies manuscript preparation and eases transfer between OSA journals. Applied Optics, JOSA A, JOSA B, Optics Letters, and Optica authors may also use the previous, legacy templates, particularly if a precise length estimate is needed. Authors will still need to adhere to article-length restrictions based on the final, published format.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%  body  %%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Adherence to the specifications listed in this template is essential for efficient review and publication of submissions. Proper reference format is especially important (see Section \ref{sec:refs}). $\etaup$

\section{Multiple corresponding authors}

There are two options for indicating multiple corresponding authorship, and they are formatted quite differently. The first format would be as follows and uses an asterisk to denote one of the authors:

\begin{verbatim}
\author{Author One\authormark{1,3} and Author Two\authormark{2,4,*}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org}

\email{\authormark{*}opex@osa.org}
\end{verbatim}

This format will generate the following appearance:

\medskip

\author{Author One\authormark{1,3} and Author Two\authormark{2,4,*}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org}

\email{\authormark{*}opex@osa.org}

\medskip

The second format forgoes the asterisk and sets all email addresses equally within the affiliations. Please note that this format does not use the \verb+\email{}+ field at all.
\begin{verbatim}
\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW,
Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America,
2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org\\
\authormark{4}opex@osa.org}
\end{verbatim}

This format will generate the following appearance:

\medskip

\author{Author One\authormark{1,3} and Author Two\authormark{2,4}}

\address{\authormark{1}Peer Review, Publications Department,
Optical Society of America, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{2}Publications Department, Optical Society of America, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}xyz@osa.org\\
\authormark{4}opex@osa.org}
\medskip
These are the preferred
%express journal
formats for multiple corresponding authorship, and either may be used.

\section{Abstract}
The abstract should be limited to approximately 100 words. If the work of another author is cited in the abstract, that citation should be written out without a number, (e.g., journal, volume, first page, and year in square brackets [Opt. Express {\bfseries 22}, 1234 (2014)]), and a separate citation should be included in the body of the text. The first reference cited in the main text must be [1]. Do not include numbers, bullets, or lists inside the abstract.

\begin{figure}[h!]
\centering\includegraphics[width=7cm]{osafig1}
\caption{Sample caption (Fig. 2, \cite{Yelin:03}).}
\end{figure}

\section{Assessing final manuscript length}
OSA's Universal Manuscript Template is based on the OSA Express layout and will provide an accurate length estimate for Optics Express, Biomedical Optics Express,  Optical Materials Express, and OSA's newest title OSA Continuum. Applied Optics, JOSAA, JOSAB, Optics Letters, Optica, and Photonics Research publish articles in a two-column layout. To estimate the final page count in a two-column layout, multiply the manuscript page count (in increments of 1/4 page) by 60\%. For example, 11.5 pages in the OSA Universal Manuscript Template are roughly equivalent to 7 composed two-column pages. Note that the estimate is only an approximation, as treatment of figure sizing, equation display, and other aspects can vary greatly across manuscripts. Authors of Letters may use the legacy template for a more accurate length estimate.

\section{Figures, tables, and supplementary materials}

\subsection{Figures and tables}
Figures and tables should be placed in the body of the manuscript. Standard \LaTeX{} environments should be used to place tables and figures:
\begin{verbatim}
\begin{figure}[htbp]
\centering\includegraphics[width=7cm]{osafig1}
\caption{Sample caption (Fig. 2, \cite{Yelin:03}).}
\end{figure}
\end{verbatim}

\subsection{Supplementary materials in OSA journals}
OSA journals allow authors to include supplementary materials as integral parts of a manuscript. Such materials are subject to peer-review procedures along with the rest of the paper and should be uploaded and described using OSA's Prism manuscript system. Please refer to the \href{https://www.opticsinfobase.org/submit/style/supplementary_materials.cfm}{Author Guidelines for Supplementary Materials in OSA Journals} for more detailed instructions on labeling supplementary materials and your manuscript. Visualizations, Data Files, Datasets, and Code must be associated with a figure, table, or equation, OR be referenced in the results section of the manuscript.

\textbf{Authors may also include Supplemental Documents} (PDF documents with expanded descriptions or methods) with the primary manuscript. At this time, supplemental PDF files are not accepted for partner titles, JOCN and Photonics Research. To reference the supplementary document, the statement ``See Supplement 1 for supporting content.'' should appear at the bottom of the manuscript (above the References heading). Please note that to create text color for supplementary materials links, use of the command \\
\verb|\textcolor{urlblue}{Visualization 1}| is preferred to using the command\\
\verb|\url{Visualization 1}|.

\begin{figure}[ht!]
\centering\includegraphics{osafig2}
\caption{(a) Three traps create three rings of magnetic nanoparticles. (b) The rings interact with one another (see \textcolor{urlblue}{Visualization 1}, \cite{Masajada:13}).}
\end{figure}

\begin{verbatim}
\begin{figure}[hbt!]
\centering\includegraphics{opexfig2}
\caption{Normalized modulus distributions of transverse electrical
field components of the TM01 mode in PWs with (a) SiO_2 core
and (b) Si core}{Visualization 1}), \cite{Masajada:13}).}
\end{figure}
\end{verbatim}

\section{Mathematical and scientific notation}

\subsection{Displayed equations} Displayed equations should be centered.
Equation numbers should appear at the right-hand margin, in
parentheses:
\begin{equation}
J(\rho) =
 \frac{\gamma^2}{2} \; \sum_{k({\rm even}) = -\infty}^{\infty}
	\frac{(1 + k \tau)}{ \left[ (1 + k \tau)^2 + (\gamma  \rho)^2  \right]^{3/2} }.
\end{equation}

All equations should be numbered in the order in which they appear
and should be referenced  from within the main text as Eq. (1),
Eq. (2), and so on [or as inequality (1), etc., as appropriate].

\section*{Funding}
Content in the funding section will be generated entirely from details submitted to Prism. Authors may add placeholder text in the manuscript to assess length, but any text added to this section in the manuscript will be replaced during production and will display official funder names along with any grant numbers provided. If additional details about a funder are required, they may be added to the Acknowledgments, even if this duplicates information in the funding section. See the example below in Acknowledgements.

\section*{Acknowledgments}
Acknowledgments should be included at the end of the document. The section title should not follow the numbering scheme of the body of the paper. Additional information crediting individuals who contributed to the work being reported, clarifying who received funding from a particular source, or other information that does not fit the criteria for the funding block may also be included; for example, ``K. Flockhart thanks the National Science Foundation for help identifying collaborators for this work.''

\section*{Disclosures}

Disclosures should be listed in a separate nonnumbered section at the end of the manuscript. List the Disclosures codes identified on OSA's \href{http://www.osapublishing.org/submit/review/conflicts-interest-policy.cfm}{Conflict of Interest policy page}, as shown in the examples below:

\medskip

\noindent ABC: 123 Corporation (I,E,P), DEF: 456 Corporation (R,S). GHI: 789 Corporation (C).

\medskip

\noindent If there are no disclosures, then list ``The authors declare no conflicts of interest.''

\section{References}
\label{sec:refs}
Proper formatting of references is extremely important, not only for consistent appearance but also for accurate electronic tagging. Please follow the guidelines provided below on formatting, callouts, and use of Bib\TeX.

\subsection{Formatting reference items}
Each source must have its own reference number. Footnotes (notes at the bottom of text pages) are not used in OSA journals. References require all author names, full titles, and inclusive pagination. Examples of common reference types can be found on the  \href{http://www.osapublishing.org/submit/style/style_traditional_journals.cfm} {Author and Reviewer Resource Center}.

The commands \verb+\begin{thebibliography}{}+ and \verb+\end{thebibliography}+ format the section according to standard style, showing the title {\bfseries References}.  Use the \verb+\bibitem{label}+ command to start each reference.

\subsection{Formatting reference citations}
References should be numbered consecutively in the order in which they are referenced in the body of the paper. Set reference callouts with standard \verb+\cite{}+ command or set manually inside square brackets [1].

To reference multiple articles at once, simply use a comma to separate the reference labels, e.g. \verb+\cite{Yelin:03,Masajada:13,Zhang:14}+, produces \cite{Yelin:03,Masajada:13,Zhang:14}.
%Using the \texttt{cite.sty} package will make these citations appear like so: [2--4].

\subsection{Bib\TeX}
\label{sec:bibtex}
Bib\TeX{} may be used to create a file containing the references, whose contents (i.e., contents of \texttt{.bbl} file) can then be pasted into the bibliography section of the \texttt{.tex} file. A Bib\TeX{} style file, \texttt{osajnl.bst}, is provided.

If your manuscript already contains a manually formatted \verb|\begin{thebibliography}|... \verb|\end{thebibliography}| list, then delete the \texttt{latexmkrc} file from your submission files. However you should ensure that your manually-formatted reference list adheres to the OSA style accurately.

\section{Conclusion}
After proofreading the manuscript, compress your .tex manuscript file and all figures (which should be in EPS or PDF format) in a ZIP, TAR or TAR-GZIP package. All files must be referenced at the root level (e.g., file \texttt{figure-1.eps}, not \texttt{/myfigs/figure-1.eps}). If there are supplementary materials, the associated files should not be included in your manuscript archive but be uploaded separately through the Prism interface.

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%

Add references with BibTeX or manually.
\cite{Zhang:14,OSA,FORSTER2007,Dean2006,testthesis,Yelin:03,Masajada:13,codeexample}

%%%%%%%%%% If using BibTeX:
\bibliography{Biblio_dbl_quantum}

%%%%%%%%%% If preparing manually:
% \begin{thebibliography}{1}
% \newcommand{\enquote}[1]{``#1''}

% \bibitem{Zhang:14}
% Y.~Zhang, S.~Qiao, L.~Sun, Q.~W. Shi, W.~Huang, L.~Li, and Z.~Yang,
%   \enquote{Photoinduced active terahertz metamaterials with nanostructured
%   vanadium dioxide film deposited by sol-gel method,}
%   {\protect\JournalTitle{Optics Express}} \textbf{22}, 11070--11078 (2014).

% \bibitem{OSA}
% {Optical Society}, \enquote{{OSA Publishing},}
%   \url{http://www.osapublishing.org}.

% \bibitem{FORSTER2007}
% P.~Forster, V.~Ramaswamy, P.~Artaxo, T.~Bernsten, R.~Betts, D.~Fahey,
%   J.~Haywood, J.~Lean, D.~Lowe, G.~Myhre, J.~Nganga, R.~Prinn, G.~Raga,
%   M.~Schulz, and R.~V. Dorland, \enquote{Changes in atmospheric consituents and
%   in radiative forcing,} in \enquote{Climate Change 2007: The Physical Science
%   Basis. Contribution of Working Group 1 to the Fourth assesment report of
%   Intergovernmental Panel on Climate Change,}  S.~Solomon, D.~Qin, M.~Manning,
%   Z.~Chen, M.~Marquis, K.~B. Averyt, M.~Tignor, and H.~L. Miler, eds.
%   (Cambridge University Press, 2007).

% \end{thebibliography}

\end{document}

### 2021-01-05 Mes papiers/CR-CVD/CVD_para.tex

\documentclass[9pt,twocolumn,twoside]{osajnl}
%% Please use 11pt if submitting to AOP
% \documentclass[11pt,twocolumn,twoside]{osajnl}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}

\journal{optica} % Choose journal (ao, aop, josaa, josab, ol, optica, pr)

% See template introduction for guidance on setting shortarticle option
\setboolean{shortarticle}{true}
% true = letter / tutorial
% false = research / review article
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

% (depending on journal).

\title{Optical detection of paramagnetic defects in a CVD-grown diamond using Nitrogen-Vacancy centers}

\author[1]{C. Pellet-Mary}
\author[1]{P. Huillery}
\author[1]{M. Perdriat}
\author[2]{A. Tallaire}
\author[1]{G. H\'etet}

\affil[1]{Laboratoire de Physique de l'Ecole normale sup\'erieure, ENS, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e Paris-Diderot, Sorbonne Paris Cit\'e, Paris, France.}

\affil[*]{Corresponding author: gabriel.hetet@ens.fr}

%% To be edited by editor
% \dates{Compiled \today}

%\ociscodes{(140.3490) Lasers, distributed feedback; (060.2420) Fibers, polarization-maintaining;(060.3735) Fiber Bragg gratings.}

%% To be edited by editor
% \doi{\url{http://dx.doi.org/10.1364/XX.XX.XXXXXX}}

\begin{abstract}
The spins of Nitrogen Vacancy centers in Chemical Vapor Deposition (CVD) grown diamonds form ideal probes of magnetic fields and temperature, as well as ideal qu-bits for quantum information processing. 
Studying and controlling the magnetic environment of NV centers in such high purity crystals is thus essential.
In this letter, we demonstrate optical detection of paramagnetic species in a CVD grown diamond, such as hydrogen-related complexes, by coupling them to NV centers.
%We also unravel the role played by double quantum processes at zero magnetic field in the NV relaxation. 
The resonant transfer of NV centers spin-polarization to the electronic spins of these defects induces a loss of the NV photoluminescence, generating conspicuous features in the NV photoluminesence. 
The resonant condition is tuned with high accuracy by scanning a magnetic field along the 100 crystal direction.
Our results offer prospects for more detailed studies of CVD grown processes as well as for coherent control of the spin of a novel class of paramagnetic species.
\end{abstract}

\setboolean{displaycopyright}{true}

\begin{document}

\maketitle

\section{Introduction}
The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science. One major reason is that it can be optically polarized and read-out and features long coherence time even at ambient conditions. 
%The NV property can also then be employed to manipulate and to detect other dark spins that are interacting with it through dipole-dipole coupling, either inside or outside the diamond. 
In order to optimize the capabilities of the NV center, its magnetic environment must be controlled carefully. 
Synthetic growth of diamond crystal now reaches a level of control that makes pristine diamond crystals almost flawless, eliminating the source of magnetic noise from other nearby impurities. The standard growth technique works through Chemical Vapor Deposition (CVD) of carbon atoms from a methane gas in an ultra clean vacuum environment. It offers the possibility to work with 99.9\% purity as well as using 12C enriched methane, hence removing spin fluctuations fro 13C atoms. More results from CVD.

The control of the purity of the material is typically done using Electron-Paramagnetic-Resonance (EPR). 
EPR operates with a very high resolution and was instrumental for detecting many species in diamond (uni of Warwick's review).
It would also be ideal if the defects were detected in the same platform that is used to for nanoscale sensing and quantum information processing, namely confocal microscopy. 
This would offer the opportunity to polarize these defects using the NV center, dynamically decouple them from the NV evolution for enhanced AC magnetometry and even to use them as ancilla qubits.

 \begin{figure}[htbp]
\centering
\fbox{\includegraphics[width=\linewidth]{Fig1.eps}}
\caption{Schematics showing various paramagnetic defects in CVD grown diamonds polarized by a Nitrogen Vacancy center. Bottom right :  crystalline structure of the diamond.}
\label{fig1}
\end{figure}

%There has been a renewal of interest recently on the production of heavily doped diamonds recently.  
%Lukin, meyer....budker. 
%Cross relaxation amongst closely-packed NV centers can take place at moderate magnetic fields \cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}, thus far, other impurities generally require by scanning large magnetic fields (more than hundreds of gauss) along the 111 direction of diamond. 
%This is generally required to probe paramagnetic impurities that have a small ZFS. 
%Most of the measurements are done at the ESLAC/GLSAC. OORT, BUDKER, HOLLENBERG, BAJAJ, MANSON.

%It is more difficult to obtain large concentration of Nitrogen-Vacancy centers in diamonds made from Chemical vapour deposition (CVD). 
%Cross-relaxation studies are more scarce using CVD doped diamonds.

Here, we use confocal microscopy and a high-density NV ensembles in a CVD grown diamond to detect paramagnetic defects from CVD grown sample.
The detection was realized using cross-correlation studies through magnetic field scans. 
Cross-relaxation typically takes place when the electronic or nuclear spins of two atomic species exchange their polarization via resonant magnetic dipole-dipole interactions. 
If the spin of an NV center is polarized and coupled to an unpolarized spin B, it will lose part of its polarization at the expense of B and thus see a drop of its PL rate. 
Tuning the frequency of both spin transitions will thus result in a drop of the NV photoluminescence when that they are co-resonant, thus enabling detection of the spin state of B.

The hamiltonian that describes typical spin-1 defects in diamond under a magnetic field be written as 
\bea
\hat{H}=\hbar D S_z^2+\hbar \gamma_e\bm S \cdot \bm B
\eea
where $S$ is...
The first term describes dipole interactions amongst the two electrons of the spin-1 system. The second term describes the Zeeman interaction.
$D$ can vary significantly from one defect to another as it is sensitive to the local crystal field. It can then be used to find the structure of the defect in EPR measurements.
Also, hyperfine interactions can add zero-field splittings that can be observed under high resolution scans. 
The angle of the magnetic field can thus also be to bring defects into resonance.

Cross relaxation amongst closely-packed NV centers can take place at moderate magnetic fields \cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}, thus far, other impurities generally require by scanning large magnetic fields (more than hundreds of gauss) along the 111 direction of diamond. 
This is generally required to probe paramagnetic impurities that have a small ZFS. 
Most of the measurements are done at the ESLAC/GLSAC. OORT, BUDKER, HOLLENBERG, BAJAJ, MANSON.
Similar cross relaxation phenomena were studied by measuring changes in the optical hole depth in the zero-phonon line and studying decay rates oin stimulated spin-echo and spin-locking signals ? (BUDKER).

This procedure was proven useful for detecting the spin of $^{13}C$ atoms and the substitutional nitrogen Vacancy center in diamonds in the past. Such measurements not only show detection of this dark spins but can also be used to polarize them (Polarisation of 13C at low fields : PINES, MERILES ?). 
To date however, most studies have however been carried out using crystals grown by the High-Pressure-High-Temperature (HPHT) process, where a poor control over the paramagnetic defect density is available. In CVD grown samples, the typical paramagnetic impurities that have been detected by EPR are hydrogen complex such as the NVH, VH, VH2, where the hydrogen comes from the methane in the reactor. Fig.  \ref{fig1} shows a schematics of the various paramagnetic defects in a CVD grown diamonds that can be coupled to the Nitrogen Vacancy center in our study.

\begin{figure}[ht]
\centering
\fbox{\includegraphics[width=\linewidth]{Carte_annotation}}
\caption{NV$^-$ photoluminescence as a function of a scanning magnetic field around the [100] crystalline directions with a fixed amplitude $|B|=115G$. The planes orthogonal to the [010],[001],[011], and [01$\bar 1$] directions are indicated by dashed lines.}
\label{map}
\end{figure}

Recent work in the doping processes has shown the possibility to reach NV centers concentrations in the 3 to 5 ppm range in CVD grown samples \cite{Edmonds, TALLAIRE2020421, MINDARAVA2020182}, without sacrificing the $T_2^*$ coherence time, opening a path towards detecting these hydrogen-related complexes using cross-relaxation with NV centers.
%Several impurities in CVD grown diamonds have been detected using Electron-Paramagnetic Resonance.  In order to purify diamond crystal and optimize their conduction and properties for prolonged qubit lifetimes. 
%Extensive work has been done in the past twenty years to detect and characterize the various  that appear in EPR. 
The sample we use in this study was grown using CVD with the addition of N$_2$0 in the methane phase. Then, high energy electron irradiation and in-situ annealing was realized in order to convert the nitrogen into NV centers. A concentration of 5 ppm of NV center was obtained.
It was shown that in this sample the $T_2^*$ was not degraded after annealing, and that the NV density was large enough to enable the observation of $T_1$ dependent concentration.  It is well-known that when the NV concentration increases above 1 ppm, the longitudinal decay of NV centers is degraded. 
Furthermore, it has been observed that cross-relaxation amongst closely pack NVs of different classes can be observed.
It was postulated that such a degradation of the $T_1$ time came from the coupling of NV centers to other NV related defects whose spin fluctuates.

To perform the cross-relaxation detection, we use a homebuilt confocal microscope that comprises a 1mW green laser, an objective with a numerical aperture of XX to focus the laser onto the sample as well as to collect the PL. The PL was filtered from the green laser using a dichroic mirror and a notch filter coupled to a multimode fiber and sent to avalanche photodiode. The magnetic field scans were realized using a C-shaped electromagnet driven by a XX generator.  We monitor the NV PL as the magnetic field is scanned.

Contrary the more commonly chosen $\langle 111 \rangle$ direction, we use magnetic field scans along the $\langle 100 \rangle$ direction. 
Looking at the diamond structure (bottom right in Fig.  \ref{fig1})-a), we see that in this direction, the four possible projections of the NV centers along the magnetic field are the same. This means that doing a B field scan along this direction will not make NV centers energies from different classes cross. At arbitrary angles, this can result is several lines that could mask cross-relaxation from other species. 
The down side of this choice of B field direction is that for large B field amplitudes, the transverse component of the B field makes the PL drop [], thus limiting the range of magnetic field amplitudes that can be employed.
In order to find the $\langle 100 \rangle$ direction we perform a preliminary angular scan of the magnetic field using a goniometer at a fix magnetic field amplitude.
Fig. \ref{map}-a)  shows the NV Photoluminescence as a function of magnetic field direction, referenced by azimuthal and polar angles $(\phi, \theta)$ in the above described CVD grown sample. The PL is seen to drop for particular values of the angle as anticipated.
%The symmetric pattern suggests that the diamond crystalline main directions play a role. Because of the anisotropy of the NV center, see \ref{Cristallo}, the NV PL drop can indeed be a signature of the crystal direction.
Such a drop of the PL along the crystal axis was already observed by many groups. LUKIN, HEMMER, BUDKER. JARMOLA. OORT
\cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017} and correspond to situation where NV centers cross. 
Using such a map enables to identify clearly the crystalline axis and, in particular, the central point on this map, the $\langle 100 \rangle$ direction.

The planes orthogonal to the [010],[001],[011], and [01$\bar 1$] directions are indicated by dashed lines and show the locus of the cross relaxation, with a smaller PL contrast. 
The width of the CR peak is compatible with the $T_2^*$ time at this B field values. The contrast is determined both by the fluctuating NV concentration and the polarized NVs.
Such a map can in itself be a useful tool in order to measure magnetic fields without microwave, as already pointed out in \cite{}.

To double check these interpretations, we run an electron spin resonances by fixing the magnetic field amplitude and by scanning the microwave frequency. 
Fig. \ref{map}-b) shows an ESR taken at the angle ... showing two degeneracies.
Fig. \ref{map}-b) shows an ESR taken at the angle .. .showing four degeneracies. This hints towards the 100 direction.

We now tune the magnetic field along the 100 direction and scan the magnetic field amplitude. 
Fig. \ref{scan} shows the PL as a function of magnetic field amplitude, along the $\langle 100 \rangle$ direction in the 15 G to 145 G range. 
Careful calibration of the magnetic field amplitude was realized using a microwave tone of known frequency ....
The angle was controlled to within XX. 
For these data, the averaging was done for XX days. 
Three features clearly appear in this spectrum : one broader feature at 20G, another one at 56 G and one last one at 122 G. 
We also note an overall drop of the PL, as a the result of state mixing in the optically excited state. 
We made attempts to fit this curve using the full 7 level model, but this only gave poor agreement with experimental data. Further work will be done in order to clarify this. 
In order to let the three salient features detach better from the spectrum, we thus resorted to a 4th order polynomial fit without the spectral bumps, and subtract it from the curve. The subtracted curve is displayed below.  The width and contrast of each peak can now be compared and the position of each peak can be determined with great accuracy.

\begin{figure}[htbp]
\centering
\fbox{\includegraphics[width=\linewidth]{Scans_fig4}}
\caption{Optical detection of cross-relaxations. \textbf{Top} NV$^-$ photoluminescence while scanning the magnetic field along the [100] crystalline direction (plain line) and polynomial fit to the fourth order (dashed line). \textbf{Bottom} Subtraction of the previous signal by the polynomial fit (circles), simulated CR magnetic field amplitude (dashed,vertical) and gaussian fits for the second and third dips (plain line)}
\label{scan}
\end{figure}

In order to attribute the the three features to their respective defects, we run a similar scan (not shown) on heavily doped HPHT diamond samples in the same 100 direction. Only the first feature appeared. 
This observation led us to search for CVD specific defects as candidates for the last two peaks. Paramagnetic defects in CVD grown diamonds have been extensively studied using EPR at the university of Warwick. These studies demonstrate that  NVH, VH, VH2 complexes can be stable in diamond and they also estimate the zero-field splitting $D$ for most of them.
Oxygen related complex may also occur here, due to our N2O are not resonant (https://core.ac.uk/download/pdf/46556.pdf)
The first peak could then be coming from dipolar coupling between NVs and paramagnetic defects that are present in all diamonds, such as the substitutional nitrogen, also called $P_1$ centers ($[P_1]\approx 10 ppm$ ?) and the ${13}^C$ atoms (${13}^C= 1\%$). 
Using the results from Budker et al. we first intended to correlate the peaks with NV centers that are coupled to $P_1$ centers. 
 $P_1$ centers have a small zero-field splitting so in order to appear a such small magnetic fields, they would first have to be coupled to an NV centers and then be coupled to a nearby polarized NV.
The obtained eigenfrequencies were inconsistent with our observed peaks and our spectral resolution error (~1G). This lead us to consider the ${13}^C$ atoms. 
In itself, the ${13}^C$ spin has a very small zero-field splitting too (=?). In order to give a contribution to the spectra, it would have to be strongly coupled to an NV center. 
In this case, the $^{13}$C-NV pair would show 4 transitions corresponding to ....

Fig. \ref{energy-levels} shows the transition frequencies of the NV$^-$, the $^{13}$C-NV pair and the VH- as a function of the magnetic field amplitude along the 100 direction.
The energy evolution of another defect, whose structure is unknown, and was coined the WAR1 defects is plotted on this graph. 
The points were the NV levels cross the other defects would in principle correspond to an enhanced double quantum cross-relaxation process between the NV and the presumably unpolarized species. 
The NV0 ?

\begin{figure}[htbp]
\centering
\fbox{\includegraphics[width=\linewidth]{Theorie_CR}}
\caption{Simulated transitions energies of the various spins considered as a function of a magnetic field aligned along the [100] axis. NV centers transitions are represented using a plain line, VH$^-$ using a dashed line, the WAR1 using a dash-dotted line and $^{13}$C-NV pair using a dotted line. The amplitudes of the magnetic field where the energy level of the NV center crosses the one of another species are represented by vertical dashed lines}
\label{energy-levels}
\end{figure}

The first peak that we observe coincides with cross-relaxation that is expected to take place with $^{13}$C atoms that is one shell away from the Nitrogen vacancy centers. 
This drop of the PL is consistent with $^{13}$C-NV pair.  Using typical values found in the literature for the unpolarized NVs (also called fluctuators), we can estimate the concentration of this $^{13}$C-NV pair to be XX, making this hypothesis very likely. 
The second peak that we observe coincides very well with the crossing between the 0 to -1 NV transition and the 0 to +1 VH- transition.
Such a defect was studied in details in XXX and was found to have a $D=$. 
Making a gaussian fit to this data show that the position of the detected peak is , which gives a $D=$.
Estimation suggest that the concentration of this defect is about 100 ppb (?).

The third peak corresponds to a spin defect that has a D=.
We found in Cruddace that it was already observed in EPR, although its structure is unknown.

NVH (H3) center (diamagnetic in ground state), paramagnetic only in excited state but requires UV light. 
No other species were observed in the optical spectrum or EPR.

The technique can be used to detect the degree of polarisation of $^{13}$C atoms at room temperature.

Table \ref{tab:shape-functions}

\begin{table}[htbp]
\centering
\caption{\bf Zero-field splitting parameter $D_z$ for the different spin-1 species}
\begin{tabular}{ccc}
\hline
$D_z$ estimation (MHz) & Cruddace's work\citep{cruddace2007magnetic} & Our work \\
\hline
NV$^-$ & 2872(7) & * \\
VH$^-$ & 2706(30) & 2694(5)  \\
WAR1 & 2466(60) & 2470(10) \\
\hline
\end{tabular}
  \label{tab:shape-functions}
\end{table}

\section{Funding}
GH acknowledges funding by the French National Research Agency (ANR) through the T-ERC project QUOVADIS.

\section{Acknowledgments}
We would like to thank Neil Manson and Carols Meriles for fruitful discussions.

\section{Disclosures}

The authors declare no conflicts of interest.'

% Bibliography
\bibliography{Biblio_CVD-CR}

% Full bibliography added automatically for Optics Letters submissions; the following line will simply be ignored if submitting to other journals.
% Note that this extra page will not count against page length
\bibliographyfullrefs{sample}

\end{document}

### 2021-01-29 Mes papiers/Méca CR/18112020/SI/Meca_cross_SI.tex
yoyo

\documentclass[preprintnumbers,amsmath,amssymb,onecolumn,12pt]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\def\sgn{\mathop{\rm sgn}}

\begin{document}

\vspace{0.2in}
{\Large \hspace{1.6in}\textsc{Supplementary Material} }\\
\
\title{Resonant Dipole-Dipole Induced Mechanical Rotation of a Diamond Crystal}

\author{C. Pellet-Mary, P. Huillery, M. Perdriat, G. H\'etet}

\affiliation{$^1$
Laboratoire de Physique de l'Ecole normale sup\'erieure, ENS, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e Paris-Diderot, Sorbonne Paris Cit\'e, Paris, France.
}

\maketitle

\tableofcontents

\section{NV$^-$ center Theory}
\subsection{NV spin hamiltonian}
The hamiltonian of the electronic spin of the negatively charged nitrogen-vacancy center in its ground state can be written as :
\begin{equation*}
  \hat{\mathcal{H}}_s=\hbar D S_z^2 + \hbar \gamma_e \textbf{B}\cdot\hat{\textbf{S}},
  \end{equation*} 
where $D = (2\pi) 2.87$ GHz is the crystal field splitting originating from spin-spin interactions, and $\gamma_e = 28 $GHz/T is the electron gyromagnetic ratio. We neglect contributions from the strain and local electric field in the hamiltonian since we are working with magnetic fields on the order of 100G. We also neglect the hyperfine interaction with the nuclear spin of the $^{14}$N atom since we are working with ensembles with a typical inhomogeneous broadening of 5 MHz. The \textbf{z} axis of the $S_z$ operator here is the axis formed by the nitrogen atom and the vacancy in the body frame.

% hence why we do have different energy transitions for different NV orientations, depending on the projection of the magnetic field on the NV axis.

\subsection{Diamond crystalline axes and degeneracy conditions}
\begin{figure}[!ht]
  \centering \scalebox{0.45}{\includegraphics{SI_Cristallo}}
  \caption{a) Representation of the four crystalline axes of the diamond b) Representation of the magnetic field in the diamond crystalline basis c) Representation of the crystalline planes in the $(\theta , \phi)$ basis. The [110]$_\perp$ family of planes are shown using dashed lines, [100]$_\perp$ using plain lines. The [100] direction is marked by red circles, the [111] direction by green squares. The path of the magnetic field in the experiment of Fig. 3 of the main text is shown with blue dots.}
	\label{cristallo}
\end{figure}
There are four possible crystalline axis for the NV centers (so-called ``classes" of NV) which are depicted in Fig. \ref{cristallo} b) and correspond to the crystalline directions [$111$], [$1\bar 1 \bar 1$], [$\bar 1 1 \bar 1$] and [$\bar 1 \bar 1 1$].

The magnetic field is represented in the diamond basis in Fig. \ref{cristallo} a) where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z} ([$001$]) direction.

For some orientations of the magnetic field, the projection of the magnetic field on two or more NV axes will be identical, and therefore the energy level of the corresponding classes will be the same. These degeneracies are represented in Fig \ref{cristallo} c), where the dashed lines are the locus of the crystalline planes orthogonal to the [110] directions (6 planes in total). When the magnetic field belongs to these planes, we observe a degeneracy between two classes of NVs, as can be seen in the Fig. 4 of the main paper.

The plain lines are the loci of the planes orthogonal to the [100] directions (3 planes total). When the magnetic field angle lies in these planes, every class is resonant with another class, as can be seen in Fig. \ref{CR_22}.
The red circles correspond to the [100] directions, for which the four classes of NV are degenerate. The green squares correspond to the [111] direction where one class is aligned with the magnetic field, and the three others are degenerate. Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 3 of the main text, where we can see that a plane from the [110]$_\perp$ family is being crossed.

\section{Experimental details}

\subsection{Experimental setup}

\begin{figure}[!ht]
  \centering \scalebox{0.3}{\includegraphics{setup}}
  \caption{Illustration of the experimental setup.}
	\label{Optics}
\end{figure}

The experimental setup illustrated in Fig.\ref{Optics} is similar to the one used in \citep{DelordPRL} with the addition of a permanent magnet and an electromagnetic (EM) coil in order to perform magnetic field scans. The diamond sample is typically illuminated with 1mW of 532 nm laser light, focused by an objective with a numerical aperture of 0.5. An acousto-optic modulator (AOM) is used to switch on and off the 532nm laser and to finely tuned its power. The photo-luminescence (PL) is collected by the objective, separated form the excitation light using a dichroic mirror (DM) and a 532nm notch filter (NF), and detected using a multimode-fiber single-photon avalanche photo-detector (APD) (SPCM-AQRH-15 from Perkin Elmer). Typically, from the heavily doped samples that we use, we can detect PL photons at a rate of 1 MHz after attenuating the signal by a factor 100 with neutral density filters. 
The Paul trap is a pseudo-ring with a diameter of approximately 200 $\mu$m, as can be seen in \citep{DelordPhD}. It acts both as trap through the high voltage (HV) and as a microwave (MW) antenna.

The magnetic field generated by the (homemade) EM coil is controlled by a programmable power supply (Rhode \& Shwartz NGE 103) performing current ramps.
While the levitating setup is located in a vacuum chamber, all the experiments presented in this article are performed at atmospheric pressure.

\subsection{$T_1$ Measurement}
\begin{figure}[!ht]
  \centering \scalebox{1.2}{\includegraphics{T1_soustraction_shema}}
  \caption{$T_1$ measurement protocol. Green bars represent laser excitation, red bar represent resonant microwave $\pi$ pulse.}
	\label{T1_protocol}
\end{figure}

As shown in the Fig. 2-c) of the main text, the spin lifetime of the NV centers is modified in the presence of cross-relaxation with other classes of NV centers.
Here we present the protocol for removing the effects of charge state transfer in the dark, which masks the PL signal decay induced solely by spin depolarization. 
The protocol described in Fig. \ref{T1_protocol} consists in using two sequences. In the first sequence the spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$. The spin state is finally read out using a 10 $\mu$s laser pulse (shorter than the polarization time of the spins).

The second sequence uses the same parameters (polarization time, dark time and readout time) than the first sequence, but uses an extra resonant microwave $\pi$ pulse tuned to a transition of one of the four classes of NV$^-$ right before the readout pulse. The latter sequence brings the population of the $m_s=\ket{0}$ state to the $m_s=\ket{\pm 1}$ state of one class of NV centers.

Writing $S_1$ the signal obtained by the first sequence and $S_2$ the one obtained by the second sequence, and then measuring the difference $S_1-S_2$ between the two, we can measure the spin state population of a single class, and at the same time remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark (which would be common to the two signals and therefore subtracted).
In order to avoid low frequency noises or drift, we alternate both sequence when performing the measurement.

\subsection{Magnetic field calibration}

A permanent magnet is placed a few cm away from the diamond sample in order to apply a uniform magnetic field to the NV centers together with an electro-magnet.

To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonance (ODMR) spectra and use two orientations to determine both the angle and the magnetic field amplitude.

\subsection{Spin-mechanical detection}

High sensitivity of the spin-torque is achieved by using a speckle pattern produced by the rough surface of the micro-diamond under coherent illumination. When the particle is stably levitating, at the particle image plane, we then focus a small area of the speckle image onto an optical fibre and detect the photons transmitted through the fibre with the APD$_1$. The detected signal is then highly sensitive to the particle position and orientation.

For the spin-torque measurements presented in Fig. 4-a), the microwave detuning is scanned in 2~MHz steps with a duration of 10 ms per points. During those 10ms, the diamond orientation has enough time to reach its equilibrium position and the spin torque effect can be observed. The average count-rate is about 1 Mega-counts/s.

\subsection{Signal drift for levitating particles}

Measurements on levitating diamonds have to be relatively short (few minutes at most) because of a slow drift on the particle orientation which changes the detection location on the specular reflection off the diamond surface. The most likely origin of this drift is the loss of charges of the diamond due to photoionization by the laser, which changes the trapping conditions over time.

%This drift is the reason why the photoluminescence scan in Fig.3 of the main text has a lower signal to noise ratio than the scan in Fig.1 of the main text, which was done on a static diamond for a few hours.

%\subsection{PL detection}
%
%In these measurements, compared to \cite{DelordNat}, to detect the PL we use another detection channel that is more resilient to motion. We therefore do not need to detect prior to ring down.

\section{Principle of the mechanical detection}

\subsection{Origin of the magnetic torque}

The magnetic torque responsible for the motion of the diamond fundamentally comes from the crystalline anisotropy of the NV centers and from the transverse field $B_\perp$ responsible for mixing the eigenstates in the stationary state. We will start by considering the torque from a single NV center.
%The hamiltonian of a single NV under a magnetic field is 
%\be
%\hat{H}=\hbar D S_z^2+\hbar \gamma_e\bm S \cdot \bm B
%\ee
We simplify the problem by assuming that the motion is in the $x-z$ plane so the magnetic part of the hamiltonian reads 
$H_{B}= \hbar \gamma_e B  (\hat{S_x} \sin\theta + \hat{S_z} \cos\theta)$
where $\theta$ is the angle between the B field and NV center quantization axis.
We thus obtain the spin torque 
\be
\hat \tau_s = -\partial \hat{H} /\partial \theta = \hbar \gamma_e B (-\cos\theta \hat{S_x}+\sin\theta \hat{S_z}).
\ee
The mean value of the torque in terms of the reduced density matrix elements $\rho_{ij}$ in the basis of the $S_z$ eigenstates ($-1,0,1$) in the body fixed frame is
\be
\langle \hat \tau_s \rangle = \hbar \gamma_e B   (\rho_{-1-1}-\rho_{11})\sin\theta - \hbar \frac{\gamma_e B}{\sqrt{2}} S \cos\theta\label{torquemean}
\ee
where we introduced $S=\rho_{0,1}+\rho_{1,0}+\rho_{0,-1}+\rho_{-1,0}$.
The reduced density matrix elements are traced over a bath consisting of laser photons used to polarized the NV at a rate $\gamma_{\rm las}$, phonons or spin-fluctuators acting on the spin populations at a rate $R_1=1/T_1$ and 
$P_1$ or nuclear spins dephazing the electronic spin at a rate $1/T_2^*$.
In the limit $\gamma_e B \ll D$, and $\gamma_{\rm las}\gg \Gamma_1$ the laser still polarizes the spins in the ground state so that $\rho_{00}\gg {\rho_{11},\rho_{-1-1}}$.
Assuming that the pure dephazing $T_2^*$ is much shorter than the sum of the population relaxation $T_1/2$ and the laser induced repolarization $\gamma_{\rm las}$ the equations of motion for the coherences are

\bea
\frac{\partial \rho_{01}}{\partial t}&=&-\frac{1}{2T_2^*} \rho_{01}-i\frac{ \gamma_e B}{\sqrt{2}}\sin\theta-i\rho_{01}D+\mathcal{O}(\frac{(\gamma_e B)^2}{D})\\
\frac{\partial \rho_{0-1}}{\partial t}&=&-\frac{1}{2T_2^*}  \rho_{0-1}-i\frac{ \gamma_e B}{\sqrt{2}}\sin\theta-i\rho_{0-1}D+\mathcal{O}(\frac{(\gamma_e B)^2}{D}).
\eea
%Taking the complex value of these equations give the two other eq coherence terms. 
%The coherences can have an impact on the angle since they are responsible for the mean torque applied to the particle. 
The motional dynamics is very slow compared to the zero-field and magnetic field splittings. The latter are also larger then the decoherence rate $1/T_2^*$ so we can adiabatically eliminate the coherences and find 
\bea
\rho_{01}=\rho_{10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D} \quad {\rm and} \quad \rho_{0-1}=\rho_{-10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D}\eea
Since 
\bea
\rho_{11}-\rho_{-1-1}=\mathcal{O}(\frac{\gamma_e B}{D}).
\eea 
Re-injecting these expressions in the expression for the mean torque, we get 
\bea
\langle \hat \tau_s \rangle = \frac{\hbar(\gamma_e B)^2}{D}\sin2\theta
 +  \mathcal{O}(\frac{(\gamma_e B)^3}{D^2}).\label{torque_anal}
\eea

It is in fact the gradient of the energy in the ground state taken at the angle $\theta$.

Indeed, supposing that $\gamma B \ll D$, so that $H_{B}$ can be treated as a perturbation to the spin-spin hamiltonian, 
the perturbed energy $\epsilon_0$ of $\ket{0}$ is 
$$ \epsilon_g=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B)^2 }{D} \sin^2\theta
$$

The formula $ -\partial \epsilon_g /\partial \theta$ then gives Eq. \ref{torque_anal}. It is the equation that is used in the core of the manuscript. 
In the approximate regime of the present study, the Feynman-Hellman theorem that relates the angular derivative of the mean energy to the torque in the pure state case, 
can in fact also be applied directly here.

\begin{figure}[ht]
\centering{ \scalebox{0.4} {\includegraphics{map_torque_1_classe}}}
  \caption{\textbf{a)} Skecth showing an NV center aligned to the \textbf{z} direction in the presence of an external magnetic field \textbf{B} at an angle $\theta$ with respect the NV axis. The resulting spin vector $\langle \mathbf{ \hat S} \rangle$ of the NV center is shown by the blue arrow.
   \textbf{b)} Longitudinal ($S_\parallel$ in dashed line) and transverse ($S_\perp$ in plain line) components of the average value of the spin operator, in $\hbar$ unit, as a function of $\theta$ and for a magnetic field amplitude $|\mathbf{B}|=100$ G.   
   \textbf{c)} Amplitude of the magnetic torque acting on a single spin as a function of $\theta$ for $|\mathbf{B}|=100$ G.
   \textbf{d)} Amplitude of the same magnetic torque in the crystalline basis with $\theta$ and $\phi$ being the polar and azimuthal angle with respect to the [100] direction.}
  		\label{Torque1classe}
\end{figure}

Another way to estimate the torque is depicted in Fig \ref{Torque1classe}. Under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, under the low magnetic fields ($<$ 200 G) we are working at, will be oriented at an angle of 90$^\circ$ from the NV axis : $\langle \hat S_z \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$. This magnetization vanishes when the magnetic field is aligned with the NV center since there is no longer a transverse field responsible for the mixing of the eigenstates.

The magnetization of the NV center is therefore not aligned with the magnetic field, except when the field is also at a 90$^\circ$ angle from the NV axis, which means that the magnetic torque $\mathbf \Gamma = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B $ will be non-zero everywhere except when the field is aligned with the center, or in the plane normal to the direction of the center. We can describe each NV center as a paramagnetic defect with the anisotropic magnetic susceptibility $\chi = \begin{pmatrix}-\chi_\perp & 0 & 0\\ 0 & -\chi_\perp & 0 \\ 0 & 0 & 0  \end{pmatrix}$ in the $(\mathbf{x},\mathbf{y},\mathbf{z})$ basis where $\mathbf{z}$ is the orientation of the NV center.

The amplitude of the torque with respect to the magnetic field orientation (at a B field amplitude of 100 G) is represented in 1D in Fig \ref{Torque1classe}-c) where we can see a behavior very close to $|\sin(2\theta)|$, which is the formula (3) of the main text that we obtained through a perturbative approach. The same torque amplitude is represented in 2D in Fig \ref{Torque1classe} d). The two purple dots in the map correspond to the [111] direction when the magnetic field is aligned with the centers. The curvy line corresponds to the $[111]_\perp$ plane. Importantly, the maximum torque value is $1\cdot 10^{-26}$ N.m for a single spin.

\subsection{Total spin torque and influence of cross-relaxation}

%\begin{figure}[ht]
%\centering{ \scalebox{0.4} {\includegraphics{map_torque_4_classe}}}
%  \caption{Magnetic torque amplitude for 4 spins, one in each possible orientation, as a function of $\theta$ and $\phi$, the polar and azimuthal angle with respect to the [100] direction. \textbf{a)} Not taking into account the CR between NV centers. \textbf{b)} Taking into account the CR between NV centers.}
%  		\label{Torque4classes}
%\end{figure}

Fig 3 a) in the main text, represents the same map as Fig \ref{Torque1classe}-c) but including the four NV centers, one in each of the possible [111] orientations. We can see that the maximum torque actually decreased to $3\cdot 10^{-28}$ N.m even though we increased the number of NV centers by four. This is due to the directional averaging of the torque generated by the four centers. The torque per NV center is decreased by more than two orders of magnitude when taking the directional averaging into consideration.
Fig 3 b) in the main text shows the same map, this time taking into account the modification of the spin lifetime due to cross-relaxations. The detailed model is presented in the section \ref{Simu}.
There are two things to note here : 
\begin{enumerate}
\item The maximum torque has increased by an order of magnitude compared to the previous case. It reached up to $3\cdot 10^{-27}$ N.m for four spins, so about $10^{-27}$ Nm per spin. Qualitatively, this is because cross-relaxation will lower the torque contribution of specific classes (the ones that get depolarized), meaning that the end result is closer to the single spin case (there is less directional averaging).
\item The change in magnetic torque is resonant, it happens only when different classes are brought to resonance. This can be seen by comparing Fig 3 b) of the main text to the [110]$_\perp$ planes that were drawn in Fig. \ref{cristallo}. The change in the signal when scanning a magnetic field across a CR will be much sharper than the sinusoidal change in the spin-torque.
\end{enumerate}

\subsection{Torque sensing with a levitating diamond}

The way we experimentally measure spin-torques applied on the levitating diamond is by measuring the induced diamond orientational displacement from equilibrium. 
We model the trap as a pure harmonic potential, both for the center of mass and for the librational degrees of freedom of the diamond with trapping frequencies $\omega_t \approx (2\pi) \cdot 1$ kHz. Considering a single librational degree of freedom, we can write the torque exerted by the trap as $\Gamma_t=-K(\theta-\theta_{eq})$, where $K=I \omega_t^2$ is the stiffness of the trap, $I$ being the moment of inertia of the diamond.

The application of an external torque $\Gamma_{\mathrm{ext}}$ to the diamond will therefore shift the angular equilibrium position  in such a way that : 
\begin{align}
-K(\theta-\theta_{\mathrm{eq}}) + \Gamma_{\mathrm{ext}} &= -K(\theta-\theta_{\mathrm{eq}}') 
\end{align}
so that
\begin{align}
\delta\theta = \theta_{\mathrm{eq}}'-\theta_{\mathrm{eq}} &= \frac{\Gamma_{\mathrm{ext}}}{K}=\frac{\Gamma_{\mathrm{ext}}}{I \omega_t^2}
\end{align}

In our case, $\Gamma_{\mathrm{ext}}$ is the magnetic torque exerted by the NV$^-$ spins on the diamond. We can write it $\Gamma_{\mathrm{ext}} = N_{NV} \langle \Gamma_{\mathrm{1 spin}} \rangle$  where $\langle \Gamma_{\mathrm{1 spin}} \rangle = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B \approx 10^{-27}$ Nm is the expected magnetic torque applied by one spin.

By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle n(NV^-)}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-4} rad$$ where $n(NV^-) \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 6.3\cdot 10^3$ rad/s is the typical value of the trap angular frequency.

It should be noted that the main uncertainty comes here from the diamond size can change the expected result by an order of magnitude.

\section{Cross-relaxation detection for another type of degeneracy}

%Faire 2 lignes avec sur la première ligne l'ESR et le trajet sur la carte du champ mag
\begin{figure}[!ht]
  \centering \scalebox{0.4}{\includegraphics{22-total}}
  \caption{Mechanical detection of a dipolar interaction wen crossing a [100]$_\perp$ plane. \textbf{a)} Path of the magnetic field angle (red dots) in the ($\theta , \phi$) basis in the measurements, with respect to the [100] direction. The [100]$_\perp$ family of planes are shown in plain blue lines. The [110]$_\perp$ family of planes is shown in orange dashed lines. \textbf{b)} ESR spectra measured at 4 different magnetic field values. Calculated evolution of the NV transition frequencies as a function of the electromagnet B field $B_{\rm EM}$, for the four $\ket{0} \to \ket{-1}$ transitions (dashed lines). \textbf{c)} Photoluminescence of the NV centers as a function of the scanning magnetic field (blue crosses) and gaussian fit (orange line). \textbf{d)} Simulated population in $\ket{m_s=0}$ for the stationary state with (plain) or without (dashed) taking into account the decrease of the T$_1$ induced by the cross-relaxations.
  \textbf{e)} Signal reflected off the diamond surface as a function of magnetic field amplitude $B_{\rm EM}$ (blue crosses). The orange line shows a double gaussian fit. \textbf{f)} Simulated torque applied by the spins on the diamond, with (plain) or without(dashed) taking into account the cross-relaxations.}
	\label{CR_22}
\end{figure}

Similarly to in Fig. 3 of the main text, we managed to mechanically detect other types of mechanically induced dipolar couplings.
Fig. \ref{CR_22}b) shows a calculation of the frequencies of the $\ket{0} \to \ket{-1}$ transitions for all four classes of NV centers and the mechanically detected ESR spectra measured using the reflected laser for various magnetic field values. Unlike the experiment reported in the main text, this time all four classes of NV are resonant with another class at B=23 G. This indicates that we are crossing a [100]$_\perp$ plane instead of a [110]$_\perp$ one, as can be seen in Fig \ref{CR_22}a).

Fig. \ref{CR_22}c) shows the recorded photoluminescence of the NV centers during the magnetic field scan. As expected, a drop in the photoluminescence rate is observed when the degeneracy occurs. The PL drop is slightly more pronounced in this case compared to the experiments presented in the main text because all classes are depolarized here, instead of only two. This is well predicted by the numerical estimates shown in Fig. \ref{CR_22}d).

Fig. \ref{CR_22}e) shows the signal of the laser reflected off the diamond surface, proportional to the angular displacement. Here, there is a clear difference between the angular response and the experiment shown in the main text. Instead of a single Gaussian drop centered on the CR resonance, two bumps are observed on both sides of the resonance. 
Fig. \ref{CR_22}f) shows the result of simulations, where we can see a dispersive profile with an almost zero torque at the resonance. 
The reason we do not observe a change of sign in the experiment (with two positive bumps instead of a positive and negative one) is likely to be because of the non-linearity of our detection : if the signal initially corresponds to a dark spot of the speckle, then a change in the motion of the diamond can only result in an increased signal.

Let us give a physical interpretation of the dispersive angular profile for this degeneracy condition. 
Here, the magnetic torque generated by the four classes of NV is not modified exactly on resonance, since all four classes are depolarized identically. Close to resonance, all classes will not be identically depolarized however : looking at Fig. \ref{CR_22}b), we can see that the two classes which have a higher frequency are always slightly closer to each other than the two classes of lower frequency. This can be since by computing the slope in the evolution of the transitions frequencies, which are found to be smaller for the two upper classes. 
This effect results in more depolarization for these two classes, except when they are exactly on resonance. This interpretation explains the overall shape of the torque and thus of the angular response.
A quantitative analysis would require knowledge about the directions of the 4-NV directions with respect to the three main directions of the trapped diamond axes as well as a calibration of the sensitivity of the speckle detection method for the three corresponding angular modes.

\section{Simulation details}
\label{Simu}
In this part we will discuss the method used to simulate the average torque as well as the population in the $\ket{m_s=0}$ state. Numerical solving of the master equation was performed using the Quantum Toolbox in Python (QuTiP) \citep{qutip1} \citep{qutip2}.

In order to describe the dynamics of our spin ensemble, we introduce a incoherent optical pumping through the jump operators $\mathcal{L}_+ = \Gamma_l \ket{0}\bra{+1} $ and $\mathcal{L}_- = \Gamma_l \ket{0}\bra{-1} $, where $\Gamma_l \approx (2\pi) 10$ kHz is the laser induced polarizing rate to the ground state.
We also introduce the usual $T_1$ jump operators $\mathcal{L}_i^j= \frac{1}{T_1}\ket{i}\bra{j}$ where $\ket{i,j}$=$\ket{0, \pm 1}$.

In order to describe the $T_1$ modification induced by the cross-relaxations, we use a phenomenological model where each class has its own T$_1^i$ ($i \in \{1,2,3,4\}$) that depends on the energy levels of the other classes with the formula :
\begin{equation*}
\frac{1}{T_1^i}=\frac{1}{T_1^0}+\sum_{j \neq i} \frac{1}{T_1^{dd}}e^{-\frac{(\nu_i-\nu_j)^2}{2(\sigma^{dd})^2}},
\end{equation*}
where $\nu_i$ and $\nu_j$ are the transition frequencies of the classes $i$ and $j$ (we are arbitrarily considering the $\ket{0} \to \ket{-1}$ transition here, since the resonance condition is the same for both transitions at the magnetic fields we are working at. This is not always true for magnetic fields greater than 592 G\citep{van_oort_cross-relaxation_1989}). 
$\sigma^{dd} $ is the width of the CR features. We measured them to be similar to the inhomogeneous broadening given by the dipolar coupling to $P_1$ centers, {\it i.e.} $\approx$ 6 MHz.

$T_1^0=1.03$ ms and $T_1^{dd}=0.38$ ms were chosen to match the $T_1$ measurements presented in Fig. 1 of the main text. We only focus on the $T_1$ without degeneracy and the one with a single degeneracy since our experiments will not have more than two resonant classes at once. Our model is probably not suited to deal with triple our quadruple resonances.

Finally, according to previous measurements performed in \citep{choi_depolarization_2017}, only the $\ket{0}\bra{\pm1}$ and $\ket{\pm1}\bra{0}$ (corresponding to a single quantum exchange in the dipole-dipole interaction) operators are modified by the cross-relaxations. Using this model, we can numerically solve the master equation and get the density matrix in the stationary state $\rho_s$. With $\rho_s$ we can directly obtain the $\ket{m_s=0}$ population, corresponding to the experimentally measured photoluminescence.

With regards to the torque estimation, we use a semi-classical formula :
\begin{equation*}
\mathbf \Gamma = N_0 \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B,
\end{equation*}
where $N_0 \approx 10^9$ is an estimate of the number of spins in our sample based on the average size and NV density of our diamonds, $\gamma_e$ is the gyromagnetic ratio of the electron and $\langle\hat{\mathbf S}\rangle =\mathrm{Tr}(\rho_s \mathbf{\hat S})$ is the averaged spin vector in the stationary sate, averaged again over the four possible orientations of NV.
This formula assumes that the spin dynamics is faster than the dynamics of the motion of the diamond, which is the case in our experiments.

In our plots in Fig. \ref{CR_22} and Fig. 3 of the main text, we only represent one spatial component (e.g. $\Gamma_x$) of the torque, because the three components behave similarly.
\bibliography{trap}

\end{document}

### 2021-02-01 Mes papiers/Méca CR/18112020/Meca_cross.tex

\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-1}
%\documentclass[preprintnumbers,amsmath,amssymb]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}

%\newcommand{|}{Y$_2$SiO$_5$}
\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{Rotating a Diamond Crystal using Interacting Spin Ensembles}

\author{C. Pellet-Mary$^1$, P. Huillery$^1$, M. Perdriat$^1$, G. H\'etet}

\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e de Paris , 24 rue Lhomond, 75231 Paris Cedex 05, France.}

\begin{abstract}
Magnetic-dipole interactions between the spins of defects in crystals play a key role in a vast range of applications. Here, we demonstrate resonant dipole-dipole induced rotation of a levitating diamond containing nitrogen-vacancy (NV) centers. Specifically, we employ cross-relaxation between the electronic spins of NV centers to increase  
the particle magnetic susceptibility by an order of magnitude.
Our approach opens a path towards the use of mechanical oscillators to detect paramagnetic defects that lack optical transitions, to investigation of angular momentum conservation in spin relaxation processes and to novel means of cooling the motion of mechanical oscillators.
\end{abstract}

\maketitle

%Defects in solid state materials with long-lived electronic or nuclear spins play a key role in the fields of quantum computing, communication and sensing. 
%Such defects can be hosted by various materials, the most studied examples being nuclear impurities in quantum dots \cite{Qdots}, dopants in silicon \cite{Zwanenburg}, or negatively charged nitrogen vacancy (NV$^-$) centers in diamond \cite{Doherty}.
%The detection of their spins is generically performed {\it via} emission or absorption of electromagnetic radiation, both in the radio/micro-wave or optical domain, or purely electrically \cite{Zwanenburg, Hopper}.
%Electron-Paramagnetic Resonance (EPR) and Optically Detected Magnetic Resonance (ODMR) are two such detection methods, both routinely operating under ambient conditions and with large signal to noise ratios.
%EPR can detect most paramagnetic species with great accuracy, while the latter reaches single spin resolution \cite{Wrachtrup1}, albeit only with optically active defects.
%Notably, both techniques can also efficiently detect spins through magnetic dipolar interactions. 
%Dedicated Pulsed-EPR \cite{Mims} and ODMR protocols such as the Double-Electron-Electron-Resonance (DEER), Electron-Nuclear-Double-Resonance (ENDOR) or Cross-Relaxation (CR) at the Hartmann Hahn resonance have all indeed been proven essential for characterizing dipolar interactions, offering ways to detect distant spins \cite{Mamin}.
%
%Another more recent route for detecting spins is to couple them to mechanical oscillators.
%Electronic spins have first been detected this way in \cite{Alzetta} and nuclear spins can be sensed using state of the art magnetic resonance force microscope (MRFM) \cite{rugar, MaminH}. MRFM is a very efficient probe for paramagnetic impurities with nanometer resolution.
%Conversely, torque sensors lends themselves naturally to the detection of solid state defects, which thanks to anisotropic terms in the spin hamiltonian can exert torques in a homogeneous magnetic field. 
%Torque sensing was employed recently to detect the long-lived spins of NV centers in diamonds and to demonstrate spin-cooling of a mechanical oscillator \cite{DelordNat}.
%In general, spin-mechanical experiments with long-lived spins also offer prospects for performing quantum mechanical experiments with large objects \cite{yin, Wan, Scala, Lee_2017}. However, while dipolar interactions can be detected using EPR or ODMR, mechanical detection of dipolar interactions between long-lived spins has not been observed. Compared to these two established techniques, it could offer the advantage of sensing angular momentum conservation during dipolar relaxation \cite{Zangara} as well as providing an alternative path for controlling the quantum motion of mechanical oscillators.
%
%Here, we report on the detection of spin-interactions using a mechanical torque sensor.
%Specifically, we employ a levitating diamond in a Paul trap with embedded NV centers and use the diamond orientation as a probe of the NV spin interactions.
%We use closely packed nitrogen-vacancy centers to ensure both large spin-torques on the mechanical oscillator together with large dipolar-interactions amongst the spins.

Controlling the motion of macroscopic oscillators in the quantum regime has been the subject of intense research over the past decades \cite{Teufel, Riedinger}.  In this direction, opto-mechanical systems have had tremendous success. Cooling mechanical oscillators to the motional ground state has for instance been achieved by coupling micro-objects to optical cavities using various platforms \cite{Aspelmeyer, Delic, Chan}. 
Similarly, it was proposed to strongly couple long-lived atomic spins, such as the electronic spin of nitrogen-vacancy (NV) centers, to mechanical oscillators \cite{Treutlein2014, Rabl, Lee_2017}. Realizing such an experiment at the single spin level would offer the formidable prospect of transferring the inherent quantum nature of electronic spins to the oscillators. In a regime where the spin decoherence rate is below the spin-mechanical coupling strength, spin-mechanical platforms will then have far reaching implications in quantum sensing and tests of quantum mechanics.

Most efforts are presently hampered by the low single-spin coupling strengths however, which are currently far below the typical spin decoherence rates \cite{Kolkowitz, Gieseler, DelordNat, Arcizet}.   
One solution to counteract this issue is to work with large ensembles of spins \cite{DelordNat}. This approach leaves on the side non-linear spin-mechanical effects, but offers a more favorable path towards ground state spin-cooling \cite{Rabl} and towards the observation of many-body effects mediated by the motion \cite{Wei, ma2016quantum}.

\begin{figure}[!ht]
  \centering \scalebox{0.1}{\includegraphics{dilutednv.pdf}}
  \caption{General principle of the resonant dipole-dipole induced mechanical rotation. The four possible directions of the nitrogen-vacancy centers in the diamond are shown in the left/right panels together with their spin-torque contributions (arrows of the corresponding colors). Left panel: the (quasi)-rotational invariance gives a small total spin torque $\tau_s$.
Right panel: A magnetic field (not shown) is tuned so that the spin class 1 and 3 point to the same direction. Cross-relaxation (CR) between these two classes of NV centers occurs, altering the rotational symmetry and increasing $\tau_s$. 
  }\label{principle}
\end{figure}

However, although the spin-mechanical coupling strength increases linearly with the number of atoms initially, new effects modify this scaling-law when the mean atom-atom distance is on the order of $\approx$ 10 nm. There, dipolar interactions amongst NV centers can become dominant, significantly enriching the physics at play. It has in fact been predicted that interaction amongst atoms can cooperatively enhance the coupling to the motion, akin to super-radiant processes \cite{Bachelard, PANAT}. 
Recent theory and experiments show that interactions between the optical dipoles NV centers enhances optical binding forces \cite{Venkatesh, Juan}. The interaction between the electronic spins of distant NV centers may also show similar spin-mechanical scaling. Further, the spin-interactions can be tuned resonantly amongst different NV orientations \cite{van_oort_cross-relaxation_1989}, offering prospects for studying the interplay between dipolar interactions and motional degrees of freedom in a controlled fashion. Increasing the density of NV centers also means that they can couple to other spins in the diamond \cite{armstrong_nvnv_2010, Alfasi, Epstein, Hall} and even transfer their polarization \cite{WangBajaj}.
When coupled to the angular degree of freedom, the latter effect would provide the unprecedented possibility to measure the exchange of spin angular momenta during cross-relaxation
%The laser-controlled spin-relaxation of NV centers enables a very high degree of control over the exchange of angular momenta between the spin and mechanical oscillators. 
%Coupling NV centers to impurities that cannot be optical polarized could then lead to interesting spectroscopic tools 
and enable controlling mechanical oscillators in the quantum regime \cite{Zangara}.

%We then discuss how to cool a mechanical oscillator using CR.

%When this occurs, the spin-torque is modified so that subsequently, the diamond rotates.
\begin{figure}[!ht]
  \centering \scalebox{0.08}{\includegraphics{setup.pdf}}
  \caption{Schematics of the experiment. A micro-diamond is levitating in a ring Paul trap enclosed in a vacuum chamber. A green laser is used both to polarize the NV centers in the levitating diamond and to detect the motion. Part of the speckle pattern formed in the image plane is sent onto APD$_1$ after passing through a beam splitter (BS). The photoluminescence from the NV centers is collected on APD$_2$ after filtering out the green laser light by a dichroic mirror (DM) and a notch filter (NF). a) Sketch showing the NV-NV depolarization process. Green arrows represent the optical pumping to the brighter $\ket{m_s=0}$ state. The two curvy blue arrows with different thicknesses represent short/long longitudinal relaxation of NV$_2$/NV$_1$. Red circles represent the population in each state and red dashed arrows represent the resonant dipole-dipole interaction between the two NV centers. c) Longitudinal relaxation from a single transition frequency. i) The class is not at resonance with any other classes : $T_1$=1.02 ms. ii) The class is at resonance with another class : $T_1$=0.28 ms. iii) The class is at resonance with the three other classes : $T_1$=0.15 ms.
  }\label{setup}
\end{figure}

Here, we employ resonant dipolar interactions between the spins of negatively charged nitrogen-vacancy centers to tune the orientation of a micro-mechanical oscillator. 
Specifically, we use NV centers inside a diamond that is levitating in a Paul trap and use resonant cross-relaxation between them to actuate
the spin-torque.
%We first present the results and conclude by discussing the potential applications behind our observations.
The key mechanism is depicted in Fig.~\ref{principle}. 
As can be seen in the left panel, NV centers can be found in four different orientations in the diamond crystalline structure. Due to quasi-rotational invariance of the problem, although each orientation could exert a significant torque to the diamond, the total spin-torque $\tau_s$ is very small.
%Our approach lies in the cross-relaxation (CR) that takes place when two spins with different orientations and polarization become resonant. 
%As depicted in Fig.~\ref{principle}-a), quasi-rotational invariance of the problem can make the sum of the spin torque of the four spin directions negligible. 
However by tuning an external magnetic field, resonant dipole-dipole interactions between the spin of NV centers of different orientations can be enhanced. 
When they become resonant, the polarization of the different orientations can then be exchanged in a process called cross-relaxation.
The right panel of Fig.~\ref{principle}, shows a cross-relaxation mechanism that partly removes the contribution from two classes of NV centers (labelled 1 and 3 in Fig.~\ref{principle}), which breaks the four-spin rotational invariance. The total spin torque $\tau_s$ can then be large enough to rotate the diamond.

%We first show that cross-relaxation can be effective in micro-diamonds that are heavily doped by NV centers.
%The NV center has two unpaired electrons in its ground state. In the triplet state, the dipolar interaction between them leads to a zero-field splitting $D=2.87$ GHz.
%The three spin states are labelled $| m_s={0}, {\pm 1}\rangle$.
%Due to an intersystem crossing in the optically excited state of the NV centers \cite{Doherty}, the state $m_s=\ket{0}$ is brighter than the state $m_s=\ket{\pm 1}$ upon green laser optical illumination \cite{Hopper}. 
%This provides a means to read out the Zeeman splitting by scanning a microwave tone around the resonance.
The diamonds that we use are in the form of a powder of particles that have a diameter of 15 $\mu$m. They are supplied by the company Adamas, which produces diamonds with a concentration of NV centers in the 3-4 ppm range. %The dipolar interactions may thus be effective since the average distance between the NV centers of about 10 nm. 
We load the micro-diamonds in a Paul trap that is similar to the one used in \cite{delordPRL} and depicted in Fig. \ref{setup}-a). One important feature of this system is the angular confinement provided by the trap, which offers the unique possibility to study the spin-induced torques.
Also, this geometry enables to apply a microwave signal directly on the trapping electrode, which provides an efficient means to excite the spins.
The photoluminescence from the NV centers is detected using standard confocal microscopy. We use about 100 $\mu$W of laser light at 532nm to polarize the NV centers. The laser is focused via a lens inside a vacuum chamber which has a numerical aperture of 0.5 and a working distance of 8mm. The focal point of the laser is kept few tens of micrometers away from the micro-diamond to mitigate the effect of radiation pressure and to enable laser excitation of the whole diamond \cite{delordPRL,delord2016}.

\begin{figure}[!ht]
  \centering \scalebox{0.24}{\includegraphics{map_torque_4_classe.pdf}}
  \caption{Numerical simulations of the spin-torques on a diamond containing one NV center per orientation as a function of $\theta$ and $\phi$, the polar and azimuthal angle with respect to the [100] direction. a) and b) show the torque with and without cross-relaxation between NV centers respectively.}
  \label{Numerics}
\end{figure}

%Let us first show that dipolar interactions are effective in our employed diamond micro-particles by studying cross-relaxation when they are simply attached to the trap. 
Fig.~\ref{setup}-b) depicts the dipolar interaction between two NV centers. In this example, the electronic spin of NV$_1$ is polarized in the ground state via the green laser, whereas NV$_2$ has a relaxation time $T_1$ that counteracts the influence of the optical polarization, reducing its total spin polarization. The spins will exchange magnetic quanta through flip-flop processes resulting in a depolarization of NV$_1$. At large enough NV densities, a few fast-decaying NV centers can then depolarize an ensemble of NV centers through dipolar interaction. This can reduce the average $T_1$ of the ensemble from the phonon-limited $T_1$ ($\approx$ ms) to a few hundreds of micro-seconds \cite{Jarmola} and lower the total photoluminescence. 
This effect has been observed by many groups \cite{van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, mrozek_longitudinal_2015, choi_depolarization_2017} in bulk materials. The origin of the fast-decaying NV centers was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}. The NV centers that undergo tunneling with other impurities (possibly with the substitutional nitrogen defect \cite{manson_nv_2018}) have a largely reduced longitudinal spin lifetime $T_1$.
%This process has recently revived interest because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}. 
%In our spin-mechanics experiment, we use micro-diamond particules subjected to a green laser field and a magnetic field.
%To observe the dipolar effects in a levitating platform, we employ heavily doped micro-diamonds.

Such a process has not been presented with nano- or micro-particles in the literature to the best of our knowledge. Smaller diamond particles in fact tend to suffer from extra parasitic surface effects such as spin depolarization due to interaction with paramagnetic dangling bonds on the surface \cite{Tetienne}, or enhanced charge transfer between the NV$^0$ and NV$^-$ charge states \cite{Dhomkar} so it is essential to verify that it can be observed with micro-particles. 
We start by searching for CR using micro-diamonds that are physically attached to the trap, by employing a fixed bias magnetic field $||\bf B_{\rm bias}||\approx$100 G and by tuning another magnetic field $\bf B_{\rm em}$ at some angle with $\bf B_{\rm bias}$ using an electromagnet (see Fig.~\ref{setup}-b)).

At specific magnetic field directions with respect to the crystalline axes, degeneracy between the spin of NV centers can be reached \cite{van_oort_cross-relaxation_1989}. %We check this systematically by searching for a reduction of the PL using APD$_2$ and using Opticall-Detected-Magnetic-Resonance (see SI) to correlated it with degeneracies. %We first characterize the depolarization of a diamond that is attached to trap ring to ensure that cross-relaxation can be observed in our samples.
%POUR LES SI 
%It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
%where, to our knowledge, such cross-relaxations have not been observed. 
%Indeed, surface effects may start to play a role and change quantitatively.
In this case, the longitudinal decay time from NV centers should be reduced. To verify this, we measured the $T_1$ time by applying a green laser that polarizes the NV centers and measure the photoluminescence at a later time.
Such a measurement however turned out to be significantly impacted by recharging of NV centers in the dark \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use the sequence presented in the SI, where a microwave pulse is applied or not prior to spin relaxation. The PL signals acquired in the two different measurements are then subtracted and shown for different degeneracy configurations in Figure \ref{setup}-c).
In the absence of degeneracy, we observe an exponentially decaying profile, from which we extract a $T_1=1.08$ ms, already shorter than the phonon limited lifetime in dilute bulk materials \cite{Tetienne}. This lifetime is even further reduced to 0.28 ms (resp. 0.15 ms) when two (resp. four) orientations are brought to resonance. This hints towards the role played by dipolar interactions, which are enhanced when classes of NV centers are resonant \cite{van_oort_cross-relaxation_1989, choi_depolarization_2017}. The effective density of NV being able to exchange spin quanta indeed increases, resulting in a stronger depolarization process.

%\begin{figure}[!ht]
%  \centering \scalebox{0.6}{\includegraphics{cross_121_200G.png}}
%  \caption{Numerics : PL change in the presence of cross-relaxation (top) and spin-mechanical detection (below). 
%  }\label{CR_deposited}
%\end{figure}

The main goal our study is to search for the mechanical action of such dipolar induced relaxations with diamonds levitating in the Paul trap.
%Following the picture presented in Fig. 1, we will seek for changes in the diamond orientation at CR conditions. 
%For a given levitating particle, we can optimize, in real time, the signal coming from the angular displacement of the particle by selecting the most favorable region of the particle image. To do this, we monitor our optical signal while switching a microwave field tuned to one of the ESR transitions, at a frequency of 1 Hz.  Alignment is realized by maximizing the variation of the coupled light as the diamond rotates between two angular positions. 
One major basic ingredient is the spin-torque coming from the NV centers when they are polarized in the ground state. 
Let us consider first the dependence of the ground state energy of a single spin as a function of the angle between a magnetic field and the NV axis.
The Hamiltonian for one NV orientation with quantization axis $z$ in the particle frame reads 
\begin{equation}\hat{H}_{\rm NV}=\hbar D \hat{S}_z^2+ \hbar \gamma_e \bf B  \cdot \bf\hat S,
\end{equation}
where $\bf\hat S$ is the spin-vector, $D=(2\pi)2.87$ GHz the zero-field splitting and $\bf B$ is the external magnetic field.
Under the condition $\gamma ||\bm B|| \ll D$, and assuming an NV center and a magnetic field in an $(x,z)$ plane, $H_{B}=  \hbar \gamma_e {\bf B}  \cdot {\bf \hat S}=\hbar\gamma_e B  ( S_x \sin\theta + S_z \cos\theta)$ can be treated as a perturbation to the anisotropic part $\hbar D \hat{S}_z^2$ of the Hamiltonian. Here, $\theta$ is the angle between the magnetic field and the body-fixed NV center axis $z$.
The perturbed energy $\epsilon_g$ of the ground state due to the transverse B field $B_\perp=B \sin(\theta)$ is then
\begin{equation} \epsilon_g=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B_\perp)^2 }{D}
\end{equation}
\begin{figure*}[!ht]
  \centering \scalebox{0.45}{\includegraphics{CRmeca}}
  \caption{a) Top :  Signal reflected off the diamond surface as a function of microwave frequency for three different magnetic field values. Bottom : sketch showing the crossing of a crystal plane when the magnetic field angle is tuned.
b) PL detection as a function of B$_{\rm em}$ across a dipole-dipole resonance. i) experimental data with gaussian fit, ii) simulation of the population in $\ket{m_s=0}$ state, taking into account cross-relaxation (plain) or not (dashed).
c) Angular detection as a function of B$_{em}$ across a co-resonance. i) experimental data with gaussian fit, ii) simulation of the magnetic torque applied to the diamond, taking into account cross-relaxation (plain) or not (dashed).
  }
  \label{data}
\end{figure*}
A direct use of the Feynman-Hellman theorem can give the torque in the ground state. We find that 
\begin{equation}\tau_s=- \frac{\partial \epsilon_g}{\partial \theta}=\hbar  \frac{(\gamma_e B)^2}{D} \sin 2\theta.\label{Eq1}
\end{equation}
A proof of the applicability of this theorem is presented in the SI. 
At an angle $\theta=\pi/4$, where the torque is maximized and at a B field of 100 G, we find $\tau_s\approx 5\times 10^{-17}$ N.m. Using $10^9$ spins polarized in the ground state and taking a librational confinement frequency of the diamond in the Paul trap to be around $\omega_\theta/(2\pi)=1$ kHz, we obtain an angular displacement 
of $\tau/I\omega_\theta^2\approx$1 mrad, which can be measured with a high signal-to-noise ratio in our set-up \cite{DelordNat}.
%We checked numerically that at these magnetic field values and angle, the NV are still polarized mostly in the ground state. 
%Note that the torque is zero at $\theta=mod (\pi/2)$ where the eigenstates of $S_z$ are not mixed. 
%The Hamiltonian ground state is the eigenstate $| S=1,m_s=0\rangle$, which is is non magnetic.

%Note also that the torque tends to zero if  tends to zero at any angle $\theta$. This is the case in the presence of strong spin-depolarisation, i.e. when $\hat\rho\rightarrow\frac{1}{3}\hat{I}$ (with $\hat{I}$ the identity operator).

As already hinted to in the explanations of Fig. 1, the contributions from the other NV classes must also be taken into account.
Fig. \ref{Numerics} presents the result of numerical calculations of the torque coming from the four classes of NV centers.   
Fig. \ref{Numerics}-a) shows the torque magnitude as a function of $\theta$ and $\phi$ without taking into account the cross-relaxations. The torque from each of the four classes appear clearly from the symmetry. Their different contributions however sum up to give a maximum torque of around $10^{-28}$ N.m, which is 20 times smaller than the torque that can be obtained for a single class due to the quasi-rotational invariance of the 4-NV classes. When two classes of NV center are resonant however, the induced cross-relaxation partly breaks this rotational invariance. 
Fig. \ref{Numerics}-b) shows the same plot, but including the effect of the cross-relaxation when classes of NV centers are degenerate (see SI).  
Here we use numbers that are deduced from our experimental observations of the CR induced change of the $T_1$.
One can see that a new pattern with larger spin-torque values is superimposed to the previous map. These larger values correspond to crossings of the crystal planes where NV degeneracies occur. 
At these coordinates, one recovers the torque estimation of Eq.~\ref{Eq1}, found for a single class, which would then imply a strong enough torque for overcoming the Paul trap confinement.

%Fig. \ref{principle}-a) is a depiction of a levitating diamond containing NV centers that have four possible directions along the $\langle 111 \rangle$ axes. 
%In the presence of a magnetic field, each NV centers has a different magnetic field projection, which means a different torque. 
%Below, we show the sum of the ground state spin-torques $\tau_{NR}$, resulting from the four classes. 
%Fig. \ref{principle}-ii) shows the same situation, with a magnetic field that is tuned so that two NV classes are co-resonant. 
%When the re-polarization rate due to the green laser is smaller than the cross-relaxation rate, the magnetization coming from these two classes (indicated by a circle) tends to cancel. The population in the three NV eigenstates will indeed equilibrate, giving a smaller magnetization.
%The resulting torque is thus changed to $\tau_{R}$ and the diamond then turns about another angle.
%This is the essence of the proposed dipolar detection. 
%shows the total torque when 1-2-3-4 NV transitions are in the ground state. It is very small because the torque from 
%2-4 and 1-3 cancel each other. Looking at each pair 2-4 and 1-3 however, one can see that the torque is large (here 20 MHz for one spin).

To observe the effect of such resonant dipolar interactions experimentally, we use similar parameters and magnetic field arrangement than when the diamonds were not levitating. The diamond crystalline direction with respect to the magnetic field direction is characterized with greater precision than ODMR by recording the spin-torques given to the diamond when a microwave drives the spin to the $m_s=-1$ state. 
The angle detection is performed by collecting the back-reflected green light from the diamond interface (see Fig. 2), separated from the excitation light using a beam splitter. 
Fig. \ref{data}-a) shows spin-mechanical detection of spin-resonances for three different $\bf B_{\rm em}$ amplitudes. 
At 10 and 25 G, one can observe 4 peaks in the spectrum that demonstrate spin induced torque on the diamond from the 4 classes of NVs.
At 17 G however, two classes merge at 2.75 GHz. This is where we expect to observe CR. 
%Note that perfect degeneracy would still make the diamond rotate when performing microwave scans since the microwave polarization is not the same for the two classes, and thus does not drive equally the two spin transitions.

Analysis (SI) suggest that here, since we observe a single degeneracy, the magnetic field crosses a plane that is perpendicular to the $(110)$ direction as shown in Fig. \ref{data}-a).
Fig. \ref{data}-b) shows the photoluminescence as a function of $\bf B_{\rm em}$ both experimentally (trace i) and numerically (trace ii).
As expected, the PL decreases across the degeneracies at around the same magnetic field value, as a consequence of the CR-induced reduction of the $T_1$ time.
Fig. \ref{data}-c), trace i) is a measurement of the diamond angular position acquired simultaneously to the PL. Trace ii) is the corresponding calculation.   
A pronounced variation of the reflected signal is also observed, demonstrating here the close correspondence between degeneracy and diamond rotation and letting us conclude that we are observing a significant change in the spin-torque as the dipolar interactions between the spins increase. Note that as opposed to the PL detection, the signal coming from the particle surface can increase or decrease depending on how the speckle is aligned to the fiber. This explains the differing shapes of the signals in the experiments and the simulations. 
Fitting this curve by a Gaussian, we deduce a width that is similar to the PL width (1.3G and 1.7G respectively). This gives a width of X MHz ? comparable to ???
Similar experiments were realized on different particles under different degeneracies and similar effects were observed.
In the SI, we show similar results taken under different bias magnetic fields and under a two-fold degeneracy. 
%Note that the present effect cannot be attributed to a change in the angular momentum, as in the Einstein de Haas effect. 
%The latter would manifest itself for much smaller particles sizes. Further, the rotation would occur dynamically and not manifest itself as a static torque as is the case in the present experiment.

Let us conclude by mentioning the applications offered by our work. 
The observed resonant dipole-dipole induced mechanical rotation can be employed to control the temperature and stiffness of mechanical oscillators. 
For this, a delayed dynamical back-action from the diamond magnetization to the mechanical oscillator is required \cite{Aspelmeyer} in order to increase the imaginary part of the spin-mechanical susceptibility \cite{DelordNat}.
At a magnetic field value corresponding to a negative detuning from the CR feature, the NV fluctuator will partly depolarize the spins and let the two other transitions apply a torque until the spins re-polarize back, thus produces a potentially efficient cooling cycle under vacuum.  % In essence, this would be the analogue of the spin-cooling shown in \cite{DelordNat}, but without using microwave and using dipolar-coupled spins.

Conversely, our demonstration can be viewed as a novel spectroscopic technic to mechanically sense dipolar interactions between NV centers and spins that cannot be polarized optically.  Using the above cross-relaxation effect, one could use a magnetic field that is oriented along the [111] direction and detect paramagnetic species that do not have 
a strong anisotropy using a magnetic field gradient.  This opens a path towards the investigation of angular momentum conservation during the relaxation processes \cite{Zangara}.
%ODMR and EPR detect energies shifts but not angular momenta.  It would be beneficial to detect the dipolar interaction using the motion of the body itself in order to gain knowledge about angular momentum exchange as proposed in .
%The dipolar interaction can readily be detected indirectly using the motion of the object itself or via the Zeeman shift or motion of close paramagnetic or ferro-magnetic objects in the presence of a magnetic field.

Last, and more prospectively, one could consider this work to be a kind of bottom-up approach for magnetism, where both the interaction and the relaxation can be tuned in order to reproduce the behaviors of a magnet. (may also form a testbed for nano-magnetism).
The detailed microscopic origin of magnetism depends a lot on the material and the relaxation mechanism is not controlled and is very short, making measurements of the  Gilbert-damping a complicated task.
%The microscopic origin of the angular momentum exchange is also not well-understood. 
Our work may thus open a path towards using NV centers to probe microscopic effects that take place in nano-scale magnetism.

%To conclude, we demonstrate dipole-dipole induced rotation of a levitating diamond containing nitrogen-vacancy (NV) centers. 
%Our demonstration opens a path towards the use of mechanical oscillators to detect paramagnetic defects that lack optical transitions, to investigation of angular momentum conservation in spin relaxation processes and to novel means of cooling the motion of mechanical oscillators.

\section*{acknowledgements}
GH acknowledges SIRTEQ for funding.

\bibliography{trap}

\end{document}

### 2021-02-01 Talks/Défaut 2020/presentation.tex

@@
 \begin{itemize}
-\item CR experiments are much simpler to setup (low $\vec B$, no microwave)
-\pause
+\item CR experiments are much simpler to setup (and less expensive!) : low $\vec B$, no microwave.
+\end{itemize}
+\includegraphics[scale=.3]{epr-converted}
+\end{frame}
+\begin{frame}{Comparison between cross-relaxation and EPR}
+Upsides :
+\begin{itemize}
+\item CR experiments are much simpler to setup (and less expensive!) : low $\vec B$, no microwave.
 \item Doable on micro and potentially nano-diamonds
@@
 \item{Optical detection of new spin defects in CVD diamond through cross-relaxations with NV centers}
-\item{Mechanical detection of NV magnetization enhanced by cross-relaxations}
+\item{Mechanical detection of NV induced magnetization enhanced by cross-relaxations}
 \end{itemize}
@@
 \end{frame}
+\begin{frame}{T1}
+\centering
+\includegraphics[scale=.28]{NV-NV_3}
+\end{frame}
 \end{document}

### 2021-02-01 abstract/IQFA 2020/Arnaque moldave/figure et caption.tex

--- 2020-10-29 abstract/IQFA 2020/Abstract_Pellet-Mary.tex
@@
-\documentclass[aps,twocolumn,showpacs]{revtex4-1}
+\documentclass[a4]{article}
 
@@
 \usepackage{ifpdf}
-\ifpdf
 \usepackage{epstopdf}   
 \usepackage{url}
-\fi
 
 \begin{document}
-\title{Mechanical and relaxation-based detection of dipolar-interactions between spins in diamond}
+\section{Toto}
+Blabla
+\begin{center}
 
-\author{Clément Pellet-Mary$^1$, Maxime Perdriat$^1$, Paul Huillery$^1$, and Gabriel Hétet$^{1*}$}
+\begin{figure*}[ht]
+\includegraphics[scale=.6]{CRmeca}
+\caption{ i) ESR spectra while scanning B$_{em}$ with some bias field applied.
+ii)Angular position of the diamond while scanning B$_{em}$ without a microwave.
+The rotation of the diamond at 17 G corresponds to the resonance between two classes of NV}
 
-\affiliation{$^1$Laboratoire de Physique de l’Ecole Normale Supérieure, ENS, Université PSL, CNRS, Sorbonne Université, Université de Paris, F-75005 Paris, France.\\$^*$ gabriel.hetet@phys.ens.fr
-}
-
-\begin{abstract}
-\normalsize
-The negatively charged nitrogen vacancy center (NV$^-$) is a well established quantum system, employed both for applications and fundamental research, thanks to its good coherence properties, long lifetimes and most importantly their ability to be optically polarized.
-We have recently found new ways to use these properties of NV centers in order to study the dipolar interactions between ensemble of spins in diamond, including NV centers as well as other spin defects.
-
-On the one hand, we have observed cross-relaxations (CR) between NV centers and other defects (see Fig.
-1), namely VH$^-$ \cite{VH}, War1 (first defect found by EPR at the university of Warwick, still chemically unknown) and $^{13}$C First-shell modified NV.
-None of these defects had been observed through CR before.
-This is a first step toward hyper-polarization of new spin defects in diamond.
-
-On the other hand, we have used the mechanical detection of a levitating diamond \cite{nature} to measure the torque applied by the NV centers on the diamond when two classes of NV with different orientation are brought into resonance(see Fig.
-2).
-This effect is due to the dipolar-mediated modification of the T$_1$ of NV centers happening with dense ensembles.\cite{Lukin}.
-This paves the way toward the mechanical detection of other spin impurities in diamond, as well as the observation of the Einstein-de Haas effect with paramagnetic systems.
-\cite{Meriles}
-
-\begin{center}
-\includegraphics[scale=.25]{CR_total}
+\end{figure*}
 \end{center}
-
-Figure 1 : Photoluminescence change while scanning the magnetic field along the crystalline [100] direction. Three dips are observed for the CR with $^{13}$C-NV, VH$^-$ and War1.
-
-\begin{center}
-\includegraphics[scale=.25]{CRmeca}
-\end{center}
-
-Figure 2 : Angular position of the diamond while scanning B$_{\mathrm{Scan}}$ without a microwave. The rotation of the diamond at 17 G corresponds to the resonance between two classes of NV
-\end{abstract}
-
-\maketitle
-
-\begin{thebibliography}{}
-
-\bibitem{VH} Glover, C., et al. Physical review letters 90.18 (2003): 185507.
-
-\bibitem{nature} Delord, T., et al. Nature 580.7801 (2020): 56-59.
-
-\bibitem{Lukin} Choi, J., et al. Physical review letters 118.9 (2017): 093601.
-
-\bibitem{Meriles} Zangara, P.R., et al. Physical Review B 100.23 (2019): 235410.
-
-\end{thebibliography}
 \end{document}

### 2021-02-01 abstract/IQFA 2020/fiche.tex

--- 2020-11-17 Talks/IQFA 2020/Script.tex
@@
 \usetikzlibrary{fit}
+\usepackage[margin=0in]{geometry}
 
@@
 \begin{document}
-\section{intro}
-Thanks to the organizer for inviting me.
-Today I'm going to present to you two of our recent results regarding dipolar spin interaction in diamond.
-The first one is the optical observation of interaction between Natrogen-Vacancy (or NV) spins and new spin species in diamond, and the second is the mechanical detection of dipolar interaction between NV centers.
-
-But first, I will have to briefly introduce the physics of NV centers.
-
-\section{NV center}
-%Slide avec un NV dans sa maille cristalline et les niveaux d'énergie (juste excité
-% et fondamental) avec un zoom sur le niveau d'énergie fondamentale.
-A Nitrogen-vacancy center, which consists of nitrogen and a vacancy in the diamond crystal lattice, is a colored center, because it emits red light when optically excited (in our case with a green laser).
-However, the main reasons NV centers are widely used in quantum technology are not its optical properties, which are rather poor, but its spin properties.
-
-An NV center is a Spin-1 specie with no spin-orbit coupling and relatively good spin properties : a lifetime of a few milliseconds at room temperature, and a coherence time that can reach ???
-for single spins.
-In our case since we are working with relatively dense ensembles, we do have an inhomogenous broadening of a few MHz.
-
-But the most important property of the NV spin is that it can be optically polarized : At zero magnetic field, because of (origin of D???), the ms=0 state has an energy level about 3 GHz lower than that of the magnetic states ms=+/-1.
-At room temperature, this means that the ms=0 state is more populated by a whole .0000001\%.
-But when you are optically exciting (Reformuler ca) the NV center,because the optical cycle is different for the ms=0 and the magnetic states, you start pumping population in the 0 state up to about 80\%, which would be equivalent to a Temperature of 0.6 K.
-Another important property, which is related, is that the ms=0 state is brighter than the magnetic state, (quantum ratios ???).
-
-So to sum up, exciting an NV center with a green laser allows us to polarize the NV spin, and looking at the photoluminescence emitted by the NV allows us to measure that polarization.
-
-\section{Magnetometry with NV centers}
-The most common use of NV centers is probably magnetometry.
-The reason is rather simple : thanks to the optical pumping and the ability to read the quantum state, you can quite easily find the energy levels of your spin states, and since these levels depend on the magnetic field though the Zeeman effect, you can deduce the magnetic field.
-The simplest protocol is called Optically Detected Magnetic Resonance (ODMR).
-This protocol consist in continuously shining a green laser on your diamond and looking at the photoluminescence (or PL) while scanning a microwave frequency.
-Whenever the microwave is resonant with a 0 to -1 or 0 to +1 transition, the spin is partially depolarized in a magnetic state and the PL drops.% Plot avec un ODMR de single spin.
-
-This would be what you observe with a single spin, but we are dealing with ensemble.
-And an important aspect I did not mention so far is that the ms=0,+1 or -1 states are implicitly the eigenstates of a Sz operator, with a particular direction z.
-And since we have a zero-field splitting term, the direction of z is not imposed by the magnetic field.
-To the fist order, as long as the zeeman splitting is lower than than the zero field splitting, we can consider that the z axis is given by the zero field splitting, and the Zeeman splitting will simply be the projection of the magnetic field along the z axis.
-The physical z axis would then simply be the one formed by the nitrogen and the vacancy.
-And since there are four possible crystalline orientation in a diamond, one for each valence bond of the carbon, there are four possible z axis for our NV centers, and therefor 4 possible projection of the magnetic field.
-%Un peu de géométrie
-
-Going back to the ODMR, when dealing with ensemble, this is what you would observe : 4 transitions of 0 to -1, one for each possible orientation, or class, of NV centers, and 4 0 to +1 transitions.
-
-\section{Polarization transfer}
-Ref : Kamp, E.
-J., Carvajal, B., \& Samarth, N.
-(2018).
-Continuous wave protocol for simultaneous polarization and optical detection of P1-center electron spin resonance.
-Physical Review B, 97(4), 045204.
-
-le science 2016 de Lukin, le PRX de Wallsworth
-
-Another thing you can do when you have a polarized spin is to transfer its polarization, through dipolar interaction,  to another spin species.
-This is an especially active area for NV application with Dynamic Nuclear Polarization where you can transfer some of the NV polarization to nuclear spins either in the diamond (papier 13C et 14N), or outside the diamond (papier Lukin et wallsworth).
-But you can also transfer its polarization to a (dark) electron spin.
-So far, as far as y know, there has only been transfer to electric spin inside the diamond, and only to one particular spin which is the substitutional Nitrogen N minus defect, often called P1 center.
-
-There are several ways to transfer polarization from one spin to another, but the simplest one is cross-relaxation : by tuning the magnetic field you can sometimes match the energy transition of diferent spins species and they will start to exchange spin quanta through RESONANT dipolar interaction, which tend to balance the spin polarization of the two spins.
-
-This leads me to our first result :
-\section{Cross-Relaxation with NV centers}
-%1ere slide : Scan + en dessous 100 et puis soustraction par dessus.
-What we see here is the change in photoluminescence of our NV centers when we scan a magnetic field along the 100 direction of the diamond. I'll start detailing what we observed.
-
-The first thing to discuss is the 100 direction.
-It is a defined by the crystaline planes of the diamond, and it correspond to a direction where the magnetic field projection is the same on all four classes of the diamond, so in a way we are back to the single spin case which will make analysis much easier.
-
-The second thing to discuss is the overall decrease of the photoluminescence with the magnetic field.
-This is because every class is misaligned with the magnetic field, which means that by increasing the field amplitude, you will increase both the longitudinal field, so the zeeman splitting Bz*Sz in the Hamiltonian, but you also increase the transverse magnetic field component, so Bx*Sx which will mix the bright ms=0 state with the dark ms=+/-1 states and decrease the overall luminosity.
-
-So we expect a decreasing enveloppe (slide), and when we subtract the enveloppe of the signal, what remains is this :(new slide avec la soustraction) three clear features, none of which at been observed before and that we attributed to cross-relaxation of the NV center with three different species.
-The first
-
+\centering
+\textbf{Abstract :} The nitrogen-vacancy center's spin is a powerful sensor used in a large variety of applications.
+Here we use Cross-Relaxations (C.R.) to detect the presence of dark spins (VH$^-$ and WAR1) in CVD-grown diamond.
+We also managed to mechanically detect cross-relaxations between two classes of NV centers by measuring the torque applied on a levitating diamond.
   \end{document}

### 2021-02-01 théorie/fiches de lecture/Ashcroft-Mermin/fiche.tex

@@
 \usepackage[utf8]{inputenc}
-\usepackage[]{amsmath}
+\usepackage[]{amsmath,amssymb}
 \usepackage[]{braket} % \bra, \ket etc
@@
 
- En oubliant pas la dégénérescence du spin, on peut donc construire l'état fondamental en remplissant les niveaux d'énergies en partant du bas (quis de l'anti-symmétrisation ?), ce qui dans l'espace des k revient à remplir une boule vu que l'énergie c'est la distance à l'origine dans l'espace des k : La sphère de Fermi, de rayon $k_f$ (pour bien se rappeler que c'est dans l'espace réciproque).
+ En oubliant pas la dégénérescence du spin, on peut donc construire l'état fondamental en remplissant les niveaux d'énergies en partant du bas (quid de l'anti-symmétrisation ?), ce qui dans l'espace des k revient à remplir une boule vu que l'énergie c'est la distance à l'origine dans l'espace des k : La sphère de Fermi, de rayon $k_f$ (pour bien se rappeler que c'est dans l'espace réciproque).
 Et la relation avec la densité volumique (dans l'espace réel) d'état $n$ est : $$n=\frac{k_f^3}{3\pi^2}$$ Et $n$ on le connait, c'est globalement le même d'un métal à l'autre (en fonction du paramètre de maille et du nombre d'électrons de Valence mais toujours le meme ordre de grandeur).
@@
   Bon il se trouve que je suis peut-etre allé un peu vite en besogne, et que le dl de Sommerfeld doit se faire autour de $\mu$ plutot que de $\mathcal{E}_f$, ce qui change légèrement le facteur numérique.
-Mais bon le scaling est bon, et par rapport à un gaz classique $c_V(T)=\frac{3}{2}nk_B$, on voit qu'on a pris un facteur $\frac{k_BT}{\mathcal{E}_f}$ dans les detns, d'ou le fait que la chaleur spécifique des $e^-$ est 100x moins importante que pour un gaz parfait.
+Mais bon le scaling est bon, et par rapport à un gaz classique $c_V(T)=\frac{3}{2}nk_B$, on voit qu'on a pris un facteur $\frac{k_BT}{\mathcal{E}_f}$ dans les dents, d'ou le fait que la chaleur spécifique des $e^-$ est 100x moins importante que pour un gaz parfait.
 Par contre ça fait un scaling linéaire avec T, alors que les solides à basse température c'est en $T^3$, donc à basse température (qq K) les métaux ont une chaleur spécifique essentiellement due aux $e^-$ de conductions.
@@
   \end{itemize}
+  \chapter{Diffraction par rayon X}
+  L'énergie d'un photon pour avoir $\lambda$ de l'ordre de l'Angstrom est $\sim 10^4$ eV.
+Donc faut du bon gros rayon X qui tache pour sonder ça.
+Du coup diffraction par un réseau périodique (tu peux aussi faire des rayons X sur un un solide amorphe mais t'auras pas des pics), tu vas avoir des pics.
+Y'a deux façons de voir ça : Bragg ou van Laue.
+  \section{Réflexion de Bragg}
+  C'est la méthode avec le réseau direct.
+On considère un faisceau incident avec une longueur d'onde $\lambda$ et une famille de plans séparés d'une distance $d$.
+En faite on s'en fout un peu que chaque plan forme un réseau d'atomes, la seule chose qui nous interesse c'est l'angle $\theta$ entre le faisceau incident et la normale des plans, et la distance $d$ entre ces plans.
+
+  On considère que chaque plan cristallin réfléchit une partie du faisceau incident avec un angle $-\theta$ (pourquoi ?
+au niveau microscopique).
+Alors tu vas observer un pic de réflexion quand les réflexions par tous les plans interfèrent constructivement, soit $$\frac{2d \sin \theta}{\lambda} = n \in \mathbb{N}$$Ou $n$ est l'ordre de diffraction.
+Du coup chaque famille de plan a des pics correspondant, ou $theta$ doit être considéré par rapport à la normale des plans.
+Donc c'est vite le bordel en vrai.
+  \section{Formulation de Van Laue}
+  Yippity Yippita, pour une onde incidente de vecteur d'onde $k$, tu auras un pic de réflexion avec un vecteur d'onde $k'$ si $K=k'-k$ appartient au réseau réciproque.
+
+  C'est un peu sioux, mais vu qu'ici on laisse libre deux vecteurs ($k$ et $k'$) alors que la famille de plans sondés ne correspond qu'au seul vecteur $K=k-k'$, ça veut dire que tu as de la redondance (plusieurs couples $k$ et $k'$ qui correspondent au même plan de diffraction).
+En faite tu peux défini les \textbf{plans de Bragg} qui sont des plans dans l'espace réciproque tels que si $k$ appartient à ce plan, alors il existe un $k'$ qui permette d'observer un maximum de réflexion pour le plan recherché (en gros il existe $k'$ tel que $K=k-k'$, ce qui n'est pas toujours vrai pour une diffusion élastique vu que $k$ et $k'$ doivent avoir la meme norme).
+  \section{Construction d'Ewald}
+  Bon c'est le dessins du futurs qui ressemblent au tricercle de ses morts.
+Mais en vrai ça va.
+L'idée c'est que les plans de Bragg par défaut ils pavent pas du tout l'espace (c'est un peu comme des plans cristallins, mais dans l'espace réciproque.
+Bref).
+Donc si tu veux voir des pics tu as en gros deux solutions :
+  \begin{itemize}
+  \item Varier la longueur d'onde/ utiliser des rayons X poly-chromatiques pour scanner une région plus large. C'est la \textbf{Méthode de Laue}
+  \item Faire tourner le cristal/utiliser des poudres (\textbf{méthode de Debye-Scherrer}).
+Étonnament faire tourner le cristal c'est équivalent à faire tourner le réseau réciproque, ça reste de la magie noire pour moi tout ça.
+  \end{itemize}
+  \section{Facteurs de structure}
+  On a fait l'hypothèse que l'onde était réfléchie par les plans cristallins, sauf que les plans cristallins c'est pas forcément des plans atomiques, c'est uniquement le centre du motif cristallin.
+Du coup pour les réseaux ou la maille n'est pas mono-atomique (genre le diamant), il faut rajouter un \textbf{facteur de structure géométrique} qui vient moduler l'amplitude de chaque pic chaque pic de diffraction.
+C'est pas hyper prédictif d'un point de vue quantitatif parce que t'as encore d'autres facteurs qui viennent se rajouter après, mais t'as des fois ou le facteur de structure fait 0, et dans ce cas la c'est niet, tu as pas de pic.
+
+  Ce qu'il faut retenir c'est que considérer des motifs plus compliqués par rapport à des motifs simple, ça ne peut que te retirer des pics de diffraction, pas t'en ajouter.
+Par exemple si tu prends un réseau CC mais que tu le considère comme un réseau CS à motif (le réseau CS ayant plus de points dans le réseau réciproque parce que les distances sont plus lointaines), tu verras que la moitié des points du réseau réciproque ne donnent pas de pic avec le facteur de structure, parce que le réseau réciproque d'un CC c'est un CFC qui a deux fois moins de points qu'un CS.
+  \chapter{Classification des réseaux, systèmes, groupes d'espaces et toutes ces conneries}
+  Bientot on fera de la physique, c'est promis. 
+  \begin{tcolorbox}
+  \textbf{Groupe de symmétrie :} Dans ce contexte, c'est l'ensemble des opérations rigides (qui laisse invariantes les distances entre chaque point, aka \textbf{isométrie}) qui transforme le réseau en lui-même (en considérant tous les points comme équivalents).
+  \end{tcolorbox}
+  \section{Classsification des réseaux de Bravais}
+  On va commencer par considérer des réseaux de Bravais "pure", c'est à dire qu'à chaque noeud du réseau on a un objet de symmétrie maximale (une sphère, un atome quoi).
+Je rappelle que l'idée avec les symmétries c'est de soustraire.
+  \subsection{Invariance par translation}
+  C'est un peu la définition du réseau de Bravais, tu est invariant par translation selon un vecteur $\mathbf{R}=n_1 \mathbf{a_1}+n_2 \mathbf{a_2}+n_3 \mathbf{a_3}$
+  \subsection{Invariance par rotation, réflexion, inversion}
+  \begin{itemize}
+  \item \textbf{Rotation :} Une rotation est définie par une \textbf{axe} (1D, orienté) et un \textbf{angle} de rotation autour de l'axe.
+  \item \textbf{Reflexion :} Une réflexion (symmétrie miroir) est définie par un \textbf{plan} (2D) : une réflexion $M'$ est à la même distance $d$ que $M$ du plan, et $MM'$ est perpendiculaire au plan.
+  \item \textbf{Inversion :} Une inversion est définie par rapport à un \textbf{point} (0D) : elle transforme tout point $M$ tel que $\vec{OM} = \vec r \to -\vec r = \vec{OM'}$.
+Tous les points du réseau de Bravais sont des centres d'inversion (les coordonnées sont algébriques)
+  \end{itemize}
+  \section{Systèmes cristallins et groupe ponctuel}
+  Le \textbf{groupe ponctuel}, c'est l'ensemble des transformations qui laissent au moins un point du réseau invariant (donc en gros les pas-translations).
+Après mure réflexion, il se trouve que le groupe ponctuel n'est pas Abelien.
+Il existe 7 groupes ponctuels distincts pour des réseaux de Bravais qui définissent les 7 \textbf{Systèmes cristallins} :
+
+   \textbf{cubique}, \textbf{tétragonal} (pavé droit base carré), \textbf{orthorombique} (pavé droit de base rectangulaire), \textbf{monoclinique} (on a plus que deux angles droits), \textbf{triclinique} (plus d'angles droit du tout, c'est le cas le plus général).
+Tu as aussi le \textbf{trigonal} qui est un cube étiré selon la diagonale (normes des $a_i$ égales et angles entre chaque $a_i$ égaux, c'est le losange en 3D en gros) et \textbf{l'hexagonal} (pavé droit de section hexagonale regulière.
+  \section{Groupes d'espace et groupes ponctuels}
+  Le groupe d'espace c'est le groupe qui inclue la translation en plus.
+Chaque groupe d'espace "appartient" à un groupe ponctuel, par exemple les 3 groupes d'espaces pour le système cubique c'est le cubique simple, le CC et le CFC.
+
+  Maintenant si on ne s'interesse plus au réseau de bravais "pur", mais à  tout réseau cristallin, il faut augmenter le nombre de groupes ponctuels et de groupes d'espace, cf la table en dessous :
+
+  \begin{center}
+  \begin{tabular}{ccc}
+\hline
+ & Groupes ponctuels & Groupes d'espace \\
+\hline
+Bravais & 7 & 14 \\
+Général & 32 & 230 \\
+\hline
+\end{tabular}
+  \end{center}
+
+ Et on s'en tiendra à ça pour l'instant.
+ \chapter{Niveaux électroniques dans un potentiel périodique}
+ Le retour de la physique
+ \section{Théorème de Bloch}
+ Pour rappel, ce qu'on avait considéré dans le chapitre 2 c'était un gaz d'électron, quantique certe, mais libre.
+C'est pour ça qu'on s'est emmerdé à décrire la structure cristalline, pour décrire le potentiel effectif des ions sur les électrons.
+Mais dans un premier temps on ne vas utiliser que deux propriétés :
+ \begin{itemize}
+ \item Les électrons sont toujours supposés indpéendants, on résoudra Shrondinger à un électron
+ \item Le potenteil électrique généré par les ions est périodique (et indépendant du temps, on considère les ions immobiles toujours), soit $ V(r+R)=V(r) $ pour tout R appartenant au réseau de Bravais.
+ \end{itemize}
+ \begin{tcolorbox}
+ \textbf{Théorème de Bloch :} La fonction d'onde des états propres à un électron dans un potentiel périodique peut s'écrire comme une onde plane multipliée par une fonction périodique sur le réseau : $$\psi_{n\mathbf{k}}(r) = e^{i\mathbf{k}\cdot\mathbf{r}} u_{n\mathbf{k}}(r) $$
+
+ Avec $u_{n\mathbf{k}}(r+R)=u_{n\mathbf{k}}(r)$ pour tout $R$ appartenant au réseau de Bravais.
+
+ Une formulation alternative est : $$ \psi_{n\mathbf{k}}(r+R) = e^{i\mathbf{k}\cdot\mathbf{R}} \psi_{n\mathbf{k}}(r)$$
+ \end{tcolorbox}
+ Au passage, comme on regarde les états propres d'un Hamiltonien ne dépendant pas du temps, les états propres (= stationnaires) ne dépendant pas du temps.
+Le $n$ correspond à la n-ième énergie propre, et le $\mathbf{k}$ au vecteur d'onde.
+
+ La démonstration n'est pas très compliquée, on utilise le fait que $\mathcal{H}$ commute avec l'opérateur de translation selon un vecteur de Bravais $T_R$, ce qui veut dire qu'il esxiste une base commune pour $\mathcal{H}$ et $T_R$.
+Or on peut montrer assez facilement que les états propres de $T_R$ vérifient tous la deuxième formulation du théorème de Bloch, donc les états propres de $\mathcal{H}$ aussi.
+
+ \subsection{Conditions aux limites de Born-Von Karman}
+ Dans l'absolu $k$ est complexe, ce qui est un peu chiant.
+Du coup on reprend les même conditions aux limites périodiques (sur l'ensemble du cristal) qu'avant, c'est à dire : $$ \psi_{n\mathbf{k}}(r+N_i \mathbf{a}_i) = \psi_{n\mathbf{k}}(r)$$ Ou les $a_i$ sont les vecteurs de la base du réseau de Bravais, et $N=N_1N_2N_3$ est le nombre total d'atomes du cristal.
+
+ On retrouve alors que la maille primitive du réseau réciproque (aka PZB) contient N modes, et que le volume de chaque mode vaut $$ \Delta \vec k = \frac{(2\pi)^3}{V} $$
     \end{document}

### 2021-02-07 Mes papiers/Méca CR/Cooling_cross.tex
toto

%\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4}
\documentclass[preprintnumbers,amsmath,amssymb]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}

%\newcommand{|}{Y$_2$SiO$_5$}
\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{Cooling the motion of a trapped diamond via NV cross-relaxation}

\maketitle

\tableofcontents

\section{Librational mode susceptibility}

TODO : Inclure couplage au cm. Décrire Sx et Sz comme le fait Arcizet (couplage dispersif et couplage résonnant...)

The equation of motion for the librational mode $\theta(t)$ is

\begin{eqnarray}I \ddot \theta&=& -I \omega_\theta^2 \theta -I\gamma \dot \theta+\Gamma_{\rm spin}+\Gamma_T,
\end{eqnarray}\label{eqnmotion}
where $I$ is the moment of inertia, $\omega_\theta$ is the angular frequency for that mode. $\gamma$ is the damping rate due to collisions with the background gaz and $\Gamma_T(t)$ is the associated Langevin torque.
$\Gamma_{\rm spin}(\theta)$ is a torque that depends on the angle $\theta$. Fourier Transforming this equation and writing each quantity as $f(\omega)=\langle f \rangle+\delta f (\omega)$ yields to first order 
$$\delta \theta(\omega)=\chi_{\rm eff}(\omega)\delta\Gamma_T(\omega)$$
where
\bea
\chi_{\rm eff}(\omega)&=&\frac{1}{I(\omega_\theta^2-\omega^2-i\omega \gamma)-K(\omega)} \quad {\rm and }\quad K(\omega)=\frac{K_s}{1+i\omega\tau}\\
\eea
Here $K(\omega)$ is a dynamical spin-stiffness. 
It was found by linearizing $\Gamma_{\rm spin}(\omega,\theta)$ about an angle $\theta_0$, yielding $\Gamma_{\rm spin}(\omega,\theta)=\Gamma_{\rm spin}(\theta_0)+K(\omega) (\theta-\theta_0)$. Such a dynamical spring effect arises in the presence of a back action from the mechanical oscillator onto the spins. Note that $K_s$ can either be positive (binding) or negative (anti-binding).
We can rewrite the susceptibility in a condensed form 
\bea
\chi_{\rm eff}(\omega)&=&\frac{1}{I(\tilde\omega_\theta^2-\omega^2-i\omega \tilde\gamma)}
\eea
with 
\bea
\tilde\omega_\theta^2=\omega_\theta^2\big[1+\frac{K_s}{K_t}\frac{1}{1+(\omega_\theta\tau)^2}\big] \quad {\rm and} \quad \tilde\gamma=\gamma\big[1-Q\frac{K_s}{K_t}\frac{\omega_\theta\tau}{1+(\omega_\theta\tau)^2}\big]
\eea
the modified damping a frequencies of the mechanical oscillator.
Here $K_s=I\omega_\theta^2$ is the trap stiffness and $Q=\omega_\theta/\gamma$ is the trap quality factor. A strong delay $\tau$ compared to the trap period $2\pi/\omega_\theta$ enables cooling or heating if $K_s$ is positive/negative.

In practice, there are several ways to achieve such a dynamical stiffness. Photothermal, spin polarisation delay (our nature), ...
Another way induce dynamical backaction : cross-relaxation from two-coupled NV centers.

We need a torque $\Gamma_{\rm spin}(\theta)$ that varies on a small angular scale to induce a strong restoring force $K$. 
Let us first discuss how to attain a static torque and then how it can be modified using cross-relaxation.

\section{Static torque}

The static torque is provided by NV centers in the presence of an off-axis magnetic field. 
Physical picture:  Under spin polarisation in the ground state, an off-axis magnetic field tilts the spins away from $z$ (where the spin-spin interaction is minimized) : a torque is then applied to the levitation object to align
back the two spins of the NV spin one triplet to the magnetic field.
Let us calculate the torque from a single NV first.  We will then estimate the total static torque from the four NV centers close to a degeneracy (where cross-relaxation leads to a modified stiffness).

\subsection{Static torque from a single NV}

The hamiltonian of a single NV under a magnetic field is 
\be
\hat{H}=\hbar D S_z^2+\hbar \gamma_e\bm S \cdot \bm B
\ee
We simplify the problem by assuming that the motion is in the $x-z$ plane and that $\gamma_e B \ll D$.
The torque operator 
\be
\hat \tau = -\partial \hat{H} /\partial \theta = \hbar \gamma_e B (\sin\theta \hat{S_x}-\cos\theta \hat{S_z}).
\ee
The mean value of the torque in terms of the reduced density matrix elements $\rho_{ij}$ in the basis of the $S_z$ eigenstates ($-1,0,1$) is
\be
\langle \hat \tau \rangle = \hbar \gamma_e B   (\rho_{-1-1}-\rho_{11})\cos\theta + \hbar \frac{\gamma_e B}{\sqrt{2}} S \sin\theta
\ee
where we introduced $S=\rho_{0,1}+\rho_{1,0}+\rho_{0,-1}+\rho_{-1,0}$.
The reduced density matrix elements are traced over a bath consisting of laser photons used to polarised the NV at a rate $\gamma_{\rm las}$, phonons or spin-fluctuators acting on the spin populations at a rate $R_1=1/T_1$ and 
$P_1$ or nuclear spins dephasing the electronic spin at a rate $1/T_2^*$.
Assuming that the pure dephasing $T_2^*$ is much shorter than the sum of the population relaxation $T_1/2$ and the laser induced repolarisation $\gamma_{\rm las}$ the equations of motion for the coherences are (I FORGOT THE SQRT(2) FACTORS)

\bea
\frac{\partial \rho_{01}}{\partial t}&=&-\frac{1}{2T_2^*} \rho_{01}-i\gamma_e B\cos\theta(\rho_{00}-\rho_{11})-i\rho_{01}(\gamma_e B \sin\theta +D)+\mathcal{O}(\frac{(\gamma_e B)^2}{D})\\
\frac{\partial \rho_{0-1}}{\partial t}&=&-\frac{1}{2T_2^*} \rho_{0-1}-i\gamma_e B\cos\theta(\rho_{00}-\rho_{-1-1})-i\rho_{0-1}(\gamma_e B \sin\theta -D)+\mathcal{O}(\frac{(\gamma_e B)^2}{D}).
\eea
Taking the complex value of these equations give the two other coherence terms. 
In writing those, we neglected the $\gamma_e B\rho_{-11}$ and $\gamma_e B \rho_{1-1}$ terms which are of order $(\gamma_e B)^2/D$. 
Writing the population difference $\pi=\rho_{00}-(\rho_{11}+\rho_{-1-1})/2$, we also have 
\bea
\frac{\partial \pi}{\partial t}=-(\frac{1}{2}R_1+\gamma_{\rm las})\pi+\gamma_{\rm las}+\mathcal{O}(\frac{(\gamma_e B)^2}{D}).
\eea
The source terms in the population dynamics is negligible since they are proportional to $\gamma_e B$ times the coherences (which are of order $\gamma_e B /D$).

The coherences will have an impact on the angle since they are responsible for the mean torque applied to the particle. 
However, the motional dynamics is very slow compared to the zero-field and magnetic field splittings. The latter are also larger then the deocherence rate $1/T_2^*$ so we can adiabatically eliminate the coherences and find 
\bea
\rho_{01}=\rho_{10}\approx-\frac{\gamma_e B \cos\theta}{D}(\rho_{00}-\rho_{11}) \quad {\rm and} \quad \rho_{0-1}=\rho_{-10}\approx-\frac{\gamma_e B \cos\theta}{D}(\rho_{00}-\rho_{-1-1})
\eea

We also have 
\bea
\rho_{11}-\rho_{-1-1}=\mathcal{O}(\frac{\gamma_e B}{D})
\eea
Note that these approximations are also valid in the presence of cross-relaxation. Re-injecting these expressions in the expression for the mean torque, we get 
\bea
\langle \hat \tau \rangle(t) = \alpha   \pi(t) +  \mathcal{O}(\frac{(\gamma_e B)^3}{D^2}).
\eea
where 
\be
 \alpha=-\frac{\hbar(\gamma_e B)^2}{D}\sin2\theta
\ee
is in fact the gradient of the energy in the ground state taken at the angle $\theta$.

Indeed, supposing that $\gamma B \ll D$, so that $H_{B}= \gamma_e B  ( S_x \cos\theta + S_z \sin\theta)$ can be treated as a perturbation to the spin-spin hamiltonian. 
The perturbed energy $\epsilon_0$ of $\ket{0}$ due to the transverse B field is 
$$ \epsilon_0=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B)^2 }{D} \cos^2\theta
$$
The angular derivative of which gives $\alpha$ (pas exactement ?! facteurs deux qui trainent).

The torque is zero at $\theta=\pi/2$ because the eigenstates of $S_z$ are not mixed so the ground state is the eigenstate $| S=1,m_s=0\rangle$ which is is non magnetic. It is maximized at $\theta=\pi/4$.
Note also that the torque tends to zero if $\pi$ tends to zero at any angle $\theta$. This is the case in the presence of strong spin-depolarisation, i.e. when $\hat\rho\rightarrow\frac{1}{3}\hat{I}$ (with $\hat{I}$ the identity operator).

At an angle $\theta=\pi/4$, using $10^9$ spins polarised in the ground state at a B field of 100 G, the static torque is comparable to the torque observed in the Nature paper and should enable a shift in the angular position to be observed.
The change in the torque due to the angular dependence in $\sin2\theta$ is fairly smooth. The associated dynamical-stiffness will thus be negligible and no cooling will be observed.
In principle, one can however use nearby spins that are coupled by dipolar interactions to increase the sensitivity of the torque to a change in angle. This can be done use cross-relaxation with dipolar-coupled NVs of a different class.
Before showing this, we will first investigate the total torque that can be observed in the presence of several NV orientations.

\subsection{Static torque from several NVs}

The above described spin-torque may actually be compensated by the torque from other NVs in the crystals. 
Let us estimate the total torque.
Fig. c) shows the total torque when 1-2-3-4 NV transitions are in the ground state. It is very small because the torque from 
2-4 and 1-3 cancel each other. Looking at each pair 2-4 and 1-3 however, one can see that the torque is large (here 20 MHz for one spin).

\begin{figure*}[!ht]
  \centering \scalebox{0.54}{\includegraphics{torque_NVs.eps}}
  \caption{a)  Energy levels in the two excited states for the 4 NV orientations under 120 Gauss. b) Energy levels in the ground state of the 4 NV orientations with their labels. The circle shows the degeneracy of the 2-4 transitions.
  c) Torque applied to the diamond when the spin are in the ground state. When all the transitions 1-2-3-4 contribute a strong frustration almost cancels the torque because the 2-4 and 1-3 apply an opposite torque.  When the degeneracy condition is met between the 2-4 depolarise due to cross-relaxation and thus a torque is restored. 
  }
\end{figure*}

\section{Cross-relaxation}

The starting point to get a dynamical back action to the libration, is thus to reach degeneracy and to kill the population in the ground state of one or several NVs in order to break the mechanical frustration and restore the torque, as depicted in Fig.1.c). 
Under small enough laser, cross-relaxation (CR) due to NV fluctuators resonant with the 2-4 transition will partly depolarise them and let the 1-3 transitions apply the torque until the spins repolarise back. This produces a cooling cycle.

In essence this forms the analogue of the spin-cooling proposal in the nature paper, but without microwave and using dipolar-coupled spins.
Modifier equation 14 pour inclure les 4 orientations et ajouter l'expression pour $R_1(\theta)$.

%
%The Langevin torque obeys the relation
%
%$$\langle \Gamma_T(\omega)\Gamma_T(\omega')\rangle=2\pi \delta(\omega+\omega')S_T(\omega)$$
%where 
%
%$$S_T(\omega)=-\frac{2kT}{\omega}{\rm Im}(\frac{1}{\chi(\omega)})$$
%is the Langevin torque spectrum.
%Using the susceptibility for the librational mode, one finds
% $$S_T(\omega)=2kT\gamma I$$.
%
%The librational spectrum is then be found to be 
%
%%$$S_\theta(\omega)=|\chi(\omega)|^2 S_T(\omega)$$
%
%\begin{eqnarray}
%S_\theta(\omega)&=&|\chi(\omega)|^2 S_T(\omega)\\
%&=&\frac{2\gamma kT}{I ((\omega_\theta^2-\omega^2)^2 +\gamma^2\omega^2)}
%\end{eqnarray}
%
%One can check that the equipartition theorem holds here so that 
%
%$$\frac{1}{2} I \omega_\theta^2 \langle \theta^2\rangle= \frac{1}{2} kT $$
%where $$\langle \theta^2\rangle=\int S_\theta(\omega) d\omega$$
%

\end{document}

### 2021-02-15 Mes papiers/Méca CR/18112020/SI/Meca_cross_SI.tex
tutu

@@
 \
-\title{Resonant Dipole-Dipole Induced Mechanical Rotation of a Diamond Crystal}
+\title{Observation of a Magnetic-Torque Enhanced by Interacting Spin Ensembles}
 
@@
 \begin{equation*}
-  \hat{\mathcal{H}}_s=\hbar D S_z^2 + \hbar \gamma_e \textbf{B}\cdot\hat{\textbf{S}},
+  \hat{\mathcal{H}}_s=\hbar D S_{z'}^2 + \hbar \gamma_e \textbf{B}\cdot\hat{\textbf{S}},
   \end{equation*} 
 where $D = (2\pi) 2.87$ GHz is the crystal field splitting originating from spin-spin interactions, and $\gamma_e = 28 $GHz/T is the electron gyromagnetic ratio. 
-We neglect contributions from the strain and local electric field in the hamiltonian since we are working with magnetic fields on the order of 100G.
-We also neglect the hyperfine interaction with the nuclear spin of the $^{14}$N atom since we are working with ensembles with a typical inhomogeneous broadening of 5 MHz.
-The \textbf{z} axis of the $S_z$ operator here is the axis formed by the nitrogen atom and the vacancy in the body frame.
+The \textbf{z'} axis in the $S_{z'}$ operator here is the axis formed by the nitrogen atom and the vacancy in the body fixed frame.
+We neglect contributions from the strain and local electric field in the hamiltonian since we are working with magnetic fields on the order of 0.1T, which induce splittings  larger than the splitting of the zero-field ESR line ($\approx 20$ MHz).
+We also neglect the hyperfine interaction with the nuclear spin of the $^{14}$N atom since we are working with ensembles with typical inhomogeneous broadening of 5~MHz.
 
@@
 The [100] direction is marked by red circles, the [111] direction by green squares.
-The path of the magnetic field in the experiment of Fig.
+The magnetic field path in the experiment of Fig.
 3 of the main text is shown with blue dots.}
@@
 \end{figure}
-There are four possible crystalline axis for the NV centers (so-called ``classes" of NV) which are depicted in Fig.
+There are four possible crystalline axes for the N-V direction (so-called ``classes" of NV) in the diamond.
+They are depicted in Fig.
 \ref{cristallo} b) and correspond to the crystalline directions [$111$], [$1\bar 1 \bar 1$], [$\bar 1 1 \bar 1$] and [$\bar 1 \bar 1 1$].
 
-The magnetic field is represented in the diamond basis in Fig.
-\ref{cristallo} a) where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z} ([$001$]) direction.
-
+The magnetic field direction is represented in Fig.
+\ref{cristallo} a), where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z'} ([$001$]) direction.
 For some orientations of the magnetic field, the projection of the magnetic field on two or more NV axes will be identical, and therefore the energy level of the corresponding classes will be the same.
-These degeneracies are represented in Fig \ref{cristallo} c), where the dashed lines are the locus of the crystalline planes orthogonal to the [110] directions (6 planes in total).
+These degeneracies are represented in Fig \ref{cristallo} c), where the dashed lines are the locii of the crystalline planes orthogonal to the [110] directions (6 planes in total).
 When the magnetic field belongs to these planes, we observe a degeneracy between two classes of NVs, as can be seen in the Fig.
@@
 
-The plain lines are the loci of the planes orthogonal to the [100] directions (3 planes total).
-When the magnetic field angle lies in these planes, every class is resonant with another class, as can be seen in Fig.
+The plain lines are the locii of the planes orthogonal to the [100] directions (3 planes in total).
+When the magnetic field lies in these planes, all classes are co-resonant, as can be seen in Fig.
 \ref{CR_22}.
@@
 
-\subsection{$T_1$ Measurement}
+\subsection{$T_1$ measurement}
 \begin{figure}[!ht]
@@
 As shown in the Fig. 2-c) of the main text, the spin lifetime of the NV centers is modified in the presence of cross-relaxation with other classes of NV centers.
-Here we present the protocol for removing the effects of charge state transfer in the dark, which masks the PL signal decay induced solely by spin depolarization. 
+Here we present the protocol for removing the effects of charge state transfer in the dark, which mask the PL signal decay induced solely by spin depolarization. 
 The protocol described in Fig.
 \ref{T1_protocol} consists in using two sequences.
-In the first sequence the spins are initially polarized in their $m_s=\ket{0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$.
+In the first sequence the spins are initially polarized in the $\ket{m_s=0}$ state through a 1 ms green laser excitation pulse and then left to evolve in the dark for a variable dark time $\tau$.
 The spin state is finally read out using a 10 $\mu$s laser pulse (shorter than the polarization time of the spins).
@@
 The second sequence uses the same parameters (polarization time, dark time and readout time) than the first sequence, but uses an extra resonant microwave $\pi$ pulse tuned to a transition of one of the four classes of NV$^-$ right before the readout pulse.
-The latter sequence brings the population of the $m_s=\ket{0}$ state to the $m_s=\ket{\pm 1}$ state of one class of NV centers.
+The latter sequence brings population from the $\ket{m_s=0}$ state to the $\ket{m_s=\pm 1}$ state for one class of NV centers.
 
-Writing $S_1$ the signal obtained by the first sequence and $S_2$ the one obtained by the second sequence, and then measuring the difference $S_1-S_2$ between the two, we can measure the spin state population of a single class, and at the same time remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark (which would be common to the two signals and therefore subtracted).
-In order to avoid low frequency noises or drift, we alternate both sequence when performing the measurement.
+By measuring the difference between the two signals obtained in these two measurements, we can extract the evolution of the spin state population from a single NV class and, at the same time, remove unwanted contributions to the photoluminescence, such as charge state transfer in the dark (which give the same background contribution to the to the measurements).
+In order to avoid low frequency noises such as laser drifts from the focal point or intensity fluctuations, we alternate both sequences while performing the measurement.
 
@@
 
-A permanent magnet is placed a few cm away from the diamond sample in order to apply a uniform magnetic field to the NV centers together with an electro-magnet.
+A neodymium permanent magnet and an electro-magnet are placed a few centimeters away from the diamond sample in order to apply a uniform and controllable magnetic field to the NV centers.
 
-To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonance (ODMR) spectra and use two orientations to determine both the angle and the magnetic field amplitude.
+To calibrate the magnetic field magnitude $B$, and its orientation $\theta$ with respect to the NV axis, we record Optically Detected Magnetic Resonance (ODMR) spectra and record two the frequency of two transitions $|0 \rangle \rightarrow |-1\rangle$ and $|0 \rangle \rightarrow |+1\rangle$ from the same class to determine both the angle of the B field with respect to this class and the magnetic field amplitude.
 
@@
 
-\subsection{Signal drift for levitating particles}
+\subsection{Angular signal drift for levitating particles}
 
@@
 
-The magnetic torque responsible for the motion of the diamond fundamentally comes from the crystalline anisotropy of the NV centers and from the transverse field $B_\perp$ responsible for mixing the eigenstates in the stationary state.
+The magnetic torque responsible for the motion of the diamond fundamentally comes from the anisotropy of the NV centers and from the transverse field $B_\perp$ responsible for mixing the eigenstates in the stationary state.
 We will start by considering the torque from a single NV center.
@@
 %\ee
-We simplify the problem by assuming that the motion is in the $x-z$ plane so the magnetic part of the hamiltonian reads 
-$H_{B}= \hbar \gamma_e B  (\hat{S_x} \sin\theta + \hat{S_z} \cos\theta)$
-where $\theta$ is the angle between the B field and NV center quantization axis.
-We thus obtain the spin torque 
+Without lack of generality, we will assume that the B field points in the $z$ direction and take the motion to be in the $x-z$ plane (in the lab frame), see Fig.~\ref{axes}.
+
+\begin{figure}[ht]
+\centering{ \scalebox{0.4} {\includegraphics{axes.eps}}}
+  \caption{Notations used to define the axes in the body fixed and laboratory frames of reference $\mathcal{R'}$ and $\mathcal{R}$ respectively.}
+  		\label{axes}
+\end{figure}
+
+In the body fixed frame, the magnetic part of the hamiltonian reads 
+$\hat {H}_{B}= \hbar \gamma_e B  (\hat{S}_{x'} \sin\theta + \hat{S}_{z'} \cos\theta)$
+where $\theta$ is the angle between the B field and NV center quantization axis $z'$.
+We thus obtain the spin torque operator
 \be
-\hat \tau_s = -\partial \hat{H} /\partial \theta = \hbar \gamma_e B (-\cos\theta \hat{S_x}+\sin\theta \hat{S_z}).
+\hat \tau_s = -\frac{\partial \hat{H}}{\partial \theta} = \hbar \gamma_e B (-\cos\theta \hat{S}_{x'}+\sin\theta \hat{S}_{z'}).
 \ee
-The mean value of the torque in terms of the reduced density matrix elements $\rho_{ij}$ in the basis of the $S_z$ eigenstates ($-1,0,1$) in the body fixed frame is
+The mean value of the torque operator in terms of the reduced density matrix elements $\rho_{ij}$ in the basis of the $S_{z'}$ eigenstates $|-1\rangle_{z'},|0\rangle_{z'},|1\rangle_{z'}$ is 
 \be
-\langle \hat \tau_s \rangle = \hbar \gamma_e B   (\rho_{-1-1}-\rho_{11})\sin\theta - \hbar \frac{\gamma_e B}{\sqrt{2}} S \cos\theta\label{torquemean}
+\langle \hat \tau_s \rangle =\mathrm{Tr}_\mathcal{B}(\hat \rho \hat \tau_s) = \hbar \gamma_e B   (\rho_{-1-1}-\rho_{11})\sin\theta - \hbar \frac{\gamma_e B}{\sqrt{2}} S \cos\theta\label{torquemean}
 \ee
 where we introduced $S=\rho_{0,1}+\rho_{1,0}+\rho_{0,-1}+\rho_{-1,0}$.
-The reduced density matrix elements are traced over a bath consisting of laser photons used to polarized the NV at a rate $\gamma_{\rm las}$, phonons or spin-fluctuators acting on the spin populations at a rate $R_1=1/T_1$ and
-$P_1$ or nuclear spins dephazing the electronic spin at a rate $1/T_2^*$.
-In the limit $\gamma_e B \ll D$, and $\gamma_{\rm las}\gg \Gamma_1$ the laser still polarizes the spins in the ground state so that $\rho_{00}\gg {\rho_{11},\rho_{-1-1}}$.
-Assuming that the pure dephazing $T_2^*$ is much shorter than the sum of the population relaxation $T_1/2$ and the laser induced repolarization $\gamma_{\rm las}$ the equations of motion for the coherences are
+The bath $\mathcal{B}$ over which the trace is performed consists of laser photons used to polarized the NV at a rate $\gamma_{\rm las}$, phonons or spin-fluctuators acting on the spin populations at a rate $\Gamma_1=1/T_1$ and
+$P_1$ centers or nuclear spins dephazing the electronic spin at a rate $1/T_2^*$.
+In the limit $\gamma_e B \ll D$, and $\gamma_{\rm las}\gg \Gamma_1$ the laser efficiently polarizes the electronic spins in the ground state so that $\rho_{00}\gg {\rho_{11},\rho_{-1-1}}$.
+The pure dephazing $T_2^*\approx 100$ns is much shorter than the sum of the population relaxation time $T_1/2 \le 1$ms and the laser induced repolarization time $1/\gamma_{\rm las} \le $ 100 $\mu$s.
+The equations of motion for the coherences thus read
 
@@
 %The coherences can have an impact on the angle since they are responsible for the mean torque applied to the particle. 
-The motional dynamics is very slow compared to the zero-field and magnetic field splittings.
-The latter are also larger then the decoherence rate $1/T_2^*$ so we can adiabatically eliminate the coherences and find
+The characteristic motional dynamics is very slow compared to the zero-field and magnetic field rates $D$ and $\gamma B$.
+The latter are also much larger then the decoherence rate $1/T_2^*$ in our experiments, so we can adiabatically eliminate the coherences and find
 \bea
-\rho_{01}=\rho_{10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D} \quad {\rm and} \quad \rho_{0-1}=\rho_{-10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D}\eea
+\rho_{01}=\rho_{10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D} \quad {\rm and} \quad \rho_{0-1}=\rho_{-10}\approx-\frac{\gamma_e B \sin\theta}{\sqrt{2} D}.\eea
 Since 
@@
 
-Indeed, supposing that $\gamma B \ll D$, so that $H_{B}$ can be treated as a perturbation to the spin-spin hamiltonian, 
+Indeed, supposing that $\gamma B \ll D$, so that $H_{B}$ can be treated as a perturbation to the spin-spin hamiltonian $D \hat S_{z'}^2$, 
 the perturbed energy $\epsilon_0$ of $\ket{0}$ is 
-$$ \epsilon_g=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B)^2 }{D} \sin^2\theta
-$$
+\bea \epsilon_g=\sum_{m_s=\pm 1}\frac{ |\bra{0} H_B \ket{\pm 1}|^2}{-\epsilon^0_{\pm 1}}=-\hbar\frac{(\gamma_e B)^2 }{D} \sin^2\theta
+\eea
 
-The formula $ -\partial \epsilon_g /\partial \theta$ then gives Eq. \ref{torque_anal}. It is the equation that is used in the core of the manuscript. 
-In the approximate regime of the present study, the Feynman-Hellman theorem that relates the angular derivative of the mean energy to the torque in the pure state case, 
-can in fact also be applied directly here.
+Taking $ -\partial \epsilon_g /\partial \theta$ then gives Eq. \ref{torque_anal}. It is the equation that is used in the core of the manuscript. 
+In the approximate regime of the present study, the Hellmann-Feynman theorem (exact for pure states) that relates the angular derivative of the mean energy to the torque is correct in the above-described limits where dissipation is negligible.
+\\
 
@@
 \centering{ \scalebox{0.4} {\includegraphics{map_torque_1_classe}}}
-  \caption{\textbf{a)} Skecth showing an NV center aligned to the \textbf{z} direction in the presence of an external magnetic field \textbf{B} at an angle $\theta$ with respect the NV axis.
+  \caption{\textbf{a)} Skecth showing an NV center aligned to the \textbf{z'} direction in the presence of an external magnetic field \textbf{B} at an angle $\theta$ with respect the NV axis.
 The resulting spin vector $\langle \mathbf{ \hat S} \rangle$ of the NV center is shown by the blue arrow.
@@
 Another way to estimate the torque is depicted in Fig \ref{Torque1classe}.
-Under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, under the low magnetic fields ($<$ 200 G) we are working at, will be oriented at an angle of 90$^\circ$ from the NV axis : $\langle \hat S_z \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$.
+Under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, under the low magnetic fields ($<$ 200 G) we are working at, will be oriented at an angle of 90$^\circ$ from the NV axis : $\langle \hat S_{z'} \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$.
 This magnetization vanishes when the magnetic field is aligned with the NV center since there is no longer a transverse field responsible for the mixing of the eigenstates.
@@
 The magnetization of the NV center is therefore not aligned with the magnetic field, except when the field is also at a 90$^\circ$ angle from the NV axis, which means that the magnetic torque $\mathbf \Gamma = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B $ will be non-zero everywhere except when the field is aligned with the center, or in the plane normal to the direction of the center.
-We can describe each NV center as a paramagnetic defect with the anisotropic magnetic susceptibility $\chi = \begin{pmatrix}-\chi_\perp & 0 & 0\\ 0 & -\chi_\perp & 0 \\ 0 & 0 & 0  \end{pmatrix}$ in the $(\mathbf{x},\mathbf{y},\mathbf{z})$ basis where $\mathbf{z}$ is the orientation of the NV center.
+We can describe each NV center as a paramagnetic defect with the anisotropic magnetic susceptibility $\chi = \begin{pmatrix}-\chi_\perp & 0 & 0\\ 0 & -\chi_\perp & 0 \\ 0 & 0 & 0  \end{pmatrix}$ in the $(\mathbf{x'},\mathbf{y'},\mathbf{z'})$ basis where $\mathbf{z'}$ is the orientation of the NV center.
 
@@
 
-By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle n(NV^-)}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-4} rad$$ where $n(NV^-) \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 6.3\cdot 10^3$ rad/s is the typical value of the trap angular frequency.
+By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle n(NV^-)}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-3} rad$$ where $n(NV^-) \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 6.3\cdot 10^3$ rad/s is the typical value of the trap angular frequency.
 

### 2021-02-15 data/20210125/Adamas 11/Passage à l'ordi 2/traitement.py

@@ -109,6 +109,7 @@ def stretch_exp_fit(x,y,Amp=None,ss=None,tau=None) :
 	popt, pcov = curve_fit(f, x, y, p0)
 	return(popt,f(x,popt[0],popt[1],popt[2]))
 
+
 def stretch_arb_fit(x,y,Amp=None,ss=None,tau=None,alpha=0.5) :
 	if not Amp :
 		Amp=max(y)-min(y)
@@ -134,27 +135,29 @@ def stretch_soustraction(x,y,Amp=None,tau=None) :
 	popt, pcov = curve_fit(f, x, y, p0)
 	return(popt,f(x,popt[0],popt[1]))
 
-def third_stretch(x,y,Amp=None,tau=None) :
+def third_stretch(x,y,Amp=None,ss=None,tau=None) :
 	if not Amp :
 		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
 	if not tau :
 		tau=x[int(len(x)/10)]-x[0]
-	def f(x,Amp,tau) :
-		return Amp*np.exp(-(x/tau)^(1/3))
-	p0=[Amp,tau]
+	def f(x,Amp,ss,tau) :
+		return Amp*np.exp(-(x/tau)**(0.3333))+ss
+	p0=[Amp,ss,tau]
 	popt, pcov = curve_fit(f, x, y, p0)
-	return(popt,f(x,popt[0],popt[1]))
+	return(popt,f(x,popt[0],popt[1],popt[2]))
 
-def stretch_et_phonon(x,y,Amp=None,tau=None) :
-	if not Amp :
-		Amp=max(y)-min(y)
-	if not tau :
-		tau=x[int(len(x)/10)]-x[0]
-	def f(x,Amp,tau) :
-		return Amp*np.exp(-np.sqrt(x/tau)-x/5E-3)
-	p0=[Amp,tau]
+def stretch_et_phonon(x,y,Amp_stretch=None,Amp_exp=0.1,tau_stretch=None,tau_exp=5e-3) :
+	if not Amp_stretch :
+		Amp_stretch=max(y)-min(y)
+	if not tau_stretch :
+		tau_stretch=x[int(len(x)/10)]-x[0]
+	def f(x,Amp_stretch,Amp_exp,tau_stretch,tau_exp) :
+		return Amp_stretch*np.exp(-np.sqrt(x/tau_stretch))+Amp_exp*np.exp(-(x/tau_exp))
[…]

### 2021-02-19 Mes papiers/Méca CR/18112020/Meca_cross.tex

@@
 When they become resonant, the polarization of the different orientations can then be exchanged in a process called cross-relaxation.
+The conditions on the magnetic field for a cross-relaxation to happen are described in Sec.
+I of the Supplementary Material (SM)\cite{SM_CR_meca}.
+\nocite{DelordPRL} \nocite{delordPhD} \nocite{DelordNat} \nocite{qutip1} \nocite{qutip2} \nocite{van_oort_cross-relaxation_1989} \nocite{choi_depolarization_2017}
 The right panel of Fig.~\ref{principle}, shows a cross-relaxation mechanism that partly removes the contribution from two classes of NV centers (labelled 1 and 3 in Fig.~\ref{principle}), which breaks the four-spin rotational invariance.
@@
 %The dipolar interactions may thus be effective since the average distance between the NV centers of about 10 nm.
-We load the micro-diamonds in a Paul trap that is similar to the one used in \cite{delordPRL} and depicted in Fig.
+We load the micro-diamonds in a Paul trap that is similar to the one used in \cite{DelordPRL} and depicted in Fig.
 \ref{setup}-a).
@@
 The laser is focused via a lens inside a vacuum chamber which has a numerical aperture of 0.5 and a working distance of 8mm.
-The focal point of the laser is kept few tens of micrometers away from the micro-diamond to mitigate the effect of radiation pressure and to enable laser excitation of the whole diamond \cite{delordPRL,delord2016}.
+The focal point of the laser is kept few tens of micrometers away from the micro-diamond to mitigate the effect of radiation pressure and to enable laser excitation of the whole diamond \cite{DelordPRL,delord2016}.
 
@@
 Smaller diamond particles in fact tend to suffer from extra parasitic surface effects such as spin depolarization due to interaction with paramagnetic dangling bonds on the surface \cite{Tetienne}, or enhanced charge transfer between the NV$^0$ and NV$^-$ charge states \cite{Dhomkar} so it is essential to verify that it can be observed with micro-particles.
-We start by searching for CR using micro-diamonds that are physically attached to the trap, by employing a fixed bias magnetic field $||\bf B_{\rm bias}||\approx$100 G and by tuning another magnetic field $\bf B_{\rm em}$ at some angle with $\bf B_{\rm bias}$ using an electromagnet (see Fig.~\ref{setup}-b)).
+We start by searching for CR using micro-diamonds that are physically attached to the trap, by employing a fixed bias magnetic field $||\bf B_{\rm bias}||\approx$100 G and by tuning another magnetic field $\bf B_{\rm em}$ at some angle with $\bf B_{\rm bias}$ using an electromagnet (see Fig.~\ref{setup}-a)).
+The change in orientation of the total magnetic field can be visualized in Sec.
+I of the SM \cite{SM_CR_meca}.
 
@@
 Such a measurement however turned out to be significantly impacted by recharging of NV centers in the dark \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
-In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use the sequence presented in the SI, where a microwave pulse is applied or not prior to spin relaxation.
+In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use the sequence presented in Sec.
+II of the SM \cite{SM_CR_meca}, where a microwave pulse is applied or not prior to spin relaxation.
 The PL signals acquired in the two different measurements are then subtracted and shown for different degeneracy configurations in Figure \ref{setup}-c).
@@
 \end{equation}
-A proof of the applicability of this theorem is presented in the SI. 
+A proof of the applicability of this theorem is presented in  Sec. III of the SM \cite{SM_CR_meca}. 
 At an angle $\theta=\pi/4$, where the torque is maximized and at a B field of 100 G, we find $\tau_s\approx 5\times 10^{-17}$ N.m.
@@
 When two classes of NV center are resonant however, the induced cross-relaxation partly breaks this rotational invariance.
-Fig. \ref{Numerics}-b) shows the same plot, but including the effect of the cross-relaxation when classes of NV centers are degenerate (see SI).  
+Fig.
+\ref{Numerics}-b) shows the same plot, but including the effect of the cross-relaxation when classes of NV centers are degenerate.
+Details on the modelization can be found in Sec.V of the SM \cite{SM_CR_meca}.
 Here we use numbers that are deduced from our experimental observations of the CR induced change of the $T_1$.
@@
 The diamond crystalline direction with respect to the magnetic field direction is characterized with greater precision than ODMR by recording the spin-torques given to the diamond when a microwave drives the spin to the $m_s=-1$ state.
-The angle detection is performed by collecting the back-reflected green light from the diamond interface (see Fig. 2), separated from the excitation light using a beam splitter. 
+The angle detection is performed by collecting the back-reflected green light from the diamond interface (see Fig. 2-a), separated from the excitation light using a beam splitter. 
 Fig. \ref{data}-a) shows spin-mechanical detection of spin-resonances for three different $\bf B_{\rm em}$ amplitudes. 
@@
 
-Analysis (SI) suggest that here, since we observe a single degeneracy, the magnetic field crosses a plane that is perpendicular to the $(110)$ direction as shown in Fig. \ref{data}-a).
+Analysis developed in Sec.I of the SM \cite{SM_CR_meca} suggest that here, since we observe a single degeneracy, the magnetic field crosses a plane that is perpendicular to the $(110)$ direction as shown in Fig.
+\ref{data}-a).
 Fig. \ref{data}-b) shows the photoluminescence as a function of $\bf B_{\rm em}$ both experimentally (trace i) and numerically (trace ii).
@@
 This explains the differing shapes of the signals in the experiments and the simulations.
-Fitting this curve by a Gaussian, we deduce a width that is similar to the PL width (1.3G and 1.7G respectively). This gives a width of X MHz ? comparable to ???
+Fitting this curve by a Gaussian, we deduce a width that is similar to the PL width (2.1G and 2.8G respectively).
+This gives a width of 9 (resp.
+12) MHz  comparable to the inhomogeneous broadening of the sample.
 Similar experiments were realized on different particles under different degeneracies and similar effects were observed.
-In the SI, we show similar results taken under different bias magnetic fields and under a two-fold degeneracy. 
+In Sec. IV of the SM, we show similar results taken under different bias magnetic fields and under a two-fold degeneracy. 
 %Note that the present effect cannot be attributed to a change in the angular momentum, as in the Einstein de Haas effect. 

### 2021-02-19 Mes papiers/Méca CR/18112020/SI/Meca_cross_SI.tex

@@
 
+\section{Depolarization induced by NV-NV cross-relaxation}
+As explained in the main text, when the density of NV$^-$ centers in the sample is high enough (typically for concentrations higher than 1 ppm), the ensemble of NV spins will lose some of its polarization through dipolar coupling between the NV centers.
+This phenomenon is at the heart of the mechanism that allows us to exalt the magnetic suceptibility of our diamond through dipolar interaction, and it has already been observed independently by many groups in bulk diamond %refs.
+
+In particular, \cite{choi_depolarization_2017} proposes a model based on "fluctuators" : a subgroup of NV centers with a very short lifetime (possibly due to their electron tunnelling in and out of the NV site) that can act as a source of classical noise with a central frequency given by the usual transition frequencies of the NV center Hamiltonian.
+A prediction of this model is that the modified lifetime of the ensemble of NV centers should have a stretch exponential profile (of the form $\exp(-\sqrt{\frac{t}{T_1}})$) which we do observe experimentally.
+
+\subsection{Stretch exponential profile of the lifetimes}
+\begin{figure}
+\label{stretch}
+\includegraphics[scale=0.7]{T1-stretch_exp}
+\caption{Lifetime measurement for a case where all four classes are degenerate. Plain orange line correspond to a stretch-exponential fit and dashed red line to a simple exponential fit}
+\end{figure}
+In the theory developed in \cite{choi_depolarization_2017}, the stretch exponential profile arises from the inhomogeneity of the distance from each NV centers to the closest fluctuators.
+If $\rho_{00}^s(t)$, the population in the $\ket{m_s=0}$ state for each NV centers, follow a law of the form $\rho_{00}^s(t) \propto \exp(-\gamma t)$, then, assuming an even spatial distribution of fluctuators, the authors of \cite{choi_depolarization_2017} show that the distribution in $\gamma$ should follow a law of the form $$\rho(\gamma) \approx \frac{e^{-\frac{1}{4 \gamma T_1}}}{\sqrt{4 \pi \gamma^3 T_1}} $$
+Averaging then over all NV centers gives the stretch exponential profile observed from the ensemble :
+$$ \rho_{00}^e(t) \propto \int_0^{+\infty} \rho(\gamma) e^{-\gamma t} d\gamma = e^{-\sqrt{\frac{t}{T_1}}} $$
+where $ \rho_{00}^e(t)$ correspond to the average population in the $\ket{m_s=0}$ state for the ensemble of spins.
+
+Fig.
+\ref{stretch} shows a lifetime measurement on a static micro diamond following the protocol described in sec.III.
+Here all four classes are resonant with the applied microwave frequency, which correspond to the maximum degree of degeneracy between the NV centers, and therefore the stronger modification of the lifetime induced by the resonant dipolar coupling.
+The signal we obtain was fitted using a stretch exponential profile  and a simple exponential profile.
+We can see that the stretch exponential profile ($R^2=0.981$) is in better agreement with the data than the exponential fit ($R^2=0.942$), in particular for the very short times (we expect the longer times to be dominated by the phonon-limited exponential lifetime).
+
+Finally it should be noted that the stretch exponential profile arising from point-like depolarization sources is a relatively general result that has for example also been observed for the depolarization of NV centers induced by substitutional nitrogen (P1) defects in diamond \cite{Hall}
+
+\subsection{•}
+
+The easiest way to probe the mechanism of dipolar-induced modification of the lifetime, as well as to justify the fact that the effects come in fact from the NV-NV dipolar interaction, is to change the degeneracy conditions between the four classes of NV centers by tuning the magnetic field.
+(ref sec.I)
+Because the NV spins can only exchange spin quanta when they are quasi-resonant, then by tuning the number of at degeneracy, you can modify the effective density of NV centers able to by reso
 \section{Experimental details}

### 2021-02-22 Mes papiers/Méca CR/18112020/SI/Meca_cross_SI.tex

--- 2021-02-15 Mes papiers/Méca CR/18112020/SI/Meca_cross_SI.tex
@@
 
-\author{C. Pellet-Mary, P. Huillery, M. Perdriat, G. H\'etet}
+\author{C. Pellet-Mary$^1$, P. Huillery$^1$, M. Perdriat$^1$, G. H\'etet$^1$}
 
@@
 The \textbf{z'} axis in the $S_{z'}$ operator here is the axis formed by the nitrogen atom and the vacancy in the body fixed frame.
-We neglect contributions from the strain and local electric field in the hamiltonian since we are working with magnetic fields on the order of 0.1T, which induce splittings  larger than the splitting of the zero-field ESR line ($\approx 20$ MHz).
+We neglect contributions from the strain and local electric field in the hamiltonian since we are working with magnetic fields on the order of 10 mT, which induce splittings larger than the splitting of the zero-field ESR line ($\approx 20$ MHz).
 We also neglect the hyperfine interaction with the nuclear spin of the $^{14}$N atom since we are working with ensembles with typical inhomogeneous broadening of 5~MHz.
@@
   \caption{a) Representation of the four crystalline axes of the diamond b) Representation of the magnetic field in the diamond crystalline basis c) Representation of the crystalline planes in the $(\theta , \phi)$ basis.
-The [110]$_\perp$ family of planes are shown using dashed lines, [100]$_\perp$ using plain lines.
+The $\{ 110 \}$ family of planes are shown using dashed lines and the $\{ 110 \}$ family using plain lines.
 The [100] direction is marked by red circles, the [111] direction by green squares.
@@
 The magnetic field direction is represented in Fig.
-\ref{cristallo} a), where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z'} ([$001$]) direction.
+\ref{cristallo} a), where the usual polar and azimuthal angles $\theta$ and $\phi$ are defined with respect to the \textbf{z'} ([$001$]) direction (we denote with ' the axes in the diamond frame).
 For some orientations of the magnetic field, the projection of the magnetic field on two or more NV axes will be identical, and therefore the energy level of the corresponding classes will be the same.
-These degeneracies are represented in Fig \ref{cristallo} c), where the dashed lines are the locii of the crystalline planes orthogonal to the [110] directions (6 planes in total).
-When the magnetic field belongs to these planes, we observe a degeneracy between two classes of NVs, as can be seen in the Fig.
+These degeneracies are represented in Fig \ref{cristallo} c), where the dashed lines are the locii of the $\{ 110 \}$ family of planes (plane normal to the [110] direction and all other equivalent directions, making 6 planes in total).
+When the magnetic field belongs to these planes, we observe a degeneracy between two classes of NVs, as can be seen in the Fig.\ref{scan_dege} or in Fig.
 4 of the main paper.
 
-The plain lines are the locii of the planes orthogonal to the [100] directions (3 planes in total).
-When the magnetic field lies in these planes, all classes are co-resonant, as can be seen in Fig.
-\ref{CR_22}.
+The plain lines are the locii of the $\{ 100 \}$ family of planes (3 planes in total).
+When the magnetic field lies in these planes, all classes are co-resonant, as can be seen in Fig.\ref{scan_dege} or in Fig.\ref{CR_22}.
 The red circles correspond to the [100] directions, for which the four classes of NV are degenerate.
 The green squares correspond to the [111] direction where one class is aligned with the magnetic field, and the three others are degenerate.
-Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 3 of the main text, where we can see that a plane from the [110]$_\perp$ family is being crossed.
+Finally the blue dots correspond to the path followed by the magnetic field in the experiment presented in Fig 3 of the main text, where we can see that a plane from the $\{ 110 \}$ family is being crossed.
+
+\section{Depolarization induced by NV-NV cross-relaxation}
+As explained in the main text, when the density of NV$^-$ centers in the sample is high enough (typically for concentrations higher than 1 ppm), the ensemble of NV spins will lose some of its polarization through dipolar coupling between the NV centers.
+This phenomenon is at the heart of the mechanism that allows us to exalt the magnetic suceptibility of our diamond through dipolar interaction, and it has already been observed independently by many groups in bulk diamond \cite{Jarmola} \cite{mrozek_longitudinal_2015} \cite{choi_depolarization_2017} \cite{akhmedzhanov_microwave-free_2017} \cite{giri_coupled_2018}
+
+In particular, \cite{choi_depolarization_2017} proposes a model based on "fluctuators" : a subgroup of NV centers with a very short lifetime (possibly due to their electron tunnelling in and out of the NV site) that can act as a source of classical noise with a central frequency given by the usual transition frequencies of the NV$^-$ spin Hamiltonian.
+A prediction of this model is that the modified lifetime of the ensemble of NV centers should have a stretch exponential profile (of the form $e^{-\sqrt{\frac{t}{T_1}}}$) which we do observe experimentally.
+
+\subsection{Stretch exponential profile of the lifetimes}
+\begin{figure}
+\includegraphics[scale=0.7]{T1-stretch_exp}
+\caption{Lifetime measurement for a case where all four classes are degenerate. Plain orange line correspond to a stretch-exponential fit and dashed red line to a simple exponential fit}
+	\label{stretch}
+\end{figure}
+In the theory developed in \cite{choi_depolarization_2017}, the stretch exponential profile arises from the inhomogeneity of the distance from each NV centers to the closest fluctuators.
+If $\rho_{00}^s(t)$, the population in the $\ket{m_s=0}$ state for each NV centers evolving in the dark, follow a law of the form $\rho_{00}^s(t) \propto \exp(-\gamma t)$ where $\gamma$ is the individual depolarization rate of the spin; then, assuming an even spatial distribution of fluctuators, the authors of \cite{choi_depolarization_2017} show that the distribution in $\gamma$ should follow a law of the form $$\rho(\gamma) \approx \frac{e^{-\frac{1}{4 \gamma T_1}}}{\sqrt{4 \pi \gamma^3 T_1}} $$ where $\rho(\gamma)$ is the density of probability of $\gamma$.
+
+Averaging then over all NV centers gives the stretch exponential profile observed from the ensemble :
+$$ \rho_{00}^e(t) \propto \int_0^{+\infty} \rho(\gamma) e^{-\gamma t} d\gamma = e^{-\sqrt{\frac{t}{T_1}}} $$
+where $ \rho_{00}^e(t)$ correspond to the average population in the $\ket{m_s=0}$ state for the ensemble of spins.
+
+Fig.
+\ref{stretch} shows a lifetime measurement on a static micro diamond following the protocol described in Sec.III.
+Here all four classes are resonant with the applied microwave frequency, which corresponds to the maximum degree of degeneracy between the NV centers, and therefore the stronger modification of the lifetime induced by the resonant dipolar coupling.
+The signal we obtain was fitted using a stretch exponential profile  and a simple exponential profile.
+We can see that the stretch exponential profile ($R^2=0.981$) is in better agreement with the data than the exponential fit ($R^2=0.942$), in particular for the very short times (we expect the longer times to be dominated by the phonon-limited exponential lifetime).
+
+Finally it should be noted that the stretch exponential profile arising from point-like depolarization sources is a relatively general result that has for example also been observed for the depolarization of NV centers induced by substitutional nitrogen (P1) defects in diamond \cite{hall_detection_2016}
+
+\subsection{Scanning the degeneracy conditions}
+\begin{figure}
+\includegraphics[scale=0.3]{scan_CR}
+\caption{Experimental results for a magnetic field scan with an electromagnet while a static magnetic field offset is applied with a permanent magnet (setup similar to the one presented in Fig.\ref{Optics}) \textbf{a)} Frequency Transitions for the $\ket{m_s=0} \to \ket{m_s=-1}$ transition of the four classes of NV centers as a function of the scanning magnetic field (data extrapolated from ODMR spectra) \textbf{b)} Change in photoluminescence of the NV$^-$ centers as a function of the scanning magnetic field.
+\textbf{c)} Example of 3 ODMR spectra for the 3 particular field values corresponding to the circled areas in subfigure a).}
+	\label{scan_dege}
+\end{figure}
+The easiest way to probe the mechanism of dipolar-induced modification of the lifetime is to change the degeneracy conditions between the four classes of NV centers by tuning the magnetic field, as explained in Sec.I.
+Because the NV spins can only exchange spin quanta when they are quasi-resonant, then by tuning the number of classes at degeneracy, you can modify the effective density of interacting NV centers, and therefore tune the depolarization effect.
+
+An example of this is given in Fig.2 of the main text with the varying lifetime depending on the degeneracy condition, but another way to probe this effect is shown in Fig.\ref{scan_dege} : in this figure, we have observed the change in photoluminescence from a static microdiamond while changing the magnetic field in order to explore different degeneracy conditions.
+In order to do this, we need two sources of magnetic field : an electromagnet to scan the field and a permanent magnet to apply a magnetic field offset in an other direction (otherwise the magnetic field orientation with respect to the diamond axes would remain the same as the field is scanned).
+
+In this particular case, we can see that as the magnetic field is scanned, it crosses three "degeneracy planes" (as described in Sec.I) : first a plane of the the $\{ 110 \}$ family at B=13 G, with a single degeneracy condition, then a plane of the $\{ 100 \}$ family at B=44 G where there is a simultaneous degeneracy condition for two pairs of classes, and then another plane of the the $\{ 110 \}$ family at B=76 G.
+We notice that each time a degeneracy between at least two classes of NV happens, we observe in Fig.\ref{scan_dege}b) a sharp decrease in photoluminescence.
+This is a signature of the change in lifetime of the ensemble of spins : indeed, the photoluminescence of NV ensembles is proportional to the average population in the $\ket{m_s=0}$ state, and the $\ket{m_s=0}$ population of the spins is the result of the competition between the polarization rate due to the green laser and the various depolarization mechanism.
+Increasing the depolarization rate of the spins will therefore decrease the overall luminosity.
 
@@
 
-Another way to estimate the torque is depicted in Fig \ref{Torque1classe}.
-Under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, under the low magnetic fields ($<$ 200 G) we are working at, will be oriented at an angle of 90$^\circ$ from the NV axis : $\langle \hat S_{z'} \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$.
+Another way to estimate the torque is to numerically solve the master equation of the system as depicted in Fig \ref{Torque1classe}.
+We find that under green excitation and in the presence of an external magnetic field, the spins will acquire a magnetization $\gamma_e \langle\hat{\mathbf S}\rangle$ which, under the low magnetic fields ($<$ 200 G) we are working at, will be oriented at an angle of 90$^\circ$ from the NV axis : $\langle \hat S_{z'} \rangle \approx 0$ and $\langle \hat S_\perp \rangle \neq 0$.
 This magnetization vanishes when the magnetic field is aligned with the NV center since there is no longer a transverse field responsible for the mixing of the eigenstates.
@@
 
-The amplitude of the torque with respect to the magnetic field orientation (at a B field amplitude of 100 G) is represented in 1D in Fig \ref{Torque1classe}-c) where we can see a behavior very close to $|\sin(2\theta)|$, which is the formula (3) of the main text that we obtained through a perturbative approach.
+The amplitude of the torque with respect to the magnetic field orientation at a B field amplitude of 100 G is represented in 1D in Fig \ref{Torque1classe}-c) where we can see a behavior very close to $|\sin(2\theta)|$, as found in Eq.\ref{torque_anal} through a perturbative approach.
 The same torque amplitude is represented in 2D in Fig \ref{Torque1classe} d).
 The two purple dots in the map correspond to the [111] direction when the magnetic field is aligned with the centers.
-The curvy line corresponds to the $[111]_\perp$ plane.
+The curvy line corresponds to the (111) plane.
 Importantly, the maximum torque value is $1\cdot 10^{-26}$ N.m for a single spin.
@@
 \item The change in magnetic torque is resonant, it happens only when different classes are brought to resonance.
-This can be seen by comparing Fig 3 b) of the main text to the [110]$_\perp$ planes that were drawn in Fig.
+This can be seen by comparing Fig 3 b) of the main text to the $\{ 110 \}$ planes that were drawn in Fig.
 \ref{cristallo}.
@@
 
-It should be noted that the main uncertainty comes here from the diamond size can change the expected result by an order of magnitude.
+It should be noted that the main uncertainty comes here from the diamond size which can change the expected result by an order of magnitude.
 

### 2021-02-23 Mes papiers/CR-CVD/Reviewer_answer.tex
tutu

\documentclass{article}
\usepackage{physics}
\usepackage[]{quoting}

\begin{document}

\section*{Summary of the changes}

We would first like to thank the referees for the time they spent reviewing of our manuscripts.

Both referees found our work interesting for the field and our experimental results to be solid. They however pointed to several points lacking from the original manuscripts, in particular in regard to specifications of the sample and the defects we study, and the possibility to quantitatively measure the concentration of the new defects that we found.

We agree with these remarks and have amended the manuscript in consequence to the best of our abilities.

\section*{Referee 1}
\subsection*{Referee 1 report}

\begin{quote}
In this paper, Pellet-Mary et al. carry out a systematic investigation of paramagnetic defects other than the nitrogen vacancy center in cvd grown diamond. Utilizing NV center fluorescence in the same sample and sweeping an external magnetic field in the [100] direction, paramagnetic defects are identified as fluorescence dips at certain fields when the Zeeman split energy levels overlap. The identification and understanding of paramagnetic defects is important as they might serve as sources of decoherence in the diamond sample, and thus the present article yields important information. I have the following questions and comments before I can recommend publication.

1) On page 1, the authors briefly refer to Fig. 1 illustrating the different types of defects detected here. I believe the readers would appreciate if the defects with their respective properties could be introduced briefly already here.

2) The authors describe the sample fabrication on page 2 and it is mentioned earlier that they address a high density NV spin ensemble. However, I believe many readers would be interested in the nitrogen and nitrogen vacancy concentration obtained in the sample. The N (or P1 center) concentration is mentioned later on page 3, but this should presented coherently on the sample description?

3) On page 2, the paragraph starting with "Most experimental studies have been carried out...." It is not exactly clear which kind or type of experiments the authors refer to.

4) On page 2, the authors write "...T*2 was not degraded after annealing, yet the NV density was large enough to enable concentration dependent longitudinal relaxation [15, 17]." Thus, is the sample here the same as the one in Refs. [16, 17]?

5) On page 3, the text explaining the model illustrated in Fig. 4: in my opinion, the authors could add some information on the model. What is the symmetry of the defects and their orientation?

6) The spectra in Fig. 3 are fitted with Gaussian functions, though resonances should have a Lorentzian lineshape. Is there a particular reason for the choice?

7) The detection of the three electron spin-1 species in their sample is convincing. Using their method, would it be possible also to estimate the species concentration and eventually the spin dynamics? I discussion of this would be appreciated.

8) A more thorough discussion about the applied growth conditions and the occurrence of the observed defects would be useful? Does the concentration of the mentioned complexes (NVH, VH, VH2) strongly change as a function of e.g. temperature?

Minor comments:

- On page 2, the reference to Fig. 1-b) has a typo.

- The magnetic field axis in Figs. 3 and 4 have no units.

- The legend in Fig. 4 should read ‘13C-NV’ instead of 13C as explained in the text.
\end{quote}

\subsection*{Our answer}
Here is our answer to the points raised by Referee 1 :
\begin{enumerate}
\item We have added a short description of the three relevant defects of this study when describing Fig.1
\item We have added the concentration of NV and N in the description of the sample.
\item We have reformulated the sentence to make clear that we mention other cross-relaxation studies based on NV centers.
\item We have rewritten the paragraph to make it clear that this is indeed the same sample.
\item We have added information on the symmetry of the different defects with respect to the [100] axis to justify that we only take into account two transitions per defect in our model (except for the 13C-NV defect which is described in the supplementary material)
\item We added a discussion on the Gaussian vs Lorentzian line-shape in the description of fig.3. Given that the approximate shorter distance between our spins is more than 10 nm, the coupling strength between our spins would be less than 100 KHz. The width of the lines in fig.3 is therefore essentially due to the inhomogeneous Gaussian broadening of the ensemble of spins.
\item We added a paragraph at the end of the manuscript where we give a tentative estimate of the VH- and WAR1 concentration based on our measurement.
\item La c'est plus tendu de répondre. Alex ?
\end{enumerate}

We corrected the typos and missing information pointed in the minor comments and thank Referee 1 for their observations.

\section*{Referee 2}
\subsection*{Referee 2 report}
\begin{quote}
Pellet-Mary et al. report an experimental study on detecting paramagnetic defects in CVD-grown diamond by performing optical scans. From the optical detection of NV’s photoluminescence (PL) in varying external magnetic field, the authors identify two spin-1 defects, namely VH- and WAR1 in diamond based on cross-relaxations between NV centers and these defects. The authors further attribute a broad dip signal in PL to the interaction between bare NV and strongly coupled NV-C13 pairs.

In general, I find this work interesting and the experiment results are solid. The manuscript is also clearly written. However, I think there are several issues that should be addressed or clarified before further assessment. Specifically, to make the arguments more solid, the following points should be addressed.
Issues and questions:

1. In plotting the subtracted normalized PL curve, the authors used a 4th-order polynomial to fit the PL curve. Can the authors explain how they choose the order of the fitting function? Specifically, a higher order of polynomial might hide the spectral bumps that we are interested in and a lower order might result to a worse signal-to-noise ratio.

2. Similar question as above: would it be possible to theoretically calculate the PL drop vs B field using the transverse component of the magnetic field that results to the NV state mixing?

3. The zero-field-splittings (ZFS) of two defects were extracted from the position of two spectral bumps (one at 56 G another at 122 G). Could other information related with these defects be extracted from the salient features in the PL curve? For example, the depth (signal contrast) of the two dips could be related with defect concentrations and the width might have a connection with inhomogeneities in interactions between NVs and the defects.

4. The authors claimed that the signal of VH- and WAR1 defects was not observed in several HPHT samples. Would it be attributed to the difference (if any) between CVD-grown and HPHT samples in the irradiation and annealing process? For clarification, it might better to list the details of NV creation process for the HPHT samples.

5. Could the authors give some brief explanation that why the VH- and WAR1 defects only observed in CVD-grown samples (from the view of chemistry/material science)?

6. What’s the concentration of 13C in the CVD sample used in the study? Comparing it with the NV concentration (about 4.6ppm), what will be the probability that a NV center would have a strongly coupled first-shell C13? Would the resonant condition be the same for 13C-NV pairs beyond first shell interactions? These questions should be better clarified in the manuscript.

7. Could the authors briefly mention whether it’s possible that the technique could be extended to detect defects that have ZFS that are very different from NV center’s?

8. What is the quantization axis of the observed defects with respect to the [100] axis? Would the picture in Fig.4 still hold if one considers possible state mixings?

Some minor issues:

1. The legend label of 13C in Fig.4 (also Fig.1 in SI) is misleading. I would suggest label it with ‘NV- 13C’ instead of ‘13C’.

2. As mentioned before, I would suggest adding more information on the CVD sample such as 13C abundancy as well as estimation of defect concentrations, if possible.
\end{quote}
\subsection*{Our answer}
Here is our answer to the points raised by Referee 1 :
\begin{enumerate}
\item We have added a discussion of the 4th order polynomial in the description of fig.3, Referee 2 was mostly right on their assumption regarding the order of the polynomial. (although we have not really seen hiding of the spectral bumps even for higher order polynomial)
\item In the same section we explain that we tried to find a numerical solution for the overall decrease of PL with the magnetic field, but the solutions we found were not in agreement with our experimental data. (the decrease in our data seem too linear compared to the predicted quadratic profile of the numerical simulations)
\item We added a discussion regarding this point (similar to question 7 of Referee 1) at the end of the manuscript. It is in fact possible to extract an estimate of the concentrations with the area of the peaks (which is probably more relevant than the amplitude of the peaks in this case), but the width of the peak is mostly due to inhomogeneous broadening and not to interactions.
\item On a pas les infos pour les adamas, je sais pas si on peut en avoir pour les sumitumo ?
\item Demander à Alex. Je pense qu'il faut juste justifier le fait qu'il y ait plus d'hydrogène dans les CVD que dans les HPHT.
\item We have added a discussion of the concentration of 13C-NV pairs in the last paragraph when quantifying the concentration of our species. As for the beyond fist-shell 13C we have clarified in the 13C section that any 13C further away than the first-shell would have an hyper-fine bonding of less than 15 MHz, making its contribution almost indistinguishable from the inhomogeneous broadening due to the ambiant spin bath.
\item We added on page 3 a clarification that scans in the [100] direction cannot be used to detect spins with ZFS too different from the NV centers, due to the transverse field depolarization of NV centers for high field values.
\item The quantization axes for the $S_z$ operator of NV, VH and WAR1 in fig.4 were taken to be one of the four $\langle 111 \rangle$ direction, which are all equivalent when the magnetic field is in the [100] direction. Although this is not clearly visible, fig.4 does take into account state mixing due to the transverse field, by diagonalizing the Hamiltonian with both the longitudinal and transverse contribution of the magnetic field. 
\end{enumerate}

As for the minor comments, we changed the label and added more information on the CVD sample when describing it, and a section on the different defects concentration. We thank Referee 2 for their observations.

\end{document}

### 2021-02-23 Mes papiers/CR-CVD/CVD_para.tex
tutu

@@
-\documentclass[9pt,twocolumn,twoside]{osajnl}
-%% Please use 11pt if submitting to AOP
-% \documentclass[11pt,twocolumn,twoside]{osajnl}
-\usepackage{graphicx}% Include figure files
-\usepackage{dcolumn}% Align table columns on decimal point
-\usepackage{bm}% bold math
+\documentclass[9pt,twocolumn,twoside]{revtex4-1}
+
+\usepackage{graphicx}
+\usepackage{dcolumn}
+\usepackage{bm}
 \usepackage{natbib}
@@
 
-\journal{optica} % Choose journal (ao, aop, josaa, josab, ol, optica, pr)
-
-% See template introduction for guidance on setting shortarticle option
-\setboolean{shortarticle}{true}
-% true = letter / tutorial
-% false = research / review article
 \newcommand{\bea}{\begin{eqnarray}}
@@
 
-% (depending on journal).
+\begin{document}
+\title{Optical Detection of Paramagnetic Defects in a CVD-grown Diamond}
 
-\title{Optical detection of paramagnetic defects in a CVD-grown diamond using Nitrogen-Vacancy centers}
+\author{C. Pellet-Mary$^{1}$}
+\author{P. Huillery $^{1}$}
+\author{M. Perdriat $^{1}$}
+\author{A. Tallaire $^{2}$}
+\author{G. H\'etet$^{1}$}
 
-\author[1]{C. Pellet-Mary}
-\author[1]{P. Huillery}
-\author[1]{M. Perdriat}
-\author[2]{A. Tallaire}
-\author[1]{G. H\'etet}
-
-\affil[1]{Laboratoire de Physique de l'Ecole normale sup\'erieure, ENS, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e Paris-Diderot, Sorbonne Paris Cit\'e, Paris, France.}
-
-\affil[*]{Corresponding author: gabriel.hetet@ens.fr}
-
-%% To be edited by editor
-% \dates{Compiled \today}
-
-%\ociscodes{(140.3490) Lasers, distributed feedback; (060.2420) Fibers, polarization-maintaining;(060.3735) Fiber Bragg gratings.}
-
-%% To be edited by editor
-% \doi{\url{http://dx.doi.org/10.1364/XX.XX.XXXXXX}}
+\affiliation{$^1$Laboratoire de Physique de l'Ecole normale sup\'erieure, ENS, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e Paris-Diderot, Sorbonne Paris Cit\'e, Paris, France.}
+\affiliation{$^2$IRCP, Ecole Nationale Supérieure de Chimie de Paris, 11, rue Pierre et Marie Curie, 75005 Paris, France}
 
 \begin{abstract}
-The spins of Nitrogen Vacancy centers in Chemical Vapor Deposition (CVD) grown diamonds form ideal probes of magnetic fields and temperature, as well as ideal qu-bits for quantum information processing.
-Studying and controlling the magnetic environment of NV centers in such high purity crystals is thus essential.
-In this letter, we demonstrate optical detection of paramagnetic species in a CVD grown diamond, such as hydrogen-related complexes, by coupling them to NV centers.
-%We also unravel the role played by double quantum processes at zero magnetic field in the NV relaxation. 
-The resonant transfer of NV centers spin-polarization to the electronic spins of these defects induces a loss of the NV photoluminescence, generating conspicuous features in the NV photoluminesence. 
-The resonant condition is tuned with high accuracy by scanning a magnetic field along the 100 crystal direction.
-Our results offer prospects for more detailed studies of CVD grown processes as well as for coherent control of the spin of a novel class of paramagnetic species.
+The electronic spins of the nitrogen-vacancy centers (NV centers) in Chemical-Vapor-Deposition (CVD) grown diamonds form ideal probes of magnetic fields and temperature, as well as promising qu-bits for quantum information processing.
+Studying and controlling the magnetic environment of NV centers in such high purity crystals is thus essential for these applications.
+We demonstrate optical detection of paramagnetic species, such as hydrogen-related complexes, in a CVD-grown diamond.
+The resonant transfer of the NV centers' polarized electronic spins to the electronic spins of these species generates conspicuous features in the NV photoluminescence by employing magnetic field scans along the [100] crystal direction.
+Our results offer prospects for more detailed studies of CVD-grown processes as well as for coherent control of the spin of novel classes of hyper-polarized paramagnetic species.
 \end{abstract}
-
-\setboolean{displaycopyright}{true}
-
-\begin{document}
 
@@
 
-\section{Introduction}
-The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science.
+The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing \cite{Rondin_2014} and quantum information science  \cite{DOHERTY20131}.
 One major reason is that it can be optically polarized and read-out and features long coherence time even at ambient conditions. 
 %The NV property can also then be employed to manipulate and to detect other dark spins that are interacting with it through dipole-dipole coupling, either inside or outside the diamond. 
-In order to optimize the capabilities of the NV center, its magnetic environment must be controlled carefully. 
-Synthetic growth of diamond crystal now reaches a level of control that makes pristine diamond crystals almost flawless, eliminating the source of magnetic noise from other nearby impurities.
-The standard growth technique works through Chemical Vapor Deposition (CVD) of carbon atoms from a methane gas in an ultra clean vacuum environment.
-It offers the possibility to work with 99.9\% purity as well as using 12C enriched methane, hence removing spin fluctuations fro 13C atoms.
-More results from CVD.
+In order to optimize the capabilities of the NV center, its magnetic environment must be very well controlled. 
+Synthetic growth of diamond crystals is now reaching a level of maturity that makes the most pristine diamond crystals almost flawless, thus eliminating the source of magnetic noise from other nearby impurities than the  NV sensor.
+The standard growth technique works through Chemical Vapor Deposition (CVD) of carbon atoms from a methane gas in an ultra-high vacuum environment.
+It offers the possibility to use crystals with 99.9\% purity, as well as using $^{12}$C enriched methane, hence removing spin fluctuations from $^{13}$C atoms and enhancing further the sensing capabilities of NV centers \cite{Achard}.
 
-The control of the purity of the material is typically done using Electron-Paramagnetic-Resonance (EPR). 
-EPR operates with a very high resolution and was instrumental for detecting many species in diamond (uni of Warwick's review).
-It would also be ideal if the defects were detected in the same platform that is used to for nanoscale sensing and quantum information processing, namely confocal microscopy. 
-This would offer the opportunity to polarize these defects using the NV center, dynamically decouple them from the NV evolution for enhanced AC magnetometry and even to use them as ancilla qubits.
+Electron-Paramagnetic-Resonance (EPR) spectroscopy is the method of choice for controlling the concentration of paramagnetic defects in materials.Thanks to their high resolution, EPR spectrometers are also essential tools for understanding the conformation of many defects that remain in
+CVD-grown diamond materials \cite{newton_epr_2007, glover_hydrogen_2004, glover_hydrogen_2003}. However, compared to confocal microscopy, this equipment is rather bulky and less cost-effective. 
+It would also be ideal if these defects were detected with the very same technology that is employed for nanoscale sensing \cite{Rondin_2014} and quantum information processing \cite{Yao}. 
+It would indeed offer the opportunity to hyper-polarize these defects using the optically polarized NV center and to employ them as quantum bits at ambient conditions.
 
@@
 \centering
-\fbox{\includegraphics[width=\linewidth]{Fig1.eps}}
-\caption{Schematics showing various paramagnetic defects in CVD grown diamonds polarized by a Nitrogen Vacancy center. Bottom right :  crystalline structure of the diamond.}
+{\includegraphics[width=\linewidth]{Fig1}}
+\caption{Schematics showing various paramagnetic defects in CVD-grown diamonds interacting with the electronic spin of a negatively charged nitrogen-vacancy center.
+\textbf{Bottom right}: Atomic positions in the diamond unit cell.}
 \label{fig1}
@@
 
-Here, we use confocal microscopy and a high-density NV ensembles in a CVD grown diamond to detect paramagnetic defects from CVD grown sample.
-The detection was realized using cross-correlation studies through magnetic field scans. 
-Cross-relaxation typically takes place when the electronic or nuclear spins of two atomic species exchange their polarization via resonant magnetic dipole-dipole interactions. 
-If the spin of an NV center is polarized and coupled to an unpolarized spin B, it will lose part of its polarization at the expense of B and thus see a drop of its PL rate. 
-Tuning the frequency of both spin transitions will thus result in a drop of the NV photoluminescence when that they are co-resonant, thus enabling detection of the spin state of B.
-
-The hamiltonian that describes typical spin-1 defects in diamond under a magnetic field be written as 
-\bea
-\hat{H}=\hbar D S_z^2+\hbar \gamma_e\bm S \cdot \bm B
-\eea
-where $S$ is...
-The first term describes dipole interactions amongst the two electrons of the spin-1 system. The second term describes the Zeeman interaction.
-$D$ can vary significantly from one defect to another as it is sensitive to the local crystal field. It can then be used to find the structure of the defect in EPR measurements.
-Also, hyperfine interactions can add zero-field splittings that can be observed under high resolution scans. 
-The angle of the magnetic field can thus also be to bring defects into resonance.
-
-Cross relaxation amongst closely-packed NV centers can take place at moderate magnetic fields \cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}, thus far, other impurities generally require by scanning large magnetic fields (more than hundreds of gauss) along the 111 direction of diamond.
-This is generally required to probe paramagnetic impurities that have a small ZFS. 
-Most of the measurements are done at the ESLAC/GLSAC. OORT, BUDKER, HOLLENBERG, BAJAJ, MANSON.
-Similar cross relaxation phenomena were studied by measuring changes in the optical hole depth in the zero-phonon line and studying decay rates oin stimulated spin-echo and spin-locking signals ?
-(BUDKER).
-
-This procedure was proven useful for detecting the spin of $^{13}C$ atoms and the substitutional nitrogen Vacancy center in diamonds in the past.
-Such measurements not only show detection of this dark spins but can also be used to polarize them (Polarisation of 13C at low fields : PINES, MERILES ?).
-To date however, most studies have however been carried out using crystals grown by the High-Pressure-High-Temperature (HPHT) process, where a poor control over the paramagnetic defect density is available.
-In CVD grown samples, the typical paramagnetic impurities that have been detected by EPR are hydrogen complex such as the NVH, VH, VH2, where the hydrogen comes from the methane in the reactor.
+Here, we use confocal laser microscopy to detect paramagnetic defects {\it via} the coupling to a high-density NV spin ensemble in a CVD-grown diamond.
 Fig.
-\ref{fig1} shows a schematics of the various paramagnetic defects in a CVD grown diamonds that can be coupled to the Nitrogen Vacancy center in our study.
+\ref{fig1} shows a schematics of the various paramagnetic defects that have been coupled to the nitrogen-vacancy center and detected optically in our study.
+These defects are the VH$^-$ and WAR1 defects, two spin-1 defects with comparable zero field splitting to the NV$^-$ center \citep{glover_hydrogen_2004, cruddace2007magnetic}, as well as $^{13}$C $1/2$-nuclear spin, through hyper-fine coupling with a secondary NV$^-$ center.
+The detection is realized by measuring the NV photoluminescence while performing magnetic field scans to resonantly enhance dipole-dipole interactions and observe cross-relaxations (CR).
+Cross-relaxation typically takes place when the electronic or nuclear spins of two atomic species exchange their polarizations via resonant magnetic dipole-dipole interactions. 
+If the spin of an NV center A is polarized and coupled to an unpolarized spin B with a much larger relaxation rate, it will lose part of its polarization at the expense of B and thus see a drop of its photoluminescence (PL) rate.
+Tuning the frequency of both spin transitions so that are co-resonant will result in a reduction of the PL of NV A, thus enabling detection of the spin energy of B.
+%It can then be used to unambiguously identify the defect. The response to a change in the angle of a magnetic field can further be used to unravel the structure of the defect in EPR measurements.
+%Hyperfine interactions can add zero-field splittings that can be observed under high resolution scans. 
+%The angle of the magnetic field can thus also be to bring defects into resonance.
 
@@
 \centering
-\fbox{\includegraphics[width=\linewidth]{Carte_annotation}}
-\caption{NV$^-$ photoluminescence as a function of a scanning magnetic field around the [100] crystalline directions with a fixed amplitude $|B|=115G$.
-The planes orthogonal to the [010],[001],[011], and [01$\bar 1$] directions are indicated by dashed lines.}
+{\includegraphics[width=\linewidth]{Carte_annotation}}
+\caption{a) NV$^-$ photoluminescence as a function of the magnetic field angle around the [100] crystalline direction, at a fixed amplitude $|\vec{B}|=115$G.
+The planes orthogonal to the [010], [001], [011], and [01$\bar 1$] directions are indicated by dashed lines.
+b) Left: Optically-detected-magnetic-resonance (ODMR) spectrum taken at the angular position $(\phi, \theta)=(3^\circ,3^\circ)$.
+Right: ODMR spectrum taken at the exact center of the map, {\it i.e.} the [100] direction.}
 \label{map}
@@
 
-Recent work in the doping processes has shown the possibility to reach NV centers concentrations in the 3 to 5 ppm range in CVD grown samples \cite{Edmonds, TALLAIRE2020421, MINDARAVA2020182}, without sacrificing the $T_2^*$ coherence time, opening a path towards detecting these hydrogen-related complexes using cross-relaxation with NV centers.
-%Several impurities in CVD grown diamonds have been detected using Electron-Paramagnetic Resonance.
+The negatively charged nitrogen-vacancy center has a zero-field splitting $D=(2\pi) 2.87$ GHz in the ground state, related to the dipole-dipole interaction between the spin of the two unpaired electrons.
+One of the most important properties of the spin of the NV$^-$ center is that it can be optically polarized and read-out and features long relaxation time ($\approx $ms) even at ambient conditions, which can be used to detect other species via CR and even to polarize them.
+Many paramagnetic defects in diamond also carry a spin 1.
+The zero-field splittings are in fact a fingerprint of the defect.
+To measure it, one can tune the angle and magnitude of an external magnetic field angle to cause cross-relaxation with the NV centers.
+The obtained CR position can then be used to trace back the zero-field splitting of the defect.
+%Four possible directions of the NV axis in the crystalline diamond structure can then lead to CR amongst NV centers at moderate magnetic fields \cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}.
+%Detecting paramagnetic impurities using the NV center often requires large magnetic fields (> 500 G) along the diamond [111] direction. 
+%This is generally required to probe impurities with an electronic spin 1/2 such as substitutional nitrogen, or the $^{13}$C nuclear spin.
+%\cite{Epstein,Wang, armstrong_nvnv_2010, Hall, Wickenbrock}.
+%Most of the measurements are done at the ESLAC/GLSAC. OORT, BUDKER, HOLLENBERG, BAJAJ, MANSON.
+%Similar cross relaxation phenomena were studied by measuring changes in the optical hole depth in the zero-phonon line and studying decay rates oin stimulated spin-echo and spin-locking signals ?
+(BUDKER).
+%This procedure was proven useful for detecting and understanding the spin of $^{13}C$ atoms and the substitutional nitrogen Vacancy center in diamonds in the past. 
+%Such measurements not only show detection of dark spins but can also be used to polarize them (Polarisation of 13C at low fields : PINES, MERILES ?).
+
+%In CVD grown samples, the typical paramagnetic impurities that have been detected by EPR are hydrogen complex such as the NVH, VH, VH2, where the hydrogen comes from the methane in the reactor.
+Fig.
+\ref{fig1} shows a schematics of the various paramagnetic defects in a CVD grown diamonds that are coupled to the Nitrogen Vacancy center in our study.
+Previous cross-relaxation studies with NV$^-$ centers have been mostly carried out using heavily doped crystals grown by the High-Pressure-High-Temperature (HPHT) process \citep{Epstein,Wang, armstrong_nvnv_2010, Hall, Wickenbrock, Alfasi}.
+Recent efforts in the doping processes have also made it possible to reach NV centers concentrations in the 3 to 5 ppm range in CVD grown samples \cite{Edmonds, TALLAIRE2020421, MINDARAVA2020182}, opening a path towards detecting other paramagnetic defects in CVD grown diamonds.
+%these hydrogen-related complexes using cross-relaxation with NV centers.%Several impurities in CVD grown diamonds have been detected using Electron-Paramagnetic Resonance.
 In order to purify diamond crystal and optimize their conduction and properties for prolonged qubit lifetimes.
 %Extensive work has been done in the past twenty years to detect and characterize the various  that appear in EPR. 
-The sample we use in this study was grown using CVD with the addition of N$_2$0 in the methane phase.
-Then, high energy electron irradiation and in-situ annealing was realized in order to convert the nitrogen into NV centers.
-A concentration of 5 ppm of NV center was obtained.
-It was shown that in this sample the $T_2^*$ was not degraded after annealing, and that the NV density was large enough to enable the observation of $T_1$ dependent concentration.
-It is well-known that when the NV concentration increases above 1 ppm, the longitudinal decay of NV centers is degraded.
-Furthermore, it has been observed that cross-relaxation amongst closely pack NVs of different classes can be observed.
-It was postulated that such a degradation of the $T_1$ time came from the coupling of NV centers to other NV related defects whose spin fluctuates.
+The sample we use in this study  correspond to the pink irradiated sample in \cite{TALLAIRE2020421}. It was grown using CVD with the addition of 500ppm of N$_2$0 to the H2/CH4 (96/4) gas phase.
+Then high energy (10MeV) electron irradiation at a fluence of $2\times18$ cm$^{-2}$ and at a temperature of 900 degrees celsius was realised, giving a final concentration of NV$^-$ of about 4.6~ppm for an initial concentration of nitrogen in the sample of about 25~ppm.
+It was shown that in this sample the $T_2^*$ was not degraded after annealing, yet the NV density was large enough to enable concentration dependent longitudinal relaxation \cite{TALLAIRE2020421, jarmola_longitudinal_2015}.
+%It is well-known that when the NV concentration increases above 1 ppm, the longitudinal decay of NV centers is degraded. 
+%Furthermore, it has been observed that cross-relaxation amongst closely pack NVs of different classes can be observed.
+%It was postulated that such a degradation of the $T_1$ time came from the coupling of NV centers to other NV-related defects.
 
-To perform the cross-relaxation detection, we use a homebuilt confocal microscope that comprises a 1mW green laser, an objective with a numerical aperture of XX to focus the laser onto the sample as well as to collect the PL.
-The PL was filtered from the green laser using a dichroic mirror and a notch filter coupled to a multimode fiber and sent to avalanche photodiode.
-The magnetic field scans were realized using a C-shaped electromagnet driven by a XX generator.
-We monitor the NV PL as the magnetic field is scanned.
-
-Contrary the more commonly chosen $\langle 111 \rangle$ direction, we use magnetic field scans along the $\langle 100 \rangle$ direction. 
-Looking at the diamond structure (bottom right in Fig.
-\ref{fig1})-a), we see that in this direction, the four possible projections of the NV centers along the magnetic field are the same.
-This means that doing a B field scan along this direction will not make NV centers energies from different classes cross.
-At arbitrary angles, this can result is several lines that could mask cross-relaxation from other species.
-The down side of this choice of B field direction is that for large B field amplitudes, the transverse component of the B field makes the PL drop [], thus limiting the range of magnetic field amplitudes that can be employed.
-In order to find the $\langle 100 \rangle$ direction we perform a preliminary angular scan of the magnetic field using a goniometer at a fix magnetic field amplitude.
-Fig.
-\ref{map}-a)  shows the NV Photoluminescence as a function of magnetic field direction, referenced by azimuthal and polar angles $(\phi, \theta)$ in the above described CVD grown sample.
-The PL is seen to drop for particular values of the angle as anticipated.
-%The symmetric pattern suggests that the diamond crystalline main directions play a role.
-Because of the anisotropy of the NV center, see \ref{Cristallo}, the NV PL drop can indeed be a signature of the crystal direction.
-Such a drop of the PL along the crystal axis was already observed by many groups. LUKIN, HEMMER, BUDKER. JARMOLA. OORT
-\cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, armstrong_nvnv_2010, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017} and correspond to situation where NV centers cross.
-Using such a map enables to identify clearly the crystalline axis and, in particular, the central point on this map, the $\langle 100 \rangle$ direction.
-
-The planes orthogonal to the [010],[001],[011], and [01$\bar 1$] directions are indicated by dashed lines and show the locus of the cross relaxation, with a smaller PL contrast. 
-The width of the CR peak is compatible with the $T_2^*$ time at this B field values. The contrast is determined both by the fluctuating NV concentration and the polarized NVs.
-Such a map can in itself be a useful tool in order to measure magnetic fields without microwave, as already pointed out in \cite{}.
-
-To double check these interpretations, we run an electron spin resonances by fixing the magnetic field amplitude and by scanning the microwave frequency. 
-Fig. \ref{map}-b) shows an ESR taken at the angle ... showing two degeneracies.
-Fig. \ref{map}-b) shows an ESR taken at the angle .. .showing four degeneracies. This hints towards the 100 direction.
-
-We now tune the magnetic field along the 100 direction and scan the magnetic field amplitude. 
-Fig. \ref{scan} shows the PL as a function of magnetic field amplitude, along the $\langle 100 \rangle$ direction in the 15 G to 145 G range. 
-Careful calibration of the magnetic field amplitude was realized using a microwave tone of known frequency ....
-The angle was controlled to within XX. 
-For these data, the averaging was done for XX days. 
-Three features clearly appear in this spectrum : one broader feature at 20G, another one at 56 G and one last one at 122 G. 
-We also note an overall drop of the PL, as a the result of state mixing in the optically excited state. 
-We made attempts to fit this curve using the full 7 level model, but this only gave poor agreement with experimental data. Further work will be done in order to clarify this. 
-In order to let the three salient features detach better from the spectrum, we thus resorted to a 4th order polynomial fit without the spectral bumps, and subtract it from the curve.
-The subtracted curve is displayed below.
-The width and contrast of each peak can now be compared and the position of each peak can be determined with great accuracy.
-
-\begin{figure}[htbp]
+In this study, we use a homebuilt confocal microscope that comprises a 1mW green laser, an objective with a numerical aperture of 0.25 to focus the laser onto the sample as well as to collect the NV photoluminescence (PL).
+The PL was filtered from the green laser using a dichroic mirror and a notch filter.
+It was then coupled to a multimode fiber and detected by an avalanche photodiode.
+The magnetic field scans were realized using a C-shaped electromagnet driven by a current generator (HP 33120A).
+We then monitor the NV PL synchronously with the changes in the magnetic field.
+\begin{figure}[ht]
 \centering
-\fbox{\includegraphics[width=\linewidth]{Scans_fig4}}
+{\includegraphics[width=\linewidth]{Scans_fig4}}
 \caption{Optical detection of cross-relaxations.
-\textbf{Top} NV$^-$ photoluminescence while scanning the magnetic field along the [100] crystalline direction (plain line) and polynomial fit to the fourth order (dashed line).
-\textbf{Bottom} Subtraction of the previous signal by the polynomial fit (circles), simulated CR magnetic field amplitude (dashed,vertical) and gaussian fits for the second and third dips (plain line)}
+{Top}: NV$^-$ photoluminescence counts as a function of the magnetic field amplitude along the [100] crystalline direction (plain line) and fourth-order polynomial fit (dashed line).
+{Bottom} Curve obtained by subtracting the polynomial fit to the above signal (circles).
+The simulated cross-relaxations positions are shown by the green dashed, vertical lines.
+Gaussian fits to the second and third dips are shown by plain orange lines.}
 \label{scan}
 \end{figure}
+Contrary to the more commonly employed $[ 111 ]$ direction, we scan the magnetic field along the diamond $[100]$ crystalline direction. 
+Looking at the diamond structure (bottom right in Fig.
+\ref{fig1}), it can be noticed that, in this direction, the four possible projections of the NV centers (or any other $C_{3v}$ defect) along the magnetic field are identical.
+This means that a B field scan along this direction cannot make the transition energies of the different classes of NV centers cross.
+Under arbitrary angles, this could otherwise result in several low field PL features \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the supplementary materials for data showing scans along the [111] direction).
+The downside of this choice of the [100] direction is that at large amplitudes, the transverse component of the magnetic field depolarizes all classes of NVs, thus limiting the magnetic field range that can be employed and therefore limiting this detection process to paramagnetic defects with zero-field splittings close to the NV center's one.
 
-In order to attribute the the three features to their respective defects, we run a similar scan (not shown) on heavily doped HPHT diamond samples in the same 100 direction.
+In order to identify the $[100]$ direction, we perform an angular scan of the magnetic field using a dual-axis goniometer (Thorlabs GNL20-Z8) that holds a permanent magnet at a fixed distance from the sample.
+Fig.
+\ref{map}-a) shows the NV photoluminescence as a function of magnetic field direction, referenced by azimuthal and polar angles $(\phi, \theta)$ with respect to the $[100]$ direction, using the above described CVD-grown sample.
+The PL is seen to drop for particular values of the magnet angular coordinates that correspond to specific crystalline axes.
+Such a drop of the PL, also observed in
+\cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}, correspond to situation where the transitions of NV centers become degenerate.
+The planes orthogonal to the [010], [001], [011], and [01$\bar 1$] directions are indicated by dashed lines and show the locus of the cross relaxation. These planes all cross on the [100] axis.
+The origin of these sharp changes in the photoluminescence was attributed to cross relaxation between polarized NV centers and rapidly decaying NV centers, so called fluctuators \cite{choi_depolarization_2017}, the precise origin of the latter remains unknown.
+The width of the CR features when the magnetic field crosses a plane perpendicularly was found to be compatible with the NV decoherence rate  ($\approx$ 6 MHz).
+The contrast is determined both by the fluctuating NV and polarized NV concentrations.
+
+Such a map can in itself be useful for measuring magnetic fields without microwave \cite{akhmedzhanov_magnetometry_2019}. 
+It also enables to identify the crystalline axes. 
+To verify that the central line is the [100] axis, we realize microwave scans around the NV transitions.
+Away from the crystal planes, we consistently observe 8 ODMR lines coming from the four $| m_s=0\rangle$ to $| m_s=\pm1\rangle$ spin transitions of the \{111\}-oriented NV centers.
+On the planes orthogonal to the [010], [001], [011], and [01$\bar 1$] directions however, we expect degeneracies.
+Fig. \ref{map}-b) shows an ODMR spectrum taken at $(\phi, \theta)=(3^\circ,3^\circ)$. As expected, at this position, two pairs of NV classes cross.
+Fig.
+\ref{map}-b) shows an ESR taken at the angle  $(\phi, \theta)=(0^\circ,0^\circ)$ showing only two features, as expected from a [100] axis.
+Using this goniometer, the magnetic field angle could be finely adjusted so that the four NV lines become fully degenerate along the $[100]$ direction with an error estimate of $\pm 0.5^\circ$.
+
+Fig. \ref{scan} shows the PL as a function of magnetic field amplitude along the $[ 100 ]$ direction, in the 15~G to 145~G range. 
+Calibration of the magnetic field amplitude was realized by applying microwave signals at varying frequencies in 2 MHz steps on several magnetic field scans.
+The microwave induced PL features in the scan are then used to relate the magnetic field to the applied voltage.
+Three features appear in the overall PL evolution in this spectrum : one feature at 20~G,  56~G and 122~G. 
+For these data, the averaging was done for 24 hours, but the features appear already with a signal to noise ratio greater than 1 after 10 minutes.  
+We also note an overall drop of the PL, as a result of state mixing in the optically excited state \cite{DOHERTY20131}.
+
+In order to let the three salient features detach better from the spectrum, we fitted a 4th-order polynomial to the PL evolution, without the spectral bumps, and subtract it from the data.
+We chose to use a 4th-order polynomial since numerical solving of the transverse-field dependent spin depolarization could not properly fit our results.
+The order of the polynomial fit was chosen to be 4 since higher order did not yield significantly better envelope fitting, and lower order were noticeably worse.
+Thanks to the subtraction, the width and contrast of each peak can now be compared and their position can be determined with greater accuracy.
+The second and third peaks were better fitted with Gaussian line shapes than Lorentzian ones, which is coherent with the fact that the inhomogeneous broadening of the various spin species dominate the coupling strength between the spins.
+
+In order to attribute the three features to their respective defects, we run a similar scan (Section 1 in the Supplementary Materials) on a type Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range, in the same $[100]$ direction.
 Only the first feature appeared.
-This observation led us to search for CVD specific defects as candidates for the last two peaks.
-Paramagnetic defects in CVD grown diamonds have been extensively studied using EPR at the university of Warwick.
-These studies demonstrate that  NVH, VH, VH2 complexes can be stable in diamond and they also estimate the zero-field splitting $D$ for most of them.
-Oxygen related complex may also occur here, due to our N2O are not resonant (https://core.ac.uk/download/pdf/46556.pdf)
-The first peak could then be coming from dipolar coupling between NVs and paramagnetic defects that are present in all diamonds, such as the substitutional nitrogen, also called $P_1$ centers ($[P_1]\approx 10 ppm$ ?) and the ${13}^C$ atoms (${13}^C= 1\%$).
-Using the results from Budker et al. we first intended to correlate the peaks with NV centers that are coupled to $P_1$ centers. 
- $P_1$ centers have a small zero-field splitting so in order to appear a such small magnetic fields, they would first have to be coupled to an NV centers and then be coupled to a nearby polarized NV.
-The obtained eigenfrequencies were inconsistent with our observed peaks and our spectral resolution error (~1G). This lead us to consider the ${13}^C$ atoms. 
-In itself, the ${13}^C$ spin has a very small zero-field splitting too (=?). In order to give a contribution to the spectra, it would have to be strongly coupled to an NV center. 
-In this case, the $^{13}$C-NV pair would show 4 transitions corresponding to ....
-
-Fig. \ref{energy-levels} shows the transition frequencies of the NV$^-$, the $^{13}$C-NV pair and the VH- as a function of the magnetic field amplitude along the 100 direction.
-The energy evolution of another defect, whose structure is unknown, and was coined the WAR1 defects is plotted on this graph. 
-The points were the NV levels cross the other defects would in principle correspond to an enhanced double quantum cross-relaxation process between the NV and the presumably unpolarized species. 
-The NV0 ?
-
-\begin{figure}[htbp]
-\centering
-\fbox{\includegraphics[width=\linewidth]{Theorie_CR}}
-\caption{Simulated transitions energies of the various spins considered as a function of a magnetic field aligned along the [100] axis.
-NV centers transitions are represented using a plain line, VH$^-$ using a dashed line, the WAR1 using a dash-dotted line and $^{13}$C-NV pair using a dotted line.
-The amplitudes of the magnetic field where the energy level of the NV center crosses the one of another species are represented by vertical dashed lines}
-\label{energy-levels}
-\end{figure}
-
-The first peak that we observe coincides with cross-relaxation that is expected to take place with $^{13}$C atoms that is one shell away from the Nitrogen vacancy centers. 
-This drop of the PL is consistent with $^{13}$C-NV pair.
-Using typical values found in the literature for the unpolarized NVs (also called fluctuators), we can estimate the concentration of this $^{13}$C-NV pair to be XX, making this hypothesis very likely.
-The second peak that we observe coincides very well with the crossing between the 0 to -1 NV transition and the 0 to +1 VH- transition.
-Such a defect was studied in details in XXX and was found to have a $D=$. 
-Making a gaussian fit to this data show that the position of the detected peak is , which gives a $D=$.
-Estimation suggest that the concentration of this defect is about 100 ppb (?).
-
-The third peak corresponds to a spin defect that has a D=.
-We found in Cruddace that it was already observed in EPR, although its structure is unknown.
-
-NVH (H3) center (diamagnetic in ground state), paramagnetic only in excited state but requires UV light. 
-No other species were observed in the optical spectrum or EPR.
-
-The technique can be used to detect the degree of polarisation of $^{13}$C atoms at room temperature.
-
-Table \ref{tab:shape-functions}
-
+This observation guided us to search for CVD-related defects as good candidates for the last two features.
+Paramagnetic defects in CVD-grown diamonds have been extensively studied using EPR  \citep{Ashfold}.
+These studies demonstrate that hydrogen-related complex such as the NVH, VH, VH2 complexes can all be stable in diamond.
+Although their composition is not always known, several zero-field splittings $D$ can be found in the literature.
+Given the small difference between the NV transitions and the observed features, we only concentrate on the reported defects with ZFS that are close to the NV$^-$.
+The middle column of table \ref{table}, shows the zero-field splitting for two such spin-1 defects found in \citep{cruddace2007magnetic}, namely the negatively charged hydrogen-vacancy (VH$^-$) and WAR1 defects.
+The latter was analyzed in EPR but its exact structure is unknown \citep{cruddace2007magnetic}.
 \begin{table}[htbp]
 \centering
-\caption{\bf Zero-field splitting parameter $D_z$ for the different spin-1 species}
+\caption{\bf Zero-field splitting parameter $D$ for the spin-1 species in our sample}
 \begin{tabular}{ccc}
@@
 \end{tabular}
-  \label{tab:shape-functions}
+  \label{table}
 \end{table}
+Fig.
+\ref{energy-levels} shows the frequencies of the NV$^-$, the VH$^-$ and WAR1 spin transitions as a function of the magnetic field amplitude along the $[100]$ direction.
+Since the NV$^-$ and VH$^-$ are $C_{3v}$ defects, and WAR1 is a pseudo-$C_{3v}$ defect \citep{cruddace2007magnetic}, all possible orientation are degenerate for this magnetic field orientation, giving only two lines corresponding to the $\ket{0} \to \ket{+1}$ and $\ket{0} \to \ket{-1}$ transitions.
+The points were the NV levels cross the other defects can give rise to cross-relaxation.
 
-\section{Funding}
-GH acknowledges funding by the French National Research Agency (ANR) through the T-ERC project QUOVADIS.
+Using this theoretical calculations, we find that the second peak in Fig.
+\ref{scan} coincides very well with a CR that would occur at the crossing between the $|m_s=0\rangle$ to $|m_s=-1\rangle$ NV transition and the $|m_s=0\rangle$ to $|m_s=+1\rangle$ VH$^-$ transition.
+A gaussian fit to the feature in Fig.
+\ref{scan} enables us to extrapolate a value for $D=2694(5)$MHz that coincides very well to that of the VH$^-$ within the error margins indicated in \citep{cruddace2007magnetic}.
+The third peak at 122 G would correspond to a spin defect that has a $D=2470(10)$MHz that also matches well the one of the WAR1 defect. 
+The values of $D$ for these two defects and the error bars from our measurements are now included in the table \ref{table}.
+The error bars are estimated by taking into account the precision on the NV$^-$ ZFS, on the angle, on the magnetic field calibration and the precision of the fits. 
+We obtain a factor of 6 improvement over the precision reported in \citep{cruddace2007magnetic}.
+We thus attribute the second and third feature to cross relaxation between the NV center and the VH$^-$ and WAR1 defect.
+Scans along the [111] direction shown in the Supplementary Material also corroborate this conclusion.
 
-\section{Acknowledgments}
-We would like to thank Neil Manson and Carols Meriles for fruitful discussions.
+\begin{figure}[htbp]
+\centering
+{\includegraphics[width=\linewidth]{Theorie_CR}}
+\caption{Simulated transition energies of the various considered spins as a function of a magnetic field aligned to the [100] axis.
+The NV centers electronic spin transitions are shown by the plain thick line, the VH$^-$ using dashed lines, the WAR1 using dash-dotted lines and $^{13}$C-NV pair using dotted lines.
+The amplitudes of the magnetic field where the energy level of the NV center crosses the one of another species are represented by vertical dashed lines.}
+\label{energy-levels}
+\end{figure}
+The remaining broader feature in the spectrum could be coming from dipolar coupling between the spin of NV centers and paramagnetic defects that are present also in HPHT-grown diamonds, such as the substitutional nitrogen, also called $P_1$ centers ($[P_1]\approx 5-20$ ppm in our sample) or the $^{13}$C atoms (natural isotopic abundance [$^{13}$C]= 1\%).
+Using the results from \cite{simanovskaia_sidebands_2013}, we first intended to correlate its spectral position to the transitions of dipolar-coupled NV and $P_1$ centers. 
+ $P_1$ centers have a zero-field splitting of only 100 MHz.
+In order to cross one of the two NV spin transitions at a magnetic field of 20 G, they would then have to be coupled off-resonantly to the spin of nearby NV centers.
+The resulting pair could then be coupled resonantly to a nearby polarized NV.
+The obtained eigenfrequencies of the dipolar coupled P$_1$-NV pair that we extracted were all inconsistent with the spectral positions of our observed peak within our spectral resolution error ($\approx$1G) (see Section 3 of the supplementary materials).
+This leads us to consider instead the nuclear spin of the $^{13}$C atoms as the most likely candidate.
 
-\section{Disclosures}
+The nuclear spin of the $^{13}$C atoms do not have a ZFS and its gyromagnetic factor is four orders of magnitude lower than that of the electron.
+Here again, in order to give a CR contribution to the spectrum at modest magnetic fields, the nuclear spin of the $^{13}$C would first have to be strongly coupled to an NV center.
+The resulting pair could then be co-resonant with the spin transitions of a bare NV center, as depicted in Fig.
+1.
+This situation can manifest itself when the nuclear spin of $^{13}$C atoms are a few shells distant from the nitrogen-vacancy centers.
+A strong hyperfine coupling rate of 130 MHz can for instance be reached with a $^{13}$C that is only one shell away from the NV center.
+In contrast, higher order shells have a maximum hyperfine coupling of 15 MHz \cite{smeltzer201113c} making them hard to distinguish from the inhomogeneous broadening.
+The modeling of this interaction is presented in the section 2 of the supplementary materials.
+The dotted lines in Fig.
+\ref{energy-levels} show the four transition frequencies of the first shell $^{13}$C coupled to an NV center.
+The crossings between the $|m_s=0\rangle$ to $|m_s=\pm 1\rangle$ NV transition and the four transitions of this  $^{13}$C-NV pair occur at magnetic field values around 20~G.
+We added these four lines to the experimental spectrum shown in Fig.
+\ref{scan}.
+A good agreement is found between the magnetic field at which of these four transitions cross the NV and the broad 20~G feature, letting us conclude that it is the result of a CR between an unpolarized $^{13}$C-NV pair and a polarized NV center.
 
-The authors declare no conflicts of interest.'
+Because the sample we use has a natural 1.1\% abundance of $^{13}$C, we can actually use the first peak as a way to calibrate the concentration of VH$^-$ and WAR1 in our sample.
+Because there are three possible sites for a first-shell $^{13}$C-NV pair, we can estimate that the concentration of $^{13}$C-NV pair is about 3.3\% of the total concentration of NV centers, i.e.
+$\sim$135~ppb.
+We can then compare the area of the second and third peak to the first one to have a first order estimation of the concentration of VH$^-$ and WAR1.
+We find [VH$^-$]$\approx$195~ppb and [WAR1]$\approx$124~ppb.
+We decided to compare the area of the peaks instead of the amplitudes since the width of the peak is mostly due to inhomogeneous spread of the spins, and not to the spin-spin interaction strength.
+It should be noted that a severe drawback of this method is the loss of contrast for higher magnetic field amplitude, meaning in this case that the WAR1 concentration is probably undervalued.
 
-% Bibliography
-\bibliography{Biblio_CVD-CR}
+In conclusion, we demonstrated all-optical detection of paramagnetic species in a diamond grown by the Chemical-Vapor-Deposition method. 
+Using magnetic field scans along the [100] direction, we identified 3-body interactions between NV centers and $^{13}$C-NV pairs, and cross-relaxations between the spin of NV centers and the VH$^-$ and the WAR1 defects, by coupling them to NV centers.
+Our results offer prospects for more detailed studies of CVD-grown processes as well as for realizing quantum networks with these newly coupled spins.
 
-% Full bibliography added automatically for Optics Letters submissions; the following line will simply be ignored if submitting to other journals.
-% Note that this extra page will not count against page length
-\bibliographyfullrefs{sample}
+\section*{Acknowledgments}
+We would like to thank Neil Manson and Carlos Meriles for fruitful discussions. 
+The authors would also like to thank SIRTEQ for funding. GH also acknowledges funding by the French National Research Agency (ANR) through the T-ERC project QUOVADIS.
+
+%merlin.mbs apsrev4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
+%Control: key (0)
+%Control: author (8) initials jnrlst
+%Control: editor formatted (1) identically to author
+%Control: production of article title (-1) disabled
+%Control: page (0) single
+%Control: year (1) truncated
+%Control: production of eprint (0) enabled
+\begin{thebibliography}{27}%
+\makeatletter
+\providecommand \@ifxundefined [1]{%
+ \@ifx{#1\undefined}
+}%
+\providecommand \@ifnum [1]{%
+ \ifnum #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \@ifx [1]{%
+ \ifx #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \natexlab [1]{#1}%
+\providecommand \enquote  [1]{``#1''}%
+\providecommand \bibnamefont  [1]{#1}%
+\providecommand \bibfnamefont [1]{#1}%
+\providecommand \citenamefont [1]{#1}%
+\providecommand \href@noop [0]{\@secondoftwo}%
+\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
+\providecommand \@href[1]{\@@startlink{#1}\@@href}%
+\providecommand \@@href[1]{\endgroup#1\@@endlink}%
+\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
+  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
+\providecommand \@@startlink[1]{}%
+\providecommand \@@endlink[0]{}%
+\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
+\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
+\providecommand \urlprefix  [0]{URL }%
+\providecommand \Eprint [0]{\href }%
+\providecommand \doibase [0]{http://dx.doi.org/}%
+\providecommand \selectlanguage [0]{\@gobble}%
+\providecommand \bibinfo  [0]{\@secondoftwo}%
+\providecommand \bibfield  [0]{\@secondoftwo}%
+\providecommand \translation [1]{[#1]}%
+\providecommand \BibitemOpen [0]{}%
+\providecommand \bibitemStop [0]{}%
+\providecommand \bibitemNoStop [0]{.\EOS\space}%
+\providecommand \EOS [0]{\spacefactor3000\relax}%
+\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
+\let\auto@bib@innerbib\@empty
+%</preamble>
+\bibitem [{\citenamefont {Rondin}\ \emph {et~al.}(2014)\citenamefont {Rondin},
+  \citenamefont {Tetienne}, \citenamefont {Hingant}, \citenamefont {Roch},
+  \citenamefont {Maletinsky},\ and\ \citenamefont {Jacques}}]{Rondin_2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Rondin}}, \bibinfo {author} {\bibfnamefont {J.-P.}\ \bibnamefont
+  {Tetienne}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Hingant}},
+  \bibinfo {author} {\bibfnamefont {J.-F.}\ \bibnamefont {Roch}}, \bibinfo
+  {author} {\bibfnamefont {P.}~\bibnamefont {Maletinsky}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {V.}~\bibnamefont {Jacques}},\ }\href {\doibase
+  10.1088/0034-4885/77/5/056503} {\bibfield  {journal} {\bibinfo  {journal}
+  {Reports on Progress in Physics}\ }\textbf {\bibinfo {volume} {77}},\
+  \bibinfo {pages} {056503} (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Doherty}\ \emph {et~al.}(2013)\citenamefont
+  {Doherty}, \citenamefont {Manson}, \citenamefont {Delaney}, \citenamefont
+  {Jelezko}, \citenamefont {Wrachtrup},\ and\ \citenamefont
+  {Hollenberg}}]{DOHERTY20131}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~W.}\ \bibnamefont
+  {Doherty}}, \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont {Manson}},
+  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Delaney}}, \bibinfo
+  {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Wrachtrup}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {L.~C.}\ \bibnamefont {Hollenberg}},\ }\href {\doibase
+  https://doi.org/10.1016/j.physrep.2013.02.001} {\bibfield  {journal}
+  {\bibinfo  {journal} {Physics Reports}\ }\textbf {\bibinfo {volume} {528}},\
+  \bibinfo {pages} {1 } (\bibinfo {year} {2013})},\ \bibinfo {note} {the
+  nitrogen-vacancy colour centre in diamond}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Achard}\ \emph {et~al.}(2020)\citenamefont {Achard},
+  \citenamefont {Jacques},\ and\ \citenamefont {Tallaire}}]{Achard}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Achard}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Jacques}}, \
+  and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Tallaire}},\ }\href
+  {\doibase 10.1088/1361-6463/ab81d1} {\bibfield  {journal} {\bibinfo
+  {journal} {Journal of Physics D: Applied Physics}\ }\textbf {\bibinfo
+  {volume} {53}},\ \bibinfo {pages} {313001} (\bibinfo {year}
+  {2020})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Newton}(2007)}]{newton_epr_2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont
+  {Newton}},\ }\href {\doibase 10.1002/chin.200747215} {\bibfield  {journal}
+  {\bibinfo  {journal} {ChemInform}\ }\textbf {\bibinfo {volume} {38}}
+  (\bibinfo {year} {2007}),\ 10.1002/chin.200747215}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Glover}\ \emph {et~al.}(2004)\citenamefont {Glover},
+  \citenamefont {Newton}, \citenamefont {Martineau}, \citenamefont {Quinn},\
+  and\ \citenamefont {Twitchen}}]{glover_hydrogen_2004}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Glover}}, \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont {Newton}},
+  \bibinfo {author} {\bibfnamefont {P.~M.}\ \bibnamefont {Martineau}}, \bibinfo
+  {author} {\bibfnamefont {S.}~\bibnamefont {Quinn}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {D.~J.}\ \bibnamefont {Twitchen}},\ }\href {\doibase
+  10.1103/PhysRevLett.92.135502} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {92}},\ \bibinfo {pages}
+  {135502} (\bibinfo {year} {2004})},\ \bibinfo {note} {publisher: American
+  Physical Society}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Glover}\ \emph {et~al.}(2003)\citenamefont {Glover},
+  \citenamefont {Newton}, \citenamefont {Martineau}, \citenamefont {Twitchen},\
+  and\ \citenamefont {Baker}}]{glover_hydrogen_2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Glover}}, \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont {Newton}},
+  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Martineau}}, \bibinfo
+  {author} {\bibfnamefont {D.~J.}\ \bibnamefont {Twitchen}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Baker}},\ }\href {\doibase
+  10.1103/PhysRevLett.90.185507} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {90}},\ \bibinfo {pages}
+  {185507} (\bibinfo {year} {2003})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Yao}\ \emph {et~al.}(2012)\citenamefont {Yao},
+  \citenamefont {Jiang}, \citenamefont {Gorshkov}, \citenamefont {Maurer},
+  \citenamefont {Giedke}, \citenamefont {Cirac},\ and\ \citenamefont
+  {Lukin}}]{Yao}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~Y.}\ \bibnamefont
+  {Yao}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Jiang}}, \bibinfo
+  {author} {\bibfnamefont {A.~V.}\ \bibnamefont {Gorshkov}}, \bibinfo {author}
+  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
+  {\bibfnamefont {G.}~\bibnamefont {Giedke}}, \bibinfo {author} {\bibfnamefont
+  {J.~I.}\ \bibnamefont {Cirac}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {M.~D.}\ \bibnamefont {Lukin}},\ }\href@noop {} {\bibfield  {journal}
+  {\bibinfo  {journal} {Nature Communications}\ }\textbf {\bibinfo {volume}
+  {3}},\ \bibinfo {pages} {800} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Cruddace}(2007)}]{cruddace2007magnetic}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Cruddace}},\ }\emph {\bibinfo {title} {Magnetic resonance and optical
+  studies of point defects in single crystal CVD diamond}},\ \href@noop {}
+  {Ph.D. thesis},\ \bibinfo  {school} {University of Warwick} (\bibinfo {year}
+  {2007})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Epstein}\ \emph {et~al.}(2005)\citenamefont
+  {Epstein}, \citenamefont {Mendoza}, \citenamefont {Kato},\ and\ \citenamefont
+  {Awschalom}}]{Epstein}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~J.}\ \bibnamefont
+  {Epstein}}, \bibinfo {author} {\bibfnamefont {F.~M.}\ \bibnamefont
+  {Mendoza}}, \bibinfo {author} {\bibfnamefont {Y.~K.}\ \bibnamefont {Kato}}, \
+  and\ \bibinfo {author} {\bibfnamefont {D.~D.}\ \bibnamefont {Awschalom}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Nature Physics}\
+  }\textbf {\bibinfo {volume} {1}},\ \bibinfo {pages} {94} (\bibinfo {year}
+  {2005})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Wang}\ \emph {et~al.}(2014)\citenamefont {Wang},
+  \citenamefont {Shin}, \citenamefont {Seltzer}, \citenamefont {Avalos},
+  \citenamefont {Pines},\ and\ \citenamefont {Bajaj}}]{Wang}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.-J.}\ \bibnamefont
+  {Wang}}, \bibinfo {author} {\bibfnamefont {C.~S.}\ \bibnamefont {Shin}},
+  \bibinfo {author} {\bibfnamefont {S.~J.}\ \bibnamefont {Seltzer}}, \bibinfo
+  {author} {\bibfnamefont {C.~E.}\ \bibnamefont {Avalos}}, \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Pines}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {V.~S.}\ \bibnamefont {Bajaj}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Nature Communications}\ }\textbf {\bibinfo
+  {volume} {5}},\ \bibinfo {pages} {4135} (\bibinfo {year} {2014})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Armstrong}\ \emph {et~al.}(2010)\citenamefont
+  {Armstrong}, \citenamefont {Rogers}, \citenamefont {McMurtrie},\ and\
+  \citenamefont {Manson}}]{armstrong_nvnv_2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Armstrong}}, \bibinfo {author} {\bibfnamefont {L.~J.}\ \bibnamefont
+  {Rogers}}, \bibinfo {author} {\bibfnamefont {R.~L.}\ \bibnamefont
+  {McMurtrie}}, \ and\ \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont
+  {Manson}},\ }\href {\doibase 10.1016/j.phpro.2010.01.223} {\bibfield
+  {journal} {\bibinfo  {journal} {Physics Procedia}\ }\textbf {\bibinfo
+  {volume} {3}},\ \bibinfo {pages} {1569} (\bibinfo {year} {2010})},\ \bibinfo
+  {note} {number: 4}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Hall}\ \emph {et~al.}(2016)\citenamefont {Hall},
+  \citenamefont {Kehayias}, \citenamefont {Simpson}, \citenamefont {Jarmola},
+  \citenamefont {Stacey}, \citenamefont {Budker},\ and\ \citenamefont
+  {Hollenberg}}]{Hall}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.~T.}\ \bibnamefont
+  {Hall}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Kehayias}},
+  \bibinfo {author} {\bibfnamefont {D.~A.}\ \bibnamefont {Simpson}}, \bibinfo
+  {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}}, \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Stacey}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Budker}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {L.~C.~L.}\ \bibnamefont {Hollenberg}},\ }\href@noop {} {\bibfield  {journal}
+  {\bibinfo  {journal} {Nature Communications}\ }\textbf {\bibinfo {volume}
+  {7}},\ \bibinfo {pages} {10211} (\bibinfo {year} {2016})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Wickenbrock}\ \emph {et~al.}(2016)\citenamefont
+  {Wickenbrock}, \citenamefont {Zheng}, \citenamefont {Bougas}, \citenamefont
+  {Leefer}, \citenamefont {Afach}, \citenamefont {Jarmola}, \citenamefont
+  {Acosta},\ and\ \citenamefont {Budker}}]{Wickenbrock}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Wickenbrock}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Zheng}},
+  \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Bougas}}, \bibinfo
+  {author} {\bibfnamefont {N.}~\bibnamefont {Leefer}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Afach}}, \bibinfo {author} {\bibfnamefont
+  {A.}~\bibnamefont {Jarmola}}, \bibinfo {author} {\bibfnamefont {V.~M.}\
+  \bibnamefont {Acosta}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Budker}},\ }\href {\doibase 10.1063/1.4960171} {\bibfield
+  {journal} {\bibinfo  {journal} {Applied Physics Letters}\ }\textbf {\bibinfo
+  {volume} {109}},\ \bibinfo {pages} {053505} (\bibinfo {year} {2016})},\
+  \Eprint {http://arxiv.org/abs/https://doi.org/10.1063/1.4960171}
+  {https://doi.org/10.1063/1.4960171} \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Alfasi}\ \emph {et~al.}(2019)\citenamefont {Alfasi},
+  \citenamefont {Masis}, \citenamefont {Shtempluck},\ and\ \citenamefont
+  {Buks}}]{Alfasi}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
+  {Alfasi}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Masis}},
+  \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Shtempluck}}, \ and\
+  \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {Buks}},\ }\href {\doibase
+  10.1103/PhysRevB.99.214111} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. B}\ }\textbf {\bibinfo {volume} {99}},\ \bibinfo {pages} {214111}
+  (\bibinfo {year} {2019})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {{Edmonds}}\ \emph {et~al.}(2020)\citenamefont
+  {{Edmonds}}, \citenamefont {{Hart}}, \citenamefont {{Turner}}, \citenamefont
+  {{Colard}}, \citenamefont {{Schloss}}, \citenamefont {{Olsson}},
+  \citenamefont {{Trubko}}, \citenamefont {{Markham}}, \citenamefont
+  {{Rathmill}}, \citenamefont {{Horne-Smith}}, \citenamefont {{Lew}},
+  \citenamefont {{Manickam}}, \citenamefont {{Bruce}}, \citenamefont {{Kaup}},
+  \citenamefont {{Russo}}, \citenamefont {{DiMario}}, \citenamefont {{South}},
+  \citenamefont {{Hansen}}, \citenamefont {{Twitchen}},\ and\ \citenamefont
+  {{Walsworth}}}]{Edmonds}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
+  {{Edmonds}}}, \bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont
+  {{Hart}}}, \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont
+  {{Turner}}}, \bibinfo {author} {\bibfnamefont {P.-O.}\ \bibnamefont
+  {{Colard}}}, \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
+  {{Schloss}}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {{Olsson}}},
+  \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {{Trubko}}}, \bibinfo
+  {author} {\bibfnamefont {M.~L.}\ \bibnamefont {{Markham}}}, \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {{Rathmill}}}, \bibinfo {author}
+  {\bibfnamefont {B.}~\bibnamefont {{Horne-Smith}}}, \bibinfo {author}
+  {\bibfnamefont {W.}~\bibnamefont {{Lew}}}, \bibinfo {author} {\bibfnamefont
+  {A.}~\bibnamefont {{Manickam}}}, \bibinfo {author} {\bibfnamefont
+  {S.}~\bibnamefont {{Bruce}}}, \bibinfo {author} {\bibfnamefont {P.~G.}\
+  \bibnamefont {{Kaup}}}, \bibinfo {author} {\bibfnamefont {J.~C.}\
+  \bibnamefont {{Russo}}}, \bibinfo {author} {\bibfnamefont {M.~J.}\
+  \bibnamefont {{DiMario}}}, \bibinfo {author} {\bibfnamefont {J.~T.}\
+  \bibnamefont {{South}}}, \bibinfo {author} {\bibfnamefont {J.~T.}\
+  \bibnamefont {{Hansen}}}, \bibinfo {author} {\bibfnamefont {D.~J.}\
+  \bibnamefont {{Twitchen}}}, \ and\ \bibinfo {author} {\bibfnamefont {R.~L.}\
+  \bibnamefont {{Walsworth}}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
+  {journal} {arXiv e-prints}\ ,\ \bibinfo {eid} {arXiv:2004.01746}} (\bibinfo
+  {year} {2020})},\ \Eprint {http://arxiv.org/abs/2004.01746} {arXiv:2004.01746
+  [cond-mat.mtrl-sci]} \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Tallaire}\ \emph {et~al.}(2020)\citenamefont
+  {Tallaire}, \citenamefont {Brinza}, \citenamefont {Huillery}, \citenamefont
+  {Delord}, \citenamefont {Pellet-Mary}, \citenamefont {Staacke}, \citenamefont
+  {Abel}, \citenamefont {Pezzagna}, \citenamefont {Meijer}, \citenamefont
+  {Touati}, \citenamefont {Binet}, \citenamefont {Ferrier}, \citenamefont
+  {Goldner}, \citenamefont {Hetet},\ and\ \citenamefont
+  {Achard}}]{TALLAIRE2020421}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Tallaire}}, \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Brinza}},
+  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Huillery}}, \bibinfo
+  {author} {\bibfnamefont {T.}~\bibnamefont {Delord}}, \bibinfo {author}
+  {\bibfnamefont {C.}~\bibnamefont {Pellet-Mary}}, \bibinfo {author}
+  {\bibfnamefont {R.}~\bibnamefont {Staacke}}, \bibinfo {author} {\bibfnamefont
+  {B.}~\bibnamefont {Abel}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Pezzagna}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Meijer}},
+  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Touati}}, \bibinfo
+  {author} {\bibfnamefont {L.}~\bibnamefont {Binet}}, \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Ferrier}}, \bibinfo {author} {\bibfnamefont
+  {P.}~\bibnamefont {Goldner}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Hetet}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Achard}},\ }\href {\doibase
+  https://doi.org/10.1016/j.carbon.2020.08.048} {\bibfield  {journal} {\bibinfo
+   {journal} {Carbon}\ }\textbf {\bibinfo {volume} {170}},\ \bibinfo {pages}
+  {421 } (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Mindarava}\ \emph {et~al.}(2020)\citenamefont
+  {Mindarava}, \citenamefont {Blinder}, \citenamefont {Laube}, \citenamefont
+  {Knolle}, \citenamefont {Abel}, \citenamefont {Jentgens}, \citenamefont
+  {Isoya}, \citenamefont {Scheuer}, \citenamefont {Lang}, \citenamefont
+  {Schwartz}, \citenamefont {Naydenov},\ and\ \citenamefont
+  {Jelezko}}]{MINDARAVA2020182}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Mindarava}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Blinder}},
+  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Laube}}, \bibinfo
+  {author} {\bibfnamefont {W.}~\bibnamefont {Knolle}}, \bibinfo {author}
+  {\bibfnamefont {B.}~\bibnamefont {Abel}}, \bibinfo {author} {\bibfnamefont
+  {C.}~\bibnamefont {Jentgens}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Scheuer}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Lang}}, \bibinfo {author} {\bibfnamefont {I.}~\bibnamefont
+  {Schwartz}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Naydenov}},
+  \ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},\ }\href
+  {\doibase https://doi.org/10.1016/j.carbon.2020.07.077} {\bibfield  {journal}
+  {\bibinfo  {journal} {Carbon}\ }\textbf {\bibinfo {volume} {170}},\ \bibinfo
+  {pages} {182 } (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Jarmola}\ \emph {et~al.}(2015)\citenamefont
+  {Jarmola}, \citenamefont {Berzins}, \citenamefont {Smits}, \citenamefont
+  {Smits}, \citenamefont {Prikulis}, \citenamefont {Gahbauer}, \citenamefont
+  {Ferber}, \citenamefont {Erts}, \citenamefont {Auzinsh},\ and\ \citenamefont
+  {Budker}}]{jarmola_longitudinal_2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Jarmola}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Berzins}},
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Smits}}, \bibinfo
+  {author} {\bibfnamefont {K.}~\bibnamefont {Smits}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Prikulis}}, \bibinfo {author}
+  {\bibfnamefont {F.}~\bibnamefont {Gahbauer}}, \bibinfo {author}
+  {\bibfnamefont {R.}~\bibnamefont {Ferber}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Erts}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Auzinsh}}, \ and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
+  {Budker}},\ }\href {\doibase 10.1063/1.4937489} {\bibfield  {journal}
+  {\bibinfo  {journal} {Appl. Phys. Lett.}\ }\textbf {\bibinfo {volume}
+  {107}},\ \bibinfo {pages} {242403} (\bibinfo {year} {2015})},\ \bibinfo
+  {note} {number: 24}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {van Oort}\ and\ \citenamefont
+  {Glasbeek}(1991)}]{van_oort_optically_1991}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {van
+  Oort}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Glasbeek}},\
+  }\href {\doibase 10.1007/BF03166042} {\bibfield  {journal} {\bibinfo
+  {journal} {Appl Magn Reson}\ }\textbf {\bibinfo {volume} {2}},\ \bibinfo
+  {pages} {291} (\bibinfo {year} {1991})},\ \bibinfo {note} {number:
+  2}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2019)\citenamefont
+  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
+  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
+  and\ \citenamefont {Hemmer}}]{akhmedzhanov_magnetometry_2019}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
+  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
+  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
+  {\bibfnamefont {I.}~\bibnamefont {Zelensky}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href {\doibase
+  10.1103/PhysRevA.100.043844} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys. Rev. A}\ }\textbf {\bibinfo {volume} {100}},\ \bibinfo {pages}
+  {043844} (\bibinfo {year} {2019})},\ \bibinfo {note} {number: 4}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {van Oort}\ and\ \citenamefont
+  {Glasbeek}(1989)}]{van_oort_cross-relaxation_1989}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {van
+  Oort}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Glasbeek}},\
+  }\href {\doibase 10.1103/PhysRevB.40.6509} {\bibfield  {journal} {\bibinfo
+  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {40}},\ \bibinfo
+  {pages} {6509} (\bibinfo {year} {1989})},\ \bibinfo {note} {number:
+  10}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2017)\citenamefont
+  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
+  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
+  and\ \citenamefont {Hemmer}}]{akhmedzhanov_microwave-free_2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
+  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
+  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
+  {\bibfnamefont {I.}~\bibnamefont {Zelensky}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href {\doibase
+  10.1103/PhysRevA.96.013806} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. A}\ }\textbf {\bibinfo {volume} {96}},\ \bibinfo {pages} {013806}
+  (\bibinfo {year} {2017})},\ \bibinfo {note} {number: 1}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Holliday}\ \emph {et~al.}(1989)\citenamefont
+  {Holliday}, \citenamefont {Manson}, \citenamefont {Glasbeek},\ and\
+  \citenamefont {Oort}}]{holliday_optical_1989}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {Holliday}}, \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont
+  {Manson}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Glasbeek}}, \
+  and\ \bibinfo {author} {\bibfnamefont {E.~v.}\ \bibnamefont {Oort}},\ }\href
+  {\doibase 10.1088/0953-8984/1/39/021} {\bibfield  {journal} {\bibinfo
+  {journal} {J. Phys.: Condens. Matter}\ }\textbf {\bibinfo {volume} {1}},\
+  \bibinfo {pages} {7093} (\bibinfo {year} {1989})},\ \bibinfo {note} {number:
+  39}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Choi}\ \emph {et~al.}(2017)\citenamefont {Choi},
+  \citenamefont {Choi}, \citenamefont {Kucsko}, \citenamefont {Maurer},
+  \citenamefont {Shields}, \citenamefont {Sumiya}, \citenamefont {Onoda},
+  \citenamefont {Isoya}, \citenamefont {Demler}, \citenamefont {Jelezko},
+  \citenamefont {Yao},\ and\ \citenamefont {Lukin}}]{choi_depolarization_2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Choi}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo
+  {author} {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author}
+  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
+  {\bibfnamefont {B.~J.}\ \bibnamefont {Shields}}, \bibinfo {author}
+  {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
+  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
+  {E.}~\bibnamefont {Demler}}, \bibinfo {author} {\bibfnamefont
+  {F.}~\bibnamefont {Jelezko}}, \bibinfo {author} {\bibfnamefont {N.~Y.}\
+  \bibnamefont {Yao}}, \ and\ \bibinfo {author} {\bibfnamefont {M.~D.}\
+  \bibnamefont {Lukin}},\ }\href {\doibase 10.1103/PhysRevLett.118.093601}
+  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf
+  {\bibinfo {volume} {118}},\ \bibinfo {pages} {093601} (\bibinfo {year}
+  {2017})},\ \bibinfo {note} {number: 9}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ashfold}\ \emph {et~al.}(2020)\citenamefont
+  {Ashfold}, \citenamefont {Goss}, \citenamefont {Green}, \citenamefont {May},
+  \citenamefont {Newton},\ and\ \citenamefont {Peaker}}]{Ashfold}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~N.~R.}\
+  \bibnamefont {Ashfold}}, \bibinfo {author} {\bibfnamefont {J.~P.}\
+  \bibnamefont {Goss}}, \bibinfo {author} {\bibfnamefont {B.~L.}\ \bibnamefont
+  {Green}}, \bibinfo {author} {\bibfnamefont {P.~W.}\ \bibnamefont {May}},
+  \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont {Newton}}, \ and\
+  \bibinfo {author} {\bibfnamefont {C.~V.}\ \bibnamefont {Peaker}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Chemical Reviews}\
+  }\textbf {\bibinfo {volume} {120}},\ \bibinfo {pages} {5745} (\bibinfo {year}
+  {2020})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Simanovskaia}\ \emph {et~al.}(2013)\citenamefont
+  {Simanovskaia}, \citenamefont {Jensen}, \citenamefont {Jarmola},
+  \citenamefont {Aulenbacher}, \citenamefont {Manson},\ and\ \citenamefont
+  {Budker}}]{simanovskaia_sidebands_2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Simanovskaia}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {Jensen}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}},
+  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Aulenbacher}}, \bibinfo
+  {author} {\bibfnamefont {N.}~\bibnamefont {Manson}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }\href {\doibase
+  10.1103/PhysRevB.87.224106} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. B}\ }\textbf {\bibinfo {volume} {87}},\ \bibinfo {pages} {224106}
+  (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Smeltzer}\ \emph {et~al.}(2011)\citenamefont
+  {Smeltzer}, \citenamefont {Childress},\ and\ \citenamefont
+  {Gali}}]{smeltzer201113c}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.}~\bibnamefont
+  {Smeltzer}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Childress}},
+  \ and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Gali}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {New Journal of
+  Physics}\ }\textbf {\bibinfo {volume} {13}},\ \bibinfo {pages} {025021}
+  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
+\end{thebibliography}%
 

### 2021-02-23 Mes papiers/CR-CVD/Old/pre_SI.tex
tutu

--- 2020-11-27 Mes papiers/Double quantum/pre_SI.tex
@@
 \label{eq_P1}
-\nu^i_{P1}(B)=\nu^{0 \to -1}_{NV}(B)+\nu^{0 \to +1}_{NV}(B)
+\nu^i_{P1}(B)=\nu^{0 \to +1}_{NV}(B)-\nu^{0 \to +1}_{NV}(B)
 \end{equation}
@@
 
-Solving eq.
+Basing the numerical values of the P1 Hamiltonian on \citep{lazda2020cross}, we solve eq.
 \ref{eq_P1} for all possible P1 transitions (see fig.
-\ref{fig_P1}) therefore gives us all the possible magnetic field amplitude in the [100] direction where we could observe NV-P1 mutual flip transitions.
+\ref{fig_P1}) therefore giving us all the possible magnetic field amplitude in the [100] direction where we could observe NV-P1 mutual flip transitions.
 The predicted magnetic field values are 0, 3.89, 5.96, 6.58, 17.90, 28.93, 35.87, 49.44, 81.33, 83.28, 137.52, 154.20 and 246.34 G.

### 2021-02-23 Mes papiers/CR-CVD/SI/CVD_para_SI.tex
tutu

\documentclass[9pt,twocolumn,twoside]{revtex4-1}

\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{natbib}
\usepackage{physics}

\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}
\title{Optical Detection of Paramagnetic Defects in a CVD-grown diamond\\-- Supplementary Material}

\author{C. Pellet-Mary$^{1}$}
\author{P. Huillery $^{1}$}
\author{M. Perdriat $^{1}$}
\author{A. Tallaire $^{2}$}
\author{G. H\'etet$^{1}$}

\affiliation{$^1$Laboratoire de Physique de l'Ecole normale sup\'erieure, ENS, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e Paris-Diderot, Sorbonne Paris Cit\'e, Paris, France.}
\affiliation{$^2$IRCP, Ecole Nationale Supérieure de Chimie de Paris, 11, rue Pierre et Marie Curie, 75005 Paris, France}

\maketitle

\section{Photoluminescence as a function of the magnetic field for a HPHT sample}

The NV$^-$ photoluminescence as a function of the magnetic field amplitude that is presented in the Fig. 3 of the main text was realized on a CVD-grown sample. 
We also performed similar scans on several samples grown via the High-Pressure-High-Temperature (HPHT) method. 
Fig. \ref{fig1SI} shows a scan realized on HPHT samples that have been irradiated in order to increase the NV yield. 
The sample we use in Fig. \ref{fig1SI} is a Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range.
The PL feature that can be observed at 20G was also observed in the CVD-grown samples. We attributed it to the resonant dipolar interaction between a standard polarized NV center and one NV strongly coupled to the nuclear spin of a $^{13}$C atom sitting one shell away.
Notably, no other features could be detected at larger magnetic field values. We realized such scans on many different highly-doped HPHT samples and, systematically, no CR features are observed at 50 and 122 G.

 \begin{figure}[htbp]
\centering
{\includegraphics[width=\linewidth]{sumi4_13C.eps}}
\caption{NV$^-$ photoluminescence rate while scanning the magnetic field along the [100] crystalline direction. Here a HPHT sample was employed.}
\label{fig1SI}
\end{figure}

\section{NV-$^{13}$C Hamiltonian}

The full spin Hamiltonian of the NV-$^{13}$C complex can be written as follows : 
\begin{equation*}
\mathcal{H}=\mathcal{H}_{NV}+\mathcal{H}_{^{13}C}+\mathcal{H}_{HF},
\end{equation*}
where $\mathcal{H}_{NV}$ is the NV$^-$ spin Hamiltonian.
For one NV orientation, with quantization axis $z$, it reads 
\begin{equation}\hat{H}_{\rm NV}=\hbar D \hat{S}_z^2+ \hbar \gamma_e \bf B  \cdot \bf\hat S.
\end{equation}
$D$ is the zero-field-splitting and $\bf\hat S$ is the electronic spin operator.
$\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian : $\mathcal{H}_{HF}= \hat{\mathbf{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\mathbf{I}}_C$.

In the case of a first shell $^{13}$C, the hyper-fine tensor $\mathcal{A}$ can be written as \citep{simanovskaia_sidebands_2013} : $$ \mathcal{A} = \begin{pmatrix}
\mathcal{A}_{xx} & 0 & \mathcal{A}_{xz} \\ 0 & \mathcal{A}_{yy} & 0 \\ \mathcal{A}_{zx} & 0 & \mathcal{A}_{zz}
\end{pmatrix},$$
where $\mathcal{A}_{xx}=190.2(2)$ MHz, $\mathcal{A}_{yy}=120.3(2)$ MHz, $\mathcal{A}_{zz}=129.1(2)$ MHz, and  $\mathcal{A}_{xz}=\mathcal{A}_{zx}=-25.0(1)$ MHz.

Diagonalizing the total Hamiltonian, we notice that the quantization axis of the nuclear spin (in the limit $\gamma_{n} B \ll A_{ij}$) is not the same in the manifold of the $\ket{m_s=0}$ state and in the $\ket{m_s=\pm 1}$ states \citep{alvarez_local_2015}, meaning that the nuclear spin is not preserved by the electron spin flip, and gives rise to a splitting of the $\ket{m_s=0} \to \ket{m_s=\pm 1}$ transition in 4 distinct lines when the magnetic field is not aligned with the NV center \citep{jiang2018estimation}.

\section{NV-P1 mutual spin flip transitions}

Transitions corresponding to a simultaneous flip of the NV$^-$ and P1 (substitutional nitrogen) spin states, mediated by the dipolar interaction between NV$^-$ and P1 electronic spins, are commonly observed on ODMR spectra \citep{simanovskaia_sidebands_2013, kamp2018continuous, alfasi2019detection, lazda2020cross}. Here, we investigate whether the cross-relaxation processes we observe could be partly due a three-body interaction between two NV centers and one P1 center.

In order for this process to be energy conserving, the P1 transition frequency has to match the energy difference between the $\ket{m_s=0} \to \ket{m_s=-1}$ and $\ket{m_s=0} \to \ket{m_s=+1}$ transitions, giving us the equation \begin{equation}
\label{eq_P1}
\nu^i_{P1}(B)=\nu^{0 \to +1}_{NV}(B)-\nu^{0 \to +1}_{NV}(B),
\end{equation}
where $\nu^i_{P1}$ is a transition between any of the P1 spin Hamiltonian eigenstates. Note that since we are scanning the magnetic field in the crystalline [100] direction, we do not have to take into account the four classes of NV centers and P1.

The P1 spin states are defined by its electron 1/2-spin and its nuclear 1-spin, giving a manifold of 6 spin states. When considering any possible transitions between the 6 eigenstates of the P1 Hamiltonian, we have therefore a total of 15 possible transitions, including the nuclear-like transitions which have been observed through mutual spin flip with NV centers \citep{alfasi2019detection}.

%Solving eq. \ref{eq_P1} for all possible P1 transitions (see fig. \ref{fig_P1}) therefore gives us all the possible magnetic field amplitude in the [100] direction where we could observe NV-P1 mutual flip transitions. The predicted magnetic field values are 0, 3.89, 5.96, 6.58, 17.90, 28.93, 35.87, 49.44, 81.33, 83.28, 137.52, 154.20 and 246.34 G. None of these values correspond to one of the feature our scans. We then concluded that the Features we observed were not due to NV-P1 mutual flips.

Using the numerical values of the P1 Hamiltonian from \citep{lazda2020cross}, we solve Eq. \ref{eq_P1} for all possible P1 transitions (see fig. \ref{fig_P1}) therefore giving us all the possible magnetic field amplitudes in the [100] direction where we could observe NV-P1 mutual spin flip transitions. The predicted magnetic field values are 0, 3.89, 5.96, 6.58, 17.90, 28.93, 35.87, 49.44, 81.33, 83.28, 137.52, 154.20 and 246.34 G. None of these values correspond to any feature in our scans. We then concluded that the features we observed were not due to NV-P1 mutual flips.

\begin{figure}
\includegraphics[width=\linewidth]{Transis_P1}
\caption{Cross-relaxation condition for mutual flip between NV and P1 centers. The orange plain lines correspond to the NV center $\ket{m_s=0} \to \ket{m_s=-1}$ and $\ket{m_s=0} \to \ket{m_s=+1}$ transition frequencies as function of a magnetic field along the [100] crystalline axis, the blue dashed lines correspond to the frequencies of the 15 theoretical transitions of the P1 centers, added to the frequency of the NV $\ket{m_s=0} \to \ket{m_s=-1}$ transition. The green circles correspond to the particular magnetic fields where Eq. \ref{eq_P1} is verified, meaning where the energy of the P1 transition matches the energy difference between the two NV transitions.}\label{fig_P1}
\end{figure}
% Bibliography
\bibliographystyle{plain}

\section{Scan in the [111] direction}
\begin{figure}
\includegraphics[width=\linewidth]{scan_111_PL}
\caption{\textbf{a)} Plain blue line : Photoluminescence as a function of the magnetic field amplitude in the [111] crystalline direction. Dashed orange line : 4th order polynomial fit of the decreasing envelope. \textbf{b)} Subtraction of the previous signal by the envelope fit. The vertical dotted orange lines correspond to the expected cross-relaxations with VH$^-$ and the vertical dashed green lines to the expected cross-relaxations with WAR1.}\label{scan_PL}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{Transis_111_VHWAR}
\caption{Predicted transition frequencies of NV$^-$ centers (plain lines), VH$^-$ (dashed lines) and WAR1 (dash-dotted lines) when the magnetic field is scanned in the [111] direction. The magnetic field values for which  a transition of the NV center crosses one of VH$^-$ or WAR1 are also shown in Fig \ref{scan_PL}.}\label{transis_VHWAR}
\end{figure}

\begin{figure}

\includegraphics[width=\linewidth]{Transis_111_13C}
\caption{Predicted transitions frequencies for NV centers (plain lines) and NV-$^{13}$C complex when the magnetic field is scanned in the [111] direction. Vertical lines represent the various degeneracy conditions.}\label{transis_13C}
\end{figure}

We have performed similar magnetic field scan as presented in Fig.3 of the main text, but scanning the magnetic field in the crystalline [111] direction instead of the [100], meaning that this time one class of NV centers is aligned with the magnetic field while the three others are misaligned and degenerate. The results in term of cross-relaxations generally means 4 times as many degeneracy conditions as the [100] case, since the two spins involved in the cross-relaxation process now have two possible orientations, and therefore two separate transitions.

It is then significantly harder to find cross-relaxations signatures, since they are both more numerous and with a smaller contrast than in the [100] case. The experimental results are shown in Fig. \ref{scan_PL}. Similarly to Fig. 3 of the main text, we have subtracted the envelope associated with the transverse field in order to better see the fine features of the scan. The predicted cross-relaxation conditions between NV centers and VH$^-$ or WAR1 are presented in fig. \ref{transis_VHWAR} and shown in Fig. \ref{scan_PL}. While the predicted values do not exactly match the experimental results, it seems like every predicted transition falls in the vicinity of an experimental line. Importantly, it should be noted that, unlike Fig.3 of the main text, the magnetic field was not calibrated with the lengthy NV ODMR at every field value in this case. Instead we have simply applied a homotethy on the current intensity in the electromagnet to match the field values at the maxima of the scan. This means that the magnetic field of the points of low field value in particular are more susceptible to the non-linearities due to the electromagnet hysteresis cycle.

Finally we have plotted in Fig.\ref{transis_13C} the expected cross-relaxations between NV and NV-$^{13}$C complex. It seems that these transitions are too diluted to appear in our experimental scan. The peak at $\sim$37 G in fig.\ref{scan_PL} could come from the resonance between NV and NV-$^{13}$C since they seem particularly dense at this particular field values.
It is indeed  the only peak that is not matching any of the NV-VH$^-$ or NV-WAR1 resonance.

\bibliography{Biblio_CVD-CR}

\end{document}

### 2021-02-23 Mes papiers/CR-CVD/SI/CVD_para_full_arxiv.tex
tutu

--- 2021-02-23 Mes papiers/CR-CVD/CVD_para.tex
@@
 The authors would also like to thank SIRTEQ for funding. GH also acknowledges funding by the French National Research Agency (ANR) through the T-ERC project QUOVADIS.
+
+\section*{Appendix}
+\subsection{Photoluminescence as a function of magnetic field in a HPHT sample}
+
+The NV$^-$ photoluminescence as a function of the magnetic field amplitude that is presented in the Fig. 3 of the main text was realized on a CVD-grown sample. 
+We also performed similar scans on several samples grown via the High-Pressure-High-Temperature (HPHT) method. 
+Fig. \ref{fig1SI} shows a scan realized on HPHT samples that have been irradiated in order to increase the NV yield. 
+The sample we use in Fig. \ref{fig1SI} is a Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range.
+The PL feature that can be observed at 20G was also observed in the CVD-grown samples.
+We attributed it to the resonant dipolar interaction between a standard polarized NV center and one NV strongly coupled to the nuclear spin of a $^{13}$C atom sitting one shell away.
+Notably, no other features could be detected at larger magnetic field values.
+We realized such scans on many different highly-doped HPHT samples and, systematically, no CR features are observed at 50 and 122 G.
+
+ \begin{figure}[htbp]
+\centering
+{\includegraphics[width=\linewidth]{sumi4_13C}}
+\caption{NV$^-$ photoluminescence rate while scanning the magnetic field along the [100] crystalline direction. Here a HPHT sample was employed.}
+\label{fig1SI}
+\end{figure}
+
+\subsection{NV-$^{13}$C Hamiltonian}
+
+The full spin Hamiltonian of the NV-$^{13}$C complex can be written as follows : 
+\begin{equation*}
+\mathcal{H}=\mathcal{H}_{NV}+\mathcal{H}_{^{13}C}+\mathcal{H}_{HF},
+\end{equation*}
+where $\mathcal{H}_{NV}$ is the NV$^-$ spin Hamiltonian.
+For one NV orientation, with quantization axis $z$, it reads 
+\begin{equation}\hat{H}_{\rm NV}=\hbar D \hat{S}_z^2+ \hbar \gamma_e \bf B  \cdot \bf\hat S.
+\end{equation}
+$D$ is the zero-field-splitting and $\bf\hat S$ is the electronic spin operator.
+$\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian : $\mathcal{H}_{HF}= \hat{\mathbf{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\mathbf{I}}_C$.
+
+In the case of a first shell $^{13}$C, the hyper-fine tensor $\mathcal{A}$ can be written as \citep{simanovskaia_sidebands_2013} : $$ \mathcal{A} = \begin{pmatrix}
+\mathcal{A}_{xx} & 0 & \mathcal{A}_{xz} \\ 0 & \mathcal{A}_{yy} & 0 \\ \mathcal{A}_{zx} & 0 & \mathcal{A}_{zz}
+\end{pmatrix},$$
+where $\mathcal{A}_{xx}=190.2(2)$ MHz, $\mathcal{A}_{yy}=120.3(2)$ MHz, $\mathcal{A}_{zz}=129.1(2)$ MHz, and  $\mathcal{A}_{xz}=\mathcal{A}_{zx}=-25.0(1)$ MHz.
+
+Diagonalizing the total Hamiltonian, we notice that the quantization axis of the nuclear spin (in the limit $\gamma_{n} B \ll A_{ij}$) is not the same in the manifold of the $\ket{m_s=0}$ state and in the $\ket{m_s=\pm 1}$ states \citep{alvarez_local_2015}, meaning that the nuclear spin is not preserved by the electron spin flip, and gives rise to a splitting of the $\ket{m_s=0} \to \ket{m_s=\pm 1}$ transition in 4 distinct lines when the magnetic field is not aligned with the NV center \citep{jiang2018estimation}.
+
+\subsection{NV-P1 mutual spin flip transitions}
+
+Transitions corresponding to a simultaneous flip of the NV$^-$ and P1 (substitutional nitrogen) spin states, mediated by the dipolar interaction between NV$^-$ and P1 electronic spins, are commonly observed on ODMR spectra \citep{simanovskaia_sidebands_2013, kamp2018continuous, alfasi2019detection, lazda2020cross}.
+Here, we investigate whether the cross-relaxation processes we observe could be partly due a three-body interaction between two NV centers and one P1 center.
+
+In order for this process to be energy conserving, the P1 transition frequency has to match the energy difference between the $\ket{m_s=0} \to \ket{m_s=-1}$ and $\ket{m_s=0} \to \ket{m_s=+1}$ transitions, giving us the equation \begin{equation}
+\label{eq_P1}
+\nu^i_{P1}(B)=\nu^{0 \to +1}_{NV}(B)-\nu^{0 \to +1}_{NV}(B),
+\end{equation}
+where $\nu^i_{P1}$ is a transition between any of the P1 spin Hamiltonian eigenstates.
+Note that since we are scanning the magnetic field in the crystalline [100] direction, we do not have to take into account the four classes of NV centers and P1.
+
+The P1 spin states are defined by its electron 1/2-spin and its nuclear 1-spin, giving a manifold of 6 spin states.
+When considering any possible transitions between the 6 eigenstates of the P1 Hamiltonian, we have therefore a total of 15 possible transitions, including the nuclear-like transitions which have been observed through mutual spin flip with NV centers \citep{alfasi2019detection}.
+
+%Solving eq.
+\ref{eq_P1} for all possible P1 transitions (see fig.
+\ref{fig_P1}) therefore gives us all the possible magnetic field amplitude in the [100] direction where we could observe NV-P1 mutual flip transitions.
+The predicted magnetic field values are 0, 3.89, 5.96, 6.58, 17.90, 28.93, 35.87, 49.44, 81.33, 83.28, 137.52, 154.20 and 246.34 G.
+None of these values correspond to one of the feature our scans.
+We then concluded that the Features we observed were not due to NV-P1 mutual flips.
+
+Using the numerical values of the P1 Hamiltonian from \citep{lazda2020cross}, we solve Eq.
+\ref{eq_P1} for all possible P1 transitions (see fig.
+\ref{fig_P1}) therefore giving us all the possible magnetic field amplitudes in the [100] direction where we could observe NV-P1 mutual spin flip transitions.
+The predicted magnetic field values are 0, 3.89, 5.96, 6.58, 17.90, 28.93, 35.87, 49.44, 81.33, 83.28, 137.52, 154.20 and 246.34 G.
+None of these values correspond to any feature in our scans.
+We then concluded that the features we observed were not due to NV-P1 mutual flips.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{Transis_P1}
+\caption{Cross-relaxation condition for mutual flip between NV and P1 centers.
+The orange plain lines correspond to the NV center $\ket{m_s=0} \to \ket{m_s=-1}$ and $\ket{m_s=0} \to \ket{m_s=+1}$ transition frequencies as function of a magnetic field along the [100] crystalline axis, the blue dashed lines correspond to the frequencies of the 15 theoretical transitions of the P1 centers, added to the frequency of the NV $\ket{m_s=0} \to \ket{m_s=-1}$ transition.
+The green circles correspond to the particular magnetic fields where Eq.
+\ref{eq_P1} is verified, meaning where the energy of the P1 transition matches the energy difference between the two NV transitions.}\label{fig_P1}
+\end{figure}
+% Bibliography
+\bibliographystyle{plain}
+
+\subsection{Magnetic field scan along the [111] direction}
+\begin{figure}
+\includegraphics[width=\linewidth]{scan_111_PL}
+\caption{\textbf{a)} Plain blue line : Photoluminescence as a function of the magnetic field amplitude in the [111] crystalline direction.
+Dashed orange line : 4th order polynomial fit of the decreasing envelope.
+\textbf{b)} Subtraction of the previous signal by the envelope fit.
+The vertical dotted orange lines correspond to the expected cross-relaxations with VH$^-$ and the vertical dashed green lines to the expected cross-relaxations with WAR1.}\label{scan_PL}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{Transis_111_VHWAR}
+\caption{Predicted transition frequencies of NV$^-$ centers (plain lines), VH$^-$ (dashed lines) and WAR1 (dash-dotted lines) when the magnetic field is scanned in the [111] direction.
+The magnetic field values for which  a transition of the NV center crosses one of VH$^-$ or WAR1 are also shown in Fig \ref{scan_PL}.}\label{transis_VHWAR}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics[width=\linewidth]{Transis_111_13C}
+\caption{Predicted transitions frequencies for NV centers (plain lines) and NV-$^{13}$C complex when the magnetic field is scanned in the [111] direction.
+Vertical lines represent the various degeneracy conditions.}\label{transis_13C}
+\end{figure}
+
+We have performed similar magnetic field scan as presented in Fig.3 of the main text, but scanning the magnetic field in the crystalline [111] direction instead of the [100], meaning that this time one class of NV centers is aligned with the magnetic field while the three others are misaligned and degenerate.
+The results in term of cross-relaxations generally means 4 times as many degeneracy conditions as the [100] case, since the two spins involved in the cross-relaxation process now have two possible orientations, and therefore two separate transitions.
+
+It is then significantly harder to find cross-relaxations signatures, since they are both more numerous and with a smaller contrast than in the [100] case.
+The experimental results are shown in Fig.
+\ref{scan_PL}.
+Similarly to Fig.
+3 of the main text, we have subtracted the envelope associated with the transverse field in order to better see the fine features of the scan.
+The predicted cross-relaxation conditions between NV centers and VH$^-$ or WAR1 are presented in fig.
+\ref{transis_VHWAR} and shown in Fig.
+\ref{scan_PL}.
+While the predicted values do not exactly match the experimental results, it seems like every predicted transition falls in the vicinity of an experimental line.
+Importantly, it should be noted that, unlike Fig.3 of the main text, the magnetic field was not calibrated with the lengthy NV ODMR at every field value in this case.
+Instead we have simply applied a homotethy on the current intensity in the electromagnet to match the field values at the maxima of the scan.
+This means that the magnetic field of the points of low field value in particular are more susceptible to the non-linearities due to the electromagnet hysteresis cycle.
+
+Finally we have plotted in Fig.\ref{transis_13C} the expected cross-relaxations between NV and NV-$^{13}$C complex.
+It seems that these transitions are too diluted to appear in our experimental scan.
+The peak at $\sim$37 G in fig.\ref{scan_PL} could come from the resonance between NV and NV-$^{13}$C since they seem particularly dense at this particular field values.
+It is indeed  the only peak that is not matching any of the NV-VH$^-$ or NV-WAR1 resonance.
 

### 2021-02-26 Mes papiers/CR-CVD/Reviewer_answer.tex
tutu

@@
 \item We have added information on the symmetry of the different defects with respect to the [100] axis to justify that we only take into account two transitions per defect in our model (except for the 13C-NV defect which is described in the supplementary material)
-\item We added a discussion on the Gaussian vs Lorentzian line-shape in the description of fig.3.
+\item We added a discussion on the Gaussian vs Lorentzian line-shape in sec.V of the supplementary material.
 Given that the approximate shorter distance between our spins is more than 10 nm, the coupling strength between our spins would be less than 100 KHz.
@@
 \item We added a paragraph at the end of the manuscript where we give a tentative estimate of the VH- and WAR1 concentration based on our measurement.
-\item La c'est plus tendu de répondre. Alex ?
+\item It would be very relevant to analyze a range of diamond samples grown under different conditions (methane concentration, temperature, growth rate, plasma power density etc.) with this new optical technique, however we currently only have access to a single CVD sample with a high enough concentration in NV centers to observe cross-relaxation effects.
+We have already found that NV concentration strongly varies with the growth temperature (see reference: A.
+Tallaire et al.
+Temperature dependent creation of nitrogen-vacancy centers in single crystal CVD diamond layers, Diam.
+\& Relat.
+Mat.
+51 (2015) 55–60.) and we suspect that it may also influence the incorporation of other defects although we have not confirmed this so far.
 \end{enumerate}
@@
 It is in fact possible to extract an estimate of the concentrations with the area of the peaks (which is probably more relevant than the amplitude of the peaks in this case), but the width of the peak is mostly due to inhomogeneous broadening and not to interactions.
-\item On a pas les infos pour les adamas, je sais pas si on peut en avoir pour les sumitumo ?
-\item Demander à Alex. Je pense qu'il faut juste justifier le fait qu'il y ait plus d'hydrogène dans les CVD que dans les HPHT.
+\item We added fabrication details for both the CVD and the HPHT sample in Sec.VI of the supplementary material.
+As for why the VH$^-$ and WAR1 defects were only found in the CVD sample, we think that although the irradiation procedure of the two samples was not perfectly identical, it would not be the reason for the observed difference in VH- and WAR1 which is mainly due to the initial starting defects in the material (HPHT vs CVD).
+CVD synthesis is performed in a hydrogen-rich environment and the presence of H contamination in the samples is very common.
+To the contrary HPHT diamonds do not generally contain significant amounts of H but the control of other type of impurities may be more problematic such N in substitutional and aggregated forms as well as metal impurities from the melted bath.
+\item As previously explained hydrogen is not normally found in HPHT-grown material due to the growth process that does not involve this gas.
+On the contrary, CVD growth requires a high fraction of hydrogen in the gas phase (95\% or more) which is dissociated in a micro-wave plasma.
+A small fraction of these reactive H species can get incorporated into the growing crystal depending on the growth conditions.
+This would explain why VH- (and also NVH) are commonly found in CVD-grown material but not in HPHT diamonds.
+As for WAR1, there is no agreement in the literature on its precise nature, however on the light of those results we believe that it is likely that it involves H impurities.
 \item We have added a discussion of the concentration of 13C-NV pairs in the last paragraph when quantifying the concentration of our species.

### 2021-02-26 Mes papiers/CR-CVD/CVD_para.tex
tutu

@@
 \usepackage{natbib}
+ \usepackage{soulutf8}
 \usepackage{physics}
+
+\usepackage{xcolor, soul}
+\sethlcolor{green}
+ \sethlcolor{green}
+    \newcommand{\hlc}[2][yellow]{{\sethlcolor{#1} \hl{#2}}}
 
@@
 
+%Document
+%Voici un \hl{texte surligné en vert}
 The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing \cite{Rondin_2014} and quantum information science  \cite{DOHERTY20131}.
@@
 The sample we use in this study  correspond to the pink irradiated sample in \cite{TALLAIRE2020421}. It was grown using CVD with the addition of 500ppm of N$_2$0 to the H2/CH4 (96/4) gas phase.
-Then high energy (10MeV) electron irradiation at a fluence of $2\times18$ cm$^{-2}$ and at a temperature of 900 degrees celsius was realised, giving a final concentration of NV$^-$ of about 4.6~ppm for an initial concentration of nitrogen in the sample of about 25~ppm.
+Then high energy (10MeV) electron irradiation at a fluence of $2\times18$ cm$^{-2}$ and at a temperature of 900 degrees celsius was realized,  giving a final concentration of NV$^-$ of about 4.6~ppm for an initial concentration of nitrogen in the sample of about 25~ppm.
 It was shown that in this sample the $T_2^*$ was not degraded after annealing, yet the NV density was large enough to enable concentration dependent longitudinal relaxation \cite{TALLAIRE2020421, jarmola_longitudinal_2015}.
@@
 This means that a B field scan along this direction cannot make the transition energies of the different classes of NV centers cross.
-Under arbitrary angles, this could otherwise result in several low field PL features \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the supplementary materials for data showing scans along the [111] direction).
-The downside of this choice of the [100] direction is that at large amplitudes, the transverse component of the magnetic field depolarizes all classes of NVs, thus limiting the magnetic field range that can be employed and therefore limiting this detection process to paramagnetic defects with zero-field splittings close to the NV center's one.
+Under arbitrary angles, this could otherwise result in several low field PL features \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the Supplementary Materials for data showing scans along the [111] direction).
+The downside of this choice of the [100] direction is that at large amplitudes, the transverse component of the magnetic field depolarizes all classes of NVs, thus limiting the magnetic field range that can be employed and restricting this detection method to paramagnetic defects with zero-field splittings close to the NV center's.
 
@@
 
-In order to let the three salient features detach better from the spectrum, we fitted a 4th-order polynomial to the PL evolution, without the spectral bumps, and subtract it from the data.
-We chose to use a 4th-order polynomial since numerical solving of the transverse-field dependent spin depolarization could not properly fit our results.
-The order of the polynomial fit was chosen to be 4 since higher order did not yield significantly better envelope fitting, and lower order were noticeably worse.
-Thanks to the subtraction, the width and contrast of each peak can now be compared and their position can be determined with greater accuracy.
-The second and third peaks were better fitted with Gaussian line shapes than Lorentzian ones, which is coherent with the fact that the inhomogeneous broadening of the various spin species dominate the coupling strength between the spins.
+To let the three salient features detach better from the spectrum, we fitted a 4th-order polynomial to the PL evolution, without the spectral bumps, and subtract it from the data.
+Further details on the fitting choices are provided in Sec.V of the Supplementary material.
 
@@
 \ref{energy-levels} shows the frequencies of the NV$^-$, the VH$^-$ and WAR1 spin transitions as a function of the magnetic field amplitude along the $[100]$ direction.
-Since the NV$^-$ and VH$^-$ are $C_{3v}$ defects, and WAR1 is a pseudo-$C_{3v}$ defect \citep{cruddace2007magnetic}, all possible orientation are degenerate for this magnetic field orientation, giving only two lines corresponding to the $\ket{0} \to \ket{+1}$ and $\ket{0} \to \ket{-1}$ transitions.
+Since the NV$^-$ and VH$^-$ are $C_{3v}$ defects, and the WAR1 is a pseudo-$C_{3v}$ defect \citep{cruddace2007magnetic}, all possible orientation are degenerate for this magnetic field orientation, giving only two lines corresponding to the $\ket{0} \to \ket{+1}$ and $\ket{0} \to \ket{-1}$ transitions.
 The points were the NV levels cross the other defects can give rise to cross-relaxation.
@@
 The resulting pair could then be coupled resonantly to a nearby polarized NV.
-The obtained eigenfrequencies of the dipolar coupled P$_1$-NV pair that we extracted were all inconsistent with the spectral positions of our observed peak within our spectral resolution error ($\approx$1G) (see Section 3 of the supplementary materials).
+The obtained eigenfrequencies of the dipolar coupled P$_1$-NV pair that we extracted were all inconsistent with the spectral positions of our observed peak within our spectral resolution error ($\approx$1G) (see Section 3 of the Supplementary Materials).
 This leads us to consider instead the nuclear spin of the $^{13}$C atoms as the most likely candidate.
@@
 In contrast, higher order shells have a maximum hyperfine coupling of 15 MHz \cite{smeltzer201113c} making them hard to distinguish from the inhomogeneous broadening.
-The modeling of this interaction is presented in the section 2 of the supplementary materials.
+The modeling of this interaction is presented in the section 2 of the Supplementary Materials.
 The dotted lines in Fig.
@@
 
-Because the sample we use has a natural 1.1\% abundance of $^{13}$C, we can actually use the first peak as a way to calibrate the concentration of VH$^-$ and WAR1 in our sample.
-Because there are three possible sites for a first-shell $^{13}$C-NV pair, we can estimate that the concentration of $^{13}$C-NV pair is about 3.3\% of the total concentration of NV centers, i.e.
+Because the sample we use has a very well determined concentration of $^{13}$C (natural 1.1\% abundance), we can use the first CR peak to calibrate the concentration of VH$^-$ and WAR1 in our sample.
+There are three possible sites for a first-shell $^{13}$C-NV pair, so the concentration of $^{13}$C-NV pair is about 3.3\% that of the NV centers, i.e.
 $\sim$135~ppb.
-We can then compare the area of the second and third peak to the first one to have a first order estimation of the concentration of VH$^-$ and WAR1.
-We find [VH$^-$]$\approx$195~ppb and [WAR1]$\approx$124~ppb.
-We decided to compare the area of the peaks instead of the amplitudes since the width of the peak is mostly due to inhomogeneous spread of the spins, and not to the spin-spin interaction strength.
-It should be noted that a severe drawback of this method is the loss of contrast for higher magnetic field amplitude, meaning in this case that the WAR1 concentration is probably undervalued.
+Comparing the area of the second and third peak to the first one, we find [VH$^-$]$\approx$195~ppb and [WAR1]$\approx$124~ppb.
+Using areas is here more relevant than using amplitudes because inhomogeneous spread dominates over the spin-spin interaction.
+Note that the  WAR1 concentration is probably undervalued due to state mixing in the NV centers at the NV-WAR1 CR.
 
@@
   (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
+
 \end{thebibliography}%

### 2021-02-26 Mes papiers/CR-CVD/SI/CVD_para_full_arxiv.tex
tutu

@@
 \section*{Appendix}
-\subsection{Photoluminescence as a function of magnetic field in a HPHT sample}
+\subsection{Photoluminescence as a function of the magnetic field for a HPHT sample}
 
@@
 \centering
-{\includegraphics[width=\linewidth]{sumi4_13C}}
+{\includegraphics[width=\linewidth]{sumi4_13C.eps}}
 \caption{NV$^-$ photoluminescence rate while scanning the magnetic field along the [100] crystalline direction. Here a HPHT sample was employed.}
@@
 
-\subsection{Magnetic field scan along the [111] direction}
+\subsection{Scan in the [111] direction}
 \begin{figure}
@@
 
-%merlin.mbs apsrev4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
-%Control: key (0)
-%Control: author (8) initials jnrlst
-%Control: editor formatted (1) identically to author
-%Control: production of article title (-1) disabled
-%Control: page (0) single
-%Control: year (1) truncated
-%Control: production of eprint (0) enabled
-\begin{thebibliography}{27}%
-\makeatletter
-\providecommand \@ifxundefined [1]{%
- \@ifx{#1\undefined}
-}%
-\providecommand \@ifnum [1]{%
- \ifnum #1\expandafter \@firstoftwo
- \else \expandafter \@secondoftwo
- \fi
-}%
-\providecommand \@ifx [1]{%
- \ifx #1\expandafter \@firstoftwo
- \else \expandafter \@secondoftwo
- \fi
-}%
-\providecommand \natexlab [1]{#1}%
-\providecommand \enquote  [1]{``#1''}%
-\providecommand \bibnamefont  [1]{#1}%
-\providecommand \bibfnamefont [1]{#1}%
-\providecommand \citenamefont [1]{#1}%
-\providecommand \href@noop [0]{\@secondoftwo}%
-\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
-\providecommand \@href[1]{\@@startlink{#1}\@@href}%
-\providecommand \@@href[1]{\endgroup#1\@@endlink}%
-\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
-  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
-\providecommand \@@startlink[1]{}%
-\providecommand \@@endlink[0]{}%
-\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
-\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
-\providecommand \urlprefix  [0]{URL }%
-\providecommand \Eprint [0]{\href }%
-\providecommand \doibase [0]{http://dx.doi.org/}%
-\providecommand \selectlanguage [0]{\@gobble}%
-\providecommand \bibinfo  [0]{\@secondoftwo}%
-\providecommand \bibfield  [0]{\@secondoftwo}%
-\providecommand \translation [1]{[#1]}%
-\providecommand \BibitemOpen [0]{}%
-\providecommand \bibitemStop [0]{}%
-\providecommand \bibitemNoStop [0]{.\EOS\space}%
-\providecommand \EOS [0]{\spacefactor3000\relax}%
-\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
-\let\auto@bib@innerbib\@empty
-%</preamble>
-\bibitem [{\citenamefont {Rondin}\ \emph {et~al.}(2014)\citenamefont {Rondin},
-  \citenamefont {Tetienne}, \citenamefont {Hingant}, \citenamefont {Roch},
-  \citenamefont {Maletinsky},\ and\ \citenamefont {Jacques}}]{Rondin_2014}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
-  {Rondin}}, \bibinfo {author} {\bibfnamefont {J.-P.}\ \bibnamefont
-  {Tetienne}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Hingant}},
-  \bibinfo {author} {\bibfnamefont {J.-F.}\ \bibnamefont {Roch}}, \bibinfo
-  {author} {\bibfnamefont {P.}~\bibnamefont {Maletinsky}}, \ and\ \bibinfo
-  {author} {\bibfnamefont {V.}~\bibnamefont {Jacques}},\ }\href {\doibase
-  10.1088/0034-4885/77/5/056503} {\bibfield  {journal} {\bibinfo  {journal}
-  {Reports on Progress in Physics}\ }\textbf {\bibinfo {volume} {77}},\
-  \bibinfo {pages} {056503} (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Doherty}\ \emph {et~al.}(2013)\citenamefont
-  {Doherty}, \citenamefont {Manson}, \citenamefont {Delaney}, \citenamefont
-  {Jelezko}, \citenamefont {Wrachtrup},\ and\ \citenamefont
-  {Hollenberg}}]{DOHERTY20131}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~W.}\ \bibnamefont
-  {Doherty}}, \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont {Manson}},
-  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Delaney}}, \bibinfo
-  {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Wrachtrup}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {L.~C.}\ \bibnamefont {Hollenberg}},\ }\href {\doibase
-  https://doi.org/10.1016/j.physrep.2013.02.001} {\bibfield  {journal}
-  {\bibinfo  {journal} {Physics Reports}\ }\textbf {\bibinfo {volume} {528}},\
-  \bibinfo {pages} {1 } (\bibinfo {year} {2013})},\ \bibinfo {note} {the
-  nitrogen-vacancy colour centre in diamond}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Achard}\ \emph {et~al.}(2020)\citenamefont {Achard},
-  \citenamefont {Jacques},\ and\ \citenamefont {Tallaire}}]{Achard}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
-  {Achard}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Jacques}}, \
-  and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Tallaire}},\ }\href
-  {\doibase 10.1088/1361-6463/ab81d1} {\bibfield  {journal} {\bibinfo
-  {journal} {Journal of Physics D: Applied Physics}\ }\textbf {\bibinfo
-  {volume} {53}},\ \bibinfo {pages} {313001} (\bibinfo {year}
-  {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Newton}(2007)}]{newton_epr_2007}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont
-  {Newton}},\ }\href {\doibase 10.1002/chin.200747215} {\bibfield  {journal}
-  {\bibinfo  {journal} {ChemInform}\ }\textbf {\bibinfo {volume} {38}}
-  (\bibinfo {year} {2007}),\ 10.1002/chin.200747215}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Glover}\ \emph {et~al.}(2004)\citenamefont {Glover},
-  \citenamefont {Newton}, \citenamefont {Martineau}, \citenamefont {Quinn},\
-  and\ \citenamefont {Twitchen}}]{glover_hydrogen_2004}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Glover}}, \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont {Newton}},
-  \bibinfo {author} {\bibfnamefont {P.~M.}\ \bibnamefont {Martineau}}, \bibinfo
-  {author} {\bibfnamefont {S.}~\bibnamefont {Quinn}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {D.~J.}\ \bibnamefont {Twitchen}},\ }\href {\doibase
-  10.1103/PhysRevLett.92.135502} {\bibfield  {journal} {\bibinfo  {journal}
-  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {92}},\ \bibinfo {pages}
-  {135502} (\bibinfo {year} {2004})},\ \bibinfo {note} {publisher: American
-  Physical Society}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Glover}\ \emph {et~al.}(2003)\citenamefont {Glover},
-  \citenamefont {Newton}, \citenamefont {Martineau}, \citenamefont {Twitchen},\
-  and\ \citenamefont {Baker}}]{glover_hydrogen_2003}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Glover}}, \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont {Newton}},
-  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Martineau}}, \bibinfo
-  {author} {\bibfnamefont {D.~J.}\ \bibnamefont {Twitchen}}, \ and\ \bibinfo
-  {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Baker}},\ }\href {\doibase
-  10.1103/PhysRevLett.90.185507} {\bibfield  {journal} {\bibinfo  {journal}
-  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {90}},\ \bibinfo {pages}
-  {185507} (\bibinfo {year} {2003})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Yao}\ \emph {et~al.}(2012)\citenamefont {Yao},
-  \citenamefont {Jiang}, \citenamefont {Gorshkov}, \citenamefont {Maurer},
-  \citenamefont {Giedke}, \citenamefont {Cirac},\ and\ \citenamefont
-  {Lukin}}]{Yao}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~Y.}\ \bibnamefont
-  {Yao}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Jiang}}, \bibinfo
-  {author} {\bibfnamefont {A.~V.}\ \bibnamefont {Gorshkov}}, \bibinfo {author}
-  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
-  {\bibfnamefont {G.}~\bibnamefont {Giedke}}, \bibinfo {author} {\bibfnamefont
-  {J.~I.}\ \bibnamefont {Cirac}}, \ and\ \bibinfo {author} {\bibfnamefont
-  {M.~D.}\ \bibnamefont {Lukin}},\ }\href@noop {} {\bibfield  {journal}
-  {\bibinfo  {journal} {Nature Communications}\ }\textbf {\bibinfo {volume}
-  {3}},\ \bibinfo {pages} {800} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Cruddace}(2007)}]{cruddace2007magnetic}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Cruddace}},\ }\emph {\bibinfo {title} {Magnetic resonance and optical
-  studies of point defects in single crystal CVD diamond}},\ \href@noop {}
-  {Ph.D. thesis},\ \bibinfo  {school} {University of Warwick} (\bibinfo {year}
-  {2007})\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Epstein}\ \emph {et~al.}(2005)\citenamefont
-  {Epstein}, \citenamefont {Mendoza}, \citenamefont {Kato},\ and\ \citenamefont
-  {Awschalom}}]{Epstein}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~J.}\ \bibnamefont
-  {Epstein}}, \bibinfo {author} {\bibfnamefont {F.~M.}\ \bibnamefont
-  {Mendoza}}, \bibinfo {author} {\bibfnamefont {Y.~K.}\ \bibnamefont {Kato}}, \
-  and\ \bibinfo {author} {\bibfnamefont {D.~D.}\ \bibnamefont {Awschalom}},\
-  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Nature Physics}\
-  }\textbf {\bibinfo {volume} {1}},\ \bibinfo {pages} {94} (\bibinfo {year}
-  {2005})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Wang}\ \emph {et~al.}(2014)\citenamefont {Wang},
-  \citenamefont {Shin}, \citenamefont {Seltzer}, \citenamefont {Avalos},
-  \citenamefont {Pines},\ and\ \citenamefont {Bajaj}}]{Wang}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.-J.}\ \bibnamefont
-  {Wang}}, \bibinfo {author} {\bibfnamefont {C.~S.}\ \bibnamefont {Shin}},
-  \bibinfo {author} {\bibfnamefont {S.~J.}\ \bibnamefont {Seltzer}}, \bibinfo
-  {author} {\bibfnamefont {C.~E.}\ \bibnamefont {Avalos}}, \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Pines}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {V.~S.}\ \bibnamefont {Bajaj}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Nature Communications}\ }\textbf {\bibinfo
-  {volume} {5}},\ \bibinfo {pages} {4135} (\bibinfo {year} {2014})}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Armstrong}\ \emph {et~al.}(2010)\citenamefont
-  {Armstrong}, \citenamefont {Rogers}, \citenamefont {McMurtrie},\ and\
-  \citenamefont {Manson}}]{armstrong_nvnv_2010}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Armstrong}}, \bibinfo {author} {\bibfnamefont {L.~J.}\ \bibnamefont
-  {Rogers}}, \bibinfo {author} {\bibfnamefont {R.~L.}\ \bibnamefont
-  {McMurtrie}}, \ and\ \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont
-  {Manson}},\ }\href {\doibase 10.1016/j.phpro.2010.01.223} {\bibfield
-  {journal} {\bibinfo  {journal} {Physics Procedia}\ }\textbf {\bibinfo
-  {volume} {3}},\ \bibinfo {pages} {1569} (\bibinfo {year} {2010})},\ \bibinfo
-  {note} {number: 4}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Hall}\ \emph {et~al.}(2016)\citenamefont {Hall},
-  \citenamefont {Kehayias}, \citenamefont {Simpson}, \citenamefont {Jarmola},
-  \citenamefont {Stacey}, \citenamefont {Budker},\ and\ \citenamefont
-  {Hollenberg}}]{Hall}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.~T.}\ \bibnamefont
-  {Hall}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Kehayias}},
-  \bibinfo {author} {\bibfnamefont {D.~A.}\ \bibnamefont {Simpson}}, \bibinfo
-  {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}}, \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Stacey}}, \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Budker}}, \ and\ \bibinfo {author} {\bibfnamefont
-  {L.~C.~L.}\ \bibnamefont {Hollenberg}},\ }\href@noop {} {\bibfield  {journal}
-  {\bibinfo  {journal} {Nature Communications}\ }\textbf {\bibinfo {volume}
-  {7}},\ \bibinfo {pages} {10211} (\bibinfo {year} {2016})}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Wickenbrock}\ \emph {et~al.}(2016)\citenamefont
-  {Wickenbrock}, \citenamefont {Zheng}, \citenamefont {Bougas}, \citenamefont
-  {Leefer}, \citenamefont {Afach}, \citenamefont {Jarmola}, \citenamefont
-  {Acosta},\ and\ \citenamefont {Budker}}]{Wickenbrock}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Wickenbrock}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Zheng}},
-  \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Bougas}}, \bibinfo
-  {author} {\bibfnamefont {N.}~\bibnamefont {Leefer}}, \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Afach}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Jarmola}}, \bibinfo {author} {\bibfnamefont {V.~M.}\
-  \bibnamefont {Acosta}}, \ and\ \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Budker}},\ }\href {\doibase 10.1063/1.4960171} {\bibfield
-  {journal} {\bibinfo  {journal} {Applied Physics Letters}\ }\textbf {\bibinfo
-  {volume} {109}},\ \bibinfo {pages} {053505} (\bibinfo {year} {2016})},\
-  \Eprint {http://arxiv.org/abs/https://doi.org/10.1063/1.4960171}
-  {https://doi.org/10.1063/1.4960171} \BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Alfasi}\ \emph {et~al.}(2019)\citenamefont {Alfasi},
-  \citenamefont {Masis}, \citenamefont {Shtempluck},\ and\ \citenamefont
-  {Buks}}]{Alfasi}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
-  {Alfasi}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Masis}},
-  \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Shtempluck}}, \ and\
-  \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {Buks}},\ }\href {\doibase
-  10.1103/PhysRevB.99.214111} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
-  Rev. B}\ }\textbf {\bibinfo {volume} {99}},\ \bibinfo {pages} {214111}
-  (\bibinfo {year} {2019})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {{Edmonds}}\ \emph {et~al.}(2020)\citenamefont
-  {{Edmonds}}, \citenamefont {{Hart}}, \citenamefont {{Turner}}, \citenamefont
-  {{Colard}}, \citenamefont {{Schloss}}, \citenamefont {{Olsson}},
-  \citenamefont {{Trubko}}, \citenamefont {{Markham}}, \citenamefont
-  {{Rathmill}}, \citenamefont {{Horne-Smith}}, \citenamefont {{Lew}},
-  \citenamefont {{Manickam}}, \citenamefont {{Bruce}}, \citenamefont {{Kaup}},
-  \citenamefont {{Russo}}, \citenamefont {{DiMario}}, \citenamefont {{South}},
-  \citenamefont {{Hansen}}, \citenamefont {{Twitchen}},\ and\ \citenamefont
-  {{Walsworth}}}]{Edmonds}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
-  {{Edmonds}}}, \bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont
-  {{Hart}}}, \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont
-  {{Turner}}}, \bibinfo {author} {\bibfnamefont {P.-O.}\ \bibnamefont
-  {{Colard}}}, \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
-  {{Schloss}}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {{Olsson}}},
-  \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {{Trubko}}}, \bibinfo
-  {author} {\bibfnamefont {M.~L.}\ \bibnamefont {{Markham}}}, \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {{Rathmill}}}, \bibinfo {author}
-  {\bibfnamefont {B.}~\bibnamefont {{Horne-Smith}}}, \bibinfo {author}
-  {\bibfnamefont {W.}~\bibnamefont {{Lew}}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {{Manickam}}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {{Bruce}}}, \bibinfo {author} {\bibfnamefont {P.~G.}\
-  \bibnamefont {{Kaup}}}, \bibinfo {author} {\bibfnamefont {J.~C.}\
-  \bibnamefont {{Russo}}}, \bibinfo {author} {\bibfnamefont {M.~J.}\
-  \bibnamefont {{DiMario}}}, \bibinfo {author} {\bibfnamefont {J.~T.}\
-  \bibnamefont {{South}}}, \bibinfo {author} {\bibfnamefont {J.~T.}\
-  \bibnamefont {{Hansen}}}, \bibinfo {author} {\bibfnamefont {D.~J.}\
-  \bibnamefont {{Twitchen}}}, \ and\ \bibinfo {author} {\bibfnamefont {R.~L.}\
-  \bibnamefont {{Walsworth}}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
-  {journal} {arXiv e-prints}\ ,\ \bibinfo {eid} {arXiv:2004.01746}} (\bibinfo
-  {year} {2020})},\ \Eprint {http://arxiv.org/abs/2004.01746} {arXiv:2004.01746
-  [cond-mat.mtrl-sci]} \BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Tallaire}\ \emph {et~al.}(2020)\citenamefont
-  {Tallaire}, \citenamefont {Brinza}, \citenamefont {Huillery}, \citenamefont
-  {Delord}, \citenamefont {Pellet-Mary}, \citenamefont {Staacke}, \citenamefont
-  {Abel}, \citenamefont {Pezzagna}, \citenamefont {Meijer}, \citenamefont
-  {Touati}, \citenamefont {Binet}, \citenamefont {Ferrier}, \citenamefont
-  {Goldner}, \citenamefont {Hetet},\ and\ \citenamefont
-  {Achard}}]{TALLAIRE2020421}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Tallaire}}, \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Brinza}},
-  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Huillery}}, \bibinfo
-  {author} {\bibfnamefont {T.}~\bibnamefont {Delord}}, \bibinfo {author}
-  {\bibfnamefont {C.}~\bibnamefont {Pellet-Mary}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Staacke}}, \bibinfo {author} {\bibfnamefont
-  {B.}~\bibnamefont {Abel}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Pezzagna}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Meijer}},
-  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Touati}}, \bibinfo
-  {author} {\bibfnamefont {L.}~\bibnamefont {Binet}}, \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Ferrier}}, \bibinfo {author} {\bibfnamefont
-  {P.}~\bibnamefont {Goldner}}, \bibinfo {author} {\bibfnamefont
-  {G.}~\bibnamefont {Hetet}}, \ and\ \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Achard}},\ }\href {\doibase
-  https://doi.org/10.1016/j.carbon.2020.08.048} {\bibfield  {journal} {\bibinfo
-   {journal} {Carbon}\ }\textbf {\bibinfo {volume} {170}},\ \bibinfo {pages}
-  {421 } (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Mindarava}\ \emph {et~al.}(2020)\citenamefont
-  {Mindarava}, \citenamefont {Blinder}, \citenamefont {Laube}, \citenamefont
-  {Knolle}, \citenamefont {Abel}, \citenamefont {Jentgens}, \citenamefont
-  {Isoya}, \citenamefont {Scheuer}, \citenamefont {Lang}, \citenamefont
-  {Schwartz}, \citenamefont {Naydenov},\ and\ \citenamefont
-  {Jelezko}}]{MINDARAVA2020182}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
-  {Mindarava}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Blinder}},
-  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Laube}}, \bibinfo
-  {author} {\bibfnamefont {W.}~\bibnamefont {Knolle}}, \bibinfo {author}
-  {\bibfnamefont {B.}~\bibnamefont {Abel}}, \bibinfo {author} {\bibfnamefont
-  {C.}~\bibnamefont {Jentgens}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Scheuer}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Lang}}, \bibinfo {author} {\bibfnamefont {I.}~\bibnamefont
-  {Schwartz}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Naydenov}},
-  \ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},\ }\href
-  {\doibase https://doi.org/10.1016/j.carbon.2020.07.077} {\bibfield  {journal}
-  {\bibinfo  {journal} {Carbon}\ }\textbf {\bibinfo {volume} {170}},\ \bibinfo
-  {pages} {182 } (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Jarmola}\ \emph {et~al.}(2015)\citenamefont
-  {Jarmola}, \citenamefont {Berzins}, \citenamefont {Smits}, \citenamefont
-  {Smits}, \citenamefont {Prikulis}, \citenamefont {Gahbauer}, \citenamefont
-  {Ferber}, \citenamefont {Erts}, \citenamefont {Auzinsh},\ and\ \citenamefont
-  {Budker}}]{jarmola_longitudinal_2015}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Jarmola}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Berzins}},
-  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Smits}}, \bibinfo
-  {author} {\bibfnamefont {K.}~\bibnamefont {Smits}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Prikulis}}, \bibinfo {author}
-  {\bibfnamefont {F.}~\bibnamefont {Gahbauer}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Ferber}}, \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Erts}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Auzinsh}}, \ and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Budker}},\ }\href {\doibase 10.1063/1.4937489} {\bibfield  {journal}
-  {\bibinfo  {journal} {Appl. Phys. Lett.}\ }\textbf {\bibinfo {volume}
-  {107}},\ \bibinfo {pages} {242403} (\bibinfo {year} {2015})},\ \bibinfo
-  {note} {number: 24}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {van Oort}\ and\ \citenamefont
-  {Glasbeek}(1991)}]{van_oort_optically_1991}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {van
-  Oort}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Glasbeek}},\
-  }\href {\doibase 10.1007/BF03166042} {\bibfield  {journal} {\bibinfo
-  {journal} {Appl Magn Reson}\ }\textbf {\bibinfo {volume} {2}},\ \bibinfo
-  {pages} {291} (\bibinfo {year} {1991})},\ \bibinfo {note} {number:
-  2}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2019)\citenamefont
-  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
-  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
-  and\ \citenamefont {Hemmer}}]{akhmedzhanov_magnetometry_2019}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
-  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
-  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
-  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
-  {\bibfnamefont {I.}~\bibnamefont {Zelensky}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href {\doibase
-  10.1103/PhysRevA.100.043844} {\bibfield  {journal} {\bibinfo  {journal}
-  {Phys. Rev. A}\ }\textbf {\bibinfo {volume} {100}},\ \bibinfo {pages}
-  {043844} (\bibinfo {year} {2019})},\ \bibinfo {note} {number: 4}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {van Oort}\ and\ \citenamefont
-  {Glasbeek}(1989)}]{van_oort_cross-relaxation_1989}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {van
-  Oort}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Glasbeek}},\
-  }\href {\doibase 10.1103/PhysRevB.40.6509} {\bibfield  {journal} {\bibinfo
-  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {40}},\ \bibinfo
-  {pages} {6509} (\bibinfo {year} {1989})},\ \bibinfo {note} {number:
-  10}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2017)\citenamefont
-  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
-  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
-  and\ \citenamefont {Hemmer}}]{akhmedzhanov_microwave-free_2017}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
-  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
-  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
-  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
-  {\bibfnamefont {I.}~\bibnamefont {Zelensky}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href {\doibase
-  10.1103/PhysRevA.96.013806} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
-  Rev. A}\ }\textbf {\bibinfo {volume} {96}},\ \bibinfo {pages} {013806}
-  (\bibinfo {year} {2017})},\ \bibinfo {note} {number: 1}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Holliday}\ \emph {et~al.}(1989)\citenamefont
-  {Holliday}, \citenamefont {Manson}, \citenamefont {Glasbeek},\ and\
-  \citenamefont {Oort}}]{holliday_optical_1989}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
-  {Holliday}}, \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont
-  {Manson}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Glasbeek}}, \
-  and\ \bibinfo {author} {\bibfnamefont {E.~v.}\ \bibnamefont {Oort}},\ }\href
-  {\doibase 10.1088/0953-8984/1/39/021} {\bibfield  {journal} {\bibinfo
-  {journal} {J. Phys.: Condens. Matter}\ }\textbf {\bibinfo {volume} {1}},\
-  \bibinfo {pages} {7093} (\bibinfo {year} {1989})},\ \bibinfo {note} {number:
-  39}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Choi}\ \emph {et~al.}(2017)\citenamefont {Choi},
-  \citenamefont {Choi}, \citenamefont {Kucsko}, \citenamefont {Maurer},
-  \citenamefont {Shields}, \citenamefont {Sumiya}, \citenamefont {Onoda},
-  \citenamefont {Isoya}, \citenamefont {Demler}, \citenamefont {Jelezko},
-  \citenamefont {Yao},\ and\ \citenamefont {Lukin}}]{choi_depolarization_2017}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
-  {Choi}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo
-  {author} {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author}
-  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
-  {\bibfnamefont {B.~J.}\ \bibnamefont {Shields}}, \bibinfo {author}
-  {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
-  {E.}~\bibnamefont {Demler}}, \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Jelezko}}, \bibinfo {author} {\bibfnamefont {N.~Y.}\
-  \bibnamefont {Yao}}, \ and\ \bibinfo {author} {\bibfnamefont {M.~D.}\
-  \bibnamefont {Lukin}},\ }\href {\doibase 10.1103/PhysRevLett.118.093601}
-  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf
-  {\bibinfo {volume} {118}},\ \bibinfo {pages} {093601} (\bibinfo {year}
-  {2017})},\ \bibinfo {note} {number: 9}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Ashfold}\ \emph {et~al.}(2020)\citenamefont
-  {Ashfold}, \citenamefont {Goss}, \citenamefont {Green}, \citenamefont {May},
-  \citenamefont {Newton},\ and\ \citenamefont {Peaker}}]{Ashfold}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~N.~R.}\
-  \bibnamefont {Ashfold}}, \bibinfo {author} {\bibfnamefont {J.~P.}\
-  \bibnamefont {Goss}}, \bibinfo {author} {\bibfnamefont {B.~L.}\ \bibnamefont
-  {Green}}, \bibinfo {author} {\bibfnamefont {P.~W.}\ \bibnamefont {May}},
-  \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont {Newton}}, \ and\
-  \bibinfo {author} {\bibfnamefont {C.~V.}\ \bibnamefont {Peaker}},\
-  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Chemical Reviews}\
-  }\textbf {\bibinfo {volume} {120}},\ \bibinfo {pages} {5745} (\bibinfo {year}
-  {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Simanovskaia}\ \emph {et~al.}(2013)\citenamefont
-  {Simanovskaia}, \citenamefont {Jensen}, \citenamefont {Jarmola},
-  \citenamefont {Aulenbacher}, \citenamefont {Manson},\ and\ \citenamefont
-  {Budker}}]{simanovskaia_sidebands_2013}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Simanovskaia}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
-  {Jensen}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}},
-  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Aulenbacher}}, \bibinfo
-  {author} {\bibfnamefont {N.}~\bibnamefont {Manson}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }\href {\doibase
-  10.1103/PhysRevB.87.224106} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
-  Rev. B}\ }\textbf {\bibinfo {volume} {87}},\ \bibinfo {pages} {224106}
-  (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Smeltzer}\ \emph {et~al.}(2011)\citenamefont
-  {Smeltzer}, \citenamefont {Childress},\ and\ \citenamefont
-  {Gali}}]{smeltzer201113c}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.}~\bibnamefont
-  {Smeltzer}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Childress}},
-  \ and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Gali}},\
-  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {New Journal of
-  Physics}\ }\textbf {\bibinfo {volume} {13}},\ \bibinfo {pages} {025021}
-  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
-\end{thebibliography}%
+\subsection{Fitting of the cross-relaxation scans}
+\begin{figure}
+\includegraphics[scale=0.5]{Gauss_vs_Lor_VH}
+\caption{Zoom-in on the VH$^-$ cross-relaxation line from Fig.3 of the main text with Gaussian (orange,plain line) and Lorentzian (red, dashed line) fit}
+\label{Gauss_vs_Lor}
+\end{figure}
+In order to extract quantitative information from Fig.3 of the main text, we had ta apply two distinct fits to the raw data : first a polynomial fit to subtract the slowly decaying PL due to sate-mixing, then a Lorentzian fit on the VH$^-$ and WAR1 resonance lines to measure the central position, width and amplitude of the lines.
+
+For the subtracting fit, we chose a polynomial fit because numerical solving of the transverse-field dependent spin depolarization could not properly fit our results.
+We had to use up a 4th order polynomial in order to obtain a tight fit to the data.
+Higher order polynomials did not significantly change the fitting envelope.
+
+The second and third peaks were fitted better on their wings by a Gaussian than by a Lorentzian lineshape, as can be seen in Fig.\ref{Gauss_vs_Lor} where the Gaussian fit yields a $R^2=0.983$ greater than the $R^2=0.974$ from the Lorentzian fit.
+This is coherent with the fact that the inhomogeneous broadening of the various spin species dominates the coupling strength between the spins.
+
+\subsection{Sample fabrication details}
+The main CVD sample of this study is presented in thorough details in \citep{TALLAIRE2020421}.
+It is a thick (100) oriented diamond grown in-house by CVD with an initial concentration [N]$\approx$25~ppm and separated from its HPHT substrate.
+It was electron irradiated at 10MeV and a fluence of $2\cdot 10^{18}$~cm$^{-2}$ and in situ annealed at 900°C during irradiation to give a final concentration [NV]$\approx$4.6~ppm.
+
+The HPHT sample used in Sec.I of the supplementary material is a thick (111)-oriented diamond from Sumitomo and was electron irradiated at an energy of 2 MeV and a fluence of $7\cdot 10^{18}$~cm$^{-2}$.
+It was then annealed in vacuum for 2h at 950°C.
+
+We also tried the same experiment on similar HPHT samples with a fluence ranging from $4\cdot 10^{18}$~cm$^{-2}$ to $1.5\cdot 10^{18}$~cm$^{-2}$, as well as on several heavily doped 15 $\mu$m diamonds bought from Adamas Nanotechnologies, and all the HPHT sample showed a feature at $\sim$20~G, but none of them showed any feature at 56 or 125~G.
+
+\begin{thebibliography}{10}
+
+\bibitem{Achard}
+J~Achard, V~Jacques, and A~Tallaire.
+\newblock Chemical vapour deposition diamond single crystals with
+  nitrogen-vacancy centres: a review of material synthesis and technology for
+  quantum sensing applications.
+\newblock {\em Journal of Physics D: Applied Physics}, 53(31):313001, may 2020.
+
+\bibitem{akhmedzhanov_microwave-free_2017}
+Rinat Akhmedzhanov, Lev Gushchin, Nikolay Nizov, Vladimir Nizov, Dmitry
+  Sobgayda, Ilya Zelensky, and Philip Hemmer.
+\newblock Microwave-free magnetometry based on cross-relaxation resonances in
+  diamond nitrogen-vacancy centers.
+\newblock {\em Phys. Rev. A}, 96(1):013806, July 2017.
+\newblock Number: 1.
+
+\bibitem{akhmedzhanov_magnetometry_2019}
+Rinat Akhmedzhanov, Lev Gushchin, Nikolay Nizov, Vladimir Nizov, Dmitry
+  Sobgayda, Ilya Zelensky, and Philip Hemmer.
+\newblock Magnetometry by cross-relaxation-resonance detection in ensembles of
+  nitrogen-vacancy centers.
+\newblock {\em Phys. Rev. A}, 100(4):043844, October 2019.
+\newblock Number: 4.
+
+\bibitem{Alfasi}
+Nir Alfasi, Sergei Masis, Oleg Shtempluck, and Eyal Buks.
+\newblock Detection of paramagnetic defects in diamond using off-resonance
+  excitation of nv centers.
+\newblock {\em Phys. Rev. B}, 99:214111, Jun 2019.
+
+\bibitem{alfasi2019detection}
+Nir Alfasi, Sergei Masis, Oleg Shtempluck, and Eyal Buks.
+\newblock Detection of paramagnetic defects in diamond using off-resonance
+  excitation of nv centers.
+\newblock {\em Physical Review B}, 99(21):214111, 2019.
+
+\bibitem{armstrong_nvnv_2010}
+Seiji Armstrong, Lachlan~J. Rogers, Roger~L. McMurtrie, and Neil~B. Manson.
+\newblock {NV}–{NV} electron–electron spin and {NV}–{NS} electron —
+  electron and electron-nuclear spin interaction in diamond.
+\newblock {\em Physics Procedia}, 3(4):1569--1575, February 2010.
+\newblock Number: 4.
+
+\bibitem{Ashfold}
+Michael N.~R. Ashfold, Jonathan~P. Goss, Ben~L. Green, Paul~W. May, Mark~E.
+  Newton, and Chloe~V. Peaker.
+\newblock Nitrogen in diamond.
+\newblock {\em Chemical Reviews}, 120(12):5745--5794, 06 2020.
+
+\bibitem{choi_depolarization_2017}
+Joonhee Choi, Soonwon Choi, Georg Kucsko, Peter~C. Maurer, Brendan~J. Shields,
+  Hitoshi Sumiya, Shinobu Onoda, Junichi Isoya, Eugene Demler, Fedor Jelezko,
+  Norman~Y. Yao, and Mikhail~D. Lukin.
+\newblock Depolarization {Dynamics} in a {Strongly} {Interacting}
+  {Solid}-{State} {Spin} {Ensemble}.
+\newblock {\em Phys. Rev. Lett.}, 118(9):093601, March 2017.
+\newblock Number: 9.
+
+\bibitem{cruddace2007magnetic}
+Robin Cruddace.
+\newblock {\em Magnetic resonance and optical studies of point defects in
+  single crystal CVD diamond}.
+\newblock PhD thesis, University of Warwick, 2007.
+
+\bibitem{DOHERTY20131}
+Marcus~W. Doherty, Neil~B. Manson, Paul Delaney, Fedor Jelezko, Jörg
+  Wrachtrup, and Lloyd~C.L. Hollenberg.
+\newblock The nitrogen-vacancy colour centre in diamond.
+\newblock {\em Physics Reports}, 528(1):1 -- 45, 2013.
+\newblock The nitrogen-vacancy colour centre in diamond.
+
+\bibitem{Edmonds}
+Andrew~M. {Edmonds}, Connor~A. {Hart}, Matthew~J. {Turner}, Pierre-Olivier
+  {Colard}, Jennifer~M. {Schloss}, Kevin {Olsson}, Raisa {Trubko}, Matthew~L.
+  {Markham}, Adam {Rathmill}, Ben {Horne-Smith}, Wilbur {Lew}, Arul {Manickam},
+  Scott {Bruce}, Peter~G. {Kaup}, Jon~C. {Russo}, Michael~J. {DiMario},
+  Joseph~T. {South}, Jay~T. {Hansen}, Daniel~J. {Twitchen}, and Ronald~L.
+  {Walsworth}.
+\newblock {Generation of nitrogen-vacancy ensembles in diamond for quantum
+  sensors: Optimization and scalability of CVD processes}.
+\newblock {\em arXiv e-prints}, page arXiv:2004.01746, April 2020.
+
+\bibitem{Epstein}
+R.~J. Epstein, F.~M. Mendoza, Y.~K. Kato, and D.~D. Awschalom.
+\newblock Anisotropic interactions of a single spin and dark-spin spectroscopy
+  in diamond.
+\newblock {\em Nature Physics}, 1(2):94--98, 2005.
+
+\bibitem{glover_hydrogen_2003}
+Claire Glover, M.~E. Newton, P.~Martineau, D.~J. Twitchen, and J.~M. Baker.
+\newblock Hydrogen {Incorporation} in {Diamond}: {The}
+  {Nitrogen}-{Vacancy}-{Hydrogen} {Complex}.
+\newblock {\em Phys. Rev. Lett.}, 90(18):185507, May 2003.
+
+\bibitem{glover_hydrogen_2004}
+Claire Glover, M.~E. Newton, P.~M. Martineau, Samantha Quinn, and D.~J.
+  Twitchen.
+\newblock Hydrogen {Incorporation} in {Diamond}: {The} {Vacancy}-{Hydrogen}
+  {Complex}.
+\newblock {\em Phys. Rev. Lett.}, 92(13):135502, March 2004.
+\newblock Publisher: American Physical Society.
+
+\bibitem{Hall}
+L.~T. Hall, P.~Kehayias, D.~A. Simpson, A.~Jarmola, A.~Stacey, D.~Budker, and
+  L.~C.~L. Hollenberg.
+\newblock Detection of nanoscale electron spin resonance spectra demonstrated
+  using nitrogen-vacancy centre probes in diamond.
+\newblock {\em Nature Communications}, 7(1):10211, 2016.
+
+\bibitem{holliday_optical_1989}
+K~Holliday, N~B Manson, M~Glasbeek, and E~van Oort.
+\newblock Optical hole-bleaching by level anti-crossing and cross relaxation in
+  the {N}-{V} centre in diamond.
+\newblock {\em J. Phys.: Condens. Matter}, 1(39):7093--7102, October 1989.
+\newblock Number: 39.
+
+\bibitem{jarmola_longitudinal_2015}
+A.~Jarmola, A.~Berzins, J.~Smits, K.~Smits, J.~Prikulis, F.~Gahbauer,
+  R.~Ferber, D.~Erts, M.~Auzinsh, and D.~Budker.
+\newblock Longitudinal spin-relaxation in nitrogen-vacancy centers in electron
+  irradiated diamond.
+\newblock {\em Appl. Phys. Lett.}, 107(24):242403, December 2015.
+\newblock Number: 24.
+
+\bibitem{jiang2018estimation}
+Feng-Jian Jiang, Jian-Feng Ye, Zheng Jiao, Zhi-Yong Huang, and Hai-Jiang Lv.
+\newblock Estimation of vector static magnetic field by a nitrogen-vacancy
+  center with a single first-shell 13c nuclear (nv--13c) spin in diamond.
+\newblock {\em Chinese Physics B}, 27(5):057601, 2018.
+
+\bibitem{kamp2018continuous}
+EJ~Kamp, B~Carvajal, and N~Samarth.
+\newblock Continuous wave protocol for simultaneous polarization and optical
+  detection of p1-center electron spin resonance.
+\newblock {\em Physical Review B}, 97(4):045204, 2018.
+
+\bibitem{lazda2020cross}
+Reinis Lazda, Laima Busaite, Andris Berzins, Janis Smits, Marcis Auzinsh,
+  Dmitry Budker, Ruvin Ferber, and Florian Gahbauer.
+\newblock Cross-relaxation studies with optically detected magnetic resonances
+  in nitrogen-vacancy centers in diamond in an external magnetic field.
+\newblock {\em arXiv preprint arXiv:2007.00473}, 2020.
+
+\bibitem{MINDARAVA2020182}
+Yuliya Mindarava, Rémi Blinder, Christian Laube, Wolfgang Knolle, Bernd Abel,
+  Christian Jentgens, Junichi Isoya, Jochen Scheuer, Johannes Lang, Ilai
+  Schwartz, Boris Naydenov, and Fedor Jelezko.
+\newblock Efficient conversion of nitrogen to nitrogen-vacancy centers in
+  diamond particles with high-temperature electron irradiation.
+\newblock {\em Carbon}, 170:182 -- 190, 2020.
+
+\bibitem{newton_epr_2007}
+M.~E. Newton.
+\newblock {EPR}, {ENDOR} and {EPR} {Imaging} of {Defects} in {Diamond}.
+\newblock {\em ChemInform}, 38(47), November 2007.
+
+\bibitem{Rondin_2014}
+L~Rondin, J-P Tetienne, T~Hingant, J-F Roch, P~Maletinsky, and V~Jacques.
+\newblock Magnetometry with nitrogen-vacancy defects in diamond.
+\newblock {\em Reports on Progress in Physics}, 77(5):056503, may 2014.
+
+\bibitem{simanovskaia_sidebands_2013}
+Maria Simanovskaia, Kasper Jensen, Andrey Jarmola, Kurt Aulenbacher, Neil
+  Manson, and Dmitry Budker.
+\newblock Sidebands in optically detected magnetic resonance signals of
+  nitrogen vacancy centers in diamond.
+\newblock {\em Phys. Rev. B}, 87(22):224106, June 2013.
+
+\bibitem{smeltzer201113c}
+Benjamin Smeltzer, Lilian Childress, and Adam Gali.
+\newblock 13c hyperfine interactions in the nitrogen-vacancy centre in diamond.
+\newblock {\em New Journal of Physics}, 13(2):025021, 2011.
+
+\bibitem{TALLAIRE2020421}
+Alexandre Tallaire, Ovidiu Brinza, Paul Huillery, Tom Delord, Clément
+  Pellet-Mary, Robert Staacke, Bernd Abel, Sébastien Pezzagna, Jan Meijer,
+  Nadia Touati, Laurent Binet, Alban Ferrier, Philippe Goldner, Gabriel Hetet,
+  and Jocelyn Achard.
+\newblock High nv density in a pink cvd diamond grown with n2o addition.
+\newblock {\em Carbon}, 170:421 -- 429, 2020.
+
+\bibitem{van_oort_optically_1991}
+E.~van Oort and M.~Glasbeek.
+\newblock Optically detected adiabatic fast passage and cross-relaxation of the
+  {N}-{V} center in diamond.
+\newblock {\em Appl Magn Reson}, 2(2):291--300, June 1991.
+\newblock Number: 2.
+
+\bibitem{van_oort_cross-relaxation_1989}
+Eric van Oort and Max Glasbeek.
+\newblock Cross-relaxation dynamics of optically excited \textit{{N}} -
+  \textit{{V}} centers in diamond.
+\newblock {\em Phys. Rev. B}, 40(10):6509--6517, October 1989.
+\newblock Number: 10.
+
+\bibitem{Wang}
+Hai-Jing Wang, Chang~S. Shin, Scott~J. Seltzer, Claudia~E. Avalos, Alexander
+  Pines, and Vikram~S. Bajaj.
+\newblock Optically detected cross-relaxation spectroscopy of electron spins in
+  diamond.
+\newblock {\em Nature Communications}, 5(1):4135, 2014.
+
+\bibitem{Wickenbrock}
+Arne Wickenbrock, Huijie Zheng, Lykourgos Bougas, Nathan Leefer, Samer Afach,
+  Andrey Jarmola, Victor~M. Acosta, and Dmitry Budker.
+\newblock Microwave-free magnetometry with nitrogen-vacancy centers in diamond.
+\newblock {\em Applied Physics Letters}, 109(5):053505, 2016.
+
+\bibitem{Yao}
+N.~Y. Yao, L.~Jiang, A.~V. Gorshkov, P.~C. Maurer, G.~Giedke, J.~I. Cirac, and
+  M.~D. Lukin.
+\newblock Scalable architecture for a room temperature solid-state quantum
+  information processor.
+\newblock {\em Nature Communications}, 3(1):800, 2012.
+
+\bibitem{alvarez_local_2015}
+Gonzalo~A. Álvarez, Christian~O. Bretschneider, Ran Fischer, Paz London, Hisao
+  Kanda, Shinobu Onoda, Junichi Isoya, David Gershoni, and Lucio Frydman.
+\newblock Local and bulk {13C} hyperpolarization in nitrogen-vacancy-centred
+  diamonds at variable fields and orientations.
+\newblock {\em Nat Commun}, 6(1):8456, December 2015.
+
+\end{thebibliography}
 

### 2021-02-26 data/20200722/traitement2.py

@@ -377,17 +377,32 @@ print("C13 area : %F, VH ratio : %f, WAR1 ratio : %F"%(C13,VH,WAR1))
 
 c1 = next(ax._get_lines.prop_cycler)['color']
 c2 = next(ax._get_lines.prop_cycler)['color']
+c3 = next(ax._get_lines.prop_cycler)['color']
+
+def R2(y,yfit):
+	avg=np.sum(y)/len(y)
+	SStot=sum((y-avg)**2)
+	SSres=sum((y-yfit)**2)
+	return 1-SSres/SStot
 
 
-popt,yfit=gauss_fit(x[95:175],y[95:175])
-plt.plot(x[95:175],yfit,lw=2,color=c1)
-print(popt[2]*1.18)
+xmin_VH=70 #95
+xmax_VH=200 #175
+popt,yfit=gauss_fit(x[xmin_VH:xmax_VH],y[xmin_VH:xmax_VH])
+plt.plot(x[xmin_VH:xmax_VH],yfit,lw=2,color=c1,label="Gaussian")
+print(R2(y[xmin_VH:xmax_VH],yfit))
+popt,yfit=lor_fit(x[xmin_VH:xmax_VH],y[xmin_VH:xmax_VH])
+plt.plot(x[xmin_VH:xmax_VH],yfit,'--',lw=2,color=c3,label="Lorentzian")
+print(R2(y[xmin_VH:xmax_VH],yfit))
+# print(popt[2]*1.18)
 x_v=popt[1]
-plt.plot([x_v,x_v],[1,-1],'--',color=c2)
+# plt.plot([x_v,x_v],[1,-1],'--',color=c2)
 
 popt,yfit=gauss_fit(x[312:380],y[312:380])
 plt.plot(x[312:380],yfit,lw=2,color=c1)
-print(popt[2]*1.18)
+# popt,yfit=lor_fit(x[312:380],y[312:380])
+# plt.plot(x[312:380],yfit,lw=2,color=c3)
+# print(popt[2]*1.18)
 x_v=popt[1]
 plt.plot([x_v,x_v],[1,-1],'--',color=c2)
 
@@ -399,5 +414,6 @@ for x_v in x_C13 :
 
 
 ax.set_ylim(ylim)
+plt.legend()
 plt.show()
 

### 2021-02-26 ensignement/TD meca/AN.py

@@ -2,7 +2,7 @@ from math import *
 
 kb=1.38E-23
 Na=6.02e23
-
+gamma=
 
 P=1e5
 T=300

### 2021-03-08 Mes papiers/CR-CVD/SI/CVD_para_SI.tex

@@
 
-\bibliography{Biblio_CVD-CR}
+\section{Fitting of the cross-relaxation scans}
+\begin{figure}
+\includegraphics[scale=0.5]{Gauss_vs_Lor_VH}
+\caption{Zoom-in on the VH$^-$ cross-relaxation line from Fig.3 of the main text with Gaussian (orange,plain line) and Lorentzian (red, dashed line) fit}
+\label{Gauss_vs_Lor}
+\end{figure}
+In order to extract quantitative information from Fig.3 of the main text, we had ta apply two distinct fits to the raw data : first a polynomial fit to subtract the slowly decaying PL due to sate-mixing, then a Lorentzian fit on the VH$^-$ and WAR1 resonance lines to measure the central position, width and amplitude of the lines.
+
+For the subtracting fit, we chose a polynomial fit because numerical solving of the transverse-field dependent spin depolarization could not properly fit our results.
+We had to use up a 4th order polynomial in order to obtain a tight fit to the data.
+Higher order polynomials did not significantly change the fitting envelope.
+
+The second and third peaks were fitted better on their wings by a Gaussian than by a Lorentzian lineshape, as can be seen in Fig.
+\ref{Gauss_vs_Lor} where the Gaussian fit yields a $R^2=0.983$ greater than the $R^2=0.974$ from the Lorentzian fit.
+This is coherent with the fact that the inhomogeneous broadening of the various spin species dominates the coupling strength between the spins.
+
+\section{Sample fabrication details}
+The main CVD sample of this study is presented in thorough details in \citep{TALLAIRE2020421}.
+It is a thick (100) oriented diamond grown in-house by CVD with an initial concentration [N]$\approx$25~ppm and separated from its HPHT substrate.
+It was electron irradiated at 10MeV and a fluence of $2\cdot 10^{18}$~cm$^{-2}$ and in situ annealed at 900°C during irradiation to give a final concentration [NV]$\approx$4.6~ppm.
+
+The HPHT sample used in Sec.I of the supplementary material is a thick (111)-oriented diamond from Sumitomo and was electron irradiated at an energy of 2 MeV and a fluence of $7\cdot 10^{18}$~cm$^{-2}$.
+It was then annealed in vacuum for 2h at 950°C.
+
+We also tried the same experiment on similar HPHT samples with a fluence ranging from $4\cdot 10^{18}$~cm$^{-2}$ to $1.5\cdot 10^{18}$~cm$^{-2}$, as well as on several heavily doped 15 $\mu$m diamonds bought from Adamas Nanotechnologies, and all the HPHT sample showed a feature at $\sim$20~G, but none of them showed any feature at 56 or 125~G.
+
+\begin{thebibliography}{1}
+
+\bibitem{alfasi2019detection}
+Nir Alfasi, Sergei Masis, Oleg Shtempluck, and Eyal Buks.
+\newblock Detection of paramagnetic defects in diamond using off-resonance
+  excitation of nv centers.
+\newblock {\em Physical Review B}, 99(21):214111, 2019.
+
+\bibitem{jiang2018estimation}
+Feng-Jian Jiang, Jian-Feng Ye, Zheng Jiao, Zhi-Yong Huang, and Hai-Jiang Lv.
+\newblock Estimation of vector static magnetic field by a nitrogen-vacancy
+  center with a single first-shell 13c nuclear (nv--13c) spin in diamond.
+\newblock {\em Chinese Physics B}, 27(5):057601, 2018.
+
+\bibitem{kamp2018continuous}
+EJ~Kamp, B~Carvajal, and N~Samarth.
+\newblock Continuous wave protocol for simultaneous polarization and optical
+  detection of p1-center electron spin resonance.
+\newblock {\em Physical Review B}, 97(4):045204, 2018.
+
+\bibitem{lazda2020cross}
+Reinis Lazda, Laima Busaite, Andris Berzins, Janis Smits, Marcis Auzinsh,
+  Dmitry Budker, Ruvin Ferber, and Florian Gahbauer.
+\newblock Cross-relaxation studies with optically detected magnetic resonances
+  in nitrogen-vacancy centers in diamond in an external magnetic field.
+\newblock {\em arXiv preprint arXiv:2007.00473}, 2020.
+
+\bibitem{simanovskaia_sidebands_2013}
+Maria Simanovskaia, Kasper Jensen, Andrey Jarmola, Kurt Aulenbacher, Neil
+  Manson, and Dmitry Budker.
+\newblock Sidebands in optically detected magnetic resonance signals of
+  nitrogen vacancy centers in diamond.
+\newblock {\em Phys. Rev. B}, 87(22):224106, June 2013.
+
+\bibitem{TALLAIRE2020421}
+Alexandre Tallaire, Ovidiu Brinza, Paul Huillery, Tom Delord, Clément
+  Pellet-Mary, Robert Staacke, Bernd Abel, Sébastien Pezzagna, Jan Meijer,
+  Nadia Touati, Laurent Binet, Alban Ferrier, Philippe Goldner, Gabriel Hetet,
+  and Jocelyn Achard.
+\newblock High nv density in a pink cvd diamond grown with n2o addition.
+\newblock {\em Carbon}, 170:421 -- 429, 2020.
+
+\bibitem{alvarez_local_2015}
+Gonzalo~A. Álvarez, Christian~O. Bretschneider, Ran Fischer, Paz London, Hisao
+  Kanda, Shinobu Onoda, Junichi Isoya, David Gershoni, and Lucio Frydman.
+\newblock Local and bulk {13C} hyperpolarization in nitrogen-vacancy-centred
+  diamonds at variable fields and orientations.
+\newblock {\em Nat Commun}, 6(1):8456, December 2015.
+
+\end{thebibliography}
 

### 2021-03-08 Mes papiers/CR-CVD/SI/pre_SI.tex

--- 2020-11-27 Mes papiers/Double quantum/pre_SI.tex
@@
 \bibliographystyle{plain}
-
-\section{Scan in the [111] direction}
-\begin{figure}
-\label{scan_PL}
-\includegraphics[scale=.48]{scan_111_PL}
-\caption{\textbf{a)} Plain blue line : Photoluminescence change as a function of the magnetic field amplitude in the [111] crystalline direction.
-Dashed orange line : order-4 polynomial fit of the decreasing envelope.
-\textbf{b)} Subtraction of the previous signal by the envelope fit.
-The vertical dotted orange lines correspond to the expected cross-relaxations with VH$^-$ and the vertical dashed green lines to the expected cross-relaxations with WAR1.}
-\end{figure}
-\begin{figure}
-\label{transis_VHWAR}
-\includegraphics[scale=.48]{Transis_111_VHWAR}
-\caption{Predicted transitions frequencies for NV centers (Plain lines), VH$^-$ (dashed lines) and WAR1 (dashdotted lines) when the magnetic field is scanned in the [111] direction.
-The magnetic field where a transition of the NV center crosses one of VH$^-$ or WAR1 are reported on fig \ref{scan_PL}}
-\end{figure}
-
-\begin{figure}
-\label{transis_13C}
-\includegraphics[scale=.55]{Transis_111_13C}
-\caption{Predicted transitions frequencies for NV centers (Plain lines) and NV-$^{13}$C complex when the magnetic field is scanned in the [111] direction.
-Vertical lines represent the various degeneracy conditions.}
-\end{figure}
-
-We have performed similar magnetic field scan as presented in Fig.3 of the main text, but this time scanning the magnetic field in the crystalline [111] direction instead of the [100], meaning that this time one class of NV centers is aligned with the magnetic field while the three others are disaligned and degenerated.
-The results in term of cross-relaxations generally means 4 times as many degeneracy conditions as the [100] case, since the two spins involved in the cross-relaxation process now have two possible orientations, and therefore two separate transitions.
-
-It is then significantly harder to find cross-relaxations signatures, since they are both more numerous and less contrasted than in the [100] case.
-The experimental results are shown in fig.
-\ref{scan_PL}.
-Similarly to fig.
-3 of the main text, we have subtracted the envelope associated with the transverse field in order to better see the fine features of the scan.
-The predicted cross-relaxation conditions between NV centers and VH$^-$ or WAR1 are presented in fig.
-\ref{transis_VHWAR} and reported on fig.
-\ref{scan_PL}.
-While the predicted values do not exactly match the experimental results, it seems like every predicted transition falls in the vicinity of an experimental line.
-Importantly, it should be noted that, unlike Fig.3 of the main text, the magnetic field was not calibrated with NV ODMR at every field value in this case.
-Instead we have simply applied an homotethy on the current intensity in the electromagnet to match the field values at the maxima of the scan.
-This means that the magnetic field of the points of low field value in particular are more susceptible to the non-linearities due to the electromagnet hysteresis cycle.
-
-Finally we have plotted on fig.\ref{transis_13C} the expected cross-relaxations between NV and NV-$^{13}$C complex.
-It seems that these transitions are too diluted to appear in our experimental scan.
-The peak at $\sim$37 G in fig.\ref{scan_PL} which in the only peak unpaired with any of the NV-VH$^-$ or NV-WAR1 resonance could come from the resonance between NV and NV-$^{13}$C since they seem particularly dense at this particular field values.
 \bibliography{Biblio_CVD-CR}

### 2021-03-12 Mes papiers/CR-CVD/CVD_para.tex

@@
 \ref{fig1} shows a schematics of the various paramagnetic defects that have been coupled to the nitrogen-vacancy center and detected optically in our study.
-These defects are the VH$^-$ and WAR1 defects, two spin-1 defects with comparable zero field splitting to the NV$^-$ center \citep{glover_hydrogen_2004, cruddace2007magnetic}, as well as $^{13}$C $1/2$-nuclear spin, through hyper-fine coupling with a secondary NV$^-$ center.
+These defects are the VH$^-$ and WAR1 defects, two spin-1 defects with comparable zero field splitting than the NV$^-$ center \citep{glover_hydrogen_2004, cruddace2007magnetic}, as well as $^{13}$C spin $1/2$-nuclear spins, through hyperfine coupling with a secondary NV$^-$ center.
 The detection is realized by measuring the NV photoluminescence while performing magnetic field scans to resonantly enhance dipole-dipole interactions and observe cross-relaxations (CR).
@@
 %Extensive work has been done in the past twenty years to detect and characterize the various  that appear in EPR. 
-The sample we use in this study  correspond to the pink irradiated sample in \cite{TALLAIRE2020421}. It was grown using CVD with the addition of 500ppm of N$_2$0 to the H2/CH4 (96/4) gas phase.
-Then high energy (10MeV) electron irradiation at a fluence of $2\times18$ cm$^{-2}$ and at a temperature of 900 degrees celsius was realized,  giving a final concentration of NV$^-$ of about 4.6~ppm for an initial concentration of nitrogen in the sample of about 25~ppm.
-It was shown that in this sample the $T_2^*$ was not degraded after annealing, yet the NV density was large enough to enable concentration dependent longitudinal relaxation \cite{TALLAIRE2020421, jarmola_longitudinal_2015}.
+The sample we use is the pink irradiated sample presented in \cite{TALLAIRE2020421}. It was grown using CVD with the addition of 500ppm of N$_2$0 to the H2/CH4 (96/4) gas phase.
+Then high energy (10MeV) electron irradiation at a fluence of $2\times18$ cm$^{-2}$ and at a temperature of 900 $^\circ$C was realized,  giving a final concentration of NV$^-$ of about 4.6~ppm.
+The initial concentration of nitrogen in the sample was about 25~ppm.
+It was shown that the $T_2^*$ was not degraded after annealing, yet the NV density was large enough to enable concentration dependent longitudinal relaxation \cite{TALLAIRE2020421, jarmola_longitudinal_2015}.
 %It is well-known that when the NV concentration increases above 1 ppm, the longitudinal decay of NV centers is degraded. 
@@
 Looking at the diamond structure (bottom right in Fig.
-\ref{fig1}), it can be noticed that, in this direction, the four possible projections of the NV centers (or any other $C_{3v}$ defect) along the magnetic field are identical.
-This means that a B field scan along this direction cannot make the transition energies of the different classes of NV centers cross.
-Under arbitrary angles, this could otherwise result in several low field PL features \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the Supplementary Materials for data showing scans along the [111] direction).
+\ref{fig1}), it can be noticed that, in this direction, the projections of NV centers (or any other $C_{3v}$ defect) along the magnetic field are identical.
+This means the transition energies of the different classes of NV centers cannot cross under a B field scan along this direction.
+This could otherwise result in several low field PL features under arbitrary angles \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the Supplementary Materials for data showing scans along the [111] direction).
 The downside of this choice of the [100] direction is that at large amplitudes, the transverse component of the magnetic field depolarizes all classes of NVs, thus limiting the magnetic field range that can be employed and restricting this detection method to paramagnetic defects with zero-field splittings close to the NV center's.
@@
 Such a drop of the PL, also observed in
-\cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}, correspond to situation where the transitions of NV centers become degenerate.
-The planes orthogonal to the [010], [001], [011], and [01$\bar 1$] directions are indicated by dashed lines and show the locus of the cross relaxation. These planes all cross on the [100] axis.
+\cite{van_oort_optically_1991, van_oort_cross-relaxation_1989, jarmola_longitudinal_2015, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, holliday_optical_1989, choi_depolarization_2017}, correspond to transitions of NV centers becoming degenerate.
+The planes orthogonal to the [010], [001], [011], and [01$\bar 1$] directions are indicated by dashed lines and show the locus of the cross relaxation. They all cross on the [100] axis.
 The origin of these sharp changes in the photoluminescence was attributed to cross relaxation between polarized NV centers and rapidly decaying NV centers, so called fluctuators \cite{choi_depolarization_2017}, the precise origin of the latter remains unknown.
-The width of the CR features when the magnetic field crosses a plane perpendicularly was found to be compatible with the NV decoherence rate  ($\approx$ 6 MHz).
+The width of the CR features when the magnetic field crosses a plane perpendicularly, was found to be compatible with the NV decoherence rate  ($\approx$ 6 MHz).
 The contrast is determined both by the fluctuating NV and polarized NV concentrations.
@@
 \ref{map}-b) shows an ESR taken at the angle  $(\phi, \theta)=(0^\circ,0^\circ)$ showing only two features, as expected from a [100] axis.
-Using this goniometer, the magnetic field angle could be finely adjusted so that the four NV lines become fully degenerate along the $[100]$ direction with an error estimate of $\pm 0.5^\circ$.
+Using this goniometer, the magnetic field angle could be finely adjusted so that the NV lines become fully degenerate along the $[100]$ direction with an error estimate of $\pm 0.5^\circ$.
 
@@
 The microwave induced PL features in the scan are then used to relate the magnetic field to the applied voltage.
-Three features appear in the overall PL evolution in this spectrum : one feature at 20~G,  56~G and 122~G. 
-For these data, the averaging was done for 24 hours, but the features appear already with a signal to noise ratio greater than 1 after 10 minutes.  
-We also note an overall drop of the PL, as a result of state mixing in the optically excited state \cite{DOHERTY20131}.
-
+Three features appear in this spectrum : at 20~G,  56~G and 122~G. 
+Data averaging was done for 24 hours, but the features appear already with a signal to noise ratio greater than 1 after 10 minutes.  
+We also observe an overall drop of the PL as a result of state mixing in the optically excited state \cite{DOHERTY20131}. 
 To let the three salient features detach better from the spectrum, we fitted a 4th-order polynomial to the PL evolution, without the spectral bumps, and subtract it from the data.
@@
 
-In order to attribute the three features to their respective defects, we run a similar scan (Section 1 in the Supplementary Materials) on a type Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range, in the same $[100]$ direction.
+In order to attribute the three features to their respective defects, we run a similar scan (Section 1 in the Supplementary Materials) on a type Ib electron irradiated HPHT diamond crystal with a NV concentration in the 5-20 ppm range.
 Only the first feature appeared.
-This observation guided us to search for CVD-related defects as good candidates for the last two features.
+This observation guided us to search for CVD-related defects as candidates for the last two features.
 Paramagnetic defects in CVD-grown diamonds have been extensively studied using EPR  \citep{Ashfold}.
-These studies demonstrate that hydrogen-related complex such as the NVH, VH, VH2 complexes can all be stable in diamond.
-Although their composition is not always known, several zero-field splittings $D$ can be found in the literature.
-Given the small difference between the NV transitions and the observed features, we only concentrate on the reported defects with ZFS that are close to the NV$^-$.
+These studies demonstrate that hydrogen-related complex such as NVH, VH, VH2 can be stable in diamond.
+Although their compositions is not always known, several zero-field splittings $D$ can be found in the literature.
+Given the small difference between the NV transitions and the observed features, we only concentrate on reported defects with ZFS that are close to the NV$^-$.
 The middle column of table \ref{table}, shows the zero-field splitting for two such spin-1 defects found in \citep{cruddace2007magnetic}, namely the negatively charged hydrogen-vacancy (VH$^-$) and WAR1 defects.
@@
 \ref{energy-levels} shows the frequencies of the NV$^-$, the VH$^-$ and WAR1 spin transitions as a function of the magnetic field amplitude along the $[100]$ direction.
-Since the NV$^-$ and VH$^-$ are $C_{3v}$ defects, and the WAR1 is a pseudo-$C_{3v}$ defect \citep{cruddace2007magnetic}, all possible orientation are degenerate for this magnetic field orientation, giving only two lines corresponding to the $\ket{0} \to \ket{+1}$ and $\ket{0} \to \ket{-1}$ transitions.
+Since the NV$^-$ and VH$^-$ are $C_{3v}$ defects, and the WAR1 is a pseudo-$C_{3v}$ defect \citep{cruddace2007magnetic}, all possible orientations are degenerate at this magnetic field orientation, giving only two lines corresponding to the $\ket{0} \to \ket{+1}$ and $\ket{0} \to \ket{-1}$ transitions.
 The points were the NV levels cross the other defects can give rise to cross-relaxation.
@@
 A gaussian fit to the feature in Fig.
-\ref{scan} enables us to extrapolate a value for $D=2694(5)$MHz that coincides very well to that of the VH$^-$ within the error margins indicated in \citep{cruddace2007magnetic}.
-The third peak at 122 G would correspond to a spin defect that has a $D=2470(10)$MHz that also matches well the one of the WAR1 defect. 
-The values of $D$ for these two defects and the error bars from our measurements are now included in the table \ref{table}.
+\ref{scan} enables us to extrapolate a value for $D=2694(5)$MHz that matches that of the VH$^-$ within the error margins indicated in \citep{cruddace2007magnetic}.
+The third peak at 122 G would correspond to a spin defect that has a $D=2470(10)$MHz that also matches the one of the WAR1 defect. 
+The values of $D$ for these two defects and the error bars from our measurements are included in the table \ref{table}.
 The error bars are estimated by taking into account the precision on the NV$^-$ ZFS, on the angle, on the magnetic field calibration and the precision of the fits. 
-We obtain a factor of 6 improvement over the precision reported in \citep{cruddace2007magnetic}.
-We thus attribute the second and third feature to cross relaxation between the NV center and the VH$^-$ and WAR1 defect.
+We obtain a factor of 6 improvement over the precision reported in \citep{cruddace2007magnetic} and attribute the second and third feature to cross relaxation between the NV center and the VH$^-$ and WAR1 defect.
 Scans along the [111] direction shown in the Supplementary Material also corroborate this conclusion.
@@
 1.
-This situation can manifest itself when the nuclear spin of $^{13}$C atoms are a few shells distant from the nitrogen-vacancy centers.
+This situation can manifest itself when the nuclear spin of $^{13}$C atoms are a few shells away from the nitrogen-vacancy centers.
 A strong hyperfine coupling rate of 130 MHz can for instance be reached with a $^{13}$C that is only one shell away from the NV center.
@@
 Comparing the area of the second and third peak to the first one, we find [VH$^-$]$\approx$195~ppb and [WAR1]$\approx$124~ppb.
-Using areas is here more relevant than using amplitudes because inhomogeneous spread dominates over the spin-spin interaction.
-Note that the  WAR1 concentration is probably undervalued due to state mixing in the NV centers at the NV-WAR1 CR.
+Using areas is here more relevant than using amplitudes because inhomogeneous spread overwhelms the spin-spin interaction.
+Note that the WAR1 concentration is probably underestimated due to state mixing in the NV centers at the NV-WAR1 CR magnetic field.
 
 In conclusion, we demonstrated all-optical detection of paramagnetic species in a diamond grown by the Chemical-Vapor-Deposition method. 
-Using magnetic field scans along the [100] direction, we identified 3-body interactions between NV centers and $^{13}$C-NV pairs, and cross-relaxations between the spin of NV centers and the VH$^-$ and the WAR1 defects, by coupling them to NV centers.
+We identified 3-body interactions between the spin of NV centers and $^{13}$C-NV pairs as well as cross-relaxations between NV centers and the VH$^-$ and WAR1 defects.
 Our results offer prospects for more detailed studies of CVD-grown processes as well as for realizing quantum networks with these newly coupled spins.
-
-\section*{Acknowledgments}
+\begin{acknowledgments}
 We would like to thank Neil Manson and Carlos Meriles for fruitful discussions. 
 The authors would also like to thank SIRTEQ for funding. GH also acknowledges funding by the French National Research Agency (ANR) through the T-ERC project QUOVADIS. 
+\end{acknowledgments}
 

### 2021-03-12 Mes papiers/CR-CVD/SI/CVD_para_SI.tex

@@
 Fig. \ref{fig1SI} shows a scan realized on HPHT samples that have been irradiated in order to increase the NV yield. 
-The sample we use in Fig. \ref{fig1SI} is a Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range.
+The sample we use for the results shown in Fig. \ref{fig1SI} is a Ib electron irradiated HPHT diamond crystal with a NV concentration expected to be in the 5-20 ppm range.
 The PL feature that can be observed at 20G was also observed in the CVD-grown samples.
-We attributed it to the resonant dipolar interaction between a standard polarized NV center and one NV strongly coupled to the nuclear spin of a $^{13}$C atom sitting one shell away.
+We attributed it to the resonant dipolar interaction between a polarized NV center and one NV strongly coupled to the nuclear spin of a $^{13}$C atom sitting one shell away.
 Notably, no other features could be detected at larger magnetic field values.
@@
 
-\section{Fitting of the cross-relaxation scans}
+\section{Fit of the cross-relaxation features}
 \begin{figure}
@@
 \end{figure}
-In order to extract quantitative information from Fig.3 of the main text, we had ta apply two distinct fits to the raw data : first a polynomial fit to subtract the slowly decaying PL due to sate-mixing, then a Lorentzian fit on the VH$^-$ and WAR1 resonance lines to measure the central position, width and amplitude of the lines.
+In order to extract quantitative information from Fig.3 of the main text, we apply two distinct fits to the raw data : first we employ a polynomial fit to subtract the slowly decaying PL due to state-mixing, then we use a Lorentzian fit on the VH$^-$ and WAR1 resonance lines to measure the central position, the width and the contrast of the features.
 
-For the subtracting fit, we chose a polynomial fit because numerical solving of the transverse-field dependent spin depolarization could not properly fit our results.
-We had to use up a 4th order polynomial in order to obtain a tight fit to the data.
+We chose a polynomial fit in Fig.
+3 because numerical solving of the transverse-field dependent spin depolarization with the 7 levels of the NV centers could not properly fit our results.
+We had to go until a 4th order polynomial in order to obtain a tight fit to the data.
 Higher order polynomials did not significantly change the fitting envelope.
 
-The second and third peaks were fitted better on their wings by a Gaussian than by a Lorentzian lineshape, as can be seen in Fig.
-\ref{Gauss_vs_Lor} where the Gaussian fit yields a $R^2=0.983$ greater than the $R^2=0.974$ from the Lorentzian fit.
-This is coherent with the fact that the inhomogeneous broadening of the various spin species dominates the coupling strength between the spins.
+The second and third peaks were fitted better when using a Gaussian rather by a Lorentzian lineshape, as can be seen in Fig.
+\ref{Gauss_vs_Lor} where the Gaussian fit yields a $R^2=0.983$ greater than the $R^2=0.974$ from the Lorentzian fit (the difference is best seen on their wings).
+This is consistent with the fact that the inhomogeneous broadening of the various spin species dominates the coupling strength between the spins.
 
@@
 It is a thick (100) oriented diamond grown in-house by CVD with an initial concentration [N]$\approx$25~ppm and separated from its HPHT substrate.
-It was electron irradiated at 10MeV and a fluence of $2\cdot 10^{18}$~cm$^{-2}$ and in situ annealed at 900°C during irradiation to give a final concentration [NV]$\approx$4.6~ppm.
+It was electron irradiated at 10MeV and a fluence of $2\cdot 10^{18}$~cm$^{-2}$ and in situ annealed at 900$^\circ$C during irradiation to give a final concentration [NV]$\approx$4.6~ppm.
 
 The HPHT sample used in Sec.I of the supplementary material is a thick (111)-oriented diamond from Sumitomo and was electron irradiated at an energy of 2 MeV and a fluence of $7\cdot 10^{18}$~cm$^{-2}$.
-It was then annealed in vacuum for 2h at 950°C.
+It was then annealed in vacuum for 2h at 950$^\circ$C.
 
@@
 \newblock Estimation of vector static magnetic field by a nitrogen-vacancy
-  center with a single first-shell 13c nuclear (nv--13c) spin in diamond.
-\newblock {\em Chinese Physics B}, 27(5):057601, 2018.
+  center with a single first-shell 13c nuclear (nv--13c) spin in diamond.\newblock {\em Chinese Physics B}, 27(5):057601, 2018.
 
@@
 \bibitem{lazda2020cross}
-Reinis Lazda, Laima Busaite, Andris Berzins, Janis Smits, Marcis Auzinsh,
-  Dmitry Budker, Ruvin Ferber, and Florian Gahbauer.
+Reinis Lazda, Laima Busaite, Andris Berzins, Janis Smits, Marcis Auzinsh,  Dmitry Budker, Ruvin Ferber, and Florian Gahbauer.
 \newblock Cross-relaxation studies with optically detected magnetic resonances
@@
 \bibitem{simanovskaia_sidebands_2013}
-Maria Simanovskaia, Kasper Jensen, Andrey Jarmola, Kurt Aulenbacher, Neil
-  Manson, and Dmitry Budker.
+Maria Simanovskaia, Kasper Jensen, Andrey Jarmola, Kurt Aulenbacher, Neil  Manson, and Dmitry Budker.
 \newblock Sidebands in optically detected magnetic resonance signals of
@@
 Gonzalo~A. Álvarez, Christian~O. Bretschneider, Ran Fischer, Paz London, Hisao
-  Kanda, Shinobu Onoda, Junichi Isoya, David Gershoni, and Lucio Frydman.
-\newblock Local and bulk {13C} hyperpolarization in nitrogen-vacancy-centred
+  Kanda, Shinobu Onoda, Junichi Isoya, David Gershoni, and Lucio Frydman.\newblock Local and bulk {13C} hyperpolarization in nitrogen-vacancy-centred
   diamonds at variable fields and orientations.

### 2021-04-13 Mes papiers/CR-CVD/eproof_answer.tex
tutu

\documentclass{article}
\usepackage{physics}
\usepackage[]{quoting}

\begin{document}

\begin{enumerate}
\item   Please clarify what “it” refers to in “is that it can be optically.” NV-?

"it" refers to the NV- spin
\item   Please clarify what “it” refers to in “It would, indeed, offer the.” “This technology”?

"it" refers to this technology
\item   Please define VH-and WAR1

VH- corresponds to the negatively charged vacancy-hydrogen center

WAR1 corresponds to the first unidentified diamond spin defect found at the University of Warwick
\item   Please clarify what “this” refers to in “This could otherwise result in.” “This projection”?

"this" refers to "this method"
\item   The asterisk was deleted from Table I because it did not seem to correspond to anything. If it had a mean, Please define.and check formating of the table

the asterisk had no meaning
\item   Correct to use 5 rather than V for the section of the SM to match the other references. Please check.

It is correct to use 5 rather than V (note pour Gabriel : en vrai le SI est écrit en chiffre romain, est-ce que ça vaut le coup de le signaler (et de changer toutes les refs sauf celle-ci))
\item   Please define NVH

NVH stands for the nitrogen-vacancy-hydrogen center
\item   Please define ZFS

ZFS stands for zero-field splitting
\item   Please provide page/citation number for Ref. [4]

The page numbers are 131-156.
\item   Please update Ref. [15] with a published article if possible

ref. [15] has not yet been published in a peer-reviewed journal
\item   Please provide a description of the Supplemental Material in Ref. [21]

"for further experimental and simulation details"
\item   Please update Ref. [32] if possible.

Ref. [32] has been accepted in PRB but has not yet been published
\item   FQ: This funding provider could not be uniquely identified during our search of the FundRef registry (or no Contract or Grantnumber was detected). Please check information and amend if incomplete or incorrect.

Je te laisse répondre Gabriel

\end{enumerate}

\end{document}

### 2021-05-21 Mes papiers/CR-CVD/CVD_para.tex

@@
 This means the transition energies of the different classes of NV centers cannot cross under a B field scan along this direction.
-This could otherwise result in several low field PL features under arbitrary angles \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the Supplementary Materials for data showing scans along the [111] direction).
+This could otherwise result in several low field PL features under arbitrary angles \cite{van_oort_optically_1991, akhmedzhanov_magnetometry_2019} that could mask cross-relaxation from other species and also reduce the CR contrast (see section 4 of the Supplementary Materials \cite{SM_CR_CVD} (SM) for data showing scans along the [111] direction).
+\nocite{simanovskaia_sidebands_2013, alvarez_local_2015, jiang2018estimation, kamp2018continuous, Alfasi, lazda2020cross, TALLAIRE2020421}
 The downside of this choice of the [100] direction is that at large amplitudes, the transverse component of the magnetic field depolarizes all classes of NVs, thus limiting the magnetic field range that can be employed and restricting this detection method to paramagnetic defects with zero-field splittings close to the NV center's.
@@
 To let the three salient features detach better from the spectrum, we fitted a 4th-order polynomial to the PL evolution, without the spectral bumps, and subtract it from the data.
-Further details on the fitting choices are provided in Sec.V of the Supplementary material.
+Further details on the fitting choices are provided in Sec.V of the SM \cite{SM_CR_CVD}.
 
-In order to attribute the three features to their respective defects, we run a similar scan (Section 1 in the Supplementary Materials) on a type Ib electron irradiated HPHT diamond crystal with a NV concentration in the 5-20 ppm range.
+In order to attribute the three features to their respective defects, we run a similar scan (Section 1 in the SM \cite{SM_CR_CVD}) on a type Ib electron irradiated HPHT diamond crystal with a NV concentration in the 5-20 ppm range.
 Only the first feature appeared.
@@
 We obtain a factor of 6 improvement over the precision reported in \citep{cruddace2007magnetic} and attribute the second and third feature to cross relaxation between the NV center and the VH$^-$ and WAR1 defect.
-Scans along the [111] direction shown in the Supplementary Material also corroborate this conclusion.
+Scans along the [111] direction shown in the SM \cite{SM_CR_CVD} also corroborate this conclusion.
 
@@
 The resulting pair could then be coupled resonantly to a nearby polarized NV.
-The obtained eigenfrequencies of the dipolar coupled P$_1$-NV pair that we extracted were all inconsistent with the spectral positions of our observed peak within our spectral resolution error ($\approx$1G) (see Section 3 of the Supplementary Materials).
+The obtained eigenfrequencies of the dipolar coupled P$_1$-NV pair that we extracted were all inconsistent with the spectral positions of our observed peak within our spectral resolution error ($\approx$1G) (see Section 3 of the SM \cite{SM_CR_CVD}).
 This leads us to consider instead the nuclear spin of the $^{13}$C atoms as the most likely candidate.
@@
 In contrast, higher order shells have a maximum hyperfine coupling of 15 MHz \cite{smeltzer201113c} making them hard to distinguish from the inhomogeneous broadening.
-The modeling of this interaction is presented in the section 2 of the Supplementary Materials.
+The modeling of this interaction is presented in the section 2 of the SM \cite{SM_CR_CVD}.
 The dotted lines in Fig.
@@
 %Control: production of eprint (0) enabled
-\begin{thebibliography}{27}%
+\begin{thebibliography}{32}%
 \makeatletter
@@
   {NoStop}%
+\bibitem [{SM_()}]{SM_CR_CVD}%
+  \BibitemOpen
+  \href@noop {} {\emph {\bibinfo {title} {See Supplemental Material at [URL
+  will be inserted by publisher], which includes Refs. [22-26]}}}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Simanovskaia}\ \emph {et~al.}(2013)\citenamefont
+  {Simanovskaia}, \citenamefont {Jensen}, \citenamefont {Jarmola},
+  \citenamefont {Aulenbacher}, \citenamefont {Manson},\ and\ \citenamefont
+  {Budker}}]{simanovskaia_sidebands_2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Simanovskaia}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {Jensen}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}},
+  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Aulenbacher}}, \bibinfo
+  {author} {\bibfnamefont {N.}~\bibnamefont {Manson}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }\href {\doibase
+  10.1103/PhysRevB.87.224106} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. B}\ }\textbf {\bibinfo {volume} {87}},\ \bibinfo {pages} {224106}
+  (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Álvarez}\ \emph {et~al.}(2015)\citenamefont
+  {Álvarez}, \citenamefont {Bretschneider}, \citenamefont {Fischer},
+  \citenamefont {London}, \citenamefont {Kanda}, \citenamefont {Onoda},
+  \citenamefont {Isoya}, \citenamefont {Gershoni},\ and\ \citenamefont
+  {Frydman}}]{alvarez_local_2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.~A.}\ \bibnamefont
+  {Álvarez}}, \bibinfo {author} {\bibfnamefont {C.~O.}\ \bibnamefont
+  {Bretschneider}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Fischer}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {London}},
+  \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Kanda}}, \bibinfo
+  {author} {\bibfnamefont {S.}~\bibnamefont {Onoda}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Gershoni}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {L.}~\bibnamefont {Frydman}},\ }\href {\doibase 10.1038/ncomms9456}
+  {\bibfield  {journal} {\bibinfo  {journal} {Nat Commun}\ }\textbf {\bibinfo
+  {volume} {6}},\ \bibinfo {pages} {8456} (\bibinfo {year} {2015})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Jiang}\ \emph {et~al.}(2018)\citenamefont {Jiang},
+  \citenamefont {Ye}, \citenamefont {Jiao}, \citenamefont {Huang},\ and\
+  \citenamefont {Lv}}]{jiang2018estimation}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.-J.}\ \bibnamefont
+  {Jiang}}, \bibinfo {author} {\bibfnamefont {J.-F.}\ \bibnamefont {Ye}},
+  \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Jiao}}, \bibinfo {author}
+  {\bibfnamefont {Z.-Y.}\ \bibnamefont {Huang}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {H.-J.}\ \bibnamefont {Lv}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Chinese Physics B}\ }\textbf {\bibinfo
+  {volume} {27}},\ \bibinfo {pages} {057601} (\bibinfo {year}
+  {2018})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kamp}\ \emph {et~al.}(2018)\citenamefont {Kamp},
+  \citenamefont {Carvajal},\ and\ \citenamefont
+  {Samarth}}]{kamp2018continuous}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
+  {Kamp}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Carvajal}}, \
+  and\ \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Samarth}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Physical Review
+  B}\ }\textbf {\bibinfo {volume} {97}},\ \bibinfo {pages} {045204} (\bibinfo
+  {year} {2018})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Lazda}\ \emph {et~al.}(2020)\citenamefont {Lazda},
+  \citenamefont {Busaite}, \citenamefont {Berzins}, \citenamefont {Smits},
+  \citenamefont {Auzinsh}, \citenamefont {Budker}, \citenamefont {Ferber},\
+  and\ \citenamefont {Gahbauer}}]{lazda2020cross}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Lazda}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Busaite}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Berzins}}, \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Smits}}, \bibinfo {author}
+  {\bibfnamefont {M.}~\bibnamefont {Auzinsh}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Budker}}, \bibinfo {author} {\bibfnamefont
+  {R.}~\bibnamefont {Ferber}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {F.}~\bibnamefont {Gahbauer}},\ }\href@noop {} {\bibfield  {journal}
+  {\bibinfo  {journal} {arXiv preprint arXiv:2007.00473}\ } (\bibinfo {year}
+  {2020})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {van Oort}\ and\ \citenamefont
@@
   {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Simanovskaia}\ \emph {et~al.}(2013)\citenamefont
-  {Simanovskaia}, \citenamefont {Jensen}, \citenamefont {Jarmola},
-  \citenamefont {Aulenbacher}, \citenamefont {Manson},\ and\ \citenamefont
-  {Budker}}]{simanovskaia_sidebands_2013}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Simanovskaia}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
-  {Jensen}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}},
-  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Aulenbacher}}, \bibinfo
-  {author} {\bibfnamefont {N.}~\bibnamefont {Manson}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }\href {\doibase
-  10.1103/PhysRevB.87.224106} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
-  Rev. B}\ }\textbf {\bibinfo {volume} {87}},\ \bibinfo {pages} {224106}
-  (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Smeltzer}\ \emph {et~al.}(2011)\citenamefont
@@
   (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
-
 \end{thebibliography}%
-
 \end{document}

### 2021-05-21 Mes papiers/Review meca/Ensemble NV.tex

\documentclass[a4paper]{article}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}

\begin{document}
	While the optimal density of NV centers is generally thought in term of a compromise between sensitivity and  the spins coherent time $T_2^*$, other more exotic effects start to appear when the spin density reaches a critical point.

	For concentration higher than $\sim$ 1~ppm (corresponding to a mean distance between spins of $\approx$ 15~nm or a dipolar coupling strength of $\approx$ 15~KHz), the dipolar coupling between the NV spins starts to play an important role. One effect of importance is the modification of the spin lifetime $T_1$ through dipolar coupling with other NV centers \cite{Jarmola_temperature_2012,mrozek_longitudinal_2015,choi_depolarization_2017}. This particular effect has been used with a levitating diamond in a Paul trap to observe a resonant change in the spins magnetic susceptibility \cite{pellet2021magnetic}.

	Other collective effects between NV centers include the cooperative enhancement of the dipole force of the NV centers, a phenomenon similar to that of supperradiance described in \cite{Bachelard,PANAT,Venkatesh}, and observed with a levitating diamond in an optical tweezer by Juan et al. \cite{Juan}.

    \bibliographystyle{plain}
	\bibliography{Biblio_ensemble}

  \end{document}

### 2021-05-21 Mes papiers/Review meca/article.tex

--- 2020-07-10 théorie/fiches/Signal et bruit/Signal et bruit.tex
@@
-\documentclass[a4paper]{report}
+\documentclass[a4paper]{article}
 \usepackage[french]{babel}
@@
 \begin{document}
+  \pagenumbering{gobble}
+  \maketitle
+  \newpage
+  \pagenumbering{arabic}  
   \tableofcontents
-\chapter{Signal et bruit}
-  \section{Conventions, définitions}
-  \subsection{Transformée de Fourier}
-  On prend la définition "du physicien". Soit $s(t)$ un signal, alors ; 
-  \begin{align}
-  \tilde{s}_1(\omega)&=\int_{-\infty}^{+\infty} s(t) e^{-i\omega t} dt \\
-  s(t)&=\int_{-\infty}^{+\infty}\tilde{s}_1(\omega) e^{i\omega t} \dfrac{d\omega}{2\pi}
-  \end{align}
-  On peut également choisir la convention :
-  \begin{align}
-  \tilde{s}_2(\nu)&=\int_{-\infty}^{+\infty} s(t) e^{-i2\pi \nu t} dt \\
-  s(t)&=\int_{-\infty}^{+\infty}\tilde{s}_2(\nu) e^{i2\pi \nu t} d\nu
-  \end{align}
-  Et dans ce cas là tu as simplement $\tilde{s}_1(\omega)=\tilde{s}_2(2\pi \nu)$. Par contre les $\sqrt{2\pi}$ on oublie.
-
-  Dans tous les cas, ton spectre à pour dimension $[\tilde{s}]=[s]\cdot[Hz]^{-1}$
-  \subsection{Propriétés de la TF}
-  \begin{itemize}
-  \item Translation : \begin{equation}
-  \tilde{f}(t-\tau)[\omega]=e^{-i\omega \tau} \tilde{f}(t)[\omega]
-  \end{equation}
-  \item Scaling : \begin{equation}
-  \tilde{f}(a\cdot t)[\omega]= \dfrac{1}{|a|} \tilde{f}(t)[\dfrac{\omega}{a}]
-  \end{equation}
-  En particulier, $\tilde{f}(-t)[\omega]=\tilde{f}(t)[-\omega]$. C'est la "time reversibility"
-  \item $f$ réelle $\implies$ $\tilde f$ Hermitienne, soit $\tilde f^*(\omega)= \tilde f(-\omega)$
-  \item Inversement, si $f$ est hermitienne (donc paire pour une fonction réelle), $\tilde{f}$ est réelle.
-  \item Convolution : 
-  \begin{equation}
-  TF[(f*g)(\tau](\omega)=\tilde{f}(\omega)\tilde{g}(\omega)
-  \end{equation}
-  \item Auto-corrélation : \begin{align*}
-  \tilde S_{ff}(\omega)&=TF[(f(t)*f^*(-t))] \\
-  &=TF[f(t)] \cdot TF[f^*(-t)] \\
-  &= TF[f(t)] \cdot TF[f(t)]^* \\
-  &= |\tilde{f}(\omega)|^2
-  \end{align*}
-
-  \end{itemize}
-  \subsection{Transformée de Fourier tronquée}
-  L'infini c'est long, surtout dans le négatif. Donc expérimentalement on va plutôt utiliser la TF tronquée : 
-  \begin{equation}
-  \hat{s}(\omega)=\dfrac{1}{\sqrt{T}}\int_{0}^{T} s(t) e^{-i\omega t} dt
-  \end{equation}
-  \subsection{Transformée de Fourier discrète}
-  \subsection{Produit de convolution}
-  \begin{equation}
-  (f*g)(\tau)=\int_{-\infty}^{+\infty} f(\tau-t) g(t) dt = \int_{-\infty}^{+\infty} f(t) g(\tau-t) dt
-  \end{equation}
-  \subsection{Fonction d'auto-corrélation}
-  La fonction d'auto-corrélation d'un signal est définie comme :
-  \begin{equation}
-  R_{ss}(t_1,t_2)=\langle s(t_1)s^*(t_2)\rangle
-  \end{equation}
-  où $\langle ...\rangle$ représente a priori une moyenne d'ensemble.
-  Mais nous on aime bien les processus Markovien, donc $R_{ss}$ n'est plus fonction que de $\tau=|t_1-t_2|$.
-En plus de ça, on fait une petite hypothèse ergodique est la moyenne d'ensemble devient une moyenne temporelle, et on oublie au passage le facteur de normalisation Alors :
-  \begin{equation}
-  R_{ss}(\tau)=\int_{-\infty}^{+\infty} s(t) s^*(t-\tau)= (s(t)*s^*(-t))(\tau)
-  \end{equation}
-  Si jamais ton signal est réel, la fonction d'auto-corrélation est paire, donc $R_{ss}(\tau)=R_{ss}(-\tau)$. Si ton signal est complexe c'est un peu plus chiant.
-
-  La \textbf{Fonction d'auto-covariance} c'est simplement la fonction d'auto-corrélation à la quelle tu soustrais $\langle s(t)\rangle ^2$.
-  \subsection{Densité spectrale de puissance}
-  On peut définir proprement la DSP à partir de la TF tronquée : 
-  \begin{equation}
-  S_{ss}(\omega)=\lim_{T \to \infty} \langle |\hat{s}(\omega)|^2 \rangle
-  \end{equation}
- Or, \begin{align*}
- \langle |\hat{s}(\omega)|^2 \rangle &=\langle \hat{s}(\omega)\hat{s^*}(\omega) \rangle \\
- &=\langle \dfrac{1}{T} \int_{0}^{T} s(t) e^{-i\omega t} dt \int_{0}^{T} s^*(t') e^{i\omega t'} dt' \rangle\\
- &=\dfrac{1}{T} \int_{0}^{T} \int_{0}^{T}\langle s(t)s^*(t') \rangle e^{i\omega(t-t')} dt\; dt'
- \end{align*}
- Et donc tu vois que, en posant $\tau=t-t'$ et en bidouillant un peu les bornes d'intégrations, tu te retrouves avec :
- \begin{equation}
- S_{ss}(\omega)=\int_{-\infty}^\infty R_{ss}(\tau) e^{-i\omega\tau}\,d \tau=\tilde R_{ss}(\omega)=|\tilde s (\omega)|^2
- \end{equation}
- La densité spectrale de puissance est la TF de la fonction d'auto-corrélation (pour un processus Markovien). C'est le théorème de \textbf{Wiener–Khinchin}.
-
- La version plus sale c'est d'écrire la DSP comme :
- \begin{equation}
- \langle \tilde{s}(\omega) \tilde{s}^*(\omega') \rangle = S_{ss}(\omega)2\pi  \delta(\omega-\omega')
- \end{equation}
-
- \section{Résolution d'un Michelson}
- Michelson mais ça marche pour toutes les interférences à deux ondes (Ramsay entre autre).
-
- \subsection{Réponse spectrale}
- On va commencer par la réponse spectrale d'un Micheslon, c'est à dire le rapport des densités spectrales de puissance en entrée ($I_{in}(\omega)$) et en sortie ($I_{out}(\omega)$) de l'interféromètre.
-On note $A$ les amplitudes correspondantes, et $\tau$ le retard dans l'interféromètre.
-Alors :
- \begin{equation}
- I_{in}(\omega)=\tilde{A}_{in}(\omega)\tilde{A}_{in}^*(\omega)=|\tilde{A}_{in}(\omega)|^2=|TF[A_{in}(t)](\omega)|^2
- \end{equation}
- et 
- \begin{align*}
-I_{out}(\omega)&= |TF[A_{out}(t)](\omega)|^2 \\
- &= |TF[\dfrac{1}{2} (A_{in}(t)+A_{in}(t-\tau)](\omega)|^2 \\
- &= \left|TF[(A_{in}(t)](\omega)\left(\dfrac{1+e^{i\omega \tau}}{2}\right)\right|^2 \\
- &= I_{in}(\omega) \left(\dfrac{1+e^{i\omega \tau}}{2}\right)^2 \\
- &= I_{in}(\omega) \left(\dfrac{1+\cos \omega tau}{2}\right)
- \end{align*}
-
- Remarque : le 1/2 dans le $A_{out}(t)$ vient du fait que tu as deux séparatrices, qui ajoutent $1/\sqrt{2}$ chacune.
-
- On trouve donc l'intervalle spectral libre : \begin{equation}
- \Delta \nu = \dfrac{1}{\tau} = \dfrac{(2)l}{c}
- \end{equation}
-
- En résonnant maintenant en nombre de particules (photon ou autre), en supposant que le flux incident comporte $N$ particules et le flux sortant (flux transmis, ou nombre d'atomes excités dans le cas de Ramsay) est $N_e$, on a tout simplement \begin{equation}
- N_e=N\left(\dfrac{1+\cos \Phi}{2}\right)
- \end{equation}
-Avec $\Phi$ le déphasage entre les deux voies.
-\subsection{Pouvoir de résolution}
-On va maintenant considérer que $N_e$ est une variable aléatoire, chaque particule ayant une chance $p_e=\dfrac{1+\cos \Phi}{2}$ de passer. On a donc :
-\begin{align}
-\langle N_e\rangle &= N\left(\dfrac{1+\cos \Phi}{2}\right) \\
-\Delta^2 N_e &= N(p_e)(1-p_e) = N\dfrac{\sin ^2 \Phi}{4} \\
-\Delta N_e &=\sqrt{N} \dfrac{|\sin \Phi|}{2}
-\end{align}
-
-Maintenant pour s'intéresser au pouvoir de résolution, on va considérer un résultat $N_{e1}$ de $N$ particules à $\omega$ (où $\Delta$ pour un Ramsay) et un autre résultat $N_{e2}$ de $N$ particules à $\omega + \delta \omega$.
-Pour pouvoir distinguer le résultat entre les deux pulsations, il faut que la différence des valeurs moyennes soit plus grande que ($\sqrt{2}$ fois) l'écart type du résultat.
-($\sqrt{2}$ car il y a deux variables aléatoires dont tu fais la différence, donc tu sommes les variances, un peu comme une mesure à la règle.
-Mais en vrai c'est du chipotage).
-D'où : \begin{equation}
-\langle N_e (\omega+ \delta \omega)\rangle - \langle N_e (\omega)\rangle \approx -N\tau\dfrac{\sin \omega \tau}{2} \geq \sqrt{2N} \dfrac{\sin \omega \tau}{2} 
-\end{equation}
-
-D'où finalement \begin{equation}
-\delta \omega \geq \dfrac{\sqrt{2}}{\sqrt{N}\tau}
-\end{equation}
-
-On constate que le pouvoir de résolution d'un Michelson ne dépend pas de la phase : Le signal et le bruit sont tous les deux plus importants quand ta phase vaut $\pi/2$
-
-On remarque également qu'on a une limite en $1/\tau$, qui vient de la transformée de Fourier / Heisenberg temps/énergie, et en $1/\sqrt{N}$, qui est la limite "classique" pour tout processus stochastique à N particules.
-En utilisant des états quantiques à N particules (intriquées donc) on peut idéalement monter à des résolutions en $1/N$.
+  \newpage
+  \section{Introduction}
+  \section{Bibliographie} % Ca marche si t'as un fichier biblio.bib dans le même dossier
+	\bibliographystyle{plain}
+	\bibliography{biblio}
   \end{document}

### 2021-06-29 data/20210614/250 PPM N2O/traitement.py
tutu

@@ -1,0 +1,357 @@
+import sys
+import glob
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit,root_scalar
+from scipy import fftpack
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1,exclude_neg=True):
+
+	if filename[-4]!='.' :
+		if glob.glob(filename+'.txt') :
+			f=glob.glob(filename+'.txt')[0]
+		elif glob.glob(filename+'.csv') :
+			f=glob.glob(filename+'.csv')[0]
+		else :
+			print('file not found')
+			quit()
+		filename=f
+	if filename[-4:] =='.txt' :
+		x=[]
+		y=[]
+		with open(filename,'r',encoding = "ISO-8859-1") as f:
+			for line in f :
+				line=line.split()
+				try :
+					if exclude_neg :
+						if float(line[xcol])!=-1 :
+							x+=[float(line[xcol])]
+							y+=[float(line[ycol])]
+					else :
+						x+=[float(line[xcol])]
+						y+=[float(line[ycol])]				
+				except :
+					pass
+		return(np.array(x),np.array(y))
+	elif filename[-4:] =='.csv' :
+		with open(filename,'r',encoding = "ISO-8859-1") as f:
+			content=f.readlines()
+			x=content[xcol]
+			x=x.split()
+			x=[np.float(item) for item in x]
+			y=content[ycol]
+			y=y.split()
+			y=[np.float(item) for item in y]
+		return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def quad_fit(x,y) :
+	A=np.vstack([x**2,x,np.ones(len(x))]).T
+	a,b,c = np.linalg.lstsq(A, y, rcond=None)[0]
[…]

### 2021-06-29 data/20210614/50 ppm N2/traitement ESR.py

@@ -1,0 +1,327 @@
+import sys
+import glob
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import curve_fit,root_scalar
+from scipy import fftpack
+from PyQt5.QtWidgets import QApplication,QFileDialog
+
+
+
+
+def extract_data(filename,xcol=0,ycol=1,exclude_neg=True):
+
+	if filename[-4]!='.' :
+		if glob.glob(filename+'.txt') :
+			f=glob.glob(filename+'.txt')[0]
+		elif glob.glob(filename+'.csv') :
+			f=glob.glob(filename+'.csv')[0]
+		else :
+			print('file not found')
+			quit()
+		filename=f
+	if filename[-4:] =='.txt' :
+		x=[]
+		y=[]
+		with open(filename,'r',encoding = "ISO-8859-1") as f:
+			for line in f :
+				line=line.split()
+				try :
+					if exclude_neg :
+						if float(line[xcol])!=-1 :
+							x+=[float(line[xcol])]
+							y+=[float(line[ycol])]
+					else :
+						x+=[float(line[xcol])]
+						y+=[float(line[ycol])]				
+				except :
+					pass
+		return(np.array(x),np.array(y))
+	elif filename[-4:] =='.csv' :
+		with open(filename,'r',encoding = "ISO-8859-1") as f:
+			content=f.readlines()
+			x=content[xcol]
+			x=x.split()
+			x=[np.float(item) for item in x]
+			y=content[ycol]
+			y=y.split()
+			y=[np.float(item) for item in y]
+		return(np.array(x),np.array(y))
+
+def lin_fit(x,y) :
+	A=np.vstack([x,np.ones(len(x))]).T
+	a,b = np.linalg.lstsq(A, y, rcond=None)[0]
+	return(a,b,a*x+b)
+
+def quad_fit(x,y) :
+	A=np.vstack([x**2,x,np.ones(len(x))]).T
+	a,b,c = np.linalg.lstsq(A, y, rcond=None)[0]
[…]

### 2021-06-29 python_clément/TDV_laser_pulsed_soustraction.py

@@ -1,0 +1,456 @@
+import sys
+import os
+import time
+import random
+import nidaqmx
+import nidaqmx.stream_writers
+import nidaqmx.stream_readers
+import nidaqmx.task
+import nidaqmx.system
+
+import numpy as np
+
+
+from PyQt5.QtGui import QFont
+from PyQt5.QtCore import (Qt, QTimer)
+from PyQt5.QtWidgets import (QWidget, QPushButton, 
+	QHBoxLayout, QVBoxLayout, QApplication, QDesktopWidget, QMainWindow, QLineEdit, QLabel, QCheckBox, QFileDialog)
+
+from matplotlib.backends.backend_qt5agg import (FigureCanvas, NavigationToolbar2QT as NavigationToolbar)
+from matplotlib.figure import Figure
+
+
+#This is a photon counter for the NI 6341 card when the TTL signal is plugged into ctr0/source which is on pin PFI8/P2.0/81
+
+class Photon_Counter(QMainWindow):
+	def __init__(self):
+		super().__init__()
+		
+
+
+		self.f_acq=1E5
+		self.t_acq=15E-3
+
+
+		self.f_pulse=5E6
+		self.t_pulse=1E-3
+		self.t_lect=4E-4
+
+		self.n_points=301
+
+		self.f_uw=2800 #MHz
+		self.level_uw=10 #dB
+		self.t_uw=10 #us
+
+
+		self.n_lect_min=0
+		self.n_lect_max=-1
+
+		self.time_last_refresh=time.time()
+		self.refresh_rate=0.1
+		
+		self.setWindowTitle("T1")
+
+		##Creation of the graphical interface##
+
+		self.main = QWidget()
+		self.setCentralWidget(self.main)
+
+		layout= QHBoxLayout()
+		Vbox = QVBoxLayout()
+		Vbox_gauche=QVBoxLayout()
+		Vbox_droite=QVBoxLayout()
+
+		layout.addLayout(Vbox_gauche)
+		layout.addLayout(Vbox)
+		layout.addLayout(Vbox_droite)
+		self.main.setLayout(layout)
+
+		#Fields on the left
+
+		self.labelf_acq=QLabel("f_acq(max 500 KHz)")
+		self.lectf_acq=QLineEdit('%3.2E'%self.f_acq)
+		Vbox_gauche.addWidget(self.labelf_acq)
+		Vbox_gauche.addWidget(self.lectf_acq)
+		
[…]

### 2021-06-29 python_clément/test.py

@@ -1212,5 +1212,14 @@ def test_laser_pulsed_2():
 			for i in range(10):
 				read.read(100)
 
+def mesure_x_et_y():
+	with nidaqmx.Task() as tension :
+		tension.ai_channels.add_ai_voltage_chan("Dev1/ai11",min_val=-10,max_val=10)
+		tension.ai_channels.add_ai_voltage_chan("Dev1/ai5",min_val=-10,max_val=10)
+		tension.timing.cfg_samp_clk_timing(1E3,sample_mode=nidaqmx.constants.AcquisitionType.FINITE,samps_per_chan=100)
+		tension.start()
+		lecture=tension.read(100)
+	print(lecture[1])
 
+mesure_x_et_y()
 

### 2021-10-08 Talks/group meeting 2021/presentation.tex
weekend

--- 2020-12-04 Talks/Défaut 2020/presentation.tex
@@
 
-\title{Mechanical and luminescence-based detection of dipolar-interactions between spins in diamond}
-\author{Clément Pellet-Mary\\ Laboratoire de physique de l'ENS\\ \textit{ENS, Paris}}
+\title{Group meeting : Cross-relaxation with NV centers ensemble in diamond}
+%\author{Clément Pellet-Mary\\ Laboratoire de physique de l'ENS\\ \textit{ENS, Paris}}
 \date\today
@@
 \end{frame}
-\begin{frame}
-\includegraphics[scale=.3]{Shémas_intro}
-\end{frame}
-\section{NV center and Cross-Relaxations}
-\begin{frame}{Outline}
-\tableofcontents[currentsection]
-\end{frame}
-\begin{frame}{Spin Hamiltonian and orientation of the centers}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\includegraphics[scale=.25]{projection_4_classes_0}
-    \onslide<2>\includegraphics[scale=.25]{projection_4_classes_1}
-    \end{overprint}
-\end{figure}
-\end{frame}
-\begin{frame}{ODMR with an ensemble of NV centers}
-\includegraphics[scale=.65]{ESR_4spin_simu}
-\end{frame}
-\begin{frame}{Non-Resonant dipolar interaction}
-\centering
-\includegraphics[scale=.35]{CR_non_resonnant}
-\end{frame}
-\begin{frame}{Resonant dipolar interaction : Cross-Relaxation}
-\centering
-\includegraphics[scale=.35]{CR_resonnant}
-\end{frame}
-\begin{frame}{Exemple of Cross-Relaxation NV-P1}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\includegraphics[scale=.30]{CR_111_0}\footfullcite{armstrong_nvnv_2010}
-    \onslide<2>\includegraphics[scale=.30]{CR_111_05}
-    \onslide<3>\includegraphics[scale=.30]{CR_111_1}
-    \onslide<4>\includegraphics[scale=.30]{CR_111_2}
-    \end{overprint}
-\end{figure}
-\end{frame}
-\section{Cross-relaxations with new spin species}
-\begin{frame}{Outline}
-\tableofcontents[currentsection]
-\end{frame}
-\begin{frame}{Motivations of the study}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\includegraphics[scale=.30]{Motivations_0}
-    \onslide<2>\includegraphics[scale=.30]{Motivations_1}\footfullcite{tallaire2020high}
-    \onslide<3>\includegraphics[scale=.30]{Motivations_2}
-    \onslide<4>\includegraphics[scale=.30]{Motivations_3}
-    \end{overprint}
-\end{figure}
-\end{frame}
-\begin{frame}{Photoluminescence change with $\vec B$ in the [100] direction}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\centering\includegraphics[scale=.45]{Scan_100_brut_0}
-    \onslide<2>\centering\includegraphics[scale=.45]{Scan_100_brut_1}
-    \onslide<3>\centering\includegraphics[scale=.45]{Scan_100_brut_2}
-    \end{overprint}
-\end{figure}
-\end{frame}
-\begin{frame}{Cross-relaxation between NV$^-$ and VH$^-$}
-\includegraphics[scale=.48]{soustraction_1}
-\end{frame}
-\begin{frame}{Cross-relaxation between NV$^-$ and WAR1}
-\includegraphics[scale=.48]{soustraction_2}
-\end{frame}
-\begin{frame}{EPR spectroscopy}
-Electron Paramagnetic Resonance : A spectroscopy technique using absorption of a microwave at a given frequency (9.5 GHz for example) as a function of magnetic field to detect paramagnetic defects.
-\bigbreak
-VH$^-$\footfullcite{glover_hydrogen_2004} and WAR1 \footfullcite{cruddace2007magnetic} have been observed by EPR spectrocopy in CVD diamonds by Mark Newton's team at the University of Warwick.
-\end{frame}
-\begin{frame}{Comparison between cross-relaxation and EPR}
-\pause
-Upsides :
+\begin{frame}{The Nitrogen Vacancy Center}
+\begin{center}
+\includegraphics[scale=.4]{Nitrogen-vacancy_center}
+\end{center}
+
 \begin{itemize}
-\item CR experiments are much simpler to setup (low $\vec B$, no microwave)
-\pause
-\item Doable on micro and potentially nano-diamonds
-\pause
-\item NV centers produce a calibration for $\vec B$ $\to$ Better precision on the ZFS measurement
-\pause
-\item Potential for hyperpolarization of new spins
-\pause
-\end{itemize}
-Downsides :
-\begin{itemize}
-\item Requires a high NV concentration
-\pause
-\item Loss of contrast in non-[100] directions
-\end{itemize}
-\end{frame}
-\section{Cross-relaxation between NV centers}
-\begin{frame}{Outline}
-\tableofcontents[currentsection]
-\end{frame}
-\begin{frame}{Experimental proofs of NV-NV cross-relaxations}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\includegraphics[scale=.28]{NV-NV_0}
-    \onslide<2>\includegraphics[scale=.28]{NV-NV_05}
-    \onslide<3>\includegraphics[scale=.28]{NV-NV_1}
-    \onslide<4>\includegraphics[scale=.28]{NV-NV_2}
-    \onslide<5>\includegraphics[scale=.28]{NV-NV_3}
-    \end{overprint}
-\end{figure}
-\end{frame}
-\begin{frame}{Microscopic explanation of the NV-NV CR}
-\centering
-\includegraphics[scale=.35]{NVNV-micro}\footfullcite{choi_depolarization_2017}
-\end{frame}
-\begin{frame}{Cross-relaxation between NV$^-$ and $^{13}$C$-$NV$^-$}
-\includegraphics[scale=.45]{soustraction_3}
-\end{frame}
-%ajouter la slide avec la carte du coup ?
-\section{Torque induced by cross-relaxation on a levitating diamond}
-\begin{frame}{Outline}
-\tableofcontents[currentsection]
-\end{frame}
-\begin{frame}{Torque measurement with a levitating diamond}
-\centering
-\includegraphics[scale=.3]{Trap_0}%"This trap not only locks the center of mass of the diamond, but it also locks its orientations, because of the anistropy of the trap and of the diamond" "The trap can be roughly modeled by an harmonic potential centered around a particular angle at equilibrium"
-\end{frame}
-\begin{frame}{Back-scattered laser detection}
-\centering
-\includegraphics[scale=.085]{Trap_3}%And how do we measure this rotation ?
-well, since we have to send a laser on the diamond to polarise the NV centers..." "We do not observe a clear reflection of the laser, instead we observe a speckle pattern, which is produced by the many interferences on the rough surface of the diamond" "Whenever the diamond moves, the speckle pattern will either shift or change its shape..."
-\end{frame}
-\begin{frame}{Measurement of the torque caused by C.R.}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\includegraphics[scale=.08]{CRmeca_121_mieux_0}
-    \onslide<2>\includegraphics[scale=.08]{CRmeca_121_mieux_1}
-    \onslide<3>\includegraphics[scale=.08]{CRmeca_121_mieux_2}
-    \end{overprint}
-\end{figure}
-\end{frame}
-\begin{frame}{Origin of the magnetic torque}
-\begin{figure}
-    \begin{overprint}
-    \onslide<1>\includegraphics[scale=.23]{Explication_torque_alalouche_0}
-    \onslide<2>\includegraphics[scale=.23]{Explication_torque_alalouche_1}
-    \onslide<3>\includegraphics[scale=.23]{Explication_torque_alalouche_2}
-    \onslide<4>\includegraphics[scale=.23]{Explication_torque_alalouche_3}
-    \onslide<5>\includegraphics[scale=.23]{Explication_torque_alalouche_4}
-    \onslide<6>\includegraphics[scale=.23]{Explication_torque_alalouche_5}
-    \end{overprint}
-\end{figure}
-\end{frame}
-\begin{frame}{Simulation of the C.R.-caused torque}
-\centering
-\includegraphics[scale=.3]{Explication_torque}
-\end{frame}
-\begin{frame}{Torque caused by CR : other configuration}
-\centering
-\includegraphics[scale=.28]{CRmeca_22}
-\end{frame}
-\begin{frame}{Conclusion}
-\begin{itemize}
-\setlength\itemsep{1em}
-\item{Detection of new spin defects in diamond through Cross-Relaxations
-\begin{itemize}
-\item Potential for hyperpolariaztion of new dark electronic spins
-\end{itemize}}
-\item{Observations of NV$-$NV Cross-Relaxations}
-\item{Mechanical detection of NV$-$NV Cross relaxations :
-\begin{itemize}
-\item Potential for mechanical detection of new spin species
-\item Potential for Resonant Einstein-De-Haas effect 
-\end{itemize} }
+\item 0D fluorescent object with ZPL at 638 nm
+\item Controllable and readable spin at room temperature (!)
+\item Working with 10$^9$ emitters (typ.)
 \end{itemize}

### 2021-10-08 python_clément/analyse.py

@@ -218,6 +218,21 @@ def stretch_et_phonons(x,y,Amp=None,tau=None,ss=None,T1ph=5E-3) :
 	popt, pcov = curve_fit(f, x, y, p0,bounds=([-np.inf,-np.inf,0],[np.inf,np.inf,np.inf]))
 	return(popt,f(x,popt[0],popt[1],popt[2]))
 
+def Rabi_fit(x,y,Amp=None,omega=None,tau=None,ss=None):
+	if not Amp :
+		Amp=max(y)-min(y)
+	if not ss :
+		ss=y[-1]
+	if not tau :
+		tau=x[int(len(x)/3)]-x[0]
+	if not omega :
+		omega=1/(x[int(len(x)/10)]-x[0])
+	def f(x,Amp,ss,tau,omega) :
+		return Amp*np.exp(-x/tau)*np.cos(omega*x)+ss
+	p0=[Amp,ss,tau,omega]
+	popt, pcov = curve_fit(f, x, y, p0)
+	return(popt,f(x,*popt))
+
 def fit_B_dipole(x,y,B0=2000,x0=10) : #x : distance en mm, y : champ mag en G
 	def f(x,B0,x0):
 		return(B0/(x+x0)**3)
@@ -325,6 +340,9 @@ def find_B_100(freq,transi='-',B_max=100) :
 
 	RR=root_scalar(f,bracket=[0,B_max])
 	return RR.root
+
+
+
 
 class NVHamiltonian(): #x,y and z axis are taken as (100) axis
 	D=2873 #Mhz

### 2021-10-09 Mes papiers/Méca CR/reviewer_answer.tex

\documentclass{article}
\usepackage{physics}
\usepackage[]{quoting}

\begin{document}

\section*{Summary of the changes}

We would first like to thank the referees for the time they spent reviewing of our manuscripts.

[Résumer la position de ref et dire qu'ils ont raison meme si c'est pas vrai]

We agree with these remarks and have amended the manuscript in consequence to the best of our abilities.

\section*{Referee 1}
\subsection*{Referee 1 report}

\begin{quote}
This manuscript describes experiments showing mechanically detected
cross-relaxation resonance of nitrogen-vacancy centers in diamond. The
measurements are carried out on NV-centers hosted in nano-diamonds
levitated in a Paul trap in vacuum. The authors motivate their work as
a means of detecting paramagnetic defects without optical transitions,
for investigating spin relaxation processes, and for cooling the
motion of mechanical oscillators.

The work appears valid and the interpretation of the measurements and
the accompanying analysis appears well-founded. Although the
presentation is sometimes difficult to follow, the authors have
included an exhaustive supplementary section spelling out the details
of the experiment.

Nevertheless, the importance and broad interest of the work --
requirements for publication in PRL -- appear to this reader to fall
below threshold. I am hard pressed to identify how this work,
"substantially advance[s] a field, open[s] a significant new area of
research, or solves[s]—or take[s] a crucial step toward solving—a
critical outstanding problem". Cross-relaxation is a well-known
phenomenon, also in NV centers. Although it appears not to have been
studied in nano-diamonds nor to have been detected mechanically, it is
not clear why these achievements solve an outstanding problem of
interest.

The authors suggest in the conclusion that the scheme can be used to
cool a nano-diamond in the Paul trap, though no such scheme is
attempted in the paper. Also, no estimates are given as to how
efficient this process would be and what the minimum achievable
temperature would be. The second motivation offered is that this
method could be used as a spectroscopic technique for sensing dipolar
interaction between NV centers and spins that are not optically
active. Again, no such demonstration experiment is presented, although
it's not clear why this should not be possible. The final motivation
given is for "bottom-up investigations of magnetism". Unfortunately,
from the description that follows this statement, it's hard to
envision what, specifically, is intended.

For these reasons, it is my opinion that the authors should resubmit
to PRB or a specialized journal of magnetic resonance.
\end{quote}

\subsection*{Our answer}
Here is our answer to the points raised by Referee 1 :

Je sais pas trop quoi répondre sachant qu'il ne parle que de la conclusion. Est-ce qu'on change la conclu (la partie bottom-up magnetism, j'avoue que je vois pas bien ce que ça cible) ? Est-ce qu'on dit qu'on a pas réussi les applications ? Sur la partie détection de spin au moins je peux justifer : les seuls autres spins qu'on sache être présent ici c'est les P1 et non seulement c'est galère d'aligner précisement un champ mag sur la 111, en plus les effets de para/diamagnétisme des NV commencent à être significatifs à 500G.

\section*{Referee 2}
\subsection*{Referee 2 report}
\begin{quote}
In this work the authors investigate using dipolar interactions
between NV centers to enhance magnetic torque and rotate
electrically-charged, paramagnetic diamond nanocrystals that are
levitated in an ion trap, through a mechanism similar to the
Einstein-de-Haas effect for ferromagnets. An external magnetic field
is tuned to produce resonant cross-relaxation, resulting in greater
paramagnetism and spin torque. Prospects for using this approach to
sense dipolar interactions between NV centers and spins that cannot be
polarized optically are interesting. Using NV centers in levitating
diamonds for emulating magnetism at the 100 $\mu_B$ level is another
potentially useful application. I feel this work is novel and
interesting however the paper requires clarifications before I can
recommend publication.

Fig 3: the caption describes “a) and b) show the torque with and
without cross-relaxation between NV centers respectively.” should this
description be reversed? I thought b) was with CR.

The experimental setup figure in the supplemental material is showing
one APD, but the figure in paper shows 2 detectors: APD1 and APD2.
This should be consistent.

A related question: how is the actual angular motion determined and
calibrated? I find the only explanation of how the angle is measured
on pg 10 of the supplemental material “focus a small area of the
speckle image onto an optical fiber and detect the photons transmitted
through the fiber with the APD1. The detected signal is then highly
sensitive to the particle position and orientation.” How exactly is it
sensitive and how is the calibration determined? This should be better
explained in the main text as well as the Supplemental Material.

The work described in the supplemental material Fig. 3 regarding
measuring the PL is done with a static micro diamond. How is the setup
arranged as compared to the work in the Paul trap.

I was surprised there was not even more drift of the orientation of a
trapped nanodiamond in the Paul trap. Can the authors account for the
physics of the drift, in the SI a suggestion is made “The most likely
origin of this drift is the loss of charges of the diamond due to
photoionization by the laser, which changes the trapping conditions
over time.” Can this somehow be made more quantitative? It seems like
for better understanding and controlling the mechanical rotation for
future applications this may be important to understand?
\end{quote}
\subsection*{Our answer}
Here is our answer to the points raised by Referee 2 :

We have inverted the description of figure 3 in the main text to accurately describe the figure. (A FAIRE !j'ai pas le fichier à jour du main text)

We have changed the experimental setup figure in the supplementary material to be consistent with the one in the main text

La calibration de l'angle : Je pense qu'on peut rediriger vers le SI du nature de Tom (il trouve une précision de 0.3 mrad/racine(Hz). Par contre je sais pas dans quel mesure on développe la réponse.

The static experiments have been performed in the same experimental setup, with the diamonds deposited on the trap electrode/antenna instead of levitating inside it.

Le drift : je vois pas trop quoi dire

\end{document}

### 2021-10-09 Mes papiers/Méca CR/18112020/SI/Meca_cross_SI.tex

@@
 An acousto-optic modulator (AOM) is used to switch on and off the 532nm laser and to finely tuned its power.
-The photo-luminescence (PL) is collected by the objective, separated form the excitation light using a dichroic mirror (DM) and a 532nm notch filter (NF), and detected using a multimode-fiber single-photon avalanche photo-detector (APD) (SPCM-AQRH-15 from Perkin Elmer).
+The photo-luminescence (PL) is collected by the objective, separated form the excitation light using a dichroic mirror (DM) and a 532nm notch filter (NF), and detected using a multimode-fiber single-photon avalanche photo-detector (APD2) (SPCM-AQRH-15 from Perkin Elmer).
 Typically, from the heavily doped samples that we use, we can detect PL photons at a rate of 1 MHz after attenuating the signal by a factor 100 with neutral density filters.
+The reflected laser off the diamond is collected on a second similar avalanche photo-detector (APD1) after going through a beam-splitter (BS).
+
 The Paul trap is a pseudo-ring with a diameter of approximately 200 $\mu$m, as can be seen in \citep{DelordPhD}. It acts both as trap through the high voltage (HV) and as a microwave (MW) antenna.

### 2021-10-09 dico/voc.tex

@@
 \end{itemize}
+
+\item \textbf{Cathodoluminescence}  : production de PL par bombardement d'électrons.
+L'idée c'est simplement que tu excites des électrons de valence avec ton jet, et que ces électrons vont se recombiner en émettant de la PL.
+Dans la pratique c'est un peu plus crade vu que ton jet d'électron est trop énergétique, tu as plein de sous produits (rayons X etc) avant d'arriver à exciter tes électrons de valence.
+
+Ca permet de faire de la microscopie fine, en combinant un dispositif pour mesurer la PL a un TEM, mais c'est aussi tout simplement le mécanisme qui génère de la lumière dans le TV à tubes cathodiques.
   \section{Matériel}
@@
 
+  \section{Échantillons}
+  \begin{itemize}
+  \item \textbf{hBN} : Nitrure de Bore hexagonal (d'après JMC17, talk de Cassabois dispo sur youtube).
+C'est un matériel 2D 50\% B 50\%N quasi similaire au graphène (les haxagones se superposent contrairement au graphite) à grand gap indirect (6 eV).
+Il sert en particulier d'isolant pour les hétérostructures (il a une lattice constant proche du graphene).
+Il peut aussi servir à émettre dans l'UV au niveau de son gap (avec une PL 10$^3$ plus grande que le diamant, ce qui est assez étonnant), et il peut héberger des défauts colorés.
+En particulier le défaut de B semble présenter de l'ODMR avec un ZFS de 3.5 GHz.
+  \item \textbf{TMD ou TMDC} : Transition metal dichalcogenide.
+Grande famille de matériaux 2D composé d'un métal de transition et d'un (di)chacogenide, cad un élément de la famille de l'oxygène (souvent on exclue l'oxygène, donc il reste S, Se et Te, les éléments qui puent).
+Ils ont toujours la formule MD$_2$ pour leur forme 2D, ex : WSe$_2$.
+Contrairement au hBN et au graphène, ils n'ont pas une strcture parfaitement plane : les atomes de chalcogenides se mettent au dessus et en dessous de la couche de métal.
+
+  Au niveau application ça va dépendre du TMD, mais certains sont à gap direct, d'autres sont supra conducteur, et enfin la plupart ont un grand couplage spin orbite (resolvable en spectro) ce qui permet visiblement de faire de la spintronique
+  \item \textbf{Van der Waals heterostructures} : C'est des empilement de couches 2D de différents matériaux. Demander à Marin pour plus d'infos.
+  \end{itemize}
+  \item \textbf{Mott insulator} : Un matériaux qui est sensé être métallique d'après la théorie des bandes, mais qui est isolant à cause de l'interaction électron-électron (milieu à électrons fortement corrélés).
+On peut en général modéliser cette interaction avec un modèle de Hubbard
+  \item \textbf{Hubbard model} : La page wikipedia est pas claire
+  \item \textbf{CMOS} :
+  \end{itemize}
+  \section{Techniques expérimentales}
+  \begin{itemize}
+  \item \textbf{Epithaxy} :
+  \item \textbf{MBE}
+  \item \textbf{CVD}
+  \item \textbf{spin-coating}
+
+  \end{itemize}
+
   \end{document}

### 2021-10-12 Talks/group meeting 2021/presentation.tex
dur lundi en prévision

@@
 \end{frame}
+\begin{frame}
+\tableofcontents
+\end{frame}
+\section{The NV center}
+\begin{frame}
+\tableofcontents[currentsection]
+\end{frame}
 \begin{frame}{The Nitrogen Vacancy Center}
@@
 \end{center}
-
 \begin{itemize}
@@
 \end{frame}
+\begin{frame}{NV center : Optical properties}
+\centering
+\includegraphics[scale=.4]{slide_NV_optical}
+\end{frame}
+\begin{frame}{NV center : 8 levels}
+\centering
+\includegraphics[scale=.3]{NV_8_niveaux}
+\end{frame}
+\begin{frame}{NV center : 3 levels}
+\centering
+\includegraphics[scale=.23]{slide_3_niveaux}
+\end{frame}
+\begin{frame}{Spin manipulation}
+\centering
+\includegraphics[scale=.25]{slide_rabi_odmr}
+\end{frame}
+\begin{frame}{Summary : Magnetometry with NV centers}
+\centering
+\includegraphics[scale=.4]{ESR_8raies_article}
+\end{frame}
+\section{Probing dark spins with cross-relaxations}
+\begin{frame}
+\tableofcontents[currentsection]
+\end{frame}
+\begin{frame}{Principle of spin cross-relaxation (CR)}
+\centering
+\includegraphics[scale=.23]{Slide_CR}
+\end{frame}
+\begin{frame}{Detection of dark spins in CVD sample}
+\centering
+\includegraphics[scale=.35]{Slide_CR_CVD}
+\end{frame}
+\begin{frame}{Side-note : effect of transverse field on the PL}
+\centering
+\includegraphics[scale=.18]{slide_champ_transverse}
+\end{frame}
+\section{NV-NV Cross-relaxation}
+\begin{frame}
+\tableofcontents[currentsection]
+\end{frame}
+\begin{frame}{NV-NV Cross-Relaxation}
+\centering
+\includegraphics[scale=.2]{Slide_CR_NV-NV}
+\end{frame}
+\begin{frame}{NV-NV Cross-Relaxation : spin lifetime}
+\centering
+\includegraphics[scale=.28]{Slide_CR_NV-NV_T1}
+\end{frame}
+\begin{frame}{Origin of the NV-NV cross-relaxation}
+Why is it weird ? Because the flip-flop process is spin preserving
+
+$\to$ They should not change the total spin polarization of the NV ensemble.
+\bigskip
+
+Hypotheses for the origin of the NV-NV Cross-relaxation :
+\begin{itemize}
+\item Spin diffusion to unpolarized spins (out of the laser spot)
+
+$\to$ The numbers are off by several order of magnitudes \& it still works with nano-diamond
+
+\item Superradiance $\to$ effect independent of temperature
+\item Fluctuators : Some NV have their spin heavily depolarized by tunneling in and out of nearby sites
+\item Polarization of the laser ?
+\end{itemize}
+\end{frame}
+\begin{frame}{Mechanically detected Cross-Relaxation}
+\centering
+\includegraphics[scale=.45]{slide_cr_meca}
+\end{frame}
+\section{magnetometry with cross-relaxations}
+\begin{frame}
+\tableofcontents[currentsection]
+\end{frame}
+\begin{frame}{Principle of the magnetometer}
+\centering
+\includegraphics[scale=.22]{Slide_principe_magnetoNVNV}
+\end{frame}
+\begin{frame}{Low field dependence of the PL}
+\centering
+\includegraphics[scale=.23]{slide_magneto_scan}
+\end{frame}
+\begin{frame}{Sensitivity of the measure}
+\centering
+\includegraphics[scale=.26]{slide_sensi_measure}
+\end{frame}
+\begin{frame}{Comparison with other magnetometer}
+\begin{itemize}
+\item NV ensemble : $\approx 1$ pT$\sqrt{\textrm{Hz}}$
+\item Microwave-less NV ensemble : $\approx 1$ nT$\sqrt{\textrm{Hz}}$
+\item SQUIDs(superconducting circuits)/vapor cells : $\approx 1$ fT$\sqrt{\textrm{Hz}}$
+\end{itemize}
+\bigskip
+
+But,
+
+\begin{itemize}
+\item Based on smaller diamonds (10 $\mu$m diamond VS few mm diamonds)
+\item Does not require well-defined crystal axis Vs magnetic field orientation
+
+$\to$ Can use poly-crystalline/powdered samples, diamond flow in solution, etc...
+
+\item Better than transverse field by several orders of magnitude
+\end{itemize}
+
+\end{frame}
+\begin{frame}{Bonus : Double quantum cross-relaxation in zero-field}
+\centering
+\includegraphics[scale=.22]{slide_DQ}
+\end{frame}
 \end{document}

### 2021-10-12 abstract/IQFA 21/abstract.tex

\documentclass[aps,twocolumn,showpacs]{revtex4-1}

\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{theorem}
\usepackage{amsmath}
\usepackage{times}
\usepackage{color}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\ifpdf
\usepackage{epstopdf}   
\usepackage{url}
\fi

\begin{document}
\title{Probing dark spins with NV centers in CVD-grown diamond}

\author{Clément Pellet-Mary$^{1}$, Maxime Perdriat$^1$, Paul Huillery$^1$, Alexandre Tallaire$^2$ and Gabriel Hétet$^1$}
%\email{gabriel.hetet@phys.ens.fr}
\affiliation{$^1$Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, 24 rue Lhomond, 75231 Paris Cedex 05, France. \\
$^2$ IRCP, Ecole Nationale Sup\'erieure de Chimie de Paris, 11 rue Pierre et Marie Curie, 75005 Paris, France
}

\begin{abstract}
\normalsize
The electronic spin of the Nitrogen Vacancy (NV) center in diamond has given rise to a wealth of application over the past ten years, in particular in the domain of magnetic field sensing. The main reason to its rapid rise in popularity is the ability of the spin to be optically polarized and read-out at room temperature, with nothing more than a confocal microscopy setup.

Because of their point-like nature, NV centers can be placed a few nanometers away from the magnetic source, making them good candidate to probe tiny fields, such as the ones produced by other spins. Electronic and nuclear spins, both inside or outside the diamond have been detected by NV center magnetometry, down to a single spin \citep{Ref1}.

In our work \citep{Ref2}, we have used a diamond grown through chemical vapor deposition (CVD) with a relatively large ensemble of NV centers ($\approx$~3 ppm) in order to probe trace amounts of other electronic spins in the diamond, in the ppb range. This technique relies on resonant dipole-dipole coupling between the spins (flip-flop) \citep{Ref3} and unlike many NV magnetometry protocols does not require a microwave field.

Compared to traditional electronic spin resonance technique such as EPR, besides its comparatively simpler apparatus and potential gain in sensibility, an NV-based method present the advantage not to rely on the total number of spins of the sample, but rather only on the spins in the vicinity of the probed NV centers, making the measurement applicable to micro or nano-diamond, with a spatial resolution given by the diffraction limit of the microscope.

Understanding and controlling the local environment of NV-centers is a key aspect to improve NV magnetometry and other quantum applications with NV centers. Our method could provide chemists and material scientists a quick and cheap purity test for the samples they grow.
\begin{figure}[h]
\includegraphics[scale=0.3]{Fig}
\caption{\textbf{Top Figure :} Transition energies of the NV center's spin and various dark spins as a function of the external magnetic field.\\ \textbf{Bottom figure : }Normalized photoluminescence of the NV center as a function of the external magnetic field. \\Resonant dipolar coupling between polarized NV centers and unpolarized dark spins induce a depolarization of the NV spin which is detected by a drop in the NV photoluminescence}
\end{figure}
\end{abstract}

\maketitle

\begin{thebibliography}{}

\bibitem{Ref1} Zhao, Nan, et al. Nature nanotechnology 7.10 (2012): 657-662.

\bibitem{Ref2} Pellet-Mary, C., Huillery, P., Perdriat, M., Tallaire, A., Hétet, G. (2021). Physical Review B, 103(10), L100411.

\bibitem{Ref3} Armstrong, S., Rogers, L. J., McMurtrie, R. L., Manson, N. B. (2010). Physics Procedia, 3(4), 1569-1575.
\end{thebibliography}
\end{document}

### 2021-10-28 abstract/IQFA 21/abstract poster.tex
tutu

\documentclass[aps]{revtex4-1}

\begin{document}
The Nitrogen Vacancy (NV$^-$) electron spin has the remarkable property of being both polarizable and readable optically at room temperature. This property makes it a strong candidate to probe tiny magnetic field at the atomic level. Here we present our results where we managed to detect other spin impurities in the diamond at the ppb range thanks to resonant coupling with NV centers. We also present observations on the NV-NV dipolar interaction and its potential use in magnetometry.
\end{document}

### 2022-01-18 théorie/fiches/Couplage dipolaire/dipole.tex
last commit ordi asus

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage{unicode-math}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations
\usepackage{changepage}
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
  \section{Intéraction dipolaire}
  \subsection{Formulation classique}
  Dans tous les cas on considère deux moment magnétiques suffisament éloignés pour être considéré comme ponctuel, dont le premier se trouve en 0 et le second en $\mathbf{r}=r\mathbf{u}$.

  En classique on considère deux moments magnétiques $\mathbf{m_1}$ et $\mathbf{m_2}$ dont la position et l'orientation sont parfaitement définies. L'énergie d'interaction entre ces deux moments (dont découle la Force et le couple via les gradients) est :

  \begin{equation}
  H=-\frac{\mu_0}{4 \pi r^3}[3(\mathbf{m_1}\cdot \mathbf{u})(\mathbf{m_2}\cdot \mathbf{u})-\mathbf{m_1}\cdot \mathbf{m_2}]
  \end{equation}
 Il ya en plus un terme en $\delta (r)$ qui ne joue pas dans ce cas de figure, mais qui est peut etre à l'origine du contact de Fermi (je verrai ça plus tard)

 \textbf{Note :} C'est directement $H=-\mathbf{m_1} \cdot \mathbf{B}(m_2)$ le produit d'un dipole par le champ généré par l'autre dipole au point $\mathbf{r}$. Le terme en $\delta$ vient en gros de ce qu'il se passe à l'intérieur du dipole (avec un comportement différent en foncttion de si c'est deux monopoles magnétiques ou une boucle de courant)

 \subsection{Formulation quantique}
 On mets des chapeaux et c'est parti. Je vais quand même me concentrer rapidement sur ce qui m'interesse, à savoir des interactions entre spins dans des solides, donc je vais considérer la position des spins comme une variable classique (c'est sans doute un peu plus chiant pour le spin-orbite ou la position du moment magnétique n'est peut-etre pas aussi bien définie ? En plus il doit y avoir des merdes relativistes).

 Je vais aussi négliger pour l'instant l'intéraction de contact de Fermi, qui si je dis pas de bétise vient du recouvrement potentiel de la fonction d'onde de l'électron avec un autre dipole (typiquement un noyau pour une orbitale s).

 On se retrouve donc avec deux dipoles quantiques $\mathbf{\hat{m}_1}= \gamma_1 \hbar \mathbf{\hat{S}_1}$ et $\mathbf{\hat{m}_2}= \gamma_2 \hbar \mathbf{\hat{S}_2}$ ou $\gamma_{1,2}$ sont les rapports gyromagnétiques des spins. D'ou : 
 \begin{equation}
  \mathcal{H}_{dd}=-\frac{\mu_0 \gamma_1 \gamma_2 \hbar ^2}{4 \pi r^3}[3(\mathbf{\hat{S}_1}\cdot \mathbf{u})(\mathbf{\hat{S}_2}\cdot \mathbf{u})-\mathbf{\hat{S}_1}\cdot \mathbf{\hat{S}_2}]
  \end{equation}

  La base de l'espace de Hilbert par défaut pour représenter ce Hamiltonien est $\hat{S}_z^1 \ast \hat{S}_z^2$, c'est à dire la base des états $\ket{m_s^1=i,m_s^2=j}$ pour un axe \textbf{z} donné.

  Tu peux en théorie choisir 2 axes \textbf{z} différents pour tes 2 spins, mais ça va vite devenir galère quand il faudra développer les termes du Hamiltonien (sinon il faut décomposer \textbf{u} sur deux bases différentes).

  Dans les papiers, on regroupe en général tout le préfacteur numérique 
  \begin{equation}
  J_0=\frac{\mu_0 \gamma_1 \gamma_2 \hbar ^2}{4 \pi}
  \end{equation}
  $J_0 \approx (2\pi)56 \textrm{MHz/nm}^3$ pour deux spins électroniques ($\gamma \approx \gamma_0$). D'ou finalement :
   \begin{equation}
  \mathcal{H}_{dd}=-\frac{J_0}{r^3}[3(\mathbf{\hat{S}_1}\cdot \mathbf{u})(\mathbf{\hat{S}_2}\cdot \mathbf{u})-\mathbf{\hat{S}_1}\cdot \mathbf{\hat{S}_2}]
  \end{equation}

  \section{Différents couplages}
  \subsection{Couplage NV-NV même classe en champs moyen}
  En champ moyen ($D >> \gamma B_z >> d_\perp E_\perp$), on peut simplifier le Hamiltonien du centre NV au premier ordre à :
  \begin{equation}
  \mathcal{H}_{NV}=D S_z^2 + \gamma B_z S_z= 
  \begin{pmatrix}
	D-\gamma B_z &0&0 \\
	0 &0&0\\
	0&0& D+ \gamma B_z
  \end{pmatrix}
  \begin{matrix}
  \bra{-1} \\
  \bra{0} \\
  \bra{+1}
  \end{matrix}
  \end{equation}

  Les opérateurs de spins s'écrivent naturellement dans cette base (normal vu qu'on a choisi la base comme les états propres de Sz) : 
  \begin{gather*}
  S_x = \begin{pmatrix}
  0&1/\sqrt{2}&0 \\
  1/\sqrt{2}&0&1/\sqrt{2} \\
  0&1/\sqrt{2}&0
  \end{pmatrix}
  \begin{matrix}
  \bra{-1} \\
  \bra{0} \\
  \bra{+1}
  \end{matrix}  
  , \quad 
  S_y = \begin{pmatrix}
  0&i/\sqrt{2}&0 \\
  -i/\sqrt{2}&0&i/\sqrt{2} \\
  0&-i/\sqrt{2}&0
  \end{pmatrix} 
  \begin{matrix}
  \bra{-1} \\
  \bra{0} \\
  \bra{+1}
  \end{matrix} 
  \\ 
  S_z = \begin{pmatrix}
  -1&0&0 \\
  0&0&0\\
  0&0&1
  \end{pmatrix}
  \begin{matrix}
  \bra{-1} \\
  \bra{0} \\
  \bra{+1}
  \end{matrix}
  \end{gather*}

  \textbf{Note} : Si on veut prendre en compte le champ transverse, il faudra appliquer une (petite) rotation au Hamiltonien pour se placer dans la base du "NV" (= Hamiltonien diagonal), mais il faudra aussi appliquer cette rotation a $S_x$, $S_y$ et $S_z$. (En particulier les transitions $\ket{-1;+1} \to \ket{+1;-1}$ deviennent possibles, comme avec la micro-onde)

  Le hamiltonien complet des deux spins s'écrit alors :
  \begin{equation}
  \mathcal{H}=\mathcal{H}_{NV1} \ast \mathbb{1}_{NV2} + \mathbb{1}_{NV1} \ast \mathcal{H}_{NV2} + \mathcal{H}_{dd}
  \end{equation}
  \begin{adjustwidth}{-100pt}{-100pt}
  $$
 \begin{pmatrix}
  \textcolor{red}{2D - 2\gamma B_z + J_{zz}} &0&0&0&0&0&0&0&0\\
  0& \textcolor{red}{D- \gamma B_z} &0&\textcolor{blue}{J_{xx}+J_{yy}}&0&0&0&0&0\\
  0&0& \textcolor{red}{2D-J_{zz}} &0&0&0&\textcolor{blue}{0}&0&0 \\
  0&\textcolor{blue}{J_{xx}+J_{yy}}&0& \textcolor{red}{D- \gamma B_z}&0&0&0&0&0\\
  0&0&0&0& \textcolor{red}{0} &0&0&0&0\\
  0&0&0&0&0& \textcolor{red}{D+ \gamma B_z} &0&\textcolor{blue}{J_{xx}+J_{yy}}&0\\
  0&0&0&0&0&0& \textcolor{red}{2D-J_{zz}} &0&0\\
  0&0&0&0&0&\textcolor{blue}{J_{xx}+J_{yy}}&0& \textcolor{red}{D+ \gamma B_z} &0\\
  0&0&0&0&0&0&0&0& \textcolor{red}{2D+ 2\gamma B_z + J_{zz}}
  \end{pmatrix}
  \begin{matrix}
  \bra{-1,-1} \\
  \bra{-1,0} \\
  \bra{-1,+1}\\
  \bra{0,-1} \\
  \bra{0,0} \\
  \bra{0,+1}\\
  \bra{1,-1} \\
  \bra{1,0} \\
  \bra{1,+1}
  \end{matrix}
  $$
  \end{adjustwidth}
  \end{document}

### 2022-01-18 python/simu Hamiltonien NV/template.py

@@ -74,15 +74,39 @@ def egvect(H) :
 	return(val,vec)
 
 def print_matrix(M,bname) :
-	headers=['']+['|'+name+'>' for name in bname]
-	table=[]
-	for i in range(len(bname)) :
-		line=[]
-		line+=['<'+bname[i]+'|']
-		values=list(M[:,i])
-		line+=values
-		table+=[line]
-	print(tabulate(table,headers))
+	
+	if M.dtype.name=='float64' :
+		headers=['']+['|'+name+'>' for name in bname]
+		table=[]
+		for i in range(len(bname)) :
+			line=[]
+			line+=['<'+bname[i]+'|']
+			values=list(M[:,i])
+			line+=values
+			table+=[line]
+		print(tabulate(table,headers))
+
+	elif M.dtype.name=='complex128' :
+		print('Real Part :')
+		headers=['']+['|'+name+'>' for name in bname]
+		table=[]
+		for i in range(len(bname)) :
+			line=[]
+			line+=['<'+bname[i]+'|']
+			line+=[v.real for v in M[:,i]]
+			table+=[line]
+		print(tabulate(table,headers),'\n')
+
+		print('Imaginary Part :')
+		headers=['']+['|'+name+'>' for name in bname]
+		table=[]
+		for i in range(len(bname)) :
+			line=[]
+			line+=['<'+bname[i]+'|']
+			line+=[v.imag for v in M[:,i]]
+			table+=[line]
+		print(tabulate(table,headers))
+
 
 def print_vector(v,bname) :
 	table=[]
@@ -1318,10 +1342,6 @@ def PL_4classes(B,CR=False,gamma_las=1E-3,gamma_phonon=2E-4,normalized=True,affi
 	return(rho_0/4)
 
 
-Bval=3025/sqrt(3)
-B=np.array([Bval,Bval,Bval])
-Bt=np.array([5,-10,5])
-print(PL_4classes(B,affiche=True))
 
 
 def PL_plot_1() : #Je pense que ce qu'il me manque c'est le niveau excité. Ca explique pas trop pourquoi on verrait pas les +1 par contre.
@@ -1724,5 +1744,17 @@ def Lukin(B,gamma_las=1E-3,gamma_phonon=3E-4,gamma_f=1): #En Mhz
[…]

### 2022-01-28 experience/Bilan manips/2021.tex
tata yoyo

\documentclass[a4paper]{article}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}

\title{Bilan manips 2021}

\begin{document}
\maketitle
\section{Adamas}
\begin{itemize}
\item \textbf{01/07} Scan 100 (sans modulation), ESR 0B et quelques T1 (pas fous) sur 2 adamas. Toujours en mode compteur de photons donc signaux pas oufs.
\item \textbf{01/25} La suite sur 8 adamas de plus (scan 100, ESR 0B (quand j'y pense) et quelques T1 pas terribles)
\item \textbf{03/18-25} Essais de magnetometrie 0B avec des FND 100 nm et adamas 1um
\item \textbf{04/28} Zigouigouis sur des adamas 1 um
\item \textbf{05/06} Scan 100 et desax sur un 15um , il y a un léger zigouigoui sur le desax
\item \textbf{05/10} Scan 111 (x2) et 100(x2) sur 4 adamas. Pas de T1 ou d'ESR 0B, mais scans détaillés (quelques scans 2f aussi)
\item \textbf{05/31} Scan 100 sur un 15um long (mais pas d'échelle)
\item \textbf{07/20} Mesures sur 6 15um, Scan 100 sur 1,3,6, mesure ESR 0B que sur 1. Scan 111 et calibration sur 6 aussi
\item \textbf{08/31 $\to$ 09/10} (Je pense que c'est sur un adamas). Mesure du contraste de C13 et DQ en fonction de la puissance. Les meilleurs résultats sont le 09/10
\item \textbf{09/13-15} Série de T1, probablement sur adamas, probablement sans soustraction, donc pas oufs.
\item \textbf{09/27} Mesure sensitivité magnétométrie Vs angle pour un adamas avec beaucoup de DQ, et un avec peu de DQ. Il y a aussi des ESR 0B et des T1
\item \textbf{10/06-15} Tests avec laser rouge résonnant, pas hyper concluant meme si on voit une différence sur les DQ
\item \textbf{10/15} Beaucoup de mesures sur le adamas du 9/27 avec bcp de DQ : T1 sur 1,2,3 et 4 classes et sur les C13!, ESR dans toutes les configurations possibles, Série de T1 selon 100. Il y a même une mesure de la polarisation des spins nucl selon la 111. 
\item \textbf{10/19} Série T1 selon la 111, petit soucis en champ faible mais assez propre quand meme. Mesure du contraste de la micro-onde au passage
\item \textbf{11/10} Scans avec des ensembles de adamas 1um et FND 100 nm. Pas de mesures de magnétométrie (ça devait pas être fameux)
\end{itemize}
\section{CVD Alex}
\begin{itemize}
\item \textbf{03/11} Scan 100 sur le rose
\item \textbf{04/01} Scan 100 rose long
\item \textbf{04/09} Scan 100 rose en fonction de la puissance laser
\item \textbf{06/02} Scan 100 et ESR 1 raie sur le rose, le 250 ppm N2O et le substrat
\item \textbf{06/14} ESR pour mesures de T2* sur tous les échantillons irradiés d'Alex (data envoyées à Alex)
\item \textbf{06/18} Mesure largeur croisement 2 raies sur 250 ppm N2O (pour mesurer la largeur des fluctuateurs). Pas de structure fine.
\item \textbf{06/23} Scans 100 calibrés sur les 3 100 ppm (non recuits à 1200).
\item \textbf{10/20} Mesures sur les échantillons recuit à 1200 + 800 + rose : Scans 100 (mesure VH- et War1), mais aussi croisement de raies (pour mesurer les SQ) et ESR pour T2
\item \textbf{10/27} Scans 100 sur les substrats recuits et pas recuits : un pic de War1 pousse avec le recuit dans le substrat.
\item \textbf{11/02} ESR pour T2*, tests avec la tranche polie en particulier
\item \textbf{11/15} Scans 100 sur la tranche polie pour voir si les War1 migraient depuis le CVD. C'est pas le cas. Premiers tests sur la face arrière aussi, notamment un test de l'élargissement de l'ESR 0B avec la puissance uW.
\item \textbf{11/17} Mise en place du piezo et premières maps (scan 100 et ESR 0B) sur la face arrière
\item \textbf{11/18} Suite, ESR plus précis (0B et non 0B) et DQ 2f. Pb de faux départ avec la micro-onde.
\item \textbf{11/22} Tentative de map SQ, mais accident d'aimant permanent pendant la nuit.
\item \textbf{11/23} Scans 100 sur un nouvel emplacement (de la face arrière, toujours) + quelques observations sur des scans sur le rose.
\item \textbf{11/24} Map ESR 0B sur l'emplacement de la veille + observation que les DQ et les C13 ne réagissent pas pareil à un changement de puissance laser
\item \textbf{11/29} Test de la puissance laser sur les C13/DQ sur le rose (C13 pas visibles), test influence de l'angle d'une lambda/2 sur différentes classes de NV à l'ESR, et surtout premières mesures sur la dernière zone de la face arrière (pas sur que ce soit le meme qu'avant : la c'est un recuit 800, avant je crois que c'était des recuit 1200) : Map scan 100, Map ESR 0B, quelques T1, quelques scan 1x4, quelques ESR 1x4
\item \textbf{12/01} Meme sample mais cette fois ci mesure uniquement sur une ligne : DQ, SQ, T1 0B et T1 1x4, ESR 0B, ESR 1x4, ESR P1
\item \textbf{12/07} Spectres et Temps de polarisation des spins, toujours sur la meme zone.
\item \textbf{12/08} Test des effets de la puissance laser sur le ratio NV-/NV0 : bouger le z du piezo et changer la puissance laser (mesure de spectres et de Tpola)
\item \textbf{12/09} ESR P1 et DQ en fonction de z
\item \textbf{12/10} Mesure ESR en fonction de la puissance uw, sur un spot High DQ et un spot low DQ. Le spot low DQ a l'air d'avoir une puissance uw inferieure (pas de Rabi malheureusement)
\end{itemize}

\section{Sumitumo/Ludo}
\begin{itemize}
\item \textbf{03/03} Quelques scans 111 sur \textbf{Sumi 2} pour tester la photodiode
\item \textbf{03/15} Scan 100 et 1x1x1x1 de \textbf{Ludo} (pas oufs)
\item \textbf{07/13} Scans 100 du \textbf{Sumi 4} pas terribles
\end{itemize}

\section{Bonus}
\begin{itemize}
\item \textbf{04/22} Première observation de Einsein-de Haas sur des diamants
\item \textbf{05/25-27} Tentative de cooling par relaxation croisée. Quelques plots de CR en lévitation ok, mais le ringdown n'a pas marché
\item \textbf{10/25} Mesure de la puissance micro-onde en fonction de la fréquence, en testant l'effet de l'ampli et du clapet anti-retour

\end{itemize}
\end{document}

### 2022-02-08 data/20211015/Scans série T1 100 soustraction/traitement.py
tata

@@ -1,5 +1,6 @@
 import sys
 sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
 def total():
@@ -23,7 +24,7 @@ def total():
 	plt.legend()
 	plt.show()
 
-total()
+# total()
 
 def single():
 	fname='V=0.716194 V'
@@ -34,4 +35,4 @@ def single():
 	print(popt[2])
 	plt.show()
 
-# single()
+single()

### 2022-02-08 manips/Double quantum sur des adamas/Calcul eta Lukin.py

@@ -1,0 +1,131 @@
+from numpy import cos,sin,sqrt,pi,arccos
+import numpy as np
+from scipy.integrate import quad, dblquad
+
+
+#base cistal :
+def x(theta,phi):
+	return(sin(theta)*cos(phi))
+
+def y(theta,phi):
+	return(sin(theta)*sin(phi))
+
+def z(theta,phi):
+	return(cos(theta))
+
+#base NV :
+#Rappel Numpy : tu rentres les matrices ligne par ligne, et M[i]=ligne(i)
+#Ici les lignes représentent les vecteurs dela base du NV (exprimés dans la base x,y,z), et les colonnes les vecteurs x,y,z exprimés dans la base du NV
+classe1Rot=np.array([[1/sqrt(2),-1/sqrt(2),0],[1/sqrt(6),1/sqrt(6),-2/(sqrt(6))],[1/sqrt(3),1/sqrt(3),1/sqrt(3)]]) #classe1Rot.dot([1,1,1])=[0,0,1]
+classe2Rot=classe1Rot.dot(np.array([[-1,0,0],[0,-1,0],[0,0,1]]))
+classe3Rot=classe1Rot.dot(np.array([[-1,0,0],[0,1,0],[0,0,-1]]))
+classe4Rot=classe1Rot.dot(np.array([[1,0,0],[0,-1,0],[0,0,-1]]))
+
+
+# print(classe3Rot.T.dot([0,1,0]))
+# print(classe3Rot[1])
+
+classesRot=[classe1Rot,classe2Rot,classe3Rot,classe4Rot]
+
+classNV1=1
+classNV2=4
+#C'est vraiment pas opti comme façon de faire, mais bon allons y quand meme
+x1=classesRot[classNV1-1][0]
+y1=classesRot[classNV1-1][1]
+z1=classesRot[classNV1-1][2]
+
+
+# 0.38490017056254233 0.8327775080569433 0.8242328246712349 0.7513109757772565
+
+# 0.7698003602248628 0.6878741656726536
+
+t=pi/4
+x2=classesRot[classNV2-1][0]*cos(t)+classesRot[classNV2-1][1]*sin(t)
[…]

### 2022-02-16 manips/Double quantum sur des adamas/Calcul eta Lukin.py
tata

@@ -3,6 +3,7 @@ import numpy as np
 from scipy.integrate import quad, dblquad, nquad
 
 
+
 #base cistal :
 def x(theta,phi):
 	return(sin(theta)*cos(phi))
@@ -49,12 +50,41 @@ def cartesian2(anglex):
 
 
 
-def xx_diff_noC(theta,phi,anglex1=0,anglex2=0):
+def xx_diff_noC(theta,phi,anglex1=0,anglex2=1):
 	r=[x(theta,phi),y(theta,phi),z(theta,phi)]
 	x1,y1,z1=cartesian1(anglex1)
 	x2,y2,z2=cartesian2(anglex2)
 	g=(3*x1.dot(r)*x2.dot(r)-x1.dot(x2))
-	integrande=1/(4*pi)*sqrt(g**2+h**2)*sin(theta)#/(2*pi)/(2*pi)
+	integrande=1/(4*pi)*abs(g)*sin(theta)#/(2*pi)#/(2*pi)
+	return(integrande)
+
+def xx_diff_fullC(theta_r,phi_r,theta_E=0,phi_E=0):
+	r=np.array([x(theta_r,phi_r),y(theta_r,phi_r),z(theta_r,phi_r)])
+	E=np.array([x(theta_E,phi_E),y(theta_E,phi_E),z(theta_E,phi_E)])
+
+	z1=classesRot[classNV1-1][2]
+	x1=E-(E.dot(z1))*z1
+	x1=x1/np.linalg.norm(x1)
+	y1=np.cross(z1,x1)
+
+	z2=classesRot[classNV2-1][2]
+	x2=E-(E.dot(z2))*z2
+	x2=x2/np.linalg.norm(x2)
+	y2=np.cross(z2,x2)
+	# x2=-x2 #Anticorrelation
+	# y2=-y2
+
+
+	g=(3*x1.dot(r)*x2.dot(r)-x1.dot(x2))
+	integrande=1/(4*pi)*abs(g)*sin(theta_r)*1/(4*pi)*sin(theta_E)
+	return(integrande)
+
+def yy_diff_noC(theta,phi,anglex1=0,anglex2=0):
+	r=[x(theta,phi),y(theta,phi),z(theta,phi)]
+	x1,y1,z1=cartesian1(anglex1)
+	x2,y2,z2=cartesian2(anglex2)
+	g=(3*y1.dot(r)*y2.dot(r)-y1.dot(y2))
[…]

### 2022-03-09 experience/Bilan manips/2022.tex
tata

\documentclass[a4paper]{article}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}

\title{Bilan manips 2022}

\begin{document}
\maketitle
\section{Adamas}
\begin{itemize}
\item \textbf{01/24} Test sensi en fonction de puissance las sur un bon adamas, ESR 0B et 1x4, scans 100 et 1x4
\item \textbf{02/18} Premiers tests T1 121 vs 22 sur un adamas 15 um. Ca marche
\item \textbf{02/22} Beaucoup de mesure sur un adamas 15 um sur la pcb : des T1 dans tous les sens (dont 121(x4) et 22 (x4), mais aussi 100,111, 1 classe, C13 en champ nul et du champ transverse (petit moyen et grand). Les ESR associés, et une mesure de la largeur des fluctuateurs en T1 (insh'Allah. Pour l'instant ça merdouille pas mal).
\end{itemize}
\section{Echantillons Alex}
\begin{itemize}
\item \textbf{01/18} Quelques tentatives de Rabi sur la face arrière (HPHT) d'un échatillon d'Alex
\item \textbf{01/21} ESR 0B et champ non nul pour différentes poistion de l'échantillon (mieux fait en 2021)
\item \textbf{01/28} Quelques ESR sur le rose dont un ESR 100 avec les 3 raies du spin nucl visibles (mais bcp d'hysteresis de l'autre coté)
\item \textbf{02/21} Test T1 121 vs 22 sur diamant rose, ça marche pas (différence trop faible)
\end{itemize}
\section{divers}
\begin{itemize}
\item \textbf{02/17} Test puissance micro-onde en fonction de la fréquence, avec et sans clapet anti-retour, sur un un ensemble de adamas 1 um.
\item \textbf{02/21} Essai de mesures EPR en transmission avec la PCB et le diamant rose. Ca marche pas. Mais j'ai test en modulant le laser, j'aurais du moduler le champ mag.
\end{itemize}
\end{document}

### 2022-03-23 Talks/group meeting 2022/presentation.tex
TOTO

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage{graphicx}
\usepackage{physics}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage[backend=biber]{biblatex}

\addbibresource{Biblio_dbl_quantum.bib}

%\bibliographystyle{stylename}
%\bibliography{Biblio_dbl_quantum}

\title{Group meeting : Cross-relaxation with NV centers ensemble in diamond}
%\author{Clément Pellet-Mary\\ Laboratoire de physique de l'ENS\\ \textit{ENS, Paris}}
\date\today

\mode<presentation> {\usetheme{Rochester}}

\begin{document}
\begin{frame}
\maketitle
\end{frame}
\begin{frame}{Outline}
\tableofcontents
\end{frame}
\section{(Quick) Reminder on NV center}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{NV$^-$ spin in the fundamental electronic state}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{slide_3_niveaux}
\end{frame}
\begin{frame}{Principle of CW ODMR}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{shéma lock in ESR}
\end{frame}
\section{Dipolar interaction and spin lifetime}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{Today's goal : understand the PL dip in zero-field}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide 1}
\end{frame}
\begin{frame}{Why does it matter ?}
\begin{itemize}
\item The concentration of NV$^-$ centers (while maintaining good spectral properties) is increasing.
\item Magnetometry in low field is already hard with standard techniques.
\item The dip itself can be used to perform low-field, microwave-less magnetometry.
\end{itemize}
\end{frame}
\begin{frame}{Link between the PL and the spin T1}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Explication T1 PL}
\end{frame}
\begin{frame}{Spin relaxation amplified by dipolar coupling}
See animation
\end{frame}
\begin{frame}{Dipolar interaction should not modify the total spin polarization...}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Flip flop inhomogénités}
\end{frame}
\begin{frame}{...except with inhomogeneities }
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Flip flop inhomogénités 2}
\end{frame}
\section{The fluctuator model}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{Theoretical frame : the fluctuator model}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{NV Vs Fluct}
\end{frame}
\begin{frame}{The dipole-dipole Hamiltonian}
\begin{center}
\includegraphics[scale=0.4]{Shéma dipole-dipole}
\end{center}

\begin{equation*}
  \mathcal{H}_{dd}=-\frac{\mu_0 \gamma_1 \gamma_2 \hbar ^2}{4 \pi r^3}[3(\mathbf{\hat{S}_1}\cdot \mathbf{u})(\mathbf{\hat{S}_2}\cdot \mathbf{u})-\mathbf{\hat{S}_1}\cdot \mathbf{\hat{S}_2}]
  \end{equation*}

  Every matrix element $\matrixel{i;\alpha}{\mathcal{H}_{dd}}{j;\beta}$ can be written as $\xi(\lVert \vec r \rVert)  \eta (\vec u)$

  \bigskip
  Example : $\matrixel{0;-1}{\mathcal{H}_{dd}}{-1;0}$ is the matrix element for a flip-flop process
\end{frame}
\begin{frame}{$\mathcal{H}_{dd}$ for two aligned spin}
In the magnetic basis $(\ket{-1},\ket{0},\ket{+1}) \otimes (\ket{-1},\ket{0},\ket{+1})$ :

\begin{align}
  \mathcal{H}_{dd}\propto &(\frac{3}{2}(n_x^2+n_y^2)-1)\left[\op{0;+1}{+1;0}+\op{-1;0}{0;-1} + h.c. \right]\\
  &+\frac{3}{2}(n_x^2-n_y^2+i2n_x n_y)\left[\op{0;+1}{-1;0}+\op{+1;0}{0;-1}\right]\\
  &+\frac{3}{2}(n_x^2-n_y^2-i2n_x n_y)\left[\op{0;-1}{+1;0}+\op{-1;0}{0;+1}\right]\\
  &+(3n_z^2-1)\hat{S_z^1}\hat{S_z^2}\\
  &+\mathcal{H}_{\mathrm{other}}
  \end{align}

Example : $\eta_{\mathrm{flip-flop}}=\frac{3}{2}(n_x^2+n_y^2)-1)$
\end{frame}

\begin{frame}{The fluctuator model : calculation steps}
\begin{itemize}
\item Decay rate induced by a single fluctuator on a single NV : $$
\Gamma_{ij}=\sum_{\alpha,\beta} \lvert \matrixel{i,\alpha}{\mathcal{H}_{dd}}{j,\beta} \rvert ^2 \frac{\gamma_f}{(\omega_{ij}-\omega_{\beta \alpha})^2 + \gamma_f^2}
$$
\pause
\item Each NV will see a total decay rate $$\gamma = \sum_\mathrm{fluct} \Gamma_{ij}(\vec r)$$
\end{itemize}
\end{frame}
\begin{frame}{The fluctuator model : calculation steps}
\begin{itemize}
\item We can compute the distribution of these individual decay rates $\gamma$ :
\begin{align*}
\rho(\gamma)&=\int \left\{ d_{\mathrm{fluct}} \right\} \delta \left( \gamma - \sum_\mathrm{fluct} \Gamma_{ij}(\vec r) \right) \\
&=\frac{e^{-1/(4\gamma T)}}{\sqrt{4 \pi \gamma^3 T}}
\end{align*}

\pause
\item The measured lifetime will be averaged over all the NVs :
$$
PL(t) = \int e^{-\gamma t} \rho(\gamma) d\gamma = e^{-\sqrt{t/T}}
$$
\end{itemize}
\end{frame}
\begin{frame}{The fluctuator model : calculation steps}
Where :
$$
T \equiv \left( \frac{4 \pi n_f J_0 \bar \eta}{3} \right) ^2 \frac{\pi}{\gamma_f}
$$

With :
\begin{itemize}
\item $n_f$ is the fluctuator density
\item $\gamma_f$ is the fluctuator decay rate
\end{itemize}

and :
$$
 \bar \eta \equiv \frac{1}{4} \sum_{\mathrm{classes}} \frac{\gamma_f^2}{(\omega_{\mathrm{NV}}-\omega_{\mathrm{fluct}})^2 + \gamma_f^2} \int_{\theta , \phi} \lvert \eta(\theta , \phi) \rvert \dd \Omega 
$$
\end{frame}

\begin{frame}{Examples}

%\begin{align*}
% \bar \eta &= \frac{1}{4} \int_{\theta , \phi} \lvert \eta(\theta , \phi) \rvert \dd \Omega \\
% &= \frac{1}{4}\times 0.3849
%\end{align*}
%
%\begin{align*}
% \bar \eta &= \frac{1}{4} \int_{\theta , \phi} \lvert \eta(\theta , \phi) \rvert \dd \Omega + \frac{1}{4} \int_{\theta , \phi} \lvert \eta ' (\theta , \phi) \rvert \dd \Omega \\
% &= \frac{1}{4}*0.3849 + \frac{1}{4}*0.8328
%\end{align*}

\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Exemple fluct 121 1x4}
\end{frame}

\begin{frame}{Non-trivial prediction of the model}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{121 VS 22 shéma}
\end{frame}
\begin{frame}{Non-trivial prediction of the model}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{121 Vs 22 combined}
\end{frame}
\section{The reasons behind the dip in zero-field}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{The reasons behind the dip in zero-field}
\begin{center}
\includegraphics[width=.8\textwidth,height=0.9\textheight,keepaspectratio]{Scan PL random dir}
\end{center}

There are 3 parts, all linked to dipolar coupling :
\begin{enumerate}
\item The lift of the degeneracies between the four classes
\item The change of the eigenstates in the avoided crossing region
\item The double-flip terms in the dipolar Hamiltonian
\end{enumerate}
\end{frame}
\begin{frame}{1. The lift of the degeneracies}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{splitting degenerescence}
\end{frame}
\begin{frame}{But it is not the only reason}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{scan 100 slide}
\end{frame}
\begin{frame}{2. The change in eigenstates of $\mathcal{H}_0$}
%$$
%\mathcal{H}_0=\begin{pmatrix}
%D-\gamma_e B_z & \gamma_e B_x & d_\perp E_x \\
%\gamma_e B_x & 0 & \gamma_e B_x \\
%d_\perp E_x & \gamma_e B_x & D+\gamma_e B_z
%\end{pmatrix}
%$$
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Coupled Basis}
\end{frame}
\begin{frame}{Dipolar Hamiltonian in the coupled basis}
\setcounter{equation}{0}
In the coupled basis $(\ket{-},\ket{0},\ket{+}) \otimes (\ket{-},\ket{0},\ket{+})$ :
\begin{align}
  \mathcal{H}_{dd}&=(3n_x^2-1)\left[\op{0;+}{+;0}+ h.c. \right]\\
  &+(3n_y^2-1)\left[\op{0;-}{-;0}+ h.c. \right]\\
  &+i3n_x n_y\left[\op{0;+}{-;0}+\op{+;0}{0;-} +h.c. \right]\\
  &+(3n_z^2-1)\left[\op{+;-}{-;+}+ h.c.\right]\\
  &+\mathcal{H}_{\mathrm{other}}
  \end{align}
  $\eta_{\mathrm{flip-flop}} = 3n_x^2-1$ or $3n_y^2-1 \neq \frac{3}{2}(n_x^2+n_y^2)-1$

  $\to \bar \eta_{\mathrm{flip-flop}}(\ket{\pm 1})=0.3849 < \bar \eta_{\mathrm{flip-flop}}(\ket{\pm})=0.7698$
\end{frame}
\begin{frame}{$\mathcal{H}_0$ under pure transverse field}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Base couplée : vec vs val}
\end{frame}
\begin{frame}{Experimental results with transverse field}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Champ transverse mesures}
\end{frame}
\section{Low-field Magnetometry}
\begin{frame}{Outline}
\tableofcontents[currentsection]
\end{frame}
\begin{frame}{Low field magnetometry experimental setup}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide magneto 1}
\end{frame}
\begin{frame}{Magnetometry protocol}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide magneto 2}
\end{frame}
\begin{frame}{Effects of the various effects on the sensitivity}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide magneto 3}
\end{frame}
\end{document}

### 2022-03-23 data/20220317/Sumi 2/traitement_largeur_f.py

@@ -1,0 +1,237 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+import matplotlib.animation as animation
+from analyse import *
+
+"""
+Calculs Eta :
+eta0=0.3849/4
+eta=(0.3849/4+2*0.6507/4+0.8328/4)/30
+
+eta c13 0B : 0.05 eta0
+eta base +/- 1 classe : 3-4 eta0
+eta 121 : 10 eta0
+eta 22 : 7.2 eta0
+eta 31 : 28.4 eta0
+eta 40 : 42.8 eta0
+eta 0B sans DQ : 51-55 eta0
+
+
+"""
+def full_animation():
+
+	fig,axes=plt.subplots(2)
+	fig.set_size_inches([10,8])
+	[ax1,ax2]=axes
+	ax1.set_ylim([0,2.7])
+	ax2.set_ylim([0,1.1])
+	ax1.set_xlabel('Frequency (MHz)',fontweight='bold')
+	ax1.set_ylabel('ODMR contrast',fontweight='bold')
+	ax2.set_xlabel('Dark time (ms)',fontweight='bold')
+	ax2.set_ylabel('T1 contrast',fontweight='bold')
+
+	nFrameMin=20
+	nFrameMax=100
+
+
+	fnames,fval=extract_glob('Série ESR',14)
+	fnames=fnames[nFrameMin:nFrameMax]
+
+
+	x,y=extract_data(fnames[0])
+	l1=ax1.plot(x,y)[0]
+
+	xESR=x
+	yESRs=[]
+	for i in range(len(fnames)):
+		x,y=extract_data(fnames[i])
+		yESRs+=[y]
+
+	fnames,fval=extract_glob('Série T1',13)
+	fnames=fnames[nFrameMin:nFrameMax]
+
+
+	x,y=extract_data(fnames[0],ycol=5)
+	nT1max=100
+	x=x[:nT1max]*1e3
+	y=y[:nT1max]/max(y[:nT1max])
+	l2=ax2.plot(x,y,'o',markerfacecolor='None')[0]
+	popt,yfit=stretch_arb_exp_fit_zero(x,y,alpha=1,fixed=True)
+	l3=ax2.plot(x,yfit,lw=2,label='tau=%.3f'%(popt[1]))[0]
+	xT1=x
+	yT1s=[]
[…]

### 2022-03-29 théorie/fiches/Couplage dipolaire/dipole.tex

@@
 \usepackage{unicode-math}
-\usepackage[]{braket} % \bra, \ket etc
+\usepackage[]{physics} % \bra, \ket etc
 \usepackage{graphicx}
@@
   \end{equation}
+  \subsection{Base Magnétique}
+  Dans la base magnétique, on peut réécrire $H_{dd}$ comme :
+  \begin{align}
+  \mathcal{H}_{dd}=&(\frac{3}{2}(x^2+y^2)-1)\left[\op{0;+1}{+1;0}+\op{-1;0}{0;-1} + h.c. \right]\\
+  &+\frac{3}{2}(x^2-y^2+i2xy)\left[\op{0;+1}{-1;0}+\op{+1;0}{0;-1}\right]\\
+  &+\frac{3}{2}(x^2-y^2-i2xy)\left[\op{0;-1}{+1;0}+\op{-1;0}{0;+1}\right]\\
+  &+(3z^2-1)\hat{S_z^1}\hat{S_z^2}\\
+  &+\mathcal{H}_{\mathrm{other}}
+  \end{align}
+
+  Ou $\mathcal{H}_{\mathrm{other}}$ contient des termes non resonnant et non diagonaux (qui ne vont modifier ni les cohérences (termes diagonaux) ni les populations (termes résonnant non diagonaux))
+   \subsection{Base Couplée}
+  Dans la base couplée, on peut réécrire $H_{dd}$ comme :
+  \begin{align}
+  \mathcal{H}_{dd}=&(3x^2-1)\left[\op{0;+}{+;0}+ h.c. \right]\\
+  &+(3y^2-1)\left[\op{0;-}{-;0}+ h.c. \right]\\
+  &+i3xy\left[\op{0;+}{-;0}+\op{+;0}{0;-} +h.c. \right]\\
+  &+(3z^2-1)\left[\op{+;-}{-;+}+ h.c.\right]\\
+  &+\mathcal{H}_{\mathrm{other}}
+  \end{align}
 

### 2022-04-01 abstract/CECAM 2022/abstract.tex
tutu

\documentclass[aps,twocolumn,showpacs]{revtex4-1}

\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{theorem}
\usepackage{amsmath}
\usepackage{times}
\usepackage{color}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\ifpdf
\usepackage{epstopdf}   
\usepackage{url}
\fi

\begin{document}
\title{Probing dark spins with NV centers in CVD-grown diamond}

\author{Clément Pellet-Mary$^{1}$, Maxime Perdriat$^1$, Paul Huillery$^1$, Alexandre Tallaire$^2$ and Gabriel Hétet$^1$}
%\email{gabriel.hetet@phys.ens.fr}
\affiliation{$^1$Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, 24 rue Lhomond, 75231 Paris Cedex 05, France. \\
$^2$ IRCP, Ecole Nationale Sup\'erieure de Chimie de Paris, 11 rue Pierre et Marie Curie, 75005 Paris, France
}

\begin{abstract}
\normalsize
The electronic spin of the Nitrogen Vacancy (NV) center in diamond has given rise to a wealth of application over the past ten years, in particular in the domain of magnetic field sensing. The main reason to its rapid rise in popularity is the ability of the spin to be optically polarized and read-out at room temperature.

In our work \citep{Ref2}, we have used a diamond grown through chemical vapor deposition (CVD) with a relatively large ensemble of NV centers ($\approx$~3 ppm) to detect trace amount ($\approx$ ppb) of other spin impurities present in the diamond (in this case VH- ans War1, both spin 1 defects).

The technique consists in scanning a magnetic field along a particular crystalline axis and monitoring the NV- photoluminescence (PL). Whenever the NV centers' spins which are polarized by the laser become resonant with another spin family which are unpolarized at room temperature, the NVs lose polarization though resonant dipole coupling (flip-flop or double flip), resulting in a drop in the PL. The crystalline axis is chosen in order to minimize the number of resonances and get a cleaner spectrum. (Fig. II)

Compared to more traditional electronic spin resonance technique such as EPR, besides its comparatively simpler apparatus and potential gain in sensitivity, an NV-based method present the advantage not to rely on the total number of spins of the sample, but rather only on the spins in the vicinity of the probed NV centers, making the measurement applicable to micro or nano-diamond, with a spatial resolution given by the diffraction limit of the microscope.

This method has been used \citep{Ref4} to study the effect of the annealing process (a necessary step in the formation of NV centers) in the creation or suppression  of other defects in the crystal. (Fig. III)
\begin{figure}[h]
\includegraphics[scale=0.3]{Fig}
\caption{\textbf{I} : Sketch of an NV- center interacting with other spin impurities in diamond.  \textbf{IIa)} Transition energies of the NV center's spin and various dark spins as a function of the external magnetic field. \textbf{IIb)} Normalized photoluminescence of the NV center as a function of the external magnetic field. PL dip lines correspond to resonant interaction between NV centers and other spin impurities in the crystal. \textbf{IIIa)} Lock-in detection of two spin impurities (VH- and War1) for a CVD diamond annealed at 800$^\circ$C. \textbf{IIIb)} Lock-in detection for a similar diamond annealed at 1200$^\circ$C, only the War1 defect remains.}
\end{figure}
\end{abstract}

\maketitle

\begin{thebibliography}{}

\bibitem{Ref2} Pellet-Mary, C., Huillery, P., Perdriat, M., Tallaire, A., Hétet, G. (2021). Physical Review B, 103(10), L100411.

%\bibitem{Ref3} Armstrong, S., Rogers, L. J., McMurtrie, R. L., Manson, N. B. (2010). Physics Procedia, 3(4), 1569-1575.

\bibitem{Ref4} Ngambou, M. W. N., Pellet-Mary, C., Brinza, O., Antonino, A., Hetet, G., Tallaire, A., Bénédic, F. \& Achard, J. (2022).  Diamond and Related Materials, 123, 108884.

\end{thebibliography}
\end{document}

### 2022-04-05 Manuscrit Thèse/These.tex
toto

--- 2022-03-09 experience/Bilan manips/2022.tex
@@
 
-\title{Bilan manips 2022}
+\title{Le titre}
 
@@
 \maketitle
-\section{Adamas}
-\begin{itemize}
-\item \textbf{01/24} Test sensi en fonction de puissance las sur un bon adamas, ESR 0B et 1x4, scans 100 et 1x4
-\item \textbf{02/18} Premiers tests T1 121 vs 22 sur un adamas 15 um. Ca marche
-\item \textbf{02/22} Beaucoup de mesure sur un adamas 15 um sur la pcb : des T1 dans tous les sens (dont 121(x4) et 22 (x4), mais aussi 100,111, 1 classe, C13 en champ nul et du champ transverse (petit moyen et grand).
-Les ESR associés, et une mesure de la largeur des fluctuateurs en T1 (insh'Allah.
-Pour l'instant ça merdouille pas mal).
-\end{itemize}
-\section{Echantillons Alex}
-\begin{itemize}
-\item \textbf{01/18} Quelques tentatives de Rabi sur la face arrière (HPHT) d'un échatillon d'Alex
-\item \textbf{01/21} ESR 0B et champ non nul pour différentes poistion de l'échantillon (mieux fait en 2021)
-\item \textbf{01/28} Quelques ESR sur le rose dont un ESR 100 avec les 3 raies du spin nucl visibles (mais bcp d'hysteresis de l'autre coté)
-\item \textbf{02/21} Test T1 121 vs 22 sur diamant rose, ça marche pas (différence trop faible)
-\end{itemize}
-\section{divers}
-\begin{itemize}
-\item \textbf{02/17} Test puissance micro-onde en fonction de la fréquence, avec et sans clapet anti-retour, sur un un ensemble de adamas 1 um.
-\item \textbf{02/21} Essai de mesures EPR en transmission avec la PCB et le diamant rose. Ca marche pas. Mais j'ai test en modulant le laser, j'aurais du moduler le champ mag.
-\end{itemize}
+\section{Section 1}
+\subsection{Subsection 1}
+
 \end{document}

### 2022-04-05 idées/To_do_list.tex
toto

@@
   \end{itemize}
+  \section{Questions centre NV}
+  \begin{itemize}
+  \item Phonon side-band
+  \item mieux comprendre le D
+  \item inter-system crossing + Jahn-Teller effect
+  \item dependence du spin en E, effet stark, et pk la suceptibilité parallèle est 50x plus petite que la perpendiculaire (Yao)
+  \item dépendence du spin en le strain, en la température
+  \end{itemize}
 
+ \section{Manips}
+ \begin{itemize}
+ \item Dépendance polarization time/distance objectif (micro/bulk)
+ \item Le T1 soustraction dépend de la distance de l'objectif : dépend aussi de la puissance du laser ? Juste pour le T1 dipole ou aussi le T1 phonon ?
+ \item Mesurer des largeurs fluctateurs avec le T1 sur plein d'échantillons (de Ludo au rose) (et vérifier que PL=T1 si possible, mais ça rajoute des contraintes pour rien...).
+ \end{itemize}
   \end{document}

### 2022-04-08 Mes papiers/Low field CR/Low field CR.tex
tata

\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-1}
%\documentclass[preprintnumbers,amsmath,amssymb]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}

%\newcommand{|}{Y$_2$SiO$_5$}
\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{TBA}

\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, G. H\'etet}

\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e Paris Cit\'e , 24 rue Lhomond, 75231 Paris Cedex 05, France.}

\begin{abstract}
C'est trop bien
\end{abstract}

\maketitle

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
\caption{Modification of the stretch lifetime for two near-resonant classes. (a) ODMR spectrum with amplitude modulation of the microwave for the $\ket{0} \to \ket{-1}$ transition of two separate NV classes. The frequency detuning between the two transitions is denoted $\Delta \nu$. (b) Stretch part of the lifetime decay for one of the two classes as function of the frequency detuning (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz . Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
\end{figure}

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig transverse field simu}
\caption{Simulated eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field. (a) Energies of the three eigenstates $\ket{g}$, $\ket{d}$ and $\ket{e}$ as a function of the magnetic field amplitude. (b) Blue dashed curve : frequency detuning between the two transitions $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$. Plain red curve : projection of $\ket{e}$ on $\ket{+}=(\ket{+1}+\ket{-1})/\sqrt{2}$ as a function the magnetic field. Is equal to the projection of $\ket{g}$ on $\ket{0}$.}
\end{figure}

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig transverse field}
\caption{Modification of the stretch lifetime in the presence of purely transverse magnetic field. (a) Stretch component of the ensemble lifetime as a function of the field amplitude. (b) Measured transition frequencies through ODMR spectrum. (c) Stretch component of the lifetime as a function of the frequency detuning between the two transistions (blue circles), fitted by a Lorentzian centered in $\Delta \nu=0$ with half width at half maximum 8MHz. The green dashed line correspond to the lifetime of a single class aligned with the magnetic field.}
\end{figure}

\section*{acknowledgements}

\bibliography{CR}

\end{document}

### 2022-04-20 Mes papiers/Low field CR/Low field CR.tex
tata

@@
 \maketitle
+Blabla sur les NV, les ensembles etc.
+
+In this letter/article, we characterize the depolarization of the spins observed for dense ensemble of NV centers in zero magnetic field and its potential application for DC magnetometry.
+While the main mechanism behind the depolarization, the lift in the degeneracy between the four classes of NV centers, is already well studied and can be exploited in a microwave-less vector magnetometry protocol, we found two other depolarization mechanisms specific to the zero-field region which could play an important role in the low-field magnetometry protocol.
+
+The $T_1$ depolarization dynamics of single or sparse NV centers at room temperature is dominated by a two-phonon Raman process, which depends on the crystal lattice temperature but does not (?) depend on the external magnetic field.
+However, it has been observed that dense NV centers ensemble have an additional spin lifetime contribution which depends greatly on the magnetic field and not on the temperature.
+This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling.
+Some inhomogeneity between the NV centers is further needed in order to explain the depolarization of the spin ensemble.
+
+The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig [X].
+The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state.
+The reason for the decrease in PL at higher magnetic field values is the mixing of the bright spin state $\ket{0}$ to the darker state $\ket{-1}$ induced by the transverse magnetic field.
+This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.
+
+The additional spin depolarization observed
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures/fig dense vs pas dense}
+\caption{Todo}
+\label{PL_NV_density}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures/fig T1}
+\caption{Todo}
+\label{T1}
+\end{figure}
+
+\begin{figure*}
+\includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
+\caption{Dependency of the magnetic field angle for the zero field depolarization.
+(a) ODMR spectrum in zero field.
+(b) ODMR spectrum for a magnetic field $\approx$ 60 G.
+The field is aligned (b2) or misaligned (b1, $\approx 24^\circ$ of misalignment) with the crystalline [100] direction.
+(c) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude for the same field orientations as fig (b).
+(d) Stretch part of the lifetime decay as a function of the magnetic field amplitude for the same field orientations as fig (b).}
+\label{100_VS_1x4}
+\end{figure*}
 
@@
 Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
+\label{largeur_fluct}
 \end{figure}
@@
 Is equal to the projection of $\ket{g}$ on $\ket{0}$.}
+\label{calculs_B_transverse}
 \end{figure}
@@
 The green dashed line correspond to the lifetime of a single class aligned with the magnetic field.}
+\label{exp_B_transverse}
 \end{figure}

### 2022-04-20 idées/To_do_list.tex
tata

@@
  \item Mesurer des largeurs fluctateurs avec le T1 sur plein d'échantillons (de Ludo au rose) (et vérifier que PL=T1 si possible, mais ça rajoute des contraintes pour rien...).
+ \item Scan de pseudo-T1 (juste 2 points : S1 au début, S2 au milieu et (S1-S2)/S1 ) en fonction de la fréquence. Je viens de réfléchir et c'est pas si malin
+
+ \item faire une courbe sensibilité (magnetometry) en fonction de l'amplitude du champ mag modulé, selon la 100 et selon une 1x1x1x1
+ \item Mettre en place le protocole de magnetometrie par feedback (avec pyrpl probablement) puis faire une carte, a terme sur des 1/10 um sur un circuit imprimé.
+ \item La dépendance PL en fonction du champ transverse est pas la meme pour des échantillons denses et pas dense ?
+Vérfier que c'est pas une blague de l'APD-compteur-de-photon mesurée en tension.
+Peut-etre aussi une blague de filtrage spatial avec la fibre.
+ \item les différentes phases pour le T1 dipolaire, l'ESR et le champ transverse.
+ \end{itemize}
+
+ \section{Done}
+ \begin{itemize}
+  \item le 12C et 15N en champ transverse n'a qu'une bosse (normal) et la bosse du + et beaucoup plus petite que la bosse du -, et ca c'est pas normal.
+Ca vient vraiment du champ transverse vu que ça dépend beaucoup de l'angle et peu de l'amplitude du champ mag (donc des fréquences : pas un pb de micro-onde)
+
+  $\to$ C'est un problème de polarisation de la micro-onde : en champ transverse (E ou B d'ailleurs) les transitions sont senseibles à une polarisation longitudinale de la micro-onde : Sx pour $\ket{0} \to \ket{+}$ et Sy pour $\ket{0} \to \ket{-}$ par exemple, alors qu'en champ mag longitudinal, elles sont sensibles à la polarisation circulaire.
+Comme ma micro-onde est polarisée linéairement mais pas circulairement, je vois une différence qu'en champ transverse.
+
+   \item T1 soustraction en fonction de la longueur d'onde de la micro-onde (à faible puissance du coup) en champ nul. Potentiellement en 121 a moitié splitté aussi.
+
+   $\to$ +/- done, j'ai surtout l'impression que ça suit le contraste ODMR.
+Faudrait que je vérifie qu'il y ait pas des bugs dans mon programme, surtout sur la procédure de fit (T1 plus faibles pour un signal moins bon).
  \end{itemize}

### 2022-04-20 data/20220324/12C et 15N le retour/traitement.py

@@ -4,23 +4,34 @@ sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
 
-x,y=extract_data('ESR B transverse 70 G 15 dB')
-n=len(y)
-x=np.linspace(x[0],x[-1],n)
+def plot_ESR():
+	x,y=extract_data('ESR B transverse 70 G 15 dB')
+	n=len(y)
+	x=np.linspace(x[0],x[-1],n)
 
 
-# x2=list(x[0:85])+list(x[95:139])
-# y2=list(y[0:85])+list(y[95:139])
-x2=x[:110]
-y2=y[:110]
+	# x2=list(x[0:85])+list(x[95:139])
+	# y2=list(y[0:85])+list(y[95:139])
+	x2=x[:110]
+	y2=y[:110]
+	plt.plot(x,y)
+	# plt.plot(x2,y2,'x')
+
+	popt,yfit=lor_fit(x2,y2,x0=2.879)
+	def f(x,amp,x0,sigma,ss) :
+		return ss+amp*1/(1+((x-x0)/(sigma))**2)
+	plt.plot(x,f(x,*popt),label='HWHM=%.0f kHz'%(popt[2]*1e6))
+	plt.legend()
+
+
+plt.figure(num=1,figsize=(3,2),dpi=80)
+plt.xticks(fontsize=11)
+plt.yticks(fontsize=12)
+x,y=extract_data('scan EM weekend',ycol=3)
+y=y/max(y)
+x=x-x[259]
+x=x*35
 plt.plot(x,y)
-# plt.plot(x2,y2,'x')
-
-popt,yfit=lor_fit(x2,y2,x0=2.879)
-def f(x,amp,x0,sigma,ss) :
-	return ss+amp*1/(1+((x-x0)/(sigma))**2)
-plt.plot(x,f(x,*popt),label='HWHM=%.0f kHz'%(popt[2]*1e6))
-plt.legend()
 plt.show()
 
 

### 2022-04-20 data/20220324/Adamas 150 um/adamas 2/T1 1x1x1x1/traitement.py

@@ -4,6 +4,13 @@ sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
 
+plt.figure(num=1,figsize=(3,2),dpi=80)
+x,y=extract_data('ESR/V=-2.040000')
+y=y/max(y)
+plt.plot(x,y)
+plt.xticks(fontsize=11)
+plt.yticks(fontsize=12)
+
 fnames,fval=extract_glob('ESR')
 n=len(fval)
 
@@ -56,26 +63,26 @@ n=len(fval)
 
 
 
-fnames,fval=extract_glob('T1')
-n=len(fval)
-taus=np.zeros(n)
+# fnames,fval=extract_glob('T1')
+# n=len(fval)
+# taus=np.zeros(n)
 
-Bs=[np.array([78.55,31.8,16.86])*(x-0.17)/2.8 for x in fval]
+# Bs=[np.array([78.55,31.8,16.86])*(x-0.17)/2.8 for x in fval]
 
-Bamps=[norm(B)*np.sign(B[0]) for B in Bs]
+# Bamps=[norm(B)*np.sign(B[0]) for B in Bs]
 
-for i in range(n):
-	fname=fnames[i]
-	x,y=extract_data(fname,ycol=5)
-	T1ph=T1ph=0.003626
-	popt,yfit=stretch_et_phonons(x,y,T1ph=T1ph)
-	taus[i]=popt[1]
+# for i in range(n):
+# 	fname=fnames[i]
+# 	x,y=extract_data(fname,ycol=5)
+# 	T1ph=T1ph=0.003626
+# 	popt,yfit=stretch_et_phonons(x,y,T1ph=T1ph)
+# 	taus[i]=popt[1]
 
-nmin=106
-x=Bamps[nmin:]
-y=1/taus[nmin:]
+# nmin=106
+# x=Bamps[nmin:]
+# y=1/taus[nmin:]
 
-plt.plot(x,y)
+# plt.plot(x,y)
 
 
 plt.show()

### 2022-04-20 Mes papiers/Low field CR/Low field CR.tex
tata

@@
 
-The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig [X].
+The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}.
 The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state.
@@
 This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.
+
+The spin lifetime protocol used here is described in Fig.
+\ref{T1} and is based on previous similar experiments.
+It consists in a classical pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$.
+Using only this sequence can result in artifacts in the final $T_1$ signal due to charge state transfer in the dark, particularly for dense ensemble.
+It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to project the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
+By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.
+
+The ensemble spin lifetime is then fitted with a formula which include both an exponential lifetime $T_1^{\exp}$ and a stretch exponential lifetime $T_1^s$ :
+\begin{equation}
+S(\tau)=A \exp (-\frac{\tau}{T_1^{\exp}} -\sqrt{\frac{\tau}{T_1^s}}),
+\end{equation}
+We assume that the exponential part of the lifetime is defined by the phonons, and the stretch part by the coupling between NVs, as detailed below.
+In order to get comparable fitting parameters between each measurement, we decided to fix the value of the exponential lifetime in the fitting process, leaving only the stretch lifetime as a free parameter.
+For Fig.\ref{T1}, \ref{largeur_fluct}, \ref{100_VS_1x4} and \ref{exp_B_transverse}, the chosen exponential lifetime was $T_1^{\exp}=3.62\~ms$.
 
@@
 \includegraphics[width=0.45\textwidth]{Figures/fig dense vs pas dense}
-\caption{Todo}
+\caption{Photoluminescence from NV centers as a function of an external magnetic field randomly oriented for two samples : (a) a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
 \label{PL_NV_density}
@@
 \includegraphics[width=0.45\textwidth]{Figures/fig T1}
-\caption{Todo}
+\caption{Spin $T_1$ measurement protocol.
+(a) Sequence used to measure the spin lifetime : the laser is turned off for a time $\tau$ twice and the spin state is read at the end of the dark time by measuring the initial photoluminescence when the laser is turned back on.
+A microwave resonant with one of the spin transition is used in one of the sequence to project the $\ket{0}$ spin sate in either $\ket{+1}$ or $\ket{-1}$.
+The sequence result is the subtraction $S_1-S_2$.
+(b) Measurement of an ensemble spin lifetime, fitted with an expression $A \exp (-\frac{\tau}{T_1^{\exp}} -\sqrt{\frac{\tau}{T_1^s}})$}
 \label{T1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
+\caption{Tuning resonant coupling through magnetic field orientation.
+(a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice.
+(b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes.
+The detuning $\Delta \nu$ between the classes $\beta$ and $\gamma$ was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_\gamma$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz .
+Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
+\label{largeur_fluct}
 \end{figure}
@@
 \end{figure*}
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
-\caption{Modification of the stretch lifetime for two near-resonant classes.
-(a) ODMR spectrum with amplitude modulation of the microwave for the $\ket{0} \to \ket{-1}$ transition of two separate NV classes.
-The frequency detuning between the two transitions is denoted $\Delta \nu$.
-(b) Stretch part of the lifetime decay for one of the two classes as function of the frequency detuning (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz .
-Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
-\label{largeur_fluct}
-\end{figure}
 

### 2022-04-20 data/20220324/12C et 15N le retour/traitement.py

@@ -27,10 +27,33 @@ def plot_ESR():
 plt.figure(num=1,figsize=(3,2),dpi=80)
 plt.xticks(fontsize=11)
 plt.yticks(fontsize=12)
-x,y=extract_data('scan EM weekend',ycol=3)
+# x,y=extract_data('scan EM weekend',ycol=3)
+# y=y/max(y)
+# x=x-x[259]
+# x=x*35
+x,y=extract_data("scan PL à l'ancienne")
+x2=list(x)
+y2=list(y)
+for i in range(len(y)):
+	e=y[i]
+	if e <0 :
+		x2.remove(x[i])
+		y2.remove(y[i])
+
+x=np.array(x2)
+y=np.array(y2)
+
+def find_B_EM():
+	x,y=extract_data("ESR EM 3V pour le 2e scan PL")
+	plt.plot(x,y)
+	peaks=[2723,2723,2758,2758,3044,3050,3071,3075]
+	B=find_B_cartesian_mesh(peaks)
+	print(B,B.amp/3)
+	# B=76.9 G @ 3V
+x=x-x[104]
+x=x*30
+x=-x
 y=y/max(y)
-x=x-x[259]
-x=x*35
 plt.plot(x,y)
 plt.show()
 

### 2022-04-25 Mes papiers/Low field CR/Low_field_CR.tex
tata yoyo

\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-2}
%\documentclass[preprintnumbers,amsmath,amssymb]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}

\bibliographystyle{plain}

%\newcommand{|}{Y$_2$SiO$_5$}
\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{Low field depolarization of electronic spins through dipole-dipole coupling}

\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, G. H\'etet}

\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e Paris Cit\'e , 24 rue Lhomond, 75231 Paris Cedex 05, France.}

\begin{abstract}
C'est trop bien
\end{abstract}

\maketitle
Blabla sur les NV, les ensembles etc.

%réécrire le début et foutre la partie sur la PL avant le T1. Peut-etre dans le blabla sur les ensembles justement

%We have identified three mechanisms for the low field depolarization, two of which have not been reported before.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig dense vs pas dense}
\caption{Photoluminescence from NV centers as a function of an external magnetic field randomly oriented for two samples : (a) a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
\label{PL_NV_density}
\end{figure}

In this letter/article, we characterize the depolarization of the spins observed for dense ensemble of NV centers in zero magnetic field and its potential application for DC magnetometry. While the main mechanism behind the depolarization, the lift in the degeneracy between the four classes of NV centers, is already well studied and can be exploited in a microwave-less vector magnetometry protocol, we found two other depolarization mechanisms specific to the zero-field region which could play an important role in the low-field magnetometry protocol.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig T1}
\caption{Spin $T_1$ measurement protocol. (a) Sequence used to measure the spin lifetime : the laser is turned off for a time $\tau$ twice and the spin state is read at the end of the dark time by measuring the initial photoluminescence when the laser is turned back on. A microwave resonant with one of the spin transition is used in one of the sequence to project the $\ket{0}$ spin sate in either $\ket{+1}$ or $\ket{-1}$. The sequence result is the subtraction $S_1-S_2$. (b) Measurement of an ensemble spin lifetime, fitted with an expression $A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$}
\label{T1}
\end{figure}

The $T_1$ depolarization dynamics of single or sparse NV centers at room temperature is dominated by a two-phonon Raman process, which depends on the crystal lattice temperature but does not (?) depend on the external magnetic field. However, it has been observed that dense NV centers ensemble have an additional spin lifetime contribution which depends greatly on the magnetic field and not on the temperature. This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling. Some inhomogeneity between the NV centers is further needed in order to explain the depolarization of the spin ensemble. We will denote $T_1^{\rm ph}$ the characteristic time associated with the phonon relaxation process and $T_1^{\rm dd}$ the characteristic time associated with the dipole-dipole relaxation process.

The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}. The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state. The reason for the decrease in PL at higher magnetic field values is the mixing of the bright spin state $\ket{0}$ to the darker state $\ket{-1}$ induced by the transverse magnetic field. This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.

The spin lifetime protocol used here is described in Fig. \ref{T1} and is based on previous similar experiments. It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$. Using only this sequence can result in artifacts in the signal, mostly due to charge state transfer in the dark \citep{giri_coupled_2018}. It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to project the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state. By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.

%Pluto commencer par : assuming that the dipolar T1 is stretch exp... La ca sort un peu de nulle part

The ensemble spin lifetime is non-exponential due to inhomogeneities between the NV$^-$ centers. We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}. A conclusion of this model is that, for an homogeneous 3D distribution of fluctuator, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$. Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$  (see SI).

For the samples used in this paper (detailed in SI), we observed that $T_1^{\rm dd} \sim T_1^{\rm ph}$, meaning that we had to include both  in our analysis. The resulting fitting formula is therefore :
\begin{equation}
S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}}),
\end{equation}
In order to get comparable fitting parameters between each measurement, we decided to fix the value of the exponential lifetime in the fitting process, leaving only the stretch lifetime as a free parameter. For Fig.\ref{T1}, \ref{largeur_fluct}, \ref{100_VS_1x4} and \ref{exp_B_transverse}, the chosen exponential lifetime was $T_1^{\rm ph}=3.62 \ \rm ms$.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
\caption{Tuning resonant coupling through magnetic field orientation. (a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice. (b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes. The detuning $\Delta \nu$ between the classes $\beta$ and $\gamma$ was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_\gamma$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz . Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
\label{largeur_fluct}
\end{figure}

Fig. \ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins. Because of the diamond symmetry, we can distinguish 4 groups ("classes") of identical NV centers, based on the four possible orientation of the NV axis in the crystal. These four classes have \textit{a priori} different spin resonance frequency based on the projection of the magnetic field on their respective NV axis. By tuning the angle of the magnetic field, we can therefore control the frequency detuning between the different classes, and as a result we can control the number of resonant NV centers by bringing in and out of resonance several classes.

Fig. \ref{largeur_fluct} (c) in particular shows the dipole-induced lifetime $T_1^{\rm dd}$ for a given class as a function of the energy detuning with another class. One can notice that the width of the $T_1^{\rm dd}$ bell shape is significantly larger than that of the frequency overlap between the two classes, modeled here by an ODMR line stretched by a factor $\sqrt{2}$ (see SI). This extra width can not be attributed to the dipole-dipole coupling strength : for a 3 ppm concentration of NV$^-$ centers, the average coupling strength between neighboring NV centers should be $\sim 27\ \rm kHz$, two order of magnitude below the observed width. We can also see that $T_1^{\rm dd}$ follows a Lorentzian shape despite the ODMR lines being non-Lorentzian. These two observations corroborate the idea developed in \cite{choi_depolarization_2017} that the interaction width is dominated by the noise spectrum of the lifetime-limited fluctuators.

\begin{figure*}
\includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
\caption{Dependency of the magnetic field angle for the zero field depolarization. (a) ODMR spectrum in zero field. (b) ODMR spectrum for a magnetic field $\approx$ 60 G. The field is aligned (b2) or misaligned (b1, $\approx 24^\circ$ of misalignment) with the crystalline [100] direction. (c) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude for the same field orientations as fig (b). (d) Stretch part of the lifetime decay as a function of the magnetic field amplitude for the same field orientations as fig (b).}
\label{100_VS_1x4}
\end{figure*}

The depolarization in low field can be explained in large part due to the lift of degeneracy between the four classes, as can be seen in Fig. \ref{100_VS_1x4} : On the top part, a magnetic field is scanned in a direction chosen in order to quickly split the transition frequencies of the four NV classes. We can observe an increase of about 4\% in the photoluminescence and an increase by a factor of $\sim$ 16 in the spin dipolar-induced lifetime ${T_1^{\rm dd}}$ as the magnetic field is increased.

However the splitting of the four classes can not explain entirely the spin depolarization in zero field. The bottom part of Fig. \ref{100_VS_1x4} shows a similar experiment when the magnetic field is aligned with the [100] crystalline axis. In this particular orientation, the projection of the magnetic field on the four NV classes is identical, meaning that the resonance condition between the NV centers remains the same in zero and non-zero field. Yet we can still observe an increase of $\sim$ 0.5\% in the photoluminescence and an increase of $\sim$ 30\% in ${T_1^{\rm dd}}$.
We attribute the remaining depolarization to two phenomena : the change of basis in the spin Hamiltonian's eigenstate in zero-field, and the near-resonance of the dipolar double-flip processes.

The decrease of spin lifetime and the associated drop in PL for $B\approx 20\ \rm G$ is due to cross-relaxations between NV $^-$ centers with a $^{13}$C as first a neighbor, and those without \cite{pellet2021optical}.

The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$reads :
\begin{equation}
\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
\end{equation}
Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$. In situations such as the one described in Fig. \ref{largeur_fluct}, the only relevant terms of $\mathcal{H}^{\rm dd}$ in term of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$. In zero magnetic field however, we have to take into account other aspects of the dipolar Hamiltonian in consideration.

First is the change of basis of the single spin Hamiltonian $\mathcal{H}_s$ : in zero external magnetic field, the eigenstates of $\mathcal{H}_s$ in the $\{\ket{+1},\ket{-1}\}$ manifold are determined by local electric and magnetic field coming from nearby impurities \cite{mittiga2018imaging}. For the samples used in this study, $\mathcal{H}_s$ was dominated by local electric field (see SI), meaning that the proper eigenstates of $\mathcal{H}_s$ in zero magnetic field are $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $

Because the splitting between $\ket{+}$ and $\ket{-}$ ($\approx 9\ \rm MHz$) is much greater than the dipole-dipole interaction ($J_0 / r^3 \approx 30\ \rm kHz$), the flip-flop terms of $\mathcal{H}^{\rm dd}$ now reads $\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}$. Averaging over all relative positions of two aligned NV centers, we found that on average $\abs{\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}}$ was greater than 
$\abs{\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}}$ by a factor $1.8 \sim 2$ depending on the correlation lengths of the local electric field.
When all four classes are resonant, the theoretical decrease in the spin lifetime due to the change of basis is $\sim 20 \%$ (See SI de ouf).

Second is the near-resonance condition of the double-flip processes such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{-1,0}$ or $\mel{0,+}{\mathcal{H}^{\rm dd}}{-,0}$. These terms usually couple non resonant states due to the energy mismatch between $\ket{+1}$ and $\ket{-1}$. However in zero field, the energy splitting between $\ket{+}$ and $\ket{-}$ is comparable with the fluctuator's spectral width measured in Fig. \ref{largeur_fluct} to be $\approx 8\ \rm MHz$, making the double-flip population transfer possible.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig transverse field simu}
\caption{Simulated eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field. (a) Energies of the three eigenstates $\ket{g}$, $\ket{d}$ and $\ket{e}$ as a function of the magnetic field amplitude. (b) Blue dashed curve : frequency detuning between the two transitions $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$. Plain red curve : projection of $\ket{e}$ on $\ket{+}=(\ket{+1}+\ket{-1})/\sqrt{2}$ as a function the magnetic field. Is equal to the projection of $\ket{g}$ on $\ket{0}$.}
\label{calculs_B_transverse}
\end{figure}

In order to evaluate the relative contribution of these two factors, we investigate the spin relaxation in the case of pure transverse magnetic field. Calling $\{ \ket{g}, \ket{d}, \ket{e} \}$ the three eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field, Fig. \ref{calculs_B_transverse} shows that for small enough magnetic field, the $\{ \ket{g}, \ket{d}, \ket{e} \}$ basis is almost equal to the $\{ \ket{0},\ket{+},\ket{-} \} $ basis : for $B=120\ \rm G$, $\braket{0}{g}\approx 0.99$, $\braket{-1}{d}=1$ and $\braket{+1}{e}\approx 0.99$. However the energy difference between $\ket{d}$ and $\ket{e}$ reaches $\approx 45\ \rm G$ which is high enough to cancel out the double-flip processes and allows us to isolate the change of basis hypothesis from the double-flip one.
%As shown in Fig. SOON, (and in SI), the $\{ \ket{0},\ket{+},\ket{-} \} $ basis correspond to the spin Hamiltonian eigenstates in two situations : either in zero magnetic field when the electric field dominates, or when the magnetic field is purely transverse, as long as it remains small before the zero-field splitting. %(rq : est-ce que je dois mettre le hamiltonien de spin dans le main text du coup ?).

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig transverse field}
\caption{Modification of the stretch lifetime in the presence of purely transverse magnetic field. (a) Stretch component of the ensemble lifetime as a function of the field amplitude. (b) Measured transition frequencies through ODMR spectrum. (c) Stretch component of the lifetime as a function of the frequency detuning between the two transistions (blue circles), fitted by a Lorentzian centered in $\Delta \nu=0$ with half width at half maximum 8MHz. The green dashed line correspond to the lifetime of a single class aligned with the magnetic field.}
\label{exp_B_transverse}
\end{figure}

Fig. \ref{exp_B_transverse} shows the dipole-dipole spin relaxation $T_1^{\rm dd}$ for a class of NV whose axis is orthogonal to the applied magnetic field. By monitoring the frequencies of the transitions $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$, we can deduce the detuning $\Delta \nu$ between the two-spin states $\ket{0,+}$ and $\ket{-,0}$. Plotting then $1/T_1^{\rm dd}$ as a function of $\Delta \nu$, we can notice two things.

First the decrease of $1/T_1^{\rm dd}$ as $\Delta \nu$ increases corresponds to the loss of effectiveness of the double flip processes. The shape of the curve matches decently well with the tail of a Lorentzian centered in $\Delta \nu=0$ with a width of 8 MHz, confirming that the double-flip process interaction range is likely determined by the fluctuators noise spectrum just like the flip-flop processes.

Second is the plateau reached for $B \gtrapprox 30\ \rm MHz$. The final $1/T_1^{\rm dd}$ value is about twice as high as that for a spin well aligned with the magnetic field, showing the increased depolarization due to the change of basis.

This experiment shows that, for the sample studied here, the depolarization due to the double-flip processes dominates the one due to the change of basis in zero magnetic field.

\section*{Magnetometry}
DC microwave-less magnetometry has already been performed with NV ensembles, using either cross-relaxations [les russes] or level anti-crossing [bubu]. Here we propose to perform a similar protocol, but using the spin depolarization in zero-field. The main difference with the previously mentioned protocols is the fact that this one doesn't rely on crystalline orientation, making this protocol usable with diamond powders or polycrystalline samples.
\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig_magneto}
\caption{Low field magnetometry protocol. (a) Demodulated photoluminescence as a function of an externally applied magnetic field with an additional oscillatory magnetic field. (b) Measured magnetic field when alternating a small external magnetic field offset. (c) Histogram of the measurement in Fig. (b) fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$}. (d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.
\label{magneto}
\end{figure}

\section*{acknowledgements}

\bibliographystyle{plain}
\bibliography{CR}{}

\end{document}

### 2022-04-25 Mes papiers/Low field CR/Figures/Champ transverse états et energies.py

@@ -3,56 +3,93 @@ sys.path.append('D:\\These Clément\\these\\python_clément')
 sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
-plt.figure(num=1,figsize=(3,2),dpi=80)
 
-n=200
-levels=np.zeros((n,3))
-Bs=np.linspace(0,200,n)
-DeltaE=np.zeros(n)
-DeltaVect=np.zeros(n)
-DeltaVect2=np.zeros(n)
+def plot_NRJ():
+	plt.figure(num=1,figsize=(3,2),dpi=80)
+	plt.xticks(fontsize=11)
+	plt.yticks(fontsize=12)
 
-
-
-
-for i in range(n):
-	theta=90*pi/180
-	B=Bs[i]
-	H=NVHamiltonian(B=[B*sin(theta),0,B*cos(theta)],c=5)
-	E=H.egval()
-	V=H.egvect()
-	levels[i,:]=E
-	DeltaE[i]=E[2]-E[1]
-	DeltaVect[i]=(V[2].dot([1/sqrt(2),0,1/sqrt(2)]))**2
-	DeltaVect2[i]=(V[0].dot([0,1,0]))**2
-	
-	# DeltaVect[i]=abs(V[2].dot([0,1,0]))
-	# DeltaVect[i]=abs(V[0].dot([0,1,0]))
-
-
-# f,(ax,ax2)=plt.subplots(2,1,sharex=True)
-
-# f.set_size_inches((3,2))
-
-# ax2.plot(Bs,levels[:,0],color=color(0))
-# ax.plot(Bs,levels[:,1],color=color(0))
-# ax.plot(Bs,levels[:,2],color=color(0))
-
-# ax.spines['bottom'].set_visible(False)
-# ax2.spines['top'].set_visible(False)
-# ax.xaxis.tick_top()
-# ax.tick_params(labeltop=False)  # don't put tick labels at the top
-# ax2.xaxis.tick_bottom()
-
-# plt.xlabel('Magnetic Field (G)',fontsize=15)
+	n=200
+	levels=np.zeros((n,3))
+	Bs=np.linspace(0,200,n)
+	DeltaE=np.zeros(n)
+	DeltaVect=np.zeros(n)
+	DeltaVect2=np.zeros(n)
 
 
 
-ax1=plt.gca()
[…]

### 2022-04-27 Mes papiers/Low field CR/Low_field_CR.tex
toto

@@
 
+The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}.
+The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state.
+%The reason for the decrease in PL at higher magnetic field values is the mixing of the bright spin state $\ket{0}$ to the darker state $\ket{-1}$ induced by the transverse magnetic field.
+This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.
+
 \begin{figure}
@@
 The $T_1$ depolarization dynamics of single or sparse NV centers at room temperature is dominated by a two-phonon Raman process, which depends on the crystal lattice temperature but does not (?) depend on the external magnetic field.
-However, it has been observed that dense NV centers ensemble have an additional spin lifetime contribution which depends greatly on the magnetic field and not on the temperature.
+However, it has been observed that dense NV centers ensemble have an additional spin decay channel which depends greatly on the magnetic field and not on the temperature.
 This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling.
@@
 We will denote $T_1^{\rm ph}$ the characteristic time associated with the phonon relaxation process and $T_1^{\rm dd}$ the characteristic time associated with the dipole-dipole relaxation process.
-
-The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}.
-The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state.
-The reason for the decrease in PL at higher magnetic field values is the mixing of the bright spin state $\ket{0}$ to the darker state $\ket{-1}$ induced by the transverse magnetic field.
-This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.
 
@@
 By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.
-
-%Pluto commencer par : assuming that the dipolar T1 is stretch exp... La ca sort un peu de nulle part
 
@@
 (a) Stretch component of the ensemble lifetime as a function of the field amplitude.
-(b) Measured transition frequencies through ODMR spectrum.
+(b) Transition frequencies for the $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$ transitions, measured through ODMR.
 (c) Stretch component of the lifetime as a function of the frequency detuning between the two transistions (blue circles), fitted by a Lorentzian centered in $\Delta \nu=0$ with half width at half maximum 8MHz.
@@
 \section*{Magnetometry}
-DC microwave-less magnetometry has already been performed with NV ensembles, using either cross-relaxations [les russes] or level anti-crossing [bubu].
-Here we propose to perform a similar protocol, but using the spin depolarization in zero-field.
-The main difference with the previously mentioned protocols is the fact that this one doesn't rely on crystalline orientation, making this protocol usable with diamond powders or polycrystalline samples.
 \begin{figure}
@@
 \end{figure}
+DC microwave-less magnetometry has already been performed with NV ensembles, using either cross-relaxations [les russes] or level anti-crossing [bubu].
+Here we propose to perform a similar protocol, but using the spin depolarization in zero-field.
+The main difference with the previously mentioned protocols is the fact that this one doesn't rely on crystalline orientation, making this protocol usable with diamond powders or polycrystalline samples.
 
+Fig.
+\ref{magneto} (a) shows the demodulated PL when a DC magnetic field is scanned in a random direction and an additional alternating magnetic field of a few Gauss at $\sim \ \rm kHz$ is added through the same electromagnet (see experimental details in SI).
+We can see a sharp and relatively linear slope in low field $\abs{B} < 5\ \rm G$.
+Once calibrated, in this case with ODMR, the slope can provide a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as seen in \cite{zheng_microwave-free_2020}.
+
+In order to measure the sensitivity of the measurement, we alternate a small DC field of $\approx 40 \rm G$ every few seconds and take an histogram of the measured fields, as shown in Fig.
+\ref{magneto} (b) and (c).
+The histogram is well fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
+The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which allows us to measure the sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nm}/\sqrt{\rm Hz}$.
+
+We can then try to measure the relative importance of the three causes of spin depolarization on the field sensitivity.
+In order to do so, we measure the sensitivity while changing the angle of the field.
+When $\vec B$ is aligned with the [100] crystalline axis, only the double flips and the change of basis are present, whereas in every other orientation the three effects are at play.
+
+The results are shown in Fig.
+\ref{magneto} (d).
+We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity is relatively flat, which means that the double-flips and change of basis effects are the dominant factors in the sensitivity of this protocol.
+%This is in agreement with the measurements in Fig.
+\ref{100_VS_1x4} : even though the contrast is much better in the not-[100] case, the slope in term of contrast percentage/magnetic field is comparable in both cases (en faite pas complètement, y'a pas loin d'un facteur 2 sur les dérivées numériques)
+It should be noted though that this observation is sample dependent, and that other samples, including from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from the double flips and change of basis.
+
+\section*{Conclusion}
+In this work we have identified three mechanisms causing the extra spin depolarization observed in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole cross-relaxations between the spins : the lift of degeneracy between the four classes caused by the magnetic field, the change of basis in the spin Hamiltonian when the transverse electric field becomes dominant, and the double-flip process allowed by the proximity in energy of the spin $\ket{\pm}$ states.
+We identified the lift in degeneracy as the main cause in the zero-field depolarization, followed by the double-flip process and then the change of basis.
+With have demonstrated a potential use of this depolarization as a DC microwave-less and orientation-free magnetometer with senstivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal.We have observed that in some cases the double-flips and change of basis depolarization play a more important role than the lift of degeneracy in the low field sensitivity.
 \section*{acknowledgements}

### 2022-04-29 Mes papiers/Low field CR/Low_field_CR.tex
tata

@@
 \documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-2}
-%\documentclass[preprintnumbers,amsmath,amssymb]{revtex4}
 \usepackage{graphicx}% Include figure files
@@
 \usepackage[caption=false]{subfig}
-
-\bibliographystyle{plain}
 
@@
 The sequence result is the subtraction $S_1-S_2$.
-(b) Measurement of an ensemble spin lifetime, fitted with an expression $A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$}
+(b) Measurement of a dense ensemble spin lifetime in zero field and non-zero field, fitted with an expression $A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$}
 \label{T1}
@@
 
-First the decrease of $1/T_1^{\rm dd}$ as $\Delta \nu$ increases corresponds to the loss of effectiveness of the double flip processes.
+First the decrease of $1/T_1^{\rm dd}$ as $\Delta \nu$ increases which corresponds to the loss in effectiveness of the double flip processes.
 The shape of the curve matches decently well with the tail of a Lorentzian centered in $\Delta \nu=0$ with a width of 8 MHz, confirming that the double-flip process interaction range is likely determined by the fluctuators noise spectrum just like the flip-flop processes.
 
-Second is the plateau reached for $B \gtrapprox 30\ \rm MHz$.
+Second is the plateau reached for $\Delta \nu \gtrapprox 30\ \rm MHz$.
 The final $1/T_1^{\rm dd}$ value is about twice as high as that for a spin well aligned with the magnetic field, showing the increased depolarization due to the change of basis.
@@
 \ref{magneto} (d).
-We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity is relatively flat, which means that the double-flips and change of basis effects are the dominant factors in the sensitivity of this protocol.
+We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity remains relatively flat, which means that the double-flips and change of basis effects are the dominant factors in the sensitivity of this protocol.
 %This is in agreement with the measurements in Fig.

### 2022-04-29 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
tata

--- 2022-04-25 Mes papiers/Low field CR/Low_field_CR.tex
@@
-\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-2}
-%\documentclass[preprintnumbers,amsmath,amssymb]{revtex4}
+\documentclass[preprintnumbers,amsmath,amssymb,onecolumn,12pt]{revtex4-2}
 \usepackage{graphicx}% Include figure files
@@
 \usepackage[caption=false]{subfig}
-
-\bibliographystyle{plain}
-
-%\newcommand{|}{Y$_2$SiO$_5$}
-\def\sgn{\mathop{\rm sgn}}
 \newcommand{\be}{\begin{equation}}
 \newcommand{\ee}{\end{equation}}
+
 \newcommand{\bea}{\begin{eqnarray}}
@@
 
+\def\sgn{\mathop{\rm sgn}}
+
 \begin{document}
-
+\vspace{0.2in}
+{\Large \hspace{1.6in}\textsc{Supplementary Material} }\\
+\
 \title{Low field depolarization of electronic spins through dipole-dipole coupling}
 
-\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, G. H\'etet}
+\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, G. H\'etet$^1$}
 
@@
 
-\begin{abstract}
-C'est trop bien
-\end{abstract}
+\maketitle
 
-\maketitle
-Blabla sur les NV, les ensembles etc.
+\tableofcontents
 
-%réécrire le début et foutre la partie sur la PL avant le T1. Peut-etre dans le blabla sur les ensembles justement
-
-%We have identified three mechanisms for the low field depolarization, two of which have not been reported before.
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig dense vs pas dense}
-\caption{Photoluminescence from NV centers as a function of an external magnetic field randomly oriented for two samples : (a) a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
-\label{PL_NV_density}
-\end{figure}
-
-In this letter/article, we characterize the depolarization of the spins observed for dense ensemble of NV centers in zero magnetic field and its potential application for DC magnetometry.
-While the main mechanism behind the depolarization, the lift in the degeneracy between the four classes of NV centers, is already well studied and can be exploited in a microwave-less vector magnetometry protocol, we found two other depolarization mechanisms specific to the zero-field region which could play an important role in the low-field magnetometry protocol.
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig T1}
-\caption{Spin $T_1$ measurement protocol.
-(a) Sequence used to measure the spin lifetime : the laser is turned off for a time $\tau$ twice and the spin state is read at the end of the dark time by measuring the initial photoluminescence when the laser is turned back on.
-A microwave resonant with one of the spin transition is used in one of the sequence to project the $\ket{0}$ spin sate in either $\ket{+1}$ or $\ket{-1}$.
-The sequence result is the subtraction $S_1-S_2$.
-(b) Measurement of an ensemble spin lifetime, fitted with an expression $A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$}
-\label{T1}
-\end{figure}
-
-The $T_1$ depolarization dynamics of single or sparse NV centers at room temperature is dominated by a two-phonon Raman process, which depends on the crystal lattice temperature but does not (?) depend on the external magnetic field.
-However, it has been observed that dense NV centers ensemble have an additional spin lifetime contribution which depends greatly on the magnetic field and not on the temperature.
-This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling.
-Some inhomogeneity between the NV centers is further needed in order to explain the depolarization of the spin ensemble.
-We will denote $T_1^{\rm ph}$ the characteristic time associated with the phonon relaxation process and $T_1^{\rm dd}$ the characteristic time associated with the dipole-dipole relaxation process.
-
-The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}.
-The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state.
-The reason for the decrease in PL at higher magnetic field values is the mixing of the bright spin state $\ket{0}$ to the darker state $\ket{-1}$ induced by the transverse magnetic field.
-This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.
-
-The spin lifetime protocol used here is described in Fig.
-\ref{T1} and is based on previous similar experiments.
-It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$.
-Using only this sequence can result in artifacts in the signal, mostly due to charge state transfer in the dark \citep{giri_coupled_2018}.
-It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to project the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
-By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.
-
-%Pluto commencer par : assuming that the dipolar T1 is stretch exp... La ca sort un peu de nulle part
-
-The ensemble spin lifetime is non-exponential due to inhomogeneities between the NV$^-$ centers.
-We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}.
-A conclusion of this model is that, for an homogeneous 3D distribution of fluctuator, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$.
-Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$  (see SI).
-
-For the samples used in this paper (detailed in SI), we observed that $T_1^{\rm dd} \sim T_1^{\rm ph}$, meaning that we had to include both  in our analysis.
-The resulting fitting formula is therefore :
-\begin{equation}
-S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}}),
-\end{equation}
-In order to get comparable fitting parameters between each measurement, we decided to fix the value of the exponential lifetime in the fitting process, leaving only the stretch lifetime as a free parameter.
-For Fig.\ref{T1}, \ref{largeur_fluct}, \ref{100_VS_1x4} and \ref{exp_B_transverse}, the chosen exponential lifetime was $T_1^{\rm ph}=3.62 \ \rm ms$.
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
-\caption{Tuning resonant coupling through magnetic field orientation.
-(a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice.
-(b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes.
-The detuning $\Delta \nu$ between the classes $\beta$ and $\gamma$ was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_\gamma$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz .
-Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
-\label{largeur_fluct}
-\end{figure}
-
-Fig.
-\ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins.
-Because of the diamond symmetry, we can distinguish 4 groups ("classes") of identical NV centers, based on the four possible orientation of the NV axis in the crystal.
-These four classes have \textit{a priori} different spin resonance frequency based on the projection of the magnetic field on their respective NV axis.
-By tuning the angle of the magnetic field, we can therefore control the frequency detuning between the different classes, and as a result we can control the number of resonant NV centers by bringing in and out of resonance several classes.
-
-Fig.
-\ref{largeur_fluct} (c) in particular shows the dipole-induced lifetime $T_1^{\rm dd}$ for a given class as a function of the energy detuning with another class.
-One can notice that the width of the $T_1^{\rm dd}$ bell shape is significantly larger than that of the frequency overlap between the two classes, modeled here by an ODMR line stretched by a factor $\sqrt{2}$ (see SI).
-This extra width can not be attributed to the dipole-dipole coupling strength : for a 3 ppm concentration of NV$^-$ centers, the average coupling strength between neighboring NV centers should be $\sim 27\ \rm kHz$, two order of magnitude below the observed width.
-We can also see that $T_1^{\rm dd}$ follows a Lorentzian shape despite the ODMR lines being non-Lorentzian.
-These two observations corroborate the idea developed in \cite{choi_depolarization_2017} that the interaction width is dominated by the noise spectrum of the lifetime-limited fluctuators.
-
-\begin{figure*}
-\includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
-\caption{Dependency of the magnetic field angle for the zero field depolarization.
-(a) ODMR spectrum in zero field.
-(b) ODMR spectrum for a magnetic field $\approx$ 60 G.
-The field is aligned (b2) or misaligned (b1, $\approx 24^\circ$ of misalignment) with the crystalline [100] direction.
-(c) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude for the same field orientations as fig (b).
-(d) Stretch part of the lifetime decay as a function of the magnetic field amplitude for the same field orientations as fig (b).}
-\label{100_VS_1x4}
-\end{figure*}
-
-The depolarization in low field can be explained in large part due to the lift of degeneracy between the four classes, as can be seen in Fig.
-\ref{100_VS_1x4} : On the top part, a magnetic field is scanned in a direction chosen in order to quickly split the transition frequencies of the four NV classes.
-We can observe an increase of about 4\% in the photoluminescence and an increase by a factor of $\sim$ 16 in the spin dipolar-induced lifetime ${T_1^{\rm dd}}$ as the magnetic field is increased.
-
-However the splitting of the four classes can not explain entirely the spin depolarization in zero field.
-The bottom part of Fig.
-\ref{100_VS_1x4} shows a similar experiment when the magnetic field is aligned with the [100] crystalline axis.
-In this particular orientation, the projection of the magnetic field on the four NV classes is identical, meaning that the resonance condition between the NV centers remains the same in zero and non-zero field.
-Yet we can still observe an increase of $\sim$ 0.5\% in the photoluminescence and an increase of $\sim$ 30\% in ${T_1^{\rm dd}}$.
-We attribute the remaining depolarization to two phenomena : the change of basis in the spin Hamiltonian's eigenstate in zero-field, and the near-resonance of the dipolar double-flip processes.
-
-The decrease of spin lifetime and the associated drop in PL for $B\approx 20\ \rm G$ is due to cross-relaxations between NV $^-$ centers with a $^{13}$C as first a neighbor, and those without \cite{pellet2021optical}.
-
-The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$reads :
-\begin{equation}
-\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
-\end{equation}
-Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$.
-In situations such as the one described in Fig.
-\ref{largeur_fluct}, the only relevant terms of $\mathcal{H}^{\rm dd}$ in term of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$.
-In zero magnetic field however, we have to take into account other aspects of the dipolar Hamiltonian in consideration.
-
-First is the change of basis of the single spin Hamiltonian $\mathcal{H}_s$ : in zero external magnetic field, the eigenstates of $\mathcal{H}_s$ in the $\{\ket{+1},\ket{-1}\}$ manifold are determined by local electric and magnetic field coming from nearby impurities \cite{mittiga2018imaging}.
-For the samples used in this study, $\mathcal{H}_s$ was dominated by local electric field (see SI), meaning that the proper eigenstates of $\mathcal{H}_s$ in zero magnetic field are $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $
-
-Because the splitting between $\ket{+}$ and $\ket{-}$ ($\approx 9\ \rm MHz$) is much greater than the dipole-dipole interaction ($J_0 / r^3 \approx 30\ \rm kHz$), the flip-flop terms of $\mathcal{H}^{\rm dd}$ now reads $\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}$.
-Averaging over all relative positions of two aligned NV centers, we found that on average $\abs{\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}}$ was greater than
-$\abs{\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}}$ by a factor $1.8 \sim 2$ depending on the correlation lengths of the local electric field.
-When all four classes are resonant, the theoretical decrease in the spin lifetime due to the change of basis is $\sim 20 \%$ (See SI de ouf).
-
-Second is the near-resonance condition of the double-flip processes such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{-1,0}$ or $\mel{0,+}{\mathcal{H}^{\rm dd}}{-,0}$.
-These terms usually couple non resonant states due to the energy mismatch between $\ket{+1}$ and $\ket{-1}$.
-However in zero field, the energy splitting between $\ket{+}$ and $\ket{-}$ is comparable with the fluctuator's spectral width measured in Fig.
-\ref{largeur_fluct} to be $\approx 8\ \rm MHz$, making the double-flip population transfer possible.
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig transverse field simu}
-\caption{Simulated eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field.
-(a) Energies of the three eigenstates $\ket{g}$, $\ket{d}$ and $\ket{e}$ as a function of the magnetic field amplitude.
-(b) Blue dashed curve : frequency detuning between the two transitions $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$.
-Plain red curve : projection of $\ket{e}$ on $\ket{+}=(\ket{+1}+\ket{-1})/\sqrt{2}$ as a function the magnetic field.
-Is equal to the projection of $\ket{g}$ on $\ket{0}$.}
-\label{calculs_B_transverse}
-\end{figure}
-
-In order to evaluate the relative contribution of these two factors, we investigate the spin relaxation in the case of pure transverse magnetic field.
-Calling $\{ \ket{g}, \ket{d}, \ket{e} \}$ the three eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field, Fig.
-\ref{calculs_B_transverse} shows that for small enough magnetic field, the $\{ \ket{g}, \ket{d}, \ket{e} \}$ basis is almost equal to the $\{ \ket{0},\ket{+},\ket{-} \} $ basis : for $B=120\ \rm G$, $\braket{0}{g}\approx 0.99$, $\braket{-1}{d}=1$ and $\braket{+1}{e}\approx 0.99$.
-However the energy difference between $\ket{d}$ and $\ket{e}$ reaches $\approx 45\ \rm G$ which is high enough to cancel out the double-flip processes and allows us to isolate the change of basis hypothesis from the double-flip one.
-%As shown in Fig.
-SOON, (and in SI), the $\{ \ket{0},\ket{+},\ket{-} \} $ basis correspond to the spin Hamiltonian eigenstates in two situations : either in zero magnetic field when the electric field dominates, or when the magnetic field is purely transverse, as long as it remains small before the zero-field splitting.
-%(rq : est-ce que je dois mettre le hamiltonien de spin dans le main text du coup ?).
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig transverse field}
-\caption{Modification of the stretch lifetime in the presence of purely transverse magnetic field.
-(a) Stretch component of the ensemble lifetime as a function of the field amplitude.
-(b) Measured transition frequencies through ODMR spectrum.
-(c) Stretch component of the lifetime as a function of the frequency detuning between the two transistions (blue circles), fitted by a Lorentzian centered in $\Delta \nu=0$ with half width at half maximum 8MHz.
-The green dashed line correspond to the lifetime of a single class aligned with the magnetic field.}
-\label{exp_B_transverse}
-\end{figure}
-
-Fig.
-\ref{exp_B_transverse} shows the dipole-dipole spin relaxation $T_1^{\rm dd}$ for a class of NV whose axis is orthogonal to the applied magnetic field.
-By monitoring the frequencies of the transitions $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$, we can deduce the detuning $\Delta \nu$ between the two-spin states $\ket{0,+}$ and $\ket{-,0}$.
-Plotting then $1/T_1^{\rm dd}$ as a function of $\Delta \nu$, we can notice two things.
-
-First the decrease of $1/T_1^{\rm dd}$ as $\Delta \nu$ increases corresponds to the loss of effectiveness of the double flip processes.
-The shape of the curve matches decently well with the tail of a Lorentzian centered in $\Delta \nu=0$ with a width of 8 MHz, confirming that the double-flip process interaction range is likely determined by the fluctuators noise spectrum just like the flip-flop processes.
-
-Second is the plateau reached for $B \gtrapprox 30\ \rm MHz$.
-The final $1/T_1^{\rm dd}$ value is about twice as high as that for a spin well aligned with the magnetic field, showing the increased depolarization due to the change of basis.
-
-This experiment shows that, for the sample studied here, the depolarization due to the double-flip processes dominates the one due to the change of basis in zero magnetic field.
-
-\section*{Magnetometry}
-DC microwave-less magnetometry has already been performed with NV ensembles, using either cross-relaxations [les russes] or level anti-crossing [bubu].
-Here we propose to perform a similar protocol, but using the spin depolarization in zero-field.
-The main difference with the previously mentioned protocols is the fact that this one doesn't rely on crystalline orientation, making this protocol usable with diamond powders or polycrystalline samples.
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig_magneto}
-\caption{Low field magnetometry protocol.
-(a) Demodulated photoluminescence as a function of an externally applied magnetic field with an additional oscillatory magnetic field.
-(b) Measured magnetic field when alternating a small external magnetic field offset.
-(c) Histogram of the measurement in Fig.
-(b) fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$}.
-(d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.
-\label{magneto}
-\end{figure}
-
-\section*{acknowledgements}
+\section{Samples}
+\section{Experimental Setup}
+\section{Data analysis}
+\subsection{$T_1$ fitting Protocol}
+\subsection{Estimation of fluctuators width}
+\section{Les détails qui tuent}
+\subsection{Effect of laser polarization}
+\subsection{Alignment of B}
+La faudrait p-e montrer ce qu'il se passe pour un décalage de quelques degrés. A voir si j'ai déja les plots
+\section{Single spin Hamiltonian}
+\section{Fluctuator et tout}
 
 \bibliographystyle{plain}
-\bibliography{CR}{}
-
+\bibliography{CR_SI}{}
 \end{document}

### 2022-04-29 idées/To_do_list.tex
toto
tata

@@
 \documentclass[a4paper]{report}
-\usepackage[utf8]{inputenc}
 \usepackage[]{amsmath}
@@
  \item les différentes phases pour le T1 dipolaire, l'ESR et le champ transverse.
+ \item effet de la polarisation du laser (après le dichroique) sur le contraste ESR (quantifier la polarisation de chaque classe si possible, même si je controlle pas à 3D), puis sur le T1 de spin.
+(cf http://dx.doi.org/10.1088/1367-2630/17/2/023040)
+ \item Les coefs beta des stretch exponential (en fittant) ont l'air d'être modifié que par les double flips/base couplée (sur /these/data/20220324/Adamas 150 um/adamas 2).
+A vérifier si c'est le cas pour un croisement 121 et pour le champs transverse ou juste un hasard.
+ \item test magnetometry en mode 2x2 proches (genre sur le coté de la lorentzienne).
+p-e avec des bobines de Helmotz + aimant permanent.
+en faite je sais pas d'ou il faudrait partir, ce serait peut-etre plus simple juste sur une 121 pour commencer.
  \end{itemize}

### 2022-04-29 data/20220324/Adamas 150 um/adamas 1/T1/traitement.py

@@ -4,6 +4,8 @@ sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
 plt.figure(num=1,figsize=(4.5,3),dpi=80)
+plt.xticks(fontsize=12)
+plt.yticks(fontsize=12)
 
 #les x et y sont inversés pour cette série
 fnames=glob.glob('*.csv')
@@ -47,13 +49,13 @@ taus=taus[:nmax]
 
 x=deltaF
 y=1/taus
-plt.plot(x,y,'o',markerfacecolor='none')
+plt.plot(x,y,'o',markerfacecolor='none',ms=5,mew=0.7)
 popt,yfit=lor_fit_fixed(x,y,x0=0,sigma=8)
 print(popt)
-plt.plot(x,yfit)
+plt.plot(x,yfit,lw=2)
 
 gammaref=32.7
-plt.plot(deltaF,[gammaref]*nmax,'--')
+plt.plot(deltaF,[gammaref]*nmax,'--',lw=2)
 ax=plt.gca()
 ylims=ax.get_ylim()
 ax.set_ylim([0,ylims[1]])

### 2022-04-29 python_clément/analyse.py

@@ -847,15 +847,18 @@ def closest_elem(l,target):
 			basis=abs(target-l[i])
 	return n
 
-def estim_error(y,yfit):
+def estim_error(y,yfit,rel=True):
 	#C'est pas si simple, si tu prends juste l'erreur relative de chaque point tu donnes beaucoup plus de poids aux valeurs proches de 0. Le je fais un truc un peu sale mais qui donne autant de poids (absolu) à chaque point
 	n=len(y)
 	assert n==len(yfit)
 	y=np.array(y)
 	yfit=np.array(yfit)
 	vAvg=sum(abs(yfit))/n #le abs est crade mais au cas ou tu aies des valeurs positives et négatives
-	errors_rel=(y-yfit)**2/vAvg**2
-	return(sum(errors_rel)/n)
+	if rel :
+		errors=(y-yfit)**2/vAvg**2
+	else :
+		errors=(y-yfit)**2
+	return(sum(errors)/n)
 
 def lissage(t,n):
 	newt=np.array([sum(t[i:i+n])/n for i in range(len(t)-n)])
@@ -868,6 +871,22 @@ def derivative(x,y):
 	y2=np.array([(y[i+1]-y[i])/dx for i in range(n-1)])
 	x2=np.array([(x[i+1]+x[i])/2 for i in range(n-1)])
 	return(x2,y2)
+
+def find_local_min(x,y,x0):
+	i=find_elem(x,x0)
+	while y[i+1]<y[i]:
+		i+=1
+	while y[i-1]<y[i]:
+		i-=1
+	return i
+
+def find_local_max(x,y,x0):
+	i=find_elem(x,x0)
+	while y[i+1]>y[i]:
+		i+=1
+	while y[i-1]>y[i]:
+		i-=1
+	return i
 
 #~~~~~~ 2D plot ~~~~~~
 def extract_2d(fname):
@@ -919,18 +938,6 @@ def ecris_gros(x,y):
 	color = next(ax._get_lines.prop_cycler)['color']
 	plt.plot(x,y,'o',markerfacecolor="None",ms=8,mew=2,color=color)
 
-def extract_glob(SubFolderName='.',FirstValIndex='default', LastValIndex=-4): #FirstValIndex=premier caractère numérique
-	fnames=glob.glob(SubFolderName+'/*.csv')
-	if FirstValIndex=='default':
-		try :
-			FirstValIndex=fnames[0].index('=')+1
-		except :
-			raise(ValueError('Could not find a = sign in the file names'))
[…]

### 2022-04-29 Mes papiers/Low field CR/Low_field_CR.tex
atat

@@
 In this letter/article, we characterize the depolarization of the spins observed for dense ensemble of NV centers in zero magnetic field and its potential application for DC magnetometry.
-While the main mechanism behind the depolarization, the lift in the degeneracy between the four classes of NV centers, is already well studied and can be exploited in a microwave-less vector magnetometry protocol, we found two other depolarization mechanisms specific to the zero-field region which could play an important role in the low-field magnetometry protocol.
+While the main mechanism behind the depolarization, the lift in the degeneracy between the four classes of NV centers, is already well studied and can be exploited in a microwave-less vector magnetometry protocol, we found two other depolarization mechanisms specific to the zero-field region which could play an important role in a low-field magnetometry protocol.
 
@@
 
-Second is the plateau reached for $\Delta \nu \gtrapprox 30\ \rm MHz$.
+Second is the plateau reached for $\Delta \nu \gtrsim 30\ \rm MHz$.
 The final $1/T_1^{\rm dd}$ value is about twice as high as that for a spin well aligned with the magnetic field, showing the increased depolarization due to the change of basis.
@@
 The histogram is well fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
-The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which allows us to measure the sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nm}/\sqrt{\rm Hz}$.
+The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which allows us to measure the sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nT}/\sqrt{\rm Hz}$.
 

### 2022-04-29 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
atat

@@
 
+\section{NV Hamiltonian}
 \section{Samples}
@@
 \subsection{$T_1$ fitting Protocol}
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_0B}
+\caption{$T_1$ measurement in zero magnetic field with purely exponential and purely stretched exponential fits}
+\label{T1_0B}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_1x1x1x1}
+\caption{$T_1$ measurement in non-zero magnetic field with purely exponential and purely stretched exponential fits}
+\label{T1_1x4}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures_SI/fig_alphas}
+\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$}
+\label{alphas}
+\end{figure}
 \subsection{Estimation of fluctuators width}
@@
 \subsection{Effect of laser polarization}
+\begin{figure}
+\includegraphics[width=0.9\textwidth]{Figures_SI/fig_Pola}
+\caption{Effect of the polarization of the incident laser.
+(a) Photoluminescence as a function of randomly oriented magnetic field for various polarization angle.
+(b) Demodulated PL in the same conditions}
+\label{Pola}
+\end{figure}
 \subsection{Alignment of B}
 La faudrait p-e montrer ce qu'il se passe pour un décalage de quelques degrés. A voir si j'ai déja les plots
-\section{Single spin Hamiltonian}
 \section{Fluctuator et tout}
+\subsection{121 VS 22}
+\subsection{100 vs 0B}
 

### 2022-04-29 data/20220324/Adamas 150 um/adamas 2/traitement.py

@@ -4,7 +4,7 @@ sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
 
-scale=1
+scale=2
 plt.figure(num=1,figsize=(3*scale,2*scale),dpi=80)
 plt.xticks(fontsize=10+1.5*scale)
 plt.yticks(fontsize=10+1.5*scale)
@@ -43,11 +43,16 @@ def plot_T1__fit_main_text():
 	plt.plot(x,yfit,lw=2,color=color(1))
 	plt.legend()
 
-plot_T1__fit_main_text()
+# plot_T1__fit_main_text()
+
+#~~~~RQ : normaliser les fits des T1 n'a pas de sense tant que je ne connais pas la vraie valeur en t=0
+#En plus y'a des p-e des blagues de tdv de l'état metastable pour les temps très courts, à voir si ça se soutrait correctement
+#Y'a aussi le fait que je suis obligé d'attendre plus longtemps que le pulse uW, sinon ça fausse le truc (genre tu peux avoir du spin lockin ou chais pas)
 
 def plot_T1_fit(i=104):
 	fnames,fval=extract_glob('T1 1x1x1x1/T1')
 	x,y=extract_data(fnames[i],ycol=5)
+	# x=x-x[0]
 	y=y/max(y)
 	x=x*1e3
 	
@@ -57,13 +62,14 @@ def plot_T1_fit(i=104):
 	# plt.plot(x,yfit)
 	T1ph=0.003626*1e3
 	plt.plot(x,y,'x',markerfacecolor='None',mew=1.2,ms=8,color=color(0),label='Experimental data')
-	popt,yfit=stretch_exp_fit_zero(x,y)
+	popt,yfit=stretch_exp_fit_zero(x,y,norm=False)
 	plt.plot(x,yfit,'--',lw=2,color=color(1),label=r'$\exp (-\sqrt{\frac{\tau}{T_1^{\rm dd}}})$')
-	popt,yfit=exp_fit_zero(x,y)
+	popt,yfit=exp_fit_zero(x,y,norm=False)
 	plt.plot(x,yfit,'-.',lw=2,color=color(2),label=r'$\exp (-\frac{\tau}{T_1^{\rm ph}} )$')
[…]

### 2022-04-29 data/20220429/Adamas 150 um/traitement.py

@@ -1,0 +1,25 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+plt.figure(num=1,figsize=(6,4),dpi=80)
+plt.xticks(fontsize=12)
+plt.yticks(fontsize=12)
+
+polas=[0,20,40,60,80,100]
+for p in polas :
+	x,y=extract_data('scan pola %i deg'%p,ycol=1)
+	# y=y/max(y)
+	x=x*30+4.2
+	plt.plot(x,y,label='%i °'%p)
+	plt.legend()
+
+def find_B():
+	x,y=extract_data('ESR -2V 0 deg')
+	peaks=find_ESR_peaks(x,y)
+	print(peaks)
+	B=find_B_cartesian_mesh(peaks)
+	print(B, B.amp, B.transitions4Classes())
+
+plt.show()

### 2022-05-06 Mes papiers/Low field CR/Low_field_CR.tex
tata yoyo

@@
 \ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins.
-Because of the diamond symmetry, we can distinguish 4 groups ("classes") of identical NV centers, based on the four possible orientation of the NV axis in the crystal.
+Because of the diamond symmetry, we can distinguish 4 groups ("classes") of NV centers, based on the four possible orientation of the NV axis in the crystal.
 These four classes have \textit{a priori} different spin resonance frequency based on the projection of the magnetic field on their respective NV axis.
@@
 
-The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$reads :
+The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
 \begin{equation}
@@
 Calling $\{ \ket{g}, \ket{d}, \ket{e} \}$ the three eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field, Fig.
-\ref{calculs_B_transverse} shows that for small enough magnetic field, the $\{ \ket{g}, \ket{d}, \ket{e} \}$ basis is almost equal to the $\{ \ket{0},\ket{+},\ket{-} \} $ basis : for $B=120\ \rm G$, $\braket{0}{g}\approx 0.99$, $\braket{-1}{d}=1$ and $\braket{+1}{e}\approx 0.99$.
+\ref{calculs_B_transverse} shows that for small enough magnetic field, the $\{ \ket{g}, \ket{d}, \ket{e} \}$ basis is almost equal to the $\{ \ket{0},\ket{+},\ket{-} \} $ basis : for $B=120\ \rm G$, $\braket{0}{g}\approx 0.99$, $\braket{-}{d}=1$ and $\braket{+}{e}\approx 0.99$.
 However the energy difference between $\ket{d}$ and $\ket{e}$ reaches $\approx 45\ \rm G$ which is high enough to cancel out the double-flip processes and allows us to isolate the change of basis hypothesis from the double-flip one.

### 2022-05-06 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
tata yoyo

--- 2022-04-29 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
@@
 
+\section{NV$^-$ ground state Hamiltonian in low magnetic field}
+\begin{figure}
+\includegraphics[width=0.9\textwidth]{Figures_SI/fig_ESR}
+\caption{ODMR measurement (a) in zero magnetic field, (b) in non-zero magnetic field when zooming on a single class transition}
+\label{ESR_single_spin}
+\end{figure}
+With zero external magnetic field, there are three possible sources of splitting of the $\{\ket{+1},\ket{-1}\}$ manifold : local electric field, crystal strain and local magnetic field.Of these three causes, only the electric field can explain the shape of the ODMR line that we observe in zero external magnetic field (Fig.
+\ref{ESR_single_spin}a) \cite{mittiga2018imaging}.
+
+Indeed, a random local magnetic field would produce a single broadened line, and the crystal strain would produce a shifting of the zero field splitting (ZFS) comparable its splitting, which would also blur the two transitions into a single line.
+On the other hand, due to the large difference between the longitudinal and transverse electric field susceptibilities ($d_\parallel = 0.35\ \rm Hz\, cm/V$ and $d_\perp = 17\ \rm Hz\, cm/V$ \cite{van1990electric}), random local electric field will, on average, cause a splitting much stronger than its ZFS shifting and result in a two peak spectrum.
+
+We will therefore neglect the contribution of the strain, local magnetic field and longitudinal electric field,  as well as the hyper-fine structure of the NV center because of the large inhomogeneous broadening of the transitions (Fig.
+\ref{ESR_single_spin}b) and we will consider the following spin Hamiltonian for the NV$^-$ ground state :
+\begin{equation}
+\mathcal{H}_s=D S_z^2 + \gamma_e \vec{B}_{\rm ext} \cdot \vec{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
+\end{equation}
+Where $D=2.87\ \rm GHz$ is the zero field splitting and $\gamma_e=2.8\ \rm MHz/G$ the gyromagnetic ratio of the electron.
+
+In the absence of an external magnetic field, the symmetry in the ($xy$) plane allows us to chose the $x$ direction along the electric field, therefore the eigenstates of $\mathcal{H}_s$ become $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
+
+In the presence of non-zero magnetic field, let us call the Hamiltonian eigenstates $\{ \ket{g},\ket{d}, \ket{e} \} $.
+Fig.
+\ref{map etats propres} shows how close the $\ket{e}$ state is to the $\ket{+1}$ and $\ket{+}$ states as a function of the external magnetic field.
+Looking at the closeness of $\ket{d}$ to $\ket{-1}$ and $\ket{-}$ would show similar results, while $\ket{g}$ is pretty much equals to $\ket{0}$ for $B<100\ \rm G$.
+
+This tells us that in most cases, the $\{ \ket{0},\ket{+},\ket{-} \}$ basis is only the good Hamiltonian basis for magnetic field smaller than a few Gauss, except in the case of pure transverse magnetic field where the $\{ \ket{0},\ket{+},\ket{-} \}$ basis remains a good basis even for sizable magnetic fields.
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{Figures_SI/map_etats_propres}
+\caption{Closeness of the Hamiltonian most excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$ as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
+\label{map etats propres}
+\end{figure}
+
 \section{Samples}
+A faire à la fin pour être sur des figures du main text.
 \section{Experimental Setup}
-\section{Data analysis}
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{Figures_SI/shema_exp}
+\caption{Experimental Setup}
+\label{setup}
+\end{figure}
+Fig. \ref{setup} shows the general purpose experimental setup used for all the experiments presented in this article.
+
+The optical polarization and readout of the spins is done by focusing a green laser on the diamond sample with an objective lens([ref]), and collecting the back-scattered red fluorescence from the NV center on an avalanche photo-diode (APD, REF).
+The laser is filtered out using a dichroic mirror (DM, ref) and a notch filter.
+The NV$^0$ fluorescence is filtered using an additional 645 nm longpass filter.
+
+The laser used here is a [REF], providing pulses of X ns at a rate of 20 MHz, with an average power of $0.5 \sim 5\ \rm mW$.
+The trigger of the pulses is generated externally in order to achieve fast  gating of the laser for the $T_1$ measurement.
+We previously did similar experiment using a continuous 532 nm laser and observed no difference in the behavior of the spins.
+
+The magnetic field is provided by a homemade electromagnet composed of a c-shape iron core and  copper wires.
+The magnet is mounted on two mechanical rotation stages, allowing a control on the polar and azimuthal angle of the magnetic field within a fraction of a degree and is alimented through a low frequency amplifier (REF)
+
+The microwave field is generated by a Rhode \& Shwarz SMB 100A and is emitted with a handmade loop antenna. The field is gated by a [REF minicircuit] switch controlled externally.
+
+A lock-in amplifier is used either to modulate the microwave amplitude for ODMR measurement, or to add an oscillatory magnetic field for the magnetometry protocol.
+In both case we use a modulation frequency $\sim 1\ \rm kHz$ and demodulate the APD signal.
+\section{Experimental details}
 \subsection{$T_1$ fitting Protocol}
-\subsection{Estimation of fluctuators width}
+The $T_1$ profile that we observe with our samples is neither purely exponential, nor purely stretched exponential, which is to be expected when the dipole-dipole spin decay is comparable to the phonon decay.
+
+Fig.
+\ref{T1_fits} shows the result of a lifetime measurement, measured with the subtraction protocol described in the main text, for zero and non-zero magnetic field.
+The result is fitted with exponential and stretched-exponential (with a stretch factor $\beta=0.5$) functions.
+We can see that for the shorter lifetime (0B), the measurement follows more closely the stretched-exponential profile, while the longer lifetime is closer to the exponential one (except at very short times).
+
+In order to directly compare the results from different magnetic fields, we therefore decided to fit each $T_1$ measurement with both a stretched and exponential decay : $S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$.
+In order to add consistency to the measurement of $T_1^{\rm dd}$, we decided to fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ as the only free parameter in the fitting procedure.
+The value chosen for samples [REF] and [REF] was $T_1^{\rm ph}=3.62\ \rm ms$.
+Fig.
+1 of the main text shows the result of fitting the same two measurement with this formula.
+
+An other possibility would be to use an arbitrary stretched factor in the fitting function : $S(\tau)=A \exp (-(\frac{\tau}{T_1^{\rm ph}})^\beta )$.
+Fig.
+\ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field, which confirms that the $T_1$ profile gets closer to a pure stretched-exponential in zero magnetic field.
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_combined}
+\caption{$T_1$ measurement with purely exponential and purely stretched exponential fits (a) in zero magnetic field (b) in non-zero magnetic field}
+\label{T1_fits}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures_SI/fig_alphas}
+\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$ as a function of a randomly oriented magnetic field amplitude}
+\label{alphas}
+\end{figure}
+\subsection{Spectral range of the dipole-dipole cross-relaxations}
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{Figures_SI/largeur_fluct_SI}
+\caption{Dipole-dipole depolarization for two near-resonant classes.
+(a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice.
+(b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes.
+The detuning $\Delta \nu$ between the classes b and c was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_c$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz.
+Single class ESR line stretched by a factor of 1,$\sqrt{2}$ and 2 are added for comparison}
+\label{largeur_fluct}
+\end{figure}
+In the absence of other
+
+An experimental signature of the fluctuator hypothesis developed in \cite{choi_depolarization_2017} is to measure the depolarization rate for two near-resonant classes : if there are indeed very fast decaying NV centers (fluctuators with lifetime $T_1^f < 100 ns$), then the spectral width of the fluctuator would be increased beyond $1/T_2^*$ (which we assume to be homogeneous among all spins in the crystal), which means that they would be able to exchange spin quanta (flip-flop) with non-resonant NV detuned by $\Delta \nu$ such that $2\pi/T_1^f > \Delta \nu > 2\pi/T_2^*$
+
+In order to verify this claim, we have to measure the spectral overlap between two classes, which in the absence of fluctuator should be proportional to the flip-flop rate, and compare it to the actual flip-flop rate which we can measure through the depolarization rate of the spins.
+
+Fig.
+\ref{largeur_fluct} shows the results of such an experiment where we measured the stretched part of the NV's decay rate, which is very well fitted by a Lorentzian of width $\sigma=8.04 MHz$, and compare it to an ODMR line of a single class of NV centers, stretched by a factor of $\sqrt{2}$ and 2 to simulate the spectral overlap of the two classes (discussed bellow).
+One should notice that, not only is $1/T_1^{\rm dd}$ significantly larger than the ODMR profile, it also doesn't have the same shape.
+
+We should not that this broadening can not be explained by the dipole-dipole interaction strength : for a sample with 3 ppm of NV centers, the average dipole-dipole interaction strength between two nearest NV centers $J_0/r^3 \sim 27\ \rm kHz$ which is several order of magnitude lower than the broadening we observe.
+
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{Figures_SI/overlap}
+\caption{Illustration of the spectral overlap for two gaussian spectra}
+\label{overlap}
+\end{figure}
+
+As illustrated on Fig.
+\ref{overlap}, we define the spectral overlap $S(\Delta \nu)$ between two spins of spectral response $S_1(\nu)$ and $S_2(\nu)$, centered respectively on the frequencies $\nu_1$ and $\nu_2$ where $\Delta \nu = \nu_2-\nu_1$ as :
+\begin{equation}
+S(\Delta \nu)=\int S_1(\nu, \nu_1)S_2(\nu, \nu_2) d\nu
+\end{equation}
+In order to approximate the spectral overlap in our experiment, we will consider the analytical solution in the Gaussian and Lorentzian case :
+
+\begin{itemize}
+\item For two gaussians of standard deviation $\sigma$, the spectral overlap as a function of the detuning $\Delta \nu = \nu_1-\nu_2$ is itself a gaussian of standard deviation $\sigma'=\sqrt{2} \sigma$.
+:
+\begin{align*}
+S(\Delta \nu)&\propto \int \exp(-\frac{(\nu-\nu_1)^2}{2\sigma^2})\exp(-\frac{(\nu-\nu_2)^2}{2\sigma^2}) d\nu \\
+&\propto\exp(-\frac{(\Delta \nu)^2}{4\sigma^2})
+\end{align*}
+
+\item For two Lorentzian profile with width $\sigma$, the overlap function is itself a Lorentizan with width $\sigma'=2\sigma$ :
+\begin{align*}
+S(\Delta \nu)&\propto \int \frac{1}{1+ \frac{(\nu-\nu_1)^2}{\sigma^2}}\cdot \frac{1}{1+ \frac{(\nu-\nu_2)^2}{\sigma^2}} d\nu \\
+&\propto\frac{1}{1+ \frac{(\Delta \nu)^2}{4\sigma^2}}
+\end{align*}
+\end{itemize}
+
+The ODMR lines that we measure are neither Lorentzian nor Gaussian (although they tend to be closer to Gaussians), and can even be asymmetric.
+Nevertheless, the overlap between two classes can most likely be approximated by a single class ODMR profile stretched by a factor between $\sqrt{2}$ and 2.
+
 \section{Les détails qui tuent}
 \subsection{Effect of laser polarization}
+\begin{figure}
+\includegraphics[width=0.9\textwidth]{Figures_SI/fig_Pola}
+\caption{Effect of the polarization of the incident laser.
+(a) Photoluminescence as a function of randomly oriented magnetic field amplitude for various polarization angle.
+(b) Demodulated PL in the same conditions}
+\label{Pola}
+\end{figure}
 \subsection{Alignment of B}
 La faudrait p-e montrer ce qu'il se passe pour un décalage de quelques degrés. A voir si j'ai déja les plots
-\section{Single spin Hamiltonian}
 \section{Fluctuator et tout}
+\subsection{121 VS 22}
+\subsection{100 vs 0B}
 

### 2022-05-06 Mes papiers/Low field CR/SI/Figures_SI/overlap.py

@@ -1,0 +1,48 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+plt.figure(num=1,figsize=(6,4),dpi=80)
+plt.xticks(fontsize=12)
+plt.yticks(fontsize=12)
+
+
+def gaussian(x,sigma,x0):
+	return np.exp(-(x-x0)**2/(2*sigma**2))
+
+def lorentzian(x,sigma,x0):
+	return 1/(1+(x-x0)**2/sigma**2)
+
+def plot_overlap():
+	x=np.linspace(2650,2750,500)
+	y1=gaussian(x,6,2692)
+	y2=gaussian(x,6,2708)
+	y3=y1*y2
+	plt.plot(x,y1,label='NV1 spectral response')
+	plt.plot(x,y2,label='NV2 spectral response')
+	l3=plt.plot(x,y3,label='spectral overlap')
+	plt.fill_between(x,0,y3,color=color(2))
+	plt.legend()
+
+
+def compute_overlap():
+	n=500
+	x=np.linspace(-10,10,n)
+	deltas=np.linspace(-10,10,n)
+	ol=np.zeros(n)
+	for i in range(n):
+		delta=deltas[i]
+		y1=lorentzian(x,1,-delta/2)
+		y2=lorentzian(x,1,+delta/2)
+		y3=y1*y2
+		ol[i]=integration(x,y3)
+
+	ol=ol/max(ol)
+	plt.plot(x,ol)
+	popt,yfit=lor_fit(x,ol)
+	plt.plot(x,yfit)
+	print(popt)
+compute_overlap()
+
+plt.show()

### 2022-05-12 Mes papiers/Low field CR/Low_field_CR.tex
tata

@@
 \end{figure}
+%
 DC microwave-less magnetometry has already been performed with NV ensembles, using either cross-relaxations [les russes] or level anti-crossing [bubu].
@@
 The main difference with the previously mentioned protocols is the fact that this one doesn't rely on crystalline orientation, making this protocol usable with diamond powders or polycrystalline samples.
+The idea of performing optical magnetometry on the zero-field photoluminescence feature of NV ensembles has been suggested before \cite{filimonenko2018weak, filimonenko2022manifestation} but had not been implemented so far.
 

### 2022-05-12 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
tata

@@
 \end{figure}
-With zero external magnetic field, there are three possible sources of splitting of the $\{\ket{+1},\ket{-1}\}$ manifold : local electric field, crystal strain and local magnetic field.Of these three causes, only the electric field can explain the shape of the ODMR line that we observe in zero external magnetic field (Fig.
+With zero external magnetic field, there are three possible sources of splitting of the $\{\ket{+1},\ket{-1}\}$ manifold : local electric field, crystal strain and local magnetic field.
+Of these three causes, only the electric field can explain the shape of the ODMR line that we observe in zero external magnetic field (Fig.
 \ref{ESR_single_spin}a) \cite{mittiga2018imaging}.
 
-Indeed, a random local magnetic field would produce a single broadened line, and the crystal strain would produce a shifting of the zero field splitting (ZFS) comparable its splitting, which would also blur the two transitions into a single line.
+Indeed, random local magnetic field would produce a single broadened line, and crystal strain would produce a shifting of the zero field splitting (ZFS) comparable its splitting, which would also blur the two transitions into a single line.
 On the other hand, due to the large difference between the longitudinal and transverse electric field susceptibilities ($d_\parallel = 0.35\ \rm Hz\, cm/V$ and $d_\perp = 17\ \rm Hz\, cm/V$ \cite{van1990electric}), random local electric field will, on average, cause a splitting much stronger than its ZFS shifting and result in a two peak spectrum.
 
-We will therefore neglect the contribution of the strain, local magnetic field and longitudinal electric field,  as well as the hyper-fine structure of the NV center because of the large inhomogeneous broadening of the transitions (Fig.
-\ref{ESR_single_spin}b) and we will consider the following spin Hamiltonian for the NV$^-$ ground state :
+In our low magnetic field dipole-dipole coupling model, we will therefore neglect the contribution of the strain, local magnetic field and longitudinal electric field.
+We also don't take into account the hyper-fine structure of the NV center due of the large inhomogeneous broadening of the transitions (Fig.
+\ref{ESR_single_spin}b).
+We will then consider the following spin Hamiltonian for the NV$^-$ ground state : 
 \begin{equation}
@@
 
-In the absence of an external magnetic field, the symmetry in the ($xy$) plane allows us to chose the $x$ direction along the electric field, therefore the eigenstates of $\mathcal{H}_s$ become $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
+In the absence of an external magnetic field, the symmetry in the ($xy$) plane allows us to chose the $x$ direction along the electric field.
+The eigenstates of $\mathcal{H}_s$ then become $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
 
@@
 \includegraphics[width=0.7\textwidth]{Figures_SI/map_etats_propres}
-\caption{Closeness of the Hamiltonian most excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$ as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
+\caption{Numerical simulations of the closeness of the Hamiltonian most excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$, as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
 \label{map etats propres}
@@
 The result is fitted with exponential and stretched-exponential (with a stretch factor $\beta=0.5$) functions.
-We can see that for the shorter lifetime (0B), the measurement follows more closely the stretched-exponential profile, while the longer lifetime is closer to the exponential one (except at very short times).
+We can see that for the shorter lifetime (zero field), the measurement follows more closely the stretched-exponential profile, while the longer lifetime is closer to the exponential one (except at very short times).
 
 In order to directly compare the results from different magnetic fields, we therefore decided to fit each $T_1$ measurement with both a stretched and exponential decay : $S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$.
-In order to add consistency to the measurement of $T_1^{\rm dd}$, we decided to fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ as the only free parameter in the fitting procedure.
+To add consistency to the measurement of $T_1^{\rm dd}$, we decided to fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ and the signal amplitude as the only free parameters in the fitting procedure.
 The value chosen for samples [REF] and [REF] was $T_1^{\rm ph}=3.62\ \rm ms$.
@@
 
-We should not that this broadening can not be explained by the dipole-dipole interaction strength : for a sample with 3 ppm of NV centers, the average dipole-dipole interaction strength between two nearest NV centers $J_0/r^3 \sim 27\ \rm kHz$ which is several order of magnitude lower than the broadening we observe.
+We should note that this broadening can not be explained by the dipole-dipole interaction strength : for a sample with 3 ppm of NV centers, the average dipole-dipole interaction strength between two nearest NV centers $J_0/r^3 \sim 27\ \rm kHz$ which is several order of magnitude lower than the broadening we observe.
 
@@
 \subsection{Effect of laser polarization}
+Previous studies \cite{anishchik2015low, filimonenko2020weak} reported the presence of a photoluminescence dip in zero magnetic field which was heavily dependent on the laser polarization angle with respect to the diamond axes and the magnetic field.
+
+We did not observe a strong dependence on the laser polarization with the samples used in this study.
+Fig.\ref{Pola} shows the photoluminescence from one of our sample as a function of the magnetic field, either directly or with a modulation of the magnetic field, for 6 randomly chosen polarization angle and saw no major difference.
+
+In particular, we did not observe the apparition of an anti-line inside the main dip, unlike what was observed in the two previously cited work.
+We expect that the main reason behind the different behaviors is that we seem to observe far grater spin depolarization in zero-field, which we attribute to dipole-dipole coupling, and that this effect could hide smaller effects such as the one related to the laser polarization.
 \begin{figure}
@@
 La faudrait p-e montrer ce qu'il se passe pour un décalage de quelques degrés. A voir si j'ai déja les plots
-\section{Fluctuator et tout}
+\section{Extension of the fluctuator model to low magnetic fields}
+\subsection{Summary of the NV-fluctuator model}
+As mentioned in the main text, we base our analysis on the NV-fluctuator model developed in \cite{choi_depolarization_2017} whose main hypothesis and conclusions we will summarize here :
+
+The NV$^-$ centers in the crystal are divided between two categories : "normal" NV centers (simply called NV) which, in the absence of dipole-dipole coupling would have a phonon-limited $T_1$ ($T_1^{NV} \sim\ \rm ms$), and fluctuators who are NV centers with an additional, fast, depolarization mechanism, such that their lifetime $T_1^f < 100\ \rm ns$.
+Having such a short lifetime, the fluctuators are pretty much unpolarized by the green laser, making them invisible in standard optical NV measurement protocol ($T_1$, ODMR, etc.).
+
+Assuming a homogeneous distribution of the fluctuators in the bulk of the crystal, the authors conclude that the fluctuators create an additionnal decay channel for the NV population, through dipole-dipole interaction, characterized by a decay rate $\gamma$ which follows the probability distribution :
+\begin{equation}
+\rho(\gamma)=\frac{e^{-1/(4\gamma T)}}{\sqrt{4\pi \gamma^3 T}}
+\end{equation}
+Where the timescale $T$ is defined as
+\begin{equation}
+\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f}
+\end{equation}
+Where $n_f$ is the fluctuator density in the crystal in nm$^{-3}$, $J_0=52\ \rm{MHz}\cdot\rm{nm}^3$ is the characteristic dipole-dipole strength between two spins, $\gamma_f$ is the fluctuator intrinsic decay rate and $\bar \eta$ is a dimension-less number which characterizes the average dipole-dipole interaction between the NV centers and the fluctuators (resonance conditions, relative orientations etc.
+Further details are given below).
+
+The polarization dynamics of the averaged ensemble of NV centers then follows :
+\begin{equation}
+P(t)=\int_0^\infty \rho(\gamma)e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}
+\end{equation}
+\subsection{Dipole-Dipole Hamiltonian between two NV$^-$ centers}
 \subsection{121 VS 22}

### 2022-05-13 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
taratata

@@
 
-Indeed, random local magnetic field would produce a single broadened line, and crystal strain would produce a shifting of the zero field splitting (ZFS) comparable its splitting, which would also blur the two transitions into a single line.
+Indeed, random local magnetic field would produce a single broadened line while crystal strain would produce a shifting of the zero field splitting (ZFS) of the same order as the splitting of the levels, which would also blur the two transitions into a single line.
 On the other hand, due to the large difference between the longitudinal and transverse electric field susceptibilities ($d_\parallel = 0.35\ \rm Hz\, cm/V$ and $d_\perp = 17\ \rm Hz\, cm/V$ \cite{van1990electric}), random local electric field will, on average, cause a splitting much stronger than its ZFS shifting and result in a two peak spectrum.
@@
 In our low magnetic field dipole-dipole coupling model, we will therefore neglect the contribution of the strain, local magnetic field and longitudinal electric field.
-We also don't take into account the hyper-fine structure of the NV center due of the large inhomogeneous broadening of the transitions (Fig.
+We also don't take into account the hyper-fine structure of the NV center due to the large inhomogeneous broadening of the transitions (Fig.
 \ref{ESR_single_spin}b).
@@
 \begin{equation}
+\label{NV Hamiltonian}
 \mathcal{H}_s=D S_z^2 + \gamma_e \vec{B}_{\rm ext} \cdot \vec{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
@@
 
-\section{Les détails qui tuent}
 \subsection{Effect of laser polarization}
@@
 
-The NV$^-$ centers in the crystal are divided between two categories : "normal" NV centers (simply called NV) which, in the absence of dipole-dipole coupling would have a phonon-limited $T_1$ ($T_1^{NV} \sim\ \rm ms$), and fluctuators who are NV centers with an additional, fast, depolarization mechanism, such that their lifetime $T_1^f < 100\ \rm ns$.
-Having such a short lifetime, the fluctuators are pretty much unpolarized by the green laser, making them invisible in standard optical NV measurement protocol ($T_1$, ODMR, etc.).
+The NV$^-$ centers in the crystal are divided between two categories : "normal" NV centers (simply called NV) which, in the absence of dipole-dipole coupling would have a phonon-limited $T_1$ ($T_1^{NV}\sim \rm ms$), and fluctuators who are NV centers with an additional, fast, depolarization mechanism, such that their lifetime $T_1^f < 100\ \rm ns$.
+Having such a short lifetime, the fluctuators are almost unpolarized by the green laser, making them invisible in standard optical NV measurement protocol ($T_1$, ODMR, etc.).
 
@@
 \begin{equation}
-P(t)=\int_0^\infty \rho(\gamma)e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}
+P(t)=\int_0^\infty \rho(\gamma)\, e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}
 \end{equation}
+Which corresponds to the stretched-exponential part of the lifetime measurement.
 \subsection{Dipole-Dipole Hamiltonian between two NV$^-$ centers}
+We consider a system composed of an NV center interacting dipolarly with a fluctuator.
+\begin{equation}
+\mathcal{H}_{\rm tot}=\mathcal{H}_{1}+\mathcal{H}_2+\mathcal{H}_{\rm dd}
+\end{equation}
+Where $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ are the single particle Hamiltonian of the NV and fluctuator, described by the equation (\ref{NV Hamiltonian}), and $\mathcal{H}_{\rm dd}$ is the dipole-dipole interaction Hamiltonian between the two spins.
+
+We will consider the spin operators $\vec{S}_1$ and
+$\vec{S}_2$ in the NV own basis : the $z$ orientation of the $S_z$ operator is chosen along the NV axis of each spin.
+In the presence of a longitudinal magnetic field, the sense of $\hat{z}$ will be chosen such that $\vec{B}\cdot \hat{z} >0$, and the $\hat{x}$ direction is chosen along the transverse magnetic or electric field, whichever dominates the single spin Hamiltonian.
+These choices mean that their will be two distinct Cartesian basis $\{\hat{x}_1,\hat{y}_1,\hat{z}_1\}$ and $\{\hat{x}_2,\hat{y}_2,\hat{z}_2\}$ describing the two spins.
+
+The dipole-dipole Hamiltonian can then be decomposed as :
+\begin{align}
+\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}&= 3\left({\vec S}_1 \cdot \hat{u} \right)\left({\vec S}_2 \cdot \hat{u} \right) - {\vec S}_1 \cdot {\vec S}_2  \\
+&=\left[3(\hat{u}\cdot\hat{x}_1)(\hat{u}\cdot\hat{x}_2) -\hat{x}_1\cdot\hat{x}_2\right] S_x^1S_x^2 \\
+&+\left[3(\hat{u}\cdot\hat{y}_1)(\hat{u}\cdot\hat{y}_2) -\hat{y}_1\cdot\hat{y}_2\right] S_y^1S_y^2 \\
+&+\left[3(\hat{u}\cdot\hat{x}_1)(\hat{u}\cdot\hat{y}_2) -\hat{x}_1\cdot\hat{y}_2\right] S_x^1S_y^2 \\
+&+\left[3(\hat{u}\cdot\hat{y}_1)(\hat{u}\cdot\hat{x}_2) -\hat{y}_1\cdot\hat{x}_2\right] S_y^1S_x^2 \\
+&+\left[3(\hat{u}\cdot\hat{z}_1)(\hat{u}\cdot\hat{z}_2) -\hat{z}_1\cdot\hat{z}_2\right] S_z^1S_z^2 \\
+&+\mathcal{H}_{\rm other}
+\end{align}
+Where $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
+
+\subsection{Flip-flops in the magnetic basis $\{ \ket{0},\ket{+1},\ket{-1} \} $}
+We will first consider the case, treated in \cite{choi_depolarization_2017}, where the longitudinal magnetic field is strong enough that the single spin Hamiltonian eigenstates of both spins are close to $\{ \ket{0},\ket{+1},\ket{-1} \} $ (since we are interested in near-resonant spins, this means that both spins see roughly the same longitudinal and transverse magnetic field).
+
+In this scenario, only flip-flop terms (that is $\mel{\pm 1,0}{\mathcal{H}_{\rm dd}}{0,\pm 1}$) can couple two resonant two-spins states.
+Following the notation in \cite{choi_depolarization_2017}, we introduce the dimensionless factor $\eta$ defined as :
+
 \subsection{121 VS 22}

### 2022-05-16 data/20211015/Le bon adamas/traitement.py
tata

@@ -1,7 +1,12 @@
 import sys
 sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
+scale=1.3
+plt.figure(num=1,figsize=(3*scale,2*scale),dpi=80)
+plt.xticks(fontsize=13)
+plt.yticks(fontsize=13)
 
 # fname='T1 NV-C13 100 3 V'
 # x,y=extract_data(fname,xcol=4,ycol=5)
@@ -14,11 +19,17 @@ from analyse import *
 
 
 # fname='Vrai ESR 111'
-fname='ESR 111 et C13'
-cs=[2980.375, 2817.625, 3115.125, 2629.5]
+# fname='ESR 111 et C13'
+# cs=[2980.375, 2817.625, 3115.125, 2629.5]
+# x,y=extract_data(fname)
+# plt.plot(x,y)
+# popt,yfit=ESR_n_pics_auto(x,y)
+# plt.plot(x,yfit)
+# print(popt)
+# plt.show()
+
+fname='Vrai ESR 100'
 x,y=extract_data(fname)
 plt.plot(x,y)
-popt,yfit=ESR_n_pics_auto(x,y)
-plt.plot(x,yfit)
-print(popt)
+
 plt.show()

### 2022-05-18 Talks/Journées de l'axe/presentation.tex
tata

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage{graphicx}
\usepackage{physics}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage[backend=biber]{biblatex}

\addbibresource{Biblio_dbl_quantum.bib}

%\bibliographystyle{stylename}
%\bibliography{Biblio_dbl_quantum}

\title{Dipolar interactions in dense ensembles of Nitrogen-Vacancy centers}
\author{Clément Pellet-Mary, Maxime Perdriat, Gabriel Hétet}
\date{Nano-optics group}

\mode<presentation> {\usetheme{Rochester}}

\begin{document}
\begin{frame}
\maketitle
\centering
\includegraphics[scale=.2]{Nitrogen-vacancy_center}
\end{frame}
%\begin{frame}{Outline}
%\tableofcontents
%\end{frame}
%\section{Physics of the NV center}
\begin{frame}{Team main activities}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide activités}
\end{frame}

\begin{frame}{Optical properties of NV$^-$ centers}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_NV_optical}
\end{frame}
%\begin{frame}{NV$^-$ center electronic structure}
%\centering
%\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{NV_8_niveaux_1}
%\end{frame}
%\begin{frame}{NV$^-$ center electronic structure}
%\centering
%\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{NV_8_niveaux_2}
%\end{frame}
\begin{frame}{NV$^-$ center electronic structure}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{NV_8_niveaux_3}
\end{frame}
\begin{frame}{NV center spin sub-levels}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_3_niveaux}
\end{frame}
\begin{frame}{Optically Detected Magnetic Resonance (ODMR)}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR}
\end{frame}
\begin{frame}{Modification of the spin $T_1$ with dense ensemble}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_1}
\end{frame}
\begin{frame}{Modification of the spin $T_1$ due to resonant dipole coupling}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_2}
\end{frame}
\begin{frame}{Dipole-dipole coupling in zero magnetic field}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide zero field}
\end{frame}
\begin{frame}{Conclusion}
\begin{itemize}
\item The NV$^-$ center is an optically active defect in diamond which allows an optical control and readout of its spin state.
\bigskip
\item The depolarization of the spins is modified in dense ensemble due to dipole-dipole coupling.
\bigskip
\item This effect is even stronger in zero-magnetic field.
\end{itemize}
\end{frame}
\begin{frame}{Bonus : Magnetometry in zero magnetic field}
\centering
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_magneto}
\end{frame}

\end{document}

### 2022-05-19 Talks/Journées de l'axe/presentation.tex
tata

@@
 %\section{Physics of the NV center}
-\begin{frame}{Team main activities}
+\begin{frame}{What are NV centers ?}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide activités}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide  NV properties}
+\end{frame}
+\begin{frame}{Common example : Optically Detected Magnetic Resonance (ODMR)}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR}
 \end{frame}
 
+\begin{frame}{What are NV centers and why are they everywhere ?}
+\begin{itemize}
+\item Good optical properties : quantum yield $\approx 1$, stable in time and temperature
+\medskip
+\item Good spin properties : $T_1\approx\ \rm ms$ and $T_2^*\sim \mu$s at 300 K.
+\medskip
+\item Optical polarization (up to 90\%) and readout (fidelity up to 0.3) of the spin at room temperature
+\end{itemize}
+\bigskip
+$\to$ One of the most versatile spin qubit at room temperature
+\end{frame}
+
+\begin{frame}{What are NV centers being used for ?}
+$\bullet$ Quantum memories :
+\begin{center}
+\includegraphics[width=0.8\textwidth,height=0.9\textheight,keepaspectratio]{mémoire}
+\end{center}
+
+$\bullet$ Intrication :
+\begin{center}
+\includegraphics[width=0.8\textwidth,height=0.9\textheight,keepaspectratio]{Intrication}
+\end{center}
+\end{frame}
+
+\begin{frame}{Magnetomtry with NV centers : big or small}
+$\bullet$ AFM nano-scale magnetomtry :
+\begin{center}
+\includegraphics[width=0.65\textwidth,height=0.9\textheight,keepaspectratio]{magneto AFM}
+\end{center}
+
+$\bullet$ Diamond magnetic microscopy :
+\begin{center}
+\includegraphics[width=0.65\textwidth,height=0.9\textheight,keepaspectratio]{magneto graphene}
+\end{center}
+\end{frame}
+
+\begin{frame}{Context of my PhD work : new physics with dense ensemble of NV centers}
+\begin{itemize}
+\item To increase the ensemble sensitivity we need to increase the density of NV centers
+\item Chemists and material scientists have made huge progress to grow NV-rich diamond with little other impurities 
+\begin{center}
+\includegraphics[width=0.65\textwidth,height=0.9\textheight,keepaspectratio]{diamant Alex}
+\end{center}
+\item When the defects density reaches a critical point, many body effects start to appear due to dipole-dipole interaction and charge tunneling (3 ppm of NV = 30 kHz dipole coupling between NV neighbors)
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Modification of the spin $T_1$ due to resonant dipole coupling}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_2}
+\end{frame}
+\begin{frame}{Conclusion}
+Acknowledgments : Christine et les autres
+
+\bigskip
+Take home messages : 
+\begin{itemize}
+\item NV centers are defects in diamond with an optically controllable and readable spin at room temperature
+\medskip
+\item Many body effects start to manifest with the new NV-rich samples being made
+\end{itemize}
+\end{frame}
+\begin{frame}{Bonus : Magnetometry in zero magnetic field}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_magneto}
+\end{frame}
 \begin{frame}{Optical properties of NV$^-$ centers}
@@
 \end{frame}
-%\begin{frame}{NV$^-$ center electronic structure}
-%\centering
-%\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{NV_8_niveaux_1}
-%\end{frame}
-%\begin{frame}{NV$^-$ center electronic structure}
-%\centering
-%\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{NV_8_niveaux_2}
-%\end{frame}
 \begin{frame}{NV$^-$ center electronic structure}
@@
 \end{frame}
-\begin{frame}{Optically Detected Magnetic Resonance (ODMR)}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR}
-\end{frame}
-\begin{frame}{Modification of the spin $T_1$ with dense ensemble}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_1}
-\end{frame}
-\begin{frame}{Modification of the spin $T_1$ due to resonant dipole coupling}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_2}
-\end{frame}
-\begin{frame}{Dipole-dipole coupling in zero magnetic field}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide zero field}
-\end{frame}
-\begin{frame}{Conclusion}
-\begin{itemize}
-\item The NV$^-$ center is an optically active defect in diamond which allows an optical control and readout of its spin state.
-\bigskip
-\item The depolarization of the spins is modified in dense ensemble due to dipole-dipole coupling.
-\bigskip
-\item This effect is even stronger in zero-magnetic field.
-\end{itemize}
-\end{frame}
-\begin{frame}{Bonus : Magnetometry in zero magnetic field}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_magneto}
-\end{frame}
-
 \end{document}

### 2022-05-20 Talks/Journées de l'axe/presentation.tex
toto

@@
 \centering
-\includegraphics[scale=.2]{Nitrogen-vacancy_center}
+Mettre logos ici
 \end{frame}
@@
 %\section{Physics of the NV center}
-\begin{frame}{What are NV centers ?}
+\begin{frame}{What are NV centers and why are they everywhere?}
 \centering
 \includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide  NV properties}
-\end{frame}
-\begin{frame}{Common example : Optically Detected Magnetic Resonance (ODMR)}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR}
 \end{frame}
@@
 \end{center}
-\item When the defects density reaches a critical point, many body effects start to appear due to dipole-dipole interaction and charge tunneling (3 ppm of NV = 30 kHz dipole coupling between NV neighbors)
+\item When the defects density reaches a critical point, many body effects start to appear due to dipole-dipole interaction and charge transfer (3 ppm of NV = 30 kHz dipole coupling between NV neighbors)
 \end{itemize}
+\end{frame}
+\begin{frame}{Basic example : Optically Detected Magnetic Resonance (ODMR)}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR}
 \end{frame}
@@
 \end{frame}
+
+\begin{frame}{Other results with dense ensemble of spins}
+$\bullet$ Optical detection of dark spins :
+\begin{center}
+\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{dark spins}
+\end{center}
+$\bullet$ Mechanical detection of resonant dipolar coupling :
+\begin{center}
+\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{Cr-meca}
+\end{center}
+\end{frame}
+
 \begin{frame}{Conclusion}
-Acknowledgments : Christine et les autres
+\textbf{Acknowledgments}
+
+The team : Gabriel, Maxime, André, the former members : Paul, Louis Tom, the nano-optics group and the technical staff : Christine, Pascal, Arnaud...
 
 \bigskip
-Take home messages : 
+\textbf{Take home messages }
 \begin{itemize}

### 2022-05-27 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
toto

@@
 
-Assuming a homogeneous distribution of the fluctuators in the bulk of the crystal, the authors conclude that the fluctuators create an additionnal decay channel for the NV population, through dipole-dipole interaction, characterized by a decay rate $\gamma$ which follows the probability distribution :
+Assuming a homogeneous distribution of the fluctuators in the bulk of the crystal, the authors of \cite{choi_depolarization_2017} conclude that the fluctuators create an additionnal decay channel for the NV population, through dipole-dipole interaction, characterized by a decay rate $\gamma$ which follows the probability distribution :
 \begin{equation}

### 2022-05-31 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
t

@@
 \section{NV$^-$ ground state Hamiltonian in low magnetic field}
+\label{sec Hamiltonian}
 \begin{figure}
@@
 \includegraphics[width=0.7\textwidth]{Figures_SI/map_etats_propres}
-\caption{Numerical simulations of the closeness of the Hamiltonian most excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$, as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
+\caption{Numerical simulations of the closeness of the Hamiltonian most excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$, as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.
+A value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen}
 \label{map etats propres}
@@
 \frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f}
+\label{eq 1/T}
 \end{equation}
@@
 In the presence of a longitudinal magnetic field, the sense of $\hat{z}$ will be chosen such that $\vec{B}\cdot \hat{z} >0$, and the $\hat{x}$ direction is chosen along the transverse magnetic or electric field, whichever dominates the single spin Hamiltonian.
-These choices mean that their will be two distinct Cartesian basis $\{\hat{x}_1,\hat{y}_1,\hat{z}_1\}$ and $\{\hat{x}_2,\hat{y}_2,\hat{z}_2\}$ describing the two spins.
+These choices mean that there will be two distinct Cartesian basis $\{\hat{x}_1,\hat{y}_1,\hat{z}_1\}$ and $\{\hat{x}_2,\hat{y}_2,\hat{z}_2\}$ describing the two spins.
 
@@
 \begin{align}
-\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}&= 3\left({\vec S}_1 \cdot \hat{u} \right)\left({\vec S}_2 \cdot \hat{u} \right) - {\vec S}_1 \cdot {\vec S}_2  \\
+-\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}&= 3\left({\vec S}_1 \cdot \hat{u} \right)\left({\vec S}_2 \cdot \hat{u} \right) - {\vec S}_1 \cdot {\vec S}_2  \\
 &=\left[3(\hat{u}\cdot\hat{x}_1)(\hat{u}\cdot\hat{x}_2) -\hat{x}_1\cdot\hat{x}_2\right] S_x^1S_x^2 \\
@@
 \end{align}
-Where $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
+Where $J_0=\frac{\mu_0 \gamma_e^2 \hbar^2}{4 \pi} = (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r= r\vec u$ is the distance between the two spins and  $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
 
@@
 Following the notation in \cite{choi_depolarization_2017}, we introduce the dimensionless factor $\eta$ defined as :
+\begin{equation}
+\eta^2=\frac{1}{3} \abs{\mel{\pm 1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,\pm 1}}^2  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}
+\end{equation}
+Where $\gamma_f$ is the fluctuator lifetime, $\omega_f$ the fluctuator angular frequency and $\omega_{NV}$ the NV angular frequency.
+%This numerical factor characterizes both the dipole-dipole strength between the NV and the fluctuator through the matrix element and the resonance condition through the Lorentzian function.
 
-\subsection{121 VS 22}
-\subsection{100 vs 0B}
+This numerical factor is linked to the previously mentioned $\bar \eta$ by averaging over every possible orientation and angular position of the fluctuators.
+Assuming that the fluctuators are evenly distributed among all four classes (NV axis orientation) and decomposing the $\vec r$ vector in the spherical basis ($r, \theta, \phi$), we can write $\bar \eta$ as :
+
+\begin{equation}
+\label{eq. eta bar}
+\bar \eta =\frac{1}{4} \sum_{i=1}^4 \int_\theta \int_\phi  \int_{\omega_f} \int_{\omega_{NV}}   \abs{\eta(\vec u,i,\omega_{NV},\omega_f)}\, d\Omega\, \rho(\omega_{NV}) d\omega_{NV}\, \rho^i(\omega_f) d\omega_f
+\end{equation}
+
+Where $d\Omega=\sin \theta d\theta d\phi$, $\rho(\omega_{NV})$ is the distribution of angular frequencies of the probed NV centers, and  $\rho^i(\omega_f)$ is the distribution of angular frequencies  for the fluctuators of class $i$.
+
+While the distributions $\rho(\omega_{NV})$ and $\rho^i(\omega_f)$ can be approximated using the data presented in Fig.
+\ref{largeur_fluct}, we will assume that these distributions are independent of the external magnetic field, and since we are interested in comparing the same sample for various values of the magnetic field, we will simplify eq.(\ref{eq.
+eta bar}) by postulating that $\frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} =1$ if the NV center's and the fluctuator's classes are resonant (or if they are from the same class), and $\frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2 +4\gamma_f^2}=0$ otherwise.
+
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{Figures_SI/121 VS 22}
+\caption{(a) Geometrical representation of two NV classes and the two possible planes where the magnetic field has the same projection on both classes.
+The positive $\hat z_i$ direction is chosen so that $\vec{B}\cdot \hat{z}_i >0$.
+(b) $T_1$ measurement on a two-class resonance following the protocol described in main text.
+8 different magnetic field were used on the same sample : 4 times in a \{100\} plane (blue unfilled symbols) and 4 times in a \{110\} plane (orange filled symbols).
+Inset : zoom-in on the 0.3-0.8 ms region.}
+\label{121 VS 22 fig}
+\end{figure}
+
+With these hypothesis, we found that there were 3 possible scenarios where the NV and fluctuator were resonant:
+\begin{itemize}
+\item The NV and fluctuator are resonant and from the same class (the angle $\widehat{z_1 z_2}=0$) : this is always true, regardless of the magnetic field.
+We can analytically compute $\bar \eta$ in this case and find : $$ \bar {\eta}_{same}=\frac{1}{4}  \int_\theta \int_\phi \abs{\eta(\vec u,\widehat{z_1 z_2}=0,\omega_{NV}=\omega_f)}\, d\Omega= \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot \frac{2}{3 \sqrt{3}} \approx 5.55\cdot10^{-2}$$
+\item The NV and fluctuator are resonant and the angle $\widehat{z_1 z_2}=\rm{arccos}(\frac{1}{3})\approx 70.5^\circ$.
+This happens when the magnetic field lies in the \{100\} crystalline planes family.
+In this case we can numerically compute $\bar \eta_{\rm close} \approx \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot 0.6507 \approx 9.39\cdot10^{-2}$
+\item The NV and fluctuator are resonant and the angle $\widehat{z_1 z_2}=\rm{arccos}(-\frac{1}{3})\approx 109.5^\circ$.
+This happens when the magnetic field lies in the \{110\} or \{$1\bar{1}0$\} crystalline planes family.
+In this case we can numerically compute $\bar \eta_{\rm far} \approx \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot 0.8328 \approx 1.20\cdot10^{-1}$.
+This last case was not taken into account by \cite{choi_depolarization_2017}.
+\end{itemize}
+
+Fig.
+\ref{121 VS 22 fig}(a) shows a graphical representation of the difference between the two last scenarios : due to the $\vec{B}\cdot \hat{z} >0$ condition, the same two classes of NV centers can have a $\widehat{z_1 z_2}$ angle equal to $70.5^\circ$ or $109.5^\circ$ depending on the external magnetic field.
+
+These predictions can be tested experimentally.
+Fig.
+\ref{121 VS 22 fig}(b) shows the subtracted $T_1$ signal measured on the same sample [REF] for 8 different values of the magnetic field, each time on a two-class resonance : in 4 cases the magnetic field was in a \{100\} plane, and in the 4 other cases it was in a \{110\} plane.
+We can see that the measured lifetime is always smaller on the \{110\} case, which corresponds the the greater $\bar \eta$ factor computed previously.
+
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{Figures_SI/Various ESR}
+\caption{ODMR spectra for different orientations of the magnetic field. Red numbers represent the number of degenerated classes for each line of the spectrum}
+\label{Various ODMR}
+\end{figure}
+
+Here is a list of the computed values of $\bar \eta^2$, which according to eq.(\ref{eq 1/T}) is proportional to the dipole induced decay rate, for various magnetic field orientation.
+An ODMR spectrum for each of these situations is present in Fig.
+\ref{Various ODMR}.
+\begin{itemize}
+\item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
+\item $\vec{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.
+For the two resonant classes : $\bar \eta = \bar {\eta}_{same} + \bar {\eta}_{far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
+\item $\vec{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{same}+\bar {\eta}_{close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
+\item $\vec{B} \parallel \langle 111 \rangle$ : For the non-resonant class (parallel to $\vec B$) : $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.
+For the triply resonant classes: $\bar \eta = \bar {\eta}_{same}+2\bar {\eta}_{far}$ and $\bar \eta^2\approx 28.4\, \bar \eta_0^2$.
+\item $\vec{B} \parallel \langle 100 \rangle$ : For the quadruple resonance : $\bar \eta = \bar {\eta}_{same}+2\bar {\eta}_{far}$ and $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
+\end{itemize}
+
+The measured increase in the decay rate is generally smaller than the predicted one : we measured an increase by a factor $\sim 4$ for a two class degeneracy instead of the predicted $10.0$, and an increase by a factor $\sim 16$ for a two class degeneracy instead of the predicted $42.8$.
+\cite{choi_depolarization_2017} measured an increase by a factor $\sim 4$ on a two-class degeneracy instead of the predicted $7.24$ or $10.0$.
+
+\subsection{Flip-flops in the non-magnetic basis $\{ \ket{0},\ket{+},\ket{-} \} $}
+
+As described in sec.
+\ref{sec Hamiltonian}, when the single spin Hamiltonian of the NV center is dominated by the electric field or the transverse magnetic field (as long as $\gamma_e B_\perp \ll D$), the eigen basis of the NV Hamiltonian is close to $\{\ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
+We therefore have to rewrite the dipole-dipole Hamiltonian using this new basis for both spins in order to perform similar calculations as the ones done in the previous section.
+
+In this new basis, the spin operators are written :
 

### 2022-06-02 Mes papiers/Low field CR/Low_field_CR.tex
tata yoyo

@@
 
-\title{Low field depolarization of electronic spins through dipole-dipole coupling}
+\title{Low field depolarization of NV$^-$ centers through dipole-dipole coupling}
 
-\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, G. H\'etet}
+\author{C. Pellet-Mary$^1$, P. Huillery$^?$,  M. Perdriat$^1$, G. H\'etet}
 
@@
 
-The ensemble spin lifetime is non-exponential due to inhomogeneities between the NV$^-$ centers.
+The ensemble spin decay is non-exponential due to inhomogeneities between the NV$^-$ centers.
 We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}.
@@
 \end{equation}
-In order to get comparable fitting parameters between each measurement, we decided to fix the value of the exponential lifetime in the fitting process, leaving only the stretch lifetime as a free parameter.
-For Fig.\ref{T1}, \ref{largeur_fluct}, \ref{100_VS_1x4} and \ref{exp_B_transverse}, the chosen exponential lifetime was $T_1^{\rm ph}=3.62 \ \rm ms$.
+Further details on the fitting procedure are available in SI. For the rest of the article we will be only interested on the stretched exponential lifetime ${T_1^{\rm dd}}$.
 
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
-\caption{Tuning resonant coupling through magnetic field orientation.
+%\begin{figure}
+%\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
+%\caption{Tuning resonant coupling through magnetic field orientation.
 (a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice.
@@
 Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
-\label{largeur_fluct}
-\end{figure}
+%\label{largeur_fluct}
+%\end{figure}
 
-Fig.
+%Fig.
 \ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins.
@@
 By tuning the angle of the magnetic field, we can therefore control the frequency detuning between the different classes, and as a result we can control the number of resonant NV centers by bringing in and out of resonance several classes.
-
-Fig.
+%
+%Fig.
 \ref{largeur_fluct} (c) in particular shows the dipole-induced lifetime $T_1^{\rm dd}$ for a given class as a function of the energy detuning with another class.
@@
 
-The depolarization in low field can be explained in large part due to the lift of degeneracy between the four classes, as can be seen in Fig.
-\ref{100_VS_1x4} : On the top part, a magnetic field is scanned in a direction chosen in order to quickly split the transition frequencies of the four NV classes.
-We can observe an increase of about 4\% in the photoluminescence and an increase by a factor of $\sim$ 16 in the spin dipolar-induced lifetime ${T_1^{\rm dd}}$ as the magnetic field is increased.
+One of the way we can isolate dipole-dipole related phenomena with NV centers is by tuning the number of NV centers resonant with each other.
+Indeed, Fig.
+\ref{100_VS_1x4} (bi) shows the characteristic 8 lines observed with ensemble NV optically detected magnetic resonance (ODMR) spectrum.
+These 8 lines correspond to the four "classes" of NV centers - that is physical orientation of the NV axis is the diamond crystal cell - multiplied by the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$.
+The transitions of the four different classes can be controlled somewhat independently by changing the amplitude and orientation of the external magnetic field, and in particular they can be brought to resonance for particular orientation of the magnetic field.
+Fig.
+\ref{100_VS_1x4} (ci) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis.
+Other example of class resonance are given in SI.
 
-However the splitting of the four classes can not explain entirely the spin depolarization in zero field.
-The bottom part of Fig.
-\ref{100_VS_1x4} shows a similar experiment when the magnetic field is aligned with the [100] crystalline axis.
-In this particular orientation, the projection of the magnetic field on the four NV classes is identical, meaning that the resonance condition between the NV centers remains the same in zero and non-zero field.
-Yet we can still observe an increase of $\sim$ 0.5\% in the photoluminescence and an increase of $\sim$ 30\% in ${T_1^{\rm dd}}$.
-We attribute the remaining depolarization to two phenomena : the change of basis in the spin Hamiltonian's eigenstate in zero-field, and the near-resonance of the dipolar double-flip processes.
+Fig.
+\ref{100_VS_1x4} (bii) and (biii) shows respectively the PL and the stretched exponential spin decay $T_1^{\rm dd}$ of an ensemble of NV centers with respect to the amplitude of the external magnetic field.
+The orientation of the magnetic field was chosen in order to lift the degeneracy between the four classes, as shown in Fig.
+\ref{100_VS_1x4} (bi).
+We observe that the drop in PL at low magnetic field is indeed correlated with an increase in the spin decay rate, as is expected due to the higher number of resonant spins in zero magnetic field, where all four classes are resonant, compared to the non-degenerate case for $\vec \neq 0$.
+We can also see that the decrease in PL for $B>40\ \rm G$ due to state mixing by the transverse magnetic field is not correlated to a modification of the spin lifetime.
 
-The decrease of spin lifetime and the associated drop in PL for $B\approx 20\ \rm G$ is due to cross-relaxations between NV $^-$ centers with a $^{13}$C as first a neighbor, and those without \cite{pellet2021optical}.
+However, the splitting of the four classes is not sole reason to the decrease of the spin lifetime in low magnetic field.
+Indeed, we performed similar measurements in Fig.
+\ref{100_VS_1x4} (cii) and (ciii) but this time fixing the magnetic field orientation along the [100] axis.
+For this particular orientation, the four classes are always resonant, regardless of the field amplitude.
+We can see that, although considerably reduced, there still is an increase in the spin decay rate and a corresponding drop in PL for low magnetic field values.
+We attribute the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20\ \rm G$ to dipolar interaction with NV having a first neighbor $^{13}C$ \cite{pellet2021optical}.
 
+The aim of this paper is to understand the remaining contribution to the spin decay in this second case.
+We identified and isolated two additional effects : the domination by the local electric field for low magnetic field value which results in a change of the spin Hamiltonian eigenstates, and the presence of double-flip processes where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.
+
+In order to distinguish these two contributions, we decided to exploit the transverse magnetic field.... (lis un peu ce que t'as déja écris avant de tout réécrire, je ter connais moi futur)
 The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :

### 2022-06-02 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
tata yoyo

@@
 \
-\title{Low field depolarization of electronic spins through dipole-dipole coupling}
+\title{Low field depolarization of NV$^-$ centers through dipole-dipole coupling}
 
@@
 \subsection{Spectral range of the dipole-dipole cross-relaxations}
+\label{fluctuator width}
 \begin{figure}
@@
 \end{figure}
-In the absence of other
 
@@
 
-These predictions can be tested experimentally.
+These computed values can be tested experimentally.
 Fig.
@@
 \begin{itemize}
-\item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
-\item $\vec{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.
-For the two resonant classes : $\bar \eta = \bar {\eta}_{same} + \bar {\eta}_{far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
-\item $\vec{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{same}+\bar {\eta}_{close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
-\item $\vec{B} \parallel \langle 111 \rangle$ : For the non-resonant class (parallel to $\vec B$) : $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.
-For the triply resonant classes: $\bar \eta = \bar {\eta}_{same}+2\bar {\eta}_{far}$ and $\bar \eta^2\approx 28.4\, \bar \eta_0^2$.
-\item $\vec{B} \parallel \langle 100 \rangle$ : For the quadruple resonance : $\bar \eta = \bar {\eta}_{same}+2\bar {\eta}_{far}$ and $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
+\item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
+\item $\vec{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2= \bar \eta_0^2$.
+For the two resonant classes : $\bar \eta = \bar {\eta}_{\rm same} + \bar {\eta}_{\rm far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
+\item $\vec{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{\rm same}+\bar {\eta}_{\rm close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
+\item $\vec{B} \parallel \langle 111 \rangle$ : For the non-resonant class (parallel to $\vec B$) : $\bar \eta = \bar {\rm \eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.
+For the triply resonant classes: $\bar \eta = \bar {\eta}_{\rm same}+2\bar {\eta}_{\rm far}$ and $\bar \eta^2\approx 28.4\, \bar \eta_0^2$.
+\item $\vec{B} \parallel \langle 100 \rangle$ : For the quadruple resonance : $\bar \eta = \bar {\eta}_{\rm same}+2\bar {\eta}_{\rm close} + \bar {\eta}_{\rm far}$  and $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
 \end{itemize}
 
-The measured increase in the decay rate is generally smaller than the predicted one : we measured an increase by a factor $\sim 4$ for a two class degeneracy instead of the predicted $10.0$, and an increase by a factor $\sim 16$ for a two class degeneracy instead of the predicted $42.8$.
+The measured increase in the decay rate is generally smaller than the predicted one : we measured an increase by a factor $\sim 4$ for a two-class degeneracy instead of the predicted $10.0$, and an increase by a factor $\sim 16$ for a four-class degeneracy instead of the predicted $42.8$.
 \cite{choi_depolarization_2017} measured an increase by a factor $\sim 4$ on a two-class degeneracy instead of the predicted $7.24$ or $10.0$.
@@
 As described in sec.
-\ref{sec Hamiltonian}, when the single spin Hamiltonian of the NV center is dominated by the electric field or the transverse magnetic field (as long as $\gamma_e B_\perp \ll D$), the eigen basis of the NV Hamiltonian is close to $\{\ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
+\ref{sec Hamiltonian}, when the single spin Hamiltonian of the NV center is dominated by the electric field or the transverse magnetic field (as long as $\gamma_e B_\perp \ll D$), the eigenbasis of the NV Hamiltonian is close to $\{\ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
 We therefore have to rewrite the dipole-dipole Hamiltonian using this new basis for both spins in order to perform similar calculations as the ones done in the previous section.
+
+Since the flip-flop processes are now of the form $\ket{0,\pm}\bra{\pm,0}$, we redefine the $\eta$ factor as :
+\begin{equation}
+\eta^2=\frac{1}{3} \abs{\mel{\pm ,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,\pm }}^2  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}
+\end{equation}
 
@@
 
+\begin{equation}
+  S_x = \begin{pmatrix}
+  0&1&0 \\
+  1&0&0 \\
+  0&0&0
+  \end{pmatrix}
+  \begin{matrix}
+  \bra{-} \\
+  \bra{0} \\
+  \bra{+}
+  \end{matrix}  
+  , \quad 
+ S_y = \begin{pmatrix}
+  0&0&0 \\
+  0&0&1 \\
+  0&1&0
+  \end{pmatrix}
+  \begin{matrix}
+  \bra{-} \\
+  \bra{0} \\
+  \bra{+}
+  \end{matrix}  
+ , \quad 
+  S_z = \begin{pmatrix}
+  0&0&1 \\
+  0&0&0\\
+  1&0&0
+  \end{pmatrix}
+  \begin{matrix}
+  \bra{-} \\
+  \bra{0} \\
+  \bra{+}
+  \end{matrix}
+  \end{equation}
+
+The symmetry in the $(xy)$ plane is broken by the presence of the transverse electric or magnetic field.
+The $x_i$ direction is defined as the direction of the transverse electric or magnetic field which dominates the single spin Hamiltonian of the particle $i$.
+
+We will consider two cases :
+\begin{itemize}
+\item The $x$ direction is defined by an external field (transverse magnetic field or strong external electric field) projected on the $(xy)$ plane of each classes.
+In particular this means that two spins from the same class have the same $(\hat x, \hat y, \hat z)$ basis.
+\item The $x$ direction is defined by the local electric field generated by the charges in the crystal.
+We will assume that there are no correlations in the electric field between the NV and fluctuator, and will sample a random angle $\psi \in [0,2\pi]$ between the axes $\hat{x}_1$ and $\hat{x}_2$.
+\end{itemize}
+
+\begin{table}
+\begin{tabular}{cccc}
+\hline
+$\bar{\eta}$ table & $\widehat{z_1 z_2}=0$ & $\widehat{z_1 z_2}=70.5^\circ$ & $\widehat{z_1 z_2}=109.5^\circ$ \\
+\hline
+$\ket{\pm 1}$ basis & $\frac{2}{3\sqrt{3}}=$ 0.3849 & 0.6507 & 0.8328 \\
+$\ket{+/-}$ basis $\hat{x}_1\neq \hat{x}_2$ & 0.7110  & 0.6828 & 0.6828 \\
+$\ket{+/-}$ basis $\hat{x}_1= \hat{x}_2$ & $\frac{4}{3\sqrt{3}}=$0.7698  & 0.6951 & 0.6951 \\
+\hline
+\end{tabular}
+\caption{Computation of $\frac{\bar{\eta}}{ \frac{1}{4} \cdot \sqrt{\frac{1}{3}}}$ for the different eignebasis of the single spin Hamiltonian, and for the different angles between the $z$ axis of the two spins.}
+\label{table eta}
+\end{table}
+
+For this two cases, which we will call $\hat{x}_1=\hat{x}_2$ and $\hat{x}_1\neq \hat{x}_2$, we can compute the $\bar \eta$ factor for the three scenarios discussed previously : $\widehat{z_1 z_2}=0$ (same class), $\widehat{z_1 z_2}=70.5^\circ$ and $\widehat{z_1 z_2}=109.5^\circ$.
+The results, as well as those in the $\ket{\pm 1}$ basis are presented in Table \ref{table eta}.
+
+There are two particular situations where these values can be experimentally tested :
+
+The first one is described in Fig.
+[REF] of the main text : in the presence of pure transverse magnetic field (for a single class, non-resonant with the three other classes), the decay rate of the class decreases with the magnetic field amplitude, due to the double-flip processes (described below), and then reaches a plateau with a decay rate value $\sim 2$ times bigger than in the longitudinal magnetic field case.
+This increase is in agreement with the fact that $\bar{\eta}_{\rm same}$ is 2 times bigger in the non-magnetic $\ket{+/-}$ basis than in the magnetic $\ket{\pm 1}$ basis.
+The measured increase is again smaller than the predicted factor of 4 given by eq.
+(\ref{eq 1/T}).
+
+The second one is the difference between the $\vec{B} \parallel \langle 100 \rangle$ case and the $\vec{B}=0$ case.
+When a magnetic field is applied in the [100] direction, all four classes are resonant and the spin Hamiltonian basis is $\{ \ket{0},\ket{+1},\ket{-1} \} $.
+The $\bar \eta$ factor in this case has been calculated in the previous section : $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
+When no external magnetic field is applied, all four classes are also resonant but the spin Hamiltonian basis is  $\{ \ket{0},\ket{+},\ket{-} \} $ (see Sec.
+\ref{NV Hamiltonian}) and double-flip processes are near-resonant.
+The decay rate due purely to flip-flop in the $\ket{+/-}$ basis is proportional to $\bar \eta^2 = (\bar{\eta}_{\rm same}^{+/-} + 3\bar{\eta}_{\rm diff}^{+/-})^2 \approx 51.4\, \bar \eta_0^2$.
+Fig.
+[REF] of the main text shows that we indeed observe a shorter lifetime when $\vec B=0$, in agreement with the higher value of $\bar \eta^2$, but this decrease could also come from the double-flip processes which are absent when $\vec B \neq 0$.
+
+\subsection{Double-flip processes}
+
+Double-flip processes, that is processes where the two spins flip in the same direction, compared to flip-flop where the two spins flip in opposite directions, are related to matrix elements  such as $\mel{+1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-1}$ in the $\ket{\pm 1}$ basis or $\mel{+,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-}$ in the $\ket{\pm}$ basis.
+These matrix elements couple two-spin states that are never fully resonant, however with low-to zero magnetic, the residual splitting due to local electric and magnetic field ($\sim 6\ \rm MHz$ with our samples) is small enough compared to the fluctuators linewidth  measured in sec.
+\ref{fluctuator width} to be $\sim 8\ \rm MHz$, so that double-flip processes may still occur.
+
+Fig.
+[REF] of the main text shows that for lower transverse field values, the spin decay rate is increased significantly.
+We attribute this increase to double-flip processes.
+We can see that the decay rate increase ($\sim$ 5 times the baseline value) is significantly higher than the increase due solely to the $\ket{+/-}$ basis ($\sim$ 2 times the baseline value).
+This lead us to believe that the double-flip process are the main reason behind the decrease in spin lifetime for $\vec B=0$ observed in Fig.
+[REF] of the main text.
+This would probably mean that double-flip processes are the dominant factor in the sensitivity of the magnetometry protocol presented in the main text.
 \bibliographystyle{plain}

### 2022-06-03 Mes papiers/Low field CR/Low_field_CR.tex
tutu

@@
 
-\title{Low field depolarization of NV$^-$ centers through dipole-dipole coupling}
+\title{Low magnetic field depolarization of NV$^-$ centers via dipole-dipole interaction}
 
@@
 \includegraphics[width=0.45\textwidth]{Figures/fig dense vs pas dense}
-\caption{Photoluminescence from NV centers as a function of an external magnetic field randomly oriented for two samples : (a) a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
+\caption{Photoluminescence measurement from NV centers ensemble as a function of an external magnetic field randomly oriented (a) for a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) for an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
 \label{PL_NV_density}
@@
 
-In this letter/article, we characterize the depolarization of the spins observed for dense ensemble of NV centers in zero magnetic field and its potential application for DC magnetometry.
+%In this letter/article, we characterize the depolarization of the spins observed for dense ensemble of NV centers in zero magnetic field and its potential application for DC magnetometry.
 While the main mechanism behind the depolarization, the lift in the degeneracy between the four classes of NV centers, is already well studied and can be exploited in a microwave-less vector magnetometry protocol, we found two other depolarization mechanisms specific to the zero-field region which could play an important role in a low-field magnetometry protocol.
-
-The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}.
+%
+%The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}.
 The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state.
@@
 This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.
+
+Fig.
+\ref{B_transverse} shows the change in photoluminescence (PL) with respect to an external magnetic field for two samples with a low and high concentration of NV$^-$ centers.
+Both samples show a decrease in PL as the magnetic field amplitude increases, due to eigenstate mixing by the transverse part of the magnetic field \citep{epstein2005anisotropic,lai2009influence}.
+We notice however a stark difference in the low magnetic field region where only the high-density sample observe a drop in PL.
+In this article we propose to explore the reasons behind this zero-field line present only for dense NV$^-$ ensemble.
+
+Faire la jonction avec le T1
 
@@
 
-The $T_1$ depolarization dynamics of single or sparse NV centers at room temperature is dominated by a two-phonon Raman process, which depends on the crystal lattice temperature but does not (?) depend on the external magnetic field.
+The $T_1$ depolarization dynamics of single or dilute NV centers at room temperature is dominated by a two-phonon Raman process, which depends on the crystal lattice temperature but does not (?) depend on the external magnetic field.
 However, it has been observed that dense NV centers ensemble have an additional spin decay channel which depends greatly on the magnetic field and not on the temperature.
@@
 By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.
+
+\begin{figure*}
+\includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
+\caption{Dependency of the magnetic field angle for the zero field depolarization.
+(a) ODMR spectrum in zero field.
+(b) ODMR spectrum for a magnetic field $\approx$ 60 G.
+The field is aligned (b2) or misaligned (b1, $\approx 24^\circ$ of misalignment) with the crystalline [100] direction.
+(c) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude for the same field orientations as fig (b).
+(d) Stretch part of the lifetime decay as a function of the magnetic field amplitude for the same field orientations as fig (b).}
+\label{100_VS_1x4}
+\end{figure*}
 
@@
 
-\begin{figure*}
-\includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
-\caption{Dependency of the magnetic field angle for the zero field depolarization.
-(a) ODMR spectrum in zero field.
-(b) ODMR spectrum for a magnetic field $\approx$ 60 G.
-The field is aligned (b2) or misaligned (b1, $\approx 24^\circ$ of misalignment) with the crystalline [100] direction.
-(c) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude for the same field orientations as fig (b).
-(d) Stretch part of the lifetime decay as a function of the magnetic field amplitude for the same field orientations as fig (b).}
-\label{100_VS_1x4}
-\end{figure*}
-
 One of the way we can isolate dipole-dipole related phenomena with NV centers is by tuning the number of NV centers resonant with each other.
@@
 
-In order to distinguish these two contributions, we decided to exploit the transverse magnetic field.... (lis un peu ce que t'as déja écris avant de tout réécrire, je ter connais moi futur)
-The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
-\begin{equation}
-\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
-\end{equation}
-Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$.
+In order to distinguish these two contributions, we decided to simulate the effect of the electric field by using a transverse magnetic field : for relatively weak transverse magnetic field, the eigen basis of the spin Hamiltonian is close to $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $, which is the the spin Hamiltonian basis when the electric field dominates, as detailed in SI.
+This property has been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures/fig transverse field V2}
+\caption{(a) Simulated energies for the spin Hamiltonian in the presence of pure transverse magnetic field $B_\perp$.
+The three spin levels are denoted $\ket{g}$, $\ket{d}$ and $\ket{e}$.
+(b) Blue dashed curve : frequency detuning $\Delta \nu$ between the $\ket{g}$, $\ket{d}$ states.
+Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\bra{e}$ and $\bra{+}$.}
+\label{B_transverse}
+\end{figure}
+
+Fig.
+\ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian with respect to a purely transverse magnetic field.
+We will denote these eigenstates in the general case $\ket{g}$, $\ket{d}$ and $\ket{e}$.
+For weak transverse magnetic field, $\ket{g} \approx \ket{0}$, $\ket{d}=\ket{-}$ and $\ket{e} \approx \ket{+}$.
+As the magnetic field increases, $\ket{g}$ and $\ket{e}$ starts to become a mixing of $\ket{0}$ and $\ket{+}$ while $\ket{d}$ remains equal to $\ket{-}$.
+
+Crucially, this mixing happens on a different scale than the associated energy splitting of the $\ket{d}$ and $\ket{e}$ states : Fig.
+\ref{B_transverse} (b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$.
+We can see that for a field value of $\sim 130 G$, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50 MHz$, far exceeding the range of the dipole-dipole resonant coupling (see details in SI), but $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
+We can therefore decouple the effects of the double-flip processes, which are no longer resonant, from the effects of the change of basis induced by the transverse electric - or in this case magnetic - field.
+
+Fig.
+\ref{B_transverse} (c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers exposed to a transverse magnetic field between 25 and 130 G.
+The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
+We attribute the decrease in the spin decay rate when the detuning increases to the loss of effectiveness of the double-flip processes.
+We can notice that the decay rate reaches a plateau for $\Delta \nu \geq 30\ \rm MHz$ with a value about twice as big as the decay rate for a longitudinal magnetic field.
+This plateau we attribute to the effect of the transverse magnetic field, which we assume to be similar to the effect of the electric field in the low magnetic field regime.
+We should note that the effect of the double-flip processes is about 4 times more important in the depolarization rate than the effect of the electric field, even though the states are not initially fully resonant.
+
+Overall we predicted, based on an extension of the fluctuator model developed in SI, an increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$.
+While the measured increase is smaller than the predicted one, as with most of the predictions based on the fluctuator model, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.
+
+Similarly, we can compute the predicted decay rate in the $\vec B \parallel \left[100\right]$ case compared to the $\vec B = 0$ case, and find a theoretical increase of $\sim 20\%$ in the $\vec B = 0$ case due to the effects of the local electric field.
+This, along with the double-flip processes explains the behavior observed in Fig.
+\ref{100_VS_1x4} (cii) and (ciii).
+
+%... (lis un peu ce que t'as déja écrit avant de tout réécrire, je te connais moi futur. En vrai c'est de la merde, j'ai tout réécrit)
+%The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
+%\begin{equation}
+%\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
+%\end{equation}
+%Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$.
 In situations such as the one described in Fig.
@@
 In zero magnetic field however, we have to take into account other aspects of the dipolar Hamiltonian in consideration.
-
-First is the change of basis of the single spin Hamiltonian $\mathcal{H}_s$ : in zero external magnetic field, the eigenstates of $\mathcal{H}_s$ in the $\{\ket{+1},\ket{-1}\}$ manifold are determined by local electric and magnetic field coming from nearby impurities \cite{mittiga2018imaging}.
+%
+%First is the change of basis of the single spin Hamiltonian $\mathcal{H}_s$ : in zero external magnetic field, the eigenstates of $\mathcal{H}_s$ in the $\{\ket{+1},\ket{-1}\}$ manifold are determined by local electric and magnetic field coming from nearby impurities \cite{mittiga2018imaging}.
 For the samples used in this study, $\mathcal{H}_s$ was dominated by local electric field (see SI), meaning that the proper eigenstates of $\mathcal{H}_s$ in zero magnetic field are $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $
-
-Because the splitting between $\ket{+}$ and $\ket{-}$ ($\approx 9\ \rm MHz$) is much greater than the dipole-dipole interaction ($J_0 / r^3 \approx 30\ \rm kHz$), the flip-flop terms of $\mathcal{H}^{\rm dd}$ now reads $\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}$.
+%
+%Because the splitting between $\ket{+}$ and $\ket{-}$ ($\approx 9\ \rm MHz$) is much greater than the dipole-dipole interaction ($J_0 / r^3 \approx 30\ \rm kHz$), the flip-flop terms of $\mathcal{H}^{\rm dd}$ now reads $\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}$.
 Averaging over all relative positions of two aligned NV centers, we found that on average $\abs{\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}}$ was greater than
-$\abs{\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}}$ by a factor $1.8 \sim 2$ depending on the correlation lengths of the local electric field.
-When all four classes are resonant, the theoretical decrease in the spin lifetime due to the change of basis is $\sim 20 \%$ (See SI de ouf).
-
-Second is the near-resonance condition of the double-flip processes such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{-1,0}$ or $\mel{0,+}{\mathcal{H}^{\rm dd}}{-,0}$.
+%$\abs{\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}}$ by a factor $1.8 \sim 2$ depending on the correlation lengths of the local electric field.
+%When all four classes are resonant, the theoretical decrease in the spin lifetime due to the change of basis is $\sim 20 \%$ (See SI de ouf).
+%
+%Second is the near-resonance condition of the double-flip processes such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{-1,0}$ or $\mel{0,+}{\mathcal{H}^{\rm dd}}{-,0}$.
 These terms usually couple non resonant states due to the energy mismatch between $\ket{+1}$ and $\ket{-1}$.
@@
 \ref{largeur_fluct} to be $\approx 8\ \rm MHz$, making the double-flip population transfer possible.
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig transverse field simu}
-\caption{Simulated eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field.
+%
+%\begin{figure}
+%\includegraphics[width=0.45\textwidth]{Figures/fig transverse field simu}
+%\caption{Simulated eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field.
 (a) Energies of the three eigenstates $\ket{g}$, $\ket{d}$ and $\ket{e}$ as a function of the magnetic field amplitude.
@@
 Is equal to the projection of $\ket{g}$ on $\ket{0}$.}
-\label{calculs_B_transverse}
-\end{figure}
-
-In order to evaluate the relative contribution of these two factors, we investigate the spin relaxation in the case of pure transverse magnetic field.
+%\label{calculs_B_transverse}
+%\end{figure}
+%
+%In order to evaluate the relative contribution of these two factors, we investigate the spin relaxation in the case of pure transverse magnetic field.
 Calling $\{ \ket{g}, \ket{d}, \ket{e} \}$ the three eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field, Fig.
@@
 However the energy difference between $\ket{d}$ and $\ket{e}$ reaches $\approx 45\ \rm G$ which is high enough to cancel out the double-flip processes and allows us to isolate the change of basis hypothesis from the double-flip one.
-%As shown in Fig.
+%%As shown in Fig.
 SOON, (and in SI), the $\{ \ket{0},\ket{+},\ket{-} \} $ basis correspond to the spin Hamiltonian eigenstates in two situations : either in zero magnetic field when the electric field dominates, or when the magnetic field is purely transverse, as long as it remains small before the zero-field splitting.
 %(rq : est-ce que je dois mettre le hamiltonien de spin dans le main text du coup ?).
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig transverse field}
-\caption{Modification of the stretch lifetime in the presence of purely transverse magnetic field.
+%
+%\begin{figure}
+%\includegraphics[width=0.45\textwidth]{Figures/fig transverse field}
+%\caption{Modification of the stretch lifetime in the presence of purely transverse magnetic field.
 (a) Stretch component of the ensemble lifetime as a function of the field amplitude.
@@
 The green dashed line correspond to the lifetime of a single class aligned with the magnetic field.}
-\label{exp_B_transverse}
-\end{figure}
-
-Fig.
+%\label{exp_B_transverse}
+%\end{figure}
+%
+%Fig.
 \ref{exp_B_transverse} shows the dipole-dipole spin relaxation $T_1^{\rm dd}$ for a class of NV whose axis is orthogonal to the applied magnetic field.
@@
 Plotting then $1/T_1^{\rm dd}$ as a function of $\Delta \nu$, we can notice two things.
-
-First the decrease of $1/T_1^{\rm dd}$ as $\Delta \nu$ increases which corresponds to the loss in effectiveness of the double flip processes.
+%
+%First the decrease of $1/T_1^{\rm dd}$ as $\Delta \nu$ increases which corresponds to the loss in effectiveness of the double flip processes.
 The shape of the curve matches decently well with the tail of a Lorentzian centered in $\Delta \nu=0$ with a width of 8 MHz, confirming that the double-flip process interaction range is likely determined by the fluctuators noise spectrum just like the flip-flop processes.
-
-Second is the plateau reached for $\Delta \nu \gtrsim 30\ \rm MHz$.
+%
+%Second is the plateau reached for $\Delta \nu \gtrsim 30\ \rm MHz$.
 The final $1/T_1^{\rm dd}$ value is about twice as high as that for a spin well aligned with the magnetic field, showing the increased depolarization due to the change of basis.
-
-This experiment shows that, for the sample studied here, the depolarization due to the double-flip processes dominates the one due to the change of basis in zero magnetic field.
+%
+%This experiment shows that, for the sample studied here, the depolarization due to the double-flip processes dominates the one due to the change of basis in zero magnetic field.
 
@@
 %
-DC microwave-less magnetometry has already been performed with NV ensembles, using either cross-relaxations [les russes] or level anti-crossing [bubu].
+DC microwave-less magnetometry has already been performed with NV ensembles, using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}.
 Here we propose to perform a similar protocol, but using the spin depolarization in zero-field.
@@
 
-In order to measure the sensitivity of the measurement, we alternate a small DC field of $\approx 40 \rm G$ every few seconds and take an histogram of the measured fields, as shown in Fig.
+In order to measure the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take an histogram of the measured fields, as shown in Fig.
 \ref{magneto} (b) and (c).
@@
 
-We can then try to measure the relative importance of the three causes of spin depolarization on the field sensitivity.
+We can then try to measure the relative importance of the three causes of spin depolarization on the field sensitivity, namely the splitting of the classes, the local electric field and the double-flip processes.
 In order to do so, we measure the sensitivity while changing the angle of the field.
-When $\vec B$ is aligned with the [100] crystalline axis, only the double flips and the change of basis are present, whereas in every other orientation the three effects are at play.
+When $\vec B$ is aligned with the [100] crystalline axis, only the double flips and the electric field cause a depolarization, whereas in every other orientation the three effects are at play.
 
@@
 \ref{magneto} (d).
-We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity remains relatively flat, which means that the double-flips and change of basis effects are the dominant factors in the sensitivity of this protocol.
+We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity remains relatively flat, which means that the double-flips and electric field effects are the dominant factors in the sensitivity of this protocol.
+While it may seem surprising that the effects with a lower contribution on the PL contrast have a higher effect on this PL-based protocol, what matters here is not the absolute contrast but the slope of the change of contrast with respect to the magnetic field, which is sharper for the electric field and double flips than it is for the lift of degeneracy.
 %This is in agreement with the measurements in Fig.
 \ref{100_VS_1x4} : even though the contrast is much better in the not-[100] case, the slope in term of contrast percentage/magnetic field is comparable in both cases (en faite pas complètement, y'a pas loin d'un facteur 2 sur les dérivées numériques)
-It should be noted though that this observation is sample dependent, and that other samples, including from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from the double flips and change of basis.
+It should be noted though that this observation is sample dependent, and that other samples, including from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from the electric field and double-flip processes.
 
 \section*{Conclusion}
-In this work we have identified three mechanisms causing the extra spin depolarization observed in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole cross-relaxations between the spins : the lift of degeneracy between the four classes caused by the magnetic field, the change of basis in the spin Hamiltonian when the transverse electric field becomes dominant, and the double-flip process allowed by the proximity in energy of the spin $\ket{\pm}$ states.
-We identified the lift in degeneracy as the main cause in the zero-field depolarization, followed by the double-flip process and then the change of basis.
-With have demonstrated a potential use of this depolarization as a DC microwave-less and orientation-free magnetometer with senstivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal.We have observed that in some cases the double-flips and change of basis depolarization play a more important role than the lift of degeneracy in the low field sensitivity.
+In this work we have identified three mechanisms causing the extra spin depolarization observed in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole cross-relaxations between the spins : the lift of degeneracy between the four classes caused by the magnetic field, the domination of the local electric field which causes a change in the Hamiltonian eigenstates, and the double-flip process allowed by the proximity in energy of the spin $\ket{\pm}$ states.
+We identified the lift in degeneracy as the main cause in the zero-field depolarization, followed by the double-flip process and then the electric field.
+With have demonstrated a potential use of this depolarization as a DC microwave-less and orientation-free magnetometer with sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal.We have observed that in some cases the double-flips and change of basis depolarization play a more important role than the lift of degeneracy in the magnetometry sensitivity.
 \section*{acknowledgements}

### 2022-06-08 Mes papiers/Low field CR/Low_field_CR.tex
tata

@@
 
-\title{Low magnetic field depolarization of NV$^-$ centers via dipole-dipole interaction}
+\title{Low magnetic field depolarization of NV$^-$ centers through dipole-dipole interaction}
 
-\author{C. Pellet-Mary$^1$, P. Huillery$^?$,  M. Perdriat$^1$, G. H\'etet}
+\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, P. Huillery$^?$,  G. H\'etet}
 
@@
 \maketitle
-Blabla sur les NV, les ensembles etc.
 
-%réécrire le début et foutre la partie sur la PL avant le T1. Peut-etre dans le blabla sur les ensembles justement
-
-%We have identified three mechanisms for the low field depolarization, two of which have not been reported before.
+Intro
 
@@
 Fig.
-\ref{B_transverse} shows the change in photoluminescence (PL) with respect to an external magnetic field for two samples with a low and high concentration of NV$^-$ centers.
-Both samples show a decrease in PL as the magnetic field amplitude increases, due to eigenstate mixing by the transverse part of the magnetic field \citep{epstein2005anisotropic,lai2009influence}.
+\ref{PL_NV_density} shows the change in photoluminescence (PL) with respect to an external magnetic field for two samples with a low and high concentration of NV$^-$ centers.
+Both samples show a decrease in PL as the magnetic field amplitude increases, due to eigenstate mixing caused by the transverse part of the magnetic field \citep{epstein2005anisotropic,lai2009influence}.
 We notice however a stark difference in the low magnetic field region where only the high-density sample observe a drop in PL.
@@
 
-Faire la jonction avec le T1
+Unlike the decrease in PL for higher magnetic field values, the drop in PL at low magnetic field is associated with a decrease of the NV's spin lifetime $T_1$.
+Indeed, since the $\ket{m_s=0}$ spin state is brighter than the $\ket{m_s=\pm 1}$ states, and since the optical pumping in the $\ket{m_s=0}$ state is in competition with the relaxation in a thermal distribution of the states, decreasing the spins lifetime results in a decrease of the NV PL \citep{finco2021imaging}.
+To understand the behavior of the PL in zero-field, we therefore need to study the dynamics of the spin in the low magnetic field region.
 
@@
 (a) Sequence used to measure the spin lifetime : the laser is turned off for a time $\tau$ twice and the spin state is read at the end of the dark time by measuring the initial photoluminescence when the laser is turned back on.
-A microwave resonant with one of the spin transition is used in one of the sequence to project the $\ket{0}$ spin sate in either $\ket{+1}$ or $\ket{-1}$.
-The sequence result is the subtraction $S_1-S_2$.
-(b) Measurement of a dense ensemble spin lifetime in zero field and non-zero field, fitted with an expression $A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$}
+A microwave $\pi$-pulse resonant with one of the spin transition is placed before the second reading.
+The  $T_1$ sequence result is the subtraction $S_1-S_2$.
+(b) Measurement of $S_1-S_2$ for a high density of NV centers in zero magnetic field field and non-zero magnetic field.
+The fit (orange plain line) is detailed in the main text}
 \label{T1}
@@
 
-The $T_1$ depolarization dynamics of single or dilute NV centers at room temperature is dominated by a two-phonon Raman process, which depends on the crystal lattice temperature but does not (?) depend on the external magnetic field.
-However, it has been observed that dense NV centers ensemble have an additional spin decay channel which depends greatly on the magnetic field and not on the temperature.
-This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling.
+The $T_1$ depolarization dynamics of single or dilute NV centers at room temperature is dominated by a two-phonon Raman process \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depends on the crystal lattice temperature but does not depend on the external magnetic field.
+However, it has been observed that dense NV centers ensemble have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi_depolarization_2017, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field and not on the temperature .
+This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi_depolarization_2017}.
 Some inhomogeneity between the NV centers is further needed in order to explain the depolarization of the spin ensemble.
@@
 The spin lifetime protocol used here is described in Fig.
-\ref{T1} and is based on previous similar experiments.
+\ref{T1} and is based on previous similar experiments \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi_depolarization_2017}.
 It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$.
-Using only this sequence can result in artifacts in the signal, mostly due to charge state transfer in the dark \citep{giri_coupled_2018}.
+Using only this sequence can result in artifacts in the signal, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019}.
 It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to project the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
@@
 \includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
-\caption{Dependency of the magnetic field angle for the zero field depolarization.
+\caption{Dependency on the magnetic field angle for the zero field depolarization.
 (a) ODMR spectrum in zero field.
-(b) ODMR spectrum for a magnetic field $\approx$ 60 G.
-The field is aligned (b2) or misaligned (b1, $\approx 24^\circ$ of misalignment) with the crystalline [100] direction.
-(c) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude for the same field orientations as fig (b).
-(d) Stretch part of the lifetime decay as a function of the magnetic field amplitude for the same field orientations as fig (b).}
+(bi) ODMR spectrum for a magnetic field $\approx$ 60 G, misaligned by $\sim  24^\circ$ with the [100] axis.
+(bii) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude.
+(biii) Stretch part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude.
+(ci), (cii) and (ciii) : same measurement with a magnetic field aligned with the [100] axis}
 \label{100_VS_1x4}
@@
 (b) Blue dashed curve : frequency detuning $\Delta \nu$ between the $\ket{g}$, $\ket{d}$ states.
-Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\bra{e}$ and $\bra{+}$.}
+Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\bra{e}$ and $\bra{+}$.
+(c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single class as a function of the measured detuning $\Delta \nu$ between the two transitions $\ket{g} \to \ket {d}$ and  $\ket{g} \to \ket {e}$ for purely transverse magnetic field.
+The green dashed line correspond to the decay rate for the same class when the magnetic field is longitudinal}
 \label{B_transverse}
@@
 (c) Histogram of the measurement in Fig.
-(b) fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$}.
-(d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.
+(b) fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
+(d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.}
 \label{magneto}
@@
 In this work we have identified three mechanisms causing the extra spin depolarization observed in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole cross-relaxations between the spins : the lift of degeneracy between the four classes caused by the magnetic field, the domination of the local electric field which causes a change in the Hamiltonian eigenstates, and the double-flip process allowed by the proximity in energy of the spin $\ket{\pm}$ states.
-We identified the lift in degeneracy as the main cause in the zero-field depolarization, followed by the double-flip process and then the electric field.
-With have demonstrated a potential use of this depolarization as a DC microwave-less and orientation-free magnetometer with sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal.We have observed that in some cases the double-flips and change of basis depolarization play a more important role than the lift of degeneracy in the magnetometry sensitivity.
+We identified the lift in degeneracy as the main cause in the zero-field depolarization, followed by the double-flip processes and then the electric field.
+We have demonstrated a potential use of this depolarization as a DC microwave-less and orientation-free magnetometer with sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal.We have observed that in some cases the double-flips and electric field play a more important role than the lift of degeneracy in the magnetometry sensitivity.
 \section*{acknowledgements}

### 2022-06-21 Talks/Journées de l'axe/presentation.tex

@@
 \maketitle
-\centering
-Mettre logos ici
+\begin{center}
+\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{logos}
+\end{center}
 \end{frame}
@@
 %\section{Physics of the NV center}
-\begin{frame}{What are NV centers and why are they everywhere?}
+\begin{frame}{What are NV centers?}
 \centering
@@
 
-\begin{frame}{What are NV centers and why are they everywhere ?}
+\begin{frame}{What are NV centers good at ?}
 \begin{itemize}
-\item Good optical properties : quantum yield $\approx 1$, stable in time and temperature
+\item Good optical properties : quantum yield $> 0.9$, photostable
+%\pause
 \medskip
 \item Good spin properties : $T_1\approx\ \rm ms$ and $T_2^*\sim \mu$s at 300 K.
+%\pause
 \medskip
@@
 \end{itemize}
+%\pause
 \bigskip
@@
 \end{center}
+%\pause
 
@@
 
+%\pause
+
 $\bullet$ Diamond magnetic microscopy :
@@
 \begin{itemize}
-\item To increase the ensemble sensitivity we need to increase the density of NV centers
+\item To increase the magnetic field sensitivity we need to increase the density of NV centers
+%\pause
 \item Chemists and material scientists have made huge progress to grow NV-rich diamond with little other impurities 
 \begin{center}
-\includegraphics[width=0.65\textwidth,height=0.9\textheight,keepaspectratio]{diamant Alex}
+\includegraphics[width=\textwidth,height=0.2\textheight,keepaspectratio]{diamant Alex}
 \end{center}
-\item When the defects density reaches a critical point, many body effects start to appear due to dipole-dipole interaction and charge transfer (3 ppm of NV = 30 kHz dipole coupling between NV neighbors)
+%\pause
+\item When the defects density reaches a critical point, many body effects start to appear due to dipole-dipole interaction and charge transfer (3 ppm of NV = 30 kHz dipole coupling rate between NV neighbors)
 \end{itemize}
@@
 
-\begin{frame}{Other results with dense ensemble of spins}
-$\bullet$ Optical detection of dark spins :
-\begin{center}
-\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{dark spins}
-\end{center}
-$\bullet$ Mechanical detection of resonant dipolar coupling :
-\begin{center}
-\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{Cr-meca}
-\end{center}
+\begin{frame}{Optical detection of dark spins}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{dark spins}
+\end{frame}
+
+\begin{frame}{Mechanical detection of dipole-dipole coupling}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Cr-meca}
 \end{frame}
@@
 \item Many body effects start to manifest with the new NV-rich samples being made
+\medskip
+\item These effects can be exploited for applications : detection of dark spins, magnetometry, mechanical cooling...
 \end{itemize}

### 2022-06-21 Mes papiers/Low field CR/Low_field_CR.tex
tutu

@@
 
-\title{Low magnetic field depolarization of NV$^-$ centers through dipole-dipole interaction}
+\title{Depolarization of Interacting NV ensembles : \\
+The Role of Local Electric Fields and Double Flip processes}
 
@@
 \begin{abstract}
-C'est trop bien
+We present experimental results on dipolar interaction processes in ensemble of spin-1 systems.
+Using the spin of highly doped ensembles of negatively charged nitrogen vacancy centers in diamonds, we identify regimes where flip-flop, double-flip processes as well as the mixing induced by local electric field dominate.
+%We analyse these results from the perspective of magnetometry applications, and show that both double-flip processes and local electric fields improve the magnetic field sensitivity.
+Our results are relevant for understanding decoherence in many-body spin systems as well as for high sensitivity magneto- and electro-metry with long-lived interacting solid-state spins.
+As a proof of principle, we present an orientation-free and microwave-free magnetometer operating with a sensitivity below $100$ nT/$\sqrt{\rm Hz}$.
 \end{abstract}
@@
 
-Intro
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig dense vs pas dense}
-\caption{Photoluminescence measurement from NV centers ensemble as a function of an external magnetic field randomly oriented (a) for a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) for an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
-\label{PL_NV_density}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=0.45\textwidth]{Figures/fig_dense_vs_pas_dense.pdf}
+%\caption{Photoluminescence measurement from NV centers ensemble as a function of an external magnetic field applied in an arbitrary direction (a) for a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) for an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
+%\label{PL_NV_density}
+%\end{figure}
 
@@
 
+The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science in part thanks to the possibility to optically polarize and read-out its spin state at ambient conditions.
+Many recent work indeed focus on ensembles of spins from NV centers with the prospect to boost the magnetic field sensing capabilities and for studying many body effects (resultats de chez BOSCH, LUKIN , FEFOR, ALEX, BIFONE)
+
+When the spin ensemble becomes dense, inhomogeneities in the relaxation rates of NVs can give rise to spin depolarisation with a very rich many body dynamics associated with disorder.
+This mechanisms however limits the efficiency of magnetometers (LUKIN PRX).
+%It was proposed to exploit these CR to perform sensitive microwave free magnetometry (REF).  
+Recent magnetometry proposals have been put forward at zero magnetic field.
+There, the depolarisation effect was shown to be much stronger, but the detailed studies of relaxation mechanisms have not been identified to the best of our knowledge.
+Some of the possible processes that lead to cross-relaxation in strongly coupled dipolar systems are depicted in Fig.~1.
+The goal of this paper is to identify regimes where the spin flip-flop within different classes of NV centers, double-flip up and down as well as the mixing induced by local electric field play a role.
+
+The electronic spin of NV center is a spin-1 system in the ground state (see Fig.
+1).
+It can be optically polarized in the $\ket{m_s=0}$ state.
+The photoluminescence of this state is also is larger than the $\ket{m_s=\pm 1}$ states enabling spin-read out at ambient temperature.
+The $\ket{m_s=\pm 1}$ spin states are separated from the $\ket{m_s=0}$ state by  $D = (2\pi) 2.87$ GHz so that when a resonant microwave or static transverse magnetic field is applied, the photoluminescence is reduced \citep{epstein2005anisotropic,lai2009influence}.
 Fig.
-\ref{PL_NV_density} shows the change in photoluminescence (PL) with respect to an external magnetic field for two samples with a low and high concentration of NV$^-$ centers.
-Both samples show a decrease in PL as the magnetic field amplitude increases, due to eigenstate mixing caused by the transverse part of the magnetic field \citep{epstein2005anisotropic,lai2009influence}.
-We notice however a stark difference in the low magnetic field region where only the high-density sample observe a drop in PL.
-In this article we propose to explore the reasons behind this zero-field line present only for dense NV$^-$ ensemble.
+\ref{T1}-a) and b) show the change in photoluminescence (PL) with respect to an external magnetic field for two samples with a low and high concentration of NV$^-$ centers (see SI).
+Both samples show a decrease in PL as the magnetic field amplitude increases.
+There is however a stark difference in the low magnetic field region where only high-density samples shows a drop in PL (CITATION).
+This effect can be observed on all our samples (see SI) whenever the NV concentration lies in the ppm range.
 
-Unlike the decrease in PL for higher magnetic field values, the drop in PL at low magnetic field is associated with a decrease of the NV's spin lifetime $T_1$.
-Indeed, since the $\ket{m_s=0}$ spin state is brighter than the $\ket{m_s=\pm 1}$ states, and since the optical pumping in the $\ket{m_s=0}$ state is in competition with the relaxation in a thermal distribution of the states, decreasing the spins lifetime results in a decrease of the NV PL \citep{finco2021imaging}.
-To understand the behavior of the PL in zero-field, we therefore need to study the dynamics of the spin in the low magnetic field region.
+The drop of the PL at low magnetic field is generally associated with a decrease of the NV's spin lifetime $T_1$ (CITE JARMOLA), which results in a decrease of the population in the bright state $0$.
+%Indeed, since the $\ket{m_s=0}$ spin state is brighter than the $\ket{m_s=\pm 1}$ states, and since the optical pumping in the $\ket{m_s=0}$ state is in competition with the relaxation in a thermal distribution of the states, decreasing the spins lifetime results in a decrease of the NV PL \citep{finco2021imaging}.
+%To understand the behavior of the PL in zero-field, we therefore need to study the dynamics of the spin in the low magnetic field region.
+The depolarization dynamics of the spin of single or dilute NV centers at room temperature is dominated by two-phonon Raman processes \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depend on the crystal lattice temperature.
+It has been observed that dense NV centers ensemble have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi_depolarization_2017, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field amplitude.
+This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi_depolarization_2017}.
+Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble.
+We will denote $T_1^{\rm ph}$ the characteristic timescale associated with the phonon relaxation process and $T_1^{\rm dd}$ the density dependent timescale associated with the dipole-dipole relaxation process.
 
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig T1}
-\caption{Spin $T_1$ measurement protocol.
-(a) Sequence used to measure the spin lifetime : the laser is turned off for a time $\tau$ twice and the spin state is read at the end of the dark time by measuring the initial photoluminescence when the laser is turned back on.
-A microwave $\pi$-pulse resonant with one of the spin transition is placed before the second reading.
-The  $T_1$ sequence result is the subtraction $S_1-S_2$.
-(b) Measurement of $S_1-S_2$ for a high density of NV centers in zero magnetic field field and non-zero magnetic field.
-The fit (orange plain line) is detailed in the main text}
+\includegraphics[width=0.45\textwidth]{Figures/shema_summary.pdf}
+\caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation. }
 \label{T1}
@@
 
-The $T_1$ depolarization dynamics of single or dilute NV centers at room temperature is dominated by a two-phonon Raman process \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depends on the crystal lattice temperature but does not depend on the external magnetic field.
-However, it has been observed that dense NV centers ensemble have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi_depolarization_2017, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field and not on the temperature .
-This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi_depolarization_2017}.
-Some inhomogeneity between the NV centers is further needed in order to explain the depolarization of the spin ensemble.
-We will denote $T_1^{\rm ph}$ the characteristic time associated with the phonon relaxation process and $T_1^{\rm dd}$ the characteristic time associated with the dipole-dipole relaxation process.
-
-The spin lifetime protocol used here is described in Fig.
-\ref{T1} and is based on previous similar experiments \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi_depolarization_2017}.
-It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$.
-Using only this sequence can result in artifacts in the signal, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019}.
-It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to project the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
-By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures/fig_T1.pdf}
+\caption{a) Photoluminescence measurement from NV centers ensemble as a function of an external magnetic field applied in an arbitrary direction (a) for a CVD sample containing $\approx4$ ppb of NV$^-$ centers, (b) for an HPHT sample containing $\approx3$ ppm of NV$^-$ centers.
+(c) Sequence used to measure the spin lifetime.
+(d) Spin relaxation $S_1-S_2$ measured for the dense sample at zero and non-zero magnetic fields.
+The fitting procedure (orange plain line) is detailed in the main text}
+\label{T1}
+\end{figure}
 
 \begin{figure*}
-\includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
-\caption{Dependency on the magnetic field angle for the zero field depolarization.
-(a) ODMR spectrum in zero field.
-(bi) ODMR spectrum for a magnetic field $\approx$ 60 G, misaligned by $\sim  24^\circ$ with the [100] axis.
-(bii) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude.
-(biii) Stretch part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude.
-(ci), (cii) and (ciii) : same measurement with a magnetic field aligned with the [100] axis}
+\includegraphics[width=.95\textwidth]{Figures/fig_100_vs_1x1x1x1.pdf}
+\caption{(a) ODMR spectrum in zero field.
+(b)-i) ODMR spectrum for a magnetic field $\approx$ 60 G, misaligned by $\sim  24^\circ$ from the [100] axis.
+(b)-ii) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude.
+(b)-iii) Stretched part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude.
+(c)-i), (c)-ii) and (c)-iii) : same measurements but with a magnetic field close to the [100] axis}
 \label{100_VS_1x4}
@@
 
-The ensemble spin decay is non-exponential due to inhomogeneities between the NV$^-$ centers.
+Fig. \ref{T1}-c) shows the sequence employed for measuring $T_1^{\rm dd}$ in the dense sample. 
+%The spin lifetime protocol used here is described in Fig. \ref{T1} and is based on previous similar experiments 
+It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$.
+In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019}.
+It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
+By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi_depolarization_2017}.
+
+Fig. \ref{T1}-b) shows the subtracted signals when $\bm B=0$ and $\bm B \neq 0$.
+The ensemble spin decay curves are not exponential, which was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi_depolarization_2017}.
 We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}.
-A conclusion of this model is that, for an homogeneous 3D distribution of fluctuator, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$.
-Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$  (see SI).
+A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$.
+Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$.
+For the samples used in this paper (detailed in SI) $T_1^{\rm dd} \sim T_1^{\rm ph}$, so both decay processes have to included in the analysis.
+Fitting the two curves by the product of an exponential and a stretched exponential, we find $T_1^{\rm ph}=3.2$ ms and $T_1^{\rm dd}= ??$ here.
+We also conducted further measurements that confirm the lifetime-limited character of the fluctuator employed in the model \citep{choi_depolarization_2017}, see SI.
+%These results show that our sample clearly manifests dipolar interactions amongst closely packed NV centers, but do not tell about the magnitude of the several possible mechanisms involved in the dipolar relaxation.
+%In spin-1 systems, there can indeed be several channels for decay at zero fields such as flip-flop, double flip,
 
-For the samples used in this paper (detailed in SI), we observed that $T_1^{\rm dd} \sim T_1^{\rm ph}$, meaning that we had to include both  in our analysis.
-The resulting fitting formula is therefore :
+%The resulting fitting formula is therefore :
+%\begin{equation}
+%S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}}),
+%\end{equation}
+%Further details on the fitting procedure are available in SI. For the rest of the article we will be only interested on the stretched exponential lifetime ${T_1^{\rm dd}}$.
+
+One of the ways to tune dipole-dipole related phenomena is by changing the number of resonant spins.
+Fig.
+\ref{100_VS_1x4} (b)-i) shows an optically detected magnetic resonance (ODMR) spectrum from the dense ensemble.
+The observed lines correspond to the projections of the magnetic fields onto the four "classes" of NV centers - that is physical orientation of the NV axis is the diamond crystal cell - together with the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$.
+The transitions frequencies of the four different classes can be controlled by changing the amplitude and orientation of the external magnetic field.
+Fig.
+\ref{100_VS_1x4} (c)-i) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis.
+%Other example of class resonances are given in the SI.
+This configuration will be essential to decipher what dipolar relaxation mechanisms play a role at low field.
+
+Fig. \ref{100_VS_1x4} (b)-ii) and (b)-iii) show the PL from an ensemble of NV centers as a function of the amplitude of a magnetic field aligned as in Fig. \ref{100_VS_1x4} (b)-i). 
+We also extract the stretched exponential spin decay time $T_1^{\rm dd}$ by running the sequence shown in Fig. 2-c) at each of the XX magnetic field values. 
+We observe that the 3 \% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.05 kHz, as the magnetic field increases.
+This result can be explained by the higher number of resonant spins in zero magnetic field, where all four classes are degenerate, as opposed when $\bm B \neq 0$.
+The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SI ?).
+Note that the decrease in PL for $B>40\ \rm G$ are related to state mixing by the transverse magnetic field, so it is not correlated to a modification of the spin lifetime.
+
+Importantly, we observed that the lifted degeneracy of the four NV classes is not the sole reason for the increase of the spin lifetime as the B field increases.
+Fig.
+\ref{100_VS_1x4} (c)-ii) and (c)-iii) show the change in the PL and in $1/T_1^{\rm dd}$ as a function of a magnetic field that is aligned along the [100] axis.
+For this particular orientation, the four classes are always resonant, regardless of the field amplitude.
+We can see that there still is an increase in the spin decay rate and a corresponding drop in the PL under low magnetic field values, although considerably reduced.
+Also, the rate of change in both the PL and $T_1^{\rm dd}$ is larger by more than a factor of 3 in the [100] configuration.
+Although the contrast is smaller, this effect can thus drastically impact zero-field magnetometry protocols based on CR.
+The main aim of this paper is to understand the physics governing the contribution to the spin decay in this second scenario.
+Note that the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20$G is related to dipolar interaction with NV centers that have a $^{13}$C as a first neighbor \cite{pellet2021optical}.
+
+We identified and isolated two additional mechanisms that could explain this effect : local electric fields and double-flip processes. 
+The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
 \begin{equation}
-S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}}),
+\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
 \end{equation}
-Further details on the fitting procedure are available in SI. For the rest of the article we will be only interested on the stretched exponential lifetime ${T_1^{\rm dd}}$.
+Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$.
+In situations such as the one described in Fig.XX, the only relevant terms of $\mathcal{H}^{\rm dd}$ in terms of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$.
+In zero magnetic field however, we have to take into account these two other mechanisms of the dipolar Hamiltonian in consideration.
+%where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.
+It was shown in  \cite{mittiga2018imaging} that local electric fields coming from charge traps such as $P_1^+$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in large NV ensembles.
+Magnetic field noise comes in to second order in zero field.
+%We also do not take into account the hyper-fine structure of the NV center due to the large inhomogeneous broadening of the ESR transitions (see SI).
+We will then consider the following spin Hamiltonian for the electric field dependent NV$^-$ ground state : 
+\begin{equation}
+\label{NV Hamiltonian}
+\mathcal{H}_{\rm elec}= d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
+\end{equation}
+where $E_{x,y}$ are the projections of local electric fields on the NV axes. 
+Under local electric fields with orientations given by the angle $\phi_E={\rm tan}(E_x/E_y)$, the eigenstates of $H_{NV}$ are $\ket{0}$ and $\ket{\pm}=\frac{1}{\sqrt{2}}(\ket{+1}\pm e^{-i\phi_E}\ket{-1})$.
+Computing the flip-flop terms $\ket{\pm,0} \bra{0,\pm} $ in the fluctuator model shows that $1/T_1^{dd}$ should be increased at zero field, even after averaging over all possible orientations $\phi_E$ (see SI).
+Another mechanism at stake when considering dipolar interactions with spin-1 systems are double-flip processes.
+These processes correspond to terms like $\ket{+1,0} \bra{0,1}$ in the anisotropic dipolar interaction, giving rise to an exchange of two units of spin-angular momentum.
+These processes can become significant when the states $\pm 1$ are degenerate and are thus likely to explain part of the zero-field dip in Fig.
+c)-ii).
+%Theoretical model predicts a factor of $\approx $5 increase in $1/T_1^{dd}$ at zero field.
 
-%\begin{figure}
-%\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
-%\caption{Tuning resonant coupling through magnetic field orientation.
-(a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice.
-(b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes.
-The detuning $\Delta \nu$ between the classes $\beta$ and $\gamma$ was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_\gamma$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz .
-Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
-%\label{largeur_fluct}
-%\end{figure}
+In order to discriminate the role of these two effects in the low-field region, we performed another similar study under a purely transverse magnetic field.
+This regime in fact emulates the effect of the electric field while turning off the resonance condition for double flip processes : indeed for relatively weak transverse magnetic field, the eigen-basis of the spin Hamiltonian is close to $\ket{0},\ket{+},\ket{-}$.
+This property has been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
+Fig.~\ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian as a function of a purely transverse magnetic field.
+We will denote these eigenstates in the general case $\ket{g}$, $\ket{d}$ and $\ket{e}$.
+For weak transverse magnetic field, $\ket{g} \approx \ket{0}$, $\ket{d}=\ket{-}$ and $\ket{e} \approx \ket{+}$.
+As the magnetic field increases, $\ket{g}$ and $\ket{e}$ starts to become a mixing of $\ket{0}$ and $\ket{+}$ while $\ket{d}$ remains equal to $\ket{-}$.
 
-%Fig.
-\ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins.
-Because of the diamond symmetry, we can distinguish 4 groups ("classes") of NV centers, based on the four possible orientation of the NV axis in the crystal.
-These four classes have \textit{a priori} different spin resonance frequency based on the projection of the magnetic field on their respective NV axis.
-By tuning the angle of the magnetic field, we can therefore control the frequency detuning between the different classes, and as a result we can control the number of resonant NV centers by bringing in and out of resonance several classes.
-%
-%Fig.
-\ref{largeur_fluct} (c) in particular shows the dipole-induced lifetime $T_1^{\rm dd}$ for a given class as a function of the energy detuning with another class.
-One can notice that the width of the $T_1^{\rm dd}$ bell shape is significantly larger than that of the frequency overlap between the two classes, modeled here by an ODMR line stretched by a factor $\sqrt{2}$ (see SI).
-This extra width can not be attributed to the dipole-dipole coupling strength : for a 3 ppm concentration of NV$^-$ centers, the average coupling strength between neighboring NV centers should be $\sim 27\ \rm kHz$, two order of magnitude below the observed width.
-We can also see that $T_1^{\rm dd}$ follows a Lorentzian shape despite the ODMR lines being non-Lorentzian.
-These two observations corroborate the idea developed in \cite{choi_depolarization_2017} that the interaction width is dominated by the noise spectrum of the lifetime-limited fluctuators.
-
-One of the way we can isolate dipole-dipole related phenomena with NV centers is by tuning the number of NV centers resonant with each other.
-Indeed, Fig.
-\ref{100_VS_1x4} (bi) shows the characteristic 8 lines observed with ensemble NV optically detected magnetic resonance (ODMR) spectrum.
-These 8 lines correspond to the four "classes" of NV centers - that is physical orientation of the NV axis is the diamond crystal cell - multiplied by the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$.
-The transitions of the four different classes can be controlled somewhat independently by changing the amplitude and orientation of the external magnetic field, and in particular they can be brought to resonance for particular orientation of the magnetic field.
-Fig.
-\ref{100_VS_1x4} (ci) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis.
-Other example of class resonance are given in SI.
+Crucially, mixing remains even when the associated energy splitting of the $\ket{d}$ and $\ket{e}$ states is large : Fig.
+\ref{B_transverse}-(b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$.
+We can see that at a field value of $\sim 130$ G, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50$ MHz, far exceeding the range of the dipole-dipole resonant coupling (see details in SI), but $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
+We can therefore decouple the effects of the double-flip processes  $\propto \ket{g,d}\bra{d,e}$, which are no longer resonant when $| \bm B | > 130$ G, from the effects of the mixing induced by the transverse electric - or in this case magnetic - field.
 
 Fig.
-\ref{100_VS_1x4} (bii) and (biii) shows respectively the PL and the stretched exponential spin decay $T_1^{\rm dd}$ of an ensemble of NV centers with respect to the amplitude of the external magnetic field.
-The orientation of the magnetic field was chosen in order to lift the degeneracy between the four classes, as shown in Fig.
-\ref{100_VS_1x4} (bi).
-We observe that the drop in PL at low magnetic field is indeed correlated with an increase in the spin decay rate, as is expected due to the higher number of resonant spins in zero magnetic field, where all four classes are resonant, compared to the non-degenerate case for $\vec \neq 0$.
-We can also see that the decrease in PL for $B>40\ \rm G$ due to state mixing by the transverse magnetic field is not correlated to a modification of the spin lifetime.
+\ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers exposed to a transverse magnetic field between 25 (??) and 130 G (FOR A DIFFERENT SAMPLE...
+which ?
+we observe a $1/T_1^{\rm dd}$ of $30$ Hz under large longitudinal magnetic fields for this class ?).
+The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
+We attribute the decrease in the spin decay rate when the detuning increases to the loss of effectiveness of the double-flip processes.
+We can notice that the decay rate reaches a plateau for $\Delta \nu \geq 30\ \rm MHz$ with a value about twice as large as the decay rate for a longitudinal magnetic field.
+This plateau we attribute to the effect of the transverse magnetic field, which we assume to be similar to the effect of the electric field in the low magnetic field regime.
+We should note that the effect of the double-flip processes is about 4 times more important in the depolarization rate than the effect of the electric field, even though the states are not initially fully resonant.
 
-However, the splitting of the four classes is not sole reason to the decrease of the spin lifetime in low magnetic field.
-Indeed, we performed similar measurements in Fig.
-\ref{100_VS_1x4} (cii) and (ciii) but this time fixing the magnetic field orientation along the [100] axis.
-For this particular orientation, the four classes are always resonant, regardless of the field amplitude.
-We can see that, although considerably reduced, there still is an increase in the spin decay rate and a corresponding drop in PL for low magnetic field values.
-We attribute the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20\ \rm G$ to dipolar interaction with NV having a first neighbor $^{13}C$ \cite{pellet2021optical}.
-
-The aim of this paper is to understand the remaining contribution to the spin decay in this second case.
-We identified and isolated two additional effects : the domination by the local electric field for low magnetic field value which results in a change of the spin Hamiltonian eigenstates, and the presence of double-flip processes where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.
-
-In order to distinguish these two contributions, we decided to simulate the effect of the electric field by using a transverse magnetic field : for relatively weak transverse magnetic field, the eigen basis of the spin Hamiltonian is close to $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $, which is the the spin Hamiltonian basis when the electric field dominates, as detailed in SI.
-This property has been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
+Based on an extension of the fluctuator model developed in SI, we predict an overall increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$.
+While the measured increase is smaller than the predicted one, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.
 
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig transverse field V2}
-\caption{(a) Simulated energies for the spin Hamiltonian in the presence of pure transverse magnetic field $B_\perp$.
+\includegraphics[width=0.45\textwidth]{Figures/fig_transverse_field_V2.pdf}
+\caption{(a) Energies for the spin Hamiltonian in the presence of pure transverse magnetic field $B_\perp$.
 The three spin levels are denoted $\ket{g}$, $\ket{d}$ and $\ket{e}$.
-(b) Blue dashed curve : frequency detuning $\Delta \nu$ between the $\ket{g}$, $\ket{d}$ states.
+(b) Blue dashed curve : frequency detuning $\Delta \nu$ between the $\ket{e}$ and $\ket{d}$ states.
 Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\bra{e}$ and $\bra{+}$.
@@
 \end{figure}
-
-Fig.
-\ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian with respect to a purely transverse magnetic field.
-We will denote these eigenstates in the general case $\ket{g}$, $\ket{d}$ and $\ket{e}$.
-For weak transverse magnetic field, $\ket{g} \approx \ket{0}$, $\ket{d}=\ket{-}$ and $\ket{e} \approx \ket{+}$.
-As the magnetic field increases, $\ket{g}$ and $\ket{e}$ starts to become a mixing of $\ket{0}$ and $\ket{+}$ while $\ket{d}$ remains equal to $\ket{-}$.
-
-Crucially, this mixing happens on a different scale than the associated energy splitting of the $\ket{d}$ and $\ket{e}$ states : Fig.
-\ref{B_transverse} (b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$.
-We can see that for a field value of $\sim 130 G$, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50 MHz$, far exceeding the range of the dipole-dipole resonant coupling (see details in SI), but $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
-We can therefore decouple the effects of the double-flip processes, which are no longer resonant, from the effects of the change of basis induced by the transverse electric - or in this case magnetic - field.
-
-Fig.
-\ref{B_transverse} (c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers exposed to a transverse magnetic field between 25 and 130 G.
-The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
-We attribute the decrease in the spin decay rate when the detuning increases to the loss of effectiveness of the double-flip processes.
-We can notice that the decay rate reaches a plateau for $\Delta \nu \geq 30\ \rm MHz$ with a value about twice as big as the decay rate for a longitudinal magnetic field.
-This plateau we attribute to the effect of the transverse magnetic field, which we assume to be similar to the effect of the electric field in the low magnetic field regime.
-We should note that the effect of the double-flip processes is about 4 times more important in the depolarization rate than the effect of the electric field, even though the states are not initially fully resonant.
-
-Overall we predicted, based on an extension of the fluctuator model developed in SI, an increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$.
-While the measured increase is smaller than the predicted one, as with most of the predictions based on the fluctuator model, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.
-
-Similarly, we can compute the predicted decay rate in the $\vec B \parallel \left[100\right]$ case compared to the $\vec B = 0$ case, and find a theoretical increase of $\sim 20\%$ in the $\vec B = 0$ case due to the effects of the local electric field.
+Similarly, we compute the predicted decay rate when $\bm B \parallel \left[100\right]$ case compared when $\bm B = 0$, and find a theoretical increase of $\sim 20\%$ when $\bm B = 0$ due to the mixing by the local electric field.
 This, along with the double-flip processes explains the behavior observed in Fig.
-\ref{100_VS_1x4} (cii) and (ciii).
+\ref{100_VS_1x4} (c)-ii) and (c)-iii).
 
@@
 %First is the change of basis of the single spin Hamiltonian $\mathcal{H}_s$ : in zero external magnetic field, the eigenstates of $\mathcal{H}_s$ in the $\{\ket{+1},\ket{-1}\}$ manifold are determined by local electric and magnetic field coming from nearby impurities \cite{mittiga2018imaging}.
-For the samples used in this study, $\mathcal{H}_s$ was dominated by local electric field (see SI), meaning that the proper eigenstates of $\mathcal{H}_s$ in zero magnetic field are $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $
-%
+For the samples used in this study, $\mathcal{H}_s$ was dominated by local electric field (see SI), meaning that the proper eigenstates of $\mathcal{H}_s$ in zero magnetic field are $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $%
 %Because the splitting between $\ket{+}$ and $\ket{-}$ ($\approx 9\ \rm MHz$) is much greater than the dipole-dipole interaction ($J_0 / r^3 \approx 30\ \rm kHz$), the flip-flop terms of $\mathcal{H}^{\rm dd}$ now reads $\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}$.
@@
 
-\section*{Magnetometry}
 \begin{figure}
@@
 %
-DC microwave-less magnetometry has already been performed with NV ensembles, using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}.
-Here we propose to perform a similar protocol, but using the spin depolarization in zero-field.
-The main difference with the previously mentioned protocols is the fact that this one doesn't rely on crystalline orientation, making this protocol usable with diamond powders or polycrystalline samples.
-The idea of performing optical magnetometry on the zero-field photoluminescence feature of NV ensembles has been suggested before \cite{filimonenko2018weak, filimonenko2022manifestation} but had not been implemented so far.
+
+Our observations have important implications for magnetometry with NV ensembles.
+DC microwave-free magnetometry has already been performed using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}.
+Here we propose to perform a similar protocol , but using the spin depolarization at zero-field as proposed in \cite{filimonenko2018weak, filimonenko2022manifestation}.
+The main advantage with the previously mentioned protocols is the fact that the sensitivity does not depend crucially on crystalline orientation, making this protocol applicable with diamond powders or polycrystalline samples.
 
 Fig.
-\ref{magneto} (a) shows the demodulated PL when a DC magnetic field is scanned in a random direction and an additional alternating magnetic field of a few Gauss at $\sim \ \rm kHz$ is added through the same electromagnet (see experimental details in SI).
-We can see a sharp and relatively linear slope in low field $\abs{B} < 5\ \rm G$.
-Once calibrated, in this case with ODMR, the slope can provide a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as seen in \cite{zheng_microwave-free_2020}.
-
-In order to measure the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take an histogram of the measured fields, as shown in Fig.
+\ref{magneto} (a) shows the demodulated PL as a DC magnetic field is scanned in a random direction and an additional alternating magnetic field of a few Gauss at $\sim \ \rm kHz$ is added through the same electromagnet (see experimental details in SI) WHICH SAMPLE ?
+.
+We can see a sharp linear slope in low field $\abs{B} < 5\ \rm G$.
+Once calibrated, in this case with ODMR, the slope can provide a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as in \cite{zheng_microwave-free_2020}.
+In order to assess the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take a histogram of the measured fields, as shown in Fig.
 \ref{magneto} (b) and (c).
 The histogram is well fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
-The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which allows us to measure the sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nT}/\sqrt{\rm Hz}$.
+The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which gives us a sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nT}/\sqrt{\rm Hz}$.
 
-We can then try to measure the relative importance of the three causes of spin depolarization on the field sensitivity, namely the splitting of the classes, the local electric field and the double-flip processes.
-In order to do so, we measure the sensitivity while changing the angle of the field.
-When $\vec B$ is aligned with the [100] crystalline axis, only the double flips and the electric field cause a depolarization, whereas in every other orientation the three effects are at play.
-
+We now evaluate the relative role of the three causes of spin depolarization in the magnetometer, namely the splitting of the classes, the local electric field and the double-flip processes.
+In order to do so, we measure the magnetometer sensitivity while changing the angle of the magnetic field.
 The results are shown in Fig.
 \ref{magneto} (d).
-We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity remains relatively flat, which means that the double-flips and electric field effects are the dominant factors in the sensitivity of this protocol.
-While it may seem surprising that the effects with a lower contribution on the PL contrast have a higher effect on this PL-based protocol, what matters here is not the absolute contrast but the slope of the change of contrast with respect to the magnetic field, which is sharper for the electric field and double flips than it is for the lift of degeneracy.
+We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity remains relatively flat.
+When $\bm B$ is aligned with the [100] crystalline axis, only the double flips and the electric field cause a depolarization, whereas in every other orientation the three effects are at play.
+The double-flips and electric field effects are thus the dominant factors in the sensitivity of this protocol.
+While it may seem surprising that the effects with a lower contribution on the PL contrast have a higher effect on this PL-based protocol, what matters here is not the absolute contrast but the slope of the change of contrast with respect to the magnetic field.
+The latter is in fact larger for the local electric field and double flips processes than for the process involving only flip-flops.
 %This is in agreement with the measurements in Fig.
 \ref{100_VS_1x4} : even though the contrast is much better in the not-[100] case, the slope in term of contrast percentage/magnetic field is comparable in both cases (en faite pas complètement, y'a pas loin d'un facteur 2 sur les dérivées numériques)
-It should be noted though that this observation is sample dependent, and that other samples, including from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from the electric field and double-flip processes.
+It should be noted that this observation is sample dependent, and that other samples, including from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from the electric field and double-flip processes.
 
-\section*{Conclusion}
-In this work we have identified three mechanisms causing the extra spin depolarization observed in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole cross-relaxations between the spins : the lift of degeneracy between the four classes caused by the magnetic field, the domination of the local electric field which causes a change in the Hamiltonian eigenstates, and the double-flip process allowed by the proximity in energy of the spin $\ket{\pm}$ states.
-We identified the lift in degeneracy as the main cause in the zero-field depolarization, followed by the double-flip processes and then the electric field.
-We have demonstrated a potential use of this depolarization as a DC microwave-less and orientation-free magnetometer with sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal.We have observed that in some cases the double-flips and electric field play a more important role than the lift of degeneracy in the magnetometry sensitivity.
+As a conclusion, we identified three mechanisms causing extra spin depolarization in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole induced cross-relaxations between the spins of NV centers.
+The lift in degeneracy of the spin state of different NV classes was found to be the main cause of zero-field depolarization, followed by double-flip processes and then the electric field induced mixing.
+We have employed cross-relaxation for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal and show that that double-flips and electric field play an important role.
+Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins. 
+%DEER, many body, critical thermalization, 2D. Double quantum
+
 \section*{acknowledgements}

### 2022-06-21 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
tutu

@@
-\documentclass[preprintnumbers,amsmath,amssymb,onecolumn,12pt]{revtex4-2}
-\usepackage{graphicx}% Include figure files
+\documentclass[preprintnumbers,amsmath,amssymb,onecolumn,12pt]{revtex4-2}\usepackage{graphicx}% Include figure files
 \usepackage{dcolumn}% Align table columns on decimal point
@@
 \
-\title{Low field depolarization of NV$^-$ centers through dipole-dipole coupling}
+\title{Low magnetic field depolarization of NV$^-$ centers through dipole-dipole interaction}
 
-\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, G. H\'etet$^1$}
+\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, P. Huillery$^?$,  G. H\'etet}
 
@@
 \label{NV Hamiltonian}
-\mathcal{H}_s=D S_z^2 + \gamma_e \vec{B}_{\rm ext} \cdot \vec{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
+\mathcal{H}_s=D S_z^2 + \gamma_e \bm{B}_{\rm ext} \cdot \bm{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
 \end{equation}
@@
 In both case we use a modulation frequency $\sim 1\ \rm kHz$ and demodulate the APD signal.
+
+We did not use magnetic shielding to protect from the earth magnetic field since the splitting due to  $B_{\rm earth}$ is lower than the splitting due to the local electric field for the samples used here : $ \gamma_e B_{\rm earth} \approx 1.5\ \rm MHz < d_\perp E_\perp \approx 4\ \rm MHz$.
+When possible, the electromagnet was roughly aligned along the earth magnetic field to better compensate it.
 \section{Experimental details}
@@
 
-In order to directly compare the results from different magnetic fields, we therefore decided to fit each $T_1$ measurement with both a stretched and exponential decay : $S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$.
-To add consistency to the measurement of $T_1^{\rm dd}$, we decided to fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ and the signal amplitude as the only free parameters in the fitting procedure.
+In order to directly compare the results from different magnetic fields, we therefore fit each $T_1$ measurement using both a stretched and exponential decay : $S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$.
+To add consistency to the measurement of $T_1^{\rm dd}$, we fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ and the signal amplitude as the only free parameters in the fitting procedure.
 The value chosen for samples [REF] and [REF] was $T_1^{\rm ph}=3.62\ \rm ms$.
@@
 Fig.
-\ref{largeur_fluct} shows the results of such an experiment where we measured the stretched part of the NV's decay rate, which is very well fitted by a Lorentzian of width $\sigma=8.04 MHz$, and compare it to an ODMR line of a single class of NV centers, stretched by a factor of $\sqrt{2}$ and 2 to simulate the spectral overlap of the two classes (discussed bellow).
+\ref{largeur_fluct} shows the results of such an experiment where we measured the stretched part of the NV's decay rate, which is very well fitted by a Lorentzian of width $\sigma=8.04 MHz$ (FW or HW ?), and compare it to an ODMR line of a single class of NV centers, stretched by a factor of $\sqrt{2}$ and 2 to simulate the spectral overlap of the two classes (discussed bellow).
 One should notice that, not only is $1/T_1^{\rm dd}$ significantly larger than the ODMR profile, it also doesn't have the same shape.
@@
 \section{Extension of the fluctuator model to low magnetic fields}
+
+\subsection{Results}
+
+Here are the results of the fluctuator model developed in \cite{choi_depolarization_2017} that we extended to include more scenarios.
+The following sections will detail the calculations that led to these results.
+\begin{itemize}
+\item The dipole-induced spin decay is at its lowest for a class of NV centers, non resonant with the three other classes, and dominated by a longitudinal magnetic field.
+In this case the dipole-dipole relaxation is limited by flip-flop processes between spins of the same class.
+We will call the corresponding stretched lifetime $T_1^{\rm dd}\equiv T_0$.
+\item For a single class, still spectrally isolated   from the three other classes, but this time dominated by the electric field or transverse magnetic field (as long as $\gamma_e B_\perp \ll D$), the change in the eigenstates of the single particle Hamiltonian results in an increase of the average flip-flop rate.
+This leads to a new theoretical stretched lifetime $T_1^{\rm dd}= 4\, T_0$.
+\item When a magnetic field is applied along the [100] crystalline axis (for $5\ \rm G < \abs{B} \ll D$), all four classes are resonant and dominated by the longitudinal part of the magnetic field.
+This results, in increase in the average flip-flop rate compared to the isolated class case, and the predicted stretched exponential lifetime is $T_1^{\rm dd}\approx 42.8\, T_0$.
+\item In zero external magnetic field, all four classes are also resonant but we assume that the single particle Hamiltonian of each spin is dominated by a random local electric field.
+The averaging of flip-flop in this case yields a stretched lifetime $T_1^{\rm dd}\approx 51.4\, T_0$, which is about 20\% higher than the predicted lifetime for $B \parallel$ [100].
+\end{itemize}
+
+The effect of the double flip processes was not studied quantitatively since the process is never fully resonant and is therefore highly dependent on the initial splitting of the $\ket{+}$ and $\ket{-} $ states.
+Nevertheless, the presence of double-flip as an additional relaxation channel can only increase the dipole-induced spin decay rate.
+
 \subsection{Summary of the NV-fluctuator model}
-As mentioned in the main text, we base our analysis on the NV-fluctuator model developed in \cite{choi_depolarization_2017} whose main hypothesis and conclusions we will summarize here :
+Here are the main hypotheses and conclusion of the NV-fluctuator model developed in \cite{choi_depolarization_2017} :
 
@@
 
-We will consider the spin operators $\vec{S}_1$ and
-$\vec{S}_2$ in the NV own basis : the $z$ orientation of the $S_z$ operator is chosen along the NV axis of each spin.
-In the presence of a longitudinal magnetic field, the sense of $\hat{z}$ will be chosen such that $\vec{B}\cdot \hat{z} >0$, and the $\hat{x}$ direction is chosen along the transverse magnetic or electric field, whichever dominates the single spin Hamiltonian.
+We will consider the spin operators $\bm{S}_1$ and
+$\bm{S}_2$ in the NV own basis : the $z$ orientation of the $S_z$ operator is chosen along the NV axis of each spin.
+In the presence of a longitudinal magnetic field, the sense of $\hat{z}$ will be chosen such that $\bm{B}\cdot \hat{z} >0$, and the $\hat{x}$ direction is chosen along the transverse magnetic or electric field, whichever dominates the single spin Hamiltonian.
 These choices mean that there will be two distinct Cartesian basis $\{\hat{x}_1,\hat{y}_1,\hat{z}_1\}$ and $\{\hat{x}_2,\hat{y}_2,\hat{z}_2\}$ describing the two spins.
@@
 \begin{align}
--\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}&= 3\left({\vec S}_1 \cdot \hat{u} \right)\left({\vec S}_2 \cdot \hat{u} \right) - {\vec S}_1 \cdot {\vec S}_2  \\
+-\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}&= 3\left({\bm S}_1 \cdot \hat{u} \right)\left({\bm S}_2 \cdot \hat{u} \right) - {\bm S}_1 \cdot {\bm S}_2  \\
 &=\left[3(\hat{u}\cdot\hat{x}_1)(\hat{u}\cdot\hat{x}_2) -\hat{x}_1\cdot\hat{x}_2\right] S_x^1S_x^2 \\
@@
 \end{align}
-Where $J_0=\frac{\mu_0 \gamma_e^2 \hbar^2}{4 \pi} = (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r= r\vec u$ is the distance between the two spins and  $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
+Where $J_0=\frac{\mu_0 \gamma_e^2 \hbar^2}{4 \pi} = (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\bm r= r\bm{\hat u}$ is the distance between the two spins and  $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
 
@@
 This numerical factor is linked to the previously mentioned $\bar \eta$ by averaging over every possible orientation and angular position of the fluctuators.
-Assuming that the fluctuators are evenly distributed among all four classes (NV axis orientation) and decomposing the $\vec r$ vector in the spherical basis ($r, \theta, \phi$), we can write $\bar \eta$ as :
+Assuming that the fluctuators are evenly distributed among all four classes (NV axis orientation) and decomposing the $\bm r$ vector in the spherical basis ($r, \theta, \phi$), we can write $\bar \eta$ as :
 
@@
 \label{eq. eta bar}
-\bar \eta =\frac{1}{4} \sum_{i=1}^4 \int_\theta \int_\phi  \int_{\omega_f} \int_{\omega_{NV}}   \abs{\eta(\vec u,i,\omega_{NV},\omega_f)}\, d\Omega\, \rho(\omega_{NV}) d\omega_{NV}\, \rho^i(\omega_f) d\omega_f
+\bar \eta =\frac{1}{4} \sum_{i=1}^4 \int_\theta \int_\phi  \int_{\omega_f} \int_{\omega_{NV}}   \abs{\eta(\bm u,i,\omega_{NV},\omega_f)}\, d\Omega\, \rho(\omega_{NV}) d\omega_{NV}\, \rho^i(\omega_f) d\omega_f
 \end{equation}
@@
 \begin{figure}
-\includegraphics[width=0.7\textwidth]{Figures_SI/121 VS 22}
+\includegraphics[width=0.7\textwidth]{Figures_SI/121_VS_22}
 \caption{(a) Geometrical representation of two NV classes and the two possible planes where the magnetic field has the same projection on both classes.
-The positive $\hat z_i$ direction is chosen so that $\vec{B}\cdot \hat{z}_i >0$.
+The positive $\hat z_i$ direction is chosen so that $\bm{B}\cdot \hat{z}_i >0$.
 (b) $T_1$ measurement on a two-class resonance following the protocol described in main text.
@@
 Inset : zoom-in on the 0.3-0.8 ms region.}
-\label{121 VS 22 fig}
+\label{ VS 22 fig}
 \end{figure}
@@
 \item The NV and fluctuator are resonant and from the same class (the angle $\widehat{z_1 z_2}=0$) : this is always true, regardless of the magnetic field.
-We can analytically compute $\bar \eta$ in this case and find : $$ \bar {\eta}_{same}=\frac{1}{4}  \int_\theta \int_\phi \abs{\eta(\vec u,\widehat{z_1 z_2}=0,\omega_{NV}=\omega_f)}\, d\Omega= \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot \frac{2}{3 \sqrt{3}} \approx 5.55\cdot10^{-2}$$
+We can analytically compute $\bar \eta$ in this case and find : $$ \bar {\eta}_{same}=\frac{1}{4}  \int_\theta \int_\phi \abs{\eta(\bm u,\widehat{z_1 z_2}=0,\omega_{NV}=\omega_f)}\, d\Omega= \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot \frac{2}{3 \sqrt{3}} \approx 5.55\cdot10^{-2}$$
 \item The NV and fluctuator are resonant and the angle $\widehat{z_1 z_2}=\rm{arccos}(\frac{1}{3})\approx 70.5^\circ$.
@@
 Fig.
-\ref{121 VS 22 fig}(a) shows a graphical representation of the difference between the two last scenarios : due to the $\vec{B}\cdot \hat{z} >0$ condition, the same two classes of NV centers can have a $\widehat{z_1 z_2}$ angle equal to $70.5^\circ$ or $109.5^\circ$ depending on the external magnetic field.
+\ref{121 VS 22 fig}(a) shows a graphical representation of the difference between the two last scenarios : due to the $\bm{B}\cdot \hat{z} >0$ condition, the same two classes of NV centers can have a $\widehat{z_1 z_2}$ angle equal to $70.5^\circ$ or $109.5^\circ$ depending on the external magnetic field.
 
@@
 \begin{figure}
-\includegraphics[width=0.7\textwidth]{Figures_SI/Various ESR}
+\includegraphics[width=0.7\textwidth]{Figures_SI/Various_ESR}
 \caption{ODMR spectra for different orientations of the magnetic field. Red numbers represent the number of degenerated classes for each line of the spectrum}
@@
 \item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
-\item $\vec{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2= \bar \eta_0^2$.
+\item $\bm{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2= \bar \eta_0^2$.
 For the two resonant classes : $\bar \eta = \bar {\eta}_{\rm same} + \bar {\eta}_{\rm far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
-\item $\vec{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{\rm same}+\bar {\eta}_{\rm close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
-\item $\vec{B} \parallel \langle 111 \rangle$ : For the non-resonant class (parallel to $\vec B$) : $\bar \eta = \bar {\rm \eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.
+\item $\bm{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{\rm same}+\bar {\eta}_{\rm close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
+\item $\bm{B} \parallel \langle 111 \rangle$ : For the non-resonant class (parallel to $\bm B$) : $\bar \eta = \bar {\rm \eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.
 For the triply resonant classes: $\bar \eta = \bar {\eta}_{\rm same}+2\bar {\eta}_{\rm far}$ and $\bar \eta^2\approx 28.4\, \bar \eta_0^2$.
-\item $\vec{B} \parallel \langle 100 \rangle$ : For the quadruple resonance : $\bar \eta = \bar {\eta}_{\rm same}+2\bar {\eta}_{\rm close} + \bar {\eta}_{\rm far}$  and $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
+\item $\bm{B} \parallel \langle 100 \rangle$ : For the quadruple resonance : $\bar \eta = \bar {\eta}_{\rm same}+2\bar {\eta}_{\rm close} + \bar {\eta}_{\rm far}$  and $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
 \end{itemize}
@@
 
-The second one is the difference between the $\vec{B} \parallel \langle 100 \rangle$ case and the $\vec{B}=0$ case.
+The second one is the difference between the $\bm{B} \parallel \langle 100 \rangle$ case and the $\bm{B}=0$ case.
 When a magnetic field is applied in the [100] direction, all four classes are resonant and the spin Hamiltonian basis is $\{ \ket{0},\ket{+1},\ket{-1} \} $.
@@
 Fig.
-[REF] of the main text shows that we indeed observe a shorter lifetime when $\vec B=0$, in agreement with the higher value of $\bar \eta^2$, but this decrease could also come from the double-flip processes which are absent when $\vec B \neq 0$.
+[REF] of the main text shows that we indeed observe a shorter lifetime when $\bm B=0$, in agreement with the higher value of $\bar \eta^2$, but this decrease could also come from the double-flip processes which are absent when $\bm B \neq 0$.
 
@@
 
-Double-flip processes, that is processes where the two spins flip in the same direction, compared to flip-flop where the two spins flip in opposite directions, are related to matrix elements  such as $\mel{+1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-1}$ in the $\ket{\pm 1}$ basis or $\mel{+,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-}$ in the $\ket{\pm}$ basis.
+Double-flip processes are processes where both spins lose or gain one unit of spin angular momentum, as ooposed to flip-flop processes where one spin loses one quantum and the other gains one.
+These are related to matrix elements  such as $\mel{+1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-1}$ in the $\ket{\pm 1}$ basis or $\mel{+,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-}$ in the $\ket{\pm}$ basis.
 These matrix elements couple two-spin states that are never fully resonant, however with low-to zero magnetic, the residual splitting due to local electric and magnetic field ($\sim 6\ \rm MHz$ with our samples) is small enough compared to the fluctuators linewidth  measured in sec.
@@
 We can see that the decay rate increase ($\sim$ 5 times the baseline value) is significantly higher than the increase due solely to the $\ket{+/-}$ basis ($\sim$ 2 times the baseline value).
-This lead us to believe that the double-flip process are the main reason behind the decrease in spin lifetime for $\vec B=0$ observed in Fig.
+This lead us to believe that the double-flip process are the main reason behind the decrease in spin lifetime for $\bm B=0$ observed in Fig.
 [REF] of the main text.
@@
 \bibliographystyle{plain}
+
 \bibliography{CR_SI}{}

### 2022-06-21 Mes papiers/Low field CR/Version pré-Gabriel/Low_field_CR.tex
tutu

\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-2}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{natbib}
\usepackage{physics}
\usepackage[caption=false]{subfig}

%\newcommand{|}{Y$_2$SiO$_5$}
\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{Low magnetic field depolarization of NV$^-$ centers through dipole-dipole interaction}

\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, P. Huillery$^?$,  G. H\'etet}

\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e Paris Cit\'e , 24 rue Lhomond, 75231 Paris Cedex 05, France.}

\begin{abstract}
C'est trop bien
\end{abstract}

\maketitle

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/shema_summary}
\caption{Shema}
\label{PL_NV_density}
\end{figure}

Intro

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig dense vs pas dense}
\caption{Photoluminescence measurement from NV centers ensemble as a function of an external magnetic field randomly oriented (a) for a CVD sample containing $\approx$ 4 ppb NV$^-$ centers, (b) for an HPHT sample containing $\approx$ 3 ppm NV$^-$ centers}
\label{PL_NV_density}
\end{figure}

%In this letter/article, we characterize the depolarization of the spins observed for dense ensemble of NV centers in zero magnetic field and its potential application for DC magnetometry. While the main mechanism behind the depolarization, the lift in the degeneracy between the four classes of NV centers, is already well studied and can be exploited in a microwave-less vector magnetometry protocol, we found two other depolarization mechanisms specific to the zero-field region which could play an important role in a low-field magnetometry protocol.
%
%The main signature of the spin depolarization in low field is the characteristic dip in photoluminescence (PL) observed only for high density ($\gtrsim 1$ ppm) of NV centers, as shown on fig \ref{PL_NV_density}. The decrease in the spins' lifetime in zero field makes the optical polarization scheme of the NV centers less effective and therefore reduce the population of the bright $\ket{0}$ spin state. %The reason for the decrease in PL at higher magnetic field values is the mixing of the bright spin state $\ket{0}$ to the darker state $\ket{-1}$ induced by the transverse magnetic field. This effect does not modify (on first approximation) the spins' lifetime, and is common to both dense and sparse samples.

Fig. \ref{PL_NV_density} shows the change in photoluminescence (PL) with respect to an external magnetic field for two samples with a low and high concentration of NV$^-$ centers. Both samples show a decrease in PL as the magnetic field amplitude increases, due to eigenstate mixing caused by the transverse part of the magnetic field \citep{epstein2005anisotropic,lai2009influence}. We notice however a stark difference in the low magnetic field region where only the high-density sample observe a drop in PL. In this article we propose to explore the reasons behind this zero-field line present only for dense NV$^-$ ensemble.

Unlike the decrease in PL for higher magnetic field values, the drop in PL at low magnetic field is associated with a decrease of the NV's spin lifetime $T_1$. Indeed, since the $\ket{m_s=0}$ spin state is brighter than the $\ket{m_s=\pm 1}$ states, and since the optical pumping in the $\ket{m_s=0}$ state is in competition with the relaxation in a thermal distribution of the states, decreasing the spins lifetime results in a decrease of the NV PL \citep{finco2021imaging}. To understand the behavior of the PL in zero-field, we therefore need to study the dynamics of the spin in the low magnetic field region.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig T1}
\caption{Spin $T_1$ measurement protocol. (a) Sequence used to measure the spin lifetime : the laser is turned off for a time $\tau$ twice and the spin state is read at the end of the dark time by measuring the initial photoluminescence when the laser is turned back on. A microwave $\pi$-pulse resonant with one of the spin transition is placed before the second reading. The  $T_1$ sequence result is the subtraction $S_1-S_2$. (b) Measurement of $S_1-S_2$ for a high density of NV centers in zero magnetic field field and non-zero magnetic field. The fit (orange plain line) is detailed in the main text}
\label{T1}
\end{figure}

The $T_1$ depolarization dynamics of single or dilute NV centers at room temperature is dominated by a two-phonon Raman process \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depends on the crystal lattice temperature but does not depend on the external magnetic field. However, it has been observed that dense NV centers ensemble have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi_depolarization_2017, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field and not on the temperature . This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi_depolarization_2017}. Some inhomogeneity between the NV centers is further needed in order to explain the depolarization of the spin ensemble. We will denote $T_1^{\rm ph}$ the characteristic time associated with the phonon relaxation process and $T_1^{\rm dd}$ the characteristic time associated with the dipole-dipole relaxation process.

The spin lifetime protocol used here is described in Fig. \ref{T1} and is based on previous similar experiments \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi_depolarization_2017}. It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$. Using only this sequence can result in artifacts in the signal, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019}. It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to project the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state. By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.

\begin{figure*}
\includegraphics[width=.95\textwidth]{Figures/fig 100 vs 1x1x1x1}
\caption{Dependency on the magnetic field angle for the zero field depolarization. (a) ODMR spectrum in zero field. (bi) ODMR spectrum for a magnetic field $\approx$ 60 G, misaligned by $\sim  24^\circ$ with the [100] axis. (bii) Normalized photoluminescence of the NV$^-$ ensemble as a function of the magnetic field amplitude. (biii) Stretch part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude. (ci), (cii) and (ciii) : same measurement with a magnetic field aligned with the [100] axis}
\label{100_VS_1x4}
\end{figure*}

The ensemble spin decay is non-exponential due to inhomogeneities between the NV$^-$ centers. We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}. A conclusion of this model is that, for an homogeneous 3D distribution of fluctuator, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$. Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$  (see SI).

For the samples used in this paper (detailed in SI), we observed that $T_1^{\rm dd} \sim T_1^{\rm ph}$, meaning that we had to include both  in our analysis. The resulting fitting formula is therefore :
\begin{equation}
S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}}),
\end{equation}
Further details on the fitting procedure are available in SI. For the rest of the article we will be only interested on the stretched exponential lifetime ${T_1^{\rm dd}}$.

%\begin{figure}
%\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
%\caption{Tuning resonant coupling through magnetic field orientation. (a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice. (b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes. The detuning $\Delta \nu$ between the classes $\beta$ and $\gamma$ was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_\gamma$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz . Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
%\label{largeur_fluct}
%\end{figure}

%Fig. \ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins. Because of the diamond symmetry, we can distinguish 4 groups ("classes") of NV centers, based on the four possible orientation of the NV axis in the crystal. These four classes have \textit{a priori} different spin resonance frequency based on the projection of the magnetic field on their respective NV axis. By tuning the angle of the magnetic field, we can therefore control the frequency detuning between the different classes, and as a result we can control the number of resonant NV centers by bringing in and out of resonance several classes.
%
%Fig. \ref{largeur_fluct} (c) in particular shows the dipole-induced lifetime $T_1^{\rm dd}$ for a given class as a function of the energy detuning with another class. One can notice that the width of the $T_1^{\rm dd}$ bell shape is significantly larger than that of the frequency overlap between the two classes, modeled here by an ODMR line stretched by a factor $\sqrt{2}$ (see SI). This extra width can not be attributed to the dipole-dipole coupling strength : for a 3 ppm concentration of NV$^-$ centers, the average coupling strength between neighboring NV centers should be $\sim 27\ \rm kHz$, two order of magnitude below the observed width. We can also see that $T_1^{\rm dd}$ follows a Lorentzian shape despite the ODMR lines being non-Lorentzian. These two observations corroborate the idea developed in \cite{choi_depolarization_2017} that the interaction width is dominated by the noise spectrum of the lifetime-limited fluctuators.

One of the way we can isolate dipole-dipole related phenomena with NV centers is by tuning the number of NV centers resonant with each other. Indeed, Fig. \ref{100_VS_1x4} (bi) shows the characteristic 8 lines observed with ensemble NV optically detected magnetic resonance (ODMR) spectrum. These 8 lines correspond to the four "classes" of NV centers - that is physical orientation of the NV axis is the diamond crystal cell - multiplied by the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$. The transitions of the four different classes can be controlled somewhat independently by changing the amplitude and orientation of the external magnetic field, and in particular they can be brought to resonance for particular orientation of the magnetic field. Fig. \ref{100_VS_1x4} (ci) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis. Other example of class resonance are given in SI.

Fig. \ref{100_VS_1x4} (bii) and (biii) shows respectively the PL and the stretched exponential spin decay $T_1^{\rm dd}$ of an ensemble of NV centers with respect to the amplitude of the external magnetic field. The orientation of the magnetic field was chosen in order to lift the degeneracy between the four classes, as shown in Fig. \ref{100_VS_1x4} (bi). We observe that the drop in PL at low magnetic field is indeed correlated with an increase in the spin decay rate, as is expected due to the higher number of resonant spins in zero magnetic field, where all four classes are resonant, compared to the non-degenerate case for $\vec \neq 0$. We can also see that the decrease in PL for $B>40\ \rm G$ due to state mixing by the transverse magnetic field is not correlated to a modification of the spin lifetime.

However, the splitting of the four classes is not sole reason to the decrease of the spin lifetime in low magnetic field. Indeed, we performed similar measurements in Fig. \ref{100_VS_1x4} (cii) and (ciii) but this time fixing the magnetic field orientation along the [100] axis. For this particular orientation, the four classes are always resonant, regardless of the field amplitude. We can see that, although considerably reduced, there still is an increase in the spin decay rate and a corresponding drop in PL for low magnetic field values. We attribute the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20\ \rm G$ to dipolar interaction with NV having a first neighbor $^{13}C$ \cite{pellet2021optical}.

The aim of this paper is to understand the remaining contribution to the spin decay in this second case. We identified and isolated two additional effects : the domination by the local electric field for low magnetic field value which results in a change of the spin Hamiltonian eigenstates, and the presence of double-flip processes where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.

In order to distinguish these two contributions, we decided to simulate the effect of the electric field by using a transverse magnetic field : for relatively weak transverse magnetic field, the eigen basis of the spin Hamiltonian is close to $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $, which is the the spin Hamiltonian basis when the electric field dominates, as detailed in SI. This property has been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig transverse field V2}
\caption{(a) Simulated energies for the spin Hamiltonian in the presence of pure transverse magnetic field $B_\perp$. The three spin levels are denoted $\ket{g}$, $\ket{d}$ and $\ket{e}$. (b) Blue dashed curve : frequency detuning $\Delta \nu$ between the $\ket{g}$, $\ket{d}$ states. Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\bra{e}$ and $\bra{+}$. (c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single class as a function of the measured detuning $\Delta \nu$ between the two transitions $\ket{g} \to \ket {d}$ and  $\ket{g} \to \ket {e}$ for purely transverse magnetic field. The green dashed line correspond to the decay rate for the same class when the magnetic field is longitudinal}
\label{B_transverse}
\end{figure}

Fig. \ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian with respect to a purely transverse magnetic field. We will denote these eigenstates in the general case $\ket{g}$, $\ket{d}$ and $\ket{e}$. For weak transverse magnetic field, $\ket{g} \approx \ket{0}$, $\ket{d}=\ket{-}$ and $\ket{e} \approx \ket{+}$. As the magnetic field increases, $\ket{g}$ and $\ket{e}$ starts to become a mixing of $\ket{0}$ and $\ket{+}$ while $\ket{d}$ remains equal to $\ket{-}$.

Crucially, this mixing happens on a different scale than the associated energy splitting of the $\ket{d}$ and $\ket{e}$ states : Fig. \ref{B_transverse} (b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$. We can see that for a field value of $\sim 130 G$, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50 MHz$, far exceeding the range of the dipole-dipole resonant coupling (see details in SI), but $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis. We can therefore decouple the effects of the double-flip processes, which are no longer resonant, from the effects of the change of basis induced by the transverse electric - or in this case magnetic - field.

Fig. \ref{B_transverse} (c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers exposed to a transverse magnetic field between 25 and 130 G. The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR. We attribute the decrease in the spin decay rate when the detuning increases to the loss of effectiveness of the double-flip processes. We can notice that the decay rate reaches a plateau for $\Delta \nu \geq 30\ \rm MHz$ with a value about twice as big as the decay rate for a longitudinal magnetic field. This plateau we attribute to the effect of the transverse magnetic field, which we assume to be similar to the effect of the electric field in the low magnetic field regime. We should note that the effect of the double-flip processes is about 4 times more important in the depolarization rate than the effect of the electric field, even though the states are not initially fully resonant.

Overall we predicted, based on an extension of the fluctuator model developed in SI, an increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$. While the measured increase is smaller than the predicted one, as with most of the predictions based on the fluctuator model, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.

Similarly, we can compute the predicted decay rate in the $\vec B \parallel \left[100\right]$ case compared to the $\vec B = 0$ case, and find a theoretical increase of $\sim 20\%$ in the $\vec B = 0$ case due to the effects of the local electric field. This, along with the double-flip processes explains the behavior observed in Fig. \ref{100_VS_1x4} (cii) and (ciii).

%... (lis un peu ce que t'as déja écrit avant de tout réécrire, je te connais moi futur. En vrai c'est de la merde, j'ai tout réécrit)
%The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
%\begin{equation}
%\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
%\end{equation}
%Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$. In situations such as the one described in Fig. \ref{largeur_fluct}, the only relevant terms of $\mathcal{H}^{\rm dd}$ in term of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$. In zero magnetic field however, we have to take into account other aspects of the dipolar Hamiltonian in consideration.
%
%First is the change of basis of the single spin Hamiltonian $\mathcal{H}_s$ : in zero external magnetic field, the eigenstates of $\mathcal{H}_s$ in the $\{\ket{+1},\ket{-1}\}$ manifold are determined by local electric and magnetic field coming from nearby impurities \cite{mittiga2018imaging}. For the samples used in this study, $\mathcal{H}_s$ was dominated by local electric field (see SI), meaning that the proper eigenstates of $\mathcal{H}_s$ in zero magnetic field are $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $
%
%Because the splitting between $\ket{+}$ and $\ket{-}$ ($\approx 9\ \rm MHz$) is much greater than the dipole-dipole interaction ($J_0 / r^3 \approx 30\ \rm kHz$), the flip-flop terms of $\mathcal{H}^{\rm dd}$ now reads $\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}$. Averaging over all relative positions of two aligned NV centers, we found that on average $\abs{\mel{0,+}{\mathcal{H}^{\rm dd}}{+,0}}$ was greater than 
%$\abs{\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}}$ by a factor $1.8 \sim 2$ depending on the correlation lengths of the local electric field.
%When all four classes are resonant, the theoretical decrease in the spin lifetime due to the change of basis is $\sim 20 \%$ (See SI de ouf).
%
%Second is the near-resonance condition of the double-flip processes such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{-1,0}$ or $\mel{0,+}{\mathcal{H}^{\rm dd}}{-,0}$. These terms usually couple non resonant states due to the energy mismatch between $\ket{+1}$ and $\ket{-1}$. However in zero field, the energy splitting between $\ket{+}$ and $\ket{-}$ is comparable with the fluctuator's spectral width measured in Fig. \ref{largeur_fluct} to be $\approx 8\ \rm MHz$, making the double-flip population transfer possible.
%
%\begin{figure}
%\includegraphics[width=0.45\textwidth]{Figures/fig transverse field simu}
%\caption{Simulated eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field. (a) Energies of the three eigenstates $\ket{g}$, $\ket{d}$ and $\ket{e}$ as a function of the magnetic field amplitude. (b) Blue dashed curve : frequency detuning between the two transitions $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$. Plain red curve : projection of $\ket{e}$ on $\ket{+}=(\ket{+1}+\ket{-1})/\sqrt{2}$ as a function the magnetic field. Is equal to the projection of $\ket{g}$ on $\ket{0}$.}
%\label{calculs_B_transverse}
%\end{figure}
%
%In order to evaluate the relative contribution of these two factors, we investigate the spin relaxation in the case of pure transverse magnetic field. Calling $\{ \ket{g}, \ket{d}, \ket{e} \}$ the three eigenstates of the spin Hamiltonian in the presence of purely transverse magnetic field, Fig. \ref{calculs_B_transverse} shows that for small enough magnetic field, the $\{ \ket{g}, \ket{d}, \ket{e} \}$ basis is almost equal to the $\{ \ket{0},\ket{+},\ket{-} \} $ basis : for $B=120\ \rm G$, $\braket{0}{g}\approx 0.99$, $\braket{-}{d}=1$ and $\braket{+}{e}\approx 0.99$. However the energy difference between $\ket{d}$ and $\ket{e}$ reaches $\approx 45\ \rm G$ which is high enough to cancel out the double-flip processes and allows us to isolate the change of basis hypothesis from the double-flip one.
%%As shown in Fig. SOON, (and in SI), the $\{ \ket{0},\ket{+},\ket{-} \} $ basis correspond to the spin Hamiltonian eigenstates in two situations : either in zero magnetic field when the electric field dominates, or when the magnetic field is purely transverse, as long as it remains small before the zero-field splitting. %(rq : est-ce que je dois mettre le hamiltonien de spin dans le main text du coup ?).
%
%\begin{figure}
%\includegraphics[width=0.45\textwidth]{Figures/fig transverse field}
%\caption{Modification of the stretch lifetime in the presence of purely transverse magnetic field. (a) Stretch component of the ensemble lifetime as a function of the field amplitude. (b) Transition frequencies for the $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$ transitions, measured through ODMR. (c) Stretch component of the lifetime as a function of the frequency detuning between the two transistions (blue circles), fitted by a Lorentzian centered in $\Delta \nu=0$ with half width at half maximum 8MHz. The green dashed line correspond to the lifetime of a single class aligned with the magnetic field.}
%\label{exp_B_transverse}
%\end{figure}
%
%Fig. \ref{exp_B_transverse} shows the dipole-dipole spin relaxation $T_1^{\rm dd}$ for a class of NV whose axis is orthogonal to the applied magnetic field. By monitoring the frequencies of the transitions $\ket{g} \leftrightarrow \ket{d}$ and $\ket{g} \leftrightarrow \ket{e}$, we can deduce the detuning $\Delta \nu$ between the two-spin states $\ket{0,+}$ and $\ket{-,0}$. Plotting then $1/T_1^{\rm dd}$ as a function of $\Delta \nu$, we can notice two things. 
%
%First the decrease of $1/T_1^{\rm dd}$ as $\Delta \nu$ increases which corresponds to the loss in effectiveness of the double flip processes. The shape of the curve matches decently well with the tail of a Lorentzian centered in $\Delta \nu=0$ with a width of 8 MHz, confirming that the double-flip process interaction range is likely determined by the fluctuators noise spectrum just like the flip-flop processes. 
%
%Second is the plateau reached for $\Delta \nu \gtrsim 30\ \rm MHz$. The final $1/T_1^{\rm dd}$ value is about twice as high as that for a spin well aligned with the magnetic field, showing the increased depolarization due to the change of basis.
%
%This experiment shows that, for the sample studied here, the depolarization due to the double-flip processes dominates the one due to the change of basis in zero magnetic field.

\section*{Magnetometry}
\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig_magneto}
\caption{Low field magnetometry protocol. (a) Demodulated photoluminescence as a function of an externally applied magnetic field with an additional oscillatory magnetic field. (b) Measured magnetic field when alternating a small external magnetic field offset. (c) Histogram of the measurement in Fig. (b) fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$. (d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.}
\label{magneto}
\end{figure}
%
DC microwave-less magnetometry has already been performed with NV ensembles, using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}. Here we propose to perform a similar protocol, but using the spin depolarization in zero-field. The main difference with the previously mentioned protocols is the fact that this one doesn't rely on crystalline orientation, making this protocol usable with diamond powders or polycrystalline samples. The idea of performing optical magnetometry on the zero-field photoluminescence feature of NV ensembles has been suggested before \cite{filimonenko2018weak, filimonenko2022manifestation} but had not been implemented so far.

Fig. \ref{magneto} (a) shows the demodulated PL when a DC magnetic field is scanned in a random direction and an additional alternating magnetic field of a few Gauss at $\sim \ \rm kHz$ is added through the same electromagnet (see experimental details in SI). We can see a sharp and relatively linear slope in low field $\abs{B} < 5\ \rm G$. Once calibrated, in this case with ODMR, the slope can provide a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as seen in \cite{zheng_microwave-free_2020}.

In order to measure the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take an histogram of the measured fields, as shown in Fig. \ref{magneto} (b) and (c). The histogram is well fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$. The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which allows us to measure the sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nT}/\sqrt{\rm Hz}$.

We can then try to measure the relative importance of the three causes of spin depolarization on the field sensitivity, namely the splitting of the classes, the local electric field and the double-flip processes. In order to do so, we measure the sensitivity while changing the angle of the field. When $\vec B$ is aligned with the [100] crystalline axis, only the double flips and the electric field cause a depolarization, whereas in every other orientation the three effects are at play.

The results are shown in Fig. \ref{magneto} (d). We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity remains relatively flat, which means that the double-flips and electric field effects are the dominant factors in the sensitivity of this protocol. While it may seem surprising that the effects with a lower contribution on the PL contrast have a higher effect on this PL-based protocol, what matters here is not the absolute contrast but the slope of the change of contrast with respect to the magnetic field, which is sharper for the electric field and double flips than it is for the lift of degeneracy. %This is in agreement with the measurements in Fig. \ref{100_VS_1x4} : even though the contrast is much better in the not-[100] case, the slope in term of contrast percentage/magnetic field is comparable in both cases (en faite pas complètement, y'a pas loin d'un facteur 2 sur les dérivées numériques)
It should be noted though that this observation is sample dependent, and that other samples, including from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from the electric field and double-flip processes.

\section*{Conclusion}
In this work we have identified three mechanisms causing the extra spin depolarization observed in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole cross-relaxations between the spins : the lift of degeneracy between the four classes caused by the magnetic field, the domination of the local electric field which causes a change in the Hamiltonian eigenstates, and the double-flip process allowed by the proximity in energy of the spin $\ket{\pm}$ states. We identified the lift in degeneracy as the main cause in the zero-field depolarization, followed by the double-flip processes and then the electric field. We have demonstrated a potential use of this depolarization as a DC microwave-less and orientation-free magnetometer with sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal.We have observed that in some cases the double-flips and electric field play a more important role than the lift of degeneracy in the magnetometry sensitivity.
\section*{acknowledgements}

\bibliographystyle{plain}
\bibliography{CR}{}

\end{document}

### 2022-06-21 Pratique/AN.py

@@ -38,8 +38,8 @@ def couplage_NV():
 
 # couplage_NV()
 
-print((1/4*sqrt(1/3)*2/(3*sqrt(3)))**2)
 
 
-print(sqrt(0.0030864197530864196))
+
+print(139+39+109+245+99+299)
 

### 2022-06-22 Mes papiers/Low field CR/Low_field_CR.tex
tutu

@@
 The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science in part thanks to the possibility to optically polarize and read-out its spin state at ambient conditions.
-Many recent work indeed focus on ensembles of spins from NV centers with the prospect to boost the magnetic field sensing capabilities and for studying many body effects (resultats de chez BOSCH, LUKIN , FEFOR, ALEX, BIFONE)
+Many recent work indeed focus on ensembles of spins from NV centers with the prospect to boost the magnetic field sensing capabilities and for studying many body effects
+\citep{TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized, kucsko2018critical, giri_coupled_2018}
 
 When the spin ensemble becomes dense, inhomogeneities in the relaxation rates of NVs can give rise to spin depolarisation with a very rich many body dynamics associated with disorder.
-This mechanisms however limits the efficiency of magnetometers (LUKIN PRX).
+This mechanisms however limits the efficiency of magnetometers \citep{zhou2020quantum}
 %It was proposed to exploit these CR to perform sensitive microwave free magnetometry (REF).  
@@
 Both samples show a decrease in PL as the magnetic field amplitude increases.
-There is however a stark difference in the low magnetic field region where only high-density samples shows a drop in PL (CITATION).
+There is however a stark difference in the low magnetic field region where only high-density samples shows a drop in PL \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}.
 This effect can be observed on all our samples (see SI) whenever the NV concentration lies in the ppm range.
+The drop of the PL at low magnetic field is associated with a decrease of the NV's spin lifetime $T_1$ \citep{jarmola_temperature-_2012}, which results in a decrease of the population in the bright state $\ket{0}$.
 
-The drop of the PL at low magnetic field is generally associated with a decrease of the NV's spin lifetime $T_1$ (CITE JARMOLA), which results in a decrease of the population in the bright state $0$.
 %Indeed, since the $\ket{m_s=0}$ spin state is brighter than the $\ket{m_s=\pm 1}$ states, and since the optical pumping in the $\ket{m_s=0}$ state is in competition with the relaxation in a thermal distribution of the states, decreasing the spins lifetime results in a decrease of the NV PL \citep{finco2021imaging}.
@@
 The depolarization dynamics of the spin of single or dilute NV centers at room temperature is dominated by two-phonon Raman processes \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depend on the crystal lattice temperature.
-It has been observed that dense NV centers ensemble have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi_depolarization_2017, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field amplitude.
-This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi_depolarization_2017}.
+It has been observed that dense NV centers ensemble have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi2017depolarization, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field amplitude.
+This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi2017depolarization}.
 Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble.
@@
 \caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation. }
-\label{T1}
+\label{schema_intro}
 \end{figure}
@@
 \includegraphics[width=0.45\textwidth]{Figures/fig_T1.pdf}
-\caption{a) Photoluminescence measurement from NV centers ensemble as a function of an external magnetic field applied in an arbitrary direction (a) for a CVD sample containing $\approx4$ ppb of NV$^-$ centers, (b) for an HPHT sample containing $\approx3$ ppm of NV$^-$ centers.
+\caption{a) Photoluminescence measurement from NV centers ensemble as a function of an external magnetic field applied in an arbitrary direction (a) for the sample CVD-PPB, with [NV$^-$]$\approx 50\ \rm ppb$ (b) for the sample HPHT-150-1, with [NV$^-$]$\approx 3\ \rm ppm$.
 (c) Sequence used to measure the spin lifetime.
@@
 It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
-By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi_depolarization_2017}.
+By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi2017depolarization}.
 
 Fig. \ref{T1}-b) shows the subtracted signals when $\bm B=0$ and $\bm B \neq 0$.
-The ensemble spin decay curves are not exponential, which was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi_depolarization_2017}.
-We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}.
+The ensemble spin decay curves are not exponential, which was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi2017depolarization}.
+We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi2017depolarization}.
 A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$.
 Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$.
-For the samples used in this paper (detailed in SI) $T_1^{\rm dd} \sim T_1^{\rm ph}$, so both decay processes have to included in the analysis.
-Fitting the two curves by the product of an exponential and a stretched exponential, we find $T_1^{\rm ph}=3.2$ ms and $T_1^{\rm dd}= ??$ here.
-We also conducted further measurements that confirm the lifetime-limited character of the fluctuator employed in the model \citep{choi_depolarization_2017}, see SI.
+For the samples used in this paper (detailed in SI) $T_1^{\rm dd} \sim T_1^{\rm ph}$, so both decay processes have to be included in the analysis.
+Fitting the two curves by the product of an exponential and a stretched exponential, we find $T_1^{\rm ph}=3.6$ ms for both and $T_1^{\rm dd}= 0.6\ \rm ms$ and $13.0\ \rm ms$ respectively.
+We also conducted further measurements that confirm the lifetime-limited character of the fluctuator employed in the model \citep{choi2017depolarization}, see SI.
 %These results show that our sample clearly manifests dipolar interactions amongst closely packed NV centers, but do not tell about the magnitude of the several possible mechanisms involved in the dipolar relaxation.
@@
 Fig.
-\ref{100_VS_1x4} (b)-i) shows an optically detected magnetic resonance (ODMR) spectrum from the dense ensemble.
+\ref{100_VS_1x4} (b)-i) shows an optically detected magnetic resonance (ODMR) spectrum from the dense ensemble HPHT-150-1.
 The observed lines correspond to the projections of the magnetic fields onto the four "classes" of NV centers - that is physical orientation of the NV axis is the diamond crystal cell - together with the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$.
@@
 Fig. \ref{100_VS_1x4} (b)-ii) and (b)-iii) show the PL from an ensemble of NV centers as a function of the amplitude of a magnetic field aligned as in Fig. \ref{100_VS_1x4} (b)-i). 
-We also extract the stretched exponential spin decay time $T_1^{\rm dd}$ by running the sequence shown in Fig. 2-c) at each of the XX magnetic field values. 
-We observe that the 3 \% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.05 kHz, as the magnetic field increases.
+We also extract the stretched exponential spin decay time $T_1^{\rm dd}$ by running the sequence shown in Fig. 2-c) at each of the 0.5 G magnetic field values. 
+We observe that the 3 \% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.08 kHz, as the magnetic field increases.
 This result can be explained by the higher number of resonant spins in zero magnetic field, where all four classes are degenerate, as opposed when $\bm B \neq 0$.
-The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SI ?).
+The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SI ?) [Je sais pas, je le fais ?].
 Note that the decrease in PL for $B>40\ \rm G$ are related to state mixing by the transverse magnetic field, so it is not correlated to a modification of the spin lifetime.
@@
 We can see that there still is an increase in the spin decay rate and a corresponding drop in the PL under low magnetic field values, although considerably reduced.
-Also, the rate of change in both the PL and $T_1^{\rm dd}$ is larger by more than a factor of 3 in the [100] configuration.
+%Also, the rate of change in both the PL and $T_1^{\rm dd}$ is larger by more than a factor of 3 in the [100] configuration.
 Although the contrast is smaller, this effect can thus drastically impact zero-field magnetometry protocols based on CR.
+C'est faux ça, c'est meme plutot le contraire ici
 The main aim of this paper is to understand the physics governing the contribution to the spin decay in this second scenario.
@@
 %where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.
-It was shown in  \cite{mittiga2018imaging} that local electric fields coming from charge traps such as $P_1^+$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in large NV ensembles.
+It was shown in  \cite{mittiga2018imaging} that local electric fields coming from charge traps such as $P_1^+$ and $NV^-$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in large NV ensembles.
 Magnetic field noise comes in to second order in zero field.
@@
 Another mechanism at stake when considering dipolar interactions with spin-1 systems are double-flip processes.
-These processes correspond to terms like $\ket{+1,0} \bra{0,1}$ in the anisotropic dipolar interaction, giving rise to an exchange of two units of spin-angular momentum.
+These processes correspond to terms like $\ket{+1,0} \bra{0,-1}$ in the anisotropic (?) dipolar interaction, giving rise to an exchange of two units of spin-angular momentum.
 These processes can become significant when the states $\pm 1$ are degenerate and are thus likely to explain part of the zero-field dip in Fig.
@@
 We can see that at a field value of $\sim 130$ G, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50$ MHz, far exceeding the range of the dipole-dipole resonant coupling (see details in SI), but $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
-We can therefore decouple the effects of the double-flip processes  $\propto \ket{g,d}\bra{d,e}$, which are no longer resonant when $| \bm B | > 130$ G, from the effects of the mixing induced by the transverse electric - or in this case magnetic - field.
+We can therefore decouple the effects of the double-flip processes  $\propto \ket{g,d}\bra{d,e}$, which are no longer resonant when $| \bm B | \geq 100$ G, from the effects of the mixing induced by the transverse electric - or in this case magnetic - field.
 
 Fig.
-\ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers exposed to a transverse magnetic field between 25 (??) and 130 G (FOR A DIFFERENT SAMPLE...
-which ?
-we observe a $1/T_1^{\rm dd}$ of $30$ Hz under large longitudinal magnetic fields for this class ?).
+\ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers exposed to a transverse magnetic field between 20 and 130 G on another sample HPHT-150-2, upon which we measured a stretched decay $1/T_1^{\rm dd}=25\pm 5\  \rm Hz$ in the case of strong longitudinal magnetic field.
 The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
 We attribute the decrease in the spin decay rate when the detuning increases to the loss of effectiveness of the double-flip processes.
-We can notice that the decay rate reaches a plateau for $\Delta \nu \geq 30\ \rm MHz$ with a value about twice as large as the decay rate for a longitudinal magnetic field.
+We can notice that the decay rate reaches a plateau for $\Delta \nu \geq 30\ \rm MHz$ with a value $1/T_1^{\rm dd}=57\pm 5\  \rm Hz$ about twice as large as the decay rate for a longitudinal magnetic field.
 This plateau we attribute to the effect of the transverse magnetic field, which we assume to be similar to the effect of the electric field in the low magnetic field regime.
-We should note that the effect of the double-flip processes is about 4 times more important in the depolarization rate than the effect of the electric field, even though the states are not initially fully resonant.
+We should note that the effect of the double-flip processes is about 5 times more important in the depolarization rate than the effect of the electric field, even though the states are not initially fully resonant.
 
@@
 Fig.
-\ref{magneto} (a) shows the demodulated PL as a DC magnetic field is scanned in a random direction and an additional alternating magnetic field of a few Gauss at $\sim \ \rm kHz$ is added through the same electromagnet (see experimental details in SI) WHICH SAMPLE ?
-.
+\ref{magneto} (a) shows the demodulated PL as a DC magnetic field is scanned in a random direction and an additional alternating magnetic field of amplitude $\sim 10\ \rm G$ and frequency  $\sim 1\ \rm kHz$ is added through the same electromagnet.
+The sample used here is HPHT-15-1 and the laser power $\sim 1\ \rm mW$.
 We can see a sharp linear slope in low field $\abs{B} < 5\ \rm G$.
@@
 The lift in degeneracy of the spin state of different NV classes was found to be the main cause of zero-field depolarization, followed by double-flip processes and then the electric field induced mixing.
-We have employed cross-relaxation for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ crystal and show that that double-flips and electric field play an important role.
+We have employed cross-relaxation for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ commercially available diamond and show that that double-flips and electric field play an important role.
 Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins. 

### 2022-06-22 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
tutu

@@
 \section{Samples}
-A faire à la fin pour être sur des figures du main text.
+%Note pour moi : j'ai inversé adamas 1 et 2 parraport aux noms des dossiers dans 20220324
+Here are the various samples used in this study :
+\begin{itemize}
+\item \textbf{HPHT-150-1} : A 150 $\mu$m HPHT 1b diamond irradiated and annealed to reach a concentration [NV$^-$] $\approx 3\ \rm ppm$.
+Bought as is from Adamas Nanotechnology (MDNV150umHi).
+This sample is used in Fig.
+2 (b), 2(d) and 3 of the main text, as well as Fig.
+\ref{ESR_single_spin}, \ref{T1_fits} and \ref{alphas} of the SI.
+\item \textbf{HPHT-150-2} : Another sample from the same batch as HPHT-150-1. This sample is used in Fig. 4 of the main text and Fig. \ref{largeur_fluct} (c) of the SI.
+\item \textbf{HPHT-150-3} : Another sample from the same batch as HPHT-150-1. This sample is used in Fig. \ref{Pola} of the SI.
+\item \textbf{HPHT-15-1}  : A 15 $\mu$m diamond with similar properties as the last ones, also bought from Adamas Nanotechnology (MDNV15umHi). This sample is used in Fig. 5 of the main text.
+\item \textbf{HPHT-15-2}  : Another sample from the same batch as HPHT-15-1. This sample is used in Fig. \ref{121 VS 22 fig} and \ref{Various ODMR} of the SI.
+\item \textbf{CVD-PPM} : A CVD bulk diamond described in \citep{TALLAIRE2020421}.
+It has been irradiated and annealed and contains [NV$^-$] $\approx 4\ \rm ppm$.
+This sample is used in Fig.
+\ref{Alignment} of the SI.
+\item \textbf{CVD-PPB} : Another CVD bulk diamond.
+This one has not been irradiated or annealed and therefore contains [NV$^-$] $\approx 50\ \rm ppb$.
+This sample is used in Fig.
+2(a) of the main text.
+
+\end{itemize}
+
 \section{Experimental Setup}
@@
 
+\subsection{Lift of the degeneracy between the four classes}
+
+\begin{figure}
+\includegraphics[width=0.9\textwidth]{Figures_SI/Fig_splitting}
+\caption{(a) : Simulation of the 8 possible transition frequencies (four $\ket{0} \to \ket{-1}$ and four $\ket{0} \to \ket{+1}$) as a function of the magnetic field amplitude for the orientation of the magnetic field shown in Fig.
+3(b) of the main text.
+(b) Difference in frequency between each pair of closest classes as a function of the magnetic field amplitude.
+The dotted line y=8.04 MHz correspond to the estimation of the dipole-dipole interaction range.}
+\label{splitting}
+\end{figure}
+We give here an estimate of the magnetic field required to lift the degeneracy of the four classes in the case of Fig.
+3(b) of the main text.
+Fig.
+\ref{splitting} shows the simulated energies for the 4 $\ket{0} \to \ket{-1}$ and the 4 $\ket{0} \to \ket{+1}$ as a function of the magnetic field, and the difference in energy between the closest pairs of transitions.
+We can see that the energy detuning between the classes $\Delta \nu$ crosses the value of the dipole-dipole interaction range (8.04 MHz, measured on sample HPHT-150-2 but likely to be similar on sample HPHT-150-1) for magnetic field values $\norm{B} = 12 \sim 15\ \rm G$, which are close to the half-width of the zero-field feature in Fig.
+3(b)(ii) and (iii) of the main text.
+
 \subsection{Effect of laser polarization}
@@
 \caption{Effect of the polarization of the incident laser.
-(a) Photoluminescence as a function of randomly oriented magnetic field amplitude for various polarization angle.
+(a) Photoluminescence of sample HPHT-150-3 as a function of randomly oriented magnetic field amplitude for various polarization angle.
 (b) Demodulated PL in the same conditions}
@@
 \end{figure}
-\subsection{Alignment of B}
-La faudrait p-e montrer ce qu'il se passe pour un décalage de quelques degrés. A voir si j'ai déja les plots
+\subsection{Alignment of the magnetic field along [100] axis}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{Figures_SI/alignement}
+\caption{PL of sample CVD-PPM as a function of the external magnetic field amplitude for different field misalignment with the [100] axis}
+\label{Alignment}
+\end{figure}
+We claim that the decrease in PL at low magnetic field in Fig.
+3 (c)(ii) of the main text is due to to the local electric field and the double flip processes.
+One other possibility would be a misalignment of the magnetic field with respect to the [100] axis, where the four classes would only be truly resonant with zero external magnetic field.
+
+Fig.
+\ref{Alignment} shows the change in PL with respect to the external magnetic field amplitude for various misalignment of the magnetic field direction compared to the [100] diamond crystalline axis for sample [REF].
+We can see that the central drop in PL is not critically affected by the misalignment of the magnetic field, and we are therefore confident that the drop in PL we observe when $B\parallel [100]$ is not an effect of misalignment.
+
+The initial alignment of the magnetic field was controlled by applying a field of $\sim 50\ \rm G$ and monitoring the overlap of the four classes with ODMR spectra (maximizing the PL drop is also an effective method).
+We estimate the initial alignment precision to be $\pm 1^\circ$.
+The misalignment was then added by rotating the electromagnet along an arbitrary axis.
+The slight asymmetry between the positive and negative value of the magnetic field could come from the earth magnetic field or from hysteresis in the magnetic core of our electromagnet.
 \section{Extension of the fluctuator model to low magnetic fields}
@@
 Inset : zoom-in on the 0.3-0.8 ms region.}
-\label{ VS 22 fig}
+\label{121 VS 22 fig}
 \end{figure}
@@
 \end{tabular}
-\caption{Computation of $\frac{\bar{\eta}}{ \frac{1}{4} \cdot \sqrt{\frac{1}{3}}}$ for the different eignebasis of the single spin Hamiltonian, and for the different angles between the $z$ axis of the two spins.}
+\caption{Computation of $\frac{\bar{\eta}}{ \frac{1}{4} \cdot \sqrt{\frac{1}{3}}}$ for the different eigenbasis of the single spin Hamiltonian, and for the different angles between the $z$ axis of the two spins.}
 \label{table eta}

### 2022-06-29 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
maxime

@@
 \includegraphics[width=0.9\textwidth]{Figures_SI/fig_ESR}
-\caption{ODMR measurement (a) in zero magnetic field, (b) in non-zero magnetic field when zooming on a single class transition}
+\caption{ODMR measurement (a) in zero magnetic field, (b) in non-zero magnetic field when zooming on a single class}
 \label{ESR_single_spin}
@@
 
-The optical polarization and readout of the spins is done by focusing a green laser on the diamond sample with an objective lens([ref]), and collecting the back-scattered red fluorescence from the NV center on an avalanche photo-diode (APD, REF).
-The laser is filtered out using a dichroic mirror (DM, ref) and a notch filter.
+The optical polarization and readout of the spins is done by focusing a green laser on the diamond sample with an objective lens (NA=0.65), and collecting the back-scattered red fluorescence from the NV center on an avalanche photo-diode (APD, Thorlabs APD410A).
+The laser is filtered out using a dichroic mirror and a notch filter.
 The NV$^0$ fluorescence is filtered using an additional 645 nm longpass filter.
 
-The laser used here is a [REF], providing pulses of X ns at a rate of 20 MHz, with an average power of $0.5 \sim 5\ \rm mW$.
+The laser used here is a Picoquant PDL 800-D with a 520 nm LDH laser head, providing pulses of 40 ps at a rate of 20 MHz, with an average power of $0.5 \sim 5\ \rm mW$.
 The trigger of the pulses is generated externally in order to achieve fast  gating of the laser for the $T_1$ measurement.
@@
 The magnetic field is provided by a homemade electromagnet composed of a c-shape iron core and  copper wires.
-The magnet is mounted on two mechanical rotation stages, allowing a control on the polar and azimuthal angle of the magnetic field within a fraction of a degree and is alimented through a low frequency amplifier (REF)
+The magnet is mounted on two mechanical rotation stages, allowing a control on the polar and azimuthal angle of the magnetic field within a fraction of a degree and is alimented through a low frequency amplifier (Leybold power function generator 522 63)
 
-The microwave field is generated by a Rhode \& Shwarz SMB 100A and is emitted with a handmade loop antenna. The field is gated by a [REF minicircuit] switch controlled externally.
+The microwave field is generated by a Rhode \& Shwarz SMB 100A and is emitted with a handmade loop antenna.
+The field is gated by a Mini-circuits ZASWA-2-50DRA+ switch controlled externally and amplified by a Mini-circuits ZHL-5W-422+ amplifier.
 
-A lock-in amplifier is used either to modulate the microwave amplitude for ODMR measurement, or to add an oscillatory magnetic field for the magnetometry protocol.
+A lock-in amplifier (SRS SR830 DSP) is used either to modulate the microwave amplitude for ODMR measurement, or to add an oscillatory magnetic field for the magnetometry protocol.
 In both case we use a modulation frequency $\sim 1\ \rm kHz$ and demodulate the APD signal.
@@
 To add consistency to the measurement of $T_1^{\rm dd}$, we fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ and the signal amplitude as the only free parameters in the fitting procedure.
-The value chosen for samples [REF] and [REF] was $T_1^{\rm ph}=3.62\ \rm ms$.
+The value chosen for samples HPHT-150-1 and HPHT-150-2 was $T_1^{\rm ph}=3.62\ \rm ms$.
 Fig.
@@
 Fig.
-\ref{Alignment} shows the change in PL with respect to the external magnetic field amplitude for various misalignment of the magnetic field direction compared to the [100] diamond crystalline axis for sample [REF].
+\ref{Alignment} shows the change in PL with respect to the external magnetic field amplitude for various misalignment of the magnetic field direction compared to the [100] diamond crystalline axis for sample CVD-PPM.
 We can see that the central drop in PL is not critically affected by the misalignment of the magnetic field, and we are therefore confident that the drop in PL we observe when $B\parallel [100]$ is not an effect of misalignment.
@@
 Fig.
-\ref{121 VS 22 fig}(b) shows the subtracted $T_1$ signal measured on the same sample [REF] for 8 different values of the magnetic field, each time on a two-class resonance : in 4 cases the magnetic field was in a \{100\} plane, and in the 4 other cases it was in a \{110\} plane.
+\ref{121 VS 22 fig}(b) shows the subtracted $T_1$ signal measured on the same sample HPHT-15-2 for 8 different values of the magnetic field, each time on a two-class resonance : in 4 cases the magnetic field was in a \{100\} plane, and in the 4 other cases it was in a \{110\} plane.
 We can see that the measured lifetime is always smaller on the \{110\} case, which corresponds the the greater $\bar \eta$ factor computed previously.
@@
 The first one is described in Fig.
-[REF] of the main text : in the presence of pure transverse magnetic field (for a single class, non-resonant with the three other classes), the decay rate of the class decreases with the magnetic field amplitude, due to the double-flip processes (described below), and then reaches a plateau with a decay rate value $\sim 2$ times bigger than in the longitudinal magnetic field case.
+4 of the main text : in the presence of pure transverse magnetic field (for a single class, non-resonant with the three other classes), the decay rate of the class decreases with the magnetic field amplitude, due to the double-flip processes (described below), and then reaches a plateau with a decay rate value $\sim 2$ times bigger than in the longitudinal magnetic field case.
 This increase is in agreement with the fact that $\bar{\eta}_{\rm same}$ is 2 times bigger in the non-magnetic $\ket{+/-}$ basis than in the magnetic $\ket{\pm 1}$ basis.
@@
 Fig.
-[REF] of the main text shows that we indeed observe a shorter lifetime when $\bm B=0$, in agreement with the higher value of $\bar \eta^2$, but this decrease could also come from the double-flip processes which are absent when $\bm B \neq 0$.
+3 of the main text shows that we indeed observe a shorter lifetime when $\bm B=0$, in agreement with the higher value of $\bar \eta^2$, but this decrease could also come from the double-flip processes which are absent when $\bm B \neq 0$.
 
@@
 Fig.
-[REF] of the main text shows that for lower transverse field values, the spin decay rate is increased significantly.
+4 of the main text shows that for lower transverse field values, the spin decay rate is increased significantly.
 We attribute this increase to double-flip processes.
@@
 This lead us to believe that the double-flip process are the main reason behind the decrease in spin lifetime for $\bm B=0$ observed in Fig.
-[REF] of the main text.
+4 of the main text.
 This would probably mean that double-flip processes are the dominant factor in the sensitivity of the magnetometry protocol presented in the main text.

### 2022-06-29 data/20220324/Adamas 150 um/adamas 1/traitement.py

@@ -6,7 +6,7 @@ from analyse import *
 # Pour la figure en fonction du décalage en fréquence, voir le traitement.py dans T1/
 
 
-plt.figure(num=1,figsize=(2,4/3),dpi=80)
+plt.figure(num=1,figsize=(6,3),dpi=80)
 plt.xticks(fontsize=12)
 plt.yticks(fontsize=12)
 

### 2022-07-05 Manuscrit Thèse/chapter 2/Appendix dipole-dipole.tex
tata

\documentclass[a4paper]{article}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage[
    %backend=biber, 
    natbib=true,
    style=numeric,
    sorting=none %Pour faire apparaitre les refs dans l'ordre
]{biblatex} %Imports biblatex package
\addbibresource{Bib_ch2.bib} %Import the bibliography file

\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{Appendix A : Dipole-dipole interaction between two spins in solid}
\begin{document}
The dipole-dipole interaction between two point-like magnetic dipoles can be derived classically by computing the static magnetic field generated by one dipole at the position of the second dipole.

For a magnetic dipole \textbf{m}, positioned on \textbf{0}, the static magnetic field generated on position $\mathbf{r}=r\mathbf{u}$ is : \cite[p.~188]{jackson1999classical} 
\begin{equation}
\mathbf{B}(\mathbf{r})=\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}\cdot\mathbf{u})\mathbf{u} - \mathbf{m}}{r^3}+\frac{8\pi}{3}\mathbf{m}\,\delta(\mathbf{r})\right]
\end{equation}

The interaction energy between two dipole $\mathbf{m}_1$ and $\mathbf{m}_2$ with relative position \textbf{r} is therefore :

\begin{equation}
U=-\mathbf{m}_1 \cdot \mathbf{B}_2(\mathbf{r})=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u}) - \mathbf{m}_1\cdot\mathbf{m}_2}{r^3}+\frac{8\pi}{3}\mathbf{m}_1\cdot\mathbf{m}_2\,\delta(\mathbf{r})\right]
\end{equation}

The term in $\delta$, which comes from the necessity to preserve $\div{\vb B}=0$, is at the origin of the Fermi contact energy. This term plays a role for hyperfine coupling when there is an overlap of the electron wave-function with the nucleus, which is the case for s-orbitals in atoms. In the case of NV centers, Fermi contact energy plays an important role to the hyper-fine coupling between the NV$^-$ electron spin and the nuclear spin of the $^{14}$N or $^{15}$N atom \cite{doherty2012theory} or for nearby $^{13}$C atoms \cite{smeltzer201113c}.

However, in the scope of this manuscript, we will mostly concern ourselves with dipolar interaction between electronic spins separated by several atomic sites, so that the overlap of the electronic wave functions can be mostly neglected. This results in the simplified dipole-dipole coupling energy :

\begin{equation}
\label{eq_dipole_dipole_classical}
U=\frac{\mu_0}{4 \pi}\left[ \frac{ \mathbf{m}_1\cdot\mathbf{m}_2 - 3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u})}{r^3}\right]
\end{equation}

For two electronic spins, we can rewrite a quantized version of \ref{eq_dipole_dipole_classical} by replacing the magnetic moments $\mathbf{m}_i=\hbar \gamma_i \hat{\mathbf{S}}_i$ where $\gamma_i=-\frac{g_e \mu_B}{\hbar}$ is the gyromagnetic ratio of the electron. For the NV$-$ electronic spin, as well as most electronic spin defects in diamond, $g_e\approx 2$ which gives a numerical value for the gyromagnetic ratio $\gamma \approx (2\pi)\,28\ \rm{GHz/T} \approx (2\pi)\,2.8\ \rm MHz/G$. $\hat{\mathbf{S}}_i=(\hat S_x, \hat S_y, \hat S_z)$ is the spin vector operator for particle $i$. The quantized version of the dipole-dipole interaction is then:

\begin{equation}
\mathcal{H}_{dd}=\frac{J_0}{r^3}\left[ \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2 - 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u})\right]
\end{equation}

Where $J_0\equiv \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$. For two electronic spins with $g-$factors close to 2, the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$
\printbibliography
\end{document}

### 2022-07-08 Manuscrit Thèse/chapter 2/Appendix dipole-dipole.tex
la tete au weekend

@@
-\documentclass[a4paper]{article}
+\documentclass[a4paper]{book}
 \usepackage[]{amsmath}
@@
 \usepackage{graphicx} %Pour les figures je crois
+\usepackage{hyperref}
 \usepackage[
@@
     style=numeric,
-    sorting=none %Pour faire apparaitre les refs dans l'ordre
+    sorting=none, %Pour faire apparaitre les refs dans l'ordre
+    hyperref=true
 ]{biblatex} %Imports biblatex package
@@
 \begin{document}
+\section{The dipole-dipole Hamiltonian}
 The dipole-dipole interaction between two point-like magnetic dipoles can be derived classically by computing the static magnetic field generated by one dipole at the position of the second dipole.
@@
 For two electronic spins, we can rewrite a quantized version of \ref{eq_dipole_dipole_classical} by replacing the magnetic moments $\mathbf{m}_i=\hbar \gamma_i \hat{\mathbf{S}}_i$ where $\gamma_i=-\frac{g_e \mu_B}{\hbar}$ is the gyromagnetic ratio of the electron.
-For the NV$-$ electronic spin, as well as most electronic spin defects in diamond, $g_e\approx 2$ which gives a numerical value for the gyromagnetic ratio $\gamma \approx (2\pi)\,28\ \rm{GHz/T} \approx (2\pi)\,2.8\ \rm MHz/G$.
+For the NV$^-$ electronic spin, as well as most electronic spin defects in diamond, $g_e\approx 2$ which gives a numerical value for the gyromagnetic ratio $\gamma \approx (2\pi)\,28\ \rm{GHz/T} \approx (2\pi)\,2.8\ \rm MHz/G$.
 $\hat{\mathbf{S}}_i=(\hat S_x, \hat S_y, \hat S_z)$ is the spin vector operator for particle $i$.
@@
 
-Where $J_0\equiv \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$. For two electronic spins with $g-$factors close to 2, the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$
+Where $J_0\equiv \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$. For two electronic spins with $g-$factors close to 2, the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$.
+
+\section{Flip-Flops, double flips and cross-relaxations}
+
+Flip-flop, in the context of dipole-dipole interaction, is the mechanism which exchanges a quantum of spin between two spins.
+For an initial state of two spins $\ket{m_s^1=i;m_s^2=j}$, a flip-flop process will result in a final state $\ket{i+1;j-1}$ or $\ket{i-1;j+1}$.
+The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$
+
 \printbibliography

### 2022-07-08 Mes papiers/Low field CR/V2/Low_field_CR.tex
la tete au weekend

--- 2022-06-22 Mes papiers/Low field CR/Low_field_CR.tex
@@
 
-When the spin ensemble becomes dense, inhomogeneities in the relaxation rates of NVs can give rise to spin depolarisation with a very rich many body dynamics associated with disorder.
+When the spin ensemble becomes dense, inhomogeneities in the relaxation rates of NVs can give rise to spin depolarisation with a very rich many body dynamics associated with disorder \citep{choi_observation_2017}.
 This mechanisms however limits the efficiency of magnetometers \citep{zhou2020quantum}
@@
 This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi2017depolarization}.
-Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble.
+Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble \citep{choi2017depolarization}.
 We will denote $T_1^{\rm ph}$ the characteristic timescale associated with the phonon relaxation process and $T_1^{\rm dd}$ the density dependent timescale associated with the dipole-dipole relaxation process.
@@
 
-Fig. \ref{T1}-b) shows the subtracted signals when $\bm B=0$ and $\bm B \neq 0$.
+Fig. \ref{T1}-b) shows the subtracted signals when $B\approx 0$ and $B \approx 50\ \rm G$.
 The ensemble spin decay curves are not exponential, which was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi2017depolarization}.
 We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi2017depolarization}.
+This model postulates the existence of very short-lived NV centers - the so-called fluctuators - which, when coupled resonantly to the other NV centers, act as a sink of polarization for the spin ensemble.
 A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$.
@@
 We observe that the 3 \% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.08 kHz, as the magnetic field increases.
-This result can be explained by the higher number of resonant spins in zero magnetic field, where all four classes are degenerate, as opposed when $\bm B \neq 0$.
-The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SI ?) [Je sais pas, je le fais ?].
-Note that the decrease in PL for $B>40\ \rm G$ are related to state mixing by the transverse magnetic field, so it is not correlated to a modification of the spin lifetime.
+This result can be explained by the higher number of resonant spins in zero magnetic field, where all four classes are degenerate, as opposed to $\bm B \neq 0$.
+The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SI).
+Note that the decrease in PL for $B>40\ \rm G$ is related to state mixing by the transverse magnetic field, so it is not correlated to a modification of the spin lifetime.
 
@@
 Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$.
-In situations such as the one described in Fig.XX, the only relevant terms of $\mathcal{H}^{\rm dd}$ in terms of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$.
+In situations such as the one described in Fig.
+\ref{100_VS_1x4} (c)(i), the only relevant terms of $\mathcal{H}^{\rm dd}$ in terms of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$.
 In zero magnetic field however, we have to take into account these two other mechanisms of the dipolar Hamiltonian in consideration.
 %where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.
-It was shown in  \cite{mittiga2018imaging} that local electric fields coming from charge traps such as $P_1^+$ and $NV^-$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in large NV ensembles.
+It was shown in  \cite{mittiga2018imaging} that local electric fields coming from charge traps such as $P_1^+$ and NV$^-$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in large NV ensembles.
 Magnetic field noise comes in to second order in zero field.
@@
 Another mechanism at stake when considering dipolar interactions with spin-1 systems are double-flip processes.
-These processes correspond to terms like $\ket{+1,0} \bra{0,-1}$ in the anisotropic (?) dipolar interaction, giving rise to an exchange of two units of spin-angular momentum.
+These processes correspond to terms like $\ket{+1,0} \bra{0,-1}$ in the anisotropic [RQ CLEMENT : pourquoi anisotropic ?] dipolar interaction, giving rise to an exchange of two units of spin-angular momentum.
 These processes can become significant when the states $\pm 1$ are degenerate and are thus likely to explain part of the zero-field dip in Fig.
@@
 (b) Blue dashed curve : frequency detuning $\Delta \nu$ between the $\ket{e}$ and $\ket{d}$ states.
-Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\bra{e}$ and $\bra{+}$.
-(c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single class as a function of the measured detuning $\Delta \nu$ between the two transitions $\ket{g} \to \ket {d}$ and  $\ket{g} \to \ket {e}$ for purely transverse magnetic field.
-The green dashed line correspond to the decay rate for the same class when the magnetic field is longitudinal}
+Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\ket{e}$ and $\ket{+}$.
+(c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single class of sample HPHT-150-2 as a function of the amplitude of a purely transverse magnetic field.
+The red dashed line corresponds to the plateau reached for high transverse field.
+The green dashed line correspond to the value found for the same sample with a longitudinal magnetic field, with error bars.
+The corresponding detuning between the states $\ket{d}$ and $\ket{e}$ is indicated on top.}
 \label{B_transverse}

### 2022-07-08 Mes papiers/Low field CR/V3/Low_field_CR.tex
la tete au weekend

\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-2}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{physics}
\usepackage[caption=false]{subfig}

\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{Depolarization of interacting NV$^-$ centers at low magnetic fields}

\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, P. Huillery$^2$,  G. H\'etet}

\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e Paris Cit\'e , 24 rue Lhomond, 75231 Paris Cedex 05, France \\ $^2$ Univ Rennes, INSA Rennes, CNRS, Institut FOTON - UMR 6082, F-35000 Rennes, France}

\begin{abstract}
We study dipolar interaction processes in dense ensembles of negatively charged nitrogen vacancy (NV$^-$) centers under small magnetic fields. Employing magnetic field scans close to zero-field along specific crystalline direction while recording the NV$^-$ spin relaxation rate, we could identify regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the NV-NV cross-relaxation .
Our results are relevant for understanding decoherence in many-body spin systems as well as for high sensitivity magneto- and electro-metry with long-lived interacting solid-state spins. As a proof of principle, we present an orientation and microwave-free magnetometer based on cross-relaxation, and operating with a sensitivity below $100$ nT/$\sqrt{\rm Hz}$.
\end{abstract}

\maketitle

The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science thanks in part to the possibility to optically polarize and read-out its spin state at ambient conditions \cite{DOHERTY20131}. In particular, ensembles of NV centers are widely studied for their enhanced magnetic field sensing capabilities \cite{Acosta, TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized} and as pristine platforms for studying many body effects \citep{kucsko2018critical, ChoiNat, ZuYao}. When the NV spin concentration reaches ppm values, spin depolarisation, or cross-relaxation (CR) takes place through a very rich many body dynamics associated with disorder \citep{choi_observation_2017}. These mechanisms however limit the efficiency of typical microwave based NV magnetometers, but can be harnessed to attain sub-heisenberg limited magnetometers \citep{zhou2020quantum}. Further, CR mechanisms can be turned as a tool for magnetic field sensing \cite{akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019}. 
Spectral features in the photoluminescence (PL) indeed appear when the magnetic field crosses specific crystal planes where dipolar interactions are enhanced, leading to CR. The projected sensitivity of such magnetometers lie in the tens of pT$/\sqrt{\rm Hz}$  \cite{akhmedzhanov_microwave-free_2017}, on a par with the most sensitive microwave based NV magnetometers \cite{Wolf, Sturner}.

CR features at zero-field have also been observed and could also be deployed for high sensitivity microwave free magnetometry.
The CR contrast was indeed shown to be much stronger \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}. However all the relaxation mechanisms at these low field levels have not been identified. 
Here, we study dipolar interaction processes in ensembles of NV centers under small magnetic fields. Specifically, by employing magnetic field scans along specific crystalline directions, we identify regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the CR. We also present an orientation and microwave-free magnetometer operating with a sensitivity below $100$ nT/$\sqrt{\rm Hz}$ at zero-field, where double-flip processes amongst NV centers dominate.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/shema_summary.pdf}
\caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation :flip-flop processes involving two different classes of NV centers, local electric field mixing and double-flip processes where two units of spin angular momentum are exchanged. }
\label{schema_intro}
\end{figure}

The electronic spin of NV center is a spin-1 system in the electronic ground state (see Fig. 1, top left panel), which can be optically polarized in the $\ket{m_s=0}$ state. The PL of this state is also larger than in the $\ket{m_s=\pm 1}$ states enabling spin-read out at ambient temperature \cite{DOHERTY20131}. The $\ket{m_s=\pm 1}$ spin states are separated from the $\ket{m_s=0}$ state by  $D = (2\pi) 2.87$ GHz so that when a resonant microwave or static transverse magnetic field is applied, the PL is reduced \citep{epstein2005anisotropic,lai2009influence}.  The processes that lead to CR in such a spin-1 system are depicted in Fig.~1. Flip-flop processes involving coupling between NV centers from identical or from different orientations are depicted in the top-right panel. They have already been shown to play a dominant role when tuning $\delta$ {\it via} longitudinal B fields scans, when $B \gtrsim 30$G \cite{choi_observation_2017}. We will show here that under smaller magnetic fields or using purely transverse fields, double-flip up and down as well as the mixing induced by local electric field give a significant contribution to the CR (two bottom panels).

Fig.  \ref{T1}-a) and b) show the change in PL as a function of an external magnetic field for two samples with a low and high concentration of NV$^-$ centers respectively, detailed in the supplementary material (SM sec. II) of this article \citep{SI_low_filed_CR} \nocite{anishchik2015low, filimonenko2020weak, van1990electric}. Both samples show a decrease in PL as the magnetic field amplitude increases.  There is however a stark difference in the low magnetic field region where only high-density samples shows a drop in PL \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}. This effect can be observed on all our samples whenever the NV concentration lies in the ppm range. The drop of the PL at low magnetic field was shown to be associated with a decrease of the NV's spin lifetime $T_1$ \citep{jarmola_temperature-_2012}, which results in a decrease of the population in the bright state $\ket{0}$.

The depolarization dynamics of the NV spins for single or dilute ensembles of NV centers at room temperature is dominated by two-phonon Raman processes \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depend on the crystal lattice temperature. It has been observed that dense NV centers ensembles in fact  have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi2017depolarization, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field amplitude. This effect has been attributed to CR between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi2017depolarization}. 
The reason is that at zero-field many NV ``classes" - that is physical orientation of the NV axis is the diamond crystal cell - becomes resonant, which increases the dipolar relaxation.
Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble \citep{choi2017depolarization}. We will denote $T_1^{\rm ph}$ the characteristic timescale associated with the phonon relaxation process and $T_1^{\rm dd}$ the density dependent timescale associated with the dipole-dipole relaxation process.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig_T1.pdf}
\caption{a) PL measurement from NV centers ensemble as a function of an external magnetic field applied in an arbitrary direction (a) for the sample CVD-PPB, with [NV$^-$]$\approx 50\ \rm ppb$ (b) for the sample HPHT-150-1, with [NV$^-$]$\approx 3\ \rm ppm$. (c) Sequence used to measure the spin lifetime. (d) Spin relaxation $S_1-S_2$ measured for the dense sample at zero and non-zero magnetic fields. The fitting procedure (orange plain line) is detailed in the main text}
\label{T1}
\end{figure}

\begin{figure*}
\includegraphics[width=.95\textwidth]{Figures/fig_100_vs_1x1x1x1.pdf}
\caption{(a) ODMR spectrum in zero field. (b)-i) ODMR spectrum for a magnetic field $\approx$ 60 G, misaligned by $\sim  24^\circ$ from the [100] axis. (b)-ii) Normalized PL of the NV$^-$ ensemble as a function of the magnetic field amplitude. (b)-iii) Stretched part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude. (c)-i), (c)-ii) and (c)-iii) :same measurements but with a magnetic field close to the [100] axis. All the measurements were realized using the sample HPHT-150-1. }
\label{100_VS_1x4}
\end{figure*}

Before analyzing in detail the cause of low-field depolarization, we present $T_1$ measurements realized on HPHT-150-1. Fig. \ref{T1}-c) shows the sequence employed for measuring $T_1^{\rm dd}$. 
It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$. In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019}. It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right on one of the eight NV spin-resonances before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state. By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi2017depolarization}.

Fig. \ref{T1}-d) shows the subtracted signals $S_1-S_2$ when $B\approx 0$ and $B \approx 50\ \rm G$.
The relaxation rate is markedly different in both magnetic field settings, with a much more pronounced depolarization at zero-field.
Notably, the ensemble spin decay curves are not exponential. This was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi2017depolarization}.
We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi2017depolarization}. This model postulates the existence of very short-lived NV centers - the so-called fluctuators - which, when coupled resonantly to the other NV centers, act as a polarization drain for the spin ensemble. A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced relaxation should be a stretched-exponential with a stretch factor $\beta=1/2$. Such a trend was indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$ \cite{choi2017depolarization}. We conducted further measurements to confirm the stretch nature of the dipole-induced decay (see SM sec. IV A), as well as confirming the lifetime-limited model of the fluctuators by measuring the resonance condition between classes of NV centers (see SM sec. IV B).
In our case however, all dense samples used in this paper (detailed in SM sec. II) show $T_1^{\rm dd} \sim T_1^{\rm ph}$, so both decay processes have to be included in the analysis. Fitting the two curves in Fig. 2 d) by the product of a simple and a stretched exponential, we find $T_1^{\rm ph}=3.6$ ms for both curves and $T_1^{\rm dd}= 0.6\ \rm ms$ and $13.0\ \rm ms$ for the $B=0$ and $B=50$G cases respectively. These results thus demonstrates a twenty-fold increase in the dipolar depolarization rate when the B field is turned to zero.
While the fluctuator model developed in \citep{choi2017depolarization} contains some of the explanation for this increase in dipolar depolarization, due to more classes being resonant in zero field than in non-zero field, the model did not include the specificity of the zero magnetic field region, such as the importance of local electric field and the double flip processes. We therefore decided to conducted more in depth experiments in order to characterize these zero field specificities.

Fig. \ref{100_VS_1x4} (a) shows an optically detected magnetic resonance (ODMR) spectrum at zero magnetic field, from the dense ensemble HPHT-150-1.
Fig. \ref{100_VS_1x4} (b)-i) shows an ODMR spectrum from the same sample, but in a magnetic field. The observed lines correspond to the projections of the magnetic fields onto the four classes of NV centers together with the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$. The transitions frequencies of the four different classes can be controlled by changing the amplitude and orientation of the external magnetic field. Fig. \ref{100_VS_1x4} (c)-i) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis. 
This configuration will be essential to decipher which dipolar relaxation mechanisms play a role at low field.

Fig. \ref{100_VS_1x4} (b)-ii) show the PL from the NV centers ensemble as a function of the amplitude of a magnetic field aligned as in Fig. \ref{100_VS_1x4} (b)-i). 
Fig. \ref{100_VS_1x4} (b)-iii) shows the stretched exponential spin decay time $T_1^{\rm dd}$ obtained using the sequence shown in Fig. 2-c) in the same range of magnetic field values, in 5 G steps. 
We observe that the 3~\% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.08 kHz, as the magnetic field increases. This result can again be explained by the higher number of resonant spins as the zero magnetic field is reduced, where all four classes are degenerate, as opposed to $\bm B \neq 0$. The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SM sec. IV C).
Note that the decrease in PL for $B>40\ \rm G$ is related to state mixing by the transverse magnetic field, so it is not associated with a modification of the spin lifetime.

Importantly, B field scans along the [100] direction, where the four classes are always resonant, still features an increase of the spin lifetime as the B field increases. Fig. \ref{100_VS_1x4} (c)-ii) and (c)-iii) show the change in the PL and in $1/T_1^{\rm dd}$ as a function of a magnetic field that is aligned in this direction. We can see that there still is an increase in the spin decay rate and a corresponding drop in the PL under low magnetic field values, although considerably reduced.

The main aim of this paper is to understand the physics governing the contribution to the spin decay in this second scenario. Note that the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20$G is related to dipolar interaction with NV centers that have a $^{13}$C as a first neighbor \cite{pellet2021optical}.

We identify and isolate two additional mechanisms that could explain this effect :local electric fields and double-flip processes. 
 In situations such as the one described in Fig.~\ref{100_VS_1x4} (c)-i), the only relevant terms for depolarisation in $\mathcal{H}^{\rm dd}$ are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$ where $\mathcal{H}^{\rm dd}$ is the dipole-dipole Hamiltonian detailed in SM sec. V. In zero magnetic field however, we have to take into account two other mechanisms of the dipolar Hamiltonian in consideration.

It was shown in  \cite{mittiga2018imaging} that local electric fields coming from $P_1^+$ or NV$^-$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in dense NV ensembles. Magnetic field noise comes in to second order in zero field. 
We will then consider the following spin Hamiltonian for the electric field dependent NV$^-$ ground state: 
\begin{equation}
\label{NV Hamiltonian}
\mathcal{H}_{\rm elec}= d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
\end{equation}
where $E_{x,y}$ are the projections of local electric fields on the NV axes. 
Under local electric fields with orientations given by the angle $\phi_E={\rm tan}(E_x/E_y)$, the eigenstates of $H_{NV}$ are $\ket{0}$ and $\ket{\pm}=\frac{1}{\sqrt{2}}(\ket{+1}\pm e^{-i\phi_E}\ket{-1})$.
Computing the flip-flop terms $\ket{\pm,0} \bra{0,\pm} $ in the fluctuator model shows that $1/T_1^{\rm dd}$ should be increased at zero field, even after averaging over all possible orientations $\phi_E$ (see SM sec. V).  Another mechanism at stake when considering dipolar interactions with spin-1 systems are double-flip processes. 
These processes correspond to terms like $\ket{+1,0} \bra{0,-1}$ in the dipolar interaction, giving rise to an exchange of two units of spin-angular momentum. These processes can become significant when the states $\ket{m_s=\pm 1}$ are degenerate and are thus likely to explain part of the zero-field dip in Fig. 3-c)-ii).

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig_transverse_field_V3.pdf}
\caption{(a) Energies of the electronic spin states $\ket{g}$, $\ket{d}$ and $\ket{e}$ under a purely transverse magnetic field $B_\perp$. (b) Blue dashed line :frequency detuning $\Delta \nu$ between the $\ket{e}$ and $\ket{d}$ states. Red plain line :Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\ket{e}$ and $\ket{+}$. (c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single NV class as a function of the amplitude of $B_\perp$. The red dashed line is the value reached for high transverse field. The green dashed line correspond to the value found for the same sample under a longitudinal magnetic field. Error bars are depicted by the shaded pink region. The corresponding detuning between the states $\ket{d}$ and $\ket{e}$ is indicated on the top x-axis. The vertical black dashed line indicates the separation between regions A and B (see main text).}
\label{B_transverse}
\end{figure}

In order to discriminate the role of these two effects in the low-field region, we make use of a purely transverse magnetic field. The transverse B field emulates the effect of the electric field, while turning off the resonance condition for the effectiveness of double-flip processes. Indeed for relatively weak transverse magnetic field, the eigen-basis of the spin Hamiltonian is close to $\ket{0},\ket{+},\ket{-}$. This property has already been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
Fig.~\ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian as a function of a purely transverse magnetic field. We will denote these eigenstates in the general case $\ket{g}$, $\ket{d}$ and $\ket{e}$. For weak transverse magnetic field, $\ket{g} \approx \ket{0}$, $\ket{d}=\ket{-}$ and $\ket{e} \approx \ket{+}$. As the magnetic field increases, $\ket{g}$ and $\ket{e}$ starts to become a mixing of $\ket{0}$ and $\ket{+}$ while $\ket{d}$ remains equal to $\ket{-}$.

Crucially, mixing remains even when the associated energy splitting of the $\ket{d}$ and $\ket{e}$ states is large :Fig. \ref{B_transverse}-(b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$. We can see that at a field value of $\sim 130$ G, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50$ MHz, exceeding the range of the dipole-dipole resonant coupling (see SM sec. IV B). Yet, $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis. We can therefore decouple the effects of the double-flip processes  $\propto \ket{g,d}\bra{d,e}$, which are no longer resonant when $| \bm B | \geq 100$ G, from the effects of the mixing induced by the transverse electric - or in this case magnetic - field.

Fig. \ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers on sample HPHT-150-2 exposed to a transverse magnetic field between 20 and 130 G.  The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR. Beforehand, we measured a stretched decay $1/T_1^{\rm dd}=25\pm 5\  \rm Hz$ under strong longitudinal magnetic field. This value is indicated by the dashed green line. Two regions can be identified in Fig. \ref{B_transverse}-(c). In  region A, the spin decay rate decreases from 175 Hz to 60 Hz.  We attribute this effect to the loss of effectiveness of the double-flip processes. In region B, $\Delta \nu \geq 30\ \rm MHz$, the decay rate plateaus to $1/T_1^{\rm dd}=57\pm 5\  \rm Hz$, which is about twice as large as the decay rate measured under a longitudinal magnetic field. This plateau can be attributed to the change of eigenstate basis in the presence of transverse magnetic field, which is the same as under low magnetic field, where local transverse electric fields dominates the spin Hamiltonian (see SM sec. I). From these measurements, we conclude that the effect of the double-flip processes is about 5 times more important than the effect of the electric field in the depolarization rate.

Based on an extension of the fluctuator model developed in SM sec. V, we predict an overall increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$. While the measured increase is smaller than the predicted one, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis. Similarly, we compute the predicted decay rate when $\bm B \parallel \left[100\right]$ case compared when $\bm B = 0$, and find a theoretical increase of $\sim 20\%$ when $\bm B = 0$ due to the mixing by the local electric field. This, along with the double-flip processes explains the behavior observed in Fig. \ref{100_VS_1x4} (c)-ii) and (c)-iii).

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig_magneto}
\caption{Low-field magnetometry protocol. (a) Demodulated PL as a function of an externally applied magnetic field with an additional oscillatory magnetic field. (b) Measured magnetic field when alternating a small external magnetic field offset. (c) Histogram of the measurement in Fig. (b) fitted by gaussians of standard deviation $\sigma=1.5\ \mu \rm T$. (d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.}
\label{magneto}
\end{figure}

Our observations have important implications for magnetometry with NV ensembles.
DC microwave-free magnetometry has already been performed using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}. Here we propose to perform a similar protocol, but using the spin depolarization at zero-field as proposed in \cite{filimonenko2018weak, filimonenko2022manifestation}. The main advantage compared to previously employed protocols is that the sensitivity does not depend crucially on crystalline orientation, making this protocol applicable with diamond powders or polycrystalline samples. Another sizable advantage is that this measurement is that it is insensitive to thermal variations. In order to optimize signal-to-noise ratio, we use a lock-in detection and add a magnetic field modulation at $\sim 1\ \rm kHz$ with an amplitude $\sim 10\ \rm G$ through the same electromagnet. Fig. \ref{magneto}-(a) shows the demodulated PL as a DC magnetic field is scanned in an arbitrary direction. The sample used here is HPHT-15-1, the laser power $\sim 1\ \rm mW$ and the collected PL is $\sim 1\ \mu\rm W$.  We can see a sharp linear slope in low field $\abs{B} < 5\ \rm G$. Once calibrated, in this case with ODMR, the slope provides a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as in \cite{zheng_microwave-free_2020}. In order to assess the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take a histogram of the measured fields (46.000 counts in 46 s), as shown in Fig. \ref{magneto} (b) and (c). The histogram is well fitted by gaussians of standard deviation $\sigma=1.5\ \mu \rm T$. The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which gives us a DC sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nT}/\sqrt{\rm Hz}$. Normalized to the volume, this gives $\eta/\sqrt{V}\approx 4.7 \ \mu \rm{T}/\mu m^{3/2}/\sqrt{\rm Hz}$. This measurement is consistent with the experimentally found $\sim 5\ \mu \rm{T}/\sqrt{\rm Hz}$ sensitivity obtained with sample HPHT-1-1 (see samples details in SM sec. II).

We now evaluate the relative role of the three causes of spin depolarization that assist the operation of the magnetometer, namely the splitting of the NV classes, the local electric field and the double-flip processes. In order to do so, we measure the magnetometer sensitivity while changing the angle of the magnetic field.  The results are shown in Fig. \ref{magneto} (d). We can see a $\sim$ 10\% declining sensitivity as we leave the [100] region, but overall the sensitivity remains relatively invariant. 
When $\bm B$ is aligned with the [100] crystalline axis, only the double flips and the electric field cause a depolarization, whereas in every other orientation the three effects are at play.
The double-flips and electric field effects are thus the dominant factors in the sensitivity of this protocol. While it may seem surprising that the effects with a lower contribution to the PL contrast have a higher effect on this PL-based protocol, what matters here is not the absolute contrast but the slope of the change of contrast with respect to the magnetic field. The latter is in fact larger for the local electric field and double flips processes than for the process involving only flip-flops. 
It should be noted that this observation is sample dependent, and that other samples, including from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from the electric field and double-flip processes.

As a conclusion, we identified three mechanisms causing extra spin depolarization in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole induced CR between the spins of NV centers. The lift in degeneracy of the spin state of different NV classes was found to be the main cause of zero-field depolarization, followed by double-flip processes and then the electric field induced mixing. We have employed CR for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $15\ \mu \rm m$ commercially available diamond and show that that double-flips and electric field play an important role.
Our studies will be important for microwave based low field magnetometry \cite{Vetter_LFM, WangRB}. 
Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins under purely transverse as well as small magnetic fields.

\section*{acknowledgements}

We would like to acknowledge support from Alexandre Tallaire and Jocelyn Achard as well as 
SIRTEQ for funding.

%\bibliographystyle{unsrtnat}
%\bibliographystyle{apsrev4-2}
%\bibliography{CR}{}

%apsrev4-2.bst 2019-01-14 (MD) hand-edited version of apsrev4-1.bst
%Control: key (0)
%Control: author (72) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (-1) disabled
%Control: page (0) single
%Control: year (1) truncated
%Control: production of eprint (0) enabled
\begin{thebibliography}{41}%
\makeatletter
\providecommand \@ifxundefined [1]{%
 \@ifx{#1\undefined}
}%
\providecommand \@ifnum [1]{%
 \ifnum #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \@ifx [1]{%
 \ifx #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \natexlab [1]{#1}%
\providecommand \enquote  [1]{``#1''}%
\providecommand \bibnamefont  [1]{#1}%
\providecommand \bibfnamefont [1]{#1}%
\providecommand \citenamefont [1]{#1}%
\providecommand \href@noop [0]{\@secondoftwo}%
\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
\providecommand \@href[1]{\@@startlink{#1}\@@href}%
\providecommand \@@href[1]{\endgroup#1\@@endlink}%
\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
\providecommand \@@startlink[1]{}%
\providecommand \@@endlink[0]{}%
\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
\providecommand \urlprefix  [0]{URL }%
\providecommand \Eprint [0]{\href }%
\providecommand \doibase [0]{https://doi.org/}%
\providecommand \selectlanguage [0]{\@gobble}%
\providecommand \bibinfo  [0]{\@secondoftwo}%
\providecommand \bibfield  [0]{\@secondoftwo}%
\providecommand \translation [1]{[#1]}%
\providecommand \BibitemOpen [0]{}%
\providecommand \bibitemStop [0]{}%
\providecommand \bibitemNoStop [0]{.\EOS\space}%
\providecommand \EOS [0]{\spacefactor3000\relax}%
\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
\let\auto@bib@innerbib\@empty
%</preamble>
\bibitem [{\citenamefont {Doherty}\ \emph {et~al.}(2013)\citenamefont
  {Doherty}, \citenamefont {Manson}, \citenamefont {Delaney}, \citenamefont
  {Jelezko}, \citenamefont {Wrachtrup},\ and\ \citenamefont
  {Hollenberg}}]{DOHERTY20131}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~W.}\ \bibnamefont
  {Doherty}}, \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont {Manson}},
  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Delaney}}, \bibinfo
  {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Wrachtrup}},\ and\ \bibinfo {author}
  {\bibfnamefont {L.~C.}\ \bibnamefont {Hollenberg}},\ }\href
  {https://doi.org/https://doi.org/10.1016/j.physrep.2013.02.001} {\bibfield
  {journal} {\bibinfo  {journal} {Physics Reports}\ }\textbf {\bibinfo {volume}
  {528}},\ \bibinfo {pages} {1 } (\bibinfo {year} {2013})},\ \bibinfo {note}
  {the nitrogen-vacancy colour centre in diamond}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Acosta}\ \emph {et~al.}(2009)\citenamefont {Acosta},
  \citenamefont {Bauch}, \citenamefont {Ledbetter}, \citenamefont {Santori},
  \citenamefont {Fu}, \citenamefont {Barclay}, \citenamefont {Beausoleil},
  \citenamefont {Linget}, \citenamefont {Roch}, \citenamefont {Treussart},
  \citenamefont {Chemerisov}, \citenamefont {Gawlik},\ and\ \citenamefont
  {Budker}}]{Acosta}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {V.~M.}\ \bibnamefont
  {Acosta}}, \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {Bauch}},
  \bibinfo {author} {\bibfnamefont {M.~P.}\ \bibnamefont {Ledbetter}}, \bibinfo
  {author} {\bibfnamefont {C.}~\bibnamefont {Santori}}, \bibinfo {author}
  {\bibfnamefont {K.-M.~C.}\ \bibnamefont {Fu}}, \bibinfo {author}
  {\bibfnamefont {P.~E.}\ \bibnamefont {Barclay}}, \bibinfo {author}
  {\bibfnamefont {R.~G.}\ \bibnamefont {Beausoleil}}, \bibinfo {author}
  {\bibfnamefont {H.}~\bibnamefont {Linget}}, \bibinfo {author} {\bibfnamefont
  {J.~F.}\ \bibnamefont {Roch}}, \bibinfo {author} {\bibfnamefont
  {F.}~\bibnamefont {Treussart}}, \bibinfo {author} {\bibfnamefont
  {S.}~\bibnamefont {Chemerisov}}, \bibinfo {author} {\bibfnamefont
  {W.}~\bibnamefont {Gawlik}},\ and\ \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Budker}},\ }\href
  {https://doi.org/10.1103/PhysRevB.80.115202} {\bibfield  {journal} {\bibinfo
  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {80}},\ \bibinfo
  {pages} {115202} (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Tallaire}\ \emph {et~al.}(2020)\citenamefont
  {Tallaire}, \citenamefont {Brinza}, \citenamefont {Huillery}, \citenamefont
  {Delord}, \citenamefont {Pellet-Mary}, \citenamefont {Staacke}, \citenamefont
  {Abel}, \citenamefont {Pezzagna}, \citenamefont {Meijer}, \citenamefont
  {Touati}, \citenamefont {Binet}, \citenamefont {Ferrier}, \citenamefont
  {Goldner}, \citenamefont {Hetet},\ and\ \citenamefont
  {Achard}}]{TALLAIRE2020421}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Tallaire}}, \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Brinza}},
  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Huillery}}, \bibinfo
  {author} {\bibfnamefont {T.}~\bibnamefont {Delord}}, \bibinfo {author}
  {\bibfnamefont {C.}~\bibnamefont {Pellet-Mary}}, \bibinfo {author}
  {\bibfnamefont {R.}~\bibnamefont {Staacke}}, \bibinfo {author} {\bibfnamefont
  {B.}~\bibnamefont {Abel}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Pezzagna}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Meijer}},
  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Touati}}, \bibinfo
  {author} {\bibfnamefont {L.}~\bibnamefont {Binet}}, \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Ferrier}}, \bibinfo {author} {\bibfnamefont
  {P.}~\bibnamefont {Goldner}}, \bibinfo {author} {\bibfnamefont
  {G.}~\bibnamefont {Hetet}},\ and\ \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Achard}},\ }\href
  {https://doi.org/https://doi.org/10.1016/j.carbon.2020.08.048} {\bibfield
  {journal} {\bibinfo  {journal} {Carbon}\ }\textbf {\bibinfo {volume} {170}},\
  \bibinfo {pages} {421 } (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Edmonds}\ \emph {et~al.}(2021)\citenamefont
  {Edmonds}, \citenamefont {Hart}, \citenamefont {Turner}, \citenamefont
  {Colard}, \citenamefont {Schloss}, \citenamefont {Olsson}, \citenamefont
  {Trubko}, \citenamefont {Markham}, \citenamefont {Rathmill}, \citenamefont
  {Horne-Smith} \emph {et~al.}}]{edmonds2021characterisation}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
  {Edmonds}}, \bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont {Hart}},
  \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont {Turner}}, \bibinfo
  {author} {\bibfnamefont {P.-O.}\ \bibnamefont {Colard}}, \bibinfo {author}
  {\bibfnamefont {J.~M.}\ \bibnamefont {Schloss}}, \bibinfo {author}
  {\bibfnamefont {K.~S.}\ \bibnamefont {Olsson}}, \bibinfo {author}
  {\bibfnamefont {R.}~\bibnamefont {Trubko}}, \bibinfo {author} {\bibfnamefont
  {M.~L.}\ \bibnamefont {Markham}}, \bibinfo {author} {\bibfnamefont
  {A.}~\bibnamefont {Rathmill}}, \bibinfo {author} {\bibfnamefont
  {B.}~\bibnamefont {Horne-Smith}}, \emph {et~al.},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Materials for Quantum Technology}\ }\textbf
  {\bibinfo {volume} {1}},\ \bibinfo {pages} {025001} (\bibinfo {year}
  {2021})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Chatzidrosos}\ \emph {et~al.}(2021)\citenamefont
  {Chatzidrosos}, \citenamefont {Rebeirro}, \citenamefont {Zheng},
  \citenamefont {Omar}, \citenamefont {Brenneis}, \citenamefont {St{\"u}rner},
  \citenamefont {Fuchs}, \citenamefont {Buck}, \citenamefont {R{\"o}lver},
  \citenamefont {Schneemann} \emph {et~al.}}]{chatzidrosos2021fiberized}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
  {Chatzidrosos}}, \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
  {Rebeirro}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Zheng}},
  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Omar}}, \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Brenneis}}, \bibinfo {author}
  {\bibfnamefont {F.~M.}\ \bibnamefont {St{\"u}rner}}, \bibinfo {author}
  {\bibfnamefont {T.}~\bibnamefont {Fuchs}}, \bibinfo {author} {\bibfnamefont
  {T.}~\bibnamefont {Buck}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
  {R{\"o}lver}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
  {Schneemann}}, \emph {et~al.},\ }\href@noop {} {\bibfield  {journal}
  {\bibinfo  {journal} {Frontiers in Photonics}\ ,\ \bibinfo {pages} {4}}
  (\bibinfo {year} {2021})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Kucsko}\ \emph {et~al.}(2018)\citenamefont {Kucsko},
  \citenamefont {Choi}, \citenamefont {Choi}, \citenamefont {Maurer},
  \citenamefont {Zhou}, \citenamefont {Landig}, \citenamefont {Sumiya},
  \citenamefont {Onoda}, \citenamefont {Isoya}, \citenamefont {Jelezko} \emph
  {et~al.}}]{kucsko2018critical}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
  {Kucsko}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}},
  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo {author}
  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
  {\bibfnamefont {H.}~\bibnamefont {Zhou}}, \bibinfo {author} {\bibfnamefont
  {R.}~\bibnamefont {Landig}}, \bibinfo {author} {\bibfnamefont
  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
  {F.}~\bibnamefont {Jelezko}}, \emph {et~al.},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Physical review letters}\ }\textbf {\bibinfo
  {volume} {121}},\ \bibinfo {pages} {023601} (\bibinfo {year}
  {2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Choi}\ \emph
  {et~al.}(2017{\natexlab{a}})\citenamefont {Choi}, \citenamefont {Choi},
  \citenamefont {Landig}, \citenamefont {Kucsko}, \citenamefont {Zhou},
  \citenamefont {Isoya}, \citenamefont {Jelezko}, \citenamefont {Onoda},
  \citenamefont {Sumiya}, \citenamefont {Khemani}, \citenamefont {von
  Keyserlingk}, \citenamefont {Yao}, \citenamefont {Demler},\ and\
  \citenamefont {Lukin}}]{ChoiNat}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Choi}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo
  {author} {\bibfnamefont {R.}~\bibnamefont {Landig}}, \bibinfo {author}
  {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author} {\bibfnamefont
  {H.}~\bibnamefont {Zhou}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {Isoya}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},
  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Onoda}}, \bibinfo
  {author} {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author}
  {\bibfnamefont {V.}~\bibnamefont {Khemani}}, \bibinfo {author} {\bibfnamefont
  {C.}~\bibnamefont {von Keyserlingk}}, \bibinfo {author} {\bibfnamefont
  {N.~Y.}\ \bibnamefont {Yao}}, \bibinfo {author} {\bibfnamefont
  {E.}~\bibnamefont {Demler}},\ and\ \bibinfo {author} {\bibfnamefont {M.~D.}\
  \bibnamefont {Lukin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
  {journal} {Nature}\ }\textbf {\bibinfo {volume} {543}},\ \bibinfo {pages}
  {221} (\bibinfo {year} {2017}{\natexlab{a}})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Zu}\ \emph {et~al.}(2021)\citenamefont {Zu},
  \citenamefont {Machado}, \citenamefont {Ye}, \citenamefont {Choi},
  \citenamefont {Kobrin}, \citenamefont {Mittiga}, \citenamefont {Hsieh},
  \citenamefont {Bhattacharyya}, \citenamefont {Markham}, \citenamefont
  {Twitchen}, \citenamefont {Jarmola}, \citenamefont {Budker}, \citenamefont
  {Laumann}, \citenamefont {Moore},\ and\ \citenamefont {Yao}}]{ZuYao}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
  {Zu}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Machado}},
  \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Ye}}, \bibinfo {author}
  {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo {author} {\bibfnamefont
  {B.}~\bibnamefont {Kobrin}}, \bibinfo {author} {\bibfnamefont
  {T.}~\bibnamefont {Mittiga}}, \bibinfo {author} {\bibfnamefont
  {S.}~\bibnamefont {Hsieh}}, \bibinfo {author} {\bibfnamefont
  {P.}~\bibnamefont {Bhattacharyya}}, \bibinfo {author} {\bibfnamefont
  {M.}~\bibnamefont {Markham}}, \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Twitchen}}, \bibinfo {author} {\bibfnamefont
  {A.}~\bibnamefont {Jarmola}}, \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Budker}}, \bibinfo {author} {\bibfnamefont {C.~R.}\
  \bibnamefont {Laumann}}, \bibinfo {author} {\bibfnamefont {J.~E.}\
  \bibnamefont {Moore}},\ and\ \bibinfo {author} {\bibfnamefont {N.~Y.}\
  \bibnamefont {Yao}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
  {journal} {Nature}\ }\textbf {\bibinfo {volume} {597}},\ \bibinfo {pages}
  {45} (\bibinfo {year} {2021})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Choi}\ \emph
  {et~al.}(2017{\natexlab{b}})\citenamefont {Choi}, \citenamefont {Choi},
  \citenamefont {Landig}, \citenamefont {Kucsko}, \citenamefont {Zhou},
  \citenamefont {Isoya}, \citenamefont {Jelezko}, \citenamefont {Onoda},
  \citenamefont {Sumiya}, \citenamefont {Khemani}, \citenamefont {von
  Keyserlingk}, \citenamefont {Yao}, \citenamefont {Demler},\ and\
  \citenamefont {Lukin}}]{choi_observation_2017}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Choi}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo
  {author} {\bibfnamefont {R.}~\bibnamefont {Landig}}, \bibinfo {author}
  {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author} {\bibfnamefont
  {H.}~\bibnamefont {Zhou}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {Isoya}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},
  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Onoda}}, \bibinfo
  {author} {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author}
  {\bibfnamefont {V.}~\bibnamefont {Khemani}}, \bibinfo {author} {\bibfnamefont
  {C.}~\bibnamefont {von Keyserlingk}}, \bibinfo {author} {\bibfnamefont
  {N.~Y.}\ \bibnamefont {Yao}}, \bibinfo {author} {\bibfnamefont
  {E.}~\bibnamefont {Demler}},\ and\ \bibinfo {author} {\bibfnamefont {M.~D.}\
  \bibnamefont {Lukin}},\ }\href {https://doi.org/10.1038/nature21426}
  {\bibfield  {journal} {\bibinfo  {journal} {Nature}\ }\textbf {\bibinfo
  {volume} {543}},\ \bibinfo {pages} {221} (\bibinfo {year}
  {2017}{\natexlab{b}})},\ \bibinfo {note} {number: 7644 Publisher: Nature
  Publishing Group}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Zhou}\ \emph {et~al.}(2020)\citenamefont {Zhou},
  \citenamefont {Choi}, \citenamefont {Choi}, \citenamefont {Landig},
  \citenamefont {Douglas}, \citenamefont {Isoya}, \citenamefont {Jelezko},
  \citenamefont {Onoda}, \citenamefont {Sumiya}, \citenamefont {Cappellaro}
  \emph {et~al.}}]{zhou2020quantum}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {Zhou}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo
  {author} {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo {author}
  {\bibfnamefont {R.}~\bibnamefont {Landig}}, \bibinfo {author} {\bibfnamefont
  {A.~M.}\ \bibnamefont {Douglas}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
  {F.}~\bibnamefont {Jelezko}}, \bibinfo {author} {\bibfnamefont
  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
  {P.}~\bibnamefont {Cappellaro}}, \emph {et~al.},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Physical review X}\ }\textbf {\bibinfo
  {volume} {10}},\ \bibinfo {pages} {031003} (\bibinfo {year}
  {2020})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2017)\citenamefont
  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
  and\ \citenamefont {Hemmer}}]{akhmedzhanov_microwave-free_2017}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
  {\bibfnamefont {I.}~\bibnamefont {Zelensky}},\ and\ \bibinfo {author}
  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href
  {https://doi.org/10.1103/PhysRevA.96.013806} {\bibfield  {journal} {\bibinfo
  {journal} {Phys. Rev. A}\ }\textbf {\bibinfo {volume} {96}},\ \bibinfo
  {pages} {013806} (\bibinfo {year} {2017})},\ \bibinfo {note} {number:
  1}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2019)\citenamefont
  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
  and\ \citenamefont {Hemmer}}]{akhmedzhanov_magnetometry_2019}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
  {\bibfnamefont {I.}~\bibnamefont {Zelensky}},\ and\ \bibinfo {author}
  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href
  {https://doi.org/10.1103/PhysRevA.100.043844} {\bibfield  {journal} {\bibinfo
   {journal} {Phys. Rev. A}\ }\textbf {\bibinfo {volume} {100}},\ \bibinfo
  {pages} {043844} (\bibinfo {year} {2019})},\ \bibinfo {note} {number:
  4}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Wolf}\ \emph {et~al.}(2015)\citenamefont {Wolf},
  \citenamefont {Neumann}, \citenamefont {Nakamura}, \citenamefont {Sumiya},
  \citenamefont {Ohshima}, \citenamefont {Isoya},\ and\ \citenamefont
  {Wrachtrup}}]{Wolf}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
  {Wolf}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Neumann}},
  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Nakamura}}, \bibinfo
  {author} {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author}
  {\bibfnamefont {T.}~\bibnamefont {Ohshima}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Isoya}},\ and\ \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Wrachtrup}},\ }\href
  {https://doi.org/10.1103/PhysRevX.5.041001} {\bibfield  {journal} {\bibinfo
  {journal} {Phys. Rev. X}\ }\textbf {\bibinfo {volume} {5}},\ \bibinfo {pages}
  {041001} (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Stürner}\ \emph {et~al.}(2021)\citenamefont
  {Stürner}, \citenamefont {Brenneis}, \citenamefont {Buck}, \citenamefont
  {Kassel}, \citenamefont {Rölver}, \citenamefont {Fuchs}, \citenamefont
  {Savitsky}, \citenamefont {Suter}, \citenamefont {Grimmel}, \citenamefont
  {Hengesbach}, \citenamefont {Förtsch}, \citenamefont {Nakamura},
  \citenamefont {Sumiya}, \citenamefont {Onoda}, \citenamefont {Isoya},\ and\
  \citenamefont {Jelezko}}]{Sturner}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.~M.}\ \bibnamefont
  {Stürner}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Brenneis}},
  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Buck}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Kassel}}, \bibinfo {author} {\bibfnamefont
  {R.}~\bibnamefont {Rölver}}, \bibinfo {author} {\bibfnamefont
  {T.}~\bibnamefont {Fuchs}}, \bibinfo {author} {\bibfnamefont
  {A.}~\bibnamefont {Savitsky}}, \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Suter}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Grimmel}}, \bibinfo {author} {\bibfnamefont
  {S.}~\bibnamefont {Hengesbach}}, \bibinfo {author} {\bibfnamefont
  {M.}~\bibnamefont {Förtsch}}, \bibinfo {author} {\bibfnamefont
  {K.}~\bibnamefont {Nakamura}}, \bibinfo {author} {\bibfnamefont
  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Isoya}},\ and\ \bibinfo {author} {\bibfnamefont
  {F.}~\bibnamefont {Jelezko}},\ }\href
  {https://doi.org/https://doi.org/10.1002/qute.202000111} {\bibfield
  {journal} {\bibinfo  {journal} {Advanced Quantum Technologies}\ }\textbf
  {\bibinfo {volume} {4}},\ \bibinfo {pages} {2000111} (\bibinfo {year}
  {2021})},\ \Eprint
  {https://arxiv.org/abs/https://onlinelibrary.wiley.com/doi/pdf/10.1002/qute.202000111}
  {https://onlinelibrary.wiley.com/doi/pdf/10.1002/qute.202000111} \BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Jarmola}\ \emph {et~al.}(2015)\citenamefont
  {Jarmola}, \citenamefont {Berzins}, \citenamefont {Smits}, \citenamefont
  {Smits}, \citenamefont {Prikulis}, \citenamefont {Gahbauer}, \citenamefont
  {Ferber}, \citenamefont {Erts}, \citenamefont {Auzinsh},\ and\ \citenamefont
  {Budker}}]{jarmola_longitudinal_2015}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Jarmola}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Berzins}},
  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Smits}}, \bibinfo
  {author} {\bibfnamefont {K.}~\bibnamefont {Smits}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Prikulis}}, \bibinfo {author}
  {\bibfnamefont {F.}~\bibnamefont {Gahbauer}}, \bibinfo {author}
  {\bibfnamefont {R.}~\bibnamefont {Ferber}}, \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Erts}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Auzinsh}},\ and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
  {Budker}},\ }\href {https://doi.org/10.1063/1.4937489} {\bibfield  {journal}
  {\bibinfo  {journal} {Appl. Phys. Lett.}\ }\textbf {\bibinfo {volume}
  {107}},\ \bibinfo {pages} {242403} (\bibinfo {year} {2015})},\ \bibinfo
  {note} {number: 24}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Mrózek}\ \emph {et~al.}(2015)\citenamefont
  {Mrózek}, \citenamefont {Rudnicki}, \citenamefont {Kehayias}, \citenamefont
  {Jarmola}, \citenamefont {Budker},\ and\ \citenamefont
  {Gawlik}}]{mrozek_longitudinal_2015}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Mrózek}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Rudnicki}},
  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Kehayias}}, \bibinfo
  {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}}, \bibinfo {author}
  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ and\ \bibinfo {author}
  {\bibfnamefont {W.}~\bibnamefont {Gawlik}},\ }\href
  {https://doi.org/10.1140/epjqt/s40507-015-0035-z} {\bibfield  {journal}
  {\bibinfo  {journal} {EPJ Quantum Technol.}\ }\textbf {\bibinfo {volume}
  {2}},\ \bibinfo {pages} {22} (\bibinfo {year} {2015})},\ \bibinfo {note}
  {number: 1}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Epstein}\ \emph {et~al.}(2005)\citenamefont
  {Epstein}, \citenamefont {Mendoza}, \citenamefont {Kato},\ and\ \citenamefont
  {Awschalom}}]{epstein2005anisotropic}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
  {Epstein}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Mendoza}},
  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Kato}},\ and\ \bibinfo
  {author} {\bibfnamefont {D.}~\bibnamefont {Awschalom}},\ }\href@noop {}
  {\bibfield  {journal} {\bibinfo  {journal} {Nature physics}\ }\textbf
  {\bibinfo {volume} {1}},\ \bibinfo {pages} {94} (\bibinfo {year}
  {2005})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Lai}\ \emph {et~al.}(2009)\citenamefont {Lai},
  \citenamefont {Zheng}, \citenamefont {Jelezko}, \citenamefont {Treussart},\
  and\ \citenamefont {Roch}}]{lai2009influence}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~D.}\ \bibnamefont
  {Lai}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Zheng}}, \bibinfo
  {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}}, \bibinfo {author}
  {\bibfnamefont {F.}~\bibnamefont {Treussart}},\ and\ \bibinfo {author}
  {\bibfnamefont {J.-F.}\ \bibnamefont {Roch}},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Applied Physics Letters}\ }\textbf {\bibinfo
  {volume} {95}},\ \bibinfo {pages} {133101} (\bibinfo {year}
  {2009})}\BibitemShut {NoStop}%
\bibitem [{SI_()}]{SI_low_filed_CR}%
  \BibitemOpen
  \href@noop {} {\bibinfo {title} {See supplemental material at [...], which
  includes refs. [20-22]}}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Anishchik}\ \emph {et~al.}(2015)\citenamefont
  {Anishchik}, \citenamefont {Vins}, \citenamefont {Yelisseyev}, \citenamefont
  {Lukzen}, \citenamefont {Lavrik},\ and\ \citenamefont
  {Bagryansky}}]{anishchik2015low}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Anishchik}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Vins}},
  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Yelisseyev}}, \bibinfo
  {author} {\bibfnamefont {N.}~\bibnamefont {Lukzen}}, \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Lavrik}},\ and\ \bibinfo {author}
  {\bibfnamefont {V.}~\bibnamefont {Bagryansky}},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {New Journal of Physics}\ }\textbf {\bibinfo
  {volume} {17}},\ \bibinfo {pages} {023040} (\bibinfo {year}
  {2015})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2020)\citenamefont
  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},
  \citenamefont {Kilin},\ and\ \citenamefont {Jelezko}}]{filimonenko2020weak}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Nizovtsev}}, \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
  {Kilin}},\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
  {Jelezko}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Semiconductors}\ }\textbf {\bibinfo {volume} {54}},\ \bibinfo {pages} {1730}
  (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Van~Oort}\ and\ \citenamefont
  {Glasbeek}(1990)}]{van1990electric}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
  {Van~Oort}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Glasbeek}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Chemical Physics Letters}\ }\textbf {\bibinfo {volume} {168}},\ \bibinfo
  {pages} {529} (\bibinfo {year} {1990})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Jarmola}\ \emph {et~al.}(2012)\citenamefont
  {Jarmola}, \citenamefont {Acosta}, \citenamefont {Jensen}, \citenamefont
  {Chemerisov},\ and\ \citenamefont {Budker}}]{jarmola_temperature-_2012}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Jarmola}}, \bibinfo {author} {\bibfnamefont {V.~M.}\ \bibnamefont {Acosta}},
  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Jensen}}, \bibinfo
  {author} {\bibfnamefont {S.}~\bibnamefont {Chemerisov}},\ and\ \bibinfo
  {author} {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }\href
  {https://doi.org/10.1103/PhysRevLett.108.197601} {\bibfield  {journal}
  {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {108}},\
  \bibinfo {pages} {197601} (\bibinfo {year} {2012})},\ \bibinfo {note}
  {number: 19}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Redman}\ \emph {et~al.}(1991)\citenamefont {Redman},
  \citenamefont {Brown}, \citenamefont {Sands},\ and\ \citenamefont
  {Rand}}]{redman1991spin}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
  {Redman}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Brown}},
  \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Sands}},\ and\ \bibinfo
  {author} {\bibfnamefont {S.}~\bibnamefont {Rand}},\ }\href@noop {} {\bibfield
   {journal} {\bibinfo  {journal} {Physical review letters}\ }\textbf {\bibinfo
  {volume} {67}},\ \bibinfo {pages} {3420} (\bibinfo {year}
  {1991})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Norambuena}\ \emph {et~al.}(2018)\citenamefont
  {Norambuena}, \citenamefont {Mu{\~n}oz}, \citenamefont {Dinani},
  \citenamefont {Jarmola}, \citenamefont {Maletinsky}, \citenamefont {Budker},\
  and\ \citenamefont {Maze}}]{norambuena2018spin}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Norambuena}}, \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
  {Mu{\~n}oz}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Dinani}},
  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}}, \bibinfo
  {author} {\bibfnamefont {P.}~\bibnamefont {Maletinsky}}, \bibinfo {author}
  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ and\ \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Maze}},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Physical Review B}\ }\textbf {\bibinfo
  {volume} {97}},\ \bibinfo {pages} {094304} (\bibinfo {year}
  {2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Choi}\ \emph
  {et~al.}(2017{\natexlab{c}})\citenamefont {Choi}, \citenamefont {Choi},
  \citenamefont {Kucsko}, \citenamefont {Maurer}, \citenamefont {Shields},
  \citenamefont {Sumiya}, \citenamefont {Onoda}, \citenamefont {Isoya},
  \citenamefont {Demler}, \citenamefont {Jelezko} \emph
  {et~al.}}]{choi2017depolarization}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {Choi}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo
  {author} {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author}
  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
  {\bibfnamefont {B.~J.}\ \bibnamefont {Shields}}, \bibinfo {author}
  {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
  {E.}~\bibnamefont {Demler}}, \bibinfo {author} {\bibfnamefont
  {F.}~\bibnamefont {Jelezko}}, \emph {et~al.},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Physical review letters}\ }\textbf {\bibinfo
  {volume} {118}},\ \bibinfo {pages} {093601} (\bibinfo {year}
  {2017}{\natexlab{c}})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Pellet-Mary}\ \emph
  {et~al.}(2021{\natexlab{a}})\citenamefont {Pellet-Mary}, \citenamefont
  {Huillery}, \citenamefont {Perdriat},\ and\ \citenamefont
  {H{\'e}tet}}]{pellet2021magnetic}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
  {Pellet-Mary}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
  {Huillery}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Perdriat}},\
  and\ \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {H{\'e}tet}},\
  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Physical Review
  B}\ }\textbf {\bibinfo {volume} {104}},\ \bibinfo {pages} {L100411} (\bibinfo
  {year} {2021}{\natexlab{a}})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Mr{\'o}zek}\ \emph {et~al.}(2021)\citenamefont
  {Mr{\'o}zek}, \citenamefont {Wojciechowski},\ and\ \citenamefont
  {Gawlik}}]{mrozek2021characterization}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Mr{\'o}zek}}, \bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
  {Wojciechowski}},\ and\ \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
  {Gawlik}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Diamond and Related Materials}\ }\textbf {\bibinfo {volume} {120}},\
  \bibinfo {pages} {108689} (\bibinfo {year} {2021})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Giri}\ \emph {et~al.}(2018)\citenamefont {Giri},
  \citenamefont {Gorrini}, \citenamefont {Dorigoni}, \citenamefont {Avalos},
  \citenamefont {Cazzanelli}, \citenamefont {Tambalo},\ and\ \citenamefont
  {Bifone}}]{giri_coupled_2018}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
  {Giri}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Gorrini}},
  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Dorigoni}}, \bibinfo
  {author} {\bibfnamefont {C.~E.}\ \bibnamefont {Avalos}}, \bibinfo {author}
  {\bibfnamefont {M.}~\bibnamefont {Cazzanelli}}, \bibinfo {author}
  {\bibfnamefont {S.}~\bibnamefont {Tambalo}},\ and\ \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Bifone}},\ }\href
  {https://doi.org/10.1103/PhysRevB.98.045401} {\bibfield  {journal} {\bibinfo
  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {98}},\ \bibinfo
  {pages} {045401} (\bibinfo {year} {2018})},\ \bibinfo {note} {number:
  4}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Giri}\ \emph {et~al.}(2019)\citenamefont {Giri},
  \citenamefont {Dorigoni}, \citenamefont {Tambalo}, \citenamefont {Gorrini},\
  and\ \citenamefont {Bifone}}]{giri_selective_2019}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
  {Giri}}, \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Dorigoni}},
  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Tambalo}}, \bibinfo
  {author} {\bibfnamefont {F.}~\bibnamefont {Gorrini}},\ and\ \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Bifone}},\ }\href
  {https://doi.org/10.1103/PhysRevB.99.155426} {\bibfield  {journal} {\bibinfo
  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {99}},\ \bibinfo
  {pages} {155426} (\bibinfo {year} {2019})},\ \bibinfo {note} {number:
  15}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Pellet-Mary}\ \emph
  {et~al.}(2021{\natexlab{b}})\citenamefont {Pellet-Mary}, \citenamefont
  {Huillery}, \citenamefont {Perdriat}, \citenamefont {Tallaire},\ and\
  \citenamefont {H{\'e}tet}}]{pellet2021optical}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
  {Pellet-Mary}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
  {Huillery}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Perdriat}},
  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Tallaire}},\ and\
  \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {H{\'e}tet}},\ }\href@noop
  {} {\bibfield  {journal} {\bibinfo  {journal} {Physical Review B}\ }\textbf
  {\bibinfo {volume} {103}},\ \bibinfo {pages} {L100411} (\bibinfo {year}
  {2021}{\natexlab{b}})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Mittiga}\ \emph {et~al.}(2018)\citenamefont
  {Mittiga}, \citenamefont {Hsieh}, \citenamefont {Zu}, \citenamefont {Kobrin},
  \citenamefont {Machado}, \citenamefont {Bhattacharyya}, \citenamefont {Rui},
  \citenamefont {Jarmola}, \citenamefont {Choi}, \citenamefont {Budker} \emph
  {et~al.}}]{mittiga2018imaging}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
  {Mittiga}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Hsieh}},
  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Zu}}, \bibinfo {author}
  {\bibfnamefont {B.}~\bibnamefont {Kobrin}}, \bibinfo {author} {\bibfnamefont
  {F.}~\bibnamefont {Machado}}, \bibinfo {author} {\bibfnamefont
  {P.}~\bibnamefont {Bhattacharyya}}, \bibinfo {author} {\bibfnamefont
  {N.}~\bibnamefont {Rui}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Jarmola}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}},
  \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Budker}}, \emph
  {et~al.},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Physical review letters}\ }\textbf {\bibinfo {volume} {121}},\ \bibinfo
  {pages} {246402} (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Dolde}\ \emph {et~al.}(2011)\citenamefont {Dolde},
  \citenamefont {Fedder}, \citenamefont {Doherty}, \citenamefont {N{\"o}bauer},
  \citenamefont {Rempp}, \citenamefont {Balasubramanian}, \citenamefont {Wolf},
  \citenamefont {Reinhard}, \citenamefont {Hollenberg}, \citenamefont {Jelezko}
  \emph {et~al.}}]{dolde2011electric}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
  {Dolde}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Fedder}},
  \bibinfo {author} {\bibfnamefont {M.~W.}\ \bibnamefont {Doherty}}, \bibinfo
  {author} {\bibfnamefont {T.}~\bibnamefont {N{\"o}bauer}}, \bibinfo {author}
  {\bibfnamefont {F.}~\bibnamefont {Rempp}}, \bibinfo {author} {\bibfnamefont
  {G.}~\bibnamefont {Balasubramanian}}, \bibinfo {author} {\bibfnamefont
  {T.}~\bibnamefont {Wolf}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
  {Reinhard}}, \bibinfo {author} {\bibfnamefont {L.~C.}\ \bibnamefont
  {Hollenberg}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},
  \emph {et~al.},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Nature Physics}\ }\textbf {\bibinfo {volume} {7}},\ \bibinfo {pages} {459}
  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Qiu}\ \emph {et~al.}(2022)\citenamefont {Qiu},
  \citenamefont {Hamo}, \citenamefont {Vool}, \citenamefont {Zhou},\ and\
  \citenamefont {Yacoby}}]{qiu2022nanoscale}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont
  {Qiu}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Hamo}}, \bibinfo
  {author} {\bibfnamefont {U.}~\bibnamefont {Vool}}, \bibinfo {author}
  {\bibfnamefont {T.~X.}\ \bibnamefont {Zhou}},\ and\ \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Yacoby}},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {arXiv preprint arXiv:2205.03952}\ } (\bibinfo
  {year} {2022})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Wickenbrock}\ \emph {et~al.}(2016)\citenamefont
  {Wickenbrock}, \citenamefont {Zheng}, \citenamefont {Bougas}, \citenamefont
  {Leefer}, \citenamefont {Afach}, \citenamefont {Jarmola}, \citenamefont
  {Acosta},\ and\ \citenamefont {Budker}}]{Wickenbrock}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Wickenbrock}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Zheng}},
  \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Bougas}}, \bibinfo
  {author} {\bibfnamefont {N.}~\bibnamefont {Leefer}}, \bibinfo {author}
  {\bibfnamefont {S.}~\bibnamefont {Afach}}, \bibinfo {author} {\bibfnamefont
  {A.}~\bibnamefont {Jarmola}}, \bibinfo {author} {\bibfnamefont {V.~M.}\
  \bibnamefont {Acosta}},\ and\ \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Budker}},\ }\href {https://doi.org/10.1063/1.4960171}
  {\bibfield  {journal} {\bibinfo  {journal} {Applied Physics Letters}\
  }\textbf {\bibinfo {volume} {109}},\ \bibinfo {pages} {053505} (\bibinfo
  {year} {2016})},\ \Eprint
  {https://arxiv.org/abs/https://doi.org/10.1063/1.4960171}
  {https://doi.org/10.1063/1.4960171} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {Zheng}\ \emph {et~al.}(2017)\citenamefont {Zheng},
  \citenamefont {Chatzidrosos}, \citenamefont {Wickenbrock}, \citenamefont
  {Bougas}, \citenamefont {Lazda}, \citenamefont {Berzins}, \citenamefont
  {Gahbauer}, \citenamefont {Auzinsh}, \citenamefont {Ferber},\ and\
  \citenamefont {Budker}}]{zheng2017level}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {Zheng}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Chatzidrosos}},
  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Wickenbrock}}, \bibinfo
  {author} {\bibfnamefont {L.}~\bibnamefont {Bougas}}, \bibinfo {author}
  {\bibfnamefont {R.}~\bibnamefont {Lazda}}, \bibinfo {author} {\bibfnamefont
  {A.}~\bibnamefont {Berzins}}, \bibinfo {author} {\bibfnamefont {F.~H.}\
  \bibnamefont {Gahbauer}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Auzinsh}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Ferber}},\
  and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }in\
  \href@noop {} {\emph {\bibinfo {booktitle} {Slow Light, Fast Light, and
  Opto-Atomic Precision Metrology X}}},\ Vol.\ \bibinfo {volume} {10119}\
  (\bibinfo {organization} {SPIE},\ \bibinfo {year} {2017})\ pp.\ \bibinfo
  {pages} {115--122}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Zheng}\ \emph {et~al.}(2020)\citenamefont {Zheng},
  \citenamefont {Sun}, \citenamefont {Chatzidrosos}, \citenamefont {Zhang},
  \citenamefont {Nakamura}, \citenamefont {Sumiya}, \citenamefont {Ohshima},
  \citenamefont {Isoya}, \citenamefont {Wrachtrup}, \citenamefont
  {Wickenbrock},\ and\ \citenamefont {Budker}}]{zheng_microwave-free_2020}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {Zheng}}, \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Sun}}, \bibinfo
  {author} {\bibfnamefont {G.}~\bibnamefont {Chatzidrosos}}, \bibinfo {author}
  {\bibfnamefont {C.}~\bibnamefont {Zhang}}, \bibinfo {author} {\bibfnamefont
  {K.}~\bibnamefont {Nakamura}}, \bibinfo {author} {\bibfnamefont
  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
  {T.}~\bibnamefont {Ohshima}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Wrachtrup}}, \bibinfo {author} {\bibfnamefont
  {A.}~\bibnamefont {Wickenbrock}},\ and\ \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Budker}},\ }\href
  {https://doi.org/10.1103/PhysRevApplied.13.044023} {\bibfield  {journal}
  {\bibinfo  {journal} {Phys. Rev. Applied}\ }\textbf {\bibinfo {volume}
  {13}},\ \bibinfo {pages} {044023} (\bibinfo {year} {2020})},\ \bibinfo {note}
  {number: 4}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2018)\citenamefont
  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},\ and\
  \citenamefont {Kilin}}]{filimonenko2018weak}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Nizovtsev}},\ and\ \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
  {Kilin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Semiconductors}\ }\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {1865}
  (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2022)\citenamefont
  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},
  \citenamefont {Kilin},\ and\ \citenamefont
  {Jelezko}}]{filimonenko2022manifestation}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.~P.}\ \bibnamefont
  {Nizovtsev}}, \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
  {Kilin}},\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
  {Jelezko}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Journal of Applied Spectroscopy}\ }\textbf {\bibinfo {volume} {88}},\
  \bibinfo {pages} {1131} (\bibinfo {year} {2022})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Vetter}\ \emph {et~al.}(2022)\citenamefont {Vetter},
  \citenamefont {Marshall}, \citenamefont {Genov}, \citenamefont {Weiss},
  \citenamefont {Striegler}, \citenamefont {Gro\ss{}mann}, \citenamefont
  {Oviedo-Casado}, \citenamefont {Cerrillo}, \citenamefont {Prior},
  \citenamefont {Neumann},\ and\ \citenamefont {Jelezko}}]{Vetter_LFM}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~J.}\ \bibnamefont
  {Vetter}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Marshall}},
  \bibinfo {author} {\bibfnamefont {G.~T.}\ \bibnamefont {Genov}}, \bibinfo
  {author} {\bibfnamefont {T.~F.}\ \bibnamefont {Weiss}}, \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Striegler}}, \bibinfo {author}
  {\bibfnamefont {E.~F.}\ \bibnamefont {Gro\ss{}mann}}, \bibinfo {author}
  {\bibfnamefont {S.}~\bibnamefont {Oviedo-Casado}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Cerrillo}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Prior}}, \bibinfo {author} {\bibfnamefont
  {P.}~\bibnamefont {Neumann}},\ and\ \bibinfo {author} {\bibfnamefont
  {F.}~\bibnamefont {Jelezko}},\ }\href
  {https://doi.org/10.1103/PhysRevApplied.17.044028} {\bibfield  {journal}
  {\bibinfo  {journal} {Phys. Rev. Applied}\ }\textbf {\bibinfo {volume}
  {17}},\ \bibinfo {pages} {044028} (\bibinfo {year} {2022})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Wang}\ \emph {et~al.}(2022)\citenamefont {Wang},
  \citenamefont {Liu}, \citenamefont {Fan}, \citenamefont {Feng}, \citenamefont
  {Leong}, \citenamefont {Finkler}, \citenamefont {Denisenko}, \citenamefont
  {Wrachtrup}, \citenamefont {Li},\ and\ \citenamefont {Liu}}]{WangRB}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
  {Wang}}, \bibinfo {author} {\bibfnamefont {C.-F.}\ \bibnamefont {Liu}},
  \bibinfo {author} {\bibfnamefont {J.-W.}\ \bibnamefont {Fan}}, \bibinfo
  {author} {\bibfnamefont {X.}~\bibnamefont {Feng}}, \bibinfo {author}
  {\bibfnamefont {W.-H.}\ \bibnamefont {Leong}}, \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Finkler}}, \bibinfo {author} {\bibfnamefont
  {A.}~\bibnamefont {Denisenko}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Wrachtrup}}, \bibinfo {author} {\bibfnamefont
  {Q.}~\bibnamefont {Li}},\ and\ \bibinfo {author} {\bibfnamefont {R.-B.}\
  \bibnamefont {Liu}},\ }\href
  {https://doi.org/10.1103/PhysRevResearch.4.013098} {\bibfield  {journal}
  {\bibinfo  {journal} {Phys. Rev. Research}\ }\textbf {\bibinfo {volume}
  {4}},\ \bibinfo {pages} {013098} (\bibinfo {year} {2022})}\BibitemShut
  {NoStop}%
\end{thebibliography}%

\end{document}

### 2022-07-08 Mes papiers/Low field CR/V3/SI/Low_field_CR_SI.tex
la tete au weekend

--- 2022-06-29 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
@@
 \
-\title{Low magnetic field depolarization of NV$^-$ centers through dipole-dipole interaction}
+\title{Depolarization of interacting NV$^-$ centers at low magnetic fields}
 
-\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, P. Huillery$^?$,  G. H\'etet}
+\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, P. Huillery$^2$,  G. H\'etet$^1$}
 
-\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e Paris Cit\'e , 24 rue Lhomond, 75231 Paris Cedex 05, France.}
+\affiliation{$^1$ Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e Paris Cit\'e , 24 rue Lhomond, 75231 Paris Cedex 05, France \\ $^2$ Univ Rennes, INSA Rennes, CNRS, Institut FOTON - UMR 6082, F-35000 Rennes, France}
 
@@
 
-\section{NV$^-$ ground state Hamiltonian in low magnetic field}
+\section{NV$^-$ ground state Hamiltonian under low magnetic field}
 \label{sec Hamiltonian}
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=0.9\textwidth]{Figures_SI/fig_ESR}
-\caption{ODMR measurement (a) in zero magnetic field, (b) in non-zero magnetic field when zooming on a single class}
+\caption{ODMR measurement (a) under zero magnetic field, (b) under a magnetic field of $~ 50 G$ when zooming on a single NV class}
 \label{ESR_single_spin}
 \end{figure}
-With zero external magnetic field, there are three possible sources of splitting of the $\{\ket{+1},\ket{-1}\}$ manifold : local electric field, crystal strain and local magnetic field.
-Of these three causes, only the electric field can explain the shape of the ODMR line that we observe in zero external magnetic field (Fig.
+Under zero external magnetic field, there are three possible processes that can cause a splitting of the $\{\ket{+1},\ket{-1}\}$ states : local electric fields, crystal strain and local magnetic fields.
+Of these three causes, only the electric field can explain the shape of the ODMR line that is observed under zero external magnetic field (Fig.
 \ref{ESR_single_spin}a) \cite{mittiga2018imaging}.
 
-Indeed, random local magnetic field would produce a single broadened line while crystal strain would produce a shifting of the zero field splitting (ZFS) of the same order as the splitting of the levels, which would also blur the two transitions into a single line.
-On the other hand, due to the large difference between the longitudinal and transverse electric field susceptibilities ($d_\parallel = 0.35\ \rm Hz\, cm/V$ and $d_\perp = 17\ \rm Hz\, cm/V$ \cite{van1990electric}), random local electric field will, on average, cause a splitting much stronger than its ZFS shifting and result in a two peak spectrum.
+Indeed, random local magnetic field would produce a single broadened line while crystal strain would shift the zero field splitting (ZFS) by the same order of magnitude as the splitting between the levels, which would merge the two transitions onto a single line.
+Due to the large difference between the longitudinal and transverse electric field susceptibilities ($d_\parallel = 0.35\ \rm Hz\, cm/V$ and $d_\perp = 17\ \rm Hz\, cm/V$ \cite{van1990electric}), random local electric field can, on average, cause a splitting that is much stronger than the ZFS shift and result in a two peak spectrum centered on 2.87 GHz.
 
-In our low magnetic field dipole-dipole coupling model, we will therefore neglect the contribution of the strain, local magnetic field and longitudinal electric field.
-We also don't take into account the hyper-fine structure of the NV center due to the large inhomogeneous broadening of the transitions (Fig.
+\begin{figure}[h]
+\includegraphics[width=0.7\textwidth]{Figures_SI/map_etats_propres}
+\caption{Numerical simulations showing the closeness of the Hamiltonian second excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$, as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.
+A value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen}
+\label{map etats propres}
+\end{figure}
+
+In our model for  dipole-dipole coupling under small magnetic fields, we will therefore neglect the contribution of the strain, local magnetic fields and longitudinal electric fields.
+We also do not take into account the hyper-fine structure of the NV center because of the large inhomogeneous broadening of the transitions in all our samples (Fig.
 \ref{ESR_single_spin}b).
@@
 \label{NV Hamiltonian}
-\mathcal{H}_s=D S_z^2 + \gamma_e \bm{B}_{\rm ext} \cdot \bm{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
+\mathcal{H}_s/h=D S_z^2 + \gamma_e \bm{B}_{\rm ext} \cdot \bm{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
 \end{equation}
@@
 
-In the absence of an external magnetic field, the symmetry in the ($xy$) plane allows us to chose the $x$ direction along the electric field.
+In the absence of an external magnetic field, the symmetry of the hamitlonian in the ($xy$) plane allows us to pick the $x$ direction along the electric field.
 The eigenstates of $\mathcal{H}_s$ then become $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
 
-In the presence of non-zero magnetic field, let us call the Hamiltonian eigenstates $\{ \ket{g},\ket{d}, \ket{e} \} $.
+In the presence of a magnetic field at an angle $\theta$ with respect to the NV axis, we denote the Hamiltonian eigenstates $\{ \ket{g},\ket{d}, \ket{e} \} $ in ascending order of energy.
 Fig.
-\ref{map etats propres} shows how close the $\ket{e}$ state is to the $\ket{+1}$ and $\ket{+}$ states as a function of the external magnetic field.
-Looking at the closeness of $\ket{d}$ to $\ket{-1}$ and $\ket{-}$ would show similar results, while $\ket{g}$ is pretty much equals to $\ket{0}$ for $B<100\ \rm G$.
+\ref{map etats propres} shows how close the $\ket{e}$ state is to the $\ket{+1}$ and $\ket{+}$ states as a function of the external magnetic field and angle $\theta$.
+A similar analysis for the  $\ket{d}$, $\ket{-1}$ and $\ket{-}$ states show similar results, while $\ket{g}$ is pretty much equal to $\ket{0}$ for $B<100\ \rm G$.
 
-This tells us that in most cases, the $\{ \ket{0},\ket{+},\ket{-} \}$ basis is only the good Hamiltonian basis for magnetic field smaller than a few Gauss, except in the case of pure transverse magnetic field where the $\{ \ket{0},\ket{+},\ket{-} \}$ basis remains a good basis even for sizable magnetic fields.
-\begin{figure}
-\includegraphics[width=0.7\textwidth]{Figures_SI/map_etats_propres}
-\caption{Numerical simulations of the closeness of the Hamiltonian most excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$, as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.
-A value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen}
-\label{map etats propres}
-\end{figure}
+This result tells us that, in most cases, the $\{ \ket{0},\ket{+},\ket{-} \}$ basis is only the good Hamiltonian basis for magnetic field smaller than a few Gauss, except in the case of pure transverse magnetic field where the $\{ \ket{0},\ket{+},\ket{-} \}$ basis remains a good basis even for sizable magnetic fields.
 
@@
 \item \textbf{HPHT-150-1} : A 150 $\mu$m HPHT 1b diamond irradiated and annealed to reach a concentration [NV$^-$] $\approx 3\ \rm ppm$.
-Bought as is from Adamas Nanotechnology (MDNV150umHi).
-This sample is used in Fig.
+This sample was bought from Adamas Nanotechnology (MDNV150umHi).
+It is used in Fig.
 2 (b), 2(d) and 3 of the main text, as well as Fig.
@@
 \item \textbf{HPHT-15-2}  : Another sample from the same batch as HPHT-15-1. This sample is used in Fig. \ref{121 VS 22 fig} and \ref{Various ODMR} of the SI.
+\item \textbf{HPHT-1-1}  : A 1 $\mu$m diamond with similar properties as the last ones, also bought from Adamas Nanotechnology (MDNV1umHi).
+This sample is used to perform a sensitivity measurement similar to the one done in Fig.
+5 of the main text.
 \item \textbf{CVD-PPM} : A CVD bulk diamond described in \citep{TALLAIRE2020421}.
@@
 \section{Experimental Setup}
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=0.7\textwidth]{Figures_SI/shema_exp}
@@
 The laser used here is a Picoquant PDL 800-D with a 520 nm LDH laser head, providing pulses of 40 ps at a rate of 20 MHz, with an average power of $0.5 \sim 5\ \rm mW$.
+These fast pulses serve no purpose in the actual experiment.
+This laser was used solely because of its high amplitude stability.
 The trigger of the pulses is generated externally in order to achieve fast gating of the laser for the $T_1$ measurement.
-We previously did similar experiment using a continuous 532 nm laser and observed no difference in the behavior of the spins.
+We previously did similar experiment using a continuous 532 nm laser and observed no difference in the spin response.
 
 The magnetic field is provided by a homemade electromagnet composed of a c-shape iron core and  copper wires.
-The magnet is mounted on two mechanical rotation stages, allowing a control on the polar and azimuthal angle of the magnetic field within a fraction of a degree and is alimented through a low frequency amplifier (Leybold power function generator 522 63)
+The magnet is mounted on two mechanical rotation stages, allowing a control on the polar and azimuthal angle of the magnetic field within a fraction of a degree and is alimented through a low frequency amplifier (Leybold power function generator 522 63).
 
 The microwave field is generated by a Rhode \& Shwarz SMB 100A and is emitted with a handmade loop antenna.
-The field is gated by a Mini-circuits ZASWA-2-50DRA+ switch controlled externally and amplified by a Mini-circuits ZHL-5W-422+ amplifier.
+The microwave field is gated by a Mini-circuits ZASWA-2-50DRA+ switch controlled externally and amplified by a Mini-circuits ZHL-5W-422+ amplifier.
 
@@
 \subsection{$T_1$ fitting Protocol}
-The $T_1$ profile that we observe with our samples is neither purely exponential, nor purely stretched exponential, which is to be expected when the dipole-dipole spin decay is comparable to the phonon decay.
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_combined}
+\caption{$T_1$ measurement with purely exponential and purely stretched exponential fits (a) in zero magnetic field (b) in non-zero magnetic field}
+\label{T1_fits}
+\end{figure}
+\begin{figure}[h]
+\includegraphics[width=0.45\textwidth]{Figures_SI/fig_alphas}
+\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$ as a function of an arbitrarely oriented magnetic field amplitude}
+\label{alphas}
+\end{figure}
+
+The $T_1$ profile that we observe with our samples is neither purely exponential, nor purely stretched exponential, which is to be expected when the dipole-dipole induced spin decay is comparable to the phonon induced decay.
 
@@
 \ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field, which confirms that the $T_1$ profile gets closer to a pure stretched-exponential in zero magnetic field.
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_combined}
-\caption{$T_1$ measurement with purely exponential and purely stretched exponential fits (a) in zero magnetic field (b) in non-zero magnetic field}
-\label{T1_fits}
-\end{figure}
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures_SI/fig_alphas}
-\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$ as a function of a randomly oriented magnetic field amplitude}
-\label{alphas}
-\end{figure}
+
 \subsection{Spectral range of the dipole-dipole cross-relaxations}
 \label{fluctuator width}
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=0.7\textwidth]{Figures_SI/largeur_fluct_SI}
 \caption{Dipole-dipole depolarization for two near-resonant classes.
-(a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice.
+(a) Sketch of the four possible NV orientations (``classes") in a single diamond.
 (b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes.
-The detuning $\Delta \nu$ between the classes b and c was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_c$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz.
-Single class ESR line stretched by a factor of 1,$\sqrt{2}$ and 2 are added for comparison}
+The detuning $\Delta \nu$ between the classes b and c was controlled by changing the orientation of the external magnetic field.
+(c) Stretched part of the lifetime decay curve for the spins resonant with $\nu_c$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz.
+Single class ESR line stretched by a factor of 1,$\sqrt{2}$ and 2 are added for comparison.}
 \label{largeur_fluct}
@@
 
-An experimental signature of the fluctuator hypothesis developed in \cite{choi_depolarization_2017} is to measure the depolarization rate for two near-resonant classes : if there are indeed very fast decaying NV centers (fluctuators with lifetime $T_1^f < 100 ns$), then the spectral width of the fluctuator would be increased beyond $1/T_2^*$ (which we assume to be homogeneous among all spins in the crystal), which means that they would be able to exchange spin quanta (flip-flop) with non-resonant NV detuned by $\Delta \nu$ such that $2\pi/T_1^f > \Delta \nu > 2\pi/T_2^*$
+An experimental signature of the fluctuator hypothesis developed in \cite{choi_depolarization_2017} is the dependence of the depolarization rate when two near-resonant classes are brought to resonance : if there are indeed very fast decaying NV centers (fluctuators with lifetime $T_1^f < 100$ ns), then the spectral width of the fluctuator would be greater than $1/T_2^*$.
+For simplicity, it is typically assumed that the latter is the same for all spins in the crystal.
+This means that NV centers would be able to exchange spin quanta (flip-flop) with non-resonant NV fluctuators detuned by $\Delta \nu$ such that $2\pi/T_1^f > \Delta \nu > 2\pi/T_2^*$.
 
-In order to verify this claim, we have to measure the spectral overlap between two classes, which in the absence of fluctuator should be proportional to the flip-flop rate, and compare it to the actual flip-flop rate which we can measure through the depolarization rate of the spins.
+In order to verify this, we measure the spectral overlap between two classes, which in the absence of fluctuator should be on the order of $1/T_2^*$, and compare it to the actual width of the depolarization rate of the spins
+as a function of detuning.
 
 Fig.
-\ref{largeur_fluct} shows the results of such an experiment where we measured the stretched part of the NV's decay rate, which is very well fitted by a Lorentzian of width $\sigma=8.04 MHz$ (FW or HW ?), and compare it to an ODMR line of a single class of NV centers, stretched by a factor of $\sqrt{2}$ and 2 to simulate the spectral overlap of the two classes (discussed bellow).
-One should notice that, not only is $1/T_1^{\rm dd}$ significantly larger than the ODMR profile, it also doesn't have the same shape.
+\ref{largeur_fluct} shows the results of such an experiment.
+We measured the stretched part of the NV's decay rate at each detuning and obtained Fig.
+\ref{largeur_fluct} -c), which is very well fitted by a Lorentzian of half-width $\sigma=8.04$ MHz.
+We should note that this broadening can not be explained by the direct coherent dipole-dipole interaction induced splitting : for a sample with 3 ppm of NV centers, the average dipole-dipole interaction strength between two nearest NV centers $J_0/r^3 \sim 27\ \rm kHz$ which is several order of magnitude lower than the broadening we observe.
 
-We should note that this broadening can not be explained by the dipole-dipole interaction strength : for a sample with 3 ppm of NV centers, the average dipole-dipole interaction strength between two nearest NV centers $J_0/r^3 \sim 27\ \rm kHz$ which is several order of magnitude lower than the broadening we observe.
-
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=0.7\textwidth]{Figures_SI/overlap}
@@
 
-As illustrated on Fig.
+We can compare this obtained width to the width of an ODMR line of a single class of NV centers, stretched by a factor of $\sqrt{2}$ and 2 to simulate the spectral overlap of two classes.
+One should notice that, not only is $1/T_1^{\rm dd}$ significantly larger than the ODMR profile, it also does not have the same shape.
+
+As illustrated in Fig.
 \ref{overlap}, we define the spectral overlap $S(\Delta \nu)$ between two spins of spectral response $S_1(\nu)$ and $S_2(\nu)$, centered respectively on the frequencies $\nu_1$ and $\nu_2$ where $\Delta \nu = \nu_2-\nu_1$ as :
@@
 
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=0.9\textwidth]{Figures_SI/Fig_splitting}
-\caption{(a) : Simulation of the 8 possible transition frequencies (four $\ket{0} \to \ket{-1}$ and four $\ket{0} \to \ket{+1}$) as a function of the magnetic field amplitude for the orientation of the magnetic field shown in Fig.
+\caption{(a) : Simulation of the 8 possible transition frequencies (four $\ket{0} \to \ket{-1}$ and four $\ket{0} \to \ket{+1}$ transitions) as a function of the magnetic field amplitude for the orientation of the magnetic field shown in Fig.
 3(b) of the main text.
 (b) Difference in frequency between each pair of closest classes as a function of the magnetic field amplitude.
-The dotted line y=8.04 MHz correspond to the estimation of the dipole-dipole interaction range.}
+The dotted line y=8.04 MHz corresponds to an estimation of the dipole-dipole interaction frequency range.}
 \label{splitting}
@@
 Fig.
-\ref{splitting} shows the simulated energies for the 4 $\ket{0} \to \ket{-1}$ and the 4 $\ket{0} \to \ket{+1}$ as a function of the magnetic field, and the difference in energy between the closest pairs of transitions.
+\ref{splitting} shows the simulated energies for the 4 $\ket{0} \to \ket{-1}$ and the 4 $\ket{0} \to \ket{+1}$ transitions as a function of the magnetic field, and the difference in energy between the closest pairs of transitions.
 We can see that the energy detuning between the classes $\Delta \nu$ crosses the value of the dipole-dipole interaction range (8.04 MHz, measured on sample HPHT-150-2 but likely to be similar on sample HPHT-150-1) for magnetic field values $\norm{B} = 12 \sim 15\ \rm G$, which are close to the half-width of the zero-field feature in Fig.
@@
 We did not observe a strong dependence on the laser polarization with the samples used in this study.
-Fig.\ref{Pola} shows the photoluminescence from one of our sample as a function of the magnetic field, either directly or with a modulation of the magnetic field, for 6 randomly chosen polarization angle and saw no major difference.
+Fig.\ref{Pola} shows the photoluminescence from one of our sample as a function of the magnetic field, either directly or with a modulation of the magnetic field, for 6 polarization angles and saw no major difference.
 
 In particular, we did not observe the apparition of an anti-line inside the main dip, unlike what was observed in the two previously cited work.
-We expect that the main reason behind the different behaviors is that we seem to observe far grater spin depolarization in zero-field, which we attribute to dipole-dipole coupling, and that this effect could hide smaller effects such as the one related to the laser polarization.
-\begin{figure}
+We expect that the main reason behind the different behaviors is that we seem to observe far greater spin depolarization in zero-field, which we attribute to dipole-dipole coupling.
+This effect could hide smaller effects such as the one related to the laser polarization.
+\begin{figure}[h]
 \includegraphics[width=0.9\textwidth]{Figures_SI/fig_Pola}
 \caption{Effect of the polarization of the incident laser.
-(a) Photoluminescence of sample HPHT-150-3 as a function of randomly oriented magnetic field amplitude for various polarization angle.
-(b) Demodulated PL in the same conditions}
+(a) Photoluminescence of sample HPHT-150-3 as a function of arbitrarely oriented magnetic field amplitude for various polarization angle.
+(b) Demodulated PL in the same conditions.}
 \label{Pola}
@@
 \subsection{Alignment of the magnetic field along [100] axis}
-\begin{figure}
+\begin{figure}[h]
 \centering
 \includegraphics[width=0.6\textwidth]{Figures_SI/alignement}
-\caption{PL of sample CVD-PPM as a function of the external magnetic field amplitude for different field misalignment with the [100] axis}
+\caption{PL of sample CVD-PPM as a function of the external magnetic field amplitude for different field misalignment with the [100] axis.}
 \label{Alignment}
@@
 We claim that the decrease in PL at low magnetic field in Fig.
-3 (c)(ii) of the main text is due to to the local electric field and the double flip processes.
-One other possibility would be a misalignment of the magnetic field with respect to the [100] axis, where the four classes would only be truly resonant with zero external magnetic field.
+3 (c)(ii) of the main text is due to the local electric field and the double flip processes.
+One other possibility would be a misalignment of the magnetic field with respect to the [100] axis as it is scanned.
+The four classes could for instance be truly resonant only with zero external magnetic field and not when the B field is large.
 
 Fig.
-\ref{Alignment} shows the change in PL with respect to the external magnetic field amplitude for various misalignment of the magnetic field direction compared to the [100] diamond crystalline axis for sample CVD-PPM.
+\ref{Alignment} shows the change in PL with respect to the external magnetic field amplitude for various misalignments of the magnetic field direction compared to the [100] diamond crystalline axis of the sample CVD-PPM.
 We can see that the central drop in PL is not critically affected by the misalignment of the magnetic field, and we are therefore confident that the drop in PL we observe when $B\parallel [100]$ is not an effect of misalignment.
@@
 
-Here are the results of the fluctuator model developed in \cite{choi_depolarization_2017} that we extended to include more scenarios.
+Here are the results of the fluctuator model developed in \cite{choi_depolarization_2017} that we extended to include double flip as well as local electric fields.
 The following sections will detail the calculations that led to these results.
+Let us first summarize our theoretical findings :
+
 \begin{itemize}
-\item The dipole-induced spin decay is at its lowest for a class of NV centers, non resonant with the three other classes, and dominated by a longitudinal magnetic field.
+\item The dipole-induced spin decay is lowest for a spectrally isolated class of NV centers when the transverse magnetic field is negligible.
 In this case the dipole-dipole relaxation is limited by flip-flop processes between spins of the same class.
@@
 
-The NV$^-$ centers in the crystal are divided between two categories : "normal" NV centers (simply called NV) which, in the absence of dipole-dipole coupling would have a phonon-limited $T_1$ ($T_1^{NV}\sim \rm ms$), and fluctuators who are NV centers with an additional, fast, depolarization mechanism, such that their lifetime $T_1^f < 100\ \rm ns$.
+The NV$^-$ centers in the crystal are divided between two categories :``normal" NV centers (simply called NV) which, in the absence of dipole-dipole coupling would have a phonon-limited $T_1$ ($T_1^{NV}\sim \rm ms$), and fluctuators who are NV centers with an additional, fast, depolarization mechanism, such that their lifetime $T_1^f < 100\ \rm ns$.
 Having such a short lifetime, the fluctuators are almost unpolarized by the green laser, making them invisible in standard optical NV measurement protocol ($T_1$, ODMR, etc.).
@@
 \begin{equation}
-\rho(\gamma)=\frac{e^{-1/(4\gamma T)}}{\sqrt{4\pi \gamma^3 T}}
+\rho(\gamma)=\frac{e^{-1/(4\gamma T)}}{\sqrt{4\pi \gamma^3 T}},
 \end{equation}
-Where the timescale $T$ is defined as
+where the timescale $T$ is defined as
 \begin{equation}
-\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f}
+\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f}.
 \label{eq 1/T}
@@
 \begin{equation}
-P(t)=\int_0^\infty \rho(\gamma)\, e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}
+P(t)=\int_0^\infty \rho(\gamma)\, e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}.
 \end{equation}
@@
 \subsection{Dipole-Dipole Hamiltonian between two NV$^-$ centers}
-We consider a system composed of an NV center interacting dipolarly with a fluctuator.
+Let us consider the hamiltonian of an NV center interacting with a fluctuator :
 \begin{equation}
-\mathcal{H}_{\rm tot}=\mathcal{H}_{1}+\mathcal{H}_2+\mathcal{H}_{\rm dd}
+\mathcal{H}_{\rm tot}=\mathcal{H}_{1}+\mathcal{H}_2+\mathcal{H}_{\rm dd}.
 \end{equation}
-Where $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ are the single particle Hamiltonian of the NV and fluctuator, described by the equation (\ref{NV Hamiltonian}), and $\mathcal{H}_{\rm dd}$ is the dipole-dipole interaction Hamiltonian between the two spins.
+$\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ are the single particle Hamiltonian of the NV and fluctuator, described by the equation (\ref{NV Hamiltonian}), and $\mathcal{H}_{\rm dd}$ is the dipole-dipole interaction Hamiltonian between the two spins.
 
 We will consider the spin operators $\bm{S}_1$ and
-$\bm{S}_2$ in the NV own basis : the $z$ orientation of the $S_z$ operator is chosen along the NV axis of each spin.
-In the presence of a longitudinal magnetic field, the sense of $\hat{z}$ will be chosen such that $\bm{B}\cdot \hat{z} >0$, and the $\hat{x}$ direction is chosen along the transverse magnetic or electric field, whichever dominates the single spin Hamiltonian.
+$\bm{S}_2$ in the NV basis : the $z$ orientation of the $S_z$ operator is chosen along the NV axis of each spin.
+In the presence of a longitudinal magnetic field, the sense of $\hat{z}$ will be chosen such that $\bm{B}\cdot \hat{z} >0$, and the $\hat{x}$ direction is chosen along the transverse magnetic or electric field, both of which dominate the single spin Hamiltonian.
 These choices mean that there will be two distinct Cartesian basis $\{\hat{x}_1,\hat{y}_1,\hat{z}_1\}$ and $\{\hat{x}_2,\hat{y}_2,\hat{z}_2\}$ describing the two spins.
@@
 &+\left[3(\hat{u}\cdot\hat{y}_1)(\hat{u}\cdot\hat{x}_2) -\hat{y}_1\cdot\hat{x}_2\right] S_y^1S_x^2 \\
-&+\left[3(\hat{u}\cdot\hat{z}_1)(\hat{u}\cdot\hat{z}_2) -\hat{z}_1\cdot\hat{z}_2\right] S_z^1S_z^2 \\
+&+\left[3(\hat{u}\cdot\hat{z}_1)(\hat{u}\cdot\hat{z}_2) -\hat{z}_1\cdot\hat{z}_2\right] S_z^1S_z^2, \\
 &+\mathcal{H}_{\rm other}
 \end{align}
-Where $J_0=\frac{\mu_0 \gamma_e^2 \hbar^2}{4 \pi} = (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\bm r= r\bm{\hat u}$ is the distance between the two spins and  $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
+where $J_0=\frac{\mu_0 \gamma_e^2 \hbar^2}{4 \pi} = (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\bm r= r\bm{\hat u}$ is the distance between the two spins and  $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
 
@@
 \begin{equation}
-\eta^2=\frac{1}{3} \abs{\mel{\pm 1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,\pm 1}}^2  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}
+\eta^2=\frac{1}{3} \abs{\mel{\pm 1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,\pm 1}}^2  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
 \end{equation}
-Where $\gamma_f$ is the fluctuator lifetime, $\omega_f$ the fluctuator angular frequency and $\omega_{NV}$ the NV angular frequency.
+where $\gamma_f$ is the fluctuator lifetime, $\omega_f/2\pi$ is the transition frequency of the fluctuator and $\omega_{NV}/2\pi$ the NV transition frequency.
 %This numerical factor characterizes both the dipole-dipole strength between the NV and the fluctuator through the matrix element and the resonance condition through the Lorentzian function.
@@
 This numerical factor is linked to the previously mentioned $\bar \eta$ by averaging over every possible orientation and angular position of the fluctuators.
-Assuming that the fluctuators are evenly distributed among all four classes (NV axis orientation) and decomposing the $\bm r$ vector in the spherical basis ($r, \theta, \phi$), we can write $\bar \eta$ as :
+Assuming that the fluctuators are evenly distributed among all four classes of NVs and decomposing the $\bm r$ vector in the spherical basis ($r, \theta, \phi$), we can write $\bar \eta$ as :
 
@@
 \ref{largeur_fluct}, we will assume that these distributions are independent of the external magnetic field, and since we are interested in comparing the same sample for various values of the magnetic field, we will simplify eq.(\ref{eq.
-eta bar}) by postulating that $\frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} =1$ if the NV center's and the fluctuator's classes are resonant (or if they are from the same class), and $\frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2 +4\gamma_f^2}=0$ otherwise.
+eta bar}) by setting $\frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}$ to 1 if the classes of NV centers and fluctuators are resonant (or if they are from the same class), and to 0 otherwise.
 
@@
 (b) $T_1$ measurement on a two-class resonance following the protocol described in main text.
-8 different magnetic field were used on the same sample : 4 times in a \{100\} plane (blue unfilled symbols) and 4 times in a \{110\} plane (orange filled symbols).
+8 different magnetic field were employed on the same sample : 4 times in a \{100\} plane (blue unfilled symbols) and 4 times in a \{110\} plane (orange filled symbols).
 Inset : zoom-in on the 0.3-0.8 ms region.}
@@
 
-With these hypothesis, we found that there were 3 possible scenarios where the NV and fluctuator were resonant:
+With these hypotheses, we find that there are 3 possible scenarios where the NV and fluctuator are resonant :
 \begin{itemize}
-\item The NV and fluctuator are resonant and from the same class (the angle $\widehat{z_1 z_2}=0$) : this is always true, regardless of the magnetic field.
+\item The NV and fluctuator are from the same class, so that the angle $\widehat{z_1 z_2}=0$ : this is always true, regardless of the magnetic field angle and amplitude.
 We can analytically compute $\bar \eta$ in this case and find : $$ \bar {\eta}_{same}=\frac{1}{4}  \int_\theta \int_\phi \abs{\eta(\bm u,\widehat{z_1 z_2}=0,\omega_{NV}=\omega_f)}\, d\Omega= \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot \frac{2}{3 \sqrt{3}} \approx 5.55\cdot10^{-2}$$
-\item The NV and fluctuator are resonant and the angle $\widehat{z_1 z_2}=\rm{arccos}(\frac{1}{3})\approx 70.5^\circ$.
-This happens when the magnetic field lies in the \{100\} crystalline planes family.
-In this case we can numerically compute $\bar \eta_{\rm close} \approx \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot 0.6507 \approx 9.39\cdot10^{-2}$
-\item The NV and fluctuator are resonant and the angle $\widehat{z_1 z_2}=\rm{arccos}(-\frac{1}{3})\approx 109.5^\circ$.
+\item The angle $\widehat{z_1 z_2}=\rm{arccos}(\frac{1}{3})\approx 70.5^\circ$.
+This is the case when the magnetic field lies in the \{100\} crystalline planes family.
+In this case, numerical simulations yield $\bar \eta_{\rm close} \approx \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot 0.6507 \approx 9.39\cdot10^{-2}$
+\item The angle $\widehat{z_1 z_2}=\rm{arccos}(-\frac{1}{3})\approx 109.5^\circ$.
 This happens when the magnetic field lies in the \{110\} or \{$1\bar{1}0$\} crystalline planes family.
-In this case we can numerically compute $\bar \eta_{\rm far} \approx \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot 0.8328 \approx 1.20\cdot10^{-1}$.
-This last case was not taken into account by \cite{choi_depolarization_2017}.
+In this case, numerically simulations give $\bar \eta_{\rm far} \approx \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot 0.8328 \approx 1.20\cdot10^{-1}$.
+This last case was not taken into account in \cite{choi_depolarization_2017}.
 \end{itemize}
@@
 Fig.
-\ref{121 VS 22 fig}(b) shows the subtracted $T_1$ signal measured on the same sample HPHT-15-2 for 8 different values of the magnetic field, each time on a two-class resonance : in 4 cases the magnetic field was in a \{100\} plane, and in the 4 other cases it was in a \{110\} plane.
+\ref{121 VS 22 fig}(b) shows the subtracted $T_1$ signal measured on the same sample HPHT-15-2 for 8 different values of the magnetic field, each time on a two-class resonance.
+In 4 cases, the magnetic field was in a \{100\} plane, and in the 4 other cases it was in a \{110\} plane.
 We can see that the measured lifetime is always smaller on the \{110\} case, which corresponds the the greater $\bar \eta$ factor computed previously.
@@
 \includegraphics[width=0.7\textwidth]{Figures_SI/Various_ESR}
-\caption{ODMR spectra for different orientations of the magnetic field. Red numbers represent the number of degenerated classes for each line of the spectrum}
+\caption{ODMR spectra for different orientations of the magnetic field. Red numbers represent the number of degenerate classes for each line of the spectrum}
 \label{Various ODMR}
@@
 \item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
-\item $\bm{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2= \bar \eta_0^2$.
+\item $\bm{B} \in$ \{110\} : For the two non-resonant classes : $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2= \bar \eta_0^2$.
 For the two resonant classes : $\bar \eta = \bar {\eta}_{\rm same} + \bar {\eta}_{\rm far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
@@
 \item The $x$ direction is defined by the local electric field generated by the charges in the crystal.
-We will assume that there are no correlations in the electric field between the NV and fluctuator, and will sample a random angle $\psi \in [0,2\pi]$ between the axes $\hat{x}_1$ and $\hat{x}_2$.
+We will assume that the NV and fluctuator do not see the same electric field, and will sample a random angle $\psi \in [0,2\pi]$ between the axes $\hat{x}_1$ and $\hat{x}_2$.
 \end{itemize}
@@
 The first one is described in Fig.
-4 of the main text : in the presence of pure transverse magnetic field (for a single class, non-resonant with the three other classes), the decay rate of the class decreases with the magnetic field amplitude, due to the double-flip processes (described below), and then reaches a plateau with a decay rate value $\sim 2$ times bigger than in the longitudinal magnetic field case.
+4 of the main text : in the presence of pure transverse magnetic field (for a single class, non-resonant with the three other classes), the decay rate of the class decreases with the magnetic field amplitude, due to the double-flip processes (described below), and then reaches a plateau with a decay rate value $\sim 2$ times larger than in the longitudinal magnetic field case.
 This increase is in agreement with the fact that $\bar{\eta}_{\rm same}$ is 2 times bigger in the non-magnetic $\ket{+/-}$ basis than in the magnetic $\ket{\pm 1}$ basis.
@@
 
-Double-flip processes are processes where both spins lose or gain one unit of spin angular momentum, as ooposed to flip-flop processes where one spin loses one quantum and the other gains one.
+Double-flip processes are processes where both spins lose or gain one unit of spin angular momentum, as opposed to flip-flop processes where one spin loses one quantum and the other gains one.
 These are related to matrix elements  such as $\mel{+1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-1}$ in the $\ket{\pm 1}$ basis or $\mel{+,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-}$ in the $\ket{\pm}$ basis.
-These matrix elements couple two-spin states that are never fully resonant, however with low-to zero magnetic, the residual splitting due to local electric and magnetic field ($\sim 6\ \rm MHz$ with our samples) is small enough compared to the fluctuators linewidth  measured in sec.
+These matrix elements couple two-spin states that are never fully resonant, however with low to zero magnetic, the residual splitting due to local electric and magnetic field ($\sim 6\ \rm MHz$ with our samples) is small enough compared to the fluctuators linewidth  measured in sec.
 \ref{fluctuator width} to be $\sim 8\ \rm MHz$, so that double-flip processes may still occur.

### 2022-07-08 Mes papiers/Low field CR/copie.tex
la tete au weekend

--- 2022-06-22 Mes papiers/Low field CR/Low_field_CR.tex
@@
 \usepackage{bm}% bold math
-\usepackage{natbib}
+\usepackage[sort&compress]{natbib}
 \usepackage{physics}
@@
 
-\title{Depolarization of Interacting NV ensembles : \\
-The Role of Local Electric Fields and Double Flip processes}
+\title{Depolarization of interacting NV$^-$ centers at low magnetic fields}
 
@@
 \begin{abstract}
-We present experimental results on dipolar interaction processes in ensemble of spin-1 systems.
-Using the spin of highly doped ensembles of negatively charged nitrogen vacancy centers in diamonds, we identify regimes where flip-flop, double-flip processes as well as the mixing induced by local electric field dominate.
+We study dipolar interaction processes in dense ensembles of negatively charged nitrogen vacancy (NV$^-$) centers under small magnetic fields.
+Employing magnetic field scans close to zero-field along specific crystalline direction while recording the NV$^-$ spin relaxation rate, we could identify regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the NV-NV cross-relaxation .
 %We analyse these results from the perspective of magnetometry applications, and show that both double-flip processes and local electric fields improve the magnetic field sensitivity.
 Our results are relevant for understanding decoherence in many-body spin systems as well as for high sensitivity magneto- and electro-metry with long-lived interacting solid-state spins.
-As a proof of principle, we present an orientation-free and microwave-free magnetometer operating with a sensitivity below $100$ nT/$\sqrt{\rm Hz}$.
+As a proof of principle, we present an orientation and microwave-free magnetometer based on cross-relaxation, and operating with a sensitivity below $100$ nT/$\sqrt{\rm Hz}$.
 \end{abstract}
@@
 
-The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science in part thanks to the possibility to optically polarize and read-out its spin state at ambient conditions.
-Many recent work indeed focus on ensembles of spins from NV centers with the prospect to boost the magnetic field sensing capabilities and for studying many body effects
-\citep{TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized, kucsko2018critical, giri_coupled_2018}
+The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science thanks in part to the possibility to optically polarize and read-out its spin state at ambient conditions \cite{DOHERTY20131}.
+In particular, ensembles of NV centers are widely studied for their enhanced magnetic field sensing capabilities \cite{Acosta, TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized} and as pristine platforms for studying many body effects \citep{kucsko2018critical, ChoiNat, ZuYao}.
+When the NV spin concentration reaches ppm values, spin depolarisation, or cross-relaxation (CR) takes place through a very rich many body dynamics associated with disorder \citep{choi_observation_2017}.
+These mechanisms however limit the efficiency of typical micro-wave based NV magnetometers, but can be harnessed to attain sub-heisenberg limited magnetometers \citep{zhou2020quantum}.
+Further, cross-relaxation mechanisms can be turned as a tool for magnetic field sensing \cite{akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019}.
+Spectral features in the photoluminescence indeed appear when the magnetic field crosses specific crystal planes where dipolar interactions are enhanced, leading to CR.
+The projected sensitivity of such magnetometers lie in the tens of pT$/\sqrt{\rm Hz}$  \cite{akhmedzhanov_microwave-free_2017}, on a par with the most sensitive micro-wave based NV magnetometers \cite{Wolf, Sturner}.
 
-When the spin ensemble becomes dense, inhomogeneities in the relaxation rates of NVs can give rise to spin depolarisation with a very rich many body dynamics associated with disorder.
-This mechanisms however limits the efficiency of magnetometers \citep{zhou2020quantum}
+%Typical NV magnetometers operate using microwave scans across the spin resonance.  
 %It was proposed to exploit these CR to perform sensitive microwave free magnetometry (REF).  
-Recent magnetometry proposals have been put forward at zero magnetic field.
-There, the depolarisation effect was shown to be much stronger, but the detailed studies of relaxation mechanisms have not been identified to the best of our knowledge.
-Some of the possible processes that lead to cross-relaxation in strongly coupled dipolar systems are depicted in Fig.~1.
-The goal of this paper is to identify regimes where the spin flip-flop within different classes of NV centers, double-flip up and down as well as the mixing induced by local electric field play a role.
+%Recent magnetometry proposals have been put forward at zero magnetic field.
 
-The electronic spin of NV center is a spin-1 system in the ground state (see Fig.
-1).
-It can be optically polarized in the $\ket{m_s=0}$ state.
-The photoluminescence of this state is also is larger than the $\ket{m_s=\pm 1}$ states enabling spin-read out at ambient temperature.
+Cross-relaxation features at zero-field have also been observed and could also be deployed for high sensitivity micro-wave free magnetometry.
+The CR contrast was indeed shown to be much stronger \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}.
+However all the relaxation mechanisms at these low field levels have not been identified.
+Here, we study dipolar interaction processes in ensembles of NV centers under small magnetic fields.
+Specifically, by employing magnetic field scans along specific crystalline directions, we identify regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the CR.
+We also present an orientation and microwave-free magnetometer operating with a sensitivity below $100$ nT/$\sqrt{\rm Hz}$ at zero-field, where double-flip processes amongst NV centers dominate.
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{Figures/shema_summary.pdf}
+\caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation : flip-flop processes involving two different classes of NV centers, local electric field mixing and double-flip processes where two units of spin angular momentum are exchanged.
+}
+\label{schema_intro}
+\end{figure}
+
+The electronic spin of NV center is a spin-1 system in the electronic ground state (see Fig.
+1, top left panel), which can be optically polarized in the $\ket{m_s=0}$ state.
+The photoluminescence of this state is also larger than in the $\ket{m_s=\pm 1}$ states enabling spin-read out at ambient temperature \cite{DOHERTY20131}.
 The $\ket{m_s=\pm 1}$ spin states are separated from the $\ket{m_s=0}$ state by  $D = (2\pi) 2.87$ GHz so that when a resonant microwave or static transverse magnetic field is applied, the photoluminescence is reduced \citep{epstein2005anisotropic,lai2009influence}.
+The processes that lead to cross-relaxation in such a spin-1 system are depicted in Fig.~1.
+Flip-flop processes involving coupling between NV centers from identical or from different orientations are depicted in the top-right panel.
+They have already been shown to play a dominant role when tuning $\delta$ {\it via} longitudinal B fields scans, when $B \gtrsim 30$G \cite{choi_observation_2017}.
+We will show here that under smaller magnetic fields or using purely transverse fields, double-flip up and down as well as the mixing induced by local electric field give a significant contribution to the CR (two bottom panels).
+
 Fig.
-\ref{T1}-a) and b) show the change in photoluminescence (PL) with respect to an external magnetic field for two samples with a low and high concentration of NV$^-$ centers (see SI).
+\ref{T1}-a) and b) show the change in photoluminescence (PL) as a function of an external magnetic field for two samples with a low and high concentration of NV$^-$ centers respectively (see SI sec.
+II \citep{SI_low_filed_CR} \nocite{anishchik2015low, filimonenko2020weak, van1990electric}).
 Both samples show a decrease in PL as the magnetic field amplitude increases.
 There is however a stark difference in the low magnetic field region where only high-density samples shows a drop in PL \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}.
-This effect can be observed on all our samples (see SI) whenever the NV concentration lies in the ppm range.
-The drop of the PL at low magnetic field is associated with a decrease of the NV's spin lifetime $T_1$ \citep{jarmola_temperature-_2012}, which results in a decrease of the population in the bright state $\ket{0}$.
+This effect can be observed on all our samples whenever the NV concentration lies in the ppm range.
+The drop of the PL at low magnetic field was shown to be associated with a decrease of the NV's spin lifetime $T_1$ \citep{jarmola_temperature-_2012}, which results in a decrease of the population in the bright state $\ket{0}$.
 
@@
 %To understand the behavior of the PL in zero-field, we therefore need to study the dynamics of the spin in the low magnetic field region.
-The depolarization dynamics of the spin of single or dilute NV centers at room temperature is dominated by two-phonon Raman processes \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depend on the crystal lattice temperature.
-It has been observed that dense NV centers ensemble have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi2017depolarization, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field amplitude.
+The depolarization dynamics of the NV spins for single or dilute ensembles of NV centers at room temperature is dominated by two-phonon Raman processes \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depend on the crystal lattice temperature.
+It has been observed that dense NV centers ensemblesin fact  have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi2017depolarization, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field amplitude.
 This effect has been attributed to cross-relaxation between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi2017depolarization}.
-Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble.
+The reason is that at zero-field many NV ``classes" - that is physical orientation of the NV axis is the diamond crystal cell - becomes resonant, which increases the dipolar relaxation.
+Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble \citep{choi2017depolarization}.
 We will denote $T_1^{\rm ph}$ the characteristic timescale associated with the phonon relaxation process and $T_1^{\rm dd}$ the density dependent timescale associated with the dipole-dipole relaxation process.
-
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/shema_summary.pdf}
-\caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation. }
-\label{schema_intro}
-\end{figure}
 
@@
 (b)-iii) Stretched part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude.
-(c)-i), (c)-ii) and (c)-iii) : same measurements but with a magnetic field close to the [100] axis}
+(c)-i), (c)-ii) and (c)-iii) : same measurements but with a magnetic field close to the [100] axis.
+All the measurements were realized using the sample HPHT-150-1.
+}
 \label{100_VS_1x4}
@@
 
-Fig. \ref{T1}-c) shows the sequence employed for measuring $T_1^{\rm dd}$ in the dense sample. 
+Before analyzing in detail the cause of low-field depolarization, we present $T_1$ measurements realized on HPHT-150-1. Fig. \ref{T1}-c) shows the sequence employed for measuring $T_1^{\rm dd}$. 
 %The spin lifetime protocol used here is described in Fig. \ref{T1} and is based on previous similar experiments 
@@
 In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019}.
-It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
+It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right on one of the eight NV spin-resonances before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
 By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi2017depolarization}.
 
-Fig. \ref{T1}-b) shows the subtracted signals when $\bm B=0$ and $\bm B \neq 0$.
-The ensemble spin decay curves are not exponential, which was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi2017depolarization}.
+Fig. \ref{T1}-d) shows the subtracted signals $S_1-S_2$ when $B\approx 0$ and $B \approx 50\ \rm G$.
+The relaxation rate is markedly different in both magnetic field settings, with a much more pronounced depolarization at zero-field.
+Notably, the ensemble spin decay curves are not exponential. This was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi2017depolarization}.
 We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi2017depolarization}.
-A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$.
-Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$.
-For the samples used in this paper (detailed in SI) $T_1^{\rm dd} \sim T_1^{\rm ph}$, so both decay processes have to be included in the analysis.
-Fitting the two curves by the product of an exponential and a stretched exponential, we find $T_1^{\rm ph}=3.6$ ms for both and $T_1^{\rm dd}= 0.6\ \rm ms$ and $13.0\ \rm ms$ respectively.
-We also conducted further measurements that confirm the lifetime-limited character of the fluctuator employed in the model \citep{choi2017depolarization}, see SI.
+This model postulates the existence of very short-lived NV centers - the so-called fluctuators - which, when coupled resonantly to the other NV centers, act as a polarization drain for the spin ensemble.
+A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced relaxation should be a stretched-exponential with a stretch factor $\beta=1/2$.
+Such a trend was indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$ \cite{choi2017depolarization}.
+We conducted further measurements to confirm the stretch nature of the dipole-induced decay (see SI sec.
+IV A), as well as confirming the lifetime-limited model of the fluctuators by measuring the resonance condition between classes of NV centers (see SI sec.
+IV B).
+In all dense samples used in this paper (detailed in SI sec.
+II) $T_1^{\rm dd} \sim T_1^{\rm ph}$, so both decay processes have to be included in the analysis.
+Fitting the two curves in Fig.
+2 d) by the product of a simple and a stretched exponential, we find $T_1^{\rm ph}=3.6$ ms for both curves and $T_1^{\rm dd}= 0.6\ \rm ms$ and $13.0\ \rm ms$ for the $B=0$ and $B=50$G cases respectively.
+These results thus demonstrates a twenty-fold increase in the dipolar depolarization rate when the B field is turned to zero.
+While the fluctuator model developed in \citep{choi2017depolarization} contains some of the explanation for this increase in dipolar depolarization, due to more classes being resonant in zero field than in non-zero field, the model did not include the specificity of the zero magnetic field region, such as the importance of local electric field and the double flip processes.
+We therefore decided to conducted more in depth experiments in order to characterize these zero field specificities.
+%We also conducted further measurements that confirm the lifetime-limited character of the fluctuator employed in the model \citep{choi2017depolarization}, see SI sec.
+IV B.
+Using this model and the increase in the single flip-flops rates through the change in the number of resonant spins, we cannot predict such as large change in the relaxation (See SI).
+We thus conducted more in depth experiments in order to extract the other mechanisms at stake.
+
 %These results show that our sample clearly manifests dipolar interactions amongst closely packed NV centers, but do not tell about the magnitude of the several possible mechanisms involved in the dipolar relaxation.
@@
 %Further details on the fitting procedure are available in SI. For the rest of the article we will be only interested on the stretched exponential lifetime ${T_1^{\rm dd}}$.
-
-One of the ways to tune dipole-dipole related phenomena is by changing the number of resonant spins.
+Fig. \ref{100_VS_1x4} (a) shows an optically detected magnetic resonance (ODMR) spectrum at zero magnetic field, from the dense ensemble HPHT-150-1.
 Fig.
-\ref{100_VS_1x4} (b)-i) shows an optically detected magnetic resonance (ODMR) spectrum from the dense ensemble HPHT-150-1.
-The observed lines correspond to the projections of the magnetic fields onto the four "classes" of NV centers - that is physical orientation of the NV axis is the diamond crystal cell - together with the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$.
+\ref{100_VS_1x4} (b)-i) shows an ODMR spectrum from the same sample, but in a magnetic field.
+The observed lines correspond to the projections of the magnetic fields onto the four classes of NV centers together with the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$.
 The transitions frequencies of the four different classes can be controlled by changing the amplitude and orientation of the external magnetic field.
@@
 %Other example of class resonances are given in the SI.
-This configuration will be essential to decipher what dipolar relaxation mechanisms play a role at low field.
+This configuration will be essential to decipher which dipolar relaxation mechanisms play a role at low field.
 
-Fig. \ref{100_VS_1x4} (b)-ii) and (b)-iii) show the PL from an ensemble of NV centers as a function of the amplitude of a magnetic field aligned as in Fig. \ref{100_VS_1x4} (b)-i). 
-We also extract the stretched exponential spin decay time $T_1^{\rm dd}$ by running the sequence shown in Fig. 2-c) at each of the 0.5 G magnetic field values. 
-We observe that the 3 \% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.08 kHz, as the magnetic field increases.
-This result can be explained by the higher number of resonant spins in zero magnetic field, where all four classes are degenerate, as opposed when $\bm B \neq 0$.
-The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SI ?) [Je sais pas, je le fais ?].
-Note that the decrease in PL for $B>40\ \rm G$ are related to state mixing by the transverse magnetic field, so it is not correlated to a modification of the spin lifetime.
+Fig. \ref{100_VS_1x4} (b)-ii) show the PL from the NV centers ensemble as a function of the amplitude of a magnetic field aligned as in Fig. \ref{100_VS_1x4} (b)-i). 
+Fig. \ref{100_VS_1x4} (b)-iii) shows the stretched exponential spin decay time $T_1^{\rm dd}$ obtained using the sequence shown in Fig. 2-c) in the same range of magnetic field values, in 5 G steps. 
+We observe that the 3~\% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.08 kHz, as the magnetic field increases.
+This result can again be explained by the higher number of resonant spins as the zero magnetic field is reduced, where all four classes are degenerate, as opposed to $\bm B \neq 0$.
+The half width of the dip ($\approx 10$G) is consistent with a fluctuator model taking into account the change in the spin resonance overlaps as the $B$ field is swept (see SI sec.
+IV B).
+Note that the decrease in PL for $B>40\ \rm G$ is related to state mixing by the transverse magnetic field, so it is not associated with a modification of the spin lifetime.
 
-Importantly, we observed that the lifted degeneracy of the four NV classes is not the sole reason for the increase of the spin lifetime as the B field increases.
+Importantly, B field scans along the [100] direction, where the four classes are always resonant, still features an increase of the spin lifetime as the B field increases.
 Fig.
-\ref{100_VS_1x4} (c)-ii) and (c)-iii) show the change in the PL and in $1/T_1^{\rm dd}$ as a function of a magnetic field that is aligned along the [100] axis.
-For this particular orientation, the four classes are always resonant, regardless of the field amplitude.
+\ref{100_VS_1x4} (c)-ii) and (c)-iii) show the change in the PL and in $1/T_1^{\rm dd}$ as a function of a magnetic field that is aligned in this direction.
 We can see that there still is an increase in the spin decay rate and a corresponding drop in the PL under low magnetic field values, although considerably reduced.
@@
 
-We identified and isolated two additional mechanisms that could explain this effect : local electric fields and double-flip processes. 
+We identify and isolate two additional mechanisms that could explain this effect : local electric fields and double-flip processes. 
 The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
@@
 \end{equation}
-Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$.
-In situations such as the one described in Fig.XX, the only relevant terms of $\mathcal{H}^{\rm dd}$ in terms of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$.
-In zero magnetic field however, we have to take into account these two other mechanisms of the dipolar Hamiltonian in consideration.
+where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\bm r$ is the relative position of the two spins and $\bm  u = \frac{\bm  r}{\norm{\bm  r}}$.
+In situations such as the one described in Fig.~\ref{100_VS_1x4} (c)-i), the only relevant terms for depolarisation in $\mathcal{H}^{\rm dd}$ are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$.
+In zero magnetic field however, we have to take into account two other mechanisms of the dipolar Hamiltonian in consideration.
 %where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.
-It was shown in  \cite{mittiga2018imaging} that local electric fields coming from charge traps such as $P_1^+$ and $NV^-$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in large NV ensembles.
+It was shown in  \cite{mittiga2018imaging} that local electric fields coming from $P_1^+$ or NV$^-$ centers are responsible for the optically detected spin resonance profile at zero magnetic field in dense NV ensembles.
 Magnetic field noise comes in to second order in zero field.
 %We also do not take into account the hyper-fine structure of the NV center due to the large inhomogeneous broadening of the ESR transitions (see SI).
-We will then consider the following spin Hamiltonian for the electric field dependent NV$^-$ ground state : 
+We will then consider the following spin Hamiltonian for the electric field dependent NV$^-$ ground state: 
 \begin{equation}
@@
 Under local electric fields with orientations given by the angle $\phi_E={\rm tan}(E_x/E_y)$, the eigenstates of $H_{NV}$ are $\ket{0}$ and $\ket{\pm}=\frac{1}{\sqrt{2}}(\ket{+1}\pm e^{-i\phi_E}\ket{-1})$.
-Computing the flip-flop terms $\ket{\pm,0} \bra{0,\pm} $ in the fluctuator model shows that $1/T_1^{dd}$ should be increased at zero field, even after averaging over all possible orientations $\phi_E$ (see SI).
+Computing the flip-flop terms $\ket{\pm,0} \bra{0,\pm} $ in the fluctuator model shows that $1/T_1^{\rm dd}$ should be increased at zero field, even after averaging over all possible orientations $\phi_E$ (see SI sec.
+V).
 Another mechanism at stake when considering dipolar interactions with spin-1 systems are double-flip processes.
-These processes correspond to terms like $\ket{+1,0} \bra{0,-1}$ in the anisotropic (?) dipolar interaction, giving rise to an exchange of two units of spin-angular momentum.
-These processes can become significant when the states $\pm 1$ are degenerate and are thus likely to explain part of the zero-field dip in Fig.
-c)-ii).
+These processes correspond to terms like $\ket{+1,0} \bra{0,-1}$ in the dipolar interaction, giving rise to an exchange of two units of spin-angular momentum.
+These processes can become significant when the states $\ket{m_s=\pm 1}$ are degenerate and are thus likely to explain part of the zero-field dip in Fig.
+3-c)-ii).
 %Theoretical model predicts a factor of $\approx $5 increase in $1/T_1^{dd}$ at zero field.
 
-In order to discriminate the role of these two effects in the low-field region, we performed another similar study under a purely transverse magnetic field.
-This regime in fact emulates the effect of the electric field while turning off the resonance condition for double flip processes : indeed for relatively weak transverse magnetic field, the eigen-basis of the spin Hamiltonian is close to $\ket{0},\ket{+},\ket{-}$.
-This property has been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
+In order to discriminate the role of these two effects in the low-field region, we make use of a purely transverse magnetic field.
+The transverse B field emulates the effect of the electric field, while turning off the resonance condition for the effectiveness of double-flip processes.
+Indeed for relatively weak transverse magnetic field, the eigen-basis of the spin Hamiltonian is close to $\ket{0},\ket{+},\ket{-}$.
+This property has already been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
 Fig.~\ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian as a function of a purely transverse magnetic field.
@@
 \ref{B_transverse}-(b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$.
-We can see that at a field value of $\sim 130$ G, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50$ MHz, far exceeding the range of the dipole-dipole resonant coupling (see details in SI), but $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
+We can see that at a field value of $\sim 130$ G, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50$ MHz, exceeding the range of the dipole-dipole resonant coupling (see SI sec.
+IV B).
+Yet, $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
 We can therefore decouple the effects of the double-flip processes  $\propto \ket{g,d}\bra{d,e}$, which are no longer resonant when $| \bm B | \geq 100$ G, from the effects of the mixing induced by the transverse electric - or in this case magnetic - field.
-
-Fig.
-\ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers exposed to a transverse magnetic field between 20 and 130 G on another sample HPHT-150-2, upon which we measured a stretched decay $1/T_1^{\rm dd}=25\pm 5\  \rm Hz$ in the case of strong longitudinal magnetic field.
-The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
-We attribute the decrease in the spin decay rate when the detuning increases to the loss of effectiveness of the double-flip processes.
-We can notice that the decay rate reaches a plateau for $\Delta \nu \geq 30\ \rm MHz$ with a value $1/T_1^{\rm dd}=57\pm 5\  \rm Hz$ about twice as large as the decay rate for a longitudinal magnetic field.
-This plateau we attribute to the effect of the transverse magnetic field, which we assume to be similar to the effect of the electric field in the low magnetic field regime.
-We should note that the effect of the double-flip processes is about 5 times more important in the depolarization rate than the effect of the electric field, even though the states are not initially fully resonant.
-
-Based on an extension of the fluctuator model developed in SI, we predict an overall increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$.
-While the measured increase is smaller than the predicted one, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.
 
@@
 \includegraphics[width=0.45\textwidth]{Figures/fig_transverse_field_V2.pdf}
-\caption{(a) Energies for the spin Hamiltonian in the presence of pure transverse magnetic field $B_\perp$.
-The three spin levels are denoted $\ket{g}$, $\ket{d}$ and $\ket{e}$.
-(b) Blue dashed curve : frequency detuning $\Delta \nu$ between the $\ket{e}$ and $\ket{d}$ states.
-Red plain curve : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\bra{e}$ and $\bra{+}$.
-(c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single class as a function of the measured detuning $\Delta \nu$ between the two transitions $\ket{g} \to \ket {d}$ and  $\ket{g} \to \ket {e}$ for purely transverse magnetic field.
-The green dashed line correspond to the decay rate for the same class when the magnetic field is longitudinal}
+\caption{(a) Energies of the electronic spin states $\ket{g}$, $\ket{d}$ and $\ket{e}$ under a purely transverse magnetic field $B_\perp$.
+(b) Blue dashed line : frequency detuning $\Delta \nu$ between the $\ket{e}$ and $\ket{d}$ states.
+Red plain line : Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\ket{e}$ and $\ket{+}$.
+(c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single NV class as a function of the amplitude of $B_\perp$.
+The red dashed line is the value reached for high transverse field.
+The green dashed line correspond to the value found for the same sample under a longitudinal magnetic field.
+Error bars are depicted by the shaded pink region.
+The corresponding detuning between the states $\ket{d}$ and $\ket{e}$ is indicated on the top x-axis.
+The vertical black dashed line indicates the separation between regions A and B (see main text).}
 \label{B_transverse}
 \end{figure}
+
+Fig.
+\ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers on sample HPHT-150-2 exposed to a transverse magnetic field between 20 and 130 G.
+The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
+Beforehand, we measured a stretched decay $1/T_1^{\rm dd}=25\pm 5\  \rm Hz$ under strong longitudinal magnetic field.
+This value is indicated by the dashed green line.
+Two regions can be identified in Fig.
+\ref{B_transverse}-(c).
+In  region A, the spin decay rate decreases from 175 Hz to 60 Hz.
+We attribute this effect to the loss of effectiveness of the double-flip processes.
+In region B, $\Delta \nu \geq 30\ \rm MHz$, the decay rate plateaus to $1/T_1^{\rm dd}=57\pm 5\  \rm Hz$, which is about twice as large as the decay rate measured under a longitudinal magnetic field.
+This plateau can be attributed to the change of eigenstate basis in the presence of transverse magnetic field, which is the same as under low magnetic field, where local transverse electric fields dominates the spin Hamiltonian (see SI sec.
+I).
+From these measurements, we conclude that the effect of the double-flip processes is about 5 times more important than the effect of the electric field in the depolarization rate.
+%even though the states are not initially fully resonant.
+
+Based on an extension of the fluctuator model developed in SI sec.
+V, we predict an overall increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$.
+While the measured increase is smaller than the predicted one, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.
 Similarly, we compute the predicted decay rate when $\bm B \parallel \left[100\right]$ case compared when $\bm B = 0$, and find a theoretical increase of $\sim 20\%$ when $\bm B = 0$ due to the mixing by the local electric field.
@@
 \includegraphics[width=0.45\textwidth]{Figures/fig_magneto}
-\caption{Low field magnetometry protocol.
+\caption{Low-field magnetometry protocol.
 (a) Demodulated photoluminescence as a function of an externally applied magnetic field with an additional oscillatory magnetic field.
@@
 (c) Histogram of the measurement in Fig.
-(b) fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
+(b) fitted by gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
 (d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.}
@@
 DC microwave-free magnetometry has already been performed using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}.
-Here we propose to perform a similar protocol , but using the spin depolarization at zero-field as proposed in \cite{filimonenko2018weak, filimonenko2022manifestation}.
-The main advantage with the previously mentioned protocols is the fact that the sensitivity does not depend crucially on crystalline orientation, making this protocol applicable with diamond powders or polycrystalline samples.
-
+Here we propose to perform a similar protocol, but using the spin depolarization at zero-field as proposed in \cite{filimonenko2018weak, filimonenko2022manifestation}.
+The main advantage compared to previously employed protocols is that the sensitivity does not depend crucially on crystalline orientation, making this protocol applicable with diamond powders or polycrystalline samples.
+In order to optimize signal-to-noise ratio, we use a lock-in detection and add a magnetic field modulation at $\sim 1\ \rm kHz$ with an amplitude $\sim 10\ \rm G$ through the same electromagnet.
 Fig.
-\ref{magneto} (a) shows the demodulated PL as a DC magnetic field is scanned in a random direction and an additional alternating magnetic field of amplitude $\sim 10\ \rm G$ and frequency  $\sim 1\ \rm kHz$ is added through the same electromagnet.
+\ref{magneto}-(a) shows the demodulated PL as a DC magnetic field is scanned in an arbitrary direction.
 The sample used here is HPHT-15-1 and the laser power $\sim 1\ \rm mW$.
 We can see a sharp linear slope in low field $\abs{B} < 5\ \rm G$.
-Once calibrated, in this case with ODMR, the slope can provide a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as in \cite{zheng_microwave-free_2020}.
+Once calibrated, in this case with ODMR, the slope provides a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as in \cite{zheng_microwave-free_2020}.
 In order to assess the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take a histogram of the measured fields, as shown in Fig.
 \ref{magneto} (b) and (c).
-The histogram is well fitted with gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
+The histogram is well fitted by gaussians of standard deviation $\sigma=1.5\ \mu \rm T$.
 The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which gives us a sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nT}/\sqrt{\rm Hz}$.
+Normalized to the volume, this gives $\eta=\sigma \sqrt{\tau}=4.7 \ \mu \rm{T}\mu m^{3/2}/\sqrt{\rm Hz}$.
+This measurement is consistent with the experimentally found $\sim 5\ \mu \rm{T}/\sqrt{\rm Hz}$ sensitivity obtained with 1 $\mu$m$^3$ diamonds from the same supplier.
 
-We now evaluate the relative role of the three causes of spin depolarization in the magnetometer, namely the splitting of the classes, the local electric field and the double-flip processes.
+We now evaluate the relative role of the three causes of spin depolarization that assist the operation the magnetometer, namely the splitting of the NV classes, the local electric field and the double-flip processes.
 In order to do so, we measure the magnetometer sensitivity while changing the angle of the magnetic field.
@@
 \ref{magneto} (d).
-We can see a slight increase of $\sim$ 10\% in the sensitivity as we leave the [100] region, but overall the sensitivity remains relatively flat.
+We can see a $\sim$ 10\% declining sensitivity as we leave the [100] region, but overall the sensitivity remains relatively invariant.
 When $\bm B$ is aligned with the [100] crystalline axis, only the double flips and the electric field cause a depolarization, whereas in every other orientation the three effects are at play.
 The double-flips and electric field effects are thus the dominant factors in the sensitivity of this protocol.
-While it may seem surprising that the effects with a lower contribution on the PL contrast have a higher effect on this PL-based protocol, what matters here is not the absolute contrast but the slope of the change of contrast with respect to the magnetic field.
+While it may seem surprising that the effects with a lower contribution to the PL contrast have a higher effect on this PL-based protocol, what matters here is not the absolute contrast but the slope of the change of contrast with respect to the magnetic field.
 The latter is in fact larger for the local electric field and double flips processes than for the process involving only flip-flops.
@@
 The lift in degeneracy of the spin state of different NV classes was found to be the main cause of zero-field depolarization, followed by double-flip processes and then the electric field induced mixing.
-We have employed cross-relaxation for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $10\ \mu \rm m$ commercially available diamond and show that that double-flips and electric field play an important role.
-Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins. 
+We have employed cross-relaxation for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $15\ \mu \rm m$ commercially available diamond and show that that double-flips and electric field play an important role.
+Our studies will be important for micro-wave based low field magnetometry \cite{Vetter_LFM, WangRB}. 
+Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins under purely transverse as well as small magnetic fields. 
 %DEER, many body, critical thermalization, 2D. Double quantum
@@
 
-\bibliographystyle{plain}
+We would like to acknowledge support from Alexandre Tallaire and Jocelyn Achard as well as 
+SIRTEQ for funding.
+
+\bibliographystyle{aipauth4-1}
 \bibliography{CR}{}

### 2022-07-11 Manuscrit Thèse/chapter 2/Appendix dipole-dipole.tex
tata yoyo

@@
 For an initial state of two spins $\ket{m_s^1=i;m_s^2=j}$, a flip-flop process will result in a final state $\ket{i+1;j-1}$ or $\ket{i-1;j+1}$.
-The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$
+The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$, as well as the resonance condition between the initial and final state.
+We should note that for two identical spins, the flip-flop process $\ket{i;i\pm 1}\bra{i\pm 1,i}$ is always resonant.
+
+Other mechanisms enabled by the dipole-dipole Hamiltonian are the double-flips up and down which couple the states $\ket{i;j}$ and $\ket{i-1;j-1}$ or $\ket{i+1;j+1}$.
+For a single spin species in a strong magnetic field, the lift of the various spin levels by the Zeeman effect means that no double-flip process can be resonant.
+However with weak magnetic field or when two different spin species are present, double flip process can be as relevant as flip-flops, the matrix elements $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \pm 1}$ typically being of the same order of magnitude as the flip-flop ones.
+
+%réécris moi ça, c'est naze
+When two spin populations of different polarization can be resonantly coupled by either flip-flop or double-flip, cross-relaxation happens, that is a macroscopic transfer of polarization from the most polarized spin group to the least polarized one.
 

### 2022-07-11 Mes papiers/Low field CR/V3/Low_field_CR.tex
tata yoyo

@@
 The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science thanks in part to the possibility to optically polarize and read-out its spin state at ambient conditions \cite{DOHERTY20131}.
-In particular, ensembles of NV centers are widely studied for their enhanced magnetic field sensing capabilities \cite{Acosta, TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized} and as pristine platforms for studying many body effects \citep{kucsko2018critical, ChoiNat, ZuYao}.
+In particular, ensembles of NV centers are widely studied for their enhanced magnetic field sensing capabilities \cite{Acosta, TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized} and as pristine platforms for studying many body effects \citep{kucsko2018critical, ChoiNat, ZuYao, dwyer2021probing}.
 When the NV spin concentration reaches ppm values, spin depolarisation, or cross-relaxation (CR) takes place through a very rich many body dynamics associated with disorder \citep{choi_observation_2017}.
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/shema_summary.pdf}
-\caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation :flip-flop processes involving two different classes of NV centers, local electric field mixing and double-flip processes where two units of spin angular momentum are exchanged.
+\includegraphics[width=0.45\textwidth]{shema_summary.pdf}
+\caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation: flip-flop processes involving two different classes of NV centers, local electric field mixing and double-flip processes where two units of spin angular momentum are exchanged.
 }
@@
 This effect has been attributed to CR between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi2017depolarization}.
-The reason is that at zero-field many NV ``classes" - that is physical orientation of the NV axis is the diamond crystal cell - becomes resonant, which increases the dipolar relaxation.
+The reason is that at zero-field many NV ``classes" - that is physical orientation of the NV axis in the diamond cell - becomes resonant, which increases the dipolar relaxation.
 Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble \citep{choi2017depolarization}.
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig_T1.pdf}
+\includegraphics[width=0.45\textwidth]{fig_T1.pdf}
 \caption{a) PL measurement from NV centers ensemble as a function of an external magnetic field applied in an arbitrary direction (a) for the sample CVD-PPB, with [NV$^-$]$\approx 50\ \rm ppb$ (b) for the sample HPHT-150-1, with [NV$^-$]$\approx 3\ \rm ppm$.
@@
 \begin{figure*}
-\includegraphics[width=.95\textwidth]{Figures/fig_100_vs_1x1x1x1.pdf}
+\includegraphics[width=.95\textwidth]{fig_100_vs_1x1x1x1.pdf}
 \caption{(a) ODMR spectrum in zero field.
@@
 (b)-iii) Stretched part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude.
-(c)-i), (c)-ii) and (c)-iii) :same measurements but with a magnetic field close to the [100] axis.
+(c)-i), (c)-ii) and (c)-iii): same measurements but with a magnetic field close to the [100] axis.
 All the measurements were realized using the sample HPHT-150-1.
@@
 Fig. \ref{100_VS_1x4} (b)-iii) shows the stretched exponential spin decay time $T_1^{\rm dd}$ obtained using the sequence shown in Fig. 2-c) in the same range of magnetic field values, in 5 G steps. 
-We observe that the 3~\% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.08 kHz, as the magnetic field increases.
+We observe that the 4~\% increase in PL under low magnetic field is indeed correlated with a drop in the spin decay rate from 1.6 kHz to 0.08 kHz, as the magnetic field increases.
 This result can again be explained by the higher number of resonant spins as the zero magnetic field is reduced, where all four classes are degenerate, as opposed to $\bm B \neq 0$.
@@
 
-We identify and isolate two additional mechanisms that could explain this effect :local electric fields and double-flip processes. 
+We identify and isolate two additional mechanisms that could explain this effect: local electric fields and double-flip processes. 
  In situations such as the one described in Fig.~\ref{100_VS_1x4} (c)-i), the only relevant terms for depolarisation in $\mathcal{H}^{\rm dd}$ are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$ where $\mathcal{H}^{\rm dd}$ is the dipole-dipole Hamiltonian detailed in SM sec.
 V.
-In zero magnetic field however, we have to take into account two other mechanisms of the dipolar Hamiltonian in consideration.
+In zero magnetic field however, we have to take two other mechanisms of the dipolar Hamiltonian into consideration.
 
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig_transverse_field_V3.pdf}
-\caption{(a) Energies of the electronic spin states $\ket{g}$, $\ket{d}$ and $\ket{e}$ under a purely transverse magnetic field $B_\perp$.
-(b) Blue dashed line :frequency detuning $\Delta \nu$ between the $\ket{e}$ and $\ket{d}$ states.
-Red plain line :Matching factor $\abs{\bra{e}\ket{+}}^2$ between the states $\ket{e}$ and $\ket{+}$.
-(c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single NV class as a function of the amplitude of $B_\perp$.
+\includegraphics[width=0.45\textwidth]{fig_transverse_field_V3.pdf}
+\caption{(a) ODMR spectrum for $B_\perp=20\ \rm G$.
+The two transition frequencies of the class orthogonal to $B$ are denoted $\nu_+$ and $\nu_-$(b) Measurement of the frequencies $\nu_+$ and $\nu_-$ through ODMR spectra as a function of the transverse field amplitude (c) Measurement of the stretched exponential decay rate $1/T_1^{\rm dd}$ for a single NV class as a function of the amplitude of $B_\perp$.
 The red dashed line is the value reached for high transverse field.
@@
 Error bars are depicted by the shaded pink region.
-The corresponding detuning between the states $\ket{d}$ and $\ket{e}$ is indicated on the top x-axis.
+The detuning $\Delta \nu = \nu_+-\nu-$ is indicated on the top x-axis.
 The vertical black dashed line indicates the separation between regions A and B (see main text).}
@@
 
-In order to discriminate the role of these two effects in the low-field region, we make use of a purely transverse magnetic field.
-The transverse B field emulates the effect of the electric field, while turning off the resonance condition for the effectiveness of double-flip processes.
-Indeed for relatively weak transverse magnetic field, the eigen-basis of the spin Hamiltonian is close to $\ket{0},\ket{+},\ket{-}$.
-This property has already been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
-Fig.~\ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian as a function of a purely transverse magnetic field.
-We will denote these eigenstates in the general case $\ket{g}$, $\ket{d}$ and $\ket{e}$.
-For weak transverse magnetic field, $\ket{g} \approx \ket{0}$, $\ket{d}=\ket{-}$ and $\ket{e} \approx \ket{+}$.
-As the magnetic field increases, $\ket{g}$ and $\ket{e}$ starts to become a mixing of $\ket{0}$ and $\ket{+}$ while $\ket{d}$ remains equal to $\ket{-}$.
+In order to experimentally discriminate the contribution from the electric field and the double flips, we would need to apply an electric field strong enough to lift the pseudo-degeneracy between the $\ket{+}$ and $\ket{-}$ states.
+We estimated that we would need a field of $\sim 10^6\ \rm{V/cm}$ far out of range of our experimental reach.
+Instead we decided to emulate the effects of the electric field by applying a purely transverse magnetic field on one of the NV class.
+Indeed, for $B_{\perp} \leq 150\ \rm G$, the eigenstates of the spin Hamiltonian are also $\ket{0}$, $\ket{+}$ and $\ket{-}$ within a precision < 2\%.
+This property has previously been used to increase the electric field sensing ability of NV centers \cite{dolde2011electric,qiu2022nanoscale}.
+We further discuss the validity of the method in SM sec.
+[REF]
 
-Crucially, mixing remains even when the associated energy splitting of the $\ket{d}$ and $\ket{e}$ states is large :Fig.
-\ref{B_transverse}-(b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$.
-We can see that at a field value of $\sim 130$ G, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50$ MHz, exceeding the range of the dipole-dipole resonant coupling (see SM sec.
-IV B).
-Yet, $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
-We can therefore decouple the effects of the double-flip processes  $\propto \ket{g,d}\bra{d,e}$, which are no longer resonant when $| \bm B | \geq 100$ G, from the effects of the mixing induced by the transverse electric - or in this case magnetic - field.
+Fig.~\ref{B_transverse} (a) shows an ODMR spectrum for $B_{\perp}=20\ \rm G$ on sample HPHT-150-2.
+The central two lines correspond to the $\ket{0}\to \ket{-}$ and $\ket{0}\to \ket{+}$ transitions for the class orthogonal to the magnetic field, the other features corresponding to the transitions of the other classes.
+We chose the initial value of 20 G so that the class of interest is far enough detuned from the other classes that we don't need to consider inter-class flip-flops.
 
 Fig.
-\ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers on sample HPHT-150-2 exposed to a transverse magnetic field between 20 and 130 G.
-The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
-Beforehand, we measured a stretched decay $1/T_1^{\rm dd}=25\pm 5\  \rm Hz$ under strong longitudinal magnetic field.
-This value is indicated by the dashed green line.
-Two regions can be identified in Fig.
-\ref{B_transverse}-(c).
-In  region A, the spin decay rate decreases from 175 Hz to 60 Hz.
-We attribute this effect to the loss of effectiveness of the double-flip processes.
-In region B, $\Delta \nu \geq 30\ \rm MHz$, the decay rate plateaus to $1/T_1^{\rm dd}=57\pm 5\  \rm Hz$, which is about twice as large as the decay rate measured under a longitudinal magnetic field.
-This plateau can be attributed to the change of eigenstate basis in the presence of transverse magnetic field, which is the same as under low magnetic field, where local transverse electric fields dominates the spin Hamiltonian (see SM sec.
-I).
-From these measurements, we conclude that the effect of the double-flip processes is about 5 times more important than the effect of the electric field in the depolarization rate.
+\ref{B_transverse}-(c) shows the measurement of the decay rate $1/T_1^{\rm dd}$ for the class orthogonal to the magnetic field, as a function of the transverse field amplitude.
+We can observe two regions on this graph: region A where the decay rate decreases with the magnetic field, and region B where it stabilizes to a value $1/T_1^{\rm dd}=60\pm 5\ \rm{s}^{-1}$.
+We also indicate on this graph the value $1/T_1^{\rm dd}=25\pm 5\ \rm{s}^{-1}$ found for the same class but with a magnetic field oriented parallel to the class instead of transverse to it.
 
-Based on an extension of the fluctuator model developed in SM sec.
-V, we predict an overall increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$.
-While the measured increase is smaller than the predicted one, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.
-Similarly, we compute the predicted decay rate when $\bm B \parallel \left[100\right]$ case compared when $\bm B = 0$, and find a theoretical increase of $\sim 20\%$ when $\bm B = 0$ due to the mixing by the local electric field.
-This, along with the double-flip processes explains the behavior observed in Fig.
-\ref{100_VS_1x4} (c)-ii) and (c)-iii).
+What happens in the A region is that, as we increase the transverse field value, we are also increasing the detuning $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states, as shown in Fig.
+\ref{B_transverse} (b).
+By increasing this detuning, we are progressively quenching the double-flips until we reach the B region where only the transverse field effect remains.
+We can therefore directly compare the value of $60\ \rm{s}^{-1}$, corresponding the intra-class flip-flops in the $\ket{\pm}$ basis, to the $25\ \rm{s}^{-1}$ value which corresponds the intra-class flip-flops in the $\ket{\pm 1}$ basis.
+This increase is corroborated by our model detailed in SM sec.
+V where we predicted an increase in the decay rate by a factor of 4 in the $\ket{\pm}$ basis compared to the $\ket{\pm 1}$ basis.
+We can also note that the double-flips under low magnetic field are responsible for an increase in the decay rate $\sim 5$ times greater than that of the electric field, and are therefore likely the main cause of zero field depolarization in Fig.
+\ref{100_VS_1x4} (c)-iii).
 
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig_magneto}
+\includegraphics[width=0.45\textwidth]{fig_magneto}
 \caption{Low-field magnetometry protocol.
@@
 The main advantage compared to previously employed protocols is that the sensitivity does not depend crucially on crystalline orientation, making this protocol applicable with diamond powders or polycrystalline samples.
-Another sizable advantage is that this measurement is that it is insensitive to thermal variations.
+Another significant advantage is that this measurement is insensitive to thermal variations.
 In order to optimize signal-to-noise ratio, we use a lock-in detection and add a magnetic field modulation at $\sim 1\ \rm kHz$ with an amplitude $\sim 10\ \rm G$ through the same electromagnet.
@@
 Our studies will be important for microwave based low field magnetometry \cite{Vetter_LFM, WangRB}. 
-Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins under purely transverse as well as small magnetic fields.
+Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins under purely transverse or small magnetic fields.
 
-\section*{acknowledgements}
+\begin{acknowledgments}
 
@@
 
+\end{acknowledgments}
 %\bibliographystyle{unsrtnat}
-%\bibliographystyle{apsrev4-2}
-%\bibliography{CR}{}
-
-%apsrev4-2.bst 2019-01-14 (MD) hand-edited version of apsrev4-1.bst
-%Control: key (0)
-%Control: author (72) initials jnrlst
-%Control: editor formatted (1) identically to author
-%Control: production of article title (-1) disabled
-%Control: page (0) single
-%Control: year (1) truncated
-%Control: production of eprint (0) enabled
-\begin{thebibliography}{41}%
-\makeatletter
-\providecommand \@ifxundefined [1]{%
- \@ifx{#1\undefined}
-}%
-\providecommand \@ifnum [1]{%
- \ifnum #1\expandafter \@firstoftwo
- \else \expandafter \@secondoftwo
- \fi
-}%
-\providecommand \@ifx [1]{%
- \ifx #1\expandafter \@firstoftwo
- \else \expandafter \@secondoftwo
- \fi
-}%
-\providecommand \natexlab [1]{#1}%
-\providecommand \enquote  [1]{``#1''}%
-\providecommand \bibnamefont  [1]{#1}%
-\providecommand \bibfnamefont [1]{#1}%
-\providecommand \citenamefont [1]{#1}%
-\providecommand \href@noop [0]{\@secondoftwo}%
-\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
-\providecommand \@href[1]{\@@startlink{#1}\@@href}%
-\providecommand \@@href[1]{\endgroup#1\@@endlink}%
-\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
-  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
-\providecommand \@@startlink[1]{}%
-\providecommand \@@endlink[0]{}%
-\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
-\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
-\providecommand \urlprefix  [0]{URL }%
-\providecommand \Eprint [0]{\href }%
-\providecommand \doibase [0]{https://doi.org/}%
-\providecommand \selectlanguage [0]{\@gobble}%
-\providecommand \bibinfo  [0]{\@secondoftwo}%
-\providecommand \bibfield  [0]{\@secondoftwo}%
-\providecommand \translation [1]{[#1]}%
-\providecommand \BibitemOpen [0]{}%
-\providecommand \bibitemStop [0]{}%
-\providecommand \bibitemNoStop [0]{.\EOS\space}%
-\providecommand \EOS [0]{\spacefactor3000\relax}%
-\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
-\let\auto@bib@innerbib\@empty
-%</preamble>
-\bibitem [{\citenamefont {Doherty}\ \emph {et~al.}(2013)\citenamefont
-  {Doherty}, \citenamefont {Manson}, \citenamefont {Delaney}, \citenamefont
-  {Jelezko}, \citenamefont {Wrachtrup},\ and\ \citenamefont
-  {Hollenberg}}]{DOHERTY20131}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~W.}\ \bibnamefont
-  {Doherty}}, \bibinfo {author} {\bibfnamefont {N.~B.}\ \bibnamefont {Manson}},
-  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Delaney}}, \bibinfo
-  {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Wrachtrup}},\ and\ \bibinfo {author}
-  {\bibfnamefont {L.~C.}\ \bibnamefont {Hollenberg}},\ }\href
-  {https://doi.org/https://doi.org/10.1016/j.physrep.2013.02.001} {\bibfield
-  {journal} {\bibinfo  {journal} {Physics Reports}\ }\textbf {\bibinfo {volume}
-  {528}},\ \bibinfo {pages} {1 } (\bibinfo {year} {2013})},\ \bibinfo {note}
-  {the nitrogen-vacancy colour centre in diamond}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Acosta}\ \emph {et~al.}(2009)\citenamefont {Acosta},
-  \citenamefont {Bauch}, \citenamefont {Ledbetter}, \citenamefont {Santori},
-  \citenamefont {Fu}, \citenamefont {Barclay}, \citenamefont {Beausoleil},
-  \citenamefont {Linget}, \citenamefont {Roch}, \citenamefont {Treussart},
-  \citenamefont {Chemerisov}, \citenamefont {Gawlik},\ and\ \citenamefont
-  {Budker}}]{Acosta}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {V.~M.}\ \bibnamefont
-  {Acosta}}, \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {Bauch}},
-  \bibinfo {author} {\bibfnamefont {M.~P.}\ \bibnamefont {Ledbetter}}, \bibinfo
-  {author} {\bibfnamefont {C.}~\bibnamefont {Santori}}, \bibinfo {author}
-  {\bibfnamefont {K.-M.~C.}\ \bibnamefont {Fu}}, \bibinfo {author}
-  {\bibfnamefont {P.~E.}\ \bibnamefont {Barclay}}, \bibinfo {author}
-  {\bibfnamefont {R.~G.}\ \bibnamefont {Beausoleil}}, \bibinfo {author}
-  {\bibfnamefont {H.}~\bibnamefont {Linget}}, \bibinfo {author} {\bibfnamefont
-  {J.~F.}\ \bibnamefont {Roch}}, \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Treussart}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Chemerisov}}, \bibinfo {author} {\bibfnamefont
-  {W.}~\bibnamefont {Gawlik}},\ and\ \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Budker}},\ }\href
-  {https://doi.org/10.1103/PhysRevB.80.115202} {\bibfield  {journal} {\bibinfo
-  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {80}},\ \bibinfo
-  {pages} {115202} (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Tallaire}\ \emph {et~al.}(2020)\citenamefont
-  {Tallaire}, \citenamefont {Brinza}, \citenamefont {Huillery}, \citenamefont
-  {Delord}, \citenamefont {Pellet-Mary}, \citenamefont {Staacke}, \citenamefont
-  {Abel}, \citenamefont {Pezzagna}, \citenamefont {Meijer}, \citenamefont
-  {Touati}, \citenamefont {Binet}, \citenamefont {Ferrier}, \citenamefont
-  {Goldner}, \citenamefont {Hetet},\ and\ \citenamefont
-  {Achard}}]{TALLAIRE2020421}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Tallaire}}, \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Brinza}},
-  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Huillery}}, \bibinfo
-  {author} {\bibfnamefont {T.}~\bibnamefont {Delord}}, \bibinfo {author}
-  {\bibfnamefont {C.}~\bibnamefont {Pellet-Mary}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Staacke}}, \bibinfo {author} {\bibfnamefont
-  {B.}~\bibnamefont {Abel}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Pezzagna}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Meijer}},
-  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Touati}}, \bibinfo
-  {author} {\bibfnamefont {L.}~\bibnamefont {Binet}}, \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Ferrier}}, \bibinfo {author} {\bibfnamefont
-  {P.}~\bibnamefont {Goldner}}, \bibinfo {author} {\bibfnamefont
-  {G.}~\bibnamefont {Hetet}},\ and\ \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Achard}},\ }\href
-  {https://doi.org/https://doi.org/10.1016/j.carbon.2020.08.048} {\bibfield
-  {journal} {\bibinfo  {journal} {Carbon}\ }\textbf {\bibinfo {volume} {170}},\
-  \bibinfo {pages} {421 } (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Edmonds}\ \emph {et~al.}(2021)\citenamefont
-  {Edmonds}, \citenamefont {Hart}, \citenamefont {Turner}, \citenamefont
-  {Colard}, \citenamefont {Schloss}, \citenamefont {Olsson}, \citenamefont
-  {Trubko}, \citenamefont {Markham}, \citenamefont {Rathmill}, \citenamefont
-  {Horne-Smith} \emph {et~al.}}]{edmonds2021characterisation}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
-  {Edmonds}}, \bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont {Hart}},
-  \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont {Turner}}, \bibinfo
-  {author} {\bibfnamefont {P.-O.}\ \bibnamefont {Colard}}, \bibinfo {author}
-  {\bibfnamefont {J.~M.}\ \bibnamefont {Schloss}}, \bibinfo {author}
-  {\bibfnamefont {K.~S.}\ \bibnamefont {Olsson}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Trubko}}, \bibinfo {author} {\bibfnamefont
-  {M.~L.}\ \bibnamefont {Markham}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Rathmill}}, \bibinfo {author} {\bibfnamefont
-  {B.}~\bibnamefont {Horne-Smith}}, \emph {et~al.},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Materials for Quantum Technology}\ }\textbf
-  {\bibinfo {volume} {1}},\ \bibinfo {pages} {025001} (\bibinfo {year}
-  {2021})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Chatzidrosos}\ \emph {et~al.}(2021)\citenamefont
-  {Chatzidrosos}, \citenamefont {Rebeirro}, \citenamefont {Zheng},
-  \citenamefont {Omar}, \citenamefont {Brenneis}, \citenamefont {St{\"u}rner},
-  \citenamefont {Fuchs}, \citenamefont {Buck}, \citenamefont {R{\"o}lver},
-  \citenamefont {Schneemann} \emph {et~al.}}]{chatzidrosos2021fiberized}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
-  {Chatzidrosos}}, \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
-  {Rebeirro}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Zheng}},
-  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Omar}}, \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Brenneis}}, \bibinfo {author}
-  {\bibfnamefont {F.~M.}\ \bibnamefont {St{\"u}rner}}, \bibinfo {author}
-  {\bibfnamefont {T.}~\bibnamefont {Fuchs}}, \bibinfo {author} {\bibfnamefont
-  {T.}~\bibnamefont {Buck}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {R{\"o}lver}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
-  {Schneemann}}, \emph {et~al.},\ }\href@noop {} {\bibfield  {journal}
-  {\bibinfo  {journal} {Frontiers in Photonics}\ ,\ \bibinfo {pages} {4}}
-  (\bibinfo {year} {2021})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Kucsko}\ \emph {et~al.}(2018)\citenamefont {Kucsko},
-  \citenamefont {Choi}, \citenamefont {Choi}, \citenamefont {Maurer},
-  \citenamefont {Zhou}, \citenamefont {Landig}, \citenamefont {Sumiya},
-  \citenamefont {Onoda}, \citenamefont {Isoya}, \citenamefont {Jelezko} \emph
-  {et~al.}}]{kucsko2018critical}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
-  {Kucsko}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}},
-  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo {author}
-  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
-  {\bibfnamefont {H.}~\bibnamefont {Zhou}}, \bibinfo {author} {\bibfnamefont
-  {R.}~\bibnamefont {Landig}}, \bibinfo {author} {\bibfnamefont
-  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Jelezko}}, \emph {et~al.},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Physical review letters}\ }\textbf {\bibinfo
-  {volume} {121}},\ \bibinfo {pages} {023601} (\bibinfo {year}
-  {2018})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Choi}\ \emph
-  {et~al.}(2017{\natexlab{a}})\citenamefont {Choi}, \citenamefont {Choi},
-  \citenamefont {Landig}, \citenamefont {Kucsko}, \citenamefont {Zhou},
-  \citenamefont {Isoya}, \citenamefont {Jelezko}, \citenamefont {Onoda},
-  \citenamefont {Sumiya}, \citenamefont {Khemani}, \citenamefont {von
-  Keyserlingk}, \citenamefont {Yao}, \citenamefont {Demler},\ and\
-  \citenamefont {Lukin}}]{ChoiNat}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Choi}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo
-  {author} {\bibfnamefont {R.}~\bibnamefont {Landig}}, \bibinfo {author}
-  {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author} {\bibfnamefont
-  {H.}~\bibnamefont {Zhou}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
-  {Isoya}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},
-  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Onoda}}, \bibinfo
-  {author} {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author}
-  {\bibfnamefont {V.}~\bibnamefont {Khemani}}, \bibinfo {author} {\bibfnamefont
-  {C.}~\bibnamefont {von Keyserlingk}}, \bibinfo {author} {\bibfnamefont
-  {N.~Y.}\ \bibnamefont {Yao}}, \bibinfo {author} {\bibfnamefont
-  {E.}~\bibnamefont {Demler}},\ and\ \bibinfo {author} {\bibfnamefont {M.~D.}\
-  \bibnamefont {Lukin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
-  {journal} {Nature}\ }\textbf {\bibinfo {volume} {543}},\ \bibinfo {pages}
-  {221} (\bibinfo {year} {2017}{\natexlab{a}})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Zu}\ \emph {et~al.}(2021)\citenamefont {Zu},
-  \citenamefont {Machado}, \citenamefont {Ye}, \citenamefont {Choi},
-  \citenamefont {Kobrin}, \citenamefont {Mittiga}, \citenamefont {Hsieh},
-  \citenamefont {Bhattacharyya}, \citenamefont {Markham}, \citenamefont
-  {Twitchen}, \citenamefont {Jarmola}, \citenamefont {Budker}, \citenamefont
-  {Laumann}, \citenamefont {Moore},\ and\ \citenamefont {Yao}}]{ZuYao}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Zu}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Machado}},
-  \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Ye}}, \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo {author} {\bibfnamefont
-  {B.}~\bibnamefont {Kobrin}}, \bibinfo {author} {\bibfnamefont
-  {T.}~\bibnamefont {Mittiga}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Hsieh}}, \bibinfo {author} {\bibfnamefont
-  {P.}~\bibnamefont {Bhattacharyya}}, \bibinfo {author} {\bibfnamefont
-  {M.}~\bibnamefont {Markham}}, \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Twitchen}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Jarmola}}, \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Budker}}, \bibinfo {author} {\bibfnamefont {C.~R.}\
-  \bibnamefont {Laumann}}, \bibinfo {author} {\bibfnamefont {J.~E.}\
-  \bibnamefont {Moore}},\ and\ \bibinfo {author} {\bibfnamefont {N.~Y.}\
-  \bibnamefont {Yao}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
-  {journal} {Nature}\ }\textbf {\bibinfo {volume} {597}},\ \bibinfo {pages}
-  {45} (\bibinfo {year} {2021})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Choi}\ \emph
-  {et~al.}(2017{\natexlab{b}})\citenamefont {Choi}, \citenamefont {Choi},
-  \citenamefont {Landig}, \citenamefont {Kucsko}, \citenamefont {Zhou},
-  \citenamefont {Isoya}, \citenamefont {Jelezko}, \citenamefont {Onoda},
-  \citenamefont {Sumiya}, \citenamefont {Khemani}, \citenamefont {von
-  Keyserlingk}, \citenamefont {Yao}, \citenamefont {Demler},\ and\
-  \citenamefont {Lukin}}]{choi_observation_2017}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Choi}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo
-  {author} {\bibfnamefont {R.}~\bibnamefont {Landig}}, \bibinfo {author}
-  {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author} {\bibfnamefont
-  {H.}~\bibnamefont {Zhou}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
-  {Isoya}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},
-  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Onoda}}, \bibinfo
-  {author} {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author}
-  {\bibfnamefont {V.}~\bibnamefont {Khemani}}, \bibinfo {author} {\bibfnamefont
-  {C.}~\bibnamefont {von Keyserlingk}}, \bibinfo {author} {\bibfnamefont
-  {N.~Y.}\ \bibnamefont {Yao}}, \bibinfo {author} {\bibfnamefont
-  {E.}~\bibnamefont {Demler}},\ and\ \bibinfo {author} {\bibfnamefont {M.~D.}\
-  \bibnamefont {Lukin}},\ }\href {https://doi.org/10.1038/nature21426}
-  {\bibfield  {journal} {\bibinfo  {journal} {Nature}\ }\textbf {\bibinfo
-  {volume} {543}},\ \bibinfo {pages} {221} (\bibinfo {year}
-  {2017}{\natexlab{b}})},\ \bibinfo {note} {number: 7644 Publisher: Nature
-  Publishing Group}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Zhou}\ \emph {et~al.}(2020)\citenamefont {Zhou},
-  \citenamefont {Choi}, \citenamefont {Choi}, \citenamefont {Landig},
-  \citenamefont {Douglas}, \citenamefont {Isoya}, \citenamefont {Jelezko},
-  \citenamefont {Onoda}, \citenamefont {Sumiya}, \citenamefont {Cappellaro}
-  \emph {et~al.}}]{zhou2020quantum}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
-  {Zhou}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Choi}}, \bibinfo
-  {author} {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Landig}}, \bibinfo {author} {\bibfnamefont
-  {A.~M.}\ \bibnamefont {Douglas}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Jelezko}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
-  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
-  {P.}~\bibnamefont {Cappellaro}}, \emph {et~al.},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Physical review X}\ }\textbf {\bibinfo
-  {volume} {10}},\ \bibinfo {pages} {031003} (\bibinfo {year}
-  {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2017)\citenamefont
-  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
-  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
-  and\ \citenamefont {Hemmer}}]{akhmedzhanov_microwave-free_2017}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
-  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
-  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
-  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
-  {\bibfnamefont {I.}~\bibnamefont {Zelensky}},\ and\ \bibinfo {author}
-  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href
-  {https://doi.org/10.1103/PhysRevA.96.013806} {\bibfield  {journal} {\bibinfo
-  {journal} {Phys. Rev. A}\ }\textbf {\bibinfo {volume} {96}},\ \bibinfo
-  {pages} {013806} (\bibinfo {year} {2017})},\ \bibinfo {note} {number:
-  1}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Akhmedzhanov}\ \emph {et~al.}(2019)\citenamefont
-  {Akhmedzhanov}, \citenamefont {Gushchin}, \citenamefont {Nizov},
-  \citenamefont {Nizov}, \citenamefont {Sobgayda}, \citenamefont {Zelensky},\
-  and\ \citenamefont {Hemmer}}]{akhmedzhanov_magnetometry_2019}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Akhmedzhanov}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
-  {Gushchin}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Nizov}},
-  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Nizov}}, \bibinfo
-  {author} {\bibfnamefont {D.}~\bibnamefont {Sobgayda}}, \bibinfo {author}
-  {\bibfnamefont {I.}~\bibnamefont {Zelensky}},\ and\ \bibinfo {author}
-  {\bibfnamefont {P.}~\bibnamefont {Hemmer}},\ }\href
-  {https://doi.org/10.1103/PhysRevA.100.043844} {\bibfield  {journal} {\bibinfo
-   {journal} {Phys. Rev. A}\ }\textbf {\bibinfo {volume} {100}},\ \bibinfo
-  {pages} {043844} (\bibinfo {year} {2019})},\ \bibinfo {note} {number:
-  4}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Wolf}\ \emph {et~al.}(2015)\citenamefont {Wolf},
-  \citenamefont {Neumann}, \citenamefont {Nakamura}, \citenamefont {Sumiya},
-  \citenamefont {Ohshima}, \citenamefont {Isoya},\ and\ \citenamefont
-  {Wrachtrup}}]{Wolf}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
-  {Wolf}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Neumann}},
-  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Nakamura}}, \bibinfo
-  {author} {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author}
-  {\bibfnamefont {T.}~\bibnamefont {Ohshima}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}},\ and\ \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Wrachtrup}},\ }\href
-  {https://doi.org/10.1103/PhysRevX.5.041001} {\bibfield  {journal} {\bibinfo
-  {journal} {Phys. Rev. X}\ }\textbf {\bibinfo {volume} {5}},\ \bibinfo {pages}
-  {041001} (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Stürner}\ \emph {et~al.}(2021)\citenamefont
-  {Stürner}, \citenamefont {Brenneis}, \citenamefont {Buck}, \citenamefont
-  {Kassel}, \citenamefont {Rölver}, \citenamefont {Fuchs}, \citenamefont
-  {Savitsky}, \citenamefont {Suter}, \citenamefont {Grimmel}, \citenamefont
-  {Hengesbach}, \citenamefont {Förtsch}, \citenamefont {Nakamura},
-  \citenamefont {Sumiya}, \citenamefont {Onoda}, \citenamefont {Isoya},\ and\
-  \citenamefont {Jelezko}}]{Sturner}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.~M.}\ \bibnamefont
-  {Stürner}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Brenneis}},
-  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Buck}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Kassel}}, \bibinfo {author} {\bibfnamefont
-  {R.}~\bibnamefont {Rölver}}, \bibinfo {author} {\bibfnamefont
-  {T.}~\bibnamefont {Fuchs}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Savitsky}}, \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Suter}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Grimmel}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Hengesbach}}, \bibinfo {author} {\bibfnamefont
-  {M.}~\bibnamefont {Förtsch}}, \bibinfo {author} {\bibfnamefont
-  {K.}~\bibnamefont {Nakamura}}, \bibinfo {author} {\bibfnamefont
-  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}},\ and\ \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Jelezko}},\ }\href
-  {https://doi.org/https://doi.org/10.1002/qute.202000111} {\bibfield
-  {journal} {\bibinfo  {journal} {Advanced Quantum Technologies}\ }\textbf
-  {\bibinfo {volume} {4}},\ \bibinfo {pages} {2000111} (\bibinfo {year}
-  {2021})},\ \Eprint
-  {https://arxiv.org/abs/https://onlinelibrary.wiley.com/doi/pdf/10.1002/qute.202000111}
-  {https://onlinelibrary.wiley.com/doi/pdf/10.1002/qute.202000111} \BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Jarmola}\ \emph {et~al.}(2015)\citenamefont
-  {Jarmola}, \citenamefont {Berzins}, \citenamefont {Smits}, \citenamefont
-  {Smits}, \citenamefont {Prikulis}, \citenamefont {Gahbauer}, \citenamefont
-  {Ferber}, \citenamefont {Erts}, \citenamefont {Auzinsh},\ and\ \citenamefont
-  {Budker}}]{jarmola_longitudinal_2015}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Jarmola}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Berzins}},
-  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Smits}}, \bibinfo
-  {author} {\bibfnamefont {K.}~\bibnamefont {Smits}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Prikulis}}, \bibinfo {author}
-  {\bibfnamefont {F.}~\bibnamefont {Gahbauer}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Ferber}}, \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Erts}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Auzinsh}},\ and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Budker}},\ }\href {https://doi.org/10.1063/1.4937489} {\bibfield  {journal}
-  {\bibinfo  {journal} {Appl. Phys. Lett.}\ }\textbf {\bibinfo {volume}
-  {107}},\ \bibinfo {pages} {242403} (\bibinfo {year} {2015})},\ \bibinfo
-  {note} {number: 24}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Mrózek}\ \emph {et~al.}(2015)\citenamefont
-  {Mrózek}, \citenamefont {Rudnicki}, \citenamefont {Kehayias}, \citenamefont
-  {Jarmola}, \citenamefont {Budker},\ and\ \citenamefont
-  {Gawlik}}]{mrozek_longitudinal_2015}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Mrózek}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Rudnicki}},
-  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Kehayias}}, \bibinfo
-  {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}}, \bibinfo {author}
-  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ and\ \bibinfo {author}
-  {\bibfnamefont {W.}~\bibnamefont {Gawlik}},\ }\href
-  {https://doi.org/10.1140/epjqt/s40507-015-0035-z} {\bibfield  {journal}
-  {\bibinfo  {journal} {EPJ Quantum Technol.}\ }\textbf {\bibinfo {volume}
-  {2}},\ \bibinfo {pages} {22} (\bibinfo {year} {2015})},\ \bibinfo {note}
-  {number: 1}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Epstein}\ \emph {et~al.}(2005)\citenamefont
-  {Epstein}, \citenamefont {Mendoza}, \citenamefont {Kato},\ and\ \citenamefont
-  {Awschalom}}]{epstein2005anisotropic}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Epstein}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Mendoza}},
-  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Kato}},\ and\ \bibinfo
-  {author} {\bibfnamefont {D.}~\bibnamefont {Awschalom}},\ }\href@noop {}
-  {\bibfield  {journal} {\bibinfo  {journal} {Nature physics}\ }\textbf
-  {\bibinfo {volume} {1}},\ \bibinfo {pages} {94} (\bibinfo {year}
-  {2005})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Lai}\ \emph {et~al.}(2009)\citenamefont {Lai},
-  \citenamefont {Zheng}, \citenamefont {Jelezko}, \citenamefont {Treussart},\
-  and\ \citenamefont {Roch}}]{lai2009influence}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~D.}\ \bibnamefont
-  {Lai}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Zheng}}, \bibinfo
-  {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}}, \bibinfo {author}
-  {\bibfnamefont {F.}~\bibnamefont {Treussart}},\ and\ \bibinfo {author}
-  {\bibfnamefont {J.-F.}\ \bibnamefont {Roch}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Applied Physics Letters}\ }\textbf {\bibinfo
-  {volume} {95}},\ \bibinfo {pages} {133101} (\bibinfo {year}
-  {2009})}\BibitemShut {NoStop}%
-\bibitem [{SI_()}]{SI_low_filed_CR}%
-  \BibitemOpen
-  \href@noop {} {\bibinfo {title} {See supplemental material at [...], which
-  includes refs. [20-22]}}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Anishchik}\ \emph {et~al.}(2015)\citenamefont
-  {Anishchik}, \citenamefont {Vins}, \citenamefont {Yelisseyev}, \citenamefont
-  {Lukzen}, \citenamefont {Lavrik},\ and\ \citenamefont
-  {Bagryansky}}]{anishchik2015low}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Anishchik}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Vins}},
-  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Yelisseyev}}, \bibinfo
-  {author} {\bibfnamefont {N.}~\bibnamefont {Lukzen}}, \bibinfo {author}
-  {\bibfnamefont {N.}~\bibnamefont {Lavrik}},\ and\ \bibinfo {author}
-  {\bibfnamefont {V.}~\bibnamefont {Bagryansky}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {New Journal of Physics}\ }\textbf {\bibinfo
-  {volume} {17}},\ \bibinfo {pages} {023040} (\bibinfo {year}
-  {2015})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2020)\citenamefont
-  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},
-  \citenamefont {Kilin},\ and\ \citenamefont {Jelezko}}]{filimonenko2020weak}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
-  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Nizovtsev}}, \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
-  {Kilin}},\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
-  {Jelezko}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Semiconductors}\ }\textbf {\bibinfo {volume} {54}},\ \bibinfo {pages} {1730}
-  (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Van~Oort}\ and\ \citenamefont
-  {Glasbeek}(1990)}]{van1990electric}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
-  {Van~Oort}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Glasbeek}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Chemical Physics Letters}\ }\textbf {\bibinfo {volume} {168}},\ \bibinfo
-  {pages} {529} (\bibinfo {year} {1990})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Jarmola}\ \emph {et~al.}(2012)\citenamefont
-  {Jarmola}, \citenamefont {Acosta}, \citenamefont {Jensen}, \citenamefont
-  {Chemerisov},\ and\ \citenamefont {Budker}}]{jarmola_temperature-_2012}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Jarmola}}, \bibinfo {author} {\bibfnamefont {V.~M.}\ \bibnamefont {Acosta}},
-  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Jensen}}, \bibinfo
-  {author} {\bibfnamefont {S.}~\bibnamefont {Chemerisov}},\ and\ \bibinfo
-  {author} {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }\href
-  {https://doi.org/10.1103/PhysRevLett.108.197601} {\bibfield  {journal}
-  {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {108}},\
-  \bibinfo {pages} {197601} (\bibinfo {year} {2012})},\ \bibinfo {note}
-  {number: 19}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Redman}\ \emph {et~al.}(1991)\citenamefont {Redman},
-  \citenamefont {Brown}, \citenamefont {Sands},\ and\ \citenamefont
-  {Rand}}]{redman1991spin}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Redman}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Brown}},
-  \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Sands}},\ and\ \bibinfo
-  {author} {\bibfnamefont {S.}~\bibnamefont {Rand}},\ }\href@noop {} {\bibfield
-   {journal} {\bibinfo  {journal} {Physical review letters}\ }\textbf {\bibinfo
-  {volume} {67}},\ \bibinfo {pages} {3420} (\bibinfo {year}
-  {1991})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Norambuena}\ \emph {et~al.}(2018)\citenamefont
-  {Norambuena}, \citenamefont {Mu{\~n}oz}, \citenamefont {Dinani},
-  \citenamefont {Jarmola}, \citenamefont {Maletinsky}, \citenamefont {Budker},\
-  and\ \citenamefont {Maze}}]{norambuena2018spin}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Norambuena}}, \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
-  {Mu{\~n}oz}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Dinani}},
-  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Jarmola}}, \bibinfo
-  {author} {\bibfnamefont {P.}~\bibnamefont {Maletinsky}}, \bibinfo {author}
-  {\bibfnamefont {D.}~\bibnamefont {Budker}},\ and\ \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Maze}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Physical Review B}\ }\textbf {\bibinfo
-  {volume} {97}},\ \bibinfo {pages} {094304} (\bibinfo {year}
-  {2018})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Choi}\ \emph
-  {et~al.}(2017{\natexlab{c}})\citenamefont {Choi}, \citenamefont {Choi},
-  \citenamefont {Kucsko}, \citenamefont {Maurer}, \citenamefont {Shields},
-  \citenamefont {Sumiya}, \citenamefont {Onoda}, \citenamefont {Isoya},
-  \citenamefont {Demler}, \citenamefont {Jelezko} \emph
-  {et~al.}}]{choi2017depolarization}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
-  {Choi}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}}, \bibinfo
-  {author} {\bibfnamefont {G.}~\bibnamefont {Kucsko}}, \bibinfo {author}
-  {\bibfnamefont {P.~C.}\ \bibnamefont {Maurer}}, \bibinfo {author}
-  {\bibfnamefont {B.~J.}\ \bibnamefont {Shields}}, \bibinfo {author}
-  {\bibfnamefont {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
-  {S.}~\bibnamefont {Onoda}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
-  {E.}~\bibnamefont {Demler}}, \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Jelezko}}, \emph {et~al.},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Physical review letters}\ }\textbf {\bibinfo
-  {volume} {118}},\ \bibinfo {pages} {093601} (\bibinfo {year}
-  {2017}{\natexlab{c}})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Pellet-Mary}\ \emph
-  {et~al.}(2021{\natexlab{a}})\citenamefont {Pellet-Mary}, \citenamefont
-  {Huillery}, \citenamefont {Perdriat},\ and\ \citenamefont
-  {H{\'e}tet}}]{pellet2021magnetic}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Pellet-Mary}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
-  {Huillery}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Perdriat}},\
-  and\ \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {H{\'e}tet}},\
-  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Physical Review
-  B}\ }\textbf {\bibinfo {volume} {104}},\ \bibinfo {pages} {L100411} (\bibinfo
-  {year} {2021}{\natexlab{a}})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Mr{\'o}zek}\ \emph {et~al.}(2021)\citenamefont
-  {Mr{\'o}zek}, \citenamefont {Wojciechowski},\ and\ \citenamefont
-  {Gawlik}}]{mrozek2021characterization}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Mr{\'o}zek}}, \bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
-  {Wojciechowski}},\ and\ \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
-  {Gawlik}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Diamond and Related Materials}\ }\textbf {\bibinfo {volume} {120}},\
-  \bibinfo {pages} {108689} (\bibinfo {year} {2021})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Giri}\ \emph {et~al.}(2018)\citenamefont {Giri},
-  \citenamefont {Gorrini}, \citenamefont {Dorigoni}, \citenamefont {Avalos},
-  \citenamefont {Cazzanelli}, \citenamefont {Tambalo},\ and\ \citenamefont
-  {Bifone}}]{giri_coupled_2018}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Giri}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Gorrini}},
-  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Dorigoni}}, \bibinfo
-  {author} {\bibfnamefont {C.~E.}\ \bibnamefont {Avalos}}, \bibinfo {author}
-  {\bibfnamefont {M.}~\bibnamefont {Cazzanelli}}, \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Tambalo}},\ and\ \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Bifone}},\ }\href
-  {https://doi.org/10.1103/PhysRevB.98.045401} {\bibfield  {journal} {\bibinfo
-  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {98}},\ \bibinfo
-  {pages} {045401} (\bibinfo {year} {2018})},\ \bibinfo {note} {number:
-  4}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Giri}\ \emph {et~al.}(2019)\citenamefont {Giri},
-  \citenamefont {Dorigoni}, \citenamefont {Tambalo}, \citenamefont {Gorrini},\
-  and\ \citenamefont {Bifone}}]{giri_selective_2019}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
-  {Giri}}, \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Dorigoni}},
-  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Tambalo}}, \bibinfo
-  {author} {\bibfnamefont {F.}~\bibnamefont {Gorrini}},\ and\ \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Bifone}},\ }\href
-  {https://doi.org/10.1103/PhysRevB.99.155426} {\bibfield  {journal} {\bibinfo
-  {journal} {Phys. Rev. B}\ }\textbf {\bibinfo {volume} {99}},\ \bibinfo
-  {pages} {155426} (\bibinfo {year} {2019})},\ \bibinfo {note} {number:
-  15}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Pellet-Mary}\ \emph
-  {et~al.}(2021{\natexlab{b}})\citenamefont {Pellet-Mary}, \citenamefont
-  {Huillery}, \citenamefont {Perdriat}, \citenamefont {Tallaire},\ and\
-  \citenamefont {H{\'e}tet}}]{pellet2021optical}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Pellet-Mary}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
-  {Huillery}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Perdriat}},
-  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Tallaire}},\ and\
-  \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {H{\'e}tet}},\ }\href@noop
-  {} {\bibfield  {journal} {\bibinfo  {journal} {Physical Review B}\ }\textbf
-  {\bibinfo {volume} {103}},\ \bibinfo {pages} {L100411} (\bibinfo {year}
-  {2021}{\natexlab{b}})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Mittiga}\ \emph {et~al.}(2018)\citenamefont
-  {Mittiga}, \citenamefont {Hsieh}, \citenamefont {Zu}, \citenamefont {Kobrin},
-  \citenamefont {Machado}, \citenamefont {Bhattacharyya}, \citenamefont {Rui},
-  \citenamefont {Jarmola}, \citenamefont {Choi}, \citenamefont {Budker} \emph
-  {et~al.}}]{mittiga2018imaging}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
-  {Mittiga}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Hsieh}},
-  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Zu}}, \bibinfo {author}
-  {\bibfnamefont {B.}~\bibnamefont {Kobrin}}, \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Machado}}, \bibinfo {author} {\bibfnamefont
-  {P.}~\bibnamefont {Bhattacharyya}}, \bibinfo {author} {\bibfnamefont
-  {N.}~\bibnamefont {Rui}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Jarmola}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Choi}},
-  \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Budker}}, \emph
-  {et~al.},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Physical review letters}\ }\textbf {\bibinfo {volume} {121}},\ \bibinfo
-  {pages} {246402} (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Dolde}\ \emph {et~al.}(2011)\citenamefont {Dolde},
-  \citenamefont {Fedder}, \citenamefont {Doherty}, \citenamefont {N{\"o}bauer},
-  \citenamefont {Rempp}, \citenamefont {Balasubramanian}, \citenamefont {Wolf},
-  \citenamefont {Reinhard}, \citenamefont {Hollenberg}, \citenamefont {Jelezko}
-  \emph {et~al.}}]{dolde2011electric}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
-  {Dolde}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Fedder}},
-  \bibinfo {author} {\bibfnamefont {M.~W.}\ \bibnamefont {Doherty}}, \bibinfo
-  {author} {\bibfnamefont {T.}~\bibnamefont {N{\"o}bauer}}, \bibinfo {author}
-  {\bibfnamefont {F.}~\bibnamefont {Rempp}}, \bibinfo {author} {\bibfnamefont
-  {G.}~\bibnamefont {Balasubramanian}}, \bibinfo {author} {\bibfnamefont
-  {T.}~\bibnamefont {Wolf}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
-  {Reinhard}}, \bibinfo {author} {\bibfnamefont {L.~C.}\ \bibnamefont
-  {Hollenberg}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Jelezko}},
-  \emph {et~al.},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Nature Physics}\ }\textbf {\bibinfo {volume} {7}},\ \bibinfo {pages} {459}
-  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Qiu}\ \emph {et~al.}(2022)\citenamefont {Qiu},
-  \citenamefont {Hamo}, \citenamefont {Vool}, \citenamefont {Zhou},\ and\
-  \citenamefont {Yacoby}}]{qiu2022nanoscale}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont
-  {Qiu}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Hamo}}, \bibinfo
-  {author} {\bibfnamefont {U.}~\bibnamefont {Vool}}, \bibinfo {author}
-  {\bibfnamefont {T.~X.}\ \bibnamefont {Zhou}},\ and\ \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Yacoby}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {arXiv preprint arXiv:2205.03952}\ } (\bibinfo
-  {year} {2022})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Wickenbrock}\ \emph {et~al.}(2016)\citenamefont
-  {Wickenbrock}, \citenamefont {Zheng}, \citenamefont {Bougas}, \citenamefont
-  {Leefer}, \citenamefont {Afach}, \citenamefont {Jarmola}, \citenamefont
-  {Acosta},\ and\ \citenamefont {Budker}}]{Wickenbrock}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Wickenbrock}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Zheng}},
-  \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Bougas}}, \bibinfo
-  {author} {\bibfnamefont {N.}~\bibnamefont {Leefer}}, \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Afach}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Jarmola}}, \bibinfo {author} {\bibfnamefont {V.~M.}\
-  \bibnamefont {Acosta}},\ and\ \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Budker}},\ }\href {https://doi.org/10.1063/1.4960171}
-  {\bibfield  {journal} {\bibinfo  {journal} {Applied Physics Letters}\
-  }\textbf {\bibinfo {volume} {109}},\ \bibinfo {pages} {053505} (\bibinfo
-  {year} {2016})},\ \Eprint
-  {https://arxiv.org/abs/https://doi.org/10.1063/1.4960171}
-  {https://doi.org/10.1063/1.4960171} \BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Zheng}\ \emph {et~al.}(2017)\citenamefont {Zheng},
-  \citenamefont {Chatzidrosos}, \citenamefont {Wickenbrock}, \citenamefont
-  {Bougas}, \citenamefont {Lazda}, \citenamefont {Berzins}, \citenamefont
-  {Gahbauer}, \citenamefont {Auzinsh}, \citenamefont {Ferber},\ and\
-  \citenamefont {Budker}}]{zheng2017level}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
-  {Zheng}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Chatzidrosos}},
-  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Wickenbrock}}, \bibinfo
-  {author} {\bibfnamefont {L.}~\bibnamefont {Bougas}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Lazda}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Berzins}}, \bibinfo {author} {\bibfnamefont {F.~H.}\
-  \bibnamefont {Gahbauer}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Auzinsh}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Ferber}},\
-  and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Budker}},\ }in\
-  \href@noop {} {\emph {\bibinfo {booktitle} {Slow Light, Fast Light, and
-  Opto-Atomic Precision Metrology X}}},\ Vol.\ \bibinfo {volume} {10119}\
-  (\bibinfo {organization} {SPIE},\ \bibinfo {year} {2017})\ pp.\ \bibinfo
-  {pages} {115--122}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Zheng}\ \emph {et~al.}(2020)\citenamefont {Zheng},
-  \citenamefont {Sun}, \citenamefont {Chatzidrosos}, \citenamefont {Zhang},
-  \citenamefont {Nakamura}, \citenamefont {Sumiya}, \citenamefont {Ohshima},
-  \citenamefont {Isoya}, \citenamefont {Wrachtrup}, \citenamefont
-  {Wickenbrock},\ and\ \citenamefont {Budker}}]{zheng_microwave-free_2020}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
-  {Zheng}}, \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Sun}}, \bibinfo
-  {author} {\bibfnamefont {G.}~\bibnamefont {Chatzidrosos}}, \bibinfo {author}
-  {\bibfnamefont {C.}~\bibnamefont {Zhang}}, \bibinfo {author} {\bibfnamefont
-  {K.}~\bibnamefont {Nakamura}}, \bibinfo {author} {\bibfnamefont
-  {H.}~\bibnamefont {Sumiya}}, \bibinfo {author} {\bibfnamefont
-  {T.}~\bibnamefont {Ohshima}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Isoya}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Wrachtrup}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Wickenbrock}},\ and\ \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Budker}},\ }\href
-  {https://doi.org/10.1103/PhysRevApplied.13.044023} {\bibfield  {journal}
-  {\bibinfo  {journal} {Phys. Rev. Applied}\ }\textbf {\bibinfo {volume}
-  {13}},\ \bibinfo {pages} {044023} (\bibinfo {year} {2020})},\ \bibinfo {note}
-  {number: 4}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2018)\citenamefont
-  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},\ and\
-  \citenamefont {Kilin}}]{filimonenko2018weak}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
-  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Nizovtsev}},\ and\ \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
-  {Kilin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Semiconductors}\ }\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {1865}
-  (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2022)\citenamefont
-  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},
-  \citenamefont {Kilin},\ and\ \citenamefont
-  {Jelezko}}]{filimonenko2022manifestation}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
-  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.~P.}\ \bibnamefont
-  {Nizovtsev}}, \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
-  {Kilin}},\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
-  {Jelezko}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Journal of Applied Spectroscopy}\ }\textbf {\bibinfo {volume} {88}},\
-  \bibinfo {pages} {1131} (\bibinfo {year} {2022})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Vetter}\ \emph {et~al.}(2022)\citenamefont {Vetter},
-  \citenamefont {Marshall}, \citenamefont {Genov}, \citenamefont {Weiss},
-  \citenamefont {Striegler}, \citenamefont {Gro\ss{}mann}, \citenamefont
-  {Oviedo-Casado}, \citenamefont {Cerrillo}, \citenamefont {Prior},
-  \citenamefont {Neumann},\ and\ \citenamefont {Jelezko}}]{Vetter_LFM}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~J.}\ \bibnamefont
-  {Vetter}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Marshall}},
-  \bibinfo {author} {\bibfnamefont {G.~T.}\ \bibnamefont {Genov}}, \bibinfo
-  {author} {\bibfnamefont {T.~F.}\ \bibnamefont {Weiss}}, \bibinfo {author}
-  {\bibfnamefont {N.}~\bibnamefont {Striegler}}, \bibinfo {author}
-  {\bibfnamefont {E.~F.}\ \bibnamefont {Gro\ss{}mann}}, \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Oviedo-Casado}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Cerrillo}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Prior}}, \bibinfo {author} {\bibfnamefont
-  {P.}~\bibnamefont {Neumann}},\ and\ \bibinfo {author} {\bibfnamefont
-  {F.}~\bibnamefont {Jelezko}},\ }\href
-  {https://doi.org/10.1103/PhysRevApplied.17.044028} {\bibfield  {journal}
-  {\bibinfo  {journal} {Phys. Rev. Applied}\ }\textbf {\bibinfo {volume}
-  {17}},\ \bibinfo {pages} {044028} (\bibinfo {year} {2022})}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Wang}\ \emph {et~al.}(2022)\citenamefont {Wang},
-  \citenamefont {Liu}, \citenamefont {Fan}, \citenamefont {Feng}, \citenamefont
-  {Leong}, \citenamefont {Finkler}, \citenamefont {Denisenko}, \citenamefont
-  {Wrachtrup}, \citenamefont {Li},\ and\ \citenamefont {Liu}}]{WangRB}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
-  {Wang}}, \bibinfo {author} {\bibfnamefont {C.-F.}\ \bibnamefont {Liu}},
-  \bibinfo {author} {\bibfnamefont {J.-W.}\ \bibnamefont {Fan}}, \bibinfo
-  {author} {\bibfnamefont {X.}~\bibnamefont {Feng}}, \bibinfo {author}
-  {\bibfnamefont {W.-H.}\ \bibnamefont {Leong}}, \bibinfo {author}
-  {\bibfnamefont {A.}~\bibnamefont {Finkler}}, \bibinfo {author} {\bibfnamefont
-  {A.}~\bibnamefont {Denisenko}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Wrachtrup}}, \bibinfo {author} {\bibfnamefont
-  {Q.}~\bibnamefont {Li}},\ and\ \bibinfo {author} {\bibfnamefont {R.-B.}\
-  \bibnamefont {Liu}},\ }\href
-  {https://doi.org/10.1103/PhysRevResearch.4.013098} {\bibfield  {journal}
-  {\bibinfo  {journal} {Phys. Rev. Research}\ }\textbf {\bibinfo {volume}
-  {4}},\ \bibinfo {pages} {013098} (\bibinfo {year} {2022})}\BibitemShut
-  {NoStop}%
-\end{thebibliography}%
+\bibliographystyle{apsrev4-2}
+\bibliography{CR}{}
 

### 2022-07-11 Mes papiers/Low field CR/V3/SI/Low_field_CR_SI.tex
tata yoyo

@@
 
-In the absence of an external magnetic field, the symmetry of the hamitlonian in the ($xy$) plane allows us to pick the $x$ direction along the electric field.
+In the absence of an external magnetic field, the symmetry of the Hamiltonian in the ($xy$) plane allows us to pick the $x$ direction along the electric field.
 The eigenstates of $\mathcal{H}_s$ then become $\{ \ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $.
@@
 The slight asymmetry between the positive and negative value of the magnetic field could come from the earth magnetic field or from hysteresis in the magnetic core of our electromagnet.
+
+\subsection{Emulation of the electric field with transverse magnetic field}
+\begin{figure}
+\includegraphics[width=0.9\textwidth]{Figures_SI/fig_transverse_field}
+\caption{(a) Simulated energy level of the three eigenstates $\ket{g},\ket{d}$ and $\ket{e}$ for an NV center under purely transverse magnetic field.
+The spin Hamiltonian is given in eq.
+\ref{NV Hamiltonian} and a value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen (b) Blue dashed curve : frequency splitting between the $\ket{d}$ and $\ket{e}$ states as a function of the transverse magnetic field, red plain line : matching factor $\abs{\bra{e}\ket{+}}^2$ between the $\ket{e}$ and $\ket{+}$ states}
+\label{theory_transverse_field}
+\end{figure}
+In the main text we use the property that, under purely transverse magnetic field, the eigenstates of the single spin Hamiltonian are close to the $\{ \ket{0},\ket{+},\ket{-} \}$ basis, described in the main text and in sec.
+\ref{sec Hamiltonian} of these supplementary.
+We will discuss here the validity of this claim.
+
+We will call the true eigenstates of the single spin Hamiltonian under purely transverse magnetic field $\{ \ket{g},\ket{d},\ket{e} \}$ in ascending order of energy.
+The $\ket{d}$ state is strictly equal to the $\ket{-}$ state.
+The $\ket{g}$ and $\ket{e}$ states can be written as a mixture of the $\ket{0}$ and $\ket{+}$ states, and their exact expression depends on the field value.
+As $B_{\perp}$ approaches 0 however, $\ket{g}$ and $\ket{e}$ tend toward $\ket{0}$ and $\ket{+}$ respectively.
+
+Fig.
+\ref{theory_transverse_field} (a) shows a simulation of the energy levels of the three eigenstates $\ket{g}$, $\ket{d}$ and $\ket{e}$ as a function of the transverse magnetic field amplitude.
+Of importance for the results in main text is the splitting $\Delta \nu$ between the $\ket{e}$ and $\ket{d}$ states, which is equal to the difference in the transition frequencies $\ket{g} \to \ket{d}$ and $\ket{g} \to \ket{e}$ (reported experimentally as $\nu_+$ and $\nu_-$ in main text).
+
+This value $\Delta \nu$ is reported on Fig.
+\ref{theory_transverse_field} (b) along with a matching factor $\abs{\bra{e}\ket{+}}^2$.
+This factor is always equal to $\abs{\bra{g}\ket{0}}^2$ since $\bra{d}\ket{-}=1$.
+We can see that indeed, for $B_{\perp}<150\ \rm G$, $\abs{\bra{e}\ket{+}}^2>0.98$.
+Crucially, for $B_{\perp}=150\ \rm G$, the splitting $\Delta \nu$ is equal to $\approx 70\ \rm MHz$, almost an order of magnitude greater than the dipole-dipole interaction range computed on sec.
+\ref{fluctuator width}.
+This means that we can effectively quench the double-flips by splitting the $\ket{e}$ and $\ket{d}$ states, while still maintaining eigenstates almost equal to $\ket{0}$, $\ket{+}$ and $\ket{-}$.
+
 \section{Extension of the fluctuator model to low magnetic fields}

### 2022-07-12 Manuscrit Thèse/chapter 2/Appendix dipole-dipole.tex
la tete de tes morts

@@
-\documentclass[a4paper]{book}
+\documentclass[a4paper]{report}
 \usepackage[]{amsmath}
@@
 \begin{document}
+\chapter{Detection of dark spins via cross-relaxations}
+In this chapter, we will see how...
 \section{The dipole-dipole Hamiltonian}
@@
 
-\section{Flip-Flops, double flips and cross-relaxations}
+\section{Flip-Flops, double flips and cross-relaxation}
+
+\begin{figure}
+\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
+\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
+\label{CR_shema}
+\end{figure}
 
@@
 
-%réécris moi ça, c'est naze
-When two spin populations of different polarization can be resonantly coupled by either flip-flop or double-flip, cross-relaxation happens, that is a macroscopic transfer of polarization from the most polarized spin group to the least polarized one.
+Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
+This process con occur either through flip-flops, or through double flips, as long as the resonance condition between the two spins is met.
+
+Fig.
+\ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin.
+When the two considered spin transitions become resonant, in this case by tuning a magnetic field, polarization from the NV center will be transferred to the second spin, meaning in this case that the NV center will end up less polarized and the second spin more polarized.
+
+\section{Detection of dark spins with NV centers via cross-relaxations}
 

### 2022-07-12 Mes papiers/Low field CR/V3/Low_field_CR.tex
la tete de tes morts

--- 2022-07-08 Mes papiers/Low field CR/V3/Low_field_CR.tex
@@
 We study dipolar interaction processes in dense ensembles of negatively charged nitrogen vacancy (NV$^-$) centers under small magnetic fields.
-Employing magnetic field scans close to zero-field along specific crystalline direction while recording the NV$^-$ spin relaxation rate, we could identify regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the NV-NV cross-relaxation .
+Employing magnetic field scans along specific crystalline direction while recording the NV$^-$ spin relaxation rate, we could identify regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the NV-NV cross-relaxation.
 Our results are relevant for understanding decoherence in many-body spin systems as well as for high sensitivity magneto- and electro-metry with long-lived interacting solid-state spins.
@@
 
-The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond has given rise to a wealth of applications in nanoscale sensing and quantum information science thanks in part to the possibility to optically polarize and read-out its spin state at ambient conditions \cite{DOHERTY20131}.
-In particular, ensembles of NV centers are widely studied for their enhanced magnetic field sensing capabilities \cite{Acosta, TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized} and as pristine platforms for studying many body effects \citep{kucsko2018critical, ChoiNat, ZuYao}.
+The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond have given rise to a wealth of applications in nanoscale sensing and quantum information science thanks in part to the possibility to optically polarize and read-out its spin state at ambient conditions \cite{DOHERTY20131}.
+In particular, ensembles of NV centers are widely studied for their enhanced magnetic field sensing capabilities \cite{Acosta, TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized} and as pristine platforms for observing many body effects \citep{kucsko2018critical, ChoiNat, ZuYao, dwyer2021probing}.
 When the NV spin concentration reaches ppm values, spin depolarisation, or cross-relaxation (CR) takes place through a very rich many body dynamics associated with disorder \citep{choi_observation_2017}.
@@
 
-CR features at zero-field have also been observed and could also be deployed for high sensitivity microwave free magnetometry.
+CR features at zero-field have also been observed and could be deployed for high sensitivity microwave free magnetometry  \cite{filimonenko2018weak, filimonenko2022manifestation}.
 The CR contrast was indeed shown to be much stronger \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}.
@@
 Here, we study dipolar interaction processes in ensembles of NV centers under small magnetic fields.
-Specifically, by employing magnetic field scans along specific crystalline directions, we identify regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the CR.
+Specifically, by employing magnetic field scans along specific crystalline directions, we find regimes where flip-flop, double-flip processes as well as mixing induced by local electric field play a role in the CR.
 We also present an orientation and microwave-free magnetometer operating with a sensitivity below $100$ nT/$\sqrt{\rm Hz}$ at zero-field, where double-flip processes amongst NV centers dominate.
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/shema_summary.pdf}
-\caption{Schematics showing a diamond with interacting spins as well as the three processes that can account for dipolar relaxation :flip-flop processes involving two different classes of NV centers, local electric field mixing and double-flip processes where two units of spin angular momentum are exchanged.
+\includegraphics[width=0.45\textwidth]{shema_summary.pdf}
+\caption{Schematics showing a diamond with interacting spins (central picture) as well as the three processes that can account for dipolar relaxation: flip-flop processes involving two different classes of NV centers, local electric field mixing and double-flip processes where two units of spin angular momentum are exchanged.
 }
@@
 Flip-flop processes involving coupling between NV centers from identical or from different orientations are depicted in the top-right panel.
-They have already been shown to play a dominant role when tuning $\delta$ {\it via} longitudinal B fields scans, when $B \gtrsim 30$G \cite{choi_observation_2017}.
+They have already been shown to play a dominant role by tuning the frequency between the different NV centers, when $B \gtrsim 30$G \cite{choi_observation_2017}.
 We will show here that under smaller magnetic fields or using purely transverse fields, double-flip up and down as well as the mixing induced by local electric field give a significant contribution to the CR (two bottom panels).
@@
 Fig.
-\ref{T1}-a) and b) show the change in PL as a function of an external magnetic field for two samples with a low and high concentration of NV$^-$ centers respectively, detailed in the supplementary material (SM sec.
-II) of this article \citep{SI_low_filed_CR} \nocite{anishchik2015low, filimonenko2020weak, van1990electric}.
-Both samples show a decrease in PL as the magnetic field amplitude increases.
+\ref{T1}-a) and b) show the change in PL as a function of an external magnetic field for two samples with a low and high concentration of NV$^-$ centers respectively.
+All samples used in this study are listed in the supplementary material (SM) sec.
+II \citep{SI_low_filed_CR} \nocite{anishchik2015low, filimonenko2020weak, van1990electric}.
+The nomenclature used here indicates first the fabrication process: high pressure high temperature (HPHT) or chemical vapor deposition (CVD), followed in the case of HPHT samples by the size of the diamond particle in $\mu$m.
+In both samples, we see a decrease in PL as the magnetic field amplitude increases.
 There is however a stark difference in the low magnetic field region where only high-density samples shows a drop in PL \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}.
@@
 This effect has been attributed to CR between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi2017depolarization}.
-The reason is that at zero-field many NV ``classes" - that is physical orientation of the NV axis is the diamond crystal cell - becomes resonant, which increases the dipolar relaxation.
-Inhomogeneity of the spin lifetimes is further needed in order to explain the depolarization of the spin ensemble \citep{choi2017depolarization}.
+The reason is that at zero-field all the NV ``classes" - the physical orientation of the NV axes - become resonant, which increases the dipolar relaxation.
+Inhomogeneity of the spin lifetimes is further needed to explain the depolarization of the spin ensemble \citep{choi2017depolarization}.
 We will denote $T_1^{\rm ph}$ the characteristic timescale associated with the phonon relaxation process and $T_1^{\rm dd}$ the density dependent timescale associated with the dipole-dipole relaxation process.
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig_T1.pdf}
-\caption{a) PL measurement from NV centers ensemble as a function of an external magnetic field applied in an arbitrary direction (a) for the sample CVD-PPB, with [NV$^-$]$\approx 50\ \rm ppb$ (b) for the sample HPHT-150-1, with [NV$^-$]$\approx 3\ \rm ppm$.
+\includegraphics[width=0.45\textwidth]{fig_T1.pdf}
+\caption{a) PL measurement from NV centers ensembles as a function of an external magnetic field applied in an arbitrary direction (a) for the sample CVD-1, with [NV$^-$]$\approx 50\ \rm ppb$ (b) for the sample HPHT-150-1, with [NV$^-$]$\approx 3\ \rm ppm$.
 (c) Sequence used to measure the spin lifetime.
 (d) Spin relaxation $S_1-S_2$ measured for the dense sample at zero and non-zero magnetic fields.
-The fitting procedure (orange plain line) is detailed in the main text}
+The fitting procedure (orange plain line) is detailed in the main text.}
 \label{T1}
@@
 \begin{figure*}
-\includegraphics[width=.95\textwidth]{Figures/fig_100_vs_1x1x1x1.pdf}
-\caption{(a) ODMR spectrum in zero field.
+\includegraphics[width=.95\textwidth]{fig_100_vs_1x1x1x1.pdf}
+\caption{(a) ODMR spectrum in zero field with microwave amplitude modulation (see experimental details in SM sec.
+III).
 (b)-i) ODMR spectrum for a magnetic field $\approx$ 60 G, misaligned by $\sim  24^\circ$ from the [100] axis.
@@
 It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$.
-In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019}.
-It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right on one of the eight NV spin-resonances before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state.
-By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers \citep{jarmola_temperature-_2012, mrozek_longitudinal_2015, choi2017depolarization}.
+In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019, choi2017depolarization}.
+It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right on one of the eight NV spin-resonances before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state \citep{jarmola_temperature-_2012,mrozek_longitudinal_2015,choi2017depolarization}.
+By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.
 
@@
 We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi2017depolarization}.
-This model postulates the existence of very short-lived NV centers - the so-called fluctuators - which, when coupled resonantly to the other NV centers, act as a polarization drain for the spin ensemble.
+This model is based on the existence of very short-lived NV centers - the so-called fluctuators - which, when coupled resonantly to the other NV centers, act as a polarization drain for the spin ensemble.
 A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced relaxation should be a stretched-exponential with a stretch factor $\beta=1/2$.
@@
 We conducted further measurements to confirm the stretch nature of the dipole-induced decay (see SM sec.
-IV A), as well as confirming the lifetime-limited model of the fluctuators by measuring the resonance condition between classes of NV centers (see SM sec.
+IV A), as well as confirming the lifetime-limited model of the fluctuators(see SM sec.
 IV B).
@@
 These results thus demonstrates a twenty-fold increase in the dipolar depolarization rate when the B field is turned to zero.
-While the fluctuator model developed in \citep{choi2017depolarization} contains some of the explanation for this increase in dipolar depolarization, due to more classes being resonant in zero field than in non-zero field, the model did not include the specificity of the zero magnetic field region, such as the importance of local electric field and the double flip processes.
-We therefore decided to conducted more in depth experiments in order to characterize these zero field specificities.
+The fluctuator model developed in \citep{choi2017depolarization} contains some of the explanation for this increase in dipolar depolarization: more classes are resonant in zero field, which increases the flip-flop rates.
+However the model did not include the specificity of the zero magnetic field region, such as the importance of local electric field and the double flip processes.
+We therefore decided to conducted more in depth experiments in order to characterize these extra relaxation processes.
 
@@
 \ref{100_VS_1x4} (b)-i) shows an ODMR spectrum from the same sample, but in a magnetic field.
-The observed lines correspond to the projections of the magnetic fields onto the four classes of NV centers together with the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$.
-The transitions frequencies of the four different classes can be controlled by changing the amplitude and orientation of the external magnetic field.
-Fig.
-\ref{100_VS_1x4} (c)-i) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis.
-This configuration will be essential to decipher which dipolar relaxation mechanisms play a role at low field.
+The observed lines correspond to the projections of the magnetic fields onto the four classes of NV centers together with the $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$ spin transitions.
+The transitions frequencies of the four different NV classes can be controlled by changing the amplitude and orientation of the external magnetic field, which allows us to tune the number of degenerate classes.
 
@@
 
-Importantly, B field scans along the [100] direction, where the four classes are always resonant, still features an increase of the spin lifetime as the B field increases.
+Fig.
+\ref{100_VS_1x4} (c)-i) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis.
+Importantly, B field scans along the [100] direction still feature an increase of the spin lifetime as the B field increases.
 Fig.
 \ref{100_VS_1x4} (c)-ii) and (c)-iii) show the change in the PL and in $1/T_1^{\rm dd}$ as a function of a magnetic field that is aligned in this direction.
-We can see that there still is an increase in the spin decay rate and a corresponding drop in the PL under low magnetic field values, although considerably reduced.
-
-The main aim of this paper is to understand the physics governing the contribution to the spin decay in this second scenario.
+We can see that there still is an increase in the spin decay rate and a corresponding drop in the PL under low magnetic field values.
+Although considerably reduced, the PL and $1/T_1^{\rm dd}$ change is much sharper in this scenario.
+
+The main aim of this paper is to understand the physics governing the contribution to the spin decay in this second scenario, and its impact on zero-field magnetometry.
 Note that the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20$G is related to dipolar interaction with NV centers that have a $^{13}$C as a first neighbor \cite{pellet2021optical}.
 
-We identify and isolate two additional mechanisms that could explain this effect :local electric fields and double-flip processes. 
- In situations such as the one described in Fig.~\ref{100_VS_1x4} (c)-i), the only relevant terms for depolarisation in $\mathcal{H}^{\rm dd}$ are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$ where $\mathcal{H}^{\rm dd}$ is the dipole-dipole Hamiltonian detailed in SM sec.
+We identify and isolate two additional mechanisms that could explain this effect: local electric fields and double-flip processes. 
+ When $\mathbf{B} \neq 0$, the only relevant terms for depolarization in $\mathcal{H}^{\rm dd}$ are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$ where $\mathcal{H}^{\rm dd}$ is the dipole-dipole Hamiltonian detailed in SM sec.
 V.
-In zero magnetic field however, we have to take into account two other mechanisms of the dipolar Hamiltonian in consideration.
+In zero magnetic field however, we have to take two other mechanisms of the dipolar Hamiltonian into consideration.
 
@@
 \label{NV Hamiltonian}
-\mathcal{H}_{\rm elec}= d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
+\mathcal{H}_{\rm elec}= d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right],
 \end{equation}
@@
 
-In order to discriminate the role of these two effects in the low-field region, we make use of a purely transverse magnetic field.
-The transverse B field emulates the effect of the electric field, while turning off the resonance condition for the effectiveness of double-flip processes.
-Indeed for relatively weak transverse magnetic field, the eigen-basis of the spin Hamiltonian is close to $\ket{0},\ket{+},\ket{-}$.
-This property has already been used to observe a linear dependence in the electric field for the spin transition frequencies by applying an external transverse magnetic field \cite{dolde2011electric,qiu2022nanoscale}.
-Fig.~\ref{B_transverse} (a) shows the energy levels for the three eigenstates of the NV spin Hamiltonian as a function of a purely transverse magnetic field.
-We will denote these eigenstates in the general case $\ket{g}$, $\ket{d}$ and $\ket{e}$.
-For weak transverse magnetic field, $\ket{g} \approx \ket{0}$, $\ket{d}=\ket{-}$ and $\ket{e} \approx \ket{+}$.
-As the magnetic field increases, $\ket{g}$ and $\ket{e}$ starts to become a mixing of $\ket{0}$ and $\ket{+}$ while $\ket{d}$ remains equal to $\ket{-}$.
-
-Crucially, mixing remains even when the associated energy splitting of the $\ket{d}$ and $\ket{e}$ states is large :Fig.
-\ref{B_transverse}-(b) shows both the splitting between the energy levels of $\ket{d}$ and $\ket{e}$, and the matching factor $\abs{\bra{e}\ket{+}}^2$ characterizing the closeness of the states $\bra{e}$ and $\bra{+}$.
-We can see that at a field value of $\sim 130$ G, the splitting between the $\ket{d}$ and $\ket{e}$ reaches $\sim 50$ MHz, exceeding the range of the dipole-dipole resonant coupling (see SM sec.
-IV B).
-Yet, $\abs{\bra{e}\ket{+}}^2 > 0.98$ meaning that the spin Hamiltonian eigenstates are still roughly in the $\ket{0}, \ket{+}, \ket{-}$ basis.
-We can therefore decouple the effects of the double-flip processes  $\propto \ket{g,d}\bra{d,e}$, which are no longer resonant when $| \bm B | \geq 100$ G, from the effects of the mixing induced by the transverse electric - or in this case magnetic - field.
+In order to experimentally discriminate the contribution from the electric field and the double flips, we would need to apply an electric field strong enough to lift the pseudo-degeneracy between the $\ket{+}$ and $\ket{-}$ states.
+We estimated that we would need a field of $\sim 10^6\ \rm{V/cm}$ far out of range of our experimental reach.
+Instead we decided to emulate the effects of the electric field by applying a purely transverse magnetic field on one of the NV class.
+Indeed, for $B_{\perp} \leq 150\ \rm G$, the eigenstates of the spin Hamiltonian are also $\ket{0}$, $\ket{+}$ and $\ket{-}$ within a precision < 2\%.
+This property has previously been used to increase the electric field sensing ability of NV centers \cite{dolde2011electric,qiu2022nanoscale}.
+We further discuss the validity of the method in SM sec.
+[REF]
+
+Fig.~\ref{B_transverse} (a) shows an ODMR spectrum for $B_{\perp}=20\ \rm G$ on sample HPHT-150-2.
+The central two lines correspond to the $\ket{0}\to \ket{-}$ and $\ket{0}\to \ket{+}$ transitions for the class orthogonal to the magnetic field, the other features corresponding to the transitions of the other classes.
+We chose the initial value of 20 G so that the class of interest is far enough detuned from the other classes that we don't need to consider inter-class flip-flops.
 
 Fig.
-\ref{B_transverse}-(c) shows the measurement of the stretched lifetime $T_1^{\rm dd}$ for a single class of NV centers on sample HPHT-150-2 exposed to a transverse magnetic field between 20 and 130 G.
-The detuning $\Delta \nu$ between the two states $\ket{d}$ and $\ket{e}$ was measured through ODMR.
-Beforehand, we measured a stretched decay $1/T_1^{\rm dd}=25\pm 5\  \rm Hz$ under strong longitudinal magnetic field.
-This value is indicated by the dashed green line.
-Two regions can be identified in Fig.
-\ref{B_transverse}-(c).
-In  region A, the spin decay rate decreases from 175 Hz to 60 Hz.
-We attribute this effect to the loss of effectiveness of the double-flip processes.
-In region B, $\Delta \nu \geq 30\ \rm MHz$, the decay rate plateaus to $1/T_1^{\rm dd}=57\pm 5\  \rm Hz$, which is about twice as large as the decay rate measured under a longitudinal magnetic field.
-This plateau can be attributed to the change of eigenstate basis in the presence of transverse magnetic field, which is the same as under low magnetic field, where local transverse electric fields dominates the spin Hamiltonian (see SM sec.
-I).
-From these measurements, we conclude that the effect of the double-flip processes is about 5 times more important than the effect of the electric field in the depolarization rate.
-
-Based on an extension of the fluctuator model developed in SM sec.
-V, we predict an overall increase by a factor of $\sim 4$ of the dipole-dipole induced decay rate in the non-magnetic basis $\ket{0}, \ket{+}, \ket{-}$ compared to the magnetic basis $\ket{0}, \ket{+1}, \ket{-1}$.
-While the measured increase is smaller than the predicted one, the measurements agrees with the theoretical increase in the flip-flop rate in the $\ket{+/-}$ basis compared to the $\ket{\pm 1}$ basis.
-Similarly, we compute the predicted decay rate when $\bm B \parallel \left[100\right]$ case compared when $\bm B = 0$, and find a theoretical increase of $\sim 20\%$ when $\bm B = 0$ due to the mixing by the local electric field.
-This, along with the double-flip processes explains the behavior observed in Fig.
-\ref{100_VS_1x4} (c)-ii) and (c)-iii).
+\ref{B_transverse}-(c) shows the measurement of the decay rate $1/T_1^{\rm dd}$ for the class orthogonal to the magnetic field, as a function of the transverse field amplitude.
+We can observe two regions on this graph: region A where the decay rate decreases with the magnetic field, and region B where it stabilizes to a value $1/T_1^{\rm dd}=60\pm 5\ \rm{s}^{-1}$.
+We also indicate on this graph the value $1/T_1^{\rm dd}=25\pm 5\ \rm{s}^{-1}$ found for the same class but with a magnetic field oriented parallel to the class instead of transverse to it.
+
+What happens in the A region is that, as we increase the transverse field value, we are also increasing the detuning $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states, as shown in Fig.
+\ref{B_transverse} (b).
+By increasing this detuning, we are progressively quenching the double-flips until we reach the B region where only the transverse field effect remains.
+We can therefore directly compare the value of $60\ \rm{s}^{-1}$, corresponding the intra-class flip-flops in the $\ket{\pm}$ basis, to the $25\ \rm{s}^{-1}$ value which corresponds the intra-class flip-flops in the $\ket{\pm 1}$ basis.
+This increase is corroborated by our model detailed in SM sec.
+V where we predicted an increase in the decay rate by a factor of 4 in the $\ket{\pm}$ basis compared to the $\ket{\pm 1}$ basis.
+We can also note that the double-flips under low magnetic field are responsible for an increase in the decay rate $\sim 5$ times greater than that of the electric field, and are therefore likely the main cause of zero field depolarization in Fig.
+\ref{100_VS_1x4} (c)-iii).
 
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{Figures/fig_magneto}
+\includegraphics[width=0.45\textwidth]{fig_magneto}
 \caption{Low-field magnetometry protocol.
@@
 DC microwave-free magnetometry has already been performed using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}.
-Here we propose to perform a similar protocol, but using the spin depolarization at zero-field as proposed in \cite{filimonenko2018weak, filimonenko2022manifestation}.
+Here we propose to perform a similar protocol, but using the spin depolarization at zero-field.
 The main advantage compared to previously employed protocols is that the sensitivity does not depend crucially on crystalline orientation, making this protocol applicable with diamond powders or polycrystalline samples.
-Another sizable advantage is that this measurement is that it is insensitive to thermal variations.
+Another significant advantage is that this measurement is insensitive to thermal variations.
 In order to optimize signal-to-noise ratio, we use a lock-in detection and add a magnetic field modulation at $\sim 1\ \rm kHz$ with an amplitude $\sim 10\ \rm G$ through the same electromagnet.
@@
 We can see a sharp linear slope in low field $\abs{B} < 5\ \rm G$.
-Once calibrated, in this case with ODMR, the slope provides a 1D magnetic field measurement, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as in \cite{zheng_microwave-free_2020}.
+Once calibrated, in this case with ODMR, the slope provides a 1D magnetometer, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as in \cite{zheng_microwave-free_2020}.
 In order to assess the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take a histogram of the measured fields (46.000 counts in 46 s), as shown in Fig.
@@
 The lift in degeneracy of the spin state of different NV classes was found to be the main cause of zero-field depolarization, followed by double-flip processes and then the electric field induced mixing.
-We have employed CR for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nm}/\sqrt{\rm Hz}$ for a single $15\ \mu \rm m$ commercially available diamond and show that that double-flips and electric field play an important role.
-Our studies will be important for microwave based low field magnetometry \cite{Vetter_LFM, WangRB}. 
-Besides magnetometry, our work offers prospects for understanding many body phenomena with strongly coupled spins under purely transverse as well as small magnetic fields.
-
-\section*{acknowledgements}
+We have employed CR for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nT}/\sqrt{\rm Hz}$ for a single $15\ \mu \rm m$ commercially available diamond and show that that double-flips and local electric field play an important role.
+Our studies will be important for microwave based low field magnetometry \cite{Vetter_LFM, WangRB} and for understanding many body phenomena with strongly coupled spins under purely transverse or small magnetic fields.
+
+\begin{acknowledgments}
 
@@
 
+\end{acknowledgments}
 %\bibliographystyle{unsrtnat}
@@
 %Control: production of eprint (0) enabled
-\begin{thebibliography}{41}%
+\begin{thebibliography}{42}%
 \makeatletter
@@
   {45} (\bibinfo {year} {2021})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Dwyer}\ \emph {et~al.}(2021)\citenamefont {Dwyer},
+  \citenamefont {Rodgers}, \citenamefont {Urbach}, \citenamefont {Bluvstein},
+  \citenamefont {Sangtawesin}, \citenamefont {Zhou}, \citenamefont {Nassab},
+  \citenamefont {Fitzpatrick}, \citenamefont {Yuan}, \citenamefont {De~Greve}
+  \emph {et~al.}}]{dwyer2021probing}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.~L.}\ \bibnamefont
+  {Dwyer}}, \bibinfo {author} {\bibfnamefont {L.~V.}\ \bibnamefont {Rodgers}},
+  \bibinfo {author} {\bibfnamefont {E.~K.}\ \bibnamefont {Urbach}}, \bibinfo
+  {author} {\bibfnamefont {D.}~\bibnamefont {Bluvstein}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Sangtawesin}}, \bibinfo {author}
+  {\bibfnamefont {H.}~\bibnamefont {Zhou}}, \bibinfo {author} {\bibfnamefont
+  {Y.}~\bibnamefont {Nassab}}, \bibinfo {author} {\bibfnamefont
+  {M.}~\bibnamefont {Fitzpatrick}}, \bibinfo {author} {\bibfnamefont
+  {Z.}~\bibnamefont {Yuan}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {De~Greve}}, \emph {et~al.},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
+  {journal} {arXiv preprint arXiv:2103.12757}\ } (\bibinfo {year}
+  {2021})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Choi}\ \emph
@@
   {NoStop}%
+\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2018)\citenamefont
+  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},\ and\
+  \citenamefont {Kilin}}]{filimonenko2018weak}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
+  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
+  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Nizovtsev}},\ and\ \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
+  {Kilin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Semiconductors}\ }\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {1865}
+  (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2022)\citenamefont
+  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},
+  \citenamefont {Kilin},\ and\ \citenamefont
+  {Jelezko}}]{filimonenko2022manifestation}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
+  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
+  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.~P.}\ \bibnamefont
+  {Nizovtsev}}, \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
+  {Kilin}},\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Jelezko}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Journal of Applied Spectroscopy}\ }\textbf {\bibinfo {volume} {88}},\
+  \bibinfo {pages} {1131} (\bibinfo {year} {2022})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Jarmola}\ \emph {et~al.}(2015)\citenamefont
@@
   {number: 4}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2018)\citenamefont
-  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},\ and\
-  \citenamefont {Kilin}}]{filimonenko2018weak}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
-  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Nizovtsev}},\ and\ \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
-  {Kilin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Semiconductors}\ }\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {1865}
-  (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Filimonenko}\ \emph {et~al.}(2022)\citenamefont
-  {Filimonenko}, \citenamefont {Yasinskii}, \citenamefont {Nizovtsev},
-  \citenamefont {Kilin},\ and\ \citenamefont
-  {Jelezko}}]{filimonenko2022manifestation}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
-  {Filimonenko}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
-  {Yasinskii}}, \bibinfo {author} {\bibfnamefont {A.~P.}\ \bibnamefont
-  {Nizovtsev}}, \bibinfo {author} {\bibfnamefont {S.~Y.}\ \bibnamefont
-  {Kilin}},\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
-  {Jelezko}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Journal of Applied Spectroscopy}\ }\textbf {\bibinfo {volume} {88}},\
-  \bibinfo {pages} {1131} (\bibinfo {year} {2022})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Vetter}\ \emph {et~al.}(2022)\citenamefont {Vetter},

### 2022-07-12 Mes papiers/Low field CR/V3/SI/Low_field_CR_SI.tex
la tete de tes morts

@@
 \includegraphics[width=0.9\textwidth]{Figures_SI/fig_ESR}
-\caption{ODMR measurement (a) under zero magnetic field, (b) under a magnetic field of $~ 50 G$ when zooming on a single NV class}
+\caption{ODMR measurement (a) under zero magnetic field, (b) under a magnetic field of $~ 50 G$ when zooming on a single NV class.}
 \label{ESR_single_spin}
@@
 \caption{Numerical simulations showing the closeness of the Hamiltonian second excited state $\ket{e}$ with the states $\ket{+1}$ and $\ket{+}$, as a function of the magnetic field amplitude and angle $\theta$ with respect to the NV axis.
-A value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen}
+A value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen.}
 \label{map etats propres}
@@
 \label{NV Hamiltonian}
-\mathcal{H}_s/h=D S_z^2 + \gamma_e \bm{B}_{\rm ext} \cdot \bm{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right]
+\mathcal{H}_s/h=D S_z^2 + \gamma_e \bm{B}_{\rm ext} \cdot \bm{S}+ d_\perp \left[ E_x(S_y^2-S_x^2) + E_y(S_xS_y+S_yS_x) \right],
 \end{equation}
-Where $D=2.87\ \rm GHz$ is the zero field splitting and $\gamma_e=2.8\ \rm MHz/G$ the gyromagnetic ratio of the electron.
+where $D=2.87\ \rm GHz$ is the zero field splitting and $\gamma_e=2.8\ \rm MHz/G$ the gyromagnetic ratio of the electron.
 
@@
 5 of the main text.
-\item \textbf{CVD-PPM} : A CVD bulk diamond described in \citep{TALLAIRE2020421}.
-It has been irradiated and annealed and contains [NV$^-$] $\approx 4\ \rm ppm$.
+\item \textbf{CVD-1} : A CVD type IIa bulk diamond. This sample has not been irradiated and contains [NV$^-$] $\approx 50\ \rm ppb$. This sample is used in Fig. 2(a) of the main text.
+\item \textbf{CVD-2} : A CVD bulk diamond described in \citep{TALLAIRE2020421}.
+This one has been irradiated and annealed and contains [NV$^-$] $\approx 4\ \rm ppm$.
 This sample is used in Fig.
 \ref{Alignment} of the SI.
-\item \textbf{CVD-PPB} : Another CVD bulk diamond.
-This one has not been irradiated or annealed and therefore contains [NV$^-$] $\approx 50\ \rm ppb$.
-This sample is used in Fig.
-2(a) of the main text.
 
@@
 \includegraphics[width=0.7\textwidth]{Figures_SI/shema_exp}
-\caption{Experimental Setup}
+\caption{Experimental Setup.}
 \label{setup}
@@
 We did not use magnetic shielding to protect from the earth magnetic field since the splitting due to  $B_{\rm earth}$ is lower than the splitting due to the local electric field for the samples used here : $ \gamma_e B_{\rm earth} \approx 1.5\ \rm MHz < d_\perp E_\perp \approx 4\ \rm MHz$.
-When possible, the electromagnet was roughly aligned along the earth magnetic field to better compensate it.
+
 \section{Experimental details}
@@
 \includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_combined}
-\caption{$T_1$ measurement with purely exponential and purely stretched exponential fits (a) in zero magnetic field (b) in non-zero magnetic field}
+\caption{$T_1$ measurement with purely exponential and purely stretched exponential fits (a) in zero magnetic field (b) in non-zero magnetic field.}
 \label{T1_fits}
@@
 \includegraphics[width=0.45\textwidth]{Figures_SI/fig_alphas}
-\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$ as a function of an arbitrarely oriented magnetic field amplitude}
+\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$ as a function of an arbitrarely oriented magnetic field amplitude.}
 \label{alphas}
@@
 The result is fitted with exponential and stretched-exponential (with a stretch factor $\beta=0.5$) functions.
-We can see that for the shorter lifetime (zero field), the measurement follows more closely the stretched-exponential profile, while the longer lifetime is closer to the exponential one (except at very short times).
+We can see that for the shorter lifetime (at zero field), the measurement follows more closely the stretched-exponential profile, while the longer lifetime is closer to the exponential one (except at very short times).
 
@@
 \includegraphics[width=0.7\textwidth]{Figures_SI/overlap}
-\caption{Illustration of the spectral overlap for two gaussian spectra}
+\caption{Illustration of the spectral overlap for two gaussian spectra.}
 \label{overlap}
@@
 \begin{equation}
-S(\Delta \nu)=\int S_1(\nu, \nu_1)S_2(\nu, \nu_2) d\nu
+S(\Delta \nu)=\int S_1(\nu, \nu_1)S_2(\nu, \nu_2) d\nu.
 \end{equation}
@@
 S(\Delta \nu)&\propto \int \exp(-\frac{(\nu-\nu_1)^2}{2\sigma^2})\exp(-\frac{(\nu-\nu_2)^2}{2\sigma^2}) d\nu \\
-&\propto\exp(-\frac{(\Delta \nu)^2}{4\sigma^2})
+&\propto\exp(-\frac{(\Delta \nu)^2}{4\sigma^2}).
 \end{align*}
@@
 S(\Delta \nu)&\propto \int \frac{1}{1+ \frac{(\nu-\nu_1)^2}{\sigma^2}}\cdot \frac{1}{1+ \frac{(\nu-\nu_2)^2}{\sigma^2}} d\nu \\
-&\propto\frac{1}{1+ \frac{(\Delta \nu)^2}{4\sigma^2}}
+&\propto\frac{1}{1+ \frac{(\Delta \nu)^2}{4\sigma^2}}.
 \end{align*}
@@
 \includegraphics[width=0.6\textwidth]{Figures_SI/alignement}
-\caption{PL of sample CVD-PPM as a function of the external magnetic field amplitude for different field misalignment with the [100] axis.}
+\caption{PL of sample CVD-PPM as a function of the external magnetic field amplitude for different magnetic field misalignments with respect to the [100] axis.}
 \label{Alignment}
@@
 Fig.
-\ref{Alignment} shows the change in PL with respect to the external magnetic field amplitude for various misalignments of the magnetic field direction compared to the [100] diamond crystalline axis of the sample CVD-PPM.
+\ref{Alignment} shows the change in PL with respect to the external magnetic field amplitude for various misalignments of the magnetic field direction compared to the [100] diamond crystalline axis of the sample CVD-2.
 We can see that the central drop in PL is not critically affected by the misalignment of the magnetic field, and we are therefore confident that the drop in PL we observe when $B\parallel [100]$ is not an effect of misalignment.
@@
 We estimate the initial alignment precision to be $\pm 1^\circ$.
-The misalignment was then added by rotating the electromagnet along an arbitrary axis.
+The misalignment was then introduced by rotating the electromagnet along an arbitrary axis.
 The slight asymmetry between the positive and negative value of the magnetic field could come from the earth magnetic field or from hysteresis in the magnetic core of our electromagnet.
@@
 The spin Hamiltonian is given in eq.
-\ref{NV Hamiltonian} and a value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen (b) Blue dashed curve : frequency splitting between the $\ket{d}$ and $\ket{e}$ states as a function of the transverse magnetic field, red plain line : matching factor $\abs{\bra{e}\ket{+}}^2$ between the $\ket{e}$ and $\ket{+}$ states}
+\ref{NV Hamiltonian} and a value of $d_\perp E_\perp = 4\ \rm MHz$ was chosen (b) Blue dashed curve : frequency splitting between the $\ket{d}$ and $\ket{e}$ states as a function of the transverse magnetic field, red plain line : matching factor $\abs{\bra{e}\ket{+}}^2$ between the $\ket{e}$ and $\ket{+}$ states.}
 \label{theory_transverse_field}
@@
 This leads to a new theoretical stretched lifetime $T_1^{\rm dd}= 4\, T_0$.
-\item When a magnetic field is applied along the [100] crystalline axis (for $5\ \rm G < \abs{B} \ll D$), all four classes are resonant and dominated by the longitudinal part of the magnetic field.
+\item When a magnetic field is applied along the [100] axis (for $5\ \rm G < \abs{B} \ll D/\gamma_e$), all four classes are resonant and dominated by the longitudinal part of the magnetic field.
 This results, in increase in the average flip-flop rate compared to the isolated class case, and the predicted stretched exponential lifetime is $T_1^{\rm dd}\approx 42.8\, T_0$.
@@
 
-Assuming a homogeneous distribution of the fluctuators in the bulk of the crystal, the authors of \cite{choi_depolarization_2017} conclude that the fluctuators create an additionnal decay channel for the NV population, through dipole-dipole interaction, characterized by a decay rate $\gamma$ which follows the probability distribution :
+Assuming a homogeneous distribution of the fluctuators in the bulk of the crystal, the authors of \cite{choi_depolarization_2017} show that the fluctuators create an additionnal decay channel for the NV population, through dipole-dipole interaction, characterized by a decay rate $\gamma$ which follows the probability distribution :
 \begin{equation}
@@
 \begin{equation}
-\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f}.
+\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f},
 \label{eq 1/T}
 \end{equation}
-Where $n_f$ is the fluctuator density in the crystal in nm$^{-3}$, $J_0=52\ \rm{MHz}\cdot\rm{nm}^3$ is the characteristic dipole-dipole strength between two spins, $\gamma_f$ is the fluctuator intrinsic decay rate and $\bar \eta$ is a dimension-less number which characterizes the average dipole-dipole interaction between the NV centers and the fluctuators (resonance conditions, relative orientations etc.
+where $n_f$ is the fluctuator density in the crystal in nm$^{-3}$, $J_0=52\ \rm{MHz}\cdot\rm{nm}^3$ is the characteristic dipole-dipole strength between two spins, $\gamma_f$ is the fluctuator intrinsic decay rate and $\bar \eta$ is a dimension-less number which characterizes the average dipole-dipole interaction between the NV centers and the fluctuators (resonance conditions, relative orientations etc.
 Further details are given below).
@@
 &+\left[3(\hat{u}\cdot\hat{z}_1)(\hat{u}\cdot\hat{z}_2) -\hat{z}_1\cdot\hat{z}_2\right] S_z^1S_z^2, \\
-&+\mathcal{H}_{\rm other}
+&+\mathcal{H}_{\rm other},
 \end{align}
-where $J_0=\frac{\mu_0 \gamma_e^2 \hbar^2}{4 \pi} = (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\bm r= r\bm{\hat u}$ is the distance between the two spins and  $\mathcal{H}_{\rm other}$ contains terms of the form $S_{x/y}^iS_z^j$ which couple states far from resonance and will be neglected here.
+where $J_0=\frac{\mu_0 \gamma_e^2 \hbar^2}{4 \pi} = (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\bm r= r\bm{\hat u}$ is the distance between the two spins and  $\mathcal{H}_{\rm other}$ contains terms of the form $S_x^iS_z^j$ and $S_y^iS_z^j$ which couple states far from resonance and will be neglected here.
 
@@
 \label{eq. eta bar}
-\bar \eta =\frac{1}{4} \sum_{i=1}^4 \int_\theta \int_\phi  \int_{\omega_f} \int_{\omega_{NV}}   \abs{\eta(\bm u,i,\omega_{NV},\omega_f)}\, d\Omega\, \rho(\omega_{NV}) d\omega_{NV}\, \rho^i(\omega_f) d\omega_f
+\bar \eta =\frac{1}{4} \sum_{i=1}^4 \int_\theta \int_\phi  \int_{\omega_f} \int_{\omega_{NV}}   \abs{\eta(\bm u,i,\omega_{NV},\omega_f)}\, d\Omega\, \rho(\omega_{NV}) d\omega_{NV}\, \rho^i(\omega_f) d\omega_f ,
 \end{equation}
 
-Where $d\Omega=\sin \theta d\theta d\phi$, $\rho(\omega_{NV})$ is the distribution of angular frequencies of the probed NV centers, and  $\rho^i(\omega_f)$ is the distribution of angular frequencies  for the fluctuators of class $i$.
+where $d\Omega=\sin \theta d\theta d\phi$, $\rho(\omega_{NV})$ is the distribution of angular frequencies of the probed NV centers, and  $\rho^i(\omega_f)$ is the distribution of angular frequencies  for the fluctuators of class $i$.
 
@@
 \item The NV and fluctuator are from the same class, so that the angle $\widehat{z_1 z_2}=0$ : this is always true, regardless of the magnetic field angle and amplitude.
-We can analytically compute $\bar \eta$ in this case and find : $$ \bar {\eta}_{same}=\frac{1}{4}  \int_\theta \int_\phi \abs{\eta(\bm u,\widehat{z_1 z_2}=0,\omega_{NV}=\omega_f)}\, d\Omega= \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot \frac{2}{3 \sqrt{3}} \approx 5.55\cdot10^{-2}$$
+We can analytically compute $\bar \eta$ in this case and find : $$ \bar {\eta}_{same}=\frac{1}{4}  \int_\theta \int_\phi \abs{\eta(\bm u,\widehat{z_1 z_2}=0,\omega_{NV}=\omega_f)}\, d\Omega= \frac{1}{4} \cdot \sqrt{\frac{1}{3}} \cdot \frac{2}{3 \sqrt{3}} \approx 5.55\cdot10^{-2}.$$
 \item The angle $\widehat{z_1 z_2}=\rm{arccos}(\frac{1}{3})\approx 70.5^\circ$.
@@
 \includegraphics[width=0.7\textwidth]{Figures_SI/Various_ESR}
-\caption{ODMR spectra for different orientations of the magnetic field. Red numbers represent the number of degenerate classes for each line of the spectrum}
+\caption{ODMR spectra for different orientations of the magnetic field. Red numbers represent the number of degenerate classes for each line of the spectrum.}
 \label{Various ODMR}
@@
 \begin{itemize}
-\item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
+\item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$.
 \item $\bm{B} \in$ \{110\} : For the two non-resonant classes : $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2= \bar \eta_0^2$.
-For the two resonant classes : $\bar \eta = \bar {\eta}_{\rm same} + \bar {\eta}_{\rm far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
+For the two resonant classes : $\bar \eta = \bar {\eta}_{\rm same} + \bar {\eta}_{\rm far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$.
 \item $\bm{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{\rm same}+\bar {\eta}_{\rm close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
@@
 \begin{equation}
-\eta^2=\frac{1}{3} \abs{\mel{\pm ,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,\pm }}^2  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}
+\eta^2=\frac{1}{3} \abs{\mel{\pm ,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,\pm }}^2  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
 \end{equation}
@@
   \bra{+}
-  \end{matrix}
+  \end{matrix}.
   \end{equation}

### 2022-07-19 Manuscrit Thèse/chapter 2/Appendix dipole-dipole.tex
la chaleur de tes morts

--- 2022-07-08 Manuscrit Thèse/chapter 2/Appendix dipole-dipole.tex
@@
 \begin{document}
-\section{The dipole-dipole Hamiltonian}
+\chapter{Appendix A : Dipole-Dipole Hamiltonian}
+\section{Derivation of the dipole-dipole Hamiltonian}
 The dipole-dipole interaction between two point-like magnetic dipoles can be derived classically by computing the static magnetic field generated by one dipole at the position of the second dipole.
@@
 \begin{equation}
-\mathcal{H}_{dd}=\frac{J_0}{r^3}\left[ \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2 - 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u})\right]
+\mathcal{H}_{dd}=\frac{J_0}{r^3}\left[ \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2 - 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u})\right],
 \end{equation}
+where $J_0\equiv \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$.
 
-Where $J_0\equiv \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$. For two electronic spins with $g-$factors close to 2, the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$.
-
-\section{Flip-Flops, double flips and cross-relaxations}
-
-Flip-flop, in the context of dipole-dipole interaction, is the mechanism which exchanges a quantum of spin between two spins.
-For an initial state of two spins $\ket{m_s^1=i;m_s^2=j}$, a flip-flop process will result in a final state $\ket{i+1;j-1}$ or $\ket{i-1;j+1}$.
-The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$
+\section{Decomposition of $\mathcal{H}_{dd}$ over two different spin basis}
 

### 2022-07-19 Manuscrit Thèse/chapter 2/Chapter 2.tex
la chaleur de tes morts

\documentclass[a4paper]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_ch2.bib} %Import the bibliography file

\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{Appendix A : Dipole-dipole interaction between two spins in solid}
\begin{document}
\chapter{Detection of dark spins via cross-relaxations}
In this chapter, we will see how...
\section{Flip-Flops, double flips and cross-relaxation}

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
\label{CR_shema}
\end{figure}

The dominant form of interaction between two distant spins is the magnetic dipole-dipole interaction. For two spins with with vector spin Hamiltonian $\hat{\mathbf{S}}_1$ and $\hat{\mathbf{S}}_2$, separated by a distance $\mathbf{r}=r\mathbf{u}$, the dipole-dipole interaction Hamiltonian reads (see Appendix A) :
\begin{equation}
\mathcal{H}_{dd}=\frac{J_0}{r^3}\left[ \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2 - 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u})\right],
\end{equation}
where $J_0= \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$. For two electronic spins with $g-$factors close to 2 (which is the case for the NV center and most spin defect in diamond), the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$.

Flip-flop, in the context of dipole-dipole interaction, is the mechanism which exchanges a quantum of spin between two spins. For an initial state of two spins $\ket{m_s^1=i;m_s^2=j}$, a flip-flop process will result in a final state $\ket{i+1;j-1}$ or $\ket{i-1;j+1}$. The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$, as well as the resonance condition between the initial and final state. We should note that for two identical spins, the flip-flop process $\ket{i;i\pm 1}\bra{i\pm 1,i}$ is always resonant.

Another mechanism authorized by the dipole-dipole Hamiltonian is the double-flip, either up or down, which couple the states $\ket{i;j}$ and $\ket{i-1;j-1}$ or $\ket{i+1;j+1}$. For a single spin species in a strong magnetic field, the lift of the various spin levels by the Zeeman effect means that no double-flip process can be resonant. However with weak magnetic field or when two different spin species are present, double flip process can be as important as flip-flops since the matrix elements $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \pm 1}$ are typically of the same order of magnitude as the flip-flop ones.

Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family). This process con occur either through flip-flops, or through double flips, as long as the resonance condition between the two spins is met.

Fig. \ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin. When the two considered spin transitions become resonant, in this case by tuning a magnetic field, polarization from the NV center will be transferred to the second spin, meaning in this case that the NV center will end up less polarized and the second spin more polarized.

The reason why CR tend to depolarize the NV center simply comes from a rate equation : since the initial NV population is higher in the $\ket{0}$ state than in the $\ket{-1}$ state, the flip-flop (or double flip) process that makes the NV center go to $\ket{-1}$ (symbolized by the two red arrows in Fig. \ref{CR_shema}) is more likely than the reverse flip-flop. If we only focus on the NV center, cross-relaxation is simply another thermalization process due to the coupling of the NV center to its environment.

\section{Dark spins in diamond}

We refer by dark spins to spin defects which do not interact with light, which in the case of diamond defects consist of every known spin defect bar the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.

By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen center known as the P1 center, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^13$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^12$C isotope.

\section{NV center relaxometry}

\subsection{Principle of NV center Relaxometry}
Detection \textit{via} cross-relaxation	with NV centers fall under the more general technique of NV relaxometry, where we are effectively measuring a modification in the NV center spin lifetime $T_1$. Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with other spins \citep{van1989cross,holliday1989optical,armstrong2010nv, hall2016detection,wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...

One of the main advantage of relaxometry compared to methods based on the spin coherence time $T_2$ or $T_2^*$ is that for NV center, $T_1$ is typically $10^2 \sim 10^3$ times longer than the spin echo time $T_2$, and even more so for the inhomogeneous coherence time $T_2^*$. This means that for a comparable change in $T_1$ and $T_2$, measurement based on $T_1$ will be much more sensitive. In the case of resonant detection of spins, the cross-relaxation protocol offers a theoretical gain by a factor of $\sqrt{\frac{T_2}{T_1}}$ \citep{hall2016detection} in its signal to noise ratio (SNR) compared to the DEER protocol based on $T_2$, as will be discussed below.

\subsection{PL and single$-\tau$ relaxometry protocols}
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{Figures/relaxo_pulsed_continuous}
\caption{Representation of the two relaxometry protocol. Left : representation of the three level system of the NV center with the polarization rate $\Gamma_{\rm las}$ and spin decay rates $\Gamma_1=1/T_1$ and the perturbation $\delta \Gamma$. Right: simulation of a $T_1$ measurement for $1/\Gamma_1=3\ \rm ms$ and $1/(\Gamma_1+\delta \Gamma)=2\ \rm ms$. $\delta \rm S$ represents the maximum difference in signal between the two curves for a dark time $\tau$.}
\label{T1 vs PL}
\end{figure}

There are two ways to measure a change in the spin $T_1$, as represented on Fig. \ref{T1 vs PL}. The first and most basic method simply consist in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, which itself results from an equilibrium between the polarization rate $\Gamma_{\rm las}$ and the spin decay rate $\Gamma_1=1/T_1$. A change in $\Gamma_1$ will therefore conduct to a new equilibrium with a different PL. For a small enough change $\delta \Gamma << \Gamma_1$, we can consider that we remain in the linear regime, hence why $\frac{\delta \rm PL}{\rm PL} \propto \frac{\delta \Gamma}{\Gamma_1}$.

The second method is a pulsed sequence referred as single$-\tau$ measurement, used for example in \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}. It simply consist in a $T_1$ measurement sequence (as seen in chapter 1) where you polarize the spins in the $\ket{0} state$ with a first laser pulse and read out the spin state with a second laser pulse after a dark time $\tau$. In this case however, instead of scanning the time $\tau$, you fix its value to maximize the sensitivity in a change of $T_1$ (typically $\tau \approx T_1$). Again, assuming you are in the linear regime, you observe a change in signal proportional to the change in $\Gamma_1$ : $\frac{\delta \rm S}{\rm S} \propto \frac{\delta \Gamma}{\Gamma_1}$.

It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the first case you need to adjust $\Gamma_{\rm las} \sim \Gamma_1$ to maximize the PL contrast, meaning that the laser power used is very far from the saturation limit (typically $\Gamma_1 \approx 10^3\ \rm s^{-1}$ and $\Gamma_{\rm sat} = 5\cdot10^6\ \rm s^{-1}$\citep{dreau2011avoiding}), while the second method needs to wait a time $\tau \sim T_1$ between each measurement, which result in a similar PL count and shot noise limit in both case.

On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and uses an overall smaller laser power, however it is more sensitive to drifts and changes in the optical setup. Most of the relaxometry measurement in this manuscript were performed using a PL-based detection.

\subsection{Comparison with DEER protocol}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{Figures/DEER}
\caption{Pulse sequence of the DEER protocol}
\label{DEER}
\end{figure}

Outside of cross-relaxations, another protocol exists to address dark spin resonance with NV centers, based on a modification of the NV $T_2$ time \citep{serbyn2014interferometric}. This protocol called double electron-electron resonance (DEER) is depicted on Fig. \ref{DEER}. The pulse sequence consist in a spin-echo sequence on the NV center, as described in chapter 1, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers. This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center wont be rephased. The newly measured $T_2$ time should therefore be slightly shorter:

\begin{equation}
\frac{1}{T_2^{\rm DEER}}=\frac{1}{T_2^{\rm echo}} + \delta \Gamma,
\end{equation}

where $\delta \Gamma$ is the contribution of the dark spin to the $T_2^*$ of the NV center through NV-dark spin dipole-dipole interaction.

By scanning the frequency of the second microwave (RF2) and monitoring the $T_2$ time of the NV centers, one can therefore find dark spin resonance without having to bring them to resonance with NV centers, which can be useful when the dark spin resonance is very far detuned from the NV one.

While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, they don't actually measure the same component of the dipole-dipole Hamiltonian: CR measure the off-diagonal terms responsible for flip-flops and double flips, which are the prefactor of the $S_{\overset{x}{y}} S_{\overset{x}{y}}$ terms. On the other hand, $T_2$ measurement are sensitive to the change in the Larmor frequency of the NV center, meaning the diagonal terms of the dipole-dipole Hamiltonian, which are the prefactor of the $S_z S_z$ term. Nevertheless, these prefactors are on average of the same order of magnitude, so that the $\delta \Gamma$ term is typically the same for DEER and CR.

This means that the signal of the CR protocol is typically greater by a factor $\frac{\delta \Gamma \cdot T_1}{\delta \Gamma \cdot T_2}=\frac{T_1}{T_2}$ compared to the DEER signal, while the time it takes to perform a CR single$-\tau$ sequence is greater by a factor $\frac{T_1}{T_2}$ compared to the DEER sequence, since in both cases the optimal waiting time $\tau$ is of the order the measured value ($T_1$ or $T_2$). Because the SNR scales as the inverse square root of the measuring time, this gives an increase in SNR by a factor $\sqrt{\frac{T_1}{T_2}}$ in favor of the CR protocol compared to the DEER one.

\printbibliography
\end{document}

### 2022-07-22 Manuscrit Thèse/chapter 2/Chapter 2.tex
toto

@@
 
+If we make the assumption that the spin bath coupled to the NV centers is Markovian, meaning that the dynamics of the bath is not affected by its coupling to the NV centers, we can then compute the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ due to its coupling to Spin 2 \citep{hall2016detection}:
+\begin{equation}
+\delta \Gamma_1=\frac{\Omega^2}{\Gamma_2^*} \frac{(\Gamma_2^*)^2}{(\delta \nu)^2+(\Gamma_2^*)^2},
+\label{delta gamma 1}
+\end{equation}
+where $\Omega=\frac{\mel{i;j}{\mathcal{H}_{\rm dd}}{j;i}}{\hbar}$ is the matrix element of the dipole-dipole Hamiltonian that couples the initial and final states, $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, and $\delta \nu$ is the detuning between the central frequencies of the NV center and spin 2.
 \section{Dark spins in diamond}
@@
 
-By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen center known as the P1 center, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^13$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^12$C isotope.
+\subsection{P1 centers and $^{13}$C}
+By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^13$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^12$C isotope.
+
+In NV-rich diamonds, these two impurities are the main causes of the inhomogeneous dephasing rate $\Gamma_2^*=1/T_2^*$ of the NV centers ensemble \citep{barry2020sensitivity}.
+The dephasing comes from the distribution of dipole-dipole coupling from each NV centers to its neighboring spin defects, which varies in space and time.
+
+The dephasing rate caused by a given impurity is directly proportional to the diagonal dipole-dipole matrix elements \citep{bauch2020decoherence}: 
+\begin{equation}
+\Gamma_2^*(P1) \propto \Omega_{\rm NV-P1}=\frac{1}{\hbar}\sqrt{\sum_j \abs{\mel{NV=i;P1=j}{\mathcal{H}_{\rm dd}}{NV=i;P1=j}}^2}.
+\label{Gamma2 P1}
+\end{equation}
+Since the dipole-dipole coupling rate scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity.
+In the case of P1 centers, this was verified experimentally by \citep{bauch2020decoherence} who found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
+
+$^{13}$C is much more abundant than P1 center ($\sim$ 10 000 ppm at natural abundance) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\, 10^3$ times smaller than that of the P1.
+This results in a contribution to $\Gamma_2^* \sim 10^3$ times smaller per impurity.
+It was found experimentally that the broadening caused by $^13$C is $\Gamma_2^*(^{13}C) \approx 16\ \rm Hz/ppm$ \citep{van1997dependences, barry2020sensitivity}.
+For natural abundance diamond, this means that $\Gamma_2^*(^{13}C) \approx 160\ \rm kHz$.
+
+\subsection{Other spin defects, VH$^-$, War1}
+
+Other spin defects are also commonly found in diamond containing NV centers, such as charged vacancy or vacancy clusters \citep{hounsome2006origin}, other nitrogen related defects \citep{newton2007epr}, transition metals \citep{isoya1990fourier}, and especially for CVD grown diamond, hydrogen related defects \citep{hartland2014study}, since the CVD plasma contains up to 95 \% hydrogen.
+
+Of particular interest in this chapter are two defects: the VH$^-$ defect \citep{glover2003hydrogen, glover2004hydrogen} and the War1 defect \citep{cruddace2007magnetic}.
+
+The VH$^-$ defect, prior to our observations, had only been observed \textit{via} electron paramagnetic resonance (EPR) in CVD grown diamond.
+It is formed by a vacancy with a hydrogen atom bonded to one of the four dangling carbon bond, with an extra electron trapped in the vacancy.
+This defect has a similar electronic structure to the NV$^-$ which results in an electronic spin-1 with a similar zero field splitting parameter (see Table [REF]).
+Although the presence of hydrogen in this defect was confirmed by isotopic study, it has been debated whether this defect was VH$^-$ or V$_2$H$^-$ due to a conflict with \textit{ab initio} calculations \citep{shaw2005importance}.
+The defect is therefore sometimes referred as V$_n$H$^-$.
+In this manuscript I will employ the original name VH$^-$.
+
+The War1 defect, named after the University of Warwick where it was first observed in a CVD diamond \citep{cruddace2007magnetic}, is another spin-1 defect whose chemical structure remains unknown.
+No hyperfine structure could be discerned in the relatively broad EPR line, and the use of deuterium instead of $^1$H did not change the shape of the line.
 
@@
 Detection \textit{via} cross-relaxation	with NV centers fall under the more general technique of NV relaxometry, where we are effectively measuring a modification in the NV center spin lifetime $T_1$.
-Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with other spins \citep{van1989cross,holliday1989optical,armstrong2010nv, hall2016detection,wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
+Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross,holliday1989optical,armstrong2010nv, hall2016detection,wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
 
-One of the main advantage of relaxometry compared to methods based on the spin coherence time $T_2$ or $T_2^*$ is that for NV center, $T_1$ is typically $10^2 \sim 10^3$ times longer than the spin echo time $T_2$, and even more so for the inhomogeneous coherence time $T_2^*$.
-This means that for a comparable change in $T_1$ and $T_2$, measurement based on $T_1$ will be much more sensitive.
-In the case of resonant detection of spins, the cross-relaxation protocol offers a theoretical gain by a factor of $\sqrt{\frac{T_2}{T_1}}$ \citep{hall2016detection} in its signal to noise ratio (SNR) compared to the DEER protocol based on $T_2$, as will be discussed below.
+There can be several advantages to use relaxometry protocols, which measures a change in the NV $T_1$, over the more common magnetometry protocols that rely on a change in the NV $T_2$ or $T_2^*$: the first and most obvious one is that relaxometry does not need to use an external microwave field.
+This however comes at the cost of a greater precision needed on the external magnetic field.
+Relaxometry also allows AC-magnetic field sensing in the GHz regime, where traditional microwave-based AC-megnetometry is limited to tens of MHz \citep{wang2022picotesla, alsid2022solid}.
+Finally, relaxometry measurement can be more sensitive than the T2-based one in some instances \citep{steinert2013magnetic}, which comes from the fact that changes in NV $T_1$ times are easier to detect than changes in $T_2$ or $T_2^*$ since $T_1$ is much longer than $T_2$ or $T_2^*$.
 
 \subsection{PL and single$-\tau$ relaxometry protocols}
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 The second method is a pulsed sequence referred as single$-\tau$ measurement, used for example in \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}.
-It simply consist in a $T_1$ measurement sequence (as seen in chapter 1) where you polarize the spins in the $\ket{0} state$ with a first laser pulse and read out the spin state with a second laser pulse after a dark time $\tau$.
+It simply consists in a $T_1$ measurement sequence (as seen in chapter 1) where you first polarize the spins in the $\ket{0}$ state with a laser pulse and read out the spin state with a second laser pulse after a dark time $\tau$.
 In this case however, instead of scanning the time $\tau$, you fix its value to maximize the sensitivity in a change of $T_1$ (typically $\tau \approx T_1$).
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 
-While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, they don't actually measure the same component of the dipole-dipole Hamiltonian: CR measure the off-diagonal terms responsible for flip-flops and double flips, which are the prefactor of the $S_{\overset{x}{y}} S_{\overset{x}{y}}$ terms.
-On the other hand, $T_2$ measurement are sensitive to the change in the Larmor frequency of the NV center, meaning the diagonal terms of the dipole-dipole Hamiltonian, which are the prefactor of the $S_z S_z$ term.
-Nevertheless, these prefactors are on average of the same order of magnitude, so that the $\delta \Gamma$ term is typically the same for DEER and CR.
+While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, there are several differences between the two protocols :
 
-This means that the signal of the CR protocol is typically greater by a factor $\frac{\delta \Gamma \cdot T_1}{\delta \Gamma \cdot T_2}=\frac{T_1}{T_2}$ compared to the DEER signal, while the time it takes to perform a CR single$-\tau$ sequence is greater by a factor $\frac{T_1}{T_2}$ compared to the DEER sequence, since in both cases the optimal waiting time $\tau$ is of the order the measured value ($T_1$ or $T_2$).
-Because the SNR scales as the inverse square root of the measuring time, this gives an increase in SNR by a factor $\sqrt{\frac{T_1}{T_2}}$ in favor of the CR protocol compared to the DEER one.
+Frist, the two protocols do not measure the same elements of the dipole-dipole Hamiltonian: CR measures the off-diagonal elements $\Omega_{\rm CR}=\mel{i,j}{\mathcal{H}_{\rm dd}}{j,i}/\hbar$ which are responsible for the flip-flops or double flips, while DEER measures the diagonal elements $\Omega_{\rm DEER}=\mel{i,j}{\mathcal{H}_{\rm dd}}{i,j}/\hbar$ which are responsible for the pure dephasing.
+These terms however are on average of the same amplitude, and when averaging over an ensemble we can consider $\Omega_{\rm CR} \sim \Omega_{\rm DEER}$
+
+Second and more importantly, the scaling is not the same in both cases : CR measures a change in the NV decay rate $\Gamma_1=1/T_1$ which is proportional to $\Omega_{\rm CR}^2$, as given by eq.
+\ref{delta gamma 1}, while DEER measures a change in $\Gamma_2=1/T_2$ which is proportional to $\Omega_{\rm DEER}$, as given by eq.
+\ref{Gamma2 P1}.
+This means that the signal of CR scales quadratically with the impurity concentration, while the DEER signal scales linearly with it.
+CR is therefore more suited to probe concentrated impurities, when the average distance between the NV and the probed spin is small.
+
+\section{Choice of the $\mathbf{B}$ scan orientation}
+
+Since relaxometry does not rely on a microwave field, the only parameter left to tune is the external magnetic field, which is often scanned with an electromagnet.
+Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect dark spins.
+
+\section{Attribution of the VH$^-$ and War1 lines}
 

### 2022-07-22 Manuscrit Thèse/chapter 2/AA_simu NRJ.py

@@ -1,0 +1,67 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+n=500
+
+Brange=np.linspace(0,1000,n)
+transi1=np.zeros(n)
+transi2=np.zeros(n)
+transi3=np.zeros(n)
+transi4=np.zeros(n)
+transi5=np.zeros(n)
+transi6=np.zeros(n)
+transi7=np.zeros(n)
+transi8=np.zeros(n)
+
+
+for i in range(n):
+	B=magneticField(x=0,y=0,z=Brange[i])
+	# B=magneticField(x=Brange[i]/sqrt(3),y=Brange[i]/sqrt(3),z=Brange[i]/sqrt(3))
+	H=NVHamiltonian(B,c=2,E=3)
+	transis=H.transitions()
+	transi1[i]=transis[0]
+	transi2[i]=transis[1]
+
+	transiP1=2.8*Brange[i]
+
+	H=NVHamiltonian(B,c=1,E=3)
+	transis=H.transitions()
+	transi3[i]=transis[0]
+	transi4[i]=transis[1]
+
+	transi5[i]=transiP1
+
+	# H=NVHamiltonian(B,c=2,E=3,D=2694)
+	# transis=H.transitions()
+	# transi5[i]=transis[0]
+	# transi6[i]=transis[1]
+
+
+	# H=NVHamiltonian(B,c=1,E=3,D=2694)
+	# transis=H.transitions()
+	# transi7[i]=transis[0]
+	# transi8[i]=transis[1]
+
+	
+
+
+plt.figure(num=1,figsize=(6,4),dpi=80)
+plt.xticks(fontsize=14)
+plt.yticks(fontsize=14)
+
+plt.plot(Brange,transi1,color=color(0),label=r'NV')
+plt.plot(Brange,transi2,color=color(0))
+# plt.plot(Brange,transi3,color=color(0),label=r'NV $\nparallel$ B',ls='--')
+# plt.plot(Brange,transi4,color=color(0),ls='--')
+
+# plt.plot(Brange,transi5,color=color(1),label=r'P1')
+
+# plt.plot(Brange,transi5,color=color(1),label=r'VH')
+# plt.plot(Brange,transi6,color=color(1))
+# plt.plot(Brange,transi7,color=color(1),label=r'VH $\nparallel$ B',ls='--')
[…]

### 2022-07-24 Mes papiers/Low field CR/V4/Low_field_CR_PRL.tex
tata

\documentclass[preprintnumbers,amsmath,amssymb,superscriptaddress,twocolumn,showpacs]{revtex4-2}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{physics}
\usepackage[caption=false]{subfig}

\def\sgn{\mathop{\rm sgn}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\begin{document}

\title{Relaxation of Dipolar-Coupled Spins in Transverse Fields : \\ The role of Double-flip Processes}

\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, P. Huillery$^2$,  G. H\'etet}

\affiliation{Laboratoire De Physique de l'\'Ecole Normale Sup\'erieure, \'Ecole Normale Sup\'erieure, PSL Research University, CNRS, Sorbonne Universit\'e, Universit\'e Paris Cit\'e , 24 rue Lhomond, 75231 Paris Cedex 05, France \\$^2$Univ Rennes, INSA Rennes, CNRS, Institut FOTON - UMR 6082, F-35000 Rennes, France}

\begin{abstract}
We study relaxation processes in dipolar-coupled negatively charged nitrogen vacancy (NV$^-$) centers under transverse electric fields and magnetic fields. Specifically, we uncover regimes where flip-flop, double-flip processes as well as mixing induced by local electric fields play a significant role in NV-NV cross-relaxation.
Our results are relevant for understanding decoherence in many-body spin systems as well as for high sensitivity magneto- and electro-metry with long-lived interacting solid-state spins. As a proof of principle, we present an orientation and microwave-free magnetometer based on cross-relaxation.
\end{abstract}

\maketitle

The electronic spin properties of the negatively charged nitrogen-vacancy (NV$^-$) center in diamond have given rise to a wealth of applications in nanoscale sensing and quantum information science thanks in part to the possibility to optically polarize and read-out its spin state under ambient conditions \cite{DOHERTY20131}. In particular, ensembles of NV centers are widely studied for their enhanced magnetic field sensing capabilities \cite{Acosta, TALLAIRE2020421,edmonds2021characterisation, chatzidrosos2021fiberized, Barry, Bauch} and as pristine platforms for observing many-body effects \citep{kucsko2018critical, ChoiNat, ZuYao, dwyer2021probing}. When the NV spin concentration reaches ppm values, spin depolarisation, or cross-relaxation (CR) takes place through a very rich many-body dynamics associated with disorder \citep{choi_observation_2017}. These mechanisms limit the efficiency of typical microwave based NV magnetometers, but quantum control techniques can be combined to surpass this interaction limit \citep{zhou2020quantum}. Further, CR mechanisms can be turned as a tool for magnetic field sensing \cite{akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019}. 
Spectral features in the photoluminescence (PL) indeed appear when the magnetic field crosses specific crystal planes where dipolar interactions are enhanced, leading to CR. The projected sensitivity of such magnetometers lie in the tens of pT$/\sqrt{\rm Hz}$  \cite{akhmedzhanov_microwave-free_2017}, on a par with the most sensitive microwave based NV magnetometers \cite{Wolf, Sturner}.

Cross-relaxation features close to zero magnetic field have also been observed and could be deployed for higher sensitivity microwave-free magnetometry \cite{filimonenko2018weak, filimonenko2022manifestation}.
The CR contrast was indeed shown to be much larger in this zero-field limit \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015} but all the relaxation mechanisms have not been identified.
Here, we study dipolar relaxation processes in ensembles of NV centers in the presence of small transverse electric and magnetic fields. Specifically, by employing magnetic field scans along specific crystalline directions, we find regimes where flip-flop, double-flip processes as well as mixing induced by local electric fields play a role. We also present an orientation and microwave-free magnetometer that employs double-flip processes in the dipolar interaction between NV centers.

\begin{figure}
\includegraphics[width=0.45\textwidth]{shema_summary.pdf}
\caption{Schematics showing a diamond with interacting spins (central picture) as well as three processes that account for dipolar relaxation: flip-flop processes involving two different classes of NV centers, local electric field mixing and double-flip processes where two units of spin angular momentum are exchanged. }
\label{schema_intro}
\end{figure}

The electronic spin of NV center is a spin-1 system in the electronic ground state (see Fig. 1, top left panel), which can be optically polarized in the $\ket{m_s=0}$ state. The PL of this state is also larger than in the $\ket{m_s=\pm 1}$ states enabling spin-read out at room temperature \cite{DOHERTY20131}. The $\ket{m_s=\pm 1}$ spin states are separated from the $\ket{m_s=0}$ state by  $D = (2\pi) 2.87$ GHz so that when a resonant microwave or static transverse magnetic field is applied, the PL is reduced \citep{epstein2005anisotropic,lai2009influence}.  The processes that lead to depolarisation in strongly coupled spin-1 systems are depicted in Fig.~1. Flip-flop processes involving coupling between spins with identical or different orientations, or ``classes", are depicted in the top-right panel. They have already been shown to play a dominant role in dense ensembles of NV centers when $B \gtrsim 30$~G : tuning the difference $\delta$ between the different NV centers' spin states indeed result in a strong $T_1$ reduction \cite{choi_observation_2017}.  We will show here that double-flip up and down as well as mixing induced by local electric fields (cf. two bottom panels) can also give a significant contribution to the spin depolarization.

The experimental apparatus and the samples used in this study together with their nomenclature are presented in the supplementary material (SM) sec. II and III \citep{SI_low_filed_CR}\nocite{anishchik2015low, filimonenko2020weak, van1990electric, giri_coupled_2018, giri_selective_2019, Hall,jarmola_temperature-_2012}. 
%The nomenclature that we use indicates first the fabrication process: high pressure high temperature (HPHT) or chemical vapor deposition (CVD). 
%In the case of HPHT grown samples, these indications are followed by the size of the diamond particle in $\mu$m. 
Fig.  \ref{T1}-a) and b) show the change in PL as a function of a magnetic field applied {\it via} a homebuilt electromagnet for two diamond samples with a low and high concentration of NV$^-$ centers respectively (CVD-1 and HPHT-150-1). In both samples, we see a decrease in PL as the magnetic field amplitude increases. There is however a stark difference in the low magnetic field region where only high-density samples shows a drop in PL \citep{jarmola_longitudinal_2015,  mrozek_longitudinal_2015}. 
\begin{figure}
\includegraphics[width=0.45\textwidth]{fig_T1.pdf}
\caption{a) and b) : Photoluminescence from NV center ensembles as a function of a magnetic field applied in an arbitrary direction for sample CVD-1 with [NV$^-$]$\approx 50\ \rm ppb$ and for sample HPHT-150-1 with [NV$^-$]$\approx 3\ \rm ppm$ respectively. (c) Sequence used to measure the spin lifetime. (d) Trace i) and ii) are the spin relaxation signals $S_1-S_2$ measured for the dense sample at zero and non-zero magnetic fields respectively. The fitting procedure (orange plain line) is described in the main text.}
\label{T1}
\end{figure}

\begin{figure*}
\includegraphics[width=.95\textwidth]{fig_100_vs_1x1x1x1.pdf}
\caption{(a) ODMR spectrum in zero field with microwave amplitude modulation (see experimental details in SM sec. III). (b)-i) ODMR spectrum for a magnetic field $\approx$ 60 G, misaligned by $\sim  24^\circ$ from the [100] axis. (b)-ii) Normalized PL of the NV$^-$ ensemble as a function of the magnetic field amplitude. (b)-iii) Stretched part of the spin decay $1/T_1^{\rm dd}$ as a function of the magnetic field amplitude. (c)-i), (c)-ii) and (c)-iii): same measurements as (b) but with a magnetic field close to the [100] axis. All the measurements were realized using the sample HPHT-150-1.}
\label{100_VS_1x4}
\end{figure*}
%This effect can be observed on all our samples whenever the NV concentration lies in the ppm range. The drop of the PL at low magnetic field was shown to be associated with a decrease of the NV's spin lifetime $T_1$ \citep{jarmola_temperature-_2012}, which results in a decrease of the population in the bright state $\ket{0}$.
%The depolarization dynamics of the NV spins for single or dilute ensembles of NV centers at room temperature is dominated by two-phonon Raman processes \citep{redman1991spin,jarmola_temperature-_2012,norambuena2018spin}, which depend on the crystal lattice temperature. It has been observed that dense NV centers ensembles in fact  have an additional spin decay channel \citep{jarmola_temperature-_2012,jarmola_longitudinal_2015,mrozek_longitudinal_2015, choi2017depolarization, akhmedzhanov_microwave-free_2017, akhmedzhanov_magnetometry_2019, pellet2021magnetic, mrozek2021characterization}, which depends greatly on the magnetic field amplitude. 
This effect has been attributed to CR between the NV centers through dipole-dipole coupling \citep{mrozek_longitudinal_2015, choi2017depolarization}. 
%The reason is that at zero-field all the NV ``classes" - the physical orientation of the NV axes - become resonant, which increases the dipolar relaxation.
%Inhomogeneity of the spin lifetimes is further needed to explain the depolarization of the spin ensemble \citep{choi2017depolarization}. 
We will denote $T_1^{\rm ph}$ the characteristic timescale of phonon-induced relaxation and $T_1^{\rm dd}$ the density dependent timescale associated with dipole-dipole  interactions.
%Before analyzing the cause of low-field depolarization, we present relaxation measurements on HPHT-150-1. Fig. \ref{T1}-c) shows the sequence employed for measuring $T_1^{\rm dd}$. 
Fig. \ref{T1}-c) shows the sequence employed to measure $T_1$ and Fig. \ref{T1}-d) shows the result of the measurement when $B\approx 0$ and $B \approx 50\ \rm G$.
%It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$. In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019, choi2017depolarization}. It is therefore convenient to repeat the sequence with an additional $\pi$ pulse right on one of the eight NV spin-resonances before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state \citep{jarmola_temperature-_2012,mrozek_longitudinal_2015,choi2017depolarization}. By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of selecting a specific class of NV centers. 
%The relaxation rate is markedly different in both magnetic field settings, with a much more pronounced depolarization at zero-field.
%Notably, the ensemble spin decay curves are not exponential. This was found to be due to inhomogeneities between the spin lifetimes of dipolar coupled NV$^-$ centers \citep{choi2017depolarization}.
%We will interprate our experimental results using the NV-fluctuator model developed in \citep{choi2017depolarization}. This model is based on the existence of very short-lived NV centers - the so-called fluctuators - which, when coupled resonantly to the other NV centers, act as a polarization drain for the spin ensemble. A conclusion of this model is that, for a homogeneous 3D distribution of fluctuators, the dipole-induced relaxation should be a stretched-exponential with a stretch factor $\beta=1/2$. Such a trend was indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$ \cite{choi2017depolarization}. 
%We conducted measurements to confirm the stretch nature of the dipole-induced decay (see SM sec. IV A), as well as confirming the lifetime-limited model of the fluctuators(see SM sec. IV B).
All dense samples used in this study show $T_1^{\rm dd} \sim T_1^{\rm ph}$ (see SM sec. IV A), so both decay processes are included in the analysis. Fitting the two traces of Fig. 2 d) by the product of a simple and a stretched exponential, we find $T_1^{\rm ph}=3.6$ ms for both curves and $T_1^{\rm dd}= 0.6\ \rm ms$ and $13.0\ \rm ms$ for trace i) and ii) respectively. These results thus demonstrates a twenty-fold increase in the dipolar depolarization rate when the B field is turned to zero.
The fluctuator model developed in \citep{choi2017depolarization} contains some of the explanation for this increase in dipolar depolarization: all four NV classes in the diamond are resonant in zero field, which increases the flip-flop rates. However, the model did not include extra specificities in the zero magnetic field region, namely the role of local electric field and double flip processes.

%We therefore decided to conduct more in depth experiments in order to characterize these potential extra relaxation processes.

Fig. \ref{100_VS_1x4}-a) and b)-i) show optically detected magnetic resonance (ODMR) spectra with and without magnetic field, from the dense ensemble HPHT-150-1. From the resonance line-widths in a magnetic field, we extract decoherence rates $T_2^*$ in the hundreds of nanosecond range, limited by the coupling between NV centers and the fluctuating spins of substitutional nitrogen atoms (also called $P_1$ centers).
%Fig. \ref{100_VS_1x4} (b)-i) shows an ODMR spectrum from the same sample, in a magnetic field. The observed lines correspond to the projections of the magnetic fields onto the four classes of NV centers together with the $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$ spin transitions.
%The transitions frequencies of the four different NV classes are determined by the amplitude and orientation of the external magnetic field.
Using the same magnetic field alignment, Fig. \ref{100_VS_1x4} (b)-ii) and iii) show the PL and the $1/ T_1^{\rm dd}$ decay rates from the NV center ensemble as a function of the amplitude of the magnetic field. 
We observe that the 4~\% increase in PL as the magnetic field increases is indeed correlated with a drop in the spin decay rate from 1600~s$^{-1}$ to 80~s$^{-1}$. This result can again be explained by the decreasing number of resonant spins as the zero magnetic field is increased. We also confirmed that the half width of the dip ($\approx 10$~G) is consistent with a fluctuator model taking into account the reduction of the two spin resonances overlap as the $B$ field is increased (see SM sec. IV C).
%Note that the decrease in PL for $B>40\ \rm G$ is related to state mixing by the transverse magnetic field, so it is not associated with a modification of the spin lifetime.

Fig. \ref{100_VS_1x4} (c)-i) instead shows an ODMR where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis. Fig. \ref{100_VS_1x4} (c)-ii) and (c)-iii) show the change in the PL and in $1/T_1^{\rm dd}$ as a function of a magnetic field that is aligned in this direction. Surprisingly, we can still observe an increase in the spin decay rate and a corresponding drop in the PL when the magnetic field tends to zero, although all classes are always resonant (see SM, sec. IV-E).
We also note that the PL and $1/T_1^{\rm dd}$ change is also much sharper in this scenario.
%One of the main results of this paper is the observation of spin decay in this second scenario and the study on its impact on zero-field magnetometry. 
%Note that the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20$G is related to dipolar interaction with NV centers that have a $^{13}$C as a first neighbor \cite{pellet2021optical}.

When $\gamma | \mathbf{B} |\gg 1/T_2^*$, the only resonant terms in the dipole-dipole coupling between NV centers are the flip-flop terms $~\ket{0,\pm 1}\bra{\pm 1,0}$ (see SM sec. V).
In zero magnetic field however, other resonant mechanisms of the dipolar Hamiltonian have to be taken into consideration, which may elucidate the above result.
It was shown in  \cite{mittiga2018imaging} that local electric fields coming from $P_1^+$ or NV$^-$ centers are responsible for ODMR profiles at zero magnetic field. %Magnetic field noise comes in to second order in zero field. 
In order to estimate their contributions in the spin relaxation, we perform numerical simulations where we add the following Hamiltonian for the electric field dependent NV$^-$ ground state: 
\begin{equation}
\label{NV Hamiltonian}
\mathcal{H}_{\rm elec}= d_\perp \left[ E_x(\hat S_y^2-\hat S_x^2) + E_y(\hat S_x \hat S_y+\hat S_y \hat S_x) \right].
\end{equation}
$E_{x,y}$ are the projections of local electric fields on the NV axes and $d_\perp=17~$Hz $\cdot$ cm/V is the transverse electric field susceptibility. 
Under local electric fields with orientations given by the angle $\phi_E={\rm tan}(E_x/E_y)$, the eigenstates of the total hamiltonian of the NV spins in the ground state are $\ket{0}$ and $\ket{\pm}=\frac{1}{\sqrt{2}}(\ket{+1}\pm e^{-i\phi_E}\ket{-1})$.
Computing the influence of the flip-flop terms $\ket{\pm,0} \bra{0,\pm} $ in the fluctuator model shows that $1/T_1^{\rm dd}$ indeed increases at zero field, even after averaging over all possible angles $\phi_E$ (see SM, sec. V-E).  Another mechanism at stake when considering dipolar interactions between spin-1 systems at low fields are double-flip processes. 
These processes correspond to the terms $\ket{+1,0} \bra{0,-1}$ in the dipolar interaction, giving rise to an exchange of two units of spin-angular momentum when the $\ket{\pm 1}$ states are resonant. Simulations (see SM, sec V-F) suggest that double-flips contribute significantly to the spin relaxation at low fields.
%These processes can become significant when the states $\ket{m_s=\pm 1}$ are degenerate and are thus likely to explain part of the zero-field dip in Fig. 3-c)-ii).

\begin{figure}
\includegraphics[width=0.45\textwidth]{fig_transverse_field_V3.pdf}
\caption{(a) ODMR spectrum for $B_\perp=20\ \rm G$. The two transition frequencies of the class that is orthogonal to $B$ are denoted $\nu_+$ and $\nu_-$. (b) Measurement of $\nu_+$ and $\nu_-$ through ODMR as a function of $B_\perp$. (c) Measurement of $1/T_1^{\rm dd}$ for a single NV class as a function of $B_\perp$. The red dashed line is the value reached for high transverse field. The green dashed line correspond to the value found for the same sample and NV class, under a longitudinal magnetic field. Error bars are depicted by the shaded pink region. The detuning $\Delta \nu = \nu_+-\nu-$ is indicated on the top $x$-axis. Vertical black dashed line indicates the separation between regions A and B (see main text).}
\label{B_transverse}
\end{figure}

%In order to experimentally discriminate the contribution from the electric field and the double flips, one solution is to apply an electric field in order to lift the degeneracy between the $\ket{+}$ and $\ket{-}$ states. We estimated that, in our samples, we would need a field of $\sim 10^6\ \rm{V/cm}$, which is technically very challenging. 
To quantify these two processes experimentally, we apply a purely transverse magnetic field on one of the NV class. Indeed, when $B_{\perp} \leq 150\ \rm G$, the eigenstates of the spin Hamiltonian are close to $\ket{0}$, $\ket{+}$ and $\ket{-}$ to $\approx 2\%$ (see SM sec. IV-F). We can thus use the transverse field to emulate the role of the electric field, a property that has previously been used to increase the electric field sensing ability of NV centers \cite{dolde2011electric,qiu2022nanoscale}.
Fig.~\ref{B_transverse} (a) shows an ODMR spectrum where $B_{\perp}=20\ \rm G$ on sample HPHT-150-2. The central two lines correspond to the $\ket{0}\to \ket{-}$ and $\ket{0}\to \ket{+}$ transitions for the class that is orthogonal to $\bm B$.
%the other features corresponding to the transitions of the other classes. 
%We chose the initial value of 20 G so that the class of interest is detuned far enough from the other classes. We can therefore safely neglect inter-class flip-flops. 
Fig. \ref{B_transverse}-(c) shows the measurement of the decay rate $1/T_1^{\rm dd}$ for that class, as a function of $B_{\perp}$. Two regions can be observed on this graph: in region A the decay rate decreases with $B_{\perp}$, while in region B it stabilizes to a value $1/T_1^{\rm dd}=60\pm 5\ \rm{s}^{-1}$. We also indicated the value $1/T_1^{\rm dd}=25\pm 5\ \rm{s}^{-1}$ found for the same class but employing a longitudinal magnetic field. % instead of transverse to it.

\begin{figure}[ht]
\includegraphics[width=0.45\textwidth]{fig_magneto}
\caption{Low-field magnetometry protocol. (a) Demodulated PL as a function of an externally applied magnetic field. (b) External magnetic field read-out. (c) Histogram of the measurement in Fig. (b) fitted by gaussians of standard deviation $\sigma=1.5\ \mu \rm T$. (d) Measured sensitivity as function of the angle between the external magnetic field and the [100] crystalline axis.}
\label{magneto}
\end{figure}

To understand these results, we plot the detuning $\Delta \nu$ between the states $\ket{+}$ and $\ket{-}$ as a function of $B_{\perp}$ in Fig. \ref{B_transverse} (b). When $\delta \nu > 1/T_2^*$, these states do not overlap any longer, so we can interpret our results as follows: double-flips are efficient in region A, up to a point where state overlap is negligible. In region B, only the flip-flops in the basis $\ket{\pm}$ play a role. We note that the decay rate in the $\{ \ket{0}, \ket{\pm} \}$ basis is more than a factor of two larger than the decay rate coming from intra-class flip-flops in the $\{\ket{0}, \ket{\pm 1} \}$ basis. This effect is corroborated by our model detailed in SM sec. V. Importantly, in region A, the double-flips processes give a maximum decay rate that is $\sim 5$ times greater than the decay caused by the change of basis alone. From these results and according to our model, the double-flips are thus the dominant cause of zero-field depolarization observed in Fig. \ref{100_VS_1x4} (c)-iii).

Our observations have important implications for magnetometry with NV ensembles.
DC microwave-free magnetometry has already been performed using either NV-NV cross-relaxations \citep{akhmedzhanov_microwave-free_2017,akhmedzhanov_magnetometry_2019} or level anti-crossing \citep{Wickenbrock, zheng2017level, zheng_microwave-free_2020}. Here we propose a similar protocol but using the spin depolarization at zero-field. 
To do so, we use a lock-in detection and add a magnetic field modulation at $\sim 1\ \rm kHz$ with an amplitude $\sim 10\ \rm G$ through the same electromagnet. Fig. \ref{magneto}-(a) shows the demodulated PL while a DC magnetic field is scanned in an arbitrary direction. Here, we use the sample HPHT-15-1 with a laser power $\sim 1\ \rm mW$. The optical power in the collected PL is $\sim 1\ \mu\rm W$.  We can see a sharp linear slope in low field $\abs{B} < 5\ \rm G$. Once calibrated, the slope provides a 1D magnetometer, which could be extended to 3D with a set of 3 coils or 3 electromagnets, as in \cite{zheng_microwave-free_2020}. In order to assess the sensitivity of the measurement, we alternate a small DC field of $\approx 40\ \mu\rm{T}$ every few seconds and take a histogram of the measured fields (with a total duration of $\approx $50 s), as shown in Fig. \ref{magneto}-(b) and (c). The histogram is well fitted by gaussians of standard deviation $\sigma=1.5\ \mu \rm T$. The measurement was performed here with an output low-pass filter of time constant $\tau=3\ \rm ms$, which gives us a DC sensitivity $\eta=\sigma \sqrt{\tau}=82\ \rm{nT}/\sqrt{\rm Hz}$. This value corresponds to $\eta/\sqrt{V}\approx 4.7 \ \mu \rm{T}/\mu m^{3/2}/\sqrt{\rm Hz}$ when normalized to the volume. Importantly, this measurement is consistent with the experimentally found $\sim 5\ \mu \rm{T}/\sqrt{\rm Hz}$ sensitivity obtained with sample HPHT-1-1 (see samples details in SM sec. II).

Our results implies bright perspectives for high-sensitivity magnetometry using NV centers in a millimeter size volumes. Furthermore, compared to previously employed protocols, the sensitivity does not crucially depend on crystalline orientation, making this magnetometer principle operational with diamond powders or polycrystalline samples. Another significant advantage is that this measurement is insensitive to thermal fluctuations and strain inhomogeneities because dipolar-relaxation at zero-field does not depend on the zero-field splitting. In other typical NV magnetometers however, the zero-field splitting dependence with temperature and strain can give systematic errors \cite{Bauch, Barry}.

We now evaluate the relative role of the three causes of spin depolarization that assist the operation of the magnetometer. To do so, we measure its sensitivity while changing the angle of the magnetic field.  The results are shown in Fig. \ref{magneto} (d). We observe only a $\sim$ 10\% improvement in the sensitivity as we leave the [100] region.
When $\bm B \parallel [100]$ only the double flips and the electric field cause a depolarization, whereas the three effects are at play in all the other orientations. It should be noted that this observation is sample dependent. Other samples, including those from the same batch, have shown a higher orientation dependence, corresponding to a lower contribution from local electric fields and double-flip processes.
Interestingly, the double-flips and electric field effects are here the dominant factors in the sensitivity of this protocol. 
%It may seem surprising at first that the effects with a lower contribution to the PL contrast have a higher effect on this PL-based protocol. The reason why this is the case is that the slope of the change of contrast with respect to the magnetic field is larger when depolarization is caused by local electric fields and double-flips processes than solely by flip-flops processes.

As a conclusion, we identified three mechanisms causing spin depolarization in zero field for dense ensemble of NV$^-$ centers, all related to an increase in the dipole-dipole induced CR between the spins of NV centers. The lift in degeneracy of the spin state of different NV classes was found to be the main cause of zero-field depolarization, followed by double-flip processes and then the electric field induced mixing. We have employed CR for microwave and orientation-free DC-magnetometry and demonstrated a sensitivity below $100\ \rm{nT}/\sqrt{\rm Hz}$ for a single $15\ \mu \rm m$ diameter commercially available diamond and show that that double-flips and local electric fields play a critical role. 
We believe that our results will also be important for microwave-based low field magnetometry \cite{Vetter_LFM, WangRB} and for understanding many-body phenomena \citep{kucsko2018critical, ChoiNat, ZuYao, dwyer2021probing} and spin-mechanical effects \cite{pellet2021magnetic} with strongly coupled spins under purely transverse or small magnetic fields.

\begin{acknowledgments}

We would like to acknowledge support from Alexandre Tallaire and Jocelyn Achard as well as 
SIRTEQ for funding.

\end{acknowledgments}
\bibliographystyle{apsrev4-2}
\bibliography{CR}{}

\end{document}

### 2022-07-24 Mes papiers/Low field CR/V4/SI/Low_field_CR_SI.tex
tata

--- 2022-07-12 Mes papiers/Low field CR/V3/SI/Low_field_CR_SI.tex
@@
 
-The $T_1$ profile that we observe with our samples is neither purely exponential, nor purely stretched exponential, which is to be expected when the dipole-dipole induced spin decay is comparable to the phonon induced decay.
+%The $T_1$ profile that we observe with our samples is neither purely exponential, nor purely stretched exponential, which is to be expected when the dipole-dipole induced spin decay is comparable to the phonon induced decay.
+
+Fig. 2-c) of the main text shows the sequence employed for measuring $T_1^{\rm dd}$. 
+It consists in a pump-probe measurement where the spins are first polarized in the $\ket{0}$ state by a green laser, and read-out optically after a variable dark time $\tau$.
+In highly doped samples, this sequence often results in artifacts, mostly due to charge state transfer in the dark \citep{giri_coupled_2018, giri_selective_2019, choi_depolarization_2017}.
+It is therefore convenient to repeat the sequence with an additional $\pi$ pulse on one of the eight NV spin-resonances right before the spin read-out to prepare the remaining $\ket{0}$ polarization into a darker $\ket{+1}$ or $\ket{-1}$ state \citep{jarmola_temperature-_2012,mrozek_longitudinal_2015,choi_depolarization_2017}.
+By subtracting the result of the two sequences, we select only the spin-dependent part of the signal, with the added benefit of being able to select a specific class of NV centers.
 
 Fig.
-\ref{T1_fits} shows the result of a lifetime measurement, measured with the subtraction protocol described in the main text, for zero and non-zero magnetic field.
-The result is fitted with exponential and stretched-exponential (with a stretch factor $\beta=0.5$) functions.
-We can see that for the shorter lifetime (at zero field), the measurement follows more closely the stretched-exponential profile, while the longer lifetime is closer to the exponential one (except at very short times).
+\ref{T1_fits} shows the result of a lifetime measurement for the two scenarios presented in main text, namely $B=0$ and $B=50\ \rm G$ in a direction for which every classes are split, on sample HPHT-150-1.
+In main text, we fitted these results with the expression
+\begin{equation}
+S(\tau)=A \exp(-\sqrt{\frac{\tau}{T_1^{\rm dd}}}-\frac{\tau}{T_1^{\rm ph}}),
+\label{T_1 formula}
+\end{equation}
+in order to extract the $T_1$ part associated with the dipole-dipole coupling, as was previously done to study the cross-relaxation between NV and substitutional nitrogen (P1) center \citep{Hall}.
 
-In order to directly compare the results from different magnetic fields, we therefore fit each $T_1$ measurement using both a stretched and exponential decay : $S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}})$.
-To add consistency to the measurement of $T_1^{\rm dd}$, we fix $T_1^{\rm ph}$ for each sample, leaving $T_1^{\rm dd}$ and the signal amplitude as the only free parameters in the fitting procedure.
-The value chosen for samples HPHT-150-1 and HPHT-150-2 was $T_1^{\rm ph}=3.62\ \rm ms$.
+In Fig.
+\ref{T1_fits}, we show what happens when we fit the $T_1$ measurements with either purely exponential or purely stretched exponential fits.
+We can see that for $B=0$, which corresponds to the regime where dipole-dipole relaxation is strong, the experimental data is closely fitted by the stretched exponential profile, whereas for $B=50\ \rm G$, the experimental data follows more closely the exponential profile.
+In order to describe both regime simultaneously, we have to take both $T_1^{\rm dd}$ and $T_1^{\rm ph}$ into account in our analysis.
+
+Since we are only interested in the stretched exponential part of the lifetime decay, and we are assuming that the phonon-limited exponential decay does not depend on the external magnetic field, we decided to fix the value $T_1^{\rm ph}$ for each sample, in order to reduce the number of free parameters in eq.
+\ref{T_1 formula}.
+The value chosen for all $T_1$ measurements on samples HPHT-150-1 and HPHT-150-2 was $T_1^{\rm ph}=3.62\ \rm ms$.
+
+Another possibility to fit our $T_1$ measurements which yields satisfactory fits is to use an arbitrary stretched factor in the fitting function : 
+\begin{equation}
+S(\tau)=A \exp (-\left(\frac{\tau}{T_1}\right)^\beta ).
+\end{equation}
  Fig.
-1 of the main text shows the result of fitting the same two measurement with this formula.
-
-An other possibility would be to use an arbitrary stretched factor in the fitting function : $S(\tau)=A \exp (-(\frac{\tau}{T_1^{\rm ph}})^\beta )$.
-Fig.
-\ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field, which confirms that the $T_1$ profile gets closer to a pure stretched-exponential in zero magnetic field.
+\ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field on sample HPHT-150-1, which confirms that the $T_1$ profile gets closer to a purely stretched exponential in zero magnetic field.
 

### 2022-07-24 Mes papiers/Low field CR/V4/SI/Figures_SI/Champ transverse ‚tats et energies.py

@@ -1,0 +1,103 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+
+def plot_NRJ():
+	plt.figure(num=1,figsize=(3,2),dpi=80)
+	plt.xticks(fontsize=11)
+	plt.yticks(fontsize=12)
+
+	n=200
+	levels=np.zeros((n,3))
+	Bs=np.linspace(0,200,n)
+	DeltaE=np.zeros(n)
+	DeltaVect=np.zeros(n)
+	DeltaVect2=np.zeros(n)
+
+
+
+
+	for i in range(n):
+		theta=90*pi/180
+		B=Bs[i]
+		H=NVHamiltonian(B=[B*sin(theta),0,B*cos(theta)],c=5)
+		E=H.egval()
+		V=H.egvect()
+		levels[i,:]=E
+		DeltaE[i]=E[2]-E[1]
+		DeltaVect[i]=(V[2].dot([1/sqrt(2),0,1/sqrt(2)]))**2
+		DeltaVect2[i]=(V[0].dot([0,1,0]))**2
+		
+		# DeltaVect[i]=abs(V[2].dot([0,1,0]))
+		# DeltaVect[i]=abs(V[0].dot([0,1,0]))
+
+
+	# f,(ax,ax2)=plt.subplots(2,1,sharex=True)
+
+	# f.set_size_inches((3,2))
+
+	# ax2.plot(Bs,levels[:,0],color=color(0))
+	# ax.plot(Bs,levels[:,1],color=color(0))
+	# ax.plot(Bs,levels[:,2],color=color(0))
+
+	# ax.spines['bottom'].set_visible(False)
+	# ax2.spines['top'].set_visible(False)
+	# ax.xaxis.tick_top()
+	# ax.tick_params(labeltop=False)  # don't put tick labels at the top
+	# ax2.xaxis.tick_bottom()
+
+	# plt.xlabel('Magnetic Field (G)',fontsize=15)
+
+
+
+	ax1=plt.gca()
+	ax2=ax1.twinx()
+	ax1.plot(Bs,DeltaE,'--',lw=2)
+	ax1.set_ylabel(r'$E_e - E_d$ (MHz)')
+	ax1.set_ylim(0,max(DeltaE)*1.05)
[…]

### 2022-07-26 Manuscrit Thèse/chapter 2/Chapter 2.tex
tata

@@
 
-While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, there are several differences between the two protocols :
+While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, there are several differences between the two protocols:
 

### 2022-07-26 Manuscrit Thèse/chapter 2/AA_simu contraste PL.py

@@ -1,0 +1,85 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+def alpha_func(B,c=1):
+	alpha=np.zeros((7,7))
+	alpha[6,6]=1
+	H=NVHamiltonian(B,c=c,E=0,D=2870,order='ascending')
+	egvect=np.array(H.egvect())
+	alpha[0:3,0:3]=egvect
+	H=NVHamiltonian(B,c=c,E=0,D=1420,order='ascending')
+	egvect=np.array(H.egvect())
+	alpha[3:6,3:6]=egvect
+	return (alpha)
+
+def k0_func(beta=1):
+	k0r=65
+	k047=8
+	k057=55
+	k071=1
+	k072=1
+	k0=np.zeros((7,7))
+	k0[4-1,1-1]=k0r
+	k0[1-1,4-1]=beta*k0r
+	k0[5-1,2-1]=k0r
+	k0[2-1,5-1]=beta*k0r
+	k0[6-1,3-1]=k0r
+	k0[3-1,6-1]=beta*k0r
+	k0[4-1,7-1]=k047
+	k0[5-1,7-1]=k057
+	k0[6-1,7-1]=k057
+	k0[7-1,1-1]=k071
+	k0[7-1,2-1]=k072
+	k0[7-1,3-1]=k072
+
+	return k0
+
+
+
+
+def k_func(B,c=1,beta=1):
+	alpha=alpha_func(B,c=c)
+	k0=k0_func(beta=beta)
+	k=np.zeros((7,7))
+	for i in range(7) :
+		for j in range(7) :
+			for p in range(7) :
+				for q in range(7) :
+					k[i,j]+=(alpha[i,p]**2)*(alpha[j,q]**2)*k0[p,q]
+
+	return k
+
+
+def M_func(B,c=1,beta=1): #La matrice des 7 equation linéR. J'en prends que 6 du coup vu qu'elles sont couplées
+	k=k_func(B,c,beta)
+	M=np.zeros((7,7))
+	M[6,:]=np.ones(7) #Correspond a n1+n2+...=1
+	for i in range(6):
+		for j in range(7):
+			if i==j:
+				M[i,j]=-sum(k[i,:])
+			else :
+				M[i,j]=k[j,i]
+
+
+	return(M)
[…]

### 2022-07-27 Manuscrit Thèse/chapter 2/Chapter 2.tex
tata

@@
 Detection \textit{via} cross-relaxation	with NV centers fall under the more general technique of NV relaxometry, where we are effectively measuring a modification in the NV center spin lifetime $T_1$.
-Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross,holliday1989optical,armstrong2010nv, hall2016detection,wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
+Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross,holliday1989optical,armstrong2010nv, epstein2005anisotropic,  hall2016detection ,wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
 
@@
 
+\subsection{The [100] and [111] magnetic field orientations}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_NV}
+\caption{Label}
+\label{121 vs 22 NV}
+\end{figure}
+
+Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\mathbf{B} \parallel [100]$ or 
+$\mathbf{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes as defined in chapter 1.
+When $\mathbf{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\mathbf{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.
+
+These changes in the field orientation result in massive changes in the NV behavior, manly due to the transverse magnetic field. The first
+
 \section{Attribution of the VH$^-$ and War1 lines}

### 2022-07-27 Manuscrit Thèse/chapter 2/AA_simu contraste PL.py

@@ -3,6 +3,10 @@ sys.path.append('D:\\These Clément\\these\\python_clément')
 sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
+
+#REF : "Magnetic-field-dependent photodynamics of single NV defects in diamond: an application to qualitative all-optical magnetic imaging"
+# unité naturelle pour k et Omega : µs-1 ; Beta est une fraction du temps de vie radiatif (beta=1 saturation de la transition)
+
 def alpha_func(B,c=1):
 	alpha=np.zeros((7,7))
 	alpha[6,6]=1
@@ -39,7 +43,7 @@ def k0_func(beta=1):
 
 
 
-def k_func(B,c=1,beta=1):
+def k_func(B,c=1,beta=1,Omega12=0):
 	alpha=alpha_func(B,c=c)
 	k0=k0_func(beta=beta)
 	k=np.zeros((7,7))
@@ -49,11 +53,14 @@ def k_func(B,c=1,beta=1):
 				for q in range(7) :
 					k[i,j]+=(alpha[i,p]**2)*(alpha[j,q]**2)*k0[p,q]
 
+	k[1-1,2-1]=Omega12
+	k[2-1,1-1]=Omega12
 	return k
 
+# print_matrix(k_func(B=[0,0,100],c=1,beta=0.5,Omega12=10))
 
-def M_func(B,c=1,beta=1): #La matrice des 7 equation linéR. J'en prends que 6 du coup vu qu'elles sont couplées
-	k=k_func(B,c,beta)
+def M_func(B,c=1,beta=1,Omega12=0): #La matrice des 7 equation linéR. J'en prends que 6 du coup vu qu'elles sont couplées
+	k=k_func(B,c,beta,Omega12)
 	M=np.zeros((7,7))
 	M[6,:]=np.ones(7) #Correspond a n1+n2+...=1
 	for i in range(6):
@@ -68,18 +75,57 @@ def M_func(B,c=1,beta=1): #La matrice des 7 equation linéR. J'en prends que 6 d
 # print_matrix(M_func(B=[0,0,100],c=5,beta=0.5))
 
 
-def pop_func(B,c=1,beta=1):
-	M=M_func(B,c,beta)
[…]

### 2022-07-28 Manuscrit Thèse/chapter 2/Chapter 2.tex
tata

@@
 where $\Omega=\frac{\mel{i;j}{\mathcal{H}_{\rm dd}}{j;i}}{\hbar}$ is the matrix element of the dipole-dipole Hamiltonian that couples the initial and final states, $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, and $\delta \nu$ is the detuning between the central frequencies of the NV center and spin 2.
-\section{Dark spins in diamond}
+\section{The dark spins in diamond}
 
@@
 
-\section{Choice of the $\mathbf{B}$ scan orientation}
+\section{Choice of the magnetic field orientation}
 
@@
 
-These changes in the field orientation result in massive changes in the NV behavior, manly due to the transverse magnetic field. The first
+These changes in the field orientation result in massive changes in the NV behavior, manly due to the transverse magnetic field: Fig.
+\ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the ground state spin Hamiltonian (REF chapter 1) - or to be more precise the transitions between the ground state of the spin Hamiltonian and the two excited states - for any of the four classes of NV centers when $\mathbf{B} \parallel [100]$.
+Notably, there is no transitions frequencies below 2500 MHz.
+
+On the other hand, Fig.
+\ref{121 vs 22 NV}-b) shows the transition frequencies when $\mathbf{B} \parallel [111]$, for both the class aligned and the three class misaligned.
+The three classes behave similarly than the four classes in the $[100]$ case, but the one class aligned with the magnetic field can get its $\ket{0}\to\ket{-1}$ transition arbitrarily small, as long as the magnetic field is perfectly aligned with the NV axis.
+
+\subsection{CR condition between NV$^-$ and P1}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_P1}
+\caption{Label}
+\label{121 vs 22 P1}
+\end{figure}
+
+P1 centers, the most abundant electronic spin in the diamonds commonly used, have a $1/2$ electronic spin.
+They do have a small magnetic field anisotropy coming from the hyper-fine coupling to the nitrogen nucleus ($\sim 100\ \rm MHz$), but this dependency on the field orientation is small compared to the NV center's zero field splitting $D=2870\ \rm MHz$.
+We can therefore, to the first order, neglect the hyper-fine coupling and consider P1 centers as ideal, isotropic, spin $1/2$.
+
+Fig.
+\ref{121 vs 22 P1} shows the transition frequencies for the NV center and such a spin $1/2$, as a function of a magnetic field parallel to [100] or [111].
+When $\mathbf{B} \parallel [100]$, there will never be a co-resonance between the NV and P1 transitions, meaning that no NV-P1 cross-relaxation can be observed when B is scanned in this direction.
+When $\mathbf{B} \parallel [111]$ however, the P1 $\ket{-1/2} \to \ket{+1/2}$ transition matches in energy the $\ket{0}\to\ket{-1}$ NV transition for B=512 G.
+This means that when B reaches $\approx 512\ \rm G$, some of the polarization of the NV centers will be transferred to the P1 centers, which will result in a drop in the NV PL.
+
+\subsection{CR condition between NV$^-$ and VH$^-$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_VH}
+\caption{Label}
+\label{121 vs 22 VH}
+\end{figure}
+
+VH$^-$, a much less abundant electronic spin than P1 center, has an electronic spin 1, with a spin Hamiltoninan very similar to that of the NV center.
+In particular it has the same symmetries being also a $C_{3v}$ defect, with a slightly smaller ZFS at $D_{VH} \approx 2700\ \rm MHz$ compared to $D_{NV}=2870\ \rm MHz$.
+
+Fig.
+\ref{121 vs 22 VH} shows the transition frequencies for the NV$^-$ and VH$^-$ ground state spin.
+Contrary to the P1 case, there is a co-resonance when $\mathbf{B} \parallel [100]$ for $B\approx 55\ \rm G$.
+When $\mathbf{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G.
+A scan along the [111] in this case means that the peak signal is $\sim 6$ times weaker than in the [100] case, and spread over multiple features which might overlap between themselves or with other existing PL features.
+
+The War1 spin mentioned previously is also an electronic spin 1 with (pseudo)-$C_{3v}$ symmetry and $D_{War1} \approx 2470\ \rm MHz$.
+This result in a similar behavior than VH$^-$ when it comes to the CR condition with NV centers.
+
+\subsection{CR contrast and transverse magnetic field}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/contrast_OMDR_111_Vs_100}
+\caption{Label}
+\label{121 vs 22 contrast}
+\end{figure}
+
+In order to optically detect CR between an NV center and a dark spin, two conditions must be met: the NV center must be polarized in order to have a population transfer between the NV and the dark spin, and there must be a difference in brightness between the two NV levels involved in order to have an optical signal of the CR.
+
+These two conditions are in fact the same needed to observe an ODMR spectrum (REF chapter 1), and while the NV centers verifies both criteria in presence of a purely longitudinal magnetic field, this is not the case anymore in presence of a transverse magnetic field (REF chapter 1).
+
+Fig. \ref{121 vs 22 contrast} shows the simulated relative contrast of an ODMR spectrum, as computed from [PUTAIN DE CITATION DE MORT]
+
+\section{Dark spin spectroscopy with NV centers}
 

### 2022-07-29 Manuscrit Thèse/chapter 2/Chapter 2.tex
toto

@@
-\documentclass[a4paper]{report}
+\documentclass[a4paper,11pt]{report}
 \usepackage[]{amsmath}
@@
 
+\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
 \usepackage{subcaption} % package pour faire des subfigures
@@
 \section{Flip-Flops, double flips and cross-relaxation}
-
+\label{Sec_CR}
 \begin{figure}
@@
 Detection \textit{via} cross-relaxation	with NV centers fall under the more general technique of NV relaxometry, where we are effectively measuring a modification in the NV center spin lifetime $T_1$.
-Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross,holliday1989optical,armstrong2010nv, epstein2005anisotropic,  hall2016detection ,wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
+Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
 
@@
 
+\subsection{CR condition between NV$^-$ and and $^{13}$C-NV$^-$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_13CNV}
+\caption{Label}
+\label{121 vs 22 13C-NV}
+\end{figure}
+
+I refer by $^{13}$C-NV$^-$ to a complex formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope, which is illustrated in Fig.
+\ref{121 vs 22 13C-NV}-a).
+
+The $^{13}$C-NV$^-$ complex still behaves as a "normal" NV center, except with shifted transition lines.
+In particular its electronic spin is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states.
+This is confirmed by the presence of sideband in ODMR spectra around the main NV lines \citep{simanovskaia2013sidebands}.
+
+Since both NV$^-$ and $^{13}$C-NV$^-$ are \textit{a priori} equally polarized, then following the argument in \ref{Sec_CR} there should be no CR between the two spin families.
+The reason why those transitions are still in some case visible, as will be shown below, and why in fact any NV-NV co-resonance gives a visible CR signal will be the object of the next chapter.
+
+The transitions of the $^{13}$C-NV$^-$ complex are given by diagonalizing the full spin Hamiltonian:
+\begin{equation*}
+\mathcal{H}=\mathcal{H}_{NV}+\mathcal{H}_{^{13}C}+\mathcal{H}_{HF},
+\end{equation*}
+where $\mathcal{H}_{NV}$ is the NV$^-$ spin Hamiltonian,
+$\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian: $$\mathcal{H}_{HF}= \hat{\mathbf{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\mathbf{I}}_C.$$
+
+When the $^{13}$C is the direct neighbor of the vacancy as represented in Fig.
+\ref{121 vs 22 13C-NV}-a) - this is also referred as a first-shell $^{13}$C - the hyper-fine tensor $\mathcal{A}$ can be written \citep{simanovskaia2013sidebands}:
+$$ \mathcal{A} = \begin{pmatrix}
+\mathcal{A}_{xx} & 0 & \mathcal{A}_{xz} \\ 0 & \mathcal{A}_{yy} & 0 \\ \mathcal{A}_{zx} & 0 & \mathcal{A}_{zz}
+\end{pmatrix},$$
+where $\mathcal{A}_{xx}=190$ MHz, $\mathcal{A}_{yy}=120$ MHz, $\mathcal{A}_{zz}=129$ MHz, and  $\mathcal{A}_{xz}=\mathcal{A}_{zx}=-25.0$ MHz.
+
+Fig.
+\ref{121 vs 22 13C-NV}-b) and c) show the simulated transition frequencies for a $^{13}$C-NV$^-$ complex when B is scanned along the [100] or [111] axis, as well as the "normal" NV transitions and the co-resonance between those two.
+When $\mathbf{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G.
+When $\mathbf{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
+This abundance of transitions for low (<100 G) magnetic field makes analyzing experimental data quickly intractable in the second case.
+
 \subsection{CR contrast and transverse magnetic field}
@@
 
-In order to optically detect CR between an NV center and a dark spin, two conditions must be met: the NV center must be polarized in order to have a population transfer between the NV and the dark spin, and there must be a difference in brightness between the two NV levels involved in order to have an optical signal of the CR.
+In order to optically detect CR between an NV center and a dark spin, two conditions must be met: the NV center must be polarized in order to have a population transfer between the NV and the dark spin, and there must be a difference in brightness between the two NV levels involved in order to have an optical signature of the CR.
 
-These two conditions are in fact the same needed to observe an ODMR spectrum (REF chapter 1), and while the NV centers verifies both criteria in presence of a purely longitudinal magnetic field, this is not the case anymore in presence of a transverse magnetic field (REF chapter 1).
+These two conditions are in fact the same needed to observe an ODMR spectrum (REF chapter 1), and while the NV centers verifies both criteria in presence of a purely longitudinal magnetic field, this is not always the case anymore in presence of a transverse magnetic field (REF chapter 1).
 
-Fig. \ref{121 vs 22 contrast} shows the simulated relative contrast of an ODMR spectrum, as computed from [PUTAIN DE CITATION DE MORT]
+Fig.
+\ref{121 vs 22 contrast} shows the simulated relative contrast of an ODMR spectrum, as computed from \citep{tetienne2012magnetic}, for a magnetic field aligned along [100] or [111].
+When the NV center is aligned with the magnetic field, the constrats remains equals to 1 regardless of the magnetic field because there is no transverse field.
+When the NV is not aligned however, either in the [111] or [100] case, the ODMR contrast quickly drops down for B$>100\ \rm G$.
+
+This tells us that we should not expect to observe a CR signal for B greater than a few hundreds G when B is scanned along the [100] axis.
+We could however observe CR for arbitrarily large magnetic field when scanned along the [111] axis, as long as the field is properly aligned with the NV axis.
 
 \section{Dark spin spectroscopy with NV centers}
+
+\subsection{Detection of P1 centers}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\textwidth]{Figures/NV_P1_libreAcces}
+\caption{Taken from \citep{armstrong2010nv}Label}
+\label{CR P1 exp}
+\end{figure}
+
+Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross} had been done on the detection of dark spins through CR with NV centers, and to our knowledge, all this studies where done on the detection of P1 centers.
+Following the discussion in the last section, these studies were therefore all done with $\mathbf{B} \parallel [111]$.
+
+An example of one of this study (\citep{armstrong2010nv}) is shown in Fig.
+\ref{CR P1 exp}.
+It shows the evolution of the NV PL with respect to a magnetic field scanned along the [111] crystalline axis.
+There are several information to gather from this graph:
+
+First, there is the slowly decaying PL envelope, which stabilizes to a plateau for $B\sim 1000\ \rm G$.
+This change in PL is caused by the transverse magnetic field acting on the three NV classes that are not aligned with the magnetic field.
+For $B \gtrsim 1000\ \rm G$, these three classes are fully depolarized which explains the plateau observed.
+
+Second, there are the three salient features denoted Ns-NV, NV-NV and NV-LAC.
+All these features come from cross-relaxations or level anti-crossing (LAC) from the NV class aligned with the magnetic field.
+
+The first feature (Ns-NV) for $B\approx 512 \ \rm G$ is the signature of the CR between NV and P1 centers.
+Its relatively complicated shape comes from the hyper-fine coupling between the P1 electronic spin and the $^14$N nucleus \citep{lazda2021cross}.
+Excited state level anti-crossing (ESLAC) also occurs in this region, but its optical signature, if it exists, is expected to be much weaker than the NV-P1 CR \citep{zheng2017level}.
+
+The second feature (NV-NV) for $B\approx 590 \ \rm G$ comes from the cross relaxation between NV centers aligned with the magnetic field and NV centers misaligned with the field (the co-resonance can be observed in Fig.
+\ref{121 vs 22 NV}-b)).
+This phenomenon will be further discussed in the next chapter.
+
+Finally the third feature (LAC) for $B\approx 1024 \ \rm G$ correspond to the level anti-crossing between the $\ket{0}$ and $\ket{-1}$ states of the NV center aligned with B.
+When the two levels get close enough in energy, the Hamiltonian off-diagonal terms (coming from strain local electric field or residual transverse magnetic field) will mix the $\ket{0}$ and $\ket{-1}$ states and prevent an efficient polarization by the green laser.
+This result in a sharp drop in the PL that can be exploited to perform microwave-less magnetometry with NV centers \citep{wickenbrock2016microwave, zheng2017level, zheng2020microwave}.
+
+\subsection{Detection of VH$^-$ and War1 defects}
+
+During my PhD, I have studied the 
+Our work \citep{pellet2021optical}
 

### 2022-08-02 Manuscrit Thèse/Glossary.tex
tata

\documentclass[a4paper]{article}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois

\begin{document}
\section{Glossary}
\begin{itemize}
\item PL
\item ODMR
\item CR
\item P1/VH-/War1
\end{itemize}
\section{Samples}
\subsection{CVD}
\begin{itemize}
\item CVD-pink : CVD rose d'alex
\item CVD-B/C : cf article de Midrel
\end{itemize}
\subsection{HPHT}
\begin{itemize}
\item SBST-B/C : substrat des CVD-B/C
\end{itemize}
\end{document}

### 2022-08-02 Manuscrit Thèse/chapter 2/Chapter 2.tex
tata

@@
 
-\title{Appendix A : Dipole-dipole interaction between two spins in solid}
+\title{}
 \begin{document}
 \chapter{Detection of dark spins via cross-relaxations}
-In this chapter, we will see how...
+In this chapter, we will see how the interplay between the spin and the optical properties of the NV centers can be transferred to other spins in the crystal by coupling them to NV centers.
+More specifically, I will discuss the first optical detection of two spin species, VH$^-$ and War1, in a CVD diamond by the mean of cross-relaxations with NV centers.
 \section{Flip-Flops, double flips and cross-relaxation}
 \label{Sec_CR}
-\begin{figure}
-\centering
-\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
-\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
-\label{CR_shema}
-\end{figure}
+
+\subsection{Resonant interaction : flip-flops and double flips}
 
@@
 
+\subsection{Cross-relaxations}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
+\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
+\label{CR_shema}
+\end{figure}
+
 Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
@@
 
-If we make the assumption that the spin bath coupled to the NV centers is Markovian, meaning that the dynamics of the bath is not affected by its coupling to the NV centers, we can then compute the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ due to its coupling to Spin 2 \citep{hall2016detection}:
+For a single NV center coupled to spin 2, assuming that spin 2 is initially depolarized and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$, we can compute the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by spin 2 as \citep{hall2016detection}:
 \begin{equation}
@@
 \subsection{P1 centers and $^{13}$C}
-By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^13$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^12$C isotope.
+By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
 
@@
 \begin{equation}
-\Gamma_2^*(P1) \propto \Omega_{\rm NV-P1}=\frac{1}{\hbar}\sqrt{\sum_j \abs{\mel{NV=i;P1=j}{\mathcal{H}_{\rm dd}}{NV=i;P1=j}}^2}.
+\Gamma_2^*(\rm P1) \propto \Omega_{\rm NV-P1}=\frac{1}{\hbar}\sqrt{\sum_j \abs{\mel{NV=i;P1=j}{\mathcal{H}_{\rm dd}}{NV=i;P1=j}}^2}.
 \label{Gamma2 P1}
 \end{equation}
-Since the dipole-dipole coupling rate scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity.
+Since the dipole-dipole coupling rate scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity: $\Gamma_2^*(\rm P1) \propto [P1]$.
 In the case of P1 centers, this was verified experimentally by \citep{bauch2020decoherence} who found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
+The typical concentrations of P1 center go from [REF] for high purity CVD diamond, up to 500 ppm for type 1B HPHT or natural diamonds.
 
-$^{13}$C is much more abundant than P1 center ($\sim$ 10 000 ppm at natural abundance) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\, 10^3$ times smaller than that of the P1.
+$^{13}$C is much more abundant than P1 center ($\sim$ 10 000 ppm at natural abundance) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\cdot 10^3$ times smaller than that of the P1.
 This results in a contribution to $\Gamma_2^* \sim 10^3$ times smaller per impurity.
-It was found experimentally that the broadening caused by $^13$C is $\Gamma_2^*(^{13}C) \approx 16\ \rm Hz/ppm$ \citep{van1997dependences, barry2020sensitivity}.
-For natural abundance diamond, this means that $\Gamma_2^*(^{13}C) \approx 160\ \rm kHz$.
+It was found experimentally that the broadening caused by $^{13}$C is $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 16\ \rm Hz/ppm$ \citep{van1997dependences, barry2020sensitivity}.
+For natural abundance diamond, this means that $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 160\ \rm kHz$.
 
 \subsection{Other spin defects, VH$^-$, War1}
-
+\label{other defects}
 Other spin defects are also commonly found in diamond containing NV centers, such as charged vacancy or vacancy clusters \citep{hounsome2006origin}, other nitrogen related defects \citep{newton2007epr}, transition metals \citep{isoya1990fourier}, and especially for CVD grown diamond, hydrogen related defects \citep{hartland2014study}, since the CVD plasma contains up to 95 \% hydrogen.
@@
 No hyperfine structure could be discerned in the relatively broad EPR line, and the use of deuterium instead of $^1$H did not change the shape of the line.
+
+\subsection{EPR spectroscopy}
+Electron paramagnetic resonance (EPR), the electronic equivalent to NMR, is the the most common way to detect electronic spins in solids and liquids.
+The principle of its detection is the absorption or reflection of a microwave field resonant with a spin transition.
+
+EPR generally relies on the thermal polarization of the spins to observe an absorption.
+In order to increase the polarization, an important magnetic field is applied (typ.
+0.35 T or 3500 G) and a microwave cavity (typ.
+9500 MHz) is used to improve the coupling of the spins to the field.
+
+At room temperature, the thermal polarization of a transition at 9500 MHz is $\sim 0.002$.
+This low value, compared to $\sim 0.8$ for the NV centers, explains in part why EPR is way less sensitive than NV ODMR.
+The current noise floor for EPR is typically of $\sim 10^{11}$ spins for electronic spins in diamond \citep{mitchell2013x}.
 
@@
 A change in $\Gamma_1$ will therefore conduct to a new equilibrium with a different PL.
-For a small enough change $\delta \Gamma << \Gamma_1$, we can consider that we remain in the linear regime, hence why $\frac{\delta \rm PL}{\rm PL} \propto \frac{\delta \Gamma}{\Gamma_1}$.
 
@@
 In this case however, instead of scanning the time $\tau$, you fix its value to maximize the sensitivity in a change of $T_1$ (typically $\tau \approx T_1$).
-Again, assuming you are in the linear regime, you observe a change in signal proportional to the change in $\Gamma_1$ : $\frac{\delta \rm S}{\rm S} \propto \frac{\delta \Gamma}{\Gamma_1}$.
 
@@
 \begin{equation}
-\frac{1}{T_2^{\rm DEER}}=\frac{1}{T_2^{\rm echo}} + \delta \Gamma,
+\frac{1}{T_2^{\rm DEER}}=\frac{1}{T_2^{\rm echo}} + \delta \Gamma_2^*,
 \end{equation}
 
-where $\delta \Gamma$ is the contribution of the dark spin to the $T_2^*$ of the NV center through NV-dark spin dipole-dipole interaction.
+where $\delta \Gamma_2^*$ is the contribution of the dark spin to the $T_2^*$ of the NV center.
 
@@
 \subsection{The [100] and [111] magnetic field orientations}
+\label{sec simu}
 
@@
 \includegraphics[width=0.9\textwidth]{Figures/121_vs_22_NV}
-\caption{Label}
+\caption{Simulated transition frequencies of the $\ket{0} \to {-1}$ and $\ket{0} \to {+1}$ transitions for the 4 classes of NV centers as a function of the magnetic field.
+a) When the magnetic field is along the diamond [100] crystalline axis, b) along the [111] axis.
+A representation of the 4 classes of NV centers, labeled 1 to 4, and the orientation of the magnetic field in both cases is put on top}
 \label{121 vs 22 NV}
@@
 \includegraphics[width=0.9\textwidth]{Figures/121_vs_22_P1}
-\caption{Label}
+\caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}
 \label{121 vs 22 P1}
@@
 \includegraphics[width=0.9\textwidth]{Figures/121_vs_22_VH}
-\caption{Label}
+\caption{Simulation of the transition frequencies of NV$^-$ centers and VH$^-$ centers for a magnetic field a) along the [100] axis and b) along the [111] axis}
 \label{121 vs 22 VH}
@@
 \includegraphics[width=0.9\textwidth]{Figures/121_vs_22_13CNV}
-\caption{Label}
+\caption{a) Representation of the $^{13}$C-NV complex (right) compared to a "normal" NV center (left) b) and c) simulation of the transition frequencies of NV$^-$ centers and $^{13}$C-NV$^-$ complex for a magnetic field b) along the [100] axis and c) along the [111] axis}
 \label{121 vs 22 13C-NV}
@@
 
-I refer by $^{13}$C-NV$^-$ to a complex formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope, which is illustrated in Fig.
+I refer by $^{13}$C-NV$^-$ to a complex formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope.
+Such a complex is illustrated in Fig.
 \ref{121 vs 22 13C-NV}-a).
@@
 When $\mathbf{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
-This abundance of transitions for low (<100 G) magnetic field makes analyzing experimental data quickly intractable in the second case.
+This abundance of transitions for low ($<100\ \rm G$) magnetic field makes analyzing experimental data quickly intractable in the second case.
 
 \subsection{CR contrast and transverse magnetic field}
+\label{CR contrast}
 
@@
 \includegraphics[width=0.9\textwidth]{Figures/contrast_OMDR_111_Vs_100}
-\caption{Label}
+\caption{Simulated ODMR contrast on all four classes of NV centers with a magnetic field a) along the [100] axis and b) along the [111] axis}
 \label{121 vs 22 contrast}
@@
 \includegraphics[width=0.7\textwidth]{Figures/NV_P1_libreAcces}
-\caption{Taken from \citep{armstrong2010nv}Label}
+\caption{Change in the PL of an ensemble of NV centers as a magnetic field is scanned along the [111] crystalline axis. Taken from \citep{armstrong2010nv}}
 \label{CR P1 exp}
@@
 The first feature (Ns-NV) for $B\approx 512 \ \rm G$ is the signature of the CR between NV and P1 centers.
-Its relatively complicated shape comes from the hyper-fine coupling between the P1 electronic spin and the $^14$N nucleus \citep{lazda2021cross}.
+Its relatively complicated shape comes from the hyper-fine coupling between the P1 electronic spin and the $^{14}$N nucleus \citep{lazda2021cross}.
 Excited state level anti-crossing (ESLAC) also occurs in this region, but its optical signature, if it exists, is expected to be much weaker than the NV-P1 CR \citep{zheng2017level}.
@@
 
-\subsection{Detection of VH$^-$ and War1 defects}
+\subsection{Detection of VH$^-$, War1 and $^{13}$C-NV$^-$}
 
-During my PhD, I have studied the 
-Our work \citep{pellet2021optical}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/fig_VH_et_co}
+\caption{a) Change in PL (normalized) of the ensemble of NV centers in sample CVD-pink as a magnetic field is scanned along the [100] axis.
+The orange line is a fit to the slowly decreasing envelope b) Subtraction of the PL signal by the envelope fit as a function of the magnetic field.
+The green lines and red and purple regions correspond to the predicted CR regions, as detailed in main text}
+\label{CR VH exp}
+\end{figure}
 
-\section{Attribution of the VH$^-$ and War1 lines}
+Our work \citep{pellet2021optical} focuses on the optical detection of VH$^-$, War1 and $^{13}$C-NV$^-$ through NV CR.
+This study was the first to detect VH$^-$ and War1 defects optically and the first CR experiment on a CVD diamond.
+It was also the first,to our knowledge, to use a [100] aligned magnetic field.
+
+The diamond used for this study is the sample CVD-pink [REF appendix] whose fabrication was detailed in \citep{tallaire2020high}.
+It is a CVD diamond containing $[\rm NV ]\approx 4\ \rm ppm$ for $[\rm N ]\approx26\ \rm ppm$, and a natural abundance (1.1\%) of  $^{13}$C.
+Being a CVD diamond it is also likely to incorporate hydrogen defects, as discussed in \ref{other defects}.
+
+Fig.
+\ref{CR VH exp}-a) shows the change in PL with respect to a magnetic field scanned along the diamond [100] axis.
+Similarly to fig.
+\ref{CR P1 exp}, we can see a slowly decaying envelope of the PL due to the transverse magnetic field (we don't see the plateau here since the magnetic field sops at 150 G).
+We can also see three deviations to the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
+
+In order to better discern the three features, I have fitted the envelope by using a polynomial fit of order 4 on the smooth part of the data.
+I previously tried to use a fit based on the expected PL rate of the NV centers, by using the model in \citep{tetienne2012magnetic}, but the results were not satisfying enough.
+
+Fig.
+\ref{CR VH exp}-b) shows the subtraction of the PL data by the envelope fit.
+We can clearly see three dips in PL around the values 20 G, 56 G and 122 G.
+I then computed the expected value of the magnetic field for the expected CR line thanks to the simulations shown in sec.
+\ref{sec simu}.
+For the VH$^-$ and War1 lines, I took the ZFS $D$ values from \citep{cruddace2007magnetic} and included the error bars.
+For the $^{13}$C-NV$^-$ lines, the error bars given in \citep{simanovskaia2013sidebands} are much smaller and I only plotted the expected value.
+We can see that the dips in PL match relatively well the predicted CR lines of $^{13}$C-NV$^-$, VH$^-$ and War1 respectively.
+
+The experimental setup used to obtain this data is the standard confocal microscopy setup shown in appendix REF.
+The magnetic field here is generated with an electromagnet (EM), aligned along the [100] diamond axis with a precision $<1^\circ$.
+
+The data shown here was acquired over 24 h.
+It should be noted however that a single-photon counting APD was used here, which considerably worsens the sensitivity of the measurement as discussed in [REF].
+Using a photodiode and modulation of the magnetic field, I estimate that I could get similar signal to noise ratio within a few minutes.
+
+To calibrate the magnetic field, I added a microwave field at a  fixed frequency $\nu > 2870\ \rm MHz$, and measured the EM voltage at which the PL dropped due to the $\ket{+1}$ state of the NV centers becoming resonant with the microwave tone.
+I could then use the values in Fig.
+\ref{121 vs 22 NV} to find the corresponding magnetic field.
+I varied the frequency in the range  $[2870,3150]\ \rm MHz$ by step of 2 MHz to get a calibration of the magnetic field in the $[0,150]\ \rm G$ region.
+Considering the uncertainty in the field angle, the NV ZFS $D$ value and the width of the PL drop, I estimate the final precision of the calibration to be within $\pm 1\ \rm G$.
+
+\section{Attribution of the $^{13}$C-NV$^-$, VH$^-$ and War1 lines}
+
+\subsection{The VH$^-$ and War1 lines}
+\begin{table}[htbp]
+\centering
+\caption{\bf Zero-field splitting parameter $D$ for NV$^-$, VH$^-$ and War1}
+\begin{tabular}{ccc}
+\hline
+$D_z$ estimation (MHz) & Cruddace's work\citep{cruddace2007magnetic} & Our work \\
+\hline
+NV$^-$ & 2872(7) & * \\
+VH$^-$ & 2706(30) & 2694(5)  \\
+WAR1 & 2466(60) & 2470(10) \\
+\hline
+\end{tabular}
+  \label{table VH et War1}
+\end{table}
+
+The attribution of the PL dips at 56 and 122 G in Fig.
+\ref{CR VH exp} to CR between NV and respectively VH$^-$ and War1 comes from two main observations: the closeness in the ZFS values, and the CVD nature of the sample.
+
+Table \ref{table VH et War1} shows the value and precision of the $D$ factor when measured through EPR \citep{cruddace2007magnetic}, and when computed from the data in Fig.
+\ref{CR VH exp}.
+In order to get to the $D$ value of VH$^-$ and War1, we had to assume that the spin structure was that of a spin 1 similar to the NV center, with a g-factor close to 2.
+Both of this assumptions are confirmed by the EPR measurements.
+We do find that that both $D$ measurement concord within their respective error bars.
+
+The second reason comes from the nature of the sample used.
+We tried similar PL scans along the [100] axis on many samples with dense ensemble of NV centers.
+We found that the dip at 55 G was present on all $\sim 5$ CVD samples used, and none of the $\sim 20$ HPHT samples.
+The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below.
+We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the process, and while we don't know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
+
+Finally, the apparent larger linewidth of the War1 line compared to the VH$^-$ one is also corroborated by the EPR measurements.
+
+For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively, with a slightly stronger confidence for the VH$^-$ since more is known about it.
+
+\subsection{The $^{13}$C-NV$^-$ lines}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.6\textwidth]{Figures/scan_100_Sumi4}
+\caption{Change in PL as a function of a magnetic field scanned along the [100] axis for sample SUMI-4}
+\label{scan sumi 4}
+\end{figure}
+
+The attribution of the $\sim 20\ \rm G$ dip to CR with $^{13}$C-NV$^-$ also come from two observations.
+First, the lines are present on every single sample with dense ensemble ($\gtrapprox 1$ ppm), regardless of the fabrication process.
+Fig.
+\ref{scan sumi 4} for example shows a PL scan along the [100] axis for sample Sumi-4, an HPHT sample with [NV]$\approx 5\ \rm ppm$ and natural abundance of $^{13}$C.
+I could unfortunately not try a sample isotopically enriched in $^{12}$C since those samples are generally reserved to low-nitrogen concentration ($< 10\ \rm ppm$), which does not yield enough NV centers to perform CR spectroscopy.
+
+The second reason is the shape of the PL dip.
+The two other dips seem be formed of a single line and are reasonably well fitted with Gaussian curves.
+The 20 G dip however has a flat bottom which would indicate the presence of several overlapping lines, and Fig.
+\ref{121 vs 22 13C-NV} indeed predicts that there would be 4 CR lines between 18 and 25 G.
+
+For this two reasons, we attribute the 20 G line to CR between $^{13}$C-NV$^-$ and "normal" NV centers.
+
+\section{Quantitative estimates of the NV-Dark spins CR}
+
+We can extract some quantitative information about the dark spins by exploiting the CR spectrum shown in Fig. \ref{CR VH exp}.
+
+\subsection{Dark spin Hamiltonian parameters}
+
+First, as discussed in the last section, we can estimate the dark spin ZFS $D$ value with greater precision than EPR.
+While it would be in theory possible to measure other properties of the dark spins Hamiltonian - such as the $g$-factor - by changing the angle of the magnetic field, this would here quickly become intractable due to the increase in co-resonance conditions and the relative closeness of each of these transitions.
+DEER might prove more useful in this regard as only one class of NV centers at a time is probed.
+
+\subsection{Dark spin linewidth}
+
+The second quantitative estimate we can make is of the dark spin inhomogeneous broadening $T_2^*$: if we assume the change in PL to be proportional to the change in the NV lifetime $T_1$, which is true for small enough changes, then the linewidth of the PL dips in Fig.
+\ref{CR VH exp} is the convolution of the NV and dark spin spectrum.
+By deconvoluting the data with the NV spectrum, known for example from ODMR, we can then recover the dark spin spectral response as was done in \citep{hall2016detection}.
+
+\subsection{Dark spin concentration}
+
+Finally, the third tentative estimate we can make is that of the dark spins concentration.
+We can indeed calibrate the measurement by using the $^{13}$C-NV$^-$ line, of which we know the concentration in natural abundance diamond: [$^{13}$C-NV$^-$] $\approx$ 3.3 \% [NV].
+We can then, knowing the NV concentration, try to estimate the concentration of the dark spins.
+There are several caveats though:
+
+\begin{itemize}
+\item Since the change in $T_1$ is proportional to the $T_2^*$ of the probed spins (eq.
+\ref{delta gamma 1}), we should compensate that by integrating over the entire feature, instead of simply looking at the maximum of the line.
+\item We have to take into account the loss of contrast of the NV centers as the transverse magnetic field grows, as explained in sec. \ref{CR contrast}.
+\item The $^{13}$C-NV$^-$ are not "dark spins", and its CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
+We will see however in the next chapter that we can consider a fraction of the NV centers as dark spins, and that this fraction can be estimated.
+\citep{choi2017depolarization} found a ratio of $\sim 1/4$ dark NV centers for $\sim 3/4$ bright ones.
+\end{itemize}
+
+With all this considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):
+
+I Find that the area under 0 for the VH$^-$ line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ line, from \ref{121 vs 22 contrast} I find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally I chose the dark to bright NV ratio value from \citep{choi2017depolarization} at 0.25.
+
+With an estimated NV concentration [NV] $\approx\ \rm 4.6 ppm$, giving us a concentration $[^{13}\rm{C-NV}^-]\approx \ \rm 130 ppb$, I then find $[\rm VH^-] \approx 58\ \rm ppb$.
+Similarly, I find $[\rm VH^-] \approx 68\ \rm ppb$.
+Given the large uncertainty on some parameters, in particular the dark to bright spin ratio, these estimates are probably only precise within an order of magnitude.
+
+If we consider that the sampled volume in this experiment is $< 100 \mu \rm{m}^3$, the total number of dark spins detected here is then $< 10^6$.
+Even with very conservative estimates, the technique employed here is still more sensitive than EPR by several orders of magnitude.
+
+\section{Temperature dependence of VH$^-$ and War1}
+
+\subsection{Mofifications on CVD samples}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/VH_CVD_chauffage}
+\caption{Demodulated PL on samples CVD-C and CVD-B as a function of a magnetic field along the [100] axis.
+An additional oscillating magnetic field along the same direction was used to perform the lock-in detection.
+The VH$^-$ and War1 lines are taken from \ref{CR VH exp}}
+\label{chauffage CVD}
+\end{figure}
+
+Given that the NV CR spectroscopy was both sensitive and relatively easy to setup, we decided in collaboration with the teams of Alexandre Tallaire at IRCP and Jocelyn Achard at LSPM to study the effect of annealing at high temperature on VH$^-$ and War1 \citep{ngambou2022improving}.
+
+Fig.
+\ref{chauffage CVD} shows the results of this experiment.
+This time, a modulation of the magnetic field was used and the PL coming from the photodiode was demodulated through a lock-in amplifier, which explains the derivative shape of the lines.
+The fact that the background is not perfectly flat still comes from the transverse magnetic field.
+The same lock-in parameter were used on both plots so that the amplitudes should in theory be comparable.
+
+Fig \ref{chauffage CVD}-a) shows the demodulated PL versus mganetic field amplitude along [100] on sample CVD-C, while \ref{chauffage CVD}-b) shows the same experiment on sample CVD-B.
+The fabrication and subsequent operations done on this two samples are detailed thoroughly in \citep{ngambou2022improving}, but the main difference between these two sample is that CVD-C was only annealed once after irradiation, for 2 hours at 800$^\circ$C, while CVD-B was annealed for for 2 hours at 800$^\circ$C and then for 1 hour at 1000$^\circ$C and 1200$^\circ$C.
+
+We can see that the VH$^-$ line in particular completely disappear on sample CVD-B, which we attribute to the additional annealing.
+This measurement also corroborate modifications in the infrared absorption spectrum, and an increase of the NV coherence time, associated with the annealing at higher temperature \citep{ngambou2022improving}.
+
+The War1 line on the other hand is still present after the high temperature annealings, albeit slightly smaller.
+This information can give us some insights on the chemistry of the War1 defect.
+%Et en vrai je crois que c'était deja vu dans la thèse de Cruddace
+
+\subsection{Modification on HPHT samples}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{Figures/VH_HPHT_chauffage}
+\caption{Demodulated PL on samples SBST-C and SBST-B as a function of a magnetic field along the [100] axis. The plot of SBST-C is offset for clarity}
+\label{chauffage HPHT}
+\end{figure}
+
+An unexpected side-study was done on HPHT samples.
+Indeed, the CVD samples used previously are grown by homoepithaxy on top of a diamond substrate.
+In this case, the substrate was a type 1B HPHT diamond, which was irradiated and annealed along the CVD layer.
+I Will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.
+
+Fig.
+\ref{chauffage HPHT} shows demodulated PL scans with B $\parallel$ [100] on both samples.
+They both show a line at $B \approx 20\ \rm G$ corresponding to the $^{13}$C-NV$^-$ line, and none of them show a line at $B = 56\ \rm G$ corresponding to the VH$^-$ line.
+There is however a stark difference for the War1 line: SBST-C seems to show a weak line, while SBST-B shows a very clear line, with an amplitude more than 5 times greater than SBST-C.
+
+We can also see that many weak lines on SBST-B seems to have disappeared in SBST-C, while an unidentified line at $B = 48\ \rm G$ seems to have grown similarly to the War1 line.
+
+I first suspected that some of the War1 defects had been migrating from the CVD layer to the HPHT substrate with the high-temperature annealing, but not only is it very unlikely with the annealing time (1 hour) and substrate thickness (several mm), I also could not find a dependence on the War1 concentration with the depth in the substrate, whereas a diffusion would have given a higher concentration near the CVD layer.
+The War1 defects seem to have been created by the annealing.
+
+These results suggest that, while some defects are removed by the high temperature annealing, such as the VH$^-$ and several unidentified defects in the HPHT substrate, some on the other hand seem to be created by the annealing process, such as the War1 or the one associated with the line at $B = 48\ \rm G$ in the HPHT substrate.
+
+\section{Conclusion and perspectives}
+
+To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations.
+This detection can be way more sensitive than the usual EPR spectroscopy, and considerably cheaper to setup.
+It is however not as versatile and should be seen as a complimentary tool for diamonds with dense NV ensemble.
+
+We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan, much less noisy for weak magnetic field.
+
+Finally we have seen an application of the CR to detect optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.
+
+There still remains a lot of open questions and there are several experiments that could be build upon these results. I will list some of them here.
+
+\subsection{Unidentified CR lines}
+
+There are still a large number of lines in CR spectra that have not been identified.
+While not every line necessarily correspond to a new defect - for example \citep{wunderlich2021magnetic} recently found lines corresponding to co-resonance between the $\ket{0} \to \ket{-1}$ transition of an NV center and the $\ket{-1} \to \ket{+1}$ of another NV center - the more likely hypothesis for each of these unidentified lines is still the presence of a new defect.
+
+Such unidentified lines can be seen in Fig.
+\ref{chauffage CVD} and \ref{chauffage HPHT}.
+In particular we can see a line at 160 G in Fig.
+\ref{chauffage CVD}-a) and a line at 48 G in Fig.
+\ref{chauffage HPHT} that detach clearly from the background.
+
+Further studies would be needed to find how common these lines are, and what could be there sources.
+
+\subsection{Measuring the spin dynamics of VH$^-$}
+
+We found that NV-VH$^-$ CR in CVD could produce a change in PL of $\sim 0.2 \%$.
+This kind of values are not too far from ODMR contrast with NV ensembles, which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin.
+The main issue comes from the fact that in order to measure the VH$^-$ spin states, it needs to be resonant with the NV center.
+Therefore, resonant microwave pulses with VH$^-$ will also be resonant with the NV centers, which will create an unwanted optical signal.
+
+There are two solution to circumvent this problem: the first is to use a fast switching magnetic field in order to bring the VH$^-$ and NV centers in resonance when you want to polarize or readout the VH$^-$ state, and out of resonance when you want to coherently manipulate the VH$^-$.
+This means that the magnetic field needs to be switched faster than the spin dynamics you want to probe.
+This might be doable for the population dynamics $T_1$ but this is probably too slow to probe the $T_2$ time.
+
+The other solution consists in using the spin-1 nature of VH$^-$: when a NV$^-$-VH$^-$ CR occurs, only one of the transitions of VH$^-$ and NV$^-$ is brought to resonance.
+The other one ($\ket{0} \to \ket{-1}$ for the VH$^-$) is out of resonance.
+The idea would therefore be to keep the NV $\ket{0} \to \ket{-1}$ transition and VH $\ket{0} \to \ket{+1}$ transition resonant, and to coherently manipulate the VH $\ket{0} \to \ket{-1}$ transition.
+While the polarization on this transition, and the back-action it would produce on the NV centers, is not as strong as if you were manipulating the $\ket{0} \to \ket{+1}$ transition, it should still produce a measurable signal.
+
+\subsection{Isolating a single VH$^-$}
+
+I estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$.
+This concentration is too high by about two orders of magnitude to isolate a single emitter with a diffraction-limited microscope.
+We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing.
+We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and isolate a single VH$^-$ center by scanning the sample with a fixed $B_{[100]}=56 \rm G$ while monitoring the NV PL or $T_1$.
+It is however unknown to me whether this detection scheme is sensitive enough to detect single dark spins.
 

### 2022-08-02 Manuscrit Thèse/chapter 3/Chapter 3.tex
tata

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_ch3.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{}
\begin{document}
\chapter{NV-NV cross-relaxations : the fluctuator model}
In this chapter,...

\section{Experimental observation of NV-NV cross-relaxation}
Before we discuss the theoretical complications related to NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR.
\subsection{NV-NV CR between nonequivalent NV centers}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/NV-NV_non_equivalent}[h]
\caption{label. Taken from \citep{armstrong2010nv}}
\label{non equivalent NV-NV}
\end{figure}
NV-NV CR was first observed more than thirty years ago \citep{holliday1989optical, van1989cross}. The first observations were between non-equivalent NV centers, meaning that the two NV centers involved in the dipole-dipole coupling were not polarized equally.

This scenario can happen for instance when two NV centers from different classes see a different transverse (and longitudinal) magnetic field. Fig \ref{non equivalent NV-NV} illustrates this in the case where the magnetic field is parallel with one of the four classes: $\mathbf{B} \parallel [111]$

When $\mathbf{B} \parallel [111]$, as we discussed in the last chapter, one class sees no transverse field and is therefore always polarized. The three other (equivalent) classes on the other hand get more and more depolarized as the magnetic field amplitude is increased.

It turns out that there is a co-resonnance at B=592 G between the class parallel to $\mathbf{B}$ and the three other classes. This co-resonance is represented in Fig. \ref{non equivalent NV-NV}-b) by an orange circle.

Fig. \ref{non equivalent NV-NV}-c), which we already saw in the last chapter, shows the change in PL of an ensemble of NV centers as $\mathbf{B}$ is scanned along the [111] axis. For B=592 G, we can see a drop in PL also circled in orange. This drop is the result of the CR between NV centers from the class parallel to $\mathbf{B}$ and NV centers from the three other classes.

While a difference in polarization is needed to observe a population transfer, it is not enough to explain the drop in PL. This PL drop is due to the difference in brightness between the $\ket{0}$ and $\ket{+1}$ from the different classes involved. The PL contrast between the two states is higher for the class parallel to $\mathbf{B}$ than for the three other ones, which is another consequence of the transverse field.

The co-resonance between two different classes of NV centers with different magnetic field projection is a relatively rare event: it can only occur for the $\ket{0} \to \ket{+1}$ transition, and only for magnetic fields greater than 592 G.

%techniquement y'a les resonances 0->-1 et -1/+1 mais c'est le delbor, j'ai pas envie d'embrouiller plus que ca.

\section{Microscopic origin of the fluctuators}
\subsection{Tunneling}
\subsection{Modulation of the J-coupling by the phonons}
"Electron Spin-Lattice Relaxation in Phosphorus-Doped Silicon"

"On the interaction of nuclear spins in a crystalline lattice"

"Concentration- and Compensation-Dependent Spin-Lattice Relaxation in n-Type Silicon"

Attention quand même, ce process dépend e la température si je dis pas de bêtises, c'est p-e pas ça chez nous. Nan ok, ça modifie gammaf mais pas le T1 directement. P-e que c'est mesurable ?

Attention bis : C'est l'exchange interaction (le terme en delta) qui est modifié, visiblement le couplage dipolaire etr trop faible pour expliquer ça (papier de Pines et "Spin Resonance of Impurity Atoms in Silicon"
\printbibliography
\end{document}

### 2022-08-04 Manuscrit Thèse/chapter 3/Chapter 3.tex
taratata

@@
 
-\section{Microscopic origin of the fluctuators}
-\subsection{Tunneling}
-\subsection{Modulation of the J-coupling by the phonons}
+\subsection{NV-NV CR between equivalent NV centers}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/NV-NV_equivalent}[h]
+\caption{CR between equivalent NV centers for sample ADM-15-1.
+(a) Representation of equivalent NV centers: either two NVs from the same class, or two classes with the same projected magnetic field.
+(b) Magnetic setup used for the experiment: a permanent magnet is used to apply a bias magnetic field, and an electromagnet is used to add a variable magnetic field.
+(c) Simulation of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers as a function of the scanned magnetic field.
+The transitions were computed based on several ODMR spectra.
+The red dotted lines correspond to inter-class resonance (d) Change in the PL of the NV centers ensemble as a function of the scanned magnetic field.
+The inter-class resonances are reported here.}
+\label{equivalent NV-NV}
+\end{figure}
+
+More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there was CR even between equivalent NV centers.
+
+I refer by equivalent NV centers to NV centers with the same spin Hamiltonian.
+This means either NV center from the same class, or NV centers from different classes with the same projection of the magnetic field along their axis.
+Equivalent NV centers are by nature resonant with each other, and therefore susceptible to CR.
+
+Fig.
+\ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV Cr.
+The key to observe NV-NV CR is to bring some NV centers in and out of resonance with other NV centers, which we can do with the different NV classes.
+To do so, we need to add an initial bias magnetic field in addition to the scanned magnetic field, as represented in Fig.
+\ref{equivalent NV-NV}-b).
+%When doing so, the magnetic field felt by the diamond won't have a fixed orientation during the scan, and when the magnetic field crosses certain crystalline planes (detailed below), its projection on two or more classes of NV centers will be the same.
+
+Fig.
+\ref{equivalent NV-NV}-c) shows the transition frequencies of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers on sample ADM-15-1, an HPHT micro-diamond with $[\rm NV]\sim 3\ \rm ppm$.
+There are three magnetic field values $B_{\rm scan} \sim$ 12, 42 and 77 G for which there is an inter-class resonance.
+Each of these resonances correspond to a "geometric" resonance where the $B$ field is in a symmetry plane between the two classes, meaning that the two resonant NV classes are indeed equivalent.
+
+Finally, Fig.
+\ref{equivalent NV-NV}-d) shows the PL contrast as the electromagnet field is scanned.
+There are 3 very clear dips when $B_{\rm scan} \sim$ 12, 42 and 77 G which coincide with inter-class resonances.
+We will see later that this PL dip is associated with a decrease of the NV centers spin $T_1$ time from both classes.
+
+This observation, and similar ones done by many groups, seem to indicate that there are CR between equivalent NV centers.
+This is however incompatible with our previous assumptions that equivalent NV centers are equally polarized and bright.
+To understand this phenomenon, we will need to introduce new hypotheses.
+
+\section{NV inhomogeneity and the fluctuator model}
+
+\subsection{CR in an inhomogeneous NV bath}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/inhomogene_2_spins}[h]
+\caption{label}
+\label{inhomogene}
+\end{figure}
+A likely explanation to the observed NV-NV CR is that the NV centers are not all equivalents.
+
+Indeed, if we assume that, for instance, the relaxation rate $\Gamma_1=1/T_1$ is not strictly the same for each NV centers, but instead follows a certain distribution $\rho(\Gamma_1)$, then CR between NV centers with different $\Gamma_1$ can explain the observations in Fig.
+\ref{equivalent NV-NV}.
+
+Fig.
+\ref{inhomogene} illustrates this case.
+We are assuming here a coupling between two NV centers with respective relaxation rate $\Gamma_1$ and $\Gamma_1^*$.
+Fig.
+\ref{equivalent NV-NV}-a) illustrates the model I chose for this system: $\Gamma_{\rm las}$ represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins.
+I consider here only the incoherent dynamics of the population which I model with various rates.
+I also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.
+
+By finding the steady state of these rate equations, I can compute the final population in the $\ket{0}$ for both spins, which I will assume be proportional to the total PL.
+Fig.
+\ref{inhomogene}-b) shows this $\ket{0}$ population as a function of the flip-flop rate $\Gamma_{\rm dd}$ for various values of $\Gamma_1^*$.
+
+We can see that when $\Gamma_1^*=\Gamma_1$, as expected, the PL is not modified by the coupling of the two spins regardless of the coupling strength.
+When $\Gamma_1^*>\Gamma_1$ however, not only is the starting PL slightly lower since NV$_2$ is less polarized, but the PL now decreases as we increase the coupling strength.
+This effect, which is the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is higher.
+
+\subsection{Presentation of the fluctuator model}
+Choi et al.
+in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: "normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and "fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism.
+We are assuming that the fluctuator relaxation rate, $\gamma_f$ is much greater than the optical pumping rate $\Gamma_{\rm las}$.
+The fluctuators are therefore unpolarized and effectively act as dark spins, similarly to what we discussed in the last chapter.
+
+We should note here that, while this simplifications seem a bit extreme, the fluctuator model is more than a toy model.
+There are actually good evidence of the presence of these dark NV centers in dense NV ensemble, as will be discussed below.
+
+We will consider the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
+With this assumption, we can compute the modification caused by the fluctuators on the rest of the NV lifetimes.
+
+%\begin{pmatrix}
+%0.33&0&0 \\
+%0&0.33&0 \\
+%0&0&0.33
+%\end{pmatrix}
+
+\subsection{Single NV center coupled to a single fluctuator}
+We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
+To compute the depolarization induced by the fluctuator bath the NV centers, we should first consider the interaction between a single NV center and a fluctuator.
+Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.
+
+We will first decompose the dipole-dipole Hamiltonian in a radial and angular part:
+\begin{equation}
+\mathcal{H}_{\rm dd} \approx - \frac{J_0}{r^3} \left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}+qS_z^1 S_z^2 \right] + h.c. ,
+\end{equation}
+where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative orientation of the two dipoles.
+
+We then introduce the dimensionless number $\eta$ defined as:
+\begin{equation}
+\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
+\end{equation}
+where $\gamma_f=1/T_1^f$ is the fluctuator decay rate.
+This number $\eta$ encapsulate both the geometry of the problem (trhough $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
+
+Finally, the additional depolarization rate induced by the fluctuator on the NV center reads:
+\begin{equation}
+\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right) \frac{\eta^2}{\gamma_f}.
+\end{equation}
+Compared to eq.
+[REF], we have replaced here $\Gamma_2^*$ by $\gamma_f$.
+This is because we assume that the pure dephasing, supposedly equal for the NV center and the flucutator, is smaller than the depolarization rate $\gamma_f$.
+We will further discuss this hypothesis later.
+
+\subsection{Ensemble of NV centers coupled to a bath of fluctuators}
+
+We now need to compute the average depolarization caused by the ensemble of fluctuators on the ensemble of NV centers.
+
+We consider each fluctuator as in independant depolarization source, therefore the depolarization on a single NV center reads: 
+\begin{equation}
+\gamma_{\rm NV}=\sum_i \gamma_s(\mathbf{r}_i).
+\end{equation}
+We then want to compute the distribution $\rho(\gamma_{\rm NV})$ defined as:
+\begin{equation}
+\rho(\gamma_{\rm NV})=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma_{\rm NV} \right).
+\end{equation}
+To do so, we need to determine the distribution of the fluctuators positions $\{r_i\}$. Assuming that they are homogeneously distributed in the bulk of the material, we find:
+\begin{equation}
+\rho(\gamma_{\rm NV})=\frac{e^{-1/(4\gamma_{\rm NV} T)}}{\sqrt{4\pi \gamma_{\rm NV}^3 T}},
+\end{equation}
+where the time constant $T$ was introduced and is defined as:
+\begin{equation}
+\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f},
+\label{eq 1/T}
+\end{equation}
+with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{eta}$: $\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta$.
+
+Finally, the polarization dynamics from the ensemble of NV centers can be computed:
+\begin{equation}
+P(t)=\int_0^\infty \rho(\gamma_{\rm NV})\, e^{-\gamma_{\rm NV} t}d\gamma= e^{-\sqrt{t/T}}.
+\end{equation}
+
+In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential, unlike the phonon-induced depolarization which is purely exponential.
+
+It should be noted that the stretched exponential nature of the decay is not proper to the fluctuator model.
+\citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
+The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).
+
+\section{Validation of the fluctuator model}
+We will now show experimental results validating the fluctuator model.
+\subsection{The stretched exponential lifetime}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/Stretch_vs_pas_stretch}[h]
+\label{stretch_or_not_stretch}
+\caption{T1 measurement following the protocol described in [REF].
+The best exponential and stretched exponential fits are given each time.
+a) Sample ADM-15-2 with B=0.
+The optimal $T_1$ values are $T_1^{\rm stretch}=150\ \mu \rm s$ and $T_1^{\rm exp}=410\ \mu \rm s$.
+b) Sample CVD-pink with B$\neq$0 and a single class probed.
+The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$}
+\end{figure}
+$T_1$ measurements with dense ensemble of NV centers and when many classes are resonant with each other (typically when B=0) indeed tend to show stretched exponential profiles.
+
+Fig.
+\ref{stretch_or_not_stretch} shows an example of that: in Fig.
+\ref{stretch_or_not_stretch}-a) the $T_1$ measurement comes from sample ADM-15-2 at B=0.
+The profile is clearly better fitted by a stretched exponential than a regular exponential, and the timescale found $T_1^{\rm stretch}=150\ \mu \rm s$ is much shorter than the expected phonon limite lifetime of a few ms.
+
+On the other hand, Fig.
+\ref{stretch_or_not_stretch}-b) shows a $T_1$ measurement coming from sample CVD-pink for B$\neq$0.
+This time the profile is more exponential, and the value $T_1^{\rm exp}=4.1\ \rm ms$ is coherent with a phonon-limited lifetime.
+Even though CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.
+
+\subsection{Magnetic field dependent lifetime, double exponential fit}
+
+For a lot of samples, the
+
+\subsection{The fluctuators linewidth}
+\subsection{Parallel with P-doped Si}
 "Electron Spin-Lattice Relaxation in Phosphorus-Doped Silicon"
@@
 "Concentration- and Compensation-Dependent Spin-Lattice Relaxation in n-Type Silicon"
+\subsection{Microscopic origin of the fluctuators}
+On va dire subsection plutot.
+Du coup ici discussion tunneling Vs Modulation of J Vs surface spins
 
@@
 
-Attention bis : C'est l'exchange interaction (le terme en delta) qui est modifié, visiblement le couplage dipolaire etr trop faible pour expliquer ça (papier de Pines et "Spin Resonance of Impurity Atoms in Silicon"
+Attention bis : C'est l'exchange interaction (le terme en delta) qui est modifié, visiblement le couplage dipolaire est trop faible pour expliquer ça (papier de Pines et "Spin Resonance of Impurity Atoms in Silicon")
+
+\section{Geometric conditions for NV-NV resonance}
+En gros la carte, les 5 ESR en SI du PRL (lol) et le 121 Vs 22
+
+\section{Limitations of the fluctuator model}
+La modification des T1, tjr plus petite que prévue (faire un tableau), et les T1 non-exponentiels (je crois que y'a des cas sur Ludo, bon courage mdr)
+
 \printbibliography

### 2022-08-04 experience/Bilan manips/2021.tex
tata
taratata

@@
 \item \textbf{09/13-15} Série de T1, probablement sur adamas, probablement sans soustraction, donc pas oufs.
-\item \textbf{09/27} Mesure sensitivité magnétométrie Vs angle pour un adamas avec beaucoup de DQ, et un avec peu de DQ. Il y a aussi des ESR 0B et des T1
+\item \textbf{09/27} Mesure sensitivité magnétométrie Vs angle pour un adamas avec beaucoup de DQ, et un avec peu de DQ.
+Il y a aussi des ESR 0B et des T1.
+Le T1 du low DQ est très stretch et celui du high DQ très exp.
+Mesure également de la pola du laser sur des scans 100 et 111 (spoiler: rien du tout)
 \item \textbf{10/06-15} Tests avec laser rouge résonnant, pas hyper concluant meme si on voit une différence sur les DQ
@@
 \item \textbf{06/23} Scans 100 calibrés sur les 3 100 ppm (non recuits à 1200).
-\item \textbf{10/20} Mesures sur les échantillons recuit à 1200 + 800 + rose : Scans 100 (mesure VH- et War1), mais aussi croisement de raies (pour mesurer les SQ) et ESR pour T2
+\item \textbf{10/20} Mesures sur les échantillons recuit à 1200 + 800 + rose : Scans 100 (mesure VH- et War1), mais aussi croisement de raies (pour mesurer les SQ) et ESR pour T2.
+Egalement mesure de la phase des CR Vs phase du champ transverse au lock-in sur le rose.
 \item \textbf{10/27} Scans 100 sur les substrats recuits et pas recuits : un pic de War1 pousse avec le recuit dans le substrat.
@@
 \item \textbf{10/25} Mesure de la puissance micro-onde en fonction de la fréquence, en testant l'effet de l'ampli et du clapet anti-retour
+\item \textbf{10/15} Vrais ESR en PL non démodulée pour à peu près toutes les configurations
 

### 2022-08-04 experience/Bilan manips/2022.tex
TOTO
taratata

@@
 Pour l'instant ça merdouille pas mal).
+\item \textbf{03/24} dossier adamas 150: le gros des mesures du papier CR low field. Si t'as des questions va voir le papier
+\item \textbf{04/29} Test polarization : Scan direction random (pour le papier) et ESRs sur quelques échantillons.
+
 \end{itemize}
@@
 \item \textbf{02/21} Test T1 121 vs 22 sur diamant rose, ça marche pas (différence trop faible)
+\item \textbf{03/10} ESR détaillés spin nucl sur des échantillons clean d'Alex (le recuit à 1200 et un 1er échantillon 12C 15N qui avait pas mal de 14N)
+\item \textbf{03/24} ESR sur un autre échantillon 12C 15N, cette fois ci sans trop de 14N
+\item \textbf{06/21} Test scans 100 désalignés sur le rose (pour le papier)
+\end{itemize}
+
+\section{Sumitumo et Ludo}
+\begin{itemize}
+\item \textbf{03/17} Sumi2 : l'animation du croisement de 121 avec la mesure de T1 en direct.
+Exploitable d'ailleurs.
+Sumi4: des ESR et des T1, c'est pas très clair.
+J'essaie de voir l'effet du champs transverse, mais aussi la différence de fréquence sur la soustraction (au milieu, à doite ou à gauche de la raie)
+\item \textbf{03/24} Ludo: des scans 100 et pas 100 et des T1 assez cleans
 \end{itemize}
@@
 \item \textbf{02/21} Essai de mesures EPR en transmission avec la PCB et le diamant rose. Ca marche pas. Mais j'ai test en modulant le laser, j'aurais du moduler le champ mag.
+\item \textbf{06/03-07-08} Des mesures mystiques de Gabriel
+\item \textbf{06/20} Tests avec le nouveau laser 250 mW qui se met à déconner mi-scan ESR
 \end{itemize}
+$\left[1 \bar 1 1 \right]$
 \end{document}

### 2022-08-05 Manuscrit Thèse/chapter 3/Chapter 3.tex
tata yoyo

@@
 
-\subsection{Magnetic field dependent lifetime, double exponential fit}
+\subsection{Fitting $T_1$ profile in the general case}
 
-For a lot of samples, the
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/various_fit_formulae}[h]
+\label{various_fit_formulae}
+\caption{Different fitting procedure on two $T_1$ measurements done on sample ADM-150-1 for $B=0$ and $B\neq0$.
+\\ a) Exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph})$.
+\\ b) Stretched exponential fit $S(\tau)=\exp (-\sqrt{\tau/T_1^{\rm dd}})$.
+\\ c) Bi-exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph} -\sqrt{\tau/T_1^{\rm dd}})$.
+\\ d) Bi-exponential fit with fixed $T_1^{\rm ph}=5\ \rm ms$: $S(\tau)=\exp (-\tau/5 -\sqrt{\tau/T_1^{\rm dd}})$.}
+\end{figure}
+
+\begin{table}[htbp]
+\centering
+\caption{Fitting parameters on Fig. \ref{various_fit_formulae}}
+\begin{tabular}{c|cc|cc}
+\toprule
+Figure &  \multicolumn{2}{c}{$B=0$} & \multicolumn{2}{c}{$B\neq0$}\\
+\midrule
+{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms)&  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)\\
+Fig. a) & 0.95 & * & 2.64 & * \\
+Fig. b) & * & 0.32 & * & 1.03  \\
+Fig. c) & 6.75 & 0.44 & 4.22 & 7.19 \\
+Fig. d) & \textbf{5} & 0.51 & \textbf{5} & 4.65 \\
+\bottomrule
+\end{tabular}
+
+Bold characters indicate parameters that were arbitrarily fixed.
+  \label{fitting table}
+\end{table}
+
+For many samples however, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$ is of the same order as the phonon limited lifetime $T_1^{\rm ph}$.
+Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, forcing us to include both aspects if we want a unique fitting formula.
+
+Fig.
+\ref{various_fit_formulae} shows four different fitting procedure on the two same measurements.
+The measurements here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
+The magnetic field was strong enough to split the four classes and only one class was probed in the second case.
+
+The fits used here are either purely exponential or purely stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$.
+The values of thes different fitting parameters used is reported in Table \ref{fitting table}.
+
+We can see that neither the purely exponential or stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is purely fitted by the stretched exponential.
+The protocols that include both exponential and stretched lifetimes correctly fit both curves.
+In the case of Fig.
+\ref{various_fit_formulae}-d), we arbitrarily fixed $T_1^{\rm ph}=5\ \rm ms$ since this is the value we typically measure on samples with low NV density, and we expect that the phonon-limited exponential lifetime is not modified by the NV concentration.
+
+Since the protocol where $T_1^{\rm ph}$ was fixed and the one where it wasn't both correctly fit our data, we decided to use the protocol where $T_1^{\rm ph}$ was fixed because there is one less free parameter in the fitting function, and the values we obtain on $T_1^{\rm dd}$ can be directly compared.
+
+We should note however that the values of $T_1^{\rm dd}$ we obtain are not absolute.
+Indeed, in the case of Fig.
+\ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
+While this technique of measuring $T_1^{\rm dd}$ is relatively sensitive and reproducible, it is not very accurate.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/betas}[h]
+\label{betas}
+\caption{a) Same two measurements as Fig.
+\ref{various_fit_formulae} with the fitting formula $S(\tau)=\exp ((-\tau/T_1)^{\beta})$.
+The fitting parameters are $\beta=0.58$ and $T_1=0.46\ \rm ms$ for $B=0$, and $\beta=0.80$ and $T_1=2.15\ \rm ms$ for $B\neq0$.
+b) Optimal $\beta$ parameter found for each $T_1$ measurement on the same sample as a function of the external magnetic field.}
+\end{figure}
+
+Finally, another fitting procedure is shown in Fig. \ref{betas}. This time
 
@@
 La modification des T1, tjr plus petite que prévue (faire un tableau), et les T1 non-exponentiels (je crois que y'a des cas sur Ludo, bon courage mdr)
+Parler aussi des alternatives au fluctuator model, ici ou dans la 2e section
 

### 2022-08-05 python_clément/lab.py

@@ -49,7 +49,13 @@ elif macAdressOfCurrentPC=='e4:b9:7a:ea:ff:e5' : #Ordi 3 (du fond)
 	computerUsed='Ordi3'
 	defaultDataPath="D:/DATA/"
 	defaultTheme='dark'
-elif macAdressOfCurrentPC=='b4:45:06:47:da:24' : #Ordi portable de bureau du labo
+
+elif macAdressOfCurrentPC=='b4:45:06:47:da:24' : #2e ordi perso
+	computerUsed='OrdiBureau'
+	defaultDataPath="../data/"
+	defaultTheme='light'
+
+elif macAdressOfCurrentPC=='1c:c1:0c:2b:bb:f5' : #Ordi portable du boulot
 	computerUsed='OrdiBureau'
 	defaultDataPath="../data/"
 	defaultTheme='light'

### 2022-08-06 Manuscrit Thèse/chapter 3/Chapter 3.tex
toto

@@
 \centering
-\includegraphics[width=\textwidth]{Figures/NV-NV_non_equivalent}[h]
+\includegraphics[width=\textwidth]{Figures/NV-NV_non_equivalent}
 \caption{label. Taken from \citep{armstrong2010nv}}
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/NV-NV_equivalent}[h]
+\includegraphics[width=\textwidth]{Figures/NV-NV_equivalent}
 \caption{CR between equivalent NV centers for sample ADM-15-1.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/inhomogene_2_spins}[h]
+\includegraphics[width=\textwidth]{Figures/inhomogene_2_spins}
 \caption{label}
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Stretch_vs_pas_stretch}[h]
+\includegraphics[width=\textwidth]{Figures/Stretch_vs_pas_stretch}
 \label{stretch_or_not_stretch}
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/various_fit_formulae}[h]
+\includegraphics[width=\textwidth]{Figures/various_fit_formulae}
 \label{various_fit_formulae}
@@
 The measurements here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
-The magnetic field was strong enough to split the four classes and only one class was probed in the second case.
+The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.
 
 The fits used here are either purely exponential or purely stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$.
-The values of thes different fitting parameters used is reported in Table \ref{fitting table}.
+The values of the different fitting parameters used is reported in Table \ref{fitting table}.
 
-We can see that neither the purely exponential or stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is purely fitted by the stretched exponential.
+We can see that neither the purely exponential nor stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is poorly fitted by the stretched exponential.
 The protocols that include both exponential and stretched lifetimes correctly fit both curves.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/betas}[h]
+\includegraphics[width=\textwidth]{Figures/betas}
 \label{betas}
@@
 The fitting parameters are $\beta=0.58$ and $T_1=0.46\ \rm ms$ for $B=0$, and $\beta=0.80$ and $T_1=2.15\ \rm ms$ for $B\neq0$.
-b) Optimal $\beta$ parameter found for each $T_1$ measurement on the same sample as a function of the external magnetic field.}
+b) Optimal $\beta$ parameter found for each $T_1$ measurement as a function of the external magnetic field (still on sample ADM-150-1).}
 \end{figure}
 
-Finally, another fitting procedure is shown in Fig. \ref{betas}. This time
+Finally, another fitting procedure is shown in Fig.
+\ref{betas}.
+This time we are using a single stretched exponential with an arbitrary stretch factor $\beta$.
+Fig.
+\ref{betas}-b) shows the optimal $\beta$ parameter as a function of the external magnetic field.
+We can then confirm that the $T_1$ profile gets closer to a stretched exponential ($\beta=0.5$) when $B$ goes to 0.
+
+Even though this method also yields satisfying fits, it needs two free parameters to work ($\beta$ and $T_1$), so we decided to go with the method of Fig. \ref{various_fit_formulae}-d) instead.
 
 \subsection{The fluctuators linewidth}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/fluctuator linewidth}
+\label{fluct linewidth}
+\caption{Measurement of the fluctuators linewidth using the same setup as in Fig.
+\ref{equivalent NV-NV}.
+a) Simulation of the $\ket{0} \to \ket{-1}$ transition frequency for the four classes of NV centers and actual frequencies of the two central classes from ODMR measurements.
+b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.78\ \rm MHz$.
+The green dashed line correspond to the ODMR spectrum of a single class of NV centers.}
+\end{figure}
+
+One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators linewidth.
+In the model, we assumed that fluctuator's lifetime $T_1^f=1/\gamma_f$ was so short that the fluctuators linewidth was lifetime-limited, whereas the normal NV centers linewidth is limited by the inhomogeneous broadening $T_2^*>T_1^f$.
+
+We can experimentally verify this hypothesis: the fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in the last chapter.
+
+Fig. \ref{fluct linewidth} shows such
+
+...Note that only the two classes ... were measured accurately since ... The central region ... overlap ... linear regression
 \subsection{Parallel with P-doped Si}

### 2022-08-08 Manuscrit Thèse/chapter 3/Chapter 3.tex
t

@@
 \section{Experimental observation of NV-NV cross-relaxation}
-Before we discuss the theoretical complications related to NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR.
+Before we discuss the theoretical complications related to NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR with dense NV ensembles ([NV]$\gtrsim$ 1 ppm).
 \subsection{NV-NV CR between nonequivalent NV centers}
@@
 
-\section{Validation of the fluctuator model}
-We will now show experimental results validating the fluctuator model.
+\section{Experimental investigation of the fluctuator model}
+We will now show experimental results related to the predictions of the fluctuator model.
 \subsection{The stretched exponential lifetime}
@@
 \includegraphics[width=\textwidth]{Figures/Stretch_vs_pas_stretch}
-\label{stretch_or_not_stretch}
 \caption{T1 measurement following the protocol described in [REF].
@@
 The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$}
+\label{stretch_or_not_stretch}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/various_fit_formulae}
-\label{various_fit_formulae}
 \caption{Different fitting procedure on two $T_1$ measurements done on sample ADM-150-1 for $B=0$ and $B\neq0$.
@@
 \\ d) Bi-exponential fit with fixed $T_1^{\rm ph}=5\ \rm ms$: $S(\tau)=\exp (-\tau/5 -\sqrt{\tau/T_1^{\rm dd}})$.}
+\label{various_fit_formulae}
 \end{figure}
@@
 \centering
-\caption{Fitting parameters on Fig. \ref{various_fit_formulae}}
+\caption{Fitting parameters for Fig.
+ \ref{various_fit_formulae}}
+ \label{fitting table}
 \begin{tabular}{c|cc|cc}
@@
 Bold characters indicate parameters that were arbitrarily fixed.
-  \label{fitting table}
+
 \end{table}
@@
 \includegraphics[width=\textwidth]{Figures/betas}
-\label{betas}
 \caption{a) Same two measurements as Fig.
@@
 b) Optimal $\beta$ parameter found for each $T_1$ measurement as a function of the external magnetic field (still on sample ADM-150-1).}
+\label{betas}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/fluctuator linewidth}
-\label{fluct linewidth}
 \caption{Measurement of the fluctuators linewidth using the same setup as in Fig.
@@
 b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.78\ \rm MHz$.
-The green dashed line correspond to the ODMR spectrum of a single class of NV centers.}
+The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor $\sqrt{2}$}
+\label{fluct linewidth}
 \end{figure}
@@
 
-Fig. \ref{fluct linewidth} shows such
+Fig.
+\ref{fluct linewidth} shows an experiment similar to the one presented in Fig.
+\ref{equivalent NV-NV}.
+A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers, but this time measuring the spin decay rate instead of the PL.
 
-...Note that only the two classes ... were measured accurately since ... The central region ... overlap ... linear regression
+Fig.
+\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
+Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of the two central lines to get a precise measurement of the Larmor frequency.
+In the region where the two ODMR lines overlap we have to use a linear regression to get the Larmor frequency of each class.
+
+Fig.
+\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
+For each magnetic field value, a $T_1$ profile was recorded and fitted following the procedure described in Fig.
+\ref{various_fit_formulae}-d) where we again fixed $T_1^{\rm ph}=5\ \rm ms$.
+We also show the ODMR spectrum of a single NV class for comparison.
+
+We can clearly see that the increase in $T_1^{\rm dd}$ is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
+If we assume that both NV classes spectral response is a gaussian of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
+Fig.
+\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $\sqrt{2}$, which is still much narrower than the $1/T_1^{\rm dd}$ line.
+
+A potential explanation for the broadening of the $1/T_1^{\rm dd}$ line could be the interaction strength between the NV centers, but for [NV]$\sim$5 ppm, $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$ which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.6\textwidth]{Figures/deconvolution}
+\caption{Blue curve $1/T_1^{\rm dd}$ data presented in Fig.
+\ref{fluct linewidth}.
+Red curve: deconvolution of the blue curve the  by the ODMR line shown on the same figure.
+Green curve: deconvolution of the red curve the  by the ODMR line.
+All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38}
+\label{deconvolution}
+\end{figure}
+
+The fluctuator model provides another explanation: the fluctuator spectral response is broaden by its very short lifetime.
+This also explains why the $1/T_1^{\rm dd}$ line is well matched by a Lorenzian, while the ODMR lines are closer to Gaussians in our sample.
+
+We can now extract the fluctuator lifetime from the data in Fig. \ref{fluct linewidth} by taking out the contribution of $T_2^*$ on both the NV and the fluctuator spectral response.
+
+The first way to do this is to follow the instructions in \citep{hall2016detection} and deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum.
+The result of the deconvolution is shown in Fig.
+\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (red curve).
+Fitting with Lorentzians, we find a final value $\Gamma_2 =\frac{\gamma_f}{2} = (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=12.5\ \rm ns$.
+
+Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\sqrt{2 \ln (2)}\sigma$ and a Lorentzian of HWHM $f_L$:
+\begin{align*}
+\Gamma_1^{\rm dd}&\propto \mathcal{G}(f_G)*(\mathcal{L}(f_L)*\mathcal{G}(f_G)) \\
+&\propto \mathcal{L}(f_L)*\mathcal{G}(\sqrt{2} f_G),
+\end{align*}
+where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
+Then We can use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{wiki:VoigtProfile}:
+\begin{equation}
+f = [f_G^5 + 2.69269 f_G^4 f_L + 2.42843 f_G^3 f_L^2 + 4.47163 f_G^2 f_L^3 + 0.07842 f_G f_L^4 + f_L^5]^{1/5}.
+\end{equation}
+
+With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, I find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 12.2\ \rm ns$, a value consistent with the one found with the deconvolution method.
+
+Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=3.3\ \rm MHz$, although given their experimental values - full width Half maximum (FWHM) of 25 MHz for the lifetime profile and 9.3 MHz for the ODMR line - I obtain using my method $\gamma_f= 15.7\ \rm MHz$ and $T_1^f=10.1\ \rm ns$.
+The main difference comes from the fact that they considered the spectral half-width of the fluctuators to be $2\gamma_f$ whereas I consider it to be $\gamma_f/2$.
+
+\section{Microscopic explanation of the fluctuators}
 \subsection{Parallel with P-doped Si}
+A similar 
 "Electron Spin-Lattice Relaxation in Phosphorus-Doped Silicon"

### 2022-08-08 data/20220324/Adamas 150 um/adamas 1/largeur fluct 5 bis/traitement.py

@@ -140,7 +140,39 @@ def get_T1_dconvoloved(x,y,nmiddle=24,lambd=2):
 
 
 
-def plot_T1_fluct_lw():
+def plot_T1_fluct_lw_and_ODMR():
+	x,y=get_T1_croisement()
+
+	ax1=plt.gca()
+
+	ax1.plot(x,y,'o',markerfacecolor='none',label=r'Experiment')
+	popt,yfit=lor_fit(x,y)
+	print(popt)
+	ax1.plot(x,yfit,lw=2,color=color(1),label='Lorentzian fit'%popt[2])
+
+
+	ax2=ax1.twinx()
+	x,y=extract_data('ESR 1 classe pas loin 111')
+	y=y/max(y)
+	c=hist_mean(x,y)
+	x=x-2740
+	ax2.plot(x,y,'--',color=color(2),label=r'ESR line',mew=0.5,ms=3)
+	x=x*sqrt(2)
+	ax2.plot(x,y,'--',color=color(3),label=r'ESR line$\times \sqrt{2}$',mew=0.5,ms=3)
+	# x=x*sqrt(2)
+	# n=len(x)
+	# x=x[n//5:n-n//5]
+	# y=y[n//5:n-n//5]
+	# ax2.plot(x,y,'--',color=color(4),label=r'ESR line$\times 2$',mew=0.5,ms=3)
+	ax2.tick_params(labelsize=15)
+	ax2.legend()
+
+	ax1.legend(loc=2)
+	
+# plot_T1_fluct_lw_and_ODMR()
+
+
+def plot_T1_fluct_deconvo():
 	x,y=get_T1_croisement()
 
 	ax1=plt.gca()
@@ -165,26 +197,19 @@ def plot_T1_fluct_lw():
 	ax1.plot(x,yfit,lw=2,color=color(2))#,label='Lorentzian fit'%popt[2])
 	print(popt)
 
-	# ax2=ax1.twinx()
-	# x,y=extract_data('ESR 1 classe pas loin 111')
-	# y=y/max(y)
-	# c=hist_mean(x,y)
-	# x=x-2740
-	# ax2.plot(x,y,'--',color=color(2),label=r'ESR line',mew=0.5,ms=3)
-	# # x=x*sqrt(2)
[…]

### 2022-08-08 python_clément/test programs/test.py

@@ -1703,6 +1703,8 @@ def exp(x,sigma):
 	y=np.exp(-(x**2)/(2*sigma**2))
 	return y
 
+
+
 def lor(x,sigma):
 	y=1/(1+(x/sigma)**2)
 	return y
@@ -1740,4 +1742,8 @@ def test_convolution():
 	plt.show()
 
 
-test_convolution()
+# test_convolution()
+
+
+
+print(1/(2*np.pi*7.83*2))

### 2022-08-09 Manuscrit Thèse/chapter 3/Chapter 3.tex
yayayaya

@@
 \subsection{NV-NV CR between nonequivalent NV centers}
+\label{non_equi_valent_CR}
 \begin{figure}[h]
@@
 
+Other possibilities for the dipole-dipole dependent relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or superradiance, but they were not considered to be viable explanations \citep{choi2017depolarization}
 %\begin{pmatrix}
@@
 \begin{equation}
-\gamma_{\rm NV}=\sum_i \gamma_s(\mathbf{r}_i).
+\gamma=\sum_i \gamma_s(\mathbf{r}_i).
 \end{equation}
-We then want to compute the distribution $\rho(\gamma_{\rm NV})$ defined as:
+We then want to compute the distribution $\rho(\gamma)$ defined as:
 \begin{equation}
-\rho(\gamma_{\rm NV})=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma_{\rm NV} \right).
+\rho(\gamma)=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma \right).
 \end{equation}
@@
 \begin{equation}
-\rho(\gamma_{\rm NV})=\frac{e^{-1/(4\gamma_{\rm NV} T)}}{\sqrt{4\pi \gamma_{\rm NV}^3 T}},
+\rho(\gamma)=\frac{e^{-1/(4\gamma T)}}{\sqrt{4\pi \gamma^3 T}},
 \end{equation}
@@
 \end{equation}
-with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{eta}$: $\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta$.
+with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{\eta}$: $\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta$.
 
@@
 \begin{equation}
-P(t)=\int_0^\infty \rho(\gamma_{\rm NV})\, e^{-\gamma_{\rm NV} t}d\gamma= e^{-\sqrt{t/T}}.
+P(t)=\int_0^\infty \rho(\gamma)\, e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}.
 \end{equation}
@@
 The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).
+
+\subsection{Possible microscopic origin of the fluctuators}
+
+We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}.
+It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.
+
+This conclusion is based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions, and the resonant energy transfer between molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through dipole-dipole interaction to non-excited acceptors, and where the stretched exponential dynamics of the donors was first observed \citep{forster1949experimentelle}.
+
+In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
+In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
+Both of these phenomena could lead to fast spin relaxation, the tunneling because the spin hopping is accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction strength $J$ by the crystal phonons \citep{honig1960electron}.
+Technically, the non-contact dipole-dipole interaction is also modulated by the phonons, but the effect is too small to account for the fast depolarization.
+
+Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
+To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate $\sim 10\ \rm ns$.
+In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
 
@@
 
-A potential explanation for the broadening of the $1/T_1^{\rm dd}$ line could be the interaction strength between the NV centers, but for [NV]$\sim$5 ppm, $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$ which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
+A potential explanation for the broadening of the $1/T_1^{\rm dd}$ line could be the interaction strength between the NV center and the fluctuator.
+However, for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
 
@@
 Green curve: deconvolution of the red curve the  by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38}
+All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38 MHz}
 \label{deconvolution}
@@
 
-The first way to do this is to follow the instructions in \citep{hall2016detection} and deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum.
+The first way to do this is to follow the instructions in \citep{hall2016detection} and deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm.
 The result of the deconvolution is shown in Fig.
-\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (red curve).
+\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve).
 Fitting with Lorentzians, we find a final value $\Gamma_2 =\frac{\gamma_f}{2} = (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=12.5\ \rm ns$.
@@
 where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
-Then We can use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{wiki:VoigtProfile}:
+We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{wiki:VoigtProfile}:
 \begin{equation}
@@
 
-Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=3.3\ \rm MHz$, although given their experimental values - full width Half maximum (FWHM) of 25 MHz for the lifetime profile and 9.3 MHz for the ODMR line - I obtain using my method $\gamma_f= 15.7\ \rm MHz$ and $T_1^f=10.1\ \rm ns$.
+Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=3.3\ \rm MHz$, although given their experimental values - full width at half maximum (FWHM) of 25 MHz for the lifetime profile and 9.3 MHz for the ODMR line - I obtain using my method $\gamma_f= 15.7\ \rm MHz$ and $T_1^f=10.1\ \rm ns$.
 The main difference comes from the fact that they considered the spectral half-width of the fluctuators to be $2\gamma_f$ whereas I consider it to be $\gamma_f/2$.
 
-\section{Microscopic explanation of the fluctuators}
-\subsection{Parallel with P-doped Si}
-A similar 
-"Electron Spin-Lattice Relaxation in Phosphorus-Doped Silicon"
+\section{Trouver un bon titre}
 
-"On the interaction of nuclear spins in a crystalline lattice"
+The main tool to probe NV-NV CR is to use the co-resonances between the NV classes.
+While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and supposedly resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances.
+We will discuss here the various ...
 
-"Concentration- and Compensation-Dependent Spin-Lattice Relaxation in n-Type Silicon"
-\subsection{Microscopic origin of the fluctuators}
-On va dire subsection plutot.
-Du coup ici discussion tunneling Vs Modulation of J Vs surface spins
+\subsection{Geometric conditions for inter-class resonance}
 
-Attention quand même, ce process dépend e la température si je dis pas de bêtises, c'est p-e pas ça chez nous. Nan ok, ça modifie gammaf mais pas le T1 directement. P-e que c'est mesurable ?
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{Figures/Resonance_geometries}
+\caption{ODMR spectra on sample ADM-15-3 as a function of the magnetic field orientation.
+The field amplitude is $\sim 60\ \rm G$.
+Red numbers represent the number of resonant classes for each ODMR line.
+The cubes represent the diamond unit cell, and the red planes/arrow the possible orientations of the magnetic field}
+\label{ODMR_geometries}
+\end{figure}
 
-Attention bis : C'est l'exchange interaction (le terme en delta) qui est modifié, visiblement le couplage dipolaire est trop faible pour expliquer ça (papier de Pines et "Spin Resonance of Impurity Atoms in Silicon")
+There are four possibilities to obtain inter-class resonance (outside of the non-equivalent CR discussed in sec. \ref{non_equi_valent_CR}). These possibilities are:
+\begin{itemize}
+\item $\mathbf{B} \in \{110\}$.
+We refer by $\{110\}$ as any plane orthogonal to a $\langle 110 \rangle$ or equivalent direction.
+In this magnetic configuration, two classes are at resonance and the two other ones are spectrally isolated.
+\item $\mathbf{B} \in \{100\}$. In this magnetic configuration, the four classes form two pairs of co-resonance.
+\item $\mathbf{B} \parallel \langle 111 \rangle$.
+In this configuration discussed in the last chapter, one class is aligned with the magnetic field and is spectrally isolated while the three other form a resonant triplet.
+\item $\mathbf{B} \parallel \langle 100 \rangle$. In this configuration also discussed in the last chapter, the four classes are resonant.
+\end{itemize}
+We should add to this list the case $\mathbf{B}=0$ where all four classes are also resonant, and any other possibility (which I will call random $\mathbf{B}$) where all four classes are spectrally separated.
 
-\section{Geometric conditions for NV-NV resonance}
-En gros la carte, les 5 ESR en SI du PRL (lol) et le 121 Vs 22
+An ODMR spectrum and a representation of the magnetic field in the diamond unit cell for each of these six possibilities is shown in Fig. \ref{ODMR_geometries}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/carte}
+\caption{PL contrast as a function of the magnetic field polar and azimuthal angle with respect to the [100] axis on sample CVD-pink.
+The field amplitude is $\abs{B} \sim 115 G$.
+The locus of the magnetic field in specific planes is noted by red dashed lines.}
+\label{Carte}
+\end{figure}
+
+This relation between NV-NV class resonances and the magnetic field being in specific crystalline planes or directions means that we can determine the crystal main axes simply by PL experiments.
+
+Fig.
+\ref{Carte} shows a map where we simply monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer.
+We have also reported the locus of $\mathbf{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].
+
+We know that when $\mathbf{B}$ belongs to any of these four planes, there is a co-resonance between at least two NV classes.
+This co-resonance leads to an increase in NV-NV CR (or NV-fluctuator CR) and a faster depolarization of the spins involved, which ultimately leads to a drop in PL.
+We can indeed see in fig.
+\ref{Carte} that there is clear correlation between the loci of $\mathbf{B}$ and drops in PL.
+We can also see that the PL is at its lowest when $B\parallel [001]$, which corresponds to a four class degeneracy.
+
+\subsection{$T_1$ dependency on magnetic field}
 
@@
 Parler aussi des alternatives au fluctuator model, ici ou dans la 2e section
+Bon du coup parler du fait que ça n'inclut pas la diffusion entre NV ni la saturation des fluctuateurs. A voir si ça mérite une full section ou une gross ouverture à la fin.
 

### 2022-08-09 python_clément/test programs/test.py

@@ -1746,4 +1746,4 @@ def test_convolution():
 
 
 
-print(1/(2*np.pi*7.83*2))
+print(200/(2*2.8/np.sqrt(3)))

### 2022-08-10 Manuscrit Thèse/chapter 3/Chapter 3.tex
yaya

@@
 Bold characters indicate parameters that were arbitrarily fixed.
-
 \end{table}
@@
 \ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
-Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of the two central lines to get a precise measurement of the Larmor frequency.
+Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of their two central lines to get a precise measurement of the Larmor frequency.
 In the region where the two ODMR lines overlap we have to use a linear regression to get the Larmor frequency of each class.
@@
 Fig.
-\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
+\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $\Gamma_1^{\rm dd}=T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
 For each magnetic field value, a $T_1$ profile was recorded and fitted following the procedure described in Fig.
@@
 
-We can clearly see that the increase in $T_1^{\rm dd}$ is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
+We can clearly see that the $\Gamma^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
 If we assume that both NV classes spectral response is a gaussian of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
@@
 The fluctuator model provides another explanation: the fluctuator spectral response is broaden by its very short lifetime.
-This also explains why the $1/T_1^{\rm dd}$ line is well matched by a Lorenzian, while the ODMR lines are closer to Gaussians in our sample.
+This also explains why the $1/T_1^{\rm dd}$ line is well matched by a Lorentzian, while the ODMR lines are closer to Gaussians in our sample.
 
@@
 
-\section{Trouver un bon titre}
+\section{Resonance between several NV classes}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
 \centering
@@
 
-\subsection{$T_1$ dependency on magnetic field}
+\subsection{$T_1$ dependency on magnetic field configuration}
 
-\section{Limitations of the fluctuator model}
+\begin{table}[htbp]
+\centering
+\caption{Theoretical and experimental values of $\Gamma_1^{\rm dd}=1/T_1^{\rm dd}$ for various magnetic field configurations, expressed as a function of the value found for an isolated class.
+The experimental values were measured on sample ADM-15-3 using the protocol of Fig.
+\ref{various_fit_formulae}-d) with $T_1^{\rm ph}\equiv 5\ \rm ms$.}
+ \label{T1 champ mag}
+\begin{tabular}{c|cc}
+\toprule
+$\Gamma_1^{\rm dd} (\mathbf{B})$ &  Theory & Experimental \\
+\midrule
+random $\mathbf{B}$ (1 class) & $\Gamma_0^{\rm th}$ & 1.53 ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & 5.20 $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & 4.22 $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & 9.07 $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & 11.3 $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & 15.4 $\Gamma_0^{\rm exp}$ \\
+\bottomrule
+\end{tabular}
+
+($^*$) The case $\mathbf{B}=0$ will be detailed in the next chapter. The theoretical value is a lower bond.
+\end{table}
+
+The fluctuator model allows quantitative predictions of the change in the spin lifetime, thanks to eq.
+\ref{eq 1/T}.
+There are however several unknown in this equation, such as the fluctuator density and lifetime which can be hard to estimate.
+It is however relatively easy to compare the predictions of eq.
+\ref{eq 1/T} for various magnetic field configurations.
+
+What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0 \equiv \Gamma_1^{\rm dd} (\rm{isolated class})$, and to express the decay rates when 2,3 or 4 classes are co-resonant as a function of $\Gamma_0$.
+Doing so allows to normalize $n_f$ and $\gamma_f$ in eq.
+\ref{eq 1/T} and to keep the geometrical factor $\bar \eta$ as the only unknown.
+
+Under certain assumptions, detailed in appendix REF, $\bar \eta$ can be analytically or numerically computed for each of the magnetic configurations detailed before.
+This allows us to predict the value of the spin decay rate in every situations as a function of the decay rate of an isolated class.
+
+Table \ref{T1 champ mag} shows these predicted values, along with experimentally measured ones on sample ADM-15-3.
+The experimental values were obtained with the protocol described in Fig.
+\ref{various_fit_formulae}-d) with a double exponential fit where we fixed again $T_1^{\rm dd}=5\ \rm ms$.
+
+While the general order of $\Gamma_1^{\rm dd} (\mathbf{B})$ is the same for the theoretical and experimental values, the theoretical values overestimate can significantly the experimental values, by more than a factor of 3 for the higher degeneracy scenarios.
+
+It should be noted that the experimental values were heavily influenced by the choice of the parameter $T_1^{\rm ph}$.
+Using for example $T_1^{\rm ph}=2.5\ \rm ms$ gave results closer to the theoretical ones.
+Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.
+\citep{choi2017depolarization} for instance found that a 2-class degeneracy (the author did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded a depolarization rate $\sim 4$ times higher than that of a single isolated class, a value comparable to the one we found in table \ref{T1 champ mag}.
+We can also see in Fig.
+\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration lead to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, compared to the predicted factor of 10.
+
+This failure of the fluctuator model to accurately predict the changes in $T_1$ is one of the reasons that led us to believe that the model is incomplete.
+
+\section{Conclusion and perspectives}
+
+\subsection{Limitations of the fluctuator model}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/Anormal_T1}
+\caption{$T_1$ measurements on sample ADM-15-4, fitted with pure exponential or stretched exponential.
+a) $T_1$ of a single isolated class when $\mathbf{B} \neq 0$.
+The optimal fit parameters are respectively $T_1^{\rm exp}=1.43\ \rm ms$ and $T_1^{\rm stretch}=0.56\ \rm ms$.
+b) $T_1$ of the four degenerate classes when $\mathbf{B}=0$.
+The optimal fit parameters are respectively $T_1^{\rm exp}=0.15\ \rm ms$ and $T_1^{\rm stretch}=0.05\ \rm ms$.}
+\label{anormal T1}
+\end{figure}
+
+I already mentioned that the discrepancy between the theoretical and experimental $T_1$  values were an issue of the current fluctuator model, although the discrepancy could come in part from the $T_1$ fitting procedure.
+Another, more glaring problem is the fact that some dipole-dipole mediated $T_1$ are not stretched exponential.
+
+Fig. \ref{anormal T1} shows
+
 La modification des T1, tjr plus petite que prévue (faire un tableau), et les T1 non-exponentiels (je crois que y'a des cas sur Ludo, bon courage mdr)

### 2022-08-10 data/20220222/Adamas 15 um/traitement.py

@@ -58,7 +58,7 @@ def random_plot():
 	plt.show()
 
 
-random_plot()
+# random_plot()
 def plot_champs_transverses():
 	fname='T1 champ transverse très faible classe 4 (DQ)'
 	x,y=extract_data(fname,ycol=5)
@@ -135,28 +135,81 @@ def fig121_Vs_22():
 	plt.show()
 
 
-def test_fits_t1_phonon():
+def test_fits_t1_phonon(fname, valRef=None,fixed=True):
 	#Bon après pas mal de test, c'est bien de la merde les fits avec stretch et phonon, le T1 stretch il danse vraiment la salsa et les valeurs c'est assez nimp
 	#C'est peut etre a cause du coté micro, il faudrait vraiment que je fasse plus de test sur les gros adamas pour en être sur. Et limite broyer des diamants aussi
 
 	#En attendant si tu veux des valeurs, mieux vaut prendre des stretch arb. alpha=0.8 fonctionne bien
-	T1ph=1/1800
+	# T1ph=1/1800
+	T1ph=2.11
 
-	fname='T1 1 classe pour le champ transverse moyen'
 
-	# fname='T1 2x2 autre raie'
-	# fname='T1 2x2'
-	# fname='T1 2x2 droite (2)'
-	# fname='T1 2x2 gauche (2)'
-	# fname='T1 121 nuit'
-	# fname='T1 121 (3)'
-	# fname='T1 121 (4)'
-	# fname='T1 1x2x1'
-	# fname='T1 0B court'
+	
 	x,y=extract_data(fname,ycol=5)
+	y=y/max(y)
+	x=x*1000
 	plt.plot(x,y)
-	# popt,yfit=stretch_et_phonons(x,y,T1ph=T1ph,fixed=False)
-	popt,yfit=stretch_arb_exp_fit_zero(x,y,alpha=0.8,fixed=True)
+	# popt,yfit=exp_fit(x,y)
+	popt,yfit=stretch_et_phonons(x,y,T1ph=T1ph,fixed=fixed)
+	# popt,yfit=stretch_arb_exp_fit_zero(x,y,alpha=0.8,fixed=True)
 	plt.plot(x,yfit)
-	print(1/popt[1])#,popt[2])
-	plt.show()
+	if valRef:
+		print(popt[1],valRef/popt[1])
+		return(popt[1],valRef/popt[1])
+	else :
[…]

### 2022-08-11 Manuscrit Thèse/chapter 3/Appendix_eta.tex
tata

--- 2022-08-08 Manuscrit Thèse/chapter 3/Chapter 3.tex
@@
 \begin{document}
-\chapter{NV-NV cross-relaxations : the fluctuator model}
-In this chapter,...
-
-\section{Experimental observation of NV-NV cross-relaxation}
-Before we discuss the theoretical complications related to NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR with dense NV ensembles ([NV]$\gtrsim$ 1 ppm).
-\subsection{NV-NV CR between nonequivalent NV centers}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/NV-NV_non_equivalent}
-\caption{label. Taken from \citep{armstrong2010nv}}
-\label{non equivalent NV-NV}
-\end{figure}
-NV-NV CR was first observed more than thirty years ago \citep{holliday1989optical, van1989cross}.
-The first observations were between non-equivalent NV centers, meaning that the two NV centers involved in the dipole-dipole coupling were not polarized equally.
-
-This scenario can happen for instance when two NV centers from different classes see a different transverse (and longitudinal) magnetic field.
-Fig \ref{non equivalent NV-NV} illustrates this in the case where the magnetic field is parallel with one of the four classes: $\mathbf{B} \parallel [111]$
-
-When $\mathbf{B} \parallel [111]$, as we discussed in the last chapter, one class sees no transverse field and is therefore always polarized.
-The three other (equivalent) classes on the other hand get more and more depolarized as the magnetic field amplitude is increased.
-
-It turns out that there is a co-resonnance at B=592 G between the class parallel to $\mathbf{B}$ and the three other classes.
-This co-resonance is represented in Fig.
-\ref{non equivalent NV-NV}-b) by an orange circle.
-
-Fig.
-\ref{non equivalent NV-NV}-c), which we already saw in the last chapter, shows the change in PL of an ensemble of NV centers as $\mathbf{B}$ is scanned along the [111] axis.
-For B=592 G, we can see a drop in PL also circled in orange.
-This drop is the result of the CR between NV centers from the class parallel to $\mathbf{B}$ and NV centers from the three other classes.
-
-While a difference in polarization is needed to observe a population transfer, it is not enough to explain the drop in PL.
-This PL drop is due to the difference in brightness between the $\ket{0}$ and $\ket{+1}$ from the different classes involved.
-The PL contrast between the two states is higher for the class parallel to $\mathbf{B}$ than for the three other ones, which is another consequence of the transverse field.
-
-The co-resonance between two different classes of NV centers with different magnetic field projection is a relatively rare event: it can only occur for the $\ket{0} \to \ket{+1}$ transition, and only for magnetic fields greater than 592 G.
-
-%techniquement y'a les resonances 0->-1 et -1/+1 mais c'est le delbor, j'ai pas envie d'embrouiller plus que ca.
-
-\subsection{NV-NV CR between equivalent NV centers}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/NV-NV_equivalent}
-\caption{CR between equivalent NV centers for sample ADM-15-1.
-(a) Representation of equivalent NV centers: either two NVs from the same class, or two classes with the same projected magnetic field.
-(b) Magnetic setup used for the experiment: a permanent magnet is used to apply a bias magnetic field, and an electromagnet is used to add a variable magnetic field.
-(c) Simulation of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers as a function of the scanned magnetic field.
-The transitions were computed based on several ODMR spectra.
-The red dotted lines correspond to inter-class resonance (d) Change in the PL of the NV centers ensemble as a function of the scanned magnetic field.
-The inter-class resonances are reported here.}
-\label{equivalent NV-NV}
-\end{figure}
-
-More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there was CR even between equivalent NV centers.
-
-I refer by equivalent NV centers to NV centers with the same spin Hamiltonian.
-This means either NV center from the same class, or NV centers from different classes with the same projection of the magnetic field along their axis.
-Equivalent NV centers are by nature resonant with each other, and therefore susceptible to CR.
-
-Fig.
-\ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV Cr.
-The key to observe NV-NV CR is to bring some NV centers in and out of resonance with other NV centers, which we can do with the different NV classes.
-To do so, we need to add an initial bias magnetic field in addition to the scanned magnetic field, as represented in Fig.
-\ref{equivalent NV-NV}-b).
-%When doing so, the magnetic field felt by the diamond won't have a fixed orientation during the scan, and when the magnetic field crosses certain crystalline planes (detailed below), its projection on two or more classes of NV centers will be the same.
-
-Fig.
-\ref{equivalent NV-NV}-c) shows the transition frequencies of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers on sample ADM-15-1, an HPHT micro-diamond with $[\rm NV]\sim 3\ \rm ppm$.
-There are three magnetic field values $B_{\rm scan} \sim$ 12, 42 and 77 G for which there is an inter-class resonance.
-Each of these resonances correspond to a "geometric" resonance where the $B$ field is in a symmetry plane between the two classes, meaning that the two resonant NV classes are indeed equivalent.
-
-Finally, Fig.
-\ref{equivalent NV-NV}-d) shows the PL contrast as the electromagnet field is scanned.
-There are 3 very clear dips when $B_{\rm scan} \sim$ 12, 42 and 77 G which coincide with inter-class resonances.
-We will see later that this PL dip is associated with a decrease of the NV centers spin $T_1$ time from both classes.
-
-This observation, and similar ones done by many groups, seem to indicate that there are CR between equivalent NV centers.
-This is however incompatible with our previous assumptions that equivalent NV centers are equally polarized and bright.
-To understand this phenomenon, we will need to introduce new hypotheses.
-
-\section{NV inhomogeneity and the fluctuator model}
-
-\subsection{CR in an inhomogeneous NV bath}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/inhomogene_2_spins}
-\caption{label}
-\label{inhomogene}
-\end{figure}
-A likely explanation to the observed NV-NV CR is that the NV centers are not all equivalents.
-
-Indeed, if we assume that, for instance, the relaxation rate $\Gamma_1=1/T_1$ is not strictly the same for each NV centers, but instead follows a certain distribution $\rho(\Gamma_1)$, then CR between NV centers with different $\Gamma_1$ can explain the observations in Fig.
-\ref{equivalent NV-NV}.
-
-Fig.
-\ref{inhomogene} illustrates this case.
-We are assuming here a coupling between two NV centers with respective relaxation rate $\Gamma_1$ and $\Gamma_1^*$.
-Fig.
-\ref{equivalent NV-NV}-a) illustrates the model I chose for this system: $\Gamma_{\rm las}$ represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins.
-I consider here only the incoherent dynamics of the population which I model with various rates.
-I also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.
-
-By finding the steady state of these rate equations, I can compute the final population in the $\ket{0}$ for both spins, which I will assume be proportional to the total PL.
-Fig.
-\ref{inhomogene}-b) shows this $\ket{0}$ population as a function of the flip-flop rate $\Gamma_{\rm dd}$ for various values of $\Gamma_1^*$.
-
-We can see that when $\Gamma_1^*=\Gamma_1$, as expected, the PL is not modified by the coupling of the two spins regardless of the coupling strength.
-When $\Gamma_1^*>\Gamma_1$ however, not only is the starting PL slightly lower since NV$_2$ is less polarized, but the PL now decreases as we increase the coupling strength.
-This effect, which is the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is higher.
-
-\subsection{Presentation of the fluctuator model}
-Choi et al.
-in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: "normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and "fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism.
-We are assuming that the fluctuator relaxation rate, $\gamma_f$ is much greater than the optical pumping rate $\Gamma_{\rm las}$.
-The fluctuators are therefore unpolarized and effectively act as dark spins, similarly to what we discussed in the last chapter.
-
-We should note here that, while this simplifications seem a bit extreme, the fluctuator model is more than a toy model.
-There are actually good evidence of the presence of these dark NV centers in dense NV ensemble, as will be discussed below.
-
-We will consider the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
-With this assumption, we can compute the modification caused by the fluctuators on the rest of the NV lifetimes.
-
-%\begin{pmatrix}
-%0.33&0&0 \\
-%0&0.33&0 \\
-%0&0&0.33
-%\end{pmatrix}
-
-\subsection{Single NV center coupled to a single fluctuator}
-We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
-To compute the depolarization induced by the fluctuator bath the NV centers, we should first consider the interaction between a single NV center and a fluctuator.
-Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.
-
-We will first decompose the dipole-dipole Hamiltonian in a radial and angular part:
-\begin{equation}
-\mathcal{H}_{\rm dd} \approx - \frac{J_0}{r^3} \left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}+qS_z^1 S_z^2 \right] + h.c. ,
-\end{equation}
-where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative orientation of the two dipoles.
-
-We then introduce the dimensionless number $\eta$ defined as:
-\begin{equation}
-\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
-\end{equation}
-where $\gamma_f=1/T_1^f$ is the fluctuator decay rate.
-This number $\eta$ encapsulate both the geometry of the problem (trhough $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
-
-Finally, the additional depolarization rate induced by the fluctuator on the NV center reads:
-\begin{equation}
-\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right) \frac{\eta^2}{\gamma_f}.
-\end{equation}
-Compared to eq.
-[REF], we have replaced here $\Gamma_2^*$ by $\gamma_f$.
-This is because we assume that the pure dephasing, supposedly equal for the NV center and the flucutator, is smaller than the depolarization rate $\gamma_f$.
-We will further discuss this hypothesis later.
-
-\subsection{Ensemble of NV centers coupled to a bath of fluctuators}
-
-We now need to compute the average depolarization caused by the ensemble of fluctuators on the ensemble of NV centers.
-
-We consider each fluctuator as in independant depolarization source, therefore the depolarization on a single NV center reads: 
-\begin{equation}
-\gamma_{\rm NV}=\sum_i \gamma_s(\mathbf{r}_i).
-\end{equation}
-We then want to compute the distribution $\rho(\gamma_{\rm NV})$ defined as:
-\begin{equation}
-\rho(\gamma_{\rm NV})=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma_{\rm NV} \right).
-\end{equation}
-To do so, we need to determine the distribution of the fluctuators positions $\{r_i\}$. Assuming that they are homogeneously distributed in the bulk of the material, we find:
-\begin{equation}
-\rho(\gamma_{\rm NV})=\frac{e^{-1/(4\gamma_{\rm NV} T)}}{\sqrt{4\pi \gamma_{\rm NV}^3 T}},
-\end{equation}
-where the time constant $T$ was introduced and is defined as:
-\begin{equation}
-\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f},
-\label{eq 1/T}
-\end{equation}
-with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{eta}$: $\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta$.
-
-Finally, the polarization dynamics from the ensemble of NV centers can be computed:
-\begin{equation}
-P(t)=\int_0^\infty \rho(\gamma_{\rm NV})\, e^{-\gamma_{\rm NV} t}d\gamma= e^{-\sqrt{t/T}}.
-\end{equation}
-
-In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential, unlike the phonon-induced depolarization which is purely exponential.
-
-It should be noted that the stretched exponential nature of the decay is not proper to the fluctuator model.
-\citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
-The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).
-
-\section{Experimental investigation of the fluctuator model}
-We will now show experimental results related to the predictions of the fluctuator model.
-\subsection{The stretched exponential lifetime}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/Stretch_vs_pas_stretch}
-\caption{T1 measurement following the protocol described in [REF].
-The best exponential and stretched exponential fits are given each time.
-a) Sample ADM-15-2 with B=0.
-The optimal $T_1$ values are $T_1^{\rm stretch}=150\ \mu \rm s$ and $T_1^{\rm exp}=410\ \mu \rm s$.
-b) Sample CVD-pink with B$\neq$0 and a single class probed.
-The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$}
-\label{stretch_or_not_stretch}
-\end{figure}
-$T_1$ measurements with dense ensemble of NV centers and when many classes are resonant with each other (typically when B=0) indeed tend to show stretched exponential profiles.
-
-Fig.
-\ref{stretch_or_not_stretch} shows an example of that: in Fig.
-\ref{stretch_or_not_stretch}-a) the $T_1$ measurement comes from sample ADM-15-2 at B=0.
-The profile is clearly better fitted by a stretched exponential than a regular exponential, and the timescale found $T_1^{\rm stretch}=150\ \mu \rm s$ is much shorter than the expected phonon limite lifetime of a few ms.
-
-On the other hand, Fig.
-\ref{stretch_or_not_stretch}-b) shows a $T_1$ measurement coming from sample CVD-pink for B$\neq$0.
-This time the profile is more exponential, and the value $T_1^{\rm exp}=4.1\ \rm ms$ is coherent with a phonon-limited lifetime.
-Even though CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.
-
-\subsection{Fitting $T_1$ profile in the general case}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/various_fit_formulae}
-\caption{Different fitting procedure on two $T_1$ measurements done on sample ADM-150-1 for $B=0$ and $B\neq0$.
-\\ a) Exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph})$.
-\\ b) Stretched exponential fit $S(\tau)=\exp (-\sqrt{\tau/T_1^{\rm dd}})$.
-\\ c) Bi-exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph} -\sqrt{\tau/T_1^{\rm dd}})$.
-\\ d) Bi-exponential fit with fixed $T_1^{\rm ph}=5\ \rm ms$: $S(\tau)=\exp (-\tau/5 -\sqrt{\tau/T_1^{\rm dd}})$.}
-\label{various_fit_formulae}
-\end{figure}
-
-\begin{table}[htbp]
-\centering
-\caption{Fitting parameters for Fig.
- \ref{various_fit_formulae}}
- \label{fitting table}
-\begin{tabular}{c|cc|cc}
-\toprule
-Figure &  \multicolumn{2}{c}{$B=0$} & \multicolumn{2}{c}{$B\neq0$}\\
-\midrule
-{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms)&  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)\\
-Fig. a) & 0.95 & * & 2.64 & * \\
-Fig. b) & * & 0.32 & * & 1.03  \\
-Fig. c) & 6.75 & 0.44 & 4.22 & 7.19 \\
-Fig. d) & \textbf{5} & 0.51 & \textbf{5} & 4.65 \\
-\bottomrule
-\end{tabular}
-
-Bold characters indicate parameters that were arbitrarily fixed.
-
-\end{table}
-
-For many samples however, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$ is of the same order as the phonon limited lifetime $T_1^{\rm ph}$.
-Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, forcing us to include both aspects if we want a unique fitting formula.
-
-Fig.
-\ref{various_fit_formulae} shows four different fitting procedure on the two same measurements.
-The measurements here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
-The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.
-
-The fits used here are either purely exponential or purely stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$.
-The values of the different fitting parameters used is reported in Table \ref{fitting table}.
-
-We can see that neither the purely exponential nor stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is poorly fitted by the stretched exponential.
-The protocols that include both exponential and stretched lifetimes correctly fit both curves.
-In the case of Fig.
-\ref{various_fit_formulae}-d), we arbitrarily fixed $T_1^{\rm ph}=5\ \rm ms$ since this is the value we typically measure on samples with low NV density, and we expect that the phonon-limited exponential lifetime is not modified by the NV concentration.
-
-Since the protocol where $T_1^{\rm ph}$ was fixed and the one where it wasn't both correctly fit our data, we decided to use the protocol where $T_1^{\rm ph}$ was fixed because there is one less free parameter in the fitting function, and the values we obtain on $T_1^{\rm dd}$ can be directly compared.
-
-We should note however that the values of $T_1^{\rm dd}$ we obtain are not absolute.
-Indeed, in the case of Fig.
-\ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
-While this technique of measuring $T_1^{\rm dd}$ is relatively sensitive and reproducible, it is not very accurate.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/betas}
-\caption{a) Same two measurements as Fig.
-\ref{various_fit_formulae} with the fitting formula $S(\tau)=\exp ((-\tau/T_1)^{\beta})$.
-The fitting parameters are $\beta=0.58$ and $T_1=0.46\ \rm ms$ for $B=0$, and $\beta=0.80$ and $T_1=2.15\ \rm ms$ for $B\neq0$.
-b) Optimal $\beta$ parameter found for each $T_1$ measurement as a function of the external magnetic field (still on sample ADM-150-1).}
-\label{betas}
-\end{figure}
-
-Finally, another fitting procedure is shown in Fig.
-\ref{betas}.
-This time we are using a single stretched exponential with an arbitrary stretch factor $\beta$.
-Fig.
-\ref{betas}-b) shows the optimal $\beta$ parameter as a function of the external magnetic field.
-We can then confirm that the $T_1$ profile gets closer to a stretched exponential ($\beta=0.5$) when $B$ goes to 0.
-
-Even though this method also yields satisfying fits, it needs two free parameters to work ($\beta$ and $T_1$), so we decided to go with the method of Fig. \ref{various_fit_formulae}-d) instead.
-
-\subsection{The fluctuators linewidth}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/fluctuator linewidth}
-\caption{Measurement of the fluctuators linewidth using the same setup as in Fig.
-\ref{equivalent NV-NV}.
-a) Simulation of the $\ket{0} \to \ket{-1}$ transition frequency for the four classes of NV centers and actual frequencies of the two central classes from ODMR measurements.
-b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.78\ \rm MHz$.
-The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor $\sqrt{2}$}
-\label{fluct linewidth}
-\end{figure}
-
-One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators linewidth.
-In the model, we assumed that fluctuator's lifetime $T_1^f=1/\gamma_f$ was so short that the fluctuators linewidth was lifetime-limited, whereas the normal NV centers linewidth is limited by the inhomogeneous broadening $T_2^*>T_1^f$.
-
-We can experimentally verify this hypothesis: the fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in the last chapter.
-
-Fig.
-\ref{fluct linewidth} shows an experiment similar to the one presented in Fig.
-\ref{equivalent NV-NV}.
-A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers, but this time measuring the spin decay rate instead of the PL.
-
-Fig.
-\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
-Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of the two central lines to get a precise measurement of the Larmor frequency.
-In the region where the two ODMR lines overlap we have to use a linear regression to get the Larmor frequency of each class.
-
-Fig.
-\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
-For each magnetic field value, a $T_1$ profile was recorded and fitted following the procedure described in Fig.
-\ref{various_fit_formulae}-d) where we again fixed $T_1^{\rm ph}=5\ \rm ms$.
-We also show the ODMR spectrum of a single NV class for comparison.
-
-We can clearly see that the increase in $T_1^{\rm dd}$ is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
-If we assume that both NV classes spectral response is a gaussian of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
-Fig.
-\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $\sqrt{2}$, which is still much narrower than the $1/T_1^{\rm dd}$ line.
-
-A potential explanation for the broadening of the $1/T_1^{\rm dd}$ line could be the interaction strength between the NV centers, but for [NV]$\sim$5 ppm, $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$ which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\textwidth]{Figures/deconvolution}
-\caption{Blue curve $1/T_1^{\rm dd}$ data presented in Fig.
-\ref{fluct linewidth}.
-Red curve: deconvolution of the blue curve the  by the ODMR line shown on the same figure.
-Green curve: deconvolution of the red curve the  by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38}
-\label{deconvolution}
-\end{figure}
-
-The fluctuator model provides another explanation: the fluctuator spectral response is broaden by its very short lifetime.
-This also explains why the $1/T_1^{\rm dd}$ line is well matched by a Lorenzian, while the ODMR lines are closer to Gaussians in our sample.
-
-We can now extract the fluctuator lifetime from the data in Fig. \ref{fluct linewidth} by taking out the contribution of $T_2^*$ on both the NV and the fluctuator spectral response.
-
-The first way to do this is to follow the instructions in \citep{hall2016detection} and deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum.
-The result of the deconvolution is shown in Fig.
-\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (red curve).
-Fitting with Lorentzians, we find a final value $\Gamma_2 =\frac{\gamma_f}{2} = (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=12.5\ \rm ns$.
-
-Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\sqrt{2 \ln (2)}\sigma$ and a Lorentzian of HWHM $f_L$:
-\begin{align*}
-\Gamma_1^{\rm dd}&\propto \mathcal{G}(f_G)*(\mathcal{L}(f_L)*\mathcal{G}(f_G)) \\
-&\propto \mathcal{L}(f_L)*\mathcal{G}(\sqrt{2} f_G),
-\end{align*}
-where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
-Then We can use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{wiki:VoigtProfile}:
-\begin{equation}
-f = [f_G^5 + 2.69269 f_G^4 f_L + 2.42843 f_G^3 f_L^2 + 4.47163 f_G^2 f_L^3 + 0.07842 f_G f_L^4 + f_L^5]^{1/5}.
-\end{equation}
-
-With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, I find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 12.2\ \rm ns$, a value consistent with the one found with the deconvolution method.
-
-Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=3.3\ \rm MHz$, although given their experimental values - full width Half maximum (FWHM) of 25 MHz for the lifetime profile and 9.3 MHz for the ODMR line - I obtain using my method $\gamma_f= 15.7\ \rm MHz$ and $T_1^f=10.1\ \rm ns$.
-The main difference comes from the fact that they considered the spectral half-width of the fluctuators to be $2\gamma_f$ whereas I consider it to be $\gamma_f/2$.
-
-\section{Microscopic explanation of the fluctuators}
-\subsection{Parallel with P-doped Si}
-A similar 
-"Electron Spin-Lattice Relaxation in Phosphorus-Doped Silicon"
-
-"On the interaction of nuclear spins in a crystalline lattice"
-
-"Concentration- and Compensation-Dependent Spin-Lattice Relaxation in n-Type Silicon"
-\subsection{Microscopic origin of the fluctuators}
-On va dire subsection plutot.
-Du coup ici discussion tunneling Vs Modulation of J Vs surface spins
-
-Attention quand même, ce process dépend e la température si je dis pas de bêtises, c'est p-e pas ça chez nous. Nan ok, ça modifie gammaf mais pas le T1 directement. P-e que c'est mesurable ?
-
-Attention bis : C'est l'exchange interaction (le terme en delta) qui est modifié, visiblement le couplage dipolaire est trop faible pour expliquer ça (papier de Pines et "Spin Resonance of Impurity Atoms in Silicon")
-
-\section{Geometric conditions for NV-NV resonance}
-En gros la carte, les 5 ESR en SI du PRL (lol) et le 121 Vs 22
-
-\section{Limitations of the fluctuator model}
-La modification des T1, tjr plus petite que prévue (faire un tableau), et les T1 non-exponentiels (je crois que y'a des cas sur Ludo, bon courage mdr)
-Parler aussi des alternatives au fluctuator model, ici ou dans la 2e section
+\chapter{Appendix: Computation of $\bar \eta$}
 

### 2022-08-11 Manuscrit Thèse/chapter 3/Chapter 3.tex
tata

@@
 \chapter{NV-NV cross-relaxations : the fluctuator model}
-In this chapter,...
+In this chapter we will look at the phenomenon of cross-relaxations between NV centers, both from an experimental and theoretical point of view.
+We will see how a model of fast and slow relaxing centers, the fluctuator model, can explain most of the experimental observations, and finally we will see some limitations of the model.
 
 \section{Experimental observation of NV-NV cross-relaxation}
-Before we discuss the theoretical complications related to NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR with dense NV ensembles ([NV]$\gtrsim$ 1 ppm).
+Before we discuss the theoretical implications of NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles ([NV]$\gtrsim$ 1 ppm).
 \subsection{NV-NV CR between nonequivalent NV centers}
@@
 \includegraphics[width=\textwidth]{Figures/NV-NV_non_equivalent}
-\caption{label. Taken from \citep{armstrong2010nv}}
+\caption{a) Representation of two non equivalent spins resonantly coupled.
+One spin is aligned with the magnetic field and therefore well polarized, while the second spin is not.
+b) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
+c) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
 \label{non equivalent NV-NV}
@@
 Fig.
-\ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV Cr.
+\ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV CR.
 The key to observe NV-NV CR is to bring some NV centers in and out of resonance with other NV centers, which we can do with the different NV classes.
@@
 \includegraphics[width=\textwidth]{Figures/inhomogene_2_spins}
-\caption{label}
+\caption{a) Schematics of the model: we consider a four level system consisting of 2 levels of 2 NV centers resonantly coupled through the flip-flop rate $\Gamma_{\rm dd}$.
+These two NV centers are each pumped in their respective $\ket{0}$ state via the green laser with a rate $\Gamma_{\rm las}$, and have their own relaxation rate $\Gamma_1$ and $\Gamma_1^*$ respectively.
+b) Average population of the two NV centers in the $\ket{0}$ state in the steady state as a function of $\Gamma_{\rm dd}$ and $\Gamma_1^*$.
+The steady state was computed by solving a classical rate equation where I fixed $\Gamma_{\rm las}=1000\ \rm Hz$ and $\Gamma_1=300\ \rm Hz$.}
 \label{inhomogene}
@@
 \ref{inhomogene} illustrates this case.
-We are assuming here a coupling between two NV centers with respective relaxation rate $\Gamma_1$ and $\Gamma_1^*$.
-Fig.
-\ref{equivalent NV-NV}-a) illustrates the model I chose for this system: $\Gamma_{\rm las}$ represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins.
-I consider here only the incoherent dynamics of the population which I model with various rates.
-I also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.
+Our model here consists of two NV centers resonantly coupled with two separate relaxation rates $\Gamma_1$ and $\Gamma_1^*$.
+To complete the model, we add the rate $\Gamma_{\rm las}$ which represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins.
+We only consider here the incoherent dynamics of the population, modeled with rates.
+We also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.
 
-By finding the steady state of these rate equations, I can compute the final population in the $\ket{0}$ for both spins, which I will assume be proportional to the total PL.
+By finding the steady state of these rate equations, we can compute the final population in the $\ket{0}$ for both spins, which we will assume to be proportional to the total PL.
 Fig.
@@
 When $\Gamma_1^*>\Gamma_1$ however, not only is the starting PL slightly lower since NV$_2$ is less polarized, but the PL now decreases as we increase the coupling strength.
-This effect, which is the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is higher.
+This effect, which is at the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is important.
 
@@
 
-Other possibilities for the dipole-dipole dependent relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or superradiance, but they were not considered to be viable explanations \citep{choi2017depolarization}
+Other possibilities to explain the change in the relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or superradiance, but they were not considered to be viable explanations \citep{choi2017depolarization}.
 %\begin{pmatrix}
@@
 where $\gamma_f=1/T_1^f$ is the fluctuator decay rate.
-This number $\eta$ encapsulate both the geometry of the problem (trhough $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
+This number $\eta$ encapsulates both the geometry of the problem (through $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
 
@@
 
-We consider each fluctuator as in independant depolarization source, therefore the depolarization on a single NV center reads: 
+We consider each fluctuator as in independent depolarization source, therefore the depolarization on a single NV center reads: 
 \begin{equation}
@@
 While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and supposedly resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances.
-We will discuss here the various ...
+We will discuss here the various scenarios where NV classes overlap, and the quantitative impact it has on the centers' spin $T_1$.
 
@@
 
-\subsection{$T_1$ dependency on magnetic field configuration}
+\subsection{Quantitative modification of $T_1$}
 
@@
 Using for example $T_1^{\rm ph}=2.5\ \rm ms$ gave results closer to the theoretical ones.
-Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.
-\citep{choi2017depolarization} for instance found that a 2-class degeneracy (the author did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded a depolarization rate $\sim 4$ times higher than that of a single isolated class, a value comparable to the one we found in table \ref{T1 champ mag}.
+Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded a depolarization rate $\sim 4$ times higher than that of a single isolated class, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one.
 We can also see in Fig.
-\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration lead to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, compared to the predicted factor of 10.
+\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration lead to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, 4 times smaller than the predicted factor of 10.
 
-This failure of the fluctuator model to accurately predict the changes in $T_1$ is one of the reasons that led us to believe that the model is incomplete.
+The failure of the fluctuator model to accurately predict the changes in $T_1$ is one of the reasons that led us to believe that the model is incomplete.
 
 \section{Conclusion and perspectives}
+
+In conclusion, we have seen in this chapter that dense NV ensemble present "anormal" CR between the NV centers, and we have shown how the fluctuator model developed by Choi et al.
+in \citep{choi2017depolarization} can explain this behavior.
+The fluctuator model not only explains the NV-NV CR, but also correctly predicts the stretched exponential lifetime profile.
+It also correctly predicts that the fluctuators have a broader spectral response than the NV centers.
+Nevertheless, the model does not describe perfectly every aspects of NV-NV relaxation, and still has some unknown such as the very nature of the fluctuators.
+
+In what follows I will describe what I think would be interesting investigation points to further understand the NV-NV relaxation and the fluctuator model.
 
@@
 
-I already mentioned that the discrepancy between the theoretical and experimental $T_1$  values were an issue of the current fluctuator model, although the discrepancy could come in part from the $T_1$ fitting procedure.
-Another, more glaring problem is the fact that some dipole-dipole mediated $T_1$ are not stretched exponential.
+I already mentioned that the discrepancy between the theoretical and experimental $T_1$  values were an issue of the current fluctuator model, although the difference could come in part from the $T_1$ fitting procedure.
+Another, more glaring problem is the fact that some $T_1$ profile are not stretched exponential.
 
-Fig. \ref{anormal T1} shows
+Fig.
+\ref{anormal T1} shows $T_1$ measurement on sample ADM-15-4, either on a single isolated class ,where the dipole-dipole interactions are at their weakest, or on all four resonant classes at $\mathbf{B}=0$, where the interactions are at their strongest.
 
-La modification des T1, tjr plus petite que prévue (faire un tableau), et les T1 non-exponentiels (je crois que y'a des cas sur Ludo, bon courage mdr)
-Parler aussi des alternatives au fluctuator model, ici ou dans la 2e section
-Bon du coup parler du fait que ça n'inclut pas la diffusion entre NV ni la saturation des fluctuateurs. A voir si ça mérite une full section ou une gross ouverture à la fin.
+For the single isolated class, the $T_1$ profile seems to be closer to a stretch exponential, which is coherent with the fact that the lifetime $T_1 \sim 1\ \rm ms$ is already smaller than the expected phonon limited lifetime $T_1^{\rm ph} \sim 5\ \rm ms$.
+However when $\mathbf{B}=0$, although the lifetime gets $\sim$ 10 times shorter, the profile becomes clearly exponential.
+
+The spin dynamics in this sample is still dominated by resonant dipole-dipole interactions: the more classes are brought to resonance, the shorter the spin $T_1$.
+But here, the profile gets closer to a pure exponential when $T_1$ decreases, which is in complete contradiction with the fluctuator model.
+
+This particular sample was a 15 $\mu$m micro-diamond, which could have a role in this observation since we know that spins near the surface have a reduced $T_1$ \citep{rosskopf2014investigation}, and micro-diamonds have proportionally more near surface spins.
+But other diamonds from the same batch, such as the one measured in Fig.
+\ref{stretch_or_not_stretch}, show clearly stretched exponential profile in the same conditions, and meanwhile some bulk diamonds show very short non-stretched exponential lifetimes, similar to the present sample.
+
+It seems therefore that the current fluctuator model does not correctly describe the spin dynamics of every NV-dense samples.
+Looking at what was previously done for P-doped SI \citep{vugmeister1978spin} or for resonant energy transfer between molecules \citep{yokota1967effects}, the fluctuator model could probably be improved by taking into account the finite lifetime of the fluctuators (i.e.
+the saturation of the NV-fluctuator CR), and the spin diffusion between NV centers (the fluctuator model only considers NV-fluctator interactions and not NV-NV interactions).
+
+\subsection{Investigation of the fluctuators nature}
+While the hypothesis of fluctuators being clusters of NV centers or NV-other spins is solidified by the precedents in P-doped Si \citep{honig1960electron} or F centers in KCl \citep{warren1964spin}, there are still many unknowns on the exact nature of the fluctuators.
+There are at least two aspects that I think should be investigated.
+
+The first one would be to study the fluctuator spectral profile on various samples showing NV-NV CR.
+In particular it would be interesting to know whether this profile is always Lorentzian, and what its linewidth is.
+In the sample I studied on Fig.
+\ref{fluct linewidth}, I found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough my sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 45 ppm$ sample used in \citep{choi2017depolarization}.
+If $\gamma_f$ was found to be somewhat of a constant among all samples showing NV-NV CR, this would give us a lot of information on the potential nature of the fluctuators.
+Studying the temperature dependency of $\gamma_f$ could also give us information on whether the depolarization mechanism is due to tunneling or modulation of the $J-$coupling by the phonons.
+
+The other aspect would be to identify potential defects that could form fluctuators when paired with NV centers.
+The main candidate for this would be substitutional nitrogen (P1 or P1$^+$ centers) since they are the most abundant electronic spin species in the diamond we use.
+The ideal way to confirm this hypothesis would be to compare the spin $T_1$ for  different samples with the same [NV] concentration, but different [P1] concentration.
+If indeed P1 centers play a role in the presence of fluctuators, we would expect a shorter NV $T_1$ time in the sample with high P1 concentration.
 

### 2022-08-12 Manuscrit Thèse/chapter 4/Chapter 4.tex
codi coda

--- 2022-08-09 Manuscrit Thèse/chapter 3/Chapter 3.tex
@@
 ]{biblatex} %Imports biblatex package
-\addbibresource{Bib_ch3.bib} %Import the bibliography file
+\addbibresource{Bib_ch4.bib} %Import the bibliography file
 
@@
 \begin{document}
-\chapter{NV-NV cross-relaxations : the fluctuator model}
-In this chapter,...
+\chapter{NV-NV CR under transverse or low fields: application to magnetometry}
 
-\section{Experimental observation of NV-NV cross-relaxation}
-Before we discuss the theoretical complications related to NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR with dense NV ensembles ([NV]$\gtrsim$ 1 ppm).
-\subsection{NV-NV CR between nonequivalent NV centers}
-\label{non_equi_valent_CR}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/NV-NV_non_equivalent}
-\caption{label. Taken from \citep{armstrong2010nv}}
-\label{non equivalent NV-NV}
-\end{figure}
-NV-NV CR was first observed more than thirty years ago \citep{holliday1989optical, van1989cross}.
-The first observations were between non-equivalent NV centers, meaning that the two NV centers involved in the dipole-dipole coupling were not polarized equally.
+\section{NV spin Hamiltonian under low and transverse fields}
 
-This scenario can happen for instance when two NV centers from different classes see a different transverse (and longitudinal) magnetic field.
-Fig \ref{non equivalent NV-NV} illustrates this in the case where the magnetic field is parallel with one of the four classes: $\mathbf{B} \parallel [111]$
+\section{Modification of NV-NV CR in the $\{ \ket{0}, \ket{+}, \ket{-} \}$ basis}
 
-When $\mathbf{B} \parallel [111]$, as we discussed in the last chapter, one class sees no transverse field and is therefore always polarized.
-The three other (equivalent) classes on the other hand get more and more depolarized as the magnetic field amplitude is increased.
+\section{Low field magnetometry}
 
-It turns out that there is a co-resonnance at B=592 G between the class parallel to $\mathbf{B}$ and the three other classes.
-This co-resonance is represented in Fig.
-\ref{non equivalent NV-NV}-b) by an orange circle.
-
-Fig.
-\ref{non equivalent NV-NV}-c), which we already saw in the last chapter, shows the change in PL of an ensemble of NV centers as $\mathbf{B}$ is scanned along the [111] axis.
-For B=592 G, we can see a drop in PL also circled in orange.
-This drop is the result of the CR between NV centers from the class parallel to $\mathbf{B}$ and NV centers from the three other classes.
-
-While a difference in polarization is needed to observe a population transfer, it is not enough to explain the drop in PL.
-This PL drop is due to the difference in brightness between the $\ket{0}$ and $\ket{+1}$ from the different classes involved.
-The PL contrast between the two states is higher for the class parallel to $\mathbf{B}$ than for the three other ones, which is another consequence of the transverse field.
-
-The co-resonance between two different classes of NV centers with different magnetic field projection is a relatively rare event: it can only occur for the $\ket{0} \to \ket{+1}$ transition, and only for magnetic fields greater than 592 G.
-
-%techniquement y'a les resonances 0->-1 et -1/+1 mais c'est le delbor, j'ai pas envie d'embrouiller plus que ca.
-
-\subsection{NV-NV CR between equivalent NV centers}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/NV-NV_equivalent}
-\caption{CR between equivalent NV centers for sample ADM-15-1.
-(a) Representation of equivalent NV centers: either two NVs from the same class, or two classes with the same projected magnetic field.
-(b) Magnetic setup used for the experiment: a permanent magnet is used to apply a bias magnetic field, and an electromagnet is used to add a variable magnetic field.
-(c) Simulation of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers as a function of the scanned magnetic field.
-The transitions were computed based on several ODMR spectra.
-The red dotted lines correspond to inter-class resonance (d) Change in the PL of the NV centers ensemble as a function of the scanned magnetic field.
-The inter-class resonances are reported here.}
-\label{equivalent NV-NV}
-\end{figure}
-
-More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there was CR even between equivalent NV centers.
-
-I refer by equivalent NV centers to NV centers with the same spin Hamiltonian.
-This means either NV center from the same class, or NV centers from different classes with the same projection of the magnetic field along their axis.
-Equivalent NV centers are by nature resonant with each other, and therefore susceptible to CR.
-
-Fig.
-\ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV Cr.
-The key to observe NV-NV CR is to bring some NV centers in and out of resonance with other NV centers, which we can do with the different NV classes.
-To do so, we need to add an initial bias magnetic field in addition to the scanned magnetic field, as represented in Fig.
-\ref{equivalent NV-NV}-b).
-%When doing so, the magnetic field felt by the diamond won't have a fixed orientation during the scan, and when the magnetic field crosses certain crystalline planes (detailed below), its projection on two or more classes of NV centers will be the same.
-
-Fig.
-\ref{equivalent NV-NV}-c) shows the transition frequencies of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers on sample ADM-15-1, an HPHT micro-diamond with $[\rm NV]\sim 3\ \rm ppm$.
-There are three magnetic field values $B_{\rm scan} \sim$ 12, 42 and 77 G for which there is an inter-class resonance.
-Each of these resonances correspond to a "geometric" resonance where the $B$ field is in a symmetry plane between the two classes, meaning that the two resonant NV classes are indeed equivalent.
-
-Finally, Fig.
-\ref{equivalent NV-NV}-d) shows the PL contrast as the electromagnet field is scanned.
-There are 3 very clear dips when $B_{\rm scan} \sim$ 12, 42 and 77 G which coincide with inter-class resonances.
-We will see later that this PL dip is associated with a decrease of the NV centers spin $T_1$ time from both classes.
-
-This observation, and similar ones done by many groups, seem to indicate that there are CR between equivalent NV centers.
-This is however incompatible with our previous assumptions that equivalent NV centers are equally polarized and bright.
-To understand this phenomenon, we will need to introduce new hypotheses.
-
-\section{NV inhomogeneity and the fluctuator model}
-
-\subsection{CR in an inhomogeneous NV bath}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/inhomogene_2_spins}
-\caption{label}
-\label{inhomogene}
-\end{figure}
-A likely explanation to the observed NV-NV CR is that the NV centers are not all equivalents.
-
-Indeed, if we assume that, for instance, the relaxation rate $\Gamma_1=1/T_1$ is not strictly the same for each NV centers, but instead follows a certain distribution $\rho(\Gamma_1)$, then CR between NV centers with different $\Gamma_1$ can explain the observations in Fig.
-\ref{equivalent NV-NV}.
-
-Fig.
-\ref{inhomogene} illustrates this case.
-We are assuming here a coupling between two NV centers with respective relaxation rate $\Gamma_1$ and $\Gamma_1^*$.
-Fig.
-\ref{equivalent NV-NV}-a) illustrates the model I chose for this system: $\Gamma_{\rm las}$ represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins.
-I consider here only the incoherent dynamics of the population which I model with various rates.
-I also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.
-
-By finding the steady state of these rate equations, I can compute the final population in the $\ket{0}$ for both spins, which I will assume be proportional to the total PL.
-Fig.
-\ref{inhomogene}-b) shows this $\ket{0}$ population as a function of the flip-flop rate $\Gamma_{\rm dd}$ for various values of $\Gamma_1^*$.
-
-We can see that when $\Gamma_1^*=\Gamma_1$, as expected, the PL is not modified by the coupling of the two spins regardless of the coupling strength.
-When $\Gamma_1^*>\Gamma_1$ however, not only is the starting PL slightly lower since NV$_2$ is less polarized, but the PL now decreases as we increase the coupling strength.
-This effect, which is the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is higher.
-
-\subsection{Presentation of the fluctuator model}
-Choi et al.
-in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: "normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and "fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism.
-We are assuming that the fluctuator relaxation rate, $\gamma_f$ is much greater than the optical pumping rate $\Gamma_{\rm las}$.
-The fluctuators are therefore unpolarized and effectively act as dark spins, similarly to what we discussed in the last chapter.
-
-We should note here that, while this simplifications seem a bit extreme, the fluctuator model is more than a toy model.
-There are actually good evidence of the presence of these dark NV centers in dense NV ensemble, as will be discussed below.
-
-We will consider the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
-With this assumption, we can compute the modification caused by the fluctuators on the rest of the NV lifetimes.
-
-Other possibilities for the dipole-dipole dependent relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or superradiance, but they were not considered to be viable explanations \citep{choi2017depolarization}
-%\begin{pmatrix}
-%0.33&0&0 \\
-%0&0.33&0 \\
-%0&0&0.33
-%\end{pmatrix}
-
-\subsection{Single NV center coupled to a single fluctuator}
-We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
-To compute the depolarization induced by the fluctuator bath the NV centers, we should first consider the interaction between a single NV center and a fluctuator.
-Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.
-
-We will first decompose the dipole-dipole Hamiltonian in a radial and angular part:
-\begin{equation}
-\mathcal{H}_{\rm dd} \approx - \frac{J_0}{r^3} \left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}+qS_z^1 S_z^2 \right] + h.c. ,
-\end{equation}
-where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative orientation of the two dipoles.
-
-We then introduce the dimensionless number $\eta$ defined as:
-\begin{equation}
-\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
-\end{equation}
-where $\gamma_f=1/T_1^f$ is the fluctuator decay rate.
-This number $\eta$ encapsulate both the geometry of the problem (trhough $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
-
-Finally, the additional depolarization rate induced by the fluctuator on the NV center reads:
-\begin{equation}
-\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right) \frac{\eta^2}{\gamma_f}.
-\end{equation}
-Compared to eq.
-[REF], we have replaced here $\Gamma_2^*$ by $\gamma_f$.
-This is because we assume that the pure dephasing, supposedly equal for the NV center and the flucutator, is smaller than the depolarization rate $\gamma_f$.
-We will further discuss this hypothesis later.
-
-\subsection{Ensemble of NV centers coupled to a bath of fluctuators}
-
-We now need to compute the average depolarization caused by the ensemble of fluctuators on the ensemble of NV centers.
-
-We consider each fluctuator as in independant depolarization source, therefore the depolarization on a single NV center reads: 
-\begin{equation}
-\gamma=\sum_i \gamma_s(\mathbf{r}_i).
-\end{equation}
-We then want to compute the distribution $\rho(\gamma)$ defined as:
-\begin{equation}
-\rho(\gamma)=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma \right).
-\end{equation}
-To do so, we need to determine the distribution of the fluctuators positions $\{r_i\}$. Assuming that they are homogeneously distributed in the bulk of the material, we find:
-\begin{equation}
-\rho(\gamma)=\frac{e^{-1/(4\gamma T)}}{\sqrt{4\pi \gamma^3 T}},
-\end{equation}
-where the time constant $T$ was introduced and is defined as:
-\begin{equation}
-\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f},
-\label{eq 1/T}
-\end{equation}
-with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{\eta}$: $\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta$.
-
-Finally, the polarization dynamics from the ensemble of NV centers can be computed:
-\begin{equation}
-P(t)=\int_0^\infty \rho(\gamma)\, e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}.
-\end{equation}
-
-In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential, unlike the phonon-induced depolarization which is purely exponential.
-
-It should be noted that the stretched exponential nature of the decay is not proper to the fluctuator model.
-\citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
-The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).
-
-\subsection{Possible microscopic origin of the fluctuators}
-
-We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}.
-It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.
-
-This conclusion is based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions, and the resonant energy transfer between molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through dipole-dipole interaction to non-excited acceptors, and where the stretched exponential dynamics of the donors was first observed \citep{forster1949experimentelle}.
-
-In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
-In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
-Both of these phenomena could lead to fast spin relaxation, the tunneling because the spin hopping is accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction strength $J$ by the crystal phonons \citep{honig1960electron}.
-Technically, the non-contact dipole-dipole interaction is also modulated by the phonons, but the effect is too small to account for the fast depolarization.
-
-Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
-To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate $\sim 10\ \rm ns$.
-In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
-
-\section{Experimental investigation of the fluctuator model}
-We will now show experimental results related to the predictions of the fluctuator model.
-\subsection{The stretched exponential lifetime}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/Stretch_vs_pas_stretch}
-\caption{T1 measurement following the protocol described in [REF].
-The best exponential and stretched exponential fits are given each time.
-a) Sample ADM-15-2 with B=0.
-The optimal $T_1$ values are $T_1^{\rm stretch}=150\ \mu \rm s$ and $T_1^{\rm exp}=410\ \mu \rm s$.
-b) Sample CVD-pink with B$\neq$0 and a single class probed.
-The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$}
-\label{stretch_or_not_stretch}
-\end{figure}
-$T_1$ measurements with dense ensemble of NV centers and when many classes are resonant with each other (typically when B=0) indeed tend to show stretched exponential profiles.
-
-Fig.
-\ref{stretch_or_not_stretch} shows an example of that: in Fig.
-\ref{stretch_or_not_stretch}-a) the $T_1$ measurement comes from sample ADM-15-2 at B=0.
-The profile is clearly better fitted by a stretched exponential than a regular exponential, and the timescale found $T_1^{\rm stretch}=150\ \mu \rm s$ is much shorter than the expected phonon limite lifetime of a few ms.
-
-On the other hand, Fig.
-\ref{stretch_or_not_stretch}-b) shows a $T_1$ measurement coming from sample CVD-pink for B$\neq$0.
-This time the profile is more exponential, and the value $T_1^{\rm exp}=4.1\ \rm ms$ is coherent with a phonon-limited lifetime.
-Even though CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.
-
-\subsection{Fitting $T_1$ profile in the general case}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/various_fit_formulae}
-\caption{Different fitting procedure on two $T_1$ measurements done on sample ADM-150-1 for $B=0$ and $B\neq0$.
-\\ a) Exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph})$.
-\\ b) Stretched exponential fit $S(\tau)=\exp (-\sqrt{\tau/T_1^{\rm dd}})$.
-\\ c) Bi-exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph} -\sqrt{\tau/T_1^{\rm dd}})$.
-\\ d) Bi-exponential fit with fixed $T_1^{\rm ph}=5\ \rm ms$: $S(\tau)=\exp (-\tau/5 -\sqrt{\tau/T_1^{\rm dd}})$.}
-\label{various_fit_formulae}
-\end{figure}
-
-\begin{table}[htbp]
-\centering
-\caption{Fitting parameters for Fig.
- \ref{various_fit_formulae}}
- \label{fitting table}
-\begin{tabular}{c|cc|cc}
-\toprule
-Figure &  \multicolumn{2}{c}{$B=0$} & \multicolumn{2}{c}{$B\neq0$}\\
-\midrule
-{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms)&  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)\\
-Fig. a) & 0.95 & * & 2.64 & * \\
-Fig. b) & * & 0.32 & * & 1.03  \\
-Fig. c) & 6.75 & 0.44 & 4.22 & 7.19 \\
-Fig. d) & \textbf{5} & 0.51 & \textbf{5} & 4.65 \\
-\bottomrule
-\end{tabular}
-
-Bold characters indicate parameters that were arbitrarily fixed.
-
-\end{table}
-
-For many samples however, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$ is of the same order as the phonon limited lifetime $T_1^{\rm ph}$.
-Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, forcing us to include both aspects if we want a unique fitting formula.
-
-Fig.
-\ref{various_fit_formulae} shows four different fitting procedure on the two same measurements.
-The measurements here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
-The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.
-
-The fits used here are either purely exponential or purely stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$.
-The values of the different fitting parameters used is reported in Table \ref{fitting table}.
-
-We can see that neither the purely exponential nor stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is poorly fitted by the stretched exponential.
-The protocols that include both exponential and stretched lifetimes correctly fit both curves.
-In the case of Fig.
-\ref{various_fit_formulae}-d), we arbitrarily fixed $T_1^{\rm ph}=5\ \rm ms$ since this is the value we typically measure on samples with low NV density, and we expect that the phonon-limited exponential lifetime is not modified by the NV concentration.
-
-Since the protocol where $T_1^{\rm ph}$ was fixed and the one where it wasn't both correctly fit our data, we decided to use the protocol where $T_1^{\rm ph}$ was fixed because there is one less free parameter in the fitting function, and the values we obtain on $T_1^{\rm dd}$ can be directly compared.
-
-We should note however that the values of $T_1^{\rm dd}$ we obtain are not absolute.
-Indeed, in the case of Fig.
-\ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
-While this technique of measuring $T_1^{\rm dd}$ is relatively sensitive and reproducible, it is not very accurate.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/betas}
-\caption{a) Same two measurements as Fig.
-\ref{various_fit_formulae} with the fitting formula $S(\tau)=\exp ((-\tau/T_1)^{\beta})$.
-The fitting parameters are $\beta=0.58$ and $T_1=0.46\ \rm ms$ for $B=0$, and $\beta=0.80$ and $T_1=2.15\ \rm ms$ for $B\neq0$.
-b) Optimal $\beta$ parameter found for each $T_1$ measurement as a function of the external magnetic field (still on sample ADM-150-1).}
-\label{betas}
-\end{figure}
-
-Finally, another fitting procedure is shown in Fig.
-\ref{betas}.
-This time we are using a single stretched exponential with an arbitrary stretch factor $\beta$.
-Fig.
-\ref{betas}-b) shows the optimal $\beta$ parameter as a function of the external magnetic field.
-We can then confirm that the $T_1$ profile gets closer to a stretched exponential ($\beta=0.5$) when $B$ goes to 0.
-
-Even though this method also yields satisfying fits, it needs two free parameters to work ($\beta$ and $T_1$), so we decided to go with the method of Fig. \ref{various_fit_formulae}-d) instead.
-
-\subsection{The fluctuators linewidth}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/fluctuator linewidth}
-\caption{Measurement of the fluctuators linewidth using the same setup as in Fig.
-\ref{equivalent NV-NV}.
-a) Simulation of the $\ket{0} \to \ket{-1}$ transition frequency for the four classes of NV centers and actual frequencies of the two central classes from ODMR measurements.
-b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.78\ \rm MHz$.
-The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor $\sqrt{2}$}
-\label{fluct linewidth}
-\end{figure}
-
-One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators linewidth.
-In the model, we assumed that fluctuator's lifetime $T_1^f=1/\gamma_f$ was so short that the fluctuators linewidth was lifetime-limited, whereas the normal NV centers linewidth is limited by the inhomogeneous broadening $T_2^*>T_1^f$.
-
-We can experimentally verify this hypothesis: the fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in the last chapter.
-
-Fig.
-\ref{fluct linewidth} shows an experiment similar to the one presented in Fig.
-\ref{equivalent NV-NV}.
-A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers, but this time measuring the spin decay rate instead of the PL.
-
-Fig.
-\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
-Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of the two central lines to get a precise measurement of the Larmor frequency.
-In the region where the two ODMR lines overlap we have to use a linear regression to get the Larmor frequency of each class.
-
-Fig.
-\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
-For each magnetic field value, a $T_1$ profile was recorded and fitted following the procedure described in Fig.
-\ref{various_fit_formulae}-d) where we again fixed $T_1^{\rm ph}=5\ \rm ms$.
-We also show the ODMR spectrum of a single NV class for comparison.
-
-We can clearly see that the increase in $T_1^{\rm dd}$ is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
-If we assume that both NV classes spectral response is a gaussian of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
-Fig.
-\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $\sqrt{2}$, which is still much narrower than the $1/T_1^{\rm dd}$ line.
-
-A potential explanation for the broadening of the $1/T_1^{\rm dd}$ line could be the interaction strength between the NV center and the fluctuator.
-However, for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\textwidth]{Figures/deconvolution}
-\caption{Blue curve $1/T_1^{\rm dd}$ data presented in Fig.
-\ref{fluct linewidth}.
-Red curve: deconvolution of the blue curve the  by the ODMR line shown on the same figure.
-Green curve: deconvolution of the red curve the  by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38 MHz}
-\label{deconvolution}
-\end{figure}
-
-The fluctuator model provides another explanation: the fluctuator spectral response is broaden by its very short lifetime.
-This also explains why the $1/T_1^{\rm dd}$ line is well matched by a Lorenzian, while the ODMR lines are closer to Gaussians in our sample.
-
-We can now extract the fluctuator lifetime from the data in Fig. \ref{fluct linewidth} by taking out the contribution of $T_2^*$ on both the NV and the fluctuator spectral response.
-
-The first way to do this is to follow the instructions in \citep{hall2016detection} and deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm.
-The result of the deconvolution is shown in Fig.
-\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve).
-Fitting with Lorentzians, we find a final value $\Gamma_2 =\frac{\gamma_f}{2} = (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=12.5\ \rm ns$.
-
-Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\sqrt{2 \ln (2)}\sigma$ and a Lorentzian of HWHM $f_L$:
-\begin{align*}
-\Gamma_1^{\rm dd}&\propto \mathcal{G}(f_G)*(\mathcal{L}(f_L)*\mathcal{G}(f_G)) \\
-&\propto \mathcal{L}(f_L)*\mathcal{G}(\sqrt{2} f_G),
-\end{align*}
-where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
-We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{wiki:VoigtProfile}:
-\begin{equation}
-f = [f_G^5 + 2.69269 f_G^4 f_L + 2.42843 f_G^3 f_L^2 + 4.47163 f_G^2 f_L^3 + 0.07842 f_G f_L^4 + f_L^5]^{1/5}.
-\end{equation}
-
-With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, I find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 12.2\ \rm ns$, a value consistent with the one found with the deconvolution method.
-
-Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=3.3\ \rm MHz$, although given their experimental values - full width at half maximum (FWHM) of 25 MHz for the lifetime profile and 9.3 MHz for the ODMR line - I obtain using my method $\gamma_f= 15.7\ \rm MHz$ and $T_1^f=10.1\ \rm ns$.
-The main difference comes from the fact that they considered the spectral half-width of the fluctuators to be $2\gamma_f$ whereas I consider it to be $\gamma_f/2$.
-
-\section{Trouver un bon titre}
-
-The main tool to probe NV-NV CR is to use the co-resonances between the NV classes.
-While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and supposedly resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances.
-We will discuss here the various ...
-
-\subsection{Geometric conditions for inter-class resonance}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{Figures/Resonance_geometries}
-\caption{ODMR spectra on sample ADM-15-3 as a function of the magnetic field orientation.
-The field amplitude is $\sim 60\ \rm G$.
-Red numbers represent the number of resonant classes for each ODMR line.
-The cubes represent the diamond unit cell, and the red planes/arrow the possible orientations of the magnetic field}
-\label{ODMR_geometries}
-\end{figure}
-
-There are four possibilities to obtain inter-class resonance (outside of the non-equivalent CR discussed in sec. \ref{non_equi_valent_CR}). These possibilities are:
-\begin{itemize}
-\item $\mathbf{B} \in \{110\}$.
-We refer by $\{110\}$ as any plane orthogonal to a $\langle 110 \rangle$ or equivalent direction.
-In this magnetic configuration, two classes are at resonance and the two other ones are spectrally isolated.
-\item $\mathbf{B} \in \{100\}$. In this magnetic configuration, the four classes form two pairs of co-resonance.
-\item $\mathbf{B} \parallel \langle 111 \rangle$.
-In this configuration discussed in the last chapter, one class is aligned with the magnetic field and is spectrally isolated while the three other form a resonant triplet.
-\item $\mathbf{B} \parallel \langle 100 \rangle$. In this configuration also discussed in the last chapter, the four classes are resonant.
-\end{itemize}
-We should add to this list the case $\mathbf{B}=0$ where all four classes are also resonant, and any other possibility (which I will call random $\mathbf{B}$) where all four classes are spectrally separated.
-
-An ODMR spectrum and a representation of the magnetic field in the diamond unit cell for each of these six possibilities is shown in Fig. \ref{ODMR_geometries}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/carte}
-\caption{PL contrast as a function of the magnetic field polar and azimuthal angle with respect to the [100] axis on sample CVD-pink.
-The field amplitude is $\abs{B} \sim 115 G$.
-The locus of the magnetic field in specific planes is noted by red dashed lines.}
-\label{Carte}
-\end{figure}
-
-This relation between NV-NV class resonances and the magnetic field being in specific crystalline planes or directions means that we can determine the crystal main axes simply by PL experiments.
-
-Fig.
-\ref{Carte} shows a map where we simply monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer.
-We have also reported the locus of $\mathbf{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].
-
-We know that when $\mathbf{B}$ belongs to any of these four planes, there is a co-resonance between at least two NV classes.
-This co-resonance leads to an increase in NV-NV CR (or NV-fluctuator CR) and a faster depolarization of the spins involved, which ultimately leads to a drop in PL.
-We can indeed see in fig.
-\ref{Carte} that there is clear correlation between the loci of $\mathbf{B}$ and drops in PL.
-We can also see that the PL is at its lowest when $B\parallel [001]$, which corresponds to a four class degeneracy.
-
-\subsection{$T_1$ dependency on magnetic field}
-
-\section{Limitations of the fluctuator model}
-La modification des T1, tjr plus petite que prévue (faire un tableau), et les T1 non-exponentiels (je crois que y'a des cas sur Ludo, bon courage mdr)
-Parler aussi des alternatives au fluctuator model, ici ou dans la 2e section
-Bon du coup parler du fait que ça n'inclut pas la diffusion entre NV ni la saturation des fluctuateurs. A voir si ça mérite une full section ou une gross ouverture à la fin.
-
+\section{Conclusion and perspectives}
 \printbibliography

### 2022-08-14 Manuscrit Thèse/chapter 4/Chapter 4.tex
salam alekoum

@@
 
+(mentionner l'article ici !)....
+The main motivation for this study was the potential to use NV-NV CR as a low field magnetometry protocol.
+Indeed, while NV-NV CR lines can be used to perform magnetometry with non zero fields [les russes], there are several advantages to use NV-NV CR as close to the zero magnetic field region, in particular the non dependence of the magnetic field orientation with respect to the crystal lattice.
+The behavior of
+
 \section{NV spin Hamiltonian under low and transverse fields}
 
-\section{Modification of NV-NV CR in the $\{ \ket{0}, \ket{+}, \ket{-} \}$ basis}
+Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates.
+
+\subsection{NV spin Hamiltonian in zero external magnetic field}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/ESR_simus}
+\caption{Simulations of inhomogeneous zero field ODMR when sampling various parameters.
+a) Simulation when sampling each components of the magnetic field over a Gaussian of deviation $\sigma=2\ \rm G$.
+b) Simulation when sampling each components of the strain tensor $\bar{\bar{\varepsilon}}$ over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
+c) Simulation when sampling each components of the electric field over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
+d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
+\label{simus ESR}
+\end{figure}
+
+In the absence of external magnetic field, we have to take into account other elements which would otherwise be of second order in the spin Hamiltonian.
+These elements are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}.
+The hyperfine splitting due to nearby nuclei will be considered separately, although to a large extent it behaves like a local magnetic field.
+
+Due to the large zero field splitting $D=2870 MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian).
+The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.
+
+The NV$^-$spin Hamiltonian in the $\{ \ket{-1}, \ket{+1} \}$ basis can be written as \citep{udvarhelyi2018spin}:
+\begin{equation}
+\mathcal{H}=\begin{pmatrix}
+D-\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}}) & f_\perp(\mathbf{E}) + g_\perp(\bar{\bar{\varepsilon}})\\
+f^*_\perp(\mathbf{E}) + g^*_\perp(\bar{\bar{\varepsilon}})&D+\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}})
+\end{pmatrix},
+\label{Hamiltonien pm1}
+\end{equation}
+where $B_\parallel$ is the component of the magnetic field along the NV axis, and $f_\parallel, f_\perp, g_\perp$, and $g_\parallel$ are functions of the electric field $\mathbf{E}$ and the strain tensor $\bar{\bar{\varepsilon}}$, whose expressions are:
+
+\begin{align}
+f_\parallel(\mathbf{E})&=d_\parallel E_z, \\
+f_\perp(\mathbf{E})&=d_\perp ( E_x + i E_y), \\
+g_\parallel(\bar{\bar{\varepsilon}})&= h_{41}(\varepsilon_{xx}+\varepsilon_{yy})+h_{43} \varepsilon_{zz}, \\
+g_\perp(\bar{\bar{\varepsilon}}) &= \frac{1}{2} \left[ h_{16}(\varepsilon_{zx}+i \varepsilon_{zy}) + h_{15}\left(\frac{\varepsilon_{yy}-\varepsilon_{xx}}{2}+i\varepsilon_{xy}\right) \right],
+\end{align}
+where $d_\parallel=0.35 \ \rm{Hz\, cm/V}$ and $d_\perp=17\ \rm{Hz\, cm/V}$ have been measured experimentally \citep{van1990electric}, and $h_{43}=2300\ \rm{MHz}$, $h_{41}=-6420\ \rm{MHz}$, $h_{15}=5700\ \rm{MHz}$ and $h_{16}=19660\ \rm{MHz}$ were computed through DFT \citep{udvarhelyi2018spin} and show reasonable agreement with experiments \citep{barson2017nanomechanical}.
+
+Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.
+
+The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting.
+For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$.
+The shift in energy will therefore tend to blur the energy split when averaging over a large number of spins.
+
+Fig.
+\ref{simus ESR} shows a simulation of how each parameters of the spin Hamiltonian - local magnetic field, local electric field and strain - affects the zero external field ODMR profile.
+To do these simulations, I sampled each parameters separately $10^6$ times and plotted the histogram of the two eigenvalues of the Hamiltonian written in eq.
+\ref{Hamiltonien pm1}.
+Fig.
+\ref{simus ESR}-d) shows an experimental zero field ODMR spectrum, typical of what we observe with dense NV ensembles.
+
+Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR.
+Given the simulation results, the only parameter that can give rise to this shape is the electric field.
+We will therefore consider that the NV Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.
+
+We will then adopt the following simplified Hamiltonian for the zero external field regime:
+\begin{equation}
+\mathcal{H}=\begin{pmatrix}
+D&0&d_\perp E_\perp^* \\
+0&0&0 \\
+d_\perp E_\perp &0&D
+\end{pmatrix},
+\end{equation}
+whose eigenvectors are $\ket{0}$ and$\ket{\pm}$ of eigenvalues 0 and $D\pm d_\perp \abs{E_\perp}$, where $\ket{\pm}$ are defined as:
+\begin{equation}
+\ket{\pm}=\frac{\ket{+1}\pm e^{-i\phi_E}\ket{-1}}{\sqrt{2}},
+\end{equation}
+where $\tan(\phi_E)=E_y/E_x$.
+
+\subsection{NV spin Hamiltonian under purely transverse magnetic field}
+
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.7\textwidth]{Figures/map_etats_propres}
+%\caption{Closeness of the Hamiltonian most excited state $\ket{3}$ with either $\ket{+1}$ or $\ket{+}$ as a function of the external magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
+%\label{map champ transverse}
+%\end{figure}
+
+We will consider here the case of purely transverse magnetic field with respect to the NV axis, i.e.
+$\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
+In practice, this generally means $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.
+
+In this regime, the NV Hamiltonian eigenstates are similar to the case dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{\pm}$\citep{qiu2021nuclear, qiu2022nanoscale}, of eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D$ and $D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
+\begin{equation}
+\ket{\pm}=\frac{\ket{+1}\pm e^{-2i\phi_B}\ket{-1}}{\sqrt{2}},
+\end{equation}
+and $\tan(\phi_B)=B_y/B_x$.
+
+For the case where $d_\perp E_\perp \sim \frac{(\gamma_e B_\perp)^2}{D}$ and $\phi_E \neq 2\phi_B$, the eigenstates of the Hamiltonian are still $\ket{0},\ket{\pm}$ with a relative angle $\phi$ in between $\phi_E$ and $2\phi_B$.
+
+In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian or $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the spin Hamiltonian is dominated by the longitudinal magnetic field, the spin eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+
+\subsection{Hyperfine coupling and inhomogeneous broadening}
+Jambonneau mais pas que. Foutre des ESR si possible. Je vais en chier en vrai.
+
+\section{Modification of NV-NV CR in the transverse field dominated regime}
+In this part, we will study the cross-relaxation between NV centers whose spin Hamiltonian is dominated by a transverse field, either magnetic or electric.
+This is an important regime because it corresponds to the low magnetic field region where the electric field dominates, which is the region for which we want to implement our magnetometry protocol.
+We will also study here the double flip CR processes which we had neglected in the last chapter.
+
+\subsection{NV-NV CR under low magnetic field}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
+\caption{tata. Changer les T1 avec T1ph=5ms}
+\label{scan 1x1x1x1}
+\end{figure}
+We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.
+
+The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust to isolate each effects.
+
+Fig.
+\ref{scan 1x1x1x1}-c) and d) for example shows the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G.
+The ODMR at the initial and final magnetic fields are shown in Fig.
+\ref{scan 1x1x1x1}-a) and b).
+
+While it is clear that the spin lifetime, as well as the PL, increases when the magnetic field, there is no clear indication that this is because of the specificity of the low field region.
+Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/scan_100}
+\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis. Changer les T1 avec T1ph=5ms}
+\label{scan 100}
+\end{figure}
+
+Fig.
+\ref{scan 100} presents a way to circumvent this issue: by applying the magnetic field along the [100] crystalline axis, we can make sure that the four classes of NV centers always stay resonant regardless of the magnetic field amplitude.
+
+We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig.
+\ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of [REF] in zero field, whereas in Fig.
+\ref{scan 100}, it was only reduced by a factor of [REF].
+The main reason for the PL and $T_1$ drop was indeed the co-resonance between the four classes.
+
+Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained by using only the inter-class resonances.
+There are some additional depolarization mechanisms which are proper to the zero field region.
+We should also note that, while the zero-field PL contrast is bigger in Fig.
+\ref{scan 1x1x1x1}-c) than in Fig.
+\ref{scan 100}-c), the slope, which is the limiting factor for sensing ability, is actually very similar in both cases with a value $\sim [REF]\ \rm G^{-1}$.
+
+\subsection{Potential causes for low field depolarization}
+
+Tutut, tu fais le hyperfin avant de continuer.
+\subsubsection{Change in eigenstates}
+
+\subsubsection{Double flips}
+
+\subsubsection{Change in $T_2^*$}
+
+\subsection{NV-NV CR under purely transverse magnetic field}
+
+Sa mère, faut que je replotte tous les T1 avec t1ph=5 ms.
+
+\subsection{Other potential causes}
+Alignment : 100 et perp (utiliser la carte)
+Pola laser
 
@@
 
+Ne pas oublier (ou alors en perspective, c'est bien aussi) : les deux adamas du 20210927 avec le petit et le gros
+
 \section{Conclusion and perspectives}
+- Improvement to the sensitivity : Scale up, material optimization (15N !)
+
+- Understand the impact of the various factors on the flip-flop CR and DQ CR: T2*, strain, local electric noise ...
+
+- Application for uneven surfaces, polycrystalline heteroepithaxy, low/no microwave environment (dimaond anvil cells, bio sensing)
+
 \printbibliography

### 2022-08-14 Manuscrit Thèse/chapter 4/AA_print_matrices.py

@@ -1,0 +1,21 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+
+H=NVHamiltonian([0,0,0],c=1,order='traditionnal')
+Sz=H.Sz
+Sx=H.Sx
+Sy=H.Sy
+Sz2=H.Sz2
+
+basis=['-1','0','+1']
+
+# print_matrix(Sx.dot(Sz)+Sz.dot(Sx)) #{Sx,Sz}
+
+# print_matrix(Sy.dot(Sz)+Sz.dot(Sy)) #{Sy,Sz}
+
+# print_matrix(Sy.dot(Sy)-Sx.dot(Sx)) #(Sy^2 -Sx^2)
+
+print_matrix(Sx.dot(Sy)+Sy.dot(Sx)) #{Sx,Sy}

### 2022-08-17 Manuscrit Thèse/chapter 3/Appendix_eta.tex
botulus

@@
 \chapter{Appendix: Computation of $\bar \eta$}
+We will discuss here the specifics to the computation of the  $\bar \eta$ factor for different geometric configurations, and how to compute the $T_1$ times in table [REF] of chapter [REF], and table [REF] of chapter [REF].
+
+As a reminder, $\bar \eta$ is defined as the averaged value of $\eta$ over all possible configurations.
+
+\begin{equation}
+\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta,
+\end{equation}
+
+\begin{equation}
+\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
+\end{equation}
+
+\section{$\bar \eta$}
 

### 2022-08-17 Manuscrit Thèse/chapter 3/Chapter 3.tex
botulus

@@
 We should note here that, while this simplifications seem a bit extreme, the fluctuator model is more than a toy model.
-There are actually good evidence of the presence of these dark NV centers in dense NV ensemble, as will be discussed below.
+There are good evidence of the presence of these dark NV centers in dense NV ensemble, as will be discussed below.
 
@@
 \begin{equation}
+\label{eq eta}
 \eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
 \end{equation}
-where $\gamma_f=1/T_1^f$ is the fluctuator decay rate.
+where $\gamma_f$ is the fluctuator decay rate for each possible channel between $\ket{0}, \ket{-1}$ and $\ket{+1}$.
 This number $\eta$ encapsulates both the geometry of the problem (through $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
+
+We can note that spectral response of a single fluctuator, for which we only consider the broadening caused by the decay rate $\gamma_f$, is a Lorentzian of half width $2\gamma_f$.
+This is different from the usual lifetime limited optical spectra which are Lorentzian of half width $1/2T_1=\Gamma_1/2$.
+The difference comes here from the fact that we consider thermalization going both ways (i.e.
+$\ket{e} \to \ket{g}$ and $\ket{g} \to \ket{e}$ instead of only $\ket{e} \to \ket{g}$) and that each state of the fluctuator can decay into two other states.
 
@@
 Compared to eq.
-[REF], we have replaced here $\Gamma_2^*$ by $\gamma_f$.
-This is because we assume that the pure dephasing, supposedly equal for the NV center and the flucutator, is smaller than the depolarization rate $\gamma_f$.
-We will further discuss this hypothesis later.
+[REF], the dependency on the inhomogeneous broadening of the spins $1/T_2^*$ is hidden in the $\eta$ factor, and we have this additional broadening $\gamma_f$ coming from the fluctuators very short lifetime.
 
@@
 \end{equation}
-with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{\eta}$: $\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta$.
+with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{\eta}$:
+
+\begin{equation}
+\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta.
+\end{equation}
 
@@
 The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).
+
+\subsection{Computation of $\bar \eta$}
+\label{sec computation eta}
+The parameter $\bar \eta$ in eq.
+\ref{eq 1/T} is of crucial importance since it encompass the dependence of $1/T$ in both the magnetic field and the inhomogeneous broadening $T_2^*$ of the NV centers and the fluctuators.
+
+We will start by separating the factor $\eta$ given in eq. \ref{eq eta} into a a geometric and a spectral part:
+\begin{align}
+\eta^2&=G^2(\psi_{NV},\psi_f,\theta,\phi) R^2(\omega_f,\omega_{NV}), \\
+G^2&=\abs{g}^2+\abs{h}^2,  \\ 
+R^2(\omega_f,\omega_{NV}) &=\frac{1}{3} \cdot \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{align}
+
+$G$ is a function which depends purely on the various angles of the problem (angles of the relative position of the NV center and the fluctuator, angle of the NV and fluctuator $z$ axis), while $R$ represents the resonance condition between the NV and the fluctuator.
+
+We will now average this $\eta$ factor for an ensemble of fluctuators and NV centers:
+\begin{align*}
+\bar \eta &= \int \rm{Prob}(\eta) \abs{\eta} d\eta, \\
+&= \left(\int \rho(\psi_{NV},\psi_f,\theta,\phi) |G(\psi_{NV},\psi_f,\theta,\phi)| d\psi_{NV} d\psi_f d\theta d\phi \right) \\
+&\, \times \left( \int \rho(\omega_{NV}) \rho(\omega_{f}) |R(\omega_f,\omega_{NV})| d\omega_{NV} d\omega_{f} \right), \\
+&= \bar G \cdot \bar R,
+\end{align*}
+where we introduced $\bar G$ and $\bar R$, the averaging of $|G|$ and $|R|$.
+
+The computation of $\bar G$ will be further discussed in sec. \ref{sec quantitative T1} and in appendix [REF]. We will focus here on $\bar R$.
+
+To compute, we need to know the distributions $\rho(\omega_{NV})$ and $\rho(\omega_{f})$.
+We can experimentally measure $\rho(\omega_{NV})$ though ODMR, and we can reasonably expect $\rho(\omega_{f})$ to follow the same probability.
+
+We unfortunately cannot analytically solve $\bar R$ for standard distributions of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ (either Gaussians or Lorentzians).
+We can however solve $\overline{R^2}$, which is relatively close to $(\bar{R})^2$ as long as $2 \gamma_f >> 1/T_2^*$.
+
+For a Lorentzian profile of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$:
+\begin{align*}
+\rho(\omega_{f})&=\frac{1}{\pi \Gamma_f} \frac{1}{1+ \left(\frac{\omega_f-\omega^0_f}{\Gamma_f}\right)^2} \\
+\rho(\omega_{NV})&=\frac{1}{\pi \Gamma_{NV}} \frac{1}{1+ \left(\frac{\omega_{NV}-\omega^0_{NV}}{\Gamma_{NV}}\right)^2},
+\end{align*}
+
+where $\Gamma_{f}=1/T_2^*(\rm fluctuator)$ and $\Gamma_{NV}=1/T_2^*(\rm NV)$. We then find that:
+
+\begin{align*}
+\bar{R^2}&=\frac{1}{3} \int \rho(\omega_{NV}) \rho(\omega_{f}) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} d\omega_{NV} d\omega_{f} \\
+&= \frac{1}{3}\cdot  \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}} \cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
+\end{align*}
+
+In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\bar {R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f >> 1/T_2^*$) in eq.
+\ref{eq 1/T}, we find that:
+\begin{equation}
+\label{eq 1/T avec T2*}
+\frac{1}{T} \propto \frac{1}{\gamma_f + 1/T_2^*},
+\end{equation}
+where we have assumed that $T_2^*$ had the same value for the NV centers and the fluctuators.
+
+If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\bar{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian ans a Gaussian, which does not have a clear analytic solution.
+
+Eq.
+\ref{eq 1/T avec T2*} is subject to many approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, we don't know whether the fluctuator inhomogeneous broadening is the same than the NV one, and $\bar {R ^2} \neq (\bar{R})^2$ in the general case.
+Nevertheless, eq.
+\ref{eq 1/T avec T2*} gives us a general idea of the evolution of $1/T$ with $T_2^*$.
 
@@
 This also explains why the $1/T_1^{\rm dd}$ line is well matched by a Lorentzian, while the ODMR lines are closer to Gaussians in our sample.
+This means that we can extract the fluctuator lifetime from the data in Fig.
+\ref{fluct linewidth} by taking out the contribution of $T_2^*$ on both the NV and the fluctuator spectral response.
 
-We can now extract the fluctuator lifetime from the data in Fig. \ref{fluct linewidth} by taking out the contribution of $T_2^*$ on both the NV and the fluctuator spectral response.
+To do this we first assume that the dipole induced decay rate is proportional to the overlap of the spectral responses of the NV centers and fluctuators $S_{NV}$ and $S_f$.
+We will take $S_{NV}( \omega)$ and $S_{f}( \omega)$ as functions centered in $\omega=0$, and consider that the spectral responses are shifted by $\omega^0_f$ and $\omega^0_{NV}$ respectively, the central frequencies of each group.
+We then have:
+\begin{align*}
+\frac{1}{T_1^{\rm dd}} &\propto \int S_{NV} (\omega-\omega_{f}^0) S_f (\omega-\omega_{NV}^0) d\omega \\
+&= \int S_{NV} (\omega-\Delta \omega) S_f (\omega)
+ d\omega \\
+ &= (S_{NV}*S_{f})(\Delta \omega),
+\end{align*}
+
+where $\Delta \omega=\omega_{NV}^0-\omega_{f}^0$.
+
+Using the approximation $\bar {R ^2} \approx (\bar{R})^2$ detailed in sec.
+\ref{sec computation eta}, we can then decompose the fluctuator spectral response between the lifetime contribution and the inhomogeneous broadening:
+\begin{equation}
+S_{f}(\omega)=\left(S_f^*(\omega)\right)*\left(\frac{1}{1+\left(\frac{\omega}{2 \gamma_f}\right)^2} \right),
+\end{equation}
+
+where $S_f^*$... Ne pas oublier les overline pour les R2 !
 
@@
 where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
-We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{wiki:VoigtProfile}:
+We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{ida2000extended}:
 \begin{equation}
@@
 \subsection{Quantitative modification of $T_1$}
-
+\label{sec quantitative T1}
 \begin{table}[htbp]

### 2022-08-17 data/20210927/mesure sensi Vs angle/adamas 2/traitement.py

@@ -4,69 +4,42 @@ sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
 
-fname="T1 0B soustraction"
-plt.figure(num=1,figsize=(6,4),dpi=80)
-plt.xticks(fontsize=15)
-plt.yticks(fontsize=15)
-plt.locator_params(axis='x', nbins=5)
-x,y=extract_data(fname,xcol=0,ycol=5)
-x=x*1e3
+def fit_T1_0B():
+	fname="T1 0B soustraction"
+	plt.figure(num=1,figsize=(6,4),dpi=80)
+	plt.xticks(fontsize=15)
+	plt.yticks(fontsize=15)
+	plt.locator_params(axis='x', nbins=5)
+	x,y=extract_data(fname,xcol=0,ycol=5)
+	x=x*1e3
+	y=y/max(y)
+	plt.plot(x,y,'o',markerfacecolor="None",ms=5,mew=1,label='Experimental Data')
+	popt,yfit=exp_fit_zero(x,y)
+	print(popt)
+	plt.plot(x,yfit,lw=3,label=r'exp($-t/\tau$)')
+	popt,yfit=stretch_exp_fit_zero(x,y)
+	print(popt)
+	plt.plot(x,yfit,lw=3,label=r'exp($-\sqrt{t/\tau}$)',color=color(2))
+	ax=plt.gca()
+	# ax.set_xlabel(r'Dark time (ms)',fontsize=20)
+	# ax.set_ylabel(r'Spin polarization (AU)' ,fontsize=20)
+	plt.legend(fontsize=15)
+	plt.show()
+
+
+x,y=extract_data('ESR 0B')
 y=y/max(y)
-plt.plot(x,y,'o',markerfacecolor="None",ms=5,mew=1,label='Experimental Data')
-popt,yfit=exp_fit_zero(x,y)
-print(popt)
-plt.plot(x,yfit,lw=3,label=r'exp($-t/\tau$)')
-popt,yfit=stretch_exp_fit_zero(x,y)
-print(popt)
-plt.plot(x,yfit,lw=3,label=r'exp($-\sqrt{t/\tau}$)',color=color(2))
-ax=plt.gca()
-# ax.set_xlabel(r'Dark time (ms)',fontsize=20)
-# ax.set_ylabel(r'Spin polarization (AU)' ,fontsize=20)
-plt.legend(fontsize=15)
[…]

### 2022-08-17 data/20220324/Adamas 150 um/adamas 1/traitement.py

@@ -113,8 +113,7 @@ def plot_ESR_0B():
 	y=y/max(y)
 	plt.plot(x,y,color=color(3))
 
-
-plot_ESR_0B()
+# plot_ESR_0B()
 
 def plot_ESR_1x4():
 	x,y=extract_data('ESR/V=5.000000 V')
@@ -126,5 +125,18 @@ def plot_ESR_1x4():
 	# plt.plot(y)
 
 # plot_ESR_1x4()
+
+def plot_ESR_0B_Vs_1classe():
+	x,y=extract_data('ESR 0B -20 et -30 dB',ycol=5)
+	y=y/max(y)
+	x=x-0
+	plt.plot(x,y)
+	x,y=extract_data('ESR 1 classe pas loin 111')
+	y=y-min(y)
+	y=y/max(y)
+	x=x-2740+2865
+	plt.plot(x,y)
+
+plot_ESR_0B_Vs_1classe()
 
 plt.show()

### 2022-08-17 python_clément/analyse.py

@@ -6,7 +6,7 @@ from scipy.optimize import curve_fit,root_scalar,minimize
 from scipy.signal import find_peaks
 from scipy import fftpack
 from PyQt5.QtWidgets import QApplication,QFileDialog
-from numpy import sqrt,pi,cos,sin
+from numpy import sqrt,pi,cos,sin,exp
 from numpy.linalg import norm
 
 

### 2022-08-18 Manuscrit Thèse/chapter 3/Chapter 3.tex
tto

@@
 \begin{align*}
-\bar{R^2}&=\frac{1}{3} \int \rho(\omega_{NV}) \rho(\omega_{f}) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} d\omega_{NV} d\omega_{f} \\
+\overline{R^2}&=\frac{1}{3} \int \rho(\omega_{NV}) \rho(\omega_{f}) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} d\omega_{NV} d\omega_{f} \\
 &= \frac{1}{3}\cdot  \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}} \cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
@@
 
-In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\bar {R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f >> 1/T_2^*$) in eq.
+In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f >> 1/T_2^*$) in eq.
 \ref{eq 1/T}, we find that:
@@
 
-If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\bar{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian ans a Gaussian, which does not have a clear analytic solution.
+If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\overline{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian ans a Gaussian, which does not have a clear analytic solution.
 
 Eq.
-\ref{eq 1/T avec T2*} is subject to many approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, we don't know whether the fluctuator inhomogeneous broadening is the same than the NV one, and $\bar {R ^2} \neq (\bar{R})^2$ in the general case.
+\ref{eq 1/T avec T2*} is subject to many approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, we don't know whether the fluctuator inhomogeneous broadening is the same than the NV one, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
 Nevertheless, eq.
@@
 
-where $S_f^*$... Ne pas oublier les overline pour les R2 !
+where $S_f^*$ is the spectral broadening caused by the inhomogeneous distribution of the fluctuators, which we will suppose to be equal to $S_{NV}$.
+With these assumptions, we are now ready to extract $\gamma_f$ from the data of Fig.
+\ref{fluct linewidth}.
 
-The first way to do this is to follow the instructions in \citep{hall2016detection} and deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm.
+The first way to do this, as was done for NV-P1 CR in \citep{hall2016detection}, is to deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm.
 The result of the deconvolution is shown in Fig.
 \ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve).
-Fitting with Lorentzians, we find a final value $\Gamma_2 =\frac{\gamma_f}{2} = (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=12.5\ \rm ns$.
+Fitting with Lorentzians, we find a final value $\Gamma_2 =2\gamma_f= (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=49.9\ \rm ns$.
 
-Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\sqrt{2 \ln (2)}\sigma$ and a Lorentzian of HWHM $f_L$:
+Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\frac{\sqrt{\ln (2)}}{\pi T_2^*}$ and a Lorentzian of HWHM $f_L=\frac{2 \gamma_f}{2 \pi}$:
 \begin{align*}
@@
 
-With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, I find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 12.2\ \rm ns$, a value consistent with the one found with the deconvolution method.
+With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, we find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 48.7\ \rm ns$, a value consistent with the one found with the deconvolution method.
 
-Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=3.3\ \rm MHz$, although given their experimental values - full width at half maximum (FWHM) of 25 MHz for the lifetime profile and 9.3 MHz for the ODMR line - I obtain using my method $\gamma_f= 15.7\ \rm MHz$ and $T_1^f=10.1\ \rm ns$.
-The main difference comes from the fact that they considered the spectral half-width of the fluctuators to be $2\gamma_f$ whereas I consider it to be $\gamma_f/2$.
+Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48.2\ \rm ns$, a value surprisingly similar to the one we got.
 

### 2022-08-18 Manuscrit Thèse/chapter 4/Chapter 4.tex
tto

@@
 
-In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian or $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the spin Hamiltonian is dominated by the longitudinal magnetic field, the spin eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian or $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
 
 \subsection{Hyperfine coupling and inhomogeneous broadening}
-Jambonneau mais pas que. Foutre des ESR si possible. Je vais en chier en vrai.
+
+Another modification caused by the presence of transverse fields is the modification of the NV centers $T_2^*$ and hyperfine splittings.
+These modifications can have an impact on the NV-NV CR processes
 

### 2022-08-18 python_clément/test programs/test.py

@@ -1744,4 +1744,8 @@ def test_convolution():
 
 a=[0,1,2]
 
+<<<<<<< Updated upstream
 print(a[-4])
+=======
+print(1/(pi*6.6*1e6))
+>>>>>>> Stashed changes

### 2022-08-19 Manuscrit Thèse/chapter 3/Chapter 3.tex
ttotara

@@
 
+\subsubsection{The fluctuators spectral response}
 \begin{figure}[h]
@@
 
-One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators linewidth.
-In the model, we assumed that fluctuator's lifetime $T_1^f=1/\gamma_f$ was so short that the fluctuators linewidth was lifetime-limited, whereas the normal NV centers linewidth is limited by the inhomogeneous broadening $T_2^*>T_1^f$.
+One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators spectral response.
+Indeed, as we will see, the fluctuator lifetime is so short that it significantly broadens the fluctuators spectral response compared to standard NV centers.
 
-We can experimentally verify this hypothesis: the fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in the last chapter.
+The fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in \citep{hall2016detection}.
 
@@
 Fig.
-\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
-Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of their two central lines to get a precise measurement of the Larmor frequency.
+\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field, and the experimentally measured central frequencies of the two classes of interest when the two lines can be clearly separated.
+Having a precise measurement of the detuning between the two classes $\Delta \nu$ is key in this experiment to get a precise value of the fluctuator linewidth.
+%Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of their two central lines to get a precise measurement of the Larmor frequency.
 In the region where the two ODMR lines overlap we have to use a linear regression to get the Larmor frequency of each class.
@@
 
-A potential explanation for the broadening of the $1/T_1^{\rm dd}$ line could be the interaction strength between the NV center and the fluctuator.
+We attribute this broadening to the fluctuators lifetime.
+Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator.
 However, for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
 
+\subsubsection{The fluctuators lifetime}
 \begin{figure}[h]
@@
 
-The fluctuator model provides another explanation: the fluctuator spectral response is broaden by its very short lifetime.
-This also explains why the $1/T_1^{\rm dd}$ line is well matched by a Lorentzian, while the ODMR lines are closer to Gaussians in our sample.
-This means that we can extract the fluctuator lifetime from the data in Fig.
-\ref{fluct linewidth} by taking out the contribution of $T_2^*$ on both the NV and the fluctuator spectral response.
+We now want to extract the fluctuators lifetime $T_1^f=1/\gamma_f$ to get further information on their potential nature.
 

### 2022-08-19 Manuscrit Thèse/chapter 4/Chapter 4.tex
ttotara

@@
 
-Another modification caused by the presence of transverse fields is the modification of the NV centers $T_2^*$ and hyperfine splittings.
-These modifications can have an impact on the NV-NV CR processes
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2*}
+\caption{tata. Changer les T1 avec T1ph=5ms}
+\label{ESR for T2*}
+\end{figure}
+
+We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
+These change are relevant to our study of NV-NV CR due to the relation between the dipole-induced relaxation rate and $T_2^*$ detailed in sec.
+[REF].
+
+A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
+In the $\{ \ket{0}, \ket{\pm 1} \}$, the eigen energies are linearly sensitive to (longitudinal) magnetic field, and only sensitive to electric fields at the second order, and vice versa for the $\{ \ket{0}, \ket{\pm} \}$ basis.
+
+These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei.
+These effects can drastically modify the ODMR lineshape in zero external magnetic field \citep{jamonneau2016competition} or purely transverse magnetic field \citep{qiu2021nuclear,qiu2022nanoscale}.
+
+We will only consider here the hyper-fine splitting caused by the $^14$N nuclei of the nitrogen atom forming the NV center.
+$^14$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin.
+The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
+\begin{equation}
+\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
+\end{equation}
+
+where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor.
+$\mathcal{H}_n$ and $\bar{\bar{A}}$ can be written:
+
+\begin{align}
+\bar{\bar{A}} &= \begin{pmatrix}
+A_{xx} & 0 & 0 \\
+0 & A_{yy} & 0 \\
+0 & 0 & A_{zz}
+\end{pmatrix} \\
+\mathcal{H}_n &= \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
+\end{align}
+
+where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.
+
+Fig.
+\ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$.
+A table with the full width at half maximum of each ODMR line is shown in Fig.
+\ref{ESR for T2*}-d)
+
+While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2, being a Type 1B HPHT sample, contains significantly more impurities besides NV centers, most likely P1 centers.
+These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.
+
+Fig.
+\ref{ESR for T2*}-a) allows us to evaluate the magnetic field noise in both samples, since the $\ket{\pm 1}$ states are not sensitive to weak electric fields.
+We indeed find that the Sumi-2 sample has much more magnetic noise, to the point where the hyper-fine structure is no longer resolved.
+The total width of the line however is mostly dominated by the hyper-fine splitting, which result in a similar total linewidth in both cases
+
+We should note that the current practice in magnetometry is to consider only of the three hyper-fine lines when they are resolved.
+This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines.
+In our case however, we have to consider the spectral overlap between NV centers and fluctuators, which have an additional broadening of $2\gamma_f \approx 6\ \rm MHz$ [REF] that completely obscures the hyper-fine structure.
+We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.
+
+Fig.
+\ref{ESR for T2*}-b) allows us to evaluate the electric field noise in both samples, since the $\ket{\pm}$ states are not sensitive to weak magnetic fields.
+Similarly, the hyper-fine structure is hidden in this configuration since all three hyper-fine levels are nearly-degenerate, provided that $\frac{(\gamma_e B_\perp)^2}{D} > A_{xx},A_{zz},Q$ which is typically the case for $B_\perp > 40\ \rm G$.
+We can note that the electric field noise is significantly stronger in the HPHT sample, which leads to an ODMR linewidth more than twice as big.
+
+Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field.
+For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction).
+For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_I=0}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$.
+The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.
+
+Overall, we are mostly interested in the linewidth of the ODMR lines for various magnetic field conditions.
+While these linewidth are quite significantly changed for sample CVD-pink, the change is less pronounced on sample Sumi-2.
+Most HPHT samples we used are from type 1B diamond, and behave similarly to sample Sumi-2.
+Since almost all measurements in this chapter will be conducted on HPHT samples, we will consider the modification of $T_2^*$ as a minor effect.
 
@@
 
-Tutut, tu fais le hyperfin avant de continuer.
+We will discuss here three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.
+
 \subsubsection{Change in eigenstates}
+The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
+
+This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
+We justify this change of basis, which only consider the single NV Hamiltonian instead of the full Hamiltonian of the two coupled NV centers, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$.
+
+To compute the decay rate in these case, similarly to what we did in sec.
+REF, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
+
+The averaging of these matrix elements, which is needed to compute the expected decay rates, is detailed in appendix [REF].
+The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane.
+This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.
+
+We therefore need to make an assumption on the distribution of the transverse field in the sample.
+In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane.
+However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq.
+[REF]), there could still be local correlations in the transverse field felt by the NV and the fluctuator.
+
+We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the $x$ and $y$ axes of every spin was the same, which correspond to $l_c=\infty$.
+
+We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{th}$ if $l_c=\infty$.
+$\Gamma_0^{th}$ has the same definition as in table [REF], it is the expected CR rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
+
+In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes are degenerate but the Hamiltonian eignebasis is $\{ \ket{0}, \ket{\pm 1} \}$.
+The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig.
+\ref{scan 100}
 
 \subsubsection{Double flips}
+Allez, j'ai bien avancé aujourd'hui, à toi de jouer maintenant. Il erste les trucs chiants aussi si t'as du temps (légendes et replotter avec T1ph=5 ms.
 
@@
 
+A la fin faire un tableau avec les potentiels increase de $\Gamma_1$ pour les trois cas. Ca veut dire traiter correctement les DQ... (ou au moins à peu près).
 \subsection{NV-NV CR under purely transverse magnetic field}
@@
 
-Ne pas oublier (ou alors en perspective, c'est bien aussi) : les deux adamas du 20210927 avec le petit et le gros
+Ne pas oublier (ou alors en perspective, c'est bien aussi) : les deux adamas du 20210927 avec le petit et le gros DQ
 

### 2022-08-19 python_clément/lab.py

@@ -90,6 +90,7 @@ class Graphical_interface(QMainWindow) :
 			qapp.setStyleSheet(qdarkstyle.load_stylesheet(qt_api='pyqt5'))
 		self.keyPressed=keyPressed
 	def keyPressEvent(self, e):
+		#See example in Plot_data
 		if self.keyPressed=='default':
 			return
 		else :
@@ -599,7 +600,7 @@ class startStopButton():
 class fitButton():
 	def __init__(self,line,fit='lin',name='fit',menu=False,spaceAbove=1,spaceBelow=0): 
 
-		self.fitLibrary=['lin','exp','stretch','arb stretch','ESR','ESR and B']
+		self.fitLibrary=['lin','exp','stretch','arb stretch','ESR','ESR and B','lorentzian','gaussian']
 
 		if menu :
 			self.menu=dropDownMenu('Chose fit','--No fit chosen--',*self.fitLibrary,action=self.itemChosenMenu,spaceAbove=spaceAbove,spaceBelow=0)
@@ -670,6 +671,16 @@ class fitButton():
 			self.func=f
 			self.paramsToShow=[0,1,2,3]
 			self.labelNames=['B amp (G)','Angle from 100 (°)', 'Angle from 111 (°)', 'Width (Mhz)']
+		elif fit=='lorentzian' :
+			from analyse import lor_fit as f
+			self.func=f
+			self.paramsToShow=[1,2]
+			self.labelNames=['x0','sigma']
+		if fit=='gaussian' :
+			from analyse import gauss_fit as f
+			self.func=f
+			self.paramsToShow=[1,2]
+			self.labelNames=['x0','sigma']
 	def itemChosenMenu(self):
 		self.setFitFunction(fit=self.menu.text())
 		self.addFitButton.setEnabled(True)

### 2022-08-22 Manuscrit Thèse/chapter 2/Chapter 2.tex
taratata

@@
     natbib=true,
-    style=numeric,
+    style=numeric-comp,
     sorting=none, %Pour faire apparaitre les refs dans l'ordre
@@
 \chapter{Detection of dark spins via cross-relaxations}
-In this chapter, we will see how the interplay between the spin and the optical properties of the NV centers can be transferred to other spins in the crystal by coupling them to NV centers.
+In this chapter, we will see how the interplay between the optical and spin properties of the NV centers can be transferred to other spins in the crystal by coupling them to NV centers.
 More specifically, I will discuss the first optical detection of two spin species, VH$^-$ and War1, in a CVD diamond by the mean of cross-relaxations with NV centers.
@@
 \ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin.
-When the two considered spin transitions become resonant, in this case by tuning a magnetic field, polarization from the NV center will be transferred to the second spin, meaning in this case that the NV center will end up less polarized and the second spin more polarized.
+When the two considered spin transitions become resonant, in this case by tuning the magnetic field, polarization from the NV center will be transferred to the second spin, making the NV center less polarized and the second spin more polarized than when they are out of resonance.
 
@@
 
-We refer by dark spins to spin defects which do not interact with light, which in the case of diamond defects consist of every known spin defect bar the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
+We refer by dark spins to spin defects which do not interact with light, which in the case of diamond defects consist of almost every known spin defect bar the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
 
@@
 It is formed by a vacancy with a hydrogen atom bonded to one of the four dangling carbon bond, with an extra electron trapped in the vacancy.
-This defect has a similar electronic structure to the NV$^-$ which results in an electronic spin-1 with a similar zero field splitting parameter (see Table [REF]).
+This defect has a similar electronic structure to the NV$^-$ which results in an electronic spin-1 with a similar zero field splitting parameter (see Table \ref{table VH et War1}).
 Although the presence of hydrogen in this defect was confirmed by isotopic study, it has been debated whether this defect was VH$^-$ or V$_2$H$^-$ due to a conflict with \textit{ab initio} calculations \citep{shaw2005importance}.
@@
 This however comes at the cost of a greater precision needed on the external magnetic field.
-Relaxometry also allows AC-magnetic field sensing in the GHz regime, where traditional microwave-based AC-megnetometry is limited to tens of MHz \citep{wang2022picotesla, alsid2022solid}.
+Relaxometry also allows AC-magnetic field sensing in the GHz regime \citep{wang2022picotesla, alsid2022solid}, while traditional microwave-based AC-megnetometry is limited to tens of MHz.
 Finally, relaxometry measurement can be more sensitive than the T2-based one in some instances \citep{steinert2013magnetic}, which comes from the fact that changes in NV $T_1$ times are easier to detect than changes in $T_2$ or $T_2^*$ since $T_1$ is much longer than $T_2$ or $T_2^*$.
@@
 When $\mathbf{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G.
-A scan along the [111] in this case means that the peak signal is $\sim 6$ times weaker than in the [100] case, and spread over multiple features which might overlap between themselves or with other existing PL features.
+A scan along the [111] in this case means that the peak signal will be $\sim 6$ times weaker than in the [100] case, and spread over multiple lines which might overlap between themselves or with other existing PL features.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_13CNV}
-\caption{a) Representation of the $^{13}$C-NV complex (right) compared to a "normal" NV center (left) b) and c) simulation of the transition frequencies of NV$^-$ centers and $^{13}$C-NV$^-$ complex for a magnetic field b) along the [100] axis and c) along the [111] axis}
+\includegraphics[width=1\textwidth]{Figures/121_vs_22_13CNV}
+\caption{a) Representation of the $^{13}$C-NV complex (right) compared to a ``normal" NV center (left) b) and c) simulation of the transition frequencies of NV$^-$ centers and $^{13}$C-NV$^-$ complex for a magnetic field b) along the [100] axis and c) along the [111] axis}
 \label{121 vs 22 13C-NV}
@@
 
-The $^{13}$C-NV$^-$ complex still behaves as a "normal" NV center, except with shifted transition lines.
+The $^{13}$C-NV$^-$ complex still behaves as a ``normal" NV center, except that its transition lines are shifted by the hyper-fin coupling.
 In particular its electronic spin is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states.
@@
 Fig.
-\ref{121 vs 22 13C-NV}-b) and c) show the simulated transition frequencies for a $^{13}$C-NV$^-$ complex when B is scanned along the [100] or [111] axis, as well as the "normal" NV transitions and the co-resonance between those two.
+\ref{121 vs 22 13C-NV}-b) and c) show the simulated transition frequencies for a $^{13}$C-NV$^-$ complex when B is scanned along the [100] or [111] axis, as well as the ``normal" NV transitions and the co-resonance between those two.
 When $\mathbf{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G.
@@
 Similarly to fig.
-\ref{CR P1 exp}, we can see a slowly decaying envelope of the PL due to the transverse magnetic field (we don't see the plateau here since the magnetic field sops at 150 G).
+\ref{CR P1 exp}, we can see a slowly decaying envelope of the PL due to the transverse magnetic field (we don't see the plateau here since the magnetic field stops at 150 G).
 We can also see three deviations to the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
 
-In order to better discern the three features, I have fitted the envelope by using a polynomial fit of order 4 on the smooth part of the data.
-I previously tried to use a fit based on the expected PL rate of the NV centers, by using the model in \citep{tetienne2012magnetic}, but the results were not satisfying enough.
+In order to better discern the three features, we fitted the envelope by using a polynomial fit of order 4 on the smooth part of the data.
+We previously tried to use a fit based on the expected PL rate of the NV centers, by using the model in \citep{tetienne2012magnetic}, but the results were not satisfying.
 
@@
 We can clearly see three dips in PL around the values 20 G, 56 G and 122 G.
-I then computed the expected value of the magnetic field for the expected CR line thanks to the simulations shown in sec.
+We then computed the expected value of the magnetic field for the expected CR line thanks to the simulations shown in sec.
 \ref{sec simu}.
-For the VH$^-$ and War1 lines, I took the ZFS $D$ values from \citep{cruddace2007magnetic} and included the error bars.
-For the $^{13}$C-NV$^-$ lines, the error bars given in \citep{simanovskaia2013sidebands} are much smaller and I only plotted the expected value.
+For the VH$^-$ and War1 lines, we took the ZFS $D$ values from \citep{cruddace2007magnetic} and included the error bars.
+For the $^{13}$C-NV$^-$ lines, the error bars given in \citep{simanovskaia2013sidebands} are much smaller and we only plotted the expected value.
 We can see that the dips in PL match relatively well the predicted CR lines of $^{13}$C-NV$^-$, VH$^-$ and War1 respectively.
@@
 It should be noted however that a single-photon counting APD was used here, which considerably worsens the sensitivity of the measurement as discussed in [REF].
-Using a photodiode and modulation of the magnetic field, I estimate that I could get similar signal to noise ratio within a few minutes.
+Using a photodiode and modulation of the magnetic field, we estimate that we could get similar signal to noise ratio within a few minutes.
 
-To calibrate the magnetic field, I added a microwave field at a  fixed frequency $\nu > 2870\ \rm MHz$, and measured the EM voltage at which the PL dropped due to the $\ket{+1}$ state of the NV centers becoming resonant with the microwave tone.
-I could then use the values in Fig.
+To calibrate the magnetic field, we added a microwave field at a  fixed frequency $\nu > 2870\ \rm MHz$, and measured the EM voltage at which the PL dropped due to the $\ket{+1}$ state of the NV centers becoming resonant with the microwave tone.
+We could then use the values in Fig.
 \ref{121 vs 22 NV} to find the corresponding magnetic field.
-I varied the frequency in the range  $[2870,3150]\ \rm MHz$ by step of 2 MHz to get a calibration of the magnetic field in the $[0,150]\ \rm G$ region.
-Considering the uncertainty in the field angle, the NV ZFS $D$ value and the width of the PL drop, I estimate the final precision of the calibration to be within $\pm 1\ \rm G$.
+We varied the frequency in the range  $[2870,3150]\ \rm MHz$ by step of 2 MHz to get a calibration of the magnetic field in the $[0,150]\ \rm G$ region.
+Considering the uncertainty in the field angle, the NV ZFS $D$ value and the width of the PL drop, we estimate the final precision of the calibration to be within $\pm 1\ \rm G$.
 
@@
 We tried similar PL scans along the [100] axis on many samples with dense ensemble of NV centers.
-We found that the dip at 55 G was present on all $\sim 5$ CVD samples used, and none of the $\sim 20$ HPHT samples.
+We found that the dip at 55 G was present on all CVD samples ($\sim$ 5 samples) used, and none of the HPHT samples ($\sim$ 20 samples).
 The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below.
@@
 
-For this two reasons, we attribute the 20 G line to CR between $^{13}$C-NV$^-$ and "normal" NV centers.
+For this two reasons, we attribute the 20 G line to CR between $^{13}$C-NV$^-$ and ``normal" NV centers.
 
@@
 
+This procedure will be discussed in more details in the next chapter. We find here that the linewidth of VH$^-$ and War1 are respectively [REF] and [REF] MHz.
+
 \subsection{Dark spin concentration}
@@
 \item We have to take into account the loss of contrast of the NV centers as the transverse magnetic field grows, as explained in sec. \ref{CR contrast}.
-\item The $^{13}$C-NV$^-$ are not "dark spins", and its CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
+\item The $^{13}$C-NV$^-$ are not ``dark spins", and its CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
 We will see however in the next chapter that we can consider a fraction of the NV centers as dark spins, and that this fraction can be estimated.
@@
 
-I Find that the area under 0 for the VH$^-$ line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ line, from \ref{121 vs 22 contrast} I find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally I chose the dark to bright NV ratio value from \citep{choi2017depolarization} at 0.25.
+For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose the dark to bright NV ratio value from \citep{choi2017depolarization} at 0.25.
 
-With an estimated NV concentration [NV] $\approx\ \rm 4.6 ppm$, giving us a concentration $[^{13}\rm{C-NV}^-]\approx \ \rm 130 ppb$, I then find $[\rm VH^-] \approx 58\ \rm ppb$.
-Similarly, I find $[\rm VH^-] \approx 68\ \rm ppb$.
+With an estimated NV concentration [NV] $\approx\ \rm 4.6 ppm$, giving us a concentration $[^{13}\rm{C-NV}^-]\approx \ \rm 130 ppb$, we then find $[\rm VH^-] \approx 58\ \rm ppb$.
+Similarly, we find $[\rm VH^-] \approx 68\ \rm ppb$.
 Given the large uncertainty on some parameters, in particular the dark to bright spin ratio, these estimates are probably only precise within an order of magnitude.
@@
 Fig \ref{chauffage CVD}-a) shows the demodulated PL versus mganetic field amplitude along [100] on sample CVD-C, while \ref{chauffage CVD}-b) shows the same experiment on sample CVD-B.
-The fabrication and subsequent operations done on this two samples are detailed thoroughly in \citep{ngambou2022improving}, but the main difference between these two sample is that CVD-C was only annealed once after irradiation, for 2 hours at 800$^\circ$C, while CVD-B was annealed for for 2 hours at 800$^\circ$C and then for 1 hour at 1000$^\circ$C and 1200$^\circ$C.
+The fabrication and subsequent operations done on this two samples are detailed thoroughly in \citep{ngambou2022improving}, but the main difference between these two sample is that CVD-C was only annealed once after irradiation, for 2 hours at 800$^\circ$C, while CVD-B was annealed for 2 hours at 800$^\circ$C and then for 1 hour at 1000$^\circ$C and 1 hour at 1200$^\circ$C.
 
@@
 In this case, the substrate was a type 1B HPHT diamond, which was irradiated and annealed along the CVD layer.
-I Will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.
+We Will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.
 
@@
 
-I first suspected that some of the War1 defects had been migrating from the CVD layer to the HPHT substrate with the high-temperature annealing, but not only is it very unlikely with the annealing time (1 hour) and substrate thickness (several mm), I also could not find a dependence on the War1 concentration with the depth in the substrate, whereas a diffusion would have given a higher concentration near the CVD layer.
+We first suspected that some of the War1 defects had been migrating from the CVD layer to the HPHT substrate with the high-temperature annealing, but not only is it very unlikely with the annealing time (1 hour) and substrate thickness (several mm), We also could not find a dependence on the War1 concentration with the depth in the substrate, whereas a diffusion would have given a higher concentration near the CVD layer.
 The War1 defects seem to have been created by the annealing.
 
-These results suggest that, while some defects are removed by the high temperature annealing, such as the VH$^-$ and several unidentified defects in the HPHT substrate, some on the other hand seem to be created by the annealing process, such as the War1 or the one associated with the line at $B = 48\ \rm G$ in the HPHT substrate.
+These results suggest that, while some defects are removed by the high temperature annealing, such as the VH$^-$ and several unidentified defects in the HPHT substrate, some seem to be created by the annealing process, such as the War1 or the one associated with the line at $B = 48\ \rm G$ in the HPHT substrate.
 
@@
 To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations.
-This detection can be way more sensitive than the usual EPR spectroscopy, and considerably cheaper to setup.
+This detection can be way more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up.
 It is however not as versatile and should be seen as a complimentary tool for diamonds with dense NV ensemble.
 
-We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan, much less noisy for weak magnetic field.
+We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan, much less noisy than the [111] scans for weak magnetic field.
 
@@
 
-I estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$.
+We estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$.
 This concentration is too high by about two orders of magnitude to isolate a single emitter with a diffraction-limited microscope.
 We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing.
-We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and isolate a single VH$^-$ center by scanning the sample with a fixed $B_{[100]}=56 \rm G$ while monitoring the NV PL or $T_1$.
-It is however unknown to me whether this detection scheme is sensitive enough to detect single dark spins.
+We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and optically isolate a single VH$^-$ center by scanning a confocal microscope with a fixed $B_{[100]}=56\ \rm G$ while monitoring the NV PL or $T_1$.
+It is however unclear to me whether this detection scheme would be sensitive enough to detect single dark spins.
 

### 2022-08-22 Manuscrit Thèse/chapter 3/Chapter 3.tex
taratata

@@
 In this chapter we will look at the phenomenon of cross-relaxations between NV centers, both from an experimental and theoretical point of view.
-We will see how a model of fast and slow relaxing centers, the fluctuator model, can explain most of the experimental observations, and finally we will see some limitations of the model.
 
-\section{Experimental observation of NV-NV cross-relaxation}
-Before we discuss the theoretical implications of NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles ([NV]$\gtrsim$ 1 ppm).
+The presence of strong NV-NV dipolar interaction in dense ensemble ([NV] $\gtrsim 1\ \rm ppm$) proved to be both an obstacle for ensemble magnetometry \citep{zhou2020quantum} and an opportunity to explore and use the many-body interaction resulting from it \citep{zhou2020quantum, choi2017observation, kucsko2018critical, zu2021emergent}.
+
+We will see here how a model of fast and slow relaxing centers, the fluctuator model \citep{choi2017depolarization}, can explain most of the experimental observations, and we will see some limitations of the model.
+
+\section{Experimental observation of NV-NV CR}
+Before we discuss the theoretical implications of NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
 \subsection{NV-NV CR between nonequivalent NV centers}
@@
 
-More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there was CR even between equivalent NV centers.
+More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, jarmola2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there were cross-relaxations even between equivalent NV centers.
 
@@
 \ref{equivalent NV-NV}-c) shows the transition frequencies of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers on sample ADM-15-1, an HPHT micro-diamond with $[\rm NV]\sim 3\ \rm ppm$.
-There are three magnetic field values $B_{\rm scan} \sim$ 12, 42 and 77 G for which there is an inter-class resonance.
-Each of these resonances correspond to a "geometric" resonance where the $B$ field is in a symmetry plane between the two classes, meaning that the two resonant NV classes are indeed equivalent.
+There are three magnetic field values $B_{\rm scan} \approx$ 12, 42 and 77 G for which there is an inter-class resonance.
+Each of these resonances correspond to a ``geometric" resonance where the $B$ field is in a symmetry plane between the two classes, meaning that the two resonant NV classes are indeed equivalent.
 
@@
 \ref{equivalent NV-NV}-d) shows the PL contrast as the electromagnet field is scanned.
-There are 3 very clear dips when $B_{\rm scan} \sim$ 12, 42 and 77 G which coincide with inter-class resonances.
+There are 3 very clear dips when $B_{\rm scan} \approx$ 12, 42 and 77 G which coincide with inter-class resonances.
 We will see later that this PL dip is associated with a decrease of the NV centers spin $T_1$ time from both classes.
 
-This observation, and similar ones done by many groups, seem to indicate that there are CR between equivalent NV centers.
+This observation, and similar ones done by many groups \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization, akhmedzhanov2017microwave, giri2018coupled}, seem to indicate that there are CR between equivalent NV centers.
 This is however incompatible with our previous assumptions that equivalent NV centers are equally polarized and bright.
@@
 Choi et al.
-in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: "normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and "fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism.
+in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism.
 We are assuming that the fluctuator relaxation rate, $\gamma_f$ is much greater than the optical pumping rate $\Gamma_{\rm las}$.
@@
 \end{equation}
-where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative orientation of the two dipoles.
+where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative position and orientation of the two dipoles.
 
@@
 
-We can note that spectral response of a single fluctuator, for which we only consider the broadening caused by the decay rate $\gamma_f$, is a Lorentzian of half width $2\gamma_f$.
+We can note that the spectral response of a single fluctuator, for which we only consider the broadening caused by the decay rate $\gamma_f$, is a Lorentzian of half width $2\gamma_f$.
 This is different from the usual lifetime limited optical spectra which are Lorentzian of half width $1/2T_1=\Gamma_1/2$.
@@
 Compared to eq.
-[REF], the dependency on the inhomogeneous broadening of the spins $1/T_2^*$ is hidden in the $\eta$ factor, and we have this additional broadening $\gamma_f$ coming from the fluctuators very short lifetime.
+[REF], the dependency on the inhomogeneous broadening of the spins $1/T_2^*$ is hidden in the $\eta$ factor, and we have the additional broadening $\gamma_f$ coming from the fluctuators very short lifetime.
 
@@
 \end{equation}
-To do so, we need to determine the distribution of the fluctuators positions $\{r_i\}$. Assuming that they are homogeneously distributed in the bulk of the material, we find:
+To do so, we need to determine the distribution of the fluctuators positions and orientations$\{r_i\}$. Assuming that they are homogeneously distributed in the bulk of the material, we find:
 \begin{equation}
@@
 \eta^2&=G^2(\psi_{NV},\psi_f,\theta,\phi) R^2(\omega_f,\omega_{NV}), \\
-G^2&=\abs{g}^2+\abs{h}^2,  \\ 
-R^2(\omega_f,\omega_{NV}) &=\frac{1}{3} \cdot \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+G^2&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
+R^2(\omega_f,\omega_{NV}) &= \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
 \end{align}
@@
 
-To compute, we need to know the distributions $\rho(\omega_{NV})$ and $\rho(\omega_{f})$.
+To perform the computation, we need to know the distributions $\rho(\omega_{NV})$ and $\rho(\omega_{f})$.
 We can experimentally measure $\rho(\omega_{NV})$ though ODMR, and we can reasonably expect $\rho(\omega_{f})$ to follow the same probability.
@@
 We unfortunately cannot analytically solve $\bar R$ for standard distributions of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ (either Gaussians or Lorentzians).
-We can however solve $\overline{R^2}$, which is relatively close to $(\bar{R})^2$ as long as $2 \gamma_f >> 1/T_2^*$.
+We can however solve $\overline{R^2}$, which is relatively close to $(\bar{R})^2$ as long as $2 \gamma_f \gg 1/T_2^*$.
 
-For a Lorentzian profile of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$:
+For a Lorentzian profile of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ defined as:
 \begin{align*}
@@
 
-where $\Gamma_{f}=1/T_2^*(\rm fluctuator)$ and $\Gamma_{NV}=1/T_2^*(\rm NV)$. We then find that:
+where $\Gamma_{f}=1/T_2^*(\rm fluctuator)$ and $\Gamma_{NV}=1/T_2^*(\rm NV)$ are the inhomogeneous broadening of the NV and fluctuator ensemble, and $\omega^0_{NV}$ and $\omega^0_{f}$ are the central angular frequencies of the NV and fluctuator ensemble.
+We then find that:
 
 \begin{align*}
-\overline{R^2}&=\frac{1}{3} \int \rho(\omega_{NV}) \rho(\omega_{f}) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} d\omega_{NV} d\omega_{f} \\
-&= \frac{1}{3}\cdot  \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}} \cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
+\overline{R^2}&= \int \rho(\omega_{NV}) \rho(\omega_{f}) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} d\omega_{NV} d\omega_{f} \\
+&=\frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}} \cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
 \end{align*}
 
-In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f >> 1/T_2^*$) in eq.
+In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f \gg 1/T_2^*$) in eq.
 \ref{eq 1/T}, we find that:
@@
 \label{eq 1/T avec T2*}
-\frac{1}{T} \propto \frac{1}{\gamma_f + 1/T_2^*},
+\frac{1}{T} \propto \frac{1}{2\gamma_f + \Gamma_f + \Gamma_{NV}}.
 \end{equation}
-where we have assumed that $T_2^*$ had the same value for the NV centers and the fluctuators.
 
@@
 Eq.
-\ref{eq 1/T avec T2*} is subject to many approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, we don't know whether the fluctuator inhomogeneous broadening is the same than the NV one, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
-Nevertheless, eq.
-\ref{eq 1/T avec T2*} gives us a general idea of the evolution of $1/T$ with $T_2^*$.
+\ref{eq 1/T avec T2*} is subject to several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
+Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$.
 
@@
 
-This conclusion is based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions, and the resonant energy transfer between molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through dipole-dipole interaction to non-excited acceptors, and where the stretched exponential dynamics of the donors was first observed \citep{forster1949experimentelle}.
+This conclusion was based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions, and also to the resonant energy transfer between molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through dipole-dipole interaction to non-excited acceptors, and where the stretched exponential dynamics of the donors was first observed \citep{forster1949experimentelle}.
 
@@
 In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
-Both of these phenomena could lead to fast spin relaxation, the tunneling because the spin hopping is accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction strength $J$ by the crystal phonons \citep{honig1960electron}.
-Technically, the non-contact dipole-dipole interaction is also modulated by the phonons, but the effect is too small to account for the fast depolarization.
+Both of these phenomena could lead to fast spin relaxation, the tunneling because the spin hopping can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction strength $J$ by the crystal phonons \citep{honig1960electron}.
+While the non-contact dipole-dipole interaction is also modulated by the phonons, this effect would be too small to account for the fast depolarization observed here.
 
 Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
-To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate $\sim 10\ \rm ns$.
+To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate between neighboring sites of $\sim 10\ \rm ns$.
 In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
@@
 This time the profile is more exponential, and the value $T_1^{\rm exp}=4.1\ \rm ms$ is coherent with a phonon-limited lifetime.
-Even though CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.
+Even though the sample CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.
 
@@
 Fig.
-\ref{various_fit_formulae} shows four different fitting procedure on the two same measurements.
-The measurements here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
+\ref{various_fit_formulae} shows four different fitting procedure on the two same experimental data.
+The data here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
 The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.
@@
 \ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
-While this technique of measuring $T_1^{\rm dd}$ is relatively sensitive and reproducible, it is not very accurate.
+While this technique of measuring $T_1^{\rm dd}$ is useful to compare different cases, because it is fairly sensitive, it is not an exact measurement of the dipole-dipole induced lifetime.
 
@@
 Fig.
-\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $\Gamma_1^{\rm dd}=T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
+\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $\Gamma_1^{\rm dd}=1/T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
 For each magnetic field value, a $T_1$ profile was recorded and fitted following the procedure described in Fig.
@@
 We can clearly see that the $\Gamma^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
-If we assume that both NV classes spectral response is a gaussian of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
+If we assume that both NV classes spectral response are gaussians of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
 Fig.
@@
 We attribute this broadening to the fluctuators lifetime.
-Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator.
-However, for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
+Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator, but for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
 
@@
 To do this we first assume that the dipole induced decay rate is proportional to the overlap of the spectral responses of the NV centers and fluctuators $S_{NV}$ and $S_f$.
-We will take $S_{NV}( \omega)$ and $S_{f}( \omega)$ as functions centered in $\omega=0$, and consider that the spectral responses are shifted by $\omega^0_f$ and $\omega^0_{NV}$ respectively, the central frequencies of each group.
+We will take $S_{NV}( \omega)$ and $S_{f}( \omega)$ as functions centered in $\omega=0$, and consider that the spectral responses are shifted by $\omega^0_{NV}$ and $\omega^0_{f}$ respectively, the central frequencies of each group.
 We then have:
 \begin{align*}
-\frac{1}{T_1^{\rm dd}} &\propto \int S_{NV} (\omega-\omega_{f}^0) S_f (\omega-\omega_{NV}^0) d\omega \\
+\frac{1}{T_1^{\rm dd}} &\propto \int S_{NV} (\omega-\omega_{NV}^0) S_f (\omega-\omega_{f}^0) d\omega \\
 &= \int S_{NV} (\omega-\Delta \omega) S_f (\omega)
@@
 We know that when $\mathbf{B}$ belongs to any of these four planes, there is a co-resonance between at least two NV classes.
-This co-resonance leads to an increase in NV-NV CR (or NV-fluctuator CR) and a faster depolarization of the spins involved, which ultimately leads to a drop in PL.
+This co-resonance leads to an increase in NV-NV CR and a faster depolarization of the spins involved, which ultimately leads to a drop in PL.
 We can indeed see in fig.
 \ref{Carte} that there is clear correlation between the loci of $\mathbf{B}$ and drops in PL.
-We can also see that the PL is at its lowest when $B\parallel [001]$, which corresponds to a four class degeneracy.
+We can also see that the PL is at its lowest when $\mathbf{B}\parallel [001]$, which corresponds to a four class degeneracy.
 
@@
 The experimental values were measured on sample ADM-15-3 using the protocol of Fig.
-\ref{various_fit_formulae}-d) with $T_1^{\rm ph}\equiv 5\ \rm ms$.}
+\ref{various_fit_formulae}-d) with $T_1^{\rm ph}= 5\ \rm ms$.}
  \label{T1 champ mag}
@@
 
-What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0 \equiv \Gamma_1^{\rm dd} (\rm{isolated class})$, and to express the decay rates when 2,3 or 4 classes are co-resonant as a function of $\Gamma_0$.
+What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators of the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2,3 or 4 classes are co-resonant as a function of $\Gamma_0$.
 Doing so allows to normalize $n_f$ and $\gamma_f$ in eq.
-\ref{eq 1/T} and to keep the geometrical factor $\bar \eta$ as the only unknown.
+\ref{eq 1/T} and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
 
-Under certain assumptions, detailed in appendix REF, $\bar \eta$ can be analytically or numerically computed for each of the magnetic configurations detailed before.
+$\bar G$ can be analytically or numerically computed for each of the magnetic configurations listed before, as detailed in appendix [REF].
 This allows us to predict the value of the spin decay rate in every situations as a function of the decay rate of an isolated class.
 
-Table \ref{T1 champ mag} shows these predicted values, along with experimentally measured ones on sample ADM-15-3.
+Table \ref{T1 champ mag} shows these predicted values, along with experimental ones on sample ADM-15-3.
 The experimental values were obtained with the protocol described in Fig.
@@
 
-While the general order of $\Gamma_1^{\rm dd} (\mathbf{B})$ is the same for the theoretical and experimental values, the theoretical values overestimate can significantly the experimental values, by more than a factor of 3 for the higher degeneracy scenarios.
+The theoretical values are in a good qualitative agreement with the experimental ones, both ranking the different scenarios in the same order.
+The quantitative predictions of the theory however are rather poor, being off by more than a factor of 3 for the shortest lifetimes.
 
@@
 Using for example $T_1^{\rm ph}=2.5\ \rm ms$ gave results closer to the theoretical ones.
-Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded a depolarization rate $\sim 4$ times higher than that of a single isolated class, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one.
+Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded an increased depolarization rate by a factor $\sim 4$, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one.
 We can also see in Fig.
-\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration lead to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, 4 times smaller than the predicted factor of 10.
+\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration leads to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, 4 times smaller than the predicted factor of 10.
 
@@
 
-In conclusion, we have seen in this chapter that dense NV ensemble present "anormal" CR between the NV centers, and we have shown how the fluctuator model developed by Choi et al.
+In conclusion, we have seen in this chapter that dense NV ensemble present ``anormal" CR between the NV centers, and we have shown how the fluctuator model developed by Choi et al.
 in \citep{choi2017depolarization} can explain this behavior.
 The fluctuator model not only explains the NV-NV CR, but also correctly predicts the stretched exponential lifetime profile.
-It also correctly predicts that the fluctuators have a broader spectral response than the NV centers.
-Nevertheless, the model does not describe perfectly every aspects of NV-NV relaxation, and still has some unknown such as the very nature of the fluctuators.
+It also correctly predicts that the fluctuators have a broader spectral response than the other NV centers.
+Nevertheless, the model does not describe perfectly every aspects of NV-NV relaxation, and there still are many aspects of this problem, both theoretical and experimental, to explore.
 
@@
 While the hypothesis of fluctuators being clusters of NV centers or NV-other spins is solidified by the precedents in P-doped Si \citep{honig1960electron} or F centers in KCl \citep{warren1964spin}, there are still many unknowns on the exact nature of the fluctuators.
-There are at least two aspects that I think should be investigated.
+There are at least two aspects that I think merit to be investigated.
 

### 2022-08-22 Manuscrit Thèse/chapter 4/Chapter 4.tex
taratata

@@
 \subsection{NV spin Hamiltonian under purely transverse magnetic field}
-
+\label{sec B transverse}
 %\begin{figure}[h]
@@
 \subsection{Hyperfine coupling and inhomogeneous broadening}
-
+\label{sec modif T2*}
 \begin{figure}[h]
@@
 
-We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the $x$ and $y$ axes of every spin was the same, which correspond to $l_c=\infty$.
+We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the $x$ and $y$ axes of every spin are the same, which correspond to $l_c=\infty$.
 
@@
 \subsubsection{Double flips}
-Allez, j'ai bien avancé aujourd'hui, à toi de jouer maintenant. Il erste les trucs chiants aussi si t'as du temps (légendes et replotter avec T1ph=5 ms.
 
+\begin{table}[htbp]
+\centering
+\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
+ \label{table double flips}
+\begin{tabular}{c|cc}
+%\toprule
+{} & $l_c=0$ & $l_c=\infty$ \\
+\midrule
+$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
+$\bar{R}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
+$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
+%\bottomrule
+\end{tabular}
+\end{table}
+
+Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is simultaneous flip of both spins in the same directions like for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
+The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
+
+We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field.
+And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field.
+But we can see from Fig.
+\ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma \approx 8.8\ \rm MHz$ measured on Fig.
+[REF].
+In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.
+
+To compute the depolarization induced by the double flips, we need to add another decay channel, on top of the flip-flop one, in the fluctuator model.
+We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
+Finally, the question about the correlation  length of the electric field also needs to be asked.
+
+The results are summed up in Table \ref{table double flips}.
+We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma $), this would be the maximum possible decay rate involving the double flips.
+$\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma$).
+This is close to be true for the experimental values we have.
+Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma $).
+We come back to the results of the last section where only considered the flip-flops.
+
+The inclusion of double flips in zero field with $\bar{R}^2=0.5$ therefore increases the theoretical  decay rate in zero field by [REF] \% compared to the previous case. 
 \subsubsection{Change in $T_2^*$}
 
-A la fin faire un tableau avec les potentiels increase de $\Gamma_1$ pour les trois cas. Ca veut dire traiter correctement les DQ... (ou au moins à peu près).
+The last aspect we considered was the change in $T_2^*$ (we include the change in the hyper-fine interaction in the $T_2^*$) for weak magnetic field, as discussed in sec. \ref{sec modif T2*}.
+
+To quantify its impact in the depolarization rate, we employ eq.
+[REF], wherche use the values $2\gamma_f = 6.5\ \rm MHz$ and we take for $\Gamma_f=\Gamma_NV$ the values of half width at half maximum found in Fig.
+\ref{ESR for T2*}.
+
+We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 6 \%$ for sample Sumi-2, and by $\sim 19 \%$ for sample CVD-pink. 
 \subsection{NV-NV CR under purely transverse magnetic field}
 
-Sa mère, faut que je replotte tous les T1 avec t1ph=5 ms.
+We now want to experimentally discriminate these different contributions and to see how close the actual decay rate is to the theoretical one.
+
+One of the reason we want to discriminate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic VS electric noise ...).
+Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.
+
+The ideal way to isolate each contributions would be to apply an electric field strong enough to split the transition of the 4 classes, and to measure the decay rate of a single class dominated by the transverse electric field.
+Such an electric field would need to be of the order of $\sim 10^6\ \rm{V/cm}$, which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.
+
+Instead of an electric field, we will use a transverse magnetic field, which as described in sec.
+\ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
 

### 2022-08-22 data/20220324/Adamas 150 um/adamas 1/traitement.py

@@ -98,12 +98,13 @@ def fit_T1_1classe():
 	# for T1ph in np.linspace(1e-3,5e-3,100):
 	# 	popt,yfit=stretch_et_phonons(x,y,T1ph=T1ph)
 	# 	print('T1ph=%f,error=%f'%(T1ph,estim_error(y,yfit)))
-	T1ph=0.003626*1e3
+	# T1ph=0.003626*1e3
+	T1ph=5
 	popt,yfit=stretch_et_phonons(x,y,T1ph=T1ph)
 	plt.plot(x,yfit,lw=2)
 	print(popt,1/popt[1]*1e3)
 
-# fit_T1_1classe()
+fit_T1_1classe()
 
 def plot_ESR_0B():
 	# x,y=extract_data('ESR 0B -20 et -30 dB',ycol=1)
@@ -137,6 +138,6 @@ def plot_ESR_0B_Vs_1classe():
 	x=x-2740+2865
 	plt.plot(x,y)
 
-plot_ESR_0B_Vs_1classe()
+# plot_ESR_0B_Vs_1classe()
 
 plt.show()

### 2022-08-22 manips/Double quantum sur des adamas/Calcul eta Lukin.py

@@ -49,7 +49,7 @@ def cartesian2(anglex):
 	return(x2,y2,z2)
 
 
-
+#Legacy
 def xx_diff_noC(theta,phi,anglex1=0,anglex2=1):
 	r=[x(theta,phi),y(theta,phi),z(theta,phi)]
 	x1,y1,z1=cartesian1(anglex1)
@@ -57,7 +57,6 @@ def xx_diff_noC(theta,phi,anglex1=0,anglex2=1):
 	g=(3*x1.dot(r)*x2.dot(r)-x1.dot(x2))
 	integrande=1/(4*pi)*abs(g)*sin(theta)#/(2*pi)#/(2*pi)
 	return(integrande)
-
 def xx_diff_fullC(theta_r,phi_r,theta_E=0,phi_E=0):
 	r=np.array([x(theta_r,phi_r),y(theta_r,phi_r),z(theta_r,phi_r)])
 	E=np.array([x(theta_E,phi_E),y(theta_E,phi_E),z(theta_E,phi_E)])
@@ -78,7 +77,6 @@ def xx_diff_fullC(theta_r,phi_r,theta_E=0,phi_E=0):
 	g=(3*x1.dot(r)*x2.dot(r)-x1.dot(x2))
 	integrande=1/(4*pi)*abs(g)*sin(theta_r)*1/(4*pi)*sin(theta_E)
 	return(integrande)
-
 def yy_diff_noC(theta,phi,anglex1=0,anglex2=0):
 	r=[x(theta,phi),y(theta,phi),z(theta,phi)]
 	x1,y1,z1=cartesian1(anglex1)
@@ -86,17 +84,33 @@ def yy_diff_noC(theta,phi,anglex1=0,anglex2=0):
 	g=(3*y1.dot(r)*y2.dot(r)-y1.dot(y2))
 	integrande=1/(4*pi)*abs(g)*sin(theta)#/(2*pi)#/(2*pi)
 	return(integrande)
-
-
-def gp_diff_noC(theta,phi,anglex1=0,anglex2=0):
+def xy_diff_noC(theta,phi,anglex1=0,anglex2=0):
 	r=[x(theta,phi),y(theta,phi),z(theta,phi)]
 	x1,y1,z1=cartesian1(anglex1)
 	x2,y2,z2=cartesian2(anglex2)
[…]

### 2022-08-23 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

@@
 
-\section{Modification of NV-NV CR in the transverse field dominated regime}
+\section{NV-NV CR in the low magnetic field regime}
 In this part, we will study the cross-relaxation between NV centers whose spin Hamiltonian is dominated by a transverse field, either magnetic or electric.
@@
 
-\subsubsection{Change in eigenstates}
+\subsubsection{Eigenstates modification}
 The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
@@
 The inclusion of double flips in zero field with $\bar{R}^2=0.5$ therefore increases the theoretical  decay rate in zero field by [REF] \% compared to the previous case. 
-\subsubsection{Change in $T_2^*$}
+\subsubsection{$T_2^*$ modification}
 
@@
 We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 6 \%$ for sample Sumi-2, and by $\sim 19 \%$ for sample CVD-pink. 
-\subsection{NV-NV CR under purely transverse magnetic field}
+\subsection{Other potential causes}
+Alignment : 100 et perp (utiliser la carte)
+Pola laser
 
-We now want to experimentally discriminate these different contributions and to see how close the actual decay rate is to the theoretical one.
+\section{NV-NV CR under purely transverse magnetic field}
+We now want to experimentally discriminate these different contributions and to see how close the actual decay rates are to the predicted ones.
 
-One of the reason we want to discriminate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic VS electric noise ...).
+One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic VS electric noise ...).
 Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.
@@
 \ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
+\subsection{Principle of the experiment}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/transis_transverse_field}
+\caption{Simulation of the eigenstates of the NV Hamiltonian for $d_\perp E_perp=4\ \rm MHz$ and a purely transverse magnetic field $B_\perp$.
+(a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$.
+The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$ (b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
+\label{eigenstates transverse field}
+\end{figure}
 
-\subsection{Other potential causes}
-Alignment : 100 et perp (utiliser la carte)
-Pola laser
+Fig.
+\ref{eigenstates transverse field} shows a simulation of the eigenstates and eigenenergies for a NV Hamiltonian subject to transverse electric field and magnetic field.
+For convenience we chose to take $E_\perp = E_x$ and $B_\perp = B_x$, taking other geometric configuration only alters slightly the results.
+We labeled the eigenstates of the Hamiltonian $\ket{1}$, $\ket{2}$ and $\ket{3}$ in ascending order of energy.
+
+Fig.
+\ref{eigenstates transverse field}-b) shows both the splitting $\Delta \nu$ between the $\ket{2}$ and $\ket{3}$ states, and how close the $\ket{3}$ state is to the $\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}}$ state via the factor $\abs{\bra{3}\ket{+}}^2$.
+We look at the $\ket{3}$ state since $\ket{2}=\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}}$ in this case.
+
+These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 since we wanted to observe the modification caused by the eigenbasis $\{\ket{0},\ket{\pm}\}$.
+This way we can isolate the $T_1^{\rm dd}$ contribution coming from the change of eigenstates to the one coming from the double flips.
+Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100,200] G seems to satisfy both these criteria.
 

### 2022-08-23 Manuscrit Thèse/chapter 4/Chapter 4.tex
taratata

@@
 
+\subsection{Experimental data}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{Figures/scan_T1_perp}
+\caption{Experimental data for the depolarization under purely transverse magnetic field on sample ADM-150-2.
+a) ODMR spectrum for $|\mathbf{B}|=20\ \rm G$.
+The transitions of the class orthogonal to the magnetic field are labeled $\nu_+$ and $\nu_-$.
+b) Dependency of $\nu_+$ and $\nu_-$ with the magnetic field, as measured through ODMR.
+c) Measurement of $T_1^{\rm dd}$ as a function of the transverse magnetic field.
+The corresponding energy splittings $\Delta \nu=\nu_+-\nu_-$ are written on top.
+We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$ indicated by a red line.
+We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$}
+\label{champ tranverse exp}
+\end{figure}
+
+We perform here a $T_1^{\rm dd}$ measurement similar to the one presented in Fig.
+[REF], on the same sample and with the same fitting parameters, but this time probing a class orthogonal to the magnetic field.
+
+Fig.
+\ref{champ tranverse exp}-a) and b) show the evolution of the transition frequencies $\nu_+=\frac{E_{\ket{3}}-E_{\ket{1}}}{h}$ and $\nu_-=\frac{E_{\ket{2}}-E_{\ket{1}}}{h}$ where $\ket{1}$, $\ket{2}$ and $\ket{3}$ are the states described in Fig.
+\ref{eigenstates transverse field}.
+The magnetic field was scanned between 20 and 130 G by increment of $\sim 0.5\ \rm G$.
+
+Fig.
+\ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ with the magnetic field.
+We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$.
+In region B, the $1/T_1^{\rm dd}$ value is stable and does not decrease further when $B_\perp$, or $\Delta \nu$ are increased.
+
+We attribute the decrease in region A to the double flips: as the magnetic field and $\Delta \nu$ are increased, the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$.
+This value is coherent with the previously measured fluctuator linewidth.
+We can then estimate the part of the double flips in the low field (20 G here) depolarization rate and find that they amount for $\sim 50\%$ in this case.
+
+Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see that they are not very far detuned in Fig.
+\ref{champ tranverse exp}-a), but we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A, which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.
+
+We now turn to region B.
+Even tough the double flips are completely quenched, the effects of the eigenbasis modification and the $T_2^*$ reduction are still present, because we can see on Fig.
+\ref{eigenstates transverse field}-b) that the eigenstates of the spin Hamiltonian are still very close to $\{ \ket{0}, \ket{\pm} \}$ for these magnetic field values.
+
+Indeed, when we compare the plateau value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$ to the value $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ found with a longitudinal magnetic field, we notice an increase by $\sim 50\%$ of the depolarization rate.
+This is in agreement with the theory which predicts an increase of the decay rate in the $\ket{\pm}$ basis compared to the $\ket{\pm 1}$ basis, although the theoretical increase, when considering only the change of eigenbasis, is a factor of 4 (see appendix [REF]).
+We again find the the theory significantly overestimates the measured decay rates, although this discrepancy could still come from the fitting procedure.
+
+We unfortunately cannot experimentally discriminate the depolarization coming from the eigenbasis change or the $T_2^*$ while using the same sample, since both of this properties are tied to the Hamiltonian eigenstates.
+We do know however that the double flips seem to play a more important role in the low field depolarization, even in a regime where the eigenbasis change should play an important role (we expect $\Gamma_1^{\ket{\pm}}=4 \Gamma_1^{\ket{\pm}}$ for a single class, but only $\Gamma_1^{\ket{\pm}}\approx 1.2 \Gamma_1^{\ket{\pm}}$ when the 4 classes are resonant).
+
 \section{Low field magnetometry}

### 2022-08-24 Manuscrit Thèse/chapter 4/Chapter 4.tex
gjrkn

@@
 
-We will only consider here the hyper-fine splitting caused by the $^14$N nuclei of the nitrogen atom forming the NV center.
-$^14$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin.
+We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center.
+$^{14}$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin.
 The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
@@
 \section{NV-NV CR in the low magnetic field regime}
-In this part, we will study the cross-relaxation between NV centers whose spin Hamiltonian is dominated by a transverse field, either magnetic or electric.
-This is an important regime because it corresponds to the low magnetic field region where the electric field dominates, which is the region for which we want to implement our magnetometry protocol.
-We will also study here the double flip CR processes which we had neglected in the last chapter.
+In this part, we will study the cross-relaxation between NV centers under low magnetic field, whose spin Hamiltonian is dominated by a transverse electric field.
+Understanding the various mechanisms operating in this regime will be important for the magnetometry protocol presented after.
 
-\subsection{NV-NV CR under low magnetic field}
+\subsection{Experimental observations}
 \begin{figure}[h]
@@
 Fig.
-\ref{scan 1x1x1x1}-c) and d) for example shows the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G.
+\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G.
 The ODMR at the initial and final magnetic fields are shown in Fig.
@@
 
-While it is clear that the spin lifetime, as well as the PL, increases when the magnetic field, there is no clear indication that this is because of the specificity of the low field region.
+While it is clear that the spin lifetime as well as the PL increases with the magnetic field, there is no clear indication that this is because of the specificity of the low field region.
 Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.
+
+We can note however that the PL is not an exact mirror of the spin dynamics: while the low field drop in PL is indeed associated with a reduced spin lifetime, the drop in PL at higher magnetic field comes from the states mixing induced by the transverse magnetic field, and is not associated with a modification of the spin dynamics (to the first order).
 
@@
 
-Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained by using only the inter-class resonances.
+Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained by considering only the inter-class resonances.
 There are some additional depolarization mechanisms which are proper to the zero field region.
@@
 \ref{scan 1x1x1x1}-c) than in Fig.
-\ref{scan 100}-c), the slope, which is the limiting factor for sensing ability, is actually very similar in both cases with a value $\sim [REF]\ \rm G^{-1}$.
+\ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually very similar in both cases with a value $\sim [REF]\ \rm G^{-1}$.
+
+We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]
 
@@
 This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
-We justify this change of basis, which only consider the single NV Hamiltonian instead of the full Hamiltonian of the two coupled NV centers, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$.
+We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively.
+To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
 
-To compute the decay rate in these case, similarly to what we did in sec.
-REF, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
-
-The averaging of these matrix elements, which is needed to compute the expected decay rates, is detailed in appendix [REF].
+The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF].
 The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane.
@@
 However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq.
-[REF]), there could still be local correlations in the transverse field felt by the NV and the fluctuator.
+[REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.
 
-We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the $x$ and $y$ axes of every spin are the same, which correspond to $l_c=\infty$.
+We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
 
-We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{th}$ if $l_c=\infty$.
-$\Gamma_0^{th}$ has the same definition as in table [REF], it is the expected CR rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
+We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$.
+$\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
 
-In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes are degenerate but the Hamiltonian eignebasis is $\{ \ket{0}, \ket{\pm 1} \}$.
+In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$.
 The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig.
@@
 \midrule
-$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-$\bar{R}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
-$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
+$\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
+$\bar{G}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
+$\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
 %\bottomrule
@@
 
-Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is simultaneous flip of both spins in the same directions like for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
+Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
 The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
@@
 But we can see from Fig.
-\ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma \approx 8.8\ \rm MHz$ measured on Fig.
+\ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig.
 [REF].
@@
 
-To compute the depolarization induced by the double flips, we need to add another decay channel, on top of the flip-flop one, in the fluctuator model.
-We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
+To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
+We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{G}$ factor defined in [REF].
 Finally, the question about the correlation  length of the electric field also needs to be asked.
@@
 The results are summed up in Table \ref{table double flips}.
-We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma $), this would be the maximum possible decay rate involving the double flips.
-$\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma$).
-This is close to be true for the experimental values we have.
-Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma $).
+We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{G}^2$ factor: $\bar{G}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
+$\bar{G}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$).
+This is close to the the experimental values we have.
+Finally $\bar{G}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $).
 We come back to the results of the last section where only considered the flip-flops.
 
-The inclusion of double flips in zero field with $\bar{R}^2=0.5$ therefore increases the theoretical  decay rate in zero field by [REF] \% compared to the previous case. 
+We therefore predict that the double flips amounts for an increase of the decay rate by [REF] \% (taking $\bar{G}^2=0.5$) compared to the previous case.
+
 \subsubsection{$T_2^*$ modification}
@@
 We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 6 \%$ for sample Sumi-2, and by $\sim 19 \%$ for sample CVD-pink.
+
+\subsubsection{Summary}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/Shema_summary_theory}
+\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
+\label{summary_theory}
+\end{figure}
+
+Fig. \ref{summary_theory} recapitulate the various mechanisms at play in the low field depolarization of NV ensemble.
+
+We start from from the simplest case where every classes are sepctrally isolated and the only dipole-induced relaxation coms from the flip-flop between NV and fluctuators from the same class.
+We then consider the case where all four classes are resonant ($B\parallel [100]$) and finally the case B=0.
+
+The three extra depolarization mechanisms in zero field are then detailed, with the predicted increase of depolarization rate for each step.
+The predicted rates in zero field depend on several parameters, we chose the following ones to compute the different numerical values: $l_c=0$, $\bar{G}^2=0.5$, $2\gamma_f = 6.5\ \rm MHz$ and $2/T_2^*=5.9\ \rm MHz$ (corresponding to sample Sumi-2 if we assume a Lorentzian ODMR profile).
+
 \subsection{Other potential causes}
-Alignment : 100 et perp (utiliser la carte)
-Pola laser
+We have also considered two other causes that we have ultimately ruled out in this case.
+
+\subsubsection{Magnetic field alignment}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{Figures/alignement_rose}
+\caption{PL of sample CVD-pink as a function of the magnetic field for various misalignment of the field direction with the [100] axis}
+\label{alignement rose}
+\end{figure}
+
+The first is a potential misalignment of the magnetic field with the [100] axis, which would cause a splitting of the 4 classes as the magnetic field increases.
+This hypothesis is not consistent with the observations in Fig.
+\ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased.
+Instead we observe a sharp decrease followed by a pseudo-plateau.
+
+We still tried to estimate the sensitivity of the low field depolarization with the field orientation.
+Fig.
+\ref{alignement rose} shows PL scans on sample CVD-pink for various alignment of the magnetic field around the [100] axis.
+We estimate our alignments to be precise within $\pm 1^\circ$.
+We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization does not come from a field misalignment.
+
+\subsubsection{Laser polarization}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{Figures/pola_laser}
+\caption{tata}
+\label{pola laser}
+\end{figure}
+
+Other studies \citep{anishchik2015low, filimonenko2020weak} previously saw a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.
+
+Fig.
+\ref{pola laser} shows PL scans on sample ADM-150-3 for various polarization angle of the incident laser.
+We chose to apply the magnetic field in the laser polarization plane, in a direction that did not match any particular crystalline planes.
+
+We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$.
+We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.
 
@@
 We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$.
-In region B, the $1/T_1^{\rm dd}$ value is stable and does not decrease further when $B_\perp$, or $\Delta \nu$ are increased.
+In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ or $\Delta \nu=\nu_+=\nu_-$ are increased.
 
@@
 
-Indeed, when we compare the plateau value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$ to the value $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ found with a longitudinal magnetic field, we notice an increase by $\sim 50\%$ of the depolarization rate.
-This is in agreement with the theory which predicts an increase of the decay rate in the $\ket{\pm}$ basis compared to the $\ket{\pm 1}$ basis, although the theoretical increase, when considering only the change of eigenbasis, is a factor of 4 (see appendix [REF]).
-We again find the the theory significantly overestimates the measured decay rates, although this discrepancy could still come from the fitting procedure.
+To evaluate the contribution of the eigenbasis and $T_2^*$, we measure the same class of NV centers, but this time with a longitudinal magnetic field ($B_\parallel \sim 100 G$) to get a baseline value of $1/T_1^{\rm dd}$ in the $\ket{\pm 1}$ basis.
+We found a value of $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ in this case, reported on Fig.
+\ref{champ tranverse exp}-c) as a green line.
+
+We do find that the decay rate is higher in the $\ket{\pm}$ basis by $\sim 50\% $ which corroborates our predictions, as both the change of eigenbasis from $\ket{\pm 1}$ to $\ket{\pm}$ and the increase of $T_2^*$ in the $\ket{\pm}$ basis are predicted to increase the depolarization rate.
+The theoretical increase however overestimates once again the actual increase: for the change of eigenbasis alone we predicted an increase of the decay rate by a factor of 4 (see appendix [REF]), over 8 times more than the actual increase.
 
 We unfortunately cannot experimentally discriminate the depolarization coming from the eigenbasis change or the $T_2^*$ while using the same sample, since both of this properties are tied to the Hamiltonian eigenstates.
-We do know however that the double flips seem to play a more important role in the low field depolarization, even in a regime where the eigenbasis change should play an important role (we expect $\Gamma_1^{\ket{\pm}}=4 \Gamma_1^{\ket{\pm}}$ for a single class, but only $\Gamma_1^{\ket{\pm}}\approx 1.2 \Gamma_1^{\ket{\pm}}$ when the 4 classes are resonant).
+But we do know that in this scenario, for a single isolated class, the contribution of both the change of eigenbasis and $T_2^*$ amounts for $\sim 3$ times less than the double flips.
 
-\section{Low field magnetometry}
+Based on these observation, we expect that difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by, in order of importance:
+\begin{enumerate}
+\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
+\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
+\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
+\end{enumerate}
 
-Ne pas oublier (ou alors en perspective, c'est bien aussi) : les deux adamas du 20210927 avec le petit et le gros DQ
+\section{Introduction to ensemble NV magnetometry}
+Je vais en chier de ouf pour trouver des images.
+\subsection{AC and DC magnetometry}
+\subsection{Low field magnetometry}
+Dans le dernier papier de Jelezko y'a des vrais raisons de faire du low field
+\subsection{Microwave-free magnetometry}
+Bubu et les russes
+\subsubsection{Orientation-free magnetometry}
+Arcizet
 
+\section{Low field CR magnetometry}
+
+\subsection{Principle of the experiment}
+
+\subsection{Characterization of the experiment}
+sensi, dépendence sensi angle, variance d'allan
+\subsection{Comparison with other magnetometry protocol}
+Budker : mesurer vite deuf les pentes
 \section{Conclusion and perspectives}
-- Improvement to the sensitivity : Scale up, material optimization (15N !)
+- Improvement to the sensitivity : Scale up (les deux papiers sur le picoTesla mW, avec les deux systèmes), material optimization (15N !, [NV] concentration, )
 
 - Understand the impact of the various factors on the flip-flop CR and DQ CR: T2*, strain, local electric noise ...
+Mettre ici les deux adamas du 20210927 avec le petit et le gros DQ, et peut etre aussi les DQ et SQ qui s'inversent sur le substrat.
+Comprendre
 
-- Application for uneven surfaces, polycrystalline heteroepithaxy, low/no microwave environment (dimaond anvil cells, bio sensing)
+- Application for uneven surfaces, polycrystalline heteroepithaxy, low/no microwave environment (dimaond anvil cells, bio sensing).
 

### 2022-08-26 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

@@
 \section{Introduction to ensemble NV magnetometry}
-Je vais en chier de ouf pour trouver des images.
+Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.
+
+NV magnetometry is generally decomposed between single NV and ensemble magnetometry.
+Single NVs,often associated with scanning probes, offer a spatial resolution limited only by how far the magnetic source can be from the NV center (typically $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
+
+Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working working with samples of several hundreds of $\mu$m.
+
+We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR.
+We will not try to give a complete overview of this field, but only to the elements close to the technique we present.
+For a general overview on microwave-based ensemble NV magnetometry, we recommend the excellent review from Barry et al \citep{barry2020sensitivity}.
+
 \subsection{AC and DC magnetometry}
+The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted in population.
+These protocols are generally distinguished between DC-broadband and AC-narrow band.
+
+For DC-broadband magnetometry, the general technique consist of a Ramsey interferometry experiment, presented in chapter 1, which is limited by the spin inhomogeneous coherence time $T_2^*$.
+The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$.
+Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry.
+The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20 pT/ \sqrt{\rm Hz}$
+AC: pas oublier. J'ai oublié. Nan en vrai j'ai pas oublié, mais je le mettrai dans la partie sans uW
 \subsection{Low field magnetometry}
@@
 
-\section{Low field CR magnetometry}
+\section{Low field depolarization magnetometry}
 

### 2022-08-27 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2*}
+\includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2}
 \caption{tata. Changer les T1 avec T1ph=5ms}
@@
 
-One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic VS electric noise ...).
+One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic and electric noise ...).
 Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.
@@
 
+\subsubsection{Region A}
+
 We attribute the decrease in region A to the double flips: as the magnetic field and $\Delta \nu$ are increased, the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$.
@@
 Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see that they are not very far detuned in Fig.
-\ref{champ tranverse exp}-a), but we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A, which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.
+\ref{champ tranverse exp}-a).
+However we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.
+
+\subsubsection{Region B}
 
@@
 
-Based on these observation, we expect that difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by, in order of importance:
+\subsubsection{Conclusion of the experimental observation}
+
+Based on these depolarization measurements under transverse magnetic field, we expect that the depolarization at low magnetic fields behave similarly, and that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
 \begin{enumerate}
@@
 
-For DC-broadband magnetometry, the general technique consist of a Ramsey interferometry experiment, presented in chapter 1, which is limited by the spin inhomogeneous coherence time $T_2^*$.
+For DC-broadband magnetometry, the general technique consist of a Ramsey interferometry experiment, presented in chapter 1, which is limited by the spin inhomogeneous coherence time $T_2^* \lesssim 1\ \rm{\mu s}$ for NV ensembles.
 The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$.
 Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry.
-The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20 pT/ \sqrt{\rm Hz}$
-AC: pas oublier. J'ai oublié. Nan en vrai j'ai pas oublié, mais je le mettrai dans la partie sans uW
+The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20 pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
+
+For AC magnetometry, the basic protocol consist of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
+Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field.
+The AC magnetic field however has to be in phase with the rephasing pulses, which limits the AC sensing to a relatively narrow band.
+The state of the art in term of AC sensing is $\sim 1\ \rm pT/\sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}.
+
+Recently, \citep{xie2021hybrid} have achieved new records in both DC ($\sim 200\ \rm fT/\sqrt{\rm Hz}$) and AC ($\sim 10\ \rm fT/\sqrt{\rm Hz}$) NV magnetometry by using flux concentrators which are ferromagnets concentrating the magnetic field in a small volume, similarly to an electromagnet.
 \subsection{Low field magnetometry}
-Dans le dernier papier de Jelezko y'a des vrais raisons de faire du low field
+Magnetometry for low magnetic field ($\gamma B < d_\perp E_\perp$) is challenging since in this regime the NV eigenstates and eigenenergies do not depend on the magnetic field in the first order.
+This complication is often lifted  by simply using a bias magnetic field, typically from a permanent magnet.
+
+Some systems however need to be kept under low to zero external magnetic fields, for example to study the J-coupling between spins \citep{sutter2012computational}, or the phase of skyrmions \citep{zazvorka2020skyrmion}.
+To study these systems with NV centers, we need to find new protocols for low field NV magnetometry.
+
+The main technique for low field magnetometry consist in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}.
+The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$.
+
+Another technique consist in using the $^{13}C-$NV complex we discussed in sec.
+[REF] which because of the magnetic field offset caused by the $^{13}C$ nucleus are frist order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
+With natural abundance $^{13}C$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}C$ concentration can lower the coherence of the NV centers.
+This technique is more suited for a carefully chosen single NV center.
+
 \subsection{Microwave-free magnetometry}
-Bubu et les russes 
-\subsubsection{Orientation-free magnetometry}
-Arcizet
+All the precedent magnetemetry protocols described relied on the use of a controllable magnetic field, to coherently manipulate the NV spin state and to measure the action of the magnetic field on the NV Larmor frequency.
+
+There are other protocols which correlate the magnetic field with the spin $T_1$ time (relaxometry).
+These protocols operate without microwave field and work for either DC or high frequency AC sensing.
+
+\subsubsection{microwave-less DC magnetometry}
+The DC microwave-free protocols rely on sharp feature in the PL for specific magnetic fields around certain cross-relaxations processes or level anti-crossing.
+Fig.
+[REF] in chapter 2 show such PL features, in particular the ground state level anti-crossing (GSLAC, simply labeled LAC on the figure), the sharpest of these features.
+
+The GSLAC occurs for a longitudinal magnetic field when $\gamma B_z = D$, which corresponds to $B=1024\ \rm G$.
+In these conditions, the normally bright $\ket{0}$ state and the dark $\ket{-1}$ state become resonant and any residual mixing between these two states (coming from the electric field, the strain or the residual transverse magnetic field) will cause a drop in the NV polarization, and in the PL \citep{broadway2016anticrossing}.
+
+This protocol has been implemented, first only for longitudinal magnetic field measurement \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.
+
+Another DC microwave-less protocol was proposed based on NV-NV CR \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
+This technique reconstructs the external magnetic field by scanning an other magnetic field, similar to Fig.
+[REF] in chapter 3, and by measuring the position of the NV-NV CR lines.
+By scanning the magnetic field in different directions, one can then recover the initial 3D magnetic field.
+The author of this paper did not measure the sensitivity of this protocol, but they predict a shot noise limited sensitivity of $24.7\ \rm nT/\sqrt{\rm Hz}$.
+
+\subsubsection{microwave-less AC magnetometry}
+The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed.
+The detection bandwidth is adjusted with an external magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.
+
+Sensitivities as low as a few $pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, although both protocols rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
+
+\subsection{Orientation-free magnetometry}
+Every protocol described so far requires a precise knowledge  of the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states versus the magnetic field.
+This knowledge implies that the direction of each NV axis in a frame of reference is known, because of the strong anisotropy in the NV Hamiltonian.
+
+For a single crystal, these axes can always be calibrated through ODMR.
+For a polycrystalline diamond, diamond powder or even a single diamond with an erratic motion however, the knowledge of these axes is not possible.
+New protocols need to be found if one wants to perform magnetometry with these materials.
+
+To our knowledge, the only orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
+This protocol however has an extremely low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic}.
+
+Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies.
+The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the larmor frequency), or via the PL \citep{finco2021imaging}.
 
 \section{Low field depolarization magnetometry}
+We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the spontaneous depolarization of dense NV ensemble at low magnetic field.
+This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional field is required.
+
+We will first describe its operation and then characterize it and compare it to the previously mentioned NV magnetometry protocols.
 
 \subsection{Principle of the experiment}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/setup_magneto}
+\caption{Principle of the LFDM protocol a) Experimental setup.
+The confocal microscopy setup is similar to the one described in [REF], the lock-in amplifier (LIA) is used here to add a modulation on the magnetic field, instead of the microwave.
+Both the PL and the LIA signal are monitored.
+b) example of LIA signal on sample ADM-15-4 when scanning the magnetic field $B_0$ between -150 and 150 G.
+c) PL signal corresponding to the LIA signal}
+\label{setup magneto}
+\end{figure}
+
+Fig.
 
@@
 
-- Application for uneven surfaces, polycrystalline heteroepithaxy, low/no microwave environment (dimaond anvil cells, bio sensing).
+- Application for uneven surfaces, polycrystalline heteroepithaxy, low/no microwave environment (dimaond anvil cells, bio sensing). Gradiometry Degen, Arcizet, MRFM
 

### 2022-08-27 data/20210927/mesure sensi Vs angle/le bon adamas/traitement.py

@@ -3,9 +3,9 @@ sys.path.append('D:\\These Clément\\these\\python_clément')
 sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
-plt.figure(num=1,figsize=(6,4),dpi=80)
-plt.xticks(fontsize=15)
-plt.yticks(fontsize=15)
+plt.figure(num=1,figsize=(3,2),dpi=80)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
 # plt.locator_params(axis='x', nbins=5)
 
 
@@ -155,18 +155,28 @@ plt.yticks(fontsize=15)
 # fname='T1 100 3V'
 # fname='T1 1x1x1x1'
 # nmax=-1
-fname='T1 0B nuit'
-nmax=100
-x,y=extract_data(fname,ycol=5)
-x=x*1e3
-y=y/max(y)
-plt.plot(x[:nmax],y[:nmax],'x',markerfacecolor="None",ms=7,mew=1.5,label='Experimental Data',color=color(0))
-popt,yfit=exp_fit_zero(x,y)
-print(popt)
-plt.plot(x[:nmax],yfit[:nmax],lw=3,label=r'exp($-t/\tau$)',color=color(1))
-popt,yfit=stretch_exp_fit_zero(x,y)
-print(popt)
-plt.plot(x[:nmax],yfit[:nmax],lw=3,label=r'exp($-\sqrt{t/\tau}$)',color=color(2))
+def plot_scan():
+	fname='scan 100 20 deg from 100'
+	x,y=extract_data(fname,ycol=3)
+	x=x*65/2
+	y=y/max(y)
+	plt.plot(x,y,lw=2)
 
-plt.legend()
+plot_scan()
+
+def plot_T1_exp():
+	fname='T1 0B nuit'
+	nmax=100
+	x,y=extract_data(fname,ycol=5)
+	x=x*1e3
+	y=y/max(y)
+	plt.plot(x[:nmax],y[:nmax],'x',markerfacecolor="None",ms=7,mew=1.5,label='Experimental Data',color=color(0))
+	popt,yfit=exp_fit_zero(x,y)
+	print(popt)
+	plt.plot(x[:nmax],yfit[:nmax],lw=3,label=r'exp($-t/\tau$)',color=color(1))
+	popt,yfit=stretch_exp_fit_zero(x,y)
+	print(popt)
[…]

### 2022-08-28 Manuscrit Thèse/chapter 4/Chapter 4.tex
lalalere

@@
 
-Based on these depolarization measurements under transverse magnetic field, we expect that the depolarization at low magnetic fields behave similarly, and that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
+We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field.
+We expect that the depolarization for low magnetic fields behave similarly, and that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
 \begin{enumerate}
@@
 Fig.
+\ref{setup magneto} shows the principle of the LFDM protocol.
+The experimental setup is similar to the one presented in Fig.
+[REF] with the exception that there are no microwaves, and that the magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\delta B$ used for the modulation.
+We typically use a modulation frequency $f_{\rm mod} \approx 1\ \rm kHz$, the same value we use for ODMR measurements.
+The main voltage $V_0$ responsible for the DC field and the oscillating voltage $\delta V$ responsible for the AC field are summed via a homemade bias tee.
+
+Fig.
+\ref{setup magneto}-b) and c) show respectively the PL measured directly on the photodiode and the lock-in signal when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G.
+The magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.
+
+Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup.
+to improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing.
+Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.
+
+Fig.
+[A FAIRE] shows how the AC and DC sensing work for LFDM.
+For DC sensing, we want to use an oscillating field with an amplitude that matches the width of the PL dip $|\delta B| \approx \sigma$, in order to maximize the slope of the lock-in signal around $B_0=0$.
+For AC sensing, we want to use a static field $B_0=\sigma$ corresponding to the region where the lock-in signal is most sensitive to AC field.
 
 \subsection{Characterization of the experiment}
+We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols.
+The characterization was only done for DC sensing, but DC and AC are supposed to be symmetrical.
+The sample used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
+\subsubsection{Sensitivity}
+
 sensi, dépendence sensi angle, variance d'allan

### 2022-08-29 Manuscrit Thèse/chapter 4/Chapter 4.tex
bvjrk

@@
 For a single crystal, these axes can always be calibrated through ODMR.
-For a polycrystalline diamond, diamond powder or even a single diamond with an erratic motion however, the knowledge of these axes is not possible.
+However, for a polycrystalline diamond, diamond powder or even a single diamond with an erratic motion, the knowledge of these axes is not possible.
 New protocols need to be found if one wants to perform magnetometry with these materials.
@@
 
-\subsection{Characterization of the experiment}
+\subsection{Characterization of the LFDM protocol}
 We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols.
-The characterization was only done for DC sensing, but DC and AC are supposed to be symmetrical.
-The sample used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
+The characterization was only done for DC sensing.
+The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
 \subsubsection{Sensitivity}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/sensi_alternee}
+\caption{Sensitivity measurement of the LFDM on sample ADM-15-4.
+a) Temporal trace of the LIA signal when applying an alternating external magnetic field of amplitude 35 $\mu$T and frequency $f=1\ \rm Hz$.
+b) Histogram of the DC measurement over 50 s.
+The data is fitted with Gaussians of width $1.6\ \rm{\mu T}$.}
+\label{sensei alternee}
+\end{figure}
+We characterize the DC sensitivity by looking at the noise floor on the LIA signal when being on the steepest slope, around $B_0=0$.
+The following parameters were used in this case: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
+
+Fig.
+\ref{sensei alternee} shows the sensitivity measurement protocol.
+First, the applied magnetic field is calibrated with ODMR spectra, then an alternating DC field of known amplitude (here $\Delta B_0=35\ \rm{\mu T}$) is applied on the sample.
+Fig.
+\ref{sensei alternee}-a) shows the temporal trace of the LIA signal in response to this alternating DC field.
+Knowing the amplitude of the alternating field, we can then convert the voltage signal into a magnetic field amplitude, with in this case a ratio $45 \pm 3\ \rm{\mu T /V}$.
+
+Fig.
+\ref{sensei alternee}-b) shows an histogram of the measured magnetic field amplitude for both the high and low value of the applied magnetic field.
+The total acquisition time was 50 s with the external field being alternated every second.
+The histograms are nicely fitted with Gaussian profiles of width $\sigma=1.5\pm 0.1\ \rm{\mu T}$ which allows us to estimate the DC sensitivity of the protocol:
+\begin{equation}
+\eta = \frac{\sigma}{\sqrt{2 \tau}}= 116 \pm 10\ \rm{nT} / \sqrt{\rm Hz},
+\end{equation}
+
+where $\tau=3\ r=\rm ms$ is the low-pass filter time constant.
+
+We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$).
+Laser amplitude fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.
+
+\subsubsection{Angular dependence of the sensitivity}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/sensi_angulaire}
+\caption{tata}
+\label{angular sensi}
+\end{figure}
+
+We will now look at the dependence of the sensitivity with respect to the magnetic field angle.
+In particular, we are interested in comparing the sensitivity when $\mathbf{B}\parallel [100]$, where there is no lift of degeneracy between the four classes as $\mathbf{B}$ increases; and a case where $\mathbf{B}\nparallel [100]$ and where there is a lift of degeneracy between the four classes.
+We previously saw in Fig.
+\ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\mathbf{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.
+
+Fig.
+\ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angle of $B_0$.
+For the blue curve, the field was scanned along the [100] axis, for the orange curve, it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis
 

### 2022-08-29 data/20220825/adamas 15/Mesures bruit/traitement.py

@@ -13,7 +13,7 @@ def plot_allan_dev(sensimag=True):
 	import allantools
 	plt.yscale('log')
 	plt.xscale('log')
-	plt.figure(num=1,figsize=(6,4),dpi=80)
+	plt.figure(num=1,figsize=(4.5,3),dpi=80)
 	plt.xticks(fontsize=16)
 	plt.yticks(fontsize=16)
 	# plt.locator_params(axis='x', nbins=5)
@@ -50,12 +50,13 @@ def plot_allan_dev(sensimag=True):
 	plt.legend()
 	plt.show()
 
-# plot_allan_dev()
+plot_allan_dev()
 
 def plot_data_raw():
-	plt.figure(num=1,figsize=(6,4),dpi=80)
+	plt.figure(num=1,figsize=(4.5,3),dpi=80)
 	plt.xticks(fontsize=16)
 	plt.yticks(fontsize=16)
+	plt.locator_params(axis='x', nbins=5)
 
 	t,data=extract_data('Mesure près V=0 plus')
 	conv0=0.06/3*35*1e-4*1e6
@@ -71,4 +72,4 @@ def plot_data_raw():
 	plt.show()
 
 
-plot_data_raw()
+# plot_data_raw()

### 2022-08-30 Manuscrit Thèse/chapter 4/Chapter 4.tex
lalilalal

@@
 \subsection{Potential causes for low field depolarization}
-
+\label{sec causes zero field}
 We will discuss here three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.
@@
 
-Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working working with samples of several hundreds of $\mu$m.
+Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of several hundreds of $\mu$m.
 
@@
 Recently, \citep{xie2021hybrid} have achieved new records in both DC ($\sim 200\ \rm fT/\sqrt{\rm Hz}$) and AC ($\sim 10\ \rm fT/\sqrt{\rm Hz}$) NV magnetometry by using flux concentrators which are ferromagnets concentrating the magnetic field in a small volume, similarly to an electromagnet.
+Because the gain from the flux concentrators is not tied to a specific magnetometry protocol, we will focus on the results obtained without flux concentrators.
 \subsection{Low field magnetometry}
@@
 
-Another technique consist in using the $^{13}C-$NV complex we discussed in sec.
-[REF] which because of the magnetic field offset caused by the $^{13}C$ nucleus are frist order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
-With natural abundance $^{13}C$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}C$ concentration can lower the coherence of the NV centers.
+Another technique consist in using the $^{13}\rm{C}-$NV complex we discussed in sec.
+[REF] which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are frist order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
+With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers.
 This technique is more suited for a carefully chosen single NV center.
@@
 \ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angle of $B_0$.
-For the blue curve, the field was scanned along the [100] axis, for the orange curve, it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis
+For the blue curve, the field was scanned along the [100] axis, for the orange curve, it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis.
+We can see that the maximum amplitude of the lock-in signal (relevant for AC sensing) is of the same order of magnitude for both case, and the slope around $B_0$ (relevant for DC sensing) is almost identical.
+Fig.
+\ref{angular sensi}-c) shows the sensitivity, measured with the protocol described in Fig.
+\ref{sensei alternee}, with respect to the magnetic field angle with the [100] axis.
+We can notice that the sensitivity depends only weakly on the magnetic field angle.
 
-sensi, dépendence sensi angle, variance d'allan
+The fact that the sensitivity is similar when $\mathbf{B}\parallel [100]$ or $\mathbf{B}\nparallel [100]$ means that the inter-class co-resonances plays a lower role in the sensitivity than the zero-field specific depolarization mechanisms, previously detailed in Sec.
+\ref{sec causes zero field}.
+We cannot be sure which of the three zero-field mechanism is the dominant one, but based on the experimental results shown in Fig.
+\ref{champ tranverse exp}, we can assume that the double-flips are the dominant factor in the LFDM sensitivity.
+
+\subsubsection{Temporal stability}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/Temporal_just_0B}
+\caption{tata}
+\label{temporal 0B}
+\end{figure}
+
+An important aspect of sensors is their stability in time.
+The sensitivity being expressed in $\rm{nT} / \sqrt{\rm Hz}$ suggests that the signal to noise ratio (SNR) given by repeating the measurement increases as $\sqrt{T}$ where $T$ is the total acquisition time.
+This is true for an ideal system, but in practice drifts and low frequency noise mean that the actual SNR for long acquisition time will be worse than the SNR at short time scaled by a factor $\sqrt{T}$.
+
+One of the main source of drift for NV magnetometry based on spin resonance is the lattice temperature change: the $D$ factor in the NV Hamiltonian (eq.
+[REF]) is sensitive to the lattice temperature because of the thermal dilatation of the crystal, which is the basis for NV thermometry \citep{kucsko2013nanometre}.
+This means that thermal fluctuations will impact the Larmor frequency of the spins and may cause a detuning with the initial measured frequency for long acquisition times.
+The LFDM protocol however does not rely on a specific resonance frequency, but rather on a co-resonnance condition between neighboring spins, which does not depend directly on the crystal temperature.
+
+Fig.
+\ref{temporal 0B} shows the temporal stability of the LFDM protocol over an acquisition time of 100 s.
+Fig.
+\ref{temporal 0B}-a) shows a temporal trace of the LIA signal where we have converted the signal into a measured magnetic field.
+Fig.
+\ref{temporal 0B}-b) shows the estimated SNR for a signal of 100 nT as a function of the total acquisition time and was computed from the temporal trace.
+
+We can see that the SNR does not deviate from the $\sqrt{T}$ scaling for times up to 50 s, and could probably be extended much longer.
+The non-linearity for shorter times comes from the LIA low-pass filter.
+
 \subsection{Comparison with other magnetometry protocol}
+We will now compare the LFDM performances with the previously mentioned magnetometry protocols.
+We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity.
+This discrepancy however come in large part to the smaller sensor volume used in the characterization, and to a relatively low PL collection efficiency compared to the other experiments.
+
+\subsubsection{Comparison with microwave based protocols}
+
+\begin{table}[htbp]
+\centering
+\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
+ \label{table sensi volumiques}
+\begin{tabular}{c|cc}
+%\toprule
+{} & $l_c=0$ & $l_c=\infty$ \\
+\midrule
+$\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
+$\bar{G}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
+$\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
+%\bottomrule
+\end{tabular}
+\end{table}
+
+comparaison microwave: le scan avec la microwave + zoom + les variances d'allan et trace temporelles.
 Budker : mesurer vite deuf les pentes
+champ tranverse : demerde toi
 \section{Conclusion and perspectives}

### 2022-08-31 Manuscrit Thèse/chapter 4/Chapter 4.tex
lala

@@
 \ref{scan 1x1x1x1}-c) than in Fig.
-\ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually very similar in both cases with a value $\sim [REF]\ \rm G^{-1}$.
+\ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually similar in both cases.
 
@@
 
-Sensitivities as low as a few $pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, although both protocols rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
+Sensitivities as low as a few $pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
 
@@
 
+It should be noted that this non angular dependence was particularly marked on sample ADM-15-4.
+Other samples, including from the same batch, showed a higher angular dependence with typically a sensitivity $\sim 2$ times worse when $B\parallel [100]$ than in the other directions.
+
 \subsubsection{Temporal stability}
@@
 
-We can see that the SNR does not deviate from the $\sqrt{T}$ scaling for times up to 50 s, and could probably be extended much longer.
+We can see that the SNR does not deviate from the $\sqrt{T}$ scaling for times up to 50 s, and could probably be extended for longer times.
 The non-linearity for shorter times comes from the LIA low-pass filter.
@@
 \centering
-\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
+\caption{Tutu}
  \label{table sensi volumiques}
-\begin{tabular}{c|cc}
-%\toprule
-{} & $l_c=0$ & $l_c=\infty$ \\
+\begin{tabular}{c|ccc}
+\toprule
+{} & Zhou et al. \citep{zhou2020quantum} (AC) & Barry et al.\citep{barry2016optical} (DC) & LFDM (DC) \\
 \midrule
-$\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-$\bar{G}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
-$\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
-%\bottomrule
+$\eta\ (\rm{nT} / \sqrt{\rm Hz})$ & 92&0.015& 116 \\
+$V\ (\mu \rm{m}^3)$ &$8.1 \cdot 10^{-3}$ &$5.2 \cdot 10^{6}$& $3.3 \cdot 10^{3}$ \\
+$\eta_v\ (\rm{nT}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ &8.3&34&6700 \\
+\bottomrule
 \end{tabular}
@@
 
-comparaison microwave: le scan avec la microwave + zoom + les variances d'allan et trace temporelles.
-Budker : mesurer vite deuf les pentes
-champ tranverse : demerde toi
+We will first compare the LFDM protocol with the best microwave-based NV magnetometry protocols.
+A better metric than the sensitivity to compare different protocols is the volume-normalized sensitivity defined in \citep{zhou2020quantum} as $\eta_v=\eta\cdot \sqrt{V}$ where $\eta$ is the sensitivity and $V$ the effective volume of the sensor.
+This scaling assumes that each NV center in an ensemble is an independent magnetic field probe, and as the total number of NV centers is proportional to $V$, the sensitivity should scale as $1/\sqrt{V}$.
+
+Table \ref{table sensi volumiques} reports the best AC and DC volume-normalized sensitivities found in the literature.
+The LFDM protocol performs $\sim 200$ times worse than the best DC magnetometer.
+It should be noted however that the PL collection in our measurement, where we use a confocal microscope using an objective with NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/compariason_0B_MW}
+\caption{tata}
+\label{comparaison 0B MW}
+\end{figure}
+
+Fig.
+\ref{comparaison 0B MW} shows a comparison between LFDM and microwave-based magnetometry using our setup.
+A fixed microwave tone (here at 2600 MHz) is sent to the diamond and a magnetic field $B_0$ close to the [111] axis is scanned along an oscillating field $\delta B$ to perform a lock-in detection.
+
+Fig.
+\ref{comparaison 0B MW}-b) shows the LIA signal over a $B_0$ scan: the profile is similar to the one in Fig.
+\ref{setup magneto}-b) with the addition to two sharp lines for $|B_0| \approx 100\ \rm G$.
+These two lines correspond to the transition $\ket{0} \to \ket{-1}$ (or $\ket{0} \to \ket{+1}$ for the negative fields) of the class aligned with the magnetic field when it comes to resonance with the 2600 MHz microwave tone.
+The microwave power used here correspond to a Rabi frequency $\Omega \approx (2\pi) 3\ \rm MHz$ and the slope on the steepest part of the microwave line is $\sim 2.5$ times higher than the slope in zero field.
+
+Fig.
+\ref{comparaison 0B MW}-c) and d) show the temporal stability of the sensor similarly to Fig.
+\ref{temporal 0B}, on both the zero field line and the microwave line.
+While we can see that the microwave signal is more sensitive thanks to the higher slope, it is also more prone to drift, most likely because of the shift of the Larmor frequency caused by the temperature fluctuations.
+We can even see that the LFDM becomes more sensitive for acquisition times beyond 10 s.
+
+We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the microwave power (although $\Omega_{Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$.
+Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.
+
+\subsubsection{Comparison with microwave-less protocols}
+The only microwave-less protocols to have measured a sensitivity are the GSLAC protocols described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
+Unfortunately the sensing volume in this studies is not known, however due to the similarity in our approaches, we should be able to directly compare the physical origin of both protocols, i.e.
+comparing the zero field PL dip and the GSLAC PL dip.
+
+In Fig.
+\ref{scan 1x1x1x1}, we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
+In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type 1B irradiated diamond, similar to the ones we use.
+In \citep{zheng2020microwave} they manage to keep a contrast of 4.5 \% while reducing the GSLAC linewidth to $0.38 G$ by using a CVD sample with a lower [NV] and [N] density and isotopically enriched in $^{12}$C.
+
+The zero field technique is therefore comparable with the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear what kind of material engineering would improve its sensitivity.
+
+%Je parle pas de la dépendence en puissance ou du champ transverse, je laisse ça pour plus tard.
 \section{Conclusion and perspectives}
+We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field.
+We have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers.
+
+We have identified three mechanisms by which the depolarization is increased in zero-field: the inter-class flip-flops, the double flips, and the change in the NV eigenstates (which includes the change in $T_2^*$).We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
+
+Finally, we have seen how this low-field depolarization can be applied to perform magnetometry experiments, and what advantages and disadvantages this technique offers compared to the established NV magnetometry protocols.
+
+I will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.
+
+\subsection{Material optimization for LFDM}
+
+Most magnetometry protocols have a relatively clear material optimization path: DC protocols want to improve the figure of merit $\rho T_2^*$ and AC protocols want to improve $\rho T_2$, where $\rho=[NV]$ is the NV concentration.
+The reason is that the number of NV centers for a given volume scales as $\rho$, and therefore the sensitivity scale as $\sqrt{\rho}$.
+Meanwhile increasing the density of defects tend to decrease the spin coherence times $T_2$ and $T_2^*$, and the AC and DC sensitivities scale respectively as $1/\sqrt{T_2}$ and $1/\sqrt{T_2^*}$.
+
+For the LFDM protocol however, the material optimization is far from obvious.
+Indeed, while increasing $T_2^*$ can only improve the sensitivity, the gain might not be as interesting as for other protocols due to the broadening caused by the fluctuator lifetime.
+On the other hand, increasing the NV density not only increases the number of sensor per volume, but also increases the dipole-dipole strength between NV centers.
+It has been my experience however that crystals with higher NV densities showed poorer LFDM performance (lower zero field PL contrast) than crystals with lower NV density.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/substrat}
+\caption{tata}
+\label{substrat chelou}
+\end{figure}
+
+Fig.
+\ref{substrat chelou} shows various measurements on the same sample SBST-C, an HPHT bulk sample containing [NV]$\sim 1\ \rm ppm$, on two spatially distinct regions A and B.
+These two regions most likely correspond to two distinct growth sectors of the diamond, meaning that these regions were not grown in the same exact environment, which can impact the creation of impurities.
+
+The lifetime for both $B=0$ and $B\neq 0$ are shorter and closer to stretched exponential profile in region B than in region A, and the zero field PL contrast is larger when B is scanned in an arbitrary direction.
+These three observations seem to indicate that region B has a higher concentration of fluctuators than region A.
+When B is scanned along the [100] however (to observe double flips and the change of basis effect), there is no PL dip for region B, but there is still one for region A.
+
+\subsection{Finding the optimal parameters for LFDM}
+
 - Improvement to the sensitivity : Scale up (les deux papiers sur le picoTesla mW, avec les deux systèmes), material optimization (15N !, [NV] concentration, )

### 2022-08-31 data/20220825/adamas 15/microwave 1 raie modul B/traitement.py

@@ -1,0 +1,16 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+# print(glob.glob('*.csv'))
+plt.figure(num=1,figsize=(4.5,3),dpi=80)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
+plt.locator_params(axis='x', nbins=5)
+plt.locator_params(axis='y', nbins=5)
+
+x,y=extract_data('2600 MHz 20 dB avec ampli.csv')
+x=(x*35)+8
+plt.plot(x,y,)
+
+plt.show()

### 2022-09-01 Manuscrit Thèse/chapter 4/Chapter 4.tex
gnuerklngj

@@
 
-\subsection{Potential causes for low field depolarization}
+\subsection{Theory of the low field depolarization}
 \label{sec causes zero field}
-We will discuss here three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.
+We have identified three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.
 
@@
 \section{Low field depolarization magnetometry}
+\label{sec LFDM}
 We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the spontaneous depolarization of dense NV ensemble at low magnetic field.
@@
 
-\subsection{Characterization of the LFDM protocol}
-We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols.
-The characterization was only done for DC sensing.
-The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
-\subsubsection{Sensitivity}
+\subsection{Sensitivity of the LFDM protocol}
 \begin{figure}[h!]
@@
 \end{figure}
+
+We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols.
+The characterization was only done for DC sensing.
+The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
+
 We characterize the DC sensitivity by looking at the noise floor on the LIA signal when being on the steepest slope, around $B_0=0$.
-The following parameters were used in this case: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
+The following parameters were used in the case presented: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
 
@@
 
-where $\tau=3\ r=\rm ms$ is the low-pass filter time constant.
+where $\tau=3\ r=\rm ms$ is the LIA low-pass filter time constant.
 
@@
 
-\subsubsection{Angular dependence of the sensitivity}
+\subsection{Angular dependence of the sensitivity}
+\label{sec angular sensi}
 \begin{figure}[h!]
@@
 \includegraphics[width=\textwidth]{Figures/sensi_angulaire}
-\caption{tata}
+\caption{Angular dependence of the sensitivity.
+a) LIA signal for $B_0$ scanned between $-$150 and +150 G along the [100] axis, or along a direction 20$^\circ$ off the [100] axis.
+b) Zoom-in of the LIA signal on the zero-field slopes.
+c) Measured sensitivity as a function of the magnetic field angle with the [100] axis.}
 \label{angular sensi}
@@
 
-It should be noted that this non angular dependence was particularly marked on sample ADM-15-4.
-Other samples, including from the same batch, showed a higher angular dependence with typically a sensitivity $\sim 2$ times worse when $B\parallel [100]$ than in the other directions.
+It should be noted that this low angular dependence was particularly marked on sample ADM-15-4.
+Other samples, including from the same batch, showed a higher angular dependence with typically a sensitivity $2 \sim 3$ times worse for $B\parallel [100]$ than for the other directions.
 
-\subsubsection{Temporal stability}
+\subsection{Temporal stability}
 \begin{figure}[h!]
@@
 
-\subsection{Comparison with other magnetometry protocol}
-We will now compare the LFDM performances with the previously mentioned magnetometry protocols.
+\section{Comparison between LFDM and other magnetometry protocol}
+We will now compare the LFDM performances we measured with the previously mentioned magnetometry protocols.
 We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity.
@@
 
-\subsubsection{Comparison with microwave based protocols}
+\subsection{Comparison with microwave based protocols}
 
@@
 
-\subsubsection{Comparison with microwave-less protocols}
+\subsection{Comparison with microwave-less protocols}
 The only microwave-less protocols to have measured a sensitivity are the GSLAC protocols described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
@@
 In Fig.
-\ref{scan 1x1x1x1}, we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
+\ref{setup magneto}-c) we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
 In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type 1B irradiated diamond, similar to the ones we use.
@@
 
-\subsection{Material optimization for LFDM}
+\subsection{LFDM optimization}
+
+Given the few parameters involved in the LFDM protocol, the main path for LFDM optimization seem to be the material optimization of the samples.
 
@@
 For the LFDM protocol however, the material optimization is far from obvious.
-Indeed, while increasing $T_2^*$ can only improve the sensitivity, the gain might not be as interesting as for other protocols due to the broadening caused by the fluctuator lifetime.
+Indeed, while increasing $T_2^*$ would improve the sensitivity, the gain might not be as interesting as for other protocols due to the broadening caused by the fluctuator lifetime.
 On the other hand, increasing the NV density not only increases the number of sensor per volume, but also increases the dipole-dipole strength between NV centers.
-It has been my experience however that crystals with higher NV densities showed poorer LFDM performance (lower zero field PL contrast) than crystals with lower NV density.
+If we assume that fluctuators are constituted of closely packed impurities, then increasing the NV density should also increase the fluctuator proportion.
+
+It seems then that increasing the NV density would be a clear upgrade of the LFDM performances.
+In practice however, I observed that the relationship between NV density, fluctuators, and LFDM performances was generally not straightforward.
+The two following examples will illustrate that.
 
@@
 Fig.
-\ref{substrat chelou} shows various measurements on the same sample SBST-C, an HPHT bulk sample containing [NV]$\sim 1\ \rm ppm$, on two spatially distinct regions A and B.
-These two regions most likely correspond to two distinct growth sectors of the diamond, meaning that these regions were not grown in the same exact environment, which can impact the creation of impurities.
+\ref{substrat chelou} shows various measurements on the same sample SBST-C, a HPHT bulk sample containing [NV]$\sim  \rm ppm$, on two spatially distinct regions A and B.
+These two regions most likely correspond to two distinct growth sectors of the diamond, meaning that these regions were not grown in the same exact environment, which can impact among other things the creation of impurities.
+Both region showed a similar PL which would indicate a similar density of NV centers.
 
-The lifetime for both $B=0$ and $B\neq 0$ are shorter and closer to stretched exponential profile in region B than in region A, and the zero field PL contrast is larger when B is scanned in an arbitrary direction.
-These three observations seem to indicate that region B has a higher concentration of fluctuators than region A.
-When B is scanned along the [100] however (to observe double flips and the change of basis effect), there is no PL dip for region B, but there is still one for region A.
+The metric we are interested in to evaluate the LFDM performance are the PL contrast when $\mathbf{B}$ is scanned in an arbitrary direction, which mainly represents the inter-class flip flop and which I will refer to as FF-contrast; and the PL contrast when B is scanned $\mathbf{B}$ is scanned along the [100] direction, which mainly represents the double flips and which I will refer to as DF-contrast.
+We are interested in the DF-contrast since we have seen in sec.
+\ref{sec angular sensi} that it could be the main contribution to the LFDM sensitivity.
 
-\subsection{Finding the optimal parameters for LFDM}
+Fig.
+\ref{substrat chelou}-a) and b) indicate that region B shows a higher FF-contrast, but a smaller DF-contrast than region A.
+Looking at the $T_1$ measurements for both regions, region B shows a shorter $T_1$ in both zero field and non-zero field, which would indicate a larger fluctuator density in region B.
+If this is true, than the fluctuator density is not directly correlated to the DF-contrast.
 
-- Improvement to the sensitivity : Scale up (les deux papiers sur le picoTesla mW, avec les deux systèmes), material optimization (15N !, [NV] concentration, )
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/double_adamas}
+\caption{tata}
+\label{double dragon}
+\end{figure}
 
-- Understand the impact of the various factors on the flip-flop CR and DQ CR: T2*, strain, local electric noise ...
-Mettre ici les deux adamas du 20210927 avec le petit et le gros DQ, et peut etre aussi les DQ et SQ qui s'inversent sur le substrat.
-Comprendre
+We now turn to Fig.
+\ref{double dragon}.
+This figure shows the same 4 measurements but this time on two distinct samples, ADM-15-1 and ADM-15-4, which came from the same batch.
+Both samples showed relatively similar PL which could again indicate a similar NV density.
 
-- Application for uneven surfaces, polycrystalline heteroepithaxy, low/no microwave environment (dimaond anvil cells, bio sensing). Gradiometry Degen, Arcizet, MRFM
+We can first notice the disparity in the  zero field PL contrasts: the FF-contrast is twice as big in sample ADM-15-4 as in sample ADM-15-1, and the DF-contrast is almost 5 times bigger.
+When looking at the $T_1$, we can notice that the lifetime for non-zero field is slightly longer for sample ADM-15-4 than for ADM-15-1, but its lifetime in zero-field is shorter (it is also exponential and not stretched-exponential as discussed in the last chapter [REF], which could have an influence on the PL contrast).
+From the $T_1$ measurements we can not determine which sample contains the most fluctuators, and therefore correlate the fluctuator or NV density with the DF and FF-contrasts.
+
+These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field.
+We have have seen a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive.
+If they were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.
+
+Other non-material parameters in the LFDM protocols should also be investigated.
+This includes the temperature of the sample, the intensity and wavelength of the laser, the frequency and amplitude of the AC magnetic field.
+For the results presented in sec.
+\ref{sec LFDM}, we chose the parameters which we found optimal within our available experimental range.
+
+\subsection{Potential applications for LFDM}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/Applications}
+\caption{tata}
+\label{applications}
+\end{figure}
+
+We will now turn to potential applications for LFDM, which could be useful even without material optimization.
+
+Fig.
+\ref{applications} illustrates two potential applications for LFDM.
+The first one is to probe a surface with irregularities in the micron-scale to perform wide-field magnetic imaging.
+The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}.
+While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism.
+And even for two perfectly flat surface, controlling the angle between the diamond and the sample can be challenging.
+
+Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always be within $\sim$ 1 $\mu$m of the sample.
+Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.
+One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are already available.
+I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ (\rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ than for the 15 $\mu$m samples.
+This sensitivity is already better than some previous applications of wide field imaging at $20 6 \ (\rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}.
+Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required.
+Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field intensity.
+
+Another potential application for LFDM is shown in Fig.
+\ref{applications}-b.
+It consists in the detection of magnetic field gradient on top of a surface with a nano-diamond probe attached to a nano-resonator.
+The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig.
+[A faire].
+This idea is based on the previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}, and a very similar protocol as recently been developed with single NV centers \citep{huxter2022scanning}.
+
+In order to use AC LFDM with a scanning probe setup, one would need to have a nano-diamond with at least one NV center and one fluctuator.
+This may not be such a rare event: a (20 nm)$^{3}$ nanodimaond with [NV]=5 ppm contains on average $\approx 4$ NV centers.
+If we take a fluctuator proportion of 30 \% as measured in \citep{choi2017depolarization}, then finding a nanodiamond with at least one NV and one fluctuator  seems feasible.
+I have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect to better understand the nature of the fluctuators, and I therefore cannot estimate the sensitivity of such a probe.
+
+Other potential applications for LFDM includes the possibility to use polycrystalline diamond, potentially grown through heteroepitaxy on top of a surface of interest, or to follow individual diamonds flowing in liquid or living organism, similar to what was done in \citep{feng2021association}, but without having to constantly calibrate the diamond axes through ODMR which significantly hinders the temporal resolution of microwave-based technique.
+
+Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous static magnetic field or microwave field.
 

### 2022-09-01 data/20210927/mesure sensi Vs angle/traitement.py

@@ -1,0 +1,71 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+
+plt.figure(num=1,figsize=(4.5,3),dpi=80)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
+plt.locator_params(axis='x', nbins=5)
+plt.locator_params(axis='y', nbins=5)
+
+# print(glob.glob('**/*.csv',recursive=True))
+
+
+def plot_PL_100():
+	xmin=230
+	xmax=285
+	x,y=extract_data("adamas 2/scan 100 j'ai compris l'astuce je suis trop refait.csv",ycol=3)
+	x=x*35-5.4
+	y=y/max(y)
+	plt.plot(x[xmin:xmax],y[xmin:xmax],label='ADM-15-1',lw=2)
+
+	x,y=extract_data('le bon adamas/scan 100.csv',ycol=3)
+	x=x*35-6.2
+	y=y/max(y)
+	plt.plot(x[xmin:xmax],y[xmin:xmax],label='ADM-15-4',lw=2)
+
+def plot_PL_1x1x1x1():
+	xmin=300
+	xmax=740
+	x,y=extract_data('adamas 2/scan 20 deg from 100.csv',ycol=3)
+	x=x*35-5.8
+	y=y/max(y)
+	plt.plot(x[xmin:xmax],y[xmin:xmax],label='ADM-15-1',lw=2)
+
+	xmin=150
+	xmax=370
+	x,y=extract_data('le bon adamas/scan 100 20 deg from 100',ycol=3)
+	x=x*35-5.2
+	y=y/max(y)
+	plt.plot(x[xmin:xmax],y[xmin:xmax],label='ADM-15-4',lw=2)
+
+
+def plot_T1_1x1x1x1():
+	x,y=extract_data('adamas 2/T1 splitté soustraction long',ycol=5)
+	x=x*1e3
+	y=y/max(y)
+	plt.plot(x,y,label='ADM-15-1',lw=2)
+
+	x,y=extract_data('le bon adamas/T1 1x1x1x1',ycol=5)	
+	x=x*1e3
+	y=y/max(y)
+	plt.plot(x,y,label='ADM-15-4',lw=2)
+plot_T1_1x1x1x1()
+
+
+def plot_T1_100():
[…]

### 2022-09-01 python_clément/test programs/test.py

@@ -1741,8 +1741,12 @@ def test_convolution():
 
 	plt.show()
 
-
-x=np.linspace(-10,10,500)
-y=(np.sin(2*pi*x)/x)**2
-plt.plot(x,y)
+x=np.linspace(-10,10,1024)
+gauss_mask=gauss(x,sigma=0.1)
+y=np.random.randn(1024)
+yTF=np.fft.fft(y)
+yTF=yTF*gauss_mask
+yback=np.fft.ifft(yTF)
+plt.plot(yback**2)
+plt.ylim(0,3)
 plt.show()

### 2022-09-02 Manuscrit Thèse/chapter 4/Chapter 4.tex
bvfhkzdbvj

@@
 
-(mentionner l'article ici !)....
-The main motivation for this study was the potential to use NV-NV CR as a low field magnetometry protocol.
-Indeed, while NV-NV CR lines can be used to perform magnetometry with non zero fields [les russes], there are several advantages to use NV-NV CR as close to the zero magnetic field region, in particular the non dependence of the magnetic field orientation with respect to the crystal lattice.
-The behavior of
+In the last chapter we saw how dense ensemble of NV centers have an intrinsic depolarization mechanism mediated by the NV-NV dipole interaction.
+In this chapter we will see how this feature can be exploited to perform magnetometry.
+
+Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
+We will focus here on low to zero field magnetometry, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the complex physics of the NV centers in low magnetic field.
+
+This chapter is constructed as follows: We will first describe the physics of NV centers under low or transverse magnetic field (the physics is similar in both cases), we will then show both theoretically and experimentally that the NV-NV CR is modified for low and transverse magnetic field.
+We will identify different mechanisms by which the relaxation is modified and we will try to quantify their relative importance.
+
+We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field.
+We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols.
+Finally we will give some perspectives on the how to improve LFDM and on potential applications.
+
+The results presented in this chapter are based in large part on the article \citep{pellet2022spin}.
 
@@
 
+Parler du champs magnétique terrestre.
+
 \subsection{NV spin Hamiltonian under purely transverse magnetic field}
@@
 \includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2}
-\caption{tata. Changer les T1 avec T1ph=5ms}
+\caption{ODMR lineshapes for longitudinal, transverse and low magnetic fields on samples CVD-pink and Sumi-2.
+a) ODMR spectrum of a single class of NV centers for a longitudinal magnetic field ($B_\parallel \sim 100\ \rm G$).
+The microwave frequency is given with respect to the ODMR line central frequency.
+b) ODMR spectrum of a single class for a purely transverse magnetic field ($B_\perp \sim 100\ \rm G$).
+c) ODMR of all 4 classes for 0 external magnetic field.
+d) Table reporting the full width at half maximum of each ODMR line.}
 \label{ESR for T2*}
@@
 
-We should note that the current practice in magnetometry is to consider only of the three hyper-fine lines when they are resolved.
+We should note that the usual practice in magnetometry is to consider only of the three hyper-fine lines when they are resolved.
 This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines.
@@
 For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction).
-For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_I=0}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$.
+For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_s=-1,m_I=+1}$ and $\ket{m_s=+1,m_I=-1}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$.
 The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.
 
-Overall, we are mostly interested in the linewidth of the ODMR lines for various magnetic field conditions.
-While these linewidth are quite significantly changed for sample CVD-pink, the change is less pronounced on sample Sumi-2.
-Most HPHT samples we used are from type 1B diamond, and behave similarly to sample Sumi-2.
-Since almost all measurements in this chapter will be conducted on HPHT samples, we will consider the modification of $T_2^*$ as a minor effect.
+The table in Fig.
+\ref{ESR for T2*}-d) reports the linewidths measured on both sample for the various magnetic field configurations.
+We can see that the linewidth of sample CVD-pink varies by more than a factor of 2, while the ones from sample Sumi-2 varied by less than 20 \%.
+We should note that  almost all samples used in this chapter are HPHT type 1b diamond, similar to sample Sumi-2 and unlike sample CVD-pink.
+We should therefore expect the modification of $T_2^*$ to have a moderate impact on the depolarization dynamics.
+
+\medskip
+
+To summarize this section, we have seen that (for the samples used in this chapter) the NV centers spin Hamiltonian is dominated by transverse electric field for low external magnetic field, and that it causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
+We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples charged in impurities.
 
@@
 \includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
-\caption{tata. Changer les T1 avec T1ph=5ms}
+\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction.
+a) ODMR spectrum for B=0.
+b) ODMR spectrum for B=60 G.
+c) PL contrast as a function of the external magnetic field.
+d) $1/T_1^{\rm dd}$ as a function of the external magnetic field.
+A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
 \label{scan 1x1x1x1}
@@
 \section{NV-NV CR under purely transverse magnetic field}
-We now want to experimentally discriminate these different contributions and to see how close the actual decay rates are to the predicted ones.
+We have seen in the last part that there was a spontaneous spin depolarization specifically in zero external magnetic field, and we have identified three potential mechanisms which are the double-flips, the change in the Hamiltonian eigenstates and the modification of $T_2^*$.
+We now want to experimentally discriminate the contribution of each of these mechanisms and to see how close the actual decay rates are to the predicted ones.
 
@@
 \ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
+
 \subsection{Principle of the experiment}
@@
 
-We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field.
-We expect that the depolarization for low magnetic fields behave similarly, and that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
+We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field, and we expect that the depolarization for low magnetic fields behave similarly.
+We therefore claim that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
 \begin{enumerate}
@@
 \section{Introduction to ensemble NV magnetometry}
+
 Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.
@@
 
-Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of several hundreds of $\mu$m.
+Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of several hundreds of up to several mm.
 
 We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR.
-We will not try to give a complete overview of this field, but only to the elements close to the technique we present.
-For a general overview on microwave-based ensemble NV magnetometry, we recommend the excellent review from Barry et al \citep{barry2020sensitivity}.
+We will not try to give a complete overview of this field, but only of the elements close to the technique we present.
+For a general overview on NV magnetometry, we recommend the excellent following reviews  \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.
 
@@
 \subsection{Orientation-free magnetometry}
-Every protocol described so far requires a precise knowledge  of the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states versus the magnetic field.
-This knowledge implies that the direction of each NV axis in a frame of reference is known, because of the strong anisotropy in the NV Hamiltonian.
+Every protocol described so far requires a precise knowledge  of the Larmor frequency between the $\ket{0}$ and $\ket{\pm 1}$ states as a function of the magnetic field.
+This knowledge implies that the direction of each NV axis in the frame of the magnetic field is known, because of the strong anisotropy in the NV Hamiltonian.
 
@@
 Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies.
-The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the larmor frequency), or via the PL \citep{finco2021imaging}.
+The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the Larmor frequency), or via the PL \citep{finco2021imaging}.
 
@@
 
-We will first describe its operation and then characterize it and compare it to the previously mentioned NV magnetometry protocols.
+We will first describe its operation and characterize it before comparing it to the previously mentioned NV magnetometry protocols.
 
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/AC_and_DC_sensing}
+\caption{DC and AC LFDM sensing protocols on sample ADM-15-5.
+The magnetic field is scanned in an arbitrary direction.
+a) DC sensing: $B_0$ is scanned between $-$80 and +80 G while an AC field of amplitude 10 G is used for the lock-in detection.
+b) AC sensing: $B_0$ scans are performed for 5 different $|\tilde B_{\rm AC}|$ values between 2 and 10 G.
+The best AC sensitivity is achieved for $B_0\approx 5\ \rm G$.}
+\label{AC and DC sensing}
+\end{figure}
+
 Fig.
-[A FAIRE] shows how the AC and DC sensing work for LFDM.
+\ref{AC and DC sensing} shows how the AC and DC sensing work for LFDM.
 For DC sensing, we want to use an oscillating field with an amplitude that matches the width of the PL dip $|\delta B| \approx \sigma$, in order to maximize the slope of the lock-in signal around $B_0=0$.
-For AC sensing, we want to use a static field $B_0=\sigma$ corresponding to the region where the lock-in signal is most sensitive to AC field.
+For AC sensing, we want to use a static field $B_0\approx \sigma/2$ corresponding to the region where the lock-in signal is most sensitive to AC field.
 
@@
 
-where $\tau=3\ r=\rm ms$ is the LIA low-pass filter time constant.
+where $\tau=3\ \rm ms$ is the LIA low-pass filter time constant.
 
 We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$).
-Laser amplitude fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.
+Laser fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.
 
@@
 \includegraphics[width=\textwidth]{Figures/Temporal_just_0B}
-\caption{tata}
+\caption{Temporal stability of DC-LFDM as measured on sample ADM-15-5.
+a) Temporal trace of the LIA signal for $B_0 \approx 0$.
+The Lock-in signal was converted to a measured magnetic field with the previously described protocol.
+b) Estimated signal to noise ratio for a signal of 100 nT as a function of the total acquisition time $T$.
+The noise value is computed by averaging the variance of the temporal trace over each time duration $T$.
+A curve $y=0.5\, T^{1/2}$ is added for comparison}
 \label{temporal 0B}
@@
 \section{Comparison between LFDM and other magnetometry protocol}
-We will now compare the LFDM performances we measured with the previously mentioned magnetometry protocols.
+Now that we have characterized the LFDM protocol, we can compare its performances with the previously mentioned magnetometry protocols.
 We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity.
@@
 \centering
-\caption{Tutu}
+\caption{Comparison of the volume-normalized sensitivity of the best AC and DC protocols with LFDM}
  \label{table sensi volumiques}
@@
 The LFDM protocol performs $\sim 200$ times worse than the best DC magnetometer.
-It should be noted however that the PL collection in our measurement, where we use a confocal microscope using an objective with NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.
+It should be noted however that the PL collection in our measurement, where we use a confocal microscope with a numerical aperture NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.
 
@@
 \includegraphics[width=\textwidth]{Figures/compariason_0B_MW}
-\caption{tata}
+\caption{Comparison between LFDM and microwave-based magnetometry on the same experimental setup.
+The sample used is ADM-15-5 and the magnetic field is aligned close to a [111] axis.
+a) Schematics of the experiment: the setup is the same as the one described in Fig.
+\ref{setup magneto} with the addition of a continuous microwave tone at the frequency $\nu=2600\ \rm MHz$.
+b) LIA signal as $B_0$ is scanned from $-$150 to +150 G.
+The two lines at $\pm 100\ \rm G$ are caused by the microwave.
+c) Temporal trace of the measured B field for $B_0 \approx 0$ (on the middle of the zero field slope) and $B_0 \approx 100\ \rm G$ (on the middle of the microwave slope).
+The orange curve is shifted by $\approx 100\ \rm G$ to sit on the same scale as the blue curve.
+d) Estimated SNR fo a signal of 100 nT for $B_0 \approx 0$ and $B_0 \approx 100\ \rm G$ as a function of the total acquisition time}
 \label{comparaison 0B MW}
@@
 
-We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the microwave power (although $\Omega_{Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$.
+We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the available microwave power (although $\Omega_{Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$.
 Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.
@@
 \includegraphics[width=\textwidth]{Figures/substrat}
-\caption{tata}
+\caption{Measurements on sample SBST-C for two different regions A and B.
+a) PL contrast as a function of a magnetic field scanned in an aribitrary direction (same direction for both regions).
+b) PL contrast as a function of a magnetic field scanned in the [100] direction.
+c) $T_1$ measurement on a single class when $B\neq 0$.
+d) $T_1$ measurement on all  four classes when $B=0$ }
 \label{substrat chelou}
@@
 \includegraphics[width=\textwidth]{Figures/double_adamas}
-\caption{tata}
+\caption{Same measurements as Fig.
+\ref{substrat chelou} on two distinct samples ADM-15-4 and ADM-15-1.
+Every parameters (laser power,optical setup , AC magnetic field) was kept identical for the two samples.}
 \label{double dragon}
@@
 These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field.
-We have have seen a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive.
-If they were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.
+We have noticed a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive.
+If these parameters were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.
 
@@
 \includegraphics[width=\textwidth]{Figures/Applications}
-\caption{tata}
+\caption{Schematics of two potential application for LFDM. The working of each application is detailed in the main text}
 \label{applications}
@@
 Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.
-One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are already available.
-I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ (\rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ than for the 15 $\mu$m samples.
-This sensitivity is already better than some previous applications of wide field imaging at $20 6 \ (\rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}.
+One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are well suited for this task.
+I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples.
+This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}.
 Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required.
-Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field intensity.
+Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field homogeneity.
 
@@
 \ref{applications}-b.
-It consists in the detection of magnetic field gradient on top of a surface with a nano-diamond probe attached to a nano-resonator.
+It consists in the detection of magnetic field gradient on top of a surface thanks to a nano-diamond probe attached to a nano-resonator.
 The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig.
-[A faire].
-This idea is based on the previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}, and a very similar protocol as recently been developed with single NV centers \citep{huxter2022scanning}.
+\ref{AC and DC sensing}-b).
+This idea is based on the previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}.
+A very similar protocol was recently developed with microwave based AC sensing \citep{huxter2022scanning}.
 
@@
 
-Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous static magnetic field or microwave field.
+Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous magnetic field or microwave field along the sample.
 

### 2022-09-04 Manuscrit Thèse/chapter 1/Chapter 1.tex
tata

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric-comp,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_ch1.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{}
\begin{document}
\chapter{Introduction to NV center and its common usage}
In this first chapter we will introduce the

\section{The NV center as a physical object}
\subsection{Diamond overview}
\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{Figures/diamant_et_NV}
\caption{a) Drawing of a diamond with red fluorescent spots symbolizing NV centers. b) Zoom-in on the crystalline structure of an NV center}
\label{diamond+NV}
\end{figure}
We will start this chapter by discussing the physical aspect of the NV center, and of its host material, the diamond. We will cover some of the basic properties of diamond and of its most common defects before addressing the creation process of both diamonds and NV centers.

Fig. \ref{diamond+NV} illustrates a diamond with red fluorescent spots symbolizing NV centers, as well as the crystalline structure of the NV centers, where two carbon atoms are missing and one of them is replaced by a nitrogen atom.
\subsection{Diamond overview}
\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{Figures/phase_diagram_carbon_wiki}
\caption{Pressure-temperature phase diagram of elemental carbon. From wikipedia, adapted from \citep{bundy1989pressure, bundy1996pressure}.}
\label{carbon phase diagram}
\end{figure}

Diamond is a remarkable material in many aspects, it is famous for having the highest hardness and thermal conductivity of any natural material, and for its high optical refractive index and optical dispersion which makes it an extremely valuable crystal in jewelry.

Diamond is also famous for being extremely rare in nature, even though its only constituent, carbon, is one of the most abundant element on the Earth surface. The reason is that diamond is not the most stable phase of carbon under atmospheric pressure and temperature. This honor goes to the graphite which is found abundantly in coal.

Fig. \ref{carbon phase diagram} shows the PT phase diagram of elemental carbon. We can see that diamond becomes the thermodynamically stable phase only for pressure above a few GPa, $10^4$ times higher than atmospheric pressure. Even though diamond is unstable under ambient conditions, the extremely high energy activation needed to break the sp$^3$ bond between two carbon atoms means that the kinematics of the $C_{\rm diamond} \to C_{\rm graphite}$ transition is almost completely frozen. In many ways diamonds are more stable than graphite under ambient condition due to their relative chemical inertness.

Where the thermodynamic equilibrium really matters is for the crystallization process. The higher pressure needed to reach the diamond-stable region can only be found naturally deep under the earth crust, typically between 150 and 240 km below sea level \citep{tappert2011diamonds}, which explain their rarity. It also explains why, despite many previous efforts, diamonds were synthetically produced for the first time only in 1953 \citep{barnard2000diamond}, more than 50 years after the first synthetic sapphires. To this day, artificial diamond synthesis remains an active field of research \citep{shenderova2019synthesis, achard2020chemical}.

Diamonds, either natural or artificial, are used in the lab and in industry for their extreme sturdiness, with applications ranging from abrasive to diamond anvil cells \citep{jayaraman1983diamond}. More recently, diamonds have been the object of optics and electronics applications. For both of these fields however, the focus is no longer solely on the diamond but also on the defects found inside it.

\subsection{Point defects in diamond}
\begin{figure}[h!]
\centering
\includegraphics[width=0.9\textwidth]{Figures/Diamond_type}
\caption{Traditional classification of diamonds, based on \citep{tappert2011diamonds}}
\label{diamond type}
\end{figure}
Diamond in its intrinsic form is an insulator with a bandgap of of $\sim 5.47\ \rm eV$ and is transparent from far infrared to deep UV. As any crystal however, diamond is prone to defects. We will concentrate here on 0D point-like defects and omit structural defects such as dislocations, even though those too can affect the optical properties of the diamond \citep{collins2000colour}.

Point-like defects can be constituted by the absence of a carbon atom (``vacancies"), an interstitial defect between two carbon atoms or the substitution of a carbon atom with another atom. These impurities result in unpaired electron or holes localized around the impurity site. This localization causes discrete energy levels for the holes or the electrons, some of which will lie inside the diamond bandgap and will impact the crystal optical and electrical properties.

Defects with electronic transitions in the optical range are called ``colored centers" as they are responsible for the coloring of the gems.  H, He, Li, B, N,
O, Ne, P, Si, As, Ti, Cr, Ni, Co, Zn, Zr, Ag, W, Xe and Tl are known to form colored centers when introduced in diamond \citep{zaitsev2013optical, shenderova2019synthesis}, and each of these elements can form several defects. Nitrogen alone can form more than 50 different colored centers \citep{dobrinets2016hpht, ashfold2020nitrogen}.

Nitrogen, and to a lesser extant boron, are the most commonly found extrinsic elements in natural diamond. The traditional classification of diamond is presented in Fig. \ref{diamond type}. It is based on the N and B concentration, with a threshold of a few ppm (part per million) for each species as this was the smaller amount easily detectable through IR absorption. Type Ia diamonds contain clustered N defects such as the B-center which is the N$_4$V defect. In contrast, type Ib diamond mostly contains C-center which are single substitutional N$_s$ defects, also referred to as P1 centers in the spin community. Type IIb diamond contains B impurities which give them a blueish-grey color, and type IIa diamond contains no detectable impurities via IR absorption \citep{ashfold2020nitrogen}.

Nitrogen-vacancy centers are a rare occurrence in both natural diamonds and untreated synthetic diamonds, but we will see later that the much more common N$_s$ defects can be converted in NV centers. When working with NV centers, the starting crystal is therefore often a type Ib if one wants to work with ensemble, or a type IIa if one wants to work with single NV centers.

The boron defects are mainly studied in the context of p-doped diamond semi-conductor: a concentration of [B] $\approx 1000\ \rm ppm$ is needed to achieve a significant overlap between the impurity sites, which turns the diamond into a semi-metal \citep{macpherson2015practical}.Other defects of interest, generally not present naturally but introduced voluntarily, include P and S defects to create n-doped diamond \citep{das2022diamond} (still an active field of research), as well as group IV-vacancy defects (SiV, GeV, SnV, PbV) for quantum optics application \citep{bradac2019quantum}.

The NV center however remains by far the most studied defect in diamond due to its unique spin properties which will be explained below.

\subsection{Creation of synthetic diamond for NV application}

\printbibliography
\end{document}

### 2022-09-04 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

@@
 Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
-We will focus here on low to zero field magnetometry, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the complex physics of the NV centers in low magnetic field.
+We will focus here on the low to zero magnetic field region, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the more complex physics of the NV centers under low magnetic field.
 
@@
 We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols.
-Finally we will give some perspectives on the how to improve LFDM and on potential applications.
+Finally we will give some perspectives on how to improve LFDM and on potential applications.
 
@@
 
-Due to the large zero field splitting $D=2870 MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian).
+Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian).
 The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.
@@
 Given the simulation results, the only parameter that can give rise to this shape is the electric field.
-We will therefore consider that the NV Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.
+
+We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample.
+In particular, we do not shield for the earth magnetic field in our experiment.
+The splitting caused by the earth magnetic field has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$ is $\sim 4$ times weaker than the splitting we typically observe.
+
+We will therefore consider that the low external magnetic field spin Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.
 
@@
 
-Parler du champs magnétique terrestre.
-
 \subsection{NV spin Hamiltonian under purely transverse magnetic field}
@@
 
-We will consider here the case of purely transverse magnetic field with respect to the NV axis, i.e.
-$\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
-In practice, this generally means $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.
+We now turn to the case of purely transverse magnetic field with respect to the NV axis, i.e.
+$\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
+In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.
 
-In this regime, the NV Hamiltonian eigenstates are similar to the case dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{\pm}$\citep{qiu2021nuclear, qiu2022nanoscale}, of eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D$ and $D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
+In this regime, the NV Hamiltonian eigenstates are similar to the case where it is dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{\pm}$\citep{qiu2021nuclear, qiu2022nanoscale}, of eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D$ and $D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
 \begin{equation}
@@
 
-In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian or $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
 
@@
 $\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-$\bar{G}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
+$\bar{G}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
 $\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
@@
 To quantify its impact in the depolarization rate, we employ eq.
-[REF], wherche use the values $2\gamma_f = 6.5\ \rm MHz$ and we take for $\Gamma_f=\Gamma_NV$ the values of half width at half maximum found in Fig.
+[REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
 \ref{ESR for T2*}.
 
-We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 6 \%$ for sample Sumi-2, and by $\sim 19 \%$ for sample CVD-pink.
+We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
 
@@
 Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup.
-to improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing.
+To improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing.
 Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.
@@
 \section{Conclusion and perspectives}
-We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field.
-We have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers.
-
+We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field, and wee have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers.
 We have identified three mechanisms by which the depolarization is increased in zero-field: the inter-class flip-flops, the double flips, and the change in the NV eigenstates (which includes the change in $T_2^*$).We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
@@
 The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}.
-While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism.
-And even for two perfectly flat surface, controlling the angle between the diamond and the sample can be challenging.
+While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism samples.
+Furthermore, controlling the angle between the diamond and the sample can be challenging even for two perfectly flat surface.
 
-Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always be within $\sim$ 1 $\mu$m of the sample.
+Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always have a sensor within $\sim$ 1 $\mu$m of the sample.
 Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.

### 2022-09-05 Manuscrit Thèse/chapter 1/Chapter 1.tex
bydr

@@
 \section{The NV center as a physical object}
-\subsection{Diamond overview}
 \begin{figure}[h!]
 \centering
-\includegraphics[width=0.6\textwidth]{Figures/diamant_et_NV}
-\caption{a) Drawing of a diamond with red fluorescent spots symbolizing NV centers. b) Zoom-in on the crystalline structure of an NV center}
+\includegraphics[width=0.9\textwidth]{Figures/diamant_et_NV}
+\caption{From the diamond to the NV center.
+a) Drawing of a bulk diamond.
+We consider a diamond to be ``bulk" if it is bigger than a few 100 $\mu$m.
+b) Drawing representing a fluorescence microscopy image of a diamond surface.
+The red fluorescent spots symbolize NV centers.
+c) Zoom-in on the crystalline structure of an NV center.
+Block dots represent carbon atoms}
 \label{diamond+NV}
@@
 Fig.
-\ref{diamond+NV} illustrates a diamond with red fluorescent spots symbolizing NV centers, as well as the crystalline structure of the NV centers, where two carbon atoms are missing and one of them is replaced by a nitrogen atom.
+\ref{diamond+NV} illustrates the various scales to consider in this study: we start with a diamond which for lab purposes is generally no bigger than a few mm.
+If we zoom-in on the surface of some diamonds, we can find bright fluorescent spots corresponding to particular crystalline defects.
+Standard confocal microscopy is limited in its resolution to the $\sim \mu$m range, but if we could zoom-in even further we would see the structure of the defect, illustrated here with the crystalline structure of an NV center.
 \subsection{Diamond overview}

### 2022-09-05 Manuscrit Thèse/chapter 4/Chapter 4.tex
bydr

@@
 
+\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
 \title{}
@@
 
-Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates.
+Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates compared to the high magnetic field regime.
 
@@
 
-Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian).
+Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian).
 The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.
@@
 Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR.
-Given the simulation results, the only parameter that can give rise to this shape is the electric field.
+Given the simulation results, the only parameter that can give rise to this shape is the local electric field.
 
@@
 In particular, we do not shield for the earth magnetic field in our experiment.
-The splitting caused by the earth magnetic field has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$ is $\sim 4$ times weaker than the splitting we typically observe.
+The splitting caused by the earth magnetic field however has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$, which is $\sim 4$ times weaker than the splitting we typically observe.
 
@@
 
-In this regime, the NV Hamiltonian eigenstates are similar to the case where it is dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{\pm}$\citep{qiu2021nuclear, qiu2022nanoscale}, of eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D$ and $D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
+In this regime, the NV Hamiltonian eigenstates are similar to the case where it is dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{-}, \ket{+}$\citep{qiu2021nuclear, qiu2022nanoscale}, of respective eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D,D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
 \begin{equation}
@@
 
-For the case where $d_\perp E_\perp \sim \frac{(\gamma_e B_\perp)^2}{D}$ and $\phi_E \neq 2\phi_B$, the eigenstates of the Hamiltonian are still $\ket{0},\ket{\pm}$ with a relative angle $\phi$ in between $\phi_E$ and $2\phi_B$.
+For the case where $d_\perp E_\perp \sim \frac{(\gamma_e B_\perp)^2}{D}$ and $\phi_E \neq 2\phi_B$, the eigenstates of the Hamiltonian are still of the form $\ket{0},\ket{\pm}$ with a relative angle $\phi$ in between $\phi_E$ and $2\phi_B$.
 
@@
 
-\begin{align}
-\bar{\bar{A}} &= \begin{pmatrix}
+\begin{equation}
+\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
+\end{equation}
+
+and
+\begin{equation}
+\bar{\bar{A}} = \begin{pmatrix}
 A_{xx} & 0 & 0 \\
@@
 0 & 0 & A_{zz}
-\end{pmatrix} \\
-\mathcal{H}_n &= \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
-\end{align}
+\end{pmatrix},
+\end{equation}
 
@@
 \ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$.
-A table with the full width at half maximum of each ODMR line is shown in Fig.
-\ref{ESR for T2*}-d)
 
-While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2, being a Type 1B HPHT sample, contains significantly more impurities besides NV centers, most likely P1 centers.
+While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers.
 These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.
@@
 
-We should note that the usual practice in magnetometry is to consider only of the three hyper-fine lines when they are resolved.
+We should note that the usual practice in magnetometry is to consider the linewidth only of the three hyper-fine lines when they are resolved.
 This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines.
@@
 Similarly, the hyper-fine structure is hidden in this configuration since all three hyper-fine levels are nearly-degenerate, provided that $\frac{(\gamma_e B_\perp)^2}{D} > A_{xx},A_{zz},Q$ which is typically the case for $B_\perp > 40\ \rm G$.
-We can note that the electric field noise is significantly stronger in the HPHT sample, which leads to an ODMR linewidth more than twice as big.
+We can note that the electric field noise is significantly stronger in sample Sumi-2, which leads to an ODMR linewidth more than twice as big.
 
@@
 To summarize this section, we have seen that (for the samples used in this chapter) the NV centers spin Hamiltonian is dominated by transverse electric field for low external magnetic field, and that it causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
-We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples charged in impurities.
+We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples with a high impurity density.
 
@@
 
-The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust to isolate each effects.
+The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust in order to isolate each effects.
 
@@
 \includegraphics[width=0.9\textwidth]{Figures/scan_100}
-\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis. Changer les T1 avec T1ph=5ms}
+\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis.}
 \label{scan 100}
@@
 We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig.
-\ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of [REF] in zero field, whereas in Fig.
-\ref{scan 100}, it was only reduced by a factor of [REF].
-The main reason for the PL and $T_1$ drop was indeed the co-resonance between the four classes.
+\ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of $\sim$ 8 in zero field, whereas in Fig.
+\ref{scan 100}, it was only reduced by a factor of $\sim$ 1.2.
+The main reason for the PL and $T_1$ drop in Fig.
+\ref{scan 1x1x1x1} was indeed the co-resonance between the four classes.
 
-Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained by considering only the inter-class resonances.
-There are some additional depolarization mechanisms which are proper to the zero field region.
+Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained if we consider only the inter-class resonances.
+This means that there are some additional depolarization mechanisms which are proper to the zero field region.
 We should also note that, while the zero-field PL contrast is bigger in Fig.
@@
 \midrule
-$\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-$\bar{G}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
-$\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
+$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
+$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
+$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
 %\bottomrule
@@
 To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
-We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{G}$ factor defined in [REF].
+We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
 Finally, the question about the correlation  length of the electric field also needs to be asked.
@@
 The results are summed up in Table \ref{table double flips}.
-We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{G}^2$ factor: $\bar{G}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
-$\bar{G}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$).
+We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
+$\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$).
 This is close to the the experimental values we have.
-Finally $\bar{G}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $).
+Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $).
 We come back to the results of the last section where only considered the flip-flops.
-
-We therefore predict that the double flips amounts for an increase of the decay rate by [REF] \% (taking $\bar{G}^2=0.5$) compared to the previous case.
 
@@
 
-\subsection{Other potential causes}
-We have also considered two other causes that we have ultimately ruled out in this case.
+\subsection{Other potential causes for zero field depolarization}
+We have also considered two other potential causes that we have ultimately ruled out in this case.
 
@@
 
-We still tried to estimate the sensitivity of the low field depolarization with the field orientation.
+We still estimated how sensitive the PL dip was with respect to the field alignment.
 Fig.
@@
 \includegraphics[width=.7\textwidth]{Figures/pola_laser}
-\caption{tata}
+\caption{PL of sample ADM-150-3 with respect to the magnetic field for various laser polarization angle.
+The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
 \label{pola laser}
@@
 \includegraphics[width=\textwidth]{Figures/transis_transverse_field}
-\caption{Simulation of the eigenstates of the NV Hamiltonian for $d_\perp E_perp=4\ \rm MHz$ and a purely transverse magnetic field $B_\perp$.
+\caption{Simulation of the eigenstates of the NV Hamiltonian for  a purely transverse magnetic field $B_\perp$.
+In addition to the transverse magnetic field, we consider a fixed transverse electric field of value $d_\perp E_\perp=4\ \rm MHz$.
 (a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$.
-The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$ (b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
+The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$.
+(b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
 \label{eigenstates transverse field}
@@
 \ref{eigenstates transverse field} shows a simulation of the eigenstates and eigenenergies for a NV Hamiltonian subject to transverse electric field and magnetic field.
-For convenience we chose to take $E_\perp = E_x$ and $B_\perp = B_x$, taking other geometric configuration only alters slightly the results.
+For convenience we chose to take $E_\perp = E_x$ and $B_\perp = B_x$, taking other geometric configuration alters only slightly the results.
 We labeled the eigenstates of the Hamiltonian $\ket{1}$, $\ket{2}$ and $\ket{3}$ in ascending order of energy.
@@
 We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$.
-In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ or $\Delta \nu=\nu_+=\nu_-$ are increased.
+In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ is increased.
 
@@
 
-We attribute the decrease in region A to the double flips: as the magnetic field and $\Delta \nu$ are increased, the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$.
+We attribute the decrease in region A to the double flips: as the magnetic field increases, so does $\Delta \nu$ and the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$.
 This value is coherent with the previously measured fluctuator linewidth.
-We can then estimate the part of the double flips in the low field (20 G here) depolarization rate and find that they amount for $\sim 50\%$ in this case.
+We can then estimate the part of the double flips in the total depolarization rate for the low field regime (20 G here) and find that they amount for $\sim 50\%$ in this case.
 
-Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see that they are not very far detuned in Fig.
-\ref{champ tranverse exp}-a).
+Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see in Fig.
+\ref{champ tranverse exp}-a) that they are not very far detuned.
 However we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.
@@
 
-Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of several hundreds of up to several mm.
+Ensemble NV centers center on the other hand offer sensitivities as low as $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of sizes up to several mm.
 
@@
 We will not try to give a complete overview of this field, but only of the elements close to the technique we present.
-For a general overview on NV magnetometry, we recommend the excellent following reviews  \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.
+For a general overview on NV magnetometry, we recommend the excellent reviews  \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.
 
 \subsection{AC and DC magnetometry}
-The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted in population.
+The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted into spin population.
 These protocols are generally distinguished between DC-broadband and AC-narrow band.
@@
 Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry.
-The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20 pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
+The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
 
-For AC magnetometry, the basic protocol consist of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
+For AC magnetometry, the basic protocol consist of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
 Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field.
@@
 The main technique for low field magnetometry consist in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}.
-The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$.
+The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$ \citep{zheng2019zero}.
 
@@
 \subsection{Microwave-free magnetometry}
-All the precedent magnetemetry protocols described relied on the use of a controllable magnetic field, to coherently manipulate the NV spin state and to measure the action of the magnetic field on the NV Larmor frequency.
+All the precedent magnetemetry protocols described relied on the use of a controllable magnetic field, to coherently manipulate the NV spin state and to measure the phase imprinted by the magnetic field in the rotating frame.
 
 There are other protocols which correlate the magnetic field with the spin $T_1$ time (relaxometry).
-These protocols operate without microwave field and work for either DC or high frequency AC sensing.
+These protocols can operate without microwave field and work for either DC or high frequency AC sensing.
 
@@
 
-The GSLAC occurs for a longitudinal magnetic field when $\gamma B_z = D$, which corresponds to $B=1024\ \rm G$.
+The GSLAC occurs for a purely longitudinal magnetic field where $\gamma B_z = D$, which corresponds to $B=1024\ \rm G$.
 In these conditions, the normally bright $\ket{0}$ state and the dark $\ket{-1}$ state become resonant and any residual mixing between these two states (coming from the electric field, the strain or the residual transverse magnetic field) will cause a drop in the NV polarization, and in the PL \citep{broadway2016anticrossing}.
 
-This protocol has been implemented, first only for longitudinal magnetic field measurement \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.
+A protocol based on the GSLAC line has been implemented, first only for longitudinal magnetic field measurement \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.
 
@@
 The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed.
-The detection bandwidth is adjusted with an external magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.
+The detection bandwidth is adjusted with an external DC magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.
 
-Sensitivities as low as a few $pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
+Sensitivities as low as a few $\rm pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
 
@@
 
-To our knowledge, the only orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
+To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
 This protocol however has an extremely low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic}.
@@
 We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the spontaneous depolarization of dense NV ensemble at low magnetic field.
-This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional field is required.
+This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional field is required to improve performances.
 
@@
 The experimental setup is similar to the one presented in Fig.
-[REF] with the exception that there are no microwaves, and that the magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\delta B$ used for the modulation.
+[REF] with the exception that there are no microwaves, and that the magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\tilde{B}_{\rm AC}$ used for the modulation.
 We typically use a modulation frequency $f_{\rm mod} \approx 1\ \rm kHz$, the same value we use for ODMR measurements.
-The main voltage $V_0$ responsible for the DC field and the oscillating voltage $\delta V$ responsible for the AC field are summed via a homemade bias tee.
+The main voltage $V_0$ responsible for the DC field and the oscillating voltage $\tilde{V}_{\rm AC}$ responsible for the AC field are summed via a homemade bias tee.
 
@@
 \ref{setup magneto}-b) and c) show respectively the PL measured directly on the photodiode and the lock-in signal when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G.
-The magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.
+For this configuration, the magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.
 
@@
 Fig.
-\ref{AC and DC sensing} shows how the AC and DC sensing work for LFDM.
-For DC sensing, we want to use an oscillating field with an amplitude that matches the width of the PL dip $|\delta B| \approx \sigma$, in order to maximize the slope of the lock-in signal around $B_0=0$.
-For AC sensing, we want to use a static field $B_0\approx \sigma/2$ corresponding to the region where the lock-in signal is most sensitive to AC field.
+\ref{AC and DC sensing} shows how the AC and DC sensing work for LFDM: for DC sensing the LIA signal around $B_0=0$ provide a sharp linear slope to convert the LIA voltage into a magnetic field.
+For AC sensing, the LIA signal is directly proportional to the amplitude of the field.
+The optimal AC field amplitude for DC sensing has to match the width of the PL dip $|\tilde{B}_{\rm AC}| = \sigma \approx 10\ \rm G$ in order to maximize the slope of the lock-in signal around $B_0=0$.
+the optimal DC field to use for AC sensing is $B_0= \sigma/2$ corresponding to the region where the lock-in signal is most sensitive to AC field.
 
@@
 
-We characterize the DC sensitivity by looking at the noise floor on the LIA signal when being on the steepest slope, around $B_0=0$.
+We characterize the DC sensitivity by looking at the noise floor on the LIA signal when sitting on the steepest slope, around $B_0=0$.
 The following parameters were used in the case presented: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
@@
 \ref{angular sensi}-c) shows the sensitivity, measured with the protocol described in Fig.
-\ref{sensei alternee}, with respect to the magnetic field angle with the [100] axis.
+\ref{sensei alternee}, with respect to the angle between the magnetic field and the [100] axis.
 We can notice that the sensitivity depends only weakly on the magnetic field angle.
@@
 
-One of the main source of drift for NV magnetometry based on spin resonance is the lattice temperature change: the $D$ factor in the NV Hamiltonian (eq.
+One of the main source of drift for NV magnetometry based on spin resonance is the crystal lattice temperature change: the $D$ factor in the NV Hamiltonian (eq.
 [REF]) is sensitive to the lattice temperature because of the thermal dilatation of the crystal, which is the basis for NV thermometry \citep{kucsko2013nanometre}.
@@
 \ref{temporal 0B}, on both the zero field line and the microwave line.
-While we can see that the microwave signal is more sensitive thanks to the higher slope, it is also more prone to drift, most likely because of the shift of the Larmor frequency caused by the temperature fluctuations.
+While we can see that the microwave signal is more sensitive thanks to the higher slope, it is also more prone to drift, most likely because of the shift of the Larmor frequency caused by temperature fluctuations.
 We can even see that the LFDM becomes more sensitive for acquisition times beyond 10 s.
 
-We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the available microwave power (although $\Omega_{Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$.
+We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the available microwave power (although in this case $\Omega_{\rm Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$.
 Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.
@@
 \ref{setup magneto}-c) we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
-In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type 1B irradiated diamond, similar to the ones we use.
-In \citep{zheng2020microwave} they manage to keep a contrast of 4.5 \% while reducing the GSLAC linewidth to $0.38 G$ by using a CVD sample with a lower [NV] and [N] density and isotopically enriched in $^{12}$C.
+In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type Ib irradiated diamond, similar to the ones we use.
+In \citep{zheng2020microwave} they manage to keep a contrast of 4.5 \% while reducing the GSLAC linewidth to $0.38\ \rm G$ by using a CVD sample isotopically enriched in $^{12}$C with a lower [NV] and [N] density.
 
@@
 
-Most magnetometry protocols have a relatively clear material optimization path: DC protocols want to improve the figure of merit $\rho T_2^*$ and AC protocols want to improve $\rho T_2$, where $\rho=[NV]$ is the NV concentration.
+Most magnetometry protocols have a relatively clear material optimization path: DC protocols want to improve the figure of merit $\rho T_2^*$ and AC protocols want to improve $\rho T_2$, where $\rho=[\rm NV]$ is the NV concentration.
 The reason is that the number of NV centers for a given volume scales as $\rho$, and therefore the sensitivity scale as $\sqrt{\rho}$.
@@
 
-It seems then that increasing the NV density would be a clear upgrade of the LFDM performances.
+It seems then that increasing the NV density would be the obvious path to improve the LFDM performances.
 In practice however, I observed that the relationship between NV density, fluctuators, and LFDM performances was generally not straightforward.
@@
 \ref{applications} illustrates two potential applications for LFDM.
-The first one is to probe a surface with irregularities in the micron-scale to perform wide-field magnetic imaging.
+The first one is to to perform wide-field magnetic imaging on a surface with micron-sized irregularities.
 The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}.
@@
 Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.
-One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are well suited for this task.
+One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$m red fluorescent diamond) are well suited for this task.
 I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples.
@@
 \ref{applications}-b.
-It consists in the detection of magnetic field gradient on top of a surface thanks to a nano-diamond probe attached to a nano-resonator.
+It consists in the detection of magnetic field gradient thanks to a nano-diamond probe attached to a nano-resonator.
 The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig.
 \ref{AC and DC sensing}-b).
-This idea is based on the previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}.
+This idea is based on previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}.
 A very similar protocol was recently developed with microwave based AC sensing \citep{huxter2022scanning}.
@@
 If we take a fluctuator proportion of 30 \% as measured in \citep{choi2017depolarization}, then finding a nanodiamond with at least one NV and one fluctuator  seems feasible.
-I have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect to better understand the nature of the fluctuators, and I therefore cannot estimate the sensitivity of such a probe.
+I have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect, and I therefore cannot estimate the sensitivity of such a probe.
 

### 2022-09-06 Manuscrit Thèse/chapter 1/Chapter 1.tex
ghker

@@
 \chapter{Introduction to NV center and its common usage}
-In this first chapter we will introduce the
+The aim of this first chapter is to introduce the tools and concepts required for the next chapters. We will start
 
@@
 \end{figure}
+Refaire ce paragraphe te foutre le summary après la description de la figure.
 We will start this chapter by discussing the physical aspect of the NV center, and of its host material, the diamond.
@@
 
-\subsection{Creation of synthetic diamond for NV application}
+\subsection{Creation of synthetic diamond for optics and electronics applications}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/HPHT_and_CVD}
+\caption{Schematics of HPHT (a) and CVD (b) growth process detailed in main text}
+\label{HPHT and CVD}
+\end{figure}
+
+We will now focus on the creation process of the diamonds.
+Almost all diamonds used in optics or electronics experiments today are synthetic, due to the higher control of the growth process required for these applications.
+There are two main competing technologies for the growth of large ($\mu \rm{m} \sim \rm{mm}$) single crystal diamonds: High pressure high temperature growth (HPHT) and chemical vapor deposition growth (CVD).
+Both of these processes have their own specificities which we will detail below.
+
+\subsubsection{HPHT growth}
+
+HPHT was the first route for synthetic diamonds, first developed in the 1950s \citep{barnard2000diamond ,bundy1955man}.
+It consists in reconstructing an environment where diamond is the thermodynamically stable phase of carbon, similarly to the growth process of natural diamonds.
+
+Fig.
+\ref{HPHT and CVD}-a) illustrates the apparatus of HPHT synthesis: a carbon source (graphite) and a solvent metal composed of transition metals (Fe, NI, Co typically \citep{bundy1963direct}) are put under $5 \sim 6$ GPa pressure in a press and heated to temperatures $\geq 1300\ ^\circ$C.
+A diamond crystal seed is put in the metal catalyst and a temperature gradient insures that the carbon migrates from the graphite to the diamond.
+
+The first synthesis of HPHT diamonds were all of type Ib due to unwanted nitrogen pollution in the metal melt.
+The first HPHT type IIa diamonds were produced in the late 90s by adding other elements to the metal melt (Ti, Al, Zr) which would react preferentially with nitrogen and prevent its incorporation in the diamond \citep{burns1999growth, sumiya2002growth}.
+
+\subsubsection{CVD growth}
+
+CVD growth for diamond was first demonstrated convincingly in the 80s \citep{matsumoto1982growth, matsumoto1982vapor, kamo1983diamond}.
+Unlike HPHT, CVD operates in a regime where diamond is not the thermodynamically stable phase.
+It relies therefore heavily on catalysts to stabilize the diamond phase versus the graphite one.
+
+The main catalyst used is atomic hydrogen, which can both create the free radicals required to form the $C-C$ diamond bonds, and which etches sp$^2$ bonds (graphite) faster than it does sp$^3$ bonds (diamond) \citep{gracio2010diamond}.
+
+Fig.
+\ref{HPHT and CVD}-b) illustrates the apparatus of a microwave plasma reactor: it consists in a modified vacuum chamber which acts as a microwave resonator.
+The microwave creates a plasma ball from the the gas introduced in the chamber (CH$_4$ and H$_2$ typically) with a plasma temperature $T \geq 3000 ^\circ$C \citep{ashfold2020nitrogen}, enough to dissociate the $H_2$ gas in atomic hydrogen \citep{balmer2009chemical}.
+The plasma acts as the carbon source for the diamond growing on top of the substrate, and as the catalyst to favor the diamond phase.
+The substrate is kept at a temperature $800 \sim 1200\ ^\circ$C during the growth.
+
+Other sources can be used to create the high temperature required for atomic hydrogen, such as a tungsten hot filament \citep{haubner1993diamond}, electrical discharge (arcjet) \citep{luque1998excited} or even an oxygen-acetylene combustion flame \citep{bachmann1991towards}.
+
+The main sources of extrinsic impurities for CVD-grown diamond comes from the source gasses \citep{balmer2009chemical}.
+The commercial availability of source gases containing less than $\sim$ ppm impurities made possible the growth of  high purity type IIa CVD diamonds \citep{kasu2003high, twitchen2004high, tallaire2006characterisation}.
+
+\medskip
+Both HPHT and CVD processes can produce diamonds of high quality for quantum optics or electronics application, each with their pros and cons.
+HPHT offers better scalability for thick crystals ($\sim \rm mm$) and allows the incorporation of heavy atoms (Ge \citep{palyanov2015germanium} , Sn \citep{ekimov2019effect}) which can be hard to incorporate from a gas phase.
+CVD on the other hand allows more versatility in the control of the impurities introduced in the crystal, including an isotopic control on both the impurities and the carbon itself.
+A particular advantage of CVD growth is the possibility of $\delta$-doping where a specific defect is grown only within a very thin layer of the crystal, as low as 5 nm \citep{ohno2012engineering, ishikawa2012optical, ohashi2013negatively}.
+
+\subsection{Optimizing NV center formation}
 

### 2022-09-06 Manuscrit Thèse/chapter 4/Chapter 4.tex
ghker

@@
 To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
-This protocol however has an extremely low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic}.
+This protocol however has an extremely low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic, chapman2013background, jones2020selective}.
 

### 2022-09-07 Manuscrit Thèse/chapter 1/Chapter 1.tex
njtgrkln

@@
 \begin{document}
-\chapter{Introduction to NV center and its common usage}
+\chapter{Introduction to NV centers and their common usage}
 The aim of this first chapter is to introduce the tools and concepts required for the next chapters. We will start
@@
 \end{figure}
-Refaire ce paragraphe te foutre le summary après la description de la figure.
+Refaire ce paragraphe te foutre le summary après la description de la figure. Et bien parler du NV- et du charge state
 We will start this chapter by discussing the physical aspect of the NV center, and of its host material, the diamond.
@@
 \includegraphics[width=\textwidth]{Figures/HPHT_and_CVD}
-\caption{Schematics of HPHT (a) and CVD (b) growth process detailed in main text}
+\caption{Schematics of HPHT (a) and CVD (b) growth process as detailed in main text}
 \label{HPHT and CVD}
@@
 
-HPHT was the first route for synthetic diamonds, first developed in the 1950s \citep{barnard2000diamond ,bundy1955man}.
-It consists in reconstructing an environment where diamond is the thermodynamically stable phase of carbon, similarly to the growth process of natural diamonds.
+HPHT was the first commercially viable process for synthetic diamonds, starting from the 1950s \citep{barnard2000diamond, bundy1955man}.
+The basic idea of HPHT synthesis is to reconstruct an environment where diamond is the thermodynamically stable phase of carbon, similarly to the growth process of natural diamond below the earth crust.
 
@@
 The main catalyst used is atomic hydrogen, which can both create the free radicals required to form the $C-C$ diamond bonds, and which etches sp$^2$ bonds (graphite) faster than it does sp$^3$ bonds (diamond) \citep{gracio2010diamond}.
+Hydrogen is by far the most abundant element in the gas phase: the precursor gases used for the growth are typically CH$_4$ and H$_2$ in proportions of $1-5\%$ and $95-99\%$ respectively \citep{achard2020chemical}.
 
@@
 \ref{HPHT and CVD}-b) illustrates the apparatus of a microwave plasma reactor: it consists in a modified vacuum chamber which acts as a microwave resonator.
-The microwave creates a plasma ball from the the gas introduced in the chamber (CH$_4$ and H$_2$ typically) with a plasma temperature $T \geq 3000 ^\circ$C \citep{ashfold2020nitrogen}, enough to dissociate the $H_2$ gas in atomic hydrogen \citep{balmer2009chemical}.
+The microwave creates a plasma ball from the the precursor gases with a temperature $T_{\rm plasma} \geq 3000 ^\circ$C \citep{ashfold2020nitrogen}, enough to dissociate the $H_2$ gas in atomic hydrogen \citep{balmer2009chemical}.
 The plasma acts as the carbon source for the diamond growing on top of the substrate, and as the catalyst to favor the diamond phase.
-The substrate is kept at a temperature $800 \sim 1200\ ^\circ$C during the growth.
+The substrate is kept at a temperature $700 \sim 1100\ ^\circ$C during the growth.
 
@@
 
-The main sources of extrinsic impurities for CVD-grown diamond comes from the source gasses \citep{balmer2009chemical}.
+For a good CVD reactor, the main sources of extrinsic impurities the diamond comes from the source gasses \citep{balmer2009chemical}.
 The commercial availability of source gases containing less than $\sim$ ppm impurities made possible the growth of  high purity type IIa CVD diamonds \citep{kasu2003high, twitchen2004high, tallaire2006characterisation}.
@@
 CVD on the other hand allows more versatility in the control of the impurities introduced in the crystal, including an isotopic control on both the impurities and the carbon itself.
-A particular advantage of CVD growth is the possibility of $\delta$-doping where a specific defect is grown only within a very thin layer of the crystal, as low as 5 nm \citep{ohno2012engineering, ishikawa2012optical, ohashi2013negatively}.
+A particular advantage of CVD growth is the possibility of $\delta$-doping where a specific defect is grown only within a thin layer of the crystal, as low as 5 nm \citep{ohno2012engineering, ishikawa2012optical, ohashi2013negatively}.
 
 \subsection{Optimizing NV center formation}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/formtion_NV_2}
+\caption{Illustration of the formation of NV centers. Black dots represents carbon atoms and green dots nitrogen atoms}
+\label{formation NV}
+\end{figure}
+
+NV centers are a relatively rare impurity in untreated diamonds.
+In CVD diamond for instance, the proportion of nitrogen atoms forming an NV center in an untreated crystal is at best of $2-3 \%$ \citep{hartland2014study} and sometimes much less.
+While this ratio can sometimes be enough  to work with single NV centers, applications such as sensing that require dense ensemble need a higher conversion rate.
+We will discuss here the usual techniques used to improve the formation of NV centers.
+
+Fig. \ref{formation NV} illustrates the steps needed to maximize the number of NV centers in the crystal while keeping a relatively high crystalline purity.
+
+The first step to create NV centers is to get individual substitutional N atoms in the crystal ($N_s$, C-centers or P1 centers depending on the literature).
+This can be achieved either \textit{in situ} by introducing (voluntarily or not) nitrogen in the crystal during the growth process \citep{tallaire2006characterisation, lobaev2017influence}, or \textit{ex situ} by implanting N$^+$ ions with ion implantation techniques \citep{meijer2005generation, smith2019colour}.
+Because this present work focuses on dense NV ensembles, we will consider the simplest case where the starting crystal is a type-Ib diamond due to the presence of nitrogen during the growth.
+A starting concentration [N$_s$]= $10 \sim 300$ ppm is typical for type Ib HPHT diamonds \citep{achard2020chemical}.
+
+The second step is to create monovacancies in the crystal lattice.
+This step might be optional for crystals which already contain a large number of monovacancies, but CVD crystals in particular can have less vacancies that they have substitutional nitrogen atoms \citep{mainwood1999point}.
+The general technique to create monovacancy is to bombard the sample with high energy particle in order to knock out carbon atoms from the crystal.
+Several high energy particle have been used to create vacancies, such as protons, electrons, ions, neutrons and even gamma rays \citep{davies1976optical, ashbaugh1988gemstone, kleinsasser2016high}.
+The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their ability to penetrate the crystal and to create vacancies uniformly.
+
+The final step is to migrate the vacancies next to the nitrogen atoms.
+The activation energy required to move vacancy is of $\sim 2.3\ \rm eV$ which corresponds to temperatures of $600\sim 700 ^\circ$C \citep{davies1992vacancy, newton2002recombination}.
+The most common method to move the vacancies is therefore to anneal the crystal at a temperature $800 \sim 1200 ^\circ$C for a few hours \citep{botsoa2011optimal}.
+NV centers anneal out at temperature of $1400 \sim 1500 ^\circ$C \citep{zaitsev2013optical, pinto2012diffusion} which explains why the vacancies tend to stick to the nitrogen atoms during the annealing phase.
+
+The final N$_s$ to NV conversion rate can be as high as $\sim 50 \%$ \citep{grezes2015storage, hartland2014study}, although this ratio tend to decrease for higher [N$_s$] concentrations.
+
+\subsection{Controlling the NV center charge state}
+
+The charge state of the NV center is of crucial importance for NV applications.
+Only the negatively charged NV$^-$ center shows the specific spin properties that will be detailed below, but the NV center can exist under three charge states in the diamond: NV$^-$, NV$^0$ and NV$^+$.
+The presence of NV$^+$ is only marginal in standard diamonds \citep{hauf2014addressing, pfender2017protecting} and we will focus here solely on the NV$^-$ and NV$^0$ charge state.
+
+The charge state of the NV centers depends not only on the growth process of the crystal, but also on the usage made of the sample.
+In particular, illumination in the optical range can promote electrons in the conduction band or holes in the valence band and modify the charge state of the NV center.
+
+On the material side, the crystal has to remain electrically neutral which means that at least as many charge donors need to be present as there are NV$^-$ centers.
+For nitrogen-rich samples, the donors are mostly N$^+$ centers, which imposes a limit on the maximum conversion rate of N to NV in order to keep a high [NV$^-$]/[NV$^0$] ratio.
+The possibility to introduce other donors, such as phosphorus, has been studied \citep{doi2016pure} but remains impractical so far \citep{barry2020sensitivity}.
+
+On the experimental side, it has been shown that specific wavelengths
 

### 2022-09-07 Manuscrit Thèse/chapter 1/Chapter 1.tex
nio

@@
 The final N$_s$ to NV conversion rate can be as high as $\sim 50 \%$ \citep{grezes2015storage, hartland2014study}, although this ratio tend to decrease for higher [N$_s$] concentrations.
+Most sample studied in this manuscript had a concentration $[\rm NV]= 1-10\ \rm ppm$
 
@@
 
-On the experimental side, it has been shown that specific wavelengths
+On the experimental side, it has been shown that the ionization of NV$^-$ to NV$^0$ and the reverse process depend on the intensity and wavelength of the laser used for illumination \citep{aslam2013photo}.
+It was found that the optimal wavelength to preserve the NV$^-$ charge state was  $510-540\ \rm nm$.
+The most common wavelength used with NV centers is 532 nm (green) due to the historical importance of Nd:YAG lasers.
+On the other hand, it is possible to semi-permanently ionize the NV$^-$ into NV$^0$ when using a red excitation, which has been used to create an optical memory \citep{dhomkar2016long}.
+
+Ccl sur le fait que osef pour les échantillon en excès de N mais que ça peut avoir de l'importance avec les samples low en impurities.
 

### 2022-09-07 Manuscrit Thèse/chapter 1/Chapter 1.tex
nuno

@@
 
-Ccl sur le fait que osef pour les échantillon en excès de N mais que ça peut avoir de l'importance avec les samples low en impurities.
+For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one.
+For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of [NV$^0$] can reach 50\% of the total [NV] concentration and even more [les refs chez achard et barry]
+
+\section{The NV$^-$ center physics}
 

### 2022-09-09 Manuscrit Thèse/chapter 1/Chapter 1.tex
gbrie

@@
 For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one.
-For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of [NV$^0$] can reach 50\% of the total [NV] concentration and even more [les refs chez achard et barry]
+For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach 50\% of the total NV concentration \citep{grezes2015storage}
 
-\section{The NV$^-$ center physics}
+\section{The NV$^-$ center optical and spin properties}
 
+We will now focus on the magnetic and optical properties of the NV$^-$ center. This section will cover...
+
+\subsection{NV center molecular orbitals}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{Figures/NV_electrons}
+\caption{(a) Representation of the nitrogen and carbon orbitals surrounding the vacancy of an NV center.
+Each black dot represents an electron.
+A sixth electron is required to form the negatively charged NV$^-$ state.
+(b) Molecular orbitals of the NV center predicted from the the $C_{3v}$ symmetry of the defect.
+The electron occupation of the NV$^-$ ground state is represented by red arrows.}
+\label{NV electrons}
+\end{figure}
+
+We will first cover the molecular aspect of the NV defect in diamond and the energy levels of the defect predicted from group theory and ab initio computations.
+
+Fig.
+\ref{NV electrons}-a) illustrates the $sp^3$ orbitals of the carbons and nitrogen atoms neighboring the vacancy, as well as the 5 electrons occupying dangling bonds from these atoms and a sixth electron captured to form the negatively charged NV$^-$ state.
+
+Fig.
+\ref{NV electrons}-b) shows the predicted molecular orbitals of the NV centers, initially computed as linear combination of the $sp^3$ orbitals respecting the $C_{3v}$ symmetry of the defect \citep{loubser1978electron}.
+This model has since been confirmed by ab initio computation and experimental observations \citep{doherty2013nitrogen}.
+In particular ab initio computations have shown that the molecular orbitals of the NV center are mostly localized on the nearest neighbors of the vacancy \citep{gali2008ab}.
+Ab initio computation has also determined that the $a'_1$ lies in the valence band and is therefore often ignored.
+
+The electron occupations shown in Fig. \ref{NV electrons}-b) corresponds to the electronic ground state of the NV$^-$ charge state.
+
+In the remaining part of this manuscript, we will only consider the negatively charged NV$^-$ state, unless specified otherwise.
+We will therefore omit the ``$-$" and simply refer to the negatively charged state as ``NV".
+\subsection{Optical properties of the NV center}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1\textwidth]{Figures/NV_optical}
+\caption{Optical properties of the NV center.
+(a) Electron occupation of the molecular orbitals for the electronic ground state $^3A_2$ and the first excited state $^3E$.
+(b) Diagram of the main optical transition between the two spin triplet states and of the IR transition between the two singlet sates.
+The value of the transition corresponds to the ZPL and are taken from \citep{doherty2013nitrogen}.
+(c) Representation of the vibronic structure for the $^3A_2$ and $3^E$ states under the Franck-Condon approximation.
+The graph axes are the energy of the states E and configuration coordinate q.
+The green and red arrow represents non-resonant (Stokes or anti-Stokes) absorption and emission.
+(d) Emission and absorption spectrum of the NV center main transition at 300 K, from PL and PLE measurements respectively.
+Taken from Wikipedia}
+\label{NV optical}
+\end{figure}
+
+Thanks to the discrete electronic transitions inside the diamond bad gap, the NV center is fluorescent: it can absorb light from $450-630$ nm and re-emits light in the $630-800$ nm band.
+
+Fig.
+\ref{NV optical} explains the mechanisms of the NV center optical activity.
+Fig.
+\ref{NV optical}-a) illustrates the electronic ground state $^3A_2$ and first excited state $^3E$, where 1 electron from the $a_1$ molecular orbital is promoted to the $e$ level.
+Both of these electron occupations have two unpaired electrons, meaning that they can exist either either in a spin triplet state ($S=1$) or a spin singlet state ($S=0$) \footnote{The two singlet states $^1E$ and $^1A_1$ represented in Fig.
+\ref{NV optical}-b) are both of configuration $a_1^2e^2$ where the two $e$ electrons are either in a the same $e$ orbital, or in an antisymmetric superposition of the two $e$ orbitals \citep{doherty2011negatively}.}.
+
+Fig.
+\ref{NV optical}-b) shows the zero phonon line (ZPL) for the main transition of the triplet ($^3A_2$ and $^3E$) and singlet ($^1E$ and $^1A_1$) states.
+It should be noted that due to an inter system crossing between the $^1E$ and $^3A_2$ levels, excitation or absorption of the singlet transition is only possible with a conjoint excitation of the triplet states.
+By default, we refer to the NV center photoluminescence (PL) as the emission between the two triplet states.
+
+Fig.
+\ref{NV optical}-c) illustrates the vibronic structure of the $^3A_2$ and $^3E$ states under the Franck-Condon approximation \citep{gali2011time}.
+The emission and absorption of light associated with the triplet transition can be accompanied by the emission or absorption of an optical phonon.
+Because the lifetime of the optical phonons is very short, the absorption or emission of light almost always start from the fundamental state of the $^3A_2$ and $^3E$ level, which explains why the absorption of light can only be done for energies above the ZPL, and the emission for energies below the ZPL (Stokes process).
+
+Finally Fig.
+\ref{NV optical}-d) shows the absorption and emission spectrum of NV centers at 300 K.
+A weak zero-phonon-line is visible at 637 nm, with large phonon sidebands in emission and absorption.
+
+\medskip
+Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages: thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and close to 1 quantum efficiency \citep{kurtsiefer2000stable}, they are very bright.
+With proper nanophotonics engineering (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}.
+The NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}.
+There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}.
+This limits the potential of NV centers for quantum communication protocols that rely on a deterministic photon source \citep{bradac2019quantum}.
+
+\subsection{Spin properties of the NV center electronic ground state}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1\textwidth]{Figures/spin_fine_hyperfine}
+\caption{tata}
+\label{NV spin}
+\end{figure}
+
+For most intent and purposes the hyper-fine structure of the NV center can be ignored, either 1 raie or ensemble.
+
+"particularly long electron and nuclear spin coherence times compared to other solid-state defect center" (Quantum nanophotonics in diamond)
 \printbibliography

### 2022-09-09 python_clément/test programs/test.py

@@ -1753,4 +1753,4 @@ def plot_irregular_surface():
 	plt.show()
 
 
-print(2*1.54*sin(109.5/2*pi/180))
+print(np.exp(-4.57/(1+(2/3)*pi**2*294**2/(1614**2))))

### 2022-09-11 Manuscrit Thèse/chapter 1/Chapter 1.tex
h,kltrh

@@
 \includegraphics[width=1\textwidth]{Figures/spin_fine_hyperfine}
-\caption{tata}
+\caption{Representation of the fine and hyper-fine structure of the $^3A_2$ state for a $^{14}N$ nucleus and $^{12}C$ nuclei.}
 \label{NV spin}
@@
 
-For most intent and purposes the hyper-fine structure of the NV center can be ignored, either 1 raie or ensemble.
+We will only focus here on the spin structure of the electronic ground state $^3A_2$, as the lifetime of the other states is too short to make effective spin qubits.
+All the numerical values in this section were found in \citep{smeltzer2009robust, doherty2013nitrogen} and the references within.
 
-"particularly long electron and nuclear spin coherence times compared to other solid-state defect center" (Quantum nanophotonics in diamond)
+The spin structure in the ground state consists in an electronic spin $S=1$ coupled to the nuclear spin of the nearby nuclei.
+Carbon's most common isotope is $^{12}C$ at 98.9 \% natural abundance.
+$^{12}C$ has a a nuclear spin $I=0$ and therefore causes no hyper-fine splitting.
+Nitrogen's most common isotope however is $^{14}N$ at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does cause a hyper-fine splitting.
+We will only cover here the most likely configuration for an NV center which is a $^{14}N$ nucleus and $^{12}C$ nuclei as closest neighbors.
+
+The spin system considered is then an electronic spin $S=1$ coupled to a nuclear spin $I=1$ with a Hilbert space of dimension $(2S+1)(2I+1)=9$. The spin Hamiltonian of the system can be written:
+\begin{equation}
+\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S}\cdot \bar{\bar{A}}\cdot \mathbf{I},
+\end{equation}
+where $\mathcal{H}_e$ is the purely electronic spin Hamiltonian, $\mathcal{H}_n$ the purely nuclear spin Hamiltonian, $\mathbf{S}$ the electronic spin operator, $\mathbf{I}$ the nuclear spin operator and $\bar{\bar{A}}$ the hyper fine tensor.
+\subsubsection{Electronic spin Hamiltonian}
+The electronic spin $\mathcal{H}_e$ can be written:
+\begin{equation}
+\label{eq. spin elec}
+\mathcal{H}_e=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} + \mathcal{H}_{elec}(\mathbf{E})+ \mathcal{H}_{strain}(\mathbf{\bar{\bar{\varepsilon}}}).
+\end{equation}
+
+The term $D S_z^2$ causes the fine structure of the electronic spin.
+It originates from the dipole-dipole coupling between the two unpaired electrons of the defect.
+Due to the strong anisotropy of the molecular orbitals of the defect, the fine structure term imposes the quantization axis $\mathbf{z}$ (via the $S_z$ operator) where $\mathbf{z}$ is the direction of the N$-$V axis in the diamond.
+$D$ is called the zero field splitting (ZFS) factor and its value is $D\approx 2.87\ \rm GHz$ at 300 K.
+
+$\gamma_e \mathbf{S} \cdot \mathbf{B}$ is the Zeeman splitting of the spin, where $\gamma_e \approx 2.8\ \rm MHz/G$ is the gyromagnetic ratio of the electronic spin (almost equal to that of a free electron in the case of the NV center).
+
+$\mathcal{H}_{elec}$ and $\mathcal{H}_{strain}$ are the dependency of the Hamiltonian on the electric field and crystal strain and will be discussed in further details in chapter 4.
+
+\subsubsection{Nuclear spin Hamiltonian and hyperfine coupling}
+The nuclear spin $\mathcal{H}_n$ can be written:
+\begin{equation}
+\mathcal{H}_e=Q I_z^2 + \gamma_n \mathbf{I} \cdot \mathbf{B},
+\end{equation}
+
+where $Q=-4.945\ \rm MHz$ is the nucleus electric quadrupole moment and $\gamma_n=0.308\ \rm kHz/G$ is the nuclear gyromagnetic ratio.
+
+Finally the hyper-fine tensor can be written diagonally:
+\begin{equation}
+\bar{\bar{A}} = \begin{pmatrix}
+A_{xx} & 0 & 0 \\
+0 & A_{yy} & 0 \\
+0 & 0 & A_{zz}
+\end{pmatrix},
+\end{equation}
+where $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$.
+
+Fig. \ref{NV spin} illustrates the fine and hyper-fine structure of the NV center in zero external field.
+
+Because $\gamma_n \ll \gamma_e$, the hyper-fine structure of the NV center can simply be thought of as a constant energy offset.
+This explains why for most intent and purposes the hyper-fine structure is ignored, along the electric and strain dependencies, which leads to the simpler spin Hamiltonian:
+
+\begin{align}
+\mathcal{H}&=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} \\
+&=\begin{pmatrix}
+D-\gamma_e B_z & \gamma_e (B_x+iB_y) & 0 \\
+\gamma_e (B_x-iB_y) & 0 & \gamma_e (B_x+iB_y) \\
+0 & \gamma_e (B_x+iB_y) & D+\gamma_e B_z
+\end{pmatrix}
+\end{align}
+
+\medskip
+The NV center's spin usage as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}.
+Part of the reason is the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
+However the main advantage of the NV center spin compared to other spin qubit (quantum dots, superconducting) is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.
+
+\subsection{Spin and optical levels intersystem crossing}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{Figures/ISC}
+\caption{Inter system crossing of the NV center.
+Blue arrows symbolize decay channels to and from the $\ket{m_s=\pm 1}$ states and red arrows the decay channels to and from the $\ket{m_s=0}$ states.
+Green arrows symbolize the optical excitation induced by the laser.
+The times indicated correspond to the inverse of the decay rates found in \citep{gupta2016efficient}.}
+\label{ISC}
+\end{figure}
+
+The optical excitation and radiative decay between the $^3A_2$ and $^3E$ levels preserve the spin projection $m_s$ \citep{robledo2011spin}.
+There is however a non-radiative decay path which involves in inter-system crossing with the singlet states $^1A_1$ and $^1E$ which depends on the the triplet states spin projection.
+
+Fig.
+\ref{ISC} summarizes the possible decay channels from the $^3E$ states.
+We assume that the electrons are initially excited from the $^3A_2$ states via a green laser with a rate $\gamma_las$.
+The $^3E$ states can then either decay radiatevely and emit a 637$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states.
+We do consider the decay channel through the singlet state as non-radiative due to the strong non radiative channels competing with the IR radiative transition (quantum efficiency $\leq 10^{-3}$ \citep{rogers2008infrared, ma2010excited, acosta2010optical}).
+
+The rates $\gamma$ of these intersystem crossings, represented in Fig.
+\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depends on the spin projection $m_s$: the likeliness to undergo a non-radiative decay is $\sim 4$\footnote{$\frac{15+91}{15+11}\approx 4$} times more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
+Similarly, the singlet state is $\sim 2.4$\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
+
+These spin selective intersystem crossings mean that under constant green excitation an optical pumping scheme in the $\ket{m_s=0}$ spin states takes place.
+At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}, which is equivalent to a thermal polarization (for B=0) with a spin temperature of $\approx 65\ \rm mK$ \footnote{\begin{equation*}
+T=\frac{h D}{k_B \ln (\frac{1-P}{2P})} \approx 65\ \rm mK,
+\end{equation*}
+where $D$ is th ZFS factor described in eq.
+(\ref{eq.
+spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.} even for a crystal lattice temperature of 300 K.
+
+Another effect of the intersystem crossing is the ability optically readout the spin state of the NV center: since the $\ket{m_s=0}$ state is more likely to undergo a radiative decay than the $\ket{m_s=\pm 1}$ states, the expected photoluminescence (PL) is higher when starting from a $\ket{m_s=0}$ state than a $\ket{m_s=\pm 1}$ state.
+While the $\ket{+1}$ and $\ket{-1}$ cannot be a priori distinguished, a direct excitation of the $\ket{+1} \to \ket{0}$ or $\ket{-1} \to \ket{0}$ transition with a resonant microwave can ambiguously reveal the initial spin state.
+
+Check le papier de vincent, ça m'a l'air tout faux ce truc.
+It should be noted that the decay rates indicated in Fig.
+\ref{ISC} cannot be directly measured.
+\citep{gupta2016efficient} infers them from short-time evolution of the PL after the laser is turned on.
+\citep{tetienne2012magnetic} used a different approach where they looked at the photodynamics as a function of the transverse magnetic field and found significantly different decay rates (by up to a factor of $\sim$ 2).
+
 \printbibliography

### 2022-09-11 Pratique/AN.py

@@ -1,5 +1,5 @@
 import numpy as np
-from numpy import cos, sin, tan, pi, arccos, arcsin, sqrt, exp
+from numpy import cos, sin, tan, pi, arccos, arcsin, sqrt, exp, log
 import matplotlib.pyplot as plt
 
 kb=1.38E-23
@@ -13,7 +13,7 @@ m_carbon=12.01/Na #g
 #Autre odg : ppm=1e5/µm3
 #Autre odg : 5 ppm = 10.4 nm de distance moyenne
 
-print(1e5*(20e-3**3)*5)
+# print(1e5*(20e-3**3)*5)
 
 P=1e5
 T=300
@@ -22,6 +22,7 @@ eV=1.6E-19
 # Theta_NV=arccos(-1/3)
 # print(Theta_NV*180/pi)
 
+print(h*2870*1e6/(kb*log((2*0.8)/(1-0.8))))
 
 
 def pola_NV(T):

### 2022-09-12 Manuscrit Thèse/chapter 1/Chapter 1.tex
gnjre

@@
 \gamma_e (B_x-iB_y) & 0 & \gamma_e (B_x+iB_y) \\
-0 & \gamma_e (B_x+iB_y) & D+\gamma_e B_z
+0 & \gamma_e (B_x-iB_y) & D+\gamma_e B_z
 \end{pmatrix}
@@
 Part of the reason is the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
-However the main advantage of the NV center spin compared to other spin qubit (quantum dots, superconducting) is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.
+However the main advantage of the NV center spin compared to other spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.
 
@@
 The optical excitation and radiative decay between the $^3A_2$ and $^3E$ levels preserve the spin projection $m_s$ \citep{robledo2011spin}.
-There is however a non-radiative decay path which involves in inter-system crossing with the singlet states $^1A_1$ and $^1E$ which depends on the the triplet states spin projection.
+There is however a non-radiative decay path which involves in inter-system crossing with the singlet states $^1A_1$ and $^1E$ that depends on the the triplet states spin projection.
 
@@
 \ref{ISC} summarizes the possible decay channels from the $^3E$ states.
-We assume that the electrons are initially excited from the $^3A_2$ states via a green laser with a rate $\gamma_las$.
+We assume that the electrons are initially excited from the $^3A_2$ states via a green laser with a rate $\gamma_{\rm las}$.
 The $^3E$ states can then either decay radiatevely and emit a 637$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states.
-We do consider the decay channel through the singlet state as non-radiative due to the strong non radiative channels competing with the IR radiative transition (quantum efficiency $\leq 10^{-3}$ \citep{rogers2008infrared, ma2010excited, acosta2010optical}).
+We do consider the decay channel through the singlet state as non-radiative due to the strong non radiative channels competing with the IR radiative transition (quantum efficiency $\leq 10^{-3}$ for the singlet transition \citep{rogers2008infrared, ma2010excited, acosta2010optical}).
 
-The rates $\gamma$ of these intersystem crossings, represented in Fig.
-\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depends on the spin projection $m_s$: the likeliness to undergo a non-radiative decay is $\sim 4$\footnote{$\frac{15+91}{15+11}\approx 4$} times more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
-Similarly, the singlet state is $\sim 2.4$\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
+The rates $\gamma$ of the various decay processes, represented in Fig.
+\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depends on the spin projection $m_s$: the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$$\frac{15+91}{15+11}\approx 4$$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
+Similarly, the singlet state is $\sim 2.4$ times\footnote{$$\frac{492}{204} \approx 2.4$$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
 
-These spin selective intersystem crossings mean that under constant green excitation an optical pumping scheme in the $\ket{m_s=0}$ spin states takes place.
+These spin selective intersystem crossings mean that an optical pumping scheme in the $\ket{m_s=0}$ spin state takes place under constant green excitation.
 At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}, which is equivalent to a thermal polarization (for B=0) with a spin temperature of $\approx 65\ \rm mK$ \footnote{\begin{equation*}
@@
 (\ref{eq.
-spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.} even for a crystal lattice temperature of 300 K.
+spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.} while the diamond itself can remain at room temperature.
 
-Another effect of the intersystem crossing is the ability optically readout the spin state of the NV center: since the $\ket{m_s=0}$ state is more likely to undergo a radiative decay than the $\ket{m_s=\pm 1}$ states, the expected photoluminescence (PL) is higher when starting from a $\ket{m_s=0}$ state than a $\ket{m_s=\pm 1}$ state.
-While the $\ket{+1}$ and $\ket{-1}$ cannot be a priori distinguished, a direct excitation of the $\ket{+1} \to \ket{0}$ or $\ket{-1} \to \ket{0}$ transition with a resonant microwave can ambiguously reveal the initial spin state.
+Another effect of the intersystem crossing is the ability to optically readout the spin state of the NV center: since the $\ket{m_s=0}$ state is more likely to undergo a radiative decay than the $\ket{m_s=\pm 1}$ states, the expected photoluminescence (PL) is higher when starting from a $\ket{m_s=0}$ state than a $\ket{m_s=\pm 1}$ state.
+While the $\ket{+1}$ and $\ket{-1}$ states cannot be a priori distinguished, a direct excitation of the $\ket{+1} \to \ket{0}$ or $\ket{-1} \to \ket{0}$ transition with a resonant microwave can unambiguously reveal the initial spin state.
 
-Check le papier de vincent, ça m'a l'air tout faux ce truc.
 It should be noted that the decay rates indicated in Fig.
 \ref{ISC} cannot be directly measured.
-\citep{gupta2016efficient} infers them from short-time evolution of the PL after the laser is turned on.
-\citep{tetienne2012magnetic} used a different approach where they looked at the photodynamics as a function of the transverse magnetic field and found significantly different decay rates (by up to a factor of $\sim$ 2).
+They have to be inferred from the PL dynamics at short times after illumination.
+Different experiments and different modeling processes have thus lead to different decay rates \citep{duarte2021effect}.
+
+\section{Basic operations with NV centers}
+
+This section will cover
+
+\subsection{Experimental setup}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{Figures/shema_exp}
+\caption{Confocal microscopy setup detailed in main text}
+\label{setup}
+\end{figure}
+The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field.
+Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.
+
+Fig.
+\ref{setup} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
+Unless specified otherwise, this is the experimental setup that we used to perform all the measurements shown in this manuscript.
+
+The setup consists of a confocal microscope where the excitation from a green laser (Thorlabs CPS532, 532 nm, 5 mW) is focused on a sample containing NV centers with a microscope objective (NA $\sim$ 0.5).
+The NV PL is then collected by the same objective and isolated from the back-scattered laser light with a dichroic mirror (DM) and additional filters (532 nm notch filter and 645 nm longpass filter to filter out the NV$^0$ PL).
+The PL is collected on an avalanche photodiode (Thorlabs APD410A or Perkin Elmer SPCM-AQRH-15) whose signal is sent to the acquisition card (NI X-series).
+
+The laser beam goes through an acousto-optic modulator (AOM, AA opto-electronic) controlled externally to create arbitrary light pulses (resolution $\sim$ ns).
+The microwave field is generated by a Rhode \& Shwarz SMB 100A generator and emitted with a home-made loop antenna positioned at $<$ 1 mm of the sample.
+A microwave switch (Mini-circuits ZASWA-2-50DRA+) is used to create microwave pulses (resolution $\sim$ 10 ns).
+The control pulses for the AOM and the microwave switch are generated either by the NI X-series card (resolution $\sim$ 1 $\mu$s), or by a pulseblaster ESR-PRO (resolution $\sim$ 10 ns).
+The microwave field can be amplified with a Mini-circuits ZHL-5W-422+ amplifier to achieve higher Rabi frequencies.
+
+Finally the magnetic field is generated either with a permanent magnet (up to $\sim$ 200 mT) or with a homemade electromagnet (up to $\sim$ 30 mT) alimented via a low freqency amplifier (Leybold power function generator 522663).
+Both the permanent magnet and the electromagnet can be  attached to a rotation stage for precise alignment.
+\subsection{Rabi Oscillations}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/Rabi}
+\caption{Rabi oscillations. (a) pulse sequence used to measure Rabi oscillations. (b) Experimental data on sample CVD-pink (cf samples [REF]).}
+\label{Rabi}
+\end{figure}
+
+The general technique for experiments involving NV spins proceeds as follow: 
+\begin{itemize}
+\item First the spins are polarized. This is done with a green laser pulse ($1-100$ $\mu$s depending on the laser intensity) which polarizes the spins in the $\ket{m_s=0}$ state at $\sim$ 80 \%.
+\item  Then the spins are manipulated. This process often involves a microwave field resonant with one of the spin transitions.
+\item Finally the spins are read out.
+The simplest spin readout for NV centers (but not the only one) consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
+For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore.
+The optimal readout time is of the same order as the spin polarization time ($1-100$ $\mu$s).
+\end{itemize}
+
+Fig.
+\ref{Rabi}-a) shows the pulse sequence used to observe Rabi oscillations with NV centers.
+In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$.
+The microwave frequency or the magnetic field need to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
+
+Fig.
+\ref{Rabi}-b) shows the result of such an experiment on a spin ensemble.
+The PL signal oscillations correspond to the oscillations of the spin population between the bright $\ket{0}$ state and the dark $\ket{-1}$ state.
+For long microwave pulses ($>$ 1 $\mu$s typically), the spins reach a steady state with  $\sim$ 50 \% population in both the $\ket{0}$ and $\ket{-1}$ states.
+
+\subsection{Optically detected magnetic resonance}
+
+\subsubsection{intro (PL)}
+
+\subsubsection{8 classes}
+
+\subsubsection{hyper-fine (?)}
+
+\subsubsection{Lock-in}
+
+\section{Measuring the NV center spin properties}
+
+\subsection{$T_1$ measurement}
+
+\subsection{$T_2^*$ measurement: Ramsey interferometry}
+
+\subsection{$T_2$ measurement: Spin-echo and dynamical decoupling}
+
+\section{Dipole-dipole interaction between two electronic spins}
 

### 2022-09-13 Manuscrit Thèse/chapter 1/Chapter 1.tex
demain labo

@@
 \begin{align}
+\label{NV spin Hamiltonian basic}
 \mathcal{H}&=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} \\
@@
 \subsection{Optically detected magnetic resonance}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/ODMR_basique}
+\caption{ODMR protocol.
+(a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$.
+(b) ODMR spectrum with no magnetic field on sample ADM-15-4.
+\label{ODMR basique}
+\end{figure}
 
-\subsubsection{intro (PL)}
+Optically detected magnetic resonance (ODMR) is the most commonly employed magnetometry protocol with NV centers due to its simplicity and the fact that it does not require prior calibration.
 
-\subsubsection{8 classes}
+The protocol for continuous wave (CW) ODMR is presented in Fig.
+\ref{ODMR basique}.
+In this simple protocol, the microwave is always turned on while the frequency spans a predefined interval $[f_{\rm start}, f_{\rm end}]$.
+The PL is recorded as the microwave is being swept, with a typical acquisition time for each PL sample $> 1\ \rm ms$.
+The sweeping frequency must be slow enough so that the PL can reach a steady state with the microwave field for each sample acquisition.
 
-\subsubsection{hyper-fine (?)}
+As we saw in the last section, the steady state of the NV center spin with a resonant microwave field is a statistical mixture of 50 \% $\ket{0}$ and 50 \% $\ket{\pm 1}$ \footnote{We are assuming here that the Rabi frequency is greater than the laser polarizing rate, which is often the case in practice}.
+If the microwave field is not resonant with a spin transition however, the steady state is predominantly $\ket{0}$ due to the spin polarization caused by the laser.
+The NV PL will therefore be lowered when the scanning microwave field becomes resonant with one of the NV spin transitions.
 
-\subsubsection{Lock-in}
+Fig.
+\ref{ODMR basique}-b) shows an example of ODMR spectrum on an NV ensemble with no external magnetic field.
+The reason for the two bumps instead of a single one centered on $D=2870\ \rm MHz$ is the splitting of the $\ket{+1}$ and $\ket{-1}$ levels occasioned by the strain and local electric field, which will be further discussed in chapter 4.
+
+\subsubsection{ODMR with lock-in detection}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/ODMR_lockin}
+\caption{ODMR protocol with lock-in detection.
+(a) Simplified confocal microscope setup with a lock-in amplifier (LIA) modulating the mircowave and demodulating the PL.
+(b) Demodulated ODMR spectrum with no magnetic field on sample ADM-15-4.}
+\label{ODMR lockin}
+\end{figure}
+
+Because of the ODMR protocol delivers a continuous signal (the PL) as a function of a periodic excitation (the microwave field), it is well suited to amplitude or frequency modulation and lock-in detection.
+
+The principle of lock-in detection will not be covered here, the key idea is that lock-in detection allows to translate the signal spectrally in a region less prone to noise.
+In this case, the ODMR signal is a DC measurement and is therefore prone to slow freqency noises such as thermal or mechanical drifts and laser amplitude fluctuations.
+The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low freqency noises, but slow enough compared to the NV photodynamics.
+All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1 kHz$.
+
+Fig. \ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig. \ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
+
+\subsection{ODMR with NV ensemble: longitudinal and transverse magnetic field}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/ODMR_8_classes}
+\caption{(a) Geometry of the NV center and definition of $B_z$ and $B_\perp$.
+(b) Representation of the 4 possibles orientations (``classes") of the N$-$V axis, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell.
+(c) Simulation of the evolution of the transition frequencies for each class of NV center as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
+(d) ODMR spectrum on sample ADM-150-1 for a magnetic field of $\sim 90\ \rm G$.}
+\label{ODMR lockin}
+\end{figure}
+
+We mentioned previously that the NV spin Hamiltonian (eq.
+(\ref{NV spin Hamiltonian basic})) was highly anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
+The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp$.
+For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of ...
+is slightly abusive as ...} can be approximated to: %cf AN.py
+\begin{align}
+\nu_0&= 
+\end{align}
 
@@
 \subsection{$T_2^*$ measurement: Ramsey interferometry}
+Parler aussi d'ODMR et pourquoi pas un exemple de différence entre 2 samples clean et pas clean
 

### 2022-09-14 Manuscrit Thèse/chapter 1/Chapter 1.tex
vujyh

@@
 Diamond is also famous for being extremely rare in nature, even though its only constituent, carbon, is one of the most abundant element on the Earth surface.
-The reason is that diamond is not the most stable phase of carbon under atmospheric pressure and temperature.
-This honor goes to the graphite which is found abundantly in coal.
+The reason being that graphite, and not diamond, is the thermodynamically stable phase of carbon in ambient conditions.
 
@@
 \end{figure}
-Diamond in its intrinsic form is an insulator with a bandgap of of $\sim 5.47\ \rm eV$ and is transparent from far infrared to deep UV.
+Diamond in its intrinsic form is an insulator with a bandgap of $\sim 5.47\ \rm eV$ and is transparent from far infrared to deep UV.
 As any crystal however, diamond is prone to defects.
-We will concentrate here on 0D point-like defects and omit structural defects such as dislocations, even though those too can affect the optical properties of the diamond \citep{collins2000colour}.
+We will focus here on 0D point-like defects and omit structural defects such as dislocations, even though those too can affect the optical properties of the diamond \citep{collins2000colour}.
 
@@
 
-We will now focus on the creation process of the diamonds.
 Almost all diamonds used in optics or electronics experiments today are synthetic, due to the higher control of the growth process required for these applications.
 There are two main competing technologies for the growth of large ($\mu \rm{m} \sim \rm{mm}$) single crystal diamonds: High pressure high temperature growth (HPHT) and chemical vapor deposition growth (CVD).
-Both of these processes have their own specificities which we will detail below.
+Both of these processes have their own specificities which are detailed below.
 
@@
 \ref{HPHT and CVD}-a) illustrates the apparatus of HPHT synthesis: a carbon source (graphite) and a solvent metal composed of transition metals (Fe, NI, Co typically \citep{bundy1963direct}) are put under $5 \sim 6$ GPa pressure in a press and heated to temperatures $\geq 1300\ ^\circ$C.
-A diamond crystal seed is put in the metal catalyst and a temperature gradient insures that the carbon migrates from the graphite to the diamond.
+A diamond crystal seed is put in the metal melt and a temperature gradient insures that the carbon migrates from the graphite to the diamond.
 
@@
 
-CVD growth for diamond was first demonstrated convincingly in the 80s \citep{matsumoto1982growth, matsumoto1982vapor, kamo1983diamond}.
+Diamond growth by CVD also started in the 50s, but the first reasonable growth rates were only achieved in the 80s \citep{matsumoto1982growth, matsumoto1982vapor, kamo1983diamond}.
 Unlike HPHT, CVD operates in a regime where diamond is not the thermodynamically stable phase.
@@
 
-The main catalyst used is atomic hydrogen, which can both create the free radicals required to form the $C-C$ diamond bonds, and which etches sp$^2$ bonds (graphite) faster than it does sp$^3$ bonds (diamond) \citep{gracio2010diamond}.
-Hydrogen is by far the most abundant element in the gas phase: the precursor gases used for the growth are typically CH$_4$ and H$_2$ in proportions of $1-5\%$ and $95-99\%$ respectively \citep{achard2020chemical}.
+The main catalyst used is atomic hydrogen because it etches the $C-C$ sp$^2$ bonds (graphite) faster than it does sp$^3$ bonds (diamond) \citep{gracio2010diamond}.
+In modern CVD growth, hydrogen is by far the most abundant element in the gas phase: the precursor gases used for the growth are typically CH$_4$ and H$_2$ in proportions of $1-5\%$ and $95-99\%$ respectively \citep{achard2020chemical}.
 
@@
 
-For a good CVD reactor, the main sources of extrinsic impurities the diamond comes from the source gasses \citep{balmer2009chemical}.
+If the CVD reactor is well made, the main sources of extrinsic impurities in the diamond comes from the source gasses \citep{balmer2009chemical}.
 The commercial availability of source gases containing less than $\sim$ ppm impurities made possible the growth of  high purity type IIa CVD diamonds \citep{kasu2003high, twitchen2004high, tallaire2006characterisation}.
@@
 The most common method to move the vacancies is therefore to anneal the crystal at a temperature $800 \sim 1200 ^\circ$C for a few hours \citep{botsoa2011optimal}.
-NV centers anneal out at temperature of $1400 \sim 1500 ^\circ$C \citep{zaitsev2013optical, pinto2012diffusion} which explains why the vacancies tend to stick to the nitrogen atoms during the annealing phase.
+When the vacancy is next to a nitrogen atom, it will only become mobile again for temperatures of $1400 \sim 1500 ^\circ$C \citep{zaitsev2013optical, pinto2012diffusion} which explains why the annealing process favors NV creation.
 
@@
 
-Fig. \ref{NV spin} illustrates the fine and hyper-fine structure of the NV center in zero external field.
+Fig. \ref{NV spin} illustrates the fine and hyper-fine structure of the NV center without any external fields.
 
@@
 &=\begin{pmatrix}
-D-\gamma_e B_z & \gamma_e (B_x+iB_y) & 0 \\
-\gamma_e (B_x-iB_y) & 0 & \gamma_e (B_x+iB_y) \\
-0 & \gamma_e (B_x-iB_y) & D+\gamma_e B_z
+D-\gamma_e B_z & \frac{\gamma_e (B_x+iB_y)}{\sqrt{2}} & 0 \\
+\frac{\gamma_e (B_x-iB_y)}{\sqrt{2}} & 0 & \frac{\gamma_e (B_x+iB_y)}{\sqrt{2}} \\
+0 & \frac{\gamma_e (B_x-iB_y)}{\sqrt{2}} & D+\gamma_e B_z
 \end{pmatrix}
@@
 The rates $\gamma$ of the various decay processes, represented in Fig.
-\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depends on the spin projection $m_s$: the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$$\frac{15+91}{15+11}\approx 4$$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
-Similarly, the singlet state is $\sim 2.4$ times\footnote{$$\frac{492}{204} \approx 2.4$$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
+\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depends on the spin projection $m_s$: the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$\frac{15+91}{15+11}\approx 4$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
+Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
 
@@
 
-This section will cover
+This section covers the two elementary bricks of almost any NV experiments which are the Rabi oscillations and optically detected magnetic resonance (ODMR).
+ODMR is the technique most commonly used to find ...
 
@@
 (a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$.
-(b) ODMR spectrum with no magnetic field on sample ADM-15-4.
+(b) ODMR spectrum with no magnetic field on sample ADM-15-4.}
 \label{ODMR basique}
@@
 
-Because of the ODMR protocol delivers a continuous signal (the PL) as a function of a periodic excitation (the microwave field), it is well suited to amplitude or frequency modulation and lock-in detection.
+Because the ODMR protocol delivers a continuous signal (the PL) as a function of a periodic excitation (the microwave field), it is well suited to amplitude or frequency modulation and lock-in detection.
 
@@
 The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low freqency noises, but slow enough compared to the NV photodynamics.
+
+Fig.
+\ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig.
+\ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
 All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1 kHz$.
-
-Fig. \ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig. \ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
 
@@
 (d) ODMR spectrum on sample ADM-150-1 for a magnetic field of $\sim 90\ \rm G$.}
-\label{ODMR lockin}
+\label{ODMR 8 classes}
 \end{figure}
@@
 (\ref{NV spin Hamiltonian basic})) was highly anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
-The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp$.
-For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of ...
-is slightly abusive as ...} can be approximated to: %cf AN.py
+The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig.
+\ref{ODMR 8 classes}-a).
+For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$.
+The true eigenstates however remain close to $\ket{0}$, $\ket{\pm 1}$ as long as $\gamma_e B_z > \frac{(\gamma_e B_\perp)^2}{D}$.} can be approximated in a perturtbative approach to: %cf AN.py
 \begin{align}
-\nu_0&= 
+\nu_0&= -\frac{(\gamma_e B_\perp)^2}{D} \\
+\nu_{-1} &= D - \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D} \\
+\nu_{+1} &= D + \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}
 \end{align}
+
+Fig.
+\ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal.
+We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian.
+For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
+
+Fig.
+\ref{ODMR 8 classes}-c) shows a simulation of the evolution of the transition frequencies $\nu_{+1} - \nu_{0}$ and $\nu_{-1} - \nu_{0}$ for all 4 classes as a function of the magnetic field amplitude, with the magnetic field orientation fixed in an arbitrary direction.
+Fig.
+\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles.
+8 lines are visible corresponding to the two transitions of the 4 classes.
+Based on the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.
 
@@
 
+One of the most important characterization of qubits is their relaxation times.
+Indeed, the potential applications of most qubits in quantum information or quantum sensing protocols is ultimately limited by their relaxation rates \citep{de2021materials, degen2017quantum}.
+
+The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ corresponding to the relaxation time of the spin populations to their thermal equilibrium, and the transverse relaxation time $T_2$ (following the NMR nomenclature) corresponding to the relaxation of the spin state coherences (non diagonal elements of the density matrix).
+
+The transverse relaxation time, also referred to as coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$.
+The $T_2^*$ time includes the inhomogeneous dephasing processes such as the in homogeneity in space (when multiple qubits are probed) and time of the external fields.
+The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}.
+While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used.
+More careful nomenclatures will label the $T_2$ times $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials}.
+The relaxation times are ordered as follows:
+\begin{equation}
+T_2^* < T_2 < 2 T_1.
+\end{equation}
+
+The relaxation times of NV centers depend in large part on their environment, and can therefore vary by several order of magnitude depending on the sample used.
+The following paragraphs will cover the measurement processes of the relaxation times and the typical values for the samples used in this manuscript.
+
 \subsection{$T_1$ measurement}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/T1_basique}
+\caption{Basic $T_1$ measurement scheme.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink for low laser intensity.
+The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
+\label{T1 basique}
+\end{figure}
+
+$T_1$ is the simplest relaxation time to measure because the spin populations are directly correlated to the PL.
+Fig.
+\ref{T1 basique} shows the pulse sequence used and an experimental result of the $T_1$ sequence.
+The spin lifetime $T_1$ is one of the only spin property that can be measured without a resonant microwave.
+
+The sequence shown in Fig.\ref{T1 basique} can be improved by implanting common mode rejection technique described in Fig.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/T1_protocole_soustraction}
+\caption{Common mode rejection $T_1$ measurement scheme.
+(a) Pulse seqeunce used.
+(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity.
+(c) Subtraction $S_1-S_2$ versus dark time $\tau$.
+The experimental data is fitted with an exponential decay of time constant $\tau=4.1\ \rm ms$.}
+\label{T1 soustraction}
+\end{figure}
 

### 2022-09-14 python_clément/Plot Data.py

@@ -8,24 +8,30 @@ def choseFileAction(filename=False):
 	lineSelection.removeAll()
 	global startPath,lastFile
 	if not filename :
-		filename,filters=QFileDialog.getOpenFileName(directory=startPath,filter='*.csv')
+		filename,filters=QFileDialog.getOpenFileName(directory=startPath,filter="data file(*.csv *.txt)")
 	GUI.changeTitle('Plot Data:'+filename)
 	lastFile=filename
 	if filename :	
 		startPath=os.path.dirname(filename)
-		with open(filename,'r',encoding = "ISO-8859-1") as f:
-			global data
-			data = list(csv.reader(f, delimiter=' ', quotechar='|', quoting=csv.QUOTE_NONNUMERIC))
-		for i in range(len(data)//2):
-			lineSelection.addItem('Line %i'%(i+1))
-		lineSelection.setEnabled(True)
-		try :
-			assert lastIndex < len(lineSelection.dic) #Sinon il est trop malin et il plante pas
-			lineSelection.setIndex(lastIndex)
-			choseLineAction()
-		except :
-			lineSelection.setIndex(0)
-			choseLineAction()
+		if filename[-4:]=='.csv' :
+			with open(filename,'r',encoding = "ISO-8859-1") as f:
+				global data
+				data = list(csv.reader(f, delimiter=' ', quotechar='|', quoting=csv.QUOTE_NONNUMERIC))
+			for i in range(len(data)//2):
+				lineSelection.addItem('Line %i'%(i+1))
+			lineSelection.setEnabled(True)
+			try :
+				assert lastIndex < len(lineSelection.dic) #Sinon il est trop malin et il plante pas
+				lineSelection.setIndex(lastIndex)
+				choseLineAction()
+			except :
+				lineSelection.setIndex(0)
+				choseLineAction()
+		elif filename[-4:]=='.txt' :
+			x,y=extract_data(filename)
+			if len(x) != len(y) :
+				x,y=extract_data(fname,decimalPoint=',')
+			gra.updateLine(l1,x,y)
 
 def choseLineAction():
 	i=lineSelection.index()

### 2022-09-14 python_clément/analyse.py

@@ -1000,6 +1000,11 @@ def lissage(t,n):
 	newt=np.array([sum(t[i:i+n])/n for i in range(len(t)-n)])
 	return newt
 
+def average(t,n):
+	m=len(t)//n
+	newt=np.array([sum(t[n*i:n*(i+1)])/n for i in range(m)])
+	return newt
+
 def derivative(x,y):
 	dx=x[1]-x[0]
 	n=len(y)

### 2022-09-15 Manuscrit Thèse/chapter 1/Chapter 1.tex
hgutidoe

@@
 For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one.
-For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach 50\% of the total NV concentration \citep{grezes2015storage}
+For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach up to 50\% of the total NV concentration \citep{grezes2015storage}
 
@@
 In particular ab initio computations have shown that the molecular orbitals of the NV center are mostly localized on the nearest neighbors of the vacancy \citep{gali2008ab}.
-Ab initio computation has also determined that the $a'_1$ lies in the valence band and is therefore often ignored.
+Ab initio computation has also determined that the $a'_1$ level lies in the valence band and is thus often ignored.
 
@@
 
-Thanks to the discrete electronic transitions inside the diamond bad gap, the NV center is fluorescent: it can absorb light from $450-630$ nm and re-emits light in the $630-800$ nm band.
+Because the NV center has discrete electronic transitions inside the diamond bad gap, it fluoresces: it can absorb light from $450-650$ nm and re-emits light in the $600-800$ nm band.
 
@@
 The NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}.
-There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}.
+There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (low Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}.
 This limits the potential of NV centers for quantum communication protocols that rely on a deterministic photon source \citep{bradac2019quantum}.
@@
 
-Because $\gamma_n \ll \gamma_e$, the hyper-fine structure of the NV center can simply be thought of as a constant energy offset.
-This explains why for most intent and purposes the hyper-fine structure is ignored, along the electric and strain dependencies, which leads to the simpler spin Hamiltonian:
+For most intent and purposes, the hyper-fine structure of the NV center can simply be thought of as an energy offset, and is thus often ignored.
+The electric and strain dependencies of the electronic Hamiltonian are also often ignored as they are of second order compared to the magnetic field dependency.
+This leads to the simpler spin Hamiltonian:
 
@@
 
-\medskip
+\bigskip
 The NV center's spin usage as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}.
-Part of the reason is the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
-However the main advantage of the NV center spin compared to other spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.
+Part of the reason are the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
+However the main advantage of the NV center spin compared to other solid state spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.
 
@@
 We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian.
-For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
+For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and that each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
 
@@
 
-\section{Measuring the NV center spin properties}
+\section{Relaxation times of the NV center spin}
 
-One of the most important characterization of qubits is their relaxation times.
-Indeed, the potential applications of most qubits in quantum information or quantum sensing protocols is ultimately limited by their relaxation rates \citep{de2021materials, degen2017quantum}.
+This section focuses on the dynamics of the NV spin, and more specifically on the various relaxation times.
+Relaxation times are very important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing potential applications \citep{de2021materials, degen2017quantum}.
 
-The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ corresponding to the relaxation time of the spin populations to their thermal equilibrium, and the transverse relaxation time $T_2$ (following the NMR nomenclature) corresponding to the relaxation of the spin state coherences (non diagonal elements of the density matrix).
+The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ which corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix), and the transverse relaxation time $T_2$\footnote{We follow here the NMR nomenclature} which corresponds to the relaxation of the spin state coherences (non diagonal elements of the density matrix).
 
-The transverse relaxation time, also referred to as coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$.
-The $T_2^*$ time includes the inhomogeneous dephasing processes such as the in homogeneity in space (when multiple qubits are probed) and time of the external fields.
+The transverse relaxation time, also called coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$.
+The $T_2^*$ time includes inhomogeneous dephasing processes such as the inhomogeneity in space (when multiple qubits are probed) and time of the external fields.
 The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}.
 While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used.
-More careful nomenclatures will label the $T_2$ times $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials}.
+More careful nomenclatures will label the $T_2$ times $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance.
 The relaxation times are ordered as follows:
@@
 
-The relaxation times of NV centers depend in large part on their environment, and can therefore vary by several order of magnitude depending on the sample used.
+The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the sample used.
 The following paragraphs will cover the measurement processes of the relaxation times and the typical values for the samples used in this manuscript.
 
-\subsection{$T_1$ measurement}
+\subsection{NV spin $T_1$}
 
+\subsubsection{Measurement protocol}
 \begin{figure}[h!]
@@
 
-The sequence shown in Fig.\ref{T1 basique} can be improved by implanting common mode rejection technique described in Fig.
+The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig.
+\ref{T1 soustraction}-a).
+This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$.
+Then the sequence is repeated with an additional $\pi$-pulse on one the NV transitions right before the readout.
+This $\pi$-pulse will project the remaining $\ket{0}$ population difference into a $\ket{\pm 1}$ state.
+The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which reject every contribution to the PL outside of the spin relaxation.
 
@@
 
+Fig.
+\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample as for Fig.
+\ref{T1 basique}.
+The difference between the $S_1$ data in Fig.
+\ref{T1 soustraction}-b) and the the data in Fig.
+\ref{T1 basique}-b) comes from the laser intensity used.
+In Fig.
+\ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant.
+Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark which is most likely due to charge tunneling between impurities \citep{choi2017depolarization, giri2018coupled, giri2019selective}.
+On the other hand, the laser intensity used in Fig.
+\ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.
+
+Fig.
+\ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal.
+Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig.
+\ref{T1 basique}.
+
+The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
+The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.
+
+\subsubsection{Physical origin and typical values}
+
+We distinguish three different origin for the the $T_1$ relaxation process:
+\begin{itemize}
+\item \textbf{Spin-phonon interaction:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which results in a phonon-induced spin relaxation.
+The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}.
+The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxtion times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at for mili-K temperatures \citep{astner2018solid}.
+
+\item \textbf{Electric and magnetic field noises:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}.
+As a result, magnetic and electric field noise can reduce the spin $T_1$ if their spectrum overlap with the NV transitions.
+This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}.
+The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effect are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
+
+\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipolar interaction betwen two spins will be detailed more thoroughly in the following section.
+Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation.
+Relaxation induced by NV-NV dipole-dipole coupling will be the object of chapter 3 and 4 while NV-other spin coupling will be the object of chapter 2.
+The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration, but it can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
+\end{itemize}
+
+All the measurements shown in this manuscript were performed at room temperature.
+The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$.
+Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses.
+The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed.
+$T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.
+
 \subsection{$T_2^*$ measurement: Ramsey interferometry}
+Allez have fun
 Parler aussi d'ODMR et pourquoi pas un exemple de différence entre 2 samples clean et pas clean

### 2022-09-15 Manuscrit Thèse/chapter 4/Chapter 4.tex
hgutidoe

--- 2022-09-05 Manuscrit Thèse/chapter 4/Chapter 4.tex
@@
 To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
-This protocol however has an extremely low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic}.
+This protocol however has an very low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective}.
 

### 2022-09-16 Manuscrit Thèse/chapter 1/Chapter 1.tex
b,ktld

@@
 Relaxation induced by NV-NV dipole-dipole coupling will be the object of chapter 3 and 4 while NV-other spin coupling will be the object of chapter 2.
-The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration, but it can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
+The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration and can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
 \end{itemize}
@@
 
-\subsection{$T_2^*$ measurement: Ramsey interferometry}
-Allez have fun
-Parler aussi d'ODMR et pourquoi pas un exemple de différence entre 2 samples clean et pas clean
+\subsection{NV spin $T_2^*$}
+\subsubsection{Measurement protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/Ramsey}
+\caption{Ramsey interferometry protocol.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink.} %The orange line corresponds to an Gaussian fit with time constant $\tau=400 \pm 100\ \rm ns$.}
+\label{Ramsey}
+\end{figure}
 
-\subsection{$T_2$ measurement: Spin-echo and dynamical decoupling}
+The $T_2^*$ time is the characteristic time for the loss of quantum coherence of a freely evolving system.
+It is generally defined as the decay time of of a Ramsey interferometry measurement \citep{barry2020sensitivity}.
+
+Fig.
+\ref{Ramsey} shows the Ramsey pulse sequence and an  example obtained on a spin ensemble.
+The microwave frequency is detuned by $\sim 300\ \rm kHz$ compared to the spin Larmor frequency which causes oscillations in the signal.
+$T_2^*$ is the decay time of the envelope of the Ramsey fringes.
+Unfortunately, it is hard to correctly fit the envelope of signal with so few fringes, but increasing the potential number of fringes mean further detuning the microwave which decreases the Rabi frequency to the point where $\pi$-pulses are of comparable length with the decoherence time (with our current microwave setup).
+
+Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
+The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity and since the diamond subtrate generally contains less impurities than NV ensemble.
+The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
+
+An alternative measurement of $T_2^*$ can be performed by looking at the linewidth of an ODMR spectrum.
+Indeed, ideally the ODMR spectrum should be the Fourier transform of the Ramsey signal (also called free induction decay in NMR literature).
+In practice however, both methods suffer from experimental noises and imperfections, with Ramsey-type methods being more sensitive to microwave inhomogeneities \citep{barry2020sensitivity} while ODMR is prone to laser and microwave power broadening \citep{dreau2011avoiding}.
+
+The method chosen to extract $T_2^*$ values in this manuscript was ODMR with low microwave power ($\Omega_{\rm Rabi} \ll 1/T_2^*$) and low laser intensity (compared to the the NV optical saturation).
+
+\subsubsection{Physical origin}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/ODMR_1_classes}
+\caption{ODMR spectra on a single NV class.
+(a) Experimental data on sample CVD-pink, fitted with three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
+(b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
+\label{ODMR 1 classe}
+\end{figure}
+
+For most samples, the leading cause of inhomogeneous dephasing is the temporal and spatial variations of magnetic fields caused by the paramagnetic impurities in the crystal \citep{barry2020sensitivity}.
+The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2.
+We will simply mention here that for sample with high NV density ([NV] $\geq$ ppm), $T_2^*$ is mostly limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
+
+Fig.
+\ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see [REF]).
+The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
+We can clearly see a higher linewidth, corresponding to a smaller $T_2^* value$ on the sample with higher nitrogen concentration.
+In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is completely hidden by the inhomogeneous broadening on the HPHT sample.
+
+Fig.
+\ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula.
+While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}.
+For Fig.
+\ref{ODMR 1 classe}-a), it was indeed found that Lorentzians fitted the data better than Gaussians.
+For Fig.
+\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, although this could be in part due to the unresolved hyperfine structure.
+
+\subsection{NV spin $T_2$}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/Echo}
+\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
+\label{Echo}
+\end{figure}
+\subsubsection{Measurement protocol}
+As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process known as as the spin-echo or Hahn-echo sequence.
+
+Fig.
+\ref{Echo} shows the pulse sequence used and an example on the same sample than the Ramsey measurement.
+The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time.
+This additional $\pi$-pulse causes static or slowly varying dephasing originating from detuning between the Larmor frequency and the microwave frequency to be rephased \citep{hahn1950spin}.
+
+This protocol extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
+Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
+This value could be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
+
+\subsubsection{Physical origin}
+The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence.
+The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}.
+For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2019decoherence}.
 

### 2022-09-16 data/20220818/Sumi 2/traitement.py

@@ -1,0 +1,19 @@
+import sys
+sys.path.append('D:\\These Clément\\these\\python_clément')
+sys.path.append('/home/pellet-mary/these/python_clément')
+from analyse import *
+
+plt.figure(num=1,figsize=(4.5,3),dpi=80)
+plt.xticks(fontsize=16)
+plt.yticks(fontsize=16)
+plt.locator_params(axis='x', nbins=5)
+plt.locator_params(axis='y', nbins=5)
+
+x,y=extract_data('ESR 1 raie')
+
+plt.plot(x,y,'x')
+popt,yfit=gauss_fit(x,y)
+plt.plot(x,yfit,lw=2)#
+print(popt,popt[2]*1.18)
+# plt.legend()
+plt.show()

### 2022-09-16 python_clément/test programs/test.py

@@ -1753,4 +1753,4 @@ def plot_irregular_surface():
 	plt.show()
 
 
-print(492/204)
+print((sqrt(log(2))/(pi*3.2)))

### 2022-09-19 Manuscrit Thèse/chapter 1/Chapter 1.tex
huio

@@
 \includegraphics[width=0.6\textwidth]{Figures/phase_diagram_carbon_wiki}
-\caption{Pressure-temperature phase diagram of elemental carbon. From wikipedia, adapted from \citep{bundy1989pressure, bundy1996pressure}.}
+\caption{Pressure-temperature phase diagram of elemental carbon. Credits: Wikimedia commons, adapted from \citep{bundy1989pressure, bundy1996pressure}.}
 \label{carbon phase diagram}
@@
 The graph axes are the energy of the states E and configuration coordinate q.
-The green and red arrow represents non-resonant (Stokes or anti-Stokes) absorption and emission.
+The green and red arrow represents non-resonant (Stokes) absorption and emission.
 (d) Emission and absorption spectrum of the NV center main transition at 300 K, from PL and PLE measurements respectively.
-Taken from Wikipedia}
+Credits: Wikimedia commons}
 \label{NV optical}
@@
 
-\section{Relaxation times of the NV center spin}
+\section{Dynamics the NV center spin}
 
-This section focuses on the dynamics of the NV spin, and more specifically on the various relaxation times.
-Relaxation times are very important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing potential applications \citep{de2021materials, degen2017quantum}.
+This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times.
+Relaxation times are very important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
 
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.4\textwidth]{Figures/shema_dipole_dipole}
+\caption{Geometric representation of two interacting spins.
+($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$) represent Cartesian coordinate systems in the spin frames of reference, where $\mathbf{z}_i$ is the axis of quantization of spin $i$.}
+\label{dipole-dipole}
+\end{figure}
+
+This final introductory section will focus on the dipole-dipole interaction between two electronic spins.
+
+We will consider the general situation represented in Fig.
+\ref{dipole-dipole}: two spins separated by a distance $r$ with their own coordinate systems ($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$).
+The coordinate system is chosen so that each spin operator can be written:
+
+\begin{equation}
+\hat{\mathbf{S}}_i=\hat{S}_x \mathbf{x}_i + \hat{S}_y \mathbf{y}_i + \hat{S}_z \mathbf{z}_i
+\end{equation}
+
+where the spin operators $S_{\{x,y,z\} }$ have their usual definitions.
+In particular, this means that the spin axis of quantization of each spin is $z_i$, which classically corresponds to the magnetization direction of the dipole.
+This axis can be defined either by the magnetic field, or by crystal field anisotropies, which is often the case for solid state spins at low magnetic fields with $S\geq 1$  (including the NV center).
+
+\subsection{Magnetic dipole-dipole Hamiltonian}
+
+Classically, the dipole-dipole interaction between two magnetic moments $\mathbf{m}_1$ and $\mathbf{m}_2$ can be derived by computing the static magnetic field generated by one dipole at the position of the second dipole \cite[p.~188]{jackson1999classical}:
+
+\begin{align}
+U&=-\mathbf{m}_1 \cdot \mathbf{B}(\mathbf{r}) \nonumber \\
+&=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u}) - \mathbf{m}_1\cdot\mathbf{m}_2}{r^3}+\frac{8\pi}{3}\mathbf{m}_1\cdot\mathbf{m}_2\,\delta(\mathbf{r})\right], \label{eq.
+dipole classique}
+\end{align}
+
+where $\mathbf{r}=r \mathbf{u}$ is the relative position of the two dipoles (see Fig. \ref{dipole-dipole}).
+
+The quantum analog of the dipole-dipole interaction between spins simply replaces the magnetic moments $\mathbf{m}_i$ in eq.
+(\ref{eq.
+dipole classique}) with $\hbar \gamma_i \hat{\mathbf{S}}_i$, where $\gamma_i$ is the gryromagnetic ratio and $\hat{\mathbf{S}}_i$ the spin vector operator for spin $i$:
+
+\begin{equation}
+\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2\,\delta(\mathbf{r})\right].
+\end{equation}
+
+The $\delta(\mathbf{r})$ term represents the Fermi contact energy.
+This term plays an important role for hyperfine coupling when there is an overlap of the electron and nuclear wavefunctions, such as the hyperfine coupling between the NV$^-$ electron spin and $^{14}$N \cite{doherty2012theory} or  $^{13}$C \cite{smeltzer201113c} nuclear spins.
+
+The focus of this manuscript is on the interaction between electronic spins which are separated by several atomic sites (density impurities in the $\sim$ ppm range).
+We will therefore assume that there is negligible overlap between the electronic wavefunctions and neglect the Fermi contact term.
+This leads the simplified dipole-dipole Hamiltonian:
+
+\begin{equation}
+\mathcal{H}_{\rm dd}= \frac{J_0}{r^3}\left[ 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2] \right],
+\end{equation}
+
+where $J_0=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}$.
+
+For two electronic spins with $g$-factors $\approx -2$, 
+\begin{equation*}
+\frac{J_0}{h} \approx 51.8\ \rm{MHz}\cdot \rm{nm}^3.
+\end{equation*}
+
+\subsection{Decomposition of the dipole-dipole Hamiltonian}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{Figures/flip-flop_double-flip}
+\caption{Representation of the two energy transfer mechanisms: flip-flop and double flips} 
+\label{flip flop double flip}
+\end{figure}
+
+In the weak coupling approximation, the dipole-dipole Hamiltonian between two spins can be decomposed between diagonal terms which are responsible for energy shifts of the individual spin levels, and the off-diagonal terms which are responsible for energy transfers between the two spins.
+
+In what follows, we will assume that both spins have the same quantization axis in order to simplify the Hamiltonian decomposition. A more general case is detailed in [REF].
+
+Following the notation introduced in Fig.
+\ref{dipole-dipole}, we will then assume that $$(\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1) = (\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2) = (\mathbf{x},\mathbf{y},\mathbf{z}).$$ We will denote $\mathbf{u}=u_x \mathbf{x} + u_y \mathbf{y} + u_z \mathbf{z}$ the decomposition of the unitary vector $\mathbf{u}$ in this basis.
+The dipole-dipole Hamiltonian can then be written:
+\begin{equation}
+\mathcal{H}_{\rm dd}=\frac{J_0}{r^3} (3 u_x^-1)
+\end{equation}
 \printbibliography

### 2022-09-19 Manuscrit Thèse/chapter 1/Chapter 1.tex
ovoleur

@@
 \chapter{Introduction to NV centers and their common usage}
-The aim of this first chapter is to introduce the tools and concepts required for the next chapters. We will start
+The aim of this first chapter is to introduce the tools and concepts needed to understand the next chapters.
+
+This chapter will first cover the material aspects of NV centers by focusing on the creation process of synthetic diamonds and the required steps to create NV defects inside them.
+
+It will then cover the fundamentals of NV physics, including the NV center optical and spin properties, basic experiments with NV centers and the various relaxations times ($T_1$,$T_2$,$T_2^*$) which characterizes the NV center's spin.
+
+Finally it will introduce dipole-dipole coupling between two spins and the concept of cross-relaxation.
+
+The three subjects that are the NV center physics, the diamond material aspects and the dipole-dipole interaction are at the heart of the work carried out in this thesis.
 
@@
 \subsection{Spin and optical levels intersystem crossing}
+\label{sec ISC}
 \begin{figure}[h!]
@@
 
+We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
+\ref{sec ISC}.
+For strong magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
+
 \section{Dynamics the NV center spin}
@@
 \subsection{NV spin $T_1$}
-
+\label{sec T1 NV}
 \subsubsection{Measurement protocol}
@@
 Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation.
-Relaxation induced by NV-NV dipole-dipole coupling will be the object of chapter 3 and 4 while NV-other spin coupling will be the object of chapter 2.
 The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration and can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
@@
 Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
-The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity and since the diamond subtrate generally contains less impurities than NV ensemble.
+The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity and since the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
 The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
@@
 
-This final introductory section will focus on the dipole-dipole interaction between two electronic spins.
+This final introductory section will focus on the dipole-dipole interaction between two electronic spins, which is the main form of spin-spin interaction and plays a key part in the properties of dense spin ensemble.
 
@@
 
-where the spin operators $S_{\{x,y,z\} }$ have their usual definitions.
+where the spin operators $S_{\{x,y,z\} }$ have their usual definitions in term of Pauli matrices.
 In particular, this means that the spin axis of quantization of each spin is $z_i$, which classically corresponds to the magnetization direction of the dipole.
-This axis can be defined either by the magnetic field, or by crystal field anisotropies, which is often the case for solid state spins at low magnetic fields with $S\geq 1$  (including the NV center).
+This axis can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (which includes the NV center).
 
@@
 \centering
-\includegraphics[width=.8\textwidth]{Figures/flip-flop_double-flip}
+\includegraphics[width=.7\textwidth]{Figures/flip-flop_double-flip}
 \caption{Representation of the two energy transfer mechanisms: flip-flop and double flips} 
@@
 \begin{equation}
-\mathcal{H}_{\rm dd}=\frac{J_0}{r^3} (3 u_x^-1)
+\mathcal{H}_{\rm dd}=\frac{J_0}{r^3}\left[ (3 u_x^2-1) S_x^1 S_x^2 + (3 u_y^2-1) S_y^1 S_y^2 + (3 u_z^2-1) S_z^1 S_z^2 \right],
 \end{equation}
+where $S_{\{x,y,z\} }^i$ are the spin operators acting on spin $i$.
+The $S_z^1 S_z^2$ term is the only diagonal term and corresponds to energy shifts while the $S_x^1 S_x^2$ and $S_y^1 S_y^2$ terms are responsible for energy transfer.
+To make the transfer mechanism more apparent, we will rewrite $S_x^i$ as $\frac{1}{2}(S_+^i+S_-^i)$ and $S_y^i$ as $\frac{1}{2i}(S_+^i-S_-^i)$:
+\begin{align}
+\mathcal{H}_{\rm dd}&=\frac{J_0}{r^3}\bigg{[} \frac{1}{4}(3 u_x^2 + 3 u_y^2 -2)(S_+^1 S_-^2 + S_-^1 S_+^2) \nonumber \\
+&\quad + \frac{1}{4}(3 u_x^2 - 3 u_y^2)(S_+^1 S_+^2 + S_-^1 S_-^2) \nonumber \\
+&\quad +(3 u_z^2-1) S_z^1 S_z^2 \bigg{]} \\
+\medskip
+&= \Omega_{\rm ff} (S_+^1 S_-^2 + S_-^1 S_+^2) + \Omega_{\rm df} (S_+^1 S_+^2 + S_-^1 S_-^2) + \Omega_{\rm shift} S_z^1 S_z^2. \label{eq. ff df}
+\end{align}
+
+We have introduced in eq.
+(\ref{eq.
+ff df}) the three relevant matrix elements $\Omega_{\rm ff}$, $\Omega_{\rm df}$, and $\Omega_{\rm shift}$ which correspond respectively to flip-flops, double flips and energy shifts.
+Flip-flops and double flips (sometimes referred as flip-flips) are represented in Fig.
+\ref{flip flop double flip}.
+Flip flops correspond to energy transfer between two spins for which the sum of the spin angular momentum is conserved ($\Delta (m_s^1+m_s^2)=0$), while double flips correspond to energy transfer for which the spin angular momentum is not conserved ($\Delta (m_s^1+m_s^2)=\pm 2$)\footnote{In order to preserve the total angular momentum, double flips transfer angular moment to the crystal, which is a form of Einstein de Haas effect \citep{einstein1915experimental} }.
+In both cases however, the energy between the initial and final states needs to be preserved for the transfer to be efficient.
+
+It should be noted that on average, each matrix element is of the same order of magnitude: $\Omega_{\rm ff} \sim \Omega_{\rm df} \sim \Omega_{\rm shift} \sim \frac{J_0}{r^3}$.
+
+\subsection{Cross-relaxations with NV centers}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
+\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
+\label{CR_shema}
+\end{figure}
+
+Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
+This process can occur through either flip-flops or double flips, as long as the resonance condition between the two spins is met.
+
+On of the condition for cross-relaxation is to have a difference in polarization between the two spins species considered.
+The NV center is therefore a good candidate to observe cross-relaxations as it can be optically polarized while most other spin species are not.
+
+Fig.
+\ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin.
+When the two considered spin transitions become resonant, in this case by tuning the magnetic field, polarization from the NV center will be transferred to the second spin, making the NV center less polarized and the second spin more polarized than when they are out of resonance.
+
+For a single NV center coupled to spin 2, assuming that spin 2 always remains depolarized\footnote{This is the case in practice if the spin 2 bath is much bigger than the NV bath and effectively acts as a thermostat, or if the lifetime of spin 2 is much smaller than that of the NV center.} and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$ (weak coupling regime), the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by the cross-relaxation with spin 2 can be analytically computed \citep{hall2016detection}:
+\begin{equation}
+\delta \Gamma_1=\frac{\Omega^2}{\Gamma_2^*} \frac{(\Gamma_2^*)^2}{(\delta \nu)^2+(\Gamma_2^*)^2},
+\label{delta gamma 1}
+\end{equation}
+where $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, $\delta \nu$ is the detuning in energy (expressed in frequency) between the initial and final states and $\Omega=\Omega_{\rm ff}$ or $\Omega_{\rm df}$ is the dipole-dipole matrix element corresponding to the flip-flop or double-flip process considered.
+T
+
+Eq.
+(\ref{delta gamma 1}) is the basis for every cross-relaxation detection performed in this manuscript.
+Because the NV center lifetime can be measured optically, as detailed in Sec.
+\ref{sec T1 NV}, cross-relaxation between NV-centers and other spin species can be measured optically.
+Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be explained in the following chapters.
+
+\section{Conclusion}
+
+We have covered in this introductory chapter every concepts needed to understand the results presented in the next chapter.
+We have seen the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them, and we have introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.
+
+The focus of chapter 2 will be on the cross-relaxation between NV centers and other spin impurities in CVD-grown diamond, while the focus of chapter 3 and 4 will be on the cross-relaxation between NV centers and other NV centers and their potential application in magnetometry.
 \printbibliography

### 2022-09-20 Manuscrit Thèse/chapter 1/Chapter 1.tex
fatigyé putain

@@
 \begin{document}
-\chapter{Introduction to NV centers and their common usage}
-The aim of this first chapter is to introduce the tools and concepts needed to understand the next chapters.
+\chapter{Introduction to Nitrogen-Vacancy centers and their common usage}
+The aim of this first chapter is to introduce concepts related to Nitrogen-vacancy (NV) centers, to the diamond material, and to the dipole-dipole interaction which are necessary to explain the work presented in the later chapters.
 
-This chapter will first cover the material aspects of NV centers by focusing on the creation process of synthetic diamonds and the required steps to create NV defects inside them.
-
-It will then cover the fundamentals of NV physics, including the NV center optical and spin properties, basic experiments with NV centers and the various relaxations times ($T_1$,$T_2$,$T_2^*$) which characterizes the NV center's spin.
-
-Finally it will introduce dipole-dipole coupling between two spins and the concept of cross-relaxation.
-
-The three subjects that are the NV center physics, the diamond material aspects and the dipole-dipole interaction are at the heart of the work carried out in this thesis.
+In this introductory chapter, we will first cover the material aspects of NV centers by focusing on the creation process of synthetic diamonds and the required steps to create NV defects inside them.
+We will then cover the fundamentals of NV physics, including the NV center optical and spin properties, basic experiments with NV centers and the various relaxations times ($T_1$,$T_2$,$T_2^*$) which characterizes the NV center's spin.
+Finally we will present the dipole-dipole interaction between two spins and the concept of cross-relaxation.
 
@@
 b) Drawing representing a fluorescence microscopy image of a diamond surface.
-The red fluorescent spots symbolize NV centers.
+The red fluorescent spots symbolize NV centers photoluminescence.
 c) Zoom-in on the crystalline structure of an NV center.
-Block dots represent carbon atoms}
+Black balls represent carbon atoms}
 \label{diamond+NV}
 \end{figure}
-Refaire ce paragraphe te foutre le summary après la description de la figure. Et bien parler du NV- et du charge state
-We will start this chapter by discussing the physical aspect of the NV center, and of its host material, the diamond.
-We will cover some of the basic properties of diamond and of its most common defects before addressing the creation process of both diamonds and NV centers.
+The NV center is a particular defect in diamond.
+Before studying the NV centers properties, we must first discuss the diamond itself and answer these questions: How are diamond created, what kind of defects are there in diamond, and how to specifically create NV defects.
+The field of material science related to diamonds and NV centers is a rich and complex one, and this section will only introduce basic notions regarding these subjects.
 
 Fig.
-\ref{diamond+NV} illustrates the various scales to consider in this study: we start with a diamond which for lab purposes is generally no bigger than a few mm.
-If we zoom-in on the surface of some diamonds, we can find bright fluorescent spots corresponding to particular crystalline defects.
-Standard confocal microscopy is limited in its resolution to the $\sim \mu$m range, but if we could zoom-in even further we would see the structure of the defect, illustrated here with the crystalline structure of an NV center.
-\subsection{Diamond overview}
+\ref{diamond+NV} illustrates the various scales to consider in the study of NV centers.
+Starting with a bulk diamond, typically few mm wide, then zooming-in on the diamond surface with an optical microscope to find fluorescent spots corresponding to the photoluminescence of NV centers.
+Standard confocal microscopy is limited in its resolution to the $\sim \mu$m range, but if it was possible to zoom-in even further, we would see the atomic structure of a few \AA \ illustrated in the last panel.
+\subsection{Diamond material overview}
 \begin{figure}[h!]
@@
 Diamond is also famous for being extremely rare in nature, even though its only constituent, carbon, is one of the most abundant element on the Earth surface.
-The reason being that graphite, and not diamond, is the thermodynamically stable phase of carbon in ambient conditions.
+The reason being that graphite, and not diamond, is the thermodynamically stable phase of carbon under ambient conditions.
 
@@
 The higher pressure needed to reach the diamond-stable region can only be found naturally deep under the earth crust, typically between 150 and 240 km below sea level \citep{tappert2011diamonds}, which explain their rarity.
-It also explains why, despite many previous efforts, diamonds were synthetically produced for the first time only in 1953 \citep{barnard2000diamond}, more than 50 years after the first synthetic sapphires.
+It also explains why, despite many previous attempts, diamonds were only synthetically produced in 1953 \citep{barnard2000diamond}, more than 50 years after the first synthetic sapphires.
 To this day, artificial diamond synthesis remains an active field of research \citep{shenderova2019synthesis, achard2020chemical}.
@@
 More recently, diamonds have been the object of optics and electronics applications.
-For both of these fields however, the focus is no longer solely on the diamond but also on the defects found inside it.
+For both of these fields however, the focus is no longer solely on the diamond but also on the defects hosted inside it.
 
@@
 As any crystal however, diamond is prone to defects.
-We will focus here on 0D point-like defects and omit structural defects such as dislocations, even though those too can affect the optical properties of the diamond \citep{collins2000colour}.
+We will focus here on point-like (0D) defects and omit structural defects such as dislocations, even though those too can affect the optical properties of the diamond \citep{collins2000colour}.
 
@@
 These impurities result in unpaired electron or holes localized around the impurity site.
-This localization causes discrete energy levels for the holes or the electrons, some of which will lie inside the diamond bandgap and will impact the crystal optical and electrical properties.
+This localization causes discrete energy levels for the holes or the electrons, some of which lie inside the diamond bandgap and impact the crystal optical and electrical properties.
 
-Defects with electronic transitions in the optical range are called ``colored centers" as they are responsible for the coloring of the gems.  H, He, Li, B, N,
-O, Ne, P, Si, As, Ti, Cr, Ni, Co, Zn, Zr, Ag, W, Xe and Tl are known to form colored centers when introduced in diamond \citep{zaitsev2013optical, shenderova2019synthesis}, and each of these elements can form several defects.
+Defects with electronic transitions in the optical range are called ``colored centers" as they are responsible for the coloring of the gems.
+Over 20 elements are known to form colored centers when introduced in diamond \citep{zaitsev2013optical, shenderova2019synthesis}, and each of these elemental atoms can form several defects.
 Nitrogen alone can form more than 50 different colored centers \citep{dobrinets2016hpht, ashfold2020nitrogen}.
@@
 It is based on the N and B concentration, with a threshold of a few ppm (part per million) for each species as this was the smaller amount easily detectable through IR absorption.
-Type Ia diamonds contain clustered N defects such as the B-center which is the N$_4$V defect.
+Type Ia diamonds contain clustered N defects such as the B-center (N$_4$V defect).
 In contrast, type Ib diamond mostly contains C-center which are single substitutional N$_s$ defects, also referred to as P1 centers in the spin community.
-Type IIb diamond contains B impurities which give them a blueish-grey color, and type IIa diamond contains no detectable impurities via IR absorption \citep{ashfold2020nitrogen}.
+Type IIb diamond contains boron impurities which give them a blueish-grey color, and type IIa diamond contains no detectable impurities via IR absorption \citep{ashfold2020nitrogen}.
 
 Nitrogen-vacancy centers are a rare occurrence in both natural diamonds and untreated synthetic diamonds, but we will see later that the much more common N$_s$ defects can be converted in NV centers.
-When working with NV centers, the starting crystal is therefore often a type Ib if one wants to work with ensemble, or a type IIa if one wants to work with single NV centers.
+When working with NV centers, the starting crystal is therefore often a type Ib diamond if one wants to work with ensemble, or a type IIa diamond if one wants to work with single NV centers \footnote{In order to isolate single NV in a confocal optical spot, the concentration needs to be [NV] $<$ 1 ppb.}.
 
-The boron defects are mainly studied in the context of p-doped diamond semi-conductor: a concentration of [B] $\approx 1000\ \rm ppm$ is needed to achieve a significant overlap between the impurity sites, which turns the diamond into a semi-metal \citep{macpherson2015practical}.Other defects of interest, generally not present naturally but introduced voluntarily, include P and S defects to create n-doped diamond \citep{das2022diamond} (still an active field of research), as well as group IV-vacancy defects (SiV, GeV, SnV, PbV) for quantum optics application \citep{bradac2019quantum}.
+The boron defects are mainly studied in the context of p-doped diamond semi-conductor: a concentration of [B] $\approx 1000\ \rm ppm$ is needed to achieve a significant overlap between the impurity sites, which turns the diamond into a semi-metal \citep{macpherson2015practical}.
+Other defects of interest, generally not present naturally but introduced voluntarily, include P and S defects to create n-doped diamond \citep{das2022diamond} (still an active field of research), as well as group IV-vacancy defects (SiV, GeV, SnV, PbV) for quantum optics application \citep{bradac2019quantum}.
 
-The NV center however remains by far the most studied defect in diamond due to its unique spin properties which will be explained below.
+The NV center however remains by far the most studied defect in diamond due to its unique spin properties which will be detailed below.
 
@@
 HPHT was the first commercially viable process for synthetic diamonds, starting from the 1950s \citep{barnard2000diamond, bundy1955man}.
-The basic idea of HPHT synthesis is to reconstruct an environment where diamond is the thermodynamically stable phase of carbon, similarly to the growth process of natural diamond below the earth crust.
+The basic idea of HPHT synthesis is to reconstruct an environment where diamond is the thermodynamically stable phase of carbon, similarly to the growth process of natural diamond below the earth surface.
 
@@
 The first synthesis of HPHT diamonds were all of type Ib due to unwanted nitrogen pollution in the metal melt.
-The first HPHT type IIa diamonds were produced in the late 90s by adding other elements to the metal melt (Ti, Al, Zr) which would react preferentially with nitrogen and prevent its incorporation in the diamond \citep{burns1999growth, sumiya2002growth}.
+The first HPHT type IIa diamonds were produced in the late 90s by adding other elements to the metal bath (Ti, Al, Zr) which would react preferentially with nitrogen and prevent its incorporation in the diamond \citep{burns1999growth, sumiya2002growth}.
 
@@
 \medskip
-Both HPHT and CVD processes can produce diamonds of high quality for quantum optics or electronics application, each with their pros and cons.
+Both HPHT and CVD processes can produce diamonds of high quality for quantum optics or electronics application, each with their own pros and cons.
 HPHT offers better scalability for thick crystals ($\sim \rm mm$) and allows the incorporation of heavy atoms (Ge \citep{palyanov2015germanium} , Sn \citep{ekimov2019effect}) which can be hard to incorporate from a gas phase.
@@
 NV centers are a relatively rare impurity in untreated diamonds.
-In CVD diamond for instance, the proportion of nitrogen atoms forming an NV center in an untreated crystal is at best of $2-3 \%$ \citep{hartland2014study} and sometimes much less.
-While this ratio can sometimes be enough  to work with single NV centers, applications such as sensing that require dense ensemble need a higher conversion rate.
-We will discuss here the usual techniques used to improve the formation of NV centers.
+In CVD diamond for instance, the proportion of nitrogen atoms forming an NV center in an untreated crystal is at best of $2-3 \%$ \citep{hartland2014study} and oftentimes much lower.
+While this ratio can be enough  to work with single NV centers, applications such as sensing that require dense ensemble need a higher conversion rate.
+We will discuss here the usual techniques employed to improve the formation of NV centers.
 
@@
 Several high energy particle have been used to create vacancies, such as protons, electrons, ions, neutrons and even gamma rays \citep{davies1976optical, ashbaugh1988gemstone, kleinsasser2016high}.
-The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their ability to penetrate the crystal and to create vacancies uniformly.
+The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their relatively frequent availability (compared to neutrons or gamma rays) and their ability to deeply penetrate the crystal.
 
@@
 The final N$_s$ to NV conversion rate can be as high as $\sim 50 \%$ \citep{grezes2015storage, hartland2014study}, although this ratio tend to decrease for higher [N$_s$] concentrations.
-Most sample studied in this manuscript had a concentration $[\rm NV]= 1-10\ \rm ppm$
+Most sample studied in this manuscript had a concentration $[\rm NV]= 1-10\ \rm ppm$ (see [REF] samples)
 
@@
 Only the negatively charged NV$^-$ center shows the specific spin properties that will be detailed below, but the NV center can exist under three charge states in the diamond: NV$^-$, NV$^0$ and NV$^+$.
-The presence of NV$^+$ is only marginal in standard diamonds \citep{hauf2014addressing, pfender2017protecting} and we will focus here solely on the NV$^-$ and NV$^0$ charge state.
+The presence of NV$^+$ is usually marginal \citep{hauf2014addressing, pfender2017protecting} and we will focus here solely on the NV$^-$ and NV$^0$ charge state.
 
-The charge state of the NV centers depends not only on the growth process of the crystal, but also on the usage made of the sample.
+The charge state of the NV centers depends not only on the growth process of the crystal, but also on the experiment performed with the sample.
 In particular, illumination in the optical range can promote electrons in the conduction band or holes in the valence band and modify the charge state of the NV center.
@@
 The most common wavelength used with NV centers is 532 nm (green) due to the historical importance of Nd:YAG lasers.
-On the other hand, it is possible to semi-permanently ionize the NV$^-$ into NV$^0$ when using a red excitation, which has been used to create an optical memory \citep{dhomkar2016long}.
+On the other hand, it is possible to semi-permanently ionize the NV$^-$ into NV$^0$ when using a red excitation, which has been used to create optical memories \citep{dhomkar2016long}.
 
@@
 
-We will now focus on the magnetic and optical properties of the NV$^-$ center. This section will cover...
+This section focuses on the NV$^-$ center properties both as an optical defect (colored center) and as a spin defect, and compare them to other solid state systems.
+It also covers the interplay between the spin and the optical levels, a property almost unique to the NV$^-$ center which is the basis for optical polarization and readout of the NV$^-$ spin state.
 
@@
 (b) Diagram of the main optical transition between the two spin triplet states and of the IR transition between the two singlet sates.
-The value of the transition corresponds to the ZPL and are taken from \citep{doherty2013nitrogen}.
+The value of the transition corresponds to the zero phonon line (ZPL) and are taken from \citep{doherty2013nitrogen}.
 (c) Representation of the vibronic structure for the $^3A_2$ and $3^E$ states under the Franck-Condon approximation.
@@
 
-Because the NV center has discrete electronic transitions inside the diamond bad gap, it fluoresces: it can absorb light from $450-650$ nm and re-emits light in the $600-800$ nm band.
+Because the NV center has discrete electronic transitions inside the diamond band gap, it fluoresces: it can absorb light from $450-650$ nm and re-emits light in the $600-800$ nm band.
 
 Fig.
-\ref{NV optical} explains the mechanisms of the NV center optical activity.
+\ref{NV optical} explains the origin of the NV center optical properties.
 Fig.
@@
 \medskip
-Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages: thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and close to 1 quantum efficiency \citep{kurtsiefer2000stable}, they are very bright.
+Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages.
+First, they are very bright thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and quantum efficiency of $0.7-0.8$ \citep{schirhagl2014nitrogen}.
 With proper nanophotonics engineering (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}.
-The NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}.
+Secondly, the NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}.
+Finally, the diamond itself is resistant to heat and chemical attacks, and presents low toxicity for biologic systems \citep{fu2007characterization}.
 There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (low Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}.
-This limits the potential of NV centers for quantum communication protocols that rely on a deterministic photon source \citep{bradac2019quantum}.
+This limits the potential of NV centers for quantum photonic devices \citep{bradac2019quantum}.
 
@@
 Carbon's most common isotope is $^{12}C$ at 98.9 \% natural abundance.
-$^{12}C$ has a a nuclear spin $I=0$ and therefore causes no hyper-fine splitting.
-Nitrogen's most common isotope however is $^{14}N$ at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does cause a hyper-fine splitting.
+$^{12}C$ has a a nuclear spin $I=0$ and therefore does not contribute to the hyper-fine structure.
+Nitrogen's most common isotope however is $^{14}N$ at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does create a hyper-fine splitting.
 We will only cover here the most likely configuration for an NV center which is a $^{14}N$ nucleus and $^{12}C$ nuclei as closest neighbors.
@@
 
-The term $D S_z^2$ causes the fine structure of the electronic spin.
+The term $D S_z^2$ corresponds to the fine structure of the electronic spin.
 It originates from the dipole-dipole coupling between the two unpaired electrons of the defect.
@@
 where $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$.
+The $z$ direction in the electronic spin Hamiltonian, the nuclear spin Hamiltonian and the hyper-fine tensor awlays refer to the direction of the N$-$V axis.
 
@@
 The NV center's spin usage as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}.
-Part of the reason are the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
+Part of the reason for this wide usage are the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
 However the main advantage of the NV center spin compared to other solid state spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.
@@
 We assume that the electrons are initially excited from the $^3A_2$ states via a green laser with a rate $\gamma_{\rm las}$.
-The $^3E$ states can then either decay radiatevely and emit a 637$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states.
-We do consider the decay channel through the singlet state as non-radiative due to the strong non radiative channels competing with the IR radiative transition (quantum efficiency $\leq 10^{-3}$ for the singlet transition \citep{rogers2008infrared, ma2010excited, acosta2010optical}).
+The $^3E$ states can then either decay radiatevely and emit a 600$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states.
+We consider the decay channel through the singlet state as non-radiative due to the strong non radiative channels competing with the IR radiative transition (quantum efficiency $\leq 10^{-3}$ for the singlet transition \citep{rogers2008infrared, ma2010excited, acosta2010optical}).
 
@@
 
-This section covers the two elementary bricks of almost any NV experiments which are the Rabi oscillations and optically detected magnetic resonance (ODMR).
-ODMR is the technique most commonly used to find ...
+This section introduces two elementary operations with NV centers that are Rabi oscillations and optically detected magnetic resonance (ODMR).
+
+Rabi oscillations is the basis for any kind of coherent qubit manipulation while ODMR is used to find the transition frequencies of the NV centers, as well as being a basic magnetometry protocol.
+
+This section also covers the general purpose experimental setup used for these two operations and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
 
@@
 
-For two electronic spins with $g$-factors $\approx -2$, 
+For two electronic spins with $g$-factors $\approx -2$\footnote{Most spin defects in diamond, including the NV center have a g-factor $\approx -2$ similarly to free electrons.}, 
 \begin{equation*}
@@
 
-We have covered in this introductory chapter every concepts needed to understand the results presented in the next chapter.
+We have covered in this introductory chapter every concepts needed to understand the results presented in the next chapters.
 We have seen the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them, and we have introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.
 
-The focus of chapter 2 will be on the cross-relaxation between NV centers and other spin impurities in CVD-grown diamond, while the focus of chapter 3 and 4 will be on the cross-relaxation between NV centers and other NV centers and their potential application in magnetometry.
+The focus of chapter 2 will be on the cross-relaxation between NV centers and other spin impurities in CVD-grown diamond, while the focus of chapter 3 and 4 will be on the cross-relaxation between NV centers and other NV centers, as well as their potential application in magnetometry.
 \printbibliography

### 2022-09-21 Manuscrit Thèse/chapter 1/Chapter 1.tex
gnrje

@@
 \usepackage{xcolor} % Pour avoir des couleurs dans les équations
+
+\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
 
@@
 
-Another effect of the intersystem crossing is the ability to optically readout the spin state of the NV center: since the $\ket{m_s=0}$ state is more likely to undergo a radiative decay than the $\ket{m_s=\pm 1}$ states, the expected photoluminescence (PL) is higher when starting from a $\ket{m_s=0}$ state than a $\ket{m_s=\pm 1}$ state.
+Another effect of the intersystem crossing is the ability to optically readout the spin state of the NV center: since the $\ket{0}$ state is more likely to undergo a radiative decay than the $\ket{\pm 1}$ states, the expected photoluminescence (PL) is higher when starting from the $\ket{0}$ state than a $\ket{\pm 1}$ state.
 While the $\ket{+1}$ and $\ket{-1}$ states cannot be a priori distinguished, a direct excitation of the $\ket{+1} \to \ket{0}$ or $\ket{-1} \to \ket{0}$ transition with a resonant microwave can unambiguously reveal the initial spin state.
@@
 They have to be inferred from the PL dynamics at short times after illumination.
-Different experiments and different modeling processes have thus lead to different decay rates \citep{duarte2021effect}.
+Different experiments and different models have thus lead to different decay rates \citep{duarte2021effect}.
 
@@
 
-This section also covers the general purpose experimental setup used for these two operations and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
+This section also covers the experimental setup used for these two operations and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
 
@@
 \ref{setup} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
-Unless specified otherwise, this is the experimental setup that we used to perform all the measurements shown in this manuscript.
+Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.
 
@@
 \item Finally the spins are read out.
-The simplest spin readout for NV centers (but not the only one) consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
+The simplest spin readout for NV centers consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
 For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore.
@@
 In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$.
-The microwave frequency or the magnetic field need to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
+The microwave frequency or the magnetic field needs to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
 
@@
 \ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
-All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1 kHz$.
+All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1\ \rm kHz$.
 
@@
 (b) Representation of the 4 possibles orientations (``classes") of the N$-$V axis, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell.
-(c) Simulation of the evolution of the transition frequencies for each class of NV center as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
+(c) Simulation of the evolution of the transition frequencies for each class of NV centers as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
 (d) ODMR spectrum on sample ADM-150-1 for a magnetic field of $\sim 90\ \rm G$.}
@@
 We mentioned previously that the NV spin Hamiltonian (eq.
-(\ref{NV spin Hamiltonian basic})) was highly anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
+(\ref{NV spin Hamiltonian basic})) was anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
 The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig.
 \ref{ODMR 8 classes}-a).
+
 For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$.
@@
 \ref{sec ISC}.
-For strong magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
+For strong transverse magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
 
-\section{Dynamics the NV center spin}
+\section{Dynamics of the NV center spin}
 
 This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times.
-Relaxation times are very important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
+Relaxation times are important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
 
@@
 While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used.
-More careful nomenclatures will label the $T_2$ times $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance.
+More careful nomenclatures will label the $T_2$ times as $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance.
 The relaxation times are ordered as follows:
@@
 The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the sample used.
-The following paragraphs will cover the measurement processes of the relaxation times and the typical values for the samples used in this manuscript.
+The following paragraphs will cover the measurement processes of the different relaxation times and the typical values for the samples used in this manuscript.
 
@@
 
-$T_1$ is the simplest relaxation time to measure because the spin populations are directly correlated to the PL.
+$T_1$ is the simplest relaxation time to measure because the spin populations are directly correlated to the NV centers PL.
 Fig.
@@
 This $\pi$-pulse will project the remaining $\ket{0}$ population difference into a $\ket{\pm 1}$ state.
-The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which reject every contribution to the PL outside of the spin relaxation.
+The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which rejects every contribution to the PL outside of the spin relaxation.
+
+Fig.
+\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample that was used in Fig.
+\ref{T1 basique}.
+The difference between the $S_1$ data in Fig.
+\ref{T1 soustraction}-b) and the the data in Fig.
+\ref{T1 basique}-b) comes from the laser intensity used.
+In Fig.
+\ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant.
+Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark \citep{giri2018coupled, giri2019selective} (most likely due to charge tunneling between impurities \citep{choi2017depolarization}) .
+On the other hand, the laser intensity used in Fig.
+\ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.
+
+Fig.
+\ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal.
+Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig.
+\ref{T1 basique}.
 
@@
 
-Fig.
-\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample as for Fig.
-\ref{T1 basique}.
-The difference between the $S_1$ data in Fig.
-\ref{T1 soustraction}-b) and the the data in Fig.
-\ref{T1 basique}-b) comes from the laser intensity used.
-In Fig.
-\ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant.
-Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark which is most likely due to charge tunneling between impurities \citep{choi2017depolarization, giri2018coupled, giri2019selective}.
-On the other hand, the laser intensity used in Fig.
-\ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.
-
-Fig.
-\ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal.
-Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig.
-\ref{T1 basique}.
-
 The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
@@
 \begin{itemize}
-\item \textbf{Spin-phonon interaction:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which results in a phonon-induced spin relaxation.
+\item \textbf{Spin-phonon relaxation:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which can result in a phonon-induced spin relaxation.
 The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}.
@@
 This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}.
-The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effect are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
+The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effects are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
 
@@
 \ref{Ramsey} shows the Ramsey pulse sequence and an  example obtained on a spin ensemble.
-The microwave frequency is detuned by $\sim 300\ \rm kHz$ compared to the spin Larmor frequency which causes oscillations in the signal.
+The microwave frequency is detuned by $\sim 300\ \rm kHz$ compared to the spin central Larmor frequency which causes oscillations in the signal.
 $T_2^*$ is the decay time of the envelope of the Ramsey fringes.
-Unfortunately, it is hard to correctly fit the envelope of signal with so few fringes, but increasing the potential number of fringes mean further detuning the microwave which decreases the Rabi frequency to the point where $\pi$-pulses are of comparable length with the decoherence time (with our current microwave setup).
+Unfortunately, it is hard to correctly fit the envelope of a signal with so few fringes.
+Increasing the number of fringes with our current microwave and optical setup might prove hard, as further microwave detuning would decrease the signal contrast, and would be ultimately limited by the microwave pulses resolution.
 
 Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
-The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity and since the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
+The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity, and because the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
 The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
@@
 The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
-We can clearly see a higher linewidth, corresponding to a smaller $T_2^* value$ on the sample with higher nitrogen concentration.
-In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is completely hidden by the inhomogeneous broadening on the HPHT sample.
+We can clearly see a higher linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
+In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
 
@@
 \subsubsection{Measurement protocol}
-As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process known as as the spin-echo or Hahn-echo sequence.
+As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.
 
@@
 The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time.
-This additional $\pi$-pulse causes static or slowly varying dephasing originating from detuning between the Larmor frequency and the microwave frequency to be rephased \citep{hahn1950spin}.
+This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased \citep{hahn1950spin}.
 
@@
 Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
-This value could be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
+This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
 
@@
 
-where the spin operators $S_{\{x,y,z\} }$ have their usual definitions in term of Pauli matrices.
+where the spin operators $\hat S_{\{x,y,z\} }$ have their usual definitions in term of Pauli matrices.
 In particular, this means that the spin axis of quantization of each spin is $z_i$, which classically corresponds to the magnetization direction of the dipole.
-This axis can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (which includes the NV center).
+This axis can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (including the NV center).
 
@@
 U&=-\mathbf{m}_1 \cdot \mathbf{B}(\mathbf{r}) \nonumber \\
-&=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u}) - \mathbf{m}_1\cdot\mathbf{m}_2}{r^3}+\frac{8\pi}{3}\mathbf{m}_1\cdot\mathbf{m}_2\,\delta(\mathbf{r})\right], \label{eq.
+&=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u}) - \mathbf{m}_1\cdot\mathbf{m}_2}{r^3}+\frac{8\pi}{3}\mathbf{m}_1\cdot\mathbf{m}_2\,\delta(r)\right], \label{eq.
 dipole classique}
@@
 \begin{equation}
-\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2\,\delta(\mathbf{r})\right].
+\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2\,\delta(r)\right].
 \end{equation}
 
-The $\delta(\mathbf{r})$ term represents the Fermi contact energy.
-This term plays an important role for hyperfine coupling when there is an overlap of the electron and nuclear wavefunctions, such as the hyperfine coupling between the NV$^-$ electron spin and $^{14}$N \cite{doherty2012theory} or  $^{13}$C \cite{smeltzer201113c} nuclear spins.
+The $\delta(r)$ term represents the Fermi contact energy.
+This term plays an important role when there is an overlap of the spin particle wavefunctions, such as the hyperfine coupling between the NV$^-$ electron spin and $^{14}$N \cite{doherty2012theory} or  $^{13}$C \cite{smeltzer201113c} nuclear spins.
 
 The focus of this manuscript is on the interaction between electronic spins which are separated by several atomic sites (density impurities in the $\sim$ ppm range).
-We will therefore assume that there is negligible overlap between the electronic wavefunctions and neglect the Fermi contact term.
+We will therefore assume that there is negligible overlap between the electronic wavefunctions and omit the Fermi contact term.
 This leads the simplified dipole-dipole Hamiltonian:
@@
 &\quad + \frac{1}{4}(3 u_x^2 - 3 u_y^2)(S_+^1 S_+^2 + S_-^1 S_-^2) \nonumber \\
-&\quad +(3 u_z^2-1) S_z^1 S_z^2 \bigg{]} \\
+&\quad +(3 u_z^2-1) S_z^1 S_z^2 \bigg{]} \nonumber \\
 \medskip
@@
 where $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, $\delta \nu$ is the detuning in energy (expressed in frequency) between the initial and final states and $\Omega=\Omega_{\rm ff}$ or $\Omega_{\rm df}$ is the dipole-dipole matrix element corresponding to the flip-flop or double-flip process considered.
-T
 

### 2022-09-21 Manuscrit Thèse/chapter 2/Chapter 2.tex
gnrje

@@
 \begin{document}
-\chapter{Detection of dark spins via cross-relaxations}
-In this chapter, we will see how the interplay between the optical and spin properties of the NV centers can be transferred to other spins in the crystal by coupling them to NV centers.
-More specifically, I will discuss the first optical detection of two spin species, VH$^-$ and War1, in a CVD diamond by the mean of cross-relaxations with NV centers.
-\section{Flip-Flops, double flips and cross-relaxation}
-\label{Sec_CR}
+\chapter{Cross-relaxations between NV centers and dark spins in CVD-grown diamond}
+Crystal defects are promising candidates for quantum information technologies, either through their optical or spin degree of freedom \citep{aharonovich2016solid, atature2018material, bassett2019quantum}.
+Spectrocopic studies are a necessary step in identifying and characterizing these defects, with electron paramagnetic resonance (EPR) being the spectrocopic tool of choice to measure the spin properties of the defects \citep{newton2007epr}.
 
-\subsection{Resonant interaction : flip-flops and double flips}
+In this chapter, we present a method to characterize the spin of diamond impurities through cross-relaxation with NV centers \citep{hall2016detection,pellet2021optical}.
+This method allows an all-optical detection of non-NV center spin defects, a method that is both cheaper and more sensitive than traditional EPR.
+We focus in particular on defects in CVD-grown diamonds which have historically been less studied than HPHT diamond, and for which no prior NV cross-relaxation had been reported.
 
-The dominant form of interaction between two distant spins is the magnetic dipole-dipole interaction.
-For two spins with with vector spin Hamiltonian $\hat{\mathbf{S}}_1$ and $\hat{\mathbf{S}}_2$, separated by a distance $\mathbf{r}=r\mathbf{u}$, the dipole-dipole interaction Hamiltonian reads (see Appendix A) :
-\begin{equation}
-\mathcal{H}_{dd}=\frac{J_0}{r^3}\left[ \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2 - 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u})\right],
-\end{equation}
-where $J_0= \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$.
-For two electronic spins with $g-$factors close to 2 (which is the case for the NV center and most spin defect in diamond), the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$.
+The chapter is organized as follow: we will first detail the most common paramagnetic impurities in the diamond (the so-called ``dark spins") before explaining the proceeding of NV cross-relaxation experiments, and more generally of NV relaxometry.
+We will then focus on the detection of two particular defects in a CVD diamond, VH$^-$ and War1, and on their characterization (density, spin parameters, temperature dependence).
+We will finally conclude on the possible improvements and application of this detection method.
 
-Flip-flop, in the context of dipole-dipole interaction, is the mechanism which exchanges a quantum of spin between two spins.
-For an initial state of two spins $\ket{m_s^1=i;m_s^2=j}$, a flip-flop process will result in a final state $\ket{i+1;j-1}$ or $\ket{i-1;j+1}$.
-The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$, as well as the resonance condition between the initial and final state.
-We should note that for two identical spins, the flip-flop process $\ket{i;i\pm 1}\bra{i\pm 1,i}$ is always resonant.
+The work presented in this chapter has already been published in large part in the articles \citep{pellet2021optical, ngambou2022improving}.
 
-Another mechanism authorized by the dipole-dipole Hamiltonian is the double-flip, either up or down, which couple the states $\ket{i;j}$ and $\ket{i-1;j-1}$ or $\ket{i+1;j+1}$.
-For a single spin species in a strong magnetic field, the lift of the various spin levels by the Zeeman effect means that no double-flip process can be resonant.
-However with weak magnetic field or when two different spin species are present, double flip process can be as important as flip-flops since the matrix elements $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \pm 1}$ are typically of the same order of magnitude as the flip-flop ones.
+\section{Dark spins in diamond}
 
-\subsection{Cross-relaxations}
+We refer by dark spins to paramagnetic impurities for which, unlike the NV center, the spin degree of freedom is not accessible optically.
+At room temperature, only two defects in diamond do not have a dark spin: the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
+Every other spin impurity studied so far can be considered to be dark.
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
-\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
-\label{CR_shema}
-\end{figure}
-
-Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
-This process con occur either through flip-flops, or through double flips, as long as the resonance condition between the two spins is met.
-
-Fig.
-\ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin.
-When the two considered spin transitions become resonant, in this case by tuning the magnetic field, polarization from the NV center will be transferred to the second spin, making the NV center less polarized and the second spin more polarized than when they are out of resonance.
-
-The reason why CR tend to depolarize the NV center simply comes from a rate equation : since the initial NV population is higher in the $\ket{0}$ state than in the $\ket{-1}$ state, the flip-flop (or double flip) process that makes the NV center go to $\ket{-1}$ (symbolized by the two red arrows in Fig.
-\ref{CR_shema}) is more likely than the reverse flip-flop.
-If we only focus on the NV center, cross-relaxation is simply another thermalization process due to the coupling of the NV center to its environment.
-
-For a single NV center coupled to spin 2, assuming that spin 2 is initially depolarized and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$, we can compute the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by spin 2 as \citep{hall2016detection}:
-\begin{equation}
-\delta \Gamma_1=\frac{\Omega^2}{\Gamma_2^*} \frac{(\Gamma_2^*)^2}{(\delta \nu)^2+(\Gamma_2^*)^2},
-\label{delta gamma 1}
-\end{equation}
-where $\Omega=\frac{\mel{i;j}{\mathcal{H}_{\rm dd}}{j;i}}{\hbar}$ is the matrix element of the dipole-dipole Hamiltonian that couples the initial and final states, $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, and $\delta \nu$ is the detuning between the central frequencies of the NV center and spin 2.
-\section{The dark spins in diamond}
-
-We refer by dark spins to spin defects which do not interact with light, which in the case of diamond defects consist of almost every known spin defect bar the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
+In this section we will present the most common dark spins encountered in diamond, which are simply the most common paramagnetic impurities aside from the NV center, and the relation between the diamond growth process and the presence of these impurities.
+We will also briefly present EPR spectroscopy which is the most common method to characterize these dark spins.
 
 \subsection{P1 centers and $^{13}$C}
-By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
+The most common spin defects in NV-rich diamond are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which represents 1.1\% of natural abundance carbon atoms, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
 
 In NV-rich diamonds, these two impurities are the main causes of the inhomogeneous dephasing rate $\Gamma_2^*=1/T_2^*$ of the NV centers ensemble \citep{barry2020sensitivity}.
-The dephasing comes from the distribution of dipole-dipole coupling from each NV centers to its neighboring spin defects, which varies in space and time.
 

### 2022-09-21 Manuscrit Thèse/these_fusion.tex
gnrje

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric-comp,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{bib_thesis.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme

\title{}
\begin{document}
\begin{refsection}

\chapter{Introduction to Nitrogen-Vacancy centers and their common usage}
The aim of this first chapter is to introduce concepts related to Nitrogen-vacancy (NV) centers, to the diamond material, and to the dipole-dipole interaction which are necessary to explain the work presented in the later chapters.

In this introductory chapter, we will first cover the material aspects of NV centers by focusing on the creation process of synthetic diamonds and the required steps to create NV defects inside them. We will then cover the fundamentals of NV physics, including the NV center optical and spin properties, basic experiments with NV centers and the various relaxations times ($T_1$,$T_2$,$T_2^*$) which characterizes the NV center's spin. Finally we will present the dipole-dipole interaction between two spins and the concept of cross-relaxation.

\section{The NV center as a physical object}
\begin{figure}[h!]
\centering
\includegraphics[width=0.9\textwidth]{chapter 1/Figures/diamant_et_NV}
\caption{From the diamond to the NV center. a) Drawing of a bulk diamond. We consider a diamond to be ``bulk" if it is bigger than a few 100 $\mu$m. b) Drawing representing a fluorescence microscopy image of a diamond surface. The red fluorescent spots symbolize NV centers photoluminescence. c) Zoom-in on the crystalline structure of an NV center. Black balls represent carbon atoms}
\label{diamond+NV}
\end{figure}
The NV center is a particular defect in diamond. Before studying the NV centers properties, we must first discuss the diamond itself and answer these questions: How are diamond created, what kind of defects are there in diamond, and how to specifically create NV defects. The field of material science related to diamonds and NV centers is a rich and complex one, and this section will only introduce basic notions regarding these subjects.

Fig. \ref{diamond+NV} illustrates the various scales to consider in the study of NV centers. Starting with a bulk diamond, typically few mm wide, then zooming-in on the diamond surface with an optical microscope to find fluorescent spots corresponding to the photoluminescence of NV centers. Standard confocal microscopy is limited in its resolution to the $\sim \mu$m range, but if it was possible to zoom-in even further, we would see the atomic structure of a few \AA \ illustrated in the last panel. 
\subsection{Diamond material overview}
\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{chapter 1/Figures/phase_diagram_carbon_wiki}
\caption{Pressure-temperature phase diagram of elemental carbon. Credits: Wikimedia commons, adapted from \citep{bundy1989pressure, bundy1996pressure}.}
\label{carbon phase diagram}
\end{figure}

Diamond is a remarkable material in many aspects, it is famous for having the highest hardness and thermal conductivity of any natural material, and for its high optical refractive index and optical dispersion which makes it an extremely valuable crystal in jewelry.

Diamond is also famous for being extremely rare in nature, even though its only constituent, carbon, is one of the most abundant element on the Earth surface. The reason being that graphite, and not diamond, is the thermodynamically stable phase of carbon under ambient conditions.

Fig. \ref{carbon phase diagram} shows the PT phase diagram of elemental carbon. We can see that diamond becomes the thermodynamically stable phase only for pressure above a few GPa, $10^4$ times higher than atmospheric pressure. Even though diamond is unstable under ambient conditions, the extremely high energy activation needed to break the sp$^3$ bond between two carbon atoms means that the kinematics of the $C_{\rm diamond} \to C_{\rm graphite}$ transition is almost completely frozen. In many ways diamonds are more stable than graphite under ambient condition due to their relative chemical inertness.

Where the thermodynamic equilibrium really matters is for the crystallization process. The higher pressure needed to reach the diamond-stable region can only be found naturally deep under the earth crust, typically between 150 and 240 km below sea level \citep{tappert2011diamonds}, which explain their rarity. It also explains why, despite many previous attempts, diamonds were only synthetically produced in 1953 \citep{barnard2000diamond}, more than 50 years after the first synthetic sapphires. To this day, artificial diamond synthesis remains an active field of research \citep{shenderova2019synthesis, achard2020chemical}.

Diamonds, either natural or artificial, are used in the lab and in industry for their extreme sturdiness, with applications ranging from abrasive to diamond anvil cells \citep{jayaraman1983diamond}. More recently, diamonds have been the object of optics and electronics applications. For both of these fields however, the focus is no longer solely on the diamond but also on the defects hosted inside it.

\subsection{Point defects in diamond}
\begin{figure}[h!]
\centering
\includegraphics[width=0.9\textwidth]{chapter 1/Figures/Diamond_type}
\caption{Traditional classification of diamonds, based on \citep{tappert2011diamonds}}
\label{diamond type}
\end{figure}
Diamond in its intrinsic form is an insulator with a bandgap of $\sim 5.47\ \rm eV$ and is transparent from far infrared to deep UV. As any crystal however, diamond is prone to defects. We will focus here on point-like (0D) defects and omit structural defects such as dislocations, even though those too can affect the optical properties of the diamond \citep{collins2000colour}.

Point-like defects can be constituted by the absence of a carbon atom (``vacancies"), an interstitial defect between two carbon atoms or the substitution of a carbon atom with another atom. These impurities result in unpaired electron or holes localized around the impurity site. This localization causes discrete energy levels for the holes or the electrons, some of which lie inside the diamond bandgap and impact the crystal optical and electrical properties.

Defects with electronic transitions in the optical range are called ``colored centers" as they are responsible for the coloring of the gems.  Over 20 elements are known to form colored centers when introduced in diamond \citep{zaitsev2013optical, shenderova2019synthesis}, and each of these elemental atoms can form several defects. Nitrogen alone can form more than 50 different colored centers \citep{dobrinets2016hpht, ashfold2020nitrogen}.

Nitrogen, and to a lesser extant boron, are the most commonly found extrinsic elements in natural diamond. The traditional classification of diamond is presented in Fig. \ref{diamond type}. It is based on the N and B concentration, with a threshold of a few ppm (part per million) for each species as this was the smaller amount easily detectable through IR absorption. Type Ia diamonds contain clustered N defects such as the B-center (N$_4$V defect). In contrast, type Ib diamond mostly contains C-center which are single substitutional N$_s$ defects, also referred to as P1 centers in the spin community. Type IIb diamond contains boron impurities which give them a blueish-grey color, and type IIa diamond contains no detectable impurities via IR absorption \citep{ashfold2020nitrogen}.

Nitrogen-vacancy centers are a rare occurrence in both natural diamonds and untreated synthetic diamonds, but we will see later that the much more common N$_s$ defects can be converted in NV centers. When working with NV centers, the starting crystal is therefore often a type Ib diamond if one wants to work with ensemble, or a type IIa diamond if one wants to work with single NV centers \footnote{In order to isolate single NV in a confocal optical spot, the concentration needs to be [NV] $<$ 1 ppb.}.

The boron defects are mainly studied in the context of p-doped diamond semi-conductor: a concentration of [B] $\approx 1000\ \rm ppm$ is needed to achieve a significant overlap between the impurity sites, which turns the diamond into a semi-metal \citep{macpherson2015practical}. Other defects of interest, generally not present naturally but introduced voluntarily, include P and S defects to create n-doped diamond \citep{das2022diamond} (still an active field of research), as well as group IV-vacancy defects (SiV, GeV, SnV, PbV) for quantum optics application \citep{bradac2019quantum}.

The NV center however remains by far the most studied defect in diamond due to its unique spin properties which will be detailed below.

\subsection{Creation of synthetic diamond for optics and electronics applications}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/HPHT_and_CVD}
\caption{Schematics of HPHT (a) and CVD (b) growth process as detailed in main text}
\label{HPHT and CVD}
\end{figure}

Almost all diamonds used in optics or electronics experiments today are synthetic, due to the higher control of the growth process required for these applications. There are two main competing technologies for the growth of large ($\mu \rm{m} \sim \rm{mm}$) single crystal diamonds: High pressure high temperature growth (HPHT) and chemical vapor deposition growth (CVD). Both of these processes have their own specificities which are detailed below.

\subsubsection{HPHT growth}

HPHT was the first commercially viable process for synthetic diamonds, starting from the 1950s \citep{barnard2000diamond, bundy1955man}. The basic idea of HPHT synthesis is to reconstruct an environment where diamond is the thermodynamically stable phase of carbon, similarly to the growth process of natural diamond below the earth surface.

Fig. \ref{HPHT and CVD}-a) illustrates the apparatus of HPHT synthesis: a carbon source (graphite) and a solvent metal composed of transition metals (Fe, NI, Co typically \citep{bundy1963direct}) are put under $5 \sim 6$ GPa pressure in a press and heated to temperatures $\geq 1300\ ^\circ$C. A diamond crystal seed is put in the metal melt and a temperature gradient insures that the carbon migrates from the graphite to the diamond.

The first synthesis of HPHT diamonds were all of type Ib due to unwanted nitrogen pollution in the metal melt. The first HPHT type IIa diamonds were produced in the late 90s by adding other elements to the metal bath (Ti, Al, Zr) which would react preferentially with nitrogen and prevent its incorporation in the diamond \citep{burns1999growth, sumiya2002growth}.

\subsubsection{CVD growth}

Diamond growth by CVD also started in the 50s, but the first reasonable growth rates were only achieved in the 80s \citep{matsumoto1982growth, matsumoto1982vapor, kamo1983diamond}. Unlike HPHT, CVD operates in a regime where diamond is not the thermodynamically stable phase. It relies therefore heavily on catalysts to stabilize the diamond phase versus the graphite one.

The main catalyst used is atomic hydrogen because it etches the $C-C$ sp$^2$ bonds (graphite) faster than it does sp$^3$ bonds (diamond) \citep{gracio2010diamond}. In modern CVD growth, hydrogen is by far the most abundant element in the gas phase: the precursor gases used for the growth are typically CH$_4$ and H$_2$ in proportions of $1-5\%$ and $95-99\%$ respectively \citep{achard2020chemical}.

Fig. \ref{HPHT and CVD}-b) illustrates the apparatus of a microwave plasma reactor: it consists in a modified vacuum chamber which acts as a microwave resonator. The microwave creates a plasma ball from the the precursor gases with a temperature $T_{\rm plasma} \geq 3000 ^\circ$C \citep{ashfold2020nitrogen}, enough to dissociate the $H_2$ gas in atomic hydrogen \citep{balmer2009chemical}. The plasma acts as the carbon source for the diamond growing on top of the substrate, and as the catalyst to favor the diamond phase. The substrate is kept at a temperature $700 \sim 1100\ ^\circ$C during the growth.

Other sources can be used to create the high temperature required for atomic hydrogen, such as a tungsten hot filament \citep{haubner1993diamond}, electrical discharge (arcjet) \citep{luque1998excited} or even an oxygen-acetylene combustion flame \citep{bachmann1991towards}.

If the CVD reactor is well made, the main sources of extrinsic impurities in the diamond comes from the source gasses \citep{balmer2009chemical}. The commercial availability of source gases containing less than $\sim$ ppm impurities made possible the growth of  high purity type IIa CVD diamonds \citep{kasu2003high, twitchen2004high, tallaire2006characterisation}.

\medskip
Both HPHT and CVD processes can produce diamonds of high quality for quantum optics or electronics application, each with their own pros and cons. HPHT offers better scalability for thick crystals ($\sim \rm mm$) and allows the incorporation of heavy atoms (Ge \citep{palyanov2015germanium} , Sn \citep{ekimov2019effect}) which can be hard to incorporate from a gas phase. CVD on the other hand allows more versatility in the control of the impurities introduced in the crystal, including an isotopic control on both the impurities and the carbon itself. A particular advantage of CVD growth is the possibility of $\delta$-doping where a specific defect is grown only within a thin layer of the crystal, as low as 5 nm \citep{ohno2012engineering, ishikawa2012optical, ohashi2013negatively}.

\subsection{Optimizing NV center formation}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/formtion_NV_2}
\caption{Illustration of the formation of NV centers. Black dots represents carbon atoms and green dots nitrogen atoms}
\label{formation NV}
\end{figure}

NV centers are a relatively rare impurity in untreated diamonds. In CVD diamond for instance, the proportion of nitrogen atoms forming an NV center in an untreated crystal is at best of $2-3 \%$ \citep{hartland2014study} and oftentimes much lower. While this ratio can be enough  to work with single NV centers, applications such as sensing that require dense ensemble need a higher conversion rate. We will discuss here the usual techniques employed to improve the formation of NV centers.

Fig. \ref{formation NV} illustrates the steps needed to maximize the number of NV centers in the crystal while keeping a relatively high crystalline purity.

The first step to create NV centers is to get individual substitutional N atoms in the crystal ($N_s$, C-centers or P1 centers depending on the literature). This can be achieved either \textit{in situ} by introducing (voluntarily or not) nitrogen in the crystal during the growth process \citep{tallaire2006characterisation, lobaev2017influence}, or \textit{ex situ} by implanting N$^+$ ions with ion implantation techniques \citep{meijer2005generation, smith2019colour}. Because this present work focuses on dense NV ensembles, we will consider the simplest case where the starting crystal is a type-Ib diamond due to the presence of nitrogen during the growth. A starting concentration [N$_s$]= $10 \sim 300$ ppm is typical for type Ib HPHT diamonds \citep{achard2020chemical}.

The second step is to create monovacancies in the crystal lattice. This step might be optional for crystals which already contain a large number of monovacancies, but CVD crystals in particular can have less vacancies that they have substitutional nitrogen atoms \citep{mainwood1999point}. The general technique to create monovacancy is to bombard the sample with high energy particle in order to knock out carbon atoms from the crystal. Several high energy particle have been used to create vacancies, such as protons, electrons, ions, neutrons and even gamma rays \citep{davies1976optical, ashbaugh1988gemstone, kleinsasser2016high}. The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their relatively frequent availability (compared to neutrons or gamma rays) and their ability to deeply penetrate the crystal.

The final step is to migrate the vacancies next to the nitrogen atoms. The activation energy required to move vacancy is of $\sim 2.3\ \rm eV$ which corresponds to temperatures of $600\sim 700 ^\circ$C \citep{davies1992vacancy, newton2002recombination}. The most common method to move the vacancies is therefore to anneal the crystal at a temperature $800 \sim 1200 ^\circ$C for a few hours \citep{botsoa2011optimal}. When the vacancy is next to a nitrogen atom, it will only become mobile again for temperatures of $1400 \sim 1500 ^\circ$C \citep{zaitsev2013optical, pinto2012diffusion} which explains why the annealing process favors NV creation.

The final N$_s$ to NV conversion rate can be as high as $\sim 50 \%$ \citep{grezes2015storage, hartland2014study}, although this ratio tend to decrease for higher [N$_s$] concentrations. Most sample studied in this manuscript had a concentration $[\rm NV]= 1-10\ \rm ppm$ (see [REF] samples)

\subsection{Controlling the NV center charge state}

The charge state of the NV center is of crucial importance for NV applications. Only the negatively charged NV$^-$ center shows the specific spin properties that will be detailed below, but the NV center can exist under three charge states in the diamond: NV$^-$, NV$^0$ and NV$^+$. The presence of NV$^+$ is usually marginal \citep{hauf2014addressing, pfender2017protecting} and we will focus here solely on the NV$^-$ and NV$^0$ charge state.

The charge state of the NV centers depends not only on the growth process of the crystal, but also on the experiment performed with the sample. In particular, illumination in the optical range can promote electrons in the conduction band or holes in the valence band and modify the charge state of the NV center.

On the material side, the crystal has to remain electrically neutral which means that at least as many charge donors need to be present as there are NV$^-$ centers. For nitrogen-rich samples, the donors are mostly N$^+$ centers, which imposes a limit on the maximum conversion rate of N to NV in order to keep a high [NV$^-$]/[NV$^0$] ratio. The possibility to introduce other donors, such as phosphorus, has been studied \citep{doi2016pure} but remains impractical so far \citep{barry2020sensitivity}.

On the experimental side, it has been shown that the ionization of NV$^-$ to NV$^0$ and the reverse process depend on the intensity and wavelength of the laser used for illumination \citep{aslam2013photo}. It was found that the optimal wavelength to preserve the NV$^-$ charge state was  $510-540\ \rm nm$. The most common wavelength used with NV centers is 532 nm (green) due to the historical importance of Nd:YAG lasers. On the other hand, it is possible to semi-permanently ionize the NV$^-$ into NV$^0$ when using a red excitation, which has been used to create optical memories \citep{dhomkar2016long}.

For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one. For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach up to 50\% of the total NV concentration \citep{grezes2015storage}

\section{The NV$^-$ center optical and spin properties}

This section focuses on the NV$^-$ center properties both as an optical defect (colored center) and as a spin defect, and compare them to other solid state systems. It also covers the interplay between the spin and the optical levels, a property almost unique to the NV$^-$ center which is the basis for optical polarization and readout of the NV$^-$ spin state.

\subsection{NV center molecular orbitals}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{chapter 1/Figures/NV_electrons}
\caption{(a) Representation of the nitrogen and carbon orbitals surrounding the vacancy of an NV center. Each black dot represents an electron. A sixth electron is required to form the negatively charged NV$^-$ state. (b) Molecular orbitals of the NV center predicted from the the $C_{3v}$ symmetry of the defect. The electron occupation of the NV$^-$ ground state is represented by red arrows.}
\label{NV electrons}
\end{figure}

We will first cover the molecular aspect of the NV defect in diamond and the energy levels of the defect predicted from group theory and ab initio computations.

Fig. \ref{NV electrons}-a) illustrates the $sp^3$ orbitals of the carbons and nitrogen atoms neighboring the vacancy, as well as the 5 electrons occupying dangling bonds from these atoms and a sixth electron captured to form the negatively charged NV$^-$ state.

Fig. \ref{NV electrons}-b) shows the predicted molecular orbitals of the NV centers, initially computed as linear combination of the $sp^3$ orbitals respecting the $C_{3v}$ symmetry of the defect \citep{loubser1978electron}. This model has since been confirmed by ab initio computation and experimental observations \citep{doherty2013nitrogen}. In particular ab initio computations have shown that the molecular orbitals of the NV center are mostly localized on the nearest neighbors of the vacancy \citep{gali2008ab}. Ab initio computation has also determined that the $a'_1$ level lies in the valence band and is thus often ignored.

The electron occupations shown in Fig. \ref{NV electrons}-b) corresponds to the electronic ground state of the NV$^-$ charge state.

In the remaining part of this manuscript, we will only consider the negatively charged NV$^-$ state, unless specified otherwise. We will therefore omit the ``$-$" and simply refer to the negatively charged state as ``NV".
\subsection{Optical properties of the NV center}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{chapter 1/Figures/NV_optical}
\caption{Optical properties of the NV center. (a) Electron occupation of the molecular orbitals for the electronic ground state $^3A_2$ and the first excited state $^3E$. (b) Diagram of the main optical transition between the two spin triplet states and of the IR transition between the two singlet sates. The value of the transition corresponds to the zero phonon line (ZPL) and are taken from \citep{doherty2013nitrogen}. (c) Representation of the vibronic structure for the $^3A_2$ and $3^E$ states under the Franck-Condon approximation. The graph axes are the energy of the states E and configuration coordinate q. The green and red arrow represents non-resonant (Stokes) absorption and emission. (d) Emission and absorption spectrum of the NV center main transition at 300 K, from PL and PLE measurements respectively. Credits: Wikimedia commons}
\label{NV optical}
\end{figure}

Because the NV center has discrete electronic transitions inside the diamond band gap, it fluoresces: it can absorb light from $450-650$ nm and re-emits light in the $600-800$ nm band.

Fig. \ref{NV optical} explains the origin of the NV center optical properties. Fig. \ref{NV optical}-a) illustrates the electronic ground state $^3A_2$ and first excited state $^3E$, where 1 electron from the $a_1$ molecular orbital is promoted to the $e$ level. Both of these electron occupations have two unpaired electrons, meaning that they can exist either either in a spin triplet state ($S=1$) or a spin singlet state ($S=0$) \footnote{The two singlet states $^1E$ and $^1A_1$ represented in Fig. \ref{NV optical}-b) are both of configuration $a_1^2e^2$ where the two $e$ electrons are either in a the same $e$ orbital, or in an antisymmetric superposition of the two $e$ orbitals \citep{doherty2011negatively}.}.

Fig. \ref{NV optical}-b) shows the zero phonon line (ZPL) for the main transition of the triplet ($^3A_2$ and $^3E$) and singlet ($^1E$ and $^1A_1$) states. It should be noted that due to an inter system crossing between the $^1E$ and $^3A_2$ levels, excitation or absorption of the singlet transition is only possible with a conjoint excitation of the triplet states. By default, we refer to the NV center photoluminescence (PL) as the emission between the two triplet states.

Fig. \ref{NV optical}-c) illustrates the vibronic structure of the $^3A_2$ and $^3E$ states under the Franck-Condon approximation \citep{gali2011time}. The emission and absorption of light associated with the triplet transition can be accompanied by the emission or absorption of an optical phonon. Because the lifetime of the optical phonons is very short, the absorption or emission of light almost always start from the fundamental state of the $^3A_2$ and $^3E$ level, which explains why the absorption of light can only be done for energies above the ZPL, and the emission for energies below the ZPL (Stokes process).

Finally Fig. \ref{NV optical}-d) shows the absorption and emission spectrum of NV centers at 300 K. A weak zero-phonon-line is visible at 637 nm, with large phonon sidebands in emission and absorption.

\medskip
Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages. First, they are very bright thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and quantum efficiency of $0.7-0.8$ \citep{schirhagl2014nitrogen}. With proper nanophotonics engineering (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}.  Secondly, the NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}. Finally, the diamond itself is resistant to heat and chemical attacks, and presents low toxicity for biologic systems \citep{fu2007characterization}. There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (low Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}. This limits the potential of NV centers for quantum photonic devices \citep{bradac2019quantum}.

\subsection{Spin properties of the NV center electronic ground state}
\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{chapter 1/Figures/spin_fine_hyperfine}
\caption{Representation of the fine and hyper-fine structure of the $^3A_2$ state for a $^{14}N$ nucleus and $^{12}C$ nuclei.}
\label{NV spin}
\end{figure}

We will only focus here on the spin structure of the electronic ground state $^3A_2$, as the lifetime of the other states is too short to make effective spin qubits. All the numerical values in this section were found in \citep{smeltzer2009robust, doherty2013nitrogen} and the references within.

The spin structure in the ground state consists in an electronic spin $S=1$ coupled to the nuclear spin of the nearby nuclei. Carbon's most common isotope is $^{12}C$ at 98.9 \% natural abundance. $^{12}C$ has a a nuclear spin $I=0$ and therefore does not contribute to the hyper-fine structure. Nitrogen's most common isotope however is $^{14}N$ at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does create a hyper-fine splitting. We will only cover here the most likely configuration for an NV center which is a $^{14}N$ nucleus and $^{12}C$ nuclei as closest neighbors.

The spin system considered is then an electronic spin $S=1$ coupled to a nuclear spin $I=1$ with a Hilbert space of dimension $(2S+1)(2I+1)=9$. The spin Hamiltonian of the system can be written:
\begin{equation}
\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S}\cdot \bar{\bar{A}}\cdot \mathbf{I},
\end{equation}
where $\mathcal{H}_e$ is the purely electronic spin Hamiltonian, $\mathcal{H}_n$ the purely nuclear spin Hamiltonian, $\mathbf{S}$ the electronic spin operator, $\mathbf{I}$ the nuclear spin operator and $\bar{\bar{A}}$ the hyper fine tensor.
\subsubsection{Electronic spin Hamiltonian}
The electronic spin $\mathcal{H}_e$ can be written:
\begin{equation}
\label{eq. spin elec}
\mathcal{H}_e=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} + \mathcal{H}_{elec}(\mathbf{E})+ \mathcal{H}_{strain}(\mathbf{\bar{\bar{\varepsilon}}}).
\end{equation}

The term $D S_z^2$ corresponds to the fine structure of the electronic spin. It originates from the dipole-dipole coupling between the two unpaired electrons of the defect. Due to the strong anisotropy of the molecular orbitals of the defect, the fine structure term imposes the quantization axis $\mathbf{z}$ (via the $S_z$ operator) where $\mathbf{z}$ is the direction of the N$-$V axis in the diamond. $D$ is called the zero field splitting (ZFS) factor and its value is $D\approx 2.87\ \rm GHz$ at 300 K.

$\gamma_e \mathbf{S} \cdot \mathbf{B}$ is the Zeeman splitting of the spin, where $\gamma_e \approx 2.8\ \rm MHz/G$ is the gyromagnetic ratio of the electronic spin (almost equal to that of a free electron in the case of the NV center).

$\mathcal{H}_{elec}$ and $\mathcal{H}_{strain}$ are the dependency of the Hamiltonian on the electric field and crystal strain and will be discussed in further details in chapter 4.

\subsubsection{Nuclear spin Hamiltonian and hyperfine coupling}
The nuclear spin $\mathcal{H}_n$ can be written:
\begin{equation}
\mathcal{H}_e=Q I_z^2 + \gamma_n \mathbf{I} \cdot \mathbf{B},
\end{equation}

where $Q=-4.945\ \rm MHz$ is the nucleus electric quadrupole moment and $\gamma_n=0.308\ \rm kHz/G$ is the nuclear gyromagnetic ratio.

Finally the hyper-fine tensor can be written diagonally:
\begin{equation}
\bar{\bar{A}} = \begin{pmatrix}
A_{xx} & 0 & 0 \\
0 & A_{yy} & 0 \\
0 & 0 & A_{zz}
\end{pmatrix},
\end{equation}
where $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$. The $z$ direction in the electronic spin Hamiltonian, the nuclear spin Hamiltonian and the hyper-fine tensor awlays refer to the direction of the N$-$V axis.

Fig. \ref{NV spin} illustrates the fine and hyper-fine structure of the NV center without any external fields.

For most intent and purposes, the hyper-fine structure of the NV center can simply be thought of as an energy offset, and is thus often ignored. The electric and strain dependencies of the electronic Hamiltonian are also often ignored as they are of second order compared to the magnetic field dependency. This leads to the simpler spin Hamiltonian:

\begin{align}
\label{NV spin Hamiltonian basic}
\mathcal{H}&=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} \\
&=\begin{pmatrix}
D-\gamma_e B_z & \frac{\gamma_e (B_x+iB_y)}{\sqrt{2}} & 0 \\
\frac{\gamma_e (B_x-iB_y)}{\sqrt{2}} & 0 & \frac{\gamma_e (B_x+iB_y)}{\sqrt{2}} \\
0 & \frac{\gamma_e (B_x-iB_y)}{\sqrt{2}} & D+\gamma_e B_z
\end{pmatrix}
\end{align}

\bigskip
The NV center's spin usage as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}. Part of the reason for this wide usage are the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}. However the main advantage of the NV center spin compared to other solid state spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.

\subsection{Spin and optical levels intersystem crossing}
\label{sec ISC}
\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{chapter 1/Figures/ISC}
\caption{Inter system crossing of the NV center. Blue arrows symbolize decay channels to and from the $\ket{m_s=\pm 1}$ states and red arrows the decay channels to and from the $\ket{m_s=0}$ states. Green arrows symbolize the optical excitation induced by the laser. The times indicated correspond to the inverse of the decay rates found in \citep{gupta2016efficient}.}
\label{ISC}
\end{figure}

The optical excitation and radiative decay between the $^3A_2$ and $^3E$ levels preserve the spin projection $m_s$ \citep{robledo2011spin}. There is however a non-radiative decay path which involves in inter-system crossing with the singlet states $^1A_1$ and $^1E$ that depends on the the triplet states spin projection.

Fig. \ref{ISC} summarizes the possible decay channels from the $^3E$ states. We assume that the electrons are initially excited from the $^3A_2$ states via a green laser with a rate $\gamma_{\rm las}$. The $^3E$ states can then either decay radiatevely and emit a 600$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states. We consider the decay channel through the singlet state as non-radiative due to the strong non radiative channels competing with the IR radiative transition (quantum efficiency $\leq 10^{-3}$ for the singlet transition \citep{rogers2008infrared, ma2010excited, acosta2010optical}).

The rates $\gamma$ of the various decay processes, represented in Fig. \ref{ISC} by the time constant $\tau=1/\gamma$, strongly depends on the spin projection $m_s$: the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$\frac{15+91}{15+11}\approx 4$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state. Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.

These spin selective intersystem crossings mean that an optical pumping scheme in the $\ket{m_s=0}$ spin state takes place under constant green excitation. At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}, which is equivalent to a thermal polarization (for B=0) with a spin temperature of $\approx 65\ \rm mK$ \footnote{\begin{equation*}
T=\frac{h D}{k_B \ln (\frac{1-P}{2P})} \approx 65\ \rm mK,
\end{equation*}
where $D$ is th ZFS factor described in eq. (\ref{eq. spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.} while the diamond itself can remain at room temperature.

Another effect of the intersystem crossing is the ability to optically readout the spin state of the NV center: since the $\ket{0}$ state is more likely to undergo a radiative decay than the $\ket{\pm 1}$ states, the expected photoluminescence (PL) is higher when starting from the $\ket{0}$ state than a $\ket{\pm 1}$ state. While the $\ket{+1}$ and $\ket{-1}$ states cannot be a priori distinguished, a direct excitation of the $\ket{+1} \to \ket{0}$ or $\ket{-1} \to \ket{0}$ transition with a resonant microwave can unambiguously reveal the initial spin state.

It should be noted that the decay rates indicated in Fig. \ref{ISC} cannot be directly measured. They have to be inferred from the PL dynamics at short times after illumination. Different experiments and different models have thus lead to different decay rates \citep{duarte2021effect}.

\section{Basic operations with NV centers}

This section introduces two elementary operations with NV centers that are Rabi oscillations and optically detected magnetic resonance (ODMR).

Rabi oscillations is the basis for any kind of coherent qubit manipulation while ODMR is used to find the transition frequencies of the NV centers, as well as being a basic magnetometry protocol.

This section also covers the experimental setup used for these two operations and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.

\subsection{Experimental setup}
\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{chapter 1/Figures/shema_exp}
\caption{Confocal microscopy setup detailed in main text}
\label{setup}
\end{figure}
The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field. Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.

Fig. \ref{setup} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms. Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.

The setup consists of a confocal microscope where the excitation from a green laser (Thorlabs CPS532, 532 nm, 5 mW) is focused on a sample containing NV centers with a microscope objective (NA $\sim$ 0.5). The NV PL is then collected by the same objective and isolated from the back-scattered laser light with a dichroic mirror (DM) and additional filters (532 nm notch filter and 645 nm longpass filter to filter out the NV$^0$ PL). The PL is collected on an avalanche photodiode (Thorlabs APD410A or Perkin Elmer SPCM-AQRH-15) whose signal is sent to the acquisition card (NI X-series).

The laser beam goes through an acousto-optic modulator (AOM, AA opto-electronic) controlled externally to create arbitrary light pulses (resolution $\sim$ ns). The microwave field is generated by a Rhode \& Shwarz SMB 100A generator and emitted with a home-made loop antenna positioned at $<$ 1 mm of the sample. A microwave switch (Mini-circuits ZASWA-2-50DRA+) is used to create microwave pulses (resolution $\sim$ 10 ns). The control pulses for the AOM and the microwave switch are generated either by the NI X-series card (resolution $\sim$ 1 $\mu$s), or by a pulseblaster ESR-PRO (resolution $\sim$ 10 ns). The microwave field can be amplified with a Mini-circuits ZHL-5W-422+ amplifier to achieve higher Rabi frequencies.

Finally the magnetic field is generated either with a permanent magnet (up to $\sim$ 200 mT) or with a homemade electromagnet (up to $\sim$ 30 mT) alimented via a low freqency amplifier (Leybold power function generator 522663). Both the permanent magnet and the electromagnet can be  attached to a rotation stage for precise alignment.
\subsection{Rabi Oscillations}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/Rabi}
\caption{Rabi oscillations. (a) pulse sequence used to measure Rabi oscillations. (b) Experimental data on sample CVD-pink (cf samples [REF]).}
\label{Rabi}
\end{figure}

The general technique for experiments involving NV spins proceeds as follow: 
\begin{itemize}
\item First the spins are polarized. This is done with a green laser pulse ($1-100$ $\mu$s depending on the laser intensity) which polarizes the spins in the $\ket{m_s=0}$ state at $\sim$ 80 \%.
\item  Then the spins are manipulated. This process often involves a microwave field resonant with one of the spin transitions.
\item Finally the spins are read out. The simplest spin readout for NV centers consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low. For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore. The optimal readout time is of the same order as the spin polarization time ($1-100$ $\mu$s).
\end{itemize}

Fig. \ref{Rabi}-a) shows the pulse sequence used to observe Rabi oscillations with NV centers. In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$. The microwave frequency or the magnetic field needs to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.

Fig. \ref{Rabi}-b) shows the result of such an experiment on a spin ensemble. The PL signal oscillations correspond to the oscillations of the spin population between the bright $\ket{0}$ state and the dark $\ket{-1}$ state. For long microwave pulses ($>$ 1 $\mu$s typically), the spins reach a steady state with  $\sim$ 50 \% population in both the $\ket{0}$ and $\ket{-1}$ states.

\subsection{Optically detected magnetic resonance}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_basique}
\caption{ODMR protocol. (a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$. (b) ODMR spectrum with no magnetic field on sample ADM-15-4.}
\label{ODMR basique}
\end{figure}

Optically detected magnetic resonance (ODMR) is the most commonly employed magnetometry protocol with NV centers due to its simplicity and the fact that it does not require prior calibration.

The protocol for continuous wave (CW) ODMR is presented in Fig. \ref{ODMR basique}. In this simple protocol, the microwave is always turned on while the frequency spans a predefined interval $[f_{\rm start}, f_{\rm end}]$. The PL is recorded as the microwave is being swept, with a typical acquisition time for each PL sample $> 1\ \rm ms$. The sweeping frequency must be slow enough so that the PL can reach a steady state with the microwave field for each sample acquisition.

As we saw in the last section, the steady state of the NV center spin with a resonant microwave field is a statistical mixture of 50 \% $\ket{0}$ and 50 \% $\ket{\pm 1}$ \footnote{We are assuming here that the Rabi frequency is greater than the laser polarizing rate, which is often the case in practice}. If the microwave field is not resonant with a spin transition however, the steady state is predominantly $\ket{0}$ due to the spin polarization caused by the laser. The NV PL will therefore be lowered when the scanning microwave field becomes resonant with one of the NV spin transitions.

Fig. \ref{ODMR basique}-b) shows an example of ODMR spectrum on an NV ensemble with no external magnetic field. The reason for the two bumps instead of a single one centered on $D=2870\ \rm MHz$ is the splitting of the $\ket{+1}$ and $\ket{-1}$ levels occasioned by the strain and local electric field, which will be further discussed in chapter 4.

\subsubsection{ODMR with lock-in detection}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_lockin}
\caption{ODMR protocol with lock-in detection. (a) Simplified confocal microscope setup with a lock-in amplifier (LIA) modulating the mircowave and demodulating the PL. (b) Demodulated ODMR spectrum with no magnetic field on sample ADM-15-4.}
\label{ODMR lockin}
\end{figure}

Because the ODMR protocol delivers a continuous signal (the PL) as a function of a periodic excitation (the microwave field), it is well suited to amplitude or frequency modulation and lock-in detection.

The principle of lock-in detection will not be covered here, the key idea is that lock-in detection allows to translate the signal spectrally in a region less prone to noise. In this case, the ODMR signal is a DC measurement and is therefore prone to slow freqency noises such as thermal or mechanical drifts and laser amplitude fluctuations. The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low freqency noises, but slow enough compared to the NV photodynamics.

Fig. \ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig. \ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation. All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1\ \rm kHz$.

\subsection{ODMR with NV ensemble: longitudinal and transverse magnetic field}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_8_classes}
\caption{(a) Geometry of the NV center and definition of $B_z$ and $B_\perp$. (b) Representation of the 4 possibles orientations (``classes") of the N$-$V axis, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell. (c) Simulation of the evolution of the transition frequencies for each class of NV centers as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily). (d) ODMR spectrum on sample ADM-150-1 for a magnetic field of $\sim 90\ \rm G$.}
\label{ODMR 8 classes}
\end{figure}

We mentioned previously that the NV spin Hamiltonian (eq. (\ref{NV spin Hamiltonian basic})) was anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal. The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig. \ref{ODMR 8 classes}-a).

For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$. The true eigenstates however remain close to $\ket{0}$, $\ket{\pm 1}$ as long as $\gamma_e B_z > \frac{(\gamma_e B_\perp)^2}{D}$.} can be approximated in a perturtbative approach to: %cf AN.py
\begin{align}
\nu_0&= -\frac{(\gamma_e B_\perp)^2}{D} \\
\nu_{-1} &= D - \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D} \\
\nu_{+1} &= D + \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}
\end{align}

Fig. \ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal. We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian. For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and that each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).

Fig. \ref{ODMR 8 classes}-c) shows a simulation of the evolution of the transition frequencies $\nu_{+1} - \nu_{0}$ and $\nu_{-1} - \nu_{0}$ for all 4 classes as a function of the magnetic field amplitude, with the magnetic field orientation fixed in an arbitrary direction. Fig. \ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles. 8 lines are visible corresponding to the two transitions of the 4 classes. Based on the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.

We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec. \ref{sec ISC}. For strong transverse magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.

\section{Dynamics of the NV center spin}

This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times. Relaxation times are important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.

The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ which corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix), and the transverse relaxation time $T_2$\footnote{We follow here the NMR nomenclature} which corresponds to the relaxation of the spin state coherences (non diagonal elements of the density matrix).

The transverse relaxation time, also called coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$. The $T_2^*$ time includes inhomogeneous dephasing processes such as the inhomogeneity in space (when multiple qubits are probed) and time of the external fields. The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}. While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used. More careful nomenclatures will label the $T_2$ times as $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance. The relaxation times are ordered as follows:
\begin{equation}
T_2^* < T_2 < 2 T_1.
\end{equation}

The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the sample used. The following paragraphs will cover the measurement processes of the different relaxation times and the typical values for the samples used in this manuscript.

\subsection{NV spin $T_1$}
\label{sec T1 NV}
\subsubsection{Measurement protocol}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/T1_basique}
\caption{Basic $T_1$ measurement scheme. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink for low laser intensity. The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
\label{T1 basique}
\end{figure}

$T_1$ is the simplest relaxation time to measure because the spin populations are directly correlated to the NV centers PL. Fig. \ref{T1 basique} shows the pulse sequence used and an experimental result of the $T_1$ sequence. The spin lifetime $T_1$ is one of the only spin property that can be measured without a resonant microwave.

The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig. \ref{T1 soustraction}-a). This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$. Then the sequence is repeated with an additional $\pi$-pulse on one the NV transitions right before the readout. This $\pi$-pulse will project the remaining $\ket{0}$ population difference into a $\ket{\pm 1}$ state. The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which rejects every contribution to the PL outside of the spin relaxation.

Fig. \ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample that was used in Fig. \ref{T1 basique}. The difference between the $S_1$ data in Fig. \ref{T1 soustraction}-b) and the the data in Fig. \ref{T1 basique}-b) comes from the laser intensity used. In Fig. \ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant. Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark \citep{giri2018coupled, giri2019selective} (most likely due to charge tunneling between impurities \citep{choi2017depolarization}) . On the other hand, the laser intensity used in Fig. \ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.

Fig. \ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal. Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig. \ref{T1 basique}.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/T1_protocole_soustraction}
\caption{Common mode rejection $T_1$ measurement scheme. (a) Pulse seqeunce used. (b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity. (c) Subtraction $S_1-S_2$ versus dark time $\tau$. The experimental data is fitted with an exponential decay of time constant $\tau=4.1\ \rm ms$.}
\label{T1 soustraction}
\end{figure}

The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}. The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.

\subsubsection{Physical origin and typical values}

We distinguish three different origin for the the $T_1$ relaxation process:
\begin{itemize}
\item \textbf{Spin-phonon relaxation:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which can result in a phonon-induced spin relaxation. The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}. The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxtion times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at for mili-K temperatures \citep{astner2018solid}.

\item \textbf{Electric and magnetic field noises:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}. As a result, magnetic and electric field noise can reduce the spin $T_1$ if their spectrum overlap with the NV transitions. This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}. The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effects are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).

\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipolar interaction betwen two spins will be detailed more thoroughly in the following section. Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation. The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration and can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
\end{itemize}

All the measurements shown in this manuscript were performed at room temperature. The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$. Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses. The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed. $T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.

\subsection{NV spin $T_2^*$}
\subsubsection{Measurement protocol}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/Ramsey}
\caption{Ramsey interferometry protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink.} %The orange line corresponds to an Gaussian fit with time constant $\tau=400 \pm 100\ \rm ns$.}
\label{Ramsey}
\end{figure}

The $T_2^*$ time is the characteristic time for the loss of quantum coherence of a freely evolving system. It is generally defined as the decay time of of a Ramsey interferometry measurement \citep{barry2020sensitivity}.

Fig. \ref{Ramsey} shows the Ramsey pulse sequence and an  example obtained on a spin ensemble. The microwave frequency is detuned by $\sim 300\ \rm kHz$ compared to the spin central Larmor frequency which causes oscillations in the signal. $T_2^*$ is the decay time of the envelope of the Ramsey fringes. Unfortunately, it is hard to correctly fit the envelope of a signal with so few fringes. Increasing the number of fringes with our current microwave and optical setup might prove hard, as further microwave detuning would decrease the signal contrast, and would be ultimately limited by the microwave pulses resolution.

Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons. The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity, and because the samples used for single NV experiments generally contains less impurities than those used for NV ensemble. The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.

An alternative measurement of $T_2^*$ can be performed by looking at the linewidth of an ODMR spectrum. Indeed, ideally the ODMR spectrum should be the Fourier transform of the Ramsey signal (also called free induction decay in NMR literature). In practice however, both methods suffer from experimental noises and imperfections, with Ramsey-type methods being more sensitive to microwave inhomogeneities \citep{barry2020sensitivity} while ODMR is prone to laser and microwave power broadening \citep{dreau2011avoiding}.

The method chosen to extract $T_2^*$ values in this manuscript was ODMR with low microwave power ($\Omega_{\rm Rabi} \ll 1/T_2^*$) and low laser intensity (compared to the the NV optical saturation).

\subsubsection{Physical origin}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_1_classes}
\caption{ODMR spectra on a single NV class. (a) Experimental data on sample CVD-pink, fitted with three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$). (b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
\label{ODMR 1 classe}
\end{figure}

For most samples, the leading cause of inhomogeneous dephasing is the temporal and spatial variations of magnetic fields caused by the paramagnetic impurities in the crystal \citep{barry2020sensitivity}. The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2. We will simply mention here that for sample with high NV density ([NV] $\geq$ ppm), $T_2^*$ is mostly limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.

Fig. \ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see [REF]). The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$. We can clearly see a higher linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration. In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.

Fig. \ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula. While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}. For Fig. \ref{ODMR 1 classe}-a), it was indeed found that Lorentzians fitted the data better than Gaussians. For Fig. \ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, although this could be in part due to the unresolved hyperfine structure.

\subsection{NV spin $T_2$}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 1/Figures/Echo}
\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
\label{Echo}
\end{figure}
\subsubsection{Measurement protocol}
As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.

Fig. \ref{Echo} shows the pulse sequence used and an example on the same sample than the Ramsey measurement. The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time. This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased \citep{hahn1950spin}.

This protocol extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10. Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}. This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.

\subsubsection{Physical origin}
The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence. The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}. For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2019decoherence}.

\section{Dipole-dipole interaction between two electronic spins}

\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth]{chapter 1/Figures/shema_dipole_dipole}
\caption{Geometric representation of two interacting spins. ($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$) represent Cartesian coordinate systems in the spin frames of reference, where $\mathbf{z}_i$ is the axis of quantization of spin $i$.} 
\label{dipole-dipole}
\end{figure}

This final introductory section will focus on the dipole-dipole interaction between two electronic spins, which is the main form of spin-spin interaction and plays a key part in the properties of dense spin ensemble.

We will consider the general situation represented in Fig. \ref{dipole-dipole}: two spins separated by a distance $r$ with their own coordinate systems ($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$). The coordinate system is chosen so that each spin operator can be written:

\begin{equation}
\hat{\mathbf{S}}_i=\hat{S}_x \mathbf{x}_i + \hat{S}_y \mathbf{y}_i + \hat{S}_z \mathbf{z}_i
\end{equation}

where the spin operators $\hat S_{\{x,y,z\} }$ have their usual definitions in term of Pauli matrices. In particular, this means that the spin axis of quantization of each spin is $z_i$, which classically corresponds to the magnetization direction of the dipole. This axis can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (including the NV center).

\subsection{Magnetic dipole-dipole Hamiltonian}

Classically, the dipole-dipole interaction between two magnetic moments $\mathbf{m}_1$ and $\mathbf{m}_2$ can be derived by computing the static magnetic field generated by one dipole at the position of the second dipole \cite[p.~188]{jackson1999classical}:

\begin{align}
U&=-\mathbf{m}_1 \cdot \mathbf{B}(\mathbf{r}) \nonumber \\
&=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u}) - \mathbf{m}_1\cdot\mathbf{m}_2}{r^3}+\frac{8\pi}{3}\mathbf{m}_1\cdot\mathbf{m}_2\,\delta(r)\right], \label{eq. dipole classique}
\end{align}

where $\mathbf{r}=r \mathbf{u}$ is the relative position of the two dipoles (see Fig. \ref{dipole-dipole}).

The quantum analog of the dipole-dipole interaction between spins simply replaces the magnetic moments $\mathbf{m}_i$ in eq. (\ref{eq. dipole classique}) with $\hbar \gamma_i \hat{\mathbf{S}}_i$, where $\gamma_i$ is the gryromagnetic ratio and $\hat{\mathbf{S}}_i$ the spin vector operator for spin $i$:

\begin{equation}
\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2\,\delta(r)\right].
\end{equation}

The $\delta(r)$ term represents the Fermi contact energy. This term plays an important role when there is an overlap of the spin particle wavefunctions, such as the hyperfine coupling between the NV$^-$ electron spin and $^{14}$N \cite{doherty2012theory} or  $^{13}$C \cite{smeltzer201113c} nuclear spins.

The focus of this manuscript is on the interaction between electronic spins which are separated by several atomic sites (density impurities in the $\sim$ ppm range). We will therefore assume that there is negligible overlap between the electronic wavefunctions and omit the Fermi contact term. This leads the simplified dipole-dipole Hamiltonian:

\begin{equation}
\mathcal{H}_{\rm dd}= \frac{J_0}{r^3}\left[ 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2] \right],
\end{equation}

where $J_0=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}$.

For two electronic spins with $g$-factors $\approx -2$\footnote{Most spin defects in diamond, including the NV center have a g-factor $\approx -2$ similarly to free electrons.}, 
\begin{equation*}
\frac{J_0}{h} \approx 51.8\ \rm{MHz}\cdot \rm{nm}^3.
\end{equation*}

\subsection{Decomposition of the dipole-dipole Hamiltonian}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{chapter 1/Figures/flip-flop_double-flip}
\caption{Representation of the two energy transfer mechanisms: flip-flop and double flips} 
\label{flip flop double flip}
\end{figure}

In the weak coupling approximation, the dipole-dipole Hamiltonian between two spins can be decomposed between diagonal terms which are responsible for energy shifts of the individual spin levels, and the off-diagonal terms which are responsible for energy transfers between the two spins.

In what follows, we will assume that both spins have the same quantization axis in order to simplify the Hamiltonian decomposition. A more general case is detailed in [REF].

Following the notation introduced in Fig. \ref{dipole-dipole}, we will then assume that $$(\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1) = (\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2) = (\mathbf{x},\mathbf{y},\mathbf{z}).$$ We will denote $\mathbf{u}=u_x \mathbf{x} + u_y \mathbf{y} + u_z \mathbf{z}$ the decomposition of the unitary vector $\mathbf{u}$ in this basis. The dipole-dipole Hamiltonian can then be written:
\begin{equation}
\mathcal{H}_{\rm dd}=\frac{J_0}{r^3}\left[ (3 u_x^2-1) S_x^1 S_x^2 + (3 u_y^2-1) S_y^1 S_y^2 + (3 u_z^2-1) S_z^1 S_z^2 \right],
\end{equation}
where $S_{\{x,y,z\} }^i$ are the spin operators acting on spin $i$. The $S_z^1 S_z^2$ term is the only diagonal term and corresponds to energy shifts while the $S_x^1 S_x^2$ and $S_y^1 S_y^2$ terms are responsible for energy transfer. To make the transfer mechanism more apparent, we will rewrite $S_x^i$ as $\frac{1}{2}(S_+^i+S_-^i)$ and $S_y^i$ as $\frac{1}{2i}(S_+^i-S_-^i)$:
\begin{align}
\mathcal{H}_{\rm dd}&=\frac{J_0}{r^3}\bigg{[} \frac{1}{4}(3 u_x^2 + 3 u_y^2 -2)(S_+^1 S_-^2 + S_-^1 S_+^2) \nonumber \\
&\quad + \frac{1}{4}(3 u_x^2 - 3 u_y^2)(S_+^1 S_+^2 + S_-^1 S_-^2) \nonumber \\
&\quad +(3 u_z^2-1) S_z^1 S_z^2 \bigg{]} \nonumber \\
\medskip
&= \Omega_{\rm ff} (S_+^1 S_-^2 + S_-^1 S_+^2) + \Omega_{\rm df} (S_+^1 S_+^2 + S_-^1 S_-^2) + \Omega_{\rm shift} S_z^1 S_z^2. \label{eq. ff df}
\end{align}

We have introduced in eq. (\ref{eq. ff df}) the three relevant matrix elements $\Omega_{\rm ff}$, $\Omega_{\rm df}$, and $\Omega_{\rm shift}$ which correspond respectively to flip-flops, double flips and energy shifts. Flip-flops and double flips (sometimes referred as flip-flips) are represented in Fig. \ref{flip flop double flip}. Flip flops correspond to energy transfer between two spins for which the sum of the spin angular momentum is conserved ($\Delta (m_s^1+m_s^2)=0$), while double flips correspond to energy transfer for which the spin angular momentum is not conserved ($\Delta (m_s^1+m_s^2)=\pm 2$)\footnote{In order to preserve the total angular momentum, double flips transfer angular moment to the crystal, which is a form of Einstein de Haas effect \citep{einstein1915experimental} }. In both cases however, the energy between the initial and final states needs to be preserved for the transfer to be efficient.

It should be noted that on average, each matrix element is of the same order of magnitude: $\Omega_{\rm ff} \sim \Omega_{\rm df} \sim \Omega_{\rm shift} \sim \frac{J_0}{r^3}$.

\subsection{Cross-relaxations with NV centers}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 1/Figures/Shema_CR}
\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
\label{CR_shema}
\end{figure}

Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family). This process can occur through either flip-flops or double flips, as long as the resonance condition between the two spins is met.

On of the condition for cross-relaxation is to have a difference in polarization between the two spins species considered. The NV center is therefore a good candidate to observe cross-relaxations as it can be optically polarized while most other spin species are not.

Fig. \ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin. When the two considered spin transitions become resonant, in this case by tuning the magnetic field, polarization from the NV center will be transferred to the second spin, making the NV center less polarized and the second spin more polarized than when they are out of resonance.

For a single NV center coupled to spin 2, assuming that spin 2 always remains depolarized\footnote{This is the case in practice if the spin 2 bath is much bigger than the NV bath and effectively acts as a thermostat, or if the lifetime of spin 2 is much smaller than that of the NV center.} and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$ (weak coupling regime), the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by the cross-relaxation with spin 2 can be analytically computed \citep{hall2016detection}:
\begin{equation}
\delta \Gamma_1=\frac{\Omega^2}{\Gamma_2^*} \frac{(\Gamma_2^*)^2}{(\delta \nu)^2+(\Gamma_2^*)^2},
\label{delta gamma 1}
\end{equation}
where $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, $\delta \nu$ is the detuning in energy (expressed in frequency) between the initial and final states and $\Omega=\Omega_{\rm ff}$ or $\Omega_{\rm df}$ is the dipole-dipole matrix element corresponding to the flip-flop or double-flip process considered.

Eq. (\ref{delta gamma 1}) is the basis for every cross-relaxation detection performed in this manuscript. Because the NV center lifetime can be measured optically, as detailed in Sec. \ref{sec T1 NV}, cross-relaxation between NV-centers and other spin species can be measured optically. Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be explained in the following chapters.

\section{Conclusion}

We have covered in this introductory chapter every concepts needed to understand the results presented in the next chapters. We have seen the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them, and we have introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.

The focus of chapter 2 will be on the cross-relaxation between NV centers and other spin impurities in CVD-grown diamond, while the focus of chapter 3 and 4 will be on the cross-relaxation between NV centers and other NV centers, as well as their potential application in magnetometry.
\printbibliography
\end{refsection}
 \begin{refsection}

\chapter{Detection of dark spins via cross-relaxations}
In this chapter, we will see how the interplay between the optical and spin properties of the NV centers can be transferred to other spins in the crystal by coupling them to NV centers. More specifically, I will discuss the first optical detection of two spin species, VH$^-$ and War1, in a CVD diamond by the mean of cross-relaxations with NV centers.
\section{Flip-Flops, double flips and cross-relaxation}
\label{Sec_CR}

\subsection{Resonant interaction : flip-flops and double flips}

The dominant form of interaction between two distant spins is the magnetic dipole-dipole interaction. For two spins with with vector spin Hamiltonian $\hat{\mathbf{S}}_1$ and $\hat{\mathbf{S}}_2$, separated by a distance $\mathbf{r}=r\mathbf{u}$, the dipole-dipole interaction Hamiltonian reads (see Appendix A) :
\begin{equation}
\mathcal{H}_{dd}=\frac{J_0}{r^3}\left[ \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2 - 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u})\right],
\end{equation}
where $J_0= \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$. For two electronic spins with $g-$factors close to 2 (which is the case for the NV center and most spin defect in diamond), the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$.

Flip-flop, in the context of dipole-dipole interaction, is the mechanism which exchanges a quantum of spin between two spins. For an initial state of two spins $\ket{m_s^1=i;m_s^2=j}$, a flip-flop process will result in a final state $\ket{i+1;j-1}$ or $\ket{i-1;j+1}$. The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$, as well as the resonance condition between the initial and final state. We should note that for two identical spins, the flip-flop process $\ket{i;i\pm 1}\bra{i\pm 1,i}$ is always resonant.

Another mechanism authorized by the dipole-dipole Hamiltonian is the double-flip, either up or down, which couple the states $\ket{i;j}$ and $\ket{i-1;j-1}$ or $\ket{i+1;j+1}$. For a single spin species in a strong magnetic field, the lift of the various spin levels by the Zeeman effect means that no double-flip process can be resonant. However with weak magnetic field or when two different spin species are present, double flip process can be as important as flip-flops since the matrix elements $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \pm 1}$ are typically of the same order of magnitude as the flip-flop ones.

\subsection{Cross-relaxations}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 2/Figures/Shema_CR}
\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
\label{CR_shema}
\end{figure}

Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family). This process con occur either through flip-flops, or through double flips, as long as the resonance condition between the two spins is met.

Fig. \ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin. When the two considered spin transitions become resonant, in this case by tuning the magnetic field, polarization from the NV center will be transferred to the second spin, making the NV center less polarized and the second spin more polarized than when they are out of resonance.

The reason why CR tend to depolarize the NV center simply comes from a rate equation : since the initial NV population is higher in the $\ket{0}$ state than in the $\ket{-1}$ state, the flip-flop (or double flip) process that makes the NV center go to $\ket{-1}$ (symbolized by the two red arrows in Fig. \ref{CR_shema}) is more likely than the reverse flip-flop. If we only focus on the NV center, cross-relaxation is simply another thermalization process due to the coupling of the NV center to its environment.

For a single NV center coupled to spin 2, assuming that spin 2 is initially depolarized and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$, we can compute the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by spin 2 as \citep{hall2016detection}:
\begin{equation}
\delta \Gamma_1=\frac{\Omega^2}{\Gamma_2^*} \frac{(\Gamma_2^*)^2}{(\delta \nu)^2+(\Gamma_2^*)^2},
\label{delta gamma 1}
\end{equation}
where $\Omega=\frac{\mel{i;j}{\mathcal{H}_{\rm dd}}{j;i}}{\hbar}$ is the matrix element of the dipole-dipole Hamiltonian that couples the initial and final states, $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, and $\delta \nu$ is the detuning between the central frequencies of the NV center and spin 2.
\section{The dark spins in diamond}

We refer by dark spins to spin defects which do not interact with light, which in the case of diamond defects consist of almost every known spin defect bar the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.

\subsection{P1 centers and $^{13}$C}
By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.

In NV-rich diamonds, these two impurities are the main causes of the inhomogeneous dephasing rate $\Gamma_2^*=1/T_2^*$ of the NV centers ensemble \citep{barry2020sensitivity}. The dephasing comes from the distribution of dipole-dipole coupling from each NV centers to its neighboring spin defects, which varies in space and time.

The dephasing rate caused by a given impurity is directly proportional to the diagonal dipole-dipole matrix elements \citep{bauch2020decoherence}: 
\begin{equation}
\Gamma_2^*(\rm P1) \propto \Omega_{\rm NV-P1}=\frac{1}{\hbar}\sqrt{\sum_j \abs{\mel{NV=i;P1=j}{\mathcal{H}_{\rm dd}}{NV=i;P1=j}}^2}.
\label{Gamma2 P1}
\end{equation}
Since the dipole-dipole coupling rate scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity: $\Gamma_2^*(\rm P1) \propto [P1]$. In the case of P1 centers, this was verified experimentally by \citep{bauch2020decoherence} who found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$. The typical concentrations of P1 center go from [REF] for high purity CVD diamond, up to 500 ppm for type 1B HPHT or natural diamonds.

$^{13}$C is much more abundant than P1 center ($\sim$ 10 000 ppm at natural abundance) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\cdot 10^3$ times smaller than that of the P1. This results in a contribution to $\Gamma_2^* \sim 10^3$ times smaller per impurity. It was found experimentally that the broadening caused by $^{13}$C is $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 16\ \rm Hz/ppm$ \citep{van1997dependences, barry2020sensitivity}. For natural abundance diamond, this means that $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 160\ \rm kHz$.

\subsection{Other spin defects, VH$^-$, War1}
\label{other defects}
Other spin defects are also commonly found in diamond containing NV centers, such as charged vacancy or vacancy clusters \citep{hounsome2006origin}, other nitrogen related defects \citep{newton2007epr}, transition metals \citep{isoya1990fourier}, and especially for CVD grown diamond, hydrogen related defects \citep{hartland2014study}, since the CVD plasma contains up to 95 \% hydrogen.

Of particular interest in this chapter are two defects: the VH$^-$ defect \citep{glover2003hydrogen, glover2004hydrogen} and the War1 defect \citep{cruddace2007magnetic}.

The VH$^-$ defect, prior to our observations, had only been observed \textit{via} electron paramagnetic resonance (EPR) in CVD grown diamond. It is formed by a vacancy with a hydrogen atom bonded to one of the four dangling carbon bond, with an extra electron trapped in the vacancy. This defect has a similar electronic structure to the NV$^-$ which results in an electronic spin-1 with a similar zero field splitting parameter (see Table \ref{table VH et War1}). Although the presence of hydrogen in this defect was confirmed by isotopic study, it has been debated whether this defect was VH$^-$ or V$_2$H$^-$ due to a conflict with \textit{ab initio} calculations \citep{shaw2005importance}. The defect is therefore sometimes referred as V$_n$H$^-$. In this manuscript I will employ the original name VH$^-$.

The War1 defect, named after the University of Warwick where it was first observed in a CVD diamond \citep{cruddace2007magnetic}, is another spin-1 defect whose chemical structure remains unknown. No hyperfine structure could be discerned in the relatively broad EPR line, and the use of deuterium instead of $^1$H did not change the shape of the line.

\subsection{EPR spectroscopy}
Electron paramagnetic resonance (EPR), the electronic equivalent to NMR, is the the most common way to detect electronic spins in solids and liquids. The principle of its detection is the absorption or reflection of a microwave field resonant with a spin transition.

EPR generally relies on the thermal polarization of the spins to observe an absorption. In order to increase the polarization, an important magnetic field is applied (typ. 0.35 T or 3500 G) and a microwave cavity (typ. 9500 MHz) is used to improve the coupling of the spins to the field.

At room temperature, the thermal polarization of a transition at 9500 MHz is $\sim 0.002$. This low value, compared to $\sim 0.8$ for the NV centers, explains in part why EPR is way less sensitive than NV ODMR. The current noise floor for EPR is typically of $\sim 10^{11}$ spins for electronic spins in diamond \citep{mitchell2013x}.

\section{NV center relaxometry}

\subsection{Principle of NV center Relaxometry}
Detection \textit{via} cross-relaxation	with NV centers fall under the more general technique of NV relaxometry, where we are effectively measuring a modification in the NV center spin lifetime $T_1$. Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...

There can be several advantages to use relaxometry protocols, which measures a change in the NV $T_1$, over the more common magnetometry protocols that rely on a change in the NV $T_2$ or $T_2^*$: the first and most obvious one is that relaxometry does not need to use an external microwave field. This however comes at the cost of a greater precision needed on the external magnetic field. Relaxometry also allows AC-magnetic field sensing in the GHz regime \citep{wang2022picotesla, alsid2022solid}, while traditional microwave-based AC-megnetometry is limited to tens of MHz. Finally, relaxometry measurement can be more sensitive than the T2-based one in some instances \citep{steinert2013magnetic}, which comes from the fact that changes in NV $T_1$ times are easier to detect than changes in $T_2$ or $T_2^*$ since $T_1$ is much longer than $T_2$ or $T_2^*$.

\subsection{PL and single$-\tau$ relaxometry protocols}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 2/Figures/relaxo_pulsed_continuous}
\caption{Representation of the two relaxometry protocol. Left : representation of the three level system of the NV center with the polarization rate $\Gamma_{\rm las}$ and spin decay rates $\Gamma_1=1/T_1$ and the perturbation $\delta \Gamma$. Right: simulation of a $T_1$ measurement for $1/\Gamma_1=3\ \rm ms$ and $1/(\Gamma_1+\delta \Gamma)=2\ \rm ms$. $\delta \rm S$ represents the maximum difference in signal between the two curves for a dark time $\tau$.}
\label{T1 vs PL}
\end{figure}

There are two ways to measure a change in the spin $T_1$, as represented on Fig. \ref{T1 vs PL}. The first and most basic method simply consist in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, which itself results from an equilibrium between the polarization rate $\Gamma_{\rm las}$ and the spin decay rate $\Gamma_1=1/T_1$. A change in $\Gamma_1$ will therefore conduct to a new equilibrium with a different PL.

The second method is a pulsed sequence referred as single$-\tau$ measurement, used for example in \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}. It simply consists in a $T_1$ measurement sequence (as seen in chapter 1) where you first polarize the spins in the $\ket{0}$ state with a laser pulse and read out the spin state with a second laser pulse after a dark time $\tau$. In this case however, instead of scanning the time $\tau$, you fix its value to maximize the sensitivity in a change of $T_1$ (typically $\tau \approx T_1$).

It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the first case you need to adjust $\Gamma_{\rm las} \sim \Gamma_1$ to maximize the PL contrast, meaning that the laser power used is very far from the saturation limit (typically $\Gamma_1 \approx 10^3\ \rm s^{-1}$ and $\Gamma_{\rm sat} = 5\cdot10^6\ \rm s^{-1}$\citep{dreau2011avoiding}), while the second method needs to wait a time $\tau \sim T_1$ between each measurement, which result in a similar PL count and shot noise limit in both case.

On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and uses an overall smaller laser power, however it is more sensitive to drifts and changes in the optical setup. Most of the relaxometry measurement in this manuscript were performed using a PL-based detection.

\subsection{Comparison with DEER protocol}

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{chapter 2/Figures/DEER}
\caption{Pulse sequence of the DEER protocol}
\label{DEER}
\end{figure}

Outside of cross-relaxations, another protocol exists to address dark spin resonance with NV centers, based on a modification of the NV $T_2$ time \citep{serbyn2014interferometric}. This protocol called double electron-electron resonance (DEER) is depicted on Fig. \ref{DEER}. The pulse sequence consist in a spin-echo sequence on the NV center, as described in chapter 1, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers. This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center wont be rephased. The newly measured $T_2$ time should therefore be slightly shorter:

\begin{equation}
\frac{1}{T_2^{\rm DEER}}=\frac{1}{T_2^{\rm echo}} + \delta \Gamma_2^*,
\end{equation}

where $\delta \Gamma_2^*$ is the contribution of the dark spin to the $T_2^*$ of the NV center.

By scanning the frequency of the second microwave (RF2) and monitoring the $T_2$ time of the NV centers, one can therefore find dark spin resonance without having to bring them to resonance with NV centers, which can be useful when the dark spin resonance is very far detuned from the NV one.

While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, there are several differences between the two protocols:

Frist, the two protocols do not measure the same elements of the dipole-dipole Hamiltonian: CR measures the off-diagonal elements $\Omega_{\rm CR}=\mel{i,j}{\mathcal{H}_{\rm dd}}{j,i}/\hbar$ which are responsible for the flip-flops or double flips, while DEER measures the diagonal elements $\Omega_{\rm DEER}=\mel{i,j}{\mathcal{H}_{\rm dd}}{i,j}/\hbar$ which are responsible for the pure dephasing. These terms however are on average of the same amplitude, and when averaging over an ensemble we can consider $\Omega_{\rm CR} \sim \Omega_{\rm DEER}$

Second and more importantly, the scaling is not the same in both cases : CR measures a change in the NV decay rate $\Gamma_1=1/T_1$ which is proportional to $\Omega_{\rm CR}^2$, as given by eq. \ref{delta gamma 1}, while DEER measures a change in $\Gamma_2=1/T_2$ which is proportional to $\Omega_{\rm DEER}$, as given by eq. \ref{Gamma2 P1}. This means that the signal of CR scales quadratically with the impurity concentration, while the DEER signal scales linearly with it. CR is therefore more suited to probe concentrated impurities, when the average distance between the NV and the probed spin is small.

\section{Choice of the magnetic field orientation}

Since relaxometry does not rely on a microwave field, the only parameter left to tune is the external magnetic field, which is often scanned with an electromagnet. Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect dark spins.

\subsection{The [100] and [111] magnetic field orientations}
\label{sec simu}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 2/Figures/121_vs_22_NV}
\caption{Simulated transition frequencies of the $\ket{0} \to {-1}$ and $\ket{0} \to {+1}$ transitions for the 4 classes of NV centers as a function of the magnetic field. a) When the magnetic field is along the diamond [100] crystalline axis, b) along the [111] axis. A representation of the 4 classes of NV centers, labeled 1 to 4, and the orientation of the magnetic field in both cases is put on top}
\label{121 vs 22 NV}
\end{figure}

Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\mathbf{B} \parallel [100]$ or 
$\mathbf{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes as defined in chapter 1. When $\mathbf{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\mathbf{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.

These changes in the field orientation result in massive changes in the NV behavior, manly due to the transverse magnetic field: Fig. \ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the ground state spin Hamiltonian (REF chapter 1) - or to be more precise the transitions between the ground state of the spin Hamiltonian and the two excited states - for any of the four classes of NV centers when $\mathbf{B} \parallel [100]$. Notably, there is no transitions frequencies below 2500 MHz.

On the other hand, Fig. \ref{121 vs 22 NV}-b) shows the transition frequencies when $\mathbf{B} \parallel [111]$, for both the class aligned and the three class misaligned. The three classes behave similarly than the four classes in the $[100]$ case, but the one class aligned with the magnetic field can get its $\ket{0}\to\ket{-1}$ transition arbitrarily small, as long as the magnetic field is perfectly aligned with the NV axis.

\subsection{CR condition between NV$^-$ and P1}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 2/Figures/121_vs_22_P1}
\caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}
\label{121 vs 22 P1}
\end{figure}

P1 centers, the most abundant electronic spin in the diamonds commonly used, have a $1/2$ electronic spin. They do have a small magnetic field anisotropy coming from the hyper-fine coupling to the nitrogen nucleus ($\sim 100\ \rm MHz$), but this dependency on the field orientation is small compared to the NV center's zero field splitting $D=2870\ \rm MHz$. We can therefore, to the first order, neglect the hyper-fine coupling and consider P1 centers as ideal, isotropic, spin $1/2$.

Fig. \ref{121 vs 22 P1} shows the transition frequencies for the NV center and such a spin $1/2$, as a function of a magnetic field parallel to [100] or [111]. When $\mathbf{B} \parallel [100]$, there will never be a co-resonance between the NV and P1 transitions, meaning that no NV-P1 cross-relaxation can be observed when B is scanned in this direction. When $\mathbf{B} \parallel [111]$ however, the P1 $\ket{-1/2} \to \ket{+1/2}$ transition matches in energy the $\ket{0}\to\ket{-1}$ NV transition for B=512 G. This means that when B reaches $\approx 512\ \rm G$, some of the polarization of the NV centers will be transferred to the P1 centers, which will result in a drop in the NV PL.

\subsection{CR condition between NV$^-$ and VH$^-$}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 2/Figures/121_vs_22_VH}
\caption{Simulation of the transition frequencies of NV$^-$ centers and VH$^-$ centers for a magnetic field a) along the [100] axis and b) along the [111] axis}
\label{121 vs 22 VH}
\end{figure}

VH$^-$, a much less abundant electronic spin than P1 center, has an electronic spin 1, with a spin Hamiltoninan very similar to that of the NV center. In particular it has the same symmetries being also a $C_{3v}$ defect, with a slightly smaller ZFS at $D_{VH} \approx 2700\ \rm MHz$ compared to $D_{NV}=2870\ \rm MHz$.

Fig. \ref{121 vs 22 VH} shows the transition frequencies for the NV$^-$ and VH$^-$ ground state spin. Contrary to the P1 case, there is a co-resonance when $\mathbf{B} \parallel [100]$ for $B\approx 55\ \rm G$. When $\mathbf{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G. A scan along the [111] in this case means that the peak signal will be $\sim 6$ times weaker than in the [100] case, and spread over multiple lines which might overlap between themselves or with other existing PL features.

The War1 spin mentioned previously is also an electronic spin 1 with (pseudo)-$C_{3v}$ symmetry and $D_{War1} \approx 2470\ \rm MHz$. This result in a similar behavior than VH$^-$ when it comes to the CR condition with NV centers.

\subsection{CR condition between NV$^-$ and and $^{13}$C-NV$^-$}

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_13CNV}
\caption{a) Representation of the $^{13}$C-NV complex (right) compared to a ``normal" NV center (left) b) and c) simulation of the transition frequencies of NV$^-$ centers and $^{13}$C-NV$^-$ complex for a magnetic field b) along the [100] axis and c) along the [111] axis}
\label{121 vs 22 13C-NV}
\end{figure}

I refer by $^{13}$C-NV$^-$ to a complex formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope. Such a complex is illustrated in Fig. \ref{121 vs 22 13C-NV}-a).

The $^{13}$C-NV$^-$ complex still behaves as a ``normal" NV center, except that its transition lines are shifted by the hyper-fin coupling. In particular its electronic spin is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states. This is confirmed by the presence of sideband in ODMR spectra around the main NV lines \citep{simanovskaia2013sidebands}.

Since both NV$^-$ and $^{13}$C-NV$^-$ are \textit{a priori} equally polarized, then following the argument in \ref{Sec_CR} there should be no CR between the two spin families. The reason why those transitions are still in some case visible, as will be shown below, and why in fact any NV-NV co-resonance gives a visible CR signal will be the object of the next chapter.

The transitions of the $^{13}$C-NV$^-$ complex are given by diagonalizing the full spin Hamiltonian:
\begin{equation*}
\mathcal{H}=\mathcal{H}_{NV}+\mathcal{H}_{^{13}C}+\mathcal{H}_{HF},
\end{equation*}
where $\mathcal{H}_{NV}$ is the NV$^-$ spin Hamiltonian,
$\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian: $$\mathcal{H}_{HF}= \hat{\mathbf{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\mathbf{I}}_C.$$

When the $^{13}$C is the direct neighbor of the vacancy as represented in Fig. \ref{121 vs 22 13C-NV}-a) - this is also referred as a first-shell $^{13}$C - the hyper-fine tensor $\mathcal{A}$ can be written \citep{simanovskaia2013sidebands}:
$$ \mathcal{A} = \begin{pmatrix}
\mathcal{A}_{xx} & 0 & \mathcal{A}_{xz} \\ 0 & \mathcal{A}_{yy} & 0 \\ \mathcal{A}_{zx} & 0 & \mathcal{A}_{zz}
\end{pmatrix},$$
where $\mathcal{A}_{xx}=190$ MHz, $\mathcal{A}_{yy}=120$ MHz, $\mathcal{A}_{zz}=129$ MHz, and  $\mathcal{A}_{xz}=\mathcal{A}_{zx}=-25.0$ MHz.

Fig. \ref{121 vs 22 13C-NV}-b) and c) show the simulated transition frequencies for a $^{13}$C-NV$^-$ complex when B is scanned along the [100] or [111] axis, as well as the ``normal" NV transitions and the co-resonance between those two. When $\mathbf{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G. When $\mathbf{B} \parallel [100]$ there are 18 of them between 10 and 60 G. This abundance of transitions for low ($<100\ \rm G$) magnetic field makes analyzing experimental data quickly intractable in the second case.

\subsection{CR contrast and transverse magnetic field}
\label{CR contrast}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 2/Figures/contrast_OMDR_111_Vs_100}
\caption{Simulated ODMR contrast on all four classes of NV centers with a magnetic field a) along the [100] axis and b) along the [111] axis}
\label{121 vs 22 contrast}
\end{figure}

In order to optically detect CR between an NV center and a dark spin, two conditions must be met: the NV center must be polarized in order to have a population transfer between the NV and the dark spin, and there must be a difference in brightness between the two NV levels involved in order to have an optical signature of the CR.

These two conditions are in fact the same needed to observe an ODMR spectrum (REF chapter 1), and while the NV centers verifies both criteria in presence of a purely longitudinal magnetic field, this is not always the case anymore in presence of a transverse magnetic field (REF chapter 1).

Fig. \ref{121 vs 22 contrast} shows the simulated relative contrast of an ODMR spectrum, as computed from \citep{tetienne2012magnetic}, for a magnetic field aligned along [100] or [111]. When the NV center is aligned with the magnetic field, the constrats remains equals to 1 regardless of the magnetic field because there is no transverse field. When the NV is not aligned however, either in the [111] or [100] case, the ODMR contrast quickly drops down for B$>100\ \rm G$.

This tells us that we should not expect to observe a CR signal for B greater than a few hundreds G when B is scanned along the [100] axis. We could however observe CR for arbitrarily large magnetic field when scanned along the [111] axis, as long as the field is properly aligned with the NV axis.

\section{Dark spin spectroscopy with NV centers}

\subsection{Detection of P1 centers}

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{chapter 2/Figures/NV_P1_libreAcces}
\caption{Change in the PL of an ensemble of NV centers as a magnetic field is scanned along the [111] crystalline axis. Taken from \citep{armstrong2010nv}}
\label{CR P1 exp}
\end{figure}

Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross} had been done on the detection of dark spins through CR with NV centers, and to our knowledge, all this studies where done on the detection of P1 centers. Following the discussion in the last section, these studies were therefore all done with $\mathbf{B} \parallel [111]$.

An example of one of this study (\citep{armstrong2010nv}) is shown in Fig. \ref{CR P1 exp}. It shows the evolution of the NV PL with respect to a magnetic field scanned along the [111] crystalline axis. There are several information to gather from this graph:

First, there is the slowly decaying PL envelope, which stabilizes to a plateau for $B\sim 1000\ \rm G$. This change in PL is caused by the transverse magnetic field acting on the three NV classes that are not aligned with the magnetic field. For $B \gtrsim 1000\ \rm G$, these three classes are fully depolarized which explains the plateau observed.

Second, there are the three salient features denoted Ns-NV, NV-NV and NV-LAC. All these features come from cross-relaxations or level anti-crossing (LAC) from the NV class aligned with the magnetic field.

The first feature (Ns-NV) for $B\approx 512 \ \rm G$ is the signature of the CR between NV and P1 centers. Its relatively complicated shape comes from the hyper-fine coupling between the P1 electronic spin and the $^{14}$N nucleus \citep{lazda2021cross}. Excited state level anti-crossing (ESLAC) also occurs in this region, but its optical signature, if it exists, is expected to be much weaker than the NV-P1 CR \citep{zheng2017level}.

The second feature (NV-NV) for $B\approx 590 \ \rm G$ comes from the cross relaxation between NV centers aligned with the magnetic field and NV centers misaligned with the field (the co-resonance can be observed in Fig. \ref{121 vs 22 NV}-b)). This phenomenon will be further discussed in the next chapter.

Finally the third feature (LAC) for $B\approx 1024 \ \rm G$ correspond to the level anti-crossing between the $\ket{0}$ and $\ket{-1}$ states of the NV center aligned with B. When the two levels get close enough in energy, the Hamiltonian off-diagonal terms (coming from strain local electric field or residual transverse magnetic field) will mix the $\ket{0}$ and $\ket{-1}$ states and prevent an efficient polarization by the green laser. This result in a sharp drop in the PL that can be exploited to perform microwave-less magnetometry with NV centers \citep{wickenbrock2016microwave, zheng2017level, zheng2020microwave}.

\subsection{Detection of VH$^-$, War1 and $^{13}$C-NV$^-$}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 2/Figures/fig_VH_et_co}
\caption{a) Change in PL (normalized) of the ensemble of NV centers in sample CVD-pink as a magnetic field is scanned along the [100] axis. The orange line is a fit to the slowly decreasing envelope b) Subtraction of the PL signal by the envelope fit as a function of the magnetic field. The green lines and red and purple regions correspond to the predicted CR regions, as detailed in main text}
\label{CR VH exp}
\end{figure}

Our work \citep{pellet2021optical} focuses on the optical detection of VH$^-$, War1 and $^{13}$C-NV$^-$ through NV CR. This study was the first to detect VH$^-$ and War1 defects optically and the first CR experiment on a CVD diamond. It was also the first,to our knowledge, to use a [100] aligned magnetic field.

The diamond used for this study is the sample CVD-pink [REF appendix] whose fabrication was detailed in \citep{tallaire2020high}. It is a CVD diamond containing $[\rm NV ]\approx 4\ \rm ppm$ for $[\rm N ]\approx26\ \rm ppm$, and a natural abundance (1.1\%) of  $^{13}$C. Being a CVD diamond it is also likely to incorporate hydrogen defects, as discussed in \ref{other defects}.

Fig. \ref{CR VH exp}-a) shows the change in PL with respect to a magnetic field scanned along the diamond [100] axis. Similarly to fig. \ref{CR P1 exp}, we can see a slowly decaying envelope of the PL due to the transverse magnetic field (we don't see the plateau here since the magnetic field stops at 150 G). We can also see three deviations to the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.

In order to better discern the three features, we fitted the envelope by using a polynomial fit of order 4 on the smooth part of the data. We previously tried to use a fit based on the expected PL rate of the NV centers, by using the model in \citep{tetienne2012magnetic}, but the results were not satisfying.

Fig. \ref{CR VH exp}-b) shows the subtraction of the PL data by the envelope fit. We can clearly see three dips in PL around the values 20 G, 56 G and 122 G. We then computed the expected value of the magnetic field for the expected CR line thanks to the simulations shown in sec. \ref{sec simu}. For the VH$^-$ and War1 lines, we took the ZFS $D$ values from \citep{cruddace2007magnetic} and included the error bars. For the $^{13}$C-NV$^-$ lines, the error bars given in \citep{simanovskaia2013sidebands} are much smaller and we only plotted the expected value. We can see that the dips in PL match relatively well the predicted CR lines of $^{13}$C-NV$^-$, VH$^-$ and War1 respectively.

The experimental setup used to obtain this data is the standard confocal microscopy setup shown in appendix REF. The magnetic field here is generated with an electromagnet (EM), aligned along the [100] diamond axis with a precision $<1^\circ$.

The data shown here was acquired over 24 h. It should be noted however that a single-photon counting APD was used here, which considerably worsens the sensitivity of the measurement as discussed in [REF]. Using a photodiode and modulation of the magnetic field, we estimate that we could get similar signal to noise ratio within a few minutes.

To calibrate the magnetic field, we added a microwave field at a  fixed frequency $\nu > 2870\ \rm MHz$, and measured the EM voltage at which the PL dropped due to the $\ket{+1}$ state of the NV centers becoming resonant with the microwave tone. We could then use the values in Fig. \ref{121 vs 22 NV} to find the corresponding magnetic field. We varied the frequency in the range  $[2870,3150]\ \rm MHz$ by step of 2 MHz to get a calibration of the magnetic field in the $[0,150]\ \rm G$ region. Considering the uncertainty in the field angle, the NV ZFS $D$ value and the width of the PL drop, we estimate the final precision of the calibration to be within $\pm 1\ \rm G$.

\section{Attribution of the $^{13}$C-NV$^-$, VH$^-$ and War1 lines}

\subsection{The VH$^-$ and War1 lines}
\begin{table}[htbp]
\centering
\caption{\bf Zero-field splitting parameter $D$ for NV$^-$, VH$^-$ and War1}
\begin{tabular}{ccc}
\hline
$D_z$ estimation (MHz) & Cruddace's work\citep{cruddace2007magnetic} & Our work \\
\hline
NV$^-$ & 2872(7) & * \\
VH$^-$ & 2706(30) & 2694(5)  \\
WAR1 & 2466(60) & 2470(10) \\
\hline
\end{tabular}
  \label{table VH et War1}
\end{table}

The attribution of the PL dips at 56 and 122 G in Fig. \ref{CR VH exp} to CR between NV and respectively VH$^-$ and War1 comes from two main observations: the closeness in the ZFS values, and the CVD nature of the sample.

Table \ref{table VH et War1} shows the value and precision of the $D$ factor when measured through EPR \citep{cruddace2007magnetic}, and when computed from the data in Fig. \ref{CR VH exp}. In order to get to the $D$ value of VH$^-$ and War1, we had to assume that the spin structure was that of a spin 1 similar to the NV center, with a g-factor close to 2. Both of this assumptions are confirmed by the EPR measurements. We do find that that both $D$ measurement concord within their respective error bars.

The second reason comes from the nature of the sample used. We tried similar PL scans along the [100] axis on many samples with dense ensemble of NV centers. We found that the dip at 55 G was present on all CVD samples ($\sim$ 5 samples) used, and none of the HPHT samples ($\sim$ 20 samples). The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below. We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the process, and while we don't know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.

Finally, the apparent larger linewidth of the War1 line compared to the VH$^-$ one is also corroborated by the EPR measurements.

For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively, with a slightly stronger confidence for the VH$^-$ since more is known about it.

\subsection{The $^{13}$C-NV$^-$ lines}

\begin{figure}[h]
\centering
\includegraphics[width=.6\textwidth]{chapter 2/Figures/scan_100_Sumi4}
\caption{Change in PL as a function of a magnetic field scanned along the [100] axis for sample SUMI-4}
\label{scan sumi 4}
\end{figure}

The attribution of the $\sim 20\ \rm G$ dip to CR with $^{13}$C-NV$^-$ also come from two observations. First, the lines are present on every single sample with dense ensemble ($\gtrapprox 1$ ppm), regardless of the fabrication process. Fig. \ref{scan sumi 4} for example shows a PL scan along the [100] axis for sample Sumi-4, an HPHT sample with [NV]$\approx 5\ \rm ppm$ and natural abundance of $^{13}$C. I could unfortunately not try a sample isotopically enriched in $^{12}$C since those samples are generally reserved to low-nitrogen concentration ($< 10\ \rm ppm$), which does not yield enough NV centers to perform CR spectroscopy.

The second reason is the shape of the PL dip. The two other dips seem be formed of a single line and are reasonably well fitted with Gaussian curves. The 20 G dip however has a flat bottom which would indicate the presence of several overlapping lines, and Fig. \ref{121 vs 22 13C-NV} indeed predicts that there would be 4 CR lines between 18 and 25 G.

For this two reasons, we attribute the 20 G line to CR between $^{13}$C-NV$^-$ and ``normal" NV centers.

\section{Quantitative estimates of the NV-Dark spins CR}

We can extract some quantitative information about the dark spins by exploiting the CR spectrum shown in Fig. \ref{CR VH exp}.

\subsection{Dark spin Hamiltonian parameters}

First, as discussed in the last section, we can estimate the dark spin ZFS $D$ value with greater precision than EPR. While it would be in theory possible to measure other properties of the dark spins Hamiltonian - such as the $g$-factor - by changing the angle of the magnetic field, this would here quickly become intractable due to the increase in co-resonance conditions and the relative closeness of each of these transitions. DEER might prove more useful in this regard as only one class of NV centers at a time is probed.

\subsection{Dark spin linewidth}

The second quantitative estimate we can make is of the dark spin inhomogeneous broadening $T_2^*$: if we assume the change in PL to be proportional to the change in the NV lifetime $T_1$, which is true for small enough changes, then the linewidth of the PL dips in Fig. \ref{CR VH exp} is the convolution of the NV and dark spin spectrum. By deconvoluting the data with the NV spectrum, known for example from ODMR, we can then recover the dark spin spectral response as was done in \citep{hall2016detection}.

This procedure will be discussed in more details in the next chapter. We find here that the linewidth of VH$^-$ and War1 are respectively [REF] and [REF] MHz.

\subsection{Dark spin concentration}

Finally, the third tentative estimate we can make is that of the dark spins concentration. We can indeed calibrate the measurement by using the $^{13}$C-NV$^-$ line, of which we know the concentration in natural abundance diamond: [$^{13}$C-NV$^-$] $\approx$ 3.3 \% [NV]. We can then, knowing the NV concentration, try to estimate the concentration of the dark spins. There are several caveats though:

\begin{itemize}
\item Since the change in $T_1$ is proportional to the $T_2^*$ of the probed spins (eq. \ref{delta gamma 1}), we should compensate that by integrating over the entire feature, instead of simply looking at the maximum of the line.
\item We have to take into account the loss of contrast of the NV centers as the transverse magnetic field grows, as explained in sec. \ref{CR contrast}.
\item The $^{13}$C-NV$^-$ are not ``dark spins", and its CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1. We will see however in the next chapter that we can consider a fraction of the NV centers as dark spins, and that this fraction can be estimated. \citep{choi2017depolarization} found a ratio of $\sim 1/4$ dark NV centers for $\sim 3/4$ bright ones.
\end{itemize}

With all this considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):

For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose the dark to bright NV ratio value from \citep{choi2017depolarization} at 0.25.

With an estimated NV concentration [NV] $\approx\ \rm 4.6 ppm$, giving us a concentration $[^{13}\rm{C-NV}^-]\approx \ \rm 130 ppb$, we then find $[\rm VH^-] \approx 58\ \rm ppb$. Similarly, we find $[\rm VH^-] \approx 68\ \rm ppb$. Given the large uncertainty on some parameters, in particular the dark to bright spin ratio, these estimates are probably only precise within an order of magnitude.

If we consider that the sampled volume in this experiment is $< 100 \mu \rm{m}^3$, the total number of dark spins detected here is then $< 10^6$. Even with very conservative estimates, the technique employed here is still more sensitive than EPR by several orders of magnitude.

\section{Temperature dependence of VH$^-$ and War1}

\subsection{Mofifications on CVD samples}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 2/Figures/VH_CVD_chauffage}
\caption{Demodulated PL on samples CVD-C and CVD-B as a function of a magnetic field along the [100] axis. An additional oscillating magnetic field along the same direction was used to perform the lock-in detection. The VH$^-$ and War1 lines are taken from \ref{CR VH exp}}
\label{chauffage CVD}
\end{figure}

Given that the NV CR spectroscopy was both sensitive and relatively easy to setup, we decided in collaboration with the teams of Alexandre Tallaire at IRCP and Jocelyn Achard at LSPM to study the effect of annealing at high temperature on VH$^-$ and War1 \citep{ngambou2022improving}.

Fig. \ref{chauffage CVD} shows the results of this experiment. This time, a modulation of the magnetic field was used and the PL coming from the photodiode was demodulated through a lock-in amplifier, which explains the derivative shape of the lines. The fact that the background is not perfectly flat still comes from the transverse magnetic field. The same lock-in parameter were used on both plots so that the amplitudes should in theory be comparable.

Fig \ref{chauffage CVD}-a) shows the demodulated PL versus mganetic field amplitude along [100] on sample CVD-C, while \ref{chauffage CVD}-b) shows the same experiment on sample CVD-B. The fabrication and subsequent operations done on this two samples are detailed thoroughly in \citep{ngambou2022improving}, but the main difference between these two sample is that CVD-C was only annealed once after irradiation, for 2 hours at 800$^\circ$C, while CVD-B was annealed for 2 hours at 800$^\circ$C and then for 1 hour at 1000$^\circ$C and 1 hour at 1200$^\circ$C.

We can see that the VH$^-$ line in particular completely disappear on sample CVD-B, which we attribute to the additional annealing. This measurement also corroborate modifications in the infrared absorption spectrum, and an increase of the NV coherence time, associated with the annealing at higher temperature \citep{ngambou2022improving}.

The War1 line on the other hand is still present after the high temperature annealings, albeit slightly smaller. This information can give us some insights on the chemistry of the War1 defect. %Et en vrai je crois que c'était deja vu dans la thèse de Cruddace

\subsection{Modification on HPHT samples}

\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{chapter 2/Figures/VH_HPHT_chauffage}
\caption{Demodulated PL on samples SBST-C and SBST-B as a function of a magnetic field along the [100] axis. The plot of SBST-C is offset for clarity}
\label{chauffage HPHT}
\end{figure}

An unexpected side-study was done on HPHT samples. Indeed, the CVD samples used previously are grown by homoepithaxy on top of a diamond substrate. In this case, the substrate was a type 1B HPHT diamond, which was irradiated and annealed along the CVD layer. We Will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.

Fig. \ref{chauffage HPHT} shows demodulated PL scans with B $\parallel$ [100] on both samples. They both show a line at $B \approx 20\ \rm G$ corresponding to the $^{13}$C-NV$^-$ line, and none of them show a line at $B = 56\ \rm G$ corresponding to the VH$^-$ line. There is however a stark difference for the War1 line: SBST-C seems to show a weak line, while SBST-B shows a very clear line, with an amplitude more than 5 times greater than SBST-C.

We can also see that many weak lines on SBST-B seems to have disappeared in SBST-C, while an unidentified line at $B = 48\ \rm G$ seems to have grown similarly to the War1 line.

We first suspected that some of the War1 defects had been migrating from the CVD layer to the HPHT substrate with the high-temperature annealing, but not only is it very unlikely with the annealing time (1 hour) and substrate thickness (several mm), We also could not find a dependence on the War1 concentration with the depth in the substrate, whereas a diffusion would have given a higher concentration near the CVD layer. The War1 defects seem to have been created by the annealing.

These results suggest that, while some defects are removed by the high temperature annealing, such as the VH$^-$ and several unidentified defects in the HPHT substrate, some seem to be created by the annealing process, such as the War1 or the one associated with the line at $B = 48\ \rm G$ in the HPHT substrate.

\section{Conclusion and perspectives}

To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations. This detection can be way more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up. It is however not as versatile and should be seen as a complimentary tool for diamonds with dense NV ensemble.

We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan, much less noisy than the [111] scans for weak magnetic field.

Finally we have seen an application of the CR to detect optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.

There still remains a lot of open questions and there are several experiments that could be build upon these results. I will list some of them here.

\subsection{Unidentified CR lines}

There are still a large number of lines in CR spectra that have not been identified. While not every line necessarily correspond to a new defect - for example \citep{wunderlich2021magnetic} recently found lines corresponding to co-resonance between the $\ket{0} \to \ket{-1}$ transition of an NV center and the $\ket{-1} \to \ket{+1}$ of another NV center - the more likely hypothesis for each of these unidentified lines is still the presence of a new defect.

Such unidentified lines can be seen in Fig. \ref{chauffage CVD} and \ref{chauffage HPHT}. In particular we can see a line at 160 G in Fig. \ref{chauffage CVD}-a) and a line at 48 G in Fig. \ref{chauffage HPHT} that detach clearly from the background.

Further studies would be needed to find how common these lines are, and what could be there sources.

\subsection{Measuring the spin dynamics of VH$^-$}

We found that NV-VH$^-$ CR in CVD could produce a change in PL of $\sim 0.2 \%$. This kind of values are not too far from ODMR contrast with NV ensembles, which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin. The main issue comes from the fact that in order to measure the VH$^-$ spin states, it needs to be resonant with the NV center. Therefore, resonant microwave pulses with VH$^-$ will also be resonant with the NV centers, which will create an unwanted optical signal.

There are two solution to circumvent this problem: the first is to use a fast switching magnetic field in order to bring the VH$^-$ and NV centers in resonance when you want to polarize or readout the VH$^-$ state, and out of resonance when you want to coherently manipulate the VH$^-$. This means that the magnetic field needs to be switched faster than the spin dynamics you want to probe. This might be doable for the population dynamics $T_1$ but this is probably too slow to probe the $T_2$ time.

The other solution consists in using the spin-1 nature of VH$^-$: when a NV$^-$-VH$^-$ CR occurs, only one of the transitions of VH$^-$ and NV$^-$ is brought to resonance. The other one ($\ket{0} \to \ket{-1}$ for the VH$^-$) is out of resonance. The idea would therefore be to keep the NV $\ket{0} \to \ket{-1}$ transition and VH $\ket{0} \to \ket{+1}$ transition resonant, and to coherently manipulate the VH $\ket{0} \to \ket{-1}$ transition. While the polarization on this transition, and the back-action it would produce on the NV centers, is not as strong as if you were manipulating the $\ket{0} \to \ket{+1}$ transition, it should still produce a measurable signal.

\subsection{Isolating a single VH$^-$}

We estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$. This concentration is too high by about two orders of magnitude to isolate a single emitter with a diffraction-limited microscope. We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing. We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and optically isolate a single VH$^-$ center by scanning a confocal microscope with a fixed $B_{[100]}=56\ \rm G$ while monitoring the NV PL or $T_1$. It is however unclear to me whether this detection scheme would be sensitive enough to detect single dark spins.

\printbibliography
\end{refsection}
 \begin{refsection}

\chapter{NV-NV cross-relaxations : the fluctuator model}
In this chapter we will look at the phenomenon of cross-relaxations between NV centers, both from an experimental and theoretical point of view.

The presence of strong NV-NV dipolar interaction in dense ensemble ([NV] $\gtrsim 1\ \rm ppm$) proved to be both an obstacle for ensemble magnetometry \citep{zhou2020quantum} and an opportunity to explore and use the many-body interaction resulting from it \citep{zhou2020quantum, choi2017observation, kucsko2018critical, zu2021emergent}.

We will see here how a model of fast and slow relaxing centers, the fluctuator model \citep{choi2017depolarization}, can explain most of the experimental observations, and we will see some limitations of the model.

\section{Experimental observation of NV-NV CR}
Before we discuss the theoretical implications of NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
\subsection{NV-NV CR between nonequivalent NV centers}
\label{non_equi_valent_CR}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/NV-NV_non_equivalent}
\caption{a) Representation of two non equivalent spins resonantly coupled. One spin is aligned with the magnetic field and therefore well polarized, while the second spin is not. b) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B. c) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
\label{non equivalent NV-NV}
\end{figure}
NV-NV CR was first observed more than thirty years ago \citep{holliday1989optical, van1989cross}. The first observations were between non-equivalent NV centers, meaning that the two NV centers involved in the dipole-dipole coupling were not polarized equally.

This scenario can happen for instance when two NV centers from different classes see a different transverse (and longitudinal) magnetic field. Fig \ref{non equivalent NV-NV} illustrates this in the case where the magnetic field is parallel with one of the four classes: $\mathbf{B} \parallel [111]$

When $\mathbf{B} \parallel [111]$, as we discussed in the last chapter, one class sees no transverse field and is therefore always polarized. The three other (equivalent) classes on the other hand get more and more depolarized as the magnetic field amplitude is increased.

It turns out that there is a co-resonnance at B=592 G between the class parallel to $\mathbf{B}$ and the three other classes. This co-resonance is represented in Fig. \ref{non equivalent NV-NV}-b) by an orange circle.

Fig. \ref{non equivalent NV-NV}-c), which we already saw in the last chapter, shows the change in PL of an ensemble of NV centers as $\mathbf{B}$ is scanned along the [111] axis. For B=592 G, we can see a drop in PL also circled in orange. This drop is the result of the CR between NV centers from the class parallel to $\mathbf{B}$ and NV centers from the three other classes.

While a difference in polarization is needed to observe a population transfer, it is not enough to explain the drop in PL. This PL drop is due to the difference in brightness between the $\ket{0}$ and $\ket{+1}$ from the different classes involved. The PL contrast between the two states is higher for the class parallel to $\mathbf{B}$ than for the three other ones, which is another consequence of the transverse field.

The co-resonance between two different classes of NV centers with different magnetic field projection is a relatively rare event: it can only occur for the $\ket{0} \to \ket{+1}$ transition, and only for magnetic fields greater than 592 G.

%techniquement y'a les resonances 0->-1 et -1/+1 mais c'est le delbor, j'ai pas envie d'embrouiller plus que ca.

\subsection{NV-NV CR between equivalent NV centers}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/NV-NV_equivalent}
\caption{CR between equivalent NV centers for sample ADM-15-1. (a) Representation of equivalent NV centers: either two NVs from the same class, or two classes with the same projected magnetic field. (b) Magnetic setup used for the experiment: a permanent magnet is used to apply a bias magnetic field, and an electromagnet is used to add a variable magnetic field. (c) Simulation of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers as a function of the scanned magnetic field. The transitions were computed based on several ODMR spectra. The red dotted lines correspond to inter-class resonance (d) Change in the PL of the NV centers ensemble as a function of the scanned magnetic field. The inter-class resonances are reported here.}
\label{equivalent NV-NV}
\end{figure}

More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, jarmola2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there were cross-relaxations even between equivalent NV centers.

I refer by equivalent NV centers to NV centers with the same spin Hamiltonian. This means either NV center from the same class, or NV centers from different classes with the same projection of the magnetic field along their axis. Equivalent NV centers are by nature resonant with each other, and therefore susceptible to CR.

Fig. \ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV CR. The key to observe NV-NV CR is to bring some NV centers in and out of resonance with other NV centers, which we can do with the different NV classes. To do so, we need to add an initial bias magnetic field in addition to the scanned magnetic field, as represented in Fig. \ref{equivalent NV-NV}-b). %When doing so, the magnetic field felt by the diamond won't have a fixed orientation during the scan, and when the magnetic field crosses certain crystalline planes (detailed below), its projection on two or more classes of NV centers will be the same.

Fig. \ref{equivalent NV-NV}-c) shows the transition frequencies of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers on sample ADM-15-1, an HPHT micro-diamond with $[\rm NV]\sim 3\ \rm ppm$. There are three magnetic field values $B_{\rm scan} \approx$ 12, 42 and 77 G for which there is an inter-class resonance. Each of these resonances correspond to a ``geometric" resonance where the $B$ field is in a symmetry plane between the two classes, meaning that the two resonant NV classes are indeed equivalent.

Finally, Fig. \ref{equivalent NV-NV}-d) shows the PL contrast as the electromagnet field is scanned. There are 3 very clear dips when $B_{\rm scan} \approx$ 12, 42 and 77 G which coincide with inter-class resonances. We will see later that this PL dip is associated with a decrease of the NV centers spin $T_1$ time from both classes.

This observation, and similar ones done by many groups \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization, akhmedzhanov2017microwave, giri2018coupled}, seem to indicate that there are CR between equivalent NV centers. This is however incompatible with our previous assumptions that equivalent NV centers are equally polarized and bright. To understand this phenomenon, we will need to introduce new hypotheses.

\section{NV inhomogeneity and the fluctuator model}

\subsection{CR in an inhomogeneous NV bath}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/inhomogene_2_spins}
\caption{a) Schematics of the model: we consider a four level system consisting of 2 levels of 2 NV centers resonantly coupled through the flip-flop rate $\Gamma_{\rm dd}$. These two NV centers are each pumped in their respective $\ket{0}$ state via the green laser with a rate $\Gamma_{\rm las}$, and have their own relaxation rate $\Gamma_1$ and $\Gamma_1^*$ respectively. b) Average population of the two NV centers in the $\ket{0}$ state in the steady state as a function of $\Gamma_{\rm dd}$ and $\Gamma_1^*$. The steady state was computed by solving a classical rate equation where I fixed $\Gamma_{\rm las}=1000\ \rm Hz$ and $\Gamma_1=300\ \rm Hz$.}
\label{inhomogene}
\end{figure}
A likely explanation to the observed NV-NV CR is that the NV centers are not all equivalents.

Indeed, if we assume that, for instance, the relaxation rate $\Gamma_1=1/T_1$ is not strictly the same for each NV centers, but instead follows a certain distribution $\rho(\Gamma_1)$, then CR between NV centers with different $\Gamma_1$ can explain the observations in Fig. \ref{equivalent NV-NV}.

Fig. \ref{inhomogene} illustrates this case. Our model here consists of two NV centers resonantly coupled with two separate relaxation rates $\Gamma_1$ and $\Gamma_1^*$. To complete the model, we add the rate $\Gamma_{\rm las}$ which represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins. We only consider here the incoherent dynamics of the population, modeled with rates. We also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.

By finding the steady state of these rate equations, we can compute the final population in the $\ket{0}$ for both spins, which we will assume to be proportional to the total PL. Fig. \ref{inhomogene}-b) shows this $\ket{0}$ population as a function of the flip-flop rate $\Gamma_{\rm dd}$ for various values of $\Gamma_1^*$.

We can see that when $\Gamma_1^*=\Gamma_1$, as expected, the PL is not modified by the coupling of the two spins regardless of the coupling strength. When $\Gamma_1^*>\Gamma_1$ however, not only is the starting PL slightly lower since NV$_2$ is less polarized, but the PL now decreases as we increase the coupling strength. This effect, which is at the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is important.

\subsection{Presentation of the fluctuator model}
Choi et al. in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism. We are assuming that the fluctuator relaxation rate, $\gamma_f$ is much greater than the optical pumping rate $\Gamma_{\rm las}$. The fluctuators are therefore unpolarized and effectively act as dark spins, similarly to what we discussed in the last chapter.

We should note here that, while this simplifications seem a bit extreme, the fluctuator model is more than a toy model. There are good evidence of the presence of these dark NV centers in dense NV ensemble, as will be discussed below.

We will consider the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers. With this assumption, we can compute the modification caused by the fluctuators on the rest of the NV lifetimes.

Other possibilities to explain the change in the relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or superradiance, but they were not considered to be viable explanations \citep{choi2017depolarization}.
%\begin{pmatrix}
%0.33&0&0 \\
%0&0.33&0 \\
%0&0&0.33
%\end{pmatrix}

\subsection{Single NV center coupled to a single fluctuator}
We will follow here the notations and calculation steps in \citep{choi2017depolarization}. To compute the depolarization induced by the fluctuator bath the NV centers, we should first consider the interaction between a single NV center and a fluctuator. Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.

We will first decompose the dipole-dipole Hamiltonian in a radial and angular part:
\begin{equation}
\mathcal{H}_{\rm dd} \approx - \frac{J_0}{r^3} \left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}+qS_z^1 S_z^2 \right] + h.c. ,
\end{equation}
where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative position and orientation of the two dipoles.

We then introduce the dimensionless number $\eta$ defined as:
\begin{equation}
\label{eq eta}
\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
\end{equation}
where $\gamma_f$ is the fluctuator decay rate for each possible channel between $\ket{0}, \ket{-1}$ and $\ket{+1}$. This number $\eta$ encapsulates both the geometry of the problem (through $g$ and $h$) and the resonance condition between the NV center and the fluctuator.

We can note that the spectral response of a single fluctuator, for which we only consider the broadening caused by the decay rate $\gamma_f$, is a Lorentzian of half width $2\gamma_f$. This is different from the usual lifetime limited optical spectra which are Lorentzian of half width $1/2T_1=\Gamma_1/2$. The difference comes here from the fact that we consider thermalization going both ways (i.e. $\ket{e} \to \ket{g}$ and $\ket{g} \to \ket{e}$ instead of only $\ket{e} \to \ket{g}$) and that each state of the fluctuator can decay into two other states.

Finally, the additional depolarization rate induced by the fluctuator on the NV center reads:
\begin{equation}
\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right) \frac{\eta^2}{\gamma_f}.
\end{equation}
Compared to eq. [REF], the dependency on the inhomogeneous broadening of the spins $1/T_2^*$ is hidden in the $\eta$ factor, and we have the additional broadening $\gamma_f$ coming from the fluctuators very short lifetime.

\subsection{Ensemble of NV centers coupled to a bath of fluctuators}

We now need to compute the average depolarization caused by the ensemble of fluctuators on the ensemble of NV centers.

We consider each fluctuator as in independent depolarization source, therefore the depolarization on a single NV center reads: 
\begin{equation}
\gamma=\sum_i \gamma_s(\mathbf{r}_i).
\end{equation}
We then want to compute the distribution $\rho(\gamma)$ defined as:
\begin{equation}
\rho(\gamma)=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma \right).
\end{equation}
To do so, we need to determine the distribution of the fluctuators positions and orientations$\{r_i\}$. Assuming that they are homogeneously distributed in the bulk of the material, we find:
\begin{equation}
\rho(\gamma)=\frac{e^{-1/(4\gamma T)}}{\sqrt{4\pi \gamma^3 T}},
\end{equation}
where the time constant $T$ was introduced and is defined as:
\begin{equation}
\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f},
\label{eq 1/T}
\end{equation}
with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{\eta}$:

\begin{equation}
\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta.
\end{equation}

Finally, the polarization dynamics from the ensemble of NV centers can be computed:
\begin{equation}
P(t)=\int_0^\infty \rho(\gamma)\, e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}.
\end{equation}

In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential, unlike the phonon-induced depolarization which is purely exponential.

It should be noted that the stretched exponential nature of the decay is not proper to the fluctuator model. \citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath. The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).

\subsection{Computation of $\bar \eta$}
\label{sec computation eta}
The parameter $\bar \eta$ in eq. \ref{eq 1/T} is of crucial importance since it encompass the dependence of $1/T$ in both the magnetic field and the inhomogeneous broadening $T_2^*$ of the NV centers and the fluctuators.

We will start by separating the factor $\eta$ given in eq. \ref{eq eta} into a a geometric and a spectral part:
\begin{align}
\eta^2&=G^2(\psi_{NV},\psi_f,\theta,\phi) R^2(\omega_f,\omega_{NV}), \\
G^2&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
R^2(\omega_f,\omega_{NV}) &= \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
\end{align}

$G$ is a function which depends purely on the various angles of the problem (angles of the relative position of the NV center and the fluctuator, angle of the NV and fluctuator $z$ axis), while $R$ represents the resonance condition between the NV and the fluctuator.

We will now average this $\eta$ factor for an ensemble of fluctuators and NV centers:
\begin{align*}
\bar \eta &= \int \rm{Prob}(\eta) \abs{\eta} d\eta, \\
&= \left(\int \rho(\psi_{NV},\psi_f,\theta,\phi) |G(\psi_{NV},\psi_f,\theta,\phi)| d\psi_{NV} d\psi_f d\theta d\phi \right) \\
&\, \times \left( \int \rho(\omega_{NV}) \rho(\omega_{f}) |R(\omega_f,\omega_{NV})| d\omega_{NV} d\omega_{f} \right), \\
&= \bar G \cdot \bar R,
\end{align*}
where we introduced $\bar G$ and $\bar R$, the averaging of $|G|$ and $|R|$.

The computation of $\bar G$ will be further discussed in sec. \ref{sec quantitative T1} and in appendix [REF]. We will focus here on $\bar R$.

To perform the computation, we need to know the distributions $\rho(\omega_{NV})$ and $\rho(\omega_{f})$. We can experimentally measure $\rho(\omega_{NV})$ though ODMR, and we can reasonably expect $\rho(\omega_{f})$ to follow the same probability.

We unfortunately cannot analytically solve $\bar R$ for standard distributions of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ (either Gaussians or Lorentzians). We can however solve $\overline{R^2}$, which is relatively close to $(\bar{R})^2$ as long as $2 \gamma_f \gg 1/T_2^*$.

For a Lorentzian profile of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ defined as:
\begin{align*}
\rho(\omega_{f})&=\frac{1}{\pi \Gamma_f} \frac{1}{1+ \left(\frac{\omega_f-\omega^0_f}{\Gamma_f}\right)^2} \\
\rho(\omega_{NV})&=\frac{1}{\pi \Gamma_{NV}} \frac{1}{1+ \left(\frac{\omega_{NV}-\omega^0_{NV}}{\Gamma_{NV}}\right)^2},
\end{align*}

where $\Gamma_{f}=1/T_2^*(\rm fluctuator)$ and $\Gamma_{NV}=1/T_2^*(\rm NV)$ are the inhomogeneous broadening of the NV and fluctuator ensemble, and $\omega^0_{NV}$ and $\omega^0_{f}$ are the central angular frequencies of the NV and fluctuator ensemble. We then find that:

\begin{align*}
\overline{R^2}&= \int \rho(\omega_{NV}) \rho(\omega_{f}) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} d\omega_{NV} d\omega_{f} \\
&=\frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}} \cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
\end{align*}

In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f \gg 1/T_2^*$) in eq. \ref{eq 1/T}, we find that:
\begin{equation}
\label{eq 1/T avec T2*}
\frac{1}{T} \propto \frac{1}{2\gamma_f + \Gamma_f + \Gamma_{NV}}.
\end{equation}

If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\overline{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian ans a Gaussian, which does not have a clear analytic solution.

Eq. \ref{eq 1/T avec T2*} is subject to several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case. Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$.

 \subsection{Possible microscopic origin of the fluctuators}

We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}. It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.

This conclusion was based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions, and also to the resonant energy transfer between molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through dipole-dipole interaction to non-excited acceptors, and where the stretched exponential dynamics of the donors was first observed \citep{forster1949experimentelle}.

In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap. In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]). Both of these phenomena could lead to fast spin relaxation, the tunneling because the spin hopping can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction strength $J$ by the crystal phonons \citep{honig1960electron}. While the non-contact dipole-dipole interaction is also modulated by the phonons, this effect would be too small to account for the fast depolarization observed here.

Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization. To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate between neighboring sites of $\sim 10\ \rm ns$. In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.

\section{Experimental investigation of the fluctuator model}
We will now show experimental results related to the predictions of the fluctuator model.
\subsection{The stretched exponential lifetime}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/Stretch_vs_pas_stretch}
\caption{T1 measurement following the protocol described in [REF]. The best exponential and stretched exponential fits are given each time. a) Sample ADM-15-2 with B=0. The optimal $T_1$ values are $T_1^{\rm stretch}=150\ \mu \rm s$ and $T_1^{\rm exp}=410\ \mu \rm s$. b) Sample CVD-pink with B$\neq$0 and a single class probed. The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$}
\label{stretch_or_not_stretch}
\end{figure}
$T_1$ measurements with dense ensemble of NV centers and when many classes are resonant with each other (typically when B=0) indeed tend to show stretched exponential profiles.

Fig. \ref{stretch_or_not_stretch} shows an example of that: in Fig. \ref{stretch_or_not_stretch}-a) the $T_1$ measurement comes from sample ADM-15-2 at B=0. The profile is clearly better fitted by a stretched exponential than a regular exponential, and the timescale found $T_1^{\rm stretch}=150\ \mu \rm s$ is much shorter than the expected phonon limite lifetime of a few ms.

On the other hand, Fig. \ref{stretch_or_not_stretch}-b) shows a $T_1$ measurement coming from sample CVD-pink for B$\neq$0. This time the profile is more exponential, and the value $T_1^{\rm exp}=4.1\ \rm ms$ is coherent with a phonon-limited lifetime. Even though the sample CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.

\subsection{Fitting $T_1$ profile in the general case}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/various_fit_formulae}
\caption{Different fitting procedure on two $T_1$ measurements done on sample ADM-150-1 for $B=0$ and $B\neq0$. \\ a) Exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph})$. \\ b) Stretched exponential fit $S(\tau)=\exp (-\sqrt{\tau/T_1^{\rm dd}})$. \\ c) Bi-exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph} -\sqrt{\tau/T_1^{\rm dd}})$. \\ d) Bi-exponential fit with fixed $T_1^{\rm ph}=5\ \rm ms$: $S(\tau)=\exp (-\tau/5 -\sqrt{\tau/T_1^{\rm dd}})$.}
\label{various_fit_formulae}
\end{figure}

\begin{table}[htbp]
\centering
\caption{Fitting parameters for Fig.
 \ref{various_fit_formulae}}
 \label{fitting table}
\begin{tabular}{c|cc|cc}
\toprule
Figure &  \multicolumn{2}{c}{$B=0$} & \multicolumn{2}{c}{$B\neq0$}\\
\midrule
{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms)&  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)\\
Fig. a) & 0.95 & * & 2.64 & * \\
Fig. b) & * & 0.32 & * & 1.03  \\
Fig. c) & 6.75 & 0.44 & 4.22 & 7.19 \\
Fig. d) & \textbf{5} & 0.51 & \textbf{5} & 4.65 \\
\bottomrule
\end{tabular}

Bold characters indicate parameters that were arbitrarily fixed.
\end{table}

For many samples however, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$ is of the same order as the phonon limited lifetime $T_1^{\rm ph}$. Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, forcing us to include both aspects if we want a unique fitting formula.

Fig. \ref{various_fit_formulae} shows four different fitting procedure on the two same experimental data. The data here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$). The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.

The fits used here are either purely exponential or purely stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$. The values of the different fitting parameters used is reported in Table \ref{fitting table}.

We can see that neither the purely exponential nor stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is poorly fitted by the stretched exponential. The protocols that include both exponential and stretched lifetimes correctly fit both curves. In the case of Fig. \ref{various_fit_formulae}-d), we arbitrarily fixed $T_1^{\rm ph}=5\ \rm ms$ since this is the value we typically measure on samples with low NV density, and we expect that the phonon-limited exponential lifetime is not modified by the NV concentration.

Since the protocol where $T_1^{\rm ph}$ was fixed and the one where it wasn't both correctly fit our data, we decided to use the protocol where $T_1^{\rm ph}$ was fixed because there is one less free parameter in the fitting function, and the values we obtain on $T_1^{\rm dd}$ can be directly compared.

We should note however that the values of $T_1^{\rm dd}$ we obtain are not absolute. Indeed, in the case of Fig. \ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$. While this technique of measuring $T_1^{\rm dd}$ is useful to compare different cases, because it is fairly sensitive, it is not an exact measurement of the dipole-dipole induced lifetime.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/betas}
\caption{a) Same two measurements as Fig. \ref{various_fit_formulae} with the fitting formula $S(\tau)=\exp ((-\tau/T_1)^{\beta})$. The fitting parameters are $\beta=0.58$ and $T_1=0.46\ \rm ms$ for $B=0$, and $\beta=0.80$ and $T_1=2.15\ \rm ms$ for $B\neq0$. b) Optimal $\beta$ parameter found for each $T_1$ measurement as a function of the external magnetic field (still on sample ADM-150-1).}
\label{betas}
\end{figure}

Finally, another fitting procedure is shown in Fig. \ref{betas}. This time we are using a single stretched exponential with an arbitrary stretch factor $\beta$. Fig. \ref{betas}-b) shows the optimal $\beta$ parameter as a function of the external magnetic field. We can then confirm that the $T_1$ profile gets closer to a stretched exponential ($\beta=0.5$) when $B$ goes to 0.

Even though this method also yields satisfying fits, it needs two free parameters to work ($\beta$ and $T_1$), so we decided to go with the method of Fig. \ref{various_fit_formulae}-d) instead.

\subsection{The fluctuators linewidth}

\subsubsection{The fluctuators spectral response}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/fluctuator linewidth}
\caption{Measurement of the fluctuators linewidth on sample ADM-150-2 using the same setup as in Fig. \ref{equivalent NV-NV}. a) Simulation of the $\ket{0} \to \ket{-1}$ transition frequency for the four classes of NV centers and actual frequencies of the two central classes from ODMR measurements. b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.78\ \rm MHz$. The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor $\sqrt{2}$}
\label{fluct linewidth}
\end{figure}

One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators spectral response. Indeed, as we will see, the fluctuator lifetime is so short that it significantly broadens the fluctuators spectral response compared to standard NV centers.

The fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in \citep{hall2016detection}.

Fig. \ref{fluct linewidth} shows an experiment similar to the one presented in Fig. \ref{equivalent NV-NV}. A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers, but this time measuring the spin decay rate instead of the PL.

Fig. \ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field, and the experimentally measured central frequencies of the two classes of interest when the two lines can be clearly separated. Having a precise measurement of the detuning between the two classes $\Delta \nu$ is key in this experiment to get a precise value of the fluctuator linewidth.
%Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of their two central lines to get a precise measurement of the Larmor frequency. In the region where the two ODMR lines overlap we have to use a linear regression to get the Larmor frequency of each class.

Fig. \ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $\Gamma_1^{\rm dd}=1/T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes. For each magnetic field value, a $T_1$ profile was recorded and fitted following the procedure described in Fig. \ref{various_fit_formulae}-d) where we again fixed $T_1^{\rm ph}=5\ \rm ms$. We also show the ODMR spectrum of a single NV class for comparison.

We can clearly see that the $\Gamma^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}. If we assume that both NV classes spectral response are gaussians of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$. Fig. \ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $\sqrt{2}$, which is still much narrower than the $1/T_1^{\rm dd}$ line.

We attribute this broadening to the fluctuators lifetime. Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator, but for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.

\subsubsection{The fluctuators lifetime}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{chapter 3/Figures/deconvolution}
\caption{Blue curve $1/T_1^{\rm dd}$ data presented in Fig. \ref{fluct linewidth}. Red curve: deconvolution of the blue curve the  by the ODMR line shown on the same figure. Green curve: deconvolution of the red curve the  by the ODMR line. All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38 MHz}
\label{deconvolution}
\end{figure}

We now want to extract the fluctuators lifetime $T_1^f=1/\gamma_f$ to get further information on their potential nature.

To do this we first assume that the dipole induced decay rate is proportional to the overlap of the spectral responses of the NV centers and fluctuators $S_{NV}$ and $S_f$. We will take $S_{NV}( \omega)$ and $S_{f}( \omega)$ as functions centered in $\omega=0$, and consider that the spectral responses are shifted by $\omega^0_{NV}$ and $\omega^0_{f}$ respectively, the central frequencies of each group. We then have:
\begin{align*}
\frac{1}{T_1^{\rm dd}} &\propto \int S_{NV} (\omega-\omega_{NV}^0) S_f (\omega-\omega_{f}^0) d\omega \\
&= \int S_{NV} (\omega-\Delta \omega) S_f (\omega)
 d\omega \\
 &= (S_{NV}*S_{f})(\Delta \omega),
\end{align*}

where $\Delta \omega=\omega_{NV}^0-\omega_{f}^0$.

Using the approximation $\bar {R ^2} \approx (\bar{R})^2$ detailed in sec. \ref{sec computation eta}, we can then decompose the fluctuator spectral response between the lifetime contribution and the inhomogeneous broadening:
\begin{equation}
S_{f}(\omega)=\left(S_f^*(\omega)\right)*\left(\frac{1}{1+\left(\frac{\omega}{2 \gamma_f}\right)^2} \right),
\end{equation}

where $S_f^*$ is the spectral broadening caused by the inhomogeneous distribution of the fluctuators, which we will suppose to be equal to $S_{NV}$. With these assumptions, we are now ready to extract $\gamma_f$ from the data of Fig. \ref{fluct linewidth}.

The first way to do this, as was done for NV-P1 CR in \citep{hall2016detection}, is to deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm. The result of the deconvolution is shown in Fig. \ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve). Fitting with Lorentzians, we find a final value $\Gamma_2 =2\gamma_f= (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=49.9\ \rm ns$.

Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\frac{\sqrt{\ln (2)}}{\pi T_2^*}$ and a Lorentzian of HWHM $f_L=\frac{2 \gamma_f}{2 \pi}$:
\begin{align*}
\Gamma_1^{\rm dd}&\propto \mathcal{G}(f_G)*(\mathcal{L}(f_L)*\mathcal{G}(f_G)) \\
&\propto \mathcal{L}(f_L)*\mathcal{G}(\sqrt{2} f_G),
\end{align*}
where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{ida2000extended}:
\begin{equation}
f = [f_G^5 + 2.69269 f_G^4 f_L + 2.42843 f_G^3 f_L^2 + 4.47163 f_G^2 f_L^3 + 0.07842 f_G f_L^4 + f_L^5]^{1/5}.
\end{equation}

With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, we find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 48.7\ \rm ns$, a value consistent with the one found with the deconvolution method.

Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48.2\ \rm ns$, a value surprisingly similar to the one we got.

\section{Resonance between several NV classes}

The main tool to probe NV-NV CR is to use the co-resonances between the NV classes. While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and supposedly resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances. We will discuss here the various scenarios where NV classes overlap, and the quantitative impact it has on the centers' spin $T_1$.

\subsection{Geometric conditions for inter-class resonance}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{chapter 3/Figures/Resonance_geometries}
\caption{ODMR spectra on sample ADM-15-3 as a function of the magnetic field orientation. The field amplitude is $\sim 60\ \rm G$. Red numbers represent the number of resonant classes for each ODMR line. The cubes represent the diamond unit cell, and the red planes/arrow the possible orientations of the magnetic field}
\label{ODMR_geometries}
\end{figure}

There are four possibilities to obtain inter-class resonance (outside of the non-equivalent CR discussed in sec. \ref{non_equi_valent_CR}). These possibilities are:
\begin{itemize}
\item $\mathbf{B} \in \{110\}$. We refer by $\{110\}$ as any plane orthogonal to a $\langle 110 \rangle$ or equivalent direction. In this magnetic configuration, two classes are at resonance and the two other ones are spectrally isolated.
\item $\mathbf{B} \in \{100\}$. In this magnetic configuration, the four classes form two pairs of co-resonance.
\item $\mathbf{B} \parallel \langle 111 \rangle$. In this configuration discussed in the last chapter, one class is aligned with the magnetic field and is spectrally isolated while the three other form a resonant triplet.
\item $\mathbf{B} \parallel \langle 100 \rangle$. In this configuration also discussed in the last chapter, the four classes are resonant.
\end{itemize}
We should add to this list the case $\mathbf{B}=0$ where all four classes are also resonant, and any other possibility (which I will call random $\mathbf{B}$) where all four classes are spectrally separated.

An ODMR spectrum and a representation of the magnetic field in the diamond unit cell for each of these six possibilities is shown in Fig. \ref{ODMR_geometries}.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/carte}
\caption{PL contrast as a function of the magnetic field polar and azimuthal angle with respect to the [100] axis on sample CVD-pink. The field amplitude is $\abs{B} \sim 115 G$. The locus of the magnetic field in specific planes is noted by red dashed lines.}
\label{Carte}
\end{figure}

This relation between NV-NV class resonances and the magnetic field being in specific crystalline planes or directions means that we can determine the crystal main axes simply by PL experiments.

Fig. \ref{Carte} shows a map where we simply monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer. We have also reported the locus of $\mathbf{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].

We know that when $\mathbf{B}$ belongs to any of these four planes, there is a co-resonance between at least two NV classes. This co-resonance leads to an increase in NV-NV CR and a faster depolarization of the spins involved, which ultimately leads to a drop in PL. We can indeed see in fig. \ref{Carte} that there is clear correlation between the loci of $\mathbf{B}$ and drops in PL. We can also see that the PL is at its lowest when $\mathbf{B}\parallel [001]$, which corresponds to a four class degeneracy.

\subsection{Quantitative modification of $T_1$}
\label{sec quantitative T1}
\begin{table}[htbp]
\centering
\caption{Theoretical and experimental values of $\Gamma_1^{\rm dd}=1/T_1^{\rm dd}$ for various magnetic field configurations, expressed as a function of the value found for an isolated class. The experimental values were measured on sample ADM-15-3 using the protocol of Fig. \ref{various_fit_formulae}-d) with $T_1^{\rm ph}= 5\ \rm ms$.}
 \label{T1 champ mag}
\begin{tabular}{c|cc}
\toprule
$\Gamma_1^{\rm dd} (\mathbf{B})$ &  Theory & Experimental \\
\midrule
random $\mathbf{B}$ (1 class) & $\Gamma_0^{\rm th}$ & 1.53 ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
$\mathbf{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & 5.20 $\Gamma_0^{\rm exp}$ \\
$\mathbf{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & 4.22 $\Gamma_0^{\rm exp}$ \\
$\mathbf{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & 9.07 $\Gamma_0^{\rm exp}$ \\
$\mathbf{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & 11.3 $\Gamma_0^{\rm exp}$ \\
$\mathbf{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & 15.4 $\Gamma_0^{\rm exp}$ \\
\bottomrule
\end{tabular}

($^*$) The case $\mathbf{B}=0$ will be detailed in the next chapter. The theoretical value is a lower bond.
\end{table}

The fluctuator model allows quantitative predictions of the change in the spin lifetime, thanks to eq. \ref{eq 1/T}. There are however several unknown in this equation, such as the fluctuator density and lifetime which can be hard to estimate. It is however relatively easy to compare the predictions of eq. \ref{eq 1/T} for various magnetic field configurations.

What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators of the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2,3 or 4 classes are co-resonant as a function of $\Gamma_0$. Doing so allows to normalize $n_f$ and $\gamma_f$ in eq. \ref{eq 1/T} and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.

$\bar G$ can be analytically or numerically computed for each of the magnetic configurations listed before, as detailed in appendix [REF]. This allows us to predict the value of the spin decay rate in every situations as a function of the decay rate of an isolated class.

Table \ref{T1 champ mag} shows these predicted values, along with experimental ones on sample ADM-15-3. The experimental values were obtained with the protocol described in Fig. \ref{various_fit_formulae}-d) with a double exponential fit where we fixed again $T_1^{\rm dd}=5\ \rm ms$.

The theoretical values are in a good qualitative agreement with the experimental ones, both ranking the different scenarios in the same order. The quantitative predictions of the theory however are rather poor, being off by more than a factor of 3 for the shortest lifetimes.

It should be noted that the experimental values were heavily influenced by the choice of the parameter $T_1^{\rm ph}$. Using for example $T_1^{\rm ph}=2.5\ \rm ms$ gave results closer to the theoretical ones. Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded an increased depolarization rate by a factor $\sim 4$, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one. We can also see in Fig. \ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration leads to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, 4 times smaller than the predicted factor of 10.

The failure of the fluctuator model to accurately predict the changes in $T_1$ is one of the reasons that led us to believe that the model is incomplete.

\section{Conclusion and perspectives}

In conclusion, we have seen in this chapter that dense NV ensemble present ``anormal" CR between the NV centers, and we have shown how the fluctuator model developed by Choi et al. in \citep{choi2017depolarization} can explain this behavior. The fluctuator model not only explains the NV-NV CR, but also correctly predicts the stretched exponential lifetime profile. It also correctly predicts that the fluctuators have a broader spectral response than the other NV centers. Nevertheless, the model does not describe perfectly every aspects of NV-NV relaxation, and there still are many aspects of this problem, both theoretical and experimental, to explore.

In what follows I will describe what I think would be interesting investigation points to further understand the NV-NV relaxation and the fluctuator model.

\subsection{Limitations of the fluctuator model}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 3/Figures/Anormal_T1}
\caption{$T_1$ measurements on sample ADM-15-4, fitted with pure exponential or stretched exponential. a) $T_1$ of a single isolated class when $\mathbf{B} \neq 0$. The optimal fit parameters are respectively $T_1^{\rm exp}=1.43\ \rm ms$ and $T_1^{\rm stretch}=0.56\ \rm ms$. b) $T_1$ of the four degenerate classes when $\mathbf{B}=0$. The optimal fit parameters are respectively $T_1^{\rm exp}=0.15\ \rm ms$ and $T_1^{\rm stretch}=0.05\ \rm ms$.}
\label{anormal T1}
\end{figure}

I already mentioned that the discrepancy between the theoretical and experimental $T_1$  values were an issue of the current fluctuator model, although the difference could come in part from the $T_1$ fitting procedure. Another, more glaring problem is the fact that some $T_1$ profile are not stretched exponential.

Fig. \ref{anormal T1} shows $T_1$ measurement on sample ADM-15-4, either on a single isolated class ,where the dipole-dipole interactions are at their weakest, or on all four resonant classes at $\mathbf{B}=0$, where the interactions are at their strongest.

For the single isolated class, the $T_1$ profile seems to be closer to a stretch exponential, which is coherent with the fact that the lifetime $T_1 \sim 1\ \rm ms$ is already smaller than the expected phonon limited lifetime $T_1^{\rm ph} \sim 5\ \rm ms$. However when $\mathbf{B}=0$, although the lifetime gets $\sim$ 10 times shorter, the profile becomes clearly exponential.

The spin dynamics in this sample is still dominated by resonant dipole-dipole interactions: the more classes are brought to resonance, the shorter the spin $T_1$. But here, the profile gets closer to a pure exponential when $T_1$ decreases, which is in complete contradiction with the fluctuator model.

This particular sample was a 15 $\mu$m micro-diamond, which could have a role in this observation since we know that spins near the surface have a reduced $T_1$ \citep{rosskopf2014investigation}, and micro-diamonds have proportionally more near surface spins. But other diamonds from the same batch, such as the one measured in Fig. \ref{stretch_or_not_stretch}, show clearly stretched exponential profile in the same conditions, and meanwhile some bulk diamonds show very short non-stretched exponential lifetimes, similar to the present sample.

It seems therefore that the current fluctuator model does not correctly describe the spin dynamics of every NV-dense samples. Looking at what was previously done for P-doped SI \citep{vugmeister1978spin} or for resonant energy transfer between molecules \citep{yokota1967effects}, the fluctuator model could probably be improved by taking into account the finite lifetime of the fluctuators (i.e. the saturation of the NV-fluctuator CR), and the spin diffusion between NV centers (the fluctuator model only considers NV-fluctator interactions and not NV-NV interactions).

\subsection{Investigation of the fluctuators nature}
While the hypothesis of fluctuators being clusters of NV centers or NV-other spins is solidified by the precedents in P-doped Si \citep{honig1960electron} or F centers in KCl \citep{warren1964spin}, there are still many unknowns on the exact nature of the fluctuators. There are at least two aspects that I think merit to be investigated.

The first one would be to study the fluctuator spectral profile on various samples showing NV-NV CR. In particular it would be interesting to know whether this profile is always Lorentzian, and what its linewidth is. In the sample I studied on Fig. \ref{fluct linewidth}, I found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough my sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 45 ppm$ sample used in \citep{choi2017depolarization}. If $\gamma_f$ was found to be somewhat of a constant among all samples showing NV-NV CR, this would give us a lot of information on the potential nature of the fluctuators. Studying the temperature dependency of $\gamma_f$ could also give us information on whether the depolarization mechanism is due to tunneling or modulation of the $J-$coupling by the phonons.

The other aspect would be to identify potential defects that could form fluctuators when paired with NV centers. The main candidate for this would be substitutional nitrogen (P1 or P1$^+$ centers) since they are the most abundant electronic spin species in the diamond we use. The ideal way to confirm this hypothesis would be to compare the spin $T_1$ for  different samples with the same [NV] concentration, but different [P1] concentration. If indeed P1 centers play a role in the presence of fluctuators, we would expect a shorter NV $T_1$ time in the sample with high P1 concentration.

\printbibliography
\end{refsection}
 \begin{refsection}

\chapter{NV-NV CR under transverse or low fields: application to magnetometry}

In the last chapter we saw how dense ensemble of NV centers have an intrinsic depolarization mechanism mediated by the NV-NV dipole interaction. In this chapter we will see how this feature can be exploited to perform magnetometry.

Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}. We will focus here on the low to zero magnetic field region, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the more complex physics of the NV centers under low magnetic field.

This chapter is constructed as follows: We will first describe the physics of NV centers under low or transverse magnetic field (the physics is similar in both cases), we will then show both theoretically and experimentally that the NV-NV CR is modified for low and transverse magnetic field. We will identify different mechanisms by which the relaxation is modified and we will try to quantify their relative importance.

We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field. We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols. Finally we will give some perspectives on how to improve LFDM and on potential applications.

The results presented in this chapter are based in large part on the article \citep{pellet2022spin}.

\section{NV spin Hamiltonian under low and transverse fields}

Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates compared to the high magnetic field regime.

\subsection{NV spin Hamiltonian in zero external magnetic field}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/ESR_simus}
\caption{Simulations of inhomogeneous zero field ODMR when sampling various parameters. a) Simulation when sampling each components of the magnetic field over a Gaussian of deviation $\sigma=2\ \rm G$. b) Simulation when sampling each components of the strain tensor $\bar{\bar{\varepsilon}}$ over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$. c) Simulation when sampling each components of the electric field over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$. d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
\label{simus ESR}
\end{figure}

In the absence of external magnetic field, we have to take into account other elements which would otherwise be of second order in the spin Hamiltonian. These elements are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}. The hyperfine splitting due to nearby nuclei will be considered separately, although to a large extent it behaves like a local magnetic field.

Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian). The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.

The NV$^-$spin Hamiltonian in the $\{ \ket{-1}, \ket{+1} \}$ basis can be written as \citep{udvarhelyi2018spin}:
\begin{equation}
\mathcal{H}=\begin{pmatrix}
D-\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}}) & f_\perp(\mathbf{E}) + g_\perp(\bar{\bar{\varepsilon}})\\
f^*_\perp(\mathbf{E}) + g^*_\perp(\bar{\bar{\varepsilon}})&D+\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}})
\end{pmatrix},
\label{Hamiltonien pm1}
\end{equation}
where $B_\parallel$ is the component of the magnetic field along the NV axis, and $f_\parallel, f_\perp, g_\perp$, and $g_\parallel$ are functions of the electric field $\mathbf{E}$ and the strain tensor $\bar{\bar{\varepsilon}}$, whose expressions are:

\begin{align}
f_\parallel(\mathbf{E})&=d_\parallel E_z, \\
f_\perp(\mathbf{E})&=d_\perp ( E_x + i E_y), \\
g_\parallel(\bar{\bar{\varepsilon}})&= h_{41}(\varepsilon_{xx}+\varepsilon_{yy})+h_{43} \varepsilon_{zz}, \\
g_\perp(\bar{\bar{\varepsilon}}) &= \frac{1}{2} \left[ h_{16}(\varepsilon_{zx}+i \varepsilon_{zy}) + h_{15}\left(\frac{\varepsilon_{yy}-\varepsilon_{xx}}{2}+i\varepsilon_{xy}\right) \right],
\end{align}
where $d_\parallel=0.35 \ \rm{Hz\, cm/V}$ and $d_\perp=17\ \rm{Hz\, cm/V}$ have been measured experimentally \citep{van1990electric}, and $h_{43}=2300\ \rm{MHz}$, $h_{41}=-6420\ \rm{MHz}$, $h_{15}=5700\ \rm{MHz}$ and $h_{16}=19660\ \rm{MHz}$ were computed through DFT \citep{udvarhelyi2018spin} and show reasonable agreement with experiments \citep{barson2017nanomechanical}.

Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.

The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting. For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$. The shift in energy will therefore tend to blur the energy split when averaging over a large number of spins.

Fig. \ref{simus ESR} shows a simulation of how each parameters of the spin Hamiltonian - local magnetic field, local electric field and strain - affects the zero external field ODMR profile. To do these simulations, I sampled each parameters separately $10^6$ times and plotted the histogram of the two eigenvalues of the Hamiltonian written in eq. \ref{Hamiltonien pm1}. Fig. \ref{simus ESR}-d) shows an experimental zero field ODMR spectrum, typical of what we observe with dense NV ensembles.

Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR. Given the simulation results, the only parameter that can give rise to this shape is the local electric field.

We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample. In particular, we do not shield for the earth magnetic field in our experiment. The splitting caused by the earth magnetic field however has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$, which is $\sim 4$ times weaker than the splitting we typically observe.

We will therefore consider that the low external magnetic field spin Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.

We will then adopt the following simplified Hamiltonian for the zero external field regime:
\begin{equation}
\mathcal{H}=\begin{pmatrix}
D&0&d_\perp E_\perp^* \\
0&0&0 \\
d_\perp E_\perp &0&D
\end{pmatrix},
\end{equation}
whose eigenvectors are $\ket{0}$ and$\ket{\pm}$ of eigenvalues 0 and $D\pm d_\perp \abs{E_\perp}$, where $\ket{\pm}$ are defined as:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-i\phi_E}\ket{-1}}{\sqrt{2}},
\end{equation}
where $\tan(\phi_E)=E_y/E_x$.

\subsection{NV spin Hamiltonian under purely transverse magnetic field}
\label{sec B transverse}
%\begin{figure}[h]
%\centering
%\includegraphics[width=0.7\textwidth]{chapter 4/Figures/map_etats_propres}
%\caption{Closeness of the Hamiltonian most excited state $\ket{3}$ with either $\ket{+1}$ or $\ket{+}$ as a function of the external magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
%\label{map champ transverse}
%\end{figure}

We now turn to the case of purely transverse magnetic field with respect to the NV axis, i.e. $\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$. In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.

In this regime, the NV Hamiltonian eigenstates are similar to the case where it is dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{-}, \ket{+}$\citep{qiu2021nuclear, qiu2022nanoscale}, of respective eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D,D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-2i\phi_B}\ket{-1}}{\sqrt{2}},
\end{equation}
and $\tan(\phi_B)=B_y/B_x$.

For the case where $d_\perp E_\perp \sim \frac{(\gamma_e B_\perp)^2}{D}$ and $\phi_E \neq 2\phi_B$, the eigenstates of the Hamiltonian are still of the form $\ket{0},\ket{\pm}$ with a relative angle $\phi$ in between $\phi_E$ and $2\phi_B$.

In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.

\subsection{Hyperfine coupling and inhomogeneous broadening}
\label{sec modif T2*}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/Comparaison_ESR_T2}
\caption{ODMR lineshapes for longitudinal, transverse and low magnetic fields on samples CVD-pink and Sumi-2. a) ODMR spectrum of a single class of NV centers for a longitudinal magnetic field ($B_\parallel \sim 100\ \rm G$). The microwave frequency is given with respect to the ODMR line central frequency. b) ODMR spectrum of a single class for a purely transverse magnetic field ($B_\perp \sim 100\ \rm G$). c) ODMR of all 4 classes for 0 external magnetic field. d) Table reporting the full width at half maximum of each ODMR line.}
\label{ESR for T2*}
\end{figure}

We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes. These change are relevant to our study of NV-NV CR due to the relation between the dipole-induced relaxation rate and $T_2^*$ detailed in sec. [REF].

A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment. In the $\{ \ket{0}, \ket{\pm 1} \}$, the eigen energies are linearly sensitive to (longitudinal) magnetic field, and only sensitive to electric fields at the second order, and vice versa for the $\{ \ket{0}, \ket{\pm} \}$ basis.

These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei. These effects can drastically modify the ODMR lineshape in zero external magnetic field \citep{jamonneau2016competition} or purely transverse magnetic field \citep{qiu2021nuclear,qiu2022nanoscale}.

We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center. $^{14}$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin. The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
\begin{equation}
\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
\end{equation}

where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor. $\mathcal{H}_n$ and $\bar{\bar{A}}$ can be written:

\begin{equation}
\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
\end{equation}

and
\begin{equation}
\bar{\bar{A}} = \begin{pmatrix}
A_{xx} & 0 & 0 \\
0 & A_{yy} & 0 \\
0 & 0 & A_{zz}
\end{pmatrix},
\end{equation}

where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.

Fig. \ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$.

While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers. These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.

Fig. \ref{ESR for T2*}-a) allows us to evaluate the magnetic field noise in both samples, since the $\ket{\pm 1}$ states are not sensitive to weak electric fields. We indeed find that the Sumi-2 sample has much more magnetic noise, to the point where the hyper-fine structure is no longer resolved. The total width of the line however is mostly dominated by the hyper-fine splitting, which result in a similar total linewidth in both cases

We should note that the usual practice in magnetometry is to consider the linewidth only of the three hyper-fine lines when they are resolved. This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines. In our case however, we have to consider the spectral overlap between NV centers and fluctuators, which have an additional broadening of $2\gamma_f \approx 6\ \rm MHz$ [REF] that completely obscures the hyper-fine structure. We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.

Fig. \ref{ESR for T2*}-b) allows us to evaluate the electric field noise in both samples, since the $\ket{\pm}$ states are not sensitive to weak magnetic fields. Similarly, the hyper-fine structure is hidden in this configuration since all three hyper-fine levels are nearly-degenerate, provided that $\frac{(\gamma_e B_\perp)^2}{D} > A_{xx},A_{zz},Q$ which is typically the case for $B_\perp > 40\ \rm G$. We can note that the electric field noise is significantly stronger in sample Sumi-2, which leads to an ODMR linewidth more than twice as big.

Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field. For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction). For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_s=-1,m_I=+1}$ and $\ket{m_s=+1,m_I=-1}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$. The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.

The table in Fig. \ref{ESR for T2*}-d) reports the linewidths measured on both sample for the various magnetic field configurations. We can see that the linewidth of sample CVD-pink varies by more than a factor of 2, while the ones from sample Sumi-2 varied by less than 20 \%. We should note that  almost all samples used in this chapter are HPHT type 1b diamond, similar to sample Sumi-2 and unlike sample CVD-pink. We should therefore expect the modification of $T_2^*$ to have a moderate impact on the depolarization dynamics.

\medskip

To summarize this section, we have seen that (for the samples used in this chapter) the NV centers spin Hamiltonian is dominated by transverse electric field for low external magnetic field, and that it causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$. We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples with a high impurity density.

\section{NV-NV CR in the low magnetic field regime}
In this part, we will study the cross-relaxation between NV centers under low magnetic field, whose spin Hamiltonian is dominated by a transverse electric field. Understanding the various mechanisms operating in this regime will be important for the magnetometry protocol presented after.

\subsection{Experimental observations}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 4/Figures/scan_1x1x1x1}
\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction. a) ODMR spectrum for B=0. b) ODMR spectrum for B=60 G. c) PL contrast as a function of the external magnetic field. d) $1/T_1^{\rm dd}$ as a function of the external magnetic field. A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
\label{scan 1x1x1x1}
\end{figure}
We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.

The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust in order to isolate each effects.

Fig. \ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G. The ODMR at the initial and final magnetic fields are shown in Fig. \ref{scan 1x1x1x1}-a) and b).

While it is clear that the spin lifetime as well as the PL increases with the magnetic field, there is no clear indication that this is because of the specificity of the low field region. Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.

We can note however that the PL is not an exact mirror of the spin dynamics: while the low field drop in PL is indeed associated with a reduced spin lifetime, the drop in PL at higher magnetic field comes from the states mixing induced by the transverse magnetic field, and is not associated with a modification of the spin dynamics (to the first order).

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{chapter 4/Figures/scan_100}
\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis.}
\label{scan 100}
\end{figure}

Fig. \ref{scan 100} presents a way to circumvent this issue: by applying the magnetic field along the [100] crystalline axis, we can make sure that the four classes of NV centers always stay resonant regardless of the magnetic field amplitude.

We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig. \ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of $\sim$ 8 in zero field, whereas in Fig. \ref{scan 100}, it was only reduced by a factor of $\sim$ 1.2. The main reason for the PL and $T_1$ drop in Fig. \ref{scan 1x1x1x1} was indeed the co-resonance between the four classes.

Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained if we consider only the inter-class resonances. This means that there are some additional depolarization mechanisms which are proper to the zero field region. We should also note that, while the zero-field PL contrast is bigger in Fig. \ref{scan 1x1x1x1}-c) than in Fig. \ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually similar in both cases.

We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]

\subsection{Theory of the low field depolarization}
\label{sec causes zero field}
We have identified three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.

\subsubsection{Eigenstates modification}
The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.

This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis. We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively. To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.

The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF]. The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane. This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.

We therefore need to make an assumption on the distribution of the transverse field in the sample. In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane. However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq. [REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.

We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.

We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$. $\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.

In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$. The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig. \ref{scan 100}

\subsubsection{Double flips}

\begin{table}[htbp]
\centering
\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
 \label{table double flips}
\begin{tabular}{c|cc}
%\toprule
{} & $l_c=0$ & $l_c=\infty$ \\
\midrule
$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
%\bottomrule
\end{tabular}
\end{table}

Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$. The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.

We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field. And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field. But we can see from Fig. \ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig. [REF]. In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.

To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one. We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF]. Finally, the question about the correlation  length of the electric field also needs to be asked.

The results are summed up in Table \ref{table double flips}. We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips. $\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$). This is close to the the experimental values we have. Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $). We come back to the results of the last section where only considered the flip-flops.

\subsubsection{$T_2^*$ modification}

The last aspect we considered was the change in $T_2^*$ (we include the change in the hyper-fine interaction in the $T_2^*$) for weak magnetic field, as discussed in sec. \ref{sec modif T2*}.

To quantify its impact in the depolarization rate, we employ eq. [REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig. \ref{ESR for T2*}.

We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.

\subsubsection{Summary}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/Shema_summary_theory}
\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
\label{summary_theory}
\end{figure}

Fig. \ref{summary_theory} recapitulate the various mechanisms at play in the low field depolarization of NV ensemble.

We start from from the simplest case where every classes are sepctrally isolated and the only dipole-induced relaxation coms from the flip-flop between NV and fluctuators from the same class. We then consider the case where all four classes are resonant ($B\parallel [100]$) and finally the case B=0.

The three extra depolarization mechanisms in zero field are then detailed, with the predicted increase of depolarization rate for each step. The predicted rates in zero field depend on several parameters, we chose the following ones to compute the different numerical values: $l_c=0$, $\bar{G}^2=0.5$, $2\gamma_f = 6.5\ \rm MHz$ and $2/T_2^*=5.9\ \rm MHz$ (corresponding to sample Sumi-2 if we assume a Lorentzian ODMR profile).

\subsection{Other potential causes for zero field depolarization}
We have also considered two other potential causes that we have ultimately ruled out in this case.

\subsubsection{Magnetic field alignment}
\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{chapter 4/Figures/alignement_rose}
\caption{PL of sample CVD-pink as a function of the magnetic field for various misalignment of the field direction with the [100] axis}
\label{alignement rose}
\end{figure}

The first is a potential misalignment of the magnetic field with the [100] axis, which would cause a splitting of the 4 classes as the magnetic field increases. This hypothesis is not consistent with the observations in Fig. \ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased. Instead we observe a sharp decrease followed by a pseudo-plateau.

We still estimated how sensitive the PL dip was with respect to the field alignment. Fig. \ref{alignement rose} shows PL scans on sample CVD-pink for various alignment of the magnetic field around the [100] axis. We estimate our alignments to be precise within $\pm 1^\circ$. We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization does not come from a field misalignment.

\subsubsection{Laser polarization}
\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{chapter 4/Figures/pola_laser}
\caption{PL of sample ADM-150-3 with respect to the magnetic field for various laser polarization angle. The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
\label{pola laser}
\end{figure}

Other studies \citep{anishchik2015low, filimonenko2020weak} previously saw a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.

Fig. \ref{pola laser} shows PL scans on sample ADM-150-3 for various polarization angle of the incident laser. We chose to apply the magnetic field in the laser polarization plane, in a direction that did not match any particular crystalline planes.

We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$. We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.

\section{NV-NV CR under purely transverse magnetic field}
We have seen in the last part that there was a spontaneous spin depolarization specifically in zero external magnetic field, and we have identified three potential mechanisms which are the double-flips, the change in the Hamiltonian eigenstates and the modification of $T_2^*$. We now want to experimentally discriminate the contribution of each of these mechanisms and to see how close the actual decay rates are to the predicted ones.

One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic and electric noise ...). Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.

The ideal way to isolate each contributions would be to apply an electric field strong enough to split the transition of the 4 classes, and to measure the decay rate of a single class dominated by the transverse electric field. Such an electric field would need to be of the order of $\sim 10^6\ \rm{V/cm}$, which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.

Instead of an electric field, we will use a transverse magnetic field, which as described in sec. \ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.

\subsection{Principle of the experiment}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/transis_transverse_field}
\caption{Simulation of the eigenstates of the NV Hamiltonian for  a purely transverse magnetic field $B_\perp$. In addition to the transverse magnetic field, we consider a fixed transverse electric field of value $d_\perp E_\perp=4\ \rm MHz$. (a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$. The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$. (b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
\label{eigenstates transverse field}
\end{figure}

Fig. \ref{eigenstates transverse field} shows a simulation of the eigenstates and eigenenergies for a NV Hamiltonian subject to transverse electric field and magnetic field. For convenience we chose to take $E_\perp = E_x$ and $B_\perp = B_x$, taking other geometric configuration alters only slightly the results. We labeled the eigenstates of the Hamiltonian $\ket{1}$, $\ket{2}$ and $\ket{3}$ in ascending order of energy.

Fig. \ref{eigenstates transverse field}-b) shows both the splitting $\Delta \nu$ between the $\ket{2}$ and $\ket{3}$ states, and how close the $\ket{3}$ state is to the $\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}}$ state via the factor $\abs{\bra{3}\ket{+}}^2$. We look at the $\ket{3}$ state since $\ket{2}=\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}}$ in this case.

These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 since we wanted to observe the modification caused by the eigenbasis $\{\ket{0},\ket{\pm}\}$. This way we can isolate the $T_1^{\rm dd}$ contribution coming from the change of eigenstates to the one coming from the double flips. Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100,200] G seems to satisfy both these criteria.

\subsection{Experimental data}

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{chapter 4/Figures/scan_T1_perp}
\caption{Experimental data for the depolarization under purely transverse magnetic field on sample ADM-150-2. a) ODMR spectrum for $|\mathbf{B}|=20\ \rm G$. The transitions of the class orthogonal to the magnetic field are labeled $\nu_+$ and $\nu_-$. b) Dependency of $\nu_+$ and $\nu_-$ with the magnetic field, as measured through ODMR. c) Measurement of $T_1^{\rm dd}$ as a function of the transverse magnetic field. The corresponding energy splittings $\Delta \nu=\nu_+-\nu_-$ are written on top. We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$ indicated by a red line. We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$}
\label{champ tranverse exp}
\end{figure}

We perform here a $T_1^{\rm dd}$ measurement similar to the one presented in Fig. [REF], on the same sample and with the same fitting parameters, but this time probing a class orthogonal to the magnetic field.

Fig. \ref{champ tranverse exp}-a) and b) show the evolution of the transition frequencies $\nu_+=\frac{E_{\ket{3}}-E_{\ket{1}}}{h}$ and $\nu_-=\frac{E_{\ket{2}}-E_{\ket{1}}}{h}$ where $\ket{1}$, $\ket{2}$ and $\ket{3}$ are the states described in Fig. \ref{eigenstates transverse field}. The magnetic field was scanned between 20 and 130 G by increment of $\sim 0.5\ \rm G$.

Fig. \ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ with the magnetic field. We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$. In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ is increased.

\subsubsection{Region A}

We attribute the decrease in region A to the double flips: as the magnetic field increases, so does $\Delta \nu$ and the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$. This value is coherent with the previously measured fluctuator linewidth. We can then estimate the part of the double flips in the total depolarization rate for the low field regime (20 G here) and find that they amount for $\sim 50\%$ in this case.

Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see in Fig. \ref{champ tranverse exp}-a) that they are not very far detuned. However we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.

\subsubsection{Region B}

We now turn to region B. Even tough the double flips are completely quenched, the effects of the eigenbasis modification and the $T_2^*$ reduction are still present, because we can see on Fig. \ref{eigenstates transverse field}-b) that the eigenstates of the spin Hamiltonian are still very close to $\{ \ket{0}, \ket{\pm} \}$ for these magnetic field values.

To evaluate the contribution of the eigenbasis and $T_2^*$, we measure the same class of NV centers, but this time with a longitudinal magnetic field ($B_\parallel \sim 100 G$) to get a baseline value of $1/T_1^{\rm dd}$ in the $\ket{\pm 1}$ basis. We found a value of $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ in this case, reported on Fig. \ref{champ tranverse exp}-c) as a green line.

We do find that the decay rate is higher in the $\ket{\pm}$ basis by $\sim 50\% $ which corroborates our predictions, as both the change of eigenbasis from $\ket{\pm 1}$ to $\ket{\pm}$ and the increase of $T_2^*$ in the $\ket{\pm}$ basis are predicted to increase the depolarization rate. The theoretical increase however overestimates once again the actual increase: for the change of eigenbasis alone we predicted an increase of the decay rate by a factor of 4 (see appendix [REF]), over 8 times more than the actual increase.

We unfortunately cannot experimentally discriminate the depolarization coming from the eigenbasis change or the $T_2^*$ while using the same sample, since both of this properties are tied to the Hamiltonian eigenstates. But we do know that in this scenario, for a single isolated class, the contribution of both the change of eigenbasis and $T_2^*$ amounts for $\sim 3$ times less than the double flips.

\subsubsection{Conclusion of the experimental observation}

We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field, and we expect that the depolarization for low magnetic fields behave similarly. We therefore claim that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
\begin{enumerate}
\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
\end{enumerate}

\section{Introduction to ensemble NV magnetometry}

Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.

NV magnetometry is generally decomposed between single NV and ensemble magnetometry. Single NVs,often associated with scanning probes, offer a spatial resolution limited only by how far the magnetic source can be from the NV center (typically $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).

Ensemble NV centers center on the other hand offer sensitivities as low as $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of sizes up to several mm.

We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR. We will not try to give a complete overview of this field, but only of the elements close to the technique we present. For a general overview on NV magnetometry, we recommend the excellent reviews  \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.

\subsection{AC and DC magnetometry}
The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted into spin population. These protocols are generally distinguished between DC-broadband and AC-narrow band.

For DC-broadband magnetometry, the general technique consist of a Ramsey interferometry experiment, presented in chapter 1, which is limited by the spin inhomogeneous coherence time $T_2^* \lesssim 1\ \rm{\mu s}$ for NV ensembles. The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$. Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry. The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.

For AC magnetometry, the basic protocol consist of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}. Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field. The AC magnetic field however has to be in phase with the rephasing pulses, which limits the AC sensing to a relatively narrow band. The state of the art in term of AC sensing is $\sim 1\ \rm pT/\sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}.

Recently, \citep{xie2021hybrid} have achieved new records in both DC ($\sim 200\ \rm fT/\sqrt{\rm Hz}$) and AC ($\sim 10\ \rm fT/\sqrt{\rm Hz}$) NV magnetometry by using flux concentrators which are ferromagnets concentrating the magnetic field in a small volume, similarly to an electromagnet. Because the gain from the flux concentrators is not tied to a specific magnetometry protocol, we will focus on the results obtained without flux concentrators.
\subsection{Low field magnetometry}
Magnetometry for low magnetic field ($\gamma B < d_\perp E_\perp$) is challenging since in this regime the NV eigenstates and eigenenergies do not depend on the magnetic field in the first order. This complication is often lifted  by simply using a bias magnetic field, typically from a permanent magnet.

Some systems however need to be kept under low to zero external magnetic fields, for example to study the J-coupling between spins \citep{sutter2012computational}, or the phase of skyrmions \citep{zazvorka2020skyrmion}. To study these systems with NV centers, we need to find new protocols for low field NV magnetometry.

The main technique for low field magnetometry consist in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}. The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$ \citep{zheng2019zero}.

Another technique consist in using the $^{13}\rm{C}-$NV complex we discussed in sec. [REF] which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are frist order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}. With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers. This technique is more suited for a carefully chosen single NV center.

\subsection{Microwave-free magnetometry}
All the precedent magnetemetry protocols described relied on the use of a controllable magnetic field, to coherently manipulate the NV spin state and to measure the phase imprinted by the magnetic field in the rotating frame.

There are other protocols which correlate the magnetic field with the spin $T_1$ time (relaxometry). These protocols can operate without microwave field and work for either DC or high frequency AC sensing.

\subsubsection{microwave-less DC magnetometry}
The DC microwave-free protocols rely on sharp feature in the PL for specific magnetic fields around certain cross-relaxations processes or level anti-crossing. Fig. [REF] in chapter 2 show such PL features, in particular the ground state level anti-crossing (GSLAC, simply labeled LAC on the figure), the sharpest of these features.

The GSLAC occurs for a purely longitudinal magnetic field where $\gamma B_z = D$, which corresponds to $B=1024\ \rm G$. In these conditions, the normally bright $\ket{0}$ state and the dark $\ket{-1}$ state become resonant and any residual mixing between these two states (coming from the electric field, the strain or the residual transverse magnetic field) will cause a drop in the NV polarization, and in the PL \citep{broadway2016anticrossing}.

A protocol based on the GSLAC line has been implemented, first only for longitudinal magnetic field measurement \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.

Another DC microwave-less protocol was proposed based on NV-NV CR \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}. This technique reconstructs the external magnetic field by scanning an other magnetic field, similar to Fig. [REF] in chapter 3, and by measuring the position of the NV-NV CR lines. By scanning the magnetic field in different directions, one can then recover the initial 3D magnetic field. The author of this paper did not measure the sensitivity of this protocol, but they predict a shot noise limited sensitivity of $24.7\ \rm nT/\sqrt{\rm Hz}$.

\subsubsection{microwave-less AC magnetometry}
The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed. The detection bandwidth is adjusted with an external DC magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.

Sensitivities as low as a few $\rm pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).

\subsection{Orientation-free magnetometry}
Every protocol described so far requires a precise knowledge  of the Larmor frequency between the $\ket{0}$ and $\ket{\pm 1}$ states as a function of the magnetic field. This knowledge implies that the direction of each NV axis in the frame of the magnetic field is known, because of the strong anisotropy in the NV Hamiltonian.

For a single crystal, these axes can always be calibrated through ODMR. However, for a polycrystalline diamond, diamond powder or even a single diamond with an erratic motion, the knowledge of these axes is not possible. New protocols need to be found if one wants to perform magnetometry with these materials.

To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes. This protocol however has an very low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective}.

Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies. The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the Larmor frequency), or via the PL \citep{finco2021imaging}.

\section{Low field depolarization magnetometry}
\label{sec LFDM}
We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the spontaneous depolarization of dense NV ensemble at low magnetic field. This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional field is required to improve performances.

We will first describe its operation and characterize it before comparing it to the previously mentioned NV magnetometry protocols.

\subsection{Principle of the experiment}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/setup_magneto}
\caption{Principle of the LFDM protocol a) Experimental setup. The confocal microscopy setup is similar to the one described in [REF], the lock-in amplifier (LIA) is used here to add a modulation on the magnetic field, instead of the microwave. Both the PL and the LIA signal are monitored. b) example of LIA signal on sample ADM-15-4 when scanning the magnetic field $B_0$ between -150 and 150 G. c) PL signal corresponding to the LIA signal}
\label{setup magneto}
\end{figure}

Fig. \ref{setup magneto} shows the principle of the LFDM protocol. The experimental setup is similar to the one presented in Fig. [REF] with the exception that there are no microwaves, and that the magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\tilde{B}_{\rm AC}$ used for the modulation. We typically use a modulation frequency $f_{\rm mod} \approx 1\ \rm kHz$, the same value we use for ODMR measurements. The main voltage $V_0$ responsible for the DC field and the oscillating voltage $\tilde{V}_{\rm AC}$ responsible for the AC field are summed via a homemade bias tee.

Fig. \ref{setup magneto}-b) and c) show respectively the PL measured directly on the photodiode and the lock-in signal when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G. For this configuration, the magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.

Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup. To improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing. Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.

\begin{figure}[h!]
\centering
\includegraphics[width=0.9\textwidth]{chapter 4/Figures/AC_and_DC_sensing}
\caption{DC and AC LFDM sensing protocols on sample ADM-15-5. The magnetic field is scanned in an arbitrary direction. a) DC sensing: $B_0$ is scanned between $-$80 and +80 G while an AC field of amplitude 10 G is used for the lock-in detection. b) AC sensing: $B_0$ scans are performed for 5 different $|\tilde B_{\rm AC}|$ values between 2 and 10 G. The best AC sensitivity is achieved for $B_0\approx 5\ \rm G$.}
\label{AC and DC sensing}
\end{figure}

Fig. \ref{AC and DC sensing} shows how the AC and DC sensing work for LFDM: for DC sensing the LIA signal around $B_0=0$ provide a sharp linear slope to convert the LIA voltage into a magnetic field. For AC sensing, the LIA signal is directly proportional to the amplitude of the field. The optimal AC field amplitude for DC sensing has to match the width of the PL dip $|\tilde{B}_{\rm AC}| = \sigma \approx 10\ \rm G$ in order to maximize the slope of the lock-in signal around $B_0=0$. the optimal DC field to use for AC sensing is $B_0= \sigma/2$ corresponding to the region where the lock-in signal is most sensitive to AC field.

\subsection{Sensitivity of the LFDM protocol}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/sensi_alternee}
\caption{Sensitivity measurement of the LFDM on sample ADM-15-4. a) Temporal trace of the LIA signal when applying an alternating external magnetic field of amplitude 35 $\mu$T and frequency $f=1\ \rm Hz$. b) Histogram of the DC measurement over 50 s. The data is fitted with Gaussians of width $1.6\ \rm{\mu T}$.}
\label{sensei alternee}
\end{figure}

We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols. The characterization was only done for DC sensing. The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).

We characterize the DC sensitivity by looking at the noise floor on the LIA signal when sitting on the steepest slope, around $B_0=0$. The following parameters were used in the case presented: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.

Fig. \ref{sensei alternee} shows the sensitivity measurement protocol. First, the applied magnetic field is calibrated with ODMR spectra, then an alternating DC field of known amplitude (here $\Delta B_0=35\ \rm{\mu T}$) is applied on the sample. Fig. \ref{sensei alternee}-a) shows the temporal trace of the LIA signal in response to this alternating DC field. Knowing the amplitude of the alternating field, we can then convert the voltage signal into a magnetic field amplitude, with in this case a ratio $45 \pm 3\ \rm{\mu T /V}$.

Fig. \ref{sensei alternee}-b) shows an histogram of the measured magnetic field amplitude for both the high and low value of the applied magnetic field. The total acquisition time was 50 s with the external field being alternated every second. The histograms are nicely fitted with Gaussian profiles of width $\sigma=1.5\pm 0.1\ \rm{\mu T}$ which allows us to estimate the DC sensitivity of the protocol:
\begin{equation}
\eta = \frac{\sigma}{\sqrt{2 \tau}}= 116 \pm 10\ \rm{nT} / \sqrt{\rm Hz},
\end{equation}

where $\tau=3\ \rm ms$ is the LIA low-pass filter time constant.

We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$). Laser fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.

\subsection{Angular dependence of the sensitivity}
\label{sec angular sensi}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/sensi_angulaire}
\caption{Angular dependence of the sensitivity. a) LIA signal for $B_0$ scanned between $-$150 and +150 G along the [100] axis, or along a direction 20$^\circ$ off the [100] axis. b) Zoom-in of the LIA signal on the zero-field slopes. c) Measured sensitivity as a function of the magnetic field angle with the [100] axis.}
\label{angular sensi}
\end{figure}

We will now look at the dependence of the sensitivity with respect to the magnetic field angle. In particular, we are interested in comparing the sensitivity when $\mathbf{B}\parallel [100]$, where there is no lift of degeneracy between the four classes as $\mathbf{B}$ increases; and a case where $\mathbf{B}\nparallel [100]$ and where there is a lift of degeneracy between the four classes. We previously saw in Fig. \ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\mathbf{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.

Fig. \ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angle of $B_0$. For the blue curve, the field was scanned along the [100] axis, for the orange curve, it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis. We can see that the maximum amplitude of the lock-in signal (relevant for AC sensing) is of the same order of magnitude for both case, and the slope around $B_0$ (relevant for DC sensing) is almost identical. Fig. \ref{angular sensi}-c) shows the sensitivity, measured with the protocol described in Fig. \ref{sensei alternee}, with respect to the angle between the magnetic field and the [100] axis. We can notice that the sensitivity depends only weakly on the magnetic field angle.

The fact that the sensitivity is similar when $\mathbf{B}\parallel [100]$ or $\mathbf{B}\nparallel [100]$ means that the inter-class co-resonances plays a lower role in the sensitivity than the zero-field specific depolarization mechanisms, previously detailed in Sec. \ref{sec causes zero field}. We cannot be sure which of the three zero-field mechanism is the dominant one, but based on the experimental results shown in Fig. \ref{champ tranverse exp}, we can assume that the double-flips are the dominant factor in the LFDM sensitivity.

It should be noted that this low angular dependence was particularly marked on sample ADM-15-4. Other samples, including from the same batch, showed a higher angular dependence with typically a sensitivity $2 \sim 3$ times worse for $B\parallel [100]$ than for the other directions.

\subsection{Temporal stability}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/Temporal_just_0B}
\caption{Temporal stability of DC-LFDM as measured on sample ADM-15-5. a) Temporal trace of the LIA signal for $B_0 \approx 0$. The Lock-in signal was converted to a measured magnetic field with the previously described protocol. b) Estimated signal to noise ratio for a signal of 100 nT as a function of the total acquisition time $T$. The noise value is computed by averaging the variance of the temporal trace over each time duration $T$. A curve $y=0.5\, T^{1/2}$ is added for comparison}
\label{temporal 0B}
\end{figure}

An important aspect of sensors is their stability in time. The sensitivity being expressed in $\rm{nT} / \sqrt{\rm Hz}$ suggests that the signal to noise ratio (SNR) given by repeating the measurement increases as $\sqrt{T}$ where $T$ is the total acquisition time. This is true for an ideal system, but in practice drifts and low frequency noise mean that the actual SNR for long acquisition time will be worse than the SNR at short time scaled by a factor $\sqrt{T}$.

One of the main source of drift for NV magnetometry based on spin resonance is the crystal lattice temperature change: the $D$ factor in the NV Hamiltonian (eq. [REF]) is sensitive to the lattice temperature because of the thermal dilatation of the crystal, which is the basis for NV thermometry \citep{kucsko2013nanometre}. This means that thermal fluctuations will impact the Larmor frequency of the spins and may cause a detuning with the initial measured frequency for long acquisition times. The LFDM protocol however does not rely on a specific resonance frequency, but rather on a co-resonnance condition between neighboring spins, which does not depend directly on the crystal temperature.

Fig. \ref{temporal 0B} shows the temporal stability of the LFDM protocol over an acquisition time of 100 s. Fig. \ref{temporal 0B}-a) shows a temporal trace of the LIA signal where we have converted the signal into a measured magnetic field. Fig. \ref{temporal 0B}-b) shows the estimated SNR for a signal of 100 nT as a function of the total acquisition time and was computed from the temporal trace.

We can see that the SNR does not deviate from the $\sqrt{T}$ scaling for times up to 50 s, and could probably be extended for longer times. The non-linearity for shorter times comes from the LIA low-pass filter.

\section{Comparison between LFDM and other magnetometry protocol}
Now that we have characterized the LFDM protocol, we can compare its performances with the previously mentioned magnetometry protocols. We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity. This discrepancy however come in large part to the smaller sensor volume used in the characterization, and to a relatively low PL collection efficiency compared to the other experiments.

\subsection{Comparison with microwave based protocols}

\begin{table}[htbp]
\centering
\caption{Comparison of the volume-normalized sensitivity of the best AC and DC protocols with LFDM}
 \label{table sensi volumiques}
\begin{tabular}{c|ccc}
\toprule
{} & Zhou et al. \citep{zhou2020quantum} (AC) & Barry et al.\citep{barry2016optical} (DC) & LFDM (DC) \\
\midrule
$\eta\ (\rm{nT} / \sqrt{\rm Hz})$ & 92&0.015& 116 \\
$V\ (\mu \rm{m}^3)$ &$8.1 \cdot 10^{-3}$ &$5.2 \cdot 10^{6}$& $3.3 \cdot 10^{3}$ \\
$\eta_v\ (\rm{nT}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ &8.3&34&6700 \\
\bottomrule
\end{tabular}
\end{table}

We will first compare the LFDM protocol with the best microwave-based NV magnetometry protocols. A better metric than the sensitivity to compare different protocols is the volume-normalized sensitivity defined in \citep{zhou2020quantum} as $\eta_v=\eta\cdot \sqrt{V}$ where $\eta$ is the sensitivity and $V$ the effective volume of the sensor. This scaling assumes that each NV center in an ensemble is an independent magnetic field probe, and as the total number of NV centers is proportional to $V$, the sensitivity should scale as $1/\sqrt{V}$.

Table \ref{table sensi volumiques} reports the best AC and DC volume-normalized sensitivities found in the literature. The LFDM protocol performs $\sim 200$ times worse than the best DC magnetometer. It should be noted however that the PL collection in our measurement, where we use a confocal microscope with a numerical aperture NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/compariason_0B_MW}
\caption{Comparison between LFDM and microwave-based magnetometry on the same experimental setup. The sample used is ADM-15-5 and the magnetic field is aligned close to a [111] axis. a) Schematics of the experiment: the setup is the same as the one described in Fig. \ref{setup magneto} with the addition of a continuous microwave tone at the frequency $\nu=2600\ \rm MHz$. b) LIA signal as $B_0$ is scanned from $-$150 to +150 G. The two lines at $\pm 100\ \rm G$ are caused by the microwave. c) Temporal trace of the measured B field for $B_0 \approx 0$ (on the middle of the zero field slope) and $B_0 \approx 100\ \rm G$ (on the middle of the microwave slope). The orange curve is shifted by $\approx 100\ \rm G$ to sit on the same scale as the blue curve. d) Estimated SNR fo a signal of 100 nT for $B_0 \approx 0$ and $B_0 \approx 100\ \rm G$ as a function of the total acquisition time}
\label{comparaison 0B MW}
\end{figure}

Fig. \ref{comparaison 0B MW} shows a comparison between LFDM and microwave-based magnetometry using our setup. A fixed microwave tone (here at 2600 MHz) is sent to the diamond and a magnetic field $B_0$ close to the [111] axis is scanned along an oscillating field $\delta B$ to perform a lock-in detection.

Fig. \ref{comparaison 0B MW}-b) shows the LIA signal over a $B_0$ scan: the profile is similar to the one in Fig. \ref{setup magneto}-b) with the addition to two sharp lines for $|B_0| \approx 100\ \rm G$. These two lines correspond to the transition $\ket{0} \to \ket{-1}$ (or $\ket{0} \to \ket{+1}$ for the negative fields) of the class aligned with the magnetic field when it comes to resonance with the 2600 MHz microwave tone. The microwave power used here correspond to a Rabi frequency $\Omega \approx (2\pi) 3\ \rm MHz$ and the slope on the steepest part of the microwave line is $\sim 2.5$ times higher than the slope in zero field.

Fig. \ref{comparaison 0B MW}-c) and d) show the temporal stability of the sensor similarly to Fig. \ref{temporal 0B}, on both the zero field line and the microwave line. While we can see that the microwave signal is more sensitive thanks to the higher slope, it is also more prone to drift, most likely because of the shift of the Larmor frequency caused by temperature fluctuations. We can even see that the LFDM becomes more sensitive for acquisition times beyond 10 s.

We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the available microwave power (although in this case $\Omega_{\rm Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$. Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.

\subsection{Comparison with microwave-less protocols}
The only microwave-less protocols to have measured a sensitivity are the GSLAC protocols described previously \citep{wickenbrock2016microwave, zheng2020microwave}. Unfortunately the sensing volume in this studies is not known, however due to the similarity in our approaches, we should be able to directly compare the physical origin of both protocols, i.e. comparing the zero field PL dip and the GSLAC PL dip.

In Fig. \ref{setup magneto}-c) we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$. In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type Ib irradiated diamond, similar to the ones we use. In \citep{zheng2020microwave} they manage to keep a contrast of 4.5 \% while reducing the GSLAC linewidth to $0.38\ \rm G$ by using a CVD sample isotopically enriched in $^{12}$C with a lower [NV] and [N] density.

The zero field technique is therefore comparable with the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear what kind of material engineering would improve its sensitivity.

%Je parle pas de la dépendence en puissance ou du champ transverse, je laisse ça pour plus tard.
\section{Conclusion and perspectives}
We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field, and wee have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers. We have identified three mechanisms by which the depolarization is increased in zero-field: the inter-class flip-flops, the double flips, and the change in the NV eigenstates (which includes the change in $T_2^*$).We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.

Finally, we have seen how this low-field depolarization can be applied to perform magnetometry experiments, and what advantages and disadvantages this technique offers compared to the established NV magnetometry protocols.

I will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.

\subsection{LFDM optimization}

Given the few parameters involved in the LFDM protocol, the main path for LFDM optimization seem to be the material optimization of the samples.

Most magnetometry protocols have a relatively clear material optimization path: DC protocols want to improve the figure of merit $\rho T_2^*$ and AC protocols want to improve $\rho T_2$, where $\rho=[\rm NV]$ is the NV concentration. The reason is that the number of NV centers for a given volume scales as $\rho$, and therefore the sensitivity scale as $\sqrt{\rho}$. Meanwhile increasing the density of defects tend to decrease the spin coherence times $T_2$ and $T_2^*$, and the AC and DC sensitivities scale respectively as $1/\sqrt{T_2}$ and $1/\sqrt{T_2^*}$.

For the LFDM protocol however, the material optimization is far from obvious. Indeed, while increasing $T_2^*$ would improve the sensitivity, the gain might not be as interesting as for other protocols due to the broadening caused by the fluctuator lifetime. On the other hand, increasing the NV density not only increases the number of sensor per volume, but also increases the dipole-dipole strength between NV centers. If we assume that fluctuators are constituted of closely packed impurities, then increasing the NV density should also increase the fluctuator proportion.

It seems then that increasing the NV density would be the obvious path to improve the LFDM performances. In practice however, I observed that the relationship between NV density, fluctuators, and LFDM performances was generally not straightforward. The two following examples will illustrate that.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/substrat}
\caption{Measurements on sample SBST-C for two different regions A and B. a) PL contrast as a function of a magnetic field scanned in an aribitrary direction (same direction for both regions). b) PL contrast as a function of a magnetic field scanned in the [100] direction. c) $T_1$ measurement on a single class when $B\neq 0$. d) $T_1$ measurement on all  four classes when $B=0$ }
\label{substrat chelou}
\end{figure}

Fig. \ref{substrat chelou} shows various measurements on the same sample SBST-C, a HPHT bulk sample containing [NV]$\sim  \rm ppm$, on two spatially distinct regions A and B. These two regions most likely correspond to two distinct growth sectors of the diamond, meaning that these regions were not grown in the same exact environment, which can impact among other things the creation of impurities. Both region showed a similar PL which would indicate a similar density of NV centers.

The metric we are interested in to evaluate the LFDM performance are the PL contrast when $\mathbf{B}$ is scanned in an arbitrary direction, which mainly represents the inter-class flip flop and which I will refer to as FF-contrast; and the PL contrast when B is scanned $\mathbf{B}$ is scanned along the [100] direction, which mainly represents the double flips and which I will refer to as DF-contrast. We are interested in the DF-contrast since we have seen in sec. \ref{sec angular sensi} that it could be the main contribution to the LFDM sensitivity.

Fig. \ref{substrat chelou}-a) and b) indicate that region B shows a higher FF-contrast, but a smaller DF-contrast than region A. Looking at the $T_1$ measurements for both regions, region B shows a shorter $T_1$ in both zero field and non-zero field, which would indicate a larger fluctuator density in region B. If this is true, than the fluctuator density is not directly correlated to the DF-contrast.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/double_adamas}
\caption{Same measurements as Fig. \ref{substrat chelou} on two distinct samples ADM-15-4 and ADM-15-1. Every parameters (laser power,optical setup , AC magnetic field) was kept identical for the two samples.}
\label{double dragon}
\end{figure}

We now turn to Fig. \ref{double dragon}. This figure shows the same 4 measurements but this time on two distinct samples, ADM-15-1 and ADM-15-4, which came from the same batch. Both samples showed relatively similar PL which could again indicate a similar NV density.

We can first notice the disparity in the  zero field PL contrasts: the FF-contrast is twice as big in sample ADM-15-4 as in sample ADM-15-1, and the DF-contrast is almost 5 times bigger. When looking at the $T_1$, we can notice that the lifetime for non-zero field is slightly longer for sample ADM-15-4 than for ADM-15-1, but its lifetime in zero-field is shorter (it is also exponential and not stretched-exponential as discussed in the last chapter [REF], which could have an influence on the PL contrast). From the $T_1$ measurements we can not determine which sample contains the most fluctuators, and therefore correlate the fluctuator or NV density with the DF and FF-contrasts.

These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field. We have noticed a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive. If these parameters were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.

Other non-material parameters in the LFDM protocols should also be investigated. This includes the temperature of the sample, the intensity and wavelength of the laser, the frequency and amplitude of the AC magnetic field. For the results presented in sec. \ref{sec LFDM}, we chose the parameters which we found optimal within our available experimental range.

\subsection{Potential applications for LFDM}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{chapter 4/Figures/Applications}
\caption{Schematics of two potential application for LFDM. The working of each application is detailed in the main text}
\label{applications}
\end{figure}

We will now turn to potential applications for LFDM, which could be useful even without material optimization.

Fig. \ref{applications} illustrates two potential applications for LFDM. The first one is to to perform wide-field magnetic imaging on a surface with micron-sized irregularities. The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}. While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism samples. Furthermore, controlling the angle between the diamond and the sample can be challenging even for two perfectly flat surface.

Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always have a sensor within $\sim$ 1 $\mu$m of the sample. Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens. One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$m red fluorescent diamond) are well suited for this task. I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples. This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}. Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required. Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field homogeneity.

Another potential application for LFDM is shown in Fig. \ref{applications}-b. It consists in the detection of magnetic field gradient thanks to a nano-diamond probe attached to a nano-resonator. The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig. \ref{AC and DC sensing}-b). This idea is based on previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}. A very similar protocol was recently developed with microwave based AC sensing \citep{huxter2022scanning}.

In order to use AC LFDM with a scanning probe setup, one would need to have a nano-diamond with at least one NV center and one fluctuator. This may not be such a rare event: a (20 nm)$^{3}$ nanodimaond with [NV]=5 ppm contains on average $\approx 4$ NV centers. If we take a fluctuator proportion of 30 \% as measured in \citep{choi2017depolarization}, then finding a nanodiamond with at least one NV and one fluctuator  seems feasible. I have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect, and I therefore cannot estimate the sensitivity of such a probe.

Other potential applications for LFDM includes the possibility to use polycrystalline diamond, potentially grown through heteroepitaxy on top of a surface of interest, or to follow individual diamonds flowing in liquid or living organism, similar to what was done in \citep{feng2021association}, but without having to constantly calibrate the diamond axes through ODMR which significantly hinders the temporal resolution of microwave-based technique.

Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous magnetic field or microwave field along the sample.

\printbibliography
\end{refsection}
 \end{document}

### 2022-09-22 Manuscrit Thèse/chapter 2/Chapter 2.tex
Le gros benj

@@
 
-The work presented in this chapter has already been published in large part in the articles \citep{pellet2021optical, ngambou2022improving}.
+The work presented in this chapter has been published in large part in the articles \citep{pellet2021optical, ngambou2022improving}.
 
@@
 
-We refer by dark spins to paramagnetic impurities for which, unlike the NV center, the spin degree of freedom is not accessible optically.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/tronche_de_spin}
+\caption{Crystalline structure of three spin defects: the NV center, the P1 enter (substituional Nitrogen) and the VH center.} 
+\label{repr NV P1 VH}
+\end{figure}
+
+We refer by dark spins to paramagnetic impurities for which the spin degree of freedom is not accessible optically.
 At room temperature, only two defects in diamond do not have a dark spin: the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
-Every other spin impurity studied so far can be considered to be dark.
+Every other spin impurity identified so far is considered to be dark.
 
@@
 \subsection{P1 centers and $^{13}$C}
-The most common spin defects in NV-rich diamond are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which represents 1.1\% of natural abundance carbon atoms, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
+The most common spin defects in NV-rich diamond are the substitutional nitrogen centers N$^0_S$ known as the P1 centers (see Fig.
+\ref{repr NV P1 VH}), which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which represents 1.1\% of natural abundance carbon atoms, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
+In NV-rich diamonds, these two impurities are the main causesof the NV center dephasing rates $\Gamma_2^*=1/T_2^*$ and $\Gamma_2=1/T_2$  \citep{barry2020sensitivity}.
 
-In NV-rich diamonds, these two impurities are the main causes of the inhomogeneous dephasing rate $\Gamma_2^*=1/T_2^*$ of the NV centers ensemble \citep{barry2020sensitivity}.
+The dephasing rate caused by a given impurity is directly proportional to the average diagonal dipole-dipole matrix elements ($\Omega_{\rm shift}$ in (eq.
+\ref{eq.
+ff df})) between the two spin species \citep{taylor2008high, bauch2020decoherence}.
+Since the dipole-dipole interaction scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity: $\Gamma_2^*(\rm P1) \propto [P1]$.
 
-The dephasing rate caused by a given impurity is directly proportional to the diagonal dipole-dipole matrix elements \citep{bauch2020decoherence}: 
-\begin{equation}
-\Gamma_2^*(\rm P1) \propto \Omega_{\rm NV-P1}=\frac{1}{\hbar}\sqrt{\sum_j \abs{\mel{NV=i;P1=j}{\mathcal{H}_{\rm dd}}{NV=i;P1=j}}^2}.
-\label{Gamma2 P1}
-\end{equation}
-Since the dipole-dipole coupling rate scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity: $\Gamma_2^*(\rm P1) \propto [P1]$.
-In the case of P1 centers, this was verified experimentally by \citep{bauch2020decoherence} who found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
-The typical concentrations of P1 center go from [REF] for high purity CVD diamond, up to 500 ppm for type 1B HPHT or natural diamonds.
+In the case of P1 centers, this was verified experimentally by looking at controlled sample with varying concentration of nitrogen \citep{bauch2020decoherence}.
+The authors of the study found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
+For the samples studied in this manuscript [REF], the P1 concentration was
 
-$^{13}$C is much more abundant than P1 center ($\sim$ 10 000 ppm at natural abundance) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\cdot 10^3$ times smaller than that of the P1.
-This results in a contribution to $\Gamma_2^* \sim 10^3$ times smaller per impurity.
+$^{13}$C nuclei are typically much more abundant than P1 center ($\sim$ 10 000 ppm without istopic engineering) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\cdot 10^3$ times smaller than that of the P1.
+As a result, the contribution of a single $^{13}$C nucleus to the NV dephasing rate is $\sim 10^3$ times that of a single P1 center.
 It was found experimentally that the broadening caused by $^{13}$C is $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 16\ \rm Hz/ppm$ \citep{van1997dependences, barry2020sensitivity}.
-For natural abundance diamond, this means that $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 160\ \rm kHz$.
+In samples with natural isotopic abundance of carbon, the inhomogeneous dephasing rate due to the $^{13}$C nuclear spins has a value $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 160\ \rm kHz$.
 
@@
 \label{other defects}
-Other spin defects are also commonly found in diamond containing NV centers, such as charged vacancy or vacancy clusters \citep{hounsome2006origin}, other nitrogen related defects \citep{newton2007epr}, transition metals \citep{isoya1990fourier}, and especially for CVD grown diamond, hydrogen related defects \citep{hartland2014study}, since the CVD plasma contains up to 95 \% hydrogen.
+Other spin defects are also commonly found in diamond containing NV centers, such as charged vacancy or vacancy clusters \citep{hounsome2006origin}, other nitrogen related defects \citep{newton2007epr}, transition metals \citep{isoya1990fourier}, and especially for CVD grown diamond, hydrogen related defects \citep{hartland2014study}, since the CVD plasma contains up to 95 \% hydrogen (see sec.
+\ref{fab HPHT}).
 
@@
 
-The VH$^-$ defect, prior to our observations, had only been observed \textit{via} electron paramagnetic resonance (EPR) in CVD grown diamond.
-It is formed by a vacancy with a hydrogen atom bonded to one of the four dangling carbon bond, with an extra electron trapped in the vacancy.
-This defect has a similar electronic structure to the NV$^-$ which results in an electronic spin-1 with a similar zero field splitting parameter (see Table \ref{table VH et War1}).
+The VH$^-$ defect, prior to our observations, had only been observed via electron paramagnetic resonance (EPR) in CVD grown diamond.
+It is formed by a vacancy with a hydrogen atom forming a bond with one of the four neighboring carbon atoms (see Fig.
+\ref{repr NV P1 VH}).
+To acquire its negtatively charged state, the VH center needs to capture an electron from a donor in the crystal, similarly to the NV center.
+
+This defect has a similar electronic structure as the NV$^-$ which results in an electronic spin-1 with a similar zero field splitting parameter (see Table \ref{table VH et War1}).
 Although the presence of hydrogen in this defect was confirmed by isotopic study, it has been debated whether this defect was VH$^-$ or V$_2$H$^-$ due to a conflict with \textit{ab initio} calculations \citep{shaw2005importance}.
-The defect is therefore sometimes referred as V$_n$H$^-$.
-In this manuscript I will employ the original name VH$^-$.
+The defect is therefore sometimes referred to as V$_n$H$^-$.
+In this manuscript we will employ the original terminology: VH$^-$.
 
@@
 \subsection{EPR spectroscopy}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{Figures/EPR_spectometer.jpg}
+\caption{EPR specrometer. Credits: Bruker Corporation} 
+\label{photo EPR}
+\end{figure}
+
 Electron paramagnetic resonance (EPR), the electronic equivalent to NMR, is the the most common way to detect electronic spins in solids and liquids.
@@
 
-At room temperature, the thermal polarization of a transition at 9500 MHz is $\sim 0.002$.
-This low value, compared to $\sim 0.8$ for the NV centers, explains in part why EPR is way less sensitive than NV ODMR.
-The current noise floor for EPR is typically of $\sim 10^{11}$ spins for electronic spins in diamond \citep{mitchell2013x}.
+The usual EPR apparatus, shown in Fig.
+\ref{photo EPR}, is a bulky and expensive lab equipment ($\gtrsim$ 1 M€).
+In contrast, optical detection such as ODMR is technically easier, less expensive, and allows better spatial resolution.
+
+ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ \citep{mitchell2013x}.
+This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and from the more efficient photon detection in the optical range than in the microwave range.
+
+EPR however remains a far more versatile tool as only very few spin defects have the properties required for ODMR.
+The key idea of this chapter is to allow the optical detection of spin defects that do not show ODMR, by making them resonant with NV centers and observing NV cross-relaxation.
 
@@
 
-\subsection{Principle of NV center Relaxometry}
-Detection \textit{via} cross-relaxation	with NV centers fall under the more general technique of NV relaxometry, where we are effectively measuring a modification in the NV center spin lifetime $T_1$.
-Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
+Detection of dark spins via NV center cross-relaxation fall under the more general technique of NV relaxometry, which consists in measuring a modification of the NV center spin lifetime $T_1$.
+Relaxometry with NV centers has been used not only to measure cross-relaxation with dark spins \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}or magnetic domains \citep{finco2021imaging}.
 
-There can be several advantages to use relaxometry protocols, which measures a change in the NV $T_1$, over the more common magnetometry protocols that rely on a change in the NV $T_2$ or $T_2^*$: the first and most obvious one is that relaxometry does not need to use an external microwave field.
-This however comes at the cost of a greater precision needed on the external magnetic field.
-Relaxometry also allows AC-magnetic field sensing in the GHz regime \citep{wang2022picotesla, alsid2022solid}, while traditional microwave-based AC-megnetometry is limited to tens of MHz.
-Finally, relaxometry measurement can be more sensitive than the T2-based one in some instances \citep{steinert2013magnetic}, which comes from the fact that changes in NV $T_1$ times are easier to detect than changes in $T_2$ or $T_2^*$ since $T_1$ is much longer than $T_2$ or $T_2^*$.
+The protocols used in NV relaxometry measure either directly the spin $T_1$, as detailed in sec. \ref{sec T1 NV}, or indirectly through the NV photoluminescence.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/relaxo_pulsed_continuous}
+\includegraphics[width=0.9\textwidth]{Figures/relaxo_pulsed_continuous_2}
 \caption{Representation of the two relaxometry protocol.
-Left : representation of the three level system of the NV center with the polarization rate $\Gamma_{\rm las}$ and spin decay rates $\Gamma_1=1/T_1$ and the perturbation $\delta \Gamma$.
-Right: simulation of a $T_1$ measurement for $1/\Gamma_1=3\ \rm ms$ and $1/(\Gamma_1+\delta \Gamma)=2\ \rm ms$.
-$\delta \rm S$ represents the maximum difference in signal between the two curves for a dark time $\tau$.}
+(a) Simple $T_1$ measurement pulse sequence.
+(b) Simulation of two $T_1$ measurement with different $T_1$ values.
+$\tau$ represents the optimal dark time in order to maximize the signal $\delta S$ (c) Representation of the optical pumping (green arrows) and depolarization (blue arrows) of the NV center spin.
+(d) Simulation of the PL modification caused by a change in the spin $T_1$.
+The spin $T_1$ profile in term of the magnetic field was arbitrarily chosen.}
 \label{T1 vs PL}
@@
 
-There are two ways to measure a change in the spin $T_1$, as represented on Fig.
+The two most common protocols to perform NV relaxometry ar represented on Fig.
 \ref{T1 vs PL}.
@@
 \end{figure}
+
+There can be several advantages to use relaxometry protocols, which measures a change in the NV $T_1$, over the more common magnetometry protocols that rely on a change in the NV $T_2$ or $T_2^*$: the first and most obvious one is that relaxometry does not need to use an external microwave field.
+This however comes at the cost of a greater precision needed on the external magnetic field.
+Relaxometry also allows AC-magnetic field sensing in the GHz regime \citep{wang2022picotesla, alsid2022solid}, while traditional microwave-based AC-megnetometry is limited to tens of MHz.
+Finally, relaxometry measurement can be more sensitive than the T2-based one in some instances \citep{steinert2013magnetic}, which comes from the fact that changes in NV $T_1$ times are easier to detect than changes in $T_2$ or $T_2^*$ since $T_1$ is much longer than $T_2$ or $T_2^*$.
 

### 2022-09-25 Manuscrit Thèse/chapter 2/Chapter 2.tex
btrg

@@
 
-The protocols used in NV relaxometry measure either directly the spin $T_1$, as detailed in sec. \ref{sec T1 NV}, or indirectly through the NV photoluminescence.
+The protocols used in NV relaxometry measure either directly the spin $T_1$ or indirectly through the NV photoluminescence.
 
@@
 
-The two most common protocols to perform NV relaxometry ar represented on Fig.
+The two most common NV relaxometry protocols are represented on Fig.
 \ref{T1 vs PL}.
-The first and most basic method simply consist in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, which itself results from an equilibrium between the polarization rate $\Gamma_{\rm las}$ and the spin decay rate $\Gamma_1=1/T_1$.
-A change in $\Gamma_1$ will therefore conduct to a new equilibrium with a different PL.
+The first method simply consists in measuring the spin lifetime, following the protocols described in \ref{sec T1 NV}.
+In Fig.
+\ref{T1 vs PL}, we picture the simpler $T_1$ measurement protocol, but the technique can be used with the common mode rejection protocol.
+In order to more efficiently measure a modification of the spin $T_1$, the ``single$-\tau$" protocol does not record the entire depolarization curve represented in Fig.
+\ref{T1 vs PL}-b), but instead probes the spins for a single dark time value $\tau$ \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}.
+This value is chosen in order to maximize the change in signal $\delta S$ for a given change in lifetime $\delta T_1$, typically $\tau \sim T_1$.
 
-The second method is a pulsed sequence referred as single$-\tau$ measurement, used for example in \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}.
-It simply consists in a $T_1$ measurement sequence (as seen in chapter 1) where you first polarize the spins in the $\ket{0}$ state with a laser pulse and read out the spin state with a second laser pulse after a dark time $\tau$.
-In this case however, instead of scanning the time $\tau$, you fix its value to maximize the sensitivity in a change of $T_1$ (typically $\tau \approx T_1$).
+ The second method simply consists in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, and since the steady state population in the $\ket{0}$ state results from the equilibrium between the optical polarization rate $\Gamma_{\rm las}$ and the depolarization rate $\Gamma_1=1/T_1$.
+A simple rate equation for the $\ket{0}$ state population $\rho_{00}$ gives:
+ \begin{equation}
+ \rho_{00}=\frac{\Gamma_{\rm las}+\Gamma_1}{\Gamma_{\rm las}+3\Gamma_1}.
+ \end{equation}
 
-It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the first case you need to adjust $\Gamma_{\rm las} \sim \Gamma_1$ to maximize the PL contrast, meaning that the laser power used is very far from the saturation limit (typically $\Gamma_1 \approx 10^3\ \rm s^{-1}$ and $\Gamma_{\rm sat} = 5\cdot10^6\ \rm s^{-1}$\citep{dreau2011avoiding}), while the second method needs to wait a time $\tau \sim T_1$ between each measurement, which result in a similar PL count and shot noise limit in both case.
+ Fig.
+\ref{T1 vs PL} shows a simulation of NV centers PL when the spin decay rate $\Gamma_1$ is modified.
+In this fictional scenario, we assume that $\Gamma_1$ resonantly increases for a given value of the magnetic field (we will see later in this manuscript situations matching this scenario).
+We can see that the PL decreases as an answer to the $\Gamma_1$ increase.
+The exact value of the PL drop ($\sim 2\%$ in the crase presented here) depends on the optical polarization rate $\Gamma_{\rm las}$, but this value generally does exceed a few \%, even for important $T_1$ modifications.
+
+It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the PL case, the laser intensity needs to be adjusted so that $\Gamma_{\rm las} \sim \Gamma_1$ in order to maximize the PL contrast, while the single$-\tau$ method requires a waiting time $\tau \sim T_1$ between each measurement, which result in a similar PL count and shot noise limit in both case.
 
@@
 
-There can be several advantages to use relaxometry protocols, which measures a change in the NV $T_1$, over the more common magnetometry protocols that rely on a change in the NV $T_2$ or $T_2^*$: the first and most obvious one is that relaxometry does not need to use an external microwave field.
-This however comes at the cost of a greater precision needed on the external magnetic field.
-Relaxometry also allows AC-magnetic field sensing in the GHz regime \citep{wang2022picotesla, alsid2022solid}, while traditional microwave-based AC-megnetometry is limited to tens of MHz.
-Finally, relaxometry measurement can be more sensitive than the T2-based one in some instances \citep{steinert2013magnetic}, which comes from the fact that changes in NV $T_1$ times are easier to detect than changes in $T_2$ or $T_2^*$ since $T_1$ is much longer than $T_2$ or $T_2^*$.
-
-Outside of cross-relaxations, another protocol exists to address dark spin resonance with NV centers, based on a modification of the NV $T_2$ time \citep{serbyn2014interferometric}.
-This protocol called double electron-electron resonance (DEER) is depicted on Fig.
+Aside from cross-relaxations, another protocol exists to address dark spin resonance with NV centers, based on a modification of the NV $T_2$ time \citep{mamin2012detecting, serbyn2014interferometric}.
+This protocol, called double electron-electron resonance (DEER), is depicted on Fig.
 \ref{DEER}.
-The pulse sequence consist in a spin-echo sequence on the NV center, as described in chapter 1, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers.
-This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center wont be rephased.
-The newly measured $T_2$ time should therefore be slightly shorter:
+The pulse sequence consists in applying a spin-echo on the NV center, as described in \ref{sec T2 echo}, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers.
+This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center will not be rephased.
+The newly measured $T_2$ time should thus be slightly shorter.
 
@@
 
-By scanning the frequency of the second microwave (RF2) and monitoring the $T_2$ time of the NV centers, one can therefore find dark spin resonance without having to bring them to resonance with NV centers, which can be useful when the dark spin resonance is very far detuned from the NV one.
+By scanning the frequency of the second microwave (RF2) and monitoring the $T_2$ time of the NV centers, one can find dark spin resonance without having to bring them to resonance with NV centers, which can be useful when the dark spin resonance is very far detuned from the NV one.
 
-While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, there are several differences between the two protocols:
+Both relaxometry and DEER protocols have their own advantages.
+While DEER allows more control on the probed impurity, allowing in particular coherent control of the dark spin \citep{sushkov2014magnetic}, relaxometry protocols are generally easier to implement as they do not require a pulsed microwave setup.
+In some cases at least, relaxometry protocols have been shown to be more sensitive than $T_2$ based protocols such as DEER due to the fact that $T_1$ is typically orders of magnitude greater than $T_2$ \citep{steinert2013magnetic}.
 
-Frist, the two protocols do not measure the same elements of the dipole-dipole Hamiltonian: CR measures the off-diagonal elements $\Omega_{\rm CR}=\mel{i,j}{\mathcal{H}_{\rm dd}}{j,i}/\hbar$ which are responsible for the flip-flops or double flips, while DEER measures the diagonal elements $\Omega_{\rm DEER}=\mel{i,j}{\mathcal{H}_{\rm dd}}{i,j}/\hbar$ which are responsible for the pure dephasing.
-These terms however are on average of the same amplitude, and when averaging over an ensemble we can consider $\Omega_{\rm CR} \sim \Omega_{\rm DEER}$
-
-Second and more importantly, the scaling is not the same in both cases : CR measures a change in the NV decay rate $\Gamma_1=1/T_1$ which is proportional to $\Omega_{\rm CR}^2$, as given by eq.
-\ref{delta gamma 1}, while DEER measures a change in $\Gamma_2=1/T_2$ which is proportional to $\Omega_{\rm DEER}$, as given by eq.
-\ref{Gamma2 P1}.
-This means that the signal of CR scales quadratically with the impurity concentration, while the DEER signal scales linearly with it.
-CR is therefore more suited to probe concentrated impurities, when the average distance between the NV and the probed spin is small.
+Relaxometry protocols do require however specific magnetic field values in order to bring in and out of resonance the two spins involved in the cross-relaxation.
+The choice of the field orientation and strength in order to best observe NV-dark spin cross-relaxations will be covered in the following section.
 
@@
 
-Since relaxometry does not rely on a microwave field, the only parameter left to tune is the external magnetic field, which is often scanned with an electromagnet.
-Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect dark spins.
+As relaxometry does not rely on a microwave field, the only external parameter left to tune is the magnetic field.
+Typically, the magnetic field is scanned in a fixed direction while the signal, either PL or T1, is recorded.
+Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect NV-dark spins CR.
+
+This section focuses on particular field orientation with respect to the diamond crystalline axes, and details the advantages and disadvantages of the two main choices that are the [100] and [111] direction in order to detect four particular dark spins: P1 centers, Vh$^-$, War1 and the $^{13}C-NV$ complex\footnote{The $^{13}C-NV$ complex is normally not dark, as discussed in its own section.}.
 
@@
 Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\mathbf{B} \parallel [100]$ or 
-$\mathbf{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes as defined in chapter 1.
+$\mathbf{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes.
 When $\mathbf{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\mathbf{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.
 
-These changes in the field orientation result in massive changes in the NV behavior, manly due to the transverse magnetic field: Fig.
+These changes in the field orientation result in important changes in the NV behavior, mainly due to the transverse magnetic field: Fig.
 \ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the ground state spin Hamiltonian (REF chapter 1) - or to be more precise the transitions between the ground state of the spin Hamiltonian and the two excited states - for any of the four classes of NV centers when $\mathbf{B} \parallel [100]$.
@@
 The three classes behave similarly than the four classes in the $[100]$ case, but the one class aligned with the magnetic field can get its $\ket{0}\to\ket{-1}$ transition arbitrarily small, as long as the magnetic field is perfectly aligned with the NV axis.
+
+We will now compare these two possible field orientation in order to observe cross-relaxation between NV centers and the four previously mentioned dark spin species.
 
@@
 
-I refer by $^{13}$C-NV$^-$ to a complex formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope.
+The $^{13}$C-NV$^-$ complex is formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope.
 Such a complex is illustrated in Fig.
@@
 
-The $^{13}$C-NV$^-$ complex still behaves as a ``normal" NV center, except that its transition lines are shifted by the hyper-fin coupling.
-In particular its electronic spin is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states.
-This is confirmed by the presence of sideband in ODMR spectra around the main NV lines \citep{simanovskaia2013sidebands}.
+The $^{13}$C-NV$^-$ complex still behaves as a ``normal" NV center, except that its transition lines are shifted by the hypere-fin coupling.
+In particular its electronic spin is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states\footnote{This is confirmed by the presence of sideband in ODMR spectra around the main NV lines \citep{simanovskaia2013sidebands}.}.
 
 Since both NV$^-$ and $^{13}$C-NV$^-$ are \textit{a priori} equally polarized, then following the argument in \ref{Sec_CR} there should be no CR between the two spin families.
-The reason why those transitions are still in some case visible, as will be shown below, and why in fact any NV-NV co-resonance gives a visible CR signal will be the object of the next chapter.
+The reason why those transitions are still visible is the presence of dark NV-centers (``fluctuators") which will be further explained in chapter 3.
 
@@
 When $\mathbf{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
-This abundance of transitions for low ($<100\ \rm G$) magnetic field makes analyzing experimental data quickly intractable in the second case.
+This abundance of transitions for low ($<100\ \rm G$) magnetic field makes the experimental analysis quickly intractable in the second case.
+
+There is one last aspect to take into account regarding the field orientation choice which is the impact of the transverse magnetic field on the cross-relaxation contrast.
 
@@
 
-In order to optically detect CR between an NV center and a dark spin, two conditions must be met: the NV center must be polarized in order to have a population transfer between the NV and the dark spin, and there must be a difference in brightness between the two NV levels involved in order to have an optical signature of the CR.
+Similarly to ODMR, there are two required conditions to observe NV cross-relaxation on the PL signal: 
+\begin{itemize}
+\item There must be a difference in population between the two NV spin states involved in the CR (e.g.
+$\ket{0}$ and $\ket{-1}$), otherwise there would be no population transfer with the unpolarized dark spin.
+\item The two NV spin states must have a difference brightness in order to be optically detectable.
+\end{itemize}
 
-These two conditions are in fact the same needed to observe an ODMR spectrum (REF chapter 1), and while the NV centers verifies both criteria in presence of a purely longitudinal magnetic field, this is not always the case anymore in presence of a transverse magnetic field (REF chapter 1).
+These two conditions are fulfilled when there is no magnetic field or purely longitudinal magnetic field, but not anymore in the presence of transverse magnetic field which mixes the NV spin Hamiltonian eigenstates \citep{tetienne2012magnetic}.
 
 Fig.
-\ref{121 vs 22 contrast} shows the simulated relative contrast of an ODMR spectrum, as computed from \citep{tetienne2012magnetic}, for a magnetic field aligned along [100] or [111].
-When the NV center is aligned with the magnetic field, the constrats remains equals to 1 regardless of the magnetic field because there is no transverse field.
-When the NV is not aligned however, either in the [111] or [100] case, the ODMR contrast quickly drops down for B$>100\ \rm G$.
+\ref{121 vs 22 contrast} shows the simulated relative contrast of a CR process \footnote{Due to the similarity between ODMR and CR, we assume here that the ODMR contrast is equal to the CR contrast} computed from the rates found in \citep{tetienne2012magnetic}, for a magnetic field aligned along either [100] or [111].
+When the NV center is aligned with the magnetic field, the contrast remains equal to 1 regardless of the magnetic field because there is no transverse field.
+When the NV is not aligned however, the CR contrast quickly drops down for B$>100\ \rm G$.
 
-This tells us that we should not expect to observe a CR signal for B greater than a few hundreds G when B is scanned along the [100] axis.
-We could however observe CR for arbitrarily large magnetic field when scanned along the [111] axis, as long as the field is properly aligned with the NV axis.
+The [100] direction is therefore not well suited to observe CR at magnetic fields higher than a few hundreds gauss.
+Fortunately, the co-resonance between NV centers and VH$^-$, War1 or $^{13}$C-NV$^-$ all happen for $|\mathbf{B}| < 150 \ \rm G$, and can thus be detected with a [100] oriented magnetic field.
 
 \section{Dark spin spectroscopy with NV centers}
+This section shows experimental observation of NV-dark spins CR.
+Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross} had been performed on the detection of dark spins through CR with NV centers, and to our knowledge, all this studies where done on the detection of P1 centers.
+
+We show an example of one this study to illustrate the principle of dark spin spectroscopy before showing our results on the detection of VH$^-$ and War1 in a CVD diamond.
 
@@
 
-Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross} had been done on the detection of dark spins through CR with NV centers, and to our knowledge, all this studies where done on the detection of P1 centers.
-Following the discussion in the last section, these studies were therefore all done with $\mathbf{B} \parallel [111]$.
-
-An example of one of this study (\citep{armstrong2010nv}) is shown in Fig.
+An example of a NV-P1 CR study (\citep{armstrong2010nv}) is shown in Fig.
 \ref{CR P1 exp}.
 It shows the evolution of the NV PL with respect to a magnetic field scanned along the [111] crystalline axis.
-There are several information to gather from this graph:
-
-First, there is the slowly decaying PL envelope, which stabilizes to a plateau for $B\sim 1000\ \rm G$.
+There are several information to gather from this graph: First, one observes a slowly decaying PL envelope, which stabilizes to a plateau for $B\sim 1000\ \rm G$.
 This change in PL is caused by the transverse magnetic field acting on the three NV classes that are not aligned with the magnetic field.
@@
 
-Second, there are the three salient features denoted Ns-NV, NV-NV and NV-LAC.
-All these features come from cross-relaxations or level anti-crossing (LAC) from the NV class aligned with the magnetic field.
+Second, there are three salient features denoted Ns-NV, NV-NV and NV-LAC. All these features come from cross-relaxations or level anti-crossing (LAC) from the NV class aligned with the magnetic field.
 
@@
 
-The diamond used for this study is the sample CVD-pink [REF appendix] whose fabrication was detailed in \citep{tallaire2020high}.
-It is a CVD diamond containing $[\rm NV ]\approx 4\ \rm ppm$ for $[\rm N ]\approx26\ \rm ppm$, and a natural abundance (1.1\%) of  $^{13}$C.
-Being a CVD diamond it is also likely to incorporate hydrogen defects, as discussed in \ref{other defects}.
+The diamond used for this study is the sample CVD-pink [REF appendix].
+The experimental setup used to obtain this data is the standard confocal microscopy setup shown in Sec \ref{Sec setup cahp 1}.
+The magnetic field here is generated with an electromagnet (EM), aligned along the [100] diamond axis with a precision $<1^\circ$.
 
@@
 Similarly to fig.
-\ref{CR P1 exp}, we can see a slowly decaying envelope of the PL due to the transverse magnetic field (we don't see the plateau here since the magnetic field stops at 150 G).
-We can also see three deviations to the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
+\ref{CR P1 exp}, there is a slowly decaying envelope of the PL due to the transverse magnetic field \footnote{The plateau is not visible in this case as the scan stops at 150 G}.
+There also are three deviations from the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
 
-In order to better discern the three features, we fitted the envelope by using a polynomial fit of order 4 on the smooth part of the data.
-We previously tried to use a fit based on the expected PL rate of the NV centers, by using the model in \citep{tetienne2012magnetic}, but the results were not satisfying.
+In order to better discern the three features, we fitted the envelope with a polynomial fit of order 4.
+We previously tried to simulate the PL dependency on the transverse field  by using the model in \citep{tetienne2012magnetic}, but the result did not correctly fit our data.
 
@@
 
-The experimental setup used to obtain this data is the standard confocal microscopy setup shown in appendix REF.
-The magnetic field here is generated with an electromagnet (EM), aligned along the [100] diamond axis with a precision $<1^\circ$.
-
 The data shown here was acquired over 24 h.
-It should be noted however that a single-photon counting APD was used here, which considerably worsens the sensitivity of the measurement as discussed in [REF].
+It should be noted however that a single-photon counting APD was used for this acquisition, which reduces significantly the sensitivity because of the high shot noise \footnote{The sample produces $\sim$ 100 nW of PL, which means that only $\sim 10^{-5}$ of the total PL is actually measured by an SPC-APD with a max counting rate of $10^7$ photons/s.}.
 Using a photodiode and modulation of the magnetic field, we estimate that we could get similar signal to noise ratio within a few minutes.
@@
 
-\section{Attribution of the $^{13}$C-NV$^-$, VH$^-$ and War1 lines}
+\section{Identification of the $^{13}$C-NV$^-$, VH$^-$ and War1 lines}
 
@@
 
-The attribution of the PL dips at 56 and 122 G in Fig.
-\ref{CR VH exp} to CR between NV and respectively VH$^-$ and War1 comes from two main observations: the closeness in the ZFS values, and the CVD nature of the sample.
+The identification of the PL dips at 56 and 122 G in Fig.
+\ref{CR VH exp} as CR between NV and respectively VH$^-$ and War1 comes from two main observations: the closeness in the ZFS values, and the CVD nature of the sample.
 
@@
 The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below.
-We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the process, and while we don't know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
+We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the process, and while we do not know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
 
@@
 
-For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively, with a slightly stronger confidence for the VH$^-$ since more is known about it.
+For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively, with a slightly stronger confidence for the VH$^-$ one.
 
@@
 
-The attribution of the $\sim 20\ \rm G$ dip to CR with $^{13}$C-NV$^-$ also come from two observations.
+The identification of the $\sim 20\ \rm G$ as CR between $^{13}$C-NV$^-$ and ``normal" NV centers also come from two observations.
 First, the lines are present on every single sample with dense ensemble ($\gtrapprox 1$ ppm), regardless of the fabrication process.
 Fig.
-\ref{scan sumi 4} for example shows a PL scan along the [100] axis for sample Sumi-4, an HPHT sample with [NV]$\approx 5\ \rm ppm$ and natural abundance of $^{13}$C.
-I could unfortunately not try a sample isotopically enriched in $^{12}$C since those samples are generally reserved to low-nitrogen concentration ($< 10\ \rm ppm$), which does not yield enough NV centers to perform CR spectroscopy.
+\ref{scan sumi 4} for example shows a PL scan along the [100] axis for sample Sumi-4, an HPHT sample with [NV]$\approx 5\ \rm ppm$ and natural abundance of $^{13}$C \footnote{We could unfortunately not try a sample isotopically enriched in $^{12}$C since those samples are generally reserved to low-nitrogen concentration ($< 10\ \rm ppm$), which does not yield enough NV centers to perform CR spectroscopy.
+}.
 
@@
 
+\bigskip
+
+The CR spectra shown in this section can be used to detect the presence of particular impurities in the sample, but more quantitative information on these impurities can also be extracted from these spectra, which is the object of the following section.
+
 \section{Quantitative estimates of the NV-Dark spins CR}
 
-We can extract some quantitative information about the dark spins by exploiting the CR spectrum shown in Fig. \ref{CR VH exp}.
+This section exploit the CR spectrum shown in Fig. \ref{CR VH exp} in order to extract information on the  .
 

### 2022-09-26 Manuscrit Thèse/Intro et conclusion/Intro.tex
ghuk

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric-comp,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_intro.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
\title{}
\begin{document}
\chapter*{Introduction}

Quantum mechanics was first developed at the beginning of the 20th century as an answer to  . In the second half of the 20th century, technological development based on the new understanding of light and matter have led to what was retroactively called the first quantum revolution \citep{thew2019focus}. Technological developments in the 50s and 60s include the semiconductor transistor, the laser, and nuclear magnetic resonance  which have since revolutionized the fields of information science, global communications and medical imaging.

A new field of quantum technology has emerged in the last two decades. This ``second quantum revolution" differs from the first one in the usage of individual quantum systems, such as single photons, atoms or electrons \citep{peil1999observing}, which have properties that can not be emulated by larger classical systems. These properties include quantum entanglement, superposition or measurement and form the basis of the newly formed quantum information science \citep{nielsen2002quantum, vedral2006introduction, hayashi2006quantum}.

Central to the quantum information science is the idea of quantum bits or qubits \citep{schumacher1996sending}, which are the building blocks of quantum information in analogy to the bits of classical information theory. Physical implementation of a qubit could be any quantum system with two well defined quantum states, as long as these states can be initialized, manipulated and readout \citep{divincenzo2000physical}. Popular qubit candidates include superconducting circuits \citep{nakamura1999coherent, orlando1999superconducting}, photons \citep{bennett1992quantum, zhong2020quantum}, quantum dots \citep{veldhorst2014addressable, zajac2018resonantly}

\printbibliography
\end{document}

### 2022-09-27 Manuscrit Thèse/Intro et conclusion/Intro.tex
gyui

@@
 
-Central to the quantum information science is the idea of quantum bits or qubits \citep{schumacher1996sending}, which are the building blocks of quantum information in analogy to the bits of classical information theory.
+Central to quantum information science is the idea of quantum bits or qubits \citep{schumacher1996sending}, which are the building blocks of quantum information in analogy to the bits of classical information theory.
 Physical implementation of a qubit could be any quantum system with two well defined quantum states, as long as these states can be initialized, manipulated and readout \citep{divincenzo2000physical}.
-Popular qubit candidates include superconducting circuits \citep{nakamura1999coherent, orlando1999superconducting}, photons \citep{bennett1992quantum, zhong2020quantum}, quantum dots \citep{veldhorst2014addressable, zajac2018resonantly}
+Popular qubit candidates include superconducting circuits \citep{nakamura1999coherent, orlando1999superconducting}, photons \citep{bennett1992quantum, zhong2020quantum}, quantum dots \citep{veldhorst2014addressable, zajac2018resonantly}, trapped ions \citep{friis2018observation, wright2019benchmarking} and single crystal defects \citep{jelezko2006single, baranov2011silicon, zhong2015optically}.
+
+Defects in large band gap crystals are sometimes referred to as artificial atoms \citep{buluta2011natural}, as the localized wave function of the electrons surrounding the impurity  give rise to discrete energy levels similar to that of an atom.
+Of all the possible crystals and defects, the most studied defect for quantum applications is the negatively charged nitrogen-vacancy center in diamond \citep{aharonovich2016solid, de2021materials}, often abbreviated to ``NV center".
+Although this particular defect of diamond had been known and correctly identified almost 50 years ago \citep{davies1976optical}, interest in the NV center really sparked after it was first isolated in 1997 \citep{gruber1997scanning}.
+The detection of single NV centers proved that NV centers fluoresce brightly \citep{gruber1997scanning}, are photostable \citep{kurtsiefer2000stable}, and that their electronic spin can be polarized and readout optically \citep{jelezko2004observation}, making them an promising solid state qubit.
+
+Since then, single NV centers have been successfully used to create quantum entanglement over several km \citep{hensen2015loophole}, to store quantum information at room temperature for more than one second \citep{maurer2012room}, to detect single proteins via nuclear magnetic resonance \citep{lovchinsky2016nuclear}, and for many other applications in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} or quantum sensing \citep{degen2017quantum}.
+
+As the interest in single NV center grew, diamond synthesis improved significantly \citep{achard2020chemical, barry2020sensitivity, edmonds2020generation}  which benefited not only single NV centers but also NV center ensemble.
+While single NV center are required for most quantum information processes \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble.
+Some of these applications are pictured in Fig.
+
+The first application for NV center ensemble is the use of nano or micro diamonds as fluorescent biomarkers.
+
+While most of the NV ensemble experiments work under the assumption that the ensemble is constituted of $N$ independent single NV centers
 

### 2022-09-29 Manuscrit Thèse/Intro et conclusion/Intro.tex
tant de temps ...

@@
 As the interest in single NV center grew, diamond synthesis improved significantly \citep{achard2020chemical, barry2020sensitivity, edmonds2020generation}  which benefited not only single NV centers but also NV center ensemble.
-While single NV center are required for most quantum information processes \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble.
-Some of these applications are pictured in Fig.
+While single NV center are required for most quantum information processing \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble.
+Four of these applications are pictured in Fig.
+[image bio : mohan2010vivo]
 
-The first application for NV center ensemble is the use of nano or micro diamonds as fluorescent biomarkers.
+The first application for NV center ensemble is the use of nano or micro diamonds as fluorescent biomarkers \citep{fu2007characterization, mohan2010vivo}.
+NV centers are good candidates for biomarkers as they emit a bright and stable photoluminesecence, and because diamond nanocrystals present low toxicity \citep{schirhagl2014nitrogen}.
+Some applications also use the spin degree of freedom the NV centers, either to perform background free fluorescence microscopy \citep{chapman2013background} or to monitor the angular motion of the nanodiamonds \citep{mcguinness2011quantum, feng2021association}.
+For most of these applications, NV ensemble provide a better visibility than single NV centers and are therefore preferred.
 
-While most of the NV ensemble experiments work under the assumption that the ensemble is constituted of $N$ independent single NV centers
+The second NV ensemble application is for quantum sensing \citep{degen2017quantum}.
+NV centers are predominantly used as magnetic field sensors \citep{rondin2014magnetometry, barry2020sensitivity}, although they can also detect electric fields \citep{dolde2011electric, michl2019robust}, temperature \citep{acosta2010temperature, kucsko2013nanometre} ,crystal strain \citep{ovartchaiyapong2014dynamic, doherty2014electronic} or rotation \citep{ledbetter2012gyroscopes, ajoy2012stable}.
+Both single and ensemble of NV centers can be used for sensing: single centers offer unmatched spatial resolution \citep{mitchell2020colloquium} down to a few nm \citep{lovchinsky2016nuclear}, while ensemble provide better sensitivities due to the larger number of probes \citep{wolf2015subpicotesla, barry2020sensitivity}.
+
+The third NV ensemble application is nuclear magnetic resonance (NMR) dynamical nuclear polarization \citep{eills2022spin}.
+Dynamic nuclear polarization consists in increasing the nuclear polarization of a sample by transferring electron polarization to the nuclei bath \citep{abragam1978principles}.
+Because the NV electronic spin can be optically polarized at room temperature, micro and nano diamonds containing NV centers are promising candidates as hyperpolarization agents \citep{tetienne2021prospects}.
+Dynamic nuclear polarization of nuclear spins inside the diamond lattice was demonstrated with single NV centers \citep{jacques2009dynamic, smeltzer2009robust} and NV ensemble \citep{king2015room, scheuer2016optically, schwartz2018robust}.
+Polarization transfer to nuclear spins outside of the diamond is an ongoing challenge \citep{eills2022spin, healey2021polarization}, but dense NV ensemble are believed to be necessary to achieve high external polarization \citep{tetienne2021prospects, rizzato2022polarization}.
+
+Finally, the fourth NV ensemble application is the recent field of spin mechanics \citep{perdriat2021spin}.
+Similarly to optomechanics \citep{aspelmeyer2014cavity}, the goal of spin mechanics is to couple a quantum system to the mechanical degree of freedom of a macroscopic object.
+Strong coupling between single NV center spin and external mechanical oscillators has been observed \citep{rabl2009strong, kolkowitz2012coherent}, but coupling of the diamond motion with NV centers spin has so far only been achieved with dense NV ensemble \citep{delord2020spin, pellet2021magnetic, perdriat2022angle}.
+Coupling several NV centers to the same mechanical oscillator could also lead to interesting magnetic phase transitions \citep{wei2015magnetic, ma2017proposal}.
+
+Most of the NV ensemble experiments discussed previously work under the assumption that the ensemble is constituted of $N$ independent single NV centers.
+However, as often in physics, \textit{more is different} \citep{anderson1972more}.
+Interaction between the NV center themselves or with their environment can lead to significantly different behavior between ensemble and single spins \citep{dobrovitski2008decoherence, choi2017observation, kucsko2018critical, giri2018coupled}.
+
+The work presented in this manuscript focuses on some of the
 

### 2022-10-01 Manuscrit Thèse/Intro et conclusion/Intro.tex
intro done

@@
 
-\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
 \title{}
@@
 
-Quantum mechanics was first developed at the beginning of the 20th century as an answer to  .
+Quantum mechanics was first developed at the beginning of the 20th century to explain physical observations such as the blackbody radiation \citep{planck1900theorie} or the photoelectric effect \citep{einstein1905erzeugung}.
+The quantum theory was subsequently developed in the first half of the 20th century and showed great success, in particular in giving a physical framework for chemistry \citep{pauling1931nature} and leading to a microscopic understanding of matter \citep{kittel1996introduction}.
 In the second half of the 20th century, technological development based on the new understanding of light and matter have led to what was retroactively called the first quantum revolution \citep{thew2019focus}.
-Technological developments in the 50s and 60s include the semiconductor transistor, the laser, and nuclear magnetic resonance  which have since revolutionized the fields of information science, global communications and medical imaging.
+Technologies developed in the 50s and 60s include the semiconductor transistor \citep{bardeen1948transistor}, the laser \citep{maiman1960stimulated}, and nuclear magnetic resonance \citep{abragam1961principles} which have since revolutionized the fields of information science, global communications and medical imaging.
 
 A new field of quantum technology has emerged in the last two decades.
-This ``second quantum revolution" differs from the first one in the usage of individual quantum systems, such as single photons, atoms or electrons \citep{peil1999observing}, which have properties that can not be emulated by larger classical systems.
+This ``second quantum revolution" differs from the first one in the usage of individual quantum systems, such as single photons \citep{gleyzes2007quantum}, atoms \citep{neuhauser1980localized} or electrons \citep{peil1999observing}, which have properties that can not be emulated by larger classical systems.
 These properties include quantum entanglement, superposition or measurement and form the basis of the newly formed quantum information science \citep{nielsen2002quantum, vedral2006introduction, hayashi2006quantum}.
@@
 Of all the possible crystals and defects, the most studied defect for quantum applications is the negatively charged nitrogen-vacancy center in diamond \citep{aharonovich2016solid, de2021materials}, often abbreviated to ``NV center".
-Although this particular defect of diamond had been known and correctly identified almost 50 years ago \citep{davies1976optical}, interest in the NV center really sparked after it was first isolated in 1997 \citep{gruber1997scanning}.
+Although this defect had been known and correctly identified almost 50 years ago \citep{davies1976optical}, interest in the NV center really sparked after it was first isolated in 1997 \citep{gruber1997scanning}.
 The detection of single NV centers proved that NV centers fluoresce brightly \citep{gruber1997scanning}, are photostable \citep{kurtsiefer2000stable}, and that their electronic spin can be polarized and readout optically \citep{jelezko2004observation}, making them an promising solid state qubit.
@@
 As the interest in single NV center grew, diamond synthesis improved significantly \citep{achard2020chemical, barry2020sensitivity, edmonds2020generation}  which benefited not only single NV centers but also NV center ensemble.
-While single NV center are required for most quantum information processing \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble.
+While single NV center are required for most quantum information processes \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble.
 Four of these applications are pictured in Fig.
-[image bio : mohan2010vivo]
+\ref{shema introduction}.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.9\textwidth]{Figures/Fig_intro}
+\caption{Example of applications with dense ensemble of NV centers.
+(a) In vivo nanodiomond biomarkers (from \citep{mohan2010vivo}).
+(b) Quantum sensing with NV ensemble (from \citep{ku2020imaging}).
+(c) Dynamic nuclear polarization of an external spin bath (from \citep{healey2021polarization}).
+(d) Spin mechanics with trapped micro diamonds (from \citep{delord2020spin}).}
+\label{shema introduction}
+\end{figure}
 
@@
 NV centers are good candidates for biomarkers as they emit a bright and stable photoluminesecence, and because diamond nanocrystals present low toxicity \citep{schirhagl2014nitrogen}.
-Some applications also use the spin degree of freedom the NV centers, either to perform background free fluorescence microscopy \citep{chapman2013background} or to monitor the angular motion of the nanodiamonds \citep{mcguinness2011quantum, feng2021association}.
+Some applications also use the spin degree of freedom of the NV centers, either to perform background free fluorescence microscopy \citep{chapman2013background} or to monitor the angular motion of the nanodiamonds \citep{mcguinness2011quantum, feng2021association}.
 For most of these applications, NV ensemble provide a better visibility than single NV centers and are therefore preferred.
@@
 NV centers are predominantly used as magnetic field sensors \citep{rondin2014magnetometry, barry2020sensitivity}, although they can also detect electric fields \citep{dolde2011electric, michl2019robust}, temperature \citep{acosta2010temperature, kucsko2013nanometre} ,crystal strain \citep{ovartchaiyapong2014dynamic, doherty2014electronic} or rotation \citep{ledbetter2012gyroscopes, ajoy2012stable}.
-Both single and ensemble of NV centers can be used for sensing: single centers offer unmatched spatial resolution \citep{mitchell2020colloquium} down to a few nm \citep{lovchinsky2016nuclear}, while ensemble provide better sensitivities due to the larger number of probes \citep{wolf2015subpicotesla, barry2020sensitivity}.
+Both single and ensemble of NV centers can be used for sensing: single centers offer unmatched spatial resolution \citep{mitchell2020colloquium} down to a few nm \citep{lovchinsky2016nuclear}, while ensemble provide better sensitivities \citep{wolf2015subpicotesla, barry2020sensitivity}.
 
@@
 Because the NV electronic spin can be optically polarized at room temperature, micro and nano diamonds containing NV centers are promising candidates as hyperpolarization agents \citep{tetienne2021prospects}.
-Dynamic nuclear polarization of nuclear spins inside the diamond lattice was demonstrated with single NV centers \citep{jacques2009dynamic, smeltzer2009robust} and NV ensemble \citep{king2015room, scheuer2016optically, schwartz2018robust}.
+Nuclear polarization of spins inside the diamond lattice was demonstrated with single NV centers \citep{jacques2009dynamic, smeltzer2009robust} and NV ensemble \citep{king2015room, scheuer2016optically, schwartz2018robust}.
 Polarization transfer to nuclear spins outside of the diamond is an ongoing challenge \citep{eills2022spin, healey2021polarization}, but dense NV ensemble are believed to be necessary to achieve high external polarization \citep{tetienne2021prospects, rizzato2022polarization}.
@@
 Similarly to optomechanics \citep{aspelmeyer2014cavity}, the goal of spin mechanics is to couple a quantum system to the mechanical degree of freedom of a macroscopic object.
-Strong coupling between single NV center spin and external mechanical oscillators has been observed \citep{rabl2009strong, kolkowitz2012coherent}, but coupling of the diamond motion with NV centers spin has so far only been achieved with dense NV ensemble \citep{delord2020spin, pellet2021magnetic, perdriat2022angle}.
+Strong coupling between single NV center and an external mechanical oscillators has been observed \citep{rabl2009strong, kolkowitz2012coherent}, but coupling between NV centers and the motion of the diamond has so far only been achieved with dense NV ensemble \citep{delord2020spin, pellet2021magnetic, perdriat2022angle}.
 Coupling several NV centers to the same mechanical oscillator could also lead to interesting magnetic phase transitions \citep{wei2015magnetic, ma2017proposal}.
@@
 However, as often in physics, \textit{more is different} \citep{anderson1972more}.
-Interaction between the NV center themselves or with their environment can lead to significantly different behavior between ensemble and single spins \citep{dobrovitski2008decoherence, choi2017observation, kucsko2018critical, giri2018coupled}.
+Interaction between the NV center themselves or with their environment can lead to significantly different behavior between ensemble and single spins.
+These difference include a modification of the charge state dynamics \citep{giri2018coupled} and of the spin dynamics \citep{dobrovitski2008decoherence, jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization}, as well as the apparition of exotic phenomena such as time crystal phase \citep{choi2017observation}, Anderson localization \citep{kucsko2018critical} or superradiance \citep{bradac2017room, angerer2018superradiant}.
 
-The work presented in this manuscript focuses on some of the
+The work presented in this manuscript focuses on dipolarly coupled dense ensemble of NV centers.
+More specifically, we look at spin polarization exchange between the NV centers themselves and with their environment.
+Polarization exchange (flip-flop, spin diffusion, cross-relaxation, ...) in dense ensemble of spins has been abundantly studied in the context of NMR \citep{abragam1978principles} and spintronics \citep{vzutic2004spintronics}.
+Because NV centers tend to be relatively dilute, the question of spin exchange between between NV themselves \citep{choi2017depolarization} or with their environment \citep{hall2016detection} has only recently been addressed.
+However, as samples with higher NV concentration become available \citep{acosta2009diamonds, tallaire2020high, shenderova2019synthesis}, these aspects are becoming more and more relevant.
+
+The goal of this present work is both to better understand the physics of dense NV ensemble, which is crucial for many of the NV ensemble applications cited previously, and to exploit some of these many-body properties for new applications.
+
+\bigskip
+This manuscript is organized in the following way:
+
+\medskip
+The first chapter introduces the theoretical and experimental concepts used in this manuscript.
+This includes a brief overview of diamond and NV centers fabrication processes, the physics of single NV centers and a presentation of magnetic dipole-dipole interaction and cross-relaxation.
+
+The second chapter covers the observation of cross-relaxation between NV centers and other spin impurities in CVD-grown diamond.
+We use measurements on the NV centers to identify these impurities and to quantify some of their attributes.
+Most of the results of this chapter have been published in  \citep{pellet2021optical} and \citep{ngambou2022improving}.
+
+The third chapter focuses on the dipole-dipole mediated spin relaxation within dense ensemble of NV centers.
+Following the work presented in \citep{choi2017depolarization}, we introduce the concept of NV-flutcuator to explain this phenomenon and present experimental and theoretical results solidifying this hypothesis.
+The results of this chapter have been partly published in \citep{pellet2022spin} and \citep{pellet2021magnetic}.
+
+Finally, the fourth chapter presents a study on NV-NV dipolar relaxation under low magnetic field.
+This chapter covers both the rich physics of NV centers under low magnetic field, and a potential application of the low field depolarization as a new magnetometry protocol, as well as comparisons with other NV magnetometry protocols.
+The results of this chapter have in large part been published in \citep{pellet2022spin}.
 

### 2022-10-03 Manuscrit Thèse/chapter 2/Chapter 2.tex
ghuk
bikldf

@@
 \centering
-\includegraphics[width=\textwidth]{Figures/tronche_de_spin}
+\includegraphics[width=.7\textwidth]{Figures/tronche_de_spin}
 \caption{Crystalline structure of three spin defects: the NV center, the P1 enter (substituional Nitrogen) and the VH center.} 
@@
 The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below.
-We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the process, and while we do not know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
+We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the growth process, and while we do not know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
 
@@
 
-For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively, with a slightly stronger confidence for the VH$^-$ one.
+For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively.
 
@@
 
-This section exploit the CR spectrum shown in Fig. \ref{CR VH exp} in order to extract information on the  .
+This section exploit the CR spectrum shown in Fig. \ref{CR VH exp} in order to extract information on the dark spins Hamiltonian parameters, linewidth and concentration.
 
@@
 
-This procedure will be discussed in more details in the next chapter. We find here that the linewidth of VH$^-$ and War1 are respectively [REF] and [REF] MHz.
+This procedure will be discussed in more details in the next chapter.
+We find here that the linewidth of VH$^-$ and War1 are respectively (half width at half maximum) $5.7\pm 0.6$ and $15.7\pm 1.6$ MHz \footnote{We follow here the pseudo-Voigt deconvolution method described in [REF].
+The linewidth $W=2.7$ MHz (half witdth at half maximum) of the NV centers was measured from ODMR (see Fig.
+[REF]).
+It includes all three hyper-fine transistions}.
 
@@
 
-Finally, the third tentative estimate we can make is that of the dark spins concentration.
+Finally, the third estimate we can make is that of the dark spins concentration.
 We can indeed calibrate the measurement by using the $^{13}$C-NV$^-$ line, of which we know the concentration in natural abundance diamond: [$^{13}$C-NV$^-$] $\approx$ 3.3 \% [NV].
 We can then, knowing the NV concentration, try to estimate the concentration of the dark spins.
-There are several caveats though:
+There are however several caveats:
 
@@
 
-With all this considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):
+From these considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):
 
-For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose the dark to bright NV ratio value from \citep{choi2017depolarization} at 0.25.
+For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose a dark to bright NV ratio of 0.25 as found in \citep{choi2017depolarization}.
 
@@
 
-We can see that the VH$^-$ line in particular completely disappear on sample CVD-B, which we attribute to the additional annealing.
+We can see that the VH$^-$ line in particular completely disappears on sample CVD-B, which we attribute to the additional annealing.
 This measurement also corroborate modifications in the infrared absorption spectrum, and an increase of the NV coherence time, associated with the annealing at higher temperature \citep{ngambou2022improving}.
@@
 In this case, the substrate was a type 1B HPHT diamond, which was irradiated and annealed along the CVD layer.
-We Will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.
+We will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.
 
@@
 To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations.
-This detection can be way more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up.
-It is however not as versatile and should be seen as a complimentary tool for diamonds with dense NV ensemble.
+This detection can be significantly more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up.
+This technique however is only applicable to a few defects in NV-rich diamond, meaning that its main application would probably be to characterize the environment of samples with high NV density.
 
-We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan, much less noisy than the [111] scans for weak magnetic field.
+We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan which provides a significantly less noisy signal than the [111] scans for weak magnetic field.
 
-Finally we have seen an application of the CR to detect optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.
+Finally we have seen an application of the CR to detect and characterize optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.
 
-There still remains a lot of open questions and there are several experiments that could be build upon these results. I will list some of them here.
+There still remains a lot of open questions and there are several experiments that could be build upon these results. Some of these questions are listed below.
 
@@
 We found that NV-VH$^-$ CR in CVD could produce a change in PL of $\sim 0.2 \%$.
-This kind of values are not too far from ODMR contrast with NV ensembles, which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin.
+This value is within one order of magnitude from ODMR contrast with NV ensembles (typically $<5 \%$), which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin.
 The main issue comes from the fact that in order to measure the VH$^-$ spin states, it needs to be resonant with the NV center.
@@
 
-There are two solution to circumvent this problem: the first is to use a fast switching magnetic field in order to bring the VH$^-$ and NV centers in resonance when you want to polarize or readout the VH$^-$ state, and out of resonance when you want to coherently manipulate the VH$^-$.
-This means that the magnetic field needs to be switched faster than the spin dynamics you want to probe.
-This might be doable for the population dynamics $T_1$ but this is probably too slow to probe the $T_2$ time.
+There are two solution to circumvent this problem: the first is to use a fast switching magnetic field in order to bring the VH$^-$ and NV centers in resonance when initializing or reading out the VH$^-$ state, and out of resonance to coherently manipulate the VH$^-$.
+This means that the magnetic field needs to be switched faster than the probed spin dynamics, which may be doable for the population dynamics $T_1$ but is likely too slow to probe the $T_2$ time.
 
 The other solution consists in using the spin-1 nature of VH$^-$: when a NV$^-$-VH$^-$ CR occurs, only one of the transitions of VH$^-$ and NV$^-$ is brought to resonance.
-The other one ($\ket{0} \to \ket{-1}$ for the VH$^-$) is out of resonance.
+The other one ($\ket{0} \to \ket{-1}$ for the VH$^-$ and $\ket{0} \to \ket{+1}$ for the NV) is out of resonance.
 The idea would therefore be to keep the NV $\ket{0} \to \ket{-1}$ transition and VH $\ket{0} \to \ket{+1}$ transition resonant, and to coherently manipulate the VH $\ket{0} \to \ket{-1}$ transition.
-While the polarization on this transition, and the back-action it would produce on the NV centers, is not as strong as if you were manipulating the $\ket{0} \to \ket{+1}$ transition, it should still produce a measurable signal.
+Further studies are needed to assess the effectiveness of this scheme.
 
@@
 We estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$.
-This concentration is too high by about two orders of magnitude to isolate a single emitter with a diffraction-limited microscope.
+This concentration is about two orders of magnitude too high to contemplate isolating a single emitter with a diffraction-limited microscope.
 We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing.
 We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and optically isolate a single VH$^-$ center by scanning a confocal microscope with a fixed $B_{[100]}=56\ \rm G$ while monitoring the NV PL or $T_1$.
-It is however unclear to me whether this detection scheme would be sensitive enough to detect single dark spins.
+It is however still unclear whether the present detection scheme is sensitive enough to detect single VH$^-$.
 

### 2022-10-03 python_clément/analyse.py

@@ -1290,32 +1290,32 @@ def sort_y_by_x(x,y):
 	x=sorted(x)
 	return(x,y)
 
-def dichotomy(f,target,xmin,xmax,precision='auto'):
+def dichotomy(f,target,xmin,xmax,precision='auto',**fargs):
 	import time
 	tmax=10 #s
 
-	assert (f(xmax)-target)*(f(xmin)-target) < 0
+	assert (f(xmax,**fargs)-target)*(f(xmin,**fargs)-target) < 0
 
-	if f(xmax)-target > 0:
+	if f(xmax,**fargs)-target > 0:
 		pass
 	else :
 		xmin,xmax=xmax,xmin #S'arrange pour que f(xmax)> target et f(xmin)< target
 
 	if precision=='auto':
-		precision=abs((f(xmax)-target))*1e-10
+		precision=abs((f(xmax,**fargs)-target))*1e-10
 
 
 	t=time.time()
 	ctr=0
-	delta=f(xmax)-f(xmin)
+	delta=f(xmax,**fargs)-f(xmin,**fargs)
 	while delta > precision:
 		xmid=(xmin+xmax)/2
-		if f(xmid)>target:
+		if f(xmid,**fargs)>target:
 			xmax=xmid
 		else :
 			xmin=xmid
 
-		delta=f(xmax)-f(xmin)
+		delta=f(xmax,**fargs)-f(xmin,**fargs)
 		ctr+=1
 		if time.time()-10>t:
 			raise ValueError('Took too long (iter=%i)'%ctr)

### 2022-10-04 Manuscrit Thèse/chapter 3/Chapter 3.tex
gjrkn
nblj

@@
 \chapter{NV-NV cross-relaxations : the fluctuator model}
-In this chapter we will look at the phenomenon of cross-relaxations between NV centers, both from an experimental and theoretical point of view.
+This chapter will cover cross-relaxations between NV centers in the non-zero magnetic field regime, both from an experimental and theoretical point of view.
+The case of zero magnetic field will be treated in the next chapter.
 
@@
 
-We will see here how a model of fast and slow relaxing centers, the fluctuator model \citep{choi2017depolarization}, can explain most of the experimental observations, and we will see some limitations of the model.
+We will see in this chapter how a model of fast and slow relaxing centers, the fluctuator model \citep{choi2017depolarization}, explains most observations related to the cross-relaxation, including the stretched exponential lifetime profile and the spectral broadening of the interaction.
 
 \section{Experimental observation of NV-NV CR}
-Before we discuss the theoretical implications of NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
+Let us start with the experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
+
+When dealing with NV-NV CR, we have to make the distinction between what we will label ``equivalent" and ``non-equivalent" NV centers.
+Two equivalent centers are NV centers with the same spin Hamiltonian ([REF] chap 1), which implies that they both see the same transverse and longitudinal magnetic field.
+In contrast, non-equivalent centers are center with different single-spin Hamiltonian, which may result in different eigenstates and spin polarization for the two spins.
+
+While the NV physics can explain cross-relaxations between non-equivalent NV centers, it cannot explain cross-relaxation between equivalent NV centers.
+Nevertheless, as we will see in this section, CR between both equivalent and non-equivalent NV centers is observed for samples with high NV density.
 \subsection{NV-NV CR between nonequivalent NV centers}
@@
 \includegraphics[width=\textwidth]{Figures/NV-NV_non_equivalent}
-\caption{a) Representation of two non equivalent spins resonantly coupled.
-One spin is aligned with the magnetic field and therefore well polarized, while the second spin is not.
-b) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
-c) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
-\label{non equivalent NV-NV}
+\caption{a) Illustration of the four possible direction of the the NV axis (four classes of NV centers) with a magnetic field along class 1 b) Representation of the polarization of two non equivalent spins, one aligned with the magnetic field (well polarized) and one misaligned with the magnetic field (unpolarized).
+c) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
+d) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
+\label{non-equivalent NV-NV}
 \end{figure}
-NV-NV CR was first observed more than thirty years ago \citep{holliday1989optical, van1989cross}.
-The first observations were between non-equivalent NV centers, meaning that the two NV centers involved in the dipole-dipole coupling were not polarized equally.
 
-This scenario can happen for instance when two NV centers from different classes see a different transverse (and longitudinal) magnetic field.
-Fig \ref{non equivalent NV-NV} illustrates this in the case where the magnetic field is parallel with one of the four classes: $\mathbf{B} \parallel [111]$
-
-When $\mathbf{B} \parallel [111]$, as we discussed in the last chapter, one class sees no transverse field and is therefore always polarized.
-The three other (equivalent) classes on the other hand get more and more depolarized as the magnetic field amplitude is increased.
-
-It turns out that there is a co-resonnance at B=592 G between the class parallel to $\mathbf{B}$ and the three other classes.
-This co-resonance is represented in Fig.
-\ref{non equivalent NV-NV}-b) by an orange circle.
+In the absence of magnetic gradient, two NV centers can only be non-equivalent if the projection of the magnetic field upon their respective NV axis is different.
+Fig.
+\ref{non-equivalent NV-NV}-a) illustrates such a case: we label here the four possible NV axis orientation (or NV classes) 1 to 4 and we assume that the magnetic field is aligned with class 1.
+In this conditions, an NV center belong to class 1 will always be non-equivalent with NV centers from classes 2-4, as the projection of the magnetic field on their respective axis will never be the same.
+On the other hand, NV centers from the classes 2-4 are all equivalents, even though they do not necessary belong to the same class.
 
 Fig.
-\ref{non equivalent NV-NV}-c), which we already saw in the last chapter, shows the change in PL of an ensemble of NV centers as $\mathbf{B}$ is scanned along the [111] axis.
-For B=592 G, we can see a drop in PL also circled in orange.
-This drop is the result of the CR between NV centers from the class parallel to $\mathbf{B}$ and NV centers from the three other classes.
+\ref{non-equivalent NV-NV}-b) shows the difference between the NV centers belonging to class 1 labeled NV$_\parallel$) and those belonging to classes 2-4 (labeled NV$_\nparallel$).
+As explained in [REF], the transverse magnetic field leads to a mixing of the eigenstates \footnote{The states of the misaligned NV on Fig.
+\ref{non-equivalent NV-NV}-b) are labeled $\ket{0},\ket{\pm 1}$ which is an abusive (but common) notation since the mixing caused by the transverse field means that the eigenstates of $S_z$ are no longer the Hamiltonian eigenstates.} and to a loss of the optical polarization mechanism.
+As a result, the NV$_\nparallel$ spins are less polarized than the NV$_\parallel$ ones, which is a necessary condition to observe cross-relaxation.
 
-While a difference in polarization is needed to observe a population transfer, it is not enough to explain the drop in PL.
-This PL drop is due to the difference in brightness between the $\ket{0}$ and $\ket{+1}$ from the different classes involved.
-The PL contrast between the two states is higher for the class parallel to $\mathbf{B}$ than for the three other ones, which is another consequence of the transverse field.
+Another necessary condition to observe cross-relaxation is the condition of co-resonance.
+In most situations, non-equivalent NV centers are not co-resonant, as the differences in the spin Hamiltonian generally leads to different eigenvalues.
+This is in contrast with equivalent NV centers which are by definition always resonant.
+Fig.
+\ref{non-equivalent NV-NV}-c) shows the transition frequencies of spin from class 1 and a spin from class 2-4 as a function of the magnetic field amplitude.
+Unintuitively, there is a co-resonance between the two $\ket{0} \to \ket{+1}$ transitions for $\abs{\mathbf{B}}=592\ \rm G$.
+For this particular value of the magnetic field, cross-relaxation between the NV centers from class 1 and those from classes 2-4 is possible.
 
-The co-resonance between two different classes of NV centers with different magnetic field projection is a relatively rare event: it can only occur for the $\ket{0} \to \ket{+1}$ transition, and only for magnetic fields greater than 592 G.
+Finally, Fig.
+\ref{non-equivalent NV-NV}-d), which was presented in the last chapter, shows the experimental value of the total NV PL as a function of the magnetic field amplitude, in a scenario similar to the one we just described.
+We can indeed observe the signature of NV-NV CR in the presence of a PL dip at $\abs{\mathbf{B}}=592\ \rm G$.
 
-%techniquement y'a les resonances 0->-1 et -1/+1 mais c'est le delbor, j'ai pas envie d'embrouiller plus que ca.
+As previously mentioned, cross-relaxation between non-equivalent NV centers is easily explained by the physics of single NV centers, and has been known for more than thirty years \citep{holliday1989optical, van1989cross}.
 
@@
  \label{fitting table}
-\begin{tabular}{c|cc|cc}
+\begin{tabular}{c|ccc|ccc}
 \toprule
-Figure &  \multicolumn{2}{c}{$B=0$} & \multicolumn{2}{c}{$B\neq0$}\\
+Figure &  \multicolumn{3}{c}{$B=0$} & \multicolumn{3}{c}{$B\neq0$}\\
 \midrule
-{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms)&  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)\\
-Fig. a) & 0.95 & * & 2.64 & * \\
-Fig. b) & * & 0.32 & * & 1.03  \\
-Fig. c) & 6.75 & 0.44 & 4.22 & 7.19 \\
-Fig. d) & \textbf{5} & 0.51 & \textbf{5} & 4.65 \\
+{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms) & $1-R^2$ &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)& $1-R^2$\\
+\midrule
+Fig. a) & 0.95 & * & $2\cdot 10^{-2}$ & 2.64 & * & $6\cdot 10^{-3}$ \\
+Fig. b) & * & 0.32 & $2\cdot 10^{-3}$& * & 1.03 & $2\cdot 10^{-2}$ \\
+Fig. c) & 6.75 & 0.44 & $6\cdot 10^{-4}$ & 4.22 & 7.19 & $2\cdot 10^{-4}$\\
+Fig. d) & \textbf{5} & 0.51 & $6\cdot 10^{-4}$ & \textbf{5} & 4.65 & $6\cdot 10^{-4}$\\
 \bottomrule
@@
 \includegraphics[width=\textwidth]{Figures/fluctuator linewidth}
-\caption{Measurement of the fluctuators linewidth using the same setup as in Fig.
+\caption{Measurement of the fluctuators linewidth on sample ADM-150-2 using the same setup as in Fig.
 \ref{equivalent NV-NV}.
 a) Simulation of the $\ket{0} \to \ket{-1}$ transition frequency for the four classes of NV centers and actual frequencies of the two central classes from ODMR measurements.
-b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.78\ \rm MHz$.
+b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.8 \pm 0.2\ \rm MHz$.
 The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor $\sqrt{2}$}
@@
 Green curve: deconvolution of the red curve the  by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38 MHz}
+All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz}
 \label{deconvolution}
@@
 \midrule
-random $\mathbf{B}$ (1 class) & $\Gamma_0^{\rm th}$ & 1.53 ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & 5.20 $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & 4.22 $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & 9.07 $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & 11.3 $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & 15.4 $\Gamma_0^{\rm exp}$ \\
+random $\mathbf{B}$ (1 class) & $\Gamma_0^{\rm th}$ & $1.53\pm 0.04$ ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & $5.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & $4.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & $11.6 \pm 0.4$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & $14.1 \pm 0.5$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & $19.9 \pm 0.8$ $\Gamma_0^{\rm exp}$ \\
 \bottomrule
@@
 
-What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators of the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2,3 or 4 classes are co-resonant as a function of $\Gamma_0$.
+What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators of the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2, 3 or 4 classes are co-resonant as a function of $\Gamma_0$.
 Doing so allows to normalize $n_f$ and $\gamma_f$ in eq.

### 2022-10-04 data/20220222/Adamas 15 um/traitement.py

@@ -141,7 +141,7 @@ def test_fits_t1_phonon(fname, valRef=None,fixed=True):
 
 	#En attendant si tu veux des valeurs, mieux vaut prendre des stretch arb. alpha=0.8 fonctionne bien
 	# T1ph=1/1800
-	T1ph=2.11
+	T1ph=5
 
 
 	
@@ -150,18 +150,18 @@ def test_fits_t1_phonon(fname, valRef=None,fixed=True):
 	x=x*1000
 	plt.plot(x,y)
 	# popt,yfit=exp_fit(x,y)
-	popt,yfit=stretch_et_phonons(x,y,T1ph=T1ph,fixed=fixed)
+	popt,yfit,pcov=stretch_et_phonons(x,y,T1ph=T1ph,fixed=fixed,err=True)
 	# popt,yfit=stretch_arb_exp_fit_zero(x,y,alpha=0.8,fixed=True)
 	plt.plot(x,yfit)
 	if valRef:
-		print(popt[1],valRef/popt[1])
+		print(popt[1],pcov[1],valRef/popt[1],valRef*pcov[1]/(popt[1]**2))
 		return(popt[1],valRef/popt[1])
 	else :
 		try :
 			print(popt[1],popt[2])
 			return(popt[1],popt[2])
 		except :
-			print(popt[1])
+			print(popt[1],pcov[1])
 			return(popt[1])
 
 
@@ -188,28 +188,31 @@ def test_fits_t1_phonon(fname, valRef=None,fixed=True):
 
 # fname='T1 0B court'
 
+
+
 # test_fits_t1_phonon('T1 C13 0B nuit',fixed=False) #T1dd=5.34, T1ph=2.11
 
-# valRef=test_fits_t1_phonon(fname='T1 1 classe pour le champ transverse moyen')
-# test_fits_t1_phonon(fname='T1 2x2 autre raie',valRef=valRef)
-# test_fits_t1_phonon(fname='T1 121 nuit',valRef=valRef)
-# test_fits_t1_phonon(fname='Vrai T1 1x3',valRef=valRef)
-# test_fits_t1_phonon(fname='T1 100',valRef=valRef)
-# test_fits_t1_phonon(fname='T1 0B long',valRef=valRef)
+valRef=test_fits_t1_phonon(fname='T1 1 classe pour le champ transverse moyen')
+print(1/0.64978,0.01475/0.64978**2)
[…]

### 2022-10-05 Manuscrit Thèse/chapter 3/Chapter 3.tex
hjitoe

@@
 (b) Magnetic setup used for the experiment: a permanent magnet is used to apply a bias magnetic field, and an electromagnet is used to add a variable magnetic field.
-(c) Simulation of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers as a function of the scanned magnetic field.
-The transitions were computed based on several ODMR spectra.
-The red dotted lines correspond to inter-class resonance (d) Change in the PL of the NV centers ensemble as a function of the scanned magnetic field.
-The inter-class resonances are reported here.}
+(c) Simulation showing the $\ket{0}\to\ket{-1}$ transition frequencies for the 4 classes of NV centers as a function of the scanned magnetic field.
+The transitions were computed based on ODMR spectra recorded for different magnetic field values.
+(d) Change in the PL of the NV centers ensemble as a function of the scanned magnetic field.}
 \label{equivalent NV-NV}
@@
 
-More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, jarmola2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there were cross-relaxations even between equivalent NV centers.
-
-I refer by equivalent NV centers to NV centers with the same spin Hamiltonian.
-This means either NV center from the same class, or NV centers from different classes with the same projection of the magnetic field along their axis.
-Equivalent NV centers are by nature resonant with each other, and therefore susceptible to CR.
+More recently, experiments on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there were cross-relaxations even between equivalent NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, jarmola2015longitudinal, choi2017depolarization}.
 
 Fig.
-\ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV CR.
-The key to observe NV-NV CR is to bring some NV centers in and out of resonance with other NV centers, which we can do with the different NV classes.
+\ref{equivalent NV-NV}-a) illustrates the two possible scenarios for equivalent NV centers: either two NVs from the same class, or two NVs from different classes with the same projection of the magnetic field upon their respective axis (i.e.
+$\mathbf{B}$ belongs to the plane of mirror symmetry of the two spins).
+
+In the second case, by moving the magnetic field in and out of the symmetry plane, we can change the resonance condition between the two NV classes and measure the contribution of NV-NV cross-relaxation to the NV PL and spin lifetime.
 To do so, we need to add an initial bias magnetic field in addition to the scanned magnetic field, as represented in Fig.
@@
 There are three magnetic field values $B_{\rm scan} \approx$ 12, 42 and 77 G for which there is an inter-class resonance.
-Each of these resonances correspond to a ``geometric" resonance where the $B$ field is in a symmetry plane between the two classes, meaning that the two resonant NV classes are indeed equivalent.
+These values are reported in the figure with red dotted lines.
+Unlike the case presented in Fig.
+\ref{non-equivalent NV-NV}, all the inter-class resonances reported in Fig.
+\ref{equivalent NV-NV}-c) are between equivalent classes.
 
@@
 
-This observation, and similar ones done by many groups \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization, akhmedzhanov2017microwave, giri2018coupled}, seem to indicate that there are CR between equivalent NV centers.
+This observation, and similar ones made by many groups \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization, akhmedzhanov2017microwave, giri2018coupled}, seem to indicate that there are CR between equivalent NV centers.
 This is however incompatible with our previous assumptions that equivalent NV centers are equally polarized and bright.
@@
 b) Average population of the two NV centers in the $\ket{0}$ state in the steady state as a function of $\Gamma_{\rm dd}$ and $\Gamma_1^*$.
-The steady state was computed by solving a classical rate equation where I fixed $\Gamma_{\rm las}=1000\ \rm Hz$ and $\Gamma_1=300\ \rm Hz$.}
+The steady state was computed by solving a classical rate equation where we fixed $\Gamma_{\rm las}=1000\ \rm Hz$ and $\Gamma_1=300\ \rm Hz$.}
 \label{inhomogene}
@@
 Fig.
-\ref{inhomogene} illustrates this case.
-Our model here consists of two NV centers resonantly coupled with two separate relaxation rates $\Gamma_1$ and $\Gamma_1^*$.
-To complete the model, we add the rate $\Gamma_{\rm las}$ which represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins.
-We only consider here the incoherent dynamics of the population, modeled with rates.
-We also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.
+\ref{inhomogene}-a) and b) illustrates this case.
+Our model here consists of two (resonantly coupled) NV centers with two different relaxation rates $\Gamma_1$ and $\Gamma_1'$.
+To complete the model, we add the rate $\Gamma_{\rm las}$ which is the optical pumping rate from the $\ket{-1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins \footnote{We consider here a two level system instead of the full three levels of the spin 1.
+The choice of $\ket{-1}$ instead of $\ket{+1}$ is arbitrary.}.
+We only consider here the incoherent dynamics of the population, modeled with the rates reported on Fig.
+\ref{inhomogene}-a).
+Solving for the steady state of these rate equations, we can compute the final population in the $\ket{0}$ state for both spins.
 
-By finding the steady state of these rate equations, we can compute the final population in the $\ket{0}$ for both spins, which we will assume to be proportional to the total PL.
 Fig.
-\ref{inhomogene}-b) shows this $\ket{0}$ population as a function of the flip-flop rate $\Gamma_{\rm dd}$ for various values of $\Gamma_1^*$.
+\ref{inhomogene}-b) shows the spin population in the $\ket{0}$ state of both spins, $\rho_{00}$, as a function of the flip-flop rate $\Gamma_{\rm dd}$ and for various values of $\Gamma_1'$.
+The first thing to notice is that, as $\Gamma_1'$ is increased,  $\rho_{00}$ decreases.
+This is because NV$_2$ becomes less polarized as its decay rate increases.
+The second thing to notice is that, when $\Gamma_1'=\Gamma_1$, i.e.
+when the two NV centers are truly equivalent, the flip-flop rate has no incidence on  $\rho_{00}$.
+However, when $\Gamma_1'>\Gamma_1$,  $\rho_{00}$ decreases when the flip-flop rate increases.
+Intuitively, this can be thought of as NV$_2$ acting as a polarization drain on NV$_1$.
 
-We can see that when $\Gamma_1^*=\Gamma_1$, as expected, the PL is not modified by the coupling of the two spins regardless of the coupling strength.
-When $\Gamma_1^*>\Gamma_1$ however, not only is the starting PL slightly lower since NV$_2$ is less polarized, but the PL now decreases as we increase the coupling strength.
-This effect, which is at the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is important.
+We can also notice that the influence of the flip-flop rate on $\rho_{00}$ increases when the difference in decay rates between NV$_1$ and NV$_2$ increases.
+As a result, we can suppose that a greater inhomogeneity between the NV centers would result in a stronger NV-NV cross-relaxation effect.
 
 \subsection{Presentation of the fluctuator model}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.8\textwidth]{Figures/NV_vs_fluctuator}
+\caption{Representation of a ``normal" NV center with spin depolarization rate $\Gamma_1 \ll \Gamma_{\rm las}$ and a fluctuator NV with spin depolarization rate $\gamma_f \gg \Gamma_{\rm las}$.
+Red circles represent the spin population in the steady state of each spin}
+\label{NV vs fluct}
+\end{figure}
 Choi et al.
-in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism.
-We are assuming that the fluctuator relaxation rate, $\gamma_f$ is much greater than the optical pumping rate $\Gamma_{\rm las}$.
-The fluctuators are therefore unpolarized and effectively act as dark spins, similarly to what we discussed in the last chapter.
+in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism leading to a new depolarization rate $\gamma_f$.
+Normal and fluctuator NV are represented in Fig.
+\ref{NV vs fluct}: in their model, Choi et al.
+assume that the fluctuator decay rate is fast enough compared to the laser polarization rate $\gamma_f \gg \Gamma_{\rm las}$ that the fluctuators are effectively always depolarized.
+Fluctuators are therefore similar to the dark spins we introduced in the last chapter.
 
@@
 
-We will consider the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
-With this assumption, we can compute the modification caused by the fluctuators on the rest of the NV lifetimes.
+In their model, the authors of \citep{choi2017depolarization} consider that the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
+With this assumption, the modification caused by the fluctuator bath on the ``normal" NV lifetimes can be analytically computed.
 
-Other possibilities to explain the change in the relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or superradiance, but they were not considered to be viable explanations \citep{choi2017depolarization}.
-%\begin{pmatrix}
-%0.33&0&0 \\
-%0&0.33&0 \\
-%0&0&0.33
-%\end{pmatrix}
+Other possibilities to explain the change in the relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or phonon superradiance, but they were not considered to be viable explanations: spin diffusion is orders of magnitude too slow to explain this phenomenon \citep{choi2017depolarization}, and the NV-NV CR seem to be independent of the crystal temperature \citep{jarmola2012temperature, mrozek2015longitudinal} which is incompatible with phonon superradiance \citep{choi2017depolarization}.
 
@@
 We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
-To compute the depolarization induced by the fluctuator bath the NV centers, we should first consider the interaction between a single NV center and a fluctuator.
+To compute the depolarization induced by the fluctuator bath on the NV centers, let us first consider the interaction between a single NV center and a fluctuator.
 Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.
 
-We will first decompose the dipole-dipole Hamiltonian in a radial and angular part:
+We will first decompose the dipole-dipole Hamiltonian as a product between radial and angular part:
 \begin{equation}
@@
 \end{equation}
-where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative position and orientation of the two dipoles.
+where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative radial positions and orientation of the two dipoles.
 
@@
 where $\gamma_f$ is the fluctuator decay rate for each possible channel between $\ket{0}, \ket{-1}$ and $\ket{+1}$.
-This number $\eta$ encapsulates both the geometry of the problem (through $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
+This number $\eta$ encapsulates both the angular dependency (through $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
 
@@
 The difference comes here from the fact that we consider thermalization going both ways (i.e.
-$\ket{e} \to \ket{g}$ and $\ket{g} \to \ket{e}$ instead of only $\ket{e} \to \ket{g}$) and that each state of the fluctuator can decay into two other states.
+$\ket{e} \to \ket{g}$ and $\ket{g} \to \ket{e}$ instead of only $\ket{e} \to \ket{g}$) and that each state of the fluctuator can decay into two other states (see Fig.
+\ref{NV vs fluct}).
 
-Finally, the additional depolarization rate induced by the fluctuator on the NV center reads:
+Finally, one finds that the additional depolarization rate induced by the fluctuator on the NV center reads:
 \begin{equation}
@@
 
-We now need to compute the average depolarization caused by the ensemble of fluctuators on the ensemble of NV centers.
+\begin{figure}[h]
+\centering
+\includegraphics[width=.4\textwidth]{Figures/NV_dans_bain_fluct}
+\caption{Schematics of an NV center in a fluctuator bath. The $\mathbf{r}_i$ vector represent the relative of position of fluctuator $i$ with respect to the NV center.}
+\label{NV + bain fluct}
+\end{figure}
 
-We consider each fluctuator as in independent depolarization source, therefore the depolarization on a single NV center reads: 
+The next step is to compute the depolarization caused by the ensemble of fluctuators on a single NV center.
+
+We consider the situation pictured in Fig.
+\ref{NV + bain fluct}: the NV center is at the center of the frame and the position of each fluctuator is labeled $\mathbf{r}_i$.
+We consider that each fluctuator is an independent depolarization source, therefore the depolarization on a single NV center reads:
 \begin{equation}

### 2022-10-10 Manuscrit Thèse/chapter 1/Chapter 1.tex
Le gros benj
gbrei

@@
 \usepackage{xcolor} % Pour avoir des couleurs dans les équations
-
-\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
 
@@
 \subsubsection{CVD growth}
+\label{fab HPHT}
 
@@
 \subsection{Experimental setup}
+\label{Sec setup cahp 1}
 \begin{figure}[h!]
@@
 \caption{Confocal microscopy setup detailed in main text}
-\label{setup}
+\label{setup ch1}
 \end{figure}
@@
 Fig.
-\ref{setup} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
+\ref{setup ch1} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
 Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.
@@
 \subsection{NV spin $T_2$}
+\label{sec T2 echo}
 \begin{figure}[h!]
@@
 \subsection{Cross-relaxations with NV centers}
-
+\label{Sec_CR}
 \begin{figure}[h]

### 2022-10-10 Manuscrit Thèse/chapter 2/Chapter 2.tex
gbrei

@@
 
-ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ \citep{mitchell2013x}.
-This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and from the more efficient photon detection in the optical range than in the microwave range.
+ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ spins \citep{mitchell2013x}.
+This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a spin with a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and in part from the more efficient photon detection in the optical range than in the microwave range.
 
-EPR however remains a far more versatile tool as only very few spin defects have the properties required for ODMR.
-The key idea of this chapter is to allow the optical detection of spin defects that do not show ODMR, by making them resonant with NV centers and observing NV cross-relaxation.
+The key idea of this chapter is to allow the optical detection of spin defects that normally do not show ODMR thanks to cross-relaxation with NV centers.
 
@@
 \ref{T1 vs PL} shows a simulation of NV centers PL when the spin decay rate $\Gamma_1$ is modified.
-In this fictional scenario, we assume that $\Gamma_1$ resonantly increases for a given value of the magnetic field (we will see later in this manuscript situations matching this scenario).
+In this fictional scenario, we assume that $\Gamma_1$ resonantly increases for a given value of the magnetic field (we will see later in this manuscript physical examples corresponding to this scenario).
 We can see that the PL decreases as an answer to the $\Gamma_1$ increase.
-The exact value of the PL drop ($\sim 2\%$ in the crase presented here) depends on the optical polarization rate $\Gamma_{\rm las}$, but this value generally does exceed a few \%, even for important $T_1$ modifications.
+The exact value of the PL drop ($\sim 2\%$ in the crase presented here) depends on the optical polarization rate $\Gamma_{\rm las}$ as well as the modification of the spin lifetime.
 
-It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the PL case, the laser intensity needs to be adjusted so that $\Gamma_{\rm las} \sim \Gamma_1$ in order to maximize the PL contrast, while the single$-\tau$ method requires a waiting time $\tau \sim T_1$ between each measurement, which result in a similar PL count and shot noise limit in both case.
+It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the PL case, the laser intensity needs to be adjusted so that $\Gamma_{\rm las} \sim \Gamma_1$ in order to maximize the PL contrast, while the single$-\tau$ method requires a waiting time $\tau \sim T_1$ between each measurement, which results in a similar PL count and shot noise limit in both case.
 
-On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and uses an overall smaller laser power, however it is more sensitive to drifts and changes in the optical setup.
+On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and because it requires a smaller laser power to work optimally.
+It is however more sensitive to drifts and changes in the optical setup.
 Most of the relaxometry measurement in this manuscript were performed using a PL-based detection.
@@
 
-This section focuses on particular field orientation with respect to the diamond crystalline axes, and details the advantages and disadvantages of the two main choices that are the [100] and [111] direction in order to detect four particular dark spins: P1 centers, Vh$^-$, War1 and the $^{13}C-NV$ complex\footnote{The $^{13}C-NV$ complex is normally not dark, as discussed in its own section.}.
+This section focuses on particular field orientation with respect to the diamond crystalline axes, and details the advantages and disadvantages of the two main choices that are the [100] and [111] direction in order to detect four particular dark spins: P1 centers, Vh$^-$, War1 and the $^{13}$C$-$NV complex\footnote{The $^{13}$C$-$NV complex is normally not dark, as will be discussed in its own section.}.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_NV}
+\includegraphics[width=1\textwidth]{Figures/121_vs_22_NV}
 \caption{Simulated transition frequencies of the $\ket{0} \to {-1}$ and $\ket{0} \to {+1}$ transitions for the 4 classes of NV centers as a function of the magnetic field.
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_P1}
+\includegraphics[width=1\textwidth]{Figures/121_vs_22_P1}
 \caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}

### 2022-10-10 Manuscrit Thèse/chapter 3/Appendix_eta.tex
gbrei

@@
 
+where $\eta$ is defined as:
+
 \begin{equation}
-\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
+\label{eq. eta debut annexe}
+\eta^2=\frac{1}{3} \left(\frac{\mel{i}{\mathcal{H_{\rm dd}}}{f}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
 \end{equation}
 
-\section{$\bar \eta$}
+and where $\mathcal{H_{\rm dd}}$ is the dipole-dipole Hamiltonian and $\ket{i}$ and $\ket{f}$ the initial and final two-qubits states of the flip-flop or double flip process.
 
+\section{$\bar \eta$ in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{Figures/Shema_annexe}
+\caption{Schematics of the ``normal" NV center NV$_s$ and the fluctuator NV$_f$ in and their respective basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$), as well as the relative position $\mathbf{r}(r,\theta,\phi)$ between the two spins.}
+\label{shema spins annexe}
+\end{figure}
+
+The computation of $\bar \eta$ when the single spin Hamiltonian of each spin is diagonal in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$ was treated in \citep{choi2017depolarization}.
+We will summarize their results here and consider a few different geometries than the ones presented in the article.
+
+Fig.
+\ref{shema spins annexe} represents the two spins and the three relevant angles in the problem $\theta,\phi$ and $\psi$.
+We label with $s$ the properties associated with the ``normal " NV center and $f$ those associated with the fluctuator.
+For instance the two-qubits state $\ket{m_s=0,m_f=+1}\equiv\ket{0,+1}$ corresponds to the convolution of the single-spin states $\ket{m=0}$ for the NV and $\ket{m=+1}$ for the fluctuator.
+
+The two Cartesian basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$) were chosen so that the spin vector Hamiltonian of each spin could be written:
+\begin{equation}
+\mathbf{S}_i=S_x \hat x_i + S_y \hat{y}_i + S_z \hat{z}_i,
+\end{equation}
+where :
+\begin{align*}
+S_x&=\begin{pmatrix}
+0&1/\sqrt{2}&0 \\
+1/\sqrt{2}&0&1/\sqrt{2} \\
+0&1/\sqrt{2}&0
+\end{pmatrix} \\
+S_y&=\begin{pmatrix}
+0&i/\sqrt{2}&0 \\
+-i/\sqrt{2}&0&i/\sqrt{2} \\
+0&-i/\sqrt{2}&0
+\end{pmatrix} \\
+S_z&=\begin{pmatrix}
+-1&0&0 \\
+0&0&0 \\
+0&0&+1
+\end{pmatrix}.
+\end{align*}
+
+The relative position between the two spins is noted with the vector $\mathbf{r}=r\hat{u}$.
+The vector's spherical coordinate are given in the Cartesian basis of the NV center ($\hat x_s,\hat y_s,\hat z_s$).
+
+The angle $\psi$ between the two spins $z$ axes is defined by the crystal lattice and can take three values within [0,$\pi$] which are: 0 (same class), $\cos^{-1}(1/3) \approx 70.5\ ^\circ$ and $\cos^{-1}(-1/3) \approx 109.5\ ^\circ$.
+Two distinct classes can either have a relative angle $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$ depending on the external magnetic field: we choose by definition that $\mathbf{B}\cdot\hat{z}_i >0$, which imposes the orientation of $z_i$ and therefore the angle between $\hat z_s$ and $\hat z_f$.
+This point will be further developed when discussing the $\{100\}$ and $\{110\}$ type resonances.
+
+The direction $\hat x_s=\hat x_f$ is chosen arbitrarily to be : $$\hat x=\frac{\hat z_s \times \hat z_f}{\norm{\hat z_s \times \hat z_f}}.$$
+The arbitrary choice of the $\hat x$ direction is justified by the symmetry of the problem in the ($\hat x, \hat y$) plane.
+We chose here to neglect the symmetry-breaking part of the transverse magnetic field which is justified by the fact that we do not take into consideration the mixing of the eigenstates caused by the transverse field.
+
+Following the notation introduced in \citep{choi2017depolarization}, we can rewrite the dipole-dipole Hamiltonian as:
+\begin{equation}
+\mathcal{H}_{\rm dd} \approx -\frac{J_0}{r^3}\left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}) + h.c. + q S_z^sS_z^f \right],
+\end{equation}
+with 
+\begin{align}
+g&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f) + 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{y}_s \cdot \hat{y}_f) \right] \\
+h&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{x}_s \cdot \hat{y}_f) - 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{x}_f) + (\hat{y}_s \cdot \hat{x}_f) \right] \\
+q&= 3(\hat{u}\cdot \hat{z}_s)(\hat{u}\cdot \hat{z}_f) - (\hat{z}_s \cdot \hat{z}_f).
+\end{align}
+
+We should note that the double flips have been omitted because we consider here the case $B\neq0$ for which the $\ket{+1}$ and $\ket{-1}$ states of each spin are far out of resonance.
+
+Eq. \ref{eq. eta debut annexe} can then be written:
+\begin{equation}
+\eta^2=\frac{1}{3} (\abs{g}^2 + \abs{h}^2) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{equation}
+
+In order to compute $\bar \eta$, we first decompose $\eta$ as a product of $R$ and $G$ as defined in [REF]:
+\begin{equation}
+\eta^2=G^2(\theta, \phi, \psi) R^2(\omega_f,\omega_{NV})
+\end{equation}
+with
+\begin{align*}
+G^2(\theta, \phi, \psi)&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
+R^2(\omega_f,\omega_{NV}) &= \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{align*}
+
+$R$ and $G$ can be averaged separately as they do not depend on the same variables, which means that:
+\begin{equation}
+\label{eq. eta bar r bar g bar}
+\bar \eta = \bar R \cdot \bar G.
+\end{equation}
+
+The computation of $\bar{R}$ has been discussed in [REF], we will focus here on $\bar{G}$ defined as:
+
+\begin{equation}
+\label{eq. eta base mag}
+\bar{G}= \iint \abs{G} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi}.
+\end{equation}
+
+Eq.
+\ref{eq.
+eta base mag} can be solved analytically for the case $\psi=0$ and numerically for $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$.
+The three values are reported in Table \ref{table G mag}.
+The values found for $\psi=0$ and $\psi=70.5 \ ^\circ$ are the same that were found by the authors of \citep{choi2017depolarization} \footnote{They differ by a factor $\sqrt{1/3}$ because of the definition of $G$.}.
+
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar G$ in the magnetic basis}
+ \label{table G mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+$\frac{2}{9}=0.222$ & 0.3757 & 0.4808 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Interclass resonance and magnetic field orientation}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.8\textwidth]{Figures/110_vs_100}
+\caption{(a) Illustration of the co-resonance condition between a NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
+The $\hat z$ direction is defined by the condition $\mathbf{B}\cdot\hat{z} >0$.
+(b) 8 individual $T_1$ measurements on sample ADM-15-3 on a two-class resonance for 8 different values of the magnetic field (4 $\mathbf{B} \in \{110\}$ and 4 $\mathbf{B} \in \{100\}$).}
+\label{110 vs 100 annexe}
+\end{figure}
+
+As discussed in sec [REF], there are 4 possibles magnetic field orientations for which at least two NV classes are co-resonant: $\mathbf{B} \in \{110\}$ (two-class resonance), $\mathbf{B} \in \{100\}$ (2$\times$two-class resonance), $\mathbf{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\mathbf{B} \parallel \langle 100\rangle$ (four-class resonance).
+
+Fig.
+\ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$.
+We consider here an NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
+Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\mathbf{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
+
+For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\mathbf{B} \in (100)$ or $\mathbf{B} \in (011)$.
+The angle $\psi$ between the $\hat z_s$ and $\hat z_f$ axes however is different in both cases.
+The same can be generalized for every $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\mathbf{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\mathbf{B} \in \{110\}$ type resonance.
+
+Because $\bar G$ is greater for $\psi=109.5 \ ^\circ$ than it is for $\psi=70.5 \ ^\circ$, then following eq.
+\ref{eq.
+eta bar r bar g bar}, we should expect a faster depolarization rate when $\mathbf{B} \in \{110\}$ than when $\mathbf{B} \in \{100\}$.
+Fig.
+\ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\mathbf{B} \in \{110\}$ scenario and four in the $\mathbf{B} \in \{100\}$ scenario.
+We can clearly see a slower relaxation rate for all four $\mathbf{B} \in \{100\}$ measurements compared to all four $\mathbf{B} \in \{110\}$, which agrees with the predictions of the model.
+
+For a perfect resonance matching (i.e.
+no detuning between the central frequencies of the different resonant classes), we can assume that $\bar R$ is a constant ($\bar R \equiv \bar R^0 \approx \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}$ according to eq.
+[REF]).
+We then have:
+
+\begin{align}
+\bar \eta(\rm{1\ class})&=\bar G(\psi=0) \bar R_0 \\[1em]
+\frac{\bar \eta(\mathbf{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
+\end{align}
+
+Finally, the values in Table [REF] are computed thanks to the relation: \begin{equation}
+\frac{\Gamma_1}{\Gamma_0}= \left(\frac{\bar \eta}{\bar \eta(\rm{1\ class})} \right)^2.
+\end{equation}
+
+Note that we only give here the relaxation rates for a single class.
+To estimate the total PL, one would need to compute the relaxation rate of all 4 classes independently, for instance in a $\mathbf{B} \in \{110\}$ scenario, two classes have a relaxation rate $\Gamma(\mathbf{B} \in \{110\})$ and two classes have a relaxation rate $\Gamma_0$.
 \printbibliography

### 2022-10-10 Manuscrit Thèse/chapter 3/Chapter 3.tex
gbrei

@@
 
+Before detailing the model's predictions, we will first contemplate the possible microscopic origin of these fluctuators, and review some physical systems were similar observations have been made.
+
+ \subsection{Possible microscopic origin of the fluctuators}
+
+We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}.
+It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.
+
+This conclusion was based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions.
+Similar observations were performed with Förster resonant energy transfer (FRET) in chromophore molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through (electric) dipole-dipole interaction to non-excited acceptor molecules.
+
+In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
+In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
+Both of these phenomena can lead to fast spin relaxation: electron tunneling can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction term ($J-$coupling) by the crystal phonons can directly excite the spin transitions \citep{honig1960electron}.
+While the dipole-dipole interaction terms (excluding the contact interaction) are also modulated by the phonons, this effect would be too small to account for the fast depolarization observed here.
+
+Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
+To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate between neighboring sites of $\sim 10\ \rm ns$.
+In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
+
 \subsection{Single NV center coupled to a single fluctuator}
-We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
-To compute the depolarization induced by the fluctuator bath on the NV centers, let us first consider the interaction between a single NV center and a fluctuator.
+We now turn to the computation of the depolarization induced by the fluctuator bath on the NV centers. We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
+
+Let us first consider the interaction between a single NV center and a fluctuator.
 Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.
@@
 
-In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential, unlike the phonon-induced depolarization which is purely exponential.
+In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential \footnote{Stretched exponential generally refer to functions of the form $\exp(-(t/\tau)^{\beta})$.
+In this manuscript, unless specified otherwise, we will always refer to the case $\beta=1/2$ when mentioning stretched exponential.}, unlike the phonon-induced depolarization which is purely exponential.
 
-It should be noted that the stretched exponential nature of the decay is not proper to the fluctuator model.
-\citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
-The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).
+It should be noted that the stretched exponential nature of the decay is not specific to the fluctuator model, but results from localized noise sources, randomly distributed in the bulk (3D).
+Indeed, \citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
+Theodor Förster in 1949 also predicted and observed stretched exponential dynamics in the context of FRET, where the same scaling laws are observed \citep{forster1949experimentelle}.
 
@@
 \label{sec computation eta}
+
 The parameter $\bar \eta$ in eq.
-\ref{eq 1/T} is of crucial importance since it encompass the dependence of $1/T$ in both the magnetic field and the inhomogeneous broadening $T_2^*$ of the NV centers and the fluctuators.
+(\ref{eq 1/T}) is of crucial importance as it encapsulates the resonance condition between the NV centers and the fluctuators.
+This resonance condition, i.e.
+the spectral overlap between the fluctuators and the NV centers, can be decomposed in two essential parameters: one is the central frequency of each subgroup ($\omega_{NV}^0$ and $\omega_{f}^0$ respectively) which themselves depends on the magnetic field, the other is the spectral broadening of both the NV and the fluctuators.
 
-We will start by separating the factor $\eta$ given in eq. \ref{eq eta} into a a geometric and a spectral part:
+The aim of this section is to give an explicit formula of $\bar \eta$ as a function of $\omega_{NV}^0$, $\omega_{f}^0$ and $T_2^*$.
+
+We will start by separating the factor $\eta$ given in eq. (\ref{eq eta}) into a a geometric and a spectral part:
 \begin{align}
-\eta^2&=G^2(\psi_{NV},\psi_f,\theta,\phi) R^2(\omega_f,\omega_{NV}), \\
+\eta^2&=G^2(\{\theta\}) R^2(\omega_f,\omega_{NV}), \\
 G^2&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
@@
 
-$G$ is a function which depends purely on the various angles of the problem (angles of the relative position of the NV center and the fluctuator, angle of the NV and fluctuator $z$ axis), while $R$ represents the resonance condition between the NV and the fluctuator.
+$G$ is a function which depends purely on the various angles of the problem $\{\theta\}$, while $R$ represents the resonance condition between the NV and the fluctuator.
 
-We will now average this $\eta$ factor for an ensemble of fluctuators and NV centers:
+We then average the $\eta$ factor for an ensemble of fluctuators and NV centers:
 \begin{align*}
-\bar \eta &= \int \rm{Prob}(\eta) \abs{\eta} d\eta, \\
-&= \left(\int \rho(\psi_{NV},\psi_f,\theta,\phi) |G(\psi_{NV},\psi_f,\theta,\phi)| d\psi_{NV} d\psi_f d\theta d\phi \right) \\
-&\, \times \left( \int \rho(\omega_{NV}) \rho(\omega_{f}) |R(\omega_f,\omega_{NV})| d\omega_{NV} d\omega_{f} \right), \\
+\bar \eta &= \int \rm{Prob}(\eta) \abs{\eta} d\eta \\
+&= \left(\int \rho(\{\theta\}) |G(\{\theta\})| d\{\theta\} \right) \\
+&\, \times \left( \int \rho(\omega_{NV}) \rho(\omega_{f}) |R(\omega_f,\omega_{NV})| d\omega_{NV} d\omega_{f} \right) \\
 &= \bar G \cdot \bar R,
 \end{align*}
-where we introduced $\bar G$ and $\bar R$, the averaging of $|G|$ and $|R|$.
+where we introduced $\rho(\{\theta\})$ the probability distribution of the angles $\{\theta\}$, $\rho(\omega_{NV})$ the inhomogeneous spectral distribution of the ensemble of NV centers and $\rho(\omega_{f})$ the inhomogeneous spectral distribution of the ensemble of fluctuators.
+We also introduced the notations $\bar G$ and $\bar R$ as the average of $|G|$ and $|R|$.
 
@@
 
-We unfortunately cannot analytically solve $\bar R$ for standard distributions of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ (either Gaussians or Lorentzians).
-We can however solve $\overline{R^2}$, which is relatively close to $(\bar{R})^2$ as long as $2 \gamma_f \gg 1/T_2^*$.
+We unfortunately cannot analytically solve for $\bar R$ with standard distributions of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ (either Gaussians or Lorentzians).
+We can however solve for $\overline{R^2}$, which is close to $(\bar{R})^2$ as long as $2 \gamma_f \gg 1/T_2^*$.
 
@@
 \end{align*}
-
-where $\Gamma_{f}=1/T_2^*(\rm fluctuator)$ and $\Gamma_{NV}=1/T_2^*(\rm NV)$ are the inhomogeneous broadening of the NV and fluctuator ensemble, and $\omega^0_{NV}$ and $\omega^0_{f}$ are the central angular frequencies of the NV and fluctuator ensemble.
+where $\Gamma_{f}=1/T_2^*(\rm fluctuator)$ and $\Gamma_{NV}=1/T_2^*(\rm NV)$ are the inhomogeneous broadening of the NV and fluctuator ensembles, and $\omega^0_{NV}$ and $\omega^0_{f}$ are the central angular frequencies of the NV and fluctuator ensembles.
 We then find that:
@@
 \label{eq 1/T avec T2*}
-\frac{1}{T} \propto \frac{1}{2\gamma_f + \Gamma_f + \Gamma_{NV}}.
+\frac{1}{T} \propto \frac{1}{2\gamma_f + \Gamma_f + \Gamma_{NV}}\cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
 \end{equation}
 
-If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\overline{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian ans a Gaussian, which does not have a clear analytic solution.
+If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\overline{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian and a Gaussian, which does not have a clear analytic solution.
 
 Eq.
-\ref{eq 1/T avec T2*} is subject to several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
-Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$.
-
- \subsection{Possible microscopic origin of the fluctuators}
-
-We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}.
-It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.
-
-This conclusion was based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions, and also to the resonant energy transfer between molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through dipole-dipole interaction to non-excited acceptors, and where the stretched exponential dynamics of the donors was first observed \citep{forster1949experimentelle}.
-
-In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
-In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
-Both of these phenomena could lead to fast spin relaxation, the tunneling because the spin hopping can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction strength $J$ by the crystal phonons \citep{honig1960electron}.
-While the non-contact dipole-dipole interaction is also modulated by the phonons, this effect would be too small to account for the fast depolarization observed here.
-
-Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
-To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate between neighboring sites of $\sim 10\ \rm ns$.
-In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
+\ref{eq 1/T avec T2*} was obtained via several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
+Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$, $\omega_{NV}^0$ and $\omega_{f}^0$, which will be crucial to compare the predictions of the model with experimental data.
 
@@
 \caption{T1 measurement following the protocol described in [REF].
-The best exponential and stretched exponential fits are given each time.
-a) Sample ADM-15-2 with B=0.
+The best exponential and stretched exponential fits are given in every case.
+a) Sample CVD-pink with B$\neq$0 and a single  probed class.
+The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$.
+b) Same measurement as a) in log scale.
+c) Sample ADM-15-2 with B=0.
 The optimal $T_1$ values are $T_1^{\rm stretch}=150\ \mu \rm s$ and $T_1^{\rm exp}=410\ \mu \rm s$.
-b) Sample CVD-pink with B$\neq$0 and a single class probed.
-The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$}
+d) Same measurement as c) in log scale.}
 \label{stretch_or_not_stretch}
 \end{figure}
-$T_1$ measurements with dense ensemble of NV centers and when many classes are resonant with each other (typically when B=0) indeed tend to show stretched exponential profiles.
+
+The first prediction of the fluctuator model which can be experimentally verified is the stretched-exponential profile of the spin relaxation.
+The protocol to measure spin relaxation with a common mode rejection was presented in [REF].
 
 Fig.
-\ref{stretch_or_not_stretch} shows an example of that: in Fig.
-\ref{stretch_or_not_stretch}-a) the $T_1$ measurement comes from sample ADM-15-2 at B=0.
-The profile is clearly better fitted by a stretched exponential than a regular exponential, and the timescale found $T_1^{\rm stretch}=150\ \mu \rm s$ is much shorter than the expected phonon limite lifetime of a few ms.
+\ref{stretch_or_not_stretch} shows the results of this protocol on two distinct samples and magnetic field configurations.
+In both cases, we show the results in linear and logarithmic scale as the log scale makes deviations from the standard exponential more apparent\footnote{The fits presented in Fig.
+\ref{stretch_or_not_stretch} and in the next figures were performed by fitting the real value of $S_1-S_2$ and not $\log(S_1-S_2)$.}.
 
-On the other hand, Fig.
-\ref{stretch_or_not_stretch}-b) shows a $T_1$ measurement coming from sample CVD-pink for B$\neq$0.
-This time the profile is more exponential, and the value $T_1^{\rm exp}=4.1\ \rm ms$ is coherent with a phonon-limited lifetime.
-Even though the sample CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.
+The spin relaxation results presented in Fig.
+\ref{stretch_or_not_stretch}-a) and b) was performed on sample CVD-pink ([REF]), which despite a relatively high NV concentration shows only moderate amount of dipole-dipole spin relaxation.
+In order to further minimize the amount of cross-relaxations, we chose a magnetic field for which all four NV classes are spectrally isolated, so that only fluctuators from the same class as the NV center could lead to relaxation.
+As a result, we can see a relaxation profile mostly exponential with a lifetime $T_1=4.1\ \rm ms$ comparable to the phonon-limited $T_1$ value of $\sim 5\ \rm ms$ at room temperature \citep{jarmola2012temperature}.
+
+The second spin relaxation presented in Fig.
+\ref{stretch_or_not_stretch}-c) and d) was performed on sample ADM-15-2 which showed a great amount of dipole-dipole spin relaxation.
+To further enhance the cross-relaxation rate, the experiment was performed in zero external magnetic field so that every class of NV centers would be resonant.
+We can observe a relaxation profile that deviates significantly from tthe exponential profile, and which is nicely fitted by a stretched exponential.
+The $T_1$ value obtained from the fit is also an order of magnetite lower than the phonon-limited value, which is coherent with the fact that dipole-dipole interactions dominate the spin relaxation in this case.
 
@@
 \caption{Fitting parameters for Fig.
- \ref{various_fit_formulae}}
+ \ref{various_fit_formulae}. The coefficient of determination $R^2$ for each fit is reported here.}
  \label{fitting table}
@@
 
-For many samples however, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$ is of the same order as the phonon limited lifetime $T_1^{\rm ph}$.
-Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, forcing us to include both aspects if we want a unique fitting formula.
+For many samples, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$, is of the same order of magnitude as the phonon limited lifetime $T_1^{\rm ph}$.
+Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, meaning that we have to include both aspects in the fitting formula.
 
 Fig.
-\ref{various_fit_formulae} shows four different fitting procedure on the two same experimental data.
-The data here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
+\ref{various_fit_formulae} shows four different fitting procedure on the two same relaxation measurements.
+The measurements were performed on the same sample ADM-150-1, with and without an external magnetic field ($\sim 50 G$).
 The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.
@@
 
-Since the protocol where $T_1^{\rm ph}$ was fixed and the one where it wasn't both correctly fit our data, we decided to use the protocol where $T_1^{\rm ph}$ was fixed because there is one less free parameter in the fitting function, and the values we obtain on $T_1^{\rm dd}$ can be directly compared.
+Both the protocol where $T_1^{\rm ph}$ was fixed and the one where it was not correctly fit our data.
+We therefore decided to use the protocol where $T_1^{\rm ph}$ was fixed in order to minimize the number of free parameter in the fitting function.
+This way, the values of $T_1^{\rm dd}$ obtained on different measurements can be directly compared.
 
@@
 \ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
-While this technique of measuring $T_1^{\rm dd}$ is useful to compare different cases, because it is fairly sensitive, it is not an exact measurement of the dipole-dipole induced lifetime.
+The values of $T_1^{\rm dd}$ obtained with this method have therefore to be be compared to values from the same method and the same fixed value of $T_1^{\rm ph}$.
 
@@
 \ref{betas}.
-This time we are using a single stretched exponential with an arbitrary stretch factor $\beta$.
+This time a single stretched exponential with an arbitrary stretch factor $\beta$ is used, and satisfyingly fits the data.
 Fig.
 \ref{betas}-b) shows the optimal $\beta$ parameter as a function of the external magnetic field.
-We can then confirm that the $T_1$ profile gets closer to a stretched exponential ($\beta=0.5$) when $B$ goes to 0.
-
-Even though this method also yields satisfying fits, it needs two free parameters to work ($\beta$ and $T_1$), so we decided to go with the method of Fig. \ref{various_fit_formulae}-d) instead.
+It confirms that the $T_1$ profile gets closer to a stretched exponential ($\beta=0.5$) when $B$ goes to 0.
+Even though this method also yields satisfying fits, it employs more free parameters ($\beta$ and $T_1$) than the method of \ref{various_fit_formulae}-d).
+We therefore decided to go with the later.
 
@@
 b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.8 \pm 0.2\ \rm MHz$.
-The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor $\sqrt{2}$}
+The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor of 2}
 \label{fluct linewidth}
@@
 
-One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators spectral response.
-Indeed, as we will see, the fluctuator lifetime is so short that it significantly broadens the fluctuators spectral response compared to standard NV centers.
+We now move to the second major prediction of the fluctuator model which is the spectral broadening of the fluctuators caused by the their relaxation rate $\gamma_f$.
+Indeed, following the arguments presented in \citep{choi2017depolarization}, the fluctuators homogeneous spectral response is a Lorentzian of half-width $2\gamma_f$.
+The inhomogeneous spectral response of the ensemble of fluctuators is then further broadened by the same electric, magnetic or strain noises than the NV centers, which is characterized by $1/T_2^*$.
 
-The fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in \citep{hall2016detection}.
+The fluctuator are effectively dark to ODMR measurement since they are not polarized, meaning that we cannot directly measure their linewidth as we would do for normal NV centers.
+We can however measure their linewidth through cross-relaxation, similarly to the detection of dark spins presented in the previous chapter.
 
@@
 \ref{equivalent NV-NV}.
-A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers, but this time measuring the spin decay rate instead of the PL.
+A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers.
+This time however, the spin relaxation was measured instead of the PL.
 
 Fig.
-\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field, and the experimentally measured central frequencies of the two classes of interest when the two lines can be clearly separated.
-Having a precise measurement of the detuning between the two classes $\Delta \nu$ is key in this experiment to get a precise value of the fluctuator linewidth.
+\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
+We took particular care in measuring the central frequencies of the two classes of interest (when the two lines could be clearly separated), since having a precise measurement of the detuning between the two classes $\Delta \nu$ is key to get a precise value of the fluctuator linewidth.
 %Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of their two central lines to get a precise measurement of the Larmor frequency.
@@
 We can clearly see that the $\Gamma^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
-If we assume that both NV classes spectral response are gaussians of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
+If we assume that both NV classes spectral response are Lorentzian of same width $\sigma$, then the spectral overlap between the two classes is also a Lorentzian of width $2 \sigma$ \footnote{For other line profile, the overlap would be smaller than $2 \sigma$.
+For instances, the overlap of two Gaussians of width $\sigma$ is a Gaussian of width $\sqrt{2} \sigma$}.
 Fig.
-\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $\sqrt{2}$, which is still much narrower than the $1/T_1^{\rm dd}$ line.
+\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $2$, which is still significantly narrower than the $1/T_1^{\rm dd}$ line.
 
 We attribute this broadening to the fluctuators lifetime.
-Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator, but for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
+Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator, but for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening observed here.
 
@@
 \ref{fluct linewidth}.
-Red curve: deconvolution of the blue curve the  by the ODMR line shown on the same figure.
-Green curve: deconvolution of the red curve the  by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz}
+Red curve: deconvolution of the blue curve by the ODMR line shown on the same figure.
+Green curve: deconvolution of the red curve   by the ODMR line.
+All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters)}
 \label{deconvolution}
@@
 Using the approximation $\bar {R ^2} \approx (\bar{R})^2$ detailed in sec.
-\ref{sec computation eta}, we can then decompose the fluctuator spectral response between the lifetime contribution and the inhomogeneous broadening:
+\ref{sec computation eta}, we can then decompose the fluctuator spectral response as a convolution of the lifetime contribution and the inhomogeneous broadening:
 \begin{equation}
@@
 \ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve).
-Fitting with Lorentzians, we find a final value $\Gamma_2 =2\gamma_f= (2\pi) 6.38\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=49.9\ \rm ns$.
+Fitting with Lorentzians, we find a final value $\Gamma_2 =2\gamma_f= (2\pi) 6.4 \pm 0.4\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=50 \pm 3\ \rm ns$\footnote{We define here the fluctuators lifetime $T_1^f$ as $\frac{1}{\gamma_f}$ for simplicity, but it should be noted that the actual lifetime of the fluctuators states is $\frac{1}{2\gamma_f}$ due to the competition between the two possible decay channels for each state (cf Fig.
+\ref{NV vs fluct}).}.
 
@@
 
-With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, we find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 48.7\ \rm ns$, a value consistent with the one found with the deconvolution method.
+With $f=8.8 \pm 0.2\ \rm MHz$ and $\sqrt{2}f_G=4.3 \pm 0.2\ \rm MHz$, we find $f_L=6.5 \pm 0.3\ \rm MHz$ for a final lifetime value $T_1^f = 49 \pm 2\ \rm ns$, a value consistent with the one found with the deconvolution method.
 
-Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48.2\ \rm ns$, a value surprisingly similar to the one we got.
+Doing similar experiments and calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48\ \rm ns$, a value very close to the ones we obtained above, despite using a completely different sample.
+
+In conclusion, we observed that the fluctuators linewidth was indeed broadened by their relaxation rate which justifies the initial hypothesis that $T_1^f \ll T_1^{NV}$.
+It even shows that $T_1^f$ is of the same order of magnitude as the optical lifetime ($\sim 15\ \rm ns$) which explains why the fluctuators cannot be optically polarized.
 
@@
 The main tool to probe NV-NV CR is to use the co-resonances between the NV classes.
-While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and supposedly resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances.
-We will discuss here the various scenarios where NV classes overlap, and the quantitative impact it has on the centers' spin $T_1$.
+While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances.
+In this section, we will discuss the various scenarios where NV classes overlap, and quantify theoretically and experimentally the dipole-induced relaxation rate in each case.
 
@@
 
-There are four possibilities to obtain inter-class resonance (outside of the non-equivalent CR discussed in sec. \ref{non_equi_valent_CR}). These possibilities are:
+There are four possibilities to obtain inter-class resonance (aside from the non-equivalent CR discussed in sec. \ref{non_equi_valent_CR}). These possibilities are:
 \begin{itemize}
@@
 \end{itemize}
-We should add to this list the case $\mathbf{B}=0$ where all four classes are also resonant, and any other possibility (which I will call random $\mathbf{B}$) where all four classes are spectrally separated.
+We should add to this list the case $\mathbf{B}=0$ where all four classes are also resonant, and any other possibility (which we will label random $\mathbf{B}$) where all four classes are spectrally separated.
 
@@
 Fig.
-\ref{Carte} shows a map where we simply monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer.
+\ref{Carte} shows a map where we monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer.
 We have also reported the locus of $\mathbf{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].
@@
 The experimental values were measured on sample ADM-15-3 using the protocol of Fig.
-\ref{various_fit_formulae}-d) with $T_1^{\rm ph}= 5\ \rm ms$.}
+\ref{various_fit_formulae}-d) with $T_1^{\rm ph}= 5\ \rm ms$.
+The uncertainty corresponds to the precision of the fit parameters.}
  \label{T1 champ mag}
@@
 \ref{eq 1/T}.
-There are however several unknown in this equation, such as the fluctuator density and lifetime which can be hard to estimate.
+There are however several unknowns in this equation, such as the fluctuator density and lifetime which can be hard to estimate.
 It is however relatively easy to compare the predictions of eq.
@@
 
-What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators of the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2, 3 or 4 classes are co-resonant as a function of $\Gamma_0$.
-Doing so allows to normalize $n_f$ and $\gamma_f$ in eq.
+We decided to take the case of an isolated class as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2, 3 or 4 classes are co-resonant as a function of $\Gamma_0$.
+Doing so allows us to normalize $n_f$ and $\gamma_f$ in eq.
 \ref{eq 1/T} and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
@@
 Using for example $T_1^{\rm ph}=2.5\ \rm ms$ gave results closer to the theoretical ones.
-Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded an increased depolarization rate by a factor $\sim 4$, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one.
-We can also see in Fig.
-\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration leads to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, 4 times smaller than the predicted factor of 10.
+Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded an increased depolarization rate by a factor $\sim 4$, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one\footnote{We can also see in Fig.
+\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration leads to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, which is 4 times smaller than the predicted factor of 10.}.
 
 The failure of the fluctuator model to accurately predict the changes in $T_1$ is one of the reasons that led us to believe that the model is incomplete.
+Other reasons will be presented in the following part.
 
@@
 
-In what follows I will describe what I think would be interesting investigation points to further understand the NV-NV relaxation and the fluctuator model.
+In what follows we will describe what could be interesting investigation points to further understand the NV-NV relaxation and the fluctuator model.
 
@@
 
-I already mentioned that the discrepancy between the theoretical and experimental $T_1$  values were an issue of the current fluctuator model, although the difference could come in part from the $T_1$ fitting procedure.
+We mentioned previously that the discrepancy between the theoretical and experimental $T_1$  values were in contradiction with the current fluctuator model, although the difference could come in part from the $T_1$ fitting procedure.
 Another, more glaring problem is the fact that some $T_1$ profile are not stretched exponential.
@@
 
-For the single isolated class, the $T_1$ profile seems to be closer to a stretch exponential, which is coherent with the fact that the lifetime $T_1 \sim 1\ \rm ms$ is already smaller than the expected phonon limited lifetime $T_1^{\rm ph} \sim 5\ \rm ms$.
-However when $\mathbf{B}=0$, although the lifetime gets $\sim$ 10 times shorter, the profile becomes clearly exponential.
+Although the spin lifetime still gets $\sim$ 10 times shorter in the $\mathbf{B}=0$ case, indicating that dipole-dipole interaction still plays an important role in the spin relaxation, this time the lifetime profile does not converge to a purely stretched exponential.
+In fact it seems to be the opposite: the lifetime profile is closer to a pure exponential for the shorter $T_1$ (especially for the very short times), and closer to a stretched exponential for the longer $T_1$, which is in complete contradiction with the fluctuator model and the previous observations in Fig.
+\ref{stretch_or_not_stretch} and \ref{various_fit_formulae}.
 
-The spin dynamics in this sample is still dominated by resonant dipole-dipole interactions: the more classes are brought to resonance, the shorter the spin $T_1$.
-But here, the profile gets closer to a pure exponential when $T_1$ decreases, which is in complete contradiction with the fluctuator model.
-
-This particular sample was a 15 $\mu$m micro-diamond, which could have a role in this observation since we know that spins near the surface have a reduced $T_1$ \citep{rosskopf2014investigation}, and micro-diamonds have proportionally more near surface spins.
-But other diamonds from the same batch, such as the one measured in Fig.
-\ref{stretch_or_not_stretch}, show clearly stretched exponential profile in the same conditions, and meanwhile some bulk diamonds show very short non-stretched exponential lifetimes, similar to the present sample.
+We should note that most sample studied follow the fluctuator model, in that the their lifetime profile gets closer to a stretched exponential when the dipole-dipole relaxation rate is increased.
+Sample ADM-15-4 seem to be an anomaly in this regard, since most of the samples from the same batch (including sample ADM-15-2 in Fig.
+\ref{stretch_or_not_stretch}) seem to follow the ``normal" behavior predicted by the fluctuator model.
+However, some other samples, including bulk HPHT samples and 150 $\mu$m Adamas micro-diamond also showed short (non stretched) exponential lifetimes.
 
 It seems therefore that the current fluctuator model does not correctly describe the spin dynamics of every NV-dense samples.
-Looking at what was previously done for P-doped SI \citep{vugmeister1978spin} or for resonant energy transfer between molecules \citep{yokota1967effects}, the fluctuator model could probably be improved by taking into account the finite lifetime of the fluctuators (i.e.
+Looking at what was previously done for P-doped SI \citep{vugmeister1978spin} or for FRET \citep{yokota1967effects}, the fluctuator model could probably be improved by taking into account the finite lifetime of the fluctuators (i.e.
 the saturation of the NV-fluctuator CR), and the spin diffusion between NV centers (the fluctuator model only considers NV-fluctator interactions and not NV-NV interactions).
@@
 \subsection{Investigation of the fluctuators nature}
-While the hypothesis of fluctuators being clusters of NV centers or NV-other spins is solidified by the precedents in P-doped Si \citep{honig1960electron} or F centers in KCl \citep{warren1964spin}, there are still many unknowns on the exact nature of the fluctuators.
-There are at least two aspects that I think merit to be investigated.
+While the hypothesis of fluctuators being clusters of NV centers or NV-other spins is strengthened by the precedents in P-doped Si \citep{honig1960electron} or F centers in KCl \citep{warren1964spin}, there are still many unknowns on the exact nature of the fluctuators.
+There are at least two aspects that merit to be investigated.
 
@@
 In particular it would be interesting to know whether this profile is always Lorentzian, and what its linewidth is.
-In the sample I studied on Fig.
-\ref{fluct linewidth}, I found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough my sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 45 ppm$ sample used in \citep{choi2017depolarization}.
+In the sample studied on Fig.
+\ref{fluct linewidth}, we found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough our sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 15 ppm$ sample used in \citep{choi2017depolarization} \footnote{Although the initial paper claimed a NV concentration $[\rm NV] \sim 45 ppm$, a more recent paper re-established the concentration at $\sim 45 ppm$ \citep{zhou2020quantum}}.
 If $\gamma_f$ was found to be somewhat of a constant among all samples showing NV-NV CR, this would give us a lot of information on the potential nature of the fluctuators.

### 2022-10-10 python_clément/test programs/test.py

@@ -1753,4 +1753,4 @@ def plot_irregular_surface():
 	plt.show()
 
 
-print((sqrt(log(2))/(pi*3.2)))
+print(0.8328*sqrt(1/3))

### 2022-10-11 Manuscrit Thèse/chapter 3/Chapter 3.tex
njtbke

@@
 Nevertheless, as we will see in this section, CR between both equivalent and non-equivalent NV centers is observed for samples with high NV density.
-\subsection{NV-NV CR between nonequivalent NV centers}
+\subsection{NV-NV CR between non-equivalent NV centers}
 \label{non_equi_valent_CR}
@@
 Fig.
-\ref{non-equivalent NV-NV}-b) shows the difference between the NV centers belonging to class 1 labeled NV$_\parallel$) and those belonging to classes 2-4 (labeled NV$_\nparallel$).
+\ref{non-equivalent NV-NV}-b) shows the difference between the NV centers belonging to class 1 (labeled NV$_\parallel$) and those belonging to classes 2-4 (labeled NV$_\nparallel$).
 As explained in [REF], the transverse magnetic field leads to a mixing of the eigenstates \footnote{The states of the misaligned NV on Fig.
@@
 \begin{equation}
-\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right) \frac{\eta^2}{\gamma_f}.
+\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right)^2 \frac{\eta^2}{\gamma_f}.
 \end{equation}
@@
 In the sample studied on Fig.
-\ref{fluct linewidth}, we found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough our sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 15 ppm$ sample used in \citep{choi2017depolarization} \footnote{Although the initial paper claimed a NV concentration $[\rm NV] \sim 45 ppm$, a more recent paper re-established the concentration at $\sim 45 ppm$ \citep{zhou2020quantum}}.
+\ref{fluct linewidth}, we found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough our sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 15 ppm$ sample used in \citep{choi2017depolarization} \footnote{Although the initial paper claimed a NV concentration $[\rm NV] \sim 45 ppm$, a more recent paper re-established the concentration at $\sim 15$ ppm \citep{zhou2020quantum}}.
 If $\gamma_f$ was found to be somewhat of a constant among all samples showing NV-NV CR, this would give us a lot of information on the potential nature of the fluctuators.

### 2022-10-11 Talks/Talk postdoc/presentation.tex
njtbke

--- 2022-05-19 Talks/Journées de l'axe/presentation.tex
@@
 \usepackage{tcolorbox} % toujours plus de boites
-\usepackage[backend=biber]{biblatex}
-
-\addbibresource{Biblio_dbl_quantum.bib}
+%\usepackage[backend=biber]{biblatex}
+%
+%
+%\addbibresource{Biblio_dbl_quantum.bib}
 
@@
 \maketitle
-\centering
-\includegraphics[scale=.2]{Nitrogen-vacancy_center}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{logos}
+\end{center}
 \end{frame}
@@
 %\section{Physics of the NV center}
-\begin{frame}{What are NV centers ?}
+\begin{frame}{Preamble : the NV center}
 \centering
@@
 \end{frame}
-\begin{frame}{Common example : Optically Detected Magnetic Resonance (ODMR)}
+
+\begin{frame}{Preamble : the 4 classes of NV centers}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR 4 classes}
 \end{frame}
 
-\begin{frame}{What are NV centers and why are they everywhere ?}
-\begin{itemize}
-\item Good optical properties : quantum yield $\approx 1$, stable in time and temperature
-\medskip
-\item Good spin properties : $T_1\approx\ \rm ms$ and $T_2^*\sim \mu$s at 300 K.
-\medskip
-\item Optical polarization (up to 90\%) and readout (fidelity up to 0.3) of the spin at room temperature
-\end{itemize}
-\bigskip
-$\to$ One of the most versatile spin qubit at room temperature
+\begin{frame}{Depolarization of dense NV ensemble at low magnetic field}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_presentation_sujet}
+
+\begin{frame}{outline}
+\tableofcontents
 \end{frame}
 
-\begin{frame}{What are NV centers being used for ?}
-$\bullet$ Quantum memories :
-\begin{center}
-\includegraphics[width=0.8\textwidth,height=0.9\textheight,keepaspectratio]{mémoire}
-\end{center}
+\section{Cross-relaxation with NV centers}
+1 slide théorie CR + relation T1 PL
+1 slide NV-VH
+1 slide NV-NV : pb
+\section{The NV-fluctuator model}
 
-$\bullet$ Intrication :
-\begin{center}
-\includegraphics[width=0.8\textwidth,height=0.9\textheight,keepaspectratio]{Intrication}
-\end{center}
+\section{Dipole-dipole interaction at low magnetic field}
+
+\section{Charcterization of the low field depolarization magnetometry}
+
 \end{frame}
 
-\begin{frame}{Magnetomtry with NV centers : big or small}
-$\bullet$ AFM nano-scale magnetomtry :
-\begin{center}
-\includegraphics[width=0.65\textwidth,height=0.9\textheight,keepaspectratio]{magneto AFM}
-\end{center}
-
-$\bullet$ Diamond magnetic microscopy :
-\begin{center}
-\includegraphics[width=0.65\textwidth,height=0.9\textheight,keepaspectratio]{magneto graphene}
-\end{center}
-\end{frame}
-
-\begin{frame}{Context of my PhD work : new physics with dense ensemble of NV centers}
-\begin{itemize}
-\item To increase the ensemble sensitivity we need to increase the density of NV centers
-\item Chemists and material scientists have made huge progress to grow NV-rich diamond with little other impurities 
-\begin{center}
-\includegraphics[width=0.65\textwidth,height=0.9\textheight,keepaspectratio]{diamant Alex}
-\end{center}
-\item When the defects density reaches a critical point, many body effects start to appear due to dipole-dipole interaction and charge tunneling (3 ppm of NV = 30 kHz dipole coupling between NV neighbors)
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Modification of the spin $T_1$ due to resonant dipole coupling}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_2}
-\end{frame}
-\begin{frame}{Conclusion}
-Acknowledgments : Christine et les autres
-
-\bigskip
-Take home messages : 
-\begin{itemize}
-\item NV centers are defects in diamond with an optically controllable and readable spin at room temperature
-\medskip
-\item Many body effects start to manifest with the new NV-rich samples being made
-\end{itemize}
-\end{frame}
-\begin{frame}{Bonus : Magnetometry in zero magnetic field}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_magneto}
-\end{frame}
-\begin{frame}{Optical properties of NV$^-$ centers}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_NV_optical}
-\end{frame}
-\begin{frame}{NV$^-$ center electronic structure}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{NV_8_niveaux_3}
-\end{frame}
-\begin{frame}{NV center spin sub-levels}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_3_niveaux}
-\end{frame}
 \end{document}

### 2022-10-11 data/20211015/Le bon adamas/traitement.py

@@ -30,7 +30,7 @@ plt.locator_params(axis='y', nbins=5)
 # print(popt)
 # plt.show()
 
-fname='ESR 0B zoom +10 dBm avec un T et 50 ohms'
+fname='Vrai ESR 1x1x1x1'
 x,y=extract_data(fname)
 plt.plot(x,y)
 

### 2022-10-12 Manuscrit Thèse/chapter 4/Chapter 4.tex
manger

@@
 \subsection{NV spin Hamiltonian in zero external magnetic field}
+\label{sec. Hamltonian 0 B}
 \begin{figure}[h]
@@
 
-\subsection{Hyperfine coupling and inhomogeneous broadening}
+\subsection{Hyperfine and inhomogeneous broadening}
 \label{sec modif T2*}
@@
 
-We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
+Réécrire tout ca en plus concis.
+%We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
 These change are relevant to our study of NV-NV CR due to the relation between the dipole-induced relaxation rate and $T_2^*$ detailed in sec.
 [REF].
-
-A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
+%
+%A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
 In the $\{ \ket{0}, \ket{\pm 1} \}$, the eigen energies are linearly sensitive to (longitudinal) magnetic field, and only sensitive to electric fields at the second order, and vice versa for the $\{ \ket{0}, \ket{\pm} \}$ basis.
-
-These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei.
+%
+%These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei.
 These effects can drastically modify the ODMR lineshape in zero external magnetic field \citep{jamonneau2016competition} or purely transverse magnetic field \citep{qiu2021nuclear,qiu2022nanoscale}.
-
-We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center.
+%
+%We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center.
 $^{14}$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin.
 The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
-\begin{equation}
-\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
-\end{equation}
-
-where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor.
+%\begin{equation}
+%\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
+%\end{equation}
+%
+%where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor.
 $\mathcal{H}_n$ and $\bar{\bar{A}}$ can be written:
-
-\begin{equation}
-\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
-\end{equation}
-
-and
-\begin{equation}
-\bar{\bar{A}} = \begin{pmatrix}
-A_{xx} & 0 & 0 \\
-0 & A_{yy} & 0 \\
-0 & 0 & A_{zz}
-\end{pmatrix},
-\end{equation}
-
-where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.
-
-Fig.
+%
+%\begin{equation}
+%\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
+%\end{equation}
+%
+%and
+%\begin{equation}
+%\bar{\bar{A}} = \begin{pmatrix}
+%A_{xx} & 0 & 0 \\
+%0 & A_{yy} & 0 \\
+%0 & 0 & A_{zz}
+%\end{pmatrix},
+%\end{equation}
+%
+%
+%where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.
+%
+%Fig.
 \ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$.
-
-While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers.
+%
+%While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers.
 These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.
-
-Fig.
+%
+%Fig.
 \ref{ESR for T2*}-a) allows us to evaluate the magnetic field noise in both samples, since the $\ket{\pm 1}$ states are not sensitive to weak electric fields.
@@
 The total width of the line however is mostly dominated by the hyper-fine splitting, which result in a similar total linewidth in both cases
-
-We should note that the usual practice in magnetometry is to consider the linewidth only of the three hyper-fine lines when they are resolved.
+%
+%We should note that the usual practice in magnetometry is to consider the linewidth only of the three hyper-fine lines when they are resolved.
 This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines.
@@
 We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.
-
-Fig.
+%
+%Fig.
 \ref{ESR for T2*}-b) allows us to evaluate the electric field noise in both samples, since the $\ket{\pm}$ states are not sensitive to weak magnetic fields.
@@
 We can note that the electric field noise is significantly stronger in sample Sumi-2, which leads to an ODMR linewidth more than twice as big.
-
-Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field.
+%
+%Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field.
 For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction).
@@
 The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.
-
-The table in Fig.
+%
+%The table in Fig.
 \ref{ESR for T2*}-d) reports the linewidths measured on both sample for the various magnetic field configurations.
@@
 We should therefore expect the modification of $T_2^*$ to have a moderate impact on the depolarization dynamics.
-
+%
 \medskip
@@
 \label{sec causes zero field}
-We have identified three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/Shema_summary_theory}
+\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
+\label{summary_theory}
+\end{figure}
+
+Fig. \ref{summary_theory} summarizes the different contribution that we have identified to the depolarization of the NV centers at low magnetic field.
+
+The first and main contribution is degeneracy between the four classes as the magnetic field decreases.
+Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 40$.
+This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.
+
+The three other contributions, which are the eigenbasis change, the presence of double flips and the increase in $T_2^*$ will be detailed below and in the appendix [REF].
+These three contributions are the reason for the depolarization at low magnetic field in Fig.
+\ref{scan 100}.
+In total, they amount for a further increase in the relaxation rate by a factor of $\sim 2$ compared to the $\mathbf{B} \parallel [100]$ case.
+Although this increase is small compared to the 4-classes degeneracy, we will see in sec.
+\ref{sec LFDM} that it is relevant for the  low field magnetometry protocol.
+
+We should note that the values indicated in Fig.
+\ref{summary_theory} for the double flips and the $T_2^*$ contributions depend on experimental parameters, while the ones for the 4-classes degeneracy and the eigenbasis change are purely based on theory.
 
 \subsubsection{Eigenstates modification}
-The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
 
-This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
+The modification of the Hamiltonian eigenstates discussed in sec.
+\ref{sec.
+Hamltonian 0 B} from the eigenbasis $\{ \ket{0}, \ket{+1},\ket{-1} \}$ when $B\neq 0$ to the eigenbasis $\{ \ket{0}, \ket{+},\ket{-} \}$ when $B\sim 0$ changes the dipole-dipole interaction between the two spins.
+
+A consequence of this change is the absence of diagonal terms  in the dipole-dipole Hamiltonian for the $\{ \ket{0}, \ket{+},\ket{-} \} \times \{ \ket{0}, \ket{+},\ket{-} \}$ basis, which explains why the NV centers states are protected from the magnetic noise at low magnetic field \citep{jamonneau2016competition}.
+
+Another consequence is a slight modification in the flip-flop rates.
+Once averaged over all possible orientation and position, as detailed in appendix [REF], this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
+%The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
+%
+%This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
 We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively.
 To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
-
-The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF].
+%
+%The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF].
 The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane.
 This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.
-
-We therefore need to make an assumption on the distribution of the transverse field in the sample.
+%
+%We therefore need to make an assumption on the distribution of the transverse field in the sample.
 In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane.
@@
 [REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.
-
-We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
-
-We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$.
+%
+%We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
+%
+%We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$.
 $\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
-
-In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$.
+%
+%In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$.
 The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig.
@@
 
-\begin{table}[htbp]
-\centering
-\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
- \label{table double flips}
-\begin{tabular}{c|cc}
-%\toprule
-{} & $l_c=0$ & $l_c=\infty$ \\
-\midrule
-$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
-$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
-%\bottomrule
-\end{tabular}
-\end{table}
+The double flip, or flip-flip, processes that we consider here are the one that couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$).
+These terms are normally not resonant because of the zeeman splitting between the $\ket{+1}$ and $\ket{-1}$ states.
 
-Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
+In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$.
+These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, however they are close enough in energy that the double flip processes are not completely
+
+%\begin{table}[htbp]
+%\centering
+%\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
+% \label{table double flips}
+%\begin{tabular}{c|cc}
+%%\toprule
+%{} & $l_c=0$ & $l_c=\infty$ \\
+%\midrule
+%$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
+%$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
+%$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
+%%\bottomrule
+%\end{tabular}
+%\end{table}
+%
+%Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
 The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
-
-We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field.
+%
+%We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field.
 And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field.
@@
 In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.
-
-To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
+%
+%To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
 We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
 Finally, the question about the correlation  length of the electric field also needs to be asked.
-
-The results are summed up in Table \ref{table double flips}.
+%
+%The results are summed up in Table \ref{table double flips}.
 We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
@@
 We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
-
-\subsubsection{Summary}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/Shema_summary_theory}
-\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
-\label{summary_theory}
-\end{figure}
-
-Fig. \ref{summary_theory} recapitulate the various mechanisms at play in the low field depolarization of NV ensemble.
-
-We start from from the simplest case where every classes are sepctrally isolated and the only dipole-induced relaxation coms from the flip-flop between NV and fluctuators from the same class.
-We then consider the case where all four classes are resonant ($B\parallel [100]$) and finally the case B=0.
-
-The three extra depolarization mechanisms in zero field are then detailed, with the predicted increase of depolarization rate for each step.
-The predicted rates in zero field depend on several parameters, we chose the following ones to compute the different numerical values: $l_c=0$, $\bar{G}^2=0.5$, $2\gamma_f = 6.5\ \rm MHz$ and $2/T_2^*=5.9\ \rm MHz$ (corresponding to sample Sumi-2 if we assume a Lorentzian ODMR profile).
 
@@
 
-\subsection{Experimental data}
+\subsection{Results}
 

### 2022-10-12 Talks/Talk postdoc/presentation.tex
manger

@@
 
-\begin{frame}{Depolarization of dense NV ensemble at low magnetic field}
+\begin{frame}{Subject of this presentation}
 \centering
 \includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_presentation_sujet}
+\end{frame}
 
-\begin{frame}{outline}
+\begin{frame}{Outline}
 \tableofcontents
@@
 \section{Cross-relaxation with NV centers}
-1 slide théorie CR + relation T1 PL
-1 slide NV-VH
-1 slide NV-NV : pb
-\section{The NV-fluctuator model}
-
-\section{Dipole-dipole interaction at low magnetic field}
-
-\section{Charcterization of the low field depolarization magnetometry}
-
+\begin{frame}{Principle of cross-relaxation with NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation}
 \end{frame}
 
+\begin{frame}{Example: Cross-relaxation between NV centers and VH$^-$}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_CR_VH}
+\end{frame}
+
+\begin{frame}{Cross-relaxation between NV centers and NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_CR_adamas}
+\end{frame}
+\section{The NV-fluctuator model and experimental verification}
+
+\begin{frame}{Presentation of the fluctuator model}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro}
+\end{frame}
+
+\begin{frame}{Predictions of the fluctuator model}
+\begin{itemize}
+\item $\Gamma_1$ increases when classes overlap spectrally (increase in the resonant fluctuator density).
+\item The dipole induced depolarization has a stretched exponential profile:
+$$ \rho_{00}(t) \propto \exp(-\sqrt{\frac{t}{T_1}})$$
+\item The Fluctuators spectral response is broadened by their decay rate $\gamma_f$ (lifetime limit).
+\end{itemize}
+\end{frame}
+
+\section{Dipole-dipole interaction under low magnetic field}
+
+\section{Charcterization of the low field depolarization magnetometry protocol}
+
 \end{document}

### 2022-10-13 Manuscrit Thèse/chapter 4/Chapter 4.tex
y'a pot

@@
 \end{figure}
+
+A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
+The energy of the $\ket{\pm 1}$ states depend linearly on the longitudinal magnetic field, and to the second  order on the transverse electric and magnetic fields.
+On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and to the second order on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp}$.}.
+
+The shape of the ODMR lines (i.e.
+the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis.
+Further details on the low magnetic field physics of the NV center are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.
+
+Fig.
+\ref{ESR for T2*} shows ODMR profile on two samples, CVD-pink which shows weak dipole-induced depolarization and Sumi-2 which shows strong dipole-induced depolarization, in three distinct regime: for a purely longitudinal magnetic field, for a purely transverse magnetic field, and without magnetic field.
+The linewidths of the spin transition \footnote{We consider here the full linewidth of the electronic levels, i.e.
+including all three hyper fine levels} for both samples are reported in the Table \ref{ESR for T2*}-d).
+In both cases, the linewidths are smaller in the $B=0$ and $B \perp NV$ cases than in the $B \parallel NV$ case, mainly due to the modification of the hyper-fine coupling.
+The linewidth modification reaches up to $\sim 50\ \%$ for the CVD sample and $\sim 20\ \%$ for the HPHT one.
+
+We are particularly interested in the spin transition linewidth since it directly impacts the dipole-dipole relaxation rate (as seen in sec.
+[REF]): the smaller the linewidth, the higher the dipole-dipole relaxation rate is.
 
@@
 In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$.
-These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, however they are close enough in energy that the double flip processes are not completely
+These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely turned off.
 

### 2022-10-13 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

@@
 The energy of the $\ket{\pm 1}$ states depend linearly on the longitudinal magnetic field, and to the second  order on the transverse electric and magnetic fields.
-On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and to the second order on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp}$.}.
+On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and to the second order on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp$.}.
 
@@
 
-Réécrire tout ca en plus concis.
 %We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
@@
 %
-\medskip
-
-To summarize this section, we have seen that (for the samples used in this chapter) the NV centers spin Hamiltonian is dominated by transverse electric field for low external magnetic field, and that it causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
-We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples with a high impurity density.
+\bigskip
+
+To summarize this section, we have seen that (for the samples used in this chapter) the NV center spin Hamiltonian is dominated by transverse electric field at low external magnetic field, which causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
+
+We have also seen that the spin Hamiltonian has the same eigenstates for purely transverse magnetic field than it does for low magnetic field.
+
+Finally we have seen that this change of eigenstates occasions a modification of the spectral profile and linewidth of the spins.
 
@@
 
+The computation of the depolarization rate including the double flips is detailed in appendix [REF].
+To obtain a numerical value, we need to input the experimental values of the dipole-dipole interaction range \footnote{By interaction range we mean the constant to be compared to the spectral detuning $\Delta \nu$ between the two groups.
+This is $2\gamma_f+\Gamma_{NV}+\Gamma_f$ in eq.
+[REF]} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+[REF], as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ state.
+We typically measure $\Delta \nu \approx 8\ \rm MHz$ on type 1b HPHT samples.
+
+For the value presented in Fig. \ref{simus ESR}, we inputted $\Gamma=\Delta \nu$.
 %\begin{table}[htbp]
@@
 
-The last aspect we considered was the change in $T_2^*$ (we include the change in the hyper-fine interaction in the $T_2^*$) for weak magnetic field, as discussed in sec. \ref{sec modif T2*}.
+The last aspect we considered was the change in the ODMR lineshape discussed in sec. \ref{sec modif T2*} for weak magnetic field (we refer to this as $T_2^*$ for conciseness).
 
@@
 \ref{ESR for T2*}.
-
 We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
+
+The value shown in Fig.
+\ref{summary_theory} corresponds to a $9 \%$ increase, as most of the samples studied in this manuscript are type 1b HPHT diamonds (therefore more similar to sample Sumi-2 than to sample CVD-pink).
 

### 2022-10-13 Talks/Talk Stuttgart/presentation.tex
tata

--- 2022-10-12 Talks/Talk postdoc/presentation.tex
@@
 
+\begin{frame}{$T_1$ measurement protocol}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_protocol}
+\end{frame}
+
+\begin{frame}{Stretched exponential decay profile}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_exp_stretch}
+\end{frame}
+
+\begin{frame}{Competition between stretched and exponential decay}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_exp_et_stretch}
+\end{frame}
+
+\begin{frame}{Spectral response of the fluctuators}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_fluct_linewidth}
+\end{frame}
+
+\begin{frame}{Geometry conditions for class resonances}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_geométries}
+\end{frame}
+
+\begin{frame}{PL mapping of the crystalline planes}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_carte}
+\end{frame}
+
+\begin{frame}{Limitations of the fluctuator model}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_limit_fluct}
+\end{frame}
+
 \section{Dipole-dipole interaction under low magnetic field}

### 2022-10-16 Manuscrit Thèse/chapter 4/Appendix_autres_causes.tex
tata

--- 2022-10-12 Manuscrit Thèse/chapter 4/Chapter 4.tex
@@
     natbib=true,
-    style=numeric-comp,
+    style=numeric,
     sorting=none, %Pour faire apparaitre les refs dans l'ordre
@@
 
-\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
 \title{}
 \begin{document}
-\chapter{NV-NV CR under transverse or low fields: application to magnetometry}
+\chapter{Appendix: Influence of the magnetic field misalignment and laser polarization on the low field relaxation}
 
-In the last chapter we saw how dense ensemble of NV centers have an intrinsic depolarization mechanism mediated by the NV-NV dipole interaction.
-In this chapter we will see how this feature can be exploited to perform magnetometry.
-
-Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
-We will focus here on the low to zero magnetic field region, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the more complex physics of the NV centers under low magnetic field.
-
-This chapter is constructed as follows: We will first describe the physics of NV centers under low or transverse magnetic field (the physics is similar in both cases), we will then show both theoretically and experimentally that the NV-NV CR is modified for low and transverse magnetic field.
-We will identify different mechanisms by which the relaxation is modified and we will try to quantify their relative importance.
-
-We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field.
-We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols.
-Finally we will give some perspectives on how to improve LFDM and on potential applications.
-
-The results presented in this chapter are based in large part on the article \citep{pellet2022spin}.
-
-\section{NV spin Hamiltonian under low and transverse fields}
-
-Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates compared to the high magnetic field regime.
-
-\subsection{NV spin Hamiltonian in zero external magnetic field}
-\label{sec. Hamltonian 0 B}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/ESR_simus}
-\caption{Simulations of inhomogeneous zero field ODMR when sampling various parameters.
-a) Simulation when sampling each components of the magnetic field over a Gaussian of deviation $\sigma=2\ \rm G$.
-b) Simulation when sampling each components of the strain tensor $\bar{\bar{\varepsilon}}$ over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
-c) Simulation when sampling each components of the electric field over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
-d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
-\label{simus ESR}
-\end{figure}
-
-In the absence of external magnetic field, we have to take into account other elements which would otherwise be of second order in the spin Hamiltonian.
-These elements are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}.
-The hyperfine splitting due to nearby nuclei will be considered separately, although to a large extent it behaves like a local magnetic field.
-
-Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian).
-The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.
-
-The NV$^-$spin Hamiltonian in the $\{ \ket{-1}, \ket{+1} \}$ basis can be written as \citep{udvarhelyi2018spin}:
-\begin{equation}
-\mathcal{H}=\begin{pmatrix}
-D-\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}}) & f_\perp(\mathbf{E}) + g_\perp(\bar{\bar{\varepsilon}})\\
-f^*_\perp(\mathbf{E}) + g^*_\perp(\bar{\bar{\varepsilon}})&D+\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}})
-\end{pmatrix},
-\label{Hamiltonien pm1}
-\end{equation}
-where $B_\parallel$ is the component of the magnetic field along the NV axis, and $f_\parallel, f_\perp, g_\perp$, and $g_\parallel$ are functions of the electric field $\mathbf{E}$ and the strain tensor $\bar{\bar{\varepsilon}}$, whose expressions are:
-
-\begin{align}
-f_\parallel(\mathbf{E})&=d_\parallel E_z, \\
-f_\perp(\mathbf{E})&=d_\perp ( E_x + i E_y), \\
-g_\parallel(\bar{\bar{\varepsilon}})&= h_{41}(\varepsilon_{xx}+\varepsilon_{yy})+h_{43} \varepsilon_{zz}, \\
-g_\perp(\bar{\bar{\varepsilon}}) &= \frac{1}{2} \left[ h_{16}(\varepsilon_{zx}+i \varepsilon_{zy}) + h_{15}\left(\frac{\varepsilon_{yy}-\varepsilon_{xx}}{2}+i\varepsilon_{xy}\right) \right],
-\end{align}
-where $d_\parallel=0.35 \ \rm{Hz\, cm/V}$ and $d_\perp=17\ \rm{Hz\, cm/V}$ have been measured experimentally \citep{van1990electric}, and $h_{43}=2300\ \rm{MHz}$, $h_{41}=-6420\ \rm{MHz}$, $h_{15}=5700\ \rm{MHz}$ and $h_{16}=19660\ \rm{MHz}$ were computed through DFT \citep{udvarhelyi2018spin} and show reasonable agreement with experiments \citep{barson2017nanomechanical}.
-
-Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.
-
-The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting.
-For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$.
-The shift in energy will therefore tend to blur the energy split when averaging over a large number of spins.
-
-Fig.
-\ref{simus ESR} shows a simulation of how each parameters of the spin Hamiltonian - local magnetic field, local electric field and strain - affects the zero external field ODMR profile.
-To do these simulations, I sampled each parameters separately $10^6$ times and plotted the histogram of the two eigenvalues of the Hamiltonian written in eq.
-\ref{Hamiltonien pm1}.
-Fig.
-\ref{simus ESR}-d) shows an experimental zero field ODMR spectrum, typical of what we observe with dense NV ensembles.
-
-Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR.
-Given the simulation results, the only parameter that can give rise to this shape is the local electric field.
-
-We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample.
-In particular, we do not shield for the earth magnetic field in our experiment.
-The splitting caused by the earth magnetic field however has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$, which is $\sim 4$ times weaker than the splitting we typically observe.
-
-We will therefore consider that the low external magnetic field spin Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.
-
-We will then adopt the following simplified Hamiltonian for the zero external field regime:
-\begin{equation}
-\mathcal{H}=\begin{pmatrix}
-D&0&d_\perp E_\perp^* \\
-0&0&0 \\
-d_\perp E_\perp &0&D
-\end{pmatrix},
-\end{equation}
-whose eigenvectors are $\ket{0}$ and$\ket{\pm}$ of eigenvalues 0 and $D\pm d_\perp \abs{E_\perp}$, where $\ket{\pm}$ are defined as:
-\begin{equation}
-\ket{\pm}=\frac{\ket{+1}\pm e^{-i\phi_E}\ket{-1}}{\sqrt{2}},
-\end{equation}
-where $\tan(\phi_E)=E_y/E_x$.
-
-\subsection{NV spin Hamiltonian under purely transverse magnetic field}
-\label{sec B transverse}
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.7\textwidth]{Figures/map_etats_propres}
-%\caption{Closeness of the Hamiltonian most excited state $\ket{3}$ with either $\ket{+1}$ or $\ket{+}$ as a function of the external magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
-%\label{map champ transverse}
-%\end{figure}
-
-We now turn to the case of purely transverse magnetic field with respect to the NV axis, i.e.
-$\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
-In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.
-
-In this regime, the NV Hamiltonian eigenstates are similar to the case where it is dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{-}, \ket{+}$\citep{qiu2021nuclear, qiu2022nanoscale}, of respective eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D,D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
-\begin{equation}
-\ket{\pm}=\frac{\ket{+1}\pm e^{-2i\phi_B}\ket{-1}}{\sqrt{2}},
-\end{equation}
-and $\tan(\phi_B)=B_y/B_x$.
-
-For the case where $d_\perp E_\perp \sim \frac{(\gamma_e B_\perp)^2}{D}$ and $\phi_E \neq 2\phi_B$, the eigenstates of the Hamiltonian are still of the form $\ket{0},\ket{\pm}$ with a relative angle $\phi$ in between $\phi_E$ and $2\phi_B$.
-
-In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
-
-\subsection{Hyperfine and inhomogeneous broadening}
-\label{sec modif T2*}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2}
-\caption{ODMR lineshapes for longitudinal, transverse and low magnetic fields on samples CVD-pink and Sumi-2.
-a) ODMR spectrum of a single class of NV centers for a longitudinal magnetic field ($B_\parallel \sim 100\ \rm G$).
-The microwave frequency is given with respect to the ODMR line central frequency.
-b) ODMR spectrum of a single class for a purely transverse magnetic field ($B_\perp \sim 100\ \rm G$).
-c) ODMR of all 4 classes for 0 external magnetic field.
-d) Table reporting the full width at half maximum of each ODMR line.}
-\label{ESR for T2*}
-\end{figure}
-
-Réécrire tout ca en plus concis.
-%We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
-These change are relevant to our study of NV-NV CR due to the relation between the dipole-induced relaxation rate and $T_2^*$ detailed in sec.
-[REF].
-%
-%A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
-In the $\{ \ket{0}, \ket{\pm 1} \}$, the eigen energies are linearly sensitive to (longitudinal) magnetic field, and only sensitive to electric fields at the second order, and vice versa for the $\{ \ket{0}, \ket{\pm} \}$ basis.
-%
-%These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei.
-These effects can drastically modify the ODMR lineshape in zero external magnetic field \citep{jamonneau2016competition} or purely transverse magnetic field \citep{qiu2021nuclear,qiu2022nanoscale}.
-%
-%We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center.
-$^{14}$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin.
-The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
-%\begin{equation}
-%\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
-%\end{equation}
-%
-%where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor.
-$\mathcal{H}_n$ and $\bar{\bar{A}}$ can be written:
-%
-%\begin{equation}
-%\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
-%\end{equation}
-%
-%and
-%\begin{equation}
-%\bar{\bar{A}} = \begin{pmatrix}
-%A_{xx} & 0 & 0 \\
-%0 & A_{yy} & 0 \\
-%0 & 0 & A_{zz}
-%\end{pmatrix},
-%\end{equation}
-%
-%
-%where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.
-%
-%Fig.
-\ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$.
-%
-%While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers.
-These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.
-%
-%Fig.
-\ref{ESR for T2*}-a) allows us to evaluate the magnetic field noise in both samples, since the $\ket{\pm 1}$ states are not sensitive to weak electric fields.
-We indeed find that the Sumi-2 sample has much more magnetic noise, to the point where the hyper-fine structure is no longer resolved.
-The total width of the line however is mostly dominated by the hyper-fine splitting, which result in a similar total linewidth in both cases
-%
-%We should note that the usual practice in magnetometry is to consider the linewidth only of the three hyper-fine lines when they are resolved.
-This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines.
-In our case however, we have to consider the spectral overlap between NV centers and fluctuators, which have an additional broadening of $2\gamma_f \approx 6\ \rm MHz$ [REF] that completely obscures the hyper-fine structure.
-We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.
-%
-%Fig.
-\ref{ESR for T2*}-b) allows us to evaluate the electric field noise in both samples, since the $\ket{\pm}$ states are not sensitive to weak magnetic fields.
-Similarly, the hyper-fine structure is hidden in this configuration since all three hyper-fine levels are nearly-degenerate, provided that $\frac{(\gamma_e B_\perp)^2}{D} > A_{xx},A_{zz},Q$ which is typically the case for $B_\perp > 40\ \rm G$.
-We can note that the electric field noise is significantly stronger in sample Sumi-2, which leads to an ODMR linewidth more than twice as big.
-%
-%Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field.
-For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction).
-For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_s=-1,m_I=+1}$ and $\ket{m_s=+1,m_I=-1}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$.
-The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.
-%
-%The table in Fig.
-\ref{ESR for T2*}-d) reports the linewidths measured on both sample for the various magnetic field configurations.
-We can see that the linewidth of sample CVD-pink varies by more than a factor of 2, while the ones from sample Sumi-2 varied by less than 20 \%.
-We should note that  almost all samples used in this chapter are HPHT type 1b diamond, similar to sample Sumi-2 and unlike sample CVD-pink.
-We should therefore expect the modification of $T_2^*$ to have a moderate impact on the depolarization dynamics.
-%
-\medskip
-
-To summarize this section, we have seen that (for the samples used in this chapter) the NV centers spin Hamiltonian is dominated by transverse electric field for low external magnetic field, and that it causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
-We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples with a high impurity density.
-
-\section{NV-NV CR in the low magnetic field regime}
-In this part, we will study the cross-relaxation between NV centers under low magnetic field, whose spin Hamiltonian is dominated by a transverse electric field.
-Understanding the various mechanisms operating in this regime will be important for the magnetometry protocol presented after.
-
-\subsection{Experimental observations}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
-\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction.
-a) ODMR spectrum for B=0.
-b) ODMR spectrum for B=60 G.
-c) PL contrast as a function of the external magnetic field.
-d) $1/T_1^{\rm dd}$ as a function of the external magnetic field.
-A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
-\label{scan 1x1x1x1}
-\end{figure}
-We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.
-
-The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust in order to isolate each effects.
-
-Fig.
-\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G.
-The ODMR at the initial and final magnetic fields are shown in Fig.
-\ref{scan 1x1x1x1}-a) and b).
-
-While it is clear that the spin lifetime as well as the PL increases with the magnetic field, there is no clear indication that this is because of the specificity of the low field region.
-Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.
-
-We can note however that the PL is not an exact mirror of the spin dynamics: while the low field drop in PL is indeed associated with a reduced spin lifetime, the drop in PL at higher magnetic field comes from the states mixing induced by the transverse magnetic field, and is not associated with a modification of the spin dynamics (to the first order).
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\textwidth]{Figures/scan_100}
-\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis.}
-\label{scan 100}
-\end{figure}
-
-Fig.
-\ref{scan 100} presents a way to circumvent this issue: by applying the magnetic field along the [100] crystalline axis, we can make sure that the four classes of NV centers always stay resonant regardless of the magnetic field amplitude.
-
-We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig.
-\ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of $\sim$ 8 in zero field, whereas in Fig.
-\ref{scan 100}, it was only reduced by a factor of $\sim$ 1.2.
-The main reason for the PL and $T_1$ drop in Fig.
-\ref{scan 1x1x1x1} was indeed the co-resonance between the four classes.
-
-Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained if we consider only the inter-class resonances.
-This means that there are some additional depolarization mechanisms which are proper to the zero field region.
-We should also note that, while the zero-field PL contrast is bigger in Fig.
-\ref{scan 1x1x1x1}-c) than in Fig.
-\ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually similar in both cases.
-
-We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]
-
-\subsection{Theory of the low field depolarization}
-\label{sec causes zero field}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/Shema_summary_theory}
-\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
-\label{summary_theory}
-\end{figure}
-
-Fig. \ref{summary_theory} summarizes the different contribution that we have identified to the depolarization of the NV centers at low magnetic field.
-
-The first and main contribution is degeneracy between the four classes as the magnetic field decreases.
-Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 40$.
-This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.
-
-The three other contributions, which are the eigenbasis change, the presence of double flips and the increase in $T_2^*$ will be detailed below and in the appendix [REF].
-These three contributions are the reason for the depolarization at low magnetic field in Fig.
-\ref{scan 100}.
-In total, they amount for a further increase in the relaxation rate by a factor of $\sim 2$ compared to the $\mathbf{B} \parallel [100]$ case.
-Although this increase is small compared to the 4-classes degeneracy, we will see in sec.
-\ref{sec LFDM} that it is relevant for the  low field magnetometry protocol.
-
-We should note that the values indicated in Fig.
-\ref{summary_theory} for the double flips and the $T_2^*$ contributions depend on experimental parameters, while the ones for the 4-classes degeneracy and the eigenbasis change are purely based on theory.
-
-\subsubsection{Eigenstates modification}
-
-The modification of the Hamiltonian eigenstates discussed in sec.
-\ref{sec.
-Hamltonian 0 B} from the eigenbasis $\{ \ket{0}, \ket{+1},\ket{-1} \}$ when $B\neq 0$ to the eigenbasis $\{ \ket{0}, \ket{+},\ket{-} \}$ when $B\sim 0$ changes the dipole-dipole interaction between the two spins.
-
-A consequence of this change is the absence of diagonal terms  in the dipole-dipole Hamiltonian for the $\{ \ket{0}, \ket{+},\ket{-} \} \times \{ \ket{0}, \ket{+},\ket{-} \}$ basis, which explains why the NV centers states are protected from the magnetic noise at low magnetic field \citep{jamonneau2016competition}.
-
-Another consequence is a slight modification in the flip-flop rates.
-Once averaged over all possible orientation and position, as detailed in appendix [REF], this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
-%The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
-%
-%This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
-We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively.
-To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
-%
-%The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF].
-The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane.
-This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.
-%
-%We therefore need to make an assumption on the distribution of the transverse field in the sample.
-In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane.
-However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq.
-[REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.
-%
-%We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
-%
-%We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$.
-$\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
-%
-%In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$.
-The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig.
-\ref{scan 100}
-
-\subsubsection{Double flips}
-
-The double flip, or flip-flip, processes that we consider here are the one that couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$).
-These terms are normally not resonant because of the zeeman splitting between the $\ket{+1}$ and $\ket{-1}$ states.
-
-In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$.
-These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, however they are close enough in energy that the double flip processes are not completely
-
-%\begin{table}[htbp]
-%\centering
-%\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
-% \label{table double flips}
-%\begin{tabular}{c|cc}
-%%\toprule
-%{} & $l_c=0$ & $l_c=\infty$ \\
-%\midrule
-%$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-%$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
-%$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
-%%\bottomrule
-%\end{tabular}
-%\end{table}
-%
-%Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
-The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
-%
-%We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field.
-And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field.
-But we can see from Fig.
-\ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig.
-[REF].
-In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.
-%
-%To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
-We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
-Finally, the question about the correlation  length of the electric field also needs to be asked.
-%
-%The results are summed up in Table \ref{table double flips}.
-We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
-$\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$).
-This is close to the the experimental values we have.
-Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $).
-We come back to the results of the last section where only considered the flip-flops.
-
-\subsubsection{$T_2^*$ modification}
-
-The last aspect we considered was the change in $T_2^*$ (we include the change in the hyper-fine interaction in the $T_2^*$) for weak magnetic field, as discussed in sec. \ref{sec modif T2*}.
-
-To quantify its impact in the depolarization rate, we employ eq.
-[REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
-\ref{ESR for T2*}.
-
-We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
-
-\subsection{Other potential causes for zero field depolarization}
-We have also considered two other potential causes that we have ultimately ruled out in this case.
-
-\subsubsection{Magnetic field alignment}
+\section{Magnetic field alignment}
 \begin{figure}[h]
@@
 \includegraphics[width=.7\textwidth]{Figures/alignement_rose}
-\caption{PL of sample CVD-pink as a function of the magnetic field for various misalignment of the field direction with the [100] axis}
+\caption{PL of sample CVD-pink as a function of the magnetic field for 7 different misalignment of the field with respect to the [100] axis}
 \label{alignement rose}
@@
 
-The first is a potential misalignment of the magnetic field with the [100] axis, which would cause a splitting of the 4 classes as the magnetic field increases.
-This hypothesis is not consistent with the observations in Fig.
+A potential cause for the observation of depolarization at low magnetic field when $\mathbf{B} \parallel [100]$ could be a misalignment of the magnetic field with the [100] axis.
+Indeed, such a misalignment would cause a splitting of the 4 classes as the magnetic field increases which would reduce the inter-class flip-flop rate.
+%This hypothesis is not consistent with the observations in Fig.
 \ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased.
 Instead we observe a sharp decrease followed by a pseudo-plateau.
+Fig.
+\ref{alignement rose} shows the PL of sample CVD-pink as the magnetic field is scanned with various misalignments compared to the [100] axis.
+We estimate our alignments to be precise within $\pm 1^\circ$.
+We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization in the $\mathbf{B} \parallel [100]$ case does not come from a field misalignment.
 
-We still estimated how sensitive the PL dip was with respect to the field alignment.
-Fig.
-\ref{alignement rose} shows PL scans on sample CVD-pink for various alignment of the magnetic field around the [100] axis.
-We estimate our alignments to be precise within $\pm 1^\circ$.
-We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization does not come from a field misalignment.
-
-\subsubsection{Laser polarization}
+\section{Laser polarization}
 \begin{figure}[h]
@@
 \includegraphics[width=.7\textwidth]{Figures/pola_laser}
-\caption{PL of sample ADM-150-3 with respect to the magnetic field for various laser polarization angle.
+\caption{PL of sample ADM-150-3 as a function of the magnetic field for six different laser polarization angles.
 The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
@@
 
-Other studies \citep{anishchik2015low, filimonenko2020weak} previously saw a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.
+Other studies \citep{anishchik2015low, filimonenko2020weak} previously observed a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.
 
@@
 
-\section{NV-NV CR under purely transverse magnetic field}
-We have seen in the last part that there was a spontaneous spin depolarization specifically in zero external magnetic field, and we have identified three potential mechanisms which are the double-flips, the change in the Hamiltonian eigenstates and the modification of $T_2^*$.
-We now want to experimentally discriminate the contribution of each of these mechanisms and to see how close the actual decay rates are to the predicted ones.
-
-One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic and electric noise ...).
-Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.
-
-The ideal way to isolate each contributions would be to apply an electric field strong enough to split the transition of the 4 classes, and to measure the decay rate of a single class dominated by the transverse electric field.
-Such an electric field would need to be of the order of $\sim 10^6\ \rm{V/cm}$, which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.
-
-Instead of an electric field, we will use a transverse magnetic field, which as described in sec.
-\ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
-
-\subsection{Principle of the experiment}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/transis_transverse_field}
-\caption{Simulation of the eigenstates of the NV Hamiltonian for  a purely transverse magnetic field $B_\perp$.
-In addition to the transverse magnetic field, we consider a fixed transverse electric field of value $d_\perp E_\perp=4\ \rm MHz$.
-(a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$.
-The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$.
-(b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
-\label{eigenstates transverse field}
-\end{figure}
-
-Fig.
-\ref{eigenstates transverse field} shows a simulation of the eigenstates and eigenenergies for a NV Hamiltonian subject to transverse electric field and magnetic field.
-For convenience we chose to take $E_\perp = E_x$ and $B_\perp = B_x$, taking other geometric configuration alters only slightly the results.
-We labeled the eigenstates of the Hamiltonian $\ket{1}$, $\ket{2}$ and $\ket{3}$ in ascending order of energy.
-
-Fig.
-\ref{eigenstates transverse field}-b) shows both the splitting $\Delta \nu$ between the $\ket{2}$ and $\ket{3}$ states, and how close the $\ket{3}$ state is to the $\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}}$ state via the factor $\abs{\bra{3}\ket{+}}^2$.
-We look at the $\ket{3}$ state since $\ket{2}=\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}}$ in this case.
-
-These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 since we wanted to observe the modification caused by the eigenbasis $\{\ket{0},\ket{\pm}\}$.
-This way we can isolate the $T_1^{\rm dd}$ contribution coming from the change of eigenstates to the one coming from the double flips.
-Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100,200] G seems to satisfy both these criteria.
-
-\subsection{Results}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=.8\textwidth]{Figures/scan_T1_perp}
-\caption{Experimental data for the depolarization under purely transverse magnetic field on sample ADM-150-2.
-a) ODMR spectrum for $|\mathbf{B}|=20\ \rm G$.
-The transitions of the class orthogonal to the magnetic field are labeled $\nu_+$ and $\nu_-$.
-b) Dependency of $\nu_+$ and $\nu_-$ with the magnetic field, as measured through ODMR.
-c) Measurement of $T_1^{\rm dd}$ as a function of the transverse magnetic field.
-The corresponding energy splittings $\Delta \nu=\nu_+-\nu_-$ are written on top.
-We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$ indicated by a red line.
-We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$}
-\label{champ tranverse exp}
-\end{figure}
-
-We perform here a $T_1^{\rm dd}$ measurement similar to the one presented in Fig.
-[REF], on the same sample and with the same fitting parameters, but this time probing a class orthogonal to the magnetic field.
-
-Fig.
-\ref{champ tranverse exp}-a) and b) show the evolution of the transition frequencies $\nu_+=\frac{E_{\ket{3}}-E_{\ket{1}}}{h}$ and $\nu_-=\frac{E_{\ket{2}}-E_{\ket{1}}}{h}$ where $\ket{1}$, $\ket{2}$ and $\ket{3}$ are the states described in Fig.
-\ref{eigenstates transverse field}.
-The magnetic field was scanned between 20 and 130 G by increment of $\sim 0.5\ \rm G$.
-
-Fig.
-\ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ with the magnetic field.
-We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$.
-In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ is increased.
-
-\subsubsection{Region A}
-
-We attribute the decrease in region A to the double flips: as the magnetic field increases, so does $\Delta \nu$ and the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$.
-This value is coherent with the previously measured fluctuator linewidth.
-We can then estimate the part of the double flips in the total depolarization rate for the low field regime (20 G here) and find that they amount for $\sim 50\%$ in this case.
-
-Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see in Fig.
-\ref{champ tranverse exp}-a) that they are not very far detuned.
-However we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.
-
-\subsubsection{Region B}
-
-We now turn to region B.
-Even tough the double flips are completely quenched, the effects of the eigenbasis modification and the $T_2^*$ reduction are still present, because we can see on Fig.
-\ref{eigenstates transverse field}-b) that the eigenstates of the spin Hamiltonian are still very close to $\{ \ket{0}, \ket{\pm} \}$ for these magnetic field values.
-
-To evaluate the contribution of the eigenbasis and $T_2^*$, we measure the same class of NV centers, but this time with a longitudinal magnetic field ($B_\parallel \sim 100 G$) to get a baseline value of $1/T_1^{\rm dd}$ in the $\ket{\pm 1}$ basis.
-We found a value of $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ in this case, reported on Fig.
-\ref{champ tranverse exp}-c) as a green line.
-
-We do find that the decay rate is higher in the $\ket{\pm}$ basis by $\sim 50\% $ which corroborates our predictions, as both the change of eigenbasis from $\ket{\pm 1}$ to $\ket{\pm}$ and the increase of $T_2^*$ in the $\ket{\pm}$ basis are predicted to increase the depolarization rate.
-The theoretical increase however overestimates once again the actual increase: for the change of eigenbasis alone we predicted an increase of the decay rate by a factor of 4 (see appendix [REF]), over 8 times more than the actual increase.
-
-We unfortunately cannot experimentally discriminate the depolarization coming from the eigenbasis change or the $T_2^*$ while using the same sample, since both of this properties are tied to the Hamiltonian eigenstates.
-But we do know that in this scenario, for a single isolated class, the contribution of both the change of eigenbasis and $T_2^*$ amounts for $\sim 3$ times less than the double flips.
-
-\subsubsection{Conclusion of the experimental observation}
-
-We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field, and we expect that the depolarization for low magnetic fields behave similarly.
-We therefore claim that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
-\begin{enumerate}
-\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
-\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
-\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
-\end{enumerate}
-
-\section{Introduction to ensemble NV magnetometry}
-
-Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.
-
-NV magnetometry is generally decomposed between single NV and ensemble magnetometry.
-Single NVs,often associated with scanning probes, offer a spatial resolution limited only by how far the magnetic source can be from the NV center (typically $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
-
-Ensemble NV centers center on the other hand offer sensitivities as low as $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of sizes up to several mm.
-
-We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR.
-We will not try to give a complete overview of this field, but only of the elements close to the technique we present.
-For a general overview on NV magnetometry, we recommend the excellent reviews  \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.
-
-\subsection{AC and DC magnetometry}
-The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted into spin population.
-These protocols are generally distinguished between DC-broadband and AC-narrow band.
-
-For DC-broadband magnetometry, the general technique consist of a Ramsey interferometry experiment, presented in chapter 1, which is limited by the spin inhomogeneous coherence time $T_2^* \lesssim 1\ \rm{\mu s}$ for NV ensembles.
-The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$.
-Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry.
-The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
-
-For AC magnetometry, the basic protocol consist of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
-Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field.
-The AC magnetic field however has to be in phase with the rephasing pulses, which limits the AC sensing to a relatively narrow band.
-The state of the art in term of AC sensing is $\sim 1\ \rm pT/\sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}.
-
-Recently, \citep{xie2021hybrid} have achieved new records in both DC ($\sim 200\ \rm fT/\sqrt{\rm Hz}$) and AC ($\sim 10\ \rm fT/\sqrt{\rm Hz}$) NV magnetometry by using flux concentrators which are ferromagnets concentrating the magnetic field in a small volume, similarly to an electromagnet.
-Because the gain from the flux concentrators is not tied to a specific magnetometry protocol, we will focus on the results obtained without flux concentrators.
-\subsection{Low field magnetometry}
-Magnetometry for low magnetic field ($\gamma B < d_\perp E_\perp$) is challenging since in this regime the NV eigenstates and eigenenergies do not depend on the magnetic field in the first order.
-This complication is often lifted  by simply using a bias magnetic field, typically from a permanent magnet.
-
-Some systems however need to be kept under low to zero external magnetic fields, for example to study the J-coupling between spins \citep{sutter2012computational}, or the phase of skyrmions \citep{zazvorka2020skyrmion}.
-To study these systems with NV centers, we need to find new protocols for low field NV magnetometry.
-
-The main technique for low field magnetometry consist in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}.
-The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$ \citep{zheng2019zero}.
-
-Another technique consist in using the $^{13}\rm{C}-$NV complex we discussed in sec.
-[REF] which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are frist order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
-With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers.
-This technique is more suited for a carefully chosen single NV center.
-
-\subsection{Microwave-free magnetometry}
-All the precedent magnetemetry protocols described relied on the use of a controllable magnetic field, to coherently manipulate the NV spin state and to measure the phase imprinted by the magnetic field in the rotating frame.
-
-There are other protocols which correlate the magnetic field with the spin $T_1$ time (relaxometry).
-These protocols can operate without microwave field and work for either DC or high frequency AC sensing.
-
-\subsubsection{microwave-less DC magnetometry}
-The DC microwave-free protocols rely on sharp feature in the PL for specific magnetic fields around certain cross-relaxations processes or level anti-crossing.
-Fig.
-[REF] in chapter 2 show such PL features, in particular the ground state level anti-crossing (GSLAC, simply labeled LAC on the figure), the sharpest of these features.
-
-The GSLAC occurs for a purely longitudinal magnetic field where $\gamma B_z = D$, which corresponds to $B=1024\ \rm G$.
-In these conditions, the normally bright $\ket{0}$ state and the dark $\ket{-1}$ state become resonant and any residual mixing between these two states (coming from the electric field, the strain or the residual transverse magnetic field) will cause a drop in the NV polarization, and in the PL \citep{broadway2016anticrossing}.
-
-A protocol based on the GSLAC line has been implemented, first only for longitudinal magnetic field measurement \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.
-
-Another DC microwave-less protocol was proposed based on NV-NV CR \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
-This technique reconstructs the external magnetic field by scanning an other magnetic field, similar to Fig.
-[REF] in chapter 3, and by measuring the position of the NV-NV CR lines.
-By scanning the magnetic field in different directions, one can then recover the initial 3D magnetic field.
-The author of this paper did not measure the sensitivity of this protocol, but they predict a shot noise limited sensitivity of $24.7\ \rm nT/\sqrt{\rm Hz}$.
-
-\subsubsection{microwave-less AC magnetometry}
-The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed.
-The detection bandwidth is adjusted with an external DC magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.
-
-Sensitivities as low as a few $\rm pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
-
-\subsection{Orientation-free magnetometry}
-Every protocol described so far requires a precise knowledge  of the Larmor frequency between the $\ket{0}$ and $\ket{\pm 1}$ states as a function of the magnetic field.
-This knowledge implies that the direction of each NV axis in the frame of the magnetic field is known, because of the strong anisotropy in the NV Hamiltonian.
-
-For a single crystal, these axes can always be calibrated through ODMR.
-However, for a polycrystalline diamond, diamond powder or even a single diamond with an erratic motion, the knowledge of these axes is not possible.
-New protocols need to be found if one wants to perform magnetometry with these materials.
-
-To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
-This protocol however has an very low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective}.
-
-Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies.
-The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the Larmor frequency), or via the PL \citep{finco2021imaging}.
-
-\section{Low field depolarization magnetometry}
-\label{sec LFDM}
-We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the spontaneous depolarization of dense NV ensemble at low magnetic field.
-This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional field is required to improve performances.
-
-We will first describe its operation and characterize it before comparing it to the previously mentioned NV magnetometry protocols.
-
-\subsection{Principle of the experiment}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/setup_magneto}
-\caption{Principle of the LFDM protocol a) Experimental setup.
-The confocal microscopy setup is similar to the one described in [REF], the lock-in amplifier (LIA) is used here to add a modulation on the magnetic field, instead of the microwave.
-Both the PL and the LIA signal are monitored.
-b) example of LIA signal on sample ADM-15-4 when scanning the magnetic field $B_0$ between -150 and 150 G.
-c) PL signal corresponding to the LIA signal}
-\label{setup magneto}
-\end{figure}
-
-Fig.
-\ref{setup magneto} shows the principle of the LFDM protocol.
-The experimental setup is similar to the one presented in Fig.
-[REF] with the exception that there are no microwaves, and that the magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\tilde{B}_{\rm AC}$ used for the modulation.
-We typically use a modulation frequency $f_{\rm mod} \approx 1\ \rm kHz$, the same value we use for ODMR measurements.
-The main voltage $V_0$ responsible for the DC field and the oscillating voltage $\tilde{V}_{\rm AC}$ responsible for the AC field are summed via a homemade bias tee.
-
-Fig.
-\ref{setup magneto}-b) and c) show respectively the PL measured directly on the photodiode and the lock-in signal when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G.
-For this configuration, the magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.
-
-Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup.
-To improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing.
-Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.9\textwidth]{Figures/AC_and_DC_sensing}
-\caption{DC and AC LFDM sensing protocols on sample ADM-15-5.
-The magnetic field is scanned in an arbitrary direction.
-a) DC sensing: $B_0$ is scanned between $-$80 and +80 G while an AC field of amplitude 10 G is used for the lock-in detection.
-b) AC sensing: $B_0$ scans are performed for 5 different $|\tilde B_{\rm AC}|$ values between 2 and 10 G.
-The best AC sensitivity is achieved for $B_0\approx 5\ \rm G$.}
-\label{AC and DC sensing}
-\end{figure}
-
-Fig.
-\ref{AC and DC sensing} shows how the AC and DC sensing work for LFDM: for DC sensing the LIA signal around $B_0=0$ provide a sharp linear slope to convert the LIA voltage into a magnetic field.
-For AC sensing, the LIA signal is directly proportional to the amplitude of the field.
-The optimal AC field amplitude for DC sensing has to match the width of the PL dip $|\tilde{B}_{\rm AC}| = \sigma \approx 10\ \rm G$ in order to maximize the slope of the lock-in signal around $B_0=0$.
-the optimal DC field to use for AC sensing is $B_0= \sigma/2$ corresponding to the region where the lock-in signal is most sensitive to AC field.
-
-\subsection{Sensitivity of the LFDM protocol}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/sensi_alternee}
-\caption{Sensitivity measurement of the LFDM on sample ADM-15-4.
-a) Temporal trace of the LIA signal when applying an alternating external magnetic field of amplitude 35 $\mu$T and frequency $f=1\ \rm Hz$.
-b) Histogram of the DC measurement over 50 s.
-The data is fitted with Gaussians of width $1.6\ \rm{\mu T}$.}
-\label{sensei alternee}
-\end{figure}
-
-We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols.
-The characterization was only done for DC sensing.
-The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
-
-We characterize the DC sensitivity by looking at the noise floor on the LIA signal when sitting on the steepest slope, around $B_0=0$.
-The following parameters were used in the case presented: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
-
-Fig.
-\ref{sensei alternee} shows the sensitivity measurement protocol.
-First, the applied magnetic field is calibrated with ODMR spectra, then an alternating DC field of known amplitude (here $\Delta B_0=35\ \rm{\mu T}$) is applied on the sample.
-Fig.
-\ref{sensei alternee}-a) shows the temporal trace of the LIA signal in response to this alternating DC field.
-Knowing the amplitude of the alternating field, we can then convert the voltage signal into a magnetic field amplitude, with in this case a ratio $45 \pm 3\ \rm{\mu T /V}$.
-
-Fig.
-\ref{sensei alternee}-b) shows an histogram of the measured magnetic field amplitude for both the high and low value of the applied magnetic field.
-The total acquisition time was 50 s with the external field being alternated every second.
-The histograms are nicely fitted with Gaussian profiles of width $\sigma=1.5\pm 0.1\ \rm{\mu T}$ which allows us to estimate the DC sensitivity of the protocol:
-\begin{equation}
-\eta = \frac{\sigma}{\sqrt{2 \tau}}= 116 \pm 10\ \rm{nT} / \sqrt{\rm Hz},
-\end{equation}
-
-where $\tau=3\ \rm ms$ is the LIA low-pass filter time constant.
-
-We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$).
-Laser fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.
-
-\subsection{Angular dependence of the sensitivity}
-\label{sec angular sensi}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/sensi_angulaire}
-\caption{Angular dependence of the sensitivity.
-a) LIA signal for $B_0$ scanned between $-$150 and +150 G along the [100] axis, or along a direction 20$^\circ$ off the [100] axis.
-b) Zoom-in of the LIA signal on the zero-field slopes.
-c) Measured sensitivity as a function of the magnetic field angle with the [100] axis.}
-\label{angular sensi}
-\end{figure}
-
-We will now look at the dependence of the sensitivity with respect to the magnetic field angle.
-In particular, we are interested in comparing the sensitivity when $\mathbf{B}\parallel [100]$, where there is no lift of degeneracy between the four classes as $\mathbf{B}$ increases; and a case where $\mathbf{B}\nparallel [100]$ and where there is a lift of degeneracy between the four classes.
-We previously saw in Fig.
-\ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\mathbf{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.
-
-Fig.
-\ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angle of $B_0$.
-For the blue curve, the field was scanned along the [100] axis, for the orange curve, it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis.
-We can see that the maximum amplitude of the lock-in signal (relevant for AC sensing) is of the same order of magnitude for both case, and the slope around $B_0$ (relevant for DC sensing) is almost identical.
-Fig.
-\ref{angular sensi}-c) shows the sensitivity, measured with the protocol described in Fig.
-\ref{sensei alternee}, with respect to the angle between the magnetic field and the [100] axis.
-We can notice that the sensitivity depends only weakly on the magnetic field angle.
-
-The fact that the sensitivity is similar when $\mathbf{B}\parallel [100]$ or $\mathbf{B}\nparallel [100]$ means that the inter-class co-resonances plays a lower role in the sensitivity than the zero-field specific depolarization mechanisms, previously detailed in Sec.
-\ref{sec causes zero field}.
-We cannot be sure which of the three zero-field mechanism is the dominant one, but based on the experimental results shown in Fig.
-\ref{champ tranverse exp}, we can assume that the double-flips are the dominant factor in the LFDM sensitivity.
-
-It should be noted that this low angular dependence was particularly marked on sample ADM-15-4.
-Other samples, including from the same batch, showed a higher angular dependence with typically a sensitivity $2 \sim 3$ times worse for $B\parallel [100]$ than for the other directions.
-
-\subsection{Temporal stability}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/Temporal_just_0B}
-\caption{Temporal stability of DC-LFDM as measured on sample ADM-15-5.
-a) Temporal trace of the LIA signal for $B_0 \approx 0$.
-The Lock-in signal was converted to a measured magnetic field with the previously described protocol.
-b) Estimated signal to noise ratio for a signal of 100 nT as a function of the total acquisition time $T$.
-The noise value is computed by averaging the variance of the temporal trace over each time duration $T$.
-A curve $y=0.5\, T^{1/2}$ is added for comparison}
-\label{temporal 0B}
-\end{figure}
-
-An important aspect of sensors is their stability in time.
-The sensitivity being expressed in $\rm{nT} / \sqrt{\rm Hz}$ suggests that the signal to noise ratio (SNR) given by repeating the measurement increases as $\sqrt{T}$ where $T$ is the total acquisition time.
-This is true for an ideal system, but in practice drifts and low frequency noise mean that the actual SNR for long acquisition time will be worse than the SNR at short time scaled by a factor $\sqrt{T}$.
-
-One of the main source of drift for NV magnetometry based on spin resonance is the crystal lattice temperature change: the $D$ factor in the NV Hamiltonian (eq.
-[REF]) is sensitive to the lattice temperature because of the thermal dilatation of the crystal, which is the basis for NV thermometry \citep{kucsko2013nanometre}.
-This means that thermal fluctuations will impact the Larmor frequency of the spins and may cause a detuning with the initial measured frequency for long acquisition times.
-The LFDM protocol however does not rely on a specific resonance frequency, but rather on a co-resonnance condition between neighboring spins, which does not depend directly on the crystal temperature.
-
-Fig.
-\ref{temporal 0B} shows the temporal stability of the LFDM protocol over an acquisition time of 100 s.
-Fig.
-\ref{temporal 0B}-a) shows a temporal trace of the LIA signal where we have converted the signal into a measured magnetic field.
-Fig.
-\ref{temporal 0B}-b) shows the estimated SNR for a signal of 100 nT as a function of the total acquisition time and was computed from the temporal trace.
-
-We can see that the SNR does not deviate from the $\sqrt{T}$ scaling for times up to 50 s, and could probably be extended for longer times.
-The non-linearity for shorter times comes from the LIA low-pass filter.
-
-\section{Comparison between LFDM and other magnetometry protocol}
-Now that we have characterized the LFDM protocol, we can compare its performances with the previously mentioned magnetometry protocols.
-We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity.
-This discrepancy however come in large part to the smaller sensor volume used in the characterization, and to a relatively low PL collection efficiency compared to the other experiments.
-
-\subsection{Comparison with microwave based protocols}
-
-\begin{table}[htbp]
-\centering
-\caption{Comparison of the volume-normalized sensitivity of the best AC and DC protocols with LFDM}
- \label{table sensi volumiques}
-\begin{tabular}{c|ccc}
-\toprule
-{} & Zhou et al. \citep{zhou2020quantum} (AC) & Barry et al.\citep{barry2016optical} (DC) & LFDM (DC) \\
-\midrule
-$\eta\ (\rm{nT} / \sqrt{\rm Hz})$ & 92&0.015& 116 \\
-$V\ (\mu \rm{m}^3)$ &$8.1 \cdot 10^{-3}$ &$5.2 \cdot 10^{6}$& $3.3 \cdot 10^{3}$ \\
-$\eta_v\ (\rm{nT}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ &8.3&34&6700 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-We will first compare the LFDM protocol with the best microwave-based NV magnetometry protocols.
-A better metric than the sensitivity to compare different protocols is the volume-normalized sensitivity defined in \citep{zhou2020quantum} as $\eta_v=\eta\cdot \sqrt{V}$ where $\eta$ is the sensitivity and $V$ the effective volume of the sensor.
-This scaling assumes that each NV center in an ensemble is an independent magnetic field probe, and as the total number of NV centers is proportional to $V$, the sensitivity should scale as $1/\sqrt{V}$.
-
-Table \ref{table sensi volumiques} reports the best AC and DC volume-normalized sensitivities found in the literature.
-The LFDM protocol performs $\sim 200$ times worse than the best DC magnetometer.
-It should be noted however that the PL collection in our measurement, where we use a confocal microscope with a numerical aperture NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/compariason_0B_MW}
-\caption{Comparison between LFDM and microwave-based magnetometry on the same experimental setup.
-The sample used is ADM-15-5 and the magnetic field is aligned close to a [111] axis.
-a) Schematics of the experiment: the setup is the same as the one described in Fig.
-\ref{setup magneto} with the addition of a continuous microwave tone at the frequency $\nu=2600\ \rm MHz$.
-b) LIA signal as $B_0$ is scanned from $-$150 to +150 G.
-The two lines at $\pm 100\ \rm G$ are caused by the microwave.
-c) Temporal trace of the measured B field for $B_0 \approx 0$ (on the middle of the zero field slope) and $B_0 \approx 100\ \rm G$ (on the middle of the microwave slope).
-The orange curve is shifted by $\approx 100\ \rm G$ to sit on the same scale as the blue curve.
-d) Estimated SNR fo a signal of 100 nT for $B_0 \approx 0$ and $B_0 \approx 100\ \rm G$ as a function of the total acquisition time}
-\label{comparaison 0B MW}
-\end{figure}
-
-Fig.
-\ref{comparaison 0B MW} shows a comparison between LFDM and microwave-based magnetometry using our setup.
-A fixed microwave tone (here at 2600 MHz) is sent to the diamond and a magnetic field $B_0$ close to the [111] axis is scanned along an oscillating field $\delta B$ to perform a lock-in detection.
-
-Fig.
-\ref{comparaison 0B MW}-b) shows the LIA signal over a $B_0$ scan: the profile is similar to the one in Fig.
-\ref{setup magneto}-b) with the addition to two sharp lines for $|B_0| \approx 100\ \rm G$.
-These two lines correspond to the transition $\ket{0} \to \ket{-1}$ (or $\ket{0} \to \ket{+1}$ for the negative fields) of the class aligned with the magnetic field when it comes to resonance with the 2600 MHz microwave tone.
-The microwave power used here correspond to a Rabi frequency $\Omega \approx (2\pi) 3\ \rm MHz$ and the slope on the steepest part of the microwave line is $\sim 2.5$ times higher than the slope in zero field.
-
-Fig.
-\ref{comparaison 0B MW}-c) and d) show the temporal stability of the sensor similarly to Fig.
-\ref{temporal 0B}, on both the zero field line and the microwave line.
-While we can see that the microwave signal is more sensitive thanks to the higher slope, it is also more prone to drift, most likely because of the shift of the Larmor frequency caused by temperature fluctuations.
-We can even see that the LFDM becomes more sensitive for acquisition times beyond 10 s.
-
-We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the available microwave power (although in this case $\Omega_{\rm Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$.
-Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.
-
-\subsection{Comparison with microwave-less protocols}
-The only microwave-less protocols to have measured a sensitivity are the GSLAC protocols described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
-Unfortunately the sensing volume in this studies is not known, however due to the similarity in our approaches, we should be able to directly compare the physical origin of both protocols, i.e.
-comparing the zero field PL dip and the GSLAC PL dip.
-
-In Fig.
-\ref{setup magneto}-c) we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
-In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type Ib irradiated diamond, similar to the ones we use.
-In \citep{zheng2020microwave} they manage to keep a contrast of 4.5 \% while reducing the GSLAC linewidth to $0.38\ \rm G$ by using a CVD sample isotopically enriched in $^{12}$C with a lower [NV] and [N] density.
-
-The zero field technique is therefore comparable with the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear what kind of material engineering would improve its sensitivity.
-
-%Je parle pas de la dépendence en puissance ou du champ transverse, je laisse ça pour plus tard.
-\section{Conclusion and perspectives}
-We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field, and wee have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers.
-We have identified three mechanisms by which the depolarization is increased in zero-field: the inter-class flip-flops, the double flips, and the change in the NV eigenstates (which includes the change in $T_2^*$).We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
-
-Finally, we have seen how this low-field depolarization can be applied to perform magnetometry experiments, and what advantages and disadvantages this technique offers compared to the established NV magnetometry protocols.
-
-I will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.
-
-\subsection{LFDM optimization}
-
-Given the few parameters involved in the LFDM protocol, the main path for LFDM optimization seem to be the material optimization of the samples.
-
-Most magnetometry protocols have a relatively clear material optimization path: DC protocols want to improve the figure of merit $\rho T_2^*$ and AC protocols want to improve $\rho T_2$, where $\rho=[\rm NV]$ is the NV concentration.
-The reason is that the number of NV centers for a given volume scales as $\rho$, and therefore the sensitivity scale as $\sqrt{\rho}$.
-Meanwhile increasing the density of defects tend to decrease the spin coherence times $T_2$ and $T_2^*$, and the AC and DC sensitivities scale respectively as $1/\sqrt{T_2}$ and $1/\sqrt{T_2^*}$.
-
-For the LFDM protocol however, the material optimization is far from obvious.
-Indeed, while increasing $T_2^*$ would improve the sensitivity, the gain might not be as interesting as for other protocols due to the broadening caused by the fluctuator lifetime.
-On the other hand, increasing the NV density not only increases the number of sensor per volume, but also increases the dipole-dipole strength between NV centers.
-If we assume that fluctuators are constituted of closely packed impurities, then increasing the NV density should also increase the fluctuator proportion.
-
-It seems then that increasing the NV density would be the obvious path to improve the LFDM performances.
-In practice however, I observed that the relationship between NV density, fluctuators, and LFDM performances was generally not straightforward.
-The two following examples will illustrate that.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/substrat}
-\caption{Measurements on sample SBST-C for two different regions A and B.
-a) PL contrast as a function of a magnetic field scanned in an aribitrary direction (same direction for both regions).
-b) PL contrast as a function of a magnetic field scanned in the [100] direction.
-c) $T_1$ measurement on a single class when $B\neq 0$.
-d) $T_1$ measurement on all  four classes when $B=0$ }
-\label{substrat chelou}
-\end{figure}
-
-Fig.
-\ref{substrat chelou} shows various measurements on the same sample SBST-C, a HPHT bulk sample containing [NV]$\sim  \rm ppm$, on two spatially distinct regions A and B.
-These two regions most likely correspond to two distinct growth sectors of the diamond, meaning that these regions were not grown in the same exact environment, which can impact among other things the creation of impurities.
-Both region showed a similar PL which would indicate a similar density of NV centers.
-
-The metric we are interested in to evaluate the LFDM performance are the PL contrast when $\mathbf{B}$ is scanned in an arbitrary direction, which mainly represents the inter-class flip flop and which I will refer to as FF-contrast; and the PL contrast when B is scanned $\mathbf{B}$ is scanned along the [100] direction, which mainly represents the double flips and which I will refer to as DF-contrast.
-We are interested in the DF-contrast since we have seen in sec.
-\ref{sec angular sensi} that it could be the main contribution to the LFDM sensitivity.
-
-Fig.
-\ref{substrat chelou}-a) and b) indicate that region B shows a higher FF-contrast, but a smaller DF-contrast than region A.
-Looking at the $T_1$ measurements for both regions, region B shows a shorter $T_1$ in both zero field and non-zero field, which would indicate a larger fluctuator density in region B.
-If this is true, than the fluctuator density is not directly correlated to the DF-contrast.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/double_adamas}
-\caption{Same measurements as Fig.
-\ref{substrat chelou} on two distinct samples ADM-15-4 and ADM-15-1.
-Every parameters (laser power,optical setup , AC magnetic field) was kept identical for the two samples.}
-\label{double dragon}
-\end{figure}
-
-We now turn to Fig.
-\ref{double dragon}.
-This figure shows the same 4 measurements but this time on two distinct samples, ADM-15-1 and ADM-15-4, which came from the same batch.
-Both samples showed relatively similar PL which could again indicate a similar NV density.
-
-We can first notice the disparity in the  zero field PL contrasts: the FF-contrast is twice as big in sample ADM-15-4 as in sample ADM-15-1, and the DF-contrast is almost 5 times bigger.
-When looking at the $T_1$, we can notice that the lifetime for non-zero field is slightly longer for sample ADM-15-4 than for ADM-15-1, but its lifetime in zero-field is shorter (it is also exponential and not stretched-exponential as discussed in the last chapter [REF], which could have an influence on the PL contrast).
-From the $T_1$ measurements we can not determine which sample contains the most fluctuators, and therefore correlate the fluctuator or NV density with the DF and FF-contrasts.
-
-These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field.
-We have noticed a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive.
-If these parameters were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.
-
-Other non-material parameters in the LFDM protocols should also be investigated.
-This includes the temperature of the sample, the intensity and wavelength of the laser, the frequency and amplitude of the AC magnetic field.
-For the results presented in sec.
-\ref{sec LFDM}, we chose the parameters which we found optimal within our available experimental range.
-
-\subsection{Potential applications for LFDM}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/Applications}
-\caption{Schematics of two potential application for LFDM. The working of each application is detailed in the main text}
-\label{applications}
-\end{figure}
-
-We will now turn to potential applications for LFDM, which could be useful even without material optimization.
-
-Fig.
-\ref{applications} illustrates two potential applications for LFDM.
-The first one is to to perform wide-field magnetic imaging on a surface with micron-sized irregularities.
-The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}.
-While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism samples.
-Furthermore, controlling the angle between the diamond and the sample can be challenging even for two perfectly flat surface.
-
-Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always have a sensor within $\sim$ 1 $\mu$m of the sample.
-Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.
-One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$m red fluorescent diamond) are well suited for this task.
-I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples.
-This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}.
-Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required.
-Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field homogeneity.
-
-Another potential application for LFDM is shown in Fig.
-\ref{applications}-b.
-It consists in the detection of magnetic field gradient thanks to a nano-diamond probe attached to a nano-resonator.
-The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig.
-\ref{AC and DC sensing}-b).
-This idea is based on previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}.
-A very similar protocol was recently developed with microwave based AC sensing \citep{huxter2022scanning}.
-
-In order to use AC LFDM with a scanning probe setup, one would need to have a nano-diamond with at least one NV center and one fluctuator.
-This may not be such a rare event: a (20 nm)$^{3}$ nanodimaond with [NV]=5 ppm contains on average $\approx 4$ NV centers.
-If we take a fluctuator proportion of 30 \% as measured in \citep{choi2017depolarization}, then finding a nanodiamond with at least one NV and one fluctuator  seems feasible.
-I have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect, and I therefore cannot estimate the sensitivity of such a probe.
-
-Other potential applications for LFDM includes the possibility to use polycrystalline diamond, potentially grown through heteroepitaxy on top of a surface of interest, or to follow individual diamonds flowing in liquid or living organism, similar to what was done in \citep{feng2021association}, but without having to constantly calibrate the diamond axes through ODMR which significantly hinders the temporal resolution of microwave-based technique.
-
-Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous magnetic field or microwave field along the sample.
-
 \printbibliography

### 2022-10-16 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

@@
 \begin{document}
-\chapter{NV-NV CR under transverse or low fields: application to magnetometry}
-
-In the last chapter we saw how dense ensemble of NV centers have an intrinsic depolarization mechanism mediated by the NV-NV dipole interaction.
-In this chapter we will see how this feature can be exploited to perform magnetometry.
-
-Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
-We will focus here on the low to zero magnetic field region, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the more complex physics of the NV centers under low magnetic field.
-
-This chapter is constructed as follows: We will first describe the physics of NV centers under low or transverse magnetic field (the physics is similar in both cases), we will then show both theoretically and experimentally that the NV-NV CR is modified for low and transverse magnetic field.
-We will identify different mechanisms by which the relaxation is modified and we will try to quantify their relative importance.
+\chapter{NV-NV cross-relaxation at low fields: application to magnetometry}
+
+In the last chapter we have identified a dipole-dipole relaxation mechanism within dense NV centers ensemble, and we have studied its influence in non-zero magnetic field.
+In this chapter we will cover the low to zero magnetic field region and see how the depolarization at low magnetic field can be used for magnetometry.
+
+This chapter is constructed as follows: We will first describe in sec. \ref{sec 4.1} the physics of NV centers under low or transverse magnetic field. In sec. \ref{sec causes zero field}
+we will present the different zero field depolarization causes that we have identified as well as their theoretical impact on the NV relaxation rates according to the fluctuator model.
+In sec.
+\ref{sec 4.3} we will confront the predictions of the fluctuator model with experiments under low and transverse magnetic field.
 
 We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field.
-We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols.
+We will start by giving a short overview of NV ensemble magnetometry in sec.
+\ref{sec 4.4} before characterizing the LFDM protocol in sec.
+\ref{sec LFDM} and comparing it to other magnetometry protocols in sec.
+\ref{sec 4.6}.
 Finally we will give some perspectives on how to improve LFDM and on potential applications.
@@
 \section{NV spin Hamiltonian under low and transverse fields}
-
-Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates compared to the high magnetic field regime.
+\label{sec 4.1}
+We will start this chapter by studying the single NV physics under low and transverse magnetic field.
+On of the main difference with the previously studied regime is that the eigenstates of the Hamiltonian are no longer $\{ \ket{0}, \ket{+1}, \ket{-1} \}$, which will then in turn modify the dipole-dipole interaction between the NV centers.
+
+We will first describe the spin Hamiltonian of the NV center under low magnetic field and determine the dominant factors in our samples, then do the same for transverse magnetic field, and finally investigate the modification of the ODMR lineshapes in those regimes.
 
@@
 \includegraphics[width=\textwidth]{Figures/ESR_simus}
-\caption{Simulations of inhomogeneous zero field ODMR when sampling various parameters.
-a) Simulation when sampling each components of the magnetic field over a Gaussian of deviation $\sigma=2\ \rm G$.
-b) Simulation when sampling each components of the strain tensor $\bar{\bar{\varepsilon}}$ over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
-c) Simulation when sampling each components of the electric field over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
+\caption{Monte-Carlo simulations of inhomogeneous zero field ODMR for different noise sources.
+a) Magnetic noise: each of the $\mathbf{B}$ components is sampled over a Gaussian of deviation $\sigma=2\ \rm G$.
+b) Strain noise: each of the components of the strain tensor $\bar{\bar{\varepsilon}}$ is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
+c) Electric noise: each of the $\mathbf{E}$ components is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
 d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
@@
 
-In the absence of external magnetic field, we have to take into account other elements which would otherwise be of second order in the spin Hamiltonian.
-These elements are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}.
-The hyperfine splitting due to nearby nuclei will be considered separately, although to a large extent it behaves like a local magnetic field.
-
-Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian).
+In the absence of a strong external magnetic field, the dominant terms of the NV spin Hamiltonian originate from local noises.
+These noises are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}.
+
+Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field.
 The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.
@@
 
-Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.
-
-The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting.
+%Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.
+%
+%The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting.
 For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$.
@@
 Fig.
-\ref{simus ESR} shows a simulation of how each parameters of the spin Hamiltonian - local magnetic field, local electric field and strain - affects the zero external field ODMR profile.
-To do these simulations, I sampled each parameters separately $10^6$ times and plotted the histogram of the two eigenvalues of the Hamiltonian written in eq.
-\ref{Hamiltonien pm1}.
-Fig.
-\ref{simus ESR}-d) shows an experimental zero field ODMR spectrum, typical of what we observe with dense NV ensembles.
-
-Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR.
-Given the simulation results, the only parameter that can give rise to this shape is the local electric field.
-
-We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample.
+\ref{simus ESR}-a) b) and c) shows the result of Monte-Carlo simulations of the ODMR of an ensemble of NV centers for different noise sources.
+The simulation work as follow: one of the three possible noise source (magnetic field, electric field or strain) is sampled randomly while the two other are kept at zero.
+The eigenvalues of the spin Hamiltonian given in eq.
+\ref{Hamiltonien pm1} are then computed and the process is repeated $10^6$ times to create the histograms shown in Fig.
+\ref{simus ESR}.
+
+Fig.
+\ref{simus ESR}-d) shows an experimental zero-field ODMR which is typical of the samples used in this manuscript.
+In particular we can notice a central splitting of the spectrum.
+This splitting is only present in the simulation when the dominant noise is the electric field.
+We then conclude that the dominant noise source in our sample (in the absence of external magnetic field) is the local electric field.
+
+The authors of \citep{mittiga2018imaging} came to the same conclusion for the samples with dense NV ensemble that they studied by using a more sophisticated simulation model.
+
+%We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample.
 In particular, we do not shield for the earth magnetic field in our experiment.
@@
 
-We will therefore consider that the low external magnetic field spin Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.
-
-We will then adopt the following simplified Hamiltonian for the zero external field regime:
+Given that $d_\perp \approx 50 d_\parallel$, we can consider that the dominant electric field noise comes from the transverse electric field.
+We will then adopt the following simplified Hamiltonian for the zero external magnetic field regime:
 \begin{equation}
@@
 
-We now turn to the case of purely transverse magnetic field with respect to the NV axis, i.e.
+We now turn to the case of a magnetic field that is orthogonal to the NV axis, i.e.
 $\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
@@
 
-In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+In conclusion, whenever the spin Hamiltonian is dominated by the local electric field or by the transverse magnetic field, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
 
@@
 A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
-The energy of the $\ket{\pm 1}$ states depend linearly on the longitudinal magnetic field, and to the second  order on the transverse electric and magnetic fields.
-On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and to the second order on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp$.}.
+The energy of the $\ket{\pm 1}$ states depends linearly on the longitudinal magnetic field, and quadratically on the transverse electric and magnetic fields.
+On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and quadratically on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp$.}.
 
 The shape of the ODMR lines (i.e.
-the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis.
-Further details on the low magnetic field physics of the NV center are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.
+the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis, as we established in the previous part.
+Further details on the low magnetic field regime are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.
 
@@
 
-\section{NV-NV CR in the low magnetic field regime}
-In this part, we will study the cross-relaxation between NV centers under low magnetic field, whose spin Hamiltonian is dominated by a transverse electric field.
-Understanding the various mechanisms operating in this regime will be important for the magnetometry protocol presented after.
-
-\subsection{Experimental observations}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
-\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction.
-a) ODMR spectrum for B=0.
-b) ODMR spectrum for B=60 G.
-c) PL contrast as a function of the external magnetic field.
-d) $1/T_1^{\rm dd}$ as a function of the external magnetic field.
-A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
-\label{scan 1x1x1x1}
-\end{figure}
-We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.
-
-The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust in order to isolate each effects.
-
-Fig.
-\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G.
-The ODMR at the initial and final magnetic fields are shown in Fig.
-\ref{scan 1x1x1x1}-a) and b).
-
-While it is clear that the spin lifetime as well as the PL increases with the magnetic field, there is no clear indication that this is because of the specificity of the low field region.
-Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.
-
-We can note however that the PL is not an exact mirror of the spin dynamics: while the low field drop in PL is indeed associated with a reduced spin lifetime, the drop in PL at higher magnetic field comes from the states mixing induced by the transverse magnetic field, and is not associated with a modification of the spin dynamics (to the first order).
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\textwidth]{Figures/scan_100}
-\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis.}
-\label{scan 100}
-\end{figure}
-
-Fig.
-\ref{scan 100} presents a way to circumvent this issue: by applying the magnetic field along the [100] crystalline axis, we can make sure that the four classes of NV centers always stay resonant regardless of the magnetic field amplitude.
-
-We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig.
-\ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of $\sim$ 8 in zero field, whereas in Fig.
-\ref{scan 100}, it was only reduced by a factor of $\sim$ 1.2.
-The main reason for the PL and $T_1$ drop in Fig.
-\ref{scan 1x1x1x1} was indeed the co-resonance between the four classes.
-
-Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained if we consider only the inter-class resonances.
-This means that there are some additional depolarization mechanisms which are proper to the zero field region.
-We should also note that, while the zero-field PL contrast is bigger in Fig.
-\ref{scan 1x1x1x1}-c) than in Fig.
-\ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually similar in both cases.
-
-We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]
-
-\subsection{Theory of the low field depolarization}
+\section{Theory of the low field depolarization}
 \label{sec causes zero field}
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Shema_summary_theory}
-\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
+\includegraphics[width=\textwidth]{Figures/Shema_summary_theory_2}
+\caption{Summary of the different depolarization effects in low magnetic field and their predicted depolarization rate. The ODMR spectra were taken on sample ADM-150-1}
 \label{summary_theory}
@@
 
-Fig. \ref{summary_theory} summarizes the different contribution that we have identified to the depolarization of the NV centers at low magnetic field.
-
-The first and main contribution is degeneracy between the four classes as the magnetic field decreases.
-Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 40$.
-This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.
-
-The three other contributions, which are the eigenbasis change, the presence of double flips and the increase in $T_2^*$ will be detailed below and in the appendix [REF].
-These three contributions are the reason for the depolarization at low magnetic field in Fig.
-\ref{scan 100}.
-In total, they amount for a further increase in the relaxation rate by a factor of $\sim 2$ compared to the $\mathbf{B} \parallel [100]$ case.
-Although this increase is small compared to the 4-classes degeneracy, we will see in sec.
-\ref{sec LFDM} that it is relevant for the  low field magnetometry protocol.
+We will focus on this section on the mechanisms responsible for the depolarization of dense NV ensemble at low magnetic fields.
+We have identified four such mechanisms that lead to an increase in the resonant dipole-dipole coupling between NV and fluctuators, which in turn results in an increased spin depolarization rate.
+
+Fig.
+\ref{summary_theory} summarizes these four mechanisms along with the theoretical increase in the depolarization rate caused by each of them.
+We will detail below the physical origin of these four mechanisms and the calculations that lead to the values presented in Fig.
+\ref{summary_theory}.
 
@@
 
-\subsubsection{Eigenstates modification}
+\subsection{Four classes degeneracy}
+
+The first and main contribution to the low field depolarization is the degeneracy between the four classes as the magnetic field decreases.
+Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 43$.
+This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.
+
+\subsection{Eigenstates modification}
 
@@
 
-A consequence of this change is the absence of diagonal terms  in the dipole-dipole Hamiltonian for the $\{ \ket{0}, \ket{+},\ket{-} \} \times \{ \ket{0}, \ket{+},\ket{-} \}$ basis, which explains why the NV centers states are protected from the magnetic noise at low magnetic field \citep{jamonneau2016competition}.
-
-Another consequence is a slight modification in the flip-flop rates.
+A consequence of this change is the absence of diagonal terms  in the dipole-dipole Hamiltonian written in the $\{ \ket{0}, \ket{+},\ket{-} \} \times \{ \ket{0}, \ket{+},\ket{-} \}$ basis, which explains why the NV centers states are protected from magnetic noises at low magnetic field \citep{jamonneau2016competition}.
+
+Another consequence is a slight modification of the flip-flop rates.
 Once averaged over all possible orientation and position, as detailed in appendix [REF], this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
@@
 
-\subsubsection{Double flips}
-
-The double flip, or flip-flip, processes that we consider here are the one that couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$).
-These terms are normally not resonant because of the zeeman splitting between the $\ket{+1}$ and $\ket{-1}$ states.
+\subsection{Double flips}
+
+The double flip processes have been previously discussed in sec.
+[REF] and naturally arise from the dipole-dipole Hamiltonian.
+The ones that we consider here couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$).
+These terms are generally not resonant because of the Zeeman splitting between the $\ket{+1}$ and $\ket{-1}$ states.
 
 In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$.
-These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely turned off.
+These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely quenched.
 
 The computation of the depolarization rate including the double flips is detailed in appendix [REF].
-To obtain a numerical value, we need to input the experimental values of the dipole-dipole interaction range \footnote{By interaction range we mean the constant to be compared to the spectral detuning $\Delta \nu$ between the two groups.
+To obtain a numerical value, we need to input the experimental values of the dipole-dipole interaction range \footnote{By interaction range we mean the value to be compared to the spectral detuning $\Delta \nu$ between the two groups.
 This is $2\gamma_f+\Gamma_{NV}+\Gamma_f$ in eq.
 [REF]} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
-[REF], as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ state.
-We typically measure $\Delta \nu \approx 8\ \rm MHz$ on type 1b HPHT samples.
-
-For the value presented in Fig. \ref{simus ESR}, we inputted $\Gamma=\Delta \nu$.
+[REF], as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states.
+We typically measure $\Delta \nu \approx 8\ \rm MHz$ on type Ib HPHT samples.
+
+For the value presented in Fig. \ref{simus ESR}, we inputted $\Gamma=\Delta \nu=8.8\ \rm MHz$.
 %\begin{table}[htbp]
@@
 
-\subsubsection{$T_2^*$ modification}
-
-The last aspect we considered was the change in the ODMR lineshape discussed in sec. \ref{sec modif T2*} for weak magnetic field (we refer to this as $T_2^*$ for conciseness).
-
-To quantify its impact in the depolarization rate, we employ eq.
+\subsection{$T_2^*$ increase}
+
+The last aspect we considered was the change in the ODMR lineshape discussed in sec. \ref{sec modif T2*} for weak magnetic field (we refer to the spectral linewidth as $T_2^*$ for conciseness).
+
+To quantify the impact of the $T_2^*$ increase in the depolarization rate, we employ eq.
 [REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
 \ref{ESR for T2*}.
-We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
-
-The value shown in Fig.
-\ref{summary_theory} corresponds to a $9 \%$ increase, as most of the samples studied in this manuscript are type 1b HPHT diamonds (therefore more similar to sample Sumi-2 than to sample CVD-pink).
-
-\subsection{Other potential causes for zero field depolarization}
-We have also considered two other potential causes that we have ultimately ruled out in this case.
-
-\subsubsection{Magnetic field alignment}
+We then find that the increase of $T_2^*$ in zero magnetic field occasions an increase of the decay rate by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
+
+The value shown in Fig. \ref{summary_theory} corresponds to a $9 \%$ increase.
+
+\bigskip
+
+In conclusion, we predict that the 4-classes degeneracy is by far the dominant causes for low field depolarization ($\sim \Gamma_1 \times 40$), followed by the double flips ($\sim \Gamma_1 \times 2$).
+We will however see in sec \ref{sec LFDM} that the double flips, along with the change of eigenbasis and increase in $T_2^*$ are determinant factor in the sensitivity of the low field magnetometry protocol.
+
+In the following section we will confront the theoretical results with experimental ones.
+
+\section{Experimental investigation of the low field depolarization}
+\label{sec 4.3}
+In this section we will try to experimentally discriminate the contribution of the four depolarization sources described in Fig.
+\ref{summary_theory} and confront the experimental values with the theoretical ones.
+
+We will first determine the contribution of the 4-classes degeneracy by comparing [100] aligned and non-[100] aligned magnetic field in the weak magnetic field regime, before studying the influence of the double-flips thanks to purely transverse magnetic fields.
+
+On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\mathbf{B}\parallel [100]$) and the influence of the laser polarization.
+Following further experiments detailed in appendix [REF], we have determined that these two effects were negligible compared to the four ones we identified before.
+
+\subsection{NV-NV CR in the low magnetic field regime}
+
+In this part, we will try to isolate the contribution of the 4-classes degeneracy from the three other ones by using the fact that all four classes can be at resonance for two distinct scenario:
+\begin{itemize}
+\item When $\mathbf{B}=0$.
+\item When $\mathbf{B}\parallel [100]$ (cf. Fig. [REF] (la figure avec toutes les config)).
+\end{itemize}
+
+\subsubsection{Magnetic scan in a random direction}
+
 \begin{figure}[h]
 \centering
-\includegraphics[width=.7\textwidth]{Figures/alignement_rose}
-\caption{PL of sample CVD-pink as a function of the magnetic field for various misalignment of the field direction with the [100] axis}
-\label{alignement rose}
-\end{figure}
-
-The first is a potential misalignment of the magnetic field with the [100] axis, which would cause a splitting of the 4 classes as the magnetic field increases.
-This hypothesis is not consistent with the observations in Fig.
-\ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased.
-Instead we observe a sharp decrease followed by a pseudo-plateau.
-
-We still estimated how sensitive the PL dip was with respect to the field alignment.
-Fig.
-\ref{alignement rose} shows PL scans on sample CVD-pink for various alignment of the magnetic field around the [100] axis.
-We estimate our alignments to be precise within $\pm 1^\circ$.
-We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization does not come from a field misalignment.
-
-\subsubsection{Laser polarization}
+\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
+\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction.
+a) ODMR spectrum for B=0.
+b) ODMR spectrum for B=60 G.
+c) PL contrast as a function of the external magnetic field.
+d) $1/T_1^{\rm dd}$ as a function of the external magnetic field.
+A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
+\label{scan 1x1x1x1}
+\end{figure}
+
+First we will run a control experiment by scanning the magnetic field in a non [100] direction.
+Fig.
+\ref{scan 1x1x1x1}-a) and b) shows ODMR spectrum at the starting point ($|\mathbf{B}|=0$) and end point ($|\mathbf{B}|\approx 60\ \rm G$) of the magnetic field scan.
+The magnetic field orientation was chosen in order to maximize the splitting of the four classes.
+
+Fig.
+\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as a function of the magnetic field.
+We can observe a decrease of $T_1^{\rm dd}$ by a factor of $\sim 10$ in zero magnetic field compared to non-zero magnetic field, with a half-width at half maximum of $\sim 10\ \rm G$.
+Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\mathbf{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$.
+This is because this drop of PL is associated with transverse magnetic field state mixing, and not with spin depolarization.}.
+
+We attribute the spin depolarization observed in Fig.
+\ref{scan 1x1x1x1}-d) to all four mechanisms described in the last part acting simultaneously.
+To determine the contribution of each mechanism individually, we will need to run dedicated experiments.
+
+\subsubsection{Magnetic scan with $\mathbf{B} \parallel [100]$}
+
 \begin{figure}[h]
 \centering
-\includegraphics[width=.7\textwidth]{Figures/pola_laser}
-\caption{PL of sample ADM-150-3 with respect to the magnetic field for various laser polarization angle.
-The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
-\label{pola laser}
-\end{figure}
-
-Other studies \citep{anishchik2015low, filimonenko2020weak} previously saw a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.
-
-Fig.
-\ref{pola laser} shows PL scans on sample ADM-150-3 for various polarization angle of the incident laser.
-We chose to apply the magnetic field in the laser polarization plane, in a direction that did not match any particular crystalline planes.
-
-We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$.
-We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.
-
-\section{NV-NV CR under purely transverse magnetic field}
-We have seen in the last part that there was a spontaneous spin depolarization specifically in zero external magnetic field, and we have identified three potential mechanisms which are the double-flips, the change in the Hamiltonian eigenstates and the modification of $T_2^*$.
-We now want to experimentally discriminate the contribution of each of these mechanisms and to see how close the actual decay rates are to the predicted ones.
-
-One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic and electric noise ...).
-Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.
-
-The ideal way to isolate each contributions would be to apply an electric field strong enough to split the transition of the 4 classes, and to measure the decay rate of a single class dominated by the transverse electric field.
-Such an electric field would need to be of the order of $\sim 10^6\ \rm{V/cm}$, which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.
-
-Instead of an electric field, we will use a transverse magnetic field, which as described in sec.
-\ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
-
-\subsection{Principle of the experiment}
+\includegraphics[width=0.9\textwidth]{Figures/scan_100}
+\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis.}
+\label{scan 100}
+\end{figure}
+
+This second experiment is similar to the last one except that the magnetic field was scanned in the [100] direction.
+Fig.
+\ref{scan 100}-a) and b) again show ODMR at the starting and ending point of the magnetic scan, while Fig.
+\ref{scan 100}-c) and d) show the PL and $1/T_1^{\rm dd}$ recorded during the scan.
+
+The main difference with the previous experiment is that in this case, the degeneracy between the four classes is never lifted.
+We therefore attribute the increase in $1/T_1^{\rm dd}$ at low magnetic field to the three other identified mechanisms.
+The increase in the depolarization rate in zero field ($\sim 20\ \%$) as well as the decrease inPL ($\sim 0.5 \%$) are still present, but significantly smaller than in the previous case.
+We can note however that the half-width at half maximum of the zero field depolarization ($\sim 2\ \rm G$) is also smaller in this case.
+
+We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF].
+We can not explain however why the depolarization rate appear to be smaller at $B \sim 10\ \rm G$ than it is for $B > 30\ \rm G$.
+
+The fact that the increase of the depolarization rate at zero field is about $\sim 50$ times weaker in Fig.
+\ref{scan 100} than it was in Fig.
+\ref{scan 1x1x1x1} confirms that the leading cause of depolarization in zero field is the degeneracy between the four classes of NV centers.
+
+\subsection{NV-NV CR under purely transverse magnetic field}
+
+Thank to the two previous experiments, we were able to discriminate the contribution of the 4-classes degeneracy from the three other mechanisms that are which are the eigenbasis change, the double-flips and the $T_2^* increase$.
+We now seek to discriminate the contribution of each of these three mechanisms.
+
+We unfortunately were not able to discriminate the change of eigenbasis from the $T_2^* increase$, which would likely require a statistical study over different samples with different $T_2^*$ values.
+We were however able to discriminate the double flip contribution from the two other ones by scanning the magnetic field in a direction orthogonal to one NV class ($\mathbf{B} \perp [111]$).
+
+\subsubsection{Principle of the experiment}
 \begin{figure}[h]
@@
 
+The idea behind the experiment is that the Hamiltonian eigenbasis is the same (or close to be the same) when $\mathbf{B}=0$ than it is when $\mathbf{B} \perp [111]$, as described in sec.
+\ref{sec B transverse}.
+For simplicity, we will assume here that both the magnetic field and the transverse electric field lies along the NV $x$ axis : $\mathbf{B}=B_x \hat{e}_x$ and $\mathbf{E}_\perp =E_x \hat{e}_x$
+
 Fig.
 \ref{eigenstates transverse field} shows a simulation of the eigenstates and eigenenergies for a NV Hamiltonian subject to transverse electric field and magnetic field.
-For convenience we chose to take $E_\perp = E_x$ and $B_\perp = B_x$, taking other geometric configuration alters only slightly the results.
 We labeled the eigenstates of the Hamiltonian $\ket{1}$, $\ket{2}$ and $\ket{3}$ in ascending order of energy.
@@
 \ref{eigenstates transverse field}-b) shows both the splitting $\Delta \nu$ between the $\ket{2}$ and $\ket{3}$ states, and how close the $\ket{3}$ state is to the $\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}}$ state via the factor $\abs{\bra{3}\ket{+}}^2$.
-We look at the $\ket{3}$ state since $\ket{2}=\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}}$ in this case.
-
-These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 since we wanted to observe the modification caused by the eigenbasis $\{\ket{0},\ket{\pm}\}$.
-This way we can isolate the $T_1^{\rm dd}$ contribution coming from the change of eigenstates to the one coming from the double flips.
-Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100,200] G seems to satisfy both these criteria.
-
-\subsection{Results}
+We should note that with the assumption that $B_\perp \parallel E_\perp$, the $\ket{2}$ state is strictly equal to the $\ket{-}$ state regardless of the magnetic field amplitude.
+
+These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in for the experiment: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 in order to preserve the eigenbasis and $T_2^*$ contribution.
+Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100 G,200 G] seems to satisfy both these criteria: $\Delta \nu\,(100\ \rm G) \approx 30\ \rm MHz \gg \gamma_f$\footnote{Technically, we want to compare $\Delta \nu$ with the dipole-dipole interaction width $2\gamma_f+\Gamma_{NV}+\Gamma_f$ which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+[REF].} and $\abs{\bra{3}\ket{+}}^2\, (200\ \rm G) \approx 0.96$.
+
+Ideally, we would have used an electric field instead of a transverse magnetic field to split the 4 classes energetically, but such an electric field would need to be $> 10^6\ \rm{V/cm}$ which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.
+
+\subsubsection{Results}
 
@@
 The corresponding energy splittings $\Delta \nu=\nu_+-\nu_-$ are written on top.
-We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$ indicated by a red line.
+We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$.
 We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$}
@@
 
-We perform here a $T_1^{\rm dd}$ measurement similar to the one presented in Fig.
-[REF], on the same sample and with the same fitting parameters, but this time probing a class orthogonal to the magnetic field.
-
-Fig.
-\ref{champ tranverse exp}-a) and b) show the evolution of the transition frequencies $\nu_+=\frac{E_{\ket{3}}-E_{\ket{1}}}{h}$ and $\nu_-=\frac{E_{\ket{2}}-E_{\ket{1}}}{h}$ where $\ket{1}$, $\ket{2}$ and $\ket{3}$ are the states described in Fig.
-\ref{eigenstates transverse field}.
-The magnetic field was scanned between 20 and 130 G by increment of $\sim 0.5\ \rm G$.
-
-Fig.
-\ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ with the magnetic field.
-We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$.
-In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ is increased.
-
-\subsubsection{Region A}
-
-We attribute the decrease in region A to the double flips: as the magnetic field increases, so does $\Delta \nu$ and the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$.
-This value is coherent with the previously measured fluctuator linewidth.
-We can then estimate the part of the double flips in the total depolarization rate for the low field regime (20 G here) and find that they amount for $\sim 50\%$ in this case.
-
-Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see in Fig.
-\ref{champ tranverse exp}-a) that they are not very far detuned.
-However we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.
-
-\subsubsection{Region B}
-
-We now turn to region B.
-Even tough the double flips are completely quenched, the effects of the eigenbasis modification and the $T_2^*$ reduction are still present, because we can see on Fig.
-\ref{eigenstates transverse field}-b) that the eigenstates of the spin Hamiltonian are still very close to $\{ \ket{0}, \ket{\pm} \}$ for these magnetic field values.
-
-To evaluate the contribution of the eigenbasis and $T_2^*$, we measure the same class of NV centers, but this time with a longitudinal magnetic field ($B_\parallel \sim 100 G$) to get a baseline value of $1/T_1^{\rm dd}$ in the $\ket{\pm 1}$ basis.
-We found a value of $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ in this case, reported on Fig.
-\ref{champ tranverse exp}-c) as a green line.
-
-We do find that the decay rate is higher in the $\ket{\pm}$ basis by $\sim 50\% $ which corroborates our predictions, as both the change of eigenbasis from $\ket{\pm 1}$ to $\ket{\pm}$ and the increase of $T_2^*$ in the $\ket{\pm}$ basis are predicted to increase the depolarization rate.
-The theoretical increase however overestimates once again the actual increase: for the change of eigenbasis alone we predicted an increase of the decay rate by a factor of 4 (see appendix [REF]), over 8 times more than the actual increase.
-
-We unfortunately cannot experimentally discriminate the depolarization coming from the eigenbasis change or the $T_2^*$ while using the same sample, since both of this properties are tied to the Hamiltonian eigenstates.
-But we do know that in this scenario, for a single isolated class, the contribution of both the change of eigenbasis and $T_2^*$ amounts for $\sim 3$ times less than the double flips.
-
-\subsubsection{Conclusion of the experimental observation}
-
-We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field, and we expect that the depolarization for low magnetic fields behave similarly.
-We therefore claim that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
-\begin{enumerate}
-\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
-\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
-\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
-\end{enumerate}
+Fig.
+\ref{champ tranverse exp}-a) shows an ODMR spectrum for $B=20\ \rm G$ with the two central lines corresponding to the $\ket{0} \to \ket{+}$ and $\ket{0} \to \ket{-}$ transitions of the class orthogonal to the magnetic field.
+We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\mathbf{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
+\ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\mathbf{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}
+
+Fig.
+\ref{champ tranverse exp}-b) shows the evolution of the $\nu_+$ and $\nu_-$ transition frequencies as a function of the magnetic field, measured via ODMR.
+The difference in frequency $\Delta \nu = \nu_+ - \nu_-$ is at first determined by the transverse electric field ($\sim 9\ \rm MHz$) and then grows qudratically with the transverse magnetic field up to a value of $\sim 50\ \rm MHz$.
+
+Fig.
+\ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ on a single class as a purely transverse magnetic field is scanned from 20 to 130 G.
+It also includes the $1/T_1^{\rm dd}$ value found for a purely longitudinal magnetic field which we will consider as a baseline, since the eigenbasis in this case is $\ket{\pm 1}$ and not $\ket{\pm}$.
+We have divided the graph in two regions labeled A and B:
+
+In region A, $1/T_1^{\rm dd}$ decreases with the magnetic field.
+The difference in the depolarizing rate between the $\mathbf{B} \perp [111]$ case and the $\mathbf{B} \parallel [111]$ case comes from the double flips, the eigenbasis change and the $T_2^*$ increase.
+We measure an increase by a factor $\sim 3$ for B=20 G compared to the baseline.
+
+In region B, $1/T_1^{\rm dd}$ reaches a plateau with a value of $\sim 180\ \rm s^{-1}$.
+We attribute this plateau to the fact that the double flips are completely quenched, which is coherent with the fact that $\Delta \nu \geq 30\ \rm MHz$.
+The remaining difference between the $\mathbf{B} \perp [111]$ and $\mathbf{B} \parallel [111]$ cases therefore comes from the eigenbasis change and the $T_2^*$ increase.
+This increase amount for $\sim$ 50\% of the relaxation rate baseline.
+
+We should note that the theoretical increase from the double flips and the change of eigenbasis\footnote{The $T_2^*$ increase should still amount for $\sim 10\%$.} are not the values shown in Fig.
+\ref{summary_theory}: those were computed for a 4-class resonance, while we are looking here at a single class.
+The theoretical increase of the depolarization rate for a single class are detailed in Appendix [REF].
+They are:
+\begin{itemize}
+\item $\Gamma_1^{\ket{\pm}}=4 \Gamma_0$ when taking into account the change of eigenbasis.
+\item $\Gamma_1^{dd}=[REF]\Gamma_1^{\ket{\pm}}=[REF]\Gamma_0$ when taking into account the double flips.
+\end{itemize}
+We have labeled here $\Gamma_0$ the relaxation rate coming from the flip-flop inside of a single class in the $\ket{\pm 1}$ basis (same definition as in table [REF] (table des valeurs chap 3).
+Experimentally, $\Gamma_0$ corresponds to the baseline shown in Fig.
+\ref{champ tranverse exp}-c).
+
+Although we find again that the fluctuator model overestimates the actual relaxation rates, these experimental observations still confirm the qualitative predictions of the model.
+In particular, they confirm that the double-flips are the second leading cause of zero-field depolarization after the 4-classes degeneracy.
+
+\subsection{Conclusion of the experimental observation}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/Shema_summary_exp}
+\caption{Summary of the different experimental investigation.
+A baseline value is indicated with the pink shaded region in all experimental plots.
+The left box indicates the relaxation mechanisms added compared to the baseline for the minimum value of B, and the right box those for the maximum value of B.
+(a) corresponds to Fig.
+\ref{scan 1x1x1x1}-d) where the baseline was measured to be $220\pm 20$ s$^{-1}$, (b) corresponds to Fig.
+\ref{scan 100}-d) with the same baseline, and (c) corresponds to Fig.
+\ref{champ tranverse exp}-c).}
+\label{summary_exp}
+\end{figure}
+
+Fig.
+\ref{summary_exp} summarizes the experimental results of this section.
+We explicitly indicates which of the four depolarization mechanisms are responsible for the observed depolarization rate at the beginning and at the end of each curve.
+
+The experiments confirm the overall predictions of the previous section: the dominant cause of depolarization in zero magnetic field is the degeneracy of the 4 NV classes, followed by the double-flips and then by the eigenbasis change and $T_2^*$ increase.
+
+We will now cover the low-field magnetometry protocol, after a quick overview of the field of NV magnetometry, and see the role that the different depolarization mechanisms play in it.
 
 \section{Introduction to ensemble NV magnetometry}
-
-Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.
-
-NV magnetometry is generally decomposed between single NV and ensemble magnetometry.
-Single NVs,often associated with scanning probes, offer a spatial resolution limited only by how far the magnetic source can be from the NV center (typically $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
+\label{sec 4.4}
+
+There are two classes of NV magnetometer: single NV or ensemble.
+Single NVs offer a spatial resolution limited only by how far the magnetic source can be from the NV center (down to $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
 
@@
 
-We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR.
+We will only focus in this section on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR.
 We will not try to give a complete overview of this field, but only of the elements close to the technique we present.
@@
 The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted into spin population.
-These protocols are generally distinguished between DC-broadband and AC-narrow band.
-
-For DC-broadband magnetometry, the general technique consist of a Ramsey interferometry experiment, presented in chapter 1, which is limited by the spin inhomogeneous coherence time $T_2^* \lesssim 1\ \rm{\mu s}$ for NV ensembles.
+These protocols are generally separated between DC-broadband and AC-narrow band.
+
+For DC-broadband magnetometry, the general technique consists of a Ramsey interferometry experiment, presented in chapter 1, the sensitivity of which is limited by the spin inhomogeneous coherence time $T_2^*$ (which can reach up to $\ \rm{\mu s}$ for NV ensembles).
 The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$.
 Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry.
-The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
-
-For AC magnetometry, the basic protocol consist of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
+The state of the art DC ensemble magnetometr operates with a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
+
+For AC magnetometry, the basic protocol consists of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
 Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field.
@@
 Recently, \citep{xie2021hybrid} have achieved new records in both DC ($\sim 200\ \rm fT/\sqrt{\rm Hz}$) and AC ($\sim 10\ \rm fT/\sqrt{\rm Hz}$) NV magnetometry by using flux concentrators which are ferromagnets concentrating the magnetic field in a small volume, similarly to an electromagnet.
-Because the gain from the flux concentrators is not tied to a specific magnetometry protocol, we will focus on the results obtained without flux concentrators.
+The technology of flux concentrator could be applied to any NV magnetometry protocol.
+We will here only compare experimental results obtained without a flux concentrator.
+
 \subsection{Low field magnetometry}
-Magnetometry for low magnetic field ($\gamma B < d_\perp E_\perp$) is challenging since in this regime the NV eigenstates and eigenenergies do not depend on the magnetic field in the first order.
-This complication is often lifted  by simply using a bias magnetic field, typically from a permanent magnet.
+Magnetometry at low magnetic fields ($\gamma B < d_\perp E_\perp$) is challenging since in this regime the NV eigenstates and eigenenergies do not depend on the magnetic field to the first order.
+This complication is often lifted  by simply using a bias magnetic field.
 
 Some systems however need to be kept under low to zero external magnetic fields, for example to study the J-coupling between spins \citep{sutter2012computational}, or the phase of skyrmions \citep{zazvorka2020skyrmion}.
-To study these systems with NV centers, we need to find new protocols for low field NV magnetometry.
-
-The main technique for low field magnetometry consist in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}.
+To study these systems with NV centers, new protocols for low field NV magnetometry are required.
+
+The main technique for low field magnetometry consists in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}.
+These states are linearly dependent on the magnetic field and can therefore be used to perform magnetometry.
 The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$ \citep{zheng2019zero}.
 
-Another technique consist in using the $^{13}\rm{C}-$NV complex we discussed in sec.
-[REF] which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are frist order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
+Another technique consists in using the $^{13}\rm{C}-$NV complex we discussed in sec.
+[REF] which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are first order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
 With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers.
-This technique is more suited for a carefully chosen single NV center.
-
-\subsection{Microwave-free magnetometry}
-All the precedent magnetemetry protocols described relied on the use of a controllable magnetic field, to coherently manipulate the NV spin state and to measure the phase imprinted by the magnetic field in the rotating frame.
-
-There are other protocols which correlate the magnetic field with the spin $T_1$ time (relaxometry).
+
+\subsection{Microwave-free DC magnetometry}
+All the previously described magnetemetry protocols relied on the use of a controllable microwave field to coherently manipulate the NV spin state.
+
+%There are other protocols which correlate the magnetic field to the spin $T_1$ time (relaxometry).
 These protocols can operate without microwave field and work for either DC or high frequency AC sensing.
 
-\subsubsection{microwave-less DC magnetometry}
-The DC microwave-free protocols rely on sharp feature in the PL for specific magnetic fields around certain cross-relaxations processes or level anti-crossing.
+%\subsubsection{microwave-less DC magnetometry}
+The DC microwave-free protocols rely on sharp features in the PL for specific magnetic fields values.
 Fig.
@@
 
-A protocol based on the GSLAC line has been implemented, first only for longitudinal magnetic field measurement \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.
+A protocol based on the GSLAC line has been implemented, first only for longitudinal magnetic field measurements \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.
 
@@
 
-\subsubsection{microwave-less AC magnetometry}
-The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed.
+%\subsubsection{microwave-less AC magnetometry}
+%The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed.
 The detection bandwidth is adjusted with an external DC magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.
-
-Sensitivities as low as a few $\rm pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
+%
+%Sensitivities as low as a few $\rm pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
 
 \subsection{Orientation-free magnetometry}
-Every protocol described so far requires a precise knowledge  of the Larmor frequency between the $\ket{0}$ and $\ket{\pm 1}$ states as a function of the magnetic field.
+All protocols described so far require precise knowledge of the Larmor frequency between the $\ket{0}$ and $\ket{\pm 1}$ states for a given magnetic field.
 This knowledge implies that the direction of each NV axis in the frame of the magnetic field is known, because of the strong anisotropy in the NV Hamiltonian.
@@
 To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
-This protocol however has an very low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective}.
+This protocol has an very low sensitivity and found limited use so far\citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective}.
 
@@
 \label{sec LFDM}
-We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the spontaneous depolarization of dense NV ensemble at low magnetic field.
-This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional field is required to improve performances.
+We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the depolarization of dense NV ensemble at low magnetic field.
+This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional AC field is required to improve performances.
 
@@
 Fig.
-\ref{setup magneto}-b) and c) show respectively the PL measured directly on the photodiode and the lock-in signal when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G.
+\ref{setup magneto}-b) and c) show the PL measured on the photodiode and on the lock-in respectively, when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G.
 For this configuration, the magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.
@@
 
-We characterize the DC sensitivity by looking at the noise floor on the LIA signal when sitting on the steepest slope, around $B_0=0$.
-The following parameters were used in the case presented: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
+We characterize the DC sensitivity by looking at the noise floor on the LIA signal when sitting on the steepest sloped ($B_0\approx 0$).
+The following parameters were used: $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$.
+The lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
 
@@
 
-We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$).
-Laser fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.
+We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\sqrt{\langle \delta V^2 \rangle} \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\sqrt{\langle \delta V^2 \rangle} < 1\ \rm mV$).
+The remaining noise could also come from laser intensity fluctuations.
 
@@
 
-We will now look at the dependence of the sensitivity with respect to the magnetic field angle.
-In particular, we are interested in comparing the sensitivity when $\mathbf{B}\parallel [100]$, where there is no lift of degeneracy between the four classes as $\mathbf{B}$ increases; and a case where $\mathbf{B}\nparallel [100]$ and where there is a lift of degeneracy between the four classes.
+We will now study the dependence of the sensitivity with respect to the magnetic field angle.
+In particular, we will compare the sensitivity for two scenarios: the case $\mathbf{B}\parallel [100]$and the four classes remains degenerate when $\mathbf{B}$ increases, and the case $\mathbf{B}\nparallel [100]$ where there is a lift of degeneracy between the four classes.
 We previously saw in Fig.
@@
 Fig.
-\ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angle of $B_0$.
-For the blue curve, the field was scanned along the [100] axis, for the orange curve, it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis.
+\ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angles of $B_0$.
+For the blue curve the field was scanned along the [100] axis, while for the orange curve it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis.
 We can see that the maximum amplitude of the lock-in signal (relevant for AC sensing) is of the same order of magnitude for both case, and the slope around $B_0$ (relevant for DC sensing) is almost identical.
@@
 
-The fact that the sensitivity is similar when $\mathbf{B}\parallel [100]$ or $\mathbf{B}\nparallel [100]$ means that the inter-class co-resonances plays a lower role in the sensitivity than the zero-field specific depolarization mechanisms, previously detailed in Sec.
+The fact that the sensitivity is similar when $\mathbf{B}\parallel [100]$ or $\mathbf{B}\nparallel [100]$ means that the degeneracy between the four classes plays a lower role in the sensitivity than the three other mechanisms detailed in sec.
 \ref{sec causes zero field}.
-We cannot be sure which of the three zero-field mechanism is the dominant one, but based on the experimental results shown in Fig.
+We cannot measure the individual impact of each of these three mechanisms, but based on the experimental results shown in Fig.
 \ref{champ tranverse exp}, we can assume that the double-flips are the dominant factor in the LFDM sensitivity.
@@
 \section{Comparison between LFDM and other magnetometry protocol}
+\label{sec 4.6}
 Now that we have characterized the LFDM protocol, we can compare its performances with the previously mentioned magnetometry protocols.
-We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity.
+The best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity.
 This discrepancy however come in large part to the smaller sensor volume used in the characterization, and to a relatively low PL collection efficiency compared to the other experiments.
@@
 
-We will first compare the LFDM protocol with the best microwave-based NV magnetometry protocols.
 A better metric than the sensitivity to compare different protocols is the volume-normalized sensitivity defined in \citep{zhou2020quantum} as $\eta_v=\eta\cdot \sqrt{V}$ where $\eta$ is the sensitivity and $V$ the effective volume of the sensor.
@@
 
-I will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.
+We will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.
 
 \subsection{LFDM optimization}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{Figures/double_adamas}
+\caption{a) PL of two distinct samples ADM-15-1 and ADM-15-4 as the magnetic field is scanned in an arbitrary direction (which splits the 4 classes similarly in both cases).
+b) Same measurement with $B\parallel [100]$ c) $T_1$ measurement on both samples on a single class for $B\approx 50\ \rm G$.
+d) $T_1$ measurement on both samples on all four classes for $B=0$.}
+\label{double dragon}
+\end{figure}
 
@@
 
-Most magnetometry protocols have a relatively clear material optimization path: DC protocols want to improve the figure of merit $\rho T_2^*$ and AC protocols want to improve $\rho T_2$, where $\rho=[\rm NV]$ is the NV concentration.
-The reason is that the number of NV centers for a given volume scales as $\rho$, and therefore the sensitivity scale as $\sqrt{\rho}$.
+Most magnetometry protocols have a relatively clear material optimization path: DC protocols use the figure of merit $\rho T_2^*$ and AC protocols $\rho T_2$, where $\rho=[\rm NV]$ is the NV concentration.
+The reason is that the number of NV centers for a given volume scales as $\rho$, and therefore the sensitivity scale as $1/\sqrt{\rho}$.
 Meanwhile increasing the density of defects tend to decrease the spin coherence times $T_2$ and $T_2^*$, and the AC and DC sensitivities scale respectively as $1/\sqrt{T_2}$ and $1/\sqrt{T_2^*}$.
+The sensitivity in both of these cases is then inversely proportional to the figure of merit.
 
 For the LFDM protocol however, the material optimization is far from obvious.
-Indeed, while increasing $T_2^*$ would improve the sensitivity, the gain might not be as interesting as for other protocols due to the broadening caused by the fluctuator lifetime.
+Indeed, while increasing $T_2^*$ would improve the sensitivity, the gain might not be as interesting as for traditional protocols due to the broadening caused by $\gamma_f$.
 On the other hand, increasing the NV density not only increases the number of sensor per volume, but also increases the dipole-dipole strength between NV centers.
@@
 
-It seems then that increasing the NV density would be the obvious path to improve the LFDM performances.
-In practice however, I observed that the relationship between NV density, fluctuators, and LFDM performances was generally not straightforward.
-The two following examples will illustrate that.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/substrat}
-\caption{Measurements on sample SBST-C for two different regions A and B.
+Increasing the NV density seem to be the path for further material optimization, but other issues may arise in the process.
+
+We should note that we observed large discrepancy in the zero field depolarization among the samples that we studied, even from samples coming from the same batch.
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=\textwidth]{Figures/substrat}
+%\caption{Measurements on sample SBST-C for two different regions A and B.
 a) PL contrast as a function of a magnetic field scanned in an aribitrary direction (same direction for both regions).
@@
 d) $T_1$ measurement on all  four classes when $B=0$ }
-\label{substrat chelou}
-\end{figure}
-
-Fig.
+%\label{substrat chelou}
+%\end{figure}
+%
+%Fig.
 \ref{substrat chelou} shows various measurements on the same sample SBST-C, a HPHT bulk sample containing [NV]$\sim  \rm ppm$, on two spatially distinct regions A and B.
@@
 Both region showed a similar PL which would indicate a similar density of NV centers.
-
-The metric we are interested in to evaluate the LFDM performance are the PL contrast when $\mathbf{B}$ is scanned in an arbitrary direction, which mainly represents the inter-class flip flop and which I will refer to as FF-contrast; and the PL contrast when B is scanned $\mathbf{B}$ is scanned along the [100] direction, which mainly represents the double flips and which I will refer to as DF-contrast.
+%
+%The metric we are interested in to evaluate the LFDM performance are the PL contrast when $\mathbf{B}$ is scanned in an arbitrary direction, which mainly represents the inter-class flip flop and which I will refer to as FF-contrast; and the PL contrast when B is scanned $\mathbf{B}$ is scanned along the [100] direction, which mainly represents the double flips and which I will refer to as DF-contrast.
 We are interested in the DF-contrast since we have seen in sec.
 \ref{sec angular sensi} that it could be the main contribution to the LFDM sensitivity.
-
-Fig.
+%
+%Fig.
 \ref{substrat chelou}-a) and b) indicate that region B shows a higher FF-contrast, but a smaller DF-contrast than region A.
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{Figures/double_adamas}
-\caption{Same measurements as Fig.
-\ref{substrat chelou} on two distinct samples ADM-15-4 and ADM-15-1.
-Every parameters (laser power,optical setup , AC magnetic field) was kept identical for the two samples.}
-\label{double dragon}
-\end{figure}
-
-We now turn to Fig.
+Fig.
+\ref{double dragon}-a) shows the PL on two distinct 15-$\mu$m Adamas microdiamonds for a magnetic field scanned in a random direction, and Fig.
+\ref{double dragon}-b) for a magnetic field scanned along the [100].
+We can notice that the PL contrast in zero field is twice as big for sample ADM-15-4 when $B \nparallel [100]$, and 5 time as big when $B \parallel [100]$, which would directly impact their respective sensitivity in LFDM.
+
+Fig.
+\ref{double dragon}-c) and d) shows $T_1$ measurement on both samples either on a single class for $B\neq 0$ or for all four classes for $B=0$.
+Interestingly, we can notice that the $T_1$ of sample ADM-15-1 is both shorter than the one of ADM-15-4 for $B\neq 0$, and longer for $B=0$.
+This could indicate that another relaxation mechanism is competing with the NV-fluctuator one on sample ADM-15-1, which would limit the LFDM sensitivity.
+
+%We now turn to Fig.
 \ref{double dragon}.
@@
 Both samples showed relatively similar PL which could again indicate a similar NV density.
-
-We can first notice the disparity in the  zero field PL contrasts: the FF-contrast is twice as big in sample ADM-15-4 as in sample ADM-15-1, and the DF-contrast is almost 5 times bigger.
+%
+%We can first notice the disparity in the  zero field PL contrasts: the FF-contrast is twice as big in sample ADM-15-4 as in sample ADM-15-1, and the DF-contrast is almost 5 times bigger.
 When looking at the $T_1$, we can notice that the lifetime for non-zero field is slightly longer for sample ADM-15-4 than for ADM-15-1, but its lifetime in zero-field is shorter (it is also exponential and not stretched-exponential as discussed in the last chapter [REF], which could have an influence on the PL contrast).
 From the $T_1$ measurements we can not determine which sample contains the most fluctuators, and therefore correlate the fluctuator or NV density with the DF and FF-contrasts.
-
-These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field.
+%
+%These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field.
 We have noticed a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive.
@@
 This includes the temperature of the sample, the intensity and wavelength of the laser, the frequency and amplitude of the AC magnetic field.
-For the results presented in sec.
-\ref{sec LFDM}, we chose the parameters which we found optimal within our available experimental range.
 

### 2022-10-17 Manuscrit Thèse/chapter 4/Chapter 4.tex
tutu

@@
 In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$.
-These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely quenched.
+These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely suppressed.
 
@@
 
-These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in for the experiment: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 in order to preserve the eigenbasis and $T_2^*$ contribution.
-Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100 G,200 G] seems to satisfy both these criteria: $\Delta \nu\,(100\ \rm G) \approx 30\ \rm MHz \gg \gamma_f$\footnote{Technically, we want to compare $\Delta \nu$ with the dipole-dipole interaction width $2\gamma_f+\Gamma_{NV}+\Gamma_f$ which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are the relevant parameters of the experiment: the goal is to increase $\Delta \nu$ to the point where the double flips are completely suppressed, while $\abs{\bra{3}\ket{+}}^2$ remains close to 1 in order to preserve the eigenbasis and $T_2^*$ contribution.
+Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100 G,200 G] seems to satisfy both these criteria: $\Delta \nu\,(100\ \rm G) \approx 30\ \rm MHz \gg \gamma_f$\footnote{Technically, $\Delta \nu$ is to be compared with the dipole-dipole interaction width $2\gamma_f+\Gamma_{NV}+\Gamma_f$ which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
 [REF].} and $\abs{\bra{3}\ket{+}}^2\, (200\ \rm G) \approx 0.96$.

### 2022-10-17 Talks/Talk Stuttgart/presentation.tex
tutu

@@
 
-\title{Dipolar interactions in dense ensembles of Nitrogen-Vacancy centers}
-\author{Clément Pellet-Mary, Maxime Perdriat, Gabriel Hétet}
+\title{Low field depolarization of dense NV centers ensemble: application to magnetometry}
+\author{Clément Pellet-Mary, Maxime Perdriat, Paul Huillery,\\ Gabriel Hétet}
 \date{Nano-optics group}
@@
 \section{Cross-relaxation with NV centers}
+\begin{frame}{Outline}
+\tableofcontents[currentsection]
+\end{frame}
 \begin{frame}{Principle of cross-relaxation with NV centers}
@@
 \section{The NV-fluctuator model and experimental verification}
-
+\begin{frame}{Outline}
+\tableofcontents[currentsection]
+\end{frame}
 \begin{frame}{Presentation of the fluctuator model}
@@
 \section{Dipole-dipole interaction under low magnetic field}
+\begin{frame}{Outline}
+\tableofcontents[currentsection]
+\end{frame}
+\begin{frame}{Zero field depolarization sources (theory)}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_0B_theorie}
+\end{frame}
 
-\section{Charcterization of the low field depolarization magnetometry protocol}
+\begin{frame}{Experiment: $\vec B$ in arbitrary direction}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_PL_100}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \perp [111]$}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_champs_transverse}
+\end{frame}
+
+\begin{frame}{Summary of the experimental observations}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{shema_summary_exp}
+\end{frame}
+
+\section{Low field depolarization magnetometry (LFDM)}
+\begin{frame}{Outline}
+\tableofcontents[currentsection]
+\end{frame}
+\begin{frame}{Principle of LFDM}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_principle_LFDM}
+\end{frame}
+
+\begin{frame}{Sensitivity of LFDM}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_sensi_LFDM}
+\end{frame}
+
+\begin{frame}{Comparison with the state of the art}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_litterature}
+\end{frame}
+
+\begin{frame}{Comparison with CW ODMR}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_microwave}
+\end{frame}
+
+\begin{frame}{Angular sensitivity of LFDM}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_angular_sensitivity}
+\end{frame}
+
+\begin{frame}{Application: wide-field magnetometry on irregular surfaces}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_applications_wide_field}
+\end{frame}
+
+\begin{frame}{Other applications}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_application_autre}
+\end{frame}
+
+\begin{frame}{Take home messages}
+\begin{itemize}
+\item The spin depolarization in dense NV ensemble is dominated by dipole-dipole interaction with fast decaying centers (fluctuators).
+\medskip
+\item The leading cause of depolarization in zero field is the spectral overlap of the NV classes, followed by the double flips.
+\medskip
+\item Low field depolarization magnetometry achieves a sensitivity $\sim 6\ \mu\rm T \cdot \mu \rm m^{3/2} / Hz^{1/2}$ with commercial samples.
+\medskip
+\item LFDM can be used with powder or polycristalline samples.
+\end{itemize}
+
+\end{frame}
 

### 2022-10-19 Manuscrit Thèse/Intro et conclusion/Summarys.tex
toto

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric-comp,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_intro.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{}
\begin{document}

\chapter*{English Summary}
The NV center in diamond is a promising candidate for emerging quantum technologies, including quantum communication, quantum computing and quantum metrology. Thanks to the recent progress in diamond synthesis, samples with high NV center concentration can now be created. This increase in the NV density comes with an increase of the interaction between them, which leads to a rich and complex physics. Numerous applications, such as NMR hyperpolarization, bio-imagery or spin mechanics rely on the use of dense NV ensembles. Understanding and mastering the properties of ensembles of NV centers will be a crucial point in the development of these technologies.

This doctoral thesis focuses on cross-relaxation between dense ensembles of NV centers, as well as on the interaction between NV centers and other diamond paramagnetic impurities. The aim of this work is to better understand the properties of dense NV centers ensembles, and to exploit them for new applications such as the detection of impurities and magnetometry.

In the first part, we will detail the properties of single NV centers and introduce notions regarding dipole-dipole spin coupling and cross-relaxations. We will then show how cross-relaxations can be used to optically detect dark spin resonances, in this case the VH- and War1 defects in CVD-grown diamond. Finally, we will present a study of cross-relaxations between subgroups of NV centers. We will first cover the influence of cross-relaxation on the spin dynamics and discuss its microscopic origin. We will then investigate the role of the magnetic field on cross-relaxations as well as their potential use in low field magnetometry.

\chapter*{Résumé français}

Le centre NV du diamant est un candidat prometteur pour de nombreuses technologies quantiques, que ce soit pour les communications quantiques, le calcul quantique ou la métrologie quantique. Grace aux progrès de synthétisation du diamant, des échantillons contenant une haute concentration en centres NV sont aujourd’hui réalisables. Cette augmentation de la densité de centres NV s’accompagne d’une augmentation des interactions entre ces derniers, ce qui donne lieu à une physique complexe et encore largement inexplorée. De nombreuses applications récentes, telles que l’hyper-polarisation en RMN, la bio-imagerie ou encore la spin-mécanique repose sur des ensembles denses de centres NV. Comprendre et maîtriser les propriétés d’ensemble de ces derniers est un point crucial pour l’établissement de ces technologies.

Cette thèse étudie les échanges de polarisation, par relaxation croisée, entre des ensembles denses de centres NV, ainsi que les interactions entre les centres NV et d’autres impuretés paramagnétiques présentes dans le diamant. Le but de ces travaux est de mieux comprendre les propriétés des ensembles de centres NV, et d’exploiter certaines de ces propriétés pour de nouvelles applications telles que la détection d’impuretés ou la magnétométrie.

Dans un premier temps, nous détaillerons les propriétés des centres NV et nous introduirons des concepts liés à l’interaction dipolaire entre spins et aux relaxations croisées. Nous verrons ensuite comment les relaxations croisées peuvent être utilisées pour détecter optiquement des résonances de spins qui ne sont pas optiquement actifs, en l’occurrence les défauts VH- et War1 dans des diamants CVD. Enfin nous étudierons les relaxations croisées entre différents sous-groupes de centre NV. Nous verrons d’abord leur influence sur la dynamique du spin et leur origine microscopique, puis nous étudierons leur dépendance avec le champ magnétique et leur potentielle utilisation pour la magnétométrie en champs faible.

\end{document}

### 2022-10-19 Manuscrit Thèse/Intro et conclusion/conclusion.tex
toto

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric-comp,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_intro.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{}
\begin{document}
\chapter*{Conclusion}

The work presented in this manuscript is part of an ongoing effort to develop quantum technologies, and NV centers, among many other systems, are a part of this new ecosystem. While the properties of single NV centers have been thoroughly studied in the past 20 years, we showed in this thesis that emerging properties of NV ensemble can still be discovered. These ensemble properties are relevant for the many fields using NV ensembles \citep{ barry2020sensitivity, eills2022spin, fu2007characterization, perdriat2021spin}. They also are relevant to explore fundamental aspects \citep{choi2017observation, kucsko2018critical, angerer2018superradiant} and for new potential applications \citep{akhmedzhanov2017microwave, pellet2021optical, pellet2022spin}.

At the heart of this manuscript is the notion of cross-relaxation with NV centers. Cross-relaxations are a powerful spectroscopic tool, and NV centers are naturally suited to cross-relaxation studies due to their ability to be optically polarized and read-out, in opposition to their environment.

We have showed how cross-relaxations can be used in a CVD-grown diamond to detect trace amount of other spin impurities, with higher sensitivity than the standard EPR apparatus, and only a fraction of the cost. These observations allowed us to study the temperature dependence of these impurities, and could eventually allow the detection and coherent manipulation of single dark spins through their interaction with an NV bath.

We have then studied the cross-relaxation between NV centers, thanks to the multiple classes of NV and the ability to tune them in and out of resonance. The observation of cross-relaxation between seemingly equivalent spins has led previous researchers to postulate the existence of dark NV centers called fluctuators. We investigated experimentally the predictions of the fluctuator model, in particular the prediction of a stretched exponential profile for the depolarization of the spins, and a broadening of the fluctuators spectra compared to those of regular NV centers. Our observations for the large part comfort the fluctuator model, although we did notice some deviations which could indicate some limitations of the model.

Dipole-dipole interaction within dense ensemble of NV centers ([NV] > ppm) can be responsible for an increase of the depolarization rate by more than a factor of 10. This changes considerably the properties of NV ensemble compared to single NV. Having a better understanding of the fluctuators and of their microscopic origin will be a crucial point in the future development of NV ensemble technologies. The further steps to determine the nature of the fluctuators could include the study of dedicated samples grown with specific concentrations of NV and other defects, as well as the study of nano-diamonds to try to isolate a single fluctuator.

Finally, we investigated the depolarization of NV ensemble at low magnetic field. We identified 4 different contributions to the zero field depolarization which we analyzed both theoretically through the fluctuator model, and experimentally by varying the magnetic field orientation. We concluded that the dominant factor in the zero field depolarization was the degeneracy of the 4 NV classes, followed by the double flips.

The depolarization of NV ensemble in zero field can in itself be used to measure magnetic field. We characterized such a magnetometer and compared it to the state of the art NV magnetometers. We found in particular that the dominant factor in the sensitivity of the magnetometer was the presence of the double flips, and not the classes degeneracy.

The low-field depolarization magnetometer is unique among the NV magnetometers in that is does not require a microwave or a precise orientation of the magnetic field to work, which makes it usable with powdered or polycrystalline samples. Applications of magnetometry with powders include wide-field imaging of uneven surfaces, background free fluorescence microscopy, and microwave-free nano-probe sensing.

\printbibliography

\end{document}

### 2022-10-19 Manuscrit Thèse/chapter 1/Chapter 1.tex
toto

@@
 \begin{document}
-\chapter{Introduction to Nitrogen-Vacancy centers and their common usage}
+\chapter{Spin properties of single and ensemble NV centers}
 The aim of this first chapter is to introduce concepts related to Nitrogen-vacancy (NV) centers, to the diamond material, and to the dipole-dipole interaction which are necessary to explain the work presented in the later chapters.

### 2022-10-19 Manuscrit Thèse/chapter 1/appendix_samples.tex
toto

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric-comp,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_ch1.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{}
\begin{document}
\chapter{Diamond samples}
This appendix document the known properties of every diamond samples used in this manuscript.

\section{CVD samples}
\begin{table}[htbp]
\centering
\caption{Concentration in nitrogen and NV centers for the CVD samples}
\label{Table samples CVD}
\begin{tabular}{c|cc}
\toprule
Name &  [NV$^-$]  & [N] \\
\midrule
CVD-pink & 4.6 ppm & 26 ppm \\
CVD-B & 0.43 ppm & 10-30 ppm \\
CVD-C & 0.43 ppm & 10-30 ppm \\

\bottomrule
\end{tabular}
\end{table}
Table \ref{Table samples CVD} summarizes the most relevant parameters in our study which are the concentration in (negatively charged) NV centers and the total concentration of nitrogen in the sample.
\subsection{CVD-pink}
The growth process and characterization of sample CVD pink are detailed in \citep{tallaire2020high}. The sample was grown by CVD in a gas environment containing 500 ppm of N$_2$O. The sample was then irradiated by a 10 MeV electron beam at a fluence of $2\cdot 10^{18}\ \rm cm^{-2}$ for 20 hours, while heated to 900 $^\circ$C during the irradiation process.

The concentration of NV centers and nitrogen impurities was determined via UV-visible absorption and FTIR.

\subsection{CVD-B and CVD-C}
The growth process and characterization of sample CVD pink are detailed in \citep{ngambou2022improving}. These two samples were grown with a N$_2$O concentration of 100 ppm in the gas phase, and were irradiated similarly to sample CVD-pink. The main difference between the two sample is the further annealing step on sample CVD-B, as detailed in chapter 2.

\section{HPHT samples}
\subsection{Adamas fluorescent microdimaonds}
\begin{table}[htbp]
\centering
\caption{Concentration in nitrogen and NV centers for the HPHT samples}
\label{Table samples adamas}
\begin{tabular}{c|cc}
\toprule
Name &  [NV$^-$] & [N] \\
\midrule
Adamas & 2.5-3.5 ppm & $\sim$100 ppm \\

\bottomrule
\end{tabular}
\end{table}

The main HPHT sample studied in this manuscript are the commercially available red fluorescent agents from Adamas nano. Samples labeled ADM-15-$n$ correspond to 15 $\mu$m micro-diamonds, and samples labeled ADM-150-$n$ to 150 $\mu$m micro-diamonds. The Adamas website indicates that the fluorescent diamonds contain $\sim 100$ ppm of nitrogen impurities, and that the micro-diamonds are irradiated with 2-3 MeV electron beams and annealed, without further information. The [NV] concentration, also indicated on the website, was measured via EPR.

\subsection{Other HPHT samples}
\subsubsection{Sumi-2}
Sample Sumi-2 is a type Ib HPHT sample bought from Sumitomo and irradiated with a 2 MeV electron beam and a fluence of $7\cdot 10^{18}\ \rm cm^{-2}$ which were further annealed at 800 $^\circ$C for two hours. The NV and N concentration were not measured on this sample.
\subsubsection{SBST-B and SBST-C}
Samples SBST-B and SBST-C are the type Ib HPHT substrates on which the samples CVD-B and CVD-C grew. They were irradiated and annealed along the CVD layers. The NV and N concentrations in these samples has not been measured.
\printbibliography
\end{document}

### 2022-10-19 Manuscrit Thèse/chapter 3/Appendix_eta.tex
toto

@@
 \begin{document}
-\chapter{Appendix: Computation of $\bar \eta$}
+\chapter{Computation of $\bar \eta$}
 We will discuss here the specifics to the computation of the  $\bar \eta$ factor for different geometric configurations, and how to compute the $T_1$ times in table [REF] of chapter [REF], and table [REF] of chapter [REF].
@@
 To estimate the total PL, one would need to compute the relaxation rate of all 4 classes independently, for instance in a $\mathbf{B} \in \{110\}$ scenario, two classes have a relaxation rate $\Gamma(\mathbf{B} \in \{110\})$ and two classes have a relaxation rate $\Gamma_0$.
+
+\section{$\bar \eta$ in the non-magnetic basis $\{ \ket{0}, \ket{+}, \ket{-}\}$}
+
+In this part we will study the averaged flip-flop and double flip rates in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis which corresponds to the various mechanisms discribed in chapter 4.
+
+We will define here the $\ket{\pm}$ states as:
+\begin{equation}
+\ket{\pm}=\frac{\ket{+1}\pm\ket{-1}}{\sqrt{2}}.
+\end{equation}
+
+In this basis, the spin operators become:
+\begin{align*}
+S_x&=\begin{pmatrix}
+0&0&0 \\
+0&0&1 \\
+0&1&0
+\end{pmatrix} \\
+S_y&=\begin{pmatrix}
+0&-i&0 \\
+i&0&0 \\
+0&0&0
+\end{pmatrix} \\
+S_z&=\begin{pmatrix}
+0&0&1 \\
+0&0&0 \\
+1&0&0
+\end{pmatrix}.
+\end{align*}
+
+Importantly, we have to note the breaking of the symmetry in the (xy) plane.
+This comes from the transverse electric or magnetic field dominating the single spin Hamiltonian.
+By convention, we define the $x$ direction as the direction of the dominant transverse electric or magnetic field.
+
+\subsection{Flip-flops in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis}
+\label{sec flip-flop +/-}
+Rewriting the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-}\} \times \{ \ket{0}, \ket{+}, \ket{-}\}$ changes the value of the flip-flop and double flip prefactors.
+For instance, the prefactor of the $\ket{+,0}\bra{0,+}$ term is: \begin{equation}
+-\frac{J_0}{r^3}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)\right]
+\end{equation}
+while the prefactor of the $\ket{+1,0}\bra{0,+1}$ term was: \begin{equation}
+-\frac{J_0}{r^3}\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f) + 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{y}_s \cdot \hat{y}_f) \right] 
+\end{equation}
+
+Because of the symmetry breaking of the $(xy)$ plane, we now have to consider the relative orientation of $x_1$ and $x_2$ on top of the previously mentioned angles $\theta$, $\phi$ and $\psi$.
+We define $\alpha_1$ and $\alpha_2$ as the relative angle formed by the $x_1$ and $x_2$ axes with an arbitrary axis in their respective $(xy)$ plane (the arbitrary axis can be defined for instance in the lab frame of reference).
+
+Because we assume that the physical orientation of the $x_1$ and $x_2$ axes in zero external magnetic field is determined by a random local electric field, we will sample than angles $\alpha_1$ and $\alpha_2$ on a uniform [0,$2\pi$] distribution independently for each spin.
+
+The integral to compute the $\bar G$ factor of the dipole-dipole interaction in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis then becomes\footnote{The flip-flop term considered here is $\ket{+,0}\bra{0,+}$.
+The $\ket{-,0}\bra{0,-}$ term gives the same numerical values once averaged.}
+\begin{equation}
+\bar G= \frac{1}{\sqrt{3}} \iiiint \abs{3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi} \frac{d \alpha_1}{2\pi} \frac{d \alpha_2}{2\pi}.
+\end{equation}
+
+The numerical values obtained are reported in table \ref{table G flip-flop non mag}.
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar G$ in the $\ket{\pm}$ basis}
+ \label{table G flip-flop non mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+0.410 & 0.394 & 0.394 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Finally the value $\Gamma_1^{\rm th}= 51.4 \Gamma_0^{\rm th}$ shown in Fig. [REF] was computed from:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar G_{\ket{\pm}}(\psi=0) + 3 \bar G_{\ket{\pm}}(\psi\neq 0) }{\bar G_{\ket{\pm 1}}(\psi=0)}\right)^2
+\end{equation}
+\subsection{Double-flips in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis}
+Including the double flips, there are two possible relaxation path for each transition of the regular NV center.
+For instance the transition $\ket{0} \to \ket{-}$ can involve either the flip flop $\ket{0,-} \bra{-,0}$ or the double-flip $\ket{0,+} \bra{-,0}$.
+
+The inclusion of the double flips require to modify the  expression of $\eta^2$ to include both relaxation path.
+According to the more general definition given in \citep{choi2017depolarization}, the $\eta^2$ term corresponding to the $\ket{0} \to \ket{-}$ transition of the NV center can be written\footnote{The $\ket{\pm}$ states in the formula can be replaced by $\ket{\pm 1}$ depending on the eigenbasis of the single spin Hamiltonian.}:
+\begin{align}
+\label{eq eta2 avec DQ}
+\eta^2&=\frac{1}{3} \left(\frac{\mel{0,-}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{\left(\Delta_{\omega_f}^{(0,-)} - \Delta_{\omega_{NV}}^{(0,-)}\right)^2+4\gamma_f^2} \nonumber\\
+&+\frac{1}{3} \left(\frac{\mel{0,+}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{\left(\Delta_{\omega_f}^{(0,+)} - \Delta_{\omega_{NV}}^{(0,-)}\right)^2+4\gamma_f^2},
+\end{align}
+where $\Delta_{\omega_{NV}}^{(0,\pm)}$ corresponds to the central frequency of the $\ket{0} \to \ket{\pm}$ transition of the NV center, and $\Delta_{\omega_{f}}^{(0,\pm)}$ that of the fluctuator.
+
+Eq.
+(\ref{eq eta2 avec DQ}) can no longer be factorized as a product of a spectral and spatial function.
+To compute the value of $\bar eta$ in zero magnetic field, we will proceed to the following assumptions: $\Delta_{\omega_{NV}}^{(0,\pm)}=\Delta_{\omega_{f}}^{(0,\pm)}$, and $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$\footnote{This second assumption is justified by experimental measurements.}.
+Eq.
+(\ref{eq eta2 avec DQ}) then become:
+\begin{equation}
+\eta^2=\frac{1}{3} \left(\frac{\mel{0,-}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 
++\frac{1}{3} \left(\frac{\mel{0,+}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{1}{2}.
+\end{equation}
+
+Following the same procedure as in the last section, we then compute $\bar \eta$ in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis and report the values in table \ref{table eta double flip non mag}.
+These values are used to compute the double-flip term in Fig.
+[REF]
+
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar \eta$ in the $\ket{\pm}$ basis including the double flips.}
+ \label{table eta double flip non mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+0.554 & 0.532 & 0.532 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Flip-flops and double flips in the $B\perp [111]$ regime}
+
+We will focus here on the case detailed in Fig.
+[REF]: a single class of NV centers whose Hamiltonian is dominated by the transverse magnetic field.
+Unlike the case presented in Sec.
+\ref{sec flip-flop +/-}, the $x$ direction is fixed for all spins and imposed by the direction of the magnetic field.
+
+The expression of $\bar G$ then becomes:
+\begin{equation}
+\bar{G}= \frac{1}{\sqrt{3}}\iint \abs{3(\hat{u}\cdot\hat{x})^2 - 1} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi}= \frac{4}{9}\approx 0.444.
+\end{equation}
+
+The increase in the decay rate of the eigenbasis when considering only the flip-flop is:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar G_{\ket{\pm}}(\psi=0)}{\bar G_{\ket{\pm 1}}(\psi=0)}\right)^2=4.
+\end{equation}
+
+To include the double-flip, we once again make the assumption that $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$ and find $\bar \eta=0.567$.
+The increase in the decay rate when including the decay rate is:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar \eta_{\ket{\pm}}(\psi=0)}{\bar \eta_{\ket{\pm 1}}(\psi=0)}\right)^2\approx 6.5.
+\end{equation}
 \printbibliography

### 2022-10-19 Manuscrit Thèse/sample.tex
toto

--- 2022-10-19 Manuscrit Thèse/Intro et conclusion/Summarys.tex
@@
-\documentclass[a4paper,11pt]{report}
+\documentclass[a4paper, 11pt]{book}
+
+\usepackage{psl-cover}
 \usepackage[]{amsmath}
@@
 ]{biblatex} %Imports biblatex package
-\addbibresource{Bib_intro.bib} %Import the bibliography file
+\addbibresource{bib_thesis.bib} %Import the bibliography file
 
@@
 
-\title{}
-\begin{document}
+\title{Cross-relaxation in dense ensembles of NV centers and application to magnetometry}
 
-\chapter*{English Summary}
+\author{Clément Pellet-Mary}
+
+\institute{ENS Paris}
+\doctoralschool{EDPIF}{564}
+\specialty{Physique}
+\date{07 Décembre 2022}
+
+\jurymember{1}{Dmitry Budker}{Professor JGU Mainz}{Rapporteur}
+\jurymember{2}{Philippe Tamarat}{Professeur, Université de Bordeaux}{Rapporteur}
+\jurymember{3}{Maria Chamarro}{Professeure, Sorbonne Université}{Examinateur}
+\jurymember{4}{Vincent Jacques}{Dirrecteur de recherche, CNRS}{Examinateur}
+\jurymember{5}{Alexandre Tallaire}{Dirrecteur de recherche, CNRS}{Examinateur}
+\jurymember{6}{Gabriel Hétet}{Maitre de conférence, ENS Paris}{Directeur de thèse}
+% \jurymember{9}{Prénom NOM}{Titre, établissement}{Invité}
+% \jurymember{10}{Prénom NOM}{Titre, établissement}{Invité}
+
+\frabstract{
+Le centre NV du diamant est un candidat prometteur pour de nombreuses technologies quantiques, que ce soit pour les communications quantiques, le calcul quantique ou la métrologie quantique.
+Grace aux progrès de synthétisation du diamant, des échantillons contenant une haute concentration en centres NV sont aujourd’hui réalisables.
+Cette augmentation de la densité de centres NV s’accompagne d’une augmentation des interactions entre ces derniers, ce qui donne lieu à une physique complexe et encore largement inexplorée.
+De nombreuses applications récentes, telles que l’hyper-polarisation en RMN, la bio-imagerie ou encore la spin-mécanique repose sur des ensembles denses de centres NV.
+Comprendre et maîtriser les propriétés d’ensemble de ces derniers est un point crucial pour l’établissement de ces technologies.
+
+Cette thèse étudie les échanges de polarisation, par relaxation croisée, entre des ensembles denses de centres NV, ainsi que les interactions entre les centres NV et d’autres impuretés paramagnétiques présentes dans le diamant.
+Le but de ces travaux est de mieux comprendre les propriétés des ensembles de centres NV, et d’exploiter certaines de ces propriétés pour de nouvelles applications telles que la détection d’impuretés ou la magnétométrie.
+
+Dans un premier temps, nous détaillerons les propriétés des centres NV et nous introduirons des concepts liés à l’interaction dipolaire entre spins et aux relaxations croisées.
+Nous verrons ensuite comment les relaxations croisées peuvent être utilisées pour détecter optiquement des résonances de spins qui ne sont pas optiquement actifs, en l’occurrence les défauts VH- et War1 dans des diamants CVD.
+Enfin nous étudierons les relaxations croisées entre différents sous-groupes de centre NV.
+Nous verrons d’abord leur influence sur la dynamique du spin et leur origine microscopique, puis nous étudierons leur dépendance avec le champ magnétique et leur potentielle utilisation pour la magnétométrie en champs faible.
+}
+
+\enabstract{
   The NV center in diamond is a promising candidate for emerging quantum technologies, including quantum communication, quantum computing and quantum metrology.
@@
 We will then investigate the role of the magnetic field on cross-relaxations as well as their potential use in low field magnetometry.
+}
 
-\chapter*{Résumé français}
+\frkeywords{centre NV, dynamique de spin, magnétométrie}
+\enkeywords{NV center, spin dynamics, magnetometry}
 
-Le centre NV du diamant est un candidat prometteur pour de nombreuses technologies quantiques, que ce soit pour les communications quantiques, le calcul quantique ou la métrologie quantique.
-Grace aux progrès de synthétisation du diamant, des échantillons contenant une haute concentration en centres NV sont aujourd’hui réalisables.
-Cette augmentation de la densité de centres NV s’accompagne d’une augmentation des interactions entre ces derniers, ce qui donne lieu à une physique complexe et encore largement inexplorée.
-De nombreuses applications récentes, telles que l’hyper-polarisation en RMN, la bio-imagerie ou encore la spin-mécanique repose sur des ensembles denses de centres NV.
-Comprendre et maîtriser les propriétés d’ensemble de ces derniers est un point crucial pour l’établissement de ces technologies.
+\begin{document}
 
-Cette thèse étudie les échanges de polarisation, par relaxation croisée, entre des ensembles denses de centres NV, ainsi que les interactions entre les centres NV et d’autres impuretés paramagnétiques présentes dans le diamant.
-Le but de ces travaux est de mieux comprendre les propriétés des ensembles de centres NV, et d’exploiter certaines de ces propriétés pour de nouvelles applications telles que la détection d’impuretés ou la magnétométrie.
-
-Dans un premier temps, nous détaillerons les propriétés des centres NV et nous introduirons des concepts liés à l’interaction dipolaire entre spins et aux relaxations croisées.
-Nous verrons ensuite comment les relaxations croisées peuvent être utilisées pour détecter optiquement des résonances de spins qui ne sont pas optiquement actifs, en l’occurrence les défauts VH- et War1 dans des diamants CVD.
-Enfin nous étudierons les relaxations croisées entre différents sous-groupes de centre NV.
-Nous verrons d’abord leur influence sur la dynamique du spin et leur origine microscopique, puis nous étudierons leur dépendance avec le champ magnétique et leur potentielle utilisation pour la magnétométrie en champs faible.
+\maketitle{}
 

### 2022-10-19 Manuscrit Thèse/these_fusion.tex
toto

@@
 
-\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
-
-\title{}
+\title{Cross-relaxation in dense ensembles of NV centers and application to magnetometry}
+
+\author{Clément Pellet-Mary}
+
+\institute{ENS Paris}
+\doctoralschool{EDPIF}{564}
+\specialty{Physique}
+\date{07 Décembre 2022}
+
+\jurymember{1}{Dmitry Budker}{Professor JGU Mainz}{Rapporteur}
+\jurymember{2}{Philippe Tamarat}{Professeur, Université de Bordeaux}{Rapporteur}
+\jurymember{3}{Maria Chamarro}{Professeure, Sorbonne Université}{Examinateur}
+\jurymember{4}{Vincent Jacques}{Dirrecteur de recherche, CNRS}{Examinateur}
+\jurymember{5}{Alexandre Tallaire}{Dirrecteur de recherche, CNRS}{Examinateur}
+\jurymember{6}{Gabriel Hétet}{Maitre de conférence, ENS Paris}{Directeur de thèse}
+% \jurymember{9}{Prénom NOM}{Titre, établissement}{Invité}
+% \jurymember{10}{Prénom NOM}{Titre, établissement}{Invité}
+
+\frabstract{
+Le centre NV du diamant est un candidat prometteur pour de nombreuses technologies quantiques, que ce soit pour les communications quantiques, le calcul quantique ou la métrologie quantique.
+Grace aux progrès de synthétisation du diamant, des échantillons contenant une haute concentration en centres NV sont aujourd’hui réalisables.
+Cette augmentation de la densité de centres NV s’accompagne d’une augmentation des interactions entre ces derniers, ce qui donne lieu à une physique complexe et encore largement inexplorée.
+De nombreuses applications récentes, telles que l’hyper-polarisation en RMN, la bio-imagerie ou encore la spin-mécanique repose sur des ensembles denses de centres NV.
+Comprendre et maîtriser les propriétés d’ensemble de ces derniers est un point crucial pour l’établissement de ces technologies.
+
+Cette thèse étudie les échanges de polarisation, par relaxation croisée, entre des ensembles denses de centres NV, ainsi que les interactions entre les centres NV et d’autres impuretés paramagnétiques présentes dans le diamant.
+Le but de ces travaux est de mieux comprendre les propriétés des ensembles de centres NV, et d’exploiter certaines de ces propriétés pour de nouvelles applications telles que la détection d’impuretés ou la magnétométrie.
+
+Dans un premier temps, nous détaillerons les propriétés des centres NV et nous introduirons des concepts liés à l’interaction dipolaire entre spins et aux relaxations croisées.
+Nous verrons ensuite comment les relaxations croisées peuvent être utilisées pour détecter optiquement des résonances de spins qui ne sont pas optiquement actifs, en l’occurrence les défauts VH- et War1 dans des diamants CVD.
+Enfin nous étudierons les relaxations croisées entre différents sous-groupes de centre NV.
+Nous verrons d’abord leur influence sur la dynamique du spin et leur origine microscopique, puis nous étudierons leur dépendance avec le champ magnétique et leur potentielle utilisation pour la magnétométrie en champs faible.
+}
+
+\enabstract{
+  The NV center in diamond is a promising candidate for emerging quantum technologies, including quantum communication, quantum computing and quantum metrology.
+Thanks to the recent progress in diamond synthesis, samples with high NV center concentration can now be created.
+This increase in the NV density comes with an increase of the interaction between them, which leads to a rich and complex physics.
+Numerous applications, such as NMR hyperpolarization, bio-imagery or spin mechanics rely on the use of dense NV ensembles.
+Understanding and mastering the properties of ensembles of NV centers will be a crucial point in the development of these technologies.
+
+This doctoral thesis focuses on cross-relaxation between dense ensembles of NV centers, as well as on the interaction between NV centers and other diamond paramagnetic impurities.
+The aim of this work is to better understand the properties of dense NV centers ensembles, and to exploit them for new applications such as the detection of impurities and magnetometry.
+
+In the first part, we will detail the properties of single NV centers and introduce notions regarding dipole-dipole spin coupling and cross-relaxations.
+We will then show how cross-relaxations can be used to optically detect dark spin resonances, in this case the VH- and War1 defects in CVD-grown diamond.
+Finally, we will present a study of cross-relaxations between subgroups of NV centers.
+We will first cover the influence of cross-relaxation on the spin dynamics and discuss its microscopic origin.
+We will then investigate the role of the magnetic field on cross-relaxations as well as their potential use in low field magnetometry.
+}
+
+\frkeywords{centre NV, dynamique de spin, magnétométrie}
+\enkeywords{NV center, spin dynamics, magnetometry}
+
 \begin{document}
-\begin{refsection}
-
-\chapter{Introduction to Nitrogen-Vacancy centers and their common usage}
+
+\maketitle{}
+\setcounter{tocdepth}{1}
+
+\tableofcontents
+
+\chapter*{Introduction}
+
+Quantum mechanics was first developed at the beginning of the 20th century to explain physical observations such as the blackbody radiation \citep{planck1900theorie} or the photoelectric effect \citep{einstein1905erzeugung}.
+The quantum theory was subsequently developed in the first half of the 20th century and showed great success, in particular in giving a physical framework for chemistry \citep{pauling1931nature} and leading to a microscopic understanding of matter \citep{kittel1996introduction}.
+In the second half of the 20th century, technological development based on the new understanding of light and matter have led to what was retroactively called the first quantum revolution \citep{thew2019focus}.
+Technologies developed in the 50s and 60s include the semiconductor transistor \citep{bardeen1948transistor}, the laser \citep{maiman1960stimulated}, and nuclear magnetic resonance \citep{abragam1961principles} which have since revolutionized the fields of information science, global communications and medical imaging.
+
+A new field of quantum technology has emerged in the last two decades.
+This ``second quantum revolution" differs from the first one in the usage of individual quantum systems, such as single photons \citep{gleyzes2007quantum}, atoms \citep{neuhauser1980localized} or electrons \citep{peil1999observing}, which have properties that can not be emulated by larger classical systems.
+These properties include quantum entanglement, superposition or measurement and form the basis of the newly formed quantum information science \citep{nielsen2002quantum, vedral2006introduction, hayashi2006quantum}.
+
+Central to quantum information science is the idea of quantum bits or qubits \citep{schumacher1996sending}, which are the building blocks of quantum information in analogy to the bits of classical information theory.
+Physical implementation of a qubit could be any quantum system with two well defined quantum states, as long as these states can be initialized, manipulated and readout \citep{divincenzo2000physical}.
+Popular qubit candidates include superconducting circuits \citep{nakamura1999coherent, orlando1999superconducting}, photons \citep{bennett1992quantum, zhong2020quantum}, quantum dots \citep{veldhorst2014addressable, zajac2018resonantly}, trapped ions \citep{friis2018observation, wright2019benchmarking} and single crystal defects \citep{jelezko2006single, baranov2011silicon, zhong2015optically}.
+
+Defects in large band gap crystals are sometimes referred to as artificial atoms \citep{buluta2011natural}, as the localized wave function of the electrons surrounding the impurity  give rise to discrete energy levels similar to that of an atom.
+Of all the possible crystals and defects, the most studied defect for quantum applications is the negatively charged nitrogen-vacancy center in diamond \citep{aharonovich2016solid, de2021materials}, often abbreviated to ``NV center".
+Although this defect had been known and correctly identified almost 50 years ago \citep{davies1976optical}, interest in the NV center really sparked after it was first isolated in 1997 \citep{gruber1997scanning}.
+The detection of single NV centers proved that NV centers fluoresce brightly \citep{gruber1997scanning}, are photostable \citep{kurtsiefer2000stable}, and that their electronic spin can be polarized and readout optically \citep{jelezko2004observation}, making them an promising solid state qubit.
+
+Since then, single NV centers have been successfully used to create quantum entanglement over several km \citep{hensen2015loophole}, to store quantum information at room temperature for more than one second \citep{maurer2012room}, to detect single proteins via nuclear magnetic resonance \citep{lovchinsky2016nuclear}, and for many other applications in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} or quantum sensing \citep{degen2017quantum}.
+
+As the interest in single NV center grew, diamond synthesis improved significantly \citep{achard2020chemical, barry2020sensitivity, edmonds2020generation}  which benefited not only single NV centers but also NV center ensemble.
+While single NV center are required for most quantum information processes \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble.
+Four of these applications are pictured in Fig.
+\ref{shema introduction}.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.9\textwidth]{Intro et conclusion/Figures/Fig_intro}
+\caption{Example of applications with dense ensemble of NV centers.
+(a) In vivo nanodiomond biomarkers (from \citep{mohan2010vivo}).
+(b) Quantum sensing with NV ensemble (from \citep{ku2020imaging}).
+(c) Dynamic nuclear polarization of an external spin bath (from \citep{healey2021polarization}).
+(d) Spin mechanics with trapped micro diamonds (from \citep{delord2020spin}).}
+\label{shema introduction}
+\end{figure}
+
+The first application for NV center ensemble is the use of nano or micro diamonds as fluorescent biomarkers \citep{fu2007characterization, mohan2010vivo}.
+NV centers are good candidates for biomarkers as they emit a bright and stable photoluminesecence, and because diamond nanocrystals present low toxicity \citep{schirhagl2014nitrogen}.
+Some applications also use the spin degree of freedom of the NV centers, either to perform background free fluorescence microscopy \citep{chapman2013background} or to monitor the angular motion of the nanodiamonds \citep{mcguinness2011quantum, feng2021association}.
+For most of these applications, NV ensemble provide a better visibility than single NV centers and are therefore preferred.
+
+The second NV ensemble application is for quantum sensing \citep{degen2017quantum}.
+NV centers are predominantly used as magnetic field sensors \citep{rondin2014magnetometry, barry2020sensitivity}, although they can also detect electric fields \citep{dolde2011electric, michl2019robust}, temperature \citep{acosta2010temperature, kucsko2013nanometre} ,crystal strain \citep{ovartchaiyapong2014dynamic, doherty2014electronic} or rotation \citep{ledbetter2012gyroscopes, ajoy2012stable}.
+Both single and ensemble of NV centers can be used for sensing: single centers offer unmatched spatial resolution \citep{mitchell2020colloquium} down to a few nm \citep{lovchinsky2016nuclear}, while ensemble provide better sensitivities \citep{wolf2015subpicotesla, barry2020sensitivity}.
+
+The third NV ensemble application is nuclear magnetic resonance (NMR) dynamical nuclear polarization \citep{eills2022spin}.
+Dynamic nuclear polarization consists in increasing the nuclear polarization of a sample by transferring electron polarization to the nuclei bath \citep{abragam1978principles}.
+Because the NV electronic spin can be optically polarized at room temperature, micro and nano diamonds containing NV centers are promising candidates as hyperpolarization agents \citep{tetienne2021prospects}.
+Nuclear polarization of spins inside the diamond lattice was demonstrated with single NV centers \citep{jacques2009dynamic, smeltzer2009robust} and NV ensemble \citep{king2015room, scheuer2016optically, schwartz2018robust}.
+Polarization transfer to nuclear spins outside of the diamond is an ongoing challenge \citep{eills2022spin, healey2021polarization}, but dense NV ensemble are believed to be necessary to achieve high external polarization \citep{tetienne2021prospects, rizzato2022polarization}.
+
+Finally, the fourth NV ensemble application is the recent field of spin mechanics \citep{perdriat2021spin}.
+Similarly to optomechanics \citep{aspelmeyer2014cavity}, the goal of spin mechanics is to couple a quantum system to the mechanical degree of freedom of a macroscopic object.
+Strong coupling between single NV center and an external mechanical oscillators has been observed \citep{rabl2009strong, kolkowitz2012coherent}, but coupling between NV centers and the motion of the diamond has so far only been achieved with dense NV ensemble \citep{delord2020spin, pellet2021magnetic, perdriat2022angle}.
+Coupling several NV centers to the same mechanical oscillator could also lead to interesting magnetic phase transitions \citep{wei2015magnetic, ma2017proposal}.
+
+Most of the NV ensemble experiments discussed previously work under the assumption that the ensemble is constituted of $N$ independent single NV centers.
+However, as often in physics, \textit{more is different} \citep{anderson1972more}.
+Interaction between the NV center themselves or with their environment can lead to significantly different behavior between ensemble and single spins.
+These difference include a modification of the charge state dynamics \citep{giri2018coupled} and of the spin dynamics \citep{dobrovitski2008decoherence, jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization}, as well as the apparition of exotic phenomena such as time crystal phase \citep{choi2017observation}, Anderson localization \citep{kucsko2018critical} or superradiance \citep{bradac2017room, angerer2018superradiant}.
+
+The work presented in this manuscript focuses on dipolarly coupled dense ensemble of NV centers.
+More specifically, we look at spin polarization exchange between the NV centers themselves and with their environment.
+Polarization exchange (flip-flop, spin diffusion, cross-relaxation, ...) in dense ensemble of spins has been abundantly studied in the context of NMR \citep{abragam1978principles} and spintronics \citep{vzutic2004spintronics}.
+Because NV centers tend to be relatively dilute, the question of spin exchange between between NV themselves \citep{choi2017depolarization} or with their environment \citep{hall2016detection} has only recently been addressed.
+However, as samples with higher NV concentration become available \citep{acosta2009diamonds, tallaire2020high, shenderova2019synthesis}, these aspects are becoming more and more relevant.
+
+The goal of this present work is both to better understand the physics of dense NV ensemble, which is crucial for many of the NV ensemble applications cited previously, and to exploit some of these many-body properties for new applications.
+
+\bigskip
+This manuscript is organized in the following way:
+
+\medskip
+The first chapter introduces the theoretical and experimental concepts used in this manuscript.
+This includes a brief overview of diamond and NV centers fabrication processes, the physics of single NV centers and a presentation of magnetic dipole-dipole interaction and cross-relaxation.
+
+The second chapter covers the observation of cross-relaxation between NV centers and other spin impurities in CVD-grown diamond.
+We use measurements on the NV centers to identify these impurities and to quantify some of their attributes.
+Most of the results of this chapter have been published in  \citep{pellet2021optical} and \citep{ngambou2022improving}.
+
+The third chapter focuses on the dipole-dipole mediated spin relaxation within dense ensemble of NV centers.
+Following the work presented in \citep{choi2017depolarization}, we introduce the concept of NV-flutcuator to explain this phenomenon and present experimental and theoretical results solidifying this hypothesis.
+The results of this chapter have been partly published in \citep{pellet2022spin} and \citep{pellet2021magnetic}.
+
+Finally, the fourth chapter presents a study on NV-NV dipolar relaxation under low magnetic field.
+This chapter covers both the rich physics of NV centers under low magnetic field, and a potential application of the low field depolarization as a new magnetometry protocol, as well as comparisons with other NV magnetometry protocols.
+The results of this chapter have in large part been published in \citep{pellet2022spin}.
+
+\chapter{Spin properties of single and ensemble NV centers}
 The aim of this first chapter is to introduce concepts related to Nitrogen-vacancy (NV) centers, to the diamond material, and to the dipole-dipole interaction which are necessary to explain the work presented in the later chapters.
@@
 The focus of chapter 2 will be on the cross-relaxation between NV centers and other spin impurities in CVD-grown diamond, while the focus of chapter 3 and 4 will be on the cross-relaxation between NV centers and other NV centers, as well as their potential application in magnetometry.
-\printbibliography
-\end{refsection}
- \begin{refsection}
-
-\chapter{Detection of dark spins via cross-relaxations}
-In this chapter, we will see how the interplay between the optical and spin properties of the NV centers can be transferred to other spins in the crystal by coupling them to NV centers.
-More specifically, I will discuss the first optical detection of two spin species, VH$^-$ and War1, in a CVD diamond by the mean of cross-relaxations with NV centers.
-\section{Flip-Flops, double flips and cross-relaxation}
-\label{Sec_CR}
-
-\subsection{Resonant interaction : flip-flops and double flips}
-
-The dominant form of interaction between two distant spins is the magnetic dipole-dipole interaction.
-For two spins with with vector spin Hamiltonian $\hat{\mathbf{S}}_1$ and $\hat{\mathbf{S}}_2$, separated by a distance $\mathbf{r}=r\mathbf{u}$, the dipole-dipole interaction Hamiltonian reads (see Appendix A) :
-\begin{equation}
-\mathcal{H}_{dd}=\frac{J_0}{r^3}\left[ \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2 - 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u})\right],
-\end{equation}
-where $J_0= \frac{\mu_0\gamma_1\gamma_2 \hbar^2}{4\pi}$.
-For two electronic spins with $g-$factors close to 2 (which is the case for the NV center and most spin defect in diamond), the numerical value of $J_0$ is $(2\pi)\,52\ \rm{MHz}\cdot\rm{nm}^3$.
-
-Flip-flop, in the context of dipole-dipole interaction, is the mechanism which exchanges a quantum of spin between two spins.
-For an initial state of two spins $\ket{m_s^1=i;m_s^2=j}$, a flip-flop process will result in a final state $\ket{i+1;j-1}$ or $\ket{i-1;j+1}$.
-The effective rate of this process depends on the matrix element $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \mp 1}$, as well as the resonance condition between the initial and final state.
-We should note that for two identical spins, the flip-flop process $\ket{i;i\pm 1}\bra{i\pm 1,i}$ is always resonant.
-
-Another mechanism authorized by the dipole-dipole Hamiltonian is the double-flip, either up or down, which couple the states $\ket{i;j}$ and $\ket{i-1;j-1}$ or $\ket{i+1;j+1}$.
-For a single spin species in a strong magnetic field, the lift of the various spin levels by the Zeeman effect means that no double-flip process can be resonant.
-However with weak magnetic field or when two different spin species are present, double flip process can be as important as flip-flops since the matrix elements $\mel{i;j}{\mathcal{H}_{dd}}{i\pm 1;j \pm 1}$ are typically of the same order of magnitude as the flip-flop ones.
-
-\subsection{Cross-relaxations}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.9\textwidth]{chapter 2/Figures/Shema_CR}
-\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
-\label{CR_shema}
-\end{figure}
-
-Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
-This process con occur either through flip-flops, or through double flips, as long as the resonance condition between the two spins is met.
-
-Fig.
-\ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin.
-When the two considered spin transitions become resonant, in this case by tuning the magnetic field, polarization from the NV center will be transferred to the second spin, making the NV center less polarized and the second spin more polarized than when they are out of resonance.
-
-The reason why CR tend to depolarize the NV center simply comes from a rate equation : since the initial NV population is higher in the $\ket{0}$ state than in the $\ket{-1}$ state, the flip-flop (or double flip) process that makes the NV center go to $\ket{-1}$ (symbolized by the two red arrows in Fig.
-\ref{CR_shema}) is more likely than the reverse flip-flop.
-If we only focus on the NV center, cross-relaxation is simply another thermalization process due to the coupling of the NV center to its environment.
-
-For a single NV center coupled to spin 2, assuming that spin 2 is initially depolarized and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$, we can compute the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by spin 2 as \citep{hall2016detection}:
-\begin{equation}
-\delta \Gamma_1=\frac{\Omega^2}{\Gamma_2^*} \frac{(\Gamma_2^*)^2}{(\delta \nu)^2+(\Gamma_2^*)^2},
-\label{delta gamma 1}
-\end{equation}
-where $\Omega=\frac{\mel{i;j}{\mathcal{H}_{\rm dd}}{j;i}}{\hbar}$ is the matrix element of the dipole-dipole Hamiltonian that couples the initial and final states, $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, and $\delta \nu$ is the detuning between the central frequencies of the NV center and spin 2.
-\section{The dark spins in diamond}
-
-We refer by dark spins to spin defects which do not interact with light, which in the case of diamond defects consist of almost every known spin defect bar the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
+
+\chapter{Cross-relaxations between NV centers and dark spins in CVD-grown diamond}
+Crystal defects are promising candidates for quantum information technologies, either through their optical or spin degree of freedom \citep{aharonovich2016solid, atature2018material, bassett2019quantum}.
+Spectrocopic studies are a necessary step in identifying and characterizing these defects, with electron paramagnetic resonance (EPR) being the spectrocopic tool of choice to measure the spin properties of the defects \citep{newton2007epr}.
+
+In this chapter, we present a method to characterize the spin of diamond impurities through cross-relaxation with NV centers \citep{hall2016detection,pellet2021optical}.
+This method allows an all-optical detection of non-NV center spin defects, a method that is both cheaper and more sensitive than traditional EPR.
+We focus in particular on defects in CVD-grown diamonds which have historically been less studied than HPHT diamond, and for which no prior NV cross-relaxation had been reported.
+
+The chapter is organized as follow: we will first detail the most common paramagnetic impurities in the diamond (the so-called ``dark spins") before explaining the proceeding of NV cross-relaxation experiments, and more generally of NV relaxometry.
+We will then focus on the detection of two particular defects in a CVD diamond, VH$^-$ and War1, and on their characterization (density, spin parameters, temperature dependence).
+We will finally conclude on the possible improvements and application of this detection method.
+
+The work presented in this chapter has been published in large part in the articles \citep{pellet2021optical, ngambou2022improving}.
+
+\section{Dark spins in diamond}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 2/Figures/tronche_de_spin}
+\caption{Crystalline structure of three spin defects: the NV center, the P1 enter (substituional Nitrogen) and the VH center.} 
+\label{repr NV P1 VH}
+\end{figure}
+
+We refer by dark spins to paramagnetic impurities for which the spin degree of freedom is not accessible optically.
+At room temperature, only two defects in diamond do not have a dark spin: the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
+Every other spin impurity identified so far is considered to be dark.
+
+In this section we will present the most common dark spins encountered in diamond, which are simply the most common paramagnetic impurities aside from the NV center, and the relation between the diamond growth process and the presence of these impurities.
+We will also briefly present EPR spectroscopy which is the most common method to characterize these dark spins.
 
 \subsection{P1 centers and $^{13}$C}
-By far the most common spin defects in NV-rich diamond, outside of NV centers, are the substitutional nitrogen centers N$^0_S$ known as the P1 centers, which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which is present at 1.1\% in natural abundance carbon, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
-
-In NV-rich diamonds, these two impurities are the main causes of the inhomogeneous dephasing rate $\Gamma_2^*=1/T_2^*$ of the NV centers ensemble \citep{barry2020sensitivity}.
-The dephasing comes from the distribution of dipole-dipole coupling from each NV centers to its neighboring spin defects, which varies in space and time.
-
-The dephasing rate caused by a given impurity is directly proportional to the diagonal dipole-dipole matrix elements \citep{bauch2020decoherence}: 
-\begin{equation}
-\Gamma_2^*(\rm P1) \propto \Omega_{\rm NV-P1}=\frac{1}{\hbar}\sqrt{\sum_j \abs{\mel{NV=i;P1=j}{\mathcal{H}_{\rm dd}}{NV=i;P1=j}}^2}.
-\label{Gamma2 P1}
-\end{equation}
-Since the dipole-dipole coupling rate scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity: $\Gamma_2^*(\rm P1) \propto [P1]$.
-In the case of P1 centers, this was verified experimentally by \citep{bauch2020decoherence} who found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
-The typical concentrations of P1 center go from [REF] for high purity CVD diamond, up to 500 ppm for type 1B HPHT or natural diamonds.
-
-$^{13}$C is much more abundant than P1 center ($\sim$ 10 000 ppm at natural abundance) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\cdot 10^3$ times smaller than that of the P1.
-This results in a contribution to $\Gamma_2^* \sim 10^3$ times smaller per impurity.
+The most common spin defects in NV-rich diamond are the substitutional nitrogen centers N$^0_S$ known as the P1 centers (see Fig.
+\ref{repr NV P1 VH}), which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which represents 1.1\% of natural abundance carbon atoms, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
+In NV-rich diamonds, these two impurities are the main causesof the NV center dephasing rates $\Gamma_2^*=1/T_2^*$ and $\Gamma_2=1/T_2$  \citep{barry2020sensitivity}.
+
+The dephasing rate caused by a given impurity is directly proportional to the average diagonal dipole-dipole matrix elements ($\Omega_{\rm shift}$ in (eq.
+\ref{eq.
+ff df})) between the two spin species \citep{taylor2008high, bauch2020decoherence}.
+Since the dipole-dipole interaction scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity: $\Gamma_2^*(\rm P1) \propto [P1]$.
+
+In the case of P1 centers, this was verified experimentally by looking at controlled sample with varying concentration of nitrogen \citep{bauch2020decoherence}.
+The authors of the study found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
+For the samples studied in this manuscript [REF], the P1 concentration was
+
+$^{13}$C nuclei are typically much more abundant than P1 center ($\sim$ 10 000 ppm without istopic engineering) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\cdot 10^3$ times smaller than that of the P1.
+As a result, the contribution of a single $^{13}$C nucleus to the NV dephasing rate is $\sim 10^3$ times that of a single P1 center.
 It was found experimentally that the broadening caused by $^{13}$C is $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 16\ \rm Hz/ppm$ \citep{van1997dependences, barry2020sensitivity}.
-For natural abundance diamond, this means that $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 160\ \rm kHz$.
+In samples with natural isotopic abundance of carbon, the inhomogeneous dephasing rate due to the $^{13}$C nuclear spins has a value $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 160\ \rm kHz$.
 
@@
 \subsection{EPR spectroscopy}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 2/Figures/EPR_spectometer.jpg}
+\caption{EPR specrometer. Credits: Bruker Corporation} 
+\label{photo EPR}
+\end{figure}
+
 Electron paramagnetic resonance (EPR), the electronic equivalent to NMR, is the the most common way to detect electronic spins in solids and liquids.
@@
 
-At room temperature, the thermal polarization of a transition at 9500 MHz is $\sim 0.002$.
-This low value, compared to $\sim 0.8$ for the NV centers, explains in part why EPR is way less sensitive than NV ODMR.
-The current noise floor for EPR is typically of $\sim 10^{11}$ spins for electronic spins in diamond \citep{mitchell2013x}.
+The usual EPR apparatus, shown in Fig.
+\ref{photo EPR}, is a bulky and expensive lab equipment ($\gtrsim$ 1 M€).
+In contrast, optical detection such as ODMR is technically easier, less expensive, and allows better spatial resolution.
+
+ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ spins \citep{mitchell2013x}.
+This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a spin with a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and in part from the more efficient photon detection in the optical range than in the microwave range.
+
+The key idea of this chapter is to allow the optical detection of spin defects that normally do not show ODMR thanks to cross-relaxation with NV centers.
 
@@
 
-\subsection{Principle of NV center Relaxometry}
-Detection \textit{via} cross-relaxation	with NV centers fall under the more general technique of NV relaxometry, where we are effectively measuring a modification in the NV center spin lifetime $T_1$.
-Relaxometry with NV centers has been used not only to measure resonant cross-relaxation with dark spins \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}, magnetic domains \citep{finco2021imaging}...
-
-There can be several advantages to use relaxometry protocols, which measures a change in the NV $T_1$, over the more common magnetometry protocols that rely on a change in the NV $T_2$ or $T_2^*$: the first and most obvious one is that relaxometry does not need to use an external microwave field.
-This however comes at the cost of a greater precision needed on the external magnetic field.
-Relaxometry also allows AC-magnetic field sensing in the GHz regime \citep{wang2022picotesla, alsid2022solid}, while traditional microwave-based AC-megnetometry is limited to tens of MHz.
-Finally, relaxometry measurement can be more sensitive than the T2-based one in some instances \citep{steinert2013magnetic}, which comes from the fact that changes in NV $T_1$ times are easier to detect than changes in $T_2$ or $T_2^*$ since $T_1$ is much longer than $T_2$ or $T_2^*$.
+Detection of dark spins via NV center cross-relaxation fall under the more general technique of NV relaxometry, which consists in measuring a modification of the NV center spin lifetime $T_1$.
+Relaxometry with NV centers has been used not only to measure cross-relaxation with dark spins \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}or magnetic domains \citep{finco2021imaging}.
+
+The protocols used in NV relaxometry measure either directly the spin $T_1$ or indirectly through the NV photoluminescence.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{chapter 2/Figures/relaxo_pulsed_continuous}
+\includegraphics[width=0.9\textwidth]{chapter 2/Figures/relaxo_pulsed_continuous_2}
 \caption{Representation of the two relaxometry protocol.
-Left : representation of the three level system of the NV center with the polarization rate $\Gamma_{\rm las}$ and spin decay rates $\Gamma_1=1/T_1$ and the perturbation $\delta \Gamma$.
-Right: simulation of a $T_1$ measurement for $1/\Gamma_1=3\ \rm ms$ and $1/(\Gamma_1+\delta \Gamma)=2\ \rm ms$.
-$\delta \rm S$ represents the maximum difference in signal between the two curves for a dark time $\tau$.}
+(a) Simple $T_1$ measurement pulse sequence.
+(b) Simulation of two $T_1$ measurement with different $T_1$ values.
+$\tau$ represents the optimal dark time in order to maximize the signal $\delta S$ (c) Representation of the optical pumping (green arrows) and depolarization (blue arrows) of the NV center spin.
+(d) Simulation of the PL modification caused by a change in the spin $T_1$.
+The spin $T_1$ profile in term of the magnetic field was arbitrarily chosen.}
 \label{T1 vs PL}
@@
 
-There are two ways to measure a change in the spin $T_1$, as represented on Fig.
+The two most common NV relaxometry protocols are represented on Fig.
 \ref{T1 vs PL}.
-The first and most basic method simply consist in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, which itself results from an equilibrium between the polarization rate $\Gamma_{\rm las}$ and the spin decay rate $\Gamma_1=1/T_1$.
-A change in $\Gamma_1$ will therefore conduct to a new equilibrium with a different PL.
-
-The second method is a pulsed sequence referred as single$-\tau$ measurement, used for example in \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}.
-It simply consists in a $T_1$ measurement sequence (as seen in chapter 1) where you first polarize the spins in the $\ket{0}$ state with a laser pulse and read out the spin state with a second laser pulse after a dark time $\tau$.
-In this case however, instead of scanning the time $\tau$, you fix its value to maximize the sensitivity in a change of $T_1$ (typically $\tau \approx T_1$).
-
-It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the first case you need to adjust $\Gamma_{\rm las} \sim \Gamma_1$ to maximize the PL contrast, meaning that the laser power used is very far from the saturation limit (typically $\Gamma_1 \approx 10^3\ \rm s^{-1}$ and $\Gamma_{\rm sat} = 5\cdot10^6\ \rm s^{-1}$\citep{dreau2011avoiding}), while the second method needs to wait a time $\tau \sim T_1$ between each measurement, which result in a similar PL count and shot noise limit in both case.
-
-On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and uses an overall smaller laser power, however it is more sensitive to drifts and changes in the optical setup.
+The first method simply consists in measuring the spin lifetime, following the protocols described in \ref{sec T1 NV}.
+In Fig.
+\ref{T1 vs PL}, we picture the simpler $T_1$ measurement protocol, but the technique can be used with the common mode rejection protocol.
+In order to more efficiently measure a modification of the spin $T_1$, the ``single$-\tau$" protocol does not record the entire depolarization curve represented in Fig.
+\ref{T1 vs PL}-b), but instead probes the spins for a single dark time value $\tau$ \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}.
+This value is chosen in order to maximize the change in signal $\delta S$ for a given change in lifetime $\delta T_1$, typically $\tau \sim T_1$.
+
+ The second method simply consists in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, and since the steady state population in the $\ket{0}$ state results from the equilibrium between the optical polarization rate $\Gamma_{\rm las}$ and the depolarization rate $\Gamma_1=1/T_1$.
+A simple rate equation for the $\ket{0}$ state population $\rho_{00}$ gives:
+ \begin{equation}
+ \rho_{00}=\frac{\Gamma_{\rm las}+\Gamma_1}{\Gamma_{\rm las}+3\Gamma_1}.
+ \end{equation}
+
+ Fig.
+\ref{T1 vs PL} shows a simulation of NV centers PL when the spin decay rate $\Gamma_1$ is modified.
+In this fictional scenario, we assume that $\Gamma_1$ resonantly increases for a given value of the magnetic field (we will see later in this manuscript physical examples corresponding to this scenario).
+We can see that the PL decreases as an answer to the $\Gamma_1$ increase.
+The exact value of the PL drop ($\sim 2\%$ in the crase presented here) depends on the optical polarization rate $\Gamma_{\rm las}$ as well as the modification of the spin lifetime.
+
+It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the PL case, the laser intensity needs to be adjusted so that $\Gamma_{\rm las} \sim \Gamma_1$ in order to maximize the PL contrast, while the single$-\tau$ method requires a waiting time $\tau \sim T_1$ between each measurement, which results in a similar PL count and shot noise limit in both case.
+
+On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and because it requires a smaller laser power to work optimally.
+It is however more sensitive to drifts and changes in the optical setup.
 Most of the relaxometry measurement in this manuscript were performed using a PL-based detection.
@@
 
-Outside of cross-relaxations, another protocol exists to address dark spin resonance with NV centers, based on a modification of the NV $T_2$ time \citep{serbyn2014interferometric}.
-This protocol called double electron-electron resonance (DEER) is depicted on Fig.
+Aside from cross-relaxations, another protocol exists to address dark spin resonance with NV centers, based on a modification of the NV $T_2$ time \citep{mamin2012detecting, serbyn2014interferometric}.
+This protocol, called double electron-electron resonance (DEER), is depicted on Fig.
 \ref{DEER}.
-The pulse sequence consist in a spin-echo sequence on the NV center, as described in chapter 1, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers.
-This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center wont be rephased.
-The newly measured $T_2$ time should therefore be slightly shorter:
+The pulse sequence consists in applying a spin-echo on the NV center, as described in \ref{sec T2 echo}, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers.
+This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center will not be rephased.
+The newly measured $T_2$ time should thus be slightly shorter.
 
@@
 
-By scanning the frequency of the second microwave (RF2) and monitoring the $T_2$ time of the NV centers, one can therefore find dark spin resonance without having to bring them to resonance with NV centers, which can be useful when the dark spin resonance is very far detuned from the NV one.
-
-While both cross-relaxation and DEER measure the dipole-dipole coupling of dark spins to an NV center, there are several differences between the two protocols:
-
-Frist, the two protocols do not measure the same elements of the dipole-dipole Hamiltonian: CR measures the off-diagonal elements $\Omega_{\rm CR}=\mel{i,j}{\mathcal{H}_{\rm dd}}{j,i}/\hbar$ which are responsible for the flip-flops or double flips, while DEER measures the diagonal elements $\Omega_{\rm DEER}=\mel{i,j}{\mathcal{H}_{\rm dd}}{i,j}/\hbar$ which are responsible for the pure dephasing.
-These terms however are on average of the same amplitude, and when averaging over an ensemble we can consider $\Omega_{\rm CR} \sim \Omega_{\rm DEER}$
-
-Second and more importantly, the scaling is not the same in both cases : CR measures a change in the NV decay rate $\Gamma_1=1/T_1$ which is proportional to $\Omega_{\rm CR}^2$, as given by eq.
-\ref{delta gamma 1}, while DEER measures a change in $\Gamma_2=1/T_2$ which is proportional to $\Omega_{\rm DEER}$, as given by eq.
-\ref{Gamma2 P1}.
-This means that the signal of CR scales quadratically with the impurity concentration, while the DEER signal scales linearly with it.
-CR is therefore more suited to probe concentrated impurities, when the average distance between the NV and the probed spin is small.
+By scanning the frequency of the second microwave (RF2) and monitoring the $T_2$ time of the NV centers, one can find dark spin resonance without having to bring them to resonance with NV centers, which can be useful when the dark spin resonance is very far detuned from the NV one.
+
+Both relaxometry and DEER protocols have their own advantages.
+While DEER allows more control on the probed impurity, allowing in particular coherent control of the dark spin \citep{sushkov2014magnetic}, relaxometry protocols are generally easier to implement as they do not require a pulsed microwave setup.
+In some cases at least, relaxometry protocols have been shown to be more sensitive than $T_2$ based protocols such as DEER due to the fact that $T_1$ is typically orders of magnitude greater than $T_2$ \citep{steinert2013magnetic}.
+
+Relaxometry protocols do require however specific magnetic field values in order to bring in and out of resonance the two spins involved in the cross-relaxation.
+The choice of the field orientation and strength in order to best observe NV-dark spin cross-relaxations will be covered in the following section.
 
@@
 
-Since relaxometry does not rely on a microwave field, the only parameter left to tune is the external magnetic field, which is often scanned with an electromagnet.
-Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect dark spins.
+As relaxometry does not rely on a microwave field, the only external parameter left to tune is the magnetic field.
+Typically, the magnetic field is scanned in a fixed direction while the signal, either PL or T1, is recorded.
+Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect NV-dark spins CR.
+
+This section focuses on particular field orientation with respect to the diamond crystalline axes, and details the advantages and disadvantages of the two main choices that are the [100] and [111] direction in order to detect four particular dark spins: P1 centers, Vh$^-$, War1 and the $^{13}$C$-$NV complex\footnote{The $^{13}$C$-$NV complex is normally not dark, as will be discussed in its own section.}.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{chapter 2/Figures/121_vs_22_NV}
+\includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_NV}
 \caption{Simulated transition frequencies of the $\ket{0} \to {-1}$ and $\ket{0} \to {+1}$ transitions for the 4 classes of NV centers as a function of the magnetic field.
@@
 
+We will now compare these two possible field orientation in order to observe cross-relaxation between NV centers and the four previously mentioned dark spin species.
+
 \subsection{CR condition between NV$^-$ and P1}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{chapter 2/Figures/121_vs_22_P1}
+\includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_P1}
 \caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}
@@
 
-In order to optically detect CR between an NV center and a dark spin, two conditions must be met: the NV center must be polarized in order to have a population transfer between the NV and the dark spin, and there must be a difference in brightness between the two NV levels involved in order to have an optical signature of the CR.
-
-These two conditions are in fact the same needed to observe an ODMR spectrum (REF chapter 1), and while the NV centers verifies both criteria in presence of a purely longitudinal magnetic field, this is not always the case anymore in presence of a transverse magnetic field (REF chapter 1).
-
-Fig.
-\ref{121 vs 22 contrast} shows the simulated relative contrast of an ODMR spectrum, as computed from \citep{tetienne2012magnetic}, for a magnetic field aligned along [100] or [111].
-When the NV center is aligned with the magnetic field, the constrats remains equals to 1 regardless of the magnetic field because there is no transverse field.
-When the NV is not aligned however, either in the [111] or [100] case, the ODMR contrast quickly drops down for B$>100\ \rm G$.
-
-This tells us that we should not expect to observe a CR signal for B greater than a few hundreds G when B is scanned along the [100] axis.
-We could however observe CR for arbitrarily large magnetic field when scanned along the [111] axis, as long as the field is properly aligned with the NV axis.
+Similarly to ODMR, there are two required conditions to observe NV cross-relaxation on the PL signal: 
+\begin{itemize}
+\item There must be a difference in population between the two NV spin states involved in the CR (e.g.
+$\ket{0}$ and $\ket{-1}$), otherwise there would be no population transfer with the unpolarized dark spin.
+\item The two NV spin states must have a difference brightness in order to be optically detectable.
+\end{itemize}
+
+These two conditions are fulfilled when there is no magnetic field or purely longitudinal magnetic field, but not anymore in the presence of transverse magnetic field which mixes the NV spin Hamiltonian eigenstates \citep{tetienne2012magnetic}.
+
+Fig.
+\ref{121 vs 22 contrast} shows the simulated relative contrast of a CR process \footnote{Due to the similarity between ODMR and CR, we assume here that the ODMR contrast is equal to the CR contrast} computed from the rates found in \citep{tetienne2012magnetic}, for a magnetic field aligned along either [100] or [111].
+When the NV center is aligned with the magnetic field, the contrast remains equal to 1 regardless of the magnetic field because there is no transverse field.
+When the NV is not aligned however, the CR contrast quickly drops down for B$>100\ \rm G$.
+
+The [100] direction is therefore not well suited to observe CR at magnetic fields higher than a few hundreds gauss.
+Fortunately, the co-resonance between NV centers and VH$^-$, War1 or $^{13}$C-NV$^-$ all happen for $|\mathbf{B}| < 150 \ \rm G$, and can thus be detected with a [100] oriented magnetic field.
 
 \section{Dark spin spectroscopy with NV centers}
+This section shows experimental observation of NV-dark spins CR.
+Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross} had been performed on the detection of dark spins through CR with NV centers, and to our knowledge, all this studies where done on the detection of P1 centers.
+
+We show an example of one this study to illustrate the principle of dark spin spectroscopy before showing our results on the detection of VH$^-$ and War1 in a CVD diamond.
 
@@
 Similarly to fig.
-\ref{CR P1 exp}, we can see a slowly decaying envelope of the PL due to the transverse magnetic field (we don't see the plateau here since the magnetic field stops at 150 G).
-We can also see three deviations to the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
-
-In order to better discern the three features, we fitted the envelope by using a polynomial fit of order 4 on the smooth part of the data.
-We previously tried to use a fit based on the expected PL rate of the NV centers, by using the model in \citep{tetienne2012magnetic}, but the results were not satisfying.
+\ref{CR P1 exp}, there is a slowly decaying envelope of the PL due to the transverse magnetic field \footnote{The plateau is not visible in this case as the scan stops at 150 G}.
+There also are three deviations from the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
+
+In order to better discern the three features, we fitted the envelope with a polynomial fit of order 4.
+We previously tried to simulate the PL dependency on the transverse field  by using the model in \citep{tetienne2012magnetic}, but the result did not correctly fit our data.
 
@@
 The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below.
-We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the process, and while we don't know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
+We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the growth process, and while we do not know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
 
@@
 
-For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively, with a slightly stronger confidence for the VH$^-$ since more is known about it.
+For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively.
 
@@
 
+\bigskip
+
+The CR spectra shown in this section can be used to detect the presence of particular impurities in the sample, but more quantitative information on these impurities can also be extracted from these spectra, which is the object of the following section.
+
 \section{Quantitative estimates of the NV-Dark spins CR}
 
-We can extract some quantitative information about the dark spins by exploiting the CR spectrum shown in Fig. \ref{CR VH exp}.
+This section exploit the CR spectrum shown in Fig. \ref{CR VH exp} in order to extract information on the dark spins Hamiltonian parameters, linewidth and concentration.
 
@@
 
-With all this considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):
-
-For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose the dark to bright NV ratio value from \citep{choi2017depolarization} at 0.25.
+From these considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):
+
+For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose a dark to bright NV ratio of 0.25 as found in \citep{choi2017depolarization}.
 
@@
 To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations.
-This detection can be way more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up.
-It is however not as versatile and should be seen as a complimentary tool for diamonds with dense NV ensemble.
-
-We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan, much less noisy than the [111] scans for weak magnetic field.
-
-Finally we have seen an application of the CR to detect optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.
-
-There still remains a lot of open questions and there are several experiments that could be build upon these results. I will list some of them here.
+This detection can be significantly more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up.
+This technique however is only applicable to a few defects in NV-rich diamond, meaning that its main application would probably be to characterize the environment of samples with high NV density.
+
+We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan which provides a significantly less noisy signal than the [111] scans for weak magnetic field.
+
+Finally we have seen an application of the CR to detect and characterize optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.
+
+There still remains a lot of open questions and there are several experiments that could be build upon these results. Some of these questions are listed below.
 
@@
 We estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$.
-This concentration is too high by about two orders of magnitude to isolate a single emitter with a diffraction-limited microscope.
+This concentration is about two orders of magnitude too high to contemplate isolating a single emitter with a diffraction-limited microscope.
 We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing.
 We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and optically isolate a single VH$^-$ center by scanning a confocal microscope with a fixed $B_{[100]}=56\ \rm G$ while monitoring the NV PL or $T_1$.
-It is however unclear to me whether this detection scheme would be sensitive enough to detect single dark spins.
-
-\printbibliography
-\end{refsection}
- \begin{refsection}
+It is however still unclear whether the present detection scheme is sensitive enough to detect single VH$^-$.
 
 \chapter{NV-NV cross-relaxations : the fluctuator model}
-In this chapter we will look at the phenomenon of cross-relaxations between NV centers, both from an experimental and theoretical point of view.
+This chapter will cover cross-relaxations between NV centers in the non-zero magnetic field regime, both from an experimental and theoretical point of view.
+The case of zero magnetic field will be treated in the next chapter.
 
@@
 
-We will see here how a model of fast and slow relaxing centers, the fluctuator model \citep{choi2017depolarization}, can explain most of the experimental observations, and we will see some limitations of the model.
+We will see in this chapter how a model of fast and slow relaxing centers, the fluctuator model \citep{choi2017depolarization}, explains most observations related to the cross-relaxation, including the stretched exponential lifetime profile and the spectral broadening of the interaction.
 
 \section{Experimental observation of NV-NV CR}
-Before we discuss the theoretical implications of NV-NV cross-relaxations (CR), let us first show the unambiguous experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
-\subsection{NV-NV CR between nonequivalent NV centers}
+Let us start with the experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
+
+When dealing with NV-NV CR, we have to make the distinction between what we will label ``equivalent" and ``non-equivalent" NV centers.
+Two equivalent centers are NV centers with the same spin Hamiltonian ([REF] chap 1), which implies that they both see the same transverse and longitudinal magnetic field.
+In contrast, non-equivalent centers are center with different single-spin Hamiltonian, which may result in different eigenstates and spin polarization for the two spins.
+
+While the NV physics can explain cross-relaxations between non-equivalent NV centers, it cannot explain cross-relaxation between equivalent NV centers.
+Nevertheless, as we will see in this section, CR between both equivalent and non-equivalent NV centers is observed for samples with high NV density.
+\subsection{NV-NV CR between non-equivalent NV centers}
 \label{non_equi_valent_CR}
@@
 \includegraphics[width=\textwidth]{chapter 3/Figures/NV-NV_non_equivalent}
-\caption{a) Representation of two non equivalent spins resonantly coupled.
-One spin is aligned with the magnetic field and therefore well polarized, while the second spin is not.
-b) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
-c) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
-\label{non equivalent NV-NV}
-\end{figure}
-NV-NV CR was first observed more than thirty years ago \citep{holliday1989optical, van1989cross}.
-The first observations were between non-equivalent NV centers, meaning that the two NV centers involved in the dipole-dipole coupling were not polarized equally.
-
-This scenario can happen for instance when two NV centers from different classes see a different transverse (and longitudinal) magnetic field.
-Fig \ref{non equivalent NV-NV} illustrates this in the case where the magnetic field is parallel with one of the four classes: $\mathbf{B} \parallel [111]$
-
-When $\mathbf{B} \parallel [111]$, as we discussed in the last chapter, one class sees no transverse field and is therefore always polarized.
-The three other (equivalent) classes on the other hand get more and more depolarized as the magnetic field amplitude is increased.
-
-It turns out that there is a co-resonnance at B=592 G between the class parallel to $\mathbf{B}$ and the three other classes.
-This co-resonance is represented in Fig.
-\ref{non equivalent NV-NV}-b) by an orange circle.
-
-Fig.
-\ref{non equivalent NV-NV}-c), which we already saw in the last chapter, shows the change in PL of an ensemble of NV centers as $\mathbf{B}$ is scanned along the [111] axis.
-For B=592 G, we can see a drop in PL also circled in orange.
-This drop is the result of the CR between NV centers from the class parallel to $\mathbf{B}$ and NV centers from the three other classes.
-
-While a difference in polarization is needed to observe a population transfer, it is not enough to explain the drop in PL.
-This PL drop is due to the difference in brightness between the $\ket{0}$ and $\ket{+1}$ from the different classes involved.
-The PL contrast between the two states is higher for the class parallel to $\mathbf{B}$ than for the three other ones, which is another consequence of the transverse field.
-
-The co-resonance between two different classes of NV centers with different magnetic field projection is a relatively rare event: it can only occur for the $\ket{0} \to \ket{+1}$ transition, and only for magnetic fields greater than 592 G.
-
-%techniquement y'a les resonances 0->-1 et -1/+1 mais c'est le delbor, j'ai pas envie d'embrouiller plus que ca.
+\caption{a) Illustration of the four possible direction of the the NV axis (four classes of NV centers) with a magnetic field along class 1 b) Representation of the polarization of two non equivalent spins, one aligned with the magnetic field (well polarized) and one misaligned with the magnetic field (unpolarized).
+c) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
+d) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
+\label{non-equivalent NV-NV}
+\end{figure}
+
+In the absence of magnetic gradient, two NV centers can only be non-equivalent if the projection of the magnetic field upon their respective NV axis is different.
+Fig.
+\ref{non-equivalent NV-NV}-a) illustrates such a case: we label here the four possible NV axis orientation (or NV classes) 1 to 4 and we assume that the magnetic field is aligned with class 1.
+In this conditions, an NV center belong to class 1 will always be non-equivalent with NV centers from classes 2-4, as the projection of the magnetic field on their respective axis will never be the same.
+On the other hand, NV centers from the classes 2-4 are all equivalents, even though they do not necessary belong to the same class.
+
+Fig.
+\ref{non-equivalent NV-NV}-b) shows the difference between the NV centers belonging to class 1 (labeled NV$_\parallel$) and those belonging to classes 2-4 (labeled NV$_\nparallel$).
+As explained in [REF], the transverse magnetic field leads to a mixing of the eigenstates \footnote{The states of the misaligned NV on Fig.
+\ref{non-equivalent NV-NV}-b) are labeled $\ket{0},\ket{\pm 1}$ which is an abusive (but common) notation since the mixing caused by the transverse field means that the eigenstates of $S_z$ are no longer the Hamiltonian eigenstates.} and to a loss of the optical polarization mechanism.
+As a result, the NV$_\nparallel$ spins are less polarized than the NV$_\parallel$ ones, which is a necessary condition to observe cross-relaxation.
+
+Another necessary condition to observe cross-relaxation is the condition of co-resonance.
+In most situations, non-equivalent NV centers are not co-resonant, as the differences in the spin Hamiltonian generally leads to different eigenvalues.
+This is in contrast with equivalent NV centers which are by definition always resonant.
+Fig.
+\ref{non-equivalent NV-NV}-c) shows the transition frequencies of spin from class 1 and a spin from class 2-4 as a function of the magnetic field amplitude.
+Unintuitively, there is a co-resonance between the two $\ket{0} \to \ket{+1}$ transitions for $\abs{\mathbf{B}}=592\ \rm G$.
+For this particular value of the magnetic field, cross-relaxation between the NV centers from class 1 and those from classes 2-4 is possible.
+
+Finally, Fig.
+\ref{non-equivalent NV-NV}-d), which was presented in the last chapter, shows the experimental value of the total NV PL as a function of the magnetic field amplitude, in a scenario similar to the one we just described.
+We can indeed observe the signature of NV-NV CR in the presence of a PL dip at $\abs{\mathbf{B}}=592\ \rm G$.
+
+As previously mentioned, cross-relaxation between non-equivalent NV centers is easily explained by the physics of single NV centers, and has been known for more than thirty years \citep{holliday1989optical, van1989cross}.
 
@@
 
-More recently, experiments \citep{jarmola2012temperature, mrozek2015longitudinal, jarmola2015longitudinal, choi2017depolarization} on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there were cross-relaxations even between equivalent NV centers.
-
-I refer by equivalent NV centers to NV centers with the same spin Hamiltonian.
-This means either NV center from the same class, or NV centers from different classes with the same projection of the magnetic field along their axis.
-Equivalent NV centers are by nature resonant with each other, and therefore susceptible to CR.
-
-Fig.
-\ref{equivalent NV-NV} shows an experimental observation of equivalent NV-NV CR.
-The key to observe NV-NV CR is to bring some NV centers in and out of resonance with other NV centers, which we can do with the different NV classes.
+More recently, experiments on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there were cross-relaxations even between equivalent NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, jarmola2015longitudinal, choi2017depolarization}.
+
+Fig.
+\ref{equivalent NV-NV}-a) illustrates the two possible scenarios for equivalent NV centers: either two NVs from the same class, or two NVs from different classes with the same projection of the magnetic field upon their respective axis (i.e.
+$\mathbf{B}$ belongs to the plane of mirror symmetry of the two spins).
+
+In the second case, by moving the magnetic field in and out of the symmetry plane, we can change the resonance condition between the two NV classes and measure the contribution of NV-NV cross-relaxation to the NV PL and spin lifetime.
 To do so, we need to add an initial bias magnetic field in addition to the scanned magnetic field, as represented in Fig.
@@
 Fig.
-\ref{inhomogene} illustrates this case.
-Our model here consists of two NV centers resonantly coupled with two separate relaxation rates $\Gamma_1$ and $\Gamma_1^*$.
-To complete the model, we add the rate $\Gamma_{\rm las}$ which represent the optical pumping rate from the $\ket{+1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins.
-We only consider here the incoherent dynamics of the population, modeled with rates.
-We also only consider the $\ket{0}$ and $\ket{+1}$ states as the results would be the same for the $\ket{0}$ and $\ket{-1}$ states.
-
-By finding the steady state of these rate equations, we can compute the final population in the $\ket{0}$ for both spins, which we will assume to be proportional to the total PL.
-Fig.
-\ref{inhomogene}-b) shows this $\ket{0}$ population as a function of the flip-flop rate $\Gamma_{\rm dd}$ for various values of $\Gamma_1^*$.
-
-We can see that when $\Gamma_1^*=\Gamma_1$, as expected, the PL is not modified by the coupling of the two spins regardless of the coupling strength.
-When $\Gamma_1^*>\Gamma_1$ however, not only is the starting PL slightly lower since NV$_2$ is less polarized, but the PL now decreases as we increase the coupling strength.
-This effect, which is at the origin of the CR contrast, is stronger when the difference between $\Gamma_1$ and $\Gamma_1^*$ is important.
+\ref{inhomogene}-a) and b) illustrates this case.
+Our model here consists of two (resonantly coupled) NV centers with two different relaxation rates $\Gamma_1$ and $\Gamma_1'$.
+To complete the model, we add the rate $\Gamma_{\rm las}$ which is the optical pumping rate from the $\ket{-1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins \footnote{We consider here a two level system instead of the full three levels of the spin 1.
+The choice of $\ket{-1}$ instead of $\ket{+1}$ is arbitrary.}.
+We only consider here the incoherent dynamics of the population, modeled with the rates reported on Fig.
+\ref{inhomogene}-a).
+Solving for the steady state of these rate equations, we can compute the final population in the $\ket{0}$ state for both spins.
+
+Fig.
+\ref{inhomogene}-b) shows the spin population in the $\ket{0}$ state of both spins, $\rho_{00}$, as a function of the flip-flop rate $\Gamma_{\rm dd}$ and for various values of $\Gamma_1'$.
+The first thing to notice is that, as $\Gamma_1'$ is increased,  $\rho_{00}$ decreases.
+This is because NV$_2$ becomes less polarized as its decay rate increases.
+The second thing to notice is that, when $\Gamma_1'=\Gamma_1$, i.e.
+when the two NV centers are truly equivalent, the flip-flop rate has no incidence on  $\rho_{00}$.
+However, when $\Gamma_1'>\Gamma_1$,  $\rho_{00}$ decreases when the flip-flop rate increases.
+Intuitively, this can be thought of as NV$_2$ acting as a polarization drain on NV$_1$.
+
+We can also notice that the influence of the flip-flop rate on $\rho_{00}$ increases when the difference in decay rates between NV$_1$ and NV$_2$ increases.
+As a result, we can suppose that a greater inhomogeneity between the NV centers would result in a stronger NV-NV cross-relaxation effect.
 
 \subsection{Presentation of the fluctuator model}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.8\textwidth]{chapter 3/Figures/NV_vs_fluctuator}
+\caption{Representation of a ``normal" NV center with spin depolarization rate $\Gamma_1 \ll \Gamma_{\rm las}$ and a fluctuator NV with spin depolarization rate $\gamma_f \gg \Gamma_{\rm las}$.
+Red circles represent the spin population in the steady state of each spin}
+\label{NV vs fluct}
+\end{figure}
 Choi et al.
-in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism.
-We are assuming that the fluctuator relaxation rate, $\gamma_f$ is much greater than the optical pumping rate $\Gamma_{\rm las}$.
-The fluctuators are therefore unpolarized and effectively act as dark spins, similarly to what we discussed in the last chapter.
+in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism leading to a new depolarization rate $\gamma_f$.
+Normal and fluctuator NV are represented in Fig.
+\ref{NV vs fluct}: in their model, Choi et al.
+assume that the fluctuator decay rate is fast enough compared to the laser polarization rate $\gamma_f \gg \Gamma_{\rm las}$ that the fluctuators are effectively always depolarized.
+Fluctuators are therefore similar to the dark spins we introduced in the last chapter.
 
@@
 
-We will consider the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
-With this assumption, we can compute the modification caused by the fluctuators on the rest of the NV lifetimes.
-
-Other possibilities to explain the change in the relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or superradiance, but they were not considered to be viable explanations \citep{choi2017depolarization}.
-%\begin{pmatrix}
-%0.33&0&0 \\
-%0&0.33&0 \\
-%0&0&0.33
-%\end{pmatrix}
+In their model, the authors of \citep{choi2017depolarization} consider that the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
+With this assumption, the modification caused by the fluctuator bath on the ``normal" NV lifetimes can be analytically computed.
+
+Other possibilities to explain the change in the relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or phonon superradiance, but they were not considered to be viable explanations: spin diffusion is orders of magnitude too slow to explain this phenomenon \citep{choi2017depolarization}, and the NV-NV CR seem to be independent of the crystal temperature \citep{jarmola2012temperature, mrozek2015longitudinal} which is incompatible with phonon superradiance \citep{choi2017depolarization}.
+
+Before detailing the model's predictions, we will first contemplate the possible microscopic origin of these fluctuators, and review some physical systems were similar observations have been made.
+
+ \subsection{Possible microscopic origin of the fluctuators}
+
+We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}.
+It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.
+
+This conclusion was based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions.
+Similar observations were performed with Förster resonant energy transfer (FRET) in chromophore molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through (electric) dipole-dipole interaction to non-excited acceptor molecules.
+
+In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
+In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
+Both of these phenomena can lead to fast spin relaxation: electron tunneling can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction term ($J-$coupling) by the crystal phonons can directly excite the spin transitions \citep{honig1960electron}.
+While the dipole-dipole interaction terms (excluding the contact interaction) are also modulated by the phonons, this effect would be too small to account for the fast depolarization observed here.
+
+Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
+To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate between neighboring sites of $\sim 10\ \rm ns$.
+In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
 
 \subsection{Single NV center coupled to a single fluctuator}
-We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
-To compute the depolarization induced by the fluctuator bath the NV centers, we should first consider the interaction between a single NV center and a fluctuator.
+We now turn to the computation of the depolarization induced by the fluctuator bath on the NV centers. We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
+
+Let us first consider the interaction between a single NV center and a fluctuator.
 Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.
 
-We will first decompose the dipole-dipole Hamiltonian in a radial and angular part:
+We will first decompose the dipole-dipole Hamiltonian as a product between radial and angular part:
 \begin{equation}
@@
 The difference comes here from the fact that we consider thermalization going both ways (i.e.
-$\ket{e} \to \ket{g}$ and $\ket{g} \to \ket{e}$ instead of only $\ket{e} \to \ket{g}$) and that each state of the fluctuator can decay into two other states.
-
-Finally, the additional depolarization rate induced by the fluctuator on the NV center reads:
-\begin{equation}
-\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right) \frac{\eta^2}{\gamma_f}.
+$\ket{e} \to \ket{g}$ and $\ket{g} \to \ket{e}$ instead of only $\ket{e} \to \ket{g}$) and that each state of the fluctuator can decay into two other states (see Fig.
+\ref{NV vs fluct}).
+
+Finally, one finds that the additional depolarization rate induced by the fluctuator on the NV center reads:
+\begin{equation}
+\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right)^2 \frac{\eta^2}{\gamma_f}.
 \end{equation}
@@
 
-We now need to compute the average depolarization caused by the ensemble of fluctuators on the ensemble of NV centers.
-
-We consider each fluctuator as in independent depolarization source, therefore the depolarization on a single NV center reads: 
+\begin{figure}[h]
+\centering
+\includegraphics[width=.4\textwidth]{chapter 3/Figures/NV_dans_bain_fluct}
+\caption{Schematics of an NV center in a fluctuator bath. The $\mathbf{r}_i$ vector represent the relative of position of fluctuator $i$ with respect to the NV center.}
+\label{NV + bain fluct}
+\end{figure}
+
+The next step is to compute the depolarization caused by the ensemble of fluctuators on a single NV center.
+
+We consider the situation pictured in Fig.
+\ref{NV + bain fluct}: the NV center is at the center of the frame and the position of each fluctuator is labeled $\mathbf{r}_i$.
+We consider that each fluctuator is an independent depolarization source, therefore the depolarization on a single NV center reads:
 \begin{equation}
@@
 
-In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential, unlike the phonon-induced depolarization which is purely exponential.
-
-It should be noted that the stretched exponential nature of the decay is not proper to the fluctuator model.
-\citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
-The stretched exponential is a result from localized noise sources, randomly distributed in the bulk (3D).
+In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential \footnote{Stretched exponential generally refer to functions of the form $\exp(-(t/\tau)^{\beta})$.
+In this manuscript, unless specified otherwise, we will always refer to the case $\beta=1/2$ when mentioning stretched exponential.}, unlike the phonon-induced depolarization which is purely exponential.
+
+It should be noted that the stretched exponential nature of the decay is not specific to the fluctuator model, but results from localized noise sources, randomly distributed in the bulk (3D).
+Indeed, \citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
+Theodor Förster in 1949 also predicted and observed stretched exponential dynamics in the context of FRET, where the same scaling laws are observed \citep{forster1949experimentelle}.
 
@@
 \label{sec computation eta}
+
 The parameter $\bar \eta$ in eq.
-\ref{eq 1/T} is of crucial importance since it encompass the dependence of $1/T$ in both the magnetic field and the inhomogeneous broadening $T_2^*$ of the NV centers and the fluctuators.
-
-We will start by separating the factor $\eta$ given in eq. \ref{eq eta} into a a geometric and a spectral part:
+(\ref{eq 1/T}) is of crucial importance as it encapsulates the resonance condition between the NV centers and the fluctuators.
+This resonance condition, i.e.
+the spectral overlap between the fluctuators and the NV centers, can be decomposed in two essential parameters: one is the central frequency of each subgroup ($\omega_{NV}^0$ and $\omega_{f}^0$ respectively) which themselves depends on the magnetic field, the other is the spectral broadening of both the NV and the fluctuators.
+
+The aim of this section is to give an explicit formula of $\bar \eta$ as a function of $\omega_{NV}^0$, $\omega_{f}^0$ and $T_2^*$.
+
+We will start by separating the factor $\eta$ given in eq. (\ref{eq eta}) into a a geometric and a spectral part:
 \begin{align}
-\eta^2&=G^2(\psi_{NV},\psi_f,\theta,\phi) R^2(\omega_f,\omega_{NV}), \\
+\eta^2&=G^2(\{\theta\}) R^2(\omega_f,\omega_{NV}), \\
 G^2&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
@@
 
-$G$ is a function which depends purely on the various angles of the problem (angles of the relative position of the NV center and the fluctuator, angle of the NV and fluctuator $z$ axis), while $R$ represents the resonance condition between the NV and the fluctuator.
-
-We will now average this $\eta$ factor for an ensemble of fluctuators and NV centers:
+$G$ is a function which depends purely on the various angles of the problem $\{\theta\}$, while $R$ represents the resonance condition between the NV and the fluctuator.
+
+We then average the $\eta$ factor for an ensemble of fluctuators and NV centers:
 \begin{align*}
-\bar \eta &= \int \rm{Prob}(\eta) \abs{\eta} d\eta, \\
-&= \left(\int \rho(\psi_{NV},\psi_f,\theta,\phi) |G(\psi_{NV},\psi_f,\theta,\phi)| d\psi_{NV} d\psi_f d\theta d\phi \right) \\
-&\, \times \left( \int \rho(\omega_{NV}) \rho(\omega_{f}) |R(\omega_f,\omega_{NV})| d\omega_{NV} d\omega_{f} \right), \\
+\bar \eta &= \int \rm{Prob}(\eta) \abs{\eta} d\eta \\
+&= \left(\int \rho(\{\theta\}) |G(\{\theta\})| d\{\theta\} \right) \\
+&\, \times \left( \int \rho(\omega_{NV}) \rho(\omega_{f}) |R(\omega_f,\omega_{NV})| d\omega_{NV} d\omega_{f} \right) \\
 &= \bar G \cdot \bar R,
 \end{align*}
-where we introduced $\bar G$ and $\bar R$, the averaging of $|G|$ and $|R|$.
+where we introduced $\rho(\{\theta\})$ the probability distribution of the angles $\{\theta\}$, $\rho(\omega_{NV})$ the inhomogeneous spectral distribution of the ensemble of NV centers and $\rho(\omega_{f})$ the inhomogeneous spectral distribution of the ensemble of fluctuators.
+We also introduced the notations $\bar G$ and $\bar R$ as the average of $|G|$ and $|R|$.
 
@@
 \label{eq 1/T avec T2*}
-\frac{1}{T} \propto \frac{1}{2\gamma_f + \Gamma_f + \Gamma_{NV}}.
-\end{equation}
-
-If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\overline{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian ans a Gaussian, which does not have a clear analytic solution.
+\frac{1}{T} \propto \frac{1}{2\gamma_f + \Gamma_f + \Gamma_{NV}}\cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
+\end{equation}
+
+If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\overline{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian and a Gaussian, which does not have a clear analytic solution.
 
 Eq.
-\ref{eq 1/T avec T2*} is subject to several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
-Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$.
-
- \subsection{Possible microscopic origin of the fluctuators}
-
-We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}.
-It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.
-
-This conclusion was based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions, and also to the resonant energy transfer between molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through dipole-dipole interaction to non-excited acceptors, and where the stretched exponential dynamics of the donors was first observed \citep{forster1949experimentelle}.
-
-In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
-In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
-Both of these phenomena could lead to fast spin relaxation, the tunneling because the spin hopping can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction strength $J$ by the crystal phonons \citep{honig1960electron}.
-While the non-contact dipole-dipole interaction is also modulated by the phonons, this effect would be too small to account for the fast depolarization observed here.
-
-Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
-To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate between neighboring sites of $\sim 10\ \rm ns$.
-In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
+\ref{eq 1/T avec T2*} was obtained via several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
+Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$, $\omega_{NV}^0$ and $\omega_{f}^0$, which will be crucial to compare the predictions of the model with experimental data.
 
@@
 \caption{T1 measurement following the protocol described in [REF].
-The best exponential and stretched exponential fits are given each time.
-a) Sample ADM-15-2 with B=0.
+The best exponential and stretched exponential fits are given in every case.
+a) Sample CVD-pink with B$\neq$0 and a single  probed class.
+The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$.
+b) Same measurement as a) in log scale.
+c) Sample ADM-15-2 with B=0.
 The optimal $T_1$ values are $T_1^{\rm stretch}=150\ \mu \rm s$ and $T_1^{\rm exp}=410\ \mu \rm s$.
-b) Sample CVD-pink with B$\neq$0 and a single class probed.
-The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$}
+d) Same measurement as c) in log scale.}
 \label{stretch_or_not_stretch}
 \end{figure}
-$T_1$ measurements with dense ensemble of NV centers and when many classes are resonant with each other (typically when B=0) indeed tend to show stretched exponential profiles.
-
-Fig.
-\ref{stretch_or_not_stretch} shows an example of that: in Fig.
-\ref{stretch_or_not_stretch}-a) the $T_1$ measurement comes from sample ADM-15-2 at B=0.
-The profile is clearly better fitted by a stretched exponential than a regular exponential, and the timescale found $T_1^{\rm stretch}=150\ \mu \rm s$ is much shorter than the expected phonon limite lifetime of a few ms.
-
-On the other hand, Fig.
-\ref{stretch_or_not_stretch}-b) shows a $T_1$ measurement coming from sample CVD-pink for B$\neq$0.
-This time the profile is more exponential, and the value $T_1^{\rm exp}=4.1\ \rm ms$ is coherent with a phonon-limited lifetime.
-Even though the sample CVD-pink shows some NV-NV CR feature, the spin dynamics in this sample is still dominated by the phonons, which explains the exponential profile.
+
+The first prediction of the fluctuator model which can be experimentally verified is the stretched-exponential profile of the spin relaxation.
+The protocol to measure spin relaxation with a common mode rejection was presented in [REF].
+
+Fig.
+\ref{stretch_or_not_stretch} shows the results of this protocol on two distinct samples and magnetic field configurations.
+In both cases, we show the results in linear and logarithmic scale as the log scale makes deviations from the standard exponential more apparent\footnote{The fits presented in Fig.
+\ref{stretch_or_not_stretch} and in the next figures were performed by fitting the real value of $S_1-S_2$ and not $\log(S_1-S_2)$.}.
+
+The spin relaxation results presented in Fig.
+\ref{stretch_or_not_stretch}-a) and b) was performed on sample CVD-pink ([REF]), which despite a relatively high NV concentration shows only moderate amount of dipole-dipole spin relaxation.
+In order to further minimize the amount of cross-relaxations, we chose a magnetic field for which all four NV classes are spectrally isolated, so that only fluctuators from the same class as the NV center could lead to relaxation.
+As a result, we can see a relaxation profile mostly exponential with a lifetime $T_1=4.1\ \rm ms$ comparable to the phonon-limited $T_1$ value of $\sim 5\ \rm ms$ at room temperature \citep{jarmola2012temperature}.
+
+The second spin relaxation presented in Fig.
+\ref{stretch_or_not_stretch}-c) and d) was performed on sample ADM-15-2 which showed a great amount of dipole-dipole spin relaxation.
+To further enhance the cross-relaxation rate, the experiment was performed in zero external magnetic field so that every class of NV centers would be resonant.
+We can observe a relaxation profile that deviates significantly from tthe exponential profile, and which is nicely fitted by a stretched exponential.
+The $T_1$ value obtained from the fit is also an order of magnetite lower than the phonon-limited value, which is coherent with the fact that dipole-dipole interactions dominate the spin relaxation in this case.
 
@@
 \caption{Fitting parameters for Fig.
- \ref{various_fit_formulae}}
+ \ref{various_fit_formulae}. The coefficient of determination $R^2$ for each fit is reported here.}
  \label{fitting table}
-\begin{tabular}{c|cc|cc}
+\begin{tabular}{c|ccc|ccc}
 \toprule
-Figure &  \multicolumn{2}{c}{$B=0$} & \multicolumn{2}{c}{$B\neq0$}\\
+Figure &  \multicolumn{3}{c}{$B=0$} & \multicolumn{3}{c}{$B\neq0$}\\
 \midrule
-{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms)&  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)\\
-Fig. a) & 0.95 & * & 2.64 & * \\
-Fig. b) & * & 0.32 & * & 1.03  \\
-Fig. c) & 6.75 & 0.44 & 4.22 & 7.19 \\
-Fig. d) & \textbf{5} & 0.51 & \textbf{5} & 4.65 \\
+{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms) & $1-R^2$ &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)& $1-R^2$\\
+\midrule
+Fig. a) & 0.95 & * & $2\cdot 10^{-2}$ & 2.64 & * & $6\cdot 10^{-3}$ \\
+Fig. b) & * & 0.32 & $2\cdot 10^{-3}$& * & 1.03 & $2\cdot 10^{-2}$ \\
+Fig. c) & 6.75 & 0.44 & $6\cdot 10^{-4}$ & 4.22 & 7.19 & $2\cdot 10^{-4}$\\
+Fig. d) & \textbf{5} & 0.51 & $6\cdot 10^{-4}$ & \textbf{5} & 4.65 & $6\cdot 10^{-4}$\\
 \bottomrule
@@
 
-For many samples however, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$ is of the same order as the phonon limited lifetime $T_1^{\rm ph}$.
-Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, forcing us to include both aspects if we want a unique fitting formula.
-
-Fig.
-\ref{various_fit_formulae} shows four different fitting procedure on the two same experimental data.
-The data here are two $T_1$ measurements done on the same sample ADM-150-1, following the protocol described in [REF], with and without an external magnetic field ($\sim 50 G$).
+For many samples, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$, is of the same order of magnitude as the phonon limited lifetime $T_1^{\rm ph}$.
+Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, meaning that we have to include both aspects in the fitting formula.
+
+Fig.
+\ref{various_fit_formulae} shows four different fitting procedure on the two same relaxation measurements.
+The measurements were performed on the same sample ADM-150-1, with and without an external magnetic field ($\sim 50 G$).
 The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.
@@
 a) Simulation of the $\ket{0} \to \ket{-1}$ transition frequency for the four classes of NV centers and actual frequencies of the two central classes from ODMR measurements.
-b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.78\ \rm MHz$.
-The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor $\sqrt{2}$}
+b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.8 \pm 0.2\ \rm MHz$.
+The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor of 2}
 \label{fluct linewidth}
@@
 
-One of the main arguments in favor of the fluctuator model is the measurement of the fluctuators spectral response.
-Indeed, as we will see, the fluctuator lifetime is so short that it significantly broadens the fluctuators spectral response compared to standard NV centers.
-
-The fluctuator are effectively dark to ODMR measurement since they are not polarized, but we can measure their linewidth through CR the same way we would do for a dark spin, as detailed in \citep{hall2016detection}.
+We now move to the second major prediction of the fluctuator model which is the spectral broadening of the fluctuators caused by the their relaxation rate $\gamma_f$.
+Indeed, following the arguments presented in \citep{choi2017depolarization}, the fluctuators homogeneous spectral response is a Lorentzian of half-width $2\gamma_f$.
+The inhomogeneous spectral response of the ensemble of fluctuators is then further broadened by the same electric, magnetic or strain noises than the NV centers, which is characterized by $1/T_2^*$.
+
+The fluctuator are effectively dark to ODMR measurement since they are not polarized, meaning that we cannot directly measure their linewidth as we would do for normal NV centers.
+We can however measure their linewidth through cross-relaxation, similarly to the detection of dark spins presented in the previous chapter.
 
@@
 \ref{equivalent NV-NV}.
-A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers, but this time measuring the spin decay rate instead of the PL.
-
-Fig.
-\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field, and the experimentally measured central frequencies of the two classes of interest when the two lines can be clearly separated.
-Having a precise measurement of the detuning between the two classes $\Delta \nu$ is key in this experiment to get a precise value of the fluctuator linewidth.
+A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers.
+This time however, the spin relaxation was measured instead of the PL.
+
+Fig.
+\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
+We took particular care in measuring the central frequencies of the two classes of interest (when the two lines could be clearly separated), since having a precise measurement of the detuning between the two classes $\Delta \nu$ is key to get a precise value of the fluctuator linewidth.
 %Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of their two central lines to get a precise measurement of the Larmor frequency.
@@
 We can clearly see that the $\Gamma^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
-If we assume that both NV classes spectral response are gaussians of same width $\sigma$, then the spectral overlap between the two classes is also a gaussian of width $\sqrt{2} \sigma$.
-Fig.
-\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $\sqrt{2}$, which is still much narrower than the $1/T_1^{\rm dd}$ line.
+If we assume that both NV classes spectral response are Lorentzian of same width $\sigma$, then the spectral overlap between the two classes is also a Lorentzian of width $2 \sigma$ \footnote{For other line profile, the overlap would be smaller than $2 \sigma$.
+For instances, the overlap of two Gaussians of width $\sigma$ is a Gaussian of width $\sqrt{2} \sigma$}.
+Fig.
+\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $2$, which is still significantly narrower than the $1/T_1^{\rm dd}$ line.
 
 We attribute this broadening to the fluctuators lifetime.
-Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator, but for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening that we observe.
+Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator, but for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening observed here.
 
@@
 \ref{fluct linewidth}.
-Red curve: deconvolution of the blue curve the  by the ODMR line shown on the same figure.
-Green curve: deconvolution of the red curve the  by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths 8.78, 7.49 and 6.38 MHz}
+Red curve: deconvolution of the blue curve by the ODMR line shown on the same figure.
+Green curve: deconvolution of the red curve   by the ODMR line.
+All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters)}
 \label{deconvolution}
@@
 
-With $f=8.78\ \rm MHz$ and $\sqrt{2}f_G=4.33\ \rm MHz$, we find $f_L=6.54\ \rm MHz$ for a final lifetime value $T_1^f = 48.7\ \rm ns$, a value consistent with the one found with the deconvolution method.
-
-Doing similar calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48.2\ \rm ns$, a value surprisingly similar to the one we got.
+With $f=8.8 \pm 0.2\ \rm MHz$ and $\sqrt{2}f_G=4.3 \pm 0.2\ \rm MHz$, we find $f_L=6.5 \pm 0.3\ \rm MHz$ for a final lifetime value $T_1^f = 49 \pm 2\ \rm ns$, a value consistent with the one found with the deconvolution method.
+
+Doing similar experiments and calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48\ \rm ns$, a value very close to the ones we obtained above, despite using a completely different sample.
+
+In conclusion, we observed that the fluctuators linewidth was indeed broadened by their relaxation rate which justifies the initial hypothesis that $T_1^f \ll T_1^{NV}$.
+It even shows that $T_1^f$ is of the same order of magnitude as the optical lifetime ($\sim 15\ \rm ns$) which explains why the fluctuators cannot be optically polarized.
 
@@
 
-What we decided to do was to take the case of an isolated class, where we consider that NV centers from said class or only affected by fluctuators of the same class, as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2,3 or 4 classes are co-resonant as a function of $\Gamma_0$.
-Doing so allows to normalize $n_f$ and $\gamma_f$ in eq.
+We decided to take the case of an isolated class as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2, 3 or 4 classes are co-resonant as a function of $\Gamma_0$.
+Doing so allows us to normalize $n_f$ and $\gamma_f$ in eq.
 \ref{eq 1/T} and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
@@
 
-For the single isolated class, the $T_1$ profile seems to be closer to a stretch exponential, which is coherent with the fact that the lifetime $T_1 \sim 1\ \rm ms$ is already smaller than the expected phonon limited lifetime $T_1^{\rm ph} \sim 5\ \rm ms$.
-However when $\mathbf{B}=0$, although the lifetime gets $\sim$ 10 times shorter, the profile becomes clearly exponential.
-
-The spin dynamics in this sample is still dominated by resonant dipole-dipole interactions: the more classes are brought to resonance, the shorter the spin $T_1$.
-But here, the profile gets closer to a pure exponential when $T_1$ decreases, which is in complete contradiction with the fluctuator model.
-
-This particular sample was a 15 $\mu$m micro-diamond, which could have a role in this observation since we know that spins near the surface have a reduced $T_1$ \citep{rosskopf2014investigation}, and micro-diamonds have proportionally more near surface spins.
-But other diamonds from the same batch, such as the one measured in Fig.
-\ref{stretch_or_not_stretch}, show clearly stretched exponential profile in the same conditions, and meanwhile some bulk diamonds show very short non-stretched exponential lifetimes, similar to the present sample.
+Although the spin lifetime still gets $\sim$ 10 times shorter in the $\mathbf{B}=0$ case, indicating that dipole-dipole interaction still plays an important role in the spin relaxation, this time the lifetime profile does not converge to a purely stretched exponential.
+In fact it seems to be the opposite: the lifetime profile is closer to a pure exponential for the shorter $T_1$ (especially for the very short times), and closer to a stretched exponential for the longer $T_1$, which is in complete contradiction with the fluctuator model and the previous observations in Fig.
+\ref{stretch_or_not_stretch} and \ref{various_fit_formulae}.
+
+We should note that most sample studied follow the fluctuator model, in that the their lifetime profile gets closer to a stretched exponential when the dipole-dipole relaxation rate is increased.
+Sample ADM-15-4 seem to be an anomaly in this regard, since most of the samples from the same batch (including sample ADM-15-2 in Fig.
+\ref{stretch_or_not_stretch}) seem to follow the ``normal" behavior predicted by the fluctuator model.
+However, some other samples, including bulk HPHT samples and 150 $\mu$m Adamas micro-diamond also showed short (non stretched) exponential lifetimes.
 
 It seems therefore that the current fluctuator model does not correctly describe the spin dynamics of every NV-dense samples.
-Looking at what was previously done for P-doped SI \citep{vugmeister1978spin} or for resonant energy transfer between molecules \citep{yokota1967effects}, the fluctuator model could probably be improved by taking into account the finite lifetime of the fluctuators (i.e.
+Looking at what was previously done for P-doped SI \citep{vugmeister1978spin} or for FRET \citep{yokota1967effects}, the fluctuator model could probably be improved by taking into account the finite lifetime of the fluctuators (i.e.
 the saturation of the NV-fluctuator CR), and the spin diffusion between NV centers (the fluctuator model only considers NV-fluctator interactions and not NV-NV interactions).
@@
 In particular it would be interesting to know whether this profile is always Lorentzian, and what its linewidth is.
-In the sample I studied on Fig.
-\ref{fluct linewidth}, I found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough my sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 45 ppm$ sample used in \citep{choi2017depolarization}.
+In the sample studied on Fig.
+\ref{fluct linewidth}, we found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough our sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 15 ppm$ sample used in \citep{choi2017depolarization} \footnote{Although the initial paper claimed a NV concentration $[\rm NV] \sim 45 ppm$, a more recent paper re-established the concentration at $\sim 15$ ppm \citep{zhou2020quantum}}.
 If $\gamma_f$ was found to be somewhat of a constant among all samples showing NV-NV CR, this would give us a lot of information on the potential nature of the fluctuators.
@@
 
-\printbibliography
-\end{refsection}
- \begin{refsection}
-
-\chapter{NV-NV CR under transverse or low fields: application to magnetometry}
-
-In the last chapter we saw how dense ensemble of NV centers have an intrinsic depolarization mechanism mediated by the NV-NV dipole interaction.
-In this chapter we will see how this feature can be exploited to perform magnetometry.
-
-Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
-We will focus here on the low to zero magnetic field region, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the more complex physics of the NV centers under low magnetic field.
-
-This chapter is constructed as follows: We will first describe the physics of NV centers under low or transverse magnetic field (the physics is similar in both cases), we will then show both theoretically and experimentally that the NV-NV CR is modified for low and transverse magnetic field.
-We will identify different mechanisms by which the relaxation is modified and we will try to quantify their relative importance.
+\chapter{NV-NV cross-relaxation at low fields: application to magnetometry}
+
+In the last chapter we have identified a dipole-dipole relaxation mechanism within dense NV centers ensemble, and we have studied its influence in non-zero magnetic field.
+In this chapter we will cover the low to zero magnetic field region and see how the depolarization at low magnetic field can be used for magnetometry.
+
+This chapter is constructed as follows: We will first describe in sec. \ref{sec 4.1} the physics of NV centers under low or transverse magnetic field. In sec. \ref{sec causes zero field}
+we will present the different zero field depolarization causes that we have identified as well as their theoretical impact on the NV relaxation rates according to the fluctuator model.
+In sec.
+\ref{sec 4.3} we will confront the predictions of the fluctuator model with experiments under low and transverse magnetic field.
 
 We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field.
-We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols.
+We will start by giving a short overview of NV ensemble magnetometry in sec.
+\ref{sec 4.4} before characterizing the LFDM protocol in sec.
+\ref{sec LFDM} and comparing it to other magnetometry protocols in sec.
+\ref{sec 4.6}.
 Finally we will give some perspectives on how to improve LFDM and on potential applications.
@@
 \section{NV spin Hamiltonian under low and transverse fields}
-
-Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates compared to the high magnetic field regime.
+\label{sec 4.1}
+We will start this chapter by studying the single NV physics under low and transverse magnetic field.
+On of the main difference with the previously studied regime is that the eigenstates of the Hamiltonian are no longer $\{ \ket{0}, \ket{+1}, \ket{-1} \}$, which will then in turn modify the dipole-dipole interaction between the NV centers.
+
+We will first describe the spin Hamiltonian of the NV center under low magnetic field and determine the dominant factors in our samples, then do the same for transverse magnetic field, and finally investigate the modification of the ODMR lineshapes in those regimes.
 
 \subsection{NV spin Hamiltonian in zero external magnetic field}
+\label{sec. Hamltonian 0 B}
 \begin{figure}[h]
@@
 \includegraphics[width=\textwidth]{chapter 4/Figures/ESR_simus}
-\caption{Simulations of inhomogeneous zero field ODMR when sampling various parameters.
-a) Simulation when sampling each components of the magnetic field over a Gaussian of deviation $\sigma=2\ \rm G$.
-b) Simulation when sampling each components of the strain tensor $\bar{\bar{\varepsilon}}$ over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
-c) Simulation when sampling each components of the electric field over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
+\caption{Monte-Carlo simulations of inhomogeneous zero field ODMR for different noise sources.
+a) Magnetic noise: each of the $\mathbf{B}$ components is sampled over a Gaussian of deviation $\sigma=2\ \rm G$.
+b) Strain noise: each of the components of the strain tensor $\bar{\bar{\varepsilon}}$ is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
+c) Electric noise: each of the $\mathbf{E}$ components is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
 d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
@@
 
-In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
-
-\subsection{Hyperfine coupling and inhomogeneous broadening}
+In conclusion, whenever the spin Hamiltonian is dominated by the local electric field or by the transverse magnetic field, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+
+\subsection{Hyperfine and inhomogeneous broadening}
 \label{sec modif T2*}
@@
 
-We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
+A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
+The energy of the $\ket{\pm 1}$ states depends linearly on the longitudinal magnetic field, and quadratically on the transverse electric and magnetic fields.
+On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and quadratically on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp$.}.
+
+The shape of the ODMR lines (i.e.
+the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis, as we established in the previous part.
+Further details on the low magnetic field regime are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.
+
+Fig.
+\ref{ESR for T2*} shows ODMR profile on two samples, CVD-pink which shows weak dipole-induced depolarization and Sumi-2 which shows strong dipole-induced depolarization, in three distinct regime: for a purely longitudinal magnetic field, for a purely transverse magnetic field, and without magnetic field.
+The linewidths of the spin transition \footnote{We consider here the full linewidth of the electronic levels, i.e.
+including all three hyper fine levels} for both samples are reported in the Table \ref{ESR for T2*}-d).
+In both cases, the linewidths are smaller in the $B=0$ and $B \perp NV$ cases than in the $B \parallel NV$ case, mainly due to the modification of the hyper-fine coupling.
+The linewidth modification reaches up to $\sim 50\ \%$ for the CVD sample and $\sim 20\ \%$ for the HPHT one.
+
+We are particularly interested in the spin transition linewidth since it directly impacts the dipole-dipole relaxation rate (as seen in sec.
+[REF]): the smaller the linewidth, the higher the dipole-dipole relaxation rate is.
+
+%We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
 These change are relevant to our study of NV-NV CR due to the relation between the dipole-induced relaxation rate and $T_2^*$ detailed in sec.
 [REF].
-
-A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
+%
+%A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
 In the $\{ \ket{0}, \ket{\pm 1} \}$, the eigen energies are linearly sensitive to (longitudinal) magnetic field, and only sensitive to electric fields at the second order, and vice versa for the $\{ \ket{0}, \ket{\pm} \}$ basis.
-
-These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei.
+%
+%These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei.
 These effects can drastically modify the ODMR lineshape in zero external magnetic field \citep{jamonneau2016competition} or purely transverse magnetic field \citep{qiu2021nuclear,qiu2022nanoscale}.
-
-We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center.
+%
+%We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center.
 $^{14}$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin.
 The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
-\begin{equation}
-\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
-\end{equation}
-
-where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor.
+%\begin{equation}
+%\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
+%\end{equation}
+%
+%where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor.
 $\mathcal{H}_n$ and $\bar{\bar{A}}$ can be written:
-
-\begin{equation}
-\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
-\end{equation}
-
-and
-\begin{equation}
-\bar{\bar{A}} = \begin{pmatrix}
-A_{xx} & 0 & 0 \\
-0 & A_{yy} & 0 \\
-0 & 0 & A_{zz}
-\end{pmatrix},
-\end{equation}
-
-where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.
-
-Fig.
+%
+%\begin{equation}
+%\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
+%\end{equation}
+%
+%and
+%\begin{equation}
+%\bar{\bar{A}} = \begin{pmatrix}
+%A_{xx} & 0 & 0 \\
+%0 & A_{yy} & 0 \\
+%0 & 0 & A_{zz}
+%\end{pmatrix},
+%\end{equation}
+%
+%
+%where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.
+%
+%Fig.
 \ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$.
-
-While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers.
+%
+%While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers.
 These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.
-
-Fig.
+%
+%Fig.
 \ref{ESR for T2*}-a) allows us to evaluate the magnetic field noise in both samples, since the $\ket{\pm 1}$ states are not sensitive to weak electric fields.
@@
 We should therefore expect the modification of $T_2^*$ to have a moderate impact on the depolarization dynamics.
-
-\medskip
-
-To summarize this section, we have seen that (for the samples used in this chapter) the NV centers spin Hamiltonian is dominated by transverse electric field for low external magnetic field, and that it causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
-We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples with a high impurity density.
-
-\section{NV-NV CR in the low magnetic field regime}
-In this part, we will study the cross-relaxation between NV centers under low magnetic field, whose spin Hamiltonian is dominated by a transverse electric field.
-Understanding the various mechanisms operating in this regime will be important for the magnetometry protocol presented after.
-
-\subsection{Experimental observations}
+%
+\bigskip
+
+To summarize this section, we have seen that (for the samples used in this chapter) the NV center spin Hamiltonian is dominated by transverse electric field at low external magnetic field, which causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
+
+We have also seen that the spin Hamiltonian has the same eigenstates for purely transverse magnetic field than it does for low magnetic field.
+
+Finally we have seen that this change of eigenstates occasions a modification of the spectral profile and linewidth of the spins.
+
+\section{Theory of the low field depolarization}
+\label{sec causes zero field}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/Shema_summary_theory_2}
+\caption{Summary of the different depolarization effects in low magnetic field and their predicted depolarization rate. The ODMR spectra were taken on sample ADM-150-1}
+\label{summary_theory}
+\end{figure}
+
+We will focus on this section on the mechanisms responsible for the depolarization of dense NV ensemble at low magnetic fields.
+We have identified four such mechanisms that lead to an increase in the resonant dipole-dipole coupling between NV and fluctuators, which in turn results in an increased spin depolarization rate.
+
+Fig.
+\ref{summary_theory} summarizes these four mechanisms along with the theoretical increase in the depolarization rate caused by each of them.
+We will detail below the physical origin of these four mechanisms and the calculations that lead to the values presented in Fig.
+\ref{summary_theory}.
+
+We should note that the values indicated in Fig.
+\ref{summary_theory} for the double flips and the $T_2^*$ contributions depend on experimental parameters, while the ones for the 4-classes degeneracy and the eigenbasis change are purely based on theory.
+
+\subsection{Four classes degeneracy}
+
+The first and main contribution to the low field depolarization is the degeneracy between the four classes as the magnetic field decreases.
+Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 43$.
+This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.
+
+\subsection{Eigenstates modification}
+
+The modification of the Hamiltonian eigenstates discussed in sec.
+\ref{sec.
+Hamltonian 0 B} from the eigenbasis $\{ \ket{0}, \ket{+1},\ket{-1} \}$ when $B\neq 0$ to the eigenbasis $\{ \ket{0}, \ket{+},\ket{-} \}$ when $B\sim 0$ changes the dipole-dipole interaction between the two spins.
+
+A consequence of this change is the absence of diagonal terms  in the dipole-dipole Hamiltonian written in the $\{ \ket{0}, \ket{+},\ket{-} \} \times \{ \ket{0}, \ket{+},\ket{-} \}$ basis, which explains why the NV centers states are protected from magnetic noises at low magnetic field \citep{jamonneau2016competition}.
+
+Another consequence is a slight modification of the flip-flop rates.
+Once averaged over all possible orientation and position, as detailed in appendix [REF], this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
+%The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
+%
+%This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
+We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively.
+To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
+%
+%The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF].
+The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane.
+This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.
+%
+%We therefore need to make an assumption on the distribution of the transverse field in the sample.
+In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane.
+However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq.
+[REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.
+%
+%We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
+%
+%We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$.
+$\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
+%
+%In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$.
+The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig.
+\ref{scan 100}
+
+\subsection{Double flips}
+
+The double flip processes have been previously discussed in sec.
+[REF] and naturally arise from the dipole-dipole Hamiltonian.
+The ones that we consider here couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$).
+These terms are generally not resonant because of the Zeeman splitting between the $\ket{+1}$ and $\ket{-1}$ states.
+
+In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$.
+These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely suppressed.
+
+The computation of the depolarization rate including the double flips is detailed in appendix [REF].
+To obtain a numerical value, we need to input the experimental values of the dipole-dipole interaction range \footnote{By interaction range we mean the value to be compared to the spectral detuning $\Delta \nu$ between the two groups.
+This is $2\gamma_f+\Gamma_{NV}+\Gamma_f$ in eq.
+[REF]} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+[REF], as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states.
+We typically measure $\Delta \nu \approx 8\ \rm MHz$ on type Ib HPHT samples.
+
+For the value presented in Fig. \ref{simus ESR}, we inputted $\Gamma=\Delta \nu=8.8\ \rm MHz$.
+%\begin{table}[htbp]
+%\centering
+%\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
+% \label{table double flips}
+%\begin{tabular}{c|cc}
+%%\toprule
+%{} & $l_c=0$ & $l_c=\infty$ \\
+%\midrule
+%$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
+%$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
+%$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
+%%\bottomrule
+%\end{tabular}
+%\end{table}
+%
+%Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
+The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
+%
+%We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field.
+And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field.
+But we can see from Fig.
+\ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig.
+[REF].
+In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.
+%
+%To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
+We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
+Finally, the question about the correlation  length of the electric field also needs to be asked.
+%
+%The results are summed up in Table \ref{table double flips}.
+We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
+$\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$).
+This is close to the the experimental values we have.
+Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $).
+We come back to the results of the last section where only considered the flip-flops.
+
+\subsection{$T_2^*$ increase}
+
+The last aspect we considered was the change in the ODMR lineshape discussed in sec. \ref{sec modif T2*} for weak magnetic field (we refer to the spectral linewidth as $T_2^*$ for conciseness).
+
+To quantify the impact of the $T_2^*$ increase in the depolarization rate, we employ eq.
+[REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
+\ref{ESR for T2*}.
+We then find that the increase of $T_2^*$ in zero magnetic field occasions an increase of the decay rate by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
+
+The value shown in Fig. \ref{summary_theory} corresponds to a $9 \%$ increase.
+
+\bigskip
+
+In conclusion, we predict that the 4-classes degeneracy is by far the dominant causes for low field depolarization ($\sim \Gamma_1 \times 40$), followed by the double flips ($\sim \Gamma_1 \times 2$).
+We will however see in sec \ref{sec LFDM} that the double flips, along with the change of eigenbasis and increase in $T_2^*$ are determinant factor in the sensitivity of the low field magnetometry protocol.
+
+In the following section we will confront the theoretical results with experimental ones.
+
+\section{Experimental investigation of the low field depolarization}
+\label{sec 4.3}
+In this section we will try to experimentally discriminate the contribution of the four depolarization sources described in Fig.
+\ref{summary_theory} and confront the experimental values with the theoretical ones.
+
+We will first determine the contribution of the 4-classes degeneracy by comparing [100] aligned and non-[100] aligned magnetic field in the weak magnetic field regime, before studying the influence of the double-flips thanks to purely transverse magnetic fields.
+
+On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\mathbf{B}\parallel [100]$) and the influence of the laser polarization.
+Following further experiments detailed in appendix [REF], we have determined that these two effects were negligible compared to the four ones we identified before.
+
+\subsection{NV-NV CR in the low magnetic field regime}
+
+In this part, we will try to isolate the contribution of the 4-classes degeneracy from the three other ones by using the fact that all four classes can be at resonance for two distinct scenario:
+\begin{itemize}
+\item When $\mathbf{B}=0$.
+\item When $\mathbf{B}\parallel [100]$ (cf. Fig. [REF] (la figure avec toutes les config)).
+\end{itemize}
+
+\subsubsection{Magnetic scan in a random direction}
+
 \begin{figure}[h]
@@
 \end{figure}
-We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.
-
-The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust in order to isolate each effects.
-
-Fig.
-\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G.
-The ODMR at the initial and final magnetic fields are shown in Fig.
-\ref{scan 1x1x1x1}-a) and b).
-
-While it is clear that the spin lifetime as well as the PL increases with the magnetic field, there is no clear indication that this is because of the specificity of the low field region.
-Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.
-
-We can note however that the PL is not an exact mirror of the spin dynamics: while the low field drop in PL is indeed associated with a reduced spin lifetime, the drop in PL at higher magnetic field comes from the states mixing induced by the transverse magnetic field, and is not associated with a modification of the spin dynamics (to the first order).
+
+First we will run a control experiment by scanning the magnetic field in a non [100] direction.
+Fig.
+\ref{scan 1x1x1x1}-a) and b) shows ODMR spectrum at the starting point ($|\mathbf{B}|=0$) and end point ($|\mathbf{B}|\approx 60\ \rm G$) of the magnetic field scan.
+The magnetic field orientation was chosen in order to maximize the splitting of the four classes.
+
+Fig.
+\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as a function of the magnetic field.
+We can observe a decrease of $T_1^{\rm dd}$ by a factor of $\sim 10$ in zero magnetic field compared to non-zero magnetic field, with a half-width at half maximum of $\sim 10\ \rm G$.
+Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\mathbf{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$.
+This is because this drop of PL is associated with transverse magnetic field state mixing, and not with spin depolarization.}.
+
+We attribute the spin depolarization observed in Fig.
+\ref{scan 1x1x1x1}-d) to all four mechanisms described in the last part acting simultaneously.
+To determine the contribution of each mechanism individually, we will need to run dedicated experiments.
+
+\subsubsection{Magnetic scan with $\mathbf{B} \parallel [100]$}
 
@@
 
-Fig.
-\ref{scan 100} presents a way to circumvent this issue: by applying the magnetic field along the [100] crystalline axis, we can make sure that the four classes of NV centers always stay resonant regardless of the magnetic field amplitude.
-
-We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig.
-\ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of $\sim$ 8 in zero field, whereas in Fig.
-\ref{scan 100}, it was only reduced by a factor of $\sim$ 1.2.
-The main reason for the PL and $T_1$ drop in Fig.
-\ref{scan 1x1x1x1} was indeed the co-resonance between the four classes.
-
-Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained if we consider only the inter-class resonances.
-This means that there are some additional depolarization mechanisms which are proper to the zero field region.
-We should also note that, while the zero-field PL contrast is bigger in Fig.
-\ref{scan 1x1x1x1}-c) than in Fig.
-\ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually similar in both cases.
-
-We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]
-
-\subsection{Theory of the low field depolarization}
-\label{sec causes zero field}
-We have identified three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.
-
-\subsubsection{Eigenstates modification}
-The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
-
-This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
-We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively.
-To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
-
-The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF].
-The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane.
-This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.
-
-We therefore need to make an assumption on the distribution of the transverse field in the sample.
-In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane.
-However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq.
-[REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.
-
-We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
-
-We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$.
-$\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
-
-In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$.
-The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig.
-\ref{scan 100}
-
-\subsubsection{Double flips}
-
-\begin{table}[htbp]
-\centering
-\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
- \label{table double flips}
-\begin{tabular}{c|cc}
-%\toprule
-{} & $l_c=0$ & $l_c=\infty$ \\
-\midrule
-$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
-$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
-%\bottomrule
-\end{tabular}
-\end{table}
-
-Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
-The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
-
-We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field.
-And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field.
-But we can see from Fig.
-\ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig.
-[REF].
-In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.
-
-To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
-We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
-Finally, the question about the correlation  length of the electric field also needs to be asked.
-
-The results are summed up in Table \ref{table double flips}.
-We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
-$\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$).
-This is close to the the experimental values we have.
-Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $).
-We come back to the results of the last section where only considered the flip-flops.
-
-\subsubsection{$T_2^*$ modification}
-
-The last aspect we considered was the change in $T_2^*$ (we include the change in the hyper-fine interaction in the $T_2^*$) for weak magnetic field, as discussed in sec. \ref{sec modif T2*}.
-
-To quantify its impact in the depolarization rate, we employ eq.
-[REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
-\ref{ESR for T2*}.
-
-We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
-
-\subsubsection{Summary}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{chapter 4/Figures/Shema_summary_theory}
-\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
-\label{summary_theory}
-\end{figure}
-
-Fig. \ref{summary_theory} recapitulate the various mechanisms at play in the low field depolarization of NV ensemble.
-
-We start from from the simplest case where every classes are sepctrally isolated and the only dipole-induced relaxation coms from the flip-flop between NV and fluctuators from the same class.
-We then consider the case where all four classes are resonant ($B\parallel [100]$) and finally the case B=0.
-
-The three extra depolarization mechanisms in zero field are then detailed, with the predicted increase of depolarization rate for each step.
-The predicted rates in zero field depend on several parameters, we chose the following ones to compute the different numerical values: $l_c=0$, $\bar{G}^2=0.5$, $2\gamma_f = 6.5\ \rm MHz$ and $2/T_2^*=5.9\ \rm MHz$ (corresponding to sample Sumi-2 if we assume a Lorentzian ODMR profile).
-
-\subsection{Other potential causes for zero field depolarization}
-We have also considered two other potential causes that we have ultimately ruled out in this case.
-
-\subsubsection{Magnetic field alignment}
-\begin{figure}[h]
-\centering
-\includegraphics[width=.7\textwidth]{chapter 4/Figures/alignement_rose}
-\caption{PL of sample CVD-pink as a function of the magnetic field for various misalignment of the field direction with the [100] axis}
-\label{alignement rose}
-\end{figure}
-
-The first is a potential misalignment of the magnetic field with the [100] axis, which would cause a splitting of the 4 classes as the magnetic field increases.
-This hypothesis is not consistent with the observations in Fig.
-\ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased.
-Instead we observe a sharp decrease followed by a pseudo-plateau.
-
-We still estimated how sensitive the PL dip was with respect to the field alignment.
-Fig.
-\ref{alignement rose} shows PL scans on sample CVD-pink for various alignment of the magnetic field around the [100] axis.
-We estimate our alignments to be precise within $\pm 1^\circ$.
-We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization does not come from a field misalignment.
-
-\subsubsection{Laser polarization}
-\begin{figure}[h]
-\centering
-\includegraphics[width=.7\textwidth]{chapter 4/Figures/pola_laser}
-\caption{PL of sample ADM-150-3 with respect to the magnetic field for various laser polarization angle.
-The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
-\label{pola laser}
-\end{figure}
-
-Other studies \citep{anishchik2015low, filimonenko2020weak} previously saw a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.
-
-Fig.
-\ref{pola laser} shows PL scans on sample ADM-150-3 for various polarization angle of the incident laser.
-We chose to apply the magnetic field in the laser polarization plane, in a direction that did not match any particular crystalline planes.
-
-We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$.
-We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.
-
-\section{NV-NV CR under purely transverse magnetic field}
-We have seen in the last part that there was a spontaneous spin depolarization specifically in zero external magnetic field, and we have identified three potential mechanisms which are the double-flips, the change in the Hamiltonian eigenstates and the modification of $T_2^*$.
-We now want to experimentally discriminate the contribution of each of these mechanisms and to see how close the actual decay rates are to the predicted ones.
-
-One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic and electric noise ...).
-Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.
-
-The ideal way to isolate each contributions would be to apply an electric field strong enough to split the transition of the 4 classes, and to measure the decay rate of a single class dominated by the transverse electric field.
-Such an electric field would need to be of the order of $\sim 10^6\ \rm{V/cm}$, which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.
-
-Instead of an electric field, we will use a transverse magnetic field, which as described in sec.
-\ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
-
-\subsection{Principle of the experiment}
+This second experiment is similar to the last one except that the magnetic field was scanned in the [100] direction.
+Fig.
+\ref{scan 100}-a) and b) again show ODMR at the starting and ending point of the magnetic scan, while Fig.
+\ref{scan 100}-c) and d) show the PL and $1/T_1^{\rm dd}$ recorded during the scan.
+
+The main difference with the previous experiment is that in this case, the degeneracy between the four classes is never lifted.
+We therefore attribute the increase in $1/T_1^{\rm dd}$ at low magnetic field to the three other identified mechanisms.
+The increase in the depolarization rate in zero field ($\sim 20\ \%$) as well as the decrease inPL ($\sim 0.5 \%$) are still present, but significantly smaller than in the previous case.
+We can note however that the half-width at half maximum of the zero field depolarization ($\sim 2\ \rm G$) is also smaller in this case.
+
+We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF].
+We can not explain however why the depolarization rate appear to be smaller at $B \sim 10\ \rm G$ than it is for $B > 30\ \rm G$.
+
+The fact that the increase of the depolarization rate at zero field is about $\sim 50$ times weaker in Fig.
+\ref{scan 100} than it was in Fig.
+\ref{scan 1x1x1x1} confirms that the leading cause of depolarization in zero field is the degeneracy between the four classes of NV centers.
+
+\subsection{NV-NV CR under purely transverse magnetic field}
+
+Thank to the two previous experiments, we were able to discriminate the contribution of the 4-classes degeneracy from the three other mechanisms that are which are the eigenbasis change, the double-flips and the $T_2^* increase$.
+We now seek to discriminate the contribution of each of these three mechanisms.
+
+We unfortunately were not able to discriminate the change of eigenbasis from the $T_2^* increase$, which would likely require a statistical study over different samples with different $T_2^*$ values.
+We were however able to discriminate the double flip contribution from the two other ones by scanning the magnetic field in a direction orthogonal to one NV class ($\mathbf{B} \perp [111]$).
+
+\subsubsection{Principle of the experiment}
 \begin{figure}[h]
@@
 \ref{eigenstates transverse field}-b) shows both the splitting $\Delta \nu$ between the $\ket{2}$ and $\ket{3}$ states, and how close the $\ket{3}$ state is to the $\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}}$ state via the factor $\abs{\bra{3}\ket{+}}^2$.
-We look at the $\ket{3}$ state since $\ket{2}=\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}}$ in this case.
-
-These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 since we wanted to observe the modification caused by the eigenbasis $\{\ket{0},\ket{\pm}\}$.
-This way we can isolate the $T_1^{\rm dd}$ contribution coming from the change of eigenstates to the one coming from the double flips.
-Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100,200] G seems to satisfy both these criteria.
-
-\subsection{Experimental data}
+We should note that with the assumption that $B_\perp \parallel E_\perp$, the $\ket{2}$ state is strictly equal to the $\ket{-}$ state regardless of the magnetic field amplitude.
+
+These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are the relevant parameters of the experiment: the goal is to increase $\Delta \nu$ to the point where the double flips are completely suppressed, while $\abs{\bra{3}\ket{+}}^2$ remains close to 1 in order to preserve the eigenbasis and $T_2^*$ contribution.
+Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100 G,200 G] seems to satisfy both these criteria: $\Delta \nu\,(100\ \rm G) \approx 30\ \rm MHz \gg \gamma_f$\footnote{Technically, $\Delta \nu$ is to be compared with the dipole-dipole interaction width $2\gamma_f+\Gamma_{NV}+\Gamma_f$ which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+[REF].} and $\abs{\bra{3}\ket{+}}^2\, (200\ \rm G) \approx 0.96$.
+
+Ideally, we would have used an electric field instead of a transverse magnetic field to split the 4 classes energetically, but such an electric field would need to be $> 10^6\ \rm{V/cm}$ which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.
+
+\subsubsection{Results}
 
@@
 
-We perform here a $T_1^{\rm dd}$ measurement similar to the one presented in Fig.
-[REF], on the same sample and with the same fitting parameters, but this time probing a class orthogonal to the magnetic field.
-
-Fig.
-\ref{champ tranverse exp}-a) and b) show the evolution of the transition frequencies $\nu_+=\frac{E_{\ket{3}}-E_{\ket{1}}}{h}$ and $\nu_-=\frac{E_{\ket{2}}-E_{\ket{1}}}{h}$ where $\ket{1}$, $\ket{2}$ and $\ket{3}$ are the states described in Fig.
-\ref{eigenstates transverse field}.
-The magnetic field was scanned between 20 and 130 G by increment of $\sim 0.5\ \rm G$.
-
-Fig.
-\ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ with the magnetic field.
-We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$.
-In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ is increased.
-
-\subsubsection{Region A}
-
-We attribute the decrease in region A to the double flips: as the magnetic field increases, so does $\Delta \nu$ and the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$.
-This value is coherent with the previously measured fluctuator linewidth.
-We can then estimate the part of the double flips in the total depolarization rate for the low field regime (20 G here) and find that they amount for $\sim 50\%$ in this case.
-
-Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see in Fig.
-\ref{champ tranverse exp}-a) that they are not very far detuned.
-However we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.
-
-\subsubsection{Region B}
-
-We now turn to region B.
-Even tough the double flips are completely quenched, the effects of the eigenbasis modification and the $T_2^*$ reduction are still present, because we can see on Fig.
-\ref{eigenstates transverse field}-b) that the eigenstates of the spin Hamiltonian are still very close to $\{ \ket{0}, \ket{\pm} \}$ for these magnetic field values.
-
-To evaluate the contribution of the eigenbasis and $T_2^*$, we measure the same class of NV centers, but this time with a longitudinal magnetic field ($B_\parallel \sim 100 G$) to get a baseline value of $1/T_1^{\rm dd}$ in the $\ket{\pm 1}$ basis.
-We found a value of $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ in this case, reported on Fig.
-\ref{champ tranverse exp}-c) as a green line.
-
-We do find that the decay rate is higher in the $\ket{\pm}$ basis by $\sim 50\% $ which corroborates our predictions, as both the change of eigenbasis from $\ket{\pm 1}$ to $\ket{\pm}$ and the increase of $T_2^*$ in the $\ket{\pm}$ basis are predicted to increase the depolarization rate.
-The theoretical increase however overestimates once again the actual increase: for the change of eigenbasis alone we predicted an increase of the decay rate by a factor of 4 (see appendix [REF]), over 8 times more than the actual increase.
-
-We unfortunately cannot experimentally discriminate the depolarization coming from the eigenbasis change or the $T_2^*$ while using the same sample, since both of this properties are tied to the Hamiltonian eigenstates.
-But we do know that in this scenario, for a single isolated class, the contribution of both the change of eigenbasis and $T_2^*$ amounts for $\sim 3$ times less than the double flips.
-
-\subsubsection{Conclusion of the experimental observation}
-
-We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field, and we expect that the depolarization for low magnetic fields behave similarly.
-We therefore claim that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
-\begin{enumerate}
-\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
-\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
-\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
-\end{enumerate}
+Fig.
+\ref{champ tranverse exp}-a) shows an ODMR spectrum for $B=20\ \rm G$ with the two central lines corresponding to the $\ket{0} \to \ket{+}$ and $\ket{0} \to \ket{-}$ transitions of the class orthogonal to the magnetic field.
+We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\mathbf{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
+\ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\mathbf{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}
+
+Fig.
+\ref{champ tranverse exp}-b) shows the evolution of the $\nu_+$ and $\nu_-$ transition frequencies as a function of the magnetic field, measured via ODMR.
+The difference in frequency $\Delta \nu = \nu_+ - \nu_-$ is at first determined by the transverse electric field ($\sim 9\ \rm MHz$) and then grows qudratically with the transverse magnetic field up to a value of $\sim 50\ \rm MHz$.
+
+Fig.
+\ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ on a single class as a purely transverse magnetic field is scanned from 20 to 130 G.
+It also includes the $1/T_1^{\rm dd}$ value found for a purely longitudinal magnetic field which we will consider as a baseline, since the eigenbasis in this case is $\ket{\pm 1}$ and not $\ket{\pm}$.
+We have divided the graph in two regions labeled A and B:
+
+In region A, $1/T_1^{\rm dd}$ decreases with the magnetic field.
+The difference in the depolarizing rate between the $\mathbf{B} \perp [111]$ case and the $\mathbf{B} \parallel [111]$ case comes from the double flips, the eigenbasis change and the $T_2^*$ increase.
+We measure an increase by a factor $\sim 3$ for B=20 G compared to the baseline.
+
+In region B, $1/T_1^{\rm dd}$ reaches a plateau with a value of $\sim 180\ \rm s^{-1}$.
+We attribute this plateau to the fact that the double flips are completely quenched, which is coherent with the fact that $\Delta \nu \geq 30\ \rm MHz$.
+The remaining difference between the $\mathbf{B} \perp [111]$ and $\mathbf{B} \parallel [111]$ cases therefore comes from the eigenbasis change and the $T_2^*$ increase.
+This increase amount for $\sim$ 50\% of the relaxation rate baseline.
+
+We should note that the theoretical increase from the double flips and the change of eigenbasis\footnote{The $T_2^*$ increase should still amount for $\sim 10\%$.} are not the values shown in Fig.
+\ref{summary_theory}: those were computed for a 4-class resonance, while we are looking here at a single class.
+The theoretical increase of the depolarization rate for a single class are detailed in Appendix [REF].
+They are:
+\begin{itemize}
+\item $\Gamma_1^{\ket{\pm}}=4 \Gamma_0$ when taking into account the change of eigenbasis.
+\item $\Gamma_1^{dd}=[REF]\Gamma_1^{\ket{\pm}}=[REF]\Gamma_0$ when taking into account the double flips.
+\end{itemize}
+We have labeled here $\Gamma_0$ the relaxation rate coming from the flip-flop inside of a single class in the $\ket{\pm 1}$ basis (same definition as in table [REF] (table des valeurs chap 3).
+Experimentally, $\Gamma_0$ corresponds to the baseline shown in Fig.
+\ref{champ tranverse exp}-c).
+
+Although we find again that the fluctuator model overestimates the actual relaxation rates, these experimental observations still confirm the qualitative predictions of the model.
+In particular, they confirm that the double-flips are the second leading cause of zero-field depolarization after the 4-classes degeneracy.
+
+\subsection{Conclusion of the experimental observation}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/Shema_summary_exp}
+\caption{Summary of the different experimental investigation.
+A baseline value is indicated with the pink shaded region in all experimental plots.
+The left box indicates the relaxation mechanisms added compared to the baseline for the minimum value of B, and the right box those for the maximum value of B.
+(a) corresponds to Fig.
+\ref{scan 1x1x1x1}-d) where the baseline was measured to be $220\pm 20$ s$^{-1}$, (b) corresponds to Fig.
+\ref{scan 100}-d) with the same baseline, and (c) corresponds to Fig.
+\ref{champ tranverse exp}-c).}
+\label{summary_exp}
+\end{figure}
+
+Fig.
+\ref{summary_exp} summarizes the experimental results of this section.
+We explicitly indicates which of the four depolarization mechanisms are responsible for the observed depolarization rate at the beginning and at the end of each curve.
+
+The experiments confirm the overall predictions of the previous section: the dominant cause of depolarization in zero magnetic field is the degeneracy of the 4 NV classes, followed by the double-flips and then by the eigenbasis change and $T_2^*$ increase.
+
+We will now cover the low-field magnetometry protocol, after a quick overview of the field of NV magnetometry, and see the role that the different depolarization mechanisms play in it.
 
 \section{Introduction to ensemble NV magnetometry}
-
-Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.
-
-NV magnetometry is generally decomposed between single NV and ensemble magnetometry.
-Single NVs,often associated with scanning probes, offer a spatial resolution limited only by how far the magnetic source can be from the NV center (typically $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
+\label{sec 4.4}
+
+There are two classes of NV magnetometer: single NV or ensemble.
+Single NVs offer a spatial resolution limited only by how far the magnetic source can be from the NV center (down to $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
 
@@
 
-I will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.
+We will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.
 
 \subsection{LFDM optimization}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/double_adamas}
+\caption{a) PL of two distinct samples ADM-15-1 and ADM-15-4 as the magnetic field is scanned in an arbitrary direction (which splits the 4 classes similarly in both cases).
+b) Same measurement with $B\parallel [100]$ c) $T_1$ measurement on both samples on a single class for $B\approx 50\ \rm G$.
+d) $T_1$ measurement on both samples on all four classes for $B=0$.}
+\label{double dragon}
+\end{figure}
 
@@
 
-It seems then that increasing the NV density would be the obvious path to improve the LFDM performances.
-In practice however, I observed that the relationship between NV density, fluctuators, and LFDM performances was generally not straightforward.
-The two following examples will illustrate that.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 4/Figures/substrat}
-\caption{Measurements on sample SBST-C for two different regions A and B.
+Increasing the NV density seem to be the path for further material optimization, but other issues may arise in the process.
+
+We should note that we observed large discrepancy in the zero field depolarization among the samples that we studied, even from samples coming from the same batch.
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=\textwidth]{chapter 4/Figures/substrat}
+%\caption{Measurements on sample SBST-C for two different regions A and B.
 a) PL contrast as a function of a magnetic field scanned in an aribitrary direction (same direction for both regions).
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 4/Figures/double_adamas}
-\caption{Same measurements as Fig.
-\ref{substrat chelou} on two distinct samples ADM-15-4 and ADM-15-1.
-Every parameters (laser power,optical setup , AC magnetic field) was kept identical for the two samples.}
-\label{double dragon}
-\end{figure}
-
-We now turn to Fig.
+Fig.
+\ref{double dragon}-a) shows the PL on two distinct 15-$\mu$m Adamas microdiamonds for a magnetic field scanned in a random direction, and Fig.
+\ref{double dragon}-b) for a magnetic field scanned along the [100].
+We can notice that the PL contrast in zero field is twice as big for sample ADM-15-4 when $B \nparallel [100]$, and 5 time as big when $B \parallel [100]$, which would directly impact their respective sensitivity in LFDM.
+
+Fig.
+\ref{double dragon}-c) and d) shows $T_1$ measurement on both samples either on a single class for $B\neq 0$ or for all four classes for $B=0$.
+Interestingly, we can notice that the $T_1$ of sample ADM-15-1 is both shorter than the one of ADM-15-4 for $B\neq 0$, and longer for $B=0$.
+This could indicate that another relaxation mechanism is competing with the NV-fluctuator one on sample ADM-15-1, which would limit the LFDM sensitivity.
+
+%We now turn to Fig.
 \ref{double dragon}.
@@
 
+\chapter*{Conclusion}
+
+The work presented in this manuscript is part of an ongoing effort to develop quantum technologies, and NV centers, among many other systems, are a part of this new ecosystem.
+While the properties of single NV centers have been thoroughly studied in the past 20 years, we showed in this thesis that emerging properties of NV ensemble can still be discovered.
+These ensemble properties are relevant for the many fields using NV ensembles \citep{ barry2020sensitivity, eills2022spin, fu2007characterization, perdriat2021spin}.
+They also are relevant to explore fundamental aspects \citep{choi2017observation, kucsko2018critical, angerer2018superradiant} and for new potential applications \citep{akhmedzhanov2017microwave, pellet2021optical, pellet2022spin}.
+
+At the heart of this manuscript is the notion of cross-relaxation with NV centers.
+Cross-relaxations are a powerful spectroscopic tool, and NV centers are naturally suited to cross-relaxation studies due to their ability to be optically polarized and read-out, in opposition to their environment.
+
+We have showed how cross-relaxations can be used in a CVD-grown diamond to detect trace amount of other spin impurities, with higher sensitivity than the standard EPR apparatus, and only a fraction of the cost.
+These observations allowed us to study the temperature dependence of these impurities, and could eventually allow the detection and coherent manipulation of single dark spins through their interaction with an NV bath.
+
+We have then studied the cross-relaxation between NV centers, thanks to the multiple classes of NV and the ability to tune them in and out of resonance.
+The observation of cross-relaxation between seemingly equivalent spins has led previous researchers to postulate the existence of dark NV centers called fluctuators.
+We investigated experimentally the predictions of the fluctuator model, in particular the prediction of a stretched exponential profile for the depolarization of the spins, and a broadening of the fluctuators spectra compared to those of regular NV centers.
+Our observations for the large part comfort the fluctuator model, although we did notice some deviations which could indicate some limitations of the model.
+
+Dipole-dipole interaction within dense ensemble of NV centers ([NV] > ppm) can be responsible for an increase of the depolarization rate by more than a factor of 10.
+This changes considerably the properties of NV ensemble compared to single NV.
+Having a better understanding of the fluctuators and of their microscopic origin will be a crucial point in the future development of NV ensemble technologies.
+The further steps to determine the nature of the fluctuators could include the study of dedicated samples grown with specific concentrations of NV and other defects, as well as the study of nano-diamonds to try to isolate a single fluctuator.
+
+Finally, we investigated the depolarization of NV ensemble at low magnetic field.
+We identified 4 different contributions to the zero field depolarization which we analyzed both theoretically through the fluctuator model, and experimentally by varying the magnetic field orientation.
+We concluded that the dominant factor in the zero field depolarization was the degeneracy of the 4 NV classes, followed by the double flips.
+
+The depolarization of NV ensemble in zero field can in itself be used to measure magnetic field.
+We characterized such a magnetometer and compared it to the state of the art NV magnetometers.
+We found in particular that the dominant factor in the sensitivity of the magnetometer was the presence of the double flips, and not the classes degeneracy.
+
+The low-field depolarization magnetometer is unique among the NV magnetometers in that is does not require a microwave or a precise orientation of the magnetic field to work, which makes it usable with powdered or polycrystalline samples.
+Applications of magnetometry with powders include wide-field imaging of uneven surfaces, background free fluorescence microscopy, and microwave-free nano-probe sensing.
+
+\appendix
+
+\chapter{Diamond samples}
+This appendix document the known properties of every diamond samples used in this manuscript.
+
+\section{CVD samples}
+\begin{table}[htbp]
+\centering
+\caption{Concentration in nitrogen and NV centers for the CVD samples}
+\label{Table samples CVD}
+\begin{tabular}{c|cc}
+\toprule
+Name &  [NV$^-$]  & [N] \\
+\midrule
+CVD-pink & 4.6 ppm & 26 ppm \\
+CVD-B & 0.43 ppm & 10-30 ppm \\
+CVD-C & 0.43 ppm & 10-30 ppm \\
+
+\bottomrule
+\end{tabular}
+\end{table}
+Table \ref{Table samples CVD} summarizes the most relevant parameters in our study which are the concentration in (negatively charged) NV centers and the total concentration of nitrogen in the sample.
+\subsection{CVD-pink}
+The growth process and characterization of sample CVD pink are detailed in \citep{tallaire2020high}.
+The sample was grown by CVD in a gas environment containing 500 ppm of N$_2$O.
+The sample was then irradiated by a 10 MeV electron beam at a fluence of $2\cdot 10^{18}\ \rm cm^{-2}$ for 20 hours, while heated to 900 $^\circ$C during the irradiation process.
+
+The concentration of NV centers and nitrogen impurities was determined via UV-visible absorption and FTIR.
+
+\subsection{CVD-B and CVD-C}
+The growth process and characterization of sample CVD pink are detailed in \citep{ngambou2022improving}.
+These two samples were grown with a N$_2$O concentration of 100 ppm in the gas phase, and were irradiated similarly to sample CVD-pink.
+The main difference between the two sample is the further annealing step on sample CVD-B, as detailed in chapter 2.
+
+\section{HPHT samples}
+\subsection{Adamas fluorescent microdimaonds}
+\begin{table}[htbp]
+\centering
+\caption{Concentration in nitrogen and NV centers for the HPHT samples}
+\label{Table samples adamas}
+\begin{tabular}{c|cc}
+\toprule
+Name &  [NV$^-$] & [N] \\
+\midrule
+Adamas & 2.5-3.5 ppm & $\sim$100 ppm \\
+
+\bottomrule
+\end{tabular}
+\end{table}
+
+The main HPHT sample studied in this manuscript are the commercially available red fluorescent agents from Adamas nano.
+Samples labeled ADM-15-$n$ correspond to 15 $\mu$m micro-diamonds, and samples labeled ADM-150-$n$ to 150 $\mu$m micro-diamonds.
+The Adamas website indicates that the fluorescent diamonds contain $\sim 100$ ppm of nitrogen impurities, and that the micro-diamonds are irradiated with 2-3 MeV electron beams and annealed, without further information.
+The [NV] concentration, also indicated on the website, was measured via EPR.
+
+\subsection{Other HPHT samples}
+\subsubsection{Sumi-2}
+Sample Sumi-2 is a type Ib HPHT sample bought from Sumitomo and irradiated with a 2 MeV electron beam and a fluence of $7\cdot 10^{18}\ \rm cm^{-2}$ which were further annealed at 800 $^\circ$C for two hours.
+The NV and N concentration were not measured on this sample.
+\subsubsection{SBST-B and SBST-C}
+Samples SBST-B and SBST-C are the type Ib HPHT substrates on which the samples CVD-B and CVD-C grew.
+They were irradiated and annealed along the CVD layers.
+The NV and N concentrations in these samples has not been measured.
+
+\chapter{Computation of $\bar \eta$}
+We will discuss here the specifics to the computation of the  $\bar \eta$ factor for different geometric configurations, and how to compute the $T_1$ times in table [REF] of chapter [REF], and table [REF] of chapter [REF].
+
+As a reminder, $\bar \eta$ is defined as the averaged value of $\eta$ over all possible configurations.
+
+\begin{equation}
+\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta,
+\end{equation}
+
+where $\eta$ is defined as:
+
+\begin{equation}
+\label{eq. eta debut annexe}
+\eta^2=\frac{1}{3} \left(\frac{\mel{i}{\mathcal{H_{\rm dd}}}{f}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
+\end{equation}
+
+and where $\mathcal{H_{\rm dd}}$ is the dipole-dipole Hamiltonian and $\ket{i}$ and $\ket{f}$ the initial and final two-qubits states of the flip-flop or double flip process.
+
+\section{$\bar \eta$ in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 3/Figures/Shema_annexe}
+\caption{Schematics of the ``normal" NV center NV$_s$ and the fluctuator NV$_f$ in and their respective basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$), as well as the relative position $\mathbf{r}(r,\theta,\phi)$ between the two spins.}
+\label{shema spins annexe}
+\end{figure}
+
+The computation of $\bar \eta$ when the single spin Hamiltonian of each spin is diagonal in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$ was treated in \citep{choi2017depolarization}.
+We will summarize their results here and consider a few different geometries than the ones presented in the article.
+
+Fig.
+\ref{shema spins annexe} represents the two spins and the three relevant angles in the problem $\theta,\phi$ and $\psi$.
+We label with $s$ the properties associated with the ``normal " NV center and $f$ those associated with the fluctuator.
+For instance the two-qubits state $\ket{m_s=0,m_f=+1}\equiv\ket{0,+1}$ corresponds to the convolution of the single-spin states $\ket{m=0}$ for the NV and $\ket{m=+1}$ for the fluctuator.
+
+The two Cartesian basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$) were chosen so that the spin vector Hamiltonian of each spin could be written:
+\begin{equation}
+\mathbf{S}_i=S_x \hat x_i + S_y \hat{y}_i + S_z \hat{z}_i,
+\end{equation}
+where :
+\begin{align*}
+S_x&=\begin{pmatrix}
+0&1/\sqrt{2}&0 \\
+1/\sqrt{2}&0&1/\sqrt{2} \\
+0&1/\sqrt{2}&0
+\end{pmatrix} \\
+S_y&=\begin{pmatrix}
+0&i/\sqrt{2}&0 \\
+-i/\sqrt{2}&0&i/\sqrt{2} \\
+0&-i/\sqrt{2}&0
+\end{pmatrix} \\
+S_z&=\begin{pmatrix}
+-1&0&0 \\
+0&0&0 \\
+0&0&+1
+\end{pmatrix}.
+\end{align*}
+
+The relative position between the two spins is noted with the vector $\mathbf{r}=r\hat{u}$.
+The vector's spherical coordinate are given in the Cartesian basis of the NV center ($\hat x_s,\hat y_s,\hat z_s$).
+
+The angle $\psi$ between the two spins $z$ axes is defined by the crystal lattice and can take three values within [0,$\pi$] which are: 0 (same class), $\cos^{-1}(1/3) \approx 70.5\ ^\circ$ and $\cos^{-1}(-1/3) \approx 109.5\ ^\circ$.
+Two distinct classes can either have a relative angle $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$ depending on the external magnetic field: we choose by definition that $\mathbf{B}\cdot\hat{z}_i >0$, which imposes the orientation of $z_i$ and therefore the angle between $\hat z_s$ and $\hat z_f$.
+This point will be further developed when discussing the $\{100\}$ and $\{110\}$ type resonances.
+
+The direction $\hat x_s=\hat x_f$ is chosen arbitrarily to be : $$\hat x=\frac{\hat z_s \times \hat z_f}{\norm{\hat z_s \times \hat z_f}}.$$
+The arbitrary choice of the $\hat x$ direction is justified by the symmetry of the problem in the ($\hat x, \hat y$) plane.
+We chose here to neglect the symmetry-breaking part of the transverse magnetic field which is justified by the fact that we do not take into consideration the mixing of the eigenstates caused by the transverse field.
+
+Following the notation introduced in \citep{choi2017depolarization}, we can rewrite the dipole-dipole Hamiltonian as:
+\begin{equation}
+\mathcal{H}_{\rm dd} \approx -\frac{J_0}{r^3}\left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}) + h.c. + q S_z^sS_z^f \right],
+\end{equation}
+with 
+\begin{align}
+g&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f) + 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{y}_s \cdot \hat{y}_f) \right] \\
+h&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{x}_s \cdot \hat{y}_f) - 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{x}_f) + (\hat{y}_s \cdot \hat{x}_f) \right] \\
+q&= 3(\hat{u}\cdot \hat{z}_s)(\hat{u}\cdot \hat{z}_f) - (\hat{z}_s \cdot \hat{z}_f).
+\end{align}
+
+We should note that the double flips have been omitted because we consider here the case $B\neq0$ for which the $\ket{+1}$ and $\ket{-1}$ states of each spin are far out of resonance.
+
+Eq. \ref{eq. eta debut annexe} can then be written:
+\begin{equation}
+\eta^2=\frac{1}{3} (\abs{g}^2 + \abs{h}^2) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{equation}
+
+In order to compute $\bar \eta$, we first decompose $\eta$ as a product of $R$ and $G$ as defined in [REF]:
+\begin{equation}
+\eta^2=G^2(\theta, \phi, \psi) R^2(\omega_f,\omega_{NV})
+\end{equation}
+with
+\begin{align*}
+G^2(\theta, \phi, \psi)&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
+R^2(\omega_f,\omega_{NV}) &= \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{align*}
+
+$R$ and $G$ can be averaged separately as they do not depend on the same variables, which means that:
+\begin{equation}
+\label{eq. eta bar r bar g bar}
+\bar \eta = \bar R \cdot \bar G.
+\end{equation}
+
+The computation of $\bar{R}$ has been discussed in [REF], we will focus here on $\bar{G}$ defined as:
+
+\begin{equation}
+\label{eq. eta base mag}
+\bar{G}= \iint \abs{G} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi}.
+\end{equation}
+
+Eq.
+\ref{eq.
+eta base mag} can be solved analytically for the case $\psi=0$ and numerically for $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$.
+The three values are reported in Table \ref{table G mag}.
+The values found for $\psi=0$ and $\psi=70.5 \ ^\circ$ are the same that were found by the authors of \citep{choi2017depolarization} \footnote{They differ by a factor $\sqrt{1/3}$ because of the definition of $G$.}.
+
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar G$ in the magnetic basis}
+ \label{table G mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+$\frac{2}{9}=0.222$ & 0.3757 & 0.4808 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Interclass resonance and magnetic field orientation}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{chapter 3/Figures/110_vs_100}
+\caption{(a) Illustration of the co-resonance condition between a NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
+The $\hat z$ direction is defined by the condition $\mathbf{B}\cdot\hat{z} >0$.
+(b) 8 individual $T_1$ measurements on sample ADM-15-3 on a two-class resonance for 8 different values of the magnetic field (4 $\mathbf{B} \in \{110\}$ and 4 $\mathbf{B} \in \{100\}$).}
+\label{110 vs 100 annexe}
+\end{figure}
+
+As discussed in sec [REF], there are 4 possibles magnetic field orientations for which at least two NV classes are co-resonant: $\mathbf{B} \in \{110\}$ (two-class resonance), $\mathbf{B} \in \{100\}$ (2$\times$two-class resonance), $\mathbf{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\mathbf{B} \parallel \langle 100\rangle$ (four-class resonance).
+
+Fig.
+\ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$.
+We consider here an NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
+Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\mathbf{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
+
+For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\mathbf{B} \in (100)$ or $\mathbf{B} \in (011)$.
+The angle $\psi$ between the $\hat z_s$ and $\hat z_f$ axes however is different in both cases.
+The same can be generalized for every $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\mathbf{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\mathbf{B} \in \{110\}$ type resonance.
+
+Because $\bar G$ is greater for $\psi=109.5 \ ^\circ$ than it is for $\psi=70.5 \ ^\circ$, then following eq.
+\ref{eq.
+eta bar r bar g bar}, we should expect a faster depolarization rate when $\mathbf{B} \in \{110\}$ than when $\mathbf{B} \in \{100\}$.
+Fig.
+\ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\mathbf{B} \in \{110\}$ scenario and four in the $\mathbf{B} \in \{100\}$ scenario.
+We can clearly see a slower relaxation rate for all four $\mathbf{B} \in \{100\}$ measurements compared to all four $\mathbf{B} \in \{110\}$, which agrees with the predictions of the model.
+
+For a perfect resonance matching (i.e.
+no detuning between the central frequencies of the different resonant classes), we can assume that $\bar R$ is a constant ($\bar R \equiv \bar R^0 \approx \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}$ according to eq.
+[REF]).
+We then have:
+
+\begin{align}
+\bar \eta(\rm{1\ class})&=\bar G(\psi=0) \bar R_0 \\[1em]
+\frac{\bar \eta(\mathbf{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
+\end{align}
+
+Finally, the values in Table [REF] are computed thanks to the relation: \begin{equation}
+\frac{\Gamma_1}{\Gamma_0}= \left(\frac{\bar \eta}{\bar \eta(\rm{1\ class})} \right)^2.
+\end{equation}
+
+Note that we only give here the relaxation rates for a single class.
+To estimate the total PL, one would need to compute the relaxation rate of all 4 classes independently, for instance in a $\mathbf{B} \in \{110\}$ scenario, two classes have a relaxation rate $\Gamma(\mathbf{B} \in \{110\})$ and two classes have a relaxation rate $\Gamma_0$.
+
+\section{$\bar \eta$ in the non-magnetic basis $\{ \ket{0}, \ket{+}, \ket{-}\}$}
+
+In this part we will study the averaged flip-flop and double flip rates in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis which corresponds to the various mechanisms discribed in chapter 4.
+
+We will define here the $\ket{\pm}$ states as:
+\begin{equation}
+\ket{\pm}=\frac{\ket{+1}\pm\ket{-1}}{\sqrt{2}}.
+\end{equation}
+
+In this basis, the spin operators become:
+\begin{align*}
+S_x&=\begin{pmatrix}
+0&0&0 \\
+0&0&1 \\
+0&1&0
+\end{pmatrix} \\
+S_y&=\begin{pmatrix}
+0&-i&0 \\
+i&0&0 \\
+0&0&0
+\end{pmatrix} \\
+S_z&=\begin{pmatrix}
+0&0&1 \\
+0&0&0 \\
+1&0&0
+\end{pmatrix}.
+\end{align*}
+
+Importantly, we have to note the breaking of the symmetry in the (xy) plane.
+This comes from the transverse electric or magnetic field dominating the single spin Hamiltonian.
+By convention, we define the $x$ direction as the direction of the dominant transverse electric or magnetic field.
+
+\subsection{Flip-flops in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis}
+\label{sec flip-flop +/-}
+Rewriting the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-}\} \times \{ \ket{0}, \ket{+}, \ket{-}\}$ changes the value of the flip-flop and double flip prefactors.
+For instance, the prefactor of the $\ket{+,0}\bra{0,+}$ term is: \begin{equation}
+-\frac{J_0}{r^3}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)\right]
+\end{equation}
+while the prefactor of the $\ket{+1,0}\bra{0,+1}$ term was: \begin{equation}
+-\frac{J_0}{r^3}\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f) + 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{y}_s \cdot \hat{y}_f) \right] 
+\end{equation}
+
+Because of the symmetry breaking of the $(xy)$ plane, we now have to consider the relative orientation of $x_1$ and $x_2$ on top of the previously mentioned angles $\theta$, $\phi$ and $\psi$.
+We define $\alpha_1$ and $\alpha_2$ as the relative angle formed by the $x_1$ and $x_2$ axes with an arbitrary axis in their respective $(xy)$ plane (the arbitrary axis can be defined for instance in the lab frame of reference).
+
+Because we assume that the physical orientation of the $x_1$ and $x_2$ axes in zero external magnetic field is determined by a random local electric field, we will sample than angles $\alpha_1$ and $\alpha_2$ on a uniform [0,$2\pi$] distribution independently for each spin.
+
+The integral to compute the $\bar G$ factor of the dipole-dipole interaction in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis then becomes\footnote{The flip-flop term considered here is $\ket{+,0}\bra{0,+}$.
+The $\ket{-,0}\bra{0,-}$ term gives the same numerical values once averaged.}
+\begin{equation}
+\bar G= \frac{1}{\sqrt{3}} \iiiint \abs{3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi} \frac{d \alpha_1}{2\pi} \frac{d \alpha_2}{2\pi}.
+\end{equation}
+
+The numerical values obtained are reported in table \ref{table G flip-flop non mag}.
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar G$ in the $\ket{\pm}$ basis}
+ \label{table G flip-flop non mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+0.410 & 0.394 & 0.394 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Finally the value $\Gamma_1^{\rm th}= 51.4 \Gamma_0^{\rm th}$ shown in Fig. [REF] was computed from:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar G_{\ket{\pm}}(\psi=0) + 3 \bar G_{\ket{\pm}}(\psi\neq 0) }{\bar G_{\ket{\pm 1}}(\psi=0)}\right)^2
+\end{equation}
+\subsection{Double-flips in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis}
+Including the double flips, there are two possible relaxation path for each transition of the regular NV center.
+For instance the transition $\ket{0} \to \ket{-}$ can involve either the flip flop $\ket{0,-} \bra{-,0}$ or the double-flip $\ket{0,+} \bra{-,0}$.
+
+The inclusion of the double flips require to modify the  expression of $\eta^2$ to include both relaxation path.
+According to the more general definition given in \citep{choi2017depolarization}, the $\eta^2$ term corresponding to the $\ket{0} \to \ket{-}$ transition of the NV center can be written\footnote{The $\ket{\pm}$ states in the formula can be replaced by $\ket{\pm 1}$ depending on the eigenbasis of the single spin Hamiltonian.}:
+\begin{align}
+\label{eq eta2 avec DQ}
+\eta^2&=\frac{1}{3} \left(\frac{\mel{0,-}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{\left(\Delta_{\omega_f}^{(0,-)} - \Delta_{\omega_{NV}}^{(0,-)}\right)^2+4\gamma_f^2} \nonumber\\
+&+\frac{1}{3} \left(\frac{\mel{0,+}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{\left(\Delta_{\omega_f}^{(0,+)} - \Delta_{\omega_{NV}}^{(0,-)}\right)^2+4\gamma_f^2},
+\end{align}
+where $\Delta_{\omega_{NV}}^{(0,\pm)}$ corresponds to the central frequency of the $\ket{0} \to \ket{\pm}$ transition of the NV center, and $\Delta_{\omega_{f}}^{(0,\pm)}$ that of the fluctuator.
+
+Eq.
+(\ref{eq eta2 avec DQ}) can no longer be factorized as a product of a spectral and spatial function.
+To compute the value of $\bar eta$ in zero magnetic field, we will proceed to the following assumptions: $\Delta_{\omega_{NV}}^{(0,\pm)}=\Delta_{\omega_{f}}^{(0,\pm)}$, and $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$\footnote{This second assumption is justified by experimental measurements.}.
+Eq.
+(\ref{eq eta2 avec DQ}) then become:
+\begin{equation}
+\eta^2=\frac{1}{3} \left(\frac{\mel{0,-}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 
++\frac{1}{3} \left(\frac{\mel{0,+}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{1}{2}.
+\end{equation}
+
+Following the same procedure as in the last section, we then compute $\bar \eta$ in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis and report the values in table \ref{table eta double flip non mag}.
+These values are used to compute the double-flip term in Fig.
+[REF]
+
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar \eta$ in the $\ket{\pm}$ basis including the double flips.}
+ \label{table eta double flip non mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+0.554 & 0.532 & 0.532 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Flip-flops and double flips in the $B\perp [111]$ regime}
+
+We will focus here on the case detailed in Fig.
+[REF]: a single class of NV centers whose Hamiltonian is dominated by the transverse magnetic field.
+Unlike the case presented in Sec.
+\ref{sec flip-flop +/-}, the $x$ direction is fixed for all spins and imposed by the direction of the magnetic field.
+
+The expression of $\bar G$ then becomes:
+\begin{equation}
+\bar{G}= \frac{1}{\sqrt{3}}\iint \abs{3(\hat{u}\cdot\hat{x})^2 - 1} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi}= \frac{4}{9}\approx 0.444.
+\end{equation}
+
+The increase in the decay rate of the eigenbasis when considering only the flip-flop is:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar G_{\ket{\pm}}(\psi=0)}{\bar G_{\ket{\pm 1}}(\psi=0)}\right)^2=4.
+\end{equation}
+
+To include the double-flip, we once again make the assumption that $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$ and find $\bar \eta=0.567$.
+The increase in the decay rate when including the decay rate is:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar \eta_{\ket{\pm}}(\psi=0)}{\bar \eta_{\ket{\pm 1}}(\psi=0)}\right)^2\approx 6.5.
+\end{equation}
+
+\chapter{Influence of the magnetic field misalignment and laser polarization on the low field relaxation}
+
+\section{Magnetic field alignment}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 4/Figures/alignement_rose}
+\caption{PL of sample CVD-pink as a function of the magnetic field for 7 different misalignment of the field with respect to the [100] axis}
+\label{alignement rose}
+\end{figure}
+
+A potential cause for the observation of depolarization at low magnetic field when $\mathbf{B} \parallel [100]$ could be a misalignment of the magnetic field with the [100] axis.
+Indeed, such a misalignment would cause a splitting of the 4 classes as the magnetic field increases which would reduce the inter-class flip-flop rate.
+%This hypothesis is not consistent with the observations in Fig.
+\ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased.
+Instead we observe a sharp decrease followed by a pseudo-plateau.
+Fig.
+\ref{alignement rose} shows the PL of sample CVD-pink as the magnetic field is scanned with various misalignments compared to the [100] axis.
+We estimate our alignments to be precise within $\pm 1^\circ$.
+We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization in the $\mathbf{B} \parallel [100]$ case does not come from a field misalignment.
+
+\section{Laser polarization}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 4/Figures/pola_laser}
+\caption{PL of sample ADM-150-3 as a function of the magnetic field for six different laser polarization angles.
+The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
+\label{pola laser}
+\end{figure}
+
+Other studies \citep{anishchik2015low, filimonenko2020weak} previously observed a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.
+
+Fig.
+\ref{pola laser} shows PL scans on sample ADM-150-3 for various polarization angle of the incident laser.
+We chose to apply the magnetic field in the laser polarization plane, in a direction that did not match any particular crystalline planes.
+
+We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$.
+We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.
+
 \printbibliography
-\end{refsection}
 \end{document}

### 2022-10-19 python_clément/test programs/test.py

@@ -1753,4 +1753,5 @@ def plot_irregular_surface():
 	plt.show()
 
 
-print(116/0.015)
+
+print(0.9816/sqrt(3))

### 2022-10-20 Manuscrit Thèse/These.tex
tata yoyo
toto

@@
-\documentclass[a4paper]{article}
-\usepackage[french]{babel}
-\usepackage[utf8]{inputenc}
+\documentclass[a4paper, 11pt]{report}
+
+\usepackage{psl-cover}
 \usepackage[]{amsmath}
-\usepackage[]{braket} % \bra, \ket etc
-\usepackage{graphicx}
+\usepackage[]{physics} % \bra, \ket etc
+\usepackage{graphicx} %Pour les figures je crois
+\usepackage{hyperref}
+\usepackage[
+    backend=biber, 
+    natbib=true,
+    style=numeric-comp,
+    sorting=none, %Pour faire apparaitre les refs dans l'ordre
+    hyperref=true
+]{biblatex} %Imports biblatex package
+\addbibresource{bib_thesis.bib} %Import the bibliography file
 
-\title{Le titre}
+\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
+\usepackage{subcaption} % package pour faire des subfigures
+\usepackage{multirow} % package pour multirow/multicolumn
+\usepackage{booktabs} % package pour top/mid/bottom rule
+\usepackage{tcolorbox} % toujours plus de boites
+\usepackage{xcolor} % Pour avoir des couleurs dans les équations
+
+\title{Cross-relaxation in dense ensembles of NV centers and application to magnetometry \\ \medskip Relaxation croisées d'ensembles de centres NV et application à la magnétométrie}
+
+\author{Clément Pellet-Mary}
+
+\institute{ENS Paris}
+\doctoralschool{EDPIF}{564}
+\specialty{Physique}
+\date{07 Décembre 2022}
+
+\jurymember{1}{Dmitry Budker}{Professor JGU Mainz}{Rapporteur}
+\jurymember{2}{Philippe Tamarat}{Professeur, Université de Bordeaux}{Rapporteur}
+\jurymember{3}{Maria Chamarro}{Professeure, Sorbonne Université}{Examinateur}
+\jurymember{4}{Vincent Jacques}{Directeur de recherche, CNRS}{Examinateur}
+\jurymember{5}{Alexandre Tallaire}{Directeur de recherche, CNRS}{Examinateur}
+\jurymember{6}{Gabriel Hétet}{Maître de conférences, ENS Paris}{Directeur de thèse}
+% \jurymember{9}{Prénom NOM}{Titre, établissement}{Invité}
+% \jurymember{10}{Prénom NOM}{Titre, établissement}{Invité}
+
+\frabstract{
+Le centre NV du diamant est un candidat prometteur pour de nombreuses technologies quantiques, que ce soit pour les communications quantiques, le calcul quantique ou la métrologie quantique.
+Grace aux progrès de synthétisation du diamant, des échantillons contenant une haute concentration en centres NV sont aujourd’hui réalisables.
+Cette augmentation de la densité de centres NV s’accompagne d’une augmentation des interactions entre ces derniers, ce qui donne lieu à une physique complexe et encore largement inexplorée.
+De nombreuses applications récentes, telles que l’hyper-polarisation en RMN, la bio-imagerie ou encore la spin-mécanique repose sur des ensembles denses de centres NV.
+Comprendre et maîtriser les propriétés d’ensemble de ces derniers est un point crucial pour l’établissement de ces technologies.
+
+Cette thèse étudie les échanges de polarisation, par relaxation croisée, entre des ensembles denses de centres NV, ainsi que les interactions entre les centres NV et d’autres impuretés paramagnétiques présentes dans le diamant.
+Le but de ces travaux est de mieux comprendre les propriétés des ensembles de centres NV, et d’exploiter certaines de ces propriétés pour de nouvelles applications telles que la détection d’impuretés ou la magnétométrie.
+
+Dans un premier temps, nous détaillerons les propriétés des centres NV et nous introduirons des concepts liés à l’interaction dipolaire entre spins et aux relaxations croisées.
+Nous verrons ensuite comment les relaxations croisées peuvent être utilisées pour détecter optiquement des résonances de spins qui ne sont pas optiquement actifs, en l’occurrence les défauts VH- et War1 dans des diamants CVD.
+Enfin nous étudierons les relaxations croisées entre différents sous-groupes de centre NV.
+Nous verrons d’abord leur influence sur la dynamique du spin et leur origine microscopique, puis nous étudierons leur dépendance avec le champ magnétique et leur potentielle utilisation pour la magnétométrie en champs faible.
+}
+
+\enabstract{
+  The NV center in diamond is a promising candidate for emerging quantum technologies, including quantum communication, quantum computing and quantum metrology.
+Thanks to the recent progress in diamond synthesis, samples with high NV center concentration can now be created.
+This increase in the NV density comes with an increase of the interaction between them, which leads to a rich and complex physics.
+Numerous applications, such as NMR hyperpolarization, bio-imagery or spin mechanics rely on the use of dense NV ensembles.
+Understanding and mastering the properties of ensembles of NV centers will be a crucial point in the development of these technologies.
+
+This doctoral thesis focuses on cross-relaxation between dense ensembles of NV centers, as well as on the interaction between NV centers and other diamond paramagnetic impurities.
+The aim of this work is to better understand the properties of dense NV centers ensembles, and to exploit them for new applications such as the detection of impurities and magnetometry.
+
+In the first part, we will detail the properties of single NV centers and introduce notions regarding dipole-dipole spin coupling and cross-relaxations.
+We will then show how cross-relaxations can be used to optically detect dark spin resonances, in this case the VH- and War1 defects in CVD-grown diamond.
+Finally, we will present a study of cross-relaxations between subgroups of NV centers.
+We will first cover the influence of cross-relaxation on the spin dynamics and discuss its microscopic origin.
+We will then investigate the role of the magnetic field on cross-relaxations as well as their potential use in low field magnetometry.
+}
+
+\frkeywords{centre NV, dynamique de spin, magnétométrie}
+\enkeywords{NV center, spin dynamics, magnetometry}
 
 \begin{document}
-\maketitle
-\section{Section 1}
-\subsection{Subsection 1}
 
+\maketitle{}
+\setcounter{tocdepth}{1}
+
+\tableofcontents
+
+\chapter*{Introduction}
+
+Quantum mechanics was first developed at the beginning of the 20th century to explain physical observations such as the blackbody radiation \citep{planck1900theorie} or the photoelectric effect \citep{einstein1905erzeugung}.
+The quantum theory was subsequently developed in the first half of the 20th century and showed great success, in particular in giving a physical framework for chemistry \citep{pauling1931nature} and leading to a microscopic understanding of matter \citep{kittel1996introduction}.
+In the second half of the 20th century, technological development based on the new understanding of light and matter have led to what was retroactively called the first quantum revolution \citep{thew2019focus}.
+Technologies developed in the 50s and 60s include the semiconductor transistor \citep{bardeen1948transistor}, the laser \citep{maiman1960stimulated}, and nuclear magnetic resonance \citep{abragam1961principles} which have since revolutionized the fields of information science, global communications and medical imaging.
+
+A new field of quantum technology has emerged in the last two decades.
+This ``second quantum revolution" differs from the first one in the usage of individual quantum systems, such as single photons \citep{gleyzes2007quantum}, atoms \citep{neuhauser1980localized} or electrons \citep{peil1999observing}, which have properties that can not be emulated by larger classical systems.
+These properties include quantum entanglement, superposition or measurement and form the basis of the newly formed quantum information science \citep{nielsen2002quantum, vedral2006introduction, hayashi2006quantum}.
+
+Central to quantum information science is the idea of quantum bits or qubits \citep{schumacher1996sending}, which are the building blocks of quantum information in analogy to the bits of classical information theory.
+Physical implementation of a qubit could be any quantum system with two well defined quantum states, as long as these states can be initialized, manipulated and readout \citep{divincenzo2000physical}.
+Popular qubit candidates include superconducting circuits \citep{nakamura1999coherent, orlando1999superconducting}, photons \citep{bennett1992quantum, zhong2020quantum}, quantum dots \citep{veldhorst2014addressable, zajac2018resonantly}, trapped ions \citep{friis2018observation, wright2019benchmarking} and single crystal defects \citep{jelezko2006single, baranov2011silicon, zhong2015optically}.
+
+Defects in large band gap crystals are sometimes referred to as artificial atoms \citep{buluta2011natural}, as the localized wave function of the electrons surrounding the impurity  give rise to discrete energy levels similar to that of an atom.
+Of all the possible crystals and defects, the most studied defect for quantum applications is the negatively charged nitrogen-vacancy center in diamond \citep{aharonovich2016solid, de2021materials}, often abbreviated to ``NV center".
+Although this defect had been known and correctly identified almost 50 years ago \citep{davies1976optical}, interest in the NV center really sparked after it was first isolated in 1997 \citep{gruber1997scanning}.
+The detection of single NV centers proved that NV centers fluoresce brightly \citep{gruber1997scanning}, are photostable \citep{kurtsiefer2000stable}, and that their electronic spin can be polarized and readout optically \citep{jelezko2004observation}, making them an promising solid state qubit.
+
+Since then, single NV centers have been successfully used to create quantum entanglement over several km \citep{hensen2015loophole}, to store quantum information at room temperature for more than one second \citep{maurer2012room}, to detect single proteins via nuclear magnetic resonance \citep{lovchinsky2016nuclear}, and for many other applications in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} or quantum sensing \citep{degen2017quantum}.
+
+As the interest in single NV center grew, diamond synthesis improved significantly \citep{achard2020chemical, barry2020sensitivity, edmonds2020generation}  which benefited not only single NV centers but also NV center ensemble.
+While single NV center are required for most quantum information processes \citep{ladd2010quantum}, several other applications benefit from a greater concentration of defects and are therefore more suited to operate with NV centers ensemble.
+Four of these applications are pictured in Fig.
+\ref{shema introduction}.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.9\textwidth]{Intro et conclusion/Figures/Fig_intro}
+\caption{Example of applications with dense ensemble of NV centers.
+(a) In vivo nanodiomond biomarkers (from \citep{mohan2010vivo}).
+(b) Quantum sensing with NV ensemble (from \citep{ku2020imaging}).
+(c) Dynamic nuclear polarization of an external spin bath (from \citep{healey2021polarization}).
+(d) Spin mechanics with trapped micro diamonds (from \citep{delord2020spin}).}
+\label{shema introduction}
+\end{figure}
+
+The first application for NV center ensemble is the use of nano or micro diamonds as fluorescent biomarkers \citep{fu2007characterization, mohan2010vivo}.
+NV centers are good candidates for biomarkers as they emit a bright and stable photoluminesecence, and because diamond nanocrystals present low toxicity \citep{schirhagl2014nitrogen}.
+Some applications also use the spin degree of freedom of the NV centers, either to perform background free fluorescence microscopy \citep{chapman2013background} or to monitor the angular motion of the nanodiamonds \citep{mcguinness2011quantum, feng2021association}.
+For most of these applications, NV ensemble provide a better visibility than single NV centers and are therefore preferred.
+
+The second NV ensemble application is for quantum sensing \citep{degen2017quantum}.
+NV centers are predominantly used as magnetic field sensors \citep{rondin2014magnetometry, barry2020sensitivity}, although they can also detect electric fields \citep{dolde2011electric, michl2019robust}, temperature \citep{acosta2010temperature, kucsko2013nanometre} ,crystal strain \citep{ovartchaiyapong2014dynamic, doherty2014electronic} or rotation \citep{ledbetter2012gyroscopes, ajoy2012stable}.
+Both single and ensemble of NV centers can be used for sensing: single centers offer unmatched spatial resolution \citep{mitchell2020colloquium} down to a few nm \citep{lovchinsky2016nuclear}, while ensemble provide better sensitivities \citep{wolf2015subpicotesla, barry2020sensitivity}.
+
+The third NV ensemble application is nuclear magnetic resonance (NMR) dynamical nuclear polarization \citep{eills2022spin}.
+Dynamic nuclear polarization consists in increasing the nuclear polarization of a sample by transferring electron polarization to the nuclei bath \citep{abragam1978principles}.
+Because the NV electronic spin can be optically polarized at room temperature, micro and nano diamonds containing NV centers are promising candidates as hyperpolarization agents \citep{tetienne2021prospects}.
+Nuclear polarization of spins inside the diamond lattice was demonstrated with single NV centers \citep{jacques2009dynamic, smeltzer2009robust} and NV ensemble \citep{king2015room, scheuer2016optically, schwartz2018robust}.
+Polarization transfer to nuclear spins outside of the diamond is an ongoing challenge \citep{eills2022spin, healey2021polarization}, but dense NV ensemble are believed to be necessary to achieve high external polarization \citep{tetienne2021prospects, rizzato2022polarization}.
+
+Finally, the fourth NV ensemble application is the recent field of spin mechanics \citep{perdriat2021spin}.
+Similarly to optomechanics \citep{aspelmeyer2014cavity}, the goal of spin mechanics is to couple a quantum system to the mechanical degree of freedom of a macroscopic object.
+Strong coupling between single NV center and an external mechanical oscillators has been observed \citep{rabl2009strong, kolkowitz2012coherent}, but coupling between NV centers and the motion of the diamond has so far only been achieved with dense NV ensemble \citep{delord2020spin, pellet2021magnetic, perdriat2022angle}.
+Coupling several NV centers to the same mechanical oscillator could also lead to interesting magnetic phase transitions \citep{wei2015magnetic, ma2017proposal}.
+
+Most of the NV ensemble experiments discussed previously work under the assumption that the ensemble is constituted of $N$ independent single NV centers.
+However, as often in physics, \textit{more is different} \citep{anderson1972more}.
+Interaction between the NV center themselves or with their environment can lead to significantly different behavior between ensemble and single spins.
+These difference include a modification of the charge state dynamics \citep{giri2018coupled} and of the spin dynamics \citep{dobrovitski2008decoherence, jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization}, as well as the apparition of exotic phenomena such as time crystal phase \citep{choi2017observation}, Anderson localization \citep{kucsko2018critical} or superradiance \citep{bradac2017room, angerer2018superradiant}.
+
+The work presented in this manuscript focuses on dipolarly coupled dense ensemble of NV centers.
+More specifically, we look at spin polarization exchange between the NV centers themselves and with their environment.
+Polarization exchange (flip-flop, spin diffusion, cross-relaxation, ...) in dense ensemble of spins has been abundantly studied in the context of NMR \citep{abragam1978principles} and spintronics \citep{vzutic2004spintronics}.
+Because NV centers tend to be relatively dilute, the question of spin exchange between between NV themselves \citep{choi2017depolarization} or with their environment \citep{hall2016detection} has only recently been addressed.
+However, as samples with higher NV concentration become available \citep{acosta2009diamonds, tallaire2020high, shenderova2019synthesis}, these aspects are becoming more and more relevant.
+
+The goal of this present work is both to better understand the physics of dense NV ensemble, which is crucial for many of the NV ensemble applications cited previously, and to exploit some of these many-body properties for new applications.
+
+\bigskip
+This manuscript is organized in the following way:
+
+\medskip
+The first chapter introduces the theoretical and experimental concepts used in this manuscript.
+This includes a brief overview of diamond and NV centers fabrication processes, the physics of single NV centers and a presentation of magnetic dipole-dipole interaction and cross-relaxation.
+
+The second chapter covers the observation of cross-relaxation between NV centers and other spin impurities in CVD-grown diamond.
+We use measurements on the NV centers to identify these impurities and to quantify some of their attributes.
+Most of the results of this chapter have been published in  \citep{pellet2021optical} and \citep{ngambou2022improving}.
+
+The third chapter focuses on the dipole-dipole mediated spin relaxation within dense ensemble of NV centers.
+Following the work presented in \citep{choi2017depolarization}, we introduce the concept of NV-flutcuator to explain this phenomenon and present experimental and theoretical results solidifying this hypothesis.
+The results of this chapter have been partly published in \citep{pellet2022spin} and \citep{pellet2021magnetic}.
+
+Finally, the fourth chapter presents a study on NV-NV dipolar relaxation under low magnetic field.
+This chapter covers both the rich physics of NV centers under low magnetic field, and a potential application of the low field depolarization as a new magnetometry protocol, as well as comparisons with other NV magnetometry protocols.
+The results of this chapter have in large part been published in \citep{pellet2022spin}.
+
+\chapter{Spin properties of single and ensemble NV centers}
+The aim of this first chapter is to introduce concepts related to Nitrogen-vacancy (NV) centers, to the diamond material, and to the dipole-dipole interaction which are necessary to explain the work presented in the later chapters.
+
+In this introductory chapter, we will first cover the material aspects of NV centers by focusing on the creation process of synthetic diamonds and the required steps to create NV defects inside them.
+We will then cover the fundamentals of NV physics, including the NV center optical and spin properties, basic experiments with NV centers and the various relaxations times ($T_1$,$T_2$,$T_2^*$) which characterizes the NV center's spin.
+Finally we will present the dipole-dipole interaction between two spins and the concept of cross-relaxation.
+
+\section{The NV center as a physical object}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 1/Figures/diamant_et_NV}
+\caption{From the diamond to the NV center.
+a) Drawing of a bulk diamond\footnote{In this manuscript, we will consider any samples above $\sim$ 100 $\mu$m to be bulk.}.
+b) Drawing representing a fluorescence microscopy image of a diamond surface.
+The red fluorescent spots symbolize NV centers photoluminescence.
+c) Zoom on the crystalline structure of an NV center.
+Black balls represent carbon atoms}
+\label{diamond+NV}
+\end{figure}
+Before studying the NV centers properties, we must first discuss the diamond itself and answer these questions: How are diamond created, what kind of defects are there in diamond, and how to specifically create NV defects.
+The field of material science related to diamonds and NV centers is a rich and complex one, and this section will only touch upon basic notions regarding these subjects.
+
+Fig.
+\ref{diamond+NV} illustrates the various scales to consider in the study of NV centers.
+Starting with a bulk diamond, typically few mm wide, then zooming-in on the diamond surface with an optical microscope to find fluorescent spots corresponding to the photoluminescence of NV centers.
+Standard confocal microscopy is limited in its resolution to the $\sim \mu$m range, but if it was possible to zoom-in even further, we would see the atomic structure of a few \AA \ illustrated in the last panel.
+\subsection{Diamond material overview}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.6\textwidth]{chapter 1/Figures/phase_diagram_carbon_wiki}
+\caption{Pressure-temperature phase diagram of elemental carbon. Credits: Wikimedia commons, adapted from \citep{bundy1989pressure, bundy1996pressure}.}
+\label{carbon phase diagram}
+\end{figure}
+
+Diamond is a remarkable material in many respects.
+It is famous for being the hardest and most thermally conductive non-synthetic material.
+It is also a material with high optical refractive index and optical dispersion which makes it an extremely valuable crystal in jewelry.
+
+Diamond is also famous for being extremely rare in nature, even though its only constituent, carbon, is one of the most abundant element on the Earth surface.
+The reason being that graphite, and not diamond, is the thermodynamically stable phase of carbon under ambient conditions.
+
+Fig.
+\ref{carbon phase diagram} shows the PT phase diagram of elemental carbon.
+We can see that diamond becomes the thermodynamically stable phase only for pressure above a few GPa, $10^4$ times higher than atmospheric pressure.
+Even though diamond is unstable under ambient conditions, the extremely high energy activation needed to break the sp$^3$ bond between two carbon atoms means that the kinematics of the $C_{\rm diamond} \to C_{\rm graphite}$ transition is almost completely frozen.
+In many ways diamonds are more stable than graphite under ambient condition due to their relative chemical inertness.
+
+Where the thermodynamic equilibrium really matters is for the crystallization process.
+The higher pressure needed to reach the diamond-stable region can only be found naturally deep under the earth crust, typically between 150 and 240 km below sea level \citep{tappert2011diamonds}, which explain their scarcity.
+It also explains why, despite many previous attempts, diamonds were only synthetically produced in 1953 \citep{barnard2000diamond}, more than 50 years after the first synthetic sapphires.
+To this day, artificial diamond synthesis remains an active field of research \citep{shenderova2019synthesis, achard2020chemical}.
+
+Diamonds grown naturally or artificially are used in the lab and in industry for their extreme hardness, with applications ranging from abrasive to diamond anvil cells \citep{jayaraman1983diamond}.
+More recently, diamonds have been the object of optics and electronics applications.
+For both of these fields however, the focus is no longer solely on the diamond but also on the defects hosted inside it.
+
+\subsection{Point defects in diamond}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 1/Figures/Diamond_type}
+\caption{Traditional classification of diamonds, based on \citep{tappert2011diamonds}.}
+\label{diamond type}
+\end{figure}
+Diamond in its intrinsic form is an insulator with a bandgap of $\sim 5.47\ \rm eV$ and is transparent from far infrared to deep UV.
+As any crystal however, diamond is prone to defects.
+We will focus here on point-like (0D) defects and omit structural defects such as dislocations, even though those too can affect the optical properties of the diamond \citep{collins2000colour}.
+
+Point-like defects can be constituted by the absence of a carbon atom (``vacancies"), an interstitial defect between two carbon atoms or the substitution of a carbon atom with another atom.
+These impurities result in unpaired electron or holes localized around the impurity site.
+This localization causes discrete energy levels for the holes or the electrons, some of which lie inside the diamond bandgap and impact the crystal optical and electrical properties.
+
+Defects with electronic transitions in the optical range are called ``colored centers" as they are responsible for the coloring of the gems.
+Over 20 elements are known to form colored centers when introduced in diamond \citep{zaitsev2013optical, shenderova2019synthesis}, and each of these elemental atoms can form several defects.
+Nitrogen alone can form more than 50 different colored centers \citep{dobrinets2016hpht, ashfold2020nitrogen}.
+
+Nitrogen, and to a lesser extant boron, are the most commonly found extrinsic elements in natural diamond.
+The traditional classification of diamond is presented in Fig.
+\ref{diamond type}.
+It is based on the N and B concentration, with a threshold of a few ppm (part per million) for each species as this was the smaller amount easily detectable through IR absorption.
+Type Ia diamonds contain clustered N defects such as the B-center (N$_4$V defect).
+In contrast, type Ib diamond mostly contains C-center which are single substitutional N$_s$ defects, also referred to as P1 centers in the spin community.
+Type IIb diamond contains boron impurities which give them a blueish-grey color, and type IIa diamond contains no detectable impurities via IR absorption \citep{ashfold2020nitrogen}.
+
+Nitrogen-vacancy centers are a rare occurrence in both natural diamonds and untreated synthetic diamonds, but we will see later that the much more common N$_s$ defects can be converted in NV centers.
+When working with NV centers, the starting crystal is therefore often a type Ib diamond if one wants to work with ensemble, or a type IIa diamond if one wants to work with single NV centers \footnote{In order to isolate single NV in a confocal optical spot, the concentration needs to be [NV] $<$ 1 ppb.}.
+
+The boron defects are mainly studied in the context of p-doped diamond semi-conductor: a concentration of [B] $\approx 1000\ \rm ppm$ is needed to achieve a significant overlap between the impurity sites, which turns the diamond into a semi-metal \citep{macpherson2015practical}.
+Other defects of interest, generally not present naturally but introduced voluntarily, include P and S defects to create n-doped diamond \citep{das2022diamond} (still an active field of research), as well as group IV-vacancy defects (SiV, GeV, SnV, PbV) for quantum optics application \citep{bradac2019quantum}.
+
+The NV center however remains by far the most studied defect in diamond due to its unique spin properties which will be detailed below.
+
+\subsection{Creation of synthetic diamond for optics and electronics applications}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/HPHT_and_CVD}
+\caption{Schematics of HPHT (a) and CVD (b) growth process as detailed in main text}
+\label{HPHT and CVD}
+\end{figure}
+
+Almost all diamonds used in optics or electronics experiments today are synthetic, due to the higher control of the growth process required for these applications.
+There are two main competing technologies for the growth of large ($\mu \rm{m} \sim \rm{mm}$) single crystal diamonds: High pressure high temperature growth (HPHT) and chemical vapor deposition growth (CVD).
+Both of these processes have their own specificities which are detailed below.
+
+\subsubsection{HPHT growth}
+
+HPHT was the first commercially viable process for synthetic diamonds, starting from the 1950s \citep{barnard2000diamond, bundy1955man}.
+The basic idea of HPHT synthesis is to reconstruct an environment where diamond is the thermodynamically stable phase of carbon, similarly to the growth process of natural diamond below the earth surface.
+
+Fig.
+\ref{HPHT and CVD}-a) illustrates the apparatus of HPHT synthesis: a carbon source (graphite) and a solvent metal composed of transition metals (Fe, NI, Co typically \citep{bundy1963direct}) are put under $5 \sim 6$ GPa pressure in a press and heated to temperatures $\geq 1300\ ^\circ$C.
+A diamond crystal seed is put in the metal melt and a temperature gradient insures that the carbon migrates from the graphite to the diamond.
+
+The first synthesis of HPHT diamonds were all of type Ib due to unwanted nitrogen pollution in the metal melt.
+The first HPHT type IIa diamonds were produced in the late 90s by adding other elements to the metal bath (Ti, Al, Zr) which would react preferentially with nitrogen and prevent its incorporation in the diamond \citep{burns1999growth, sumiya2002growth}.
+
+\subsubsection{CVD growth}
+\label{fab HPHT}
+
+Diamond growth by CVD also started in the 50s, but only really took off in the 80s \citep{matsumoto1982growth, matsumoto1982vapor, kamo1983diamond} with the introduction of atomic hydrogen catalysis.
+Unlike HPHT, CVD operates in a regime where diamond is not the thermodynamically stable phase.
+It relies therefore heavily on catalysts to stabilize the diamond phase versus the graphite one.
+
+The main catalyst used is atomic hydrogen because it etches the $C-C$ sp$^2$ bonds (graphite) faster than it does sp$^3$ bonds (diamond) \citep{gracio2010diamond}.
+In modern CVD growth, hydrogen is by far the most abundant element in the gas phase: the precursor gases used for the growth are typically CH$_4$ and H$_2$ in proportions of $1-5\%$ and $95-99\%$ respectively \citep{achard2020chemical}.
+
+Fig.
+\ref{HPHT and CVD}-b) illustrates the apparatus of a microwave plasma reactor: it consists in a modified vacuum chamber which acts as a microwave resonator.
+The microwave creates a plasma ball from the the precursor gases with a temperature $T_{\rm plasma} \geq 3000 ^\circ$C \citep{ashfold2020nitrogen}, enough to dissociate the $H_2$ gas in atomic hydrogen \citep{balmer2009chemical}.
+The plasma acts as the carbon source for the diamond growing on top of the substrate, and as the catalyst to favor the diamond phase.
+%The substrate is kept at a temperature $700 \sim 1100\ ^\circ$C during the growth.
+
+Other sources can be used to create the high temperature required for atomic hydrogen, such as a tungsten hot filament \citep{haubner1993diamond}, electrical discharge (arcjet) \citep{luque1998excited} or even an oxygen-acetylene combustion flame \citep{bachmann1991towards}.
+
+If the CVD reactor is well built, the main sources of extrinsic impurities in the diamond comes from the source gasses \citep{balmer2009chemical}.
+The commercial availability of source gases containing less than $\sim$ ppm impurities made possible the growth of  high purity type IIa CVD diamonds \citep{kasu2003high, twitchen2004high, tallaire2006characterisation}.
+
+\medskip
+Both HPHT and CVD processes can produce diamonds of high quality for quantum optics or electronics application, each with their own pros and cons.
+HPHT offers better scalability for thick crystals ($\sim \rm mm$) and allows the incorporation of heavy atoms (Ge \citep{palyanov2015germanium} , Sn \citep{ekimov2019effect}) which can be hard to incorporate from a gas phase.
+CVD on the other hand allows more versatility in the control of the impurities introduced in the crystal, including an isotopic control on both the impurities and the carbon itself.
+A particular advantage of CVD growth is the possibility of $\delta$-doping where a specific defect is grown only within a thin layer of the crystal, as low as 5 nm \citep{ohno2012engineering, ishikawa2012optical, ohashi2013negatively}.
+
+\subsection{Optimizing NV center formation}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/formtion_NV_2}
+\caption{Illustration of the formation of NV centers. Black and green dots represents carbon and nitrogen atoms respectively}
+\label{formation NV}
+\end{figure}
+
+NV centers are a relatively rare impurity in untreated diamonds.
+In CVD diamond for instance, the proportion of nitrogen atoms forming an NV center in an untreated crystal is at best of $2-3 \%$ \citep{hartland2014study}.
+While this ratio can be enough  to work with single NV centers, applications that require dense ensemble need a higher conversion rate.
+We will discuss here the usual techniques employed to improve the formation of NV centers.
+
+Fig. \ref{formation NV} illustrates the steps needed to maximize the number of NV centers in the crystal while keeping a relatively high crystalline purity.
+
+The first step to create NV centers is to get individual substitutional N atoms in the crystal ($N_s$, C-centers or P1 centers depending on the literature).
+This can be achieved either \textit{in situ} by introducing (voluntarily or not) nitrogen in the crystal during the growth process \citep{tallaire2006characterisation, lobaev2017influence}, or \textit{ex situ} by implanting N$^+$ ions with ion implantation techniques \citep{meijer2005generation, smith2019colour}.
+Because this present work focuses on dense NV ensembles, we will consider the simplest case where the starting crystal is a type-Ib diamond because of the presence of nitrogen during the growth.
+A starting concentration [N$_s$]= $10 \sim 300$ ppm is typical for type Ib HPHT diamonds \citep{achard2020chemical}.
+
+The second step is to create monovacancies in the crystal lattice.
+This step might be optional for crystals which already contain a large number of monovacancies, but CVD crystals in particular can have less vacancies that they have substitutional nitrogen atoms \citep{mainwood1999point}.
+The general technique to create monovacancy is to bombard the sample with high energy particle in order to knock carbon atoms out of the crystal.
+Several high energy particles have been used to create vacancies, such as protons, electrons, ions, neutrons and even gamma rays \citep{davies1976optical, ashbaugh1988gemstone, kleinsasser2016high}.
+The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their higher availability (compared to neutrons or gamma rays) and their ability to deeply penetrate the crystal.
+
+The final step is to move the vacancies next to the nitrogen atoms.
+The activation energy required to move vacancy is of $\sim 2.3\ \rm eV$ which corresponds to temperatures of $600\sim 700 ^\circ$C \citep{davies1992vacancy, newton2002recombination}.
+The most common method to move the vacancies is therefore to anneal the crystal at a temperature $800 \sim 1200 ^\circ$C for a few hours \citep{botsoa2011optimal}.
+When the vacancy is next to a nitrogen atom, it will only become mobile again for temperatures of $1400 \sim 1500 ^\circ$C \citep{zaitsev2013optical, pinto2012diffusion} which explains why the annealing process favors NV creation.
+
+The final N$_s$ to NV conversion rate can be as high as $\sim 50 \%$ \citep{grezes2015storage, hartland2014study}, although this ratio tends to decrease for higher [N$_s$] concentrations.
+Most sample studied in this manuscript had a concentration $[\rm NV]= 1-10\ \rm ppm$ (see sample details in Appendix \ref{Appendix samples})
+
+%\subsection{Controlling the NV center charge state}
+%
+%The charge state of the NV center is of crucial importance for NV applications.
+Only the negatively charged NV$^-$ center shows the specific spin properties that will be detailed below, but the NV center can exist under three charge states in the diamond: NV$^-$, NV$^0$ and NV$^+$.
+The presence of NV$^+$ is usually marginal \citep{hauf2014addressing, pfender2017protecting} and we will focus here solely on the NV$^-$ and NV$^0$ charge state.
+%
+%The charge state of the NV centers depends not only on the growth process of the crystal, but also on the experiment performed with the sample.
+In particular, illumination in the optical range can promote electrons in the conduction band or holes in the valence band and modify the charge state of the NV center.
+%
+%On the material side, the crystal has to remain electrically neutral which means that at least as many charge donors need to be present as there are NV$^-$ centers.
+For nitrogen-rich samples, the donors are mostly N$^+$ centers, which imposes a limit on the maximum conversion rate of N to NV in order to keep a high [NV$^-$]/[NV$^0$] ratio.
+The possibility to introduce other donors, such as phosphorus, has been studied \citep{doi2016pure} but remains impractical so far \citep{barry2020sensitivity}.
+%
+%On the experimental side, it has been shown that the ionization of NV$^-$ to NV$^0$ and the reverse process depend on the intensity and wavelength of the laser used for illumination \citep{aslam2013photo}.
+It was found that the optimal wavelength to preserve the NV$^-$ charge state was  $510-540\ \rm nm$.
+The most common wavelength used with NV centers is 532 nm (green) due to the historical importance of Nd:YAG lasers.
+On the other hand, it is possible to semi-permanently ionize the NV$^-$ into NV$^0$ when using a red excitation, which has been used to create optical memories \citep{dhomkar2016long}.
+%
+%For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one.
+For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach up to 50\% of the total NV concentration \citep{grezes2015storage}
+
+\section{The NV$^-$ center optical and spin properties}
+
+This section focuses on the NV$^-$ center properties both as an optical defect (colored center) and as a spin defect, and compare them to other solid state systems.
+It also covers the interplay between the spin and the optical levels, a property almost unique to the NV$^-$ center which is the basis for optical polarization and readout of the NV$^-$ spin state.
+
+\subsection{NV center molecular orbitals}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{chapter 1/Figures/NV_electrons}
+\caption{(a) Representation of the nitrogen and carbon orbitals surrounding the vacancy of an NV center.
+Each black dot represents an electron.
+A sixth electron is required to form the negatively charged NV$^-$ state.
+(b) Molecular orbitals of the NV center predicted from the the $C_{3v}$ symmetry of the defect.
+The electron occupation of the NV$^-$ ground state is represented by red arrows.}
+\label{NV electrons}
+\end{figure}
+
+We will first cover the molecular aspect of the NV defect in diamond and the energy levels of the defect predicted from group theory and ab initio computations.
+
+Fig.
+\ref{NV electrons}-a) illustrates the $sp^3$ orbitals of the carbons and nitrogen atoms neighboring the vacancy, as well as the 5 electrons occupying dangling bonds from these atoms and a sixth electron captured to form the negatively charged NV$^-$ state.
+
+Fig.
+\ref{NV electrons}-b) shows the predicted molecular orbitals of the NV centers, initially computed as linear combination of the $sp^3$ orbitals respecting the $C_{3v}$ symmetry of the defect \citep{loubser1978electron}.
+This model has since been confirmed by ab initio computation and experimental observations \citep{doherty2013nitrogen}.
+In particular ab initio computations have shown that the molecular orbitals of the NV center are mostly localized on the nearest neighbors of the vacancy \citep{gali2008ab}.
+Ab initio computation has also determined that the $a'_1$ level lies in the valence band and is thus often ignored.
+
+The electron occupations shown in Fig. \ref{NV electrons}-b) corresponds to the electronic ground state of the NV$^-$ charge state.
+
+In the remaining part of this manuscript, we will only consider the negatively charged NV$^-$ state, unless specified otherwise.
+We will therefore omit the ``$-$" from now on and simply refer to the negatively charged state as ``NV".
+\subsection{Optical properties of the NV center}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1\textwidth]{chapter 1/Figures/NV_optical}
+\caption{Optical properties of the NV center.
+(a) Electron occupation of the molecular orbitals for the electronic ground state $^3A_2$ and the first excited state $^3E$.
+(b) Diagram of the main optical transition between the two spin triplet states and of the IR transition between the two singlet sates.
+The value of the transition corresponds to the zero phonon line (ZPL) and are taken from \citep{doherty2013nitrogen}.
+(c) Representation of the vibronic structure for the $^3A_2$ and $E^3$ states under the Franck-Condon approximation.
+The graph axes are the energy of the states $E$ and configuration coordinate $q$.
+The green and red arrow represents non-resonant (Stokes) absorption and emission processes.
+(d) Emission and absorption spectrum of the NV center main transition at 300 K, from PL and PLE measurements respectively.
+Credits: Wikimedia commons}
+\label{NV optical}
+\end{figure}
+
+Because the NV center has discrete electronic transitions inside the diamond band gap, it fluoresces: it can absorb light from $450-650$ nm and re-emit light in the $600-800$ nm band.
+
+Fig.
+\ref{NV optical}-a) shows the electronic ground state $^3A_2$ and first excited state $^3E$, where 1 electron from the $a_1$ molecular orbital is promoted to the $e$ level.
+Both of these electron occupations have two unpaired electrons, meaning that they can exist either either in a spin triplet state ($S=1$) or a spin singlet state ($S=0$) \footnote{The two singlet states $^1E$ and $^1A_1$ represented in Fig.
+\ref{NV optical}-b) are both of configuration $a_1^2e^2$ where the two $e$ electrons are either in a the same $e$ orbital, or in an antisymmetric superposition of the two $e$ orbitals \citep{doherty2011negatively}.}.
+
+Fig.
+\ref{NV optical}-b) shows the zero phonon line (ZPL) for the main transition of the triplet ($^3A_2$ and $^3E$) and singlet ($^1E$ and $^1A_1$) states.
+It should be noted that due to an inter system crossing between the $^1E$ and $^3A_2$ levels, excitation or absorption of the singlet transition is only possible with a conjoint excitation of the triplet states.
+By default, we refer to the NV center photoluminescence (PL) as the emission between the two triplet states.
+
+Fig.
+\ref{NV optical}-c) illustrates the vibronic structure of the $^3A_2$ and $^3E$ states under the Franck-Condon approximation \citep{gali2011time}.
+The emission and absorption of light associated with the triplet transition can be accompanied by the emission or absorption of an optical phonon.
+Because the lifetime of the optical phonons is very short, the absorption or emission of light almost always starts from the fundamental state of the $^3A_2$ and $^3E$ level, which explains why the absorption of light can only be done for energies above the ZPL, and the emission for energies below the ZPL (Stokes process).
+
+Finally Fig.
+\ref{NV optical}-d) shows the absorption and emission spectrum of NV centers at 300 K.
+A weak zero-phonon-line is visible at 637 nm, with large phonon sidebands in emission and absorption.
+
+\medskip
+Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages.
+First, they are very bright thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and thanks to a quantum efficiency of $0.7-0.8$ \citep{schirhagl2014nitrogen}.
+With modern nanophotonics methods (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}.
+Secondly, the NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}.
+Finally, the diamond itself is resistant to heat and chemical attacks, and presents low toxicity for biologic systems \citep{fu2007characterization}.
+There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (low Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}.
+This limits the potential of NV centers for quantum photonic devices \citep{bradac2019quantum}.
+
+\subsection{Spin properties of the NV center electronic ground state}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1\textwidth]{chapter 1/Figures/spin_fine_hyperfine}
+\caption{Representation of the fine and hyper-fine structure of the $^3A_2$ state for a $^{14}$N nucleus and $^{12}$C nuclei.}
+\label{NV spin}
+\end{figure}
+
+We will only focus here on the spin structure of the electronic ground state $^3A_2$, as the lifetime of the other states is too short to make effective spin qubits.
+All the numerical values in this section were found in \citep{smeltzer2009robust, doherty2013nitrogen} and the references within.
+
+The spin structure in the ground state consists in an electronic spin $S=1$ coupled to the nuclear spin of the nearby nuclei.
+Carbon's most common isotope is $^{12}$C at 98.9 \% natural abundance.
+$^{12}$C has a a nuclear spin $I=0$ and therefore does not contribute to the hyper-fine structure.
+Nitrogen's most common isotope however is $^{14}$N at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does create a hyper-fine splitting.
+We will only cover here the most likely configuration for an NV center which is a $^{14}$N nucleus and $^{12}$C nuclei as closest neighbors.
+
+The spin system considered is then an electronic spin $S=1$ coupled to a nuclear spin $I=1$ with a Hilbert space of dimension $(2S+1)(2I+1)=9$. The spin Hamiltonian of the system can be written:
+\begin{equation}
+\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S}\cdot \bar{\bar{A}}\cdot \mathbf{I},
+\end{equation}
+where $\mathcal{H}_e$ is the purely electronic spin Hamiltonian, $\mathcal{H}_n$ the purely nuclear spin Hamiltonian, $\mathbf{S}$ the electronic spin operator, $\mathbf{I}$ the nuclear spin operator and $\bar{\bar{A}}$ the hyper fine tensor.
+\subsubsection{Electronic spin Hamiltonian}
+The electronic spin Hamiltonian $\mathcal{H}_e$ can be written:
+\begin{equation}
+\label{eq. spin elec}
+\mathcal{H}_e=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} + \mathcal{H}_{elec}(\mathbf{E})+ \mathcal{H}_{strain}(\mathbf{\bar{\bar{\varepsilon}}}).
+\end{equation}
+
+The term $D S_z^2$ corresponds to the fine structure of the electronic spin.
+It originates from the dipole-dipole coupling between the two unpaired electrons of the defect.
+Due to the strong anisotropy of the molecular orbitals of the defect, the fine structure term imposes the quantization axis $\mathbf{z}$ (via the $S_z$ operator) where $\mathbf{z}$ is the direction of the N$-$V axis in the diamond.
+$D$ is called the zero field splitting (ZFS) factor and its value is $D\approx 2.87\ \rm GHz$ at 300 K.
+
+$\gamma_e \mathbf{S} \cdot \mathbf{B}$ is the Zeeman splitting of the spin, where $\gamma_e \approx 2.8\ \rm MHz/G$ is the gyromagnetic ratio of the electronic spin (almost equal to that of a free electron in the case of the NV center).
+
+$\mathcal{H}_{elec}$ and $\mathcal{H}_{strain}$ are the dependency of the Hamiltonian on the electric field and crystal strain and will be discussed in further details in chapter 4.
+
+\subsubsection{Nuclear spin Hamiltonian and hyperfine coupling}
+The nuclear spin $\mathcal{H}_n$ can be written:
+\begin{equation}
+\mathcal{H}_e=Q I_z^2 + \gamma_n \mathbf{I} \cdot \mathbf{B},
+\end{equation}
+
+where $Q=-4.945\ \rm MHz$ is the nucleus electric quadrupole moment and $\gamma_n=-0.308\ \rm kHz/G$ is the nuclear gyromagnetic ratio.
+
+Finally the hyper-fine tensor can be written diagonally:
+\begin{equation}
+\bar{\bar{A}} = \begin{pmatrix}
+A_{xx} & 0 & 0 \\
+0 & A_{yy} & 0 \\
+0 & 0 & A_{zz}
+\end{pmatrix},
+\end{equation}
+where $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$.
+The $z$ direction in the electronic spin Hamiltonian, the nuclear spin Hamiltonian and the hyper-fine tensor awlays refer to the direction of the N$-$V axis.
+
+Fig. \ref{NV spin} shows fine and hyper-fine structure of the NV center without any external fields.
+
+The hyper-fine splitting as well as the electric and strain dependency are of second order compared to the Zeeman and zero field splitting terms.
+Unless specified otherwise, we will from now on refer to the following simplified Hamiltonian as the NV spin Hamiltonian:
+
+\begin{align}
+\label{NV spin Hamiltonian basic}
+\mathcal{H}&=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} \\
+&=\begin{pmatrix}
+D-\gamma_e B_z & \frac{\gamma_e (B_x+iB_y)}{\sqrt{2}} & 0 \\
+\frac{\gamma_e (B_x-iB_y)}{\sqrt{2}} & 0 & \frac{\gamma_e (B_x+iB_y)}{\sqrt{2}} \\
+0 & \frac{\gamma_e (B_x-iB_y)}{\sqrt{2}} & D+\gamma_e B_z
+\end{pmatrix}
+\end{align}
+
+\bigskip
+The NV center's spin use as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}.
+Part of the reason for this wide usage are the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
+However the main advantage of the NV center spin compared to other solid state spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.
+
+\subsection{Intersystem crossing of the electronic levels}
+\label{sec ISC}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{chapter 1/Figures/ISC}
+\caption{Inter system crossing of the NV center.
+Blue arrows symbolize decay channels to and from the $\ket{m_s=\pm 1}$ states and red arrows the decay channels to and from the $\ket{m_s=0}$ states.
+Green arrows symbolize the optical excitation induced by the laser.
+The indicated durations correspond to the inverse of the decay rates found in \citep{gupta2016efficient}.}
+\label{ISC}
+\end{figure}
+
+The optical excitation and radiative decay between the $^3A_2$ and $^3E$ levels preserve the spin projection $m_s$ \citep{robledo2011spin}.
+There is however a non-radiative decay path which involves an inter-system crossing with the singlet states $^1A_1$ and $^1E$ that depends on the the triplet states spin projection.
+
+Fig.
+\ref{ISC} summarizes the possible decay channels from the $^3E$ states.
+Once excited to the $^3E$ states, for instance by photoexcitation, the electrons can either decay radiatively and emit a 600$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states.
+The decay though the singlet state is often referred to as non-radiative, due to both the difference in wavelength between the triplet and the singlet transitions, and due to the strong non-radiative competition in the singlet decay (quentum efficiency $\leq 10^{-3}$ \citep{rogers2008infrared, ma2010excited, acosta2010optical}.
+
+The rates $\gamma$ of the various decay processes, represented in Fig.
+\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depend on the spin projection $m_s$.
+Importantly, the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$\frac{15+91}{15+11}\approx 4$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
+Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
+
+These spin selective intersystem crossings mean that an optical pumping scheme in the $\ket{m_s=0}$ spin state takes place under constant green excitation.
+At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}, which is equivalent to a thermal polarization (for B=0) with a spin temperature of $\approx 65\ \rm mK$ \footnote{\begin{equation*}
+T=\frac{h D}{k_B \ln (\frac{1-P}{2P})} \approx 65\ \rm mK,
+\end{equation*}
+where $D$ is th ZFS factor described in eq.
+(\ref{eq.
+spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.} while the diamond itself can remain at room temperature.
+
+Another effect of the intersystem crossing is the ability to optically readout the spin state of the NV center: since the $\ket{0}$ state is more likely to undergo a radiative decay than the $\ket{\pm 1}$ states, the expected photoluminescence (PL) is higher when starting from the $\ket{0}$ state than a $\ket{\pm 1}$ state.
+While the $\ket{+1}$ and $\ket{-1}$ states cannot be a priori distinguished, a direct excitation of the $\ket{+1} \to \ket{0}$ or $\ket{-1} \to \ket{0}$ transition with a resonant microwave can unambiguously reveal the initial spin state.
+
+It should be noted that the decay rates indicated in Fig.
+\ref{ISC} cannot be directly measured.
+They have to be inferred from the PL dynamics at short times after illumination.
+Different experiments and different models have thus lead to different decay rates \citep{duarte2021effect}.
+
+\section{Basic operations with NV centers}
+
+This section introduces two elementary operations with NV centers that are Rabi oscillations and optically detected magnetic resonance (ODMR).
+
+Rabi oscillations is the basis for any kind of coherent qubit manipulation while ODMR is used to find the transition frequencies of the NV centers, as well as being a basic magnetometry protocol.
+
+This section also covers the experimental setup used for these two operations and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
+
+\subsection{Experimental setup}
+\label{Sec setup cahp 1}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{chapter 1/Figures/shema_exp}
+\caption{Confocal microscopy setup detailed in main text}
+\label{setup ch1}
+\end{figure}
+The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field.
+Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.
+
+Fig.
+\ref{setup ch1} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
+Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.
+
+The setup consists of a confocal microscope where the excitation from a green laser (Thorlabs CPS532, 532 nm, 5 mW) is focused on a sample containing NV centers with a microscope objective (NA $\sim$ 0.5).
+The NV PL is then collected by the same objective and isolated from the back-scattered laser light with a dichroic mirror (DM) and additional filters (532 nm notch filter and 645 nm longpass filter to filter out the NV$^0$ PL).
+The PL is collected on an avalanche photodiode (Thorlabs APD410A or Perkin Elmer SPCM-AQRH-15) whose signal is sent to the acquisition card (NI X-series).
+
+The laser beam goes through an acousto-optic modulator (AOM, AA opto-electronic) controlled externally to create arbitrary light pulses (resolution $\sim$ ns).
+The microwave field is generated by a Rhode \& Shwarz SMB 100A generator and emitted with a home-made loop antenna positioned at $<$ 1 mm of the sample.
+A microwave switch (Mini-circuits ZASWA-2-50DRA+) is used to create microwave pulses (resolution $\sim$ 10 ns).
+The control pulses for the AOM and the microwave switch are generated either by the NI X-series card (resolution $\sim$ 1 $\mu$s), or by a pulseblaster ESR-PRO (resolution $\sim$ 10 ns).
+The microwave field can be amplified with a Mini-circuits ZHL-5W-422+ amplifier to achieve higher Rabi frequencies.
+
+Finally the magnetic field is generated either with a permanent magnet (up to $\sim$ 200 mT) or with a homemade electromagnet (up to $\sim$ 30 mT) alimented via a low freqency amplifier (Leybold power function generator 522663).
+Both the permanent magnet and the electromagnet can be  attached to a rotation stage for precise alignment.
+\subsection{Rabi Oscillations}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/Rabi}
+\caption{Rabi oscillations. (a) pulse sequence used to measure Rabi oscillations. (b) Experimental data on sample CVD-pink (see sample details in appendix \ref{Appendix samples}).}
+\label{Rabi}
+\end{figure}
+
+The general technique for experiments involving NV spins proceeds as follow: 
+\begin{itemize}
+\item First the spins are polarized. This is done with a green laser pulse ($1-100$ $\mu$s depending on the laser intensity) which polarizes the spins in the $\ket{m_s=0}$ state at $\sim$ 80 \%.
+\item  Then the spins are manipulated. This process often involves a microwave field resonant with one of the spin transitions.
+\item Finally the spins are read out.
+The simplest spin readout for NV centers consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
+For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore.
+The optimal readout time is of the same order as the spin polarization time ($1-100$ $\mu$s).
+\end{itemize}
+
+Fig.
+\ref{Rabi}-a) shows the pulse sequence used to observe Rabi oscillations with NV centers.
+In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$.
+The microwave frequency or the magnetic field needs to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
+
+Fig.
+\ref{Rabi}-b) shows the result of such an experiment on a spin ensemble.
+The PL signal oscillations correspond to the oscillations of the spin population between the bright $\ket{0}$ state and the dark $\ket{-1}$ state.
+For long microwave pulses ($>$ 1 $\mu$s typically), the spins reach a steady state with  $\sim$ 50 \% population in both the $\ket{0}$ and $\ket{-1}$ states.
+
+\subsection{Optically detected magnetic resonance}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_basique}
+\caption{ODMR protocol.
+(a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$.
+(b) ODMR spectrum with no magnetic field on sample ADM-15-4.}
+\label{ODMR basique}
+\end{figure}
+
+Optically detected magnetic resonance (ODMR) is the most commonly employed magnetometry protocol with NV centers due to its simplicity and the fact that it does not require prior calibration.
+
+The protocol for continuous wave (CW) ODMR is presented in Fig.
+\ref{ODMR basique}.
+In this simple protocol, the microwave is always turned on while the frequency spans a predefined interval $[f_{\rm start}, f_{\rm end}]$.
+The PL is recorded as the microwave is being swept, with a typical acquisition time for each PL sample $> 1\ \rm ms$.
+The sweeping frequency must be slow enough so that the PL can reach a steady state with the microwave field for each sample acquisition.
+
+As we saw in the last section, the steady state of the NV center spin with a resonant microwave field is a statistical mixture of 50 \% $\ket{0}$ and 50 \% $\ket{\pm 1}$ \footnote{We are assuming here that the Rabi frequency is greater than the laser polarizing rate, which is often the case in practice}.
+If the microwave field is not resonant with a spin transition however, the steady state is predominantly $\ket{0}$ due to the spin polarization caused by the laser.
+The NV PL will therefore be lowered when the scanning microwave field becomes resonant with one of the NV spin transitions.
+
+Fig.
+\ref{ODMR basique}-b) shows an example of ODMR spectrum on an NV ensemble with no external magnetic field.
+The reason for the two bumps instead of a single one centered on $D=2870\ \rm MHz$ is the splitting of the $\ket{+1}$ and $\ket{-1}$ levels occasioned by the strain and local electric field, which will be further discussed in chapter 4.
+
+\subsubsection{ODMR with lock-in detection}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_lockin}
+\caption{ODMR protocol with lock-in detection.
+(a) Simplified confocal microscope setup with a lock-in amplifier (LIA) modulating the mircowave and demodulating the PL.
+(b) Demodulated ODMR spectrum with no magnetic field on sample ADM-15-4.}
+\label{ODMR lockin}
+\end{figure}
+
+Because the ODMR protocol delivers a continuous signal (the PL) as a function of a periodic excitation (the microwave field), it is well suited to amplitude or frequency modulation and lock-in detection.
+
+The principle of lock-in detection will not be covered here, the key idea is that lock-in detection allows to translate the signal spectrally in a region less prone to noise.
+In this case, the ODMR signal is a DC measurement and is therefore prone to slow freqency noises such as thermal or mechanical drifts and laser amplitude fluctuations.
+The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low freqency noises, but slow enough compared to the NV photodynamics.
+
+Fig.
+\ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig.
+\ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
+All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1\ \rm kHz$.
+
+\subsection{ODMR with NV ensemble: longitudinal and transverse magnetic field}
+\label{sec champs transverse}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_8_classes}
+\caption{(a) Geometry of the NV center and definition of $B_z$ and $B_\perp$.
+(b) Representation of the 4 possibles orientations (``classes") of the N$-$V axis, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell.
+(c) Simulation of the evolution of the transition frequencies for each class of NV centers as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
+(d) ODMR spectrum on sample ADM-150-1 for a magnetic field of $\sim 90\ \rm G$.}
+\label{ODMR 8 classes}
+\end{figure}
+
+We mentioned previously that the NV spin Hamiltonian (eq.
+(\ref{NV spin Hamiltonian basic})) was anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
+The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig.
+\ref{ODMR 8 classes}-a).
+
+For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$.
+The true eigenstates however remain close to $\ket{0}$, $\ket{\pm 1}$ as long as $\gamma_e B_z > \frac{(\gamma_e B_\perp)^2}{D}$.} can be approximated in a perturtbative approach to: %cf AN.py
+\begin{align}
+\nu_0&= -\frac{(\gamma_e B_\perp)^2}{D} \\
+\nu_{-1} &= D - \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D} \\
+\nu_{+1} &= D + \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}
+\end{align}
+
+Fig.
+\ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal.
+We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian.
+For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and that each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
+
+Fig.
+\ref{ODMR 8 classes}-c) shows a simulation of the evolution of the transition frequencies $\nu_{+1} - \nu_{0}$ and $\nu_{-1} - \nu_{0}$ for all 4 classes as a function of the magnetic field amplitude, with the magnetic field orientation fixed in an arbitrary direction.
+Fig.
+\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles.
+8 lines are visible corresponding to the two transitions of the 4 classes.
+Based on the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.
+
+We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
+\ref{sec ISC}.
+For strong transverse magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
+
+\section{Dynamics of the NV center spin}
+
+This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times.
+Relaxation times are important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
+
+The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ which corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix), and the transverse relaxation time $T_2$\footnote{We follow here the NMR nomenclature} which corresponds to the relaxation of the spin state coherences (non diagonal elements of the density matrix).
+
+The transverse relaxation time, also called coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$.
+The $T_2^*$ time includes inhomogeneous dephasing processes such as the inhomogeneity in space (when multiple qubits are probed) and time of the external fields.
+The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}.
+While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used.
+More careful nomenclatures will label the $T_2$ times as $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance.
+The relaxation times are ordered as follows:
+\begin{equation}
+T_2^* < T_2 < 2 T_1.
+\end{equation}
+
+The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the sample used.
+The following paragraphs will cover the measurement processes of the different relaxation times and the typical values for the samples used in this manuscript.
+
+\subsection{NV spin $T_1$}
+\label{sec T1 NV}
+\subsubsection{Measurement protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/T1_basique}
+\caption{Basic $T_1$ measurement scheme.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink for low laser intensity.
+The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
+\label{T1 basique}
+\end{figure}
+
+$T_1$ is the simplest relaxation time to measure because the spin populations are directly correlated to the NV centers PL.
+Fig.
+\ref{T1 basique} shows the pulse sequence used and an experimental result of the $T_1$ sequence.
+The spin lifetime $T_1$ is one of the only spin property that can be measured without a resonant microwave.
+
+The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig.
+\ref{T1 soustraction}-a).
+This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$.
+Then the sequence is repeated with an additional $\pi$-pulse on one the NV transitions right before the readout.
+This $\pi$-pulse will project the remaining $\ket{0}$ population difference into a $\ket{\pm 1}$ state.
+The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which rejects every contribution to the PL outside of the spin relaxation.
+
+Fig.
+\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample that was used in Fig.
+\ref{T1 basique}.
+The difference between the $S_1$ data in Fig.
+\ref{T1 soustraction}-b) and the the data in Fig.
+\ref{T1 basique}-b) comes from the laser intensity used.
+In Fig.
+\ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant.
+Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark \citep{giri2018coupled, giri2019selective} (most likely due to charge tunneling between impurities \citep{choi2017depolarization}) .
+On the other hand, the laser intensity used in Fig.
+\ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.
+
+Fig.
+\ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal.
+Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig.
+\ref{T1 basique}.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/T1_protocole_soustraction}
+\caption{Common mode rejection $T_1$ measurement scheme.
+(a) Pulse seqeunce used.
+(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity.
+(c) Subtraction $S_1-S_2$ versus dark time $\tau$.
+The experimental data is fitted with an exponential decay of time constant $\tau=4.1\ \rm ms$.}
+\label{T1 soustraction}
+\end{figure}
+
+The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
+The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.
+
+\subsubsection{Physical origin and typical values}
+
+We distinguish three different origin for the the $T_1$ relaxation process:
+\begin{itemize}
+\item \textbf{Spin-phonon relaxation:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which can result in a phonon-induced spin relaxation.
+The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}.
+The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxtion times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at for mili-K temperatures \citep{astner2018solid}.
+
+\item \textbf{Electric and magnetic field noises:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}.
+As a result, magnetic and electric field noise can reduce the spin $T_1$ if their spectrum overlap with the NV transitions.
+This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}.
+The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effects are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
+
+\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipolar interaction betwen two spins will be detailed more thoroughly in the following section.
+Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation.
+The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration and can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
+\end{itemize}
+
+All the measurements shown in this manuscript were performed at room temperature.
+The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$.
+Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses.
+The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed.
+$T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.
+
+\subsection{NV spin $T_2^*$}
+\subsubsection{Measurement protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/Ramsey}
+\caption{Ramsey interferometry protocol.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink.} %The orange line corresponds to an Gaussian fit with time constant $\tau=400 \pm 100\ \rm ns$.}
+\label{Ramsey}
+\end{figure}
+
+The $T_2^*$ time is the characteristic time for the loss of quantum coherence of a freely evolving system.
+It is generally defined as the decay time of of a Ramsey interferometry measurement \citep{barry2020sensitivity}.
+
+Fig.
+\ref{Ramsey} shows the Ramsey pulse sequence and an  example obtained on a spin ensemble.
+The microwave frequency is detuned by $\sim 300\ \rm kHz$ compared to the spin central Larmor frequency which causes oscillations in the signal.
+$T_2^*$ is the decay time of the envelope of the Ramsey fringes.
+Unfortunately, it is hard to correctly fit the envelope of a signal with so few fringes.
+Increasing the number of fringes with our current microwave and optical setup might prove hard, as further microwave detuning would decrease the signal contrast, and would be ultimately limited by the microwave pulses resolution.
+
+Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
+The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity, and because the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
+The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
+
+An alternative measurement of $T_2^*$ can be performed by looking at the linewidth of an ODMR spectrum.
+Indeed, ideally the ODMR spectrum should be the Fourier transform of the Ramsey signal (also called free induction decay in NMR literature).
+In practice however, both methods suffer from experimental noises and imperfections, with Ramsey-type methods being more sensitive to microwave inhomogeneities \citep{barry2020sensitivity} while ODMR is prone to laser and microwave power broadening \citep{dreau2011avoiding}.
+
+The method chosen to extract $T_2^*$ values in this manuscript was ODMR with low microwave power ($\Omega_{\rm Rabi} \ll 1/T_2^*$) and low laser intensity (compared to the the NV optical saturation).
+
+\subsubsection{Physical origin}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_1_classes}
+\caption{ODMR spectra on a single NV class.
+(a) Experimental data on sample CVD-pink, fitted with three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
+(b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
+\label{ODMR 1 classe}
+\end{figure}
+
+For most samples, the leading cause of inhomogeneous dephasing is the temporal and spatial variations of magnetic fields caused by the paramagnetic impurities in the crystal \citep{barry2020sensitivity}.
+The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2.
+We will simply mention here that for sample with high NV density ([NV] $\geq$ ppm), $T_2^*$ is mostly limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
+
+Fig.
+\ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see sample details in appendix \ref{Appendix samples}).
+The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
+We can clearly see a higher linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
+In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
+
+Fig.
+\ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula.
+While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}.
+For Fig.
+\ref{ODMR 1 classe}-a), it was indeed found that Lorentzians fitted the data better than Gaussians.
+For Fig.
+\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, although this could be in part due to the unresolved hyperfine structure.
+
+\subsection{NV spin $T_2$}
+\label{sec T2 echo}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/Echo}
+\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
+\label{Echo}
+\end{figure}
+\subsubsection{Measurement protocol}
+As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.
+
+Fig.
+\ref{Echo} shows the pulse sequence used and an example on the same sample than the Ramsey measurement.
+The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time.
+This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased \citep{hahn1950spin}.
+
+This protocol extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
+Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
+This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
+
+\subsubsection{Physical origin}
+The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence.
+The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}.
+For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2019decoherence}.
+
+\section{Dipole-dipole interaction between two electronic spins}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.4\textwidth]{chapter 1/Figures/shema_dipole_dipole}
+\caption{Geometric representation of two interacting spins.
+($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$) represent Cartesian coordinate systems in the spin frames of reference, where $\mathbf{z}_i$ is the axis of quantization of spin $i$.}
+\label{dipole-dipole}
+\end{figure}
+
+This final introductory section will focus on the dipole-dipole interaction between two electronic spins, which is the main form of spin-spin interaction and plays a key part in the properties of dense spin ensemble.
+
+We will consider the general situation represented in Fig.
+\ref{dipole-dipole}: two spins separated by a distance $r$ with their own coordinate systems ($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$).
+The coordinate system is chosen so that each spin operator can be written:
+
+\begin{equation}
+\hat{\mathbf{S}}_i=\hat{S}_x \mathbf{x}_i + \hat{S}_y \mathbf{y}_i + \hat{S}_z \mathbf{z}_i
+\end{equation}
+
+where the spin operators $\hat S_{\{x,y,z\} }$ have their usual definitions in term of Pauli matrices.
+In particular, this means that the spin axis of quantization of each spin is $z_i$, which classically corresponds to the magnetization direction of the dipole.
+This axis can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (including the NV center).
+
+\subsection{Magnetic dipole-dipole Hamiltonian}
+
+Classically, the dipole-dipole interaction between two magnetic moments $\mathbf{m}_1$ and $\mathbf{m}_2$ can be derived by computing the static magnetic field generated by one dipole at the position of the second dipole \cite[p.~188]{jackson1999classical}:
+
+\begin{align}
+U&=-\mathbf{m}_1 \cdot \mathbf{B}(\mathbf{r}) \nonumber \\
+&=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u}) - \mathbf{m}_1\cdot\mathbf{m}_2}{r^3}+\frac{8\pi}{3}\mathbf{m}_1\cdot\mathbf{m}_2\,\delta(r)\right], \label{eq.
+dipole classique}
+\end{align}
+
+where $\mathbf{r}=r \mathbf{u}$ is the relative position of the two dipoles (see Fig. \ref{dipole-dipole}).
+
+The quantum analog of the dipole-dipole interaction between spins simply replaces the magnetic moments $\mathbf{m}_i$ in eq.
+(\ref{eq.
+dipole classique}) with $\hbar \gamma_i \hat{\mathbf{S}}_i$, where $\gamma_i$ is the gryromagnetic ratio and $\hat{\mathbf{S}}_i$ the spin vector operator for spin $i$:
+
+\begin{equation}
+\label{eq. dipole quantique J coupling}
+\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2\,\delta(r)\right].
+\end{equation}
+
+The $\delta(r)$ term represents the Fermi contact energy.
+This term plays an important role when there is an overlap of the spin particle wavefunctions, such as the hyperfine coupling between the NV$^-$ electron spin and $^{14}$N \cite{doherty2012theory} or  $^{13}$C \cite{smeltzer201113c} nuclear spins.
+
+The focus of this manuscript is on the interaction between electronic spins which are separated by several atomic sites (density impurities in the $\sim$ ppm range).
+We will therefore assume that there is negligible overlap between the electronic wavefunctions and omit the Fermi contact term.
+This leads the simplified dipole-dipole Hamiltonian:
+
+\begin{equation}
+\mathcal{H}_{\rm dd}= \frac{J_0}{r^3}\left[ 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2] \right],
+\end{equation}
+
+where $J_0=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}$.
+
+For two electronic spins with $g$-factors $\approx -2$\footnote{Most spin defects in diamond, including the NV center have a g-factor $\approx -2$ similarly to free electrons.}, 
+\begin{equation*}
+\frac{J_0}{h} \approx 51.8\ \rm{MHz}\cdot \rm{nm}^3.
+\end{equation*}
+
+\subsection{Decomposition of the dipole-dipole Hamiltonian}
+\label{sec flip-flop et double flip}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 1/Figures/flip-flop_double-flip}
+\caption{Representation of the two energy transfer mechanisms: flip-flop and double flips} 
+\label{flip flop double flip}
+\end{figure}
+
+In the weak coupling approximation, the dipole-dipole Hamiltonian between two spins can be decomposed between diagonal terms which are responsible for energy shifts of the individual spin levels, and the off-diagonal terms which are responsible for energy transfers between the two spins.
+
+In what follows, we will assume that both spins have the same quantization axis in order to simplify the Hamiltonian decomposition. A more general case is detailed in appendix \ref{Appendix eta}.
+
+Following the notation introduced in Fig.
+\ref{dipole-dipole}, we will then assume that $$(\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1) = (\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2) = (\mathbf{x},\mathbf{y},\mathbf{z}).$$ We will denote $\mathbf{u}=u_x \mathbf{x} + u_y \mathbf{y} + u_z \mathbf{z}$ the decomposition of the unitary vector $\mathbf{u}$ in this basis.
+The dipole-dipole Hamiltonian can then be written:
+\begin{equation}
+\mathcal{H}_{\rm dd}=\frac{J_0}{r^3}\left[ (3 u_x^2-1) S_x^1 S_x^2 + (3 u_y^2-1) S_y^1 S_y^2 + (3 u_z^2-1) S_z^1 S_z^2 \right],
+\end{equation}
+where $S_{\{x,y,z\} }^i$ are the spin operators acting on spin $i$.
+The $S_z^1 S_z^2$ term is the only diagonal term and corresponds to energy shifts while the $S_x^1 S_x^2$ and $S_y^1 S_y^2$ terms are responsible for energy transfer.
+To make the transfer mechanism more apparent, we will rewrite $S_x^i$ as $\frac{1}{2}(S_+^i+S_-^i)$ and $S_y^i$ as $\frac{1}{2i}(S_+^i-S_-^i)$:
+\begin{align}
+\mathcal{H}_{\rm dd}&=\frac{J_0}{r^3}\bigg{[} \frac{1}{4}(3 u_x^2 + 3 u_y^2 -2)(S_+^1 S_-^2 + S_-^1 S_+^2) \nonumber \\
+&\quad + \frac{1}{4}(3 u_x^2 - 3 u_y^2)(S_+^1 S_+^2 + S_-^1 S_-^2) \nonumber \\
+&\quad +(3 u_z^2-1) S_z^1 S_z^2 \bigg{]} \nonumber \\
+\medskip
+&= \Omega_{\rm ff} (S_+^1 S_-^2 + S_-^1 S_+^2) + \Omega_{\rm df} (S_+^1 S_+^2 + S_-^1 S_-^2) + \Omega_{\rm shift} S_z^1 S_z^2. \label{eq. ff df}
+\end{align}
+
+We have introduced in eq.
+(\ref{eq.
+ff df}) the three relevant matrix elements $\Omega_{\rm ff}$, $\Omega_{\rm df}$, and $\Omega_{\rm shift}$ which correspond respectively to flip-flops, double flips and energy shifts.
+Flip-flops and double flips (sometimes referred as flip-flips) are represented in Fig.
+\ref{flip flop double flip}.
+Flip flops correspond to energy transfer between two spins for which the sum of the spin angular momentum is conserved ($\Delta (m_s^1+m_s^2)=0$), while double flips correspond to energy transfer for which the spin angular momentum is not conserved ($\Delta (m_s^1+m_s^2)=\pm 2$)\footnote{In order to preserve the total angular momentum, double flips transfer angular moment to the crystal, which is a form of Einstein de Haas effect \citep{einstein1915experimental} }.
+In both cases however, the energy between the initial and final states needs to be preserved for the transfer to be efficient.
+
+It should be noted that on average, each matrix element is of the same order of magnitude: $\Omega_{\rm ff} \sim \Omega_{\rm df} \sim \Omega_{\rm shift} \sim \frac{J_0}{r^3}$.
+
+\subsection{Cross-relaxations with NV centers}
+\label{Sec_CR}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 1/Figures/Shema_CR}
+\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
+\label{CR_shema}
+\end{figure}
+
+Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
+This process can occur through either flip-flops or double flips, as long as the resonance condition between the two spins is met.
+
+On of the condition for cross-relaxation is to have a difference in polarization between the two spins species considered.
+The NV center is therefore a good candidate to observe cross-relaxations as it can be optically polarized while most other spin species are not.
+
+Fig.
+\ref{CR_shema} illustrates CR between a polarized NV center and an unpolarized second spin.
+When the two considered spin transitions become resonant, in this case by tuning the magnetic field, polarization from the NV center will be transferred to the second spin, making the NV center less polarized and the second spin more polarized than when they are out of resonance.
+
+For a single NV center coupled to spin 2, assuming that spin 2 always remains depolarized\footnote{This is the case in practice if the spin 2 bath is much bigger than the NV bath and effectively acts as a thermostat, or if the lifetime of spin 2 is much smaller than that of the NV center.} and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$ (weak coupling regime), the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by the cross-relaxation with spin 2 can be analytically computed \citep{hall2016detection}:
+\begin{equation}
+\delta \Gamma_1=\frac{\Omega^2}{\Gamma_2^*} \frac{(\Gamma_2^*)^2}{(\delta \nu)^2+(\Gamma_2^*)^2},
+\label{delta gamma 1}
+\end{equation}
+where $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, $\delta \nu$ is the detuning in energy (expressed in frequency) between the initial and final states and $\Omega=\Omega_{\rm ff}$ or $\Omega_{\rm df}$ is the dipole-dipole matrix element corresponding to the flip-flop or double-flip process considered.
+
+Eq.
+(\ref{delta gamma 1}) is the basis for every cross-relaxation detection performed in this manuscript.
+Because the NV center lifetime can be measured optically, as detailed in Sec.
+\ref{sec T1 NV}, cross-relaxation between NV-centers and other spin species can be measured optically.
+Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be explained in the following chapters.
+
+\section{Conclusion}
+
+We have covered in this introductory chapter every concepts needed to understand the results presented in the next chapters.
+We have seen the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them, and we have introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.
+
+The focus of chapter 2 will be on the cross-relaxation between NV centers and other spin impurities in CVD-grown diamond, while the focus of chapter 3 and 4 will be on the cross-relaxation between NV centers and other NV centers, as well as their potential application in magnetometry.
+
+\chapter{Cross-relaxations between NV centers and dark spins in CVD-grown diamond}
+Crystal defects are promising candidates for quantum information technologies, either through their optical or spin degree of freedom \citep{aharonovich2016solid, atature2018material, bassett2019quantum}.
+Spectrocopic studies are a necessary step in identifying and characterizing these defects, with electron paramagnetic resonance (EPR) being the spectrocopic tool of choice to measure the spin properties of the defects \citep{newton2007epr}.
+
+In this chapter, we present a method to characterize the spin of diamond impurities through cross-relaxation with NV centers \citep{hall2016detection,pellet2021optical}.
+This method allows an all-optical detection of non-NV center spin defects, a method that is both cheaper and more sensitive than traditional EPR.
+We focus in particular on defects in CVD-grown diamonds which have historically been less studied than HPHT diamond, and for which no prior NV cross-relaxation had been reported.
+
+The chapter is organized as follow: we will first detail the most common paramagnetic impurities in the diamond (the so-called ``dark spins") before explaining the proceeding of NV cross-relaxation experiments, and more generally of NV relaxometry.
+We will then focus on the detection of two particular defects in a CVD diamond, VH$^-$ and War1, and on their characterization (density, spin parameters, temperature dependence).
+We will finally conclude on the possible improvements and application of this detection method.
+
+The work presented in this chapter has been published in large part in the articles \citep{pellet2021optical, ngambou2022improving}.
+
+\section{Dark spins in diamond}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 2/Figures/tronche_de_spin}
+\caption{Crystalline structure of three spin defects: the NV center, the P1 enter (substituional Nitrogen) and the VH center.} 
+\label{repr NV P1 VH}
+\end{figure}
+
+We refer by dark spins to paramagnetic impurities for which the spin degree of freedom is not accessible optically.
+At room temperature, only two defects in diamond do not have a dark spin: the NV center and the more recently found ST1 center \citep{lee2013readout, john2017bright}.
+Every other spin impurity identified so far is considered to be dark.
+
+In this section we will present the most common dark spins encountered in diamond, which are simply the most common paramagnetic impurities aside from the NV center, and the relation between the diamond growth process and the presence of these impurities.
+We will also briefly present EPR spectroscopy which is the most common method to characterize these dark spins.
+
+\subsection{P1 centers and $^{13}$C}
+The most common spin defects in NV-rich diamond are the substitutional nitrogen centers N$^0_S$ known as the P1 centers (see Fig.
+\ref{repr NV P1 VH}), which have an electronic spin $1/2$ and a nuclear spin 1, and the $^{13}$C $1/2$ nuclear spin which represents 1.1\% of natural abundance carbon atoms, the remaining carbon atoms being almost exclusively the spinless $^{12}$C isotope.
+In NV-rich diamonds, these two impurities are the main causesof the NV center dephasing rates $\Gamma_2^*=1/T_2^*$ and $\Gamma_2=1/T_2$  \citep{barry2020sensitivity}.
+
+The dephasing rate caused by a given impurity is directly proportional to the average diagonal dipole-dipole matrix elements ($\Omega_{\rm shift}$ in (eq.
+\ref{eq.
+ff df})) between the two spin species \citep{taylor2008high, bauch2020decoherence}.
+Since the dipole-dipole interaction scales as $1/r^3$, and the distance $r$ between each spin scales as the inverse cubic root of the concentration,  it is expected that the dephasing rate associated with a particular impurity scales linearly with the concentration of the impurity: $\Gamma_2^*(\rm P1) \propto [P1]$.
+
+In the case of P1 centers, this was verified experimentally by looking at controlled sample with varying concentration of nitrogen \citep{bauch2020decoherence}.
+The authors of the study found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
+
+$^{13}$C nuclei are typically much more abundant than P1 center ($\sim$ 10 000 ppm without istopic engineering) but being a nuclear spin, it has a gyromagnetic ratio $\approx 2.8\cdot 10^3$ times smaller than that of the P1.
+As a result, the contribution of a single $^{13}$C nucleus to the NV dephasing rate is $\sim 10^3$ times that of a single P1 center.
+It was found experimentally that the broadening caused by $^{13}$C is $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 16\ \rm Hz/ppm$ \citep{van1997dependences, barry2020sensitivity}.
+In samples with natural isotopic abundance of carbon, the inhomogeneous dephasing rate due to the $^{13}$C nuclear spins has a value $\Gamma_2^*(^{13}\rm C ) \approx (2 \pi) 160\ \rm kHz$.
+
+\subsection{Other spin defects, VH$^-$, War1}
+\label{other defects}
+Other spin defects are also commonly found in diamond containing NV centers, such as charged vacancy or vacancy clusters \citep{hounsome2006origin}, other nitrogen related defects \citep{newton2007epr}, transition metals \citep{isoya1990fourier}, and especially for CVD grown diamond, hydrogen related defects \citep{hartland2014study}, since the CVD plasma contains up to 95 \% hydrogen (see sec.
+\ref{fab HPHT}).
+
+Of particular interest in this chapter are two defects: the VH$^-$ defect \citep{glover2003hydrogen, glover2004hydrogen} and the War1 defect \citep{cruddace2007magnetic}.
+
+The VH$^-$ defect, prior to our observations, had only been observed via electron paramagnetic resonance (EPR) in CVD grown diamond.
+It is formed by a vacancy with a hydrogen atom forming a bond with one of the four neighboring carbon atoms (see Fig.
+\ref{repr NV P1 VH}).
+To acquire its negtatively charged state, the VH center needs to capture an electron from a donor in the crystal, similarly to the NV center.
+
+This defect has a similar electronic structure as the NV$^-$ which results in an electronic spin-1 with a similar zero field splitting parameter (see Table \ref{table VH et War1}).
+Although the presence of hydrogen in this defect was confirmed by isotopic study, it has been debated whether this defect was VH$^-$ or V$_2$H$^-$ due to a conflict with \textit{ab initio} calculations \citep{shaw2005importance}.
+The defect is therefore sometimes referred to as V$_n$H$^-$.
+In this manuscript we will employ the original terminology: VH$^-$.
+
+The War1 defect, named after the University of Warwick where it was first observed in a CVD diamond \citep{cruddace2007magnetic}, is another spin-1 defect whose chemical structure remains unknown.
+No hyperfine structure could be discerned in the relatively broad EPR line, and the use of deuterium instead of $^1$H did not change the shape of the line.
+
+\subsection{EPR spectroscopy}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 2/Figures/EPR_spectometer.jpg}
+\caption{EPR specrometer. Credits: Bruker Corporation} 
+\label{photo EPR}
+\end{figure}
+
+Electron paramagnetic resonance (EPR), the electronic equivalent to NMR, is the the most common way to detect electronic spins in solids and liquids.
+The principle of its detection is the absorption or reflection of a microwave field resonant with a spin transition.
+
+EPR generally relies on the thermal polarization of the spins to observe an absorption.
+In order to increase the polarization, an important magnetic field is applied (typ.
+0.35 T or 3500 G) and a microwave cavity (typ.
+9500 MHz) is used to improve the coupling of the spins to the field.
+
+The usual EPR apparatus, shown in Fig.
+\ref{photo EPR}, is a bulky and expensive lab equipment ($\gtrsim$ 1 M€).
+In contrast, optical detection such as ODMR is technically easier, less expensive, and allows better spatial resolution.
+
+ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ spins \citep{mitchell2013x}.
+This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a spin with a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and in part from the more efficient photon detection in the optical range than in the microwave range.
+
+The key idea of this chapter is to allow the optical detection of spin defects that normally do not show ODMR thanks to cross-relaxation with NV centers.
+
+\section{NV center relaxometry}
+
+Detection of dark spins via NV center cross-relaxation fall under the more general technique of NV relaxometry, which consists in measuring a modification of the NV center spin lifetime $T_1$.
+Relaxometry with NV centers has been used not only to measure cross-relaxation with dark spins \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross}, but also to probe the magnetic noise in the diamond environment coming from spins \citep{steinert2013magnetic}, ions\citep{tetienne2013spin}, free radicals \citep{nie2021quantum}or magnetic domains \citep{finco2021imaging}.
+
+The protocols used in NV relaxometry measure either directly the spin $T_1$ or indirectly through the NV photoluminescence.
+
+\subsection{PL and single$-\tau$ relaxometry protocols}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 2/Figures/relaxo_pulsed_continuous_2}
+\caption{Representation of the two relaxometry protocol.
+(a) Simple $T_1$ measurement pulse sequence.
+(b) Simulation of two $T_1$ measurement with different $T_1$ values.
+$\tau$ represents the optimal dark time in order to maximize the signal $\delta S$ (c) Representation of the optical pumping (green arrows) and depolarization (blue arrows) of the NV center spin.
+(d) Simulation of the PL modification caused by a change in the spin $T_1$.
+The spin $T_1$ profile in term of the magnetic field was arbitrarily chosen.}
+\label{T1 vs PL}
+\end{figure}
+
+The two most common NV relaxometry protocols are represented on Fig.
+\ref{T1 vs PL}.
+The first method simply consists in measuring the spin lifetime, following the protocols described in \ref{sec T1 NV}.
+In Fig.
+\ref{T1 vs PL}, we picture the simpler $T_1$ measurement protocol, but the technique can be used with the common mode rejection protocol.
+In order to more efficiently measure a modification of the spin $T_1$, the ``single$-\tau$" protocol does not record the entire depolarization curve represented in Fig.
+\ref{T1 vs PL}-b), but instead probes the spins for a single dark time value $\tau$ \citep{pelliccione2014two, schmid2015relaxometry, tetienne2016scanning}.
+This value is chosen in order to maximize the change in signal $\delta S$ for a given change in lifetime $\delta T_1$, typically $\tau \sim T_1$.
+
+ The second method simply consists in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, and since the steady state population in the $\ket{0}$ state results from the equilibrium between the optical polarization rate $\Gamma_{\rm las}$ and the depolarization rate $\Gamma_1=1/T_1$.
+A simple rate equation for the $\ket{0}$ state population $\rho_{00}$ gives:
+ \begin{equation}
+ \rho_{00}=\frac{\Gamma_{\rm las}+\Gamma_1}{\Gamma_{\rm las}+3\Gamma_1}.
+ \end{equation}
+
+ Fig.
+\ref{T1 vs PL} shows a simulation of NV centers PL when the spin decay rate $\Gamma_1$ is modified.
+In this fictional scenario, we assume that $\Gamma_1$ resonantly increases for a given value of the magnetic field (we will see later in this manuscript physical examples corresponding to this scenario).
+We can see that the PL decreases as an answer to the $\Gamma_1$ increase.
+The exact value of the PL drop ($\sim 2\%$ in the crase presented here) depends on the optical polarization rate $\Gamma_{\rm las}$ as well as the modification of the spin lifetime.
+
+It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the PL case, the laser intensity needs to be adjusted so that $\Gamma_{\rm las} \sim \Gamma_1$ in order to maximize the PL contrast, while the single$-\tau$ method requires a waiting time $\tau \sim T_1$ between each measurement, which results in a similar PL count and shot noise limit in both case.
+
+On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and because it requires a smaller laser power to work optimally.
+It is however more sensitive to drifts and changes in the optical setup.
+Most of the relaxometry measurement in this manuscript were performed using a PL-based detection.
+
+\subsection{Comparison with DEER protocol}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{chapter 2/Figures/DEER}
+\caption{Pulse sequence of the DEER protocol}
+\label{DEER}
+\end{figure}
+
+Aside from cross-relaxations, another protocol exists to address dark spin resonance with NV centers, based on a modification of the NV $T_2$ time \citep{mamin2012detecting, serbyn2014interferometric}.
+This protocol, called double electron-electron resonance (DEER), is depicted on Fig.
+\ref{DEER}.
+The pulse sequence consists in applying a spin-echo on the NV center, as described in \ref{sec T2 echo}, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers.
+This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center will not be rephased.
+The newly measured $T_2$ time should thus be slightly shorter.
+
+\begin{equation}
+\frac{1}{T_2^{\rm DEER}}=\frac{1}{T_2^{\rm echo}} + \delta \Gamma_2^*,
+\end{equation}
+
+where $\delta \Gamma_2^*$ is the contribution of the dark spin to the $T_2^*$ of the NV center.
+
+By scanning the frequency of the second microwave (RF2) and monitoring the $T_2$ time of the NV centers, one can find dark spin resonance without having to bring them to resonance with NV centers, which can be useful when the dark spin resonance is very far detuned from the NV one.
+
+Both relaxometry and DEER protocols have their own advantages.
+While DEER allows more control on the probed impurity, allowing in particular coherent control of the dark spin \citep{sushkov2014magnetic}, relaxometry protocols are generally easier to implement as they do not require a pulsed microwave setup.
+In some cases at least, relaxometry protocols have been shown to be more sensitive than $T_2$ based protocols such as DEER due to the fact that $T_1$ is typically orders of magnitude greater than $T_2$ \citep{steinert2013magnetic}.
+
+Relaxometry protocols do require however specific magnetic field values in order to bring in and out of resonance the two spins involved in the cross-relaxation.
+The choice of the field orientation and strength in order to best observe NV-dark spin cross-relaxations will be covered in the following section.
+
+\section{Choice of the magnetic field orientation}
+
+As relaxometry does not rely on a microwave field, the only external parameter left to tune is the magnetic field.
+Typically, the magnetic field is scanned in a fixed direction while the signal, either PL or T1, is recorded.
+Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect NV-dark spins CR.
+
+This section focuses on particular field orientation with respect to the diamond crystalline axes, and details the advantages and disadvantages of the two main choices that are the [100] and [111] direction in order to detect four particular dark spins: P1 centers, Vh$^-$, War1 and the $^{13}$C$-$NV complex\footnote{The $^{13}$C$-$NV complex is normally not dark, as will be discussed in its own section.}.
+
+\subsection{The [100] and [111] magnetic field orientations}
+\label{sec simu}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_NV}
+\caption{Simulated transition frequencies of the $\ket{0} \to {-1}$ and $\ket{0} \to {+1}$ transitions for the 4 classes of NV centers as a function of the magnetic field.
+a) When the magnetic field is along the diamond [100] crystalline axis, b) along the [111] axis.
+A representation of the 4 classes of NV centers, labeled 1 to 4, and the orientation of the magnetic field in both cases is put on top}
+\label{121 vs 22 NV}
+\end{figure}
+
+Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\mathbf{B} \parallel [100]$ or 
+$\mathbf{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes.
+When $\mathbf{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\mathbf{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.
+
+These changes in the field orientation result in important changes in the NV behavior, mainly due to the transverse magnetic field: Fig.
+\ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the ground state spin Hamiltonian (REF chapter 1) - or to be more precise the transitions between the ground state of the spin Hamiltonian and the two excited states - for any of the four classes of NV centers when $\mathbf{B} \parallel [100]$.
+Notably, there is no transitions frequencies below 2500 MHz.
+
+On the other hand, Fig.
+\ref{121 vs 22 NV}-b) shows the transition frequencies when $\mathbf{B} \parallel [111]$, for both the class aligned and the three class misaligned.
+The three classes behave similarly than the four classes in the $[100]$ case, but the one class aligned with the magnetic field can get its $\ket{0}\to\ket{-1}$ transition arbitrarily small, as long as the magnetic field is perfectly aligned with the NV axis.
+
+We will now compare these two possible field orientation in order to observe cross-relaxation between NV centers and the four previously mentioned dark spin species.
+
+\subsection{CR condition between NV$^-$ and P1}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_P1}
+\caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}
+\label{121 vs 22 P1}
+\end{figure}
+
+P1 centers, the most abundant electronic spin in the diamonds commonly used, have a $1/2$ electronic spin.
+They do have a small magnetic field anisotropy coming from the hyper-fine coupling to the nitrogen nucleus ($\sim 100\ \rm MHz$), but this dependency on the field orientation is small compared to the NV center's zero field splitting $D=2870\ \rm MHz$.
+We can therefore, to the first order, neglect the hyper-fine coupling and consider P1 centers as ideal, isotropic, spin $1/2$.
+
+Fig.
+\ref{121 vs 22 P1} shows the transition frequencies for the NV center and such a spin $1/2$, as a function of a magnetic field parallel to [100] or [111].
+When $\mathbf{B} \parallel [100]$, there will never be a co-resonance between the NV and P1 transitions, meaning that no NV-P1 cross-relaxation can be observed when B is scanned in this direction.
+When $\mathbf{B} \parallel [111]$ however, the P1 $\ket{-1/2} \to \ket{+1/2}$ transition matches in energy the $\ket{0}\to\ket{-1}$ NV transition for B=512 G.
+This means that when B reaches $\approx 512\ \rm G$, some of the polarization of the NV centers will be transferred to the P1 centers, which will result in a drop in the NV PL.
+
+\subsection{CR condition between NV$^-$ and VH$^-$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 2/Figures/121_vs_22_VH}
+\caption{Simulation of the transition frequencies of NV$^-$ centers and VH$^-$ centers for a magnetic field a) along the [100] axis and b) along the [111] axis}
+\label{121 vs 22 VH}
+\end{figure}
+
+VH$^-$, a much less abundant electronic spin than P1 center, has an electronic spin 1, with a spin Hamiltoninan very similar to that of the NV center.
+In particular it has the same symmetries being also a $C_{3v}$ defect, with a slightly smaller ZFS at $D_{VH} \approx 2700\ \rm MHz$ compared to $D_{NV}=2870\ \rm MHz$.
+
+Fig.
+\ref{121 vs 22 VH} shows the transition frequencies for the NV$^-$ and VH$^-$ ground state spin.
+Contrary to the P1 case, there is a co-resonance when $\mathbf{B} \parallel [100]$ for $B\approx 55\ \rm G$.
+When $\mathbf{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G.
+A scan along the [111] in this case means that the peak signal will be $\sim 6$ times weaker than in the [100] case, and spread over multiple lines which might overlap between themselves or with other existing PL features.
+
+The War1 spin mentioned previously is also an electronic spin 1 with (pseudo)-$C_{3v}$ symmetry and $D_{War1} \approx 2470\ \rm MHz$.
+This result in a similar behavior than VH$^-$ when it comes to the CR condition with NV centers.
+
+\subsection{CR condition between NV$^-$ and and $^{13}$C-NV$^-$}
+\label{sec NV-13c-NV}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_13CNV}
+\caption{a) Representation of the $^{13}$C-NV complex (right) compared to a ``normal" NV center (left) b) and c) simulation of the transition frequencies of NV$^-$ centers and $^{13}$C-NV$^-$ complex for a magnetic field b) along the [100] axis and c) along the [111] axis}
+\label{121 vs 22 13C-NV}
+\end{figure}
+
+The $^{13}$C-NV$^-$ complex is formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope.
+Such a complex is illustrated in Fig.
+\ref{121 vs 22 13C-NV}-a).
+
+The $^{13}$C-NV$^-$ complex still behaves as a ``normal" NV center, except that its transition lines are shifted by the hypere-fin coupling.
+In particular its electronic spin is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states\footnote{This is confirmed by the presence of sideband in ODMR spectra around the main NV lines \citep{simanovskaia2013sidebands}.}.
+
+Since both NV$^-$ and $^{13}$C-NV$^-$ are \textit{a priori} equally polarized, then following the argument in \ref{Sec_CR} there should be no CR between the two spin families.
+The reason why those transitions are still visible is the presence of dark NV-centers (``fluctuators") which will be further explained in chapter 3.
+
+The transitions of the $^{13}$C-NV$^-$ complex are given by diagonalizing the full spin Hamiltonian:
+\begin{equation*}
+\mathcal{H}=\mathcal{H}_{NV}+\mathcal{H}_{^{13}C}+\mathcal{H}_{HF},
+\end{equation*}
+where $\mathcal{H}_{NV}$ is the NV$^-$ spin Hamiltonian,
+$\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian: $$\mathcal{H}_{HF}= \hat{\mathbf{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\mathbf{I}}_C.$$
+
+When the $^{13}$C is the direct neighbor of the vacancy as represented in Fig.
+\ref{121 vs 22 13C-NV}-a) - this is also referred as a first-shell $^{13}$C - the hyper-fine tensor $\mathcal{A}$ can be written \citep{simanovskaia2013sidebands}:
+$$ \mathcal{A} = \begin{pmatrix}
+\mathcal{A}_{xx} & 0 & \mathcal{A}_{xz} \\ 0 & \mathcal{A}_{yy} & 0 \\ \mathcal{A}_{zx} & 0 & \mathcal{A}_{zz}
+\end{pmatrix},$$
+where $\mathcal{A}_{xx}=190$ MHz, $\mathcal{A}_{yy}=120$ MHz, $\mathcal{A}_{zz}=129$ MHz, and  $\mathcal{A}_{xz}=\mathcal{A}_{zx}=-25.0$ MHz.
+
+Fig.
+\ref{121 vs 22 13C-NV}-b) and c) show the simulated transition frequencies for a $^{13}$C-NV$^-$ complex when B is scanned along the [100] or [111] axis, as well as the ``normal" NV transitions and the co-resonance between those two.
+When $\mathbf{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G.
+When $\mathbf{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
+This abundance of transitions for low ($<100\ \rm G$) magnetic field makes the experimental analysis quickly intractable in the second case.
+
+There is one last aspect to take into account regarding the field orientation choice which is the impact of the transverse magnetic field on the cross-relaxation contrast.
+
+\subsection{CR contrast and transverse magnetic field}
+\label{CR contrast}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 2/Figures/contrast_OMDR_111_Vs_100}
+\caption{Simulated ODMR contrast on all four classes of NV centers with a magnetic field a) along the [100] axis and b) along the [111] axis}
+\label{121 vs 22 contrast}
+\end{figure}
+
+Similarly to ODMR, there are two required conditions to observe NV cross-relaxation on the PL signal: 
+\begin{itemize}
+\item There must be a difference in population between the two NV spin states involved in the CR (e.g.
+$\ket{0}$ and $\ket{-1}$), otherwise there would be no population transfer with the unpolarized dark spin.
+\item The two NV spin states must have a difference brightness in order to be optically detectable.
+\end{itemize}
+
+These two conditions are fulfilled when there is no magnetic field or purely longitudinal magnetic field, but not anymore in the presence of transverse magnetic field which mixes the NV spin Hamiltonian eigenstates \citep{tetienne2012magnetic}.
+
+Fig.
+\ref{121 vs 22 contrast} shows the simulated relative contrast of a CR process \footnote{Due to the similarity between ODMR and CR, we assume here that the ODMR contrast is equal to the CR contrast} computed from the rates found in \citep{tetienne2012magnetic}, for a magnetic field aligned along either [100] or [111].
+When the NV center is aligned with the magnetic field, the contrast remains equal to 1 regardless of the magnetic field because there is no transverse field.
+When the NV is not aligned however, the CR contrast quickly drops down for B$>100\ \rm G$.
+
+The [100] direction is therefore not well suited to observe CR at magnetic fields higher than a few hundreds gauss.
+Fortunately, the co-resonance between NV centers and VH$^-$, War1 or $^{13}$C-NV$^-$ all happen for $|\mathbf{B}| < 150 \ \rm G$, and can thus be detected with a [100] oriented magnetic field.
+
+\section{Dark spin spectroscopy with NV centers}
+This section shows experimental observation of NV-dark spins CR.
+Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross} had been performed on the detection of dark spins through CR with NV centers, and to our knowledge, all this studies where done on the detection of P1 centers.
+
+We show an example of one this study to illustrate the principle of dark spin spectroscopy before showing our results on the detection of VH$^-$ and War1 in a CVD diamond.
+
+\subsection{Detection of P1 centers}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\textwidth]{chapter 2/Figures/NV_P1_libreAcces}
+\caption{Change in the PL of an ensemble of NV centers as a magnetic field is scanned along the [111] crystalline axis. Taken from \citep{armstrong2010nv}}
+\label{CR P1 exp}
+\end{figure}
+
+An example of a NV-P1 CR study (\citep{armstrong2010nv}) is shown in Fig.
+\ref{CR P1 exp}.
+It shows the evolution of the NV PL with respect to a magnetic field scanned along the [111] crystalline axis.
+There are several information to gather from this graph: First, one observes a slowly decaying PL envelope, which stabilizes to a plateau for $B\sim 1000\ \rm G$.
+This change in PL is caused by the transverse magnetic field acting on the three NV classes that are not aligned with the magnetic field.
+For $B \gtrsim 1000\ \rm G$, these three classes are fully depolarized which explains the plateau observed.
+
+Second, there are three salient features denoted Ns-NV, NV-NV and NV-LAC. All these features come from cross-relaxations or level anti-crossing (LAC) from the NV class aligned with the magnetic field.
+
+The first feature (Ns-NV) for $B\approx 512 \ \rm G$ is the signature of the CR between NV and P1 centers.
+Its relatively complicated shape comes from the hyper-fine coupling between the P1 electronic spin and the $^{14}$N nucleus \citep{lazda2021cross}.
+Excited state level anti-crossing (ESLAC) also occurs in this region, but its optical signature, if it exists, is expected to be much weaker than the NV-P1 CR \citep{zheng2017level}.
+
+The second feature (NV-NV) for $B\approx 590 \ \rm G$ comes from the cross relaxation between NV centers aligned with the magnetic field and NV centers misaligned with the field (the co-resonance can be observed in Fig.
+\ref{121 vs 22 NV}-b)).
+This phenomenon will be further discussed in the next chapter.
+
+Finally the third feature (LAC) for $B\approx 1024 \ \rm G$ correspond to the level anti-crossing between the $\ket{0}$ and $\ket{-1}$ states of the NV center aligned with B.
+When the two levels get close enough in energy, the Hamiltonian off-diagonal terms (coming from strain local electric field or residual transverse magnetic field) will mix the $\ket{0}$ and $\ket{-1}$ states and prevent an efficient polarization by the green laser.
+This result in a sharp drop in the PL that can be exploited to perform microwave-less magnetometry with NV centers \citep{wickenbrock2016microwave, zheng2017level, zheng2020microwave}.
+
+\subsection{Detection of VH$^-$, War1 and $^{13}$C-NV$^-$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 2/Figures/fig_VH_et_co}
+\caption{a) Change in PL (normalized) of the ensemble of NV centers in sample CVD-pink as a magnetic field is scanned along the [100] axis.
+The orange line is a fit to the slowly decreasing envelope b) Subtraction of the PL signal by the envelope fit as a function of the magnetic field.
+The green lines and red and purple regions correspond to the predicted CR regions, as detailed in main text}
+\label{CR VH exp}
+\end{figure}
+
+Our work \citep{pellet2021optical} focuses on the optical detection of VH$^-$, War1 and $^{13}$C-NV$^-$ through NV CR.
+This study was the first to detect VH$^-$ and War1 defects optically and the first CR experiment on a CVD diamond.
+It was also the first,to our knowledge, to use a [100] aligned magnetic field.
+
+The diamond used for this study is the sample CVD-pink (see sample details in appendix \ref{Appendix samples}).
+The experimental setup used to obtain this data is the standard confocal microscopy setup shown in Sec \ref{Sec setup cahp 1}.
+The magnetic field here is generated with an electromagnet (EM), aligned along the [100] diamond axis with a precision $<1^\circ$.
+
+Fig.
+\ref{CR VH exp}-a) shows the change in PL with respect to a magnetic field scanned along the diamond [100] axis.
+Similarly to fig.
+\ref{CR P1 exp}, there is a slowly decaying envelope of the PL due to the transverse magnetic field \footnote{The plateau is not visible in this case as the scan stops at 150 G}.
+There also are three deviations from the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
+
+In order to better discern the three features, we fitted the envelope with a polynomial fit of order 4.
+We previously tried to simulate the PL dependency on the transverse field  by using the model in \citep{tetienne2012magnetic}, but the result did not correctly fit our data.
+
+Fig.
+\ref{CR VH exp}-b) shows the subtraction of the PL data by the envelope fit.
+We can clearly see three dips in PL around the values 20 G, 56 G and 122 G.
+We then computed the expected value of the magnetic field for the expected CR line thanks to the simulations shown in sec.
+\ref{sec simu}.
+For the VH$^-$ and War1 lines, we took the ZFS $D$ values from \citep{cruddace2007magnetic} and included the error bars.
+For the $^{13}$C-NV$^-$ lines, the error bars given in \citep{simanovskaia2013sidebands} are much smaller and we only plotted the expected value.
+We can see that the dips in PL match relatively well the predicted CR lines of $^{13}$C-NV$^-$, VH$^-$ and War1 respectively.
+
+The data shown here was acquired over 24 h.
+It should be noted however that a single-photon counting APD was used for this acquisition, which reduces significantly the sensitivity because of the high shot noise \footnote{The sample produces $\sim$ 100 nW of PL, which means that only $\sim 10^{-5}$ of the total PL is actually measured by an SPC-APD with a max counting rate of $10^7$ photons/s.}.
+Using a photodiode and modulation of the magnetic field, we estimate that we could get similar signal to noise ratio within a few minutes.
+
+To calibrate the magnetic field, we added a microwave field at a  fixed frequency $\nu > 2870\ \rm MHz$, and measured the EM voltage at which the PL dropped due to the $\ket{+1}$ state of the NV centers becoming resonant with the microwave tone.
+We could then use the values in Fig.
+\ref{121 vs 22 NV} to find the corresponding magnetic field.
+We varied the frequency in the range  $[2870,3150]\ \rm MHz$ by step of 2 MHz to get a calibration of the magnetic field in the $[0,150]\ \rm G$ region.
+Considering the uncertainty in the field angle, the NV ZFS $D$ value and the width of the PL drop, we estimate the final precision of the calibration to be within $\pm 1\ \rm G$.
+
+\section{Identification of the $^{13}$C-NV$^-$, VH$^-$ and War1 lines}
+
+\subsection{The VH$^-$ and War1 lines}
+\begin{table}[htbp]
+\centering
+\caption{\bf Zero-field splitting parameter $D$ for NV$^-$, VH$^-$ and War1}
+\begin{tabular}{ccc}
+\hline
+$D_z$ estimation (MHz) & Cruddace's work\citep{cruddace2007magnetic} & Our work \\
+\hline
+NV$^-$ & 2872(7) & * \\
+VH$^-$ & 2706(30) & 2694(5)  \\
+WAR1 & 2466(60) & 2470(10) \\
+\hline
+\end{tabular}
+  \label{table VH et War1}
+\end{table}
+
+The identification of the PL dips at 56 and 122 G in Fig.
+\ref{CR VH exp} as CR between NV and respectively VH$^-$ and War1 comes from two main observations: the closeness in the ZFS values, and the CVD nature of the sample.
+
+Table \ref{table VH et War1} shows the value and precision of the $D$ factor when measured through EPR \citep{cruddace2007magnetic}, and when computed from the data in Fig.
+\ref{CR VH exp}.
+In order to get to the $D$ value of VH$^-$ and War1, we had to assume that the spin structure was that of a spin 1 similar to the NV center, with a g-factor close to 2.
+Both of this assumptions are confirmed by the EPR measurements.
+We do find that that both $D$ measurement concord within their respective error bars.
+
+The second reason comes from the nature of the sample used.
+We tried similar PL scans along the [100] axis on many samples with dense ensemble of NV centers.
+We found that the dip at 55 G was present on all CVD samples ($\sim$ 5 samples) used, and none of the HPHT samples ($\sim$ 20 samples).
+The same is true of the 122 G line, except that it did appear on some HPHT samples, as will be discussed below.
+We know that the VH$^-$ defect is much more likely to appear in a CVD sample due to the large quantity of hydrogen used in the growth process, and while we do not know the chemical nature of the War1 defect, the only time it was observed through EPR was on a CVD diamond.
+
+Finally, the apparent larger linewidth of the War1 line compared to the VH$^-$ one is also corroborated by the EPR measurements.
+
+For all of this reasons we attribute the 55 and 122 G dip to CR with VH$^-$ and War1 respectively.
+
+\subsection{The $^{13}$C-NV$^-$ lines}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.6\textwidth]{chapter 2/Figures/scan_100_Sumi4}
+\caption{Change in PL as a function of a magnetic field scanned along the [100] axis for sample SUMI-4}
+\label{scan sumi 4}
+\end{figure}
+
+The identification of the $\sim 20\ \rm G$ as CR between $^{13}$C-NV$^-$ and ``normal" NV centers also come from two observations.
+First, the lines are present on every single sample with dense ensemble ($\gtrapprox 1$ ppm), regardless of the fabrication process.
+Fig.
+\ref{scan sumi 4} for example shows a PL scan along the [100] axis for sample Sumi-4, an HPHT sample with [NV]$\approx 5\ \rm ppm$ and natural abundance of $^{13}$C \footnote{We could unfortunately not try a sample isotopically enriched in $^{12}$C since those samples are generally reserved to low-nitrogen concentration ($< 10\ \rm ppm$), which does not yield enough NV centers to perform CR spectroscopy.
+}.
+
+The second reason is the shape of the PL dip.
+The two other dips seem be formed of a single line and are reasonably well fitted with Gaussian curves.
+The 20 G dip however has a flat bottom which would indicate the presence of several overlapping lines, and Fig.
+\ref{121 vs 22 13C-NV} indeed predicts that there would be 4 CR lines between 18 and 25 G.
+
+For this two reasons, we attribute the 20 G line to CR between $^{13}$C-NV$^-$ and ``normal" NV centers.
+
+\bigskip
+
+The CR spectra shown in this section can be used to detect the presence of particular impurities in the sample, but more quantitative information on these impurities can also be extracted from these spectra, which is the object of the following section.
+
+\section{Quantitative estimates of the NV-Dark spins CR}
+
+This section exploit the CR spectrum shown in Fig. \ref{CR VH exp} in order to extract information on the dark spins Hamiltonian parameters, linewidth and concentration.
+
+\subsection{Dark spin Hamiltonian parameters}
+
+First, as discussed in the last section, we can estimate the dark spin ZFS $D$ value with greater precision than EPR.
+While it would be in theory possible to measure other properties of the dark spins Hamiltonian - such as the $g$-factor - by changing the angle of the magnetic field, this would here quickly become intractable due to the increase in co-resonance conditions and the relative closeness of each of these transitions.
+DEER might prove more useful in this regard as only one class of NV centers at a time is probed.
+
+\subsection{Dark spin linewidth}
+
+The second quantitative estimate we can make is of the dark spin inhomogeneous broadening $T_2^*$: if we assume the change in PL to be proportional to the change in the NV lifetime $T_1$, which is true for small enough changes, then the linewidth of the PL dips in Fig.
+\ref{CR VH exp} is the convolution of the NV and dark spin spectrum.
+By deconvoluting the data with the NV spectrum, known for example from ODMR, we can then recover the dark spin spectral response as was done in \citep{hall2016detection}.
+
+This procedure will be discussed in more details in the next chapter.
+We find here that the linewidth of VH$^-$ and War1 are respectively (half width at half maximum) $5.7\pm 0.6$ and $15.7\pm 1.6$ MHz \footnote{We follow here the pseudo-Voigt deconvolution method described in the following chapter.
+The linewidth $W=2.7$ MHz (half witdth at half maximum) of the NV centers was measured from ODMR (see Fig.
+\ref{ODMR 1 classe}).
+It includes all three hyper-fine transistions}.
+
+\subsection{Dark spin concentration}
+
+Finally, the third estimate we can make is that of the dark spins concentration.
+We can indeed calibrate the measurement by using the $^{13}$C-NV$^-$ line, of which we know the concentration in natural abundance diamond: [$^{13}$C-NV$^-$] $\approx$ 3.3 \% [NV].
+We can then, knowing the NV concentration, try to estimate the concentration of the dark spins.
+There are however several caveats:
+
+\begin{itemize}
+\item Since the change in $T_1$ is proportional to the $T_2^*$ of the probed spins (eq.
+\ref{delta gamma 1}), we should compensate that by integrating over the entire feature, instead of simply looking at the maximum of the line.
+\item We have to take into account the loss of contrast of the NV centers as the transverse magnetic field grows, as explained in sec. \ref{CR contrast}.
+\item The $^{13}$C-NV$^-$ are not ``dark spins", and its CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
+We will see however in the next chapter that we can consider a fraction of the NV centers as dark spins, and that this fraction can be estimated.
+\citep{choi2017depolarization} found a ratio of $\sim 1/4$ dark NV centers for $\sim 3/4$ bright ones.
+\end{itemize}
+
+From these considerations, we can give an estimate of the concentrations of VH$^-$ and War1 based on the data on Fig. \ref{121 vs 22 VH}-b):
+
+For the VH$^-$ line we find that the area of the line is $\approx 1.4$ times bigger than that of the $^{13}$C-NV$^-$ one, from \ref{121 vs 22 contrast} we find that the expected CR contrast is $\approx 1.2$ times smaller at 55 G than at 20 G, and finally we chose a dark to bright NV ratio of 0.25 as found in \citep{choi2017depolarization}.
+
+With an estimated NV concentration [NV] $\approx\ \rm 4.6 ppm$, giving us a concentration $[^{13}\rm{C-NV}^-]\approx \ \rm 130 ppb$, we then find $[\rm VH^-] \approx 58\ \rm ppb$.
+Similarly, we find $[\rm VH^-] \approx 68\ \rm ppb$.
+Given the large uncertainty on some parameters, in particular the dark to bright spin ratio, these estimates are probably only precise within an order of magnitude.
+
+If we consider that the sampled volume in this experiment is $< 100 \mu \rm{m}^3$, the total number of dark spins detected here is then $< 10^6$.
+Even with very conservative estimates, the technique employed here is still more sensitive than EPR by several orders of magnitude.
+
+\section{Temperature dependence of VH$^-$ and War1}
+
+\subsection{Mofifications on CVD samples}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 2/Figures/VH_CVD_chauffage}
+\caption{Demodulated PL on samples CVD-C and CVD-B as a function of a magnetic field along the [100] axis.
+An additional oscillating magnetic field along the same direction was used to perform the lock-in detection.
+The VH$^-$ and War1 lines are taken from \ref{CR VH exp}}
+\label{chauffage CVD}
+\end{figure}
+
+Given that the NV CR spectroscopy was both sensitive and relatively easy to setup, we decided in collaboration with the teams of Alexandre Tallaire at IRCP and Jocelyn Achard at LSPM to study the effect of annealing at high temperature on VH$^-$ and War1 \citep{ngambou2022improving}.
+
+Fig.
+\ref{chauffage CVD} shows the results of this experiment.
+This time, a modulation of the magnetic field was used and the PL coming from the photodiode was demodulated through a lock-in amplifier, which explains the derivative shape of the lines.
+The fact that the background is not perfectly flat still comes from the transverse magnetic field.
+The same lock-in parameter were used on both plots so that the amplitudes should in theory be comparable.
+
+Fig \ref{chauffage CVD}-a) shows the demodulated PL versus mganetic field amplitude along [100] on sample CVD-C, while \ref{chauffage CVD}-b) shows the same experiment on sample CVD-B.
+The fabrication and subsequent operations done on this two samples are detailed thoroughly in \citep{ngambou2022improving}, but the main difference between these two sample is that CVD-C was only annealed once after irradiation, for 2 hours at 800$^\circ$C, while CVD-B was annealed for 2 hours at 800$^\circ$C and then for 1 hour at 1000$^\circ$C and 1 hour at 1200$^\circ$C.
+
+We can see that the VH$^-$ line in particular completely disappears on sample CVD-B, which we attribute to the additional annealing.
+This measurement also corroborate modifications in the infrared absorption spectrum, and an increase of the NV coherence time, associated with the annealing at higher temperature \citep{ngambou2022improving}.
+
+The War1 line on the other hand is still present after the high temperature annealings, albeit slightly smaller.
+This information can give us some insights on the chemistry of the War1 defect.
+%Et en vrai je crois que c'était deja vu dans la thèse de Cruddace
+
+\subsection{Modification on HPHT samples}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 2/Figures/VH_HPHT_chauffage}
+\caption{Demodulated PL on samples SBST-C and SBST-B as a function of a magnetic field along the [100] axis. The plot of SBST-C is offset for clarity}
+\label{chauffage HPHT}
+\end{figure}
+
+An unexpected side-study was done on HPHT samples.
+Indeed, the CVD samples used previously are grown by homoepithaxy on top of a diamond substrate.
+In this case, the substrate was a type 1B HPHT diamond, which was irradiated and annealed along the CVD layer.
+We will name SBST-C and SBST-B the respective HPHT substrate of CVD-C and CVD-B.
+
+Fig.
+\ref{chauffage HPHT} shows demodulated PL scans with B $\parallel$ [100] on both samples.
+They both show a line at $B \approx 20\ \rm G$ corresponding to the $^{13}$C-NV$^-$ line, and none of them show a line at $B = 56\ \rm G$ corresponding to the VH$^-$ line.
+There is however a stark difference for the War1 line: SBST-C seems to show a weak line, while SBST-B shows a very clear line, with an amplitude more than 5 times greater than SBST-C.
+
+We can also see that many weak lines on SBST-B seems to have disappeared in SBST-C, while an unidentified line at $B = 48\ \rm G$ seems to have grown similarly to the War1 line.
+
+We first suspected that some of the War1 defects had been migrating from the CVD layer to the HPHT substrate with the high-temperature annealing, but not only is it very unlikely with the annealing time (1 hour) and substrate thickness (several mm), We also could not find a dependence on the War1 concentration with the depth in the substrate, whereas a diffusion would have given a higher concentration near the CVD layer.
+The War1 defects seem to have been created by the annealing.
+
+These results suggest that, while some defects are removed by the high temperature annealing, such as the VH$^-$ and several unidentified defects in the HPHT substrate, some seem to be created by the annealing process, such as the War1 or the one associated with the line at $B = 48\ \rm G$ in the HPHT substrate.
+
+\section{Conclusion and perspectives}
+
+To conclude this chapter, we have seen how NV centers can be used to detect dark spins via cross-relaxations.
+This detection can be significantly more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up.
+This technique however is only applicable to a few defects in NV-rich diamond, meaning that its main application would probably be to characterize the environment of samples with high NV density.
+
+We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan which provides a significantly less noisy signal than the [111] scans for weak magnetic field.
+
+Finally we have seen an application of the CR to detect and characterize optically two dark spins in CVD diamond, and to find out the response of these defects to high temperature annealing.
+
+There still remains a lot of open questions and there are several experiments that could be build upon these results. Some of these questions are listed below.
+
+\subsection{Unidentified CR lines}
+
+There are still a large number of lines in CR spectra that have not been identified.
+While not every line necessarily correspond to a new defect - for example \citep{wunderlich2021magnetic} recently found lines corresponding to co-resonance between the $\ket{0} \to \ket{-1}$ transition of an NV center and the $\ket{-1} \to \ket{+1}$ of another NV center - the more likely hypothesis for each of these unidentified lines is still the presence of a new defect.
+
+Such unidentified lines can be seen in Fig.
+\ref{chauffage CVD} and \ref{chauffage HPHT}.
+In particular we can see a line at 160 G in Fig.
+\ref{chauffage CVD}-a) and a line at 48 G in Fig.
+\ref{chauffage HPHT} that detach clearly from the background.
+
+Further studies would be needed to find how common these lines are, and what could be there sources.
+
+\subsection{Measuring the spin dynamics of VH$^-$}
+
+We found that NV-VH$^-$ CR in CVD could produce a change in PL of $\sim 0.2 \%$.
+This value is within one order of magnitude from ODMR contrast with NV ensembles (typically $<5 \%$), which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin.
+The main issue comes from the fact that in order to measure the VH$^-$ spin states, it needs to be resonant with the NV center.
+Therefore, resonant microwave pulses with VH$^-$ will also be resonant with the NV centers, which will create an unwanted optical signal.
+
+There are two solution to circumvent this problem: the first is to use a fast switching magnetic field in order to bring the VH$^-$ and NV centers in resonance when initializing or reading out the VH$^-$ state, and out of resonance to coherently manipulate the VH$^-$.
+This means that the magnetic field needs to be switched faster than the probed spin dynamics, which may be doable for the population dynamics $T_1$ but is likely too slow to probe the $T_2$ time.
+
+The other solution consists in using the spin-1 nature of VH$^-$: when a NV$^-$-VH$^-$ CR occurs, only one of the transitions of VH$^-$ and NV$^-$ is brought to resonance.
+The other one ($\ket{0} \to \ket{-1}$ for the VH$^-$ and $\ket{0} \to \ket{+1}$ for the NV) is out of resonance.
+The idea would therefore be to keep the NV $\ket{0} \to \ket{-1}$ transition and VH $\ket{0} \to \ket{+1}$ transition resonant, and to coherently manipulate the VH $\ket{0} \to \ket{-1}$ transition.
+Further studies are needed to assess the effectiveness of this scheme.
+
+\subsection{Isolating a single VH$^-$}
+
+We estimated the concentration of VH$^-$ in CVD-pink to be $\sim 60\ \rm ppb$.
+This concentration is about two orders of magnitude too high to contemplate isolating a single emitter with a diffraction-limited microscope.
+We know however that we can reduce the VH$^-$ concentration arbitrarily by applying high temperature annealing.
+We could therefore reduce the VH$^-$ concentration to be in the $0.1 \sim 1\ \rm ppb$ range and optically isolate a single VH$^-$ center by scanning a confocal microscope with a fixed $B_{[100]}=56\ \rm G$ while monitoring the NV PL or $T_1$.
+It is however still unclear whether the present detection scheme is sensitive enough to detect single VH$^-$.
+
+\chapter{NV-NV cross-relaxations : the fluctuator model}
+This chapter will cover cross-relaxations between NV centers in the non-zero magnetic field regime, both from an experimental and theoretical point of view.
+The case of zero magnetic field will be treated in the next chapter.
+
+The presence of strong NV-NV dipolar interaction in dense ensemble ([NV] $\gtrsim 1\ \rm ppm$) proved to be both an obstacle for ensemble magnetometry \citep{zhou2020quantum} and an opportunity to explore and use the many-body interaction resulting from it \citep{zhou2020quantum, choi2017observation, kucsko2018critical, zu2021emergent}.
+
+We will see in this chapter how a model of fast and slow relaxing centers, the fluctuator model \citep{choi2017depolarization}, explains most observations related to the cross-relaxation, including the stretched exponential lifetime profile and the spectral broadening of the interaction.
+
+\section{Experimental observation of NV-NV CR}
+Let us start with the experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
+
+We make the distinction between what we will label ``equivalent" and ``non-equivalent" NV centers.
+Two equivalent centers are NV centers with the same spin Hamiltonian (cf eq.
+(\ref{eq.
+spin elec})), which implies that they both see the same transverse and longitudinal magnetic field.
+In contrast, non-equivalent centers are center with different single-spin Hamiltonian, which may result in different eigenstates and spin polarization for the two spins.
+
+While the NV physics can explain cross-relaxations between non-equivalent NV centers, it cannot explain cross-relaxation between equivalent NV centers.
+Nevertheless, as we will see in this section, CR between both equivalent and non-equivalent NV centers is observed for samples with high NV density.
+\subsection{NV-NV CR between non-equivalent NV centers}
+\label{non_equi_valent_CR}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/NV-NV_non_equivalent}
+\caption{a) Illustration of the four possible direction of the the NV axis (four classes of NV centers) with a magnetic field along class 1 b) Representation of the polarization of two non equivalent spins, one aligned with the magnetic field (well polarized) and one misaligned with the magnetic field (unpolarized).
+c) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
+d) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
+\label{non-equivalent NV-NV}
+\end{figure}
+
+In the absence of magnetic gradient, two NV centers can only be non-equivalent if the projection of the magnetic field upon their respective NV axis is different.
+Fig.
+\ref{non-equivalent NV-NV}-a) illustrates such a case: we label here the four possible NV axis orientation (or NV classes) 1 to 4 and we assume that the magnetic field is aligned with class 1.
+In this conditions, an NV center belong to class 1 will always be non-equivalent with NV centers from classes 2-4, as the projection of the magnetic field on their respective axis will never be the same.
+On the other hand, NV centers from the classes 2-4 are all equivalents, even though they do not necessary belong to the same class.
+
+Fig.
+\ref{non-equivalent NV-NV}-b) shows the difference between the NV centers belonging to class 1 (labeled NV$_\parallel$) and those belonging to classes 2-4 (labeled NV$_\nparallel$).
+As explained in sec.
+\ref{sec champs transverse}, the transverse magnetic field leads to a mixing of the eigenstates \footnote{The states of the misaligned NV on Fig.
+\ref{non-equivalent NV-NV}-b) are labeled $\ket{0},\ket{\pm 1}$ which is an abusive (but common) notation since the mixing caused by the transverse field means that the eigenstates of $S_z$ are no longer the Hamiltonian eigenstates.} and to a loss of the optical polarization mechanism.
+As a result, the NV$_\nparallel$ spins are less polarized than the NV$_\parallel$ ones, which is a necessary condition to observe cross-relaxation.
+
+Another necessary condition to observe cross-relaxation is the condition of co-resonance.
+In most situations, non-equivalent NV centers are not co-resonant, as the differences in the spin Hamiltonian generally leads to different eigenvalues.
+This is in contrast with equivalent NV centers which are by definition always resonant.
+Fig.
+\ref{non-equivalent NV-NV}-c) shows the transition frequencies of spin from class 1 and a spin from class 2-4 as a function of the magnetic field amplitude.
+Unintuitively, there is a co-resonance between the two $\ket{0} \to \ket{+1}$ transitions for $\abs{\mathbf{B}}=592\ \rm G$.
+For this particular value of the magnetic field, cross-relaxation between the NV centers from class 1 and those from classes 2-4 is possible.
+
+Finally, Fig.
+\ref{non-equivalent NV-NV}-d), which was presented in the last chapter, shows the experimental value of the total NV PL as a function of the magnetic field amplitude, in a scenario similar to the one we just described.
+We can indeed observe the signature of NV-NV CR in the presence of a PL dip at $\abs{\mathbf{B}}=592\ \rm G$.
+
+As previously mentioned, cross-relaxation between non-equivalent NV centers is easily explained by the physics of single NV centers, and has been known for more than thirty years \citep{holliday1989optical, van1989cross}.
+
+\subsection{NV-NV CR between equivalent NV centers}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/NV-NV_equivalent}
+\caption{CR between equivalent NV centers for sample ADM-15-1.
+(a) Representation of equivalent NV centers: either two NVs from the same class, or two classes with the same projected magnetic field.
+(b) Magnetic setup used for the experiment: a permanent magnet is used to apply a bias magnetic field, and an electromagnet is used to add a variable magnetic field.
+(c) Simulation showing the $\ket{0}\to\ket{-1}$ transition frequencies for the 4 classes of NV centers as a function of the scanned magnetic field.
+The transitions were computed based on ODMR spectra recorded for different magnetic field values.
+(d) Change in the PL of the NV centers ensemble as a function of the scanned magnetic field.}
+\label{equivalent NV-NV}
+\end{figure}
+
+More recently, experiments on dense NV ensemble ($[\rm NV ] >1\ \rm ppm$) or at low temperature have shown that there were cross-relaxations even between equivalent NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, jarmola2015longitudinal, choi2017depolarization}.
+
+Fig.
+\ref{equivalent NV-NV}-a) illustrates the two possible scenarios for equivalent NV centers: either two NVs from the same class, or two NVs from different classes with the same projection of the magnetic field upon their respective axis (i.e.
+$\mathbf{B}$ belongs to the plane of mirror symmetry of the two spins).
+
+In the second case, by moving the magnetic field in and out of the symmetry plane, we can change the resonance condition between the two NV classes and measure the contribution of NV-NV cross-relaxation to the NV PL and spin lifetime.
+To do so, we need to add an initial bias magnetic field in addition to the scanned magnetic field, as represented in Fig.
+\ref{equivalent NV-NV}-b).
+%When doing so, the magnetic field felt by the diamond won't have a fixed orientation during the scan, and when the magnetic field crosses certain crystalline planes (detailed below), its projection on two or more classes of NV centers will be the same.
+
+Fig.
+\ref{equivalent NV-NV}-c) shows the transition frequencies of the $\ket{0}\to\ket{-1}$ transition for the 4 classes of NV centers on sample ADM-15-1, an HPHT micro-diamond with $[\rm NV]\sim 3\ \rm ppm$.
+There are three magnetic field values $B_{\rm scan} \approx$ 12, 42 and 77 G for which there is an inter-class resonance.
+These values are reported in the figure with red dotted lines.
+Unlike the case presented in Fig.
+\ref{non-equivalent NV-NV}, all the inter-class resonances reported in Fig.
+\ref{equivalent NV-NV}-c) are between equivalent classes.
+
+Finally, Fig.
+\ref{equivalent NV-NV}-d) shows the PL contrast as the electromagnet field is scanned.
+There are 3 very clear dips when $B_{\rm scan} \approx$ 12, 42 and 77 G which coincide with inter-class resonances.
+We will see later that this PL dip is associated with a decrease of the NV centers spin $T_1$ time from both classes.
+
+This observation, and similar ones made by many groups \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization, akhmedzhanov2017microwave, giri2018coupled}, seem to indicate that there are CR between equivalent NV centers.
+This is however incompatible with our previous assumptions that equivalent NV centers are equally polarized and bright.
+To understand this phenomenon, we will need to introduce new hypotheses.
+
+\section{NV inhomogeneity and the fluctuator model}
+
+\subsection{CR in an inhomogeneous NV bath}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/inhomogene_2_spins}
+\caption{a) Schematics of the model: we consider a four level system consisting of 2 levels of 2 NV centers resonantly coupled through the flip-flop rate $\Gamma_{\rm dd}$.
+These two NV centers are each pumped in their respective $\ket{0}$ state via the green laser with a rate $\Gamma_{\rm las}$, and have their own relaxation rate $\Gamma_1$ and $\Gamma_1^*$ respectively.
+b) Average population of the two NV centers in the $\ket{0}$ state in the steady state as a function of $\Gamma_{\rm dd}$ and $\Gamma_1^*$.
+The steady state was computed by solving a classical rate equation where we fixed $\Gamma_{\rm las}=1000\ \rm Hz$ and $\Gamma_1=300\ \rm Hz$.}
+\label{inhomogene}
+\end{figure}
+A likely explanation to the observed NV-NV CR is that the NV centers are not all equivalents.
+
+Indeed, if we assume that, for instance, the relaxation rate $\Gamma_1=1/T_1$ is not strictly the same for each NV centers, but instead follows a certain distribution $\rho(\Gamma_1)$, then CR between NV centers with different $\Gamma_1$ can explain the observations in Fig.
+\ref{equivalent NV-NV}.
+
+Fig.
+\ref{inhomogene}-a) and b) illustrates this case.
+Our model here consists of two (resonantly coupled) NV centers with two different relaxation rates $\Gamma_1$ and $\Gamma_1'$.
+To complete the model, we add the rate $\Gamma_{\rm las}$ which is the optical pumping rate from the $\ket{-1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins \footnote{We consider here a two level system instead of the full three levels of the spin 1.
+The choice of $\ket{-1}$ instead of $\ket{+1}$ is arbitrary.}.
+We only consider here the incoherent dynamics of the population, modeled with the rates reported on Fig.
+\ref{inhomogene}-a).
+Solving for the steady state of these rate equations, we can compute the final population in the $\ket{0}$ state for both spins.
+
+Fig.
+\ref{inhomogene}-b) shows the spin population in the $\ket{0}$ state of both spins, $\rho_{00}$, as a function of the flip-flop rate $\Gamma_{\rm dd}$ and for various values of $\Gamma_1'$.
+The first thing to notice is that, as $\Gamma_1'$ is increased,  $\rho_{00}$ decreases.
+This is because NV$_2$ becomes less polarized as its decay rate increases.
+The second thing to notice is that, when $\Gamma_1'=\Gamma_1$, i.e.
+when the two NV centers are truly equivalent, the flip-flop rate has no incidence on  $\rho_{00}$.
+However, when $\Gamma_1'>\Gamma_1$,  $\rho_{00}$ decreases when the flip-flop rate increases.
+Intuitively, this can be thought of as NV$_2$ acting as a polarization drain on NV$_1$.
+
+We can also notice that the influence of the flip-flop rate on $\rho_{00}$ increases when the difference in decay rates between NV$_1$ and NV$_2$ increases.
+As a result, we can suppose that a greater inhomogeneity between the NV centers would result in a stronger NV-NV cross-relaxation effect.
+
+\subsection{Presentation of the fluctuator model}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.8\textwidth]{chapter 3/Figures/NV_vs_fluctuator}
+\caption{Representation of a ``normal" NV center with spin depolarization rate $\Gamma_1 \ll \Gamma_{\rm las}$ and a fluctuator NV with spin depolarization rate $\gamma_f \gg \Gamma_{\rm las}$.
+Red circles represent the spin population in the steady state of each spin}
+\label{NV vs fluct}
+\end{figure}
+Choi et al.
+in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism leading to a new depolarization rate $\gamma_f$.
+Normal and fluctuator NV are represented in Fig.
+\ref{NV vs fluct}: in their model, Choi et al.
+assume that the fluctuator decay rate is fast enough compared to the laser polarization rate $\gamma_f \gg \Gamma_{\rm las}$ that the fluctuators are effectively always depolarized.
+Fluctuators are therefore similar to the dark spins we introduced in the last chapter.
+
+We should note here that, while this simplifications seem a bit extreme, the fluctuator model is more than a toy model.
+There are good evidence of the presence of these dark NV centers in dense NV ensemble, as will be discussed below.
+
+In their model, the authors of \citep{choi2017depolarization} consider that the fluctuators act as a Markovian bath, meaning that the fluctuator density matrix will always read $\rho = \frac{1}{3} I$, regardless of its interaction with NV centers.
+With this assumption, the modification caused by the fluctuator bath on the ``normal" NV lifetimes can be analytically computed.
+
+Other possibilities to explain the change in the relaxation rate were considered, such as spin diffusion to non-polarized NV centers (outside of the laser spot) or phonon superradiance, but they were not considered to be viable explanations: spin diffusion is orders of magnitude too slow to explain this phenomenon \citep{choi2017depolarization}, and the NV-NV CR seem to be independent of the crystal temperature \citep{jarmola2012temperature, mrozek2015longitudinal} which is incompatible with phonon superradiance \citep{choi2017depolarization}.
+
+Before detailing the model's predictions, we will first contemplate the possible microscopic origin of these fluctuators, and review some physical systems were similar observations have been made.
+
+ \subsection{Possible microscopic origin of the fluctuators}
+
+We should note first that a similar problem of spin ensemble relaxation rate increasing with the spin concentration was observed more than six decades ago with phosphorus doped silicon \citep{feher1959electron,honig1960electron}.
+It was proposed that the cause of the relaxation was the presence of fast relaxing centers coupled to the slow relaxing centers through spin diffusion \citep{honig1960electron, sugihara1963spin, yang1968concentration, vugmeister1978spin, berman2005spin}.
+
+This conclusion was based on previous observations on the relaxation of nuclear spins \citep{bloembergen1949interaction, de1958relaxation, blumberg1960nuclear}, where some nuclear spins had a considerably reduced lifetime due to their proximity to paramagnetic ions.
+Similar observations were performed with Förster resonant energy transfer (FRET) in chromophore molecules \citep{forster1949experimentelle, eisenthal1964influence, yokota1967effects}, where excited donor molecules are coupled through (electric) dipole-dipole interaction to non-excited acceptor molecules.
+
+In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
+In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (cf eq.
+(\ref{eq.
+dipole quantique J coupling})).
+Both of these phenomena can lead to fast spin relaxation: electron tunneling can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction term ($J-$coupling) by the crystal phonons can directly excite the spin transitions \citep{honig1960electron}.
+While the dipole-dipole interaction terms (excluding the contact interaction) are also modulated by the phonons, this effect would be too small to account for the fast depolarization observed here.
+
+Similarly, \citep{choi2017depolarization} suggests that the origin of fluctuators in NV centers are closely packed NV centers or NV-impurity pairs which undergo rapid electron hopping and spin depolarization.
+To prove their point, they look at the charge dynamics in their sample and find that there is charge recombination in the dark with a corresponding tunneling rate between neighboring sites of $\sim 10\ \rm ns$.
+In contrast to P-doped Si, the NV centers are not the most abundant electronic defects in the crystal, this is why the NV$^--$N$^+$ pair in particular is thought to be a likely candidate for the fluctuators \citep{manson2018nv}.
+
+\subsection{Single NV center coupled to a single fluctuator}
+We now turn to the computation of the depolarization induced by the fluctuator bath on the NV centers. We will follow here the notations and calculation steps in \citep{choi2017depolarization}.
+
+Let us first consider the interaction between a single NV center and a fluctuator.
+Since we assume the fluctuators to be always depolarized, this step is similar to the coupling of an NV center to a dark spin done in the last chapter.
+
+We will first decompose the dipole-dipole Hamiltonian as a product between radial and angular part:
+\begin{equation}
+\mathcal{H}_{\rm dd} \approx - \frac{J_0}{r^3} \left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}+qS_z^1 S_z^2 \right] + h.c. ,
+\end{equation}
+where the expression of $J_0, g, h$ and $q$ is given in appendix \ref{Appendix eta}.
+$g, h$ and $q$ are dimensionless factors that are function of the relative radial positions and orientation of the two dipoles.
+
+We then introduce the dimensionless number $\eta$ defined as:
+\begin{equation}
+\label{eq eta}
+\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
+\end{equation}
+where $\gamma_f$ is the fluctuator decay rate for each possible channel between $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+This number $\eta$ encapsulates both the angular dependency (through $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
+
+We can note that the spectral response of a single fluctuator, for which we only consider the broadening caused by the decay rate $\gamma_f$, is a Lorentzian of half width $2\gamma_f$.
+This is different from the usual lifetime limited optical spectra which are Lorentzian of half width $1/2T_1=\Gamma_1/2$.
+The difference comes here from the fact that we consider thermalization going both ways (i.e.
+$\ket{e} \to \ket{g}$ and $\ket{g} \to \ket{e}$ instead of only $\ket{e} \to \ket{g}$) and that each state of the fluctuator can decay into two other states (see Fig.
+\ref{NV vs fluct}).
+
+Finally, one finds that the additional depolarization rate induced by the fluctuator on the NV center reads:
+\begin{equation}
+\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right)^2 \frac{\eta^2}{\gamma_f}.
+\end{equation}
+Compared to eq.
+(\ref{delta gamma 1}), the dependency on the inhomogeneous broadening of the spins $1/T_2^*$ is hidden in the $\eta$ factor, and we have the additional broadening $\gamma_f$ coming from the fluctuators very short lifetime.
+
+\subsection{Ensemble of NV centers coupled to a bath of fluctuators}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.4\textwidth]{chapter 3/Figures/NV_dans_bain_fluct}
+\caption{Schematics of an NV center in a fluctuator bath. The $\mathbf{r}_i$ vector represent the relative of position of fluctuator $i$ with respect to the NV center.}
+\label{NV + bain fluct}
+\end{figure}
+
+The next step is to compute the depolarization caused by the ensemble of fluctuators on a single NV center.
+
+We consider the situation pictured in Fig.
+\ref{NV + bain fluct}: the NV center is at the center of the frame and the position of each fluctuator is labeled $\mathbf{r}_i$.
+We consider that each fluctuator is an independent depolarization source, therefore the depolarization on a single NV center reads:
+\begin{equation}
+\gamma=\sum_i \gamma_s(\mathbf{r}_i).
+\end{equation}
+We then want to compute the distribution $\rho(\gamma)$ defined as:
+\begin{equation}
+\rho(\gamma)=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma \right).
+\end{equation}
+To do so, we need to determine the distribution of the fluctuators positions and orientations$\{r_i\}$. Assuming that they are homogeneously distributed in the bulk of the material, we find:
+\begin{equation}
+\rho(\gamma)=\frac{e^{-1/(4\gamma T)}}{\sqrt{4\pi \gamma^3 T}},
+\end{equation}
+where the time constant $T$ was introduced and is defined as:
+\begin{equation}
+\frac{1}{T}=\left(\frac{4\pi n_fJ_0\bar \eta}{3}\right)^2 \frac{\pi}{\gamma_f},
+\label{eq 1/T}
+\end{equation}
+with $n$ the fluctuator density and $\bar \eta$ the averaged value of $\abs{\eta}$:
+
+\begin{equation}
+\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta.
+\end{equation}
+
+Finally, the polarization dynamics from the ensemble of NV centers can be computed:
+\begin{equation}
+P(t)=\int_0^\infty \rho(\gamma)\, e^{-\gamma t}d\gamma= e^{-\sqrt{t/T}}.
+\end{equation}
+
+In conclusion, the fluctuators causes a depolarization on the NV ensembles with a timsecale $T$ given by \ref{eq 1/T}, and the dynamics of this depolarization is that of a stretched exponential \footnote{Stretched exponential generally refer to functions of the form $\exp(-(t/\tau)^{\beta})$.
+In this manuscript, unless specified otherwise, we will always refer to the case $\beta=1/2$ when mentioning stretched exponential.}, unlike the phonon-induced depolarization which is purely exponential.
+
+It should be noted that the stretched exponential nature of the decay is not specific to the fluctuator model, but results from localized noise sources, randomly distributed in the bulk (3D).
+Indeed, \citep{hall2016detection} came to the same conclusion by analyzing the resonant coupling of NV centers with a P1 bath.
+Theodor Förster in 1949 also predicted and observed stretched exponential dynamics in the context of FRET, where the same scaling laws are observed \citep{forster1949experimentelle}.
+
+\subsection{Computation of $\bar \eta$}
+\label{sec computation eta}
+
+The parameter $\bar \eta$ in eq.
+(\ref{eq 1/T}) is of crucial importance as it encapsulates the resonance condition between the NV centers and the fluctuators.
+This resonance condition, i.e.
+the spectral overlap between the fluctuators and the NV centers, can be decomposed in two essential parameters: one is the central frequency of each subgroup ($\omega_{NV}^0$ and $\omega_{f}^0$ respectively) which themselves depends on the magnetic field, the other is the spectral broadening of both the NV and the fluctuators.
+
+The aim of this section is to give an explicit formula of $\bar \eta$ as a function of $\omega_{NV}^0$, $\omega_{f}^0$ and $T_2^*$.
+
+We will start by separating the factor $\eta$ given in eq. (\ref{eq eta}) into a a geometric and a spectral part:
+\begin{align}
+\eta^2&=G^2(\{\theta\}) R^2(\omega_f,\omega_{NV}), \\
+G^2&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
+R^2(\omega_f,\omega_{NV}) &= \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{align}
+
+$G$ is a function which depends purely on the various angles of the problem $\{\theta\}$, while $R$ represents the resonance condition between the NV and the fluctuator.
+
+We then average the $\eta$ factor for an ensemble of fluctuators and NV centers:
+\begin{align*}
+\bar \eta &= \int \rm{Prob}(\eta) \abs{\eta} d\eta \\
+&= \left(\int \rho(\{\theta\}) |G(\{\theta\})| d\{\theta\} \right) \\
+&\, \times \left( \int \rho(\omega_{NV}) \rho(\omega_{f}) |R(\omega_f,\omega_{NV})| d\omega_{NV} d\omega_{f} \right) \\
+&= \bar G \cdot \bar R,
+\end{align*}
+where we introduced $\rho(\{\theta\})$ the probability distribution of the angles $\{\theta\}$, $\rho(\omega_{NV})$ the inhomogeneous spectral distribution of the ensemble of NV centers and $\rho(\omega_{f})$ the inhomogeneous spectral distribution of the ensemble of fluctuators.
+We also introduced the notations $\bar G$ and $\bar R$ as the average of $|G|$ and $|R|$.
+
+The computation of $\bar G$ will be further discussed in sec. \ref{sec quantitative T1} and in appendix \ref{delta gamma 1}. We will focus here on $\bar R$.
+
+To perform the computation, we need to know the distributions $\rho(\omega_{NV})$ and $\rho(\omega_{f})$.
+We can experimentally measure $\rho(\omega_{NV})$ though ODMR, and we can reasonably expect $\rho(\omega_{f})$ to follow the same probability.
+
+We unfortunately cannot analytically solve for $\bar R$ with standard distributions of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ (either Gaussians or Lorentzians).
+We can however solve for $\overline{R^2}$, which is close to $(\bar{R})^2$ as long as $2 \gamma_f \gg 1/T_2^*$.
+
+For a Lorentzian profile of $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ defined as:
+\begin{align*}
+\rho(\omega_{f})&=\frac{1}{\pi \Gamma_f} \frac{1}{1+ \left(\frac{\omega_f-\omega^0_f}{\Gamma_f}\right)^2} \\
+\rho(\omega_{NV})&=\frac{1}{\pi \Gamma_{NV}} \frac{1}{1+ \left(\frac{\omega_{NV}-\omega^0_{NV}}{\Gamma_{NV}}\right)^2},
+\end{align*}
+where $\Gamma_{f}=1/T_2^*(\rm fluctuator)$ and $\Gamma_{NV}=1/T_2^*(\rm NV)$ are the inhomogeneous broadening of the NV and fluctuator ensembles, and $\omega^0_{NV}$ and $\omega^0_{f}$ are the central angular frequencies of the NV and fluctuator ensembles.
+We then find that:
+
+\begin{align*}
+\overline{R^2}&= \int \rho(\omega_{NV}) \rho(\omega_{f}) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2} d\omega_{NV} d\omega_{f} \\
+&=\frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}} \cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
+\end{align*}
+
+In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f \gg 1/T_2^*$) in eq.
+\ref{eq 1/T}, we find that:
+\begin{equation}
+\label{eq 1/T avec T2*}
+\frac{1}{T} \propto \frac{1}{2\gamma_f + \Gamma_f + \Gamma_{NV}}\cdot \frac{1}{1+\left(\frac{\omega^0_{NV}-\omega^0_{f}}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}\right)^2}.
+\end{equation}
+
+If $\rho(\omega_{f})$ and $\rho(\omega_{NV})$ follow Gaussian profiles instead, $\overline{R^2}$ is the integral of a Voigt profile, the convolution of a Lorentzian and a Gaussian, which does not have a clear analytic solution.
+
+Eq.
+\ref{eq 1/T avec T2*} was obtained via several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
+Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$, $\omega_{NV}^0$ and $\omega_{f}^0$, which will be crucial to compare the predictions of the model with experimental data.
+
+\section{Experimental investigation of the fluctuator model}
+We will now show experimental results related to the predictions of the fluctuator model.
+\subsection{The stretched exponential lifetime}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/Stretch_vs_pas_stretch}
+\caption{T1 measurement following the protocol described in Fig.
+\ref{T1 soustraction}.
+The best exponential and stretched exponential fits are given in every case.
+a) Sample CVD-pink with B$\neq$0 and a single  probed class.
+The optimal $T_1$ values are $T_1^{\rm stretch}=1.9\ \rm ms$ and $T_1^{\rm exp}=4.1\ \rm ms$.
+b) Same measurement as a) in log scale.
+c) Sample ADM-15-2 with B=0.
+The optimal $T_1$ values are $T_1^{\rm stretch}=150\ \mu \rm s$ and $T_1^{\rm exp}=410\ \mu \rm s$.
+d) Same measurement as c) in log scale.}
+\label{stretch_or_not_stretch}
+\end{figure}
+
+The first prediction of the fluctuator model which can be experimentally verified is the stretched-exponential profile of the spin relaxation.
+The protocol to measure spin relaxation with a common mode rejection was presented in Fig.
+\ref{T1 soustraction}.
+
+Fig.
+\ref{stretch_or_not_stretch} shows the results of this protocol on two distinct samples and magnetic field configurations.
+In both cases, we show the results in linear and logarithmic scale as the log scale makes deviations from the standard exponential more apparent\footnote{The fits presented in Fig.
+\ref{stretch_or_not_stretch} and in the next figures were performed by fitting the real value of $S_1-S_2$ and not $\log(S_1-S_2)$.}.
+
+The spin relaxation results presented in Fig.
+\ref{stretch_or_not_stretch}-a) and b) was performed on sample CVD-pink (see sample details in appendix \ref{Appendix samples}), which despite a relatively high NV concentration shows only moderate amount of dipole-dipole spin relaxation.
+In order to further minimize the amount of cross-relaxations, we chose a magnetic field for which all four NV classes are spectrally isolated, so that only fluctuators from the same class as the NV center could lead to relaxation.
+As a result, we can see a relaxation profile mostly exponential with a lifetime $T_1=4.1\ \rm ms$ comparable to the phonon-limited $T_1$ value of $\sim 5\ \rm ms$ at room temperature \citep{jarmola2012temperature}.
+
+The second spin relaxation presented in Fig.
+\ref{stretch_or_not_stretch}-c) and d) was performed on sample ADM-15-2 which showed a great amount of dipole-dipole spin relaxation.
+To further enhance the cross-relaxation rate, the experiment was performed in zero external magnetic field so that every class of NV centers would be resonant.
+We can observe a relaxation profile that deviates significantly from tthe exponential profile, and which is nicely fitted by a stretched exponential.
+The $T_1$ value obtained from the fit is also an order of magnetite lower than the phonon-limited value, which is coherent with the fact that dipole-dipole interactions dominate the spin relaxation in this case.
+
+\subsection{Fitting $T_1$ profile in the general case}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/various_fit_formulae}
+\caption{Different fitting procedure on two $T_1$ measurements done on sample ADM-150-1 for $B=0$ and $B\neq0$.
+\\ a) Exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph})$.
+\\ b) Stretched exponential fit $S(\tau)=\exp (-\sqrt{\tau/T_1^{\rm dd}})$.
+\\ c) Bi-exponential fit $S(\tau)=\exp (-\tau/T_1^{\rm ph} -\sqrt{\tau/T_1^{\rm dd}})$.
+\\ d) Bi-exponential fit with fixed $T_1^{\rm ph}=5\ \rm ms$: $S(\tau)=\exp (-\tau/5 -\sqrt{\tau/T_1^{\rm dd}})$.}
+\label{various_fit_formulae}
+\end{figure}
+
+\begin{table}[htbp]
+\centering
+\caption{Fitting parameters for Fig.
+ \ref{various_fit_formulae}. The coefficient of determination $R^2$ for each fit is reported here.}
+ \label{fitting table}
+\begin{tabular}{c|ccc|ccc}
+\toprule
+Figure &  \multicolumn{3}{c}{$B=0$} & \multicolumn{3}{c}{$B\neq0$}\\
+\midrule
+{} &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$ (ms) & $1-R^2$ &  $T_1^{\rm ph}$ (ms)& $T_1^{\rm dd}$(ms)& $1-R^2$\\
+\midrule
+Fig. a) & 0.95 & * & $2\cdot 10^{-2}$ & 2.64 & * & $6\cdot 10^{-3}$ \\
+Fig. b) & * & 0.32 & $2\cdot 10^{-3}$& * & 1.03 & $2\cdot 10^{-2}$ \\
+Fig. c) & 6.75 & 0.44 & $6\cdot 10^{-4}$ & 4.22 & 7.19 & $2\cdot 10^{-4}$\\
+Fig. d) & \textbf{5} & 0.51 & $6\cdot 10^{-4}$ & \textbf{5} & 4.65 & $6\cdot 10^{-4}$\\
+\bottomrule
+\end{tabular}
+
+Bold characters indicate parameters that were arbitrarily fixed.
+\end{table}
+
+For many samples, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$, is of the same order of magnitude as the phonon limited lifetime $T_1^{\rm ph}$.
+Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, meaning that we have to include both aspects in the fitting formula.
+
+Fig.
+\ref{various_fit_formulae} shows four different fitting procedure on the two same relaxation measurements.
+The measurements were performed on the same sample ADM-150-1, with and without an external magnetic field ($\sim 50 G$).
+The magnetic field was strong enough to split the four classes and only one class was probed in the case $B\neq0$ , whereas all 4 classes were resonant in the case $B=0$, which strongly increases the NV-NV CR.
+
+The fits used here are either purely exponential or purely stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$.
+The values of the different fitting parameters used is reported in Table \ref{fitting table}.
+
+We can see that neither the purely exponential nor stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is poorly fitted by the stretched exponential.
+The protocols that include both exponential and stretched lifetimes correctly fit both curves.
+In the case of Fig.
+\ref{various_fit_formulae}-d), we arbitrarily fixed $T_1^{\rm ph}=5\ \rm ms$ since this is the value we typically measure on samples with low NV density, and we expect that the phonon-limited exponential lifetime is not modified by the NV concentration.
+
+Both the protocol where $T_1^{\rm ph}$ was fixed and the one where it was not correctly fit our data.
+We therefore decided to use the protocol where $T_1^{\rm ph}$ was fixed in order to minimize the number of free parameter in the fitting function.
+This way, the values of $T_1^{\rm dd}$ obtained on different measurements can be directly compared.
+
+We should note however that the values of $T_1^{\rm dd}$ we obtain are not absolute.
+Indeed, in the case of Fig.
+\ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
+The values of $T_1^{\rm dd}$ obtained with this method have therefore to be be compared to values from the same method and the same fixed value of $T_1^{\rm ph}$.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/betas}
+\caption{a) Same two measurements as Fig.
+\ref{various_fit_formulae} with the fitting formula $S(\tau)=\exp ((-\tau/T_1)^{\beta})$.
+The fitting parameters are $\beta=0.58$ and $T_1=0.46\ \rm ms$ for $B=0$, and $\beta=0.80$ and $T_1=2.15\ \rm ms$ for $B\neq0$.
+b) Optimal $\beta$ parameter found for each $T_1$ measurement as a function of the external magnetic field (still on sample ADM-150-1).}
+\label{betas}
+\end{figure}
+
+Finally, another fitting procedure is shown in Fig.
+\ref{betas}.
+This time a single stretched exponential with an arbitrary stretch factor $\beta$ is used, and satisfyingly fits the data.
+Fig.
+\ref{betas}-b) shows the optimal $\beta$ parameter as a function of the external magnetic field.
+It confirms that the $T_1$ profile gets closer to a stretched exponential ($\beta=0.5$) when $B$ goes to 0.
+Even though this method also yields satisfying fits, it employs more free parameters ($\beta$ and $T_1$) than the method of \ref{various_fit_formulae}-d).
+We therefore decided to go with the later.
+
+\subsection{The fluctuators linewidth}
+
+\subsubsection{The fluctuators spectral response}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/fluctuator linewidth}
+\caption{Measurement of the fluctuators linewidth on sample ADM-150-2 using the same setup as in Fig.
+\ref{equivalent NV-NV}.
+a) Simulation of the $\ket{0} \to \ket{-1}$ transition frequency for the four classes of NV centers and actual frequencies of the two central classes from ODMR measurements.
+b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.8 \pm 0.2\ \rm MHz$.
+The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor of 2}
+\label{fluct linewidth}
+\end{figure}
+
+We now move to the second major prediction of the fluctuator model which is the spectral broadening of the fluctuators caused by the their relaxation rate $\gamma_f$.
+Indeed, following the arguments presented in \citep{choi2017depolarization}, the fluctuators homogeneous spectral response is a Lorentzian of half-width $2\gamma_f$.
+The inhomogeneous spectral response of the ensemble of fluctuators is then further broadened by the same electric, magnetic or strain noises than the NV centers, which is characterized by $1/T_2^*$.
+
+The fluctuator are effectively dark to ODMR measurement since they are not polarized, meaning that we cannot directly measure their linewidth as we would do for normal NV centers.
+We can however measure their linewidth through cross-relaxation, similarly to the detection of dark spins presented in the previous chapter.
+
+Fig.
+\ref{fluct linewidth} shows an experiment similar to the one presented in Fig.
+\ref{equivalent NV-NV}.
+A variable magnetic $B_{\rm scan}$ in addition to an offset magnetic field $B_{\rm DC} \sim 100\ \rm G$ are used in order to create a crossing between two classes of NV centers.
+This time however, the spin relaxation was measured instead of the PL.
+
+Fig.
+\ref{fluct linewidth}-a) shows the simulated transitions of the four classes of NV centers as a function of the magnetic field.
+We took particular care in measuring the central frequencies of the two classes of interest (when the two lines could be clearly separated), since having a precise measurement of the detuning between the two classes $\Delta \nu$ is key to get a precise value of the fluctuator linewidth.
+
+Fig.
+\ref{fluct linewidth}-b) shows the dipole-dipole induced relaxation rate $\Gamma_1^{\rm dd}=1/T_1^{\rm dd}$ as a function of the detuning $\Delta \nu$ between the two central classes.
+For each magnetic field value, a $T_1$ profile was recorded and fitted following the procedure described in Fig.
+\ref{various_fit_formulae}-d) where we again fixed $T_1^{\rm ph}=5\ \rm ms$.
+We also show the ODMR spectrum of a single NV class for comparison.
+
+We can clearly see that the $\Gamma^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
+If we assume that both NV classes spectral response are Lorentzian of same width $\sigma$, then the spectral overlap between the two classes is also a Lorentzian of width $2 \sigma$ \footnote{For other line profile, the overlap would be smaller than $2 \sigma$.
+For instances, the overlap of two Gaussians of width $\sigma$ is a Gaussian of width $\sqrt{2} \sigma$}.
+Fig.
+\ref{fluct linewidth}-b) also shows the ODMR line scaled by a factor $2$, which is still significantly narrower than the $1/T_1^{\rm dd}$ line.
+
+We attribute this broadening to the fluctuators lifetime.
+Another potential explanation for the broadening could be the interaction strength between the NV center and the fluctuator, but for [NV] $\sim$ 5 ppm, the average dipole-dipole coupling strength between closest NV neighbor is $\expval{\mathcal{H}_{\rm dd}} \sim 46\ \rm kHz$, which cannot explain the $\sim 4\ \rm MHz$ broadening observed here.
+
+\subsubsection{The fluctuators lifetime}
+\label{sec fluctuators lifetime}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.6\textwidth]{chapter 3/Figures/deconvolution}
+\caption{Blue curve $1/T_1^{\rm dd}$ data presented in Fig.
+\ref{fluct linewidth}.
+Red curve: deconvolution of the blue curve by the ODMR line shown on the same figure.
+Green curve: deconvolution of the red curve   by the ODMR line.
+All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters)}
+\label{deconvolution}
+\end{figure}
+
+We now want to extract the fluctuators lifetime $T_1^f=1/\gamma_f$ to get further information on their potential nature.
+
+To do this we first assume that the dipole induced decay rate is proportional to the overlap of the spectral responses of the NV centers and fluctuators $S_{NV}$ and $S_f$.
+We will take $S_{NV}( \omega)$ and $S_{f}( \omega)$ as functions centered in $\omega=0$, and consider that the spectral responses are shifted by $\omega^0_{NV}$ and $\omega^0_{f}$ respectively, the central frequencies of each group.
+We then have:
+\begin{align*}
+\frac{1}{T_1^{\rm dd}} &\propto \int S_{NV} (\omega-\omega_{NV}^0) S_f (\omega-\omega_{f}^0) d\omega \\
+&= \int S_{NV} (\omega-\Delta \omega) S_f (\omega)
+ d\omega \\
+ &= (S_{NV}*S_{f})(\Delta \omega),
+\end{align*}
+
+where $\Delta \omega=\omega_{NV}^0-\omega_{f}^0$.
+
+Using the approximation $\bar {R ^2} \approx (\bar{R})^2$ detailed in sec.
+\ref{sec computation eta}, we can then decompose the fluctuator spectral response as a convolution of the lifetime contribution and the inhomogeneous broadening:
+\begin{equation}
+S_{f}(\omega)=\left(S_f^*(\omega)\right)*\left(\frac{1}{1+\left(\frac{\omega}{2 \gamma_f}\right)^2} \right),
+\end{equation}
+
+where $S_f^*$ is the spectral broadening caused by the inhomogeneous distribution of the fluctuators, which we will suppose to be equal to $S_{NV}$.
+With these assumptions, we are now ready to extract $\gamma_f$ from the data of Fig.
+\ref{fluct linewidth}.
+
+The first way to do this, as was done for NV-P1 CR in \citep{hall2016detection}, is to deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm.
+The result of the deconvolution is shown in Fig.
+\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve).
+Fitting with Lorentzians, we find a final value $\Gamma_2 =2\gamma_f= (2\pi) 6.4 \pm 0.4\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=50 \pm 3\ \rm ns$\footnote{We define here the fluctuators lifetime $T_1^f$ as $\frac{1}{\gamma_f}$ for simplicity, but it should be noted that the actual lifetime of the fluctuators states is $\frac{1}{2\gamma_f}$ due to the competition between the two possible decay channels for each state (cf Fig.
+\ref{NV vs fluct}).}.
+
+Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\frac{\sqrt{\ln (2)}}{\pi T_2^*}$ and a Lorentzian of HWHM $f_L=\frac{2 \gamma_f}{2 \pi}$:
+\begin{align*}
+\Gamma_1^{\rm dd}&\propto \mathcal{G}(f_G)*(\mathcal{L}(f_L)*\mathcal{G}(f_G)) \\
+&\propto \mathcal{L}(f_L)*\mathcal{G}(\sqrt{2} f_G),
+\end{align*}
+where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
+We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{ida2000extended}:
+\begin{equation}
+f = [f_G^5 + 2.69269 f_G^4 f_L + 2.42843 f_G^3 f_L^2 + 4.47163 f_G^2 f_L^3 + 0.07842 f_G f_L^4 + f_L^5]^{1/5}.
+\end{equation}
+
+With $f=8.8 \pm 0.2\ \rm MHz$ and $\sqrt{2}f_G=4.3 \pm 0.2\ \rm MHz$, we find $f_L=6.5 \pm 0.3\ \rm MHz$ for a final lifetime value $T_1^f = 49 \pm 2\ \rm ns$, a value consistent with the one found with the deconvolution method.
+
+Doing similar experiments and calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48\ \rm ns$, a value very close to the ones we obtained above, despite using a completely different sample.
+
+In conclusion, we observed that the fluctuators linewidth was indeed broadened by their relaxation rate which justifies the initial hypothesis that $T_1^f \ll T_1^{NV}$.
+It even shows that $T_1^f$ is of the same order of magnitude as the optical lifetime ($\sim 15\ \rm ns$) which explains why the fluctuators cannot be optically polarized.
+
+\section{Resonance between several NV classes}
+
+The main tool to probe NV-NV CR is to use the co-resonances between the NV classes.
+While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances.
+In this section, we will discuss the various scenarios where NV classes overlap, and quantify theoretically and experimentally the dipole-induced relaxation rate in each case.
+
+\subsection{Geometric conditions for inter-class resonance}
+\label{sec geometrie resonance}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{chapter 3/Figures/Resonance_geometries}
+\caption{ODMR spectra on sample ADM-15-3 as a function of the magnetic field orientation.
+The field amplitude is $\sim 60\ \rm G$.
+Red numbers represent the number of resonant classes for each ODMR line.
+The cubes represent the diamond unit cell, and the red planes/arrow the possible orientations of the magnetic field}
+\label{ODMR_geometries}
+\end{figure}
+
+There are four possibilities to obtain inter-class resonance (aside from the non-equivalent CR discussed in sec. \ref{non_equi_valent_CR}). These possibilities are:
+\begin{itemize}
+\item $\mathbf{B} \in \{110\}$.
+We refer by $\{110\}$ as any plane orthogonal to a $\langle 110 \rangle$ or equivalent direction.
+In this magnetic configuration, two classes are at resonance and the two other ones are spectrally isolated.
+\item $\mathbf{B} \in \{100\}$. In this magnetic configuration, the four classes form two pairs of co-resonance.
+\item $\mathbf{B} \parallel \langle 111 \rangle$.
+In this configuration discussed in the last chapter, one class is aligned with the magnetic field and is spectrally isolated while the three other form a resonant triplet.
+\item $\mathbf{B} \parallel \langle 100 \rangle$. In this configuration also discussed in the last chapter, the four classes are resonant.
+\end{itemize}
+We should add to this list the case $\mathbf{B}=0$ where all four classes are also resonant, and any other possibility (which we will label random $\mathbf{B}$) where all four classes are spectrally separated.
+
+An ODMR spectrum and a representation of the magnetic field in the diamond unit cell for each of these six possibilities is shown in Fig. \ref{ODMR_geometries}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/carte}
+\caption{PL contrast as a function of the magnetic field polar and azimuthal angle with respect to the [100] axis on sample CVD-pink.
+The field amplitude is $\abs{B} \sim 115 G$.
+The locus of the magnetic field in specific planes is noted by red dashed lines.}
+\label{Carte}
+\end{figure}
+
+This relation between NV-NV class resonances and the magnetic field being in specific crystalline planes or directions means that we can determine the crystal main axes simply by PL experiments.
+
+Fig.
+\ref{Carte} shows a map where we monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer.
+We have also reported the locus of $\mathbf{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].
+
+We know that when $\mathbf{B}$ belongs to any of these four planes, there is a co-resonance between at least two NV classes.
+This co-resonance leads to an increase in NV-NV CR and a faster depolarization of the spins involved, which ultimately leads to a drop in PL.
+We can indeed see in fig.
+\ref{Carte} that there is clear correlation between the loci of $\mathbf{B}$ and drops in PL.
+We can also see that the PL is at its lowest when $\mathbf{B}\parallel [001]$, which corresponds to a four class degeneracy.
+
+\subsection{Quantitative modification of $T_1$}
+\label{sec quantitative T1}
+\begin{table}[htbp]
+\centering
+\caption{Theoretical and experimental values of $\Gamma_1^{\rm dd}=1/T_1^{\rm dd}$ for various magnetic field configurations, expressed as a function of the value found for an isolated class.
+The experimental values were measured on sample ADM-15-3 using the protocol of Fig.
+\ref{various_fit_formulae}-d) with $T_1^{\rm ph}= 5\ \rm ms$.
+The uncertainty corresponds to the precision of the fit parameters.}
+ \label{T1 champ mag}
+\begin{tabular}{c|cc}
+\toprule
+$\Gamma_1^{\rm dd} (\mathbf{B})$ &  Theory & Experimental \\
+\midrule
+random $\mathbf{B}$ (1 class) & $\Gamma_0^{\rm th}$ & $1.53\pm 0.04$ ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & $5.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & $4.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & $11.6 \pm 0.4$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & $14.1 \pm 0.5$ $\Gamma_0^{\rm exp}$ \\
+$\mathbf{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & $19.9 \pm 0.8$ $\Gamma_0^{\rm exp}$ \\
+\bottomrule
+\end{tabular}
+
+($^*$) The case $\mathbf{B}=0$ will be detailed in the next chapter. The theoretical value is a lower bond.
+\end{table}
+
+The fluctuator model allows quantitative predictions of the change in the spin lifetime, thanks to eq.
+\ref{eq 1/T}.
+There are however several unknowns in this equation, such as the fluctuator density and lifetime which can be hard to estimate.
+It is however relatively easy to compare the predictions of eq.
+\ref{eq 1/T} for various magnetic field configurations.
+
+We decided to take the case of an isolated class as the baseline value for the dipole-dipole decay rate $\Gamma_0$, and to express the decay rates when 2, 3 or 4 classes are co-resonant as a function of $\Gamma_0$.
+Doing so allows us to normalize $n_f$ and $\gamma_f$ in eq.
+\ref{eq 1/T} and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
+
+$\bar G$ can be analytically or numerically computed for each of the magnetic configurations listed before, as detailed in appendix \ref{Appendix eta}.
+This allows us to predict the value of the spin decay rate in every situations as a function of the decay rate of an isolated class.
+
+Table \ref{T1 champ mag} shows these predicted values, along with experimental ones on sample ADM-15-3.
+The experimental values were obtained with the protocol described in Fig.
+\ref{various_fit_formulae}-d) with a double exponential fit where we fixed again $T_1^{\rm dd}=5\ \rm ms$.
+
+The theoretical values are in a good qualitative agreement with the experimental ones, both ranking the different scenarios in the same order.
+The quantitative predictions of the theory however are rather poor, being off by more than a factor of 3 for the shortest lifetimes.
+
+It should be noted that the experimental values were heavily influenced by the choice of the parameter $T_1^{\rm ph}$.
+Using for example $T_1^{\rm ph}=2.5\ \rm ms$ gave results closer to the theoretical ones.
+Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded an increased depolarization rate by a factor $\sim 4$, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one\footnote{We can also see in Fig.
+\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration leads to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, which is 4 times smaller than the predicted factor of 10.}.
+
+The failure of the fluctuator model to accurately predict the changes in $T_1$ is one of the reasons that led us to believe that the model is incomplete.
+Other reasons will be presented in the following part.
+
+\section{Conclusion and perspectives}
+
+In conclusion, we have seen in this chapter that dense NV ensemble present ``anormal" CR between the NV centers, and we have shown how the fluctuator model developed by Choi et al.
+in \citep{choi2017depolarization} can explain this behavior.
+The fluctuator model not only explains the NV-NV CR, but also correctly predicts the stretched exponential lifetime profile.
+It also correctly predicts that the fluctuators have a broader spectral response than the other NV centers.
+Nevertheless, the model does not describe perfectly every aspects of NV-NV relaxation, and there still are many aspects of this problem, both theoretical and experimental, to explore.
+
+In what follows we will describe what could be interesting investigation points to further understand the NV-NV relaxation and the fluctuator model.
+
+\subsection{Limitations of the fluctuator model}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 3/Figures/Anormal_T1}
+\caption{$T_1$ measurements on sample ADM-15-4, fitted with pure exponential or stretched exponential.
+a) $T_1$ of a single isolated class when $\mathbf{B} \neq 0$.
+The optimal fit parameters are respectively $T_1^{\rm exp}=1.43\ \rm ms$ and $T_1^{\rm stretch}=0.56\ \rm ms$.
+b) $T_1$ of the four degenerate classes when $\mathbf{B}=0$.
+The optimal fit parameters are respectively $T_1^{\rm exp}=0.15\ \rm ms$ and $T_1^{\rm stretch}=0.05\ \rm ms$.}
+\label{anormal T1}
+\end{figure}
+
+We mentioned previously that the discrepancy between the theoretical and experimental $T_1$  values were in contradiction with the current fluctuator model, although the difference could come in part from the $T_1$ fitting procedure.
+Another, more glaring problem is the fact that some $T_1$ profile are not stretched exponential.
+
+Fig.
+\ref{anormal T1} shows $T_1$ measurement on sample ADM-15-4, either on a single isolated class ,where the dipole-dipole interactions are at their weakest, or on all four resonant classes at $\mathbf{B}=0$, where the interactions are at their strongest.
+
+Although the spin lifetime still gets $\sim$ 10 times shorter in the $\mathbf{B}=0$ case, indicating that dipole-dipole interaction still plays an important role in the spin relaxation, this time the lifetime profile does not converge to a purely stretched exponential.
+In fact it seems to be the opposite: the lifetime profile is closer to a pure exponential for the shorter $T_1$ (especially for the very short times), and closer to a stretched exponential for the longer $T_1$, which is in complete contradiction with the fluctuator model and the previous observations in Fig.
+\ref{stretch_or_not_stretch} and \ref{various_fit_formulae}.
+
+We should note that most sample studied follow the fluctuator model, in that the their lifetime profile gets closer to a stretched exponential when the dipole-dipole relaxation rate is increased.
+Sample ADM-15-4 seem to be an anomaly in this regard, since most of the samples from the same batch (including sample ADM-15-2 in Fig.
+\ref{stretch_or_not_stretch}) seem to follow the ``normal" behavior predicted by the fluctuator model.
+However, some other samples, including bulk HPHT samples and 150 $\mu$m Adamas micro-diamond also showed short (non stretched) exponential lifetimes.
+
+It seems therefore that the current fluctuator model does not correctly describe the spin dynamics of every NV-dense samples.
+Looking at what was previously done for P-doped SI \citep{vugmeister1978spin} or for FRET \citep{yokota1967effects}, the fluctuator model could probably be improved by taking into account the finite lifetime of the fluctuators (i.e.
+the saturation of the NV-fluctuator CR), and the spin diffusion between NV centers (the fluctuator model only considers NV-fluctator interactions and not NV-NV interactions).
+
+\subsection{Investigation of the fluctuators nature}
+While the hypothesis of fluctuators being clusters of NV centers or NV-other spins is strengthened by the precedents in P-doped Si \citep{honig1960electron} or F centers in KCl \citep{warren1964spin}, there are still many unknowns on the exact nature of the fluctuators.
+There are at least two aspects that merit to be investigated.
+
+The first one would be to study the fluctuator spectral profile on various samples showing NV-NV CR.
+In particular it would be interesting to know whether this profile is always Lorentzian, and what its linewidth is.
+In the sample studied on Fig.
+\ref{fluct linewidth}, we found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough our sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 15 ppm$ sample used in \citep{choi2017depolarization} \footnote{Although the initial paper claimed a NV concentration $[\rm NV] \sim 45 ppm$, a more recent paper re-established the concentration at $\sim 15$ ppm \citep{zhou2020quantum}}.
+If $\gamma_f$ was found to be somewhat of a constant among all samples showing NV-NV CR, this would give us a lot of information on the potential nature of the fluctuators.
+Studying the temperature dependency of $\gamma_f$ could also give us information on whether the depolarization mechanism is due to tunneling or modulation of the $J-$coupling by the phonons.
+
+The other aspect would be to identify potential defects that could form fluctuators when paired with NV centers.
+The main candidate for this would be substitutional nitrogen (P1 or P1$^+$ centers) since they are the most abundant electronic spin species in the diamond we use.
+The ideal way to confirm this hypothesis would be to compare the spin $T_1$ for  different samples with the same [NV] concentration, but different [P1] concentration.
+If indeed P1 centers play a role in the presence of fluctuators, we would expect a shorter NV $T_1$ time in the sample with high P1 concentration.
+
+\chapter{NV-NV cross-relaxation at low magnetic field: application to magnetometry}
+
+In the last chapter we have identified a dipole-dipole relaxation mechanism within dense NV centers ensemble, and we have studied its influence in non-zero magnetic field.
+In this chapter we will cover the low to zero magnetic field region and see how the depolarization at low magnetic field can be used for magnetometry.
+
+This chapter is constructed as follows: We will first describe in sec. \ref{sec 4.1} the physics of NV centers under low or transverse magnetic field. In sec. \ref{sec causes zero field}
+we will present the different zero field depolarization causes that we have identified as well as their theoretical impact on the NV relaxation rates according to the fluctuator model.
+In sec.
+\ref{sec 4.3} we will confront the predictions of the fluctuator model with experiments under low and transverse magnetic field.
+
+We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field.
+We will start by giving a short overview of NV ensemble magnetometry in sec.
+\ref{sec 4.4} before characterizing the LFDM protocol in sec.
+\ref{sec LFDM} and comparing it to other magnetometry protocols in sec.
+\ref{sec 4.6}.
+Finally we will give some perspectives on how to improve LFDM and on potential applications.
+
+The results presented in this chapter are based in large part on the article \citep{pellet2022spin}.
+
+\section{NV spin Hamiltonian under low and transverse fields}
+\label{sec 4.1}
+We will start this chapter by studying the single NV physics under low and transverse magnetic field.
+On of the main difference with the previously studied regime is that the eigenstates of the Hamiltonian are no longer $\{ \ket{0}, \ket{+1}, \ket{-1} \}$, which will then in turn modify the dipole-dipole interaction between the NV centers.
+
+We will first describe the spin Hamiltonian of the NV center under low magnetic field and determine the dominant factors in our samples, then do the same for transverse magnetic field, and finally investigate the modification of the ODMR lineshapes in those regimes.
+
+\subsection{NV spin Hamiltonian in zero external magnetic field}
+\label{sec. Hamltonian 0 B}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/ESR_simus}
+\caption{Monte-Carlo simulations of inhomogeneous zero field ODMR for different noise sources.
+a) Magnetic noise: each of the $\mathbf{B}$ components is sampled over a Gaussian of deviation $\sigma=2\ \rm G$.
+b) Strain noise: each of the components of the strain tensor $\bar{\bar{\varepsilon}}$ is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
+c) Electric noise: each of the $\mathbf{E}$ components is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
+d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
+\label{simus ESR}
+\end{figure}
+
+In the absence of a strong external magnetic field, the dominant terms of the NV spin Hamiltonian originate from local noises.
+These noises are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}.
+
+Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field.
+The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.
+
+The NV$^-$spin Hamiltonian in the $\{ \ket{-1}, \ket{+1} \}$ basis can be written as \citep{udvarhelyi2018spin}:
+\begin{equation}
+\mathcal{H}=\begin{pmatrix}
+D-\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}}) & f_\perp(\mathbf{E}) + g_\perp(\bar{\bar{\varepsilon}})\\
+f^*_\perp(\mathbf{E}) + g^*_\perp(\bar{\bar{\varepsilon}})&D+\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}})
+\end{pmatrix},
+\label{Hamiltonien pm1}
+\end{equation}
+where $B_\parallel$ is the component of the magnetic field along the NV axis, and $f_\parallel, f_\perp, g_\perp$, and $g_\parallel$ are functions of the electric field $\mathbf{E}$ and the strain tensor $\bar{\bar{\varepsilon}}$, whose expressions are:
+
+\begin{align}
+f_\parallel(\mathbf{E})&=d_\parallel E_z, \\
+f_\perp(\mathbf{E})&=d_\perp ( E_x + i E_y), \\
+g_\parallel(\bar{\bar{\varepsilon}})&= h_{41}(\varepsilon_{xx}+\varepsilon_{yy})+h_{43} \varepsilon_{zz}, \\
+g_\perp(\bar{\bar{\varepsilon}}) &= \frac{1}{2} \left[ h_{16}(\varepsilon_{zx}+i \varepsilon_{zy}) + h_{15}\left(\frac{\varepsilon_{yy}-\varepsilon_{xx}}{2}+i\varepsilon_{xy}\right) \right],
+\end{align}
+where $d_\parallel=0.35 \ \rm{Hz\, cm/V}$ and $d_\perp=17\ \rm{Hz\, cm/V}$ have been measured experimentally \citep{van1990electric}, and $h_{43}=2300\ \rm{MHz}$, $h_{41}=-6420\ \rm{MHz}$, $h_{15}=5700\ \rm{MHz}$ and $h_{16}=19660\ \rm{MHz}$ were computed through DFT \citep{udvarhelyi2018spin} and show reasonable agreement with experiments \citep{barson2017nanomechanical}.
+
+%Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.
+%
+%The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting.
+For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$.
+The shift in energy will therefore tend to blur the energy split when averaging over a large number of spins.
+
+Fig.
+\ref{simus ESR}-a) b) and c) shows the result of Monte-Carlo simulations of the ODMR of an ensemble of NV centers for different noise sources.
+The simulation work as follow: one of the three possible noise source (magnetic field, electric field or strain) is sampled randomly while the two other are kept at zero.
+The eigenvalues of the spin Hamiltonian given in eq.
+\ref{Hamiltonien pm1} are then computed and the process is repeated $10^6$ times to create the histograms shown in Fig.
+\ref{simus ESR}.
+
+Fig.
+\ref{simus ESR}-d) shows an experimental zero-field ODMR which is typical of the samples used in this manuscript.
+In particular we can notice a central splitting of the spectrum.
+This splitting is only present in the simulation when the dominant noise is the electric field.
+We then conclude that the dominant noise source in our sample (in the absence of external magnetic field) is the local electric field.
+
+The authors of \citep{mittiga2018imaging} came to the same conclusion for the samples with dense NV ensemble that they studied by using a more sophisticated simulation model.
+
+%We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample.
+In particular, we do not shield for the earth magnetic field in our experiment.
+The splitting caused by the earth magnetic field however has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$, which is $\sim 4$ times weaker than the splitting we typically observe.
+
+Given that $d_\perp \approx 50 d_\parallel$, we can consider that the dominant electric field noise comes from the transverse electric field.
+We will then adopt the following simplified Hamiltonian for the zero external magnetic field regime:
+\begin{equation}
+\mathcal{H}=\begin{pmatrix}
+D&0&d_\perp E_\perp^* \\
+0&0&0 \\
+d_\perp E_\perp &0&D
+\end{pmatrix},
+\end{equation}
+whose eigenvectors are $\ket{0}$ and$\ket{\pm}$ of eigenvalues 0 and $D\pm d_\perp \abs{E_\perp}$, where $\ket{\pm}$ are defined as:
+\begin{equation}
+\ket{\pm}=\frac{\ket{+1}\pm e^{-i\phi_E}\ket{-1}}{\sqrt{2}},
+\end{equation}
+where $\tan(\phi_E)=E_y/E_x$.
+
+\subsection{NV spin Hamiltonian under purely transverse magnetic field}
+\label{sec B transverse}
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.7\textwidth]{chapter 4/Figures/map_etats_propres}
+%\caption{Closeness of the Hamiltonian most excited state $\ket{3}$ with either $\ket{+1}$ or $\ket{+}$ as a function of the external magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
+%\label{map champ transverse}
+%\end{figure}
+
+We now turn to the case of a magnetic field that is orthogonal to the NV axis, i.e.
+$\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
+In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.
+
+In this regime, the NV Hamiltonian eigenstates are similar to the case where it is dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{-}, \ket{+}$\citep{qiu2021nuclear, qiu2022nanoscale}, of respective eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D,D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
+\begin{equation}
+\ket{\pm}=\frac{\ket{+1}\pm e^{-2i\phi_B}\ket{-1}}{\sqrt{2}},
+\end{equation}
+and $\tan(\phi_B)=B_y/B_x$.
+
+For the case where $d_\perp E_\perp \sim \frac{(\gamma_e B_\perp)^2}{D}$ and $\phi_E \neq 2\phi_B$, the eigenstates of the Hamiltonian are still of the form $\ket{0},\ket{\pm}$ with a relative angle $\phi$ in between $\phi_E$ and $2\phi_B$.
+
+In conclusion, whenever the spin Hamiltonian is dominated by the local electric field or by the transverse magnetic field, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+
+\subsection{Hyperfine and inhomogeneous broadening}
+\label{sec modif T2*}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/Comparaison_ESR_T2}
+\caption{ODMR lineshapes for longitudinal, transverse and low magnetic fields on samples CVD-pink and Sumi-2.
+a) ODMR spectrum of a single class of NV centers for a longitudinal magnetic field ($B_\parallel \sim 100\ \rm G$).
+The microwave frequency is given with respect to the ODMR line central frequency.
+b) ODMR spectrum of a single class for a purely transverse magnetic field ($B_\perp \sim 100\ \rm G$).
+c) ODMR of all 4 classes for 0 external magnetic field.
+d) Table reporting the full width at half maximum of each ODMR line.}
+\label{ESR for T2*}
+\end{figure}
+
+A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
+The energy of the $\ket{\pm 1}$ states depends linearly on the longitudinal magnetic field, and quadratically on the transverse electric and magnetic fields.
+On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and quadratically on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp$.}.
+
+The shape of the ODMR lines (i.e.
+the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis, as we established in the previous part.
+Further details on the low magnetic field regime are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.
+
+Fig.
+\ref{ESR for T2*} shows ODMR profile on two samples, CVD-pink which shows weak dipole-induced depolarization and Sumi-2 which shows strong dipole-induced depolarization, in three distinct regime: for a purely longitudinal magnetic field, for a purely transverse magnetic field, and without magnetic field.
+The linewidths of the spin transition \footnote{We consider here the full linewidth of the electronic levels, i.e.
+including all three hyper fine levels} for both samples are reported in the Table \ref{ESR for T2*}-d).
+In both cases, the linewidths are smaller in the $B=0$ and $B \perp NV$ cases than in the $B \parallel NV$ case, mainly due to the modification of the hyper-fine coupling.
+The linewidth modification reaches up to $\sim 50\ \%$ for the CVD sample and $\sim 20\ \%$ for the HPHT one.
+
+We are particularly interested in the spin transition linewidth since it directly impacts the dipole-dipole relaxation rate (as seen in sec.
+\ref{sec computation eta}): the smaller the linewidth, the higher the dipole-dipole relaxation rate is.
+
+\bigskip
+
+To summarize this section, we have seen that (for the samples used in this chapter) the NV center spin Hamiltonian is dominated by transverse electric field at low external magnetic field, which causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$.
+
+We have also seen that the spin Hamiltonian has the same eigenstates for purely transverse magnetic field than it does for low magnetic field.
+
+Finally we have seen that this change of eigenstates occasions a modification of the spectral profile and linewidth of the spins.
+
+\section{Theory of the low field depolarization}
+\label{sec causes zero field}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/Shema_summary_theory_2}
+\caption{Summary of the different depolarization effects in low magnetic field and their predicted depolarization rate. The ODMR spectra were taken on sample ADM-150-1}
+\label{summary_theory}
+\end{figure}
+
+We will focus on this section on the mechanisms responsible for the depolarization of dense NV ensemble at low magnetic fields.
+We have identified four such mechanisms that lead to an increase in the resonant dipole-dipole coupling between NV and fluctuators, which in turn results in an increased spin depolarization rate.
+
+Fig.
+\ref{summary_theory} summarizes these four mechanisms along with the theoretical increase in the depolarization rate caused by each of them.
+We will detail below the physical origin of these four mechanisms and the calculations that lead to the values presented in Fig.
+\ref{summary_theory}.
+
+We should note that the values indicated in Fig.
+\ref{summary_theory} for the double flips and the $T_2^*$ contributions depend on experimental parameters, while the ones for the 4-classes degeneracy and the eigenbasis change are purely based on theory.
+
+\subsection{Four classes degeneracy}
+
+The first and main contribution to the low field depolarization is the degeneracy between the four classes as the magnetic field decreases.
+Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 43$.
+This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.
+
+\subsection{Eigenstates modification}
+
+The modification of the Hamiltonian eigenstates discussed in sec.
+\ref{sec.
+Hamltonian 0 B} from the eigenbasis $\{ \ket{0}, \ket{+1},\ket{-1} \}$ when $B\neq 0$ to the eigenbasis $\{ \ket{0}, \ket{+},\ket{-} \}$ when $B\sim 0$ changes the dipole-dipole interaction between the two spins.
+
+A consequence of this change is the absence of diagonal terms  in the dipole-dipole Hamiltonian written in the $\{ \ket{0}, \ket{+},\ket{-} \} \times \{ \ket{0}, \ket{+},\ket{-} \}$ basis, which explains why the NV centers states are protected from magnetic noises at low magnetic field \citep{jamonneau2016competition}.
+
+Another consequence is a slight modification of the flip-flop rates.
+Once averaged over all possible orientation and position, as detailed in appendix \ref{Appendix eta}, this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
+
+\subsection{Double flips}
+
+The double flip processes have been previously discussed in sec.
+\ref{sec flip-flop et double flip} and naturally arise from the dipole-dipole Hamiltonian.
+The ones that we consider here couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$).
+These terms are generally not resonant because of the Zeeman splitting between the $\ket{+1}$ and $\ket{-1}$ states.
+
+In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$.
+These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely suppressed.
+
+The computation of the depolarization rate including the double flips is detailed in appendix \ref{Appendix eta}.
+To obtain a numerical value, we need to input the experimental values of the dipole-dipole interaction range \footnote{By interaction range we mean the value to be compared to the spectral detuning $\Delta \nu$ between the two groups.
+This is $2\gamma_f+\Gamma_{NV}+\Gamma_f$ in eq.
+(\ref{eq 1/T avec T2*})} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+\ref{fluct linewidth}, as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states.
+We typically measure $\Delta \nu \approx 8\ \rm MHz$ on type Ib HPHT samples.
+
+For the value presented in Fig. \ref{simus ESR}, we inputted $\Gamma=\Delta \nu=8.8\ \rm MHz$.
+
+\subsection{$T_2^*$ increase}
+
+The last aspect we considered was the change in the ODMR lineshape discussed in sec. \ref{sec modif T2*} for weak magnetic field (we refer to the spectral linewidth as $T_2^*$ for conciseness).
+
+To quantify the impact of the $T_2^*$ increase in the depolarization rate, we employ eq.
+(\ref{eq 1/T avec T2*}), where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
+\ref{ESR for T2*}.
+We then find that the increase of $T_2^*$ in zero magnetic field occasions an increase of the decay rate by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.
+
+The value shown in Fig. \ref{summary_theory} corresponds to a $9 \%$ increase.
+
+\bigskip
+
+In conclusion, we predict that the 4-classes degeneracy is by far the dominant causes for low field depolarization ($\sim \Gamma_1 \times 40$), followed by the double flips ($\sim \Gamma_1 \times 2$).
+We will however see in sec \ref{sec LFDM} that the double flips, along with the change of eigenbasis and increase in $T_2^*$ are determinant factor in the sensitivity of the low field magnetometry protocol.
+
+In the following section we will confront the theoretical results with experimental ones.
+
+\section{Experimental investigation of the low field depolarization}
+\label{sec 4.3}
+In this section we will try to experimentally discriminate the contribution of the four depolarization sources described in Fig.
+\ref{summary_theory} and confront the experimental values with the theoretical ones.
+
+We will first determine the contribution of the 4-classes degeneracy by comparing [100] aligned and non-[100] aligned magnetic field in the weak magnetic field regime, before studying the influence of the double-flips thanks to purely transverse magnetic fields.
+
+On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\mathbf{B}\parallel [100]$) and the influence of the laser polarization.
+Following further experiments detailed in appendix \ref{Appendix eta}, we have determined that these two effects were negligible compared to the four ones we identified before.
+
+\subsection{NV-NV CR in the low magnetic field regime}
+
+In this part, we will try to isolate the contribution of the 4-classes degeneracy from the three other ones by using the fact that all four classes can be at resonance for two distinct scenario:
+\begin{itemize}
+\item When $\mathbf{B}=0$.
+\item When $\mathbf{B}\parallel [100]$ (cf. Fig. \ref{ODMR_geometries}).
+\end{itemize}
+
+\subsubsection{Magnetic scan in a random direction}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 4/Figures/scan_1x1x1x1}
+\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction.
+a) ODMR spectrum for B=0.
+b) ODMR spectrum for B=60 G.
+c) PL contrast as a function of the external magnetic field.
+d) $1/T_1^{\rm dd}$ as a function of the external magnetic field.
+A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in Fig.
+\ref{various_fit_formulae} with $T_1^{\rm ph}=5\ \rm ms$}
+\label{scan 1x1x1x1}
+\end{figure}
+
+First we will run a control experiment by scanning the magnetic field in a non [100] direction.
+Fig.
+\ref{scan 1x1x1x1}-a) and b) shows ODMR spectrum at the starting point ($|\mathbf{B}|=0$) and end point ($|\mathbf{B}|\approx 60\ \rm G$) of the magnetic field scan.
+The magnetic field orientation was chosen in order to maximize the splitting of the four classes.
+
+Fig.
+\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter, as a function of the magnetic field.
+We can observe a decrease of $T_1^{\rm dd}$ by a factor of $\sim 10$ in zero magnetic field compared to non-zero magnetic field, with a half-width at half maximum of $\sim 10\ \rm G$.
+Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\mathbf{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$.
+This is because this drop of PL is associated with transverse magnetic field state mixing, and not with spin depolarization.}.
+
+We attribute the spin depolarization observed in Fig.
+\ref{scan 1x1x1x1}-d) to all four mechanisms described in the last part acting simultaneously.
+To determine the contribution of each mechanism individually, we will need to run dedicated experiments.
+
+\subsubsection{Magnetic scan with $\mathbf{B} \parallel [100]$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 4/Figures/scan_100}
+\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis.}
+\label{scan 100}
+\end{figure}
+
+This second experiment is similar to the last one except that the magnetic field was scanned in the [100] direction.
+Fig.
+\ref{scan 100}-a) and b) again show ODMR at the starting and ending point of the magnetic scan, while Fig.
+\ref{scan 100}-c) and d) show the PL and $1/T_1^{\rm dd}$ recorded during the scan.
+
+The main difference with the previous experiment is that in this case, the degeneracy between the four classes is never lifted.
+We therefore attribute the increase in $1/T_1^{\rm dd}$ at low magnetic field to the three other identified mechanisms.
+The increase in the depolarization rate in zero field ($\sim 20\ \%$) as well as the decrease inPL ($\sim 0.5 \%$) are still present, but significantly smaller than in the previous case.
+We can note however that the half-width at half maximum of the zero field depolarization ($\sim 2\ \rm G$) is also smaller in this case.
+
+We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in sec.
+\ref{sec NV-13c-NV}.
+We can not explain however why the depolarization rate appear to be smaller at $B \sim 10\ \rm G$ than it is for $B > 30\ \rm G$.
+
+The fact that the increase of the depolarization rate at zero field is about $\sim 50$ times weaker in Fig.
+\ref{scan 100} than it was in Fig.
+\ref{scan 1x1x1x1} confirms that the leading cause of depolarization in zero field is the degeneracy between the four classes of NV centers.
+
+\subsection{NV-NV CR under purely transverse magnetic field}
+
+Thank to the two previous experiments, we were able to discriminate the contribution of the 4-classes degeneracy from the three other mechanisms that are which are the eigenbasis change, the double-flips and the $T_2^* increase$.
+We now seek to discriminate the contribution of each of these three mechanisms.
+
+We unfortunately were not able to discriminate the change of eigenbasis from the $T_2^* increase$, which would likely require a statistical study over different samples with different $T_2^*$ values.
+We were however able to discriminate the double flip contribution from the two other ones by scanning the magnetic field in a direction orthogonal to one NV class ($\mathbf{B} \perp [111]$).
+
+\subsubsection{Principle of the experiment}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/transis_transverse_field}
+\caption{Simulation of the eigenstates of the NV Hamiltonian for  a purely transverse magnetic field $B_\perp$.
+In addition to the transverse magnetic field, we consider a fixed transverse electric field of value $d_\perp E_\perp=4\ \rm MHz$.
+(a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$.
+The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$.
+(b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
+\label{eigenstates transverse field}
+\end{figure}
+
+The idea behind the experiment is that the Hamiltonian eigenbasis is the same (or close to be the same) when $\mathbf{B}=0$ than it is when $\mathbf{B} \perp [111]$, as described in sec.
+\ref{sec B transverse}.
+For simplicity, we will assume here that both the magnetic field and the transverse electric field lies along the NV $x$ axis : $\mathbf{B}=B_x \hat{e}_x$ and $\mathbf{E}_\perp =E_x \hat{e}_x$
+
+Fig.
+\ref{eigenstates transverse field} shows a simulation of the eigenstates and eigenenergies for a NV Hamiltonian subject to transverse electric field and magnetic field.
+We labeled the eigenstates of the Hamiltonian $\ket{1}$, $\ket{2}$ and $\ket{3}$ in ascending order of energy.
+
+Fig.
+\ref{eigenstates transverse field}-b) shows both the splitting $\Delta \nu$ between the $\ket{2}$ and $\ket{3}$ states, and how close the $\ket{3}$ state is to the $\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}}$ state via the factor $\abs{\bra{3}\ket{+}}^2$.
+We should note that with the assumption that $B_\perp \parallel E_\perp$, the $\ket{2}$ state is strictly equal to the $\ket{-}$ state regardless of the magnetic field amplitude.
+
+These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are the relevant parameters of the experiment: the goal is to increase $\Delta \nu$ to the point where the double flips are completely suppressed, while $\abs{\bra{3}\ket{+}}^2$ remains close to 1 in order to preserve the eigenbasis and $T_2^*$ contribution.
+Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100 G,200 G] seems to satisfy both these criteria: $\Delta \nu\,(100\ \rm G) \approx 30\ \rm MHz \gg \gamma_f$\footnote{Technically, $\Delta \nu$ is to be compared with the dipole-dipole interaction width $2\gamma_f+\Gamma_{NV}+\Gamma_f$ which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+\ref{fluct linewidth}.} and $\abs{\bra{3}\ket{+}}^2\, (200\ \rm G) \approx 0.96$.
+
+Ideally, we would have used an electric field instead of a transverse magnetic field to split the 4 classes energetically, but such an electric field would need to be $> 10^6\ \rm{V/cm}$ which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.
+
+\subsubsection{Results}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{chapter 4/Figures/scan_T1_perp}
+\caption{Experimental data for the depolarization under purely transverse magnetic field on sample ADM-150-2.
+a) ODMR spectrum for $|\mathbf{B}|=20\ \rm G$.
+The transitions of the class orthogonal to the magnetic field are labeled $\nu_+$ and $\nu_-$.
+b) Dependency of $\nu_+$ and $\nu_-$ with the magnetic field, as measured through ODMR.
+c) Measurement of $T_1^{\rm dd}$ as a function of the transverse magnetic field.
+The corresponding energy splittings $\Delta \nu=\nu_+-\nu_-$ are written on top.
+We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$.
+We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$}
+\label{champ tranverse exp}
+\end{figure}
+
+Fig.
+\ref{champ tranverse exp}-a) shows an ODMR spectrum for $B=20\ \rm G$ with the two central lines corresponding to the $\ket{0} \to \ket{+}$ and $\ket{0} \to \ket{-}$ transitions of the class orthogonal to the magnetic field.
+We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\mathbf{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
+\ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\mathbf{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}
+
+Fig.
+\ref{champ tranverse exp}-b) shows the evolution of the $\nu_+$ and $\nu_-$ transition frequencies as a function of the magnetic field, measured via ODMR.
+The difference in frequency $\Delta \nu = \nu_+ - \nu_-$ is at first determined by the transverse electric field ($\sim 9\ \rm MHz$) and then grows qudratically with the transverse magnetic field up to a value of $\sim 50\ \rm MHz$.
+
+Fig.
+\ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ on a single class as a purely transverse magnetic field is scanned from 20 to 130 G.
+It also includes the $1/T_1^{\rm dd}$ value found for a purely longitudinal magnetic field which we will consider as a baseline, since the eigenbasis in this case is $\ket{\pm 1}$ and not $\ket{\pm}$.
+We have divided the graph in two regions labeled A and B:
+
+In region A, $1/T_1^{\rm dd}$ decreases with the magnetic field.
+The difference in the depolarizing rate between the $\mathbf{B} \perp [111]$ case and the $\mathbf{B} \parallel [111]$ case comes from the double flips, the eigenbasis change and the $T_2^*$ increase.
+We measure an increase by a factor $\sim 3$ for B=20 G compared to the baseline.
+
+In region B, $1/T_1^{\rm dd}$ reaches a plateau with a value of $\sim 180\ \rm s^{-1}$.
+We attribute this plateau to the fact that the double flips are completely quenched, which is coherent with the fact that $\Delta \nu \geq 30\ \rm MHz$.
+The remaining difference between the $\mathbf{B} \perp [111]$ and $\mathbf{B} \parallel [111]$ cases therefore comes from the eigenbasis change and the $T_2^*$ increase.
+This increase amount for $\sim$ 50\% of the relaxation rate baseline.
+
+We should note that the theoretical increase from the double flips and the change of eigenbasis\footnote{The $T_2^*$ increase should still amount for $\sim 10\%$.} are not the values shown in Fig.
+\ref{summary_theory}: those were computed for a 4-class resonance, while we are looking here at a single class.
+The theoretical increase of the depolarization rate for a single class are detailed in Appendix \ref{Appendix eta}.
+They are:
+\begin{itemize}
+\item $\Gamma_1^{\ket{\pm}}=4 \Gamma_0$ when taking into account the change of eigenbasis.
+\item $\Gamma_1^{dd}=1.6\Gamma_1^{\ket{\pm}}=6.5\Gamma_0$ when taking into account the double flips.
+\end{itemize}
+We have labeled here $\Gamma_0$ the relaxation rate coming from the flip-flop inside of a single class in the $\ket{\pm 1}$ basis (same definition as in table \ref{T1 champ mag} (table des valeurs chap 3).
+Experimentally, $\Gamma_0$ corresponds to the baseline shown in Fig.
+\ref{champ tranverse exp}-c).
+
+Although we find again that the fluctuator model overestimates the actual relaxation rates, these experimental observations still confirm the qualitative predictions of the model.
+In particular, they confirm that the double-flips are the second leading cause of zero-field depolarization after the 4-classes degeneracy.
+
+\subsection{Conclusion of the experimental observation}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/Shema_summary_exp}
+\caption{Summary of the different experimental investigation.
+A baseline value is indicated with the pink shaded region in all experimental plots.
+The left box indicates the relaxation mechanisms added compared to the baseline for the minimum value of B, and the right box those for the maximum value of B.
+(a) corresponds to Fig.
+\ref{scan 1x1x1x1}-d) where the baseline was measured to be $220\pm 20$ s$^{-1}$, (b) corresponds to Fig.
+\ref{scan 100}-d) with the same baseline, and (c) corresponds to Fig.
+\ref{champ tranverse exp}-c).}
+\label{summary_exp}
+\end{figure}
+
+Fig.
+\ref{summary_exp} summarizes the experimental results of this section.
+We explicitly indicates which of the four depolarization mechanisms are responsible for the observed depolarization rate at the beginning and at the end of each curve.
+
+The experiments confirm the overall predictions of the previous section: the dominant cause of depolarization in zero magnetic field is the degeneracy of the 4 NV classes, followed by the double-flips and then by the eigenbasis change and $T_2^*$ increase.
+
+We will now cover the low-field magnetometry protocol, after a quick overview of the field of NV magnetometry, and see the role that the different depolarization mechanisms play in it.
+
+\section{Introduction to ensemble NV magnetometry}
+\label{sec 4.4}
+
+There are two classes of NV magnetometer: single NV or ensemble.
+Single NVs offer a spatial resolution limited only by how far the magnetic source can be from the NV center (down to $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
+
+Ensemble NV centers center on the other hand offer sensitivities as low as $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of sizes up to several mm.
+
+We will only focus in this section on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR.
+We will not try to give a complete overview of this field, but only of the elements close to the technique we present.
+For a general overview on NV magnetometry, we recommend the excellent reviews  \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.
+
+\subsection{AC and DC magnetometry}
+The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted into spin population.
+These protocols are generally separated between DC-broadband and AC-narrow band.
+
+For DC-broadband magnetometry, the general technique consists of a Ramsey interferometry experiment, presented in chapter 1, the sensitivity of which is limited by the spin inhomogeneous coherence time $T_2^*$ (which can reach up to $\ \rm{\mu s}$ for NV ensembles).
+The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$.
+Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry.
+The state of the art DC ensemble magnetometr operates with a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
+
+For AC magnetometry, the basic protocol consists of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
+Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field.
+The AC magnetic field however has to be in phase with the rephasing pulses, which limits the AC sensing to a relatively narrow band.
+The state of the art in term of AC sensing is $\sim 1\ \rm pT/\sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}.
+
+Recently, \citep{xie2021hybrid} have achieved new records in both DC ($\sim 200\ \rm fT/\sqrt{\rm Hz}$) and AC ($\sim 10\ \rm fT/\sqrt{\rm Hz}$) NV magnetometry by using flux concentrators which are ferromagnets concentrating the magnetic field in a small volume, similarly to an electromagnet.
+The technology of flux concentrator could be applied to any NV magnetometry protocol.
+We will here only compare experimental results obtained without a flux concentrator.
+
+\subsection{Low field magnetometry}
+Magnetometry at low magnetic fields ($\gamma B < d_\perp E_\perp$) is challenging since in this regime the NV eigenstates and eigenenergies do not depend on the magnetic field to the first order.
+This complication is often lifted  by simply using a bias magnetic field.
+
+Some systems however need to be kept under low to zero external magnetic fields, for example to study the J-coupling between spins \citep{sutter2012computational}, or the phase of skyrmions \citep{zazvorka2020skyrmion}.
+To study these systems with NV centers, new protocols for low field NV magnetometry are required.
+
+The main technique for low field magnetometry consists in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}.
+These states are linearly dependent on the magnetic field and can therefore be used to perform magnetometry.
+The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$ \citep{zheng2019zero}.
+
+Another technique consists in using the $^{13}\rm{C}-$NV complex we discussed in sec.
+\ref{sec NV-13c-NV} which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are first order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
+With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers.
+
+\subsection{Microwave-free DC magnetometry}
+All the previously described magnetemetry protocols relied on the use of a controllable microwave field to coherently manipulate the NV spin state.
+
+%There are other protocols which correlate the magnetic field to the spin $T_1$ time (relaxometry).
+These protocols can operate without microwave field and work for either DC or high frequency AC sensing.
+
+%\subsubsection{microwave-less DC magnetometry}
+The DC microwave-free protocols rely on sharp features in the PL for specific magnetic fields values.
+Fig.
+\ref{CR P1 exp} in chapter 2 show such PL features, in particular the ground state level anti-crossing (GSLAC, simply labeled LAC on the figure), the sharpest of these features.
+
+The GSLAC occurs for a purely longitudinal magnetic field where $\gamma B_z = D$, which corresponds to $B=1024\ \rm G$.
+In these conditions, the normally bright $\ket{0}$ state and the dark $\ket{-1}$ state become resonant and any residual mixing between these two states (coming from the electric field, the strain or the residual transverse magnetic field) will cause a drop in the NV polarization, and in the PL \citep{broadway2016anticrossing}.
+
+A protocol based on the GSLAC line has been implemented, first only for longitudinal magnetic field measurements \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.
+
+Another DC microwave-less protocol was proposed based on NV-NV CR \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}.
+This technique reconstructs the external magnetic field by scanning an other magnetic field, similar to Fig.
+\ref{equivalent NV-NV} in chapter 3, and by measuring the position of the NV-NV CR lines.
+The 3D external magnetic field can then be recovered by scanning the magnetic field in different directions.
+The author of this paper did not measure the sensitivity of this protocol, but they predict a shot noise limited sensitivity of $24.7\ \rm nT/\sqrt{\rm Hz}$.
+
+%\subsubsection{microwave-less AC magnetometry}
+%The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed.
+The detection bandwidth is adjusted with an external DC magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.
+%
+%Sensitivities as low as a few $\rm pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, but both protocols are not microwave-free as they rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).
+
+\subsection{Orientation-free magnetometry}
+All protocols described so far require precise knowledge of the Larmor frequency between the $\ket{0}$ and $\ket{\pm 1}$ states for a given magnetic field.
+This knowledge implies that the direction of each NV axis in the frame of the magnetic field is known, because of the strong anisotropy in the NV Hamiltonian.
+
+For a single crystal, these axes can always be calibrated through ODMR.
+However, for a polycrystalline diamond, diamond powder or even a single diamond with an erratic motion, the knowledge of these axes is not possible.
+New protocols need to be found if one wants to perform magnetometry with these materials.
+
+To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
+This protocol has a very low sensitivity but is still sometimes used in the absence of alternative \citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective, zhu2022sunlight}.
+
+Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies.
+The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the Larmor frequency), or via the PL \citep{finco2021imaging}.
+
+\section{Low field depolarization magnetometry}
+\label{sec LFDM}
+We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the depolarization of dense NV ensemble at low magnetic field.
+This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional AC field is required to improve performances.
+
+We will first describe its operation and characterize it before comparing it to the previously mentioned NV magnetometry protocols.
+
+\subsection{Principle of the experiment}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/setup_magneto}
+\caption{Principle of the LFDM protocol a) Experimental setup.
+The confocal microscopy setup is similar to the one described in sec \ref{Sec setup cahp 1}, the lock-in amplifier (LIA) is used here to add a modulation on the magnetic field, instead of the microwave.
+Both the PL and the LIA signal are monitored.
+b) example of LIA signal on sample ADM-15-4 when scanning the magnetic field $B_0$ between -150 and 150 G.
+c) PL signal corresponding to the LIA signal}
+\label{setup magneto}
+\end{figure}
+
+Fig.
+\ref{setup magneto} shows the principle of the LFDM protocol.
+The experimental setup is similar to the one presented in chapter 1 without any microwave field.
+The magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\tilde{B}_{\rm AC}$ used for the modulation.
+We typically use a modulation frequency $f_{\rm mod} \approx 1\ \rm kHz$, the same value we use for ODMR measurements.
+The main voltage $V_0$ responsible for the DC field and the oscillating voltage $\tilde{V}_{\rm AC}$ responsible for the AC field are summed via a homemade bias tee.
+
+Fig.
+\ref{setup magneto}-b) and c) show the PL measured on the photodiode and on the lock-in respectively, when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G.
+For this configuration, the magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.
+
+Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup.
+To improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing.
+Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.9\textwidth]{chapter 4/Figures/AC_and_DC_sensing}
+\caption{DC and AC LFDM sensing protocols on sample ADM-15-5.
+The magnetic field is scanned in an arbitrary direction.
+a) DC sensing: $B_0$ is scanned between $-$80 and +80 G while an AC field of amplitude 10 G is used for the lock-in detection.
+b) AC sensing: $B_0$ scans are performed for 5 different $|\tilde B_{\rm AC}|$ values between 2 and 10 G.
+The best AC sensitivity is achieved for $B_0\approx 5\ \rm G$.}
+\label{AC and DC sensing}
+\end{figure}
+
+Fig.
+\ref{AC and DC sensing} shows how the AC and DC sensing work for LFDM: for DC sensing the LIA signal around $B_0=0$ provide a sharp linear slope to convert the LIA voltage into a magnetic field.
+For AC sensing, the LIA signal is directly proportional to the amplitude of the field.
+The optimal AC field amplitude for DC sensing has to match the width of the PL dip $|\tilde{B}_{\rm AC}| = \sigma \approx 10\ \rm G$ in order to maximize the slope of the lock-in signal around $B_0=0$.
+the optimal DC field to use for AC sensing is $B_0= \sigma/2$ corresponding to the region where the lock-in signal is most sensitive to AC field.
+
+\subsection{Sensitivity of the LFDM protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/sensi_alternee}
+\caption{Sensitivity measurement of the LFDM on sample ADM-15-4.
+a) Temporal trace of the LIA signal when applying an alternating external magnetic field of amplitude 35 $\mu$T and frequency $f=1\ \rm Hz$.
+b) Histogram of the DC measurement over 50 s.
+The data is fitted with Gaussians of width $1.6\ \rm{\mu T}$.}
+\label{sensei alternee}
+\end{figure}
+
+We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols.
+The characterization was only done for DC sensing.
+The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see sample details in appendix \ref{Appendix samples}).
+
+We characterize the DC sensitivity by looking at the noise floor on the LIA signal when sitting on the steepest sloped ($B_0\approx 0$).
+The following parameters were used: $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$.
+The lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.
+
+Fig.
+\ref{sensei alternee} shows the sensitivity measurement protocol.
+First, the applied magnetic field is calibrated with ODMR spectra, then an alternating DC field of known amplitude (here $\Delta B_0=35\ \rm{\mu T}$) is applied on the sample.
+Fig.
+\ref{sensei alternee}-a) shows the temporal trace of the LIA signal in response to this alternating DC field.
+Knowing the amplitude of the alternating field, we can then convert the voltage signal into a magnetic field amplitude, with in this case a ratio $45 \pm 3\ \rm{\mu T /V}$.
+
+Fig.
+\ref{sensei alternee}-b) shows an histogram of the measured magnetic field amplitude for both the high and low value of the applied magnetic field.
+The total acquisition time was 50 s with the external field being alternated every second.
+The histograms are nicely fitted with Gaussian profiles of width $\sigma=1.5\pm 0.1\ \rm{\mu T}$ which allows us to estimate the DC sensitivity of the protocol:
+\begin{equation}
+\eta = \frac{\sigma}{\sqrt{2 \tau}}= 116 \pm 10\ \rm{nT} / \sqrt{\rm Hz},
+\end{equation}
+
+where $\tau=3\ \rm ms$ is the LIA low-pass filter time constant.
+
+We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\sqrt{\langle \delta V^2 \rangle} \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\sqrt{\langle \delta V^2 \rangle} < 1\ \rm mV$).
+The remaining noise could also come from laser intensity fluctuations.
+
+\subsection{Angular dependence of the sensitivity}
+\label{sec angular sensi}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/sensi_angulaire}
+\caption{Angular dependence of the sensitivity.
+a) LIA signal for $B_0$ scanned between $-$150 and +150 G along the [100] axis, or along a direction 20$^\circ$ off the [100] axis.
+b) Zoom-in of the LIA signal on the zero-field slopes.
+c) Measured sensitivity as a function of the magnetic field angle with the [100] axis.}
+\label{angular sensi}
+\end{figure}
+
+We will now study the dependence of the sensitivity with respect to the magnetic field angle.
+In particular, we will compare the sensitivity for two scenarios: the case $\mathbf{B}\parallel [100]$and the four classes remains degenerate when $\mathbf{B}$ increases, and the case $\mathbf{B}\nparallel [100]$ where there is a lift of degeneracy between the four classes.
+We previously saw in Fig.
+\ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\mathbf{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.
+
+Fig.
+\ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angles of $B_0$.
+For the blue curve the field was scanned along the [100] axis, while for the orange curve it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis.
+We can see that the maximum amplitude of the lock-in signal (relevant for AC sensing) is of the same order of magnitude for both case, and the slope around $B_0$ (relevant for DC sensing) is almost identical.
+Fig.
+\ref{angular sensi}-c) shows the sensitivity, measured with the protocol described in Fig.
+\ref{sensei alternee}, with respect to the angle between the magnetic field and the [100] axis.
+We can notice that the sensitivity depends only weakly on the magnetic field angle.
+
+The fact that the sensitivity is similar when $\mathbf{B}\parallel [100]$ or $\mathbf{B}\nparallel [100]$ means that the degeneracy between the four classes plays a lower role in the sensitivity than the three other mechanisms detailed in sec.
+\ref{sec causes zero field}.
+We cannot measure the individual impact of each of these three mechanisms, but based on the experimental results shown in Fig.
+\ref{champ tranverse exp}, we can assume that the double-flips are the dominant factor in the LFDM sensitivity.
+
+It should be noted that this low angular dependence was particularly marked on sample ADM-15-4.
+Other samples, including from the same batch, showed a higher angular dependence with typically a sensitivity $2 \sim 3$ times worse for $B\parallel [100]$ than for the other directions.
+
+\subsection{Temporal stability}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/Temporal_just_0B}
+\caption{Temporal stability of DC-LFDM as measured on sample ADM-15-5.
+a) Temporal trace of the LIA signal for $B_0 \approx 0$.
+The Lock-in signal was converted to a measured magnetic field with the previously described protocol.
+b) Estimated signal to noise ratio for a signal of 100 nT as a function of the total acquisition time $T$.
+The noise value is computed by averaging the variance of the temporal trace over each time duration $T$.
+A curve $y=0.5\, T^{1/2}$ is added for comparison}
+\label{temporal 0B}
+\end{figure}
+
+An important aspect of sensors is their stability in time.
+The sensitivity being expressed in $\rm{nT} / \sqrt{\rm Hz}$ suggests that the signal to noise ratio (SNR) given by repeating the measurement increases as $\sqrt{T}$ where $T$ is the total acquisition time.
+This is true for an ideal system, but in practice drifts and low frequency noise mean that the actual SNR for long acquisition time will be worse than the SNR at short time scaled by a factor $\sqrt{T}$.
+
+One of the main source of drift for NV magnetometry based on spin resonance is the crystal lattice temperature change: the $D$ factor in the NV Hamiltonian (eq.
+(\ref{eq.
+spin elec})) is sensitive to the lattice temperature because of the thermal dilatation of the crystal, which is the basis for NV thermometry \citep{kucsko2013nanometre}.
+This means that thermal fluctuations will impact the Larmor frequency of the spins and may cause a detuning with the initial measured frequency for long acquisition times.
+The LFDM protocol however does not rely on a specific resonance frequency, but rather on a co-resonnance condition between neighboring spins, which does not depend directly on the crystal temperature.
+
+Fig.
+\ref{temporal 0B} shows the temporal stability of the LFDM protocol over an acquisition time of 100 s.
+Fig.
+\ref{temporal 0B}-a) shows a temporal trace of the LIA signal where we have converted the signal into a measured magnetic field.
+Fig.
+\ref{temporal 0B}-b) shows the estimated SNR for a signal of 100 nT as a function of the total acquisition time and was computed from the temporal trace.
+
+We can see that the SNR does not deviate from the $\sqrt{T}$ scaling for times up to 50 s, and could probably be extended for longer times.
+The non-linearity for shorter times comes from the LIA low-pass filter.
+
+\section{Comparison between LFDM and other magnetometry protocol}
+\label{sec 4.6}
+Now that we have characterized the LFDM protocol, we can compare its performances with the previously mentioned magnetometry protocols.
+The best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity.
+This discrepancy however come in large part to the smaller sensor volume used in the characterization, and to a relatively low PL collection efficiency compared to the other experiments.
+
+\subsection{Comparison with microwave based protocols}
+
+\begin{table}[htbp]
+\centering
+\caption{Comparison of the volume-normalized sensitivity of the best AC and DC protocols with LFDM}
+ \label{table sensi volumiques}
+\begin{tabular}{c|ccc}
+\toprule
+{} & Zhou et al. \citep{zhou2020quantum} (AC) & Barry et al.\citep{barry2016optical} (DC) & LFDM (DC) \\
+\midrule
+$\eta\ (\rm{nT} / \sqrt{\rm Hz})$ & 92&0.015& 116 \\
+$V\ (\mu \rm{m}^3)$ &$8.1 \cdot 10^{-3}$ &$5.2 \cdot 10^{6}$& $3.3 \cdot 10^{3}$ \\
+$\eta_v\ (\rm{nT}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ &8.3&34&6700 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+A better metric than the sensitivity to compare different protocols is the volume-normalized sensitivity defined in \citep{zhou2020quantum} as $\eta_v=\eta\cdot \sqrt{V}$ where $\eta$ is the sensitivity and $V$ the effective volume of the sensor.
+This scaling assumes that each NV center in an ensemble is an independent magnetic field probe, and as the total number of NV centers is proportional to $V$, the sensitivity should scale as $1/\sqrt{V}$.
+
+Table \ref{table sensi volumiques} reports the best AC and DC volume-normalized sensitivities found in the literature.
+The LFDM protocol performs $\sim 200$ times worse than the best DC magnetometer.
+It should be noted however that the PL collection in our measurement, where we use a confocal microscope with a numerical aperture NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/compariason_0B_MW}
+\caption{Comparison between LFDM and microwave-based magnetometry on the same experimental setup.
+The sample used is ADM-15-5 and the magnetic field is aligned close to a [111] axis.
+a) Schematics of the experiment: the setup is the same as the one described in Fig.
+\ref{setup magneto} with the addition of a continuous microwave tone at the frequency $\nu=2600\ \rm MHz$.
+b) LIA signal as $B_0$ is scanned from $-$150 to +150 G.
+The two lines at $\pm 100\ \rm G$ are caused by the microwave.
+c) Temporal trace of the measured B field for $B_0 \approx 0$ (on the middle of the zero field slope) and $B_0 \approx 100\ \rm G$ (on the middle of the microwave slope).
+The orange curve is shifted by $\approx 100\ \rm G$ to sit on the same scale as the blue curve.
+d) Estimated SNR fo a signal of 100 nT for $B_0 \approx 0$ and $B_0 \approx 100\ \rm G$ as a function of the total acquisition time}
+\label{comparaison 0B MW}
+\end{figure}
+
+Fig.
+\ref{comparaison 0B MW} shows a comparison between LFDM and microwave-based magnetometry using our setup.
+A fixed microwave tone (here at 2600 MHz) is sent to the diamond and a magnetic field $B_0$ close to the [111] axis is scanned along an oscillating field $\delta B$ to perform a lock-in detection.
+
+Fig.
+\ref{comparaison 0B MW}-b) shows the LIA recorded during a $B_0$ scan: the profile is similar to the one in Fig.
+\ref{setup magneto}-b) with the addition to two sharp lines for $|B_0| \approx 100\ \rm G$.
+These two lines correspond to the resonance condition between the microwave tone and the $\ket{0} \to \ket{\pm 1}$ transition of the class parallel to the magnetic field\footnote{We would need to increase the amplitude of the magnetic field in order to see the three other classes.}.
+The microwave power used here correspond to a Rabi frequency $\Omega \approx (2\pi) 3\ \rm MHz$ comparable with $1/T_2^*$ and the slope on the steepest part of the microwave line is $\sim 2.5$ times higher than the slope in zero field.
+
+Fig.
+\ref{comparaison 0B MW}-c) and d) show the temporal stability of the sensor similarly to Fig.
+\ref{temporal 0B}, on both the zero field line and the microwave line.
+While we can see that the microwave signal is more sensitive thanks to the higher slope, it is also more prone to drift, most likely because of the shift of the Larmor frequency caused by temperature fluctuations.
+As a consequence, we can see that the LFDM becomes more sensitive than the microwave-based measurement for acquisition times beyond 10 s.
+
+We should stress that this microwave-based detection is not optimal: the technique used here is analogous to CW ODMR which is not the most sensitive DC magnetometer \citep{barry2020sensitivity}.
+In particular techniques based on double-quantum transition can be both more sensitive and less affected by thermal drifts.
+The sample used here is also poorly suited for traditional microwave magnetometry due to its low $T_2^*$ value.
+
+\subsection{Comparison with microwave-less protocols}
+The most studied NV-based microwave-less protocols are the GSLAC protocols that we described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
+Unfortunately the sensing volume in this studies is not known, meaning that we can not directly compare the volume normalized sensitivity of both approaches.
+Both GSLAC protocols and LFDM however work under the same basis, which is to exploit the dependency of the NV PL with the magnetic field.
+In order to compare both approach, we can simply focus on the characteristics (width and depth) of the GSLAC PL feature and the zero-field PL feature.
+
+In Fig.
+\ref{setup magneto}-c) we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
+In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type Ib irradiated diamond, similar to the ones we use.
+In \citep{zheng2020microwave} they manage to keep a contrast of 4.5 \% while reducing the GSLAC linewidth to $0.38\ \rm G$ by using a CVD sample isotopically enriched in $^{12}$C with a lower [NV] and [N] density.
+
+The zero field technique is therefore comparable to the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear what kind of material engineering would improve its sensitivity.
+
+%Je parle pas de la dépendence en puissance ou du champ transverse, je laisse ça pour plus tard.
+\section{Conclusion and perspectives}
+We have explored in this chapter the reasons behind the depolarization of dense NV ensemble at low magnetic field.
+We have identified four mechanisms by which the depolarization is increased in zero-field: the degeneracy of the classes, the double flips, the change in the NV eigenstates and the modification of the spins $T_2^*$.We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
+
+We have also shown that this low-field depolarization can be applied to perform magnetometry experiments.
+We have characterized the sensitivity, temporal stability, and angular dependence of such a magnetometer, and we have compared its performances to the established NV magnetometry protocols.
+
+We will now describe what could be interesting investigation points to further improve the LFDM performance, as well as potential applications.
+
+\subsection{LFDM optimization}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/double_adamas}
+\caption{a) PL of two distinct samples ADM-15-1 and ADM-15-4 as the magnetic field is scanned in an arbitrary direction (which splits the 4 classes similarly in both cases).
+b) Same measurement with $B\parallel [100]$ c) $T_1$ measurement on both samples on a single class for $B\approx 50\ \rm G$.
+d) $T_1$ measurement on both samples on all four classes for $B=0$.}
+\label{double dragon}
+\end{figure}
+
+One of the main path for LFDM optimization is be the material optimization of the samples.
+
+For traditional DC microwave-based magnetometry, the material optimization path is relatively straightforward: if we assume every NV centers to be independent the sensitivity of a magnetometer for a given sample volume V scales as $1/\sqrt{\rho}$.
+Similarly, according to theory \citep{barry2020sensitivity}, the sensitivity of DC microwave based magnetometers scale as $1/\sqrt{T_2^*}$.
+In order to improve the performance of sample, one should aim to increase the figure of merit $\rho T_2*$.
+
+The same figure of merit cannot be used to characterize the LFDM: the sensitivity of the protocol does not scale as $1/\sqrt{T_2^*}$ due to the spectral broadening of $\gamma_f$, and the density of fluctuators not only impacts the number of sensor per unit volume, but also the density and properties of fluctuators, as well as the average flip-flop rate between NV and fluctuators.
+Further studies with higher NV concentrations than those used in this manuscript will be required to find the correct material optimization path.
+
+We should note that we observed large discrepancy in the zero field depolarization among the samples that we studied, even from samples coming from the same batch.
+
+Fig.
+\ref{double dragon}-a) shows the PL on two distinct 15-$\mu$m Adamas microdiamonds for a magnetic field scanned in a random direction, and Fig.
+\ref{double dragon}-b) for a magnetic field scanned along the [100].
+We can notice that the PL contrast in zero field is twice as big for sample ADM-15-4 when $B \nparallel [100]$, and 5 time as big when $B \parallel [100]$, which would directly impact their respective sensitivity in LFDM.
+
+Fig.
+\ref{double dragon}-c) and d) shows $T_1$ measurement on both samples either on a single class for $B\neq 0$ or for all four classes for $B=0$.
+Interestingly, we can notice that the $T_1$ of sample ADM-15-1 is both shorter than the one of ADM-15-4 for $B\neq 0$, and longer for $B=0$.
+This could indicate that another relaxation mechanism is competing with the NV-fluctuator one on sample ADM-15-1, which would limit the LFDM sensitivity on this sample.
+
+Other non-material parameters in the LFDM protocols should also be investigated.
+This includes the temperature of the sample, the intensity and wavelength of the laser and the frequency and amplitude of the AC magnetic field.
+
+\subsection{Potential applications for LFDM}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 4/Figures/Applications}
+\caption{Schematics of two potential application for LFDM. The two application are detailed in the main text}
+\label{applications}
+\end{figure}
+
+We will now turn to potential applications for LFDM, which could be useful even without material optimization.
+
+Fig.
+\ref{applications} illustrates two potential applications for LFDM.
+The first one consists in the wide-field magnetic imaging on a surface with micron-sized irregularities.
+The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}.
+While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biologic samples \citep{scholten2021widefield}.
+
+Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always keep a sensor within $\sim$ 1 $\mu$m of the sample.
+Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.
+One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$m red fluorescent diamond) are well suited for this task.
+We have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples.
+This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}.
+Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required.
+Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field homogeneity.
+
+Another potential application for LFDM is shown in Fig.
+\ref{applications}-b.
+It consists in the detection of magnetic field gradient thanks to a nano-diamond probe attached to a nano-resonator.
+The resonator motion transforms the DC gradient in an AC field which can then be probed following the protocol described in Fig.
+\ref{AC and DC sensing}-b).
+This idea is based on previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}.
+A very similar protocol was recently developed with microwave based AC sensing \citep{huxter2022scanning}.
+
+In order to use AC LFDM with a scanning probe setup, one would need to have a nano-diamond with at least one NV center and one fluctuator.
+This may not be such a rare event: a (20 nm)$^{3}$ nanodimaond with [NV]=5 ppm contains on average $\approx 4$ NV centers.
+If we take a fluctuator proportion of 30 \% as measured in \citep{choi2017depolarization}, then finding a nanodiamond with at least one NV and one fluctuator  seems feasible.
+We have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect, and we therefore cannot estimate the sensitivity of such a probe.
+
+LFDM could also find applications in biosensing.
+The lack of microwave is a significant advantage in this field to prevent heating of the living organisms, and nano and micro-diamonds are already readily used as trackers for fluorescence microscopy \citep{mohan2010vivo}.
+The use of LFDM could allow to perform background free fluorescence imaging (with a higher sensitivity than what was done with transverse magnetic field \citep{chapman2013background}).
+It could also be used to perform real time magnetometry with diamonds floating in a liquid environment, as was done in \citep{feng2021association}, but without having to constantly calibrate the diamond axes through ODMR which significantly hinders the temporal resolution of microwave-based technique.
+
+Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations ,does not require a homogeneous magnetic field or microwave field along the sample, and can be used with polycrystalline samples which are easier to produce in large size.
+A recent paper demonstrates an energy-efficient and cost-effective NV magnetometer which uses sunlight as the the optical excitation source, and exploits the transverse field dependence of the NV PL to measure the magnetic field\citep{zhu2022sunlight}.
+The use of LFDM instead of transverse field could improve the sensitivity of this magnetometer by an order of magnitude.
+
+\chapter*{Conclusion}
+
+The work presented in this manuscript is part of an ongoing effort to develop quantum technologies, and NV centers, among many other systems, are a part of this new ecosystem.
+While the properties of single NV centers have been thoroughly studied in the past 20 years, we showed in this thesis that emerging properties of NV ensemble can still be discovered.
+These ensemble properties are relevant for the many fields using NV ensembles \citep{ barry2020sensitivity, eills2022spin, fu2007characterization, perdriat2021spin}.
+They also are relevant to explore fundamental aspects \citep{choi2017observation, kucsko2018critical, angerer2018superradiant} and for new potential applications \citep{akhmedzhanov2017microwave, pellet2021optical, pellet2022spin}.
+
+At the heart of this manuscript is the notion of cross-relaxation with NV centers.
+Cross-relaxations are a powerful spectroscopic tool, and NV centers are naturally suited to cross-relaxation studies due to their ability to be optically polarized and read-out, in opposition to their environment.
+
+We have showed how cross-relaxations can be used in a CVD-grown diamond to detect trace amount of other spin impurities, with higher sensitivity than the standard EPR apparatus, and only a fraction of the cost.
+These observations allowed us to study the temperature dependence of these impurities, and could eventually allow the detection and coherent manipulation of single dark spins through their interaction with an NV bath.
+
+We have then studied the cross-relaxation between NV centers, thanks to the multiple classes of NV and the ability to tune them in and out of resonance.
+The observation of cross-relaxation between seemingly equivalent spins has led previous researchers to postulate the existence of dark NV centers called fluctuators.
+We investigated experimentally the predictions of the fluctuator model, in particular the prediction of a stretched exponential profile for the depolarization of the spins, and a broadening of the fluctuators spectra compared to those of regular NV centers.
+Our observations for the large part comfort the fluctuator model, although we did notice some deviations which could indicate some limitations of the model.
+
+Dipole-dipole interaction within dense ensemble of NV centers ([NV] > ppm) can be responsible for an increase of the depolarization rate by more than a factor of 10.
+This changes considerably the properties of NV ensemble compared to single NV.
+Having a better understanding of the fluctuators and of their microscopic origin will be a crucial point in the future development of NV ensemble technologies.
+The further steps to determine the nature of the fluctuators could include the study of dedicated samples grown with specific concentrations of NV and other defects, as well as the study of nano-diamonds to try to isolate a single fluctuator.
+
+Finally, we investigated the depolarization of NV ensemble at low magnetic field.
+We identified 4 different contributions to the zero field depolarization which we analyzed both theoretically through the fluctuator model, and experimentally by varying the magnetic field orientation.
+We concluded that the dominant factor in the zero field depolarization was the degeneracy of the 4 NV classes, followed by the double flips.
+
+The depolarization of NV ensemble in zero field can in itself be used to measure magnetic field.
+We characterized such a magnetometer and compared it to the state of the art NV magnetometers.
+We found in particular that the dominant factor in the sensitivity of the magnetometer was the presence of the double flips, and not the classes degeneracy.
+
+The low-field depolarization magnetometer is unique among the NV magnetometers in that is does not require a microwave or a precise orientation of the magnetic field to work, which makes it usable with powdered or polycrystalline samples.
+Applications of magnetometry with powders include wide-field imaging of uneven surfaces, background free fluorescence microscopy, and microwave-free nano-probe sensing.
+
+\appendix
+
+\chapter{Diamond samples}
+\label{Appendix samples}
+This appendix document the known properties of every diamond samples used in this manuscript.
+
+\section{CVD samples}
+\begin{table}[htbp]
+\centering
+\caption{Concentration in nitrogen and NV centers for the CVD samples}
+\label{Table samples CVD}
+\begin{tabular}{c|cc}
+\toprule
+Name &  [NV$^-$]  & [N] \\
+\midrule
+CVD-pink & 4.6 ppm & 26 ppm \\
+CVD-B & 0.43 ppm & 10-30 ppm \\
+CVD-C & 0.43 ppm & 10-30 ppm \\
+
+\bottomrule
+\end{tabular}
+\end{table}
+Table \ref{Table samples CVD} summarizes the most relevant parameters in our study which are the concentration in (negatively charged) NV centers and the total concentration of nitrogen in the sample.
+\subsection{CVD-pink}
+The growth process and characterization of sample CVD pink are detailed in \citep{tallaire2020high}.
+The sample was grown by CVD in a gas environment containing 500 ppm of N$_2$O.
+The sample was then irradiated by a 10 MeV electron beam at a fluence of $2\cdot 10^{18}\ \rm cm^{-2}$ for 20 hours, while heated to 900 $^\circ$C during the irradiation process.
+
+The concentration of NV centers and nitrogen impurities was determined via UV-visible absorption and FTIR.
+
+\subsection{CVD-B and CVD-C}
+The growth process and characterization of sample CVD pink are detailed in \citep{ngambou2022improving}.
+These two samples were grown with a N$_2$O concentration of 100 ppm in the gas phase, and were irradiated similarly to sample CVD-pink.
+The main difference between the two sample is the further annealing step on sample CVD-B, as detailed in chapter 2.
+
+\section{HPHT samples}
+\subsection{Adamas fluorescent microdimaonds}
+\begin{table}[htbp]
+\centering
+\caption{Concentration in nitrogen and NV centers for the HPHT samples}
+\label{Table samples adamas}
+\begin{tabular}{c|cc}
+\toprule
+Name &  [NV$^-$] & [N] \\
+\midrule
+Adamas & 2.5-3.5 ppm & $\sim$100 ppm \\
+
+\bottomrule
+\end{tabular}
+\end{table}
+
+The main HPHT sample studied in this manuscript are the commercially available red fluorescent agents from Adamas nano.
+Samples labeled ADM-15-$n$ correspond to 15 $\mu$m micro-diamonds, and samples labeled ADM-150-$n$ to 150 $\mu$m micro-diamonds.
+The Adamas website indicates that the fluorescent diamonds contain $\sim 100$ ppm of nitrogen impurities, and that the micro-diamonds are irradiated with 2-3 MeV electron beams and annealed, without further information.
+The [NV] concentration, also indicated on the website, was measured via EPR.
+
+\subsection{Other HPHT samples}
+\subsubsection{Sumi-2}
+Sample Sumi-2 is a type Ib HPHT sample bought from Sumitomo and irradiated with a 2 MeV electron beam and a fluence of $7\cdot 10^{18}\ \rm cm^{-2}$ which were further annealed at 800 $^\circ$C for two hours.
+The NV and N concentration were not measured on this sample.
+\subsubsection{SBST-B and SBST-C}
+Samples SBST-B and SBST-C are the type Ib HPHT substrates on which the samples CVD-B and CVD-C grew.
+They were irradiated and annealed along the CVD layers.
+The NV and N concentrations in these samples has not been measured.
+
+\chapter{Computation of $\bar \eta$}
+\label{Appendix eta}
+We will discuss here the specifics to the computation of the  $\bar \eta$ factor for different geometric configurations, and how to compute the $\Gamma_1$ rates in table \ref{T1 champ mag} of chapter 3, and table Fig.
+\ref{summary_theory} of chapter 4.
+
+As a reminder, $\bar \eta$ is defined as the averaged value of $\eta$ over all possible configurations.
+
+\begin{equation}
+\bar \eta = \int \rm{Prob}(\eta) \abs{\eta} d\eta,
+\end{equation}
+
+where $\eta$ is defined as:
+
+\begin{equation}
+\label{eq. eta debut annexe}
+\eta^2=\frac{1}{3} \left(\frac{\mel{i}{\mathcal{H_{\rm dd}}}{f}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
+\end{equation}
+
+and where $\mathcal{H_{\rm dd}}$ is the dipole-dipole Hamiltonian and $\ket{i}$ and $\ket{f}$ the initial and final two-qubits states of the flip-flop or double flip process.
+
+\section{$\bar \eta$ in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 3/Figures/Shema_annexe}
+\caption{Schematics of the ``normal" NV center NV$_s$ and the fluctuator NV$_f$ in and their respective basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$), as well as the relative position $\mathbf{r}(r,\theta,\phi)$ between the two spins.}
+\label{shema spins annexe}
+\end{figure}
+
+The computation of $\bar \eta$ when the single spin Hamiltonian of each spin is diagonal in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$ was treated in \citep{choi2017depolarization}.
+We will summarize their results here and consider a few different geometries than the ones presented in the article.
+
+Fig.
+\ref{shema spins annexe} represents the two spins and the three relevant angles in the problem $\theta,\phi$ and $\psi$.
+We label with $s$ the properties associated with the ``normal " NV center and $f$ those associated with the fluctuator.
+For instance the two-qubits state $\ket{m_s=0,m_f=+1}\equiv\ket{0,+1}$ corresponds to the convolution of the single-spin states $\ket{m=0}$ for the NV and $\ket{m=+1}$ for the fluctuator.
+
+The two Cartesian basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$) were chosen so that the spin vector Hamiltonian of each spin could be written:
+\begin{equation}
+\mathbf{S}_i=S_x \hat x_i + S_y \hat{y}_i + S_z \hat{z}_i,
+\end{equation}
+where :
+\begin{align*}
+S_x&=\begin{pmatrix}
+0&1/\sqrt{2}&0 \\
+1/\sqrt{2}&0&1/\sqrt{2} \\
+0&1/\sqrt{2}&0
+\end{pmatrix} \\
+S_y&=\begin{pmatrix}
+0&i/\sqrt{2}&0 \\
+-i/\sqrt{2}&0&i/\sqrt{2} \\
+0&-i/\sqrt{2}&0
+\end{pmatrix} \\
+S_z&=\begin{pmatrix}
+-1&0&0 \\
+0&0&0 \\
+0&0&+1
+\end{pmatrix}.
+\end{align*}
+
+The relative position between the two spins is noted with the vector $\mathbf{r}=r\hat{u}$.
+The vector's spherical coordinate are given in the Cartesian basis of the NV center ($\hat x_s,\hat y_s,\hat z_s$).
+
+The angle $\psi$ between the two spins $z$ axes is defined by the crystal lattice and can take three values within [0,$\pi$] which are: 0 (same class), $\cos^{-1}(1/3) \approx 70.5\ ^\circ$ and $\cos^{-1}(-1/3) \approx 109.5\ ^\circ$.
+Two distinct classes can either have a relative angle $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$ depending on the external magnetic field: we choose by definition that $\mathbf{B}\cdot\hat{z}_i >0$, which imposes the orientation of $z_i$ and therefore the angle between $\hat z_s$ and $\hat z_f$.
+This point will be further developed when discussing the $\{100\}$ and $\{110\}$ type resonances.
+
+The direction $\hat x_s=\hat x_f$ is chosen arbitrarily to be : $$\hat x=\frac{\hat z_s \times \hat z_f}{\norm{\hat z_s \times \hat z_f}}.$$
+The arbitrary choice of the $\hat x$ direction is justified by the symmetry of the problem in the ($\hat x, \hat y$) plane.
+We chose here to neglect the symmetry-breaking part of the transverse magnetic field which is justified by the fact that we do not take into consideration the mixing of the eigenstates caused by the transverse field.
+
+Following the notation introduced in \citep{choi2017depolarization}, we can rewrite the dipole-dipole Hamiltonian as:
+\begin{equation}
+\mathcal{H}_{\rm dd} \approx -\frac{J_0}{r^3}\left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}) + h.c. + q S_z^sS_z^f \right],
+\end{equation}
+with 
+\begin{align}
+g&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f) + 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{y}_s \cdot \hat{y}_f) \right] \\
+h&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{x}_s \cdot \hat{y}_f) - 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{x}_f) + (\hat{y}_s \cdot \hat{x}_f) \right] \\
+q&= 3(\hat{u}\cdot \hat{z}_s)(\hat{u}\cdot \hat{z}_f) - (\hat{z}_s \cdot \hat{z}_f).
+\end{align}
+
+We should note that the double flips have been omitted because we consider here the case $B\neq0$ for which the $\ket{+1}$ and $\ket{-1}$ states of each spin are far out of resonance.
+
+Eq. \ref{eq. eta debut annexe} can then be written:
+\begin{equation}
+\eta^2=\frac{1}{3} (\abs{g}^2 + \abs{h}^2) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{equation}
+
+In order to compute $\bar \eta$, we first decompose $\eta$ as a product of $R$ and $G$ as defined in sec. \ref{sec computation eta}:
+\begin{equation}
+\eta^2=G^2(\theta, \phi, \psi) R^2(\omega_f,\omega_{NV})
+\end{equation}
+with
+\begin{align*}
+G^2(\theta, \phi, \psi)&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
+R^2(\omega_f,\omega_{NV}) &= \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
+\end{align*}
+
+$R$ and $G$ can be averaged separately as they do not depend on the same variables, which means that:
+\begin{equation}
+\label{eq. eta bar r bar g bar}
+\bar \eta = \bar R \cdot \bar G.
+\end{equation}
+
+The computation of $\bar{R}$ has been discussed in sec. \ref{sec computation eta}, we will focus here on $\bar{G}$ defined as:
+
+\begin{equation}
+\label{eq. eta base mag}
+\bar{G}= \iint \abs{G} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi}.
+\end{equation}
+
+Eq.
+\ref{eq.
+eta base mag} can be solved analytically for the case $\psi=0$ and numerically for $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$.
+The three values are reported in Table \ref{table G mag}.
+The values found for $\psi=0$ and $\psi=70.5 \ ^\circ$ are the same that were found by the authors of \citep{choi2017depolarization} \footnote{They differ by a factor $\sqrt{1/3}$ because of the definition of $G$.}.
+
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar G$ in the magnetic basis}
+ \label{table G mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+$\frac{2}{9}=0.222$ & 0.3757 & 0.4808 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Interclass resonance and magnetic field orientation}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{chapter 3/Figures/110_vs_100}
+\caption{(a) Illustration of the co-resonance condition between a NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
+The $\hat z$ direction is defined by the condition $\mathbf{B}\cdot\hat{z} >0$.
+(b) 8 individual $T_1$ measurements on sample ADM-15-3 on a two-class resonance for 8 different values of the magnetic field (4 $\mathbf{B} \in \{110\}$ and 4 $\mathbf{B} \in \{100\}$).}
+\label{110 vs 100 annexe}
+\end{figure}
+
+As discussed in sec.
+\ref{sec geometrie resonance}, there are 4 classes of magnetic field orientations for which at least two NV classes are co-resonant: $\mathbf{B} \in \{110\}$ (two-class resonance), $\mathbf{B} \in \{100\}$ (2$\times$two-class resonance), $\mathbf{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\mathbf{B} \parallel \langle 100\rangle$ (four-class resonance).
+
+Fig.
+\ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$.
+We consider here an NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
+Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\mathbf{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
+
+For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\mathbf{B} \in (100)$ or $\mathbf{B} \in (011)$.
+The angle $\psi$ between the $\hat z_s$ and $\hat z_f$ axes however is different in both cases.
+The same can be generalized for every $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\mathbf{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\mathbf{B} \in \{110\}$ type resonance.
+
+Because $\bar G$ is greater for $\psi=109.5 \ ^\circ$ than it is for $\psi=70.5 \ ^\circ$, then following eq.
+\ref{eq.
+eta bar r bar g bar}, we should expect a faster depolarization rate when $\mathbf{B} \in \{110\}$ than when $\mathbf{B} \in \{100\}$.
+Fig.
+\ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\mathbf{B} \in \{110\}$ scenario and four in the $\mathbf{B} \in \{100\}$ scenario.
+We can clearly see a slower relaxation rate for all four $\mathbf{B} \in \{100\}$ measurements compared to all four $\mathbf{B} \in \{110\}$, which agrees with the predictions of the model.
+
+For a perfect resonance matching (i.e.
+no detuning between the central frequencies of the different resonant classes), we can assume that $\bar R$ is a constant ($\bar R \equiv \bar R^0 \approx \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}$ according to eq.
+(\ref{eq 1/T avec T2*})).
+We then have:
+
+\begin{align}
+\bar \eta(\rm{1\ class})&=\bar G(\psi=0) \bar R_0 \\[1em]
+\frac{\bar \eta(\mathbf{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\mathbf{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
+\end{align}
+
+Finally, the values in Table \ref{T1 champ mag} are computed thanks to the relation: \begin{equation}
+\frac{\Gamma_1}{\Gamma_0}= \left(\frac{\bar \eta}{\bar \eta(\rm{1\ class})} \right)^2.
+\end{equation}
+
+Note that we only give here the relaxation rates for a single class.
+To estimate the total PL, one would need to compute the relaxation rate of all 4 classes independently, for instance in a $\mathbf{B} \in \{110\}$ scenario, two classes have a relaxation rate $\Gamma(\mathbf{B} \in \{110\})$ and two classes have a relaxation rate $\Gamma_0$.
+
+\section{$\bar \eta$ in the non-magnetic basis $\{ \ket{0}, \ket{+}, \ket{-}\}$}
+
+In this part we will study the averaged flip-flop and double flip rates in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis which corresponds to the various mechanisms discribed in chapter 4.
+
+We will define here the $\ket{\pm}$ states as:
+\begin{equation}
+\ket{\pm}=\frac{\ket{+1}\pm\ket{-1}}{\sqrt{2}}.
+\end{equation}
+
+In this basis, the spin operators become:
+\begin{align*}
+S_x&=\begin{pmatrix}
+0&0&0 \\
+0&0&1 \\
+0&1&0
+\end{pmatrix} \\
+S_y&=\begin{pmatrix}
+0&-i&0 \\
+i&0&0 \\
+0&0&0
+\end{pmatrix} \\
+S_z&=\begin{pmatrix}
+0&0&1 \\
+0&0&0 \\
+1&0&0
+\end{pmatrix}.
+\end{align*}
+
+Importantly, we have to note the breaking of the symmetry in the (xy) plane.
+This comes from the transverse electric or magnetic field dominating the single spin Hamiltonian.
+By convention, we define the $x$ direction as the direction of the dominant transverse electric or magnetic field.
+
+\subsection{Flip-flops in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis}
+\label{sec flip-flop +/-}
+Rewriting the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-}\} \times \{ \ket{0}, \ket{+}, \ket{-}\}$ changes the value of the flip-flop and double flip prefactors.
+For instance, the prefactor of the $\ket{+,0}\bra{0,+}$ term is: \begin{equation}
+-\frac{J_0}{r^3}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)\right]
+\end{equation}
+while the prefactor of the $\ket{+1,0}\bra{0,+1}$ term was: \begin{equation}
+-\frac{J_0}{r^3}\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f) + 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{y}_s \cdot \hat{y}_f) \right] 
+\end{equation}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\textwidth]{chapter 3/Figures/appendix_alpha_s_et_f}
+\caption{Illustration of then angles $\alpha_s$ and $\alpha_f$ defined by the angle formed by $\hat x_s$ and $\hat x_f$ with the reference axes $\hat x_s^0$ and $\hat x_f^0$.
+Blue arrows symbolize the local transverse electric field acting on the NV and the fluctuator respectively.}
+\label{alpha s et f}
+\end{figure}
+
+Because of the symmetry breaking of the $(xy)$ plane, we now have to consider the relative orientation of $x_s$ and $x_f$ on top of the previously mentioned angles $\theta$, $\phi$ and $\psi$.
+We define $\alpha_s$ and $\alpha_f$ in Fig.
+\ref{alpha s et f} as the relative angle formed by the $x_s$ and $x_f$ axes with the arbitrary axes $\hat x_s^0$ and $\hat x_f^0$ in their respective $(xy)$ plane (the arbitrary axis can be defined for instance in the lab frame of reference).
+
+Because we assume that the physical orientation of the $x_s$ and $x_f$ axes in zero external magnetic field is determined by a random local electric field, we will sample than angles $\alpha_s$ and $\alpha_f$ on a uniform [0,$2\pi$] distribution independently for each spin.
+
+The integral to compute the $\bar G$ factor of the dipole-dipole interaction in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis then becomes\footnote{The flip-flop term considered here is $\ket{+,0}\bra{0,+}$.
+The $\ket{-,0}\bra{0,-}$ term gives the same numerical values once averaged.}
+\begin{equation}
+\bar G= \frac{1}{\sqrt{3}} \iiiint \abs{3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi} \frac{d \alpha_s}{2\pi} \frac{d \alpha_f}{2\pi}.
+\end{equation}
+
+The numerical values obtained are reported in table \ref{table G flip-flop non mag}.
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar G$ in the $\ket{\pm}$ basis}
+ \label{table G flip-flop non mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+0.410 & 0.394 & 0.394 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Finally the value $\Gamma_1^{\rm th}= 51.4 \Gamma_0^{\rm th}$ shown in Fig. \ref{summary_theory} was computed from:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar G_{\ket{\pm}}(\psi=0) + 3 \bar G_{\ket{\pm}}(\psi\neq 0) }{\bar G_{\ket{\pm 1}}(\psi=0)}\right)^2
+\end{equation}
+\subsection{Double-flips in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis}
+Including the double flips, there are two possible relaxation path for each transition of the regular NV center.
+For instance the transition $\ket{0} \to \ket{-}$ can involve either the flip flop $\ket{0,-} \bra{-,0}$ or the double-flip $\ket{0,+} \bra{-,0}$.
+
+The inclusion of the double flips require to modify the  expression of $\eta^2$ to include both relaxation path.
+According to the more general definition given in \citep{choi2017depolarization}, the $\eta^2$ term corresponding to the $\ket{0} \to \ket{-}$ transition of the NV center can be written\footnote{The $\ket{\pm}$ states in the formula can be replaced by $\ket{\pm 1}$ depending on the eigenbasis of the single spin Hamiltonian.}:
+\begin{align}
+\label{eq eta2 avec DQ}
+\eta^2&=\frac{1}{3} \left(\frac{\mel{0,-}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{\left(\Delta_{\omega_f}^{(0,-)} - \Delta_{\omega_{NV}}^{(0,-)}\right)^2+4\gamma_f^2} \nonumber\\
+&+\frac{1}{3} \left(\frac{\mel{0,+}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{\left(\Delta_{\omega_f}^{(0,+)} - \Delta_{\omega_{NV}}^{(0,-)}\right)^2+4\gamma_f^2},
+\end{align}
+where $\Delta_{\omega_{NV}}^{(0,\pm)}$ corresponds to the central frequency of the $\ket{0} \to \ket{\pm}$ transition of the NV center, and $\Delta_{\omega_{f}}^{(0,\pm)}$ that of the fluctuator.
+
+Eq.
+(\ref{eq eta2 avec DQ}) can no longer be factorized as a product of a spectral and spatial function.
+To compute the value of $\bar eta$ in zero magnetic field, we will proceed to the following assumptions: $\Delta_{\omega_{NV}}^{(0,\pm)}=\Delta_{\omega_{f}}^{(0,\pm)}$, and $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$\footnote{This second assumption is justified by experimental measurements.}.
+Eq.
+(\ref{eq eta2 avec DQ}) then become:
+\begin{equation}
+\eta^2=\frac{1}{3} \left(\frac{\mel{0,-}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 
++\frac{1}{3} \left(\frac{\mel{0,+}{\mathcal{H_{\rm dd}}}{-,0}}{\frac{J_0}{r^3}}\right)^2 \frac{1}{2}.
+\end{equation}
+
+Following the same procedure as in the last section, we then compute $\bar \eta$ in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis and report the values in table \ref{table eta double flip non mag}.
+These values are used to compute the double-flip term in Fig.
+\ref{summary_theory}
+
+\begin{table}[htbp]
+\centering
+\caption{Computation of $\bar \eta$ in the $\ket{\pm}$ basis including the double flips.}
+ \label{table eta double flip non mag}
+\begin{tabular}{c|c|c}
+\toprule
+$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\
+
+0.554 & 0.532 & 0.532 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Flip-flops and double flips in the $B\perp [111]$ regime}
+
+We will focus here on the case detailed in Fig.
+\ref{champ tranverse exp}: a single class of NV centers whose Hamiltonian is dominated by the transverse magnetic field.
+Unlike the case presented in Sec.
+\ref{sec flip-flop +/-}, the $x$ direction is fixed for all spins and imposed by the direction of the magnetic field.
+
+The expression of $\bar G$ then becomes:
+\begin{equation}
+\bar{G}= \frac{1}{\sqrt{3}}\iint \abs{3(\hat{u}\cdot\hat{x})^2 - 1} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi}= \frac{4}{9}\approx 0.444.
+\end{equation}
+
+The increase in the decay rate of the eigenbasis when considering only the flip-flop is:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar G_{\ket{\pm}}(\psi=0)}{\bar G_{\ket{\pm 1}}(\psi=0)}\right)^2=4.
+\end{equation}
+
+To include the double-flip, we once again make the assumption that $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$ and find $\bar \eta=0.567$.
+The increase in the decay rate when including the decay rate is:
+\begin{equation}
+\frac{\Gamma_1^{\rm th}}{\Gamma_0^{\rm th}}=\left(\frac{\bar \eta_{\ket{\pm}}(\psi=0)}{\bar \eta_{\ket{\pm 1}}(\psi=0)}\right)^2\approx 6.5.
+\end{equation}
+
+\chapter{Influence of the magnetic field misalignment and laser polarization on the low field relaxation}
+\label{Appendix autres causes}
+\section{Magnetic field alignment}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 4/Figures/alignement_rose}
+\caption{PL of sample CVD-pink as a function of the magnetic field for 7 different misalignment of the field with respect to the [100] axis}
+\label{alignement rose}
+\end{figure}
+
+A potential cause for the observation of depolarization at low magnetic field when $\mathbf{B} \parallel [100]$ could be a misalignment of the magnetic field with the [100] axis.
+Indeed, such a misalignment would cause a splitting of the 4 classes as the magnetic field increases which would reduce the inter-class flip-flop rate.
+%This hypothesis is not consistent with the observations in Fig.
+\ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased.
+Instead we observe a sharp decrease followed by a pseudo-plateau.
+Fig.
+\ref{alignement rose} shows the PL of sample CVD-pink as the magnetic field is scanned with various misalignments compared to the [100] axis.
+We estimate our alignments to be precise within $\pm 1^\circ$.
+We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization in the $\mathbf{B} \parallel [100]$ case does not come from a field misalignment.
+
+\section{Laser polarization}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 4/Figures/pola_laser}
+\caption{PL of sample ADM-150-3 as a function of the magnetic field for six different laser polarization angles.
+The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
+\label{pola laser}
+\end{figure}
+
+Other studies \citep{anishchik2015low, filimonenko2020weak} previously observed a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.
+
+Fig.
+\ref{pola laser} shows PL scans on sample ADM-150-3 for various polarization angle of the incident laser.
+We chose to apply the magnetic field in the laser polarization plane, in a direction that did not match any particular crystalline planes.
+
+We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$.
+We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.
+
+\printbibliography
 \end{document}

### 2022-10-20 Manuscrit Thèse/These_copie.tex
toto

--- 2022-10-19 Manuscrit Thèse/these_fusion.tex
@@
 \jurymember{3}{Maria Chamarro}{Professeure, Sorbonne Université}{Examinateur}
-\jurymember{4}{Vincent Jacques}{Dirrecteur de recherche, CNRS}{Examinateur}
-\jurymember{5}{Alexandre Tallaire}{Dirrecteur de recherche, CNRS}{Examinateur}
-\jurymember{6}{Gabriel Hétet}{Maitre de conférence, ENS Paris}{Directeur de thèse}
+\jurymember{4}{Vincent Jacques}{Directeur de recherche, CNRS}{Examinateur}
+\jurymember{5}{Alexandre Tallaire}{Directeur de recherche, CNRS}{Examinateur}
+\jurymember{6}{Gabriel Hétet}{Maître de conférences, ENS Paris}{Directeur de thèse}
 % \jurymember{9}{Prénom NOM}{Titre, établissement}{Invité}
@@
 \chapter{Diamond samples}
+\label{Appendix samples}
 This appendix document the known properties of every diamond samples used in this manuscript.
@@
 \chapter{Computation of $\bar \eta$}
+\label{Appendix eta}
 We will discuss here the specifics to the computation of the  $\bar \eta$ factor for different geometric configurations, and how to compute the $T_1$ times in table [REF] of chapter [REF], and table [REF] of chapter [REF].
@@
 \chapter{Influence of the magnetic field misalignment and laser polarization on the low field relaxation}
-
+\label{Appendix autres causes}
 \section{Magnetic field alignment}

### 2022-10-20 Manuscrit Thèse/These.tex
tata

@@
 In this introductory chapter, we will first cover the material aspects of NV centers by focusing on the creation process of synthetic diamonds and the required steps to create NV defects inside them.
-We will then cover the fundamentals of NV physics, including the NV center optical and spin properties, basic experiments with NV centers and the various relaxations times ($T_1$,$T_2$,$T_2^*$) which characterizes the NV center's spin.
-Finally we will present the dipole-dipole interaction between two spins and the concept of cross-relaxation.
+We will then cover the optical and spin properties of single NV centers, and introduce the notions of dipole-dipole interaction and cross-relaxation.
+We will then cover basic experiments with NV centers and introduce the experimental tools required to characterize the NV center's spin.
 
@@
 \caption{From the diamond to the NV center.
-a) Drawing of a bulk diamond\footnote{In this manuscript, we will consider any samples above $\sim$ 100 $\mu$m to be bulk.}.
+a) Drawing of a ``bulk" (macroscopic) diamond.
 b) Drawing representing a fluorescence microscopy image of a diamond surface.
@@
 Importantly, the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$\frac{15+91}{15+11}\approx 4$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
-Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
+Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay to the $\ket{0}$ state than to the $\ket{\pm 1}$ states.
 
 These spin selective intersystem crossings mean that an optical pumping scheme in the $\ket{m_s=0}$ spin state takes place under constant green excitation.
-At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}, which is equivalent to a thermal polarization (for B=0) with a spin temperature of $\approx 65\ \rm mK$ \footnote{\begin{equation*}
+At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}.
+This is equivalent to a spin-temperature $\approx 65\ \rm mK$ \footnote{\begin{equation*}
 T=\frac{h D}{k_B \ln (\frac{1-P}{2P})} \approx 65\ \rm mK,
 \end{equation*}
-where $D$ is th ZFS factor described in eq.
-(\ref{eq.
-spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.} while the diamond itself can remain at room temperature.
+where $D$ is th ZFS factor described in eq. (\ref{eq. spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.}.
 
@@
 Different experiments and different models have thus lead to different decay rates \citep{duarte2021effect}.
-
-\section{Basic operations with NV centers}
-
-This section introduces two elementary operations with NV centers that are Rabi oscillations and optically detected magnetic resonance (ODMR).
-
-Rabi oscillations is the basis for any kind of coherent qubit manipulation while ODMR is used to find the transition frequencies of the NV centers, as well as being a basic magnetometry protocol.
-
-This section also covers the experimental setup used for these two operations and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
-
-\subsection{Experimental setup}
-\label{Sec setup cahp 1}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.8\textwidth]{chapter 1/Figures/shema_exp}
-\caption{Confocal microscopy setup detailed in main text}
-\label{setup ch1}
-\end{figure}
-The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field.
-Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.
-
-Fig.
-\ref{setup ch1} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
-Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.
-
-The setup consists of a confocal microscope where the excitation from a green laser (Thorlabs CPS532, 532 nm, 5 mW) is focused on a sample containing NV centers with a microscope objective (NA $\sim$ 0.5).
-The NV PL is then collected by the same objective and isolated from the back-scattered laser light with a dichroic mirror (DM) and additional filters (532 nm notch filter and 645 nm longpass filter to filter out the NV$^0$ PL).
-The PL is collected on an avalanche photodiode (Thorlabs APD410A or Perkin Elmer SPCM-AQRH-15) whose signal is sent to the acquisition card (NI X-series).
-
-The laser beam goes through an acousto-optic modulator (AOM, AA opto-electronic) controlled externally to create arbitrary light pulses (resolution $\sim$ ns).
-The microwave field is generated by a Rhode \& Shwarz SMB 100A generator and emitted with a home-made loop antenna positioned at $<$ 1 mm of the sample.
-A microwave switch (Mini-circuits ZASWA-2-50DRA+) is used to create microwave pulses (resolution $\sim$ 10 ns).
-The control pulses for the AOM and the microwave switch are generated either by the NI X-series card (resolution $\sim$ 1 $\mu$s), or by a pulseblaster ESR-PRO (resolution $\sim$ 10 ns).
-The microwave field can be amplified with a Mini-circuits ZHL-5W-422+ amplifier to achieve higher Rabi frequencies.
-
-Finally the magnetic field is generated either with a permanent magnet (up to $\sim$ 200 mT) or with a homemade electromagnet (up to $\sim$ 30 mT) alimented via a low freqency amplifier (Leybold power function generator 522663).
-Both the permanent magnet and the electromagnet can be  attached to a rotation stage for precise alignment.
-\subsection{Rabi Oscillations}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/Rabi}
-\caption{Rabi oscillations. (a) pulse sequence used to measure Rabi oscillations. (b) Experimental data on sample CVD-pink (see sample details in appendix \ref{Appendix samples}).}
-\label{Rabi}
-\end{figure}
-
-The general technique for experiments involving NV spins proceeds as follow: 
-\begin{itemize}
-\item First the spins are polarized. This is done with a green laser pulse ($1-100$ $\mu$s depending on the laser intensity) which polarizes the spins in the $\ket{m_s=0}$ state at $\sim$ 80 \%.
-\item  Then the spins are manipulated. This process often involves a microwave field resonant with one of the spin transitions.
-\item Finally the spins are read out.
-The simplest spin readout for NV centers consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
-For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore.
-The optimal readout time is of the same order as the spin polarization time ($1-100$ $\mu$s).
-\end{itemize}
-
-Fig.
-\ref{Rabi}-a) shows the pulse sequence used to observe Rabi oscillations with NV centers.
-In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$.
-The microwave frequency or the magnetic field needs to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
-
-Fig.
-\ref{Rabi}-b) shows the result of such an experiment on a spin ensemble.
-The PL signal oscillations correspond to the oscillations of the spin population between the bright $\ket{0}$ state and the dark $\ket{-1}$ state.
-For long microwave pulses ($>$ 1 $\mu$s typically), the spins reach a steady state with  $\sim$ 50 \% population in both the $\ket{0}$ and $\ket{-1}$ states.
-
-\subsection{Optically detected magnetic resonance}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_basique}
-\caption{ODMR protocol.
-(a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$.
-(b) ODMR spectrum with no magnetic field on sample ADM-15-4.}
-\label{ODMR basique}
-\end{figure}
-
-Optically detected magnetic resonance (ODMR) is the most commonly employed magnetometry protocol with NV centers due to its simplicity and the fact that it does not require prior calibration.
-
-The protocol for continuous wave (CW) ODMR is presented in Fig.
-\ref{ODMR basique}.
-In this simple protocol, the microwave is always turned on while the frequency spans a predefined interval $[f_{\rm start}, f_{\rm end}]$.
-The PL is recorded as the microwave is being swept, with a typical acquisition time for each PL sample $> 1\ \rm ms$.
-The sweeping frequency must be slow enough so that the PL can reach a steady state with the microwave field for each sample acquisition.
-
-As we saw in the last section, the steady state of the NV center spin with a resonant microwave field is a statistical mixture of 50 \% $\ket{0}$ and 50 \% $\ket{\pm 1}$ \footnote{We are assuming here that the Rabi frequency is greater than the laser polarizing rate, which is often the case in practice}.
-If the microwave field is not resonant with a spin transition however, the steady state is predominantly $\ket{0}$ due to the spin polarization caused by the laser.
-The NV PL will therefore be lowered when the scanning microwave field becomes resonant with one of the NV spin transitions.
-
-Fig.
-\ref{ODMR basique}-b) shows an example of ODMR spectrum on an NV ensemble with no external magnetic field.
-The reason for the two bumps instead of a single one centered on $D=2870\ \rm MHz$ is the splitting of the $\ket{+1}$ and $\ket{-1}$ levels occasioned by the strain and local electric field, which will be further discussed in chapter 4.
-
-\subsubsection{ODMR with lock-in detection}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_lockin}
-\caption{ODMR protocol with lock-in detection.
-(a) Simplified confocal microscope setup with a lock-in amplifier (LIA) modulating the mircowave and demodulating the PL.
-(b) Demodulated ODMR spectrum with no magnetic field on sample ADM-15-4.}
-\label{ODMR lockin}
-\end{figure}
-
-Because the ODMR protocol delivers a continuous signal (the PL) as a function of a periodic excitation (the microwave field), it is well suited to amplitude or frequency modulation and lock-in detection.
-
-The principle of lock-in detection will not be covered here, the key idea is that lock-in detection allows to translate the signal spectrally in a region less prone to noise.
-In this case, the ODMR signal is a DC measurement and is therefore prone to slow freqency noises such as thermal or mechanical drifts and laser amplitude fluctuations.
-The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low freqency noises, but slow enough compared to the NV photodynamics.
-
-Fig.
-\ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig.
-\ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
-All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1\ \rm kHz$.
-
-\subsection{ODMR with NV ensemble: longitudinal and transverse magnetic field}
-\label{sec champs transverse}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_8_classes}
-\caption{(a) Geometry of the NV center and definition of $B_z$ and $B_\perp$.
-(b) Representation of the 4 possibles orientations (``classes") of the N$-$V axis, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell.
-(c) Simulation of the evolution of the transition frequencies for each class of NV centers as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
-(d) ODMR spectrum on sample ADM-150-1 for a magnetic field of $\sim 90\ \rm G$.}
-\label{ODMR 8 classes}
-\end{figure}
-
-We mentioned previously that the NV spin Hamiltonian (eq.
-(\ref{NV spin Hamiltonian basic})) was anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
-The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig.
-\ref{ODMR 8 classes}-a).
-
-For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$.
-The true eigenstates however remain close to $\ket{0}$, $\ket{\pm 1}$ as long as $\gamma_e B_z > \frac{(\gamma_e B_\perp)^2}{D}$.} can be approximated in a perturtbative approach to: %cf AN.py
-\begin{align}
-\nu_0&= -\frac{(\gamma_e B_\perp)^2}{D} \\
-\nu_{-1} &= D - \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D} \\
-\nu_{+1} &= D + \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}
-\end{align}
-
-Fig.
-\ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal.
-We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian.
-For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and that each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
-
-Fig.
-\ref{ODMR 8 classes}-c) shows a simulation of the evolution of the transition frequencies $\nu_{+1} - \nu_{0}$ and $\nu_{-1} - \nu_{0}$ for all 4 classes as a function of the magnetic field amplitude, with the magnetic field orientation fixed in an arbitrary direction.
-Fig.
-\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles.
-8 lines are visible corresponding to the two transitions of the 4 classes.
-Based on the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.
-
-We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
-\ref{sec ISC}.
-For strong transverse magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
-
-\section{Dynamics of the NV center spin}
-
-This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times.
-Relaxation times are important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
-
-The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ which corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix), and the transverse relaxation time $T_2$\footnote{We follow here the NMR nomenclature} which corresponds to the relaxation of the spin state coherences (non diagonal elements of the density matrix).
-
-The transverse relaxation time, also called coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$.
-The $T_2^*$ time includes inhomogeneous dephasing processes such as the inhomogeneity in space (when multiple qubits are probed) and time of the external fields.
-The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}.
-While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used.
-More careful nomenclatures will label the $T_2$ times as $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance.
-The relaxation times are ordered as follows:
-\begin{equation}
-T_2^* < T_2 < 2 T_1.
-\end{equation}
-
-The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the sample used.
-The following paragraphs will cover the measurement processes of the different relaxation times and the typical values for the samples used in this manuscript.
-
-\subsection{NV spin $T_1$}
-\label{sec T1 NV}
-\subsubsection{Measurement protocol}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/T1_basique}
-\caption{Basic $T_1$ measurement scheme.
-(a) Pulse seqeunce used.
-(b) Experimental data obtained on sample CVD-pink for low laser intensity.
-The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
-\label{T1 basique}
-\end{figure}
-
-$T_1$ is the simplest relaxation time to measure because the spin populations are directly correlated to the NV centers PL.
-Fig.
-\ref{T1 basique} shows the pulse sequence used and an experimental result of the $T_1$ sequence.
-The spin lifetime $T_1$ is one of the only spin property that can be measured without a resonant microwave.
-
-The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig.
-\ref{T1 soustraction}-a).
-This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$.
-Then the sequence is repeated with an additional $\pi$-pulse on one the NV transitions right before the readout.
-This $\pi$-pulse will project the remaining $\ket{0}$ population difference into a $\ket{\pm 1}$ state.
-The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which rejects every contribution to the PL outside of the spin relaxation.
-
-Fig.
-\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample that was used in Fig.
-\ref{T1 basique}.
-The difference between the $S_1$ data in Fig.
-\ref{T1 soustraction}-b) and the the data in Fig.
-\ref{T1 basique}-b) comes from the laser intensity used.
-In Fig.
-\ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant.
-Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark \citep{giri2018coupled, giri2019selective} (most likely due to charge tunneling between impurities \citep{choi2017depolarization}) .
-On the other hand, the laser intensity used in Fig.
-\ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.
-
-Fig.
-\ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal.
-Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig.
-\ref{T1 basique}.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/T1_protocole_soustraction}
-\caption{Common mode rejection $T_1$ measurement scheme.
-(a) Pulse seqeunce used.
-(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity.
-(c) Subtraction $S_1-S_2$ versus dark time $\tau$.
-The experimental data is fitted with an exponential decay of time constant $\tau=4.1\ \rm ms$.}
-\label{T1 soustraction}
-\end{figure}
-
-The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
-The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.
-
-\subsubsection{Physical origin and typical values}
-
-We distinguish three different origin for the the $T_1$ relaxation process:
-\begin{itemize}
-\item \textbf{Spin-phonon relaxation:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which can result in a phonon-induced spin relaxation.
-The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}.
-The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxtion times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at for mili-K temperatures \citep{astner2018solid}.
-
-\item \textbf{Electric and magnetic field noises:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}.
-As a result, magnetic and electric field noise can reduce the spin $T_1$ if their spectrum overlap with the NV transitions.
-This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}.
-The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effects are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
-
-\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipolar interaction betwen two spins will be detailed more thoroughly in the following section.
-Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation.
-The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration and can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
-\end{itemize}
-
-All the measurements shown in this manuscript were performed at room temperature.
-The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$.
-Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses.
-The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed.
-$T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.
-
-\subsection{NV spin $T_2^*$}
-\subsubsection{Measurement protocol}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/Ramsey}
-\caption{Ramsey interferometry protocol.
-(a) Pulse seqeunce used.
-(b) Experimental data obtained on sample CVD-pink.} %The orange line corresponds to an Gaussian fit with time constant $\tau=400 \pm 100\ \rm ns$.}
-\label{Ramsey}
-\end{figure}
-
-The $T_2^*$ time is the characteristic time for the loss of quantum coherence of a freely evolving system.
-It is generally defined as the decay time of of a Ramsey interferometry measurement \citep{barry2020sensitivity}.
-
-Fig.
-\ref{Ramsey} shows the Ramsey pulse sequence and an  example obtained on a spin ensemble.
-The microwave frequency is detuned by $\sim 300\ \rm kHz$ compared to the spin central Larmor frequency which causes oscillations in the signal.
-$T_2^*$ is the decay time of the envelope of the Ramsey fringes.
-Unfortunately, it is hard to correctly fit the envelope of a signal with so few fringes.
-Increasing the number of fringes with our current microwave and optical setup might prove hard, as further microwave detuning would decrease the signal contrast, and would be ultimately limited by the microwave pulses resolution.
-
-Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
-The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity, and because the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
-The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
-
-An alternative measurement of $T_2^*$ can be performed by looking at the linewidth of an ODMR spectrum.
-Indeed, ideally the ODMR spectrum should be the Fourier transform of the Ramsey signal (also called free induction decay in NMR literature).
-In practice however, both methods suffer from experimental noises and imperfections, with Ramsey-type methods being more sensitive to microwave inhomogeneities \citep{barry2020sensitivity} while ODMR is prone to laser and microwave power broadening \citep{dreau2011avoiding}.
-
-The method chosen to extract $T_2^*$ values in this manuscript was ODMR with low microwave power ($\Omega_{\rm Rabi} \ll 1/T_2^*$) and low laser intensity (compared to the the NV optical saturation).
-
-\subsubsection{Physical origin}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_1_classes}
-\caption{ODMR spectra on a single NV class.
-(a) Experimental data on sample CVD-pink, fitted with three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
-(b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
-\label{ODMR 1 classe}
-\end{figure}
-
-For most samples, the leading cause of inhomogeneous dephasing is the temporal and spatial variations of magnetic fields caused by the paramagnetic impurities in the crystal \citep{barry2020sensitivity}.
-The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2.
-We will simply mention here that for sample with high NV density ([NV] $\geq$ ppm), $T_2^*$ is mostly limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
-
-Fig.
-\ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see sample details in appendix \ref{Appendix samples}).
-The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
-We can clearly see a higher linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
-In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
-
-Fig.
-\ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula.
-While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}.
-For Fig.
-\ref{ODMR 1 classe}-a), it was indeed found that Lorentzians fitted the data better than Gaussians.
-For Fig.
-\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, although this could be in part due to the unresolved hyperfine structure.
-
-\subsection{NV spin $T_2$}
-\label{sec T2 echo}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/Echo}
-\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
-\label{Echo}
-\end{figure}
-\subsubsection{Measurement protocol}
-As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.
-
-Fig.
-\ref{Echo} shows the pulse sequence used and an example on the same sample than the Ramsey measurement.
-The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time.
-This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased \citep{hahn1950spin}.
-
-This protocol extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
-Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
-This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
-
-\subsubsection{Physical origin}
-The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence.
-The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}.
-For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2019decoherence}.
 
@@
 Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be explained in the following chapters.
+
+\section{Optically detected magnetic resonance (ODMR)}
+
+This section introduces one of the building block of almost any NV experiment which is the optically detected magnetic resonance (ODMR).
+ODMR is used to find the spin resonances of the NV center, which is a preliminary step to any coherent manipulation of the NV spin, as well as a powerful sensing tool in itself.
+
+This section also covers the experimental setup used for this operation and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
+
+\subsection{Experimental setup}
+\label{Sec setup cahp 1}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{chapter 1/Figures/shema_exp}
+\caption{Confocal microscopy setup (details in main text).}
+\label{setup ch1}
+\end{figure}
+The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field.
+Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.
+
+Fig.
+\ref{setup ch1} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
+Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.
+
+The setup consists of a confocal microscope where the excitation from a green laser (Thorlabs CPS532, 532 nm, 5 mW) is focused on a sample containing NV centers with a microscope objective (NA $\sim$ 0.5).
+The NV PL is then collected by the same objective and isolated from the back-scattered laser light with a dichroic mirror (DM) and additional filters (532 nm notch filter and 645 nm longpass filter to filter out the NV$^0$ PL).
+The PL is collected on an avalanche photodiode (Thorlabs APD410A or Perkin Elmer SPCM-AQRH-15) whose signal is sent to the acquisition card (NI X-series).
+
+The laser beam goes through an acousto-optic modulator (AOM, AA opto-electronic) controlled externally to create arbitrary light pulses (resolution $\sim$ ns).
+The microwave field is generated by a Rhode \& Shwarz SMB 100A generator and emitted with a home-made loop antenna positioned at $<$ 1 mm of the sample.
+A microwave switch (Mini-circuits ZASWA-2-50DRA+) is used to create microwave pulses (resolution $\sim$ 10 ns).
+The control pulses for the AOM and the microwave switch are generated either by the NI X-series card (resolution $\sim$ 1 $\mu$s), or by a pulseblaster ESR-PRO (resolution $\sim$ 10 ns).
+The microwave field can be amplified with a Mini-circuits ZHL-5W-422+ amplifier to achieve higher Rabi frequencies.
+
+Finally the magnetic field is generated either with a permanent magnet (up to $\sim$ 200 mT) or with a homemade electromagnet (up to $\sim$ 30 mT) alimented via a low freqency amplifier (Leybold power function generator 522663).
+Both the permanent magnet and the electromagnet can be  attached to a rotation stage for precise alignment.
+
+The X-series and the pulseblaster, as well as several instruments (microwave generator, motorized stages) are piloted by python scripts which I developed during my PhD.
+%Salaud de Gabriel
+%\subsection{Rabi Oscillations}
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=\textwidth]{chapter 1/Figures/Rabi}
+%\caption{Rabi oscillations. (a) pulse sequence used to measure Rabi oscillations. (b) Experimental data on sample CVD-pink (see sample details in appendix \ref{Appendix samples}).}
+%\label{Rabi}
+%\end{figure}
+%
+%The general technique for experiments involving NV spins proceeds as follow: 
+%\begin{itemize}
+%\item First the spins are polarized. This is done with a green laser pulse ($1-100$ $\mu$s depending on the laser intensity) which polarizes the spins in the $\ket{m_s=0}$ state at $\sim$ 80 \%.
+%\item  Then the spins are manipulated. This process often involves a microwave field resonant with one of the spin transitions.
+%\item Finally the spins are read out.
+The simplest spin readout for NV centers consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
+For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore.
+The optimal readout time is of the same order as the spin polarization time ($1-100$ $\mu$s).
+%\end{itemize}
+%
+%Fig.
+\ref{Rabi}-a) shows the pulse sequence used to observe Rabi oscillations with NV centers.
+In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$.
+The microwave frequency or the magnetic field needs to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
+%
+%Fig.
+\ref{Rabi}-b) shows the result of such an experiment on a spin ensemble.
+The PL signal oscillations correspond to the oscillations of the spin population between the bright $\ket{0}$ state and the dark $\ket{-1}$ state.
+For long microwave pulses ($>$ 1 $\mu$s typically), the spins reach a steady state with  $\sim$ 50 \% population in both the $\ket{0}$ and $\ket{-1}$ states.
+
+\subsection{Optically detected magnetic resonance}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_basique}
+\caption{ODMR protocol.
+(a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$.
+(b) ODMR spectrum with no magnetic field on sample ADM-15-4.
+(see sample details in appendix \ref{Appendix samples}).}
+\label{ODMR basique}
+\end{figure}
+
+Optically detected magnetic resonance (ODMR) is the most commonly employed magnetometry protocol with NV centers due to its simplicity and the fact that it does not require prior calibration.
+
+The protocol for continuous wave (CW) ODMR is presented in Fig.
+\ref{ODMR basique}.
+In this simple protocol, the microwave is always turned on while the frequency spans a predefined interval $[f_{\rm start}, f_{\rm end}]$.
+The PL is recorded as the microwave is being swept, with a typical acquisition time for each PL sample $> 1\ \rm ms$.
+The sweeping frequency must be slow enough so that the PL can reach a steady state with the microwave field for each sample acquisition.
+
+Whenever the microwave becomes resonant with one of the NV transition, it induces a depolarization of the NV center which results in a drop of PL.
+This allows an optical reading of the NV spin spectral response.
+
+Fig.
+\ref{ODMR basique}-b) shows an example of ODMR spectrum on an NV ensemble with no external magnetic field.
+The reason as to why there are two peaks instead of a single one centered on $D=2870\ \rm MHz$ is the splitting of the $\ket{+1}$ and $\ket{-1}$ levels occasioned by the strain and local electric field, which will be further discussed in chapter 4.
+The asymmetry between the two peaks could come either from the microwave polarization or from the uneven spectral response of the antenna \footnote{Our antennas are not impedance matched.}.
+
+ODMR contrast of up to 30 \% of the PL have been recorded with single spins \citep{dreau2011avoiding}. We typically observe ODMR contrast $\leq 5\ \%$ with NV ensemble.
+
+\subsubsection{ODMR with lock-in detection}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_lockin}
+\caption{ODMR protocol with lock-in detection.
+(a) Simplified confocal microscope setup with a lock-in amplifier (LIA) modulating the mircowave and demodulating the PL.
+(b) Demodulated ODMR spectrum with no magnetic field on sample ADM-15-4 (see sample details in appendix \ref{Appendix samples}).}
+\label{ODMR lockin}
+\end{figure}
+
+The ODMR protocol is well suited for lock-in detection with a modulation of the microwave amplitude or frequency modulation \footnote{These are preferred to laser modulation as only the spin part of the NV centers is affected by the microwave field}.
+
+The principle of lock-in detection will not be covered here, the key idea is that lock-in detection allows to translate the signal spectrally in a region less prone to noise.
+In this case, the ODMR signal is a DC measurement and is therefore prone to slow freqency noise such as electric noise, thermal or mechanical drifts and laser amplitude fluctuations.
+The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low frequency noise, but slow enough compared to the NV photodynamics.
+
+Fig.
+\ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig.
+\ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
+All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1\ \rm kHz$.
+
+\subsection{ODMR with NV ensemble: longitudinal and transverse magnetic field}
+\label{sec champs transverse}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_8_classes}
+\caption{(a) Geometry of the NV center and definition of $B_z$ and $B_\perp$.
+(b) Representation of the 4 possibles orientations (``classes") of the N$-$V axes, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell.
+(c) Simulation of the evolution of the transition frequencies for each class of NV centers as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
+(d) ODMR spectrum on sample ADM-150-1 (see sample details in appendix \ref{Appendix samples}) for a magnetic field of $\sim 90\ \rm G$.}
+\label{ODMR 8 classes}
+\end{figure}
+
+We mentioned previously that the NV spin Hamiltonian (eq.
+(\ref{NV spin Hamiltonian basic})) was anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
+The magnetic field is thus often decomposed as a sum of a longitudinal magnetic field $B_z$ and a transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig.
+\ref{ODMR 8 classes}-a).
+
+For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$.
+The true eigenstates however remain close to $\ket{0}$, $\ket{\pm 1}$ as long as $\gamma_e B_z > \frac{(\gamma_e B_\perp)^2}{D}$.} can be approximated in a perturbative approach by: %cf AN.py
+\begin{align}
+\nu_0&= -\frac{(\gamma_e B_\perp)^2}{D}, \nonumber \\
+\nu_{-1} &= D - \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}, \nonumber \\
+\nu_{+1} &= D + \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}.
+\end{align}
+
+Fig.
+\ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal.
+We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian.
+For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond.} and that each class gives rise to two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
+
+Fig.
+\ref{ODMR 8 classes}-c) shows a simulation of the evolution of the transition frequencies $\nu_{+1} - \nu_{0}$ and $\nu_{-1} - \nu_{0}$ for all 4 classes as a function of the magnetic field amplitude, the magnetic field orientation being held to an arbitrary direction.
+Fig.
+\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles.
+8 lines are visible corresponding to the two transitions of the 4 classes.
+From the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.
+
+We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
+\ref{sec ISC}.
+For strong transverse magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
+
+\section{Dynamics of the NV center spin}
+
+This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times.
+Relaxation times are important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
+
+There are different relaxation times corresponding to different physical aspects.
+The longitudinal relaxation time $T_1$ corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix).
+The transverse relaxation times $T_2$ and $T_2^*$ correspond to the relaxation of the coherences  (off-diagonal terms of the density matrix).
+
+$T_2$ and $T_2^*$ are called the homogeneous and inhomogeneous dephasing times\footnote{We follow here the NMR nomenclature.}.
+$T_2^*$ represents to the absolute dephasing time of the spin (or spin ensemble) coherence, while $T_2$ represent the dephasing time measured when a dynamical decoupling scheme is used (Hahn echo \citep{hahn1950spin}, XY-n \citep{maudsley1986modified}, CPMG \citep{carr1954effects} , ...).
+While the $T_2^*$ value is absolute, the $T_2$ value depend on the precise decoupling scheme used.
+$T_2$ and $T_2^*$ values are ultimately limited by the spin lifetime:
+\begin{equation}
+T_2^* < T_2 < 2 T_1.
+\end{equation}
+
+The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the employed sample.
+The following paragraphs will present the measurement protocols of the different relaxation times and the typical values for the samples used in this manuscript.
+
+\subsection{NV spin $T_1$}
+\label{sec T1 NV}
+\subsubsection{Measurement protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/T1_basique}
+\caption{Basic $T_1$ measurement scheme.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink (see sample details in appendix \ref{Appendix samples}.) for low laser intensity.
+The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
+\label{T1 basique}
+\end{figure}
+
+Fig.
+\ref{T1 basique} shows a basic protocol to measure the NV $T_1$ time.
+Fig.
+\ref{T1 basique}-a) shows the pulse sequence used, which in this case does not involve a microwave field, and Fig.
+\ref{T1 basique}-b) shows the measured PL obtained by varying the dark time $\tau$.
+The curve can then be fitted to obtain the spin ensemble $T_1$ value.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/T1_protocole_soustraction}
+\caption{Common mode rejection $T_1$ measurement scheme.
+(a) Pulse sequence.
+(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity ($I_{\rm las}=5\ \rm mW$).
+(c) Subtraction $S_1-S_2$ versus dark time $\tau$.
+The experimental data is fitted with an exponential decay of time constant $\tau=4.1\ \rm ms$.}
+\label{T1 soustraction}
+\end{figure}
+
+The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig.
+\ref{T1 soustraction}-a).
+This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$.
+Then the sequence is repeated with the addition of a $\pi$-pulse on one the NV transitions right before the readout.
+This $\pi$-pulse inverts the population of the $\ket{0}$ and $\ket{\pm 1}$ states.
+The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which rejects every contribution to the PL besides the spin relaxation.
+
+Fig.
+\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample that was used in Fig.
+\ref{T1 basique}, with a higher laser intensity.
+We can notice a stark difference between the $S_1$ curve in Fig.
+\ref{T1 soustraction}-b) and the curve in Fig.
+\ref{T1 basique}-b), despite both of them coming from the same sample and with the same measurement protocol.
+The reason for the difference is that the higher laser intensity used in Fig.
+\ref{T1 soustraction} photoionizes some of the $NV^-$ into NV$^0$ via a two-photon absorption scheme.
+The NV$^0$ can then capture an electron during the dark time \citep{giri2018coupled, giri2019selective} which changes the number of NV$^-$ centers probed depending on the dark time length.
+The goal of the subtraction scheme is to suppress this charge state recombination noise which is common to the two sequences $S_1$ and $S_2$.
+
+Fig.
+\ref{T1 soustraction}-c) shows the subtraction of the $S_1$ and $S_2$ signal.
+Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is satisfyingly fitted by an exponential decay similar to the results found in Fig.
+\ref{T1 basique}.
+
+The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
+The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.
+
+\subsubsection{Physical origin and typical values}
+
+We distinguish three different origin for the the $T_1$ relaxation process:
+\begin{itemize}
+\item \textbf{Spin-phonon relaxation:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which can result in a phonon-induced spin relaxation.
+The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}.
+The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxation times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at  mK temperatures \citep{astner2018solid}.
+
+\item \textbf{Electric and magnetic field noise:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}.
+As a result, magnetic and electric field noise can reduce the spin $T_1$ if a significant portion of their power spectral density overlaps with the NV spin transitions.
+This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}.
+The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effects are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
+
+\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation.
+The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration.
+\end{itemize}
+
+All the measurements shown in this manuscript were performed at room temperature.
+The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$.
+Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses.
+The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed.
+$T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.
+
+\subsection{NV spin $T_2^*$}
+\subsubsection{Measurement protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/Ramsey}
+\caption{Ramsey interferometry protocol.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink.} %The orange line corresponds to an Gaussian fit with time constant $\tau=400 \pm 100\ \rm ns$.}
+\label{Ramsey}
+\end{figure}
+
+$T_2^*$ is the characteristic timescale of the loss of quantum coherence for a freely evolving system.
+It is generally defined as the decay time in a Ramsey interferometry measurement \citep{barry2020sensitivity}.
+
+Fig.
+\ref{Ramsey}-a) shows the Ramsey pulse sequence: after being initialized in the $\ket{0}$ state, the NV spin is transformed with a $\pi/2$ pulse in a coherent superposition of $\ket{0}$ and $\ket{\pm 1}$ which freely precesses at the Larmor frequency.
+After a free precession time $\tau$, the remaining coherence is projected back into spin population with a second $\pi/2$ pulse.
+Fig.
+\ref{Ramsey}-b) shows an experimental example obtained on a spin ensemble.
+The microwave frequency is detuned by $\sim 300\ \rm kHz$ with respect to the spin central Larmor frequency which causes oscillations in the signal.
+$T_2^*$ is the decay time of the envelope of the Ramsey fringes.
+It is unfortunately difficult to correctly fit the envelope of a signal with so few fringes, a recurring issue with our current microwave setup.
+
+Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
+The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity, and because the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
+The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
+
+An alternative measurement of $T_2^*$ can be performed by measuring the linewidth of an ODMR spectrum.
+Indeed, ideally the ODMR spectrum should be the Fourier transform of the Ramsey signal (also called free induction decay in NMR literature).
+In practice however, both methods suffer from experimental noise and imperfections, with Ramsey-type methods being more sensitive to microwave inhomogeneities \citep{barry2020sensitivity} while ODMR is prone to laser and microwave power broadening \citep{dreau2011avoiding}.
+
+The method chosen to extract $T_2^*$ values in this manuscript was ODMR at low microwave power ($\Omega_{\rm Rabi} \ll 1/T_2^*$) and low laser intensity (compared to the NV optical saturation).
+
+\subsubsection{Physical origin}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_1_classes}
+\caption{ODMR spectra on a single NV class.
+(a) Experimental data on sample CVD-pink, fitted with three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
+(b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
+\label{ODMR 1 classe}
+\end{figure}
+
+For most samples, the leading cause of inhomogeneous dephasing is the temporal and spatial variations of magnetic fields caused by the paramagnetic impurities in the crystal \citep{barry2020sensitivity}.
+The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2.
+We will simply mention here that for sample with high NV density ([NV] $\geq$ ppm), $T_2^*$ is mostly limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
+
+Fig.
+\ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see sample details in appendix \ref{Appendix samples}).
+The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
+We can clearly see a higher linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
+In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
+
+Fig.
+\ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula.
+While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}.
+For Fig.
+\ref{ODMR 1 classe}-a), it was indeed found that Lorentzians fitted the data better than Gaussians.
+For Fig.
+\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, although this could be in part due to the unresolved hyperfine structure.
+
+\subsection{NV spin $T_2$}
+\label{sec T2 echo}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/Echo}
+\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
+\label{Echo}
+\end{figure}
+\subsubsection{Measurement protocol}
+As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.
+
+Fig.
+\ref{Echo} shows the pulse sequence used and an example on the same sample than the Ramsey measurement.
+The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time.
+This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased \citep{hahn1950spin}.
+
+This protocol extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
+Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
+This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
+
+\subsubsection{Physical origin}
+The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence.
+The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}.
+For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2019decoherence}.
 
@@
 On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\mathbf{B}\parallel [100]$) and the influence of the laser polarization.
-Following further experiments detailed in appendix \ref{Appendix eta}, we have determined that these two effects were negligible compared to the four ones we identified before.
+Following further experiments detailed in appendix \ref{Appendix autres causes}, we have determined that these two effects were negligible compared to the four ones we identified before.
 

### 2022-10-21 Manuscrit Thèse/These.tex
tutu

@@
 Of all the possible crystals and defects, the most studied defect for quantum applications is the negatively charged nitrogen-vacancy center in diamond \citep{aharonovich2016solid, de2021materials}, often abbreviated to ``NV center".
-Although this defect had been known and correctly identified almost 50 years ago \citep{davies1976optical}, interest in the NV center really sparked after it was first isolated in 1997 \citep{gruber1997scanning}.
+Although this defect has been known and correctly identified almost 50 years ago \citep{davies1976optical}, interest in the NV center really sparked after it was first isolated in 1997 \citep{gruber1997scanning}.
 The detection of single NV centers proved that NV centers fluoresce brightly \citep{gruber1997scanning}, are photostable \citep{kurtsiefer2000stable}, and that their electronic spin can be polarized and readout optically \citep{jelezko2004observation}, making them an promising solid state qubit.
@@
 More specifically, we look at spin polarization exchange between the NV centers themselves and with their environment.
-Polarization exchange (flip-flop, spin diffusion, cross-relaxation, ...) in dense ensemble of spins has been abundantly studied in the context of NMR \citep{abragam1978principles} and spintronics \citep{vzutic2004spintronics}.
-Because NV centers tend to be relatively dilute, the question of spin exchange between between NV themselves \citep{choi2017depolarization} or with their environment \citep{hall2016detection} has only recently been addressed.
+Spin exchange (flip-flop, spin diffusion, cross-relaxation, ...) in dense ensemble of spins has been abundantly studied in the context of NMR \citep{abragam1978principles} and spintronics \citep{vzutic2004spintronics}.
+Because NV centers tend to be relatively dilute, the question of spin exchange between NV themselves \citep{choi2017depolarization} or with their environment \citep{hall2016detection} has only recently been addressed.
 However, as samples with higher NV concentration become available \citep{acosta2009diamonds, tallaire2020high, shenderova2019synthesis}, these aspects are becoming more and more relevant.
@@
 We will then cover the optical and spin properties of single NV centers, and introduce the notions of dipole-dipole interaction and cross-relaxation.
-We will then cover basic experiments with NV centers and introduce the experimental tools required to characterize the NV center's spin.
+We will finally present basic experiments with NV center ensembles and introduce the experimental tools required to characterize the NV center's spin.
 
@@
 c) Zoom on the crystalline structure of an NV center.
-Black balls represent carbon atoms}
+Black balls represent carbon atoms.}
 \label{diamond+NV}
 \end{figure}
-Before studying the NV centers properties, we must first discuss the diamond itself and answer these questions: How are diamond created, what kind of defects are there in diamond, and how to specifically create NV defects.
+Before studying the NV centers properties, we must first discuss the diamond itself and answer these questions: how are diamond created, what kind of defects are there in diamond, and how to specifically create NV defects.
 The field of material science related to diamonds and NV centers is a rich and complex one, and this section will only touch upon basic notions regarding these subjects.
@@
 Diamond is a remarkable material in many respects.
-It is famous for being the hardest and most thermally conductive non-synthetic material.
-It is also a material with high optical refractive index and optical dispersion which makes it an extremely valuable crystal in jewelry.
+It is famous for being the hardest and most thermally conductive non-synthetic material, and to have high optical refractive index and optical dispersion which makes it especially valuable in jewelry.
 
@@
 \includegraphics[width=\textwidth]{chapter 1/Figures/HPHT_and_CVD}
-\caption{Schematics of HPHT (a) and CVD (b) growth process as detailed in main text}
+\caption{Schematics of HPHT (a) and CVD (b) growth process as detailed in main text.}
 \label{HPHT and CVD}
@@
 \ref{HPHT and CVD}-b) illustrates the apparatus of a microwave plasma reactor: it consists in a modified vacuum chamber which acts as a microwave resonator.
-The microwave creates a plasma ball from the the precursor gases with a temperature $T_{\rm plasma} \geq 3000 ^\circ$C \citep{ashfold2020nitrogen}, enough to dissociate the $H_2$ gas in atomic hydrogen \citep{balmer2009chemical}.
+The microwave creates a plasma ball from the precursor gases with a temperature $T_{\rm plasma} \geq 3000 ^\circ$C \citep{ashfold2020nitrogen}, enough to dissociate the $H_2$ gas in atomic hydrogen \citep{balmer2009chemical}.
 The plasma acts as the carbon source for the diamond growing on top of the substrate, and as the catalyst to favor the diamond phase.
@@
 \includegraphics[width=\textwidth]{chapter 1/Figures/formtion_NV_2}
-\caption{Illustration of the formation of NV centers. Black and green dots represents carbon and nitrogen atoms respectively}
+\caption{Illustration of the formation of NV centers. Black and green dots represents carbon and nitrogen atoms respectively.}
 \label{formation NV}
@@
 Several high energy particles have been used to create vacancies, such as protons, electrons, ions, neutrons and even gamma rays \citep{davies1976optical, ashbaugh1988gemstone, kleinsasser2016high}.
-The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their higher availability (compared to neutrons or gamma rays) and their ability to deeply penetrate the crystal.
+The most common irradiation source for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their higher availability (compared to neutrons or gamma rays) and their ability to deeply penetrate the crystal.
 
@@
 A sixth electron is required to form the negatively charged NV$^-$ state.
-(b) Molecular orbitals of the NV center predicted from the the $C_{3v}$ symmetry of the defect.
+(b) Molecular orbitals of the NV center predicted from the $C_{3v}$ symmetry of the defect.
 The electron occupation of the NV$^-$ ground state is represented by red arrows.}
@@
 \ref{NV optical}-a) shows the electronic ground state $^3A_2$ and first excited state $^3E$, where 1 electron from the $a_1$ molecular orbital is promoted to the $e$ level.
-Both of these electron occupations have two unpaired electrons, meaning that they can exist either either in a spin triplet state ($S=1$) or a spin singlet state ($S=0$) \footnote{The two singlet states $^1E$ and $^1A_1$ represented in Fig.
+Both of these electron occupations have two unpaired electrons, meaning that they can exist either in a spin triplet state ($S=1$) or a spin singlet state ($S=0$) \footnote{The two singlet states $^1E$ and $^1A_1$ represented in Fig.
 \ref{NV optical}-b) are both of configuration $a_1^2e^2$ where the two $e$ electrons are either in a the same $e$ orbital, or in an antisymmetric superposition of the two $e$ orbitals \citep{doherty2011negatively}.}.
@@
 Carbon's most common isotope is $^{12}$C at 98.9 \% natural abundance.
-$^{12}$C has a a nuclear spin $I=0$ and therefore does not contribute to the hyper-fine structure.
+$^{12}$C has a nuclear spin $I=0$ and therefore does not contribute to the hyper-fine structure.
 Nitrogen's most common isotope however is $^{14}$N at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does create a hyper-fine splitting.
@@
 \begin{equation}
-\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S}\cdot \bar{\bar{A}}\cdot \mathbf{I},
-\end{equation}
-where $\mathcal{H}_e$ is the purely electronic spin Hamiltonian, $\mathcal{H}_n$ the purely nuclear spin Hamiltonian, $\mathbf{S}$ the electronic spin operator, $\mathbf{I}$ the nuclear spin operator and $\bar{\bar{A}}$ the hyper fine tensor.
+\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \vec{S}\cdot \bar{\bar{A}}\cdot \vec{I},
+\end{equation}
+where $\mathcal{H}_e$ is the purely electronic spin Hamiltonian, $\mathcal{H}_n$ the purely nuclear spin Hamiltonian, $\vec{S}$ the electronic spin operator, $\vec{I}$ the nuclear spin operator and $\bar{\bar{A}}$ the hyper fine tensor.
 \subsubsection{Electronic spin Hamiltonian}
@@
 \label{eq. spin elec}
-\mathcal{H}_e=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} + \mathcal{H}_{elec}(\mathbf{E})+ \mathcal{H}_{strain}(\mathbf{\bar{\bar{\varepsilon}}}).
+\mathcal{H}_e=D S_z^2 + \gamma_e \vec{S} \cdot \vec{B} + \mathcal{H}_{elec}(\vec{E})+ \mathcal{H}_{strain}(\vec{\bar{\bar{\varepsilon}}}).
 \end{equation}
@@
 It originates from the dipole-dipole coupling between the two unpaired electrons of the defect.
-Due to the strong anisotropy of the molecular orbitals of the defect, the fine structure term imposes the quantization axis $\mathbf{z}$ (via the $S_z$ operator) where $\mathbf{z}$ is the direction of the N$-$V axis in the diamond.
+Due to the strong anisotropy of the molecular orbitals of the defect, the fine structure term imposes the quantization axis $\vec{z}$ (via the $S_z$ operator) where $\vec{z}$ is the direction of the N$-$V axis in the diamond.
 $D$ is called the zero field splitting (ZFS) factor and its value is $D\approx 2.87\ \rm GHz$ at 300 K.
 
-$\gamma_e \mathbf{S} \cdot \mathbf{B}$ is the Zeeman splitting of the spin, where $\gamma_e \approx 2.8\ \rm MHz/G$ is the gyromagnetic ratio of the electronic spin (almost equal to that of a free electron in the case of the NV center).
+$\gamma_e \vec{S} \cdot \vec{B}$ is the Zeeman splitting of the spin, where $\gamma_e \approx 2.8\ \rm MHz/G$ is the gyromagnetic ratio of the electronic spin (almost equal to that of a free electron in the case of the NV center).
 
@@
 \begin{equation}
-\mathcal{H}_e=Q I_z^2 + \gamma_n \mathbf{I} \cdot \mathbf{B},
+\mathcal{H}_n=Q I_z^2 + \gamma_n \vec{I} \cdot \vec{B},
 \end{equation}
@@
 
-Finally the hyper-fine tensor can be written diagonally:
+The hyper-fine tensor can be written diagonally:
 \begin{equation}
@@
 
-Fig. \ref{NV spin} shows fine and hyper-fine structure of the NV center without any external fields.
+Fig. \ref{NV spin} shows the fine and hyper-fine structure of the NV center without any external fields.
 
@@
 \label{NV spin Hamiltonian basic}
-\mathcal{H}&=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} \\
+\mathcal{H}&=D S_z^2 + \gamma_e \vec{S} \cdot \vec{B} \\
 &=\begin{pmatrix}
@@
 0 & \frac{\gamma_e (B_x-iB_y)}{\sqrt{2}} & D+\gamma_e B_z
-\end{pmatrix}
+\end{pmatrix} \nonumber
 \end{align}
@@
 The optical excitation and radiative decay between the $^3A_2$ and $^3E$ levels preserve the spin projection $m_s$ \citep{robledo2011spin}.
-There is however a non-radiative decay path which involves an inter-system crossing with the singlet states $^1A_1$ and $^1E$ that depends on the the triplet states spin projection.
+There is however a non-radiative decay path which involves an inter-system crossing with the singlet states $^1A_1$ and $^1E$ that depends on the triplet states spin projection.
 
@@
 Importantly, the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$\frac{15+91}{15+11}\approx 4$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
-Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay to the $\ket{0}$ state than to the $\ket{\pm 1}$ states.
+Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$.} more likely to decay to the $\ket{0}$ state than to the $\ket{\pm 1}$ states.
 
@@
 \end{equation*}
-where $D$ is th ZFS factor described in eq. (\ref{eq. spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.}.
+where $D$ is th ZFS factor described in eq. (\ref{eq. spin elec}.) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.}.
 
@@
 \caption{Geometric representation of two interacting spins.
-($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$) represent Cartesian coordinate systems in the spin frames of reference, where $\mathbf{z}_i$ is the axis of quantization of spin $i$.}
+($\vec{x}_1,\vec{y}_1,\vec{z}_1$) and ($\vec{x}_2,\vec{y}_2,\vec{z}_2$) represent Cartesian coordinate systems in the spin frames of reference, where $\vec{z}_i$ is the axis of quantization of spin $i$.}
 \label{dipole-dipole}
@@
 
-This final introductory section will focus on the dipole-dipole interaction between two electronic spins, which is the main form of spin-spin interaction and plays a key part in the properties of dense spin ensemble.
+This section will focus on the dipole-dipole interaction between two electronic spins, a key part of the properties of dense spin ensemble.
 
 We will consider the general situation represented in Fig.
-\ref{dipole-dipole}: two spins separated by a distance $r$ with their own coordinate systems ($\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1$) and ($\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2$).
+\ref{dipole-dipole}: two spins separated by a distance $r$ with their own coordinate systems ($\vec{x}_1,\vec{y}_1,\vec{z}_1$) and ($\vec{x}_2,\vec{y}_2,\vec{z}_2$).
 The coordinate system is chosen so that each spin operator can be written:
@@
 \begin{equation}
-\hat{\mathbf{S}}_i=\hat{S}_x \mathbf{x}_i + \hat{S}_y \mathbf{y}_i + \hat{S}_z \mathbf{z}_i
+\hat{\vec{S}}_i=\hat{S}_x \vec{x}_i + \hat{S}_y \vec{y}_i + \hat{S}_z \vec{z}_i
 \end{equation}
@@
 where the spin operators $\hat S_{\{x,y,z\} }$ have their usual definitions in term of Pauli matrices.
-In particular, this means that the spin axis of quantization of each spin is $z_i$, which classically corresponds to the magnetization direction of the dipole.
-This axis can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (including the NV center).
+The axis of quantization of each spin $z_i$ can be defined either by the magnetic field or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (including the NV center).
 
@@
 
-Classically, the dipole-dipole interaction between two magnetic moments $\mathbf{m}_1$ and $\mathbf{m}_2$ can be derived by computing the static magnetic field generated by one dipole at the position of the second dipole \cite[p.~188]{jackson1999classical}:
+Classically, the dipole-dipole interaction between two magnetic moments $\vec{m}_1$ and $\vec{m}_2$ can be derived by computing the static magnetic field generated by one dipole at the position of the second dipole \cite[p.~188]{jackson1999classical}:
 
 \begin{align}
-U&=-\mathbf{m}_1 \cdot \mathbf{B}(\mathbf{r}) \nonumber \\
-&=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\mathbf{m}_1\cdot\mathbf{u})(\mathbf{m}_2\cdot\mathbf{u}) - \mathbf{m}_1\cdot\mathbf{m}_2}{r^3}+\frac{8\pi}{3}\mathbf{m}_1\cdot\mathbf{m}_2\,\delta(r)\right], \label{eq.
+U&=-\vec{m}_1 \cdot \vec{B}(\vec{r}) \nonumber \\
+&=-\frac{\mu_0}{4 \pi}\left[ \frac{3 (\vec{m}_1\cdot\vec{u})(\vec{m}_2\cdot\vec{u}) - \vec{m}_1\cdot\vec{m}_2}{r^3}+\frac{8\pi}{3}\vec{m}_1\cdot\vec{m}_2\,\delta(r)\right], \label{eq.
 dipole classique}
@@
 
-where $\mathbf{r}=r \mathbf{u}$ is the relative position of the two dipoles (see Fig. \ref{dipole-dipole}).
-
-The quantum analog of the dipole-dipole interaction between spins simply replaces the magnetic moments $\mathbf{m}_i$ in eq.
+where $\vec{r}=r \vec{u}$ is the relative position of the two dipoles (see Fig. \ref{dipole-dipole}).
+
+The quantum analog of the dipole-dipole interaction between spins simply replaces the magnetic moments $\vec{m}_i$ in eq.
 (\ref{eq.
-dipole classique}) with $\hbar \gamma_i \hat{\mathbf{S}}_i$, where $\gamma_i$ is the gryromagnetic ratio and $\hat{\mathbf{S}}_i$ the spin vector operator for spin $i$:
+dipole classique}) with $\hbar \gamma_i \hat{\vec{S}}_i$, where $\gamma_i$ is the gryromagnetic ratio and $\hat{\vec{S}}_i$ the spin vector operator for spin $i$:
 
@@
 \label{eq. dipole quantique J coupling}
-\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2\,\delta(r)\right].
+\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\vec{S}}_1\cdot\vec{u})(\hat{\vec{S}}_2\cdot\vec{u}) - \hat{\vec{S}}_1\cdot\hat{\vec{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\vec{S}}_1\cdot\hat{\vec{S}}_2\,\delta(r)\right].
 \end{equation}
@@
 \begin{equation}
-\mathcal{H}_{\rm dd}= \frac{J_0}{r^3}\left[ 3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2] \right],
+\mathcal{H}_{\rm dd}= \frac{J_0}{r^3}\left[ 3 (\hat{\vec{S}}_1\cdot\vec{u})(\hat{\vec{S}}_2\cdot\vec{u}) - \hat{\vec{S}}_1\cdot\hat{\vec{S}}_2] \right],
 \end{equation}
@@
 
-For two electronic spins with $g$-factors $\approx -2$\footnote{Most spin defects in diamond, including the NV center have a g-factor $\approx -2$ similarly to free electrons.}, 
+For two electronic spins with $g$-factors $\approx -2$ \footnote{Most spin defects in diamond, including the NV center have a g-factor $\approx -2$ similarly to free electrons.}, 
 \begin{equation*}
@@
 
-\subsection{Decomposition of the dipole-dipole Hamiltonian}
+\subsection{Spin exchange mechanisms}
 \label{sec flip-flop et double flip}
@@
 \includegraphics[width=.7\textwidth]{chapter 1/Figures/flip-flop_double-flip}
-\caption{Representation of the two energy transfer mechanisms: flip-flop and double flips} 
+\caption{Representation of the two spin exchange mechanisms: flip-flop and double flips.} 
 \label{flip flop double flip}
@@
 
-In the weak coupling approximation, the dipole-dipole Hamiltonian between two spins can be decomposed between diagonal terms which are responsible for energy shifts of the individual spin levels, and the off-diagonal terms which are responsible for energy transfers between the two spins.
+We will here make the difference between the dipole-dipole Hamiltonian diagonal terms which are responsible for energy shifts of the individual spin levels, and the off-diagonal terms which are responsible for spin exchange \footnote{We consider here the weak coupling regime where we neglect the spin-mixing induced by the dipole-dipole interaction.
+This is justified by the experimental values in our system ($\expval{\mathcal{H}_{dd}} \ll \hbar/T_2^*.$).}.
 
@@
 Following the notation introduced in Fig.
-\ref{dipole-dipole}, we will then assume that $$(\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1) = (\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2) = (\mathbf{x},\mathbf{y},\mathbf{z}).$$ We will denote $\mathbf{u}=u_x \mathbf{x} + u_y \mathbf{y} + u_z \mathbf{z}$ the decomposition of the unitary vector $\mathbf{u}$ in this basis.
+\ref{dipole-dipole}, we will then assume that $$(\vec{x}_1,\vec{y}_1,\vec{z}_1) = (\vec{x}_2,\vec{y}_2,\vec{z}_2) = (\vec{x},\vec{y},\vec{z}).$$ We will denote $\vec{u}=u_x \vec{x} + u_y \vec{y} + u_z \vec{z}$ the decomposition of the unitary vector $\vec{u}$ in this basis.
 The dipole-dipole Hamiltonian can then be written:
@@
 where $S_{\{x,y,z\} }^i$ are the spin operators acting on spin $i$.
-The $S_z^1 S_z^2$ term is the only diagonal term and corresponds to energy shifts while the $S_x^1 S_x^2$ and $S_y^1 S_y^2$ terms are responsible for energy transfer.
-To make the transfer mechanism more apparent, we will rewrite $S_x^i$ as $\frac{1}{2}(S_+^i+S_-^i)$ and $S_y^i$ as $\frac{1}{2i}(S_+^i-S_-^i)$:
+The $S_z^1 S_z^2$ term is the only diagonal term and corresponds to energy shifts while the $S_x^1 S_x^2$ and $S_y^1 S_y^2$ terms corresponds to spin exchange.
+To make the exchange mechanisms more apparent, we will rewrite $S_x^i$ as $\frac{1}{2}(S_+^i+S_-^i)$ and $S_y^i$ as $\frac{1}{2i}(S_+^i-S_-^i)$:
 \begin{align}
@@
 (\ref{eq.
-ff df}) the three relevant matrix elements $\Omega_{\rm ff}$, $\Omega_{\rm df}$, and $\Omega_{\rm shift}$ which correspond respectively to flip-flops, double flips and energy shifts.
-Flip-flops and double flips (sometimes referred as flip-flips) are represented in Fig.
-\ref{flip flop double flip}.
-Flip flops correspond to energy transfer between two spins for which the sum of the spin angular momentum is conserved ($\Delta (m_s^1+m_s^2)=0$), while double flips correspond to energy transfer for which the spin angular momentum is not conserved ($\Delta (m_s^1+m_s^2)=\pm 2$)\footnote{In order to preserve the total angular momentum, double flips transfer angular moment to the crystal, which is a form of Einstein de Haas effect \citep{einstein1915experimental} }.
-In both cases however, the energy between the initial and final states needs to be preserved for the transfer to be efficient.
-
-It should be noted that on average, each matrix element is of the same order of magnitude: $\Omega_{\rm ff} \sim \Omega_{\rm df} \sim \Omega_{\rm shift} \sim \frac{J_0}{r^3}$.
+ff df}) the three matrix elements $\Omega_{\rm ff}$, $\Omega_{\rm df}$, and $\Omega_{\rm shift}$ which correspond respectively to flip-flops, double flips (or flip-flip) and energy shifts.
+It should be noted that each matrix element is typically of the same order of magnitude: $\Omega_{\rm ff} \sim \Omega_{\rm df} \sim \Omega_{\rm shift} \sim \frac{J_0}{r^3}$.
+
+Fig.
+\ref{flip flop double flip} illustrates the difference between flip-flops and double flips: flip-flop preserve the total spin angular momentum ($\Delta (m_s^1+m_s^2)=0$) while double-flips do not ($\Delta (m_s^1+m_s^2)=\pm 2$)\footnote{In order to preserve the total angular momentum, double flips transfer angular moment to the crystal, which is a form of Einstein de Haas effect \citep{einstein1915experimental}.}.
+Both double-flips and flip-flop are energy conservative processes, which means that they require the initial and final two-spin states to be nearly resonant (within $\sim \Omega$) in order to be efficient.
 
@@
 \label{Sec_CR}
-\begin{figure}[h]
+\begin{figure}[h!]
 \centering
 \includegraphics[width=0.9\textwidth]{chapter 1/Figures/Shema_CR}
-\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
+\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin.
+Red dots represent the population in each states and the arrows represent the various population transfer mechanisms.}
 \label{CR_shema}
@@
 Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
-This process can occur through either flip-flops or double flips, as long as the resonance condition between the two spins is met.
-
-On of the condition for cross-relaxation is to have a difference in polarization between the two spins species considered.
-The NV center is therefore a good candidate to observe cross-relaxations as it can be optically polarized while most other spin species are not.
+This spin exchange mechanism can be either flip-flop or double flips.
+
+There are two conditions required to observe cross-relaxation: 
+\begin{itemize}
+\item There must be a difference of polarization between the two spins.
+\item The two spin transitions must be resonant.
+\end{itemize}
+
+The NV center is therefore a good candidate to observe cross-relaxations as it can be optically polarized which creates a difference of polarization with the other spin species in the diamond.
 
@@
 
-For a single NV center coupled to spin 2, assuming that spin 2 always remains depolarized\footnote{This is the case in practice if the spin 2 bath is much bigger than the NV bath and effectively acts as a thermostat, or if the lifetime of spin 2 is much smaller than that of the NV center.} and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$ (weak coupling regime), the modification in the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ caused by the cross-relaxation with spin 2 can be analytically computed \citep{hall2016detection}:
+For a single NV center coupled to spin 2, assuming that spin 2 always remains depolarized\footnote{This is the case in practice if the spin 2 bath is much larger than the NV bath and effectively acts as a thermostat, or if the lifetime of spin 2 is much smaller than that of the NV center.} and that the coupling strength between the two spins is smaller than their respective dephasing time $T_2^*$ (weak coupling regime), the modification $\delta \Gamma_1$ to the NV relaxation rate $\Gamma_1=\frac{1}{T_1}$ is \citep{hall2016detection}:
 \begin{equation}
@@
 \end{equation}
-where $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin, $\delta \nu$ is the detuning in energy (expressed in frequency) between the initial and final states and $\Omega=\Omega_{\rm ff}$ or $\Omega_{\rm df}$ is the dipole-dipole matrix element corresponding to the flip-flop or double-flip process considered.
+where $\Gamma_2^*=\frac{1}{T_2^*(\rm NV)}+\frac{1}{T_2^*(\rm Spin 2)}$ is the joint dephasing rate of the NV center and the second spin \footnote{This formula assume that both NV and spin 2 have a Lorentzian spectral response with width $T_2^*$.}, $\delta \nu$ is the detuning in energy (expressed in frequency) between the initial and final states and $\Omega=\Omega_{\rm ff}$ or $\Omega_{\rm df}$ is the dipole-dipole matrix element corresponding to the flip-flop or double-flip process involved.
 
@@
 (\ref{delta gamma 1}) is the basis for every cross-relaxation detection performed in this manuscript.
-Because the NV center lifetime can be measured optically, as detailed in Sec.
+Because the NV center lifetime can be measured optically, as will be detailed in Sec.
 \ref{sec T1 NV}, cross-relaxation between NV-centers and other spin species can be measured optically.
-Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be explained in the following chapters.
+Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be shown in the following chapters.
+
+The last two sections of this chapter are dedicated to basic experiments involving NV centers.
 
@@
 This section introduces one of the building block of almost any NV experiment which is the optically detected magnetic resonance (ODMR).
-ODMR is used to find the spin resonances of the NV center, which is a preliminary step to any coherent manipulation of the NV spin, as well as a powerful sensing tool in itself.
+ODMR is used to find the spin resonances of the NV center, which is a preliminary step to any coherent manipulation of the NV spin.
 
@@
 
-The X-series and the pulseblaster, as well as several instruments (microwave generator, motorized stages) are piloted by python scripts which I developed during my PhD.
+The X-series card and the pulseblaster, as well as several instruments (microwave generator, motorized stages) are piloted by python scripts which I developed during my PhD.
 %Salaud de Gabriel
@@
 
-The ODMR protocol is well suited for lock-in detection with a modulation of the microwave amplitude or frequency modulation \footnote{These are preferred to laser modulation as only the spin part of the NV centers is affected by the microwave field}.
+The ODMR protocol is well suited for lock-in detection with a modulation of the microwave amplitude or frequency modulation \footnote{These are preferred to laser modulation as only the spin part of the NV centers is affected by the microwave field.}.
 
@@
 Fig.
-\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles.
-8 lines are visible corresponding to the two transitions of the 4 classes.
+\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles: 8 lines are visible corresponding to the two transitions of the 4 classes.
 From the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.
 
-We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
+We should also note that the state mixing induced by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
 \ref{sec ISC}.
@@
 
-There are different relaxation times corresponding to different physical aspects.
+There are different relaxation times corresponding to different spin properties.
 The longitudinal relaxation time $T_1$ corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix).
-The transverse relaxation times $T_2$ and $T_2^*$ correspond to the relaxation of the coherences  (off-diagonal terms of the density matrix).
+The transverse relaxation times $T_2$ and $T_2^*$ correspond to the relaxation of the coherences (off-diagonal elements of the density matrix).
 
 $T_2$ and $T_2^*$ are called the homogeneous and inhomogeneous dephasing times\footnote{We follow here the NMR nomenclature.}.
-$T_2^*$ represents to the absolute dephasing time of the spin (or spin ensemble) coherence, while $T_2$ represent the dephasing time measured when a dynamical decoupling scheme is used (Hahn echo \citep{hahn1950spin}, XY-n \citep{maudsley1986modified}, CPMG \citep{carr1954effects} , ...).
+$T_2^*$ represents the absolute dephasing time of the spin (or ensemble of spins) coherence, while $T_2$ represent the dephasing time measured when a dynamical decoupling scheme is used (Hahn echo \citep{hahn1950spin}, XY-n \citep{maudsley1986modified}, CPMG \citep{carr1954effects} , ...).
 While the $T_2^*$ value is absolute, the $T_2$ value depend on the precise decoupling scheme used.
@@
 (a) Pulse seqeunce used.
-(b) Experimental data obtained on sample CVD-pink (see sample details in appendix \ref{Appendix samples}.) for low laser intensity.
+(b) Experimental data obtained on sample CVD-pink (see sample details in appendix \ref{Appendix samples}.) for low laser intensity ($I_{\rm las} \approx 50\ \mu\rm W$).
 The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
@@
 Fig.
-\ref{T1 basique}-a) shows the pulse sequence used, which in this case does not involve a microwave field, and Fig.
-\ref{T1 basique}-b) shows the measured PL obtained by varying the dark time $\tau$.
+\ref{T1 basique}-a) shows the pulse sequence of the measurement: the spin is initially polarized in the $\ket{0}$ state with a first laser pulse, the system is then left to freely relax for a duration (``dark time") $\tau$ before the spin state is read out with a second laser pulse.
+Fig.
+\ref{T1 basique}-b) shows the PL values obtained by varying the dark time $\tau$: the PL decreases as $\tau$ increases due to the spin relaxation and the depolarization of the NV center.
 The curve can then be fitted to obtain the spin ensemble $T_1$ value.
@@
 (a) Pulse sequence.
-(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity ($I_{\rm las}=5\ \rm mW$).
+(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity ($I_{\rm las}\approx 5\ \rm mW$).
 (c) Subtraction $S_1-S_2$ versus dark time $\tau$.
@@
 The reason for the difference is that the higher laser intensity used in Fig.
-\ref{T1 soustraction} photoionizes some of the $NV^-$ into NV$^0$ via a two-photon absorption scheme.
+\ref{T1 soustraction} photoionizes some of the NV$^-$ into NV$^0$ via a two-photon absorption scheme.
 The NV$^0$ can then capture an electron during the dark time \citep{giri2018coupled, giri2019selective} which changes the number of NV$^-$ centers probed depending on the dark time length.
-The goal of the subtraction scheme is to suppress this charge state recombination noise which is common to the two sequences $S_1$ and $S_2$.
+The goal of the subtraction scheme is to suppress this charge state noise which is common to the two sequences $S_1$ and $S_2$.
 
@@
 The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
-The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.
-
-\subsubsection{Physical origin and typical values}
-
-We distinguish three different origin for the the $T_1$ relaxation process:
+The common mode rejection scheme described here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics in the $T_1$ analysis.
+
+\subsubsection{Physical origin and typical values of $T_1$}
+
+We distinguish three different origin for the $T_1$ relaxation process:
 \begin{itemize}
@@
 The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration.
+Value as low as $25\ \rm{\mu s}$ have been recorded for NV-P1 resonances in a type Ib diamond \citep{hall2016detection}.
 \end{itemize}
@@
 
-\subsubsection{Physical origin}
+\subsubsection{Physical origin of $T_2^*$}
 
@@
 \caption{ODMR spectra on a single NV class.
-(a) Experimental data on sample CVD-pink, fitted with three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
+(a) Experimental data on sample CVD-pink (see sample details in appendix \ref{Appendix samples}), fitted by three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
 (b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
@@
 The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2.
-We will simply mention here that for sample with high NV density ([NV] $\geq$ ppm), $T_2^*$ is mostly limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
+We will simply mention here that for samples with high NV density ([NV] $\geq$ ppm), $T_2^*$ is typically limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
 
@@
 \ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see sample details in appendix \ref{Appendix samples}).
-The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
-We can clearly see a higher linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
-In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
+The first sample is a CVD sample with a total substitutional nitrogen concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
+We can clearly see a larger linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
+In particular, we can resolve the hyperfine structure caused by the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
 
@@
 \ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula.
-While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}.
+While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic} \footnote{If the dominant broadening mechanism is the dipole-dipole coupling to paramagnetic impurities}.
 For Fig.
@@
 For Fig.
-\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, although this could be in part due to the unresolved hyperfine structure.
+\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, which could be due to the unresolved hyperfine structure or to other broadening sources such as crystal strain or magnetic field inhomogeneity.
 
@@
 \includegraphics[width=\textwidth]{chapter 1/Figures/Echo}
-\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
+\caption{Spin-echo protocol. (a) Pulse sequence. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
 \label{Echo}
@@
 \subsubsection{Measurement protocol}
-As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.
-
-Fig.
-\ref{Echo} shows the pulse sequence used and an example on the same sample than the Ramsey measurement.
-The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time.
-This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased \citep{hahn1950spin}.
-
-This protocol extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
+As precised in the introduction of this section, the exact $T_2$ definition depends on the employed rephasing scheme.
+We will focus here on the simplest rephasing process which is the spin-echo or Hahn-echo sequence.
+
+Fig.
+\ref{Echo}-a) shows the spin-echo pulse sequence: it differs from the Ramsey one by the addition of a $\pi$-pulse in the middle of the free precession time.
+This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased which extends the spin coherence time \citep{hahn1950spin}.
+
+Fig.
+\ref{Echo}-b) shows an experimental result of a spin echo on the same sample that was used for the Ramsay sequence.
+The echo extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
 Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
-This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
-
-\subsubsection{Physical origin}
+%This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
+
+\subsubsection{Physical origin of $T_2$}
 The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence.
 The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}.
-For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2019decoherence}.
+For samples with high nitrogen concentration, it has been measured that the $T_2$ time was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2020decoherence}.
 
@@
 
-We have covered in this introductory chapter every concepts needed to understand the results presented in the next chapters.
-We have seen the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them, and we have introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.
+In this introductory chapter, we have covered the concepts needed to understand the results presented in the next chapters.
+We have detailed the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them.
+We have also introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.
 
@@
 
-The dephasing rate caused by a given impurity is directly proportional to the average diagonal dipole-dipole matrix elements ($\Omega_{\rm shift}$ in (eq.
-\ref{eq.
+The dephasing rate caused by a given impurity is directly proportional to the average diagonal dipole-dipole matrix elements ($\Omega_{\rm shift}$ in eq.
+(\ref{eq.
 ff df})) between the two spin species \citep{taylor2008high, bauch2020decoherence}.
@@
 
-Electron paramagnetic resonance (EPR), the electronic equivalent to NMR, is the the most common way to detect electronic spins in solids and liquids.
+Electron paramagnetic resonance (EPR), the electronic equivalent to NMR, is the most common way to detect electronic spins in solids and liquids.
 The principle of its detection is the absorption or reflection of a microwave field resonant with a spin transition.
@@
 
-ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ spins \citep{mitchell2013x}.
+ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ spins \citep{mitchell2013x}.
 This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a spin with a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and in part from the more efficient photon detection in the optical range than in the microwave range.
@@
 
- The second method simply consists in monitoring the PL of the NV centers : since the PL is proportional to the population in the $\ket{0}$ state, and since the steady state population in the $\ket{0}$ state results from the equilibrium between the optical polarization rate $\Gamma_{\rm las}$ and the depolarization rate $\Gamma_1=1/T_1$.
-A simple rate equation for the $\ket{0}$ state population $\rho_{00}$ gives:
+ The second method simply consists in monitoring the PL of the NV centers : the PL is proportional to the population in the $\ket{0}$ state, which itself depends on the equilibrium on the depolarization rate $\Gamma_1=1/T_1$.
+A simple rate equation for the $\ket{0}$ steady state population $\rho_{00}$ gives:
  \begin{equation}
@@
 \includegraphics[width=0.5\textwidth]{chapter 2/Figures/DEER}
-\caption{Pulse sequence of the DEER protocol}
+\caption{Pulse sequence of the DEER protocol.}
 \label{DEER}
@@
 \ref{DEER}.
-The pulse sequence consists in applying a spin-echo on the NV center, as described in \ref{sec T2 echo}, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers.
+The pulse sequence consists in applying a spin-echo sequence on the NV center, as described in \ref{sec T2 echo}, with a simultaneous $\pi-$pulse on the probed dark spins during the rephasing pulse on the NV centers.
 This additional pulse means that the static (or slowly variable) contribution of the probed spins to the $T_2^*$ of the NV center will not be rephased.
@@
 Typically, the magnetic field is scanned in a fixed direction while the signal, either PL or T1, is recorded.
-Due to the strong NV anisotropy however, the direction along which $\mathbf{B}$ is scanned plays a crucial role in the ability to detect NV-dark spins CR.
+Due to the strong NV anisotropy however, the direction along which $\vec{B}$ is scanned plays a crucial role in the ability to detect NV-dark spins CR.
 
@@
 a) When the magnetic field is along the diamond [100] crystalline axis, b) along the [111] axis.
-A representation of the 4 classes of NV centers, labeled 1 to 4, and the orientation of the magnetic field in both cases is put on top}
+A representation of the 4 classes of NV centers, labeled 1 to 4, and the orientation of the magnetic field is put on top in both cases.}
 \label{121 vs 22 NV}
@@
 
-Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\mathbf{B} \parallel [100]$ or 
-$\mathbf{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes.
-When $\mathbf{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\mathbf{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.
+Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\vec{B} \parallel [100]$ or 
+$\vec{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes.
+When $\vec{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\vec{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.
 
 These changes in the field orientation result in important changes in the NV behavior, mainly due to the transverse magnetic field: Fig.
-\ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the ground state spin Hamiltonian (REF chapter 1) - or to be more precise the transitions between the ground state of the spin Hamiltonian and the two excited states - for any of the four classes of NV centers when $\mathbf{B} \parallel [100]$.
-Notably, there is no transitions frequencies below 2500 MHz.
+\ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the NV spin Hamiltonian for any of the four classes of NV centers when $\vec{B} \parallel [100]$.
+Notably, there is no transitions frequencies below 2500 MHz regardless of the field strength.
 
 On the other hand, Fig.
-\ref{121 vs 22 NV}-b) shows the transition frequencies when $\mathbf{B} \parallel [111]$, for both the class aligned and the three class misaligned.
-The three classes behave similarly than the four classes in the $[100]$ case, but the one class aligned with the magnetic field can get its $\ket{0}\to\ket{-1}$ transition arbitrarily small, as long as the magnetic field is perfectly aligned with the NV axis.
+\ref{121 vs 22 NV}-b) shows the transition frequencies when $\vec{B} \parallel [111]$, for both the class aligned and the three classes misaligned.
+The three classes behave similarly than the four classes in the $[100]$ case, but the $\ket{0}\to\ket{-1}$ transition of the class aligned with the magnetic field can become (ideally) arbitrarily small.
 
@@
 
-\subsection{CR condition between NV$^-$ and P1}
+\subsection{Co-resonance between NV$^-$ and P1}
 
@@
 \includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_P1}
-\caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}
+\caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis.}
 \label{121 vs 22 P1}
@@
 
-P1 centers, the most abundant electronic spin in the diamonds commonly used, have a $1/2$ electronic spin.
-They do have a small magnetic field anisotropy coming from the hyper-fine coupling to the nitrogen nucleus ($\sim 100\ \rm MHz$), but this dependency on the field orientation is small compared to the NV center's zero field splitting $D=2870\ \rm MHz$.
-We can therefore, to the first order, neglect the hyper-fine coupling and consider P1 centers as ideal, isotropic, spin $1/2$.
-
-Fig.
-\ref{121 vs 22 P1} shows the transition frequencies for the NV center and such a spin $1/2$, as a function of a magnetic field parallel to [100] or [111].
-When $\mathbf{B} \parallel [100]$, there will never be a co-resonance between the NV and P1 transitions, meaning that no NV-P1 cross-relaxation can be observed when B is scanned in this direction.
-When $\mathbf{B} \parallel [111]$ however, the P1 $\ket{-1/2} \to \ket{+1/2}$ transition matches in energy the $\ket{0}\to\ket{-1}$ NV transition for B=512 G.
-This means that when B reaches $\approx 512\ \rm G$, some of the polarization of the NV centers will be transferred to the P1 centers, which will result in a drop in the NV PL.
-
-\subsection{CR condition between NV$^-$ and VH$^-$}
+P1 centers, the most abundant electronic spin in the samples employed in this manuscript, have a $1/2$ electronic spin.
+Due to the hyper-fine coupling to the $^{14}$N nuclei and to a Jahn-Teller effect, the P1 spin possess a small magnetic anisotropy ($\sim 100\ \rm MHz$).
+This dependency on the field orientation is small compared to the NV center's zero field splitting $D=2870\ \rm MHz$, and we will here neglect it and consider the P1 centers as ideal, isotropic, spin $1/2$.
+
+Fig.
+\ref{121 vs 22 P1} shows the transition frequencies for the NV center and a P1 center, as a function of the amplitude of a magnetic field parallel to [100] or [111].
+When $\vec{B} \parallel [100]$, there are no possible co-resonance between the NV and P1 transitions.
+When $\vec{B} \parallel [111]$ however, the P1 $\ket{-1/2} \to \ket{+1/2}$ transition matches in energy the $\ket{0}\to\ket{-1}$ NV transition for B=512 G.
+NV-P1 cross-relaxation can therefore only be observed by scanning the magnetic field in the [111] direction.
+
+\subsection{Co-resonance between NV$^-$ and VH$^-$}
 
@@
 \includegraphics[width=0.9\textwidth]{chapter 2/Figures/121_vs_22_VH}
-\caption{Simulation of the transition frequencies of NV$^-$ centers and VH$^-$ centers for a magnetic field a) along the [100] axis and b) along the [111] axis}
+\caption{Simulation of the transition frequencies of NV$^-$ centers and VH$^-$ centers for a magnetic field a) along the [100] axis and b) along the [111] axis.}
 \label{121 vs 22 VH}
@@
 Fig.
-\ref{121 vs 22 VH} shows the transition frequencies for the NV$^-$ and VH$^-$ ground state spin.
-Contrary to the P1 case, there is a co-resonance when $\mathbf{B} \parallel [100]$ for $B\approx 55\ \rm G$.
-When $\mathbf{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G.
+\ref{121 vs 22 VH} shows the transition frequencies of the NV$^-$ and VH$^-$ spin.
+Contrary to the P1 case, there is a co-resonance when $\vec{B} \parallel [100]$ for $B\approx 55\ \rm G$.
+When $\vec{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G.
 A scan along the [111] in this case means that the peak signal will be $\sim 6$ times weaker than in the [100] case, and spread over multiple lines which might overlap between themselves or with other existing PL features.
@@
 
-\subsection{CR condition between NV$^-$ and and $^{13}$C-NV$^-$}
+\subsection{Co-resonance between NV$^-$ and $^{13}$C-NV$^-$}
 \label{sec NV-13c-NV}
@@
 \includegraphics[width=1\textwidth]{chapter 2/Figures/121_vs_22_13CNV}
-\caption{a) Representation of the $^{13}$C-NV complex (right) compared to a ``normal" NV center (left) b) and c) simulation of the transition frequencies of NV$^-$ centers and $^{13}$C-NV$^-$ complex for a magnetic field b) along the [100] axis and c) along the [111] axis}
+\caption{a) Representation of the $^{13}$C-NV complex (right) compared to a ``normal" NV center (left) b) and c) simulation of the transition frequencies of NV$^-$ centers and $^{13}$C-NV$^-$ complex for a magnetic field b) along the [100] axis and c) along the [111] axis.}
 \label{121 vs 22 13C-NV}
@@
 
-The $^{13}$C-NV$^-$ complex is formed by an NV center where one of the three carbon atoms neighboring the vacancy is a $^{13}$C isotope.
+The $^{13}$C-NV$^-$ complex is formed by an NV center where one of the three carbon atoms neighboring the vacancy (first shell carbon atom) is a $^{13}$C isotope.
 Such a complex is illustrated in Fig.
@@
 
-The $^{13}$C-NV$^-$ complex still behaves as a ``normal" NV center, except that its transition lines are shifted by the hypere-fin coupling.
-In particular its electronic spin is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states\footnote{This is confirmed by the presence of sideband in ODMR spectra around the main NV lines \citep{simanovskaia2013sidebands}.}.
-
-Since both NV$^-$ and $^{13}$C-NV$^-$ are \textit{a priori} equally polarized, then following the argument in \ref{Sec_CR} there should be no CR between the two spin families.
+The $^{13}$C-NV$^-$ complex still behaves as a ``normal" NV center, with a modified hyper-fine structure (there are now 18 possible spin states considering the electronic spin and the $^{13}$C and $^{14}$N nuclear spins).
+In particular, the electronic spin of $^{13}$C-NV$^-$ is still polarized in the $\ket{0}$ state and the $\ket{0}$ state is still brighter than the $\ket{\pm 1}$ states\footnote{This is confirmed by the presence of sideband in ODMR spectra around the main NV lines \citep{simanovskaia2013sidebands}.}.
+
+Since both NV$^-$ and $^{13}$C-NV$^-$ are \textit{a priori} equally polarized, then following the argument laid in \ref{Sec_CR} there should be no CR between the two spin families.
 The reason why those transitions are still visible is the presence of dark NV-centers (``fluctuators") which will be further explained in chapter 3.
@@
 where $\mathcal{H}_{NV}$ is the NV$^-$ spin Hamiltonian,
-$\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian: $$\mathcal{H}_{HF}= \hat{\mathbf{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\mathbf{I}}_C.$$
+$\mathcal{H}_{^{13}C}$ is the $^{13}$C nuclear spin Hamiltonian for a $1/2$ spin : $\mathcal{H}_{^{13}C}=\gamma_{n} B I_z$ where $\gamma_{n}=$10.7 MHz/T is the $^{13}$C gyromagnetic ratio, and $\mathcal{H}_{HF}$ is the hyper-fine interaction Hamiltonian: $$\mathcal{H}_{HF}= \hat{\vec{S}}_{NV} \cdot \mathcal{A} \cdot \hat{\vec{I}}_C.$$
 
@@
 \ref{121 vs 22 13C-NV}-b) and c) show the simulated transition frequencies for a $^{13}$C-NV$^-$ complex when B is scanned along the [100] or [111] axis, as well as the ``normal" NV transitions and the co-resonance between those two.
-When $\mathbf{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G.
-When $\mathbf{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
+When $\vec{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G.
+When $\vec{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
 This abundance of transitions for low ($<100\ \rm G$) magnetic field makes the experimental analysis quickly intractable in the second case.
 
-There is one last aspect to take into account regarding the field orientation choice which is the impact of the transverse magnetic field on the cross-relaxation contrast.
+There is one last aspect to take into consideration regarding the field orientation which is the impact of the transverse magnetic field on the cross-relaxation contrast.
 
@@
 \includegraphics[width=0.9\textwidth]{chapter 2/Figures/contrast_OMDR_111_Vs_100}
-\caption{Simulated ODMR contrast on all four classes of NV centers with a magnetic field a) along the [100] axis and b) along the [111] axis}
+\caption{Simulated ODMR contrast on all four classes of NV centers with a magnetic field a) along the [100] axis and b) along the [111] axis.}
 \label{121 vs 22 contrast}
@@
 
-These two conditions are fulfilled when there is no magnetic field or purely longitudinal magnetic field, but not anymore in the presence of transverse magnetic field which mixes the NV spin Hamiltonian eigenstates \citep{tetienne2012magnetic}.
-
-Fig.
-\ref{121 vs 22 contrast} shows the simulated relative contrast of a CR process \footnote{Due to the similarity between ODMR and CR, we assume here that the ODMR contrast is equal to the CR contrast} computed from the rates found in \citep{tetienne2012magnetic}, for a magnetic field aligned along either [100] or [111].
+These two conditions are no longer fulfilled in the presence of transverse magnetic field which mixes the NV spin Hamiltonian eigenstates \citep{tetienne2012magnetic}.
+
+Fig.
+\ref{121 vs 22 contrast} shows the simulated relative contrast of a CR process \footnote{Due to the similarity between ODMR and CR, we assume here that the ODMR contrast is equal to the CR contrast.} computed from the rates found in \citep{tetienne2012magnetic}, for a magnetic field aligned along either [100] or [111].
 When the NV center is aligned with the magnetic field, the contrast remains equal to 1 regardless of the magnetic field because there is no transverse field.
@@
 The [100] direction is therefore not well suited to observe CR at magnetic fields higher than a few hundreds gauss.
-Fortunately, the co-resonance between NV centers and VH$^-$, War1 or $^{13}$C-NV$^-$ all happen for $|\mathbf{B}| < 150 \ \rm G$, and can thus be detected with a [100] oriented magnetic field.
+Fortunately, the co-resonance between NV centers and VH$^-$, War1 or $^{13}$C-NV$^-$ all happen for $|\vec{B}| < 150 \ \rm G$, and can thus be detected with a [100] oriented magnetic field.
 
@@
 This section shows experimental observation of NV-dark spins CR.
-Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv,   hall2016detection, wickenbrock2016microwave,  wood2016wide,  alfasi2019detection, lazda2021cross} had been performed on the detection of dark spins through CR with NV centers, and to our knowledge, all this studies where done on the detection of P1 centers.
-
-We show an example of one this study to illustrate the principle of dark spin spectroscopy before showing our results on the detection of VH$^-$ and War1 in a CVD diamond.
+Previously to our work, several studies \citep{van1989cross, holliday1989optical, epstein2005anisotropic, armstrong2010nv, hall2016detection, wickenbrock2016microwave, wood2016wide, alfasi2019detection, lazda2021cross} had been performed on the detection of dark spins via NV cross-relaxations, and to our knowledge, all this studies where performed on the detection of P1 centers.
+
+We first show an example of one this study to illustrate the principle of dark spin spectroscopy before showing our results on the detection of VH$^-$ and War1 in a CVD diamond.
 
@@
 It shows the evolution of the NV PL with respect to a magnetic field scanned along the [111] crystalline axis.
-There are several information to gather from this graph: First, one observes a slowly decaying PL envelope, which stabilizes to a plateau for $B\sim 1000\ \rm G$.
+We can gather several information from this graph: First, we can observe a slowly decaying PL envelope, which stabilizes to a plateau for $B\sim 1000\ \rm G$.
 This change in PL is caused by the transverse magnetic field acting on the three NV classes that are not aligned with the magnetic field.
@@
 The first feature (Ns-NV) for $B\approx 512 \ \rm G$ is the signature of the CR between NV and P1 centers.
-Its relatively complicated shape comes from the hyper-fine coupling between the P1 electronic spin and the $^{14}$N nucleus \citep{lazda2021cross}.
-Excited state level anti-crossing (ESLAC) also occurs in this region, but its optical signature, if it exists, is expected to be much weaker than the NV-P1 CR \citep{zheng2017level}.
+Its shape comes from the hyper-fine coupling between the P1 electronic spin and the $^{14}$N nucleus \citep{lazda2021cross}.
+Excited state level anti-crossing (ESLAC) also occurs in this magnetic region, but its optical signature, if it exists, is expected to be much weaker than the NV-P1 CR \citep{zheng2017level}.
 
@@
 The orange line is a fit to the slowly decreasing envelope b) Subtraction of the PL signal by the envelope fit as a function of the magnetic field.
-The green lines and red and purple regions correspond to the predicted CR regions, as detailed in main text}
+The green lines and red and purple regions correspond to the predicted CR regions, as detailed in main text.}
 \label{CR VH exp}
@@
 Similarly to fig.
-\ref{CR P1 exp}, there is a slowly decaying envelope of the PL due to the transverse magnetic field \footnote{The plateau is not visible in this case as the scan stops at 150 G}.
+\ref{CR P1 exp}, there is a slowly decaying envelope of the PL due to the transverse magnetic field \footnote{The plateau is not visible in this case as the scan stops at 150 G.}.
 There also are three deviations from the smooth decreasing curve at $\sim$ 20 G, 55 G and 120 G.
@@
 \ref{CR VH exp}-b) shows the subtraction of the PL data by the envelope fit.
-We can clearly see three dips in PL around the values 20 G, 56 G and 122 G.
-We then computed the expected value of the magnetic field for the expected CR line thanks to the simulations shown in sec.
+We can see three salient dips in PL for the magnetic field values 20 G, 56 G and 122 G.
+We then computed the expected position of the different CR lines based on the simulations shown in sec.
 \ref{sec simu}.
-For the VH$^-$ and War1 lines, we took the ZFS $D$ values from \citep{cruddace2007magnetic} and included the error bars.
-For the $^{13}$C-NV$^-$ lines, the error bars given in \citep{simanovskaia2013sidebands} are much smaller and we only plotted the expected value.
+For the VH$^-$ and War1 lines, we took the ZFS $D$ values from \citep{cruddace2007magnetic} and included the provided error bars.
+For the $^{13}$C-NV$^-$ lines, the error bars given in \citep{simanovskaia2013sidebands} were much smaller and we only plotted the expected $B$ values.
 We can see that the dips in PL match relatively well the predicted CR lines of $^{13}$C-NV$^-$, VH$^-$ and War1 respectively.
@@
 It should be noted however that a single-photon counting APD was used for this acquisition, which reduces significantly the sensitivity because of the high shot noise \footnote{The sample produces $\sim$ 100 nW of PL, which means that only $\sim 10^{-5}$ of the total PL is actually measured by an SPC-APD with a max counting rate of $10^7$ photons/s.}.
-Using a photodiode and modulation of the magnetic field, we estimate that we could get similar signal to noise ratio within a few minutes.
+Using a photodiode and a modulation of the magnetic field, we estimate that we could get similar signal to noise ratio within a few minutes.
 
@@
 \ref{CR VH exp}.
-In order to get to the $D$ value of VH$^-$ and War1, we had to assume that the spin structure was that of a spin 1 similar to the NV center, with a g-factor close to 2.
+In order to get to the $D$ value of VH$^-$ and War1, we had to assume that the spin structure was that of a spin 1 similar to the NV center, with a g-factor close to -2.
 Both of this assumptions are confirmed by the EPR measurements.
-We do find that that both $D$ measurement concord within their respective error bars.
+We do find that both $D$ measurement concord within their respective error bars.
 
@@
 \includegraphics[width=.6\textwidth]{chapter 2/Figures/scan_100_Sumi4}
-\caption{Change in PL as a function of a magnetic field scanned along the [100] axis for sample SUMI-4}
+\caption{Change in PL as a function of a magnetic field scanned along the [100] axis for sample SUMI-4 (see sample details in appendix \ref{Appendix samples}).}
 \label{scan sumi 4}
@@
 Fig.
-\ref{scan sumi 4} for example shows a PL scan along the [100] axis for sample Sumi-4, an HPHT sample with [NV]$\approx 5\ \rm ppm$ and natural abundance of $^{13}$C \footnote{We could unfortunately not try a sample isotopically enriched in $^{12}$C since those samples are generally reserved to low-nitrogen concentration ($< 10\ \rm ppm$), which does not yield enough NV centers to perform CR spectroscopy.
-}.
+\ref{scan sumi 4} for example shows a PL scan along the [100] axis for sample Sumi-4, an HPHT sample with [NV]$>1 \rm ppm$ and natural abundance of $^{13}$C \footnote{We could unfortunately not try a sample isotopically enriched in $^{12}$C since those samples are generally reserved to low-nitrogen concentration ($< 10\ \rm ppm$), which does not yield enough NV centers to perform CR spectroscopy.}: only the 22 G line remain present compared to the CVD sample.
 
@@
 While it would be in theory possible to measure other properties of the dark spins Hamiltonian - such as the $g$-factor - by changing the angle of the magnetic field, this would here quickly become intractable due to the increase in co-resonance conditions and the relative closeness of each of these transitions.
-DEER might prove more useful in this regard as only one class of NV centers at a time is probed.
+DEER might prove more useful in this regard as only one class of NV centers at a time is being probed.
 
@@
 This procedure will be discussed in more details in the next chapter.
-We find here that the linewidth of VH$^-$ and War1 are respectively (half width at half maximum) $5.7\pm 0.6$ and $15.7\pm 1.6$ MHz \footnote{We follow here the pseudo-Voigt deconvolution method described in the following chapter.
-The linewidth $W=2.7$ MHz (half witdth at half maximum) of the NV centers was measured from ODMR (see Fig.
-\ref{ODMR 1 classe}).
-It includes all three hyper-fine transistions}.
-
+We find here that the linewidth of VH$^-$ and War1 are respectively (half width at half maximum) $5.7\pm 0.6$ and $15.7\pm 1.6$ MHz.
 \subsection{Dark spin concentration}
@@
 \begin{itemize}
-\item Since the change in $T_1$ is proportional to the $T_2^*$ of the probed spins (eq.
-\ref{delta gamma 1}), we should compensate that by integrating over the entire feature, instead of simply looking at the maximum of the line.
+\item Because the change in $T_1$ is inversely proportional to the $T_2^*$ of the probed spins (cf eq.
+(\ref{delta gamma 1})), we should compensate this effect by integrating over the entire area of the peak, instead of simply taking the maximum value of the line.
 \item We have to take into account the loss of contrast of the NV centers as the transverse magnetic field grows, as explained in sec. \ref{CR contrast}.
-\item The $^{13}$C-NV$^-$ are not ``dark spins", and its CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
+\item The $^{13}$C-NV$^-$ are not ``dark spins", and their CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
 We will see however in the next chapter that we can consider a fraction of the NV centers as dark spins, and that this fraction can be estimated.
 \citep{choi2017depolarization} found a ratio of $\sim 1/4$ dark NV centers for $\sim 3/4$ bright ones.
+\item Finally the lifetime of the dark spins may play a role in the NV depolarization rate. We will not consider this aspect here.
 \end{itemize}
@@
 An additional oscillating magnetic field along the same direction was used to perform the lock-in detection.
-The VH$^-$ and War1 lines are taken from \ref{CR VH exp}}
+The VH$^-$ and War1 lines are taken from \ref{CR VH exp}.}
 \label{chauffage CVD}
@@
 
-Given that the NV CR spectroscopy was both sensitive and relatively easy to setup, we decided in collaboration with the teams of Alexandre Tallaire at IRCP and Jocelyn Achard at LSPM to study the effect of annealing at high temperature on VH$^-$ and War1 \citep{ngambou2022improving}.
-
-Fig.
-\ref{chauffage CVD} shows the results of this experiment.
+We decided in collaboration with the teams of Alexandre Tallaire at IRCP and Jocelyn Achard at LSPM to investigate the effect of high temperature annealing on the stability of CVD defects, including VH$^-$ and War1 \citep{ngambou2022improving}.
+NV cross-relaxations proved to be a useful tool in this study.
+
+Fig.
+\ref{chauffage CVD} shows the results of the NV-CR experiments.
 This time, a modulation of the magnetic field was used and the PL coming from the photodiode was demodulated through a lock-in amplifier, which explains the derivative shape of the lines.
 The fact that the background is not perfectly flat still comes from the transverse magnetic field.
-The same lock-in parameter were used on both plots so that the amplitudes should in theory be comparable.
+The same lock-in parameter were used on both plots so that the lock-in signal could be comparable.
 
@@
 We can see that the VH$^-$ line in particular completely disappears on sample CVD-B, which we attribute to the additional annealing.
-This measurement also corroborate modifications in the infrared absorption spectrum, and an increase of the NV coherence time, associated with the annealing at higher temperature \citep{ngambou2022improving}.
+This measurement is corroborated by modifications in the infrared absorption spectrum, and an increase of the NV $T_2^*$ after the high temperature annealing \citep{ngambou2022improving}.
 
 The War1 line on the other hand is still present after the high temperature annealings, albeit slightly smaller.
-This information can give us some insights on the chemistry of the War1 defect.
+This measurement could give some insights on the chemistry of the War1 defect.
 %Et en vrai je crois que c'était deja vu dans la thèse de Cruddace
@@
 \includegraphics[width=.7\textwidth]{chapter 2/Figures/VH_HPHT_chauffage}
-\caption{Demodulated PL on samples SBST-C and SBST-B as a function of a magnetic field along the [100] axis. The plot of SBST-C is offset for clarity}
+\caption{Demodulated PL on samples SBST-C and SBST-B as a function of a magnetic field along the [100] axis. The plot of SBST-C is offset for clarity.}
 \label{chauffage HPHT}
@@
 
-We first suspected that some of the War1 defects had been migrating from the CVD layer to the HPHT substrate with the high-temperature annealing, but not only is it very unlikely with the annealing time (1 hour) and substrate thickness (several mm), We also could not find a dependence on the War1 concentration with the depth in the substrate, whereas a diffusion would have given a higher concentration near the CVD layer.
-The War1 defects seem to have been created by the annealing.
-
-These results suggest that, while some defects are removed by the high temperature annealing, such as the VH$^-$ and several unidentified defects in the HPHT substrate, some seem to be created by the annealing process, such as the War1 or the one associated with the line at $B = 48\ \rm G$ in the HPHT substrate.
+We first suspected that some of the War1 defects had been migrating from the CVD layer to the HPHT substrate with the high-temperature annealing, but not only is it very unlikely with the annealing time (1 hour) and substrate thickness (several mm), we also could not find a dependence on the War1 concentration with the depth in the substrate.
+The War1 defects seem to have been created in situ by the annealing.
+
+These results suggest that, while many defects, including VH$^-$, are removed by the high temperature annealing some other, such as War1, seem to be created by the annealing process.
 
@@
 This detection can be significantly more sensitive than the usual EPR spectroscopy, and considerably cheaper to set up.
-This technique however is only applicable to a few defects in NV-rich diamond, meaning that its main application would probably be to characterize the environment of samples with high NV density.
-
-We have also seen the importance of the magnetic field direction, and the underused aspect of the [100] magnetic field scan which provides a significantly less noisy signal than the [111] scans for weak magnetic field.
+This technique however is only applicable to a few defects in NV-rich diamond, meaning that its main application would likely be to characterize the environment of samples with high NV density.
+
+We have also seen the importance of the magnetic field direction, and the potential to use [100] magnetic scans for CR detection at weak magnetic fields.
 
@@
 
-There are still a large number of lines in CR spectra that have not been identified.
-While not every line necessarily correspond to a new defect - for example \citep{wunderlich2021magnetic} recently found lines corresponding to co-resonance between the $\ket{0} \to \ket{-1}$ transition of an NV center and the $\ket{-1} \to \ket{+1}$ of another NV center - the more likely hypothesis for each of these unidentified lines is still the presence of a new defect.
+There are still a large number of lines in CR spectra that have not been identified, and we could postulate that many of these lines correspond to cross-relaxations between an NV center and a particular dark spin.
+%While not every line necessarily correspond to a new defect - for example \citep{wunderlich2021magnetic} recently found lines corresponding to co-resonance between the $\ket{0} \to \ket{-1}$ transition of an NV center and the $\ket{-1} \to \ket{+1}$ of another NV center - the more likely hypothesis for each of these unidentified lines is still the presence of a new defect.
 
@@
 This value is within one order of magnitude from ODMR contrast with NV ensembles (typically $<5 \%$), which means that it would be reasonable to probe the dynamics ($T_1$, $T_2$) of the VH$^-$ spin.
-The main issue comes from the fact that in order to measure the VH$^-$ spin states, it needs to be resonant with the NV center.
-Therefore, resonant microwave pulses with VH$^-$ will also be resonant with the NV centers, which will create an unwanted optical signal.
+The main issue comes from the fact that in order to measure the VH$^-$ spin needs to be resonant with the NV center to be probed.
+Therefore, resonant microwave pulses with VH$^-$ will also be resonant with the NV centers, which will create an unwanted optical response.
 
@@
 \section{Experimental observation of NV-NV CR}
-Let us start with the experimental proof of the presence of NV-NV CR in the samples with dense NV ensembles.
+Let us start with the experimental proof of the presence of NV-NV cross-relaxations in the samples with dense NV ensembles.
 
@@
 Two equivalent centers are NV centers with the same spin Hamiltonian (cf eq.
-(\ref{eq.
-spin elec})), which implies that they both see the same transverse and longitudinal magnetic field.
+(\ref{NV spin Hamiltonian basic})), which implies that they both see the same transverse and longitudinal magnetic field.
 In contrast, non-equivalent centers are center with different single-spin Hamiltonian, which may result in different eigenstates and spin polarization for the two spins.
@@
 \includegraphics[width=\textwidth]{chapter 3/Figures/NV-NV_non_equivalent}
-\caption{a) Illustration of the four possible direction of the the NV axis (four classes of NV centers) with a magnetic field along class 1 b) Representation of the polarization of two non equivalent spins, one aligned with the magnetic field (well polarized) and one misaligned with the magnetic field (unpolarized).
-c) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
+\caption{a) Illustration of the four possible direction of the NV axis (four classes of NV centers) with a magnetic field along class 1 b) Representation of the polarization of two non equivalent spins, one aligned with the magnetic field (well polarized) and one misaligned with the magnetic field (unpolarized).
+c) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\vec{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
 d) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
@@
 \ref{non-equivalent NV-NV}-c) shows the transition frequencies of spin from class 1 and a spin from class 2-4 as a function of the magnetic field amplitude.
-Unintuitively, there is a co-resonance between the two $\ket{0} \to \ket{+1}$ transitions for $\abs{\mathbf{B}}=592\ \rm G$.
+Unintuitively, there is a co-resonance between the two $\ket{0} \to \ket{+1}$ transitions for $\abs{\vec{B}}=592\ \rm G$.
 For this particular value of the magnetic field, cross-relaxation between the NV centers from class 1 and those from classes 2-4 is possible.
@@
 \ref{non-equivalent NV-NV}-d), which was presented in the last chapter, shows the experimental value of the total NV PL as a function of the magnetic field amplitude, in a scenario similar to the one we just described.
-We can indeed observe the signature of NV-NV CR in the presence of a PL dip at $\abs{\mathbf{B}}=592\ \rm G$.
+We can indeed observe the signature of NV-NV CR in the presence of a PL dip at $\abs{\vec{B}}=592\ \rm G$.
 
@@
 \ref{equivalent NV-NV}-a) illustrates the two possible scenarios for equivalent NV centers: either two NVs from the same class, or two NVs from different classes with the same projection of the magnetic field upon their respective axis (i.e.
-$\mathbf{B}$ belongs to the plane of mirror symmetry of the two spins).
+$\vec{B}$ belongs to the plane of mirror symmetry of the two spins).
 
@@
 \caption{Representation of a ``normal" NV center with spin depolarization rate $\Gamma_1 \ll \Gamma_{\rm las}$ and a fluctuator NV with spin depolarization rate $\gamma_f \gg \Gamma_{\rm las}$.
-Red circles represent the spin population in the steady state of each spin}
+Red circles represent the spin population in the steady state of each spin.}
 \label{NV vs fluct}
@@
 \begin{equation}
-\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right)^2 \frac{\eta^2}{\gamma_f}.
+\gamma_s(\vec{r})=\left(\frac{J_0}{r^3}\right)^2 \frac{\eta^2}{\gamma_f}.
 \end{equation}
@@
 \includegraphics[width=.4\textwidth]{chapter 3/Figures/NV_dans_bain_fluct}
-\caption{Schematics of an NV center in a fluctuator bath. The $\mathbf{r}_i$ vector represent the relative of position of fluctuator $i$ with respect to the NV center.}
+\caption{Schematics of an NV center in a fluctuator bath. The $\vec{r}_i$ vector represent the relative of position of fluctuator $i$ with respect to the NV center.}
 \label{NV + bain fluct}
@@
 We consider the situation pictured in Fig.
-\ref{NV + bain fluct}: the NV center is at the center of the frame and the position of each fluctuator is labeled $\mathbf{r}_i$.
+\ref{NV + bain fluct}: the NV center is at the center of the frame and the position of each fluctuator is labeled $\vec{r}_i$.
 We consider that each fluctuator is an independent depolarization source, therefore the depolarization on a single NV center reads:
 \begin{equation}
-\gamma=\sum_i \gamma_s(\mathbf{r}_i).
+\gamma=\sum_i \gamma_s(\vec{r}_i).
 \end{equation}
@@
 \begin{equation}
-\rho(\gamma)=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\mathbf{r}_i) - \gamma \right).
+\rho(\gamma)=\int d\{r_i\} \rho(\{r_i\})\, \delta \left( \sum_i \gamma_s(\vec{r}_i) - \gamma \right).
 \end{equation}
@@
 
-We will start by separating the factor $\eta$ given in eq. (\ref{eq eta}) into a a geometric and a spectral part:
+We will start by separating the factor $\eta$ given in eq. (\ref{eq eta}) into a geometric and a spectral part:
 \begin{align}
@@
 In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f \gg 1/T_2^*$) in eq.
-\ref{eq 1/T}, we find that:
+(\ref{eq 1/T}), we find that:
 \begin{equation}
@@
 Eq.
-\ref{eq 1/T avec T2*} was obtained via several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
+(\ref{eq 1/T avec T2*}) was obtained via several approximations: the inhomogeneous distribution of the NV centers is generally not Lorentzian, and $\overline{R^2} \neq (\bar{R})^2$ in the general case.
 Nevertheless, it gives us a general idea of the evolution of $1/T$ with $T_2^*$, $\omega_{NV}^0$ and $\omega_{f}^0$, which will be crucial to compare the predictions of the model with experimental data.
@@
 \ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
-The values of $T_1^{\rm dd}$ obtained with this method have therefore to be be compared to values from the same method and the same fixed value of $T_1^{\rm ph}$.
+The values of $T_1^{\rm dd}$ obtained with this method have therefore to be compared to values from the same method and the same fixed value of $T_1^{\rm ph}$.
 
@@
 b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.8 \pm 0.2\ \rm MHz$.
-The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed one to the same line scaled up by a factor of 2}
+The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed line correspond to the same spectrm scaled up by a factor of 2.}
 \label{fluct linewidth}
@@
 If we assume that both NV classes spectral response are Lorentzian of same width $\sigma$, then the spectral overlap between the two classes is also a Lorentzian of width $2 \sigma$ \footnote{For other line profile, the overlap would be smaller than $2 \sigma$.
-For instances, the overlap of two Gaussians of width $\sigma$ is a Gaussian of width $\sqrt{2} \sigma$}.
+For instances, the overlap of two Gaussians of width $\sigma$ is a Gaussian of width $\sqrt{2} \sigma$.}.
 Fig.
@@
 Green curve: deconvolution of the red curve by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters)}
+All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters).}
 \label{deconvolution}
@@
 Red numbers represent the number of resonant classes for each ODMR line.
-The cubes represent the diamond unit cell, and the red planes/arrow the possible orientations of the magnetic field}
+The cubes represent the diamond unit cell, and the red planes/arrow the possible orientations of the magnetic field.}
 \label{ODMR_geometries}
@@
 \begin{itemize}
-\item $\mathbf{B} \in \{110\}$.
+\item $\vec{B} \in \{110\}$.
 We refer by $\{110\}$ as any plane orthogonal to a $\langle 110 \rangle$ or equivalent direction.
 In this magnetic configuration, two classes are at resonance and the two other ones are spectrally isolated.
-\item $\mathbf{B} \in \{100\}$. In this magnetic configuration, the four classes form two pairs of co-resonance.
-\item $\mathbf{B} \parallel \langle 111 \rangle$.
+\item $\vec{B} \in \{100\}$. In this magnetic configuration, the four classes form two pairs of co-resonance.
+\item $\vec{B} \parallel \langle 111 \rangle$.
 In this configuration discussed in the last chapter, one class is aligned with the magnetic field and is spectrally isolated while the three other form a resonant triplet.
-\item $\mathbf{B} \parallel \langle 100 \rangle$. In this configuration also discussed in the last chapter, the four classes are resonant.
+\item $\vec{B} \parallel \langle 100 \rangle$. In this configuration also discussed in the last chapter, the four classes are resonant.
 \end{itemize}
-We should add to this list the case $\mathbf{B}=0$ where all four classes are also resonant, and any other possibility (which we will label random $\mathbf{B}$) where all four classes are spectrally separated.
+We should add to this list the case $\vec{B}=0$ where all four classes are also resonant, and any other possibility (which we will label random $\vec{B}$) where all four classes are spectrally separated.
 
@@
 \ref{Carte} shows a map where we monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer.
-We have also reported the locus of $\mathbf{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].
-
-We know that when $\mathbf{B}$ belongs to any of these four planes, there is a co-resonance between at least two NV classes.
+We have also reported the locus of $\vec{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].
+
+We know that when $\vec{B}$ belongs to any of these four planes, there is a co-resonance between at least two NV classes.
 This co-resonance leads to an increase in NV-NV CR and a faster depolarization of the spins involved, which ultimately leads to a drop in PL.
 We can indeed see in fig.
-\ref{Carte} that there is clear correlation between the loci of $\mathbf{B}$ and drops in PL.
-We can also see that the PL is at its lowest when $\mathbf{B}\parallel [001]$, which corresponds to a four class degeneracy.
+\ref{Carte} that there is clear correlation between the loci of $\vec{B}$ and drops in PL.
+We can also see that the PL is at its lowest when $\vec{B}\parallel [001]$, which corresponds to a four class degeneracy.
 
@@
 \toprule
-$\Gamma_1^{\rm dd} (\mathbf{B})$ &  Theory & Experimental \\
+$\Gamma_1^{\rm dd} (\vec{B})$ & Theory & Experimental \\
 \midrule
-random $\mathbf{B}$ (1 class) & $\Gamma_0^{\rm th}$ & $1.53\pm 0.04$ ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & $5.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & $4.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & $11.6 \pm 0.4$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & $14.1 \pm 0.5$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & $19.9 \pm 0.8$ $\Gamma_0^{\rm exp}$ \\
+random $\vec{B}$ (1 class) & $\Gamma_0^{\rm th}$ & $1.53\pm 0.04$ ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
+$\vec{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & $5.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & $4.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & $11.6 \pm 0.4$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & $14.1 \pm 0.5$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & $19.9 \pm 0.8$ $\Gamma_0^{\rm exp}$ \\
 \bottomrule
@@
 
-($^*$) The case $\mathbf{B}=0$ will be detailed in the next chapter. The theoretical value is a lower bond.
+($^*$) The case $\vec{B}=0$ will be detailed in the next chapter. The theoretical value is a lower bond.
 \end{table}
@@
 The fluctuator model allows quantitative predictions of the change in the spin lifetime, thanks to eq.
-\ref{eq 1/T}.
+(\ref{eq 1/T}).
 There are however several unknowns in this equation, such as the fluctuator density and lifetime which can be hard to estimate.
 It is however relatively easy to compare the predictions of eq.
-\ref{eq 1/T} for various magnetic field configurations.
+(\ref{eq 1/T}) for various magnetic field configurations.
 
@@
 Doing so allows us to normalize $n_f$ and $\gamma_f$ in eq.
-\ref{eq 1/T} and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
+(\ref{eq 1/T}) and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
 
@@
 Using for example $T_1^{\rm ph}=2.5\ \rm ms$ gave results closer to the theoretical ones.
-Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\mathbf{B} \in \{110\}$ or $\mathbf{B} \in \{100\}$ configuration) yielded an increased depolarization rate by a factor $\sim 4$, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one\footnote{We can also see in Fig.
-\ref{fluct linewidth} that a $\mathbf{B} \in \{110\}$ configuration leads to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, which is 4 times smaller than the predicted factor of 10.}.
+Nevertheless, the fluctuator model generally tends to overestimate the depolarization rates.The authors of \citep{choi2017depolarization} for instance found that a 2-class degeneracy (the authors did not precise whether this was a $\vec{B} \in \{110\}$ or $\vec{B} \in \{100\}$ configuration) yielded an increased depolarization rate by a factor $\sim 4$, a value comparable to the one we found in table \ref{T1 champ mag}, and two times smaller than the predicted one\footnote{We can also see in Fig.
+\ref{fluct linewidth} that a $\vec{B} \in \{110\}$ configuration leads to a $\sim 2.5$ times increase in $\Gamma_1^{\rm dd}$, which is 4 times smaller than the predicted factor of 10.}.
 
@@
 \caption{$T_1$ measurements on sample ADM-15-4, fitted with pure exponential or stretched exponential.
-a) $T_1$ of a single isolated class when $\mathbf{B} \neq 0$.
+a) $T_1$ of a single isolated class when $\vec{B} \neq 0$.
 The optimal fit parameters are respectively $T_1^{\rm exp}=1.43\ \rm ms$ and $T_1^{\rm stretch}=0.56\ \rm ms$.
-b) $T_1$ of the four degenerate classes when $\mathbf{B}=0$.
+b) $T_1$ of the four degenerate classes when $\vec{B}=0$.
 The optimal fit parameters are respectively $T_1^{\rm exp}=0.15\ \rm ms$ and $T_1^{\rm stretch}=0.05\ \rm ms$.}
@@
 Fig.
-\ref{anormal T1} shows $T_1$ measurement on sample ADM-15-4, either on a single isolated class ,where the dipole-dipole interactions are at their weakest, or on all four resonant classes at $\mathbf{B}=0$, where the interactions are at their strongest.
-
-Although the spin lifetime still gets $\sim$ 10 times shorter in the $\mathbf{B}=0$ case, indicating that dipole-dipole interaction still plays an important role in the spin relaxation, this time the lifetime profile does not converge to a purely stretched exponential.
+\ref{anormal T1} shows $T_1$ measurement on sample ADM-15-4, either on a single isolated class ,where the dipole-dipole interactions are at their weakest, or on all four resonant classes at $\vec{B}=0$, where the interactions are at their strongest.
+
+Although the spin lifetime still gets $\sim$ 10 times shorter in the $\vec{B}=0$ case, indicating that dipole-dipole interaction still plays an important role in the spin relaxation, this time the lifetime profile does not converge to a purely stretched exponential.
 In fact it seems to be the opposite: the lifetime profile is closer to a pure exponential for the shorter $T_1$ (especially for the very short times), and closer to a stretched exponential for the longer $T_1$, which is in complete contradiction with the fluctuator model and the previous observations in Fig.
@@
 In the sample studied on Fig.
-\ref{fluct linewidth}, we found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough our sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 15 ppm$ sample used in \citep{choi2017depolarization} \footnote{Although the initial paper claimed a NV concentration $[\rm NV] \sim 45 ppm$, a more recent paper re-established the concentration at $\sim 15$ ppm \citep{zhou2020quantum}}.
+\ref{fluct linewidth}, we found a value for $\gamma_f$ surprisingly similar to the one found in \citep{choi2017depolarization}, even tough our sample only had $[\rm NV] \sim 3 ppm$ compared to the $[\rm NV] \sim 15 ppm$ sample used in \citep{choi2017depolarization} \footnote{Although the initial paper claimed a NV concentration $[\rm NV] \sim 45 ppm$, a more recent paper re-established the concentration at $\sim 15$ ppm \citep{zhou2020quantum}.}.
 If $\gamma_f$ was found to be somewhat of a constant among all samples showing NV-NV CR, this would give us a lot of information on the potential nature of the fluctuators.
@@
 \caption{Monte-Carlo simulations of inhomogeneous zero field ODMR for different noise sources.
-a) Magnetic noise: each of the $\mathbf{B}$ components is sampled over a Gaussian of deviation $\sigma=2\ \rm G$.
+a) Magnetic noise: each of the $\vec{B}$ components is sampled over a Gaussian of deviation $\sigma=2\ \rm G$.
 b) Strain noise: each of the components of the strain tensor $\bar{\bar{\varepsilon}}$ is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$.
-c) Electric noise: each of the $\mathbf{E}$ components is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
-d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
+c) Electric noise: each of the $\vec{E}$ components is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
+d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2.}
 \label{simus ESR}
@@
 \mathcal{H}=\begin{pmatrix}
-D-\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}}) & f_\perp(\mathbf{E}) + g_\perp(\bar{\bar{\varepsilon}})\\
-f^*_\perp(\mathbf{E}) + g^*_\perp(\bar{\bar{\varepsilon}})&D+\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}})
+D-\gamma_e B_\parallel + f_\parallel(\vec{E}) + g_\parallel(\bar{\bar{\varepsilon}}) & f_\perp(\vec{E}) + g_\perp(\bar{\bar{\varepsilon}})\\
+f^*_\perp(\vec{E}) + g^*_\perp(\bar{\bar{\varepsilon}})&D+\gamma_e B_\parallel + f_\parallel(\vec{E}) + g_\parallel(\bar{\bar{\varepsilon}})
 \end{pmatrix},
@@
 \end{equation}
-where $B_\parallel$ is the component of the magnetic field along the NV axis, and $f_\parallel, f_\perp, g_\perp$, and $g_\parallel$ are functions of the electric field $\mathbf{E}$ and the strain tensor $\bar{\bar{\varepsilon}}$, whose expressions are:
+where $B_\parallel$ is the component of the magnetic field along the NV axis, and $f_\parallel, f_\perp, g_\perp$, and $g_\parallel$ are functions of the electric field $\vec{E}$ and the strain tensor $\bar{\bar{\varepsilon}}$, whose expressions are:
 
 \begin{align}
-f_\parallel(\mathbf{E})&=d_\parallel E_z, \\
-f_\perp(\mathbf{E})&=d_\perp ( E_x + i E_y), \\
+f_\parallel(\vec{E})&=d_\parallel E_z, \\
+f_\perp(\vec{E})&=d_\perp ( E_x + i E_y), \\
 g_\parallel(\bar{\bar{\varepsilon}})&= h_{41}(\varepsilon_{xx}+\varepsilon_{yy})+h_{43} \varepsilon_{zz}, \\
@@
 The eigenvalues of the spin Hamiltonian given in eq.
-\ref{Hamiltonien pm1} are then computed and the process is repeated $10^6$ times to create the histograms shown in Fig.
+(\ref{Hamiltonien pm1}) are then computed and the process is repeated $10^6$ times to create the histograms shown in Fig.
 \ref{simus ESR}.
@@
 We now turn to the case of a magnetic field that is orthogonal to the NV axis, i.e.
-$\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
+$\vec{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
 In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.
@@
 The linewidths of the spin transition \footnote{We consider here the full linewidth of the electronic levels, i.e.
-including all three hyper fine levels} for both samples are reported in the Table \ref{ESR for T2*}-d).
+including all three hyper fine levels.} for both samples are reported in the Table \ref{ESR for T2*}-d).
 In both cases, the linewidths are smaller in the $B=0$ and $B \perp NV$ cases than in the $B \parallel NV$ case, mainly due to the modification of the hyper-fine coupling.
@@
 \includegraphics[width=\textwidth]{chapter 4/Figures/Shema_summary_theory_2}
-\caption{Summary of the different depolarization effects in low magnetic field and their predicted depolarization rate. The ODMR spectra were taken on sample ADM-150-1}
+\caption{Summary of the different depolarization effects in low magnetic field and their predicted depolarization rate. The ODMR spectra were taken on sample ADM-150-1.}
 \label{summary_theory}
@@
 Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 43$.
-This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.
+This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\vec{B} \parallel [100]$.
 
@@
 This is $2\gamma_f+\Gamma_{NV}+\Gamma_f$ in eq.
-(\ref{eq 1/T avec T2*})} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+(\ref{eq 1/T avec T2*}).} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
 \ref{fluct linewidth}, as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states.
@@
 
-On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\mathbf{B}\parallel [100]$) and the influence of the laser polarization.
+On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\vec{B}\parallel [100]$) and the influence of the laser polarization.
 Following further experiments detailed in appendix \ref{Appendix autres causes}, we have determined that these two effects were negligible compared to the four ones we identified before.
@@
 \begin{itemize}
-\item When $\mathbf{B}=0$.
-\item When $\mathbf{B}\parallel [100]$ (cf. Fig. \ref{ODMR_geometries}).
+\item When $\vec{B}=0$.
+\item When $\vec{B}\parallel [100]$ (cf. Fig. \ref{ODMR_geometries}).
 \end{itemize}
@@
 \includegraphics[width=0.9\textwidth]{chapter 4/Figures/scan_1x1x1x1}
-\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction.
+\caption{Low field depolarization on sample ADM-150-1 for $\vec{B}$ in an arbitrary direction.
 a) ODMR spectrum for B=0.
@@
 A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in Fig.
-\ref{various_fit_formulae} with $T_1^{\rm ph}=5\ \rm ms$}
+\ref{various_fit_formulae} with $T_1^{\rm ph}=5\ \rm ms$.}
 \label{scan 1x1x1x1}
@@
 Fig.
-\ref{scan 1x1x1x1}-a) and b) shows ODMR spectrum at the starting point ($|\mathbf{B}|=0$) and end point ($|\mathbf{B}|\approx 60\ \rm G$) of the magnetic field scan.
+\ref{scan 1x1x1x1}-a) and b) shows ODMR spectrum at the starting point ($|\vec{B}|=0$) and end point ($|\vec{B}|\approx 60\ \rm G$) of the magnetic field scan.
 The magnetic field orientation was chosen in order to maximize the splitting of the four classes.
@@
 We can observe a decrease of $T_1^{\rm dd}$ by a factor of $\sim 10$ in zero magnetic field compared to non-zero magnetic field, with a half-width at half maximum of $\sim 10\ \rm G$.
-Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\mathbf{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$.
+Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\vec{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$.
 This is because this drop of PL is associated with transverse magnetic field state mixing, and not with spin depolarization.}.
@@
 
-\subsubsection{Magnetic scan with $\mathbf{B} \parallel [100]$}
+\subsubsection{Magnetic scan with $\vec{B} \parallel [100]$}
 
@@
 \includegraphics[width=0.9\textwidth]{chapter 4/Figures/scan_100}
-\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis.}
+\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\vec{B}$ along the [100] axis.}
 \label{scan 100}
@@
 We unfortunately were not able to discriminate the change of eigenbasis from the $T_2^* increase$, which would likely require a statistical study over different samples with different $T_2^*$ values.
-We were however able to discriminate the double flip contribution from the two other ones by scanning the magnetic field in a direction orthogonal to one NV class ($\mathbf{B} \perp [111]$).
+We were however able to discriminate the double flip contribution from the two other ones by scanning the magnetic field in a direction orthogonal to one NV class ($\vec{B} \perp [111]$).
 
@@
 The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$.
-(b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
+(b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve).}
 \label{eigenstates transverse field}
@@
 
-The idea behind the experiment is that the Hamiltonian eigenbasis is the same (or close to be the same) when $\mathbf{B}=0$ than it is when $\mathbf{B} \perp [111]$, as described in sec.
+The idea behind the experiment is that the Hamiltonian eigenbasis is the same (or close to be the same) when $\vec{B}=0$ than it is when $\vec{B} \perp [111]$, as described in sec.
 \ref{sec B transverse}.
-For simplicity, we will assume here that both the magnetic field and the transverse electric field lies along the NV $x$ axis : $\mathbf{B}=B_x \hat{e}_x$ and $\mathbf{E}_\perp =E_x \hat{e}_x$
+For simplicity, we will assume here that both the magnetic field and the transverse electric field lies along the NV $x$ axis : $\vec{B}=B_x \hat{e}_x$ and $\vec{E}_\perp =E_x \hat{e}_x$
 
@@
 \caption{Experimental data for the depolarization under purely transverse magnetic field on sample ADM-150-2.
-a) ODMR spectrum for $|\mathbf{B}|=20\ \rm G$.
+a) ODMR spectrum for $|\vec{B}|=20\ \rm G$.
 The transitions of the class orthogonal to the magnetic field are labeled $\nu_+$ and $\nu_-$.
@@
 We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$.
-We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$}
+We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$.}
 \label{champ tranverse exp}
@@
 \ref{champ tranverse exp}-a) shows an ODMR spectrum for $B=20\ \rm G$ with the two central lines corresponding to the $\ket{0} \to \ket{+}$ and $\ket{0} \to \ket{-}$ transitions of the class orthogonal to the magnetic field.
-We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\mathbf{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
-\ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\mathbf{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}
+We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\vec{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
+\ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\vec{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}.
 
@@
 In region A, $1/T_1^{\rm dd}$ decreases with the magnetic field.
-The difference in the depolarizing rate between the $\mathbf{B} \perp [111]$ case and the $\mathbf{B} \parallel [111]$ case comes from the double flips, the eigenbasis change and the $T_2^*$ increase.
+The difference in the depolarizing rate between the $\vec{B} \perp [111]$ case and the $\vec{B} \parallel [111]$ case comes from the double flips, the eigenbasis change and the $T_2^*$ increase.
 We measure an increase by a factor $\sim 3$ for B=20 G compared to the baseline.
@@
 We attribute this plateau to the fact that the double flips are completely quenched, which is coherent with the fact that $\Delta \nu \geq 30\ \rm MHz$.
-The remaining difference between the $\mathbf{B} \perp [111]$ and $\mathbf{B} \parallel [111]$ cases therefore comes from the eigenbasis change and the $T_2^*$ increase.
+The remaining difference between the $\vec{B} \perp [111]$ and $\vec{B} \parallel [111]$ cases therefore comes from the eigenbasis change and the $T_2^*$ increase.
 This increase amount for $\sim$ 50\% of the relaxation rate baseline.
@@
 
-Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies.
+Orientation-free magnetometry can also be performed for noise sensing, provided that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies.
 The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the Larmor frequency), or via the PL \citep{finco2021imaging}.
@@
 b) example of LIA signal on sample ADM-15-4 when scanning the magnetic field $B_0$ between -150 and 150 G.
-c) PL signal corresponding to the LIA signal}
+c) PL signal corresponding to the LIA signal.}
 \label{setup magneto}
@@
 We will now study the dependence of the sensitivity with respect to the magnetic field angle.
-In particular, we will compare the sensitivity for two scenarios: the case $\mathbf{B}\parallel [100]$and the four classes remains degenerate when $\mathbf{B}$ increases, and the case $\mathbf{B}\nparallel [100]$ where there is a lift of degeneracy between the four classes.
+In particular, we will compare the sensitivity for two scenarios: the case $\vec{B}\parallel [100]$and the four classes remains degenerate when $\vec{B}$ increases, and the case $\vec{B}\nparallel [100]$ where there is a lift of degeneracy between the four classes.
 We previously saw in Fig.
-\ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\mathbf{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.
+\ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\vec{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.
 
@@
 
-The fact that the sensitivity is similar when $\mathbf{B}\parallel [100]$ or $\mathbf{B}\nparallel [100]$ means that the degeneracy between the four classes plays a lower role in the sensitivity than the three other mechanisms detailed in sec.
+The fact that the sensitivity is similar when $\vec{B}\parallel [100]$ or $\vec{B}\nparallel [100]$ means that the degeneracy between the four classes plays a lower role in the sensitivity than the three other mechanisms detailed in sec.
 \ref{sec causes zero field}.
@@
 The noise value is computed by averaging the variance of the temporal trace over each time duration $T$.
-A curve $y=0.5\, T^{1/2}$ is added for comparison}
+A curve $y=0.5\, T^{1/2}$ is added for comparison.}
 \label{temporal 0B}
@@
 \centering
-\caption{Comparison of the volume-normalized sensitivity of the best AC and DC protocols with LFDM}
+\caption{Comparison of the volume-normalized sensitivity of the best AC and DC protocols with LFDM.}
  \label{table sensi volumiques}
@@
 The orange curve is shifted by $\approx 100\ \rm G$ to sit on the same scale as the blue curve.
-d) Estimated SNR fo a signal of 100 nT for $B_0 \approx 0$ and $B_0 \approx 100\ \rm G$ as a function of the total acquisition time}
+d) Estimated SNR for a signal of 100 nT for $B_0 \approx 0$ and $B_0 \approx 100\ \rm G$ as a function of the total acquisition time.}
 \label{comparaison 0B MW}
@@
 \includegraphics[width=\textwidth]{chapter 4/Figures/Applications}
-\caption{Schematics of two potential application for LFDM. The two application are detailed in the main text}
+\caption{Schematics of two potential application for LFDM. The two application are detailed in the main text.}
 \label{applications}
@@
 
-Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always keep a sensor within $\sim$ 1 $\mu$m of the sample.
+Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always keep a sensor within $\sim$ 1 $\mu$m of the sample.
 Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.
@@
 Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations ,does not require a homogeneous magnetic field or microwave field along the sample, and can be used with polycrystalline samples which are easier to produce in large size.
-A recent paper demonstrates an energy-efficient and cost-effective NV magnetometer which uses sunlight as the the optical excitation source, and exploits the transverse field dependence of the NV PL to measure the magnetic field\citep{zhu2022sunlight}.
+A recent paper demonstrates an energy-efficient and cost-effective NV magnetometer which uses sunlight as the optical excitation source, and exploits the transverse field dependence of the NV PL to measure the magnetic field\citep{zhu2022sunlight}.
 The use of LFDM instead of transverse field could improve the sensitivity of this magnetometer by an order of magnitude.
@@
 \subsection{Other HPHT samples}
-\subsubsection{Sumi-2}
-Sample Sumi-2 is a type Ib HPHT sample bought from Sumitomo and irradiated with a 2 MeV electron beam and a fluence of $7\cdot 10^{18}\ \rm cm^{-2}$ which were further annealed at 800 $^\circ$C for two hours.
+\subsubsection{Sumi-2 and Sumi-4}
+Sample Sumi-2 is a type Ib HPHT sample bought from Sumitomo and irradiated with a 2 MeV electron beam and a fluence of $7\cdot 10^{18}\ \rm cm^{-2}$ for Sumi-2 and $71.5\cdot 10^{19}\ \rm cm^{-2}$ for Sumi-4.
+The two samples were further annealed at 800 $^\circ$C for two hours.
 The NV and N concentration were not measured on this sample.
@@
 \includegraphics[width=.7\textwidth]{chapter 3/Figures/Shema_annexe}
-\caption{Schematics of the ``normal" NV center NV$_s$ and the fluctuator NV$_f$ in and their respective basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$), as well as the relative position $\mathbf{r}(r,\theta,\phi)$ between the two spins.}
+\caption{Schematics of the ``normal" NV center NV$_s$ and the fluctuator NV$_f$ in and their respective basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$), as well as the relative position $\vec{r}(r,\theta,\phi)$ between the two spins.}
 \label{shema spins annexe}
@@
 \begin{equation}
-\mathbf{S}_i=S_x \hat x_i + S_y \hat{y}_i + S_z \hat{z}_i,
+\vec{S}_i=S_x \hat x_i + S_y \hat{y}_i + S_z \hat{z}_i,
 \end{equation}
@@
 
-The relative position between the two spins is noted with the vector $\mathbf{r}=r\hat{u}$.
+The relative position between the two spins is noted with the vector $\vec{r}=r\hat{u}$.
 The vector's spherical coordinate are given in the Cartesian basis of the NV center ($\hat x_s,\hat y_s,\hat z_s$).
@@
 The angle $\psi$ between the two spins $z$ axes is defined by the crystal lattice and can take three values within [0,$\pi$] which are: 0 (same class), $\cos^{-1}(1/3) \approx 70.5\ ^\circ$ and $\cos^{-1}(-1/3) \approx 109.5\ ^\circ$.
-Two distinct classes can either have a relative angle $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$ depending on the external magnetic field: we choose by definition that $\mathbf{B}\cdot\hat{z}_i >0$, which imposes the orientation of $z_i$ and therefore the angle between $\hat z_s$ and $\hat z_f$.
+Two distinct classes can either have a relative angle $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$ depending on the external magnetic field: we choose by definition that $\vec{B}\cdot\hat{z}_i >0$, which imposes the orientation of $z_i$ and therefore the angle between $\hat z_s$ and $\hat z_f$.
 This point will be further developed when discussing the $\{100\}$ and $\{110\}$ type resonances.
@@
 
-Eq. \ref{eq. eta debut annexe} can then be written:
+Eq. (\ref{eq. eta debut annexe}) can then be written:
 \begin{equation}
@@
 Eq.
-\ref{eq.
-eta base mag} can be solved analytically for the case $\psi=0$ and numerically for $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$.
+(\ref{eq.
+eta base mag}) can be solved analytically for the case $\psi=0$ and numerically for $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$.
 The three values are reported in Table \ref{table G mag}.
@@
 \caption{(a) Illustration of the co-resonance condition between a NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
-The $\hat z$ direction is defined by the condition $\mathbf{B}\cdot\hat{z} >0$.
-(b) 8 individual $T_1$ measurements on sample ADM-15-3 on a two-class resonance for 8 different values of the magnetic field (4 $\mathbf{B} \in \{110\}$ and 4 $\mathbf{B} \in \{100\}$).}
+The $\hat z$ direction is defined by the condition $\vec{B}\cdot\hat{z} >0$.
+(b) 8 individual $T_1$ measurements on sample ADM-15-3 on a two-class resonance for 8 different values of the magnetic field (4 $\vec{B} \in \{110\}$ and 4 $\vec{B} \in \{100\}$).}
 \label{110 vs 100 annexe}
@@
 As discussed in sec.
-\ref{sec geometrie resonance}, there are 4 classes of magnetic field orientations for which at least two NV classes are co-resonant: $\mathbf{B} \in \{110\}$ (two-class resonance), $\mathbf{B} \in \{100\}$ (2$\times$two-class resonance), $\mathbf{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\mathbf{B} \parallel \langle 100\rangle$ (four-class resonance).
-
-Fig.
-\ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$.
+\ref{sec geometrie resonance}, there are 4 classes of magnetic field orientations for which at least two NV classes are co-resonant: $\vec{B} \in \{110\}$ (two-class resonance), $\vec{B} \in \{100\}$ (2$\times$two-class resonance), $\vec{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\vec{B} \parallel \langle 100\rangle$ (four-class resonance).
+
+Fig.
+\ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\vec{B} \in \{110\}$ and $\vec{B} \in \{100\}$.
 We consider here an NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
-Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\mathbf{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
-
-For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\mathbf{B} \in (100)$ or $\mathbf{B} \in (011)$.
+Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\vec{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
+
+For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\vec{B} \in (100)$ or $\vec{B} \in (011)$.
 The angle $\psi$ between the $\hat z_s$ and $\hat z_f$ axes however is different in both cases.
-The same can be generalized for every $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\mathbf{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\mathbf{B} \in \{110\}$ type resonance.
+The same can be generalized for every $\vec{B} \in \{110\}$ and $\vec{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\vec{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\vec{B} \in \{110\}$ type resonance.
 
 Because $\bar G$ is greater for $\psi=109.5 \ ^\circ$ than it is for $\psi=70.5 \ ^\circ$, then following eq.
-\ref{eq.
-eta bar r bar g bar}, we should expect a faster depolarization rate when $\mathbf{B} \in \{110\}$ than when $\mathbf{B} \in \{100\}$.
-Fig.
-\ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\mathbf{B} \in \{110\}$ scenario and four in the $\mathbf{B} \in \{100\}$ scenario.
-We can clearly see a slower relaxation rate for all four $\mathbf{B} \in \{100\}$ measurements compared to all four $\mathbf{B} \in \{110\}$, which agrees with the predictions of the model.
+(\ref{eq.
+eta bar r bar g bar}), we should expect a faster depolarization rate when $\vec{B} \in \{110\}$ than when $\vec{B} \in \{100\}$.
+Fig.
+\ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\vec{B} \in \{110\}$ scenario and four in the $\vec{B} \in \{100\}$ scenario.
+We can clearly see a slower relaxation rate for all four $\vec{B} \in \{100\}$ measurements compared to all four $\vec{B} \in \{110\}$, which agrees with the predictions of the model.
 
@@
 \bar \eta(\rm{1\ class})&=\bar G(\psi=0) \bar R_0 \\[1em]
-\frac{\bar \eta(\mathbf{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
-\frac{\bar \eta(\mathbf{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
-\frac{\bar \eta(\mathbf{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
-\frac{\bar \eta(\mathbf{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
+\frac{\bar \eta(\vec{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\vec{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\vec{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\vec{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
 \end{align}
@@
 Note that we only give here the relaxation rates for a single class.
-To estimate the total PL, one would need to compute the relaxation rate of all 4 classes independently, for instance in a $\mathbf{B} \in \{110\}$ scenario, two classes have a relaxation rate $\Gamma(\mathbf{B} \in \{110\})$ and two classes have a relaxation rate $\Gamma_0$.
+To estimate the total PL, one would need to compute the relaxation rate of all 4 classes independently, for instance in a $\vec{B} \in \{110\}$ scenario, two classes have a relaxation rate $\Gamma(\vec{B} \in \{110\})$ and two classes have a relaxation rate $\Gamma_0$.
 
@@
 \includegraphics[width=0.7\textwidth]{chapter 3/Figures/appendix_alpha_s_et_f}
-\caption{Illustration of then angles $\alpha_s$ and $\alpha_f$ defined by the angle formed by $\hat x_s$ and $\hat x_f$ with the reference axes $\hat x_s^0$ and $\hat x_f^0$.
+\caption{Illustration of $\alpha_s$ and $\alpha_f$ defined as the angles formed by $\hat x_s$ and $\hat x_f$ with the reference axes $\hat x_s^0$ and $\hat x_f^0$.
 Blue arrows symbolize the local transverse electric field acting on the NV and the fluctuator respectively.}
@@
 The integral to compute the $\bar G$ factor of the dipole-dipole interaction in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis then becomes\footnote{The flip-flop term considered here is $\ket{+,0}\bra{0,+}$.
-The $\ket{-,0}\bra{0,-}$ term gives the same numerical values once averaged.}
+The $\ket{-,0}\bra{0,-}$ term gives the same numerical values once averaged.}:
 \begin{equation}
@@
 \includegraphics[width=.7\textwidth]{chapter 4/Figures/alignement_rose}
-\caption{PL of sample CVD-pink as a function of the magnetic field for 7 different misalignment of the field with respect to the [100] axis}
+\caption{PL of sample CVD-pink as a function of the magnetic field for 7 different misalignment of the field with respect to the [100] axis.}
 \label{alignement rose}
@@
 
-A potential cause for the observation of depolarization at low magnetic field when $\mathbf{B} \parallel [100]$ could be a misalignment of the magnetic field with the [100] axis.
+A potential cause for the observation of depolarization at low magnetic field when $\vec{B} \parallel [100]$ could be a misalignment of the magnetic field with the [100] axis.
 Indeed, such a misalignment would cause a splitting of the 4 classes as the magnetic field increases which would reduce the inter-class flip-flop rate.
@@
 We estimate our alignments to be precise within $\pm 1^\circ$.
-We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization in the $\mathbf{B} \parallel [100]$ case does not come from a field misalignment.
+We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization in the $\vec{B} \parallel [100]$ case does not come from a field misalignment.
 
@@
 \caption{PL of sample ADM-150-3 as a function of the magnetic field for six different laser polarization angles.
-The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field}
+The magnetic field orientation was chosen arbitrarily in the plane of polarization of the laser field.}
 \label{pola laser}

### 2022-10-21 Manuscrit Thèse/These_copie.tex
tutu

@@
-\documentclass[a4paper, 11pt]{book}
+\documentclass[a4paper, 11pt]{report}
 
@@
 
-\title{Cross-relaxation in dense ensembles of NV centers and application to magnetometry}
+\title{Cross-relaxation in dense ensembles of NV centers and application to magnetometry \\ \medskip Relaxation croisées d'ensembles de centres NV et application à la magnétométrie}
 
@@
 In this introductory chapter, we will first cover the material aspects of NV centers by focusing on the creation process of synthetic diamonds and the required steps to create NV defects inside them.
-We will then cover the fundamentals of NV physics, including the NV center optical and spin properties, basic experiments with NV centers and the various relaxations times ($T_1$,$T_2$,$T_2^*$) which characterizes the NV center's spin.
-Finally we will present the dipole-dipole interaction between two spins and the concept of cross-relaxation.
+We will then cover the optical and spin properties of single NV centers, and introduce the notions of dipole-dipole interaction and cross-relaxation.
+We will finally present basic experiments with NV centers and introduce the experimental tools required to characterize the NV center's spin.
 
@@
 \caption{From the diamond to the NV center.
-a) Drawing of a bulk diamond.
-We consider a diamond to be ``bulk" if it is bigger than a few 100 $\mu$m.
+a) Drawing of a ``bulk" (macroscopic) diamond.
 b) Drawing representing a fluorescence microscopy image of a diamond surface.
 The red fluorescent spots symbolize NV centers photoluminescence.
-c) Zoom-in on the crystalline structure of an NV center.
-Black balls represent carbon atoms}
+c) Zoom on the crystalline structure of an NV center.
+Black balls represent carbon atoms.}
 \label{diamond+NV}
 \end{figure}
-The NV center is a particular defect in diamond.
 Before studying the NV centers properties, we must first discuss the diamond itself and answer these questions: How are diamond created, what kind of defects are there in diamond, and how to specifically create NV defects.
-The field of material science related to diamonds and NV centers is a rich and complex one, and this section will only introduce basic notions regarding these subjects.
+The field of material science related to diamonds and NV centers is a rich and complex one, and this section will only touch upon basic notions regarding these subjects.
 
@@
 
-Diamond is a remarkable material in many aspects, it is famous for having the highest hardness and thermal conductivity of any natural material, and for its high optical refractive index and optical dispersion which makes it an extremely valuable crystal in jewelry.
+Diamond is a remarkable material in many respects.
+It is famous for being the hardest and most thermally conductive non-synthetic material.
+It is also a material with high optical refractive index and optical dispersion which makes it an extremely valuable crystal in jewelry.
 
@@
 Where the thermodynamic equilibrium really matters is for the crystallization process.
-The higher pressure needed to reach the diamond-stable region can only be found naturally deep under the earth crust, typically between 150 and 240 km below sea level \citep{tappert2011diamonds}, which explain their rarity.
+The higher pressure needed to reach the diamond-stable region can only be found naturally deep under the earth crust, typically between 150 and 240 km below sea level \citep{tappert2011diamonds}, which explain their scarcity.
 It also explains why, despite many previous attempts, diamonds were only synthetically produced in 1953 \citep{barnard2000diamond}, more than 50 years after the first synthetic sapphires.
@@
 
-Diamonds, either natural or artificial, are used in the lab and in industry for their extreme sturdiness, with applications ranging from abrasive to diamond anvil cells \citep{jayaraman1983diamond}.
+Diamonds grown naturally or artificially are used in the lab and in industry for their extreme hardness, with applications ranging from abrasive to diamond anvil cells \citep{jayaraman1983diamond}.
 More recently, diamonds have been the object of optics and electronics applications.
@@
 \includegraphics[width=0.9\textwidth]{chapter 1/Figures/Diamond_type}
-\caption{Traditional classification of diamonds, based on \citep{tappert2011diamonds}}
+\caption{Traditional classification of diamonds, based on \citep{tappert2011diamonds}.}
 \label{diamond type}
@@
 
-Diamond growth by CVD also started in the 50s, but the first reasonable growth rates were only achieved in the 80s \citep{matsumoto1982growth, matsumoto1982vapor, kamo1983diamond}.
+Diamond growth by CVD also started in the 50s, but only really took off in the 80s \citep{matsumoto1982growth, matsumoto1982vapor, kamo1983diamond} with the introduction of atomic hydrogen catalysis.
 Unlike HPHT, CVD operates in a regime where diamond is not the thermodynamically stable phase.
@@
 The plasma acts as the carbon source for the diamond growing on top of the substrate, and as the catalyst to favor the diamond phase.
-The substrate is kept at a temperature $700 \sim 1100\ ^\circ$C during the growth.
+%The substrate is kept at a temperature $700 \sim 1100\ ^\circ$C during the growth.
 
@@
 
-If the CVD reactor is well made, the main sources of extrinsic impurities in the diamond comes from the source gasses \citep{balmer2009chemical}.
+If the CVD reactor is well built, the main sources of extrinsic impurities in the diamond comes from the source gasses \citep{balmer2009chemical}.
 The commercial availability of source gases containing less than $\sim$ ppm impurities made possible the growth of  high purity type IIa CVD diamonds \citep{kasu2003high, twitchen2004high, tallaire2006characterisation}.
@@
 \includegraphics[width=\textwidth]{chapter 1/Figures/formtion_NV_2}
-\caption{Illustration of the formation of NV centers. Black dots represents carbon atoms and green dots nitrogen atoms}
+\caption{Illustration of the formation of NV centers. Black and green dots represents carbon and nitrogen atoms respectively.}
 \label{formation NV}
@@
 NV centers are a relatively rare impurity in untreated diamonds.
-In CVD diamond for instance, the proportion of nitrogen atoms forming an NV center in an untreated crystal is at best of $2-3 \%$ \citep{hartland2014study} and oftentimes much lower.
-While this ratio can be enough  to work with single NV centers, applications such as sensing that require dense ensemble need a higher conversion rate.
+In CVD diamond for instance, the proportion of nitrogen atoms forming an NV center in an untreated crystal is at best of $2-3 \%$ \citep{hartland2014study}.
+While this ratio can be enough  to work with single NV centers, applications that require dense ensemble need a higher conversion rate.
 We will discuss here the usual techniques employed to improve the formation of NV centers.
@@
 This can be achieved either \textit{in situ} by introducing (voluntarily or not) nitrogen in the crystal during the growth process \citep{tallaire2006characterisation, lobaev2017influence}, or \textit{ex situ} by implanting N$^+$ ions with ion implantation techniques \citep{meijer2005generation, smith2019colour}.
-Because this present work focuses on dense NV ensembles, we will consider the simplest case where the starting crystal is a type-Ib diamond due to the presence of nitrogen during the growth.
+Because this present work focuses on dense NV ensembles, we will consider the simplest case where the starting crystal is a type-Ib diamond because of the presence of nitrogen during the growth.
 A starting concentration [N$_s$]= $10 \sim 300$ ppm is typical for type Ib HPHT diamonds \citep{achard2020chemical}.
@@
 This step might be optional for crystals which already contain a large number of monovacancies, but CVD crystals in particular can have less vacancies that they have substitutional nitrogen atoms \citep{mainwood1999point}.
-The general technique to create monovacancy is to bombard the sample with high energy particle in order to knock out carbon atoms from the crystal.
-Several high energy particle have been used to create vacancies, such as protons, electrons, ions, neutrons and even gamma rays \citep{davies1976optical, ashbaugh1988gemstone, kleinsasser2016high}.
-The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their relatively frequent availability (compared to neutrons or gamma rays) and their ability to deeply penetrate the crystal.
-
-The final step is to migrate the vacancies next to the nitrogen atoms.
+The general technique to create monovacancy is to bombard the sample with high energy particle in order to knock carbon atoms out of the crystal.
+Several high energy particles have been used to create vacancies, such as protons, electrons, ions, neutrons and even gamma rays \citep{davies1976optical, ashbaugh1988gemstone, kleinsasser2016high}.
+The most common irradiation source for for bulk diamond however is electron irradiation (e-beam) \citep{acosta2009diamonds}, thanks to their higher availability (compared to neutrons or gamma rays) and their ability to deeply penetrate the crystal.
+
+The final step is to move the vacancies next to the nitrogen atoms.
 The activation energy required to move vacancy is of $\sim 2.3\ \rm eV$ which corresponds to temperatures of $600\sim 700 ^\circ$C \citep{davies1992vacancy, newton2002recombination}.
@@
 
-The final N$_s$ to NV conversion rate can be as high as $\sim 50 \%$ \citep{grezes2015storage, hartland2014study}, although this ratio tend to decrease for higher [N$_s$] concentrations.
-Most sample studied in this manuscript had a concentration $[\rm NV]= 1-10\ \rm ppm$ (see [REF] samples)
-
-\subsection{Controlling the NV center charge state}
-
-The charge state of the NV center is of crucial importance for NV applications.
+The final N$_s$ to NV conversion rate can be as high as $\sim 50 \%$ \citep{grezes2015storage, hartland2014study}, although this ratio tends to decrease for higher [N$_s$] concentrations.
+Most sample studied in this manuscript had a concentration $[\rm NV]= 1-10\ \rm ppm$ (see sample details in Appendix \ref{Appendix samples})
+
+%\subsection{Controlling the NV center charge state}
+%
+%The charge state of the NV center is of crucial importance for NV applications.
 Only the negatively charged NV$^-$ center shows the specific spin properties that will be detailed below, but the NV center can exist under three charge states in the diamond: NV$^-$, NV$^0$ and NV$^+$.
 The presence of NV$^+$ is usually marginal \citep{hauf2014addressing, pfender2017protecting} and we will focus here solely on the NV$^-$ and NV$^0$ charge state.
-
-The charge state of the NV centers depends not only on the growth process of the crystal, but also on the experiment performed with the sample.
+%
+%The charge state of the NV centers depends not only on the growth process of the crystal, but also on the experiment performed with the sample.
 In particular, illumination in the optical range can promote electrons in the conduction band or holes in the valence band and modify the charge state of the NV center.
-
-On the material side, the crystal has to remain electrically neutral which means that at least as many charge donors need to be present as there are NV$^-$ centers.
+%
+%On the material side, the crystal has to remain electrically neutral which means that at least as many charge donors need to be present as there are NV$^-$ centers.
 For nitrogen-rich samples, the donors are mostly N$^+$ centers, which imposes a limit on the maximum conversion rate of N to NV in order to keep a high [NV$^-$]/[NV$^0$] ratio.
 The possibility to introduce other donors, such as phosphorus, has been studied \citep{doi2016pure} but remains impractical so far \citep{barry2020sensitivity}.
-
-On the experimental side, it has been shown that the ionization of NV$^-$ to NV$^0$ and the reverse process depend on the intensity and wavelength of the laser used for illumination \citep{aslam2013photo}.
+%
+%On the experimental side, it has been shown that the ionization of NV$^-$ to NV$^0$ and the reverse process depend on the intensity and wavelength of the laser used for illumination \citep{aslam2013photo}.
 It was found that the optimal wavelength to preserve the NV$^-$ charge state was  $510-540\ \rm nm$.
@@
 On the other hand, it is possible to semi-permanently ionize the NV$^-$ into NV$^0$ when using a red excitation, which has been used to create optical memories \citep{dhomkar2016long}.
-
-For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one.
+%
+%For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one.
 For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach up to 50\% of the total NV concentration \citep{grezes2015storage}
@@
 In the remaining part of this manuscript, we will only consider the negatively charged NV$^-$ state, unless specified otherwise.
-We will therefore omit the ``$-$" and simply refer to the negatively charged state as ``NV".
+We will therefore omit the ``$-$" from now on and simply refer to the negatively charged state as ``NV".
 \subsection{Optical properties of the NV center}
@@
 The value of the transition corresponds to the zero phonon line (ZPL) and are taken from \citep{doherty2013nitrogen}.
-(c) Representation of the vibronic structure for the $^3A_2$ and $3^E$ states under the Franck-Condon approximation.
-The graph axes are the energy of the states E and configuration coordinate q.
-The green and red arrow represents non-resonant (Stokes) absorption and emission.
+(c) Representation of the vibronic structure for the $^3A_2$ and $E^3$ states under the Franck-Condon approximation.
+The graph axes are the energy of the states $E$ and configuration coordinate $q$.
+The green and red arrow represents non-resonant (Stokes) absorption and emission processes.
 (d) Emission and absorption spectrum of the NV center main transition at 300 K, from PL and PLE measurements respectively.
@@
 
-Because the NV center has discrete electronic transitions inside the diamond band gap, it fluoresces: it can absorb light from $450-650$ nm and re-emits light in the $600-800$ nm band.
-
-Fig.
-\ref{NV optical} explains the origin of the NV center optical properties.
-Fig.
-\ref{NV optical}-a) illustrates the electronic ground state $^3A_2$ and first excited state $^3E$, where 1 electron from the $a_1$ molecular orbital is promoted to the $e$ level.
+Because the NV center has discrete electronic transitions inside the diamond band gap, it fluoresces: it can absorb light from $450-650$ nm and re-emit light in the $600-800$ nm band.
+
+Fig.
+\ref{NV optical}-a) shows the electronic ground state $^3A_2$ and first excited state $^3E$, where 1 electron from the $a_1$ molecular orbital is promoted to the $e$ level.
 Both of these electron occupations have two unpaired electrons, meaning that they can exist either either in a spin triplet state ($S=1$) or a spin singlet state ($S=0$) \footnote{The two singlet states $^1E$ and $^1A_1$ represented in Fig.
@@
 The emission and absorption of light associated with the triplet transition can be accompanied by the emission or absorption of an optical phonon.
-Because the lifetime of the optical phonons is very short, the absorption or emission of light almost always start from the fundamental state of the $^3A_2$ and $^3E$ level, which explains why the absorption of light can only be done for energies above the ZPL, and the emission for energies below the ZPL (Stokes process).
+Because the lifetime of the optical phonons is very short, the absorption or emission of light almost always starts from the fundamental state of the $^3A_2$ and $^3E$ level, which explains why the absorption of light can only be done for energies above the ZPL, and the emission for energies below the ZPL (Stokes process).
 
@@
 Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages.
-First, they are very bright thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and quantum efficiency of $0.7-0.8$ \citep{schirhagl2014nitrogen}.
-With proper nanophotonics engineering (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}.
+First, they are very bright thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and thanks to a quantum efficiency of $0.7-0.8$ \citep{schirhagl2014nitrogen}.
+With modern nanophotonics methods (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}.
 Secondly, the NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}.
@@
 \includegraphics[width=1\textwidth]{chapter 1/Figures/spin_fine_hyperfine}
-\caption{Representation of the fine and hyper-fine structure of the $^3A_2$ state for a $^{14}N$ nucleus and $^{12}C$ nuclei.}
+\caption{Representation of the fine and hyper-fine structure of the $^3A_2$ state for a $^{14}$N nucleus and $^{12}$C nuclei.}
 \label{NV spin}
@@
 The spin structure in the ground state consists in an electronic spin $S=1$ coupled to the nuclear spin of the nearby nuclei.
-Carbon's most common isotope is $^{12}C$ at 98.9 \% natural abundance.
-$^{12}C$ has a a nuclear spin $I=0$ and therefore does not contribute to the hyper-fine structure.
-Nitrogen's most common isotope however is $^{14}N$ at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does create a hyper-fine splitting.
-We will only cover here the most likely configuration for an NV center which is a $^{14}N$ nucleus and $^{12}C$ nuclei as closest neighbors.
+Carbon's most common isotope is $^{12}$C at 98.9 \% natural abundance.
+$^{12}$C has a a nuclear spin $I=0$ and therefore does not contribute to the hyper-fine structure.
+Nitrogen's most common isotope however is $^{14}$N at 99.6 \% natural abundance which has a nuclear spin $I=1$ and does create a hyper-fine splitting.
+We will only cover here the most likely configuration for an NV center which is a $^{14}$N nucleus and $^{12}$C nuclei as closest neighbors.
 
@@
 \begin{equation}
-\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S}\cdot \bar{\bar{A}}\cdot \mathbf{I},
-\end{equation}
-where $\mathcal{H}_e$ is the purely electronic spin Hamiltonian, $\mathcal{H}_n$ the purely nuclear spin Hamiltonian, $\mathbf{S}$ the electronic spin operator, $\mathbf{I}$ the nuclear spin operator and $\bar{\bar{A}}$ the hyper fine tensor.
+\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \vec{S}\cdot \bar{\bar{A}}\cdot \vec{I},
+\end{equation}
+where $\mathcal{H}_e$ is the purely electronic spin Hamiltonian, $\mathcal{H}_n$ the purely nuclear spin Hamiltonian, $\vec{S}$ the electronic spin operator, $\vec{I}$ the nuclear spin operator and $\bar{\bar{A}}$ the hyper fine tensor.
 \subsubsection{Electronic spin Hamiltonian}
-The electronic spin $\mathcal{H}_e$ can be written:
+The electronic spin Hamiltonian $\mathcal{H}_e$ can be written:
 \begin{equation}
 \label{eq. spin elec}
-\mathcal{H}_e=D S_z^2 + \gamma_e \mathbf{S} \cdot \mathbf{B} + \mathcal{H}_{elec}(\mathbf{E})+ \mathcal{H}_{strain}(\mathbf{\bar{\bar{\varepsilon}}}).
+\mathcal{H}_e=D S_z^2 + \gamma_e \vec{S} \cdot \vec{B} + \mathcal{H}_{elec}(\vec{E})+ \mathcal{H}_{strain}(\vec{\bar{\bar{\varepsilon}}}).
 \end{equation}
@@
 \begin{equation}
-\mathcal{H}_e=Q I_z^2 + \gamma_n \mathbf{I} \cdot \mathbf{B},
-\end{equation}
-
-where $Q=-4.945\ \rm MHz$ is the nucleus electric quadrupole moment and $\gamma_n=0.308\ \rm kHz/G$ is the nuclear gyromagnetic ratio.
+\mathcal{H}_e=Q I_z^2 + \gamma_n \vec{I} \cdot \vec{B},
+\end{equation}
+
+where $Q=-4.945\ \rm MHz$ is the nucleus electric quadrupole moment and $\gamma_n=-0.308\ \rm kHz/G$ is the nuclear gyromagnetic ratio.
 
@@
 
-Fig. \ref{NV spin} illustrates the fine and hyper-fine structure of the NV center without any external fields.
-
-For most intent and purposes, the hyper-fine structure of the NV center can simply be thought of as an energy offset, and is thus often ignored.
-The electric and strain dependencies of the electronic Hamiltonian are also often ignored as they are of second order compared to the magnetic field dependency.
-This leads to the simpler spin Hamiltonian:
+Fig. \ref{NV spin} shows fine and hyper-fine structure of the NV center without any external fields.
+
+The hyper-fine splitting as well as the electric and strain dependency are of second order compared to the Zeeman and zero field splitting terms.
+Unless specified otherwise, we will from now on refer to the following simplified Hamiltonian as the NV spin Hamiltonian:
 
@@
 \bigskip
-The NV center's spin usage as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}.
+The NV center's spin use as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}.
 Part of the reason for this wide usage are the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}.
@@
 
-\subsection{Spin and optical levels intersystem crossing}
+\subsection{Intersystem crossing of the electronic levels}
 \label{sec ISC}
@@
 Green arrows symbolize the optical excitation induced by the laser.
-The times indicated correspond to the inverse of the decay rates found in \citep{gupta2016efficient}.}
+The indicated durations correspond to the inverse of the decay rates found in \citep{gupta2016efficient}.}
 \label{ISC}
@@
 The optical excitation and radiative decay between the $^3A_2$ and $^3E$ levels preserve the spin projection $m_s$ \citep{robledo2011spin}.
-There is however a non-radiative decay path which involves in inter-system crossing with the singlet states $^1A_1$ and $^1E$ that depends on the the triplet states spin projection.
+There is however a non-radiative decay path which involves an inter-system crossing with the singlet states $^1A_1$ and $^1E$ that depends on the the triplet states spin projection.
 
@@
 \ref{ISC} summarizes the possible decay channels from the $^3E$ states.
-We assume that the electrons are initially excited from the $^3A_2$ states via a green laser with a rate $\gamma_{\rm las}$.
-The $^3E$ states can then either decay radiatevely and emit a 600$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states.
-We consider the decay channel through the singlet state as non-radiative due to the strong non radiative channels competing with the IR radiative transition (quantum efficiency $\leq 10^{-3}$ for the singlet transition \citep{rogers2008infrared, ma2010excited, acosta2010optical}).
+Once excited to the $^3E$ states, for instance by photoexcitation, the electrons can either decay radiatively and emit a 600$-$800 nm photon, or undergo an intersystem crossing and decay through the singlet states.
+The decay though the singlet state is often referred to as non-radiative, due to both the difference in wavelength between the triplet and the singlet transitions, and due to the strong non-radiative competition in the singlet decay (quentum efficiency $\leq 10^{-3}$ \citep{rogers2008infrared, ma2010excited, acosta2010optical}.
 
 The rates $\gamma$ of the various decay processes, represented in Fig.
-\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depends on the spin projection $m_s$: the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$\frac{15+91}{15+11}\approx 4$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
-Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay in the $\ket{0}$ state than in the $\ket{\pm 1}$ states.
+\ref{ISC} by the time constant $\tau=1/\gamma$, strongly depend on the spin projection $m_s$.
+Importantly, the likeliness to undergo a non-radiative decay is $\sim 4$ times\footnote{$\frac{15+91}{15+11}\approx 4$} more likely when starting from the $\ket{\pm 1}$ states than when starting from the $\ket{0}$ state.
+Similarly, the singlet state is $\sim 2.4$ times\footnote{$\frac{492}{204} \approx 2.4$} more likely to decay to the $\ket{0}$ state than to the $\ket{\pm 1}$ states.
 
 These spin selective intersystem crossings mean that an optical pumping scheme in the $\ket{m_s=0}$ spin state takes place under constant green excitation.
-At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}, which is equivalent to a thermal polarization (for B=0) with a spin temperature of $\approx 65\ \rm mK$ \footnote{\begin{equation*}
+At optical saturation, the polarization of the $\ket{0}$ state can reach $70-90\ \%$ \citep{gupta2016efficient}.
+This is equivalent to a spin-temperature $\approx 65\ \rm mK$ \footnote{\begin{equation*}
 T=\frac{h D}{k_B \ln (\frac{1-P}{2P})} \approx 65\ \rm mK,
 \end{equation*}
-where $D$ is th ZFS factor described in eq.
-(\ref{eq.
-spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.} while the diamond itself can remain at room temperature.
+where $D$ is th ZFS factor described in eq. (\ref{eq. spin elec}) and $P\approx 0.8$ is the spin polarization defined as the population in the $\ket{0}$ state.}.
 
@@
 
-\section{Basic operations with NV centers}
-
-This section introduces two elementary operations with NV centers that are Rabi oscillations and optically detected magnetic resonance (ODMR).
-
-Rabi oscillations is the basis for any kind of coherent qubit manipulation while ODMR is used to find the transition frequencies of the NV centers, as well as being a basic magnetometry protocol.
-
-This section also covers the experimental setup used for these two operations and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
-
-\subsection{Experimental setup}
-\label{Sec setup cahp 1}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.8\textwidth]{chapter 1/Figures/shema_exp}
-\caption{Confocal microscopy setup detailed in main text}
-\label{setup ch1}
-\end{figure}
-The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field.
-Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.
-
-Fig.
-\ref{setup ch1} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
-Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.
-
-The setup consists of a confocal microscope where the excitation from a green laser (Thorlabs CPS532, 532 nm, 5 mW) is focused on a sample containing NV centers with a microscope objective (NA $\sim$ 0.5).
-The NV PL is then collected by the same objective and isolated from the back-scattered laser light with a dichroic mirror (DM) and additional filters (532 nm notch filter and 645 nm longpass filter to filter out the NV$^0$ PL).
-The PL is collected on an avalanche photodiode (Thorlabs APD410A or Perkin Elmer SPCM-AQRH-15) whose signal is sent to the acquisition card (NI X-series).
-
-The laser beam goes through an acousto-optic modulator (AOM, AA opto-electronic) controlled externally to create arbitrary light pulses (resolution $\sim$ ns).
-The microwave field is generated by a Rhode \& Shwarz SMB 100A generator and emitted with a home-made loop antenna positioned at $<$ 1 mm of the sample.
-A microwave switch (Mini-circuits ZASWA-2-50DRA+) is used to create microwave pulses (resolution $\sim$ 10 ns).
-The control pulses for the AOM and the microwave switch are generated either by the NI X-series card (resolution $\sim$ 1 $\mu$s), or by a pulseblaster ESR-PRO (resolution $\sim$ 10 ns).
-The microwave field can be amplified with a Mini-circuits ZHL-5W-422+ amplifier to achieve higher Rabi frequencies.
-
-Finally the magnetic field is generated either with a permanent magnet (up to $\sim$ 200 mT) or with a homemade electromagnet (up to $\sim$ 30 mT) alimented via a low freqency amplifier (Leybold power function generator 522663).
-Both the permanent magnet and the electromagnet can be  attached to a rotation stage for precise alignment.
-\subsection{Rabi Oscillations}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/Rabi}
-\caption{Rabi oscillations. (a) pulse sequence used to measure Rabi oscillations. (b) Experimental data on sample CVD-pink (cf samples [REF]).}
-\label{Rabi}
-\end{figure}
-
-The general technique for experiments involving NV spins proceeds as follow: 
-\begin{itemize}
-\item First the spins are polarized. This is done with a green laser pulse ($1-100$ $\mu$s depending on the laser intensity) which polarizes the spins in the $\ket{m_s=0}$ state at $\sim$ 80 \%.
-\item  Then the spins are manipulated. This process often involves a microwave field resonant with one of the spin transitions.
-\item Finally the spins are read out.
-The simplest spin readout for NV centers consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
-For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore.
-The optimal readout time is of the same order as the spin polarization time ($1-100$ $\mu$s).
-\end{itemize}
-
-Fig.
-\ref{Rabi}-a) shows the pulse sequence used to observe Rabi oscillations with NV centers.
-In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$.
-The microwave frequency or the magnetic field needs to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
-
-Fig.
-\ref{Rabi}-b) shows the result of such an experiment on a spin ensemble.
-The PL signal oscillations correspond to the oscillations of the spin population between the bright $\ket{0}$ state and the dark $\ket{-1}$ state.
-For long microwave pulses ($>$ 1 $\mu$s typically), the spins reach a steady state with  $\sim$ 50 \% population in both the $\ket{0}$ and $\ket{-1}$ states.
-
-\subsection{Optically detected magnetic resonance}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_basique}
-\caption{ODMR protocol.
-(a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$.
-(b) ODMR spectrum with no magnetic field on sample ADM-15-4.}
-\label{ODMR basique}
-\end{figure}
-
-Optically detected magnetic resonance (ODMR) is the most commonly employed magnetometry protocol with NV centers due to its simplicity and the fact that it does not require prior calibration.
-
-The protocol for continuous wave (CW) ODMR is presented in Fig.
-\ref{ODMR basique}.
-In this simple protocol, the microwave is always turned on while the frequency spans a predefined interval $[f_{\rm start}, f_{\rm end}]$.
-The PL is recorded as the microwave is being swept, with a typical acquisition time for each PL sample $> 1\ \rm ms$.
-The sweeping frequency must be slow enough so that the PL can reach a steady state with the microwave field for each sample acquisition.
-
-As we saw in the last section, the steady state of the NV center spin with a resonant microwave field is a statistical mixture of 50 \% $\ket{0}$ and 50 \% $\ket{\pm 1}$ \footnote{We are assuming here that the Rabi frequency is greater than the laser polarizing rate, which is often the case in practice}.
-If the microwave field is not resonant with a spin transition however, the steady state is predominantly $\ket{0}$ due to the spin polarization caused by the laser.
-The NV PL will therefore be lowered when the scanning microwave field becomes resonant with one of the NV spin transitions.
-
-Fig.
-\ref{ODMR basique}-b) shows an example of ODMR spectrum on an NV ensemble with no external magnetic field.
-The reason for the two bumps instead of a single one centered on $D=2870\ \rm MHz$ is the splitting of the $\ket{+1}$ and $\ket{-1}$ levels occasioned by the strain and local electric field, which will be further discussed in chapter 4.
-
-\subsubsection{ODMR with lock-in detection}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_lockin}
-\caption{ODMR protocol with lock-in detection.
-(a) Simplified confocal microscope setup with a lock-in amplifier (LIA) modulating the mircowave and demodulating the PL.
-(b) Demodulated ODMR spectrum with no magnetic field on sample ADM-15-4.}
-\label{ODMR lockin}
-\end{figure}
-
-Because the ODMR protocol delivers a continuous signal (the PL) as a function of a periodic excitation (the microwave field), it is well suited to amplitude or frequency modulation and lock-in detection.
-
-The principle of lock-in detection will not be covered here, the key idea is that lock-in detection allows to translate the signal spectrally in a region less prone to noise.
-In this case, the ODMR signal is a DC measurement and is therefore prone to slow freqency noises such as thermal or mechanical drifts and laser amplitude fluctuations.
-The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low freqency noises, but slow enough compared to the NV photodynamics.
-
-Fig.
-\ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig.
-\ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
-All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1\ \rm kHz$.
-
-\subsection{ODMR with NV ensemble: longitudinal and transverse magnetic field}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_8_classes}
-\caption{(a) Geometry of the NV center and definition of $B_z$ and $B_\perp$.
-(b) Representation of the 4 possibles orientations (``classes") of the N$-$V axis, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell.
-(c) Simulation of the evolution of the transition frequencies for each class of NV centers as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
-(d) ODMR spectrum on sample ADM-150-1 for a magnetic field of $\sim 90\ \rm G$.}
-\label{ODMR 8 classes}
-\end{figure}
-
-We mentioned previously that the NV spin Hamiltonian (eq.
-(\ref{NV spin Hamiltonian basic})) was anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
-The magnetic field is thus often decomposed between the longitudinal magnetic field $B_z$ and the transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig.
-\ref{ODMR 8 classes}-a).
-
-For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation of $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$.
-The true eigenstates however remain close to $\ket{0}$, $\ket{\pm 1}$ as long as $\gamma_e B_z > \frac{(\gamma_e B_\perp)^2}{D}$.} can be approximated in a perturtbative approach to: %cf AN.py
-\begin{align}
-\nu_0&= -\frac{(\gamma_e B_\perp)^2}{D} \\
-\nu_{-1} &= D - \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D} \\
-\nu_{+1} &= D + \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}
-\end{align}
-
-Fig.
-\ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal.
-We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian.
-For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and that each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
-
-Fig.
-\ref{ODMR 8 classes}-c) shows a simulation of the evolution of the transition frequencies $\nu_{+1} - \nu_{0}$ and $\nu_{-1} - \nu_{0}$ for all 4 classes as a function of the magnetic field amplitude, with the magnetic field orientation fixed in an arbitrary direction.
-Fig.
-\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles.
-8 lines are visible corresponding to the two transitions of the 4 classes.
-Based on the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.
-
-We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
-\ref{sec ISC}.
-For strong transverse magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
-
-\section{Dynamics of the NV center spin}
-
-This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times.
-Relaxation times are important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
-
-The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ which corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix), and the transverse relaxation time $T_2$\footnote{We follow here the NMR nomenclature} which corresponds to the relaxation of the spin state coherences (non diagonal elements of the density matrix).
-
-The transverse relaxation time, also called coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$.
-The $T_2^*$ time includes inhomogeneous dephasing processes such as the inhomogeneity in space (when multiple qubits are probed) and time of the external fields.
-The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}.
-While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used.
-More careful nomenclatures will label the $T_2$ times as $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance.
-The relaxation times are ordered as follows:
-\begin{equation}
-T_2^* < T_2 < 2 T_1.
-\end{equation}
-
-The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the sample used.
-The following paragraphs will cover the measurement processes of the different relaxation times and the typical values for the samples used in this manuscript.
-
-\subsection{NV spin $T_1$}
-\label{sec T1 NV}
-\subsubsection{Measurement protocol}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/T1_basique}
-\caption{Basic $T_1$ measurement scheme.
-(a) Pulse seqeunce used.
-(b) Experimental data obtained on sample CVD-pink for low laser intensity.
-The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
-\label{T1 basique}
-\end{figure}
-
-$T_1$ is the simplest relaxation time to measure because the spin populations are directly correlated to the NV centers PL.
-Fig.
-\ref{T1 basique} shows the pulse sequence used and an experimental result of the $T_1$ sequence.
-The spin lifetime $T_1$ is one of the only spin property that can be measured without a resonant microwave.
-
-The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig.
-\ref{T1 soustraction}-a).
-This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$.
-Then the sequence is repeated with an additional $\pi$-pulse on one the NV transitions right before the readout.
-This $\pi$-pulse will project the remaining $\ket{0}$ population difference into a $\ket{\pm 1}$ state.
-The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which rejects every contribution to the PL outside of the spin relaxation.
-
-Fig.
-\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample that was used in Fig.
-\ref{T1 basique}.
-The difference between the $S_1$ data in Fig.
-\ref{T1 soustraction}-b) and the the data in Fig.
-\ref{T1 basique}-b) comes from the laser intensity used.
-In Fig.
-\ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant.
-Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark \citep{giri2018coupled, giri2019selective} (most likely due to charge tunneling between impurities \citep{choi2017depolarization}) .
-On the other hand, the laser intensity used in Fig.
-\ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.
-
-Fig.
-\ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal.
-Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig.
-\ref{T1 basique}.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/T1_protocole_soustraction}
-\caption{Common mode rejection $T_1$ measurement scheme.
-(a) Pulse seqeunce used.
-(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity.
-(c) Subtraction $S_1-S_2$ versus dark time $\tau$.
-The experimental data is fitted with an exponential decay of time constant $\tau=4.1\ \rm ms$.}
-\label{T1 soustraction}
-\end{figure}
-
-The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
-The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.
-
-\subsubsection{Physical origin and typical values}
-
-We distinguish three different origin for the the $T_1$ relaxation process:
-\begin{itemize}
-\item \textbf{Spin-phonon relaxation:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which can result in a phonon-induced spin relaxation.
-The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}.
-The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxtion times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at for mili-K temperatures \citep{astner2018solid}.
-
-\item \textbf{Electric and magnetic field noises:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}.
-As a result, magnetic and electric field noise can reduce the spin $T_1$ if their spectrum overlap with the NV transitions.
-This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}.
-The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effects are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
-
-\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipolar interaction betwen two spins will be detailed more thoroughly in the following section.
-Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation.
-The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration and can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
-\end{itemize}
-
-All the measurements shown in this manuscript were performed at room temperature.
-The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$.
-Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses.
-The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed.
-$T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.
-
-\subsection{NV spin $T_2^*$}
-\subsubsection{Measurement protocol}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/Ramsey}
-\caption{Ramsey interferometry protocol.
-(a) Pulse seqeunce used.
-(b) Experimental data obtained on sample CVD-pink.} %The orange line corresponds to an Gaussian fit with time constant $\tau=400 \pm 100\ \rm ns$.}
-\label{Ramsey}
-\end{figure}
-
-The $T_2^*$ time is the characteristic time for the loss of quantum coherence of a freely evolving system.
-It is generally defined as the decay time of of a Ramsey interferometry measurement \citep{barry2020sensitivity}.
-
-Fig.
-\ref{Ramsey} shows the Ramsey pulse sequence and an  example obtained on a spin ensemble.
-The microwave frequency is detuned by $\sim 300\ \rm kHz$ compared to the spin central Larmor frequency which causes oscillations in the signal.
-$T_2^*$ is the decay time of the envelope of the Ramsey fringes.
-Unfortunately, it is hard to correctly fit the envelope of a signal with so few fringes.
-Increasing the number of fringes with our current microwave and optical setup might prove hard, as further microwave detuning would decrease the signal contrast, and would be ultimately limited by the microwave pulses resolution.
-
-Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
-The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity, and because the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
-The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
-
-An alternative measurement of $T_2^*$ can be performed by looking at the linewidth of an ODMR spectrum.
-Indeed, ideally the ODMR spectrum should be the Fourier transform of the Ramsey signal (also called free induction decay in NMR literature).
-In practice however, both methods suffer from experimental noises and imperfections, with Ramsey-type methods being more sensitive to microwave inhomogeneities \citep{barry2020sensitivity} while ODMR is prone to laser and microwave power broadening \citep{dreau2011avoiding}.
-
-The method chosen to extract $T_2^*$ values in this manuscript was ODMR with low microwave power ($\Omega_{\rm Rabi} \ll 1/T_2^*$) and low laser intensity (compared to the the NV optical saturation).
-
-\subsubsection{Physical origin}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_1_classes}
-\caption{ODMR spectra on a single NV class.
-(a) Experimental data on sample CVD-pink, fitted with three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
-(b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
-\label{ODMR 1 classe}
-\end{figure}
-
-For most samples, the leading cause of inhomogeneous dephasing is the temporal and spatial variations of magnetic fields caused by the paramagnetic impurities in the crystal \citep{barry2020sensitivity}.
-The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2.
-We will simply mention here that for sample with high NV density ([NV] $\geq$ ppm), $T_2^*$ is mostly limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
-
-Fig.
-\ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see [REF]).
-The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a likely nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
-We can clearly see a higher linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
-In particular, we can notice the hyperfine structure of the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
-
-Fig.
-\ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula.
-While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}.
-For Fig.
-\ref{ODMR 1 classe}-a), it was indeed found that Lorentzians fitted the data better than Gaussians.
-For Fig.
-\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, although this could be in part due to the unresolved hyperfine structure.
-
-\subsection{NV spin $T_2$}
-\label{sec T2 echo}
-\begin{figure}[h!]
-\centering
-\includegraphics[width=\textwidth]{chapter 1/Figures/Echo}
-\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
-\label{Echo}
-\end{figure}
-\subsubsection{Measurement protocol}
-As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.
-
-Fig.
-\ref{Echo} shows the pulse sequence used and an example on the same sample than the Ramsey measurement.
-The pulse sequence differs from the Ramsey one by adding a a $\pi$-pulse in the middle of the free precession time.
-This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased \citep{hahn1950spin}.
-
-This protocol extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
-Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
-This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
-
-\subsubsection{Physical origin}
-The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence.
-The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}.
-For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2019decoherence}.
-
 \section{Dipole-dipole interaction between two electronic spins}
@@
 where the spin operators $\hat S_{\{x,y,z\} }$ have their usual definitions in term of Pauli matrices.
-In particular, this means that the spin axis of quantization of each spin is $z_i$, which classically corresponds to the magnetization direction of the dipole.
-This axis can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (including the NV center).
+The axis of quantization of each spin $z_i$ can be defined either by the magnetic field, or by crystal field anisotropies, as is often the case for solid state spins with $S\geq 1$ (including the NV center).
 
@@
 
-where $\mathbf{r}=r \mathbf{u}$ is the relative position of the two dipoles (see Fig. \ref{dipole-dipole}).
-
-The quantum analog of the dipole-dipole interaction between spins simply replaces the magnetic moments $\mathbf{m}_i$ in eq.
+where $\vec{r}=r \vec{u}$ is the relative position of the two dipoles (see Fig. \ref{dipole-dipole}).
+
+The quantum analog of the dipole-dipole interaction between spins simply replaces the magnetic moments $\vec{m}_i$ in eq.
 (\ref{eq.
-dipole classique}) with $\hbar \gamma_i \hat{\mathbf{S}}_i$, where $\gamma_i$ is the gryromagnetic ratio and $\hat{\mathbf{S}}_i$ the spin vector operator for spin $i$:
-
-\begin{equation}
-\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\mathbf{S}}_1\cdot\mathbf{u})(\hat{\mathbf{S}}_2\cdot\mathbf{u}) - \hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\mathbf{S}}_1\cdot\hat{\mathbf{S}}_2\,\delta(r)\right].
+dipole classique}) with $\hbar \gamma_i \hat{\vec{S}}_i$, where $\gamma_i$ is the gryromagnetic ratio and $\hat{\vec{S}}_i$ the spin vector operator for spin $i$:
+
+\begin{equation}
+\label{eq. dipole quantique J coupling}
+\mathcal{H}_{\rm dd}=\frac{\mu_0 \gamma_1 \gamma_2 \hbar^2}{4 \pi}\left[ \frac{3 (\hat{\vec{S}}_1\cdot\vec{u})(\hat{\vec{S}}_2\cdot\vec{u}) - \hat{\vec{S}}_1\cdot\hat{\vec{S}}_2}{r^3}+\frac{8\pi}{3}\hat{\vec{S}}_1\cdot\hat{\vec{S}}_2\,\delta(r)\right].
 \end{equation}
@@
 
-\subsection{Decomposition of the dipole-dipole Hamiltonian}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=.7\textwidth]{chapter 1/Figures/flip-flop_double-flip}
-\caption{Representation of the two energy transfer mechanisms: flip-flop and double flips} 
-\label{flip flop double flip}
-\end{figure}
-
-In the weak coupling approximation, the dipole-dipole Hamiltonian between two spins can be decomposed between diagonal terms which are responsible for energy shifts of the individual spin levels, and the off-diagonal terms which are responsible for energy transfers between the two spins.
-
-In what follows, we will assume that both spins have the same quantization axis in order to simplify the Hamiltonian decomposition. A more general case is detailed in [REF].
+\subsection{Spin exchange mechanisms}
+\label{sec flip-flop et double flip}
+
+We will here make the difference between the dipole-dipole Hamiltonian diagonal diagonal terms which are responsible for energy shifts of the individual spin levels, and the off-diagonal terms which are responsible for spin exchange \footnote{We consider here the weak coupling regime where we neglect the spin-mixing induced by the dipole-dipole interaction.
+This is justified by the experimental values in our system.}.
+
+In what follows, we will assume that both spins have the same quantization axis in order to simplify the Hamiltonian decomposition. A more general case is detailed in appendix \ref{Appendix eta}.
 
 Following the notation introduced in Fig.
-\ref{dipole-dipole}, we will then assume that $$(\mathbf{x}_1,\mathbf{y}_1,\mathbf{z}_1) = (\mathbf{x}_2,\mathbf{y}_2,\mathbf{z}_2) = (\mathbf{x},\mathbf{y},\mathbf{z}).$$ We will denote $\mathbf{u}=u_x \mathbf{x} + u_y \mathbf{y} + u_z \mathbf{z}$ the decomposition of the unitary vector $\mathbf{u}$ in this basis.
+\ref{dipole-dipole}, we will then assume that $$(\vec{x}_1,\vec{y}_1,\vec{z}_1) = (\vec{x}_2,\vec{y}_2,\vec{z}_2) = (\vec{x},\vec{y},\vec{z}).$$ We will denote $\vec{u}=u_x \vec{x} + u_y \vec{y} + u_z \vec{z}$ the decomposition of the unitary vector $\vec{u}$ in this basis.
 The dipole-dipole Hamiltonian can then be written:
@@
 (\ref{eq.
-ff df}) the three relevant matrix elements $\Omega_{\rm ff}$, $\Omega_{\rm df}$, and $\Omega_{\rm shift}$ which correspond respectively to flip-flops, double flips and energy shifts.
-Flip-flops and double flips (sometimes referred as flip-flips) are represented in Fig.
-\ref{flip flop double flip}.
-Flip flops correspond to energy transfer between two spins for which the sum of the spin angular momentum is conserved ($\Delta (m_s^1+m_s^2)=0$), while double flips correspond to energy transfer for which the spin angular momentum is not conserved ($\Delta (m_s^1+m_s^2)=\pm 2$)\footnote{In order to preserve the total angular momentum, double flips transfer angular moment to the crystal, which is a form of Einstein de Haas effect \citep{einstein1915experimental} }.
-In both cases however, the energy between the initial and final states needs to be preserved for the transfer to be efficient.
-
-It should be noted that on average, each matrix element is of the same order of magnitude: $\Omega_{\rm ff} \sim \Omega_{\rm df} \sim \Omega_{\rm shift} \sim \frac{J_0}{r^3}$.
+ff df}) the three relevant matrix elements $\Omega_{\rm ff}$, $\Omega_{\rm df}$, and $\Omega_{\rm shift}$ which correspond respectively to flip-flops, double flips (or flip-flip) and energy shifts.
+It should be noted that each matrix element is typically of the same order of magnitude: $\Omega_{\rm ff} \sim \Omega_{\rm df} \sim \Omega_{\rm shift} \sim \frac{J_0}{r^3}$.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{chapter 1/Figures/flip-flop_double-flip}
+\caption{Representation of the two spin exchange mechanisms: flip-flop and double flips.} 
+\label{flip flop double flip}
+\end{figure}
+
+Fig.
+\ref{flip flop double flip} illustrates the difference between flip-flops and double flips: flip-flop preserve the total spin angular momentum ($\Delta (m_s^1+m_s^2)=0$) while double-flips do not ($\Delta (m_s^1+m_s^2)=\pm 2$)\footnote{In order to preserve the total angular momentum, double flips transfer angular moment to the crystal, which is a form of Einstein de Haas effect \citep{einstein1915experimental}.}.
+Both double-flips and flip-flop are energy conservative processes, which means that they require the initial and final two-spin states to be nearly resonant (within $\sim \Omega$) in order to be efficient.
 
@@
 \includegraphics[width=0.9\textwidth]{chapter 1/Figures/Shema_CR}
-\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
+\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin.
+Red dots represent the population in each states and the arrows represent the various population transfer mechanisms.}
 \label{CR_shema}
@@
 Cross-relaxation (CR) is the transfer of polarization from one spin to another (or more generally from one family of spins to another family).
-This process can occur through either flip-flops or double flips, as long as the resonance condition between the two spins is met.
-
-On of the condition for cross-relaxation is to have a difference in polarization between the two spins species considered.
-The NV center is therefore a good candidate to observe cross-relaxations as it can be optically polarized while most other spin species are not.
+This process can occur through either flip-flops or double flips.
+
+There are two conditions to observe cross-relaxation: 
+\begin{itemize}
+\item There must be a difference of polarization between the two spins.
+\item The two spin transitions must be resonant.
+\end{itemize}
+
+The NV center is therefore a good candidate to observe cross-relaxations as it can be optically polarized which creates a difference of polarization with the other spin species in the diamond.
 
@@
 (\ref{delta gamma 1}) is the basis for every cross-relaxation detection performed in this manuscript.
-Because the NV center lifetime can be measured optically, as detailed in Sec.
+Because the NV center lifetime can be measured optically, as will be detailed in Sec.
 \ref{sec T1 NV}, cross-relaxation between NV-centers and other spin species can be measured optically.
-Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be explained in the following chapters.
+Measuring the cross-relaxation depolarization can then be used to gain information on the second spin or on the dipole-dipole interaction mechanisms, as will be shown in the following chapters.
+
+The last two sections of this chapter are dedicated to basic experiments involving NV centers.
+
+\section{Optically detected magnetic resonance (ODMR)}
+
+This section introduces one of the building block of almost any NV experiment which is the optically detected magnetic resonance (ODMR).
+ODMR is used to find the spin resonances of the NV center, which is a preliminary step to any coherent manipulation of the NV spin, as well as a powerful sensing tool in itself.
+
+This section also covers the experimental setup used for this operation and for many other NV experiments, as well as the concept of longitudinal and transverse magnetic field with regard to the NV principal axis.
+
+\subsection{Experimental setup}
+\label{Sec setup cahp 1}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{chapter 1/Figures/shema_exp}
+\caption{Confocal microscopy setup (details in main text).}
+\label{setup ch1}
+\end{figure}
+The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field.
+Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.
+
+Fig.
+\ref{setup ch1} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms.
+Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.
+
+The setup consists of a confocal microscope where the excitation from a green laser (Thorlabs CPS532, 532 nm, 5 mW) is focused on a sample containing NV centers with a microscope objective (NA $\sim$ 0.5).
+The NV PL is then collected by the same objective and isolated from the back-scattered laser light with a dichroic mirror (DM) and additional filters (532 nm notch filter and 645 nm longpass filter to filter out the NV$^0$ PL).
+The PL is collected on an avalanche photodiode (Thorlabs APD410A or Perkin Elmer SPCM-AQRH-15) whose signal is sent to the acquisition card (NI X-series).
+
+The laser beam goes through an acousto-optic modulator (AOM, AA opto-electronic) controlled externally to create arbitrary light pulses (resolution $\sim$ ns).
+The microwave field is generated by a Rhode \& Shwarz SMB 100A generator and emitted with a home-made loop antenna positioned at $<$ 1 mm of the sample.
+A microwave switch (Mini-circuits ZASWA-2-50DRA+) is used to create microwave pulses (resolution $\sim$ 10 ns).
+The control pulses for the AOM and the microwave switch are generated either by the NI X-series card (resolution $\sim$ 1 $\mu$s), or by a pulseblaster ESR-PRO (resolution $\sim$ 10 ns).
+The microwave field can be amplified with a Mini-circuits ZHL-5W-422+ amplifier to achieve higher Rabi frequencies.
+
+Finally the magnetic field is generated either with a permanent magnet (up to $\sim$ 200 mT) or with a homemade electromagnet (up to $\sim$ 30 mT) alimented via a low freqency amplifier (Leybold power function generator 522663).
+Both the permanent magnet and the electromagnet can be  attached to a rotation stage for precise alignment.
+
+The X-series and the pulseblaster, as well as several instruments (microwave generator, motorized stages) are piloted by python scripts which I developed during my PhD.
+%Salaud de Gabriel
+%\subsection{Rabi Oscillations}
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=\textwidth]{chapter 1/Figures/Rabi}
+%\caption{Rabi oscillations. (a) pulse sequence used to measure Rabi oscillations. (b) Experimental data on sample CVD-pink (see sample details in appendix \ref{Appendix samples}).}
+%\label{Rabi}
+%\end{figure}
+%
+%The general technique for experiments involving NV spins proceeds as follow: 
+%\begin{itemize}
+%\item First the spins are polarized. This is done with a green laser pulse ($1-100$ $\mu$s depending on the laser intensity) which polarizes the spins in the $\ket{m_s=0}$ state at $\sim$ 80 \%.
+%\item  Then the spins are manipulated. This process often involves a microwave field resonant with one of the spin transitions.
+%\item Finally the spins are read out.
+The simplest spin readout for NV centers consists in looking at the PL right after the laser is turned on: if the spin was in the $\ket{m_s=0}$ state, the initial PL will be high, whereas if the spin was in a $\ket{m_s=\pm 1}$ state, the initial PL will be low.
+For longer exposure time, the spins end up re-polarized in the $\ket{m_s=0}$ state and no information can be gained on the initial spin state anymore.
+The optimal readout time is of the same order as the spin polarization time ($1-100$ $\mu$s).
+%\end{itemize}
+%
+%Fig.
+\ref{Rabi}-a) shows the pulse sequence used to observe Rabi oscillations with NV centers.
+In this case, the spins manipulation simply consists in a microwave pulse of variable duration $\tau$.
+The microwave frequency or the magnetic field needs to be adjusted so that at least one NV spin transition (either $\ket{0} \to \ket{+1}$ or $\ket{0} \to \ket{-1}$) is resonant with the microwave field.
+%
+%Fig.
+\ref{Rabi}-b) shows the result of such an experiment on a spin ensemble.
+The PL signal oscillations correspond to the oscillations of the spin population between the bright $\ket{0}$ state and the dark $\ket{-1}$ state.
+For long microwave pulses ($>$ 1 $\mu$s typically), the spins reach a steady state with  $\sim$ 50 \% population in both the $\ket{0}$ and $\ket{-1}$ states.
+
+\subsection{Optically detected magnetic resonance}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_basique}
+\caption{ODMR protocol.
+(a) Pulse sequence used for CW ODMR: the microwave is kept on during the whole sequence while the frequency is scanned from $f_{\rm start}$ to $f_{\rm end}$.
+(b) ODMR spectrum with no magnetic field on sample ADM-15-4.
+(see sample details in appendix \ref{Appendix samples}).}
+\label{ODMR basique}
+\end{figure}
+
+Optically detected magnetic resonance (ODMR) is the most commonly employed magnetometry protocol with NV centers due to its simplicity and the fact that it does not require prior calibration.
+
+The protocol for continuous wave (CW) ODMR is presented in Fig.
+\ref{ODMR basique}.
+In this simple protocol, the microwave is always turned on while the frequency spans a predefined interval $[f_{\rm start}, f_{\rm end}]$.
+The PL is recorded as the microwave is being swept, with a typical acquisition time for each PL sample $> 1\ \rm ms$.
+The sweeping frequency must be slow enough so that the PL can reach a steady state with the microwave field for each sample acquisition.
+
+Whenever the microwave becomes resonant with one of the NV transition, it induces a depolarization of the NV center which results in a drop of PL.
+This allows an optical reading of the NV spin spectral response.
+
+Fig.
+\ref{ODMR basique}-b) shows an example of ODMR spectrum on an NV ensemble with no external magnetic field.
+The reason as to why there are two peaks instead of a single one centered on $D=2870\ \rm MHz$ is the splitting of the $\ket{+1}$ and $\ket{-1}$ levels occasioned by the strain and local electric field, which will be further discussed in chapter 4.
+The asymmetry between the two peaks could come either from the microwave polarization or from the uneven spectral response of the antenna \footnote{Our antennas are not impedance matched.}.
+
+ODMR contrast of up to 30 \% of the PL have been recorded with single spins \citep{dreau2011avoiding}. We typically observe ODMR contrast $\leq 5\ \%$ with NV ensemble.
+
+\subsubsection{ODMR with lock-in detection}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_lockin}
+\caption{ODMR protocol with lock-in detection.
+(a) Simplified confocal microscope setup with a lock-in amplifier (LIA) modulating the mircowave and demodulating the PL.
+(b) Demodulated ODMR spectrum with no magnetic field on sample ADM-15-4 (see sample details in appendix \ref{Appendix samples}).}
+\label{ODMR lockin}
+\end{figure}
+
+The ODMR protocol is well suited for lock-in detection with a modulation of the microwave amplitude or frequency modulation \footnote{These are preferred to laser modulation as only the spin part of the NV centers is affected by the microwave field}.
+
+The principle of lock-in detection will not be covered here, the key idea is that lock-in detection allows to translate the signal spectrally in a region less prone to noise.
+In this case, the ODMR signal is a DC measurement and is therefore prone to slow freqency noise such as electric noise, thermal or mechanical drifts and laser amplitude fluctuations.
+The microwave field can thus be modulated (in amplitude, frequency or phase) at a frequency fast enough to mitigate the low frequency noise, but slow enough compared to the NV photodynamics.
+
+Fig.
+\ref{ODMR lockin}-a) shows the ODMR lock-in setup, and Fig.
+\ref{ODMR lockin}-b) shows an example of ODMR spectrum with microwave amplitude modulation.
+All the lock-in spectra shown in this manuscript were produced with 100 \% amplitude modulation at a frequency $f_{\rm mod} \sim 1\ \rm kHz$.
+
+\subsection{ODMR with NV ensemble: longitudinal and transverse magnetic field}
+\label{sec champs transverse}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_8_classes}
+\caption{(a) Geometry of the NV center and definition of $B_z$ and $B_\perp$.
+(b) Representation of the 4 possibles orientations (``classes") of the N$-$V axes, corresponding to the 4 $\langle 111 \rangle$ directions of the diamond cubic cell.
+(c) Simulation of the evolution of the transition frequencies for each class of NV centers as a function of the amplitude of the applied magnetic field ($\textbf{B}$ orientation chosen arbitrarily).
+(d) ODMR spectrum on sample ADM-150-1 (see sample details in appendix \ref{Appendix samples}) for a magnetic field of $\sim 90\ \rm G$.}
+\label{ODMR 8 classes}
+\end{figure}
+
+We mentioned previously that the NV spin Hamiltonian (eq.
+(\ref{NV spin Hamiltonian basic})) was anisotropic due to the fine structure term $D S_z^2$ where $z$ is the direction of the N$-$V axis in the crystal.
+The magnetic field is thus often decomposed as a sum of a longitudinal magnetic field $B_z$ and a transverse magnetic field $B_\perp=\sqrt{B_x^2+B_y^2}$ (see Fig.
+\ref{ODMR 8 classes}-a).
+
+For weak magnetic fields ($\gamma |\textbf{B}| \ll D$), the energy of each of the three spin Hamiltonian eigenstates $\ket{0}$, $\ket{+1}$ and $\ket{-1}$\footnote{The notation $\ket{0}$, $\ket{\pm 1}$ to designate the spin Hamiltonian eigenstates is slightly abusive as $S_z$ does not commute with $\mathcal{H}$ when $B_\perp \neq 0$.
+The true eigenstates however remain close to $\ket{0}$, $\ket{\pm 1}$ as long as $\gamma_e B_z > \frac{(\gamma_e B_\perp)^2}{D}$.} can be approximated in a perturbative approach by: %cf AN.py
+\begin{align}
+\nu_0&= -\frac{(\gamma_e B_\perp)^2}{D}, \nonumber \\
+\nu_{-1} &= D - \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}, \nonumber \\
+\nu_{+1} &= D + \gamma_e B_z + \frac{(\gamma_e B_\perp)^2}{2D}.
+\end{align}
+
+Fig.
+\ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal.
+We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian.
+For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond.} and that each class gives rise to two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).
+
+Fig.
+\ref{ODMR 8 classes}-c) shows a simulation of the evolution of the transition frequencies $\nu_{+1} - \nu_{0}$ and $\nu_{-1} - \nu_{0}$ for all 4 classes as a function of the magnetic field amplitude, the magnetic field orientation being held to an arbitrary direction.
+Fig.
+\ref{ODMR 8 classes}-d) shows an ODMR spectrum with NV ensembles.
+8 lines are visible corresponding to the two transitions of the 4 classes.
+From the position of these 8 lines, it is possible to determine the projection of the magnetic field on each of the 4 N$-$V axes, and thus to reconstruct the full 3D magnetic field with respect to the diamond crystalline axes.
+
+We should also note that the state mixing caused by the transverse magnetic field deteriorates the polarization scheme detailed in Sec.
+\ref{sec ISC}.
+For strong transverse magnetic fields ($\gamma_e B_\perp \sim D$), the optical polarization is almost completely quenched, which prevents any kind of coherent NV manipulation \citep{tetienne2012magnetic}.
+
+\section{Dynamics of the NV center spin}
+
+This section focuses on the dynamics of the NV spin, and more specifically on its various relaxation times.
+Relaxation times are important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing applications \citep{de2021materials, degen2017quantum}.
+
+There are different relaxation times corresponding to different physical aspects.
+The longitudinal relaxation time $T_1$ corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix).
+The transverse relaxation times $T_2$ and $T_2^*$ correspond to the relaxation of the coherences  (off-diagonal terms of the density matrix).
+
+$T_2$ and $T_2^*$ are called the homogeneous and inhomogeneous dephasing times\footnote{We follow here the NMR nomenclature.}.
+$T_2^*$ represents to the absolute dephasing time of the spin (or spin ensemble) coherence, while $T_2$ represent the dephasing time measured when a dynamical decoupling scheme is used (Hahn echo \citep{hahn1950spin}, XY-n \citep{maudsley1986modified}, CPMG \citep{carr1954effects} , ...).
+While the $T_2^*$ value is absolute, the $T_2$ value depend on the precise decoupling scheme used.
+$T_2$ and $T_2^*$ values are ultimately limited by the spin lifetime:
+\begin{equation}
+T_2^* < T_2 < 2 T_1.
+\end{equation}
+
+The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the employed sample.
+The following paragraphs will present the measurement protocols of the different relaxation times and the typical values for the samples used in this manuscript.
+
+\subsection{NV spin $T_1$}
+\label{sec T1 NV}
+\subsubsection{Measurement protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/T1_basique}
+\caption{Basic $T_1$ measurement scheme.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink (see sample details in appendix \ref{Appendix samples}.) for low laser intensity.
+The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
+\label{T1 basique}
+\end{figure}
+
+Fig.
+\ref{T1 basique} shows a basic protocol to measure the NV $T_1$ time.
+Fig.
+\ref{T1 basique}-a) shows the pulse sequence used, which in this case does not involve a microwave field, and Fig.
+\ref{T1 basique}-b) shows the measured PL obtained by varying the dark time $\tau$.
+The curve can then be fitted to obtain the spin ensemble $T_1$ value.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/T1_protocole_soustraction}
+\caption{Common mode rejection $T_1$ measurement scheme.
+(a) Pulse sequence.
+(b) PL signals $S_1$ and $S_2$ obtained on sample CVD-pink with a high laser intensity ($I_{\rm las}=5\ \rm mW$).
+(c) Subtraction $S_1-S_2$ versus dark time $\tau$.
+The experimental data is fitted with an exponential decay of time constant $\tau=4.1\ \rm ms$.}
+\label{T1 soustraction}
+\end{figure}
+
+The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig.
+\ref{T1 soustraction}-a).
+This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$.
+Then the sequence is repeated with the addition of a $\pi$-pulse on one the NV transitions right before the readout.
+This $\pi$-pulse inverts the population of the $\ket{0}$ and $\ket{\pm 1}$ states.
+The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which rejects every contribution to the PL besides the spin relaxation.
+
+Fig.
+\ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample that was used in Fig.
+\ref{T1 basique}, with a higher laser intensity.
+We can notice a stark difference between the $S_1$ curve in Fig.
+\ref{T1 soustraction}-b) and the curve in Fig.
+\ref{T1 basique}-b), despite both of them coming from the same sample and with the same measurement protocol.
+The reason for the difference is that the higher laser intensity used in Fig.
+\ref{T1 soustraction} photoionizes some of the $NV^-$ into NV$^0$ via a two-photon absorption scheme.
+The NV$^0$ can then capture an electron during the dark time \citep{giri2018coupled, giri2019selective} which changes the number of NV$^-$ centers probed depending on the dark time length.
+The goal of the subtraction scheme is to suppress this charge state recombination noise which is common to the two sequences $S_1$ and $S_2$.
+
+Fig.
+\ref{T1 soustraction}-c) shows the subtraction of the $S_1$ and $S_2$ signal.
+Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is satisfyingly fitted by an exponential decay similar to the results found in Fig.
+\ref{T1 basique}.
+
+The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}.
+The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.
+
+\subsubsection{Physical origin and typical values of $T_1$}
+
+We distinguish three different origin for the the $T_1$ relaxation process:
+\begin{itemize}
+\item \textbf{Spin-phonon relaxation:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which can result in a phonon-induced spin relaxation.
+The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}.
+The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxation times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at  mK temperatures \citep{astner2018solid}.
+
+\item \textbf{Electric and magnetic field noise:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}.
+As a result, magnetic and electric field noise can reduce the spin $T_1$ if a significant portion of their power spectral density overlaps with the NV spin transitions.
+This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}.
+The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effects are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).
+
+\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation.
+The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration.
+Value as low as as $25\ \rm{\mu s}$ have been recorded for NV-P1 resonances in a type Ib diamond \citep{hall2016detection}.
+\end{itemize}
+
+All the measurements shown in this manuscript were performed at room temperature.
+The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$.
+Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses.
+The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed.
+$T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.
+
+\subsection{NV spin $T_2^*$}
+\subsubsection{Measurement protocol}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/Ramsey}
+\caption{Ramsey interferometry protocol.
+(a) Pulse seqeunce used.
+(b) Experimental data obtained on sample CVD-pink.} %The orange line corresponds to an Gaussian fit with time constant $\tau=400 \pm 100\ \rm ns$.}
+\label{Ramsey}
+\end{figure}
+
+$T_2^*$ is the characteristic timescale of the loss of quantum coherence for a freely evolving system.
+It is generally defined as the decay time in a Ramsey interferometry measurement \citep{barry2020sensitivity}.
+
+Fig.
+\ref{Ramsey}-a) shows the Ramsey pulse sequence: after being initialized in the $\ket{0}$ state, the NV spin is transformed with a $\pi/2$ pulse in a coherent superposition of $\ket{0}$ and $\ket{\pm 1}$ which freely precesses at the Larmor frequency.
+After a free precession time $\tau$, the remaining coherence is projected back into spin population with a second $\pi/2$ pulse.
+Fig.
+\ref{Ramsey}-b) shows an experimental example obtained on a spin ensemble.
+The microwave frequency is detuned by $\sim 300\ \rm kHz$ with respect to the spin central Larmor frequency which causes oscillations in the signal.
+$T_2^*$ is the decay time of the envelope of the Ramsey fringes.
+It is unfortunately difficult to correctly fit the envelope of a signal with so few fringes, a recurring issue with our current microwave setup.
+
+Ramsey sequences are generally harder to perform on NV ensemble than they are on single NV for two main reasons.
+The first one is that the $T_2^*$ times tend to be longer for single NV centers as they do not suffer from spatial inhomogeneity, and because the samples used for single NV experiments generally contains less impurities than those used for NV ensemble.
+The second one is the inhomogeneous microwave power which leads to a variation of the Rabi frequencies among the spins \citep{barry2016optical, zhou2020quantum} and lowers the fringe contrast.
+
+An alternative measurement of $T_2^*$ can be performed by measuring the linewidth of an ODMR spectrum.
+Indeed, ideally the ODMR spectrum should be the Fourier transform of the Ramsey signal (also called free induction decay in NMR literature).
+In practice however, both methods suffer from experimental noise and imperfections, with Ramsey-type methods being more sensitive to microwave inhomogeneities \citep{barry2020sensitivity} while ODMR is prone to laser and microwave power broadening \citep{dreau2011avoiding}.
+
+The method chosen to extract $T_2^*$ values in this manuscript was ODMR at low microwave power ($\Omega_{\rm Rabi} \ll 1/T_2^*$) and low laser intensity (compared to the NV optical saturation).
+
+\subsubsection{Physical origin of $T_2^*$}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/ODMR_1_classes}
+\caption{ODMR spectra on a single NV class.
+(a) Experimental data on sample CVD-pink (see sample details in appendix \ref{Appendix samples}), fitted by three Lorentzians with HWHM $530$ kHz ($T_2^*=300\ \rm ns$).
+(b) Experimental data on sample Sumi-2, fitted with a single Gaussian with HWHM $3.2$ MHz ($T_2^*=83 ns$).}
+\label{ODMR 1 classe}
+\end{figure}
+
+For most samples, the leading cause of inhomogeneous dephasing is the temporal and spatial variations of magnetic fields caused by the paramagnetic impurities in the crystal \citep{barry2020sensitivity}.
+The nature of these different paramagnetic impurities will be further discussed in the section dedicated to dark spins in diamond in chapter 2.
+We will simply mention here that for samples with high NV density ([NV] $\geq$ ppm), $T_2^*$ is typically limited by the electronic spins of nitrogen impurities \citep{bauch2018ultralong}.
+
+Fig.
+\ref{ODMR 1 classe} shows ODMR spectra of a single NV class on two distinct samples (see sample details in appendix \ref{Appendix samples}).
+The first sample is a CVD sample with a total substitutional nitrogen  concentration $[N_s] \approx 26\ \rm ppm$, while the second is a type-Ib HPHT sample with a nitrogen concentration $[N_s] \geq 100\ \rm ppm$.
+We can clearly see a larger linewidth, corresponding to a smaller $T_2^*$ value on the sample with higher nitrogen concentration.
+In particular, we can resolve the hyperfine structure caused by the $^{14}$N nuclei on the CVD sample, whereas it is hidden by the inhomogeneous broadening on the HPHT sample.
+
+Fig.
+\ref{ODMR 1 classe} illustrates another difficulty of $T_2^*$ measurement which is to find the correct fitting formula.
+While the spectral profile of single NV centers is expected to be Gaussian, the spectral profile of an ensemble of NV centers  is expected to be Lorentzian \citep{dobrovitski2008decoherence, hall2014analytic}.
+For Fig.
+\ref{ODMR 1 classe}-a), it was indeed found that Lorentzians fitted the data better than Gaussians.
+For Fig.
+\ref{ODMR 1 classe}-b) however, a Gaussian was found to be a better fit than a Lorentzian, which could be due to the unresolved hyperfine structure or to other broadening sources such as crystal strain or magnetic field inhomogeneity.
+
+\subsection{NV spin $T_2$}
+\label{sec T2 echo}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=\textwidth]{chapter 1/Figures/Echo}
+\caption{Spin-echo protocol. (a) Pulse sequence. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
+\label{Echo}
+\end{figure}
+\subsubsection{Measurement protocol}
+As precised in the introduction of this section, the exact $T_2$ definition depends on the employed rephasing scheme.
+We will focus here on the simplest rephasing process which is the spin-echo or Hahn-echo sequence.
+
+Fig.
+\ref{Echo}-a) shows the spin-echo pulse sequence: it differs from the Ramsey one by the addition of a $\pi$-pulse in the middle of the free precession time.
+This additional $\pi$-pulse causes static or slowly varying dephasing to be rephased which extends the spin coherence time \citep{hahn1950spin}.
+
+Fig.
+\ref{Echo}-b) shows an experimental result of a spin echo on the same sample that was used for the Ramsay sequence.
+The echo extends the coherence time on this particular sample from $T_2^* \approx 300\ \rm ns$ to $T_2 \approx 4\ \rm{\mu s}$, an increase by more than a factor of 10.
+Improvement by a factor of $10 \sim 100$ or typical for NV ensembles \citep{barry2020sensitivity}.
+%This value could likely be extended even further by using a more complex dynamical decoupling scheme from the CPMG or XY sequence families \citep{naydenov2011dynamical, bar2012suppression}.
+
+\subsubsection{Physical origin of $T_2$}
+The origin of the relaxation time $T_2$ are fluctuating noises which are too fast to be efficiently rephased by the sequence.
+The origin of these noises is generally attributed to the flipping of spins in the vicinity of the NV center (either through spin-lattice or spin-spin mechanisms) \citep{hall2014analytic}.
+For samples with high nitrogen concentration, it has been measured that the $T_2$ time  was directly correlated to the nitrogen concentration, with the same scaling than the $T_2^*$ time \citep{bauch2020decoherence}.
 
@@
 
-We have covered in this introductory chapter every concepts needed to understand the results presented in the next chapters.
-We have seen the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them, and we have introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.
+in this introductory chapter, we have covered the concepts needed to understand the results presented in the next chapters.
+We have detailed the main steps in the creation of artificial diamonds and NV centers, we have covered the optical and spin properties of the NV center as well as the ways to measure them.
+We have also introduced the notion of dipole-dipole coupling and cross-relaxation between two spins.
 
@@
 The authors of the study found a value of $\Gamma_2^*(\rm P1)= (2 \pi) 16\ \rm kHz/ppm$.
-For the samples studied in this manuscript [REF], the P1 concentration was
 
@@
 
-Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\mathbf{B} \parallel [100]$ or 
-$\mathbf{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes.
-When $\mathbf{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\mathbf{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.
+Fig. \ref{121 vs 22 NV} shows the two main choices of scanning orientation : either $\vec{B} \parallel [100]$ or 
+$\vec{B} \parallel [111]$ where [100] and [111] refer to the diamond crystalline axes.
+When $\vec{B} \parallel [100]$, all four classes (the different possible NV orientations represented at the top of the figure) are equivalent, whereas for $\vec{B} \parallel [111]$, one class is perfectly aligned with the magnetic field while the three others are equivalently unaligned.
 
 These changes in the field orientation result in important changes in the NV behavior, mainly due to the transverse magnetic field: Fig.
-\ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the ground state spin Hamiltonian (REF chapter 1) - or to be more precise the transitions between the ground state of the spin Hamiltonian and the two excited states - for any of the four classes of NV centers when $\mathbf{B} \parallel [100]$.
+\ref{121 vs 22 NV}-a) shows the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states of the ground state spin Hamiltonian (REF chapter 1) - or to be more precise the transitions between the ground state of the spin Hamiltonian and the two excited states - for any of the four classes of NV centers when $\vec{B} \parallel [100]$.
 Notably, there is no transitions frequencies below 2500 MHz.
@@
 On the other hand, Fig.
-\ref{121 vs 22 NV}-b) shows the transition frequencies when $\mathbf{B} \parallel [111]$, for both the class aligned and the three class misaligned.
+\ref{121 vs 22 NV}-b) shows the transition frequencies when $\vec{B} \parallel [111]$, for both the class aligned and the three class misaligned.
 The three classes behave similarly than the four classes in the $[100]$ case, but the one class aligned with the magnetic field can get its $\ket{0}\to\ket{-1}$ transition arbitrarily small, as long as the magnetic field is perfectly aligned with the NV axis.
@@
 \ref{121 vs 22 P1} shows the transition frequencies for the NV center and such a spin $1/2$, as a function of a magnetic field parallel to [100] or [111].
-When $\mathbf{B} \parallel [100]$, there will never be a co-resonance between the NV and P1 transitions, meaning that no NV-P1 cross-relaxation can be observed when B is scanned in this direction.
-When $\mathbf{B} \parallel [111]$ however, the P1 $\ket{-1/2} \to \ket{+1/2}$ transition matches in energy the $\ket{0}\to\ket{-1}$ NV transition for B=512 G.
+When $\vec{B} \parallel [100]$, there will never be a co-resonance between the NV and P1 transitions, meaning that no NV-P1 cross-relaxation can be observed when B is scanned in this direction.
+When $\vec{B} \parallel [111]$ however, the P1 $\ket{-1/2} \to \ket{+1/2}$ transition matches in energy the $\ket{0}\to\ket{-1}$ NV transition for B=512 G.
 This means that when B reaches $\approx 512\ \rm G$, some of the polarization of the NV centers will be transferred to the P1 centers, which will result in a drop in the NV PL.
@@
 \ref{121 vs 22 VH} shows the transition frequencies for the NV$^-$ and VH$^-$ ground state spin.
-Contrary to the P1 case, there is a co-resonance when $\mathbf{B} \parallel [100]$ for $B\approx 55\ \rm G$.
-When $\mathbf{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G.
+Contrary to the P1 case, there is a co-resonance when $\vec{B} \parallel [100]$ for $B\approx 55\ \rm G$.
+When $\vec{B} \parallel [111]$, there are 6 co-resonance conditions between 30 and 130 G.
 A scan along the [111] in this case means that the peak signal will be $\sim 6$ times weaker than in the [100] case, and spread over multiple lines which might overlap between themselves or with other existing PL features.
@@
 \subsection{CR condition between NV$^-$ and and $^{13}$C-NV$^-$}
+\label{sec NV-13c-NV}
 
@@
 \ref{121 vs 22 13C-NV}-b) and c) show the simulated transition frequencies for a $^{13}$C-NV$^-$ complex when B is scanned along the [100] or [111] axis, as well as the ``normal" NV transitions and the co-resonance between those two.
-When $\mathbf{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G.
-When $\mathbf{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
+When $\vec{B} \parallel [100]$, there are 4 co-resonances between 18 and 25 G.
+When $\vec{B} \parallel [100]$ there are 18 of them between 10 and 60 G.
 This abundance of transitions for low ($<100\ \rm G$) magnetic field makes the experimental analysis quickly intractable in the second case.
@@
 
-The diamond used for this study is the sample CVD-pink [REF appendix].
+The diamond used for this study is the sample CVD-pink (see sample details in appendix \ref{Appendix samples}).
 The experimental setup used to obtain this data is the standard confocal microscopy setup shown in Sec \ref{Sec setup cahp 1}.
@@
 \includegraphics[width=.6\textwidth]{chapter 2/Figures/scan_100_Sumi4}
-\caption{Change in PL as a function of a magnetic field scanned along the [100] axis for sample SUMI-4}
+\caption{Change in PL as a function of a magnetic field scanned along the [100] axis for sample SUMI-4.}
 \label{scan sumi 4}
@@
 This procedure will be discussed in more details in the next chapter.
-We find here that the linewidth of VH$^-$ and War1 are respectively (half width at half maximum) $5.7\pm 0.6$ and $15.7\pm 1.6$ MHz \footnote{We follow here the pseudo-Voigt deconvolution method described in [REF].
+We find here that the linewidth of VH$^-$ and War1 are respectively (half width at half maximum) $5.7\pm 0.6$ and $15.7\pm 1.6$ MHz \footnote{We follow here the pseudo-Voigt deconvolution method described in the following chapter.
 The linewidth $W=2.7$ MHz (half witdth at half maximum) of the NV centers was measured from ODMR (see Fig.
-[REF]).
+\ref{ODMR 1 classe}).
 It includes all three hyper-fine transistions}.
@@
 \begin{itemize}
-\item Since the change in $T_1$ is proportional to the $T_2^*$ of the probed spins (eq.
-\ref{delta gamma 1}), we should compensate that by integrating over the entire feature, instead of simply looking at the maximum of the line.
+\item Because the change in $T_1$ is inversely proportional to the $T_2^*$ of the probed spins (cf eq.
+(\ref{delta gamma 1})), we should compensate this effect by integrating over the entire area of the peak, instead of simply taking the maximum value of the line.
 \item We have to take into account the loss of contrast of the NV centers as the transverse magnetic field grows, as explained in sec. \ref{CR contrast}.
-\item The $^{13}$C-NV$^-$ are not ``dark spins", and its CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
+\item The $^{13}$C-NV$^-$ are not ``dark spins", and their CR feature cannot be \textit{a priori} compared to the one from VH$^-$ and War1.
 We will see however in the next chapter that we can consider a fraction of the NV centers as dark spins, and that this fraction can be estimated.
@@
 
-When dealing with NV-NV CR, we have to make the distinction between what we will label ``equivalent" and ``non-equivalent" NV centers.
-Two equivalent centers are NV centers with the same spin Hamiltonian ([REF] chap 1), which implies that they both see the same transverse and longitudinal magnetic field.
+We make the distinction between what we will label ``equivalent" and ``non-equivalent" NV centers.
+Two equivalent centers are NV centers with the same spin Hamiltonian (cf eq.
+(\ref{eq.
+spin elec})), which implies that they both see the same transverse and longitudinal magnetic field.
 In contrast, non-equivalent centers are center with different single-spin Hamiltonian, which may result in different eigenstates and spin polarization for the two spins.
@@
 \caption{a) Illustration of the four possible direction of the the NV axis (four classes of NV centers) with a magnetic field along class 1 b) Representation of the polarization of two non equivalent spins, one aligned with the magnetic field (well polarized) and one misaligned with the magnetic field (unpolarized).
-c) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\mathbf{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
+c) Simulated frequencies for the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions when $\vec{B}\parallel [111]$ for the class aligned with B, and the 3 equivalent classes not aligned with B.
 d) Taken from \citep{armstrong2010nv}, PL change of an ensemble of NV centers as a magnetic field is scanned along the [111] axis.}
@@
 \ref{non-equivalent NV-NV}-b) shows the difference between the NV centers belonging to class 1 (labeled NV$_\parallel$) and those belonging to classes 2-4 (labeled NV$_\nparallel$).
-As explained in [REF], the transverse magnetic field leads to a mixing of the eigenstates \footnote{The states of the misaligned NV on Fig.
+As explained in sec.
+\ref{sec champs transverse}, the transverse magnetic field leads to a mixing of the eigenstates \footnote{The states of the misaligned NV on Fig.
 \ref{non-equivalent NV-NV}-b) are labeled $\ket{0},\ket{\pm 1}$ which is an abusive (but common) notation since the mixing caused by the transverse field means that the eigenstates of $S_z$ are no longer the Hamiltonian eigenstates.} and to a loss of the optical polarization mechanism.
@@
 In the case of P-doped Si, the origin of the fast relaxing centers is thought to be closely packed P impurities, so that the electronic wavefunctions of (at least) two P centers overlap.
-In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (see appendix [REF]).
+In that case two phenomena occur: the possibility of electron tunneling, and the apparition of a contact interaction term in the dipole-dipole Hamiltonian (cf eq.
+(\ref{eq.
+dipole quantique J coupling})).
 Both of these phenomena can lead to fast spin relaxation: electron tunneling can be accompanied by a spin reversal \citep{sugihara1963spin}, and the modulation of the contact interaction term ($J-$coupling) by the crystal phonons can directly excite the spin transitions \citep{honig1960electron}.
@@
 \end{equation}
-where the expression of $J_0, g, h$ and $q$ is given in appendix [REF]. $g, h$ and $q$ are dimensionless factors that are function of the relative radial positions and orientation of the two dipoles.
+where the expression of $J_0, g, h$ and $q$ is given in appendix \ref{Appendix eta}.
+$g, h$ and $q$ are dimensionless factors that are function of the relative radial positions and orientation of the two dipoles.
 
@@
 \begin{equation}
-\gamma_s(\mathbf{r})=\left(\frac{J_0}{r^3}\right)^2 \frac{\eta^2}{\gamma_f}.
+\gamma_s(\vec{r})=\left(\frac{J_0}{r^3}\right)^2 \frac{\eta^2}{\gamma_f}.
 \end{equation}
 Compared to eq.
-[REF], the dependency on the inhomogeneous broadening of the spins $1/T_2^*$ is hidden in the $\eta$ factor, and we have the additional broadening $\gamma_f$ coming from the fluctuators very short lifetime.
+(\ref{delta gamma 1}), the dependency on the inhomogeneous broadening of the spins $1/T_2^*$ is hidden in the $\eta$ factor, and we have the additional broadening $\gamma_f$ coming from the fluctuators very short lifetime.
 
@@
 
-The computation of $\bar G$ will be further discussed in sec. \ref{sec quantitative T1} and in appendix [REF]. We will focus here on $\bar R$.
+The computation of $\bar G$ will be further discussed in sec. \ref{sec quantitative T1} and in appendix \ref{delta gamma 1}. We will focus here on $\bar R$.
 
@@
 \includegraphics[width=\textwidth]{chapter 3/Figures/Stretch_vs_pas_stretch}
-\caption{T1 measurement following the protocol described in [REF].
+\caption{T1 measurement following the protocol described in Fig.
+\ref{T1 soustraction}.
 The best exponential and stretched exponential fits are given in every case.
@@
 The first prediction of the fluctuator model which can be experimentally verified is the stretched-exponential profile of the spin relaxation.
-The protocol to measure spin relaxation with a common mode rejection was presented in [REF].
+The protocol to measure spin relaxation with a common mode rejection was presented in Fig.
+\ref{T1 soustraction}.
 
@@
 The spin relaxation results presented in Fig.
-\ref{stretch_or_not_stretch}-a) and b) was performed on sample CVD-pink ([REF]), which despite a relatively high NV concentration shows only moderate amount of dipole-dipole spin relaxation.
+\ref{stretch_or_not_stretch}-a) and b) was performed on sample CVD-pink (see sample details in appendix \ref{Appendix samples}), which despite a relatively high NV concentration shows only moderate amount of dipole-dipole spin relaxation.
 In order to further minimize the amount of cross-relaxations, we chose a magnetic field for which all four NV classes are spectrally isolated, so that only fluctuators from the same class as the NV center could lead to relaxation.
@@
 We took particular care in measuring the central frequencies of the two classes of interest (when the two lines could be clearly separated), since having a precise measurement of the detuning between the two classes $\Delta \nu$ is key to get a precise value of the fluctuator linewidth.
-%Because the $T_1$ measurement protocol described in [REF] requires a resonant microwave pulse with the NV probed, we specifically record ODMR spectra of their two central lines to get a precise measurement of the Larmor frequency.
-In the region where the two ODMR lines overlap we have to use a linear regression to get the Larmor frequency of each class.
 
@@
 \subsubsection{The fluctuators lifetime}
+\label{sec fluctuators lifetime}
 \begin{figure}[h]
@@
 Green curve: deconvolution of the red curve   by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters)}
+All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters).}
 \label{deconvolution}
@@
 \subsection{Geometric conditions for inter-class resonance}
+\label{sec geometrie resonance}
 
@@
 \toprule
-$\Gamma_1^{\rm dd} (\mathbf{B})$ &  Theory & Experimental \\
+$\Gamma_1^{\rm dd} (\vec{B})$ &  Theory & Experimental \\
 \midrule
-random $\mathbf{B}$ (1 class) & $\Gamma_0^{\rm th}$ & $1.53\pm 0.04$ ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & $5.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & $4.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & $11.6 \pm 0.4$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & $14.1 \pm 0.5$ $\Gamma_0^{\rm exp}$ \\
-$\mathbf{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & $19.9 \pm 0.8$ $\Gamma_0^{\rm exp}$ \\
+random $\vec{B}$ (1 class) & $\Gamma_0^{\rm th}$ & $1.53\pm 0.04$ ms$^{-1} \equiv \Gamma_0^{\rm exp}$ \\
+$\vec{B} \in \{110\}$ (2 classes) & 10.0 $\Gamma_0^{\rm th}$ & $5.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B} \in \{100\}$ (2 classes) & 7.24 $\Gamma_0^{\rm th}$ & $4.2 \pm 0.1$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B} \parallel \langle 111 \rangle$ (3 classes) & 28.4 $\Gamma_0^{\rm th}$ & $11.6 \pm 0.4$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B} \parallel \langle 100 \rangle$ (4 classes) & 42.8 $\Gamma_0^{\rm th}$ & $14.1 \pm 0.5$ $\Gamma_0^{\rm exp}$ \\
+$\vec{B}=0$ (4 classes) & 51+($^*$) $\Gamma_0^{\rm th}$ & $19.9 \pm 0.8$ $\Gamma_0^{\rm exp}$ \\
 \bottomrule
@@
 Doing so allows us to normalize $n_f$ and $\gamma_f$ in eq.
-\ref{eq 1/T} and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
-
-$\bar G$ can be analytically or numerically computed for each of the magnetic configurations listed before, as detailed in appendix [REF].
+(\ref{eq 1/T}) and to keep $\bar G$ the geometrical part of $\bar \eta$ as the only unknown.
+
+$\bar G$ can be analytically or numerically computed for each of the magnetic configurations listed before, as detailed in appendix \ref{Appendix eta}.
 This allows us to predict the value of the spin decay rate in every situations as a function of the decay rate of an isolated class.
@@
 
-\chapter{NV-NV cross-relaxation at low fields: application to magnetometry}
+\chapter{NV-NV cross-relaxation at low magnetic field: application to magnetometry}
 
@@
 We are particularly interested in the spin transition linewidth since it directly impacts the dipole-dipole relaxation rate (as seen in sec.
-[REF]): the smaller the linewidth, the higher the dipole-dipole relaxation rate is.
-
-%We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes.
-These change are relevant to our study of NV-NV CR due to the relation between the dipole-induced relaxation rate and $T_2^*$ detailed in sec.
-[REF].
-%
-%A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment.
-In the $\{ \ket{0}, \ket{\pm 1} \}$, the eigen energies are linearly sensitive to (longitudinal) magnetic field, and only sensitive to electric fields at the second order, and vice versa for the $\{ \ket{0}, \ket{\pm} \}$ basis.
-%
-%These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei.
-These effects can drastically modify the ODMR lineshape in zero external magnetic field \citep{jamonneau2016competition} or purely transverse magnetic field \citep{qiu2021nuclear,qiu2022nanoscale}.
-%
-%We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center.
-$^{14}$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin.
-The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
-%\begin{equation}
-%\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
-%\end{equation}
-%
-%where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor.
-$\mathcal{H}_n$ and $\bar{\bar{A}}$ can be written:
-%
-%\begin{equation}
-%\mathcal{H}_n = \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
-%\end{equation}
-%
-%and
-%\begin{equation}
-%\bar{\bar{A}} = \begin{pmatrix}
-%A_{xx} & 0 & 0 \\
-%0 & A_{yy} & 0 \\
-%0 & 0 & A_{zz}
-%\end{pmatrix},
-%\end{equation}
-%
-%
-%where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.
-%
-%Fig.
-\ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$.
-%
-%While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2 contains significantly more impurities besides NV centers, most likely P1 centers.
-These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.
-%
-%Fig.
-\ref{ESR for T2*}-a) allows us to evaluate the magnetic field noise in both samples, since the $\ket{\pm 1}$ states are not sensitive to weak electric fields.
-We indeed find that the Sumi-2 sample has much more magnetic noise, to the point where the hyper-fine structure is no longer resolved.
-The total width of the line however is mostly dominated by the hyper-fine splitting, which result in a similar total linewidth in both cases
-%
-%We should note that the usual practice in magnetometry is to consider the linewidth only of the three hyper-fine lines when they are resolved.
-This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines.
-In our case however, we have to consider the spectral overlap between NV centers and fluctuators, which have an additional broadening of $2\gamma_f \approx 6\ \rm MHz$ [REF] that completely obscures the hyper-fine structure.
-We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.
-%
-%Fig.
-\ref{ESR for T2*}-b) allows us to evaluate the electric field noise in both samples, since the $\ket{\pm}$ states are not sensitive to weak magnetic fields.
-Similarly, the hyper-fine structure is hidden in this configuration since all three hyper-fine levels are nearly-degenerate, provided that $\frac{(\gamma_e B_\perp)^2}{D} > A_{xx},A_{zz},Q$ which is typically the case for $B_\perp > 40\ \rm G$.
-We can note that the electric field noise is significantly stronger in sample Sumi-2, which leads to an ODMR linewidth more than twice as big.
-%
-%Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field.
-For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction).
-For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_s=-1,m_I=+1}$ and $\ket{m_s=+1,m_I=-1}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$.
-The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.
-%
-%The table in Fig.
-\ref{ESR for T2*}-d) reports the linewidths measured on both sample for the various magnetic field configurations.
-We can see that the linewidth of sample CVD-pink varies by more than a factor of 2, while the ones from sample Sumi-2 varied by less than 20 \%.
-We should note that  almost all samples used in this chapter are HPHT type 1b diamond, similar to sample Sumi-2 and unlike sample CVD-pink.
-We should therefore expect the modification of $T_2^*$ to have a moderate impact on the depolarization dynamics.
-%
+\ref{sec computation eta}): the smaller the linewidth, the higher the dipole-dipole relaxation rate is.
+
 \bigskip
@@
 Another consequence is a slight modification of the flip-flop rates.
-Once averaged over all possible orientation and position, as detailed in appendix [REF], this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
-%The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
-%
-%This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis.
-We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively.
-To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
-%
-%The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF].
-The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane.
-This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.
-%
-%We therefore need to make an assumption on the distribution of the transverse field in the sample.
-In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane.
-However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq.
-[REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.
-%
-%We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
-%
-%We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$.
-$\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
-%
-%In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$.
-The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig.
-\ref{scan 100}
+Once averaged over all possible orientation and position, as detailed in appendix \ref{Appendix eta}, this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
 
@@
 The double flip processes have been previously discussed in sec.
-[REF] and naturally arise from the dipole-dipole Hamiltonian.
+\ref{sec flip-flop et double flip} and naturally arise from the dipole-dipole Hamiltonian.
 The ones that we consider here couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$).
@@
 
-The computation of the depolarization rate including the double flips is detailed in appendix [REF].
+The computation of the depolarization rate including the double flips is detailed in appendix \ref{Appendix eta}.
 To obtain a numerical value, we need to input the experimental values of the dipole-dipole interaction range \footnote{By interaction range we mean the value to be compared to the spectral detuning $\Delta \nu$ between the two groups.
 This is $2\gamma_f+\Gamma_{NV}+\Gamma_f$ in eq.
-[REF]} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
-[REF], as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states.
+(\ref{eq 1/T avec T2*})} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
+\ref{fluct linewidth}, as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states.
 We typically measure $\Delta \nu \approx 8\ \rm MHz$ on type Ib HPHT samples.
@@
 For the value presented in Fig. \ref{simus ESR}, we inputted $\Gamma=\Delta \nu=8.8\ \rm MHz$.
-%\begin{table}[htbp]
-%\centering
-%\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
-% \label{table double flips}
-%\begin{tabular}{c|cc}
-%%\toprule
-%{} & $l_c=0$ & $l_c=\infty$ \\
-%\midrule
-%$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
-%$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
-%$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
-%%\bottomrule
-%\end{tabular}
-%\end{table}
-%
-%Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$.
-The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
-%
-%We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field.
-And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field.
-But we can see from Fig.
-\ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig.
-[REF].
-In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.
-%
-%To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one.
-We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF].
-Finally, the question about the correlation  length of the electric field also needs to be asked.
-%
-%The results are summed up in Table \ref{table double flips}.
-We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips.
-$\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$).
-This is close to the the experimental values we have.
-Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $).
-We come back to the results of the last section where only considered the flip-flops.
 
@@
 To quantify the impact of the $T_2^*$ increase in the depolarization rate, we employ eq.
-[REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
+(\ref{eq 1/T avec T2*}), where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig.
 \ref{ESR for T2*}.
@@
 
-On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\mathbf{B}\parallel [100]$) and the influence of the laser polarization.
-Following further experiments detailed in appendix [REF], we have determined that these two effects were negligible compared to the four ones we identified before.
+On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\vec{B}\parallel [100]$) and the influence of the laser polarization.
+Following further experiments detailed in appendix \ref{Appendix autres causes}, we have determined that these two effects were negligible compared to the four ones we identified before.
 
@@
 \begin{itemize}
-\item When $\mathbf{B}=0$.
-\item When $\mathbf{B}\parallel [100]$ (cf. Fig. [REF] (la figure avec toutes les config)).
+\item When $\vec{B}=0$.
+\item When $\vec{B}\parallel [100]$ (cf. Fig. \ref{ODMR_geometries}).
 \end{itemize}
@@
 d) $1/T_1^{\rm dd}$ as a function of the external magnetic field.
-A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
+A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in Fig.
+\ref{various_fit_formulae} with $T_1^{\rm ph}=5\ \rm ms$.}
 \label{scan 1x1x1x1}
@@
 Fig.
-\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as a function of the magnetic field.
+\ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter, as a function of the magnetic field.
 We can observe a decrease of $T_1^{\rm dd}$ by a factor of $\sim 10$ in zero magnetic field compared to non-zero magnetic field, with a half-width at half maximum of $\sim 10\ \rm G$.
-Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\mathbf{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$.
+Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\vec{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$.
 This is because this drop of PL is associated with transverse magnetic field state mixing, and not with spin depolarization.}.
@@
 
-We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF].
+We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in sec.
+\ref{sec NV-13c-NV}.
 We can not explain however why the depolarization rate appear to be smaller at $B \sim 10\ \rm G$ than it is for $B > 30\ \rm G$.
@@
 Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100 G,200 G] seems to satisfy both these criteria: $\Delta \nu\,(100\ \rm G) \approx 30\ \rm MHz \gg \gamma_f$\footnote{Technically, $\Delta \nu$ is to be compared with the dipole-dipole interaction width $2\gamma_f+\Gamma_{NV}+\Gamma_f$ which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig.
-[REF].} and $\abs{\bra{3}\ket{+}}^2\, (200\ \rm G) \approx 0.96$.
+\ref{fluct linewidth}.} and $\abs{\bra{3}\ket{+}}^2\, (200\ \rm G) \approx 0.96$.
 
@@
 \ref{champ tranverse exp}-a) shows an ODMR spectrum for $B=20\ \rm G$ with the two central lines corresponding to the $\ket{0} \to \ket{+}$ and $\ket{0} \to \ket{-}$ transitions of the class orthogonal to the magnetic field.
-We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\mathbf{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
-\ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\mathbf{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}
+We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\vec{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
+\ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\vec{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}
 
@@
 \ref{summary_theory}: those were computed for a 4-class resonance, while we are looking here at a single class.
-The theoretical increase of the depolarization rate for a single class are detailed in Appendix [REF].
+The theoretical increase of the depolarization rate for a single class are detailed in Appendix \ref{Appendix eta}.
 They are:
@@
 \item $\Gamma_1^{\ket{\pm}}=4 \Gamma_0$ when taking into account the change of eigenbasis.
-\item $\Gamma_1^{dd}=[REF]\Gamma_1^{\ket{\pm}}=[REF]\Gamma_0$ when taking into account the double flips.
+\item $\Gamma_1^{dd}=1.6\Gamma_1^{\ket{\pm}}=6.5\Gamma_0$ when taking into account the double flips.
 \end{itemize}
-We have labeled here $\Gamma_0$ the relaxation rate coming from the flip-flop inside of a single class in the $\ket{\pm 1}$ basis (same definition as in table [REF] (table des valeurs chap 3).
+We have labeled here $\Gamma_0$ the relaxation rate coming from the flip-flop inside of a single class in the $\ket{\pm 1}$ basis (same definition as in table \ref{T1 champ mag} (table des valeurs chap 3).
 Experimentally, $\Gamma_0$ corresponds to the baseline shown in Fig.
@@
 Another technique consists in using the $^{13}\rm{C}-$NV complex we discussed in sec.
-[REF] which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are first order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
+\ref{sec NV-13c-NV} which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are first order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
 With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers.
@@
 Fig.
-[REF] in chapter 2 show such PL features, in particular the ground state level anti-crossing (GSLAC, simply labeled LAC on the figure), the sharpest of these features.
+\ref{CR P1 exp} in chapter 2 show such PL features, in particular the ground state level anti-crossing (GSLAC, simply labeled LAC on the figure), the sharpest of these features.
 
@@
 This technique reconstructs the external magnetic field by scanning an other magnetic field, similar to Fig.
-[REF] in chapter 3, and by measuring the position of the NV-NV CR lines.
-By scanning the magnetic field in different directions, one can then recover the initial 3D magnetic field.
+\ref{equivalent NV-NV} in chapter 3, and by measuring the position of the NV-NV CR lines.
+The 3D external magnetic field can then be recovered by scanning the magnetic field in different directions.
 The author of this paper did not measure the sensitivity of this protocol, but they predict a shot noise limited sensitivity of $24.7\ \rm nT/\sqrt{\rm Hz}$.
@@
 To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes.
-This protocol has an very low sensitivity and found limited use so far\citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective}.
+This protocol has a very low sensitivity but is still sometimes used in the absence of alternative \citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective, zhu2022sunlight}.
 
@@
 \caption{Principle of the LFDM protocol a) Experimental setup.
-The confocal microscopy setup is similar to the one described in [REF], the lock-in amplifier (LIA) is used here to add a modulation on the magnetic field, instead of the microwave.
+The confocal microscopy setup is similar to the one described in sec \ref{Sec setup cahp 1}, the lock-in amplifier (LIA) is used here to add a modulation on the magnetic field, instead of the microwave.
 Both the PL and the LIA signal are monitored.
 b) example of LIA signal on sample ADM-15-4 when scanning the magnetic field $B_0$ between -150 and 150 G.
-c) PL signal corresponding to the LIA signal}
+c) PL signal corresponding to the LIA signal.}
 \label{setup magneto}
@@
 \ref{setup magneto} shows the principle of the LFDM protocol.
-The experimental setup is similar to the one presented in Fig.
-[REF] with the exception that there are no microwaves, and that the magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\tilde{B}_{\rm AC}$ used for the modulation.
+The experimental setup is similar to the one presented in chapter 1 without any microwave field.
+The magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\tilde{B}_{\rm AC}$ used for the modulation.
 We typically use a modulation frequency $f_{\rm mod} \approx 1\ \rm kHz$, the same value we use for ODMR measurements.
@@
 The characterization was only done for DC sensing.
-The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
+The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see sample details in appendix \ref{Appendix samples}).
 
@@
 One of the main source of drift for NV magnetometry based on spin resonance is the crystal lattice temperature change: the $D$ factor in the NV Hamiltonian (eq.
-[REF]) is sensitive to the lattice temperature because of the thermal dilatation of the crystal, which is the basis for NV thermometry \citep{kucsko2013nanometre}.
+(\ref{eq.
+spin elec})) is sensitive to the lattice temperature because of the thermal dilatation of the crystal, which is the basis for NV thermometry \citep{kucsko2013nanometre}.
 This means that thermal fluctuations will impact the Larmor frequency of the spins and may cause a detuning with the initial measured frequency for long acquisition times.
@@
 Fig.
-\ref{comparaison 0B MW}-b) shows the LIA signal over a $B_0$ scan: the profile is similar to the one in Fig.
+\ref{comparaison 0B MW}-b) shows the LIA recorded during a $B_0$ scan: the profile is similar to the one in Fig.
 \ref{setup magneto}-b) with the addition to two sharp lines for $|B_0| \approx 100\ \rm G$.
-These two lines correspond to the transition $\ket{0} \to \ket{-1}$ (or $\ket{0} \to \ket{+1}$ for the negative fields) of the class aligned with the magnetic field when it comes to resonance with the 2600 MHz microwave tone.
-The microwave power used here correspond to a Rabi frequency $\Omega \approx (2\pi) 3\ \rm MHz$ and the slope on the steepest part of the microwave line is $\sim 2.5$ times higher than the slope in zero field.
+These two lines correspond to the resonance condition between the microwave tone and the $\ket{0} \to \ket{\pm 1}$ transition of the class parallel to the magnetic field\footnote{We would need to increase the amplitude of the magnetic field in order to see the three other classes.}.
+The microwave power used here correspond to a Rabi frequency $\Omega \approx (2\pi) 3\ \rm MHz$ comparable with $1/T_2^*$ and the slope on the steepest part of the microwave line is $\sim 2.5$ times higher than the slope in zero field.
 
@@
 While we can see that the microwave signal is more sensitive thanks to the higher slope, it is also more prone to drift, most likely because of the shift of the Larmor frequency caused by temperature fluctuations.
-We can even see that the LFDM becomes more sensitive for acquisition times beyond 10 s.
-
-We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the available microwave power (although in this case $\Omega_{\rm Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$.
-Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.
+As a consequence, we can see that the LFDM becomes more sensitive than the microwave-based measurement for acquisition times beyond 10 s.
+
+We should stress that this microwave-based detection is not optimal: the technique used here is analogous to CW ODMR which is not the most sensitive DC magnetometer \citep{barry2020sensitivity}.
+In particular techniques based on double-quantum transition can be both more sensitive and less affected by thermal drifts.
+The sample used here is also poorly suited for traditional microwave magnetometry due to its low $T_2^*$ value.
 
 \subsection{Comparison with microwave-less protocols}
-The only microwave-less protocols to have measured a sensitivity are the GSLAC protocols described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
-Unfortunately the sensing volume in this studies is not known, however due to the similarity in our approaches, we should be able to directly compare the physical origin of both protocols, i.e.
-comparing the zero field PL dip and the GSLAC PL dip.
+The most studied NV-based microwave-less protocols are the GSLAC protocols that we described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
+Unfortunately the sensing volume in this studies is not known, meaning that we can not directly compare the volume normalized sensitivity of both approaches.
+Both GSLAC protocols and LFDM however work under the same basis, which is to exploit the dependency of the NV PL with the magnetic field.
+In order to compare both approach, we can simply focus on the characteristics (width and depth) of the GSLAC PL feature and the zero-field PL feature.
 
@@
 
-The zero field technique is therefore comparable with the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear what kind of material engineering would improve its sensitivity.
+The zero field technique is therefore comparable to the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear what kind of material engineering would improve its sensitivity.
 
@@
 \section{Conclusion and perspectives}
-We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field, and wee have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers.
-We have identified three mechanisms by which the depolarization is increased in zero-field: the inter-class flip-flops, the double flips, and the change in the NV eigenstates (which includes the change in $T_2^*$).We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
-
-Finally, we have seen how this low-field depolarization can be applied to perform magnetometry experiments, and what advantages and disadvantages this technique offers compared to the established NV magnetometry protocols.
+We have explored in this chapter the reasons behind the depolarization of dense NV ensemble at low magnetic field.
+We have identified four mechanisms by which the depolarization is increased in zero-field: the degeneracy of the classes, the double flips, the change in the NV eigenstates and the modification of the spins $T_2^*$.We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
+
+We have also shown that this low-field depolarization can be applied to perform magnetometry experiments.
+We have characterized the sensitivity, temporal stability, and angular dependence of such a magnetometer, and we have compared its performances to the established NV magnetometry protocols.
 
@@
 
-Given the few parameters involved in the LFDM protocol, the main path for LFDM optimization seem to be the material optimization of the samples.
-
-Most magnetometry protocols have a relatively clear material optimization path: DC protocols use the figure of merit $\rho T_2^*$ and AC protocols $\rho T_2$, where $\rho=[\rm NV]$ is the NV concentration.
-The reason is that the number of NV centers for a given volume scales as $\rho$, and therefore the sensitivity scale as $1/\sqrt{\rho}$.
-Meanwhile increasing the density of defects tend to decrease the spin coherence times $T_2$ and $T_2^*$, and the AC and DC sensitivities scale respectively as $1/\sqrt{T_2}$ and $1/\sqrt{T_2^*}$.
-The sensitivity in both of these cases is then inversely proportional to the figure of merit.
-
-For the LFDM protocol however, the material optimization is far from obvious.
-Indeed, while increasing $T_2^*$ would improve the sensitivity, the gain might not be as interesting as for traditional protocols due to the broadening caused by $\gamma_f$.
-On the other hand, increasing the NV density not only increases the number of sensor per volume, but also increases the dipole-dipole strength between NV centers.
-If we assume that fluctuators are constituted of closely packed impurities, then increasing the NV density should also increase the fluctuator proportion.
-
-Increasing the NV density seem to be the path for further material optimization, but other issues may arise in the process.
+One of the main path for LFDM optimization is be the material optimization of the samples.
+
+For traditional DC microwave-based magnetometry, the material optimization path is relatively straightforward: if we assume every NV centers to be independent the sensitivity of a magnetometer for a given sample volume V scales as $1/\sqrt{\rho}$.
+Similarly, according to theory \citep{barry2020sensitivity}, the sensitivity of DC microwave based magnetometers scale as $1/\sqrt{T_2^*}$.
+In order to improve the performance of sample, one should aim to increase the figure of merit $\rho T_2*$.
+
+The same figure of merit cannot be used to characterize the LFDM: the sensitivity of the protocol does not scale as $1/\sqrt{T_2^*}$ due to the spectral broadening of $\gamma_f$, and the density of fluctuators not only impacts the number of sensor per unit volume, but also the density and properties of fluctuators, as well as the average flip-flop rate between NV and fluctuators.
+Further studies with higher NV concentrations than those used in this manuscript will be required to find the correct material optimization path.
 
 We should note that we observed large discrepancy in the zero field depolarization among the samples that we studied, even from samples coming from the same batch.
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=\textwidth]{chapter 4/Figures/substrat}
-%\caption{Measurements on sample SBST-C for two different regions A and B.
-a) PL contrast as a function of a magnetic field scanned in an aribitrary direction (same direction for both regions).
-b) PL contrast as a function of a magnetic field scanned in the [100] direction.
-c) $T_1$ measurement on a single class when $B\neq 0$.
-d) $T_1$ measurement on all  four classes when $B=0$ }
-%\label{substrat chelou}
-%\end{figure}
-%
-%Fig.
-\ref{substrat chelou} shows various measurements on the same sample SBST-C, a HPHT bulk sample containing [NV]$\sim  \rm ppm$, on two spatially distinct regions A and B.
-These two regions most likely correspond to two distinct growth sectors of the diamond, meaning that these regions were not grown in the same exact environment, which can impact among other things the creation of impurities.
-Both region showed a similar PL which would indicate a similar density of NV centers.
-%
-%The metric we are interested in to evaluate the LFDM performance are the PL contrast when $\mathbf{B}$ is scanned in an arbitrary direction, which mainly represents the inter-class flip flop and which I will refer to as FF-contrast; and the PL contrast when B is scanned $\mathbf{B}$ is scanned along the [100] direction, which mainly represents the double flips and which I will refer to as DF-contrast.
-We are interested in the DF-contrast since we have seen in sec.
-\ref{sec angular sensi} that it could be the main contribution to the LFDM sensitivity.
-%
-%Fig.
-\ref{substrat chelou}-a) and b) indicate that region B shows a higher FF-contrast, but a smaller DF-contrast than region A.
-Looking at the $T_1$ measurements for both regions, region B shows a shorter $T_1$ in both zero field and non-zero field, which would indicate a larger fluctuator density in region B.
-If this is true, than the fluctuator density is not directly correlated to the DF-contrast.
 
@@
 Interestingly, we can notice that the $T_1$ of sample ADM-15-1 is both shorter than the one of ADM-15-4 for $B\neq 0$, and longer for $B=0$.
-This could indicate that another relaxation mechanism is competing with the NV-fluctuator one on sample ADM-15-1, which would limit the LFDM sensitivity.
-
-%We now turn to Fig.
-\ref{double dragon}.
-This figure shows the same 4 measurements but this time on two distinct samples, ADM-15-1 and ADM-15-4, which came from the same batch.
-Both samples showed relatively similar PL which could again indicate a similar NV density.
-%
-%We can first notice the disparity in the  zero field PL contrasts: the FF-contrast is twice as big in sample ADM-15-4 as in sample ADM-15-1, and the DF-contrast is almost 5 times bigger.
-When looking at the $T_1$, we can notice that the lifetime for non-zero field is slightly longer for sample ADM-15-4 than for ADM-15-1, but its lifetime in zero-field is shorter (it is also exponential and not stretched-exponential as discussed in the last chapter [REF], which could have an influence on the PL contrast).
-From the $T_1$ measurements we can not determine which sample contains the most fluctuators, and therefore correlate the fluctuator or NV density with the DF and FF-contrasts.
-%
-%These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field.
-We have noticed a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive.
-If these parameters were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.
+This could indicate that another relaxation mechanism is competing with the NV-fluctuator one on sample ADM-15-1, which would limit the LFDM sensitivity on this sample.
 
 Other non-material parameters in the LFDM protocols should also be investigated.
-This includes the temperature of the sample, the intensity and wavelength of the laser, the frequency and amplitude of the AC magnetic field.
+This includes the temperature of the sample, the intensity and wavelength of the laser and the frequency and amplitude of the AC magnetic field.
 
@@
 \includegraphics[width=\textwidth]{chapter 4/Figures/Applications}
-\caption{Schematics of two potential application for LFDM. The working of each application is detailed in the main text}
+\caption{Schematics of two potential application for LFDM. The two application are detailed in the main text.}
 \label{applications}
@@
 \ref{applications} illustrates two potential applications for LFDM.
-The first one is to to perform wide-field magnetic imaging on a surface with micron-sized irregularities.
+The first one consists in the wide-field magnetic imaging on a surface with micron-sized irregularities.
 The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}.
-While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism samples.
-Furthermore, controlling the angle between the diamond and the sample can be challenging even for two perfectly flat surface.
-
-Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always have a sensor within $\sim$ 1 $\mu$m of the sample.
+While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biologic samples \citep{scholten2021widefield}.
+
+Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always keep a sensor within $\sim$ 1 $\mu$m of the sample.
 Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens.
 One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$m red fluorescent diamond) are well suited for this task.
-I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples.
+We have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples.
 This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}.
@@
 It consists in the detection of magnetic field gradient thanks to a nano-diamond probe attached to a nano-resonator.
-The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig.
+The resonator motion transforms the DC gradient in an AC field which can then be probed following the protocol described in Fig.
 \ref{AC and DC sensing}-b).
@@
 If we take a fluctuator proportion of 30 \% as measured in \citep{choi2017depolarization}, then finding a nanodiamond with at least one NV and one fluctuator  seems feasible.
-I have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect, and I therefore cannot estimate the sensitivity of such a probe.
-
-Other potential applications for LFDM includes the possibility to use polycrystalline diamond, potentially grown through heteroepitaxy on top of a surface of interest, or to follow individual diamonds flowing in liquid or living organism, similar to what was done in \citep{feng2021association}, but without having to constantly calibrate the diamond axes through ODMR which significantly hinders the temporal resolution of microwave-based technique.
-
-Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous magnetic field or microwave field along the sample.
+We have not studied the regime with few or single NV and fluctuator, which is in itself an interesting prospect, and we therefore cannot estimate the sensitivity of such a probe.
+
+LFDM could also find applications in biosensing.
+The lack of microwave is a significant advantage in this field to prevent heating of the living organisms, and nano and micro-diamonds are already readily used as trackers for fluorescence microscopy \citep{mohan2010vivo}.
+The use of LFDM could allow to perform background free fluorescence imaging (with a higher sensitivity than what was done with transverse magnetic field \citep{chapman2013background}).
+It could also be used to perform real time magnetometry with diamonds floating in a liquid environment, as was done in \citep{feng2021association}, but without having to constantly calibrate the diamond axes through ODMR which significantly hinders the temporal resolution of microwave-based technique.
+
+Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations ,does not require a homogeneous magnetic field or microwave field along the sample, and can be used with polycrystalline samples which are easier to produce in large size.
+A recent paper demonstrates an energy-efficient and cost-effective NV magnetometer which uses sunlight as the the optical excitation source, and exploits the transverse field dependence of the NV PL to measure the magnetic field\citep{zhu2022sunlight}.
+The use of LFDM instead of transverse field could improve the sensitivity of this magnetometer by an order of magnitude.
 
@@
 \label{Appendix eta}
-We will discuss here the specifics to the computation of the  $\bar \eta$ factor for different geometric configurations, and how to compute the $T_1$ times in table [REF] of chapter [REF], and table [REF] of chapter [REF].
+We will discuss here the specifics to the computation of the  $\bar \eta$ factor for different geometric configurations, and how to compute the $\Gamma_1$ rates in table \ref{T1 champ mag} of chapter 3, and table Fig.
+\ref{summary_theory} of chapter 4.
 
@@
 
-In order to compute $\bar \eta$, we first decompose $\eta$ as a product of $R$ and $G$ as defined in [REF]:
+In order to compute $\bar \eta$, we first decompose $\eta$ as a product of $R$ and $G$ as defined in sec. \ref{sec computation eta}:
 \begin{equation}
@@
 
-The computation of $\bar{R}$ has been discussed in [REF], we will focus here on $\bar{G}$ defined as:
+The computation of $\bar{R}$ has been discussed in sec. \ref{sec computation eta}, we will focus here on $\bar{G}$ defined as:
 
@@
 
-As discussed in sec [REF], there are 4 possibles magnetic field orientations for which at least two NV classes are co-resonant: $\mathbf{B} \in \{110\}$ (two-class resonance), $\mathbf{B} \in \{100\}$ (2$\times$two-class resonance), $\mathbf{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\mathbf{B} \parallel \langle 100\rangle$ (four-class resonance).
-
-Fig.
-\ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$.
+As discussed in sec.
+\ref{sec geometrie resonance}, there are 4 classes of magnetic field orientations for which at least two NV classes are co-resonant: $\vec{B} \in \{110\}$ (two-class resonance), $\vec{B} \in \{100\}$ (2$\times$two-class resonance), $\vec{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\vec{B} \parallel \langle 100\rangle$ (four-class resonance).
+
+Fig.
+\ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\vec{B} \in \{110\}$ and $\vec{B} \in \{100\}$.
 We consider here an NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
-Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\mathbf{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
-
-For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\mathbf{B} \in (100)$ or $\mathbf{B} \in (011)$.
+Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\vec{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
+
+For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\vec{B} \in (100)$ or $\vec{B} \in (011)$.
 The angle $\psi$ between the $\hat z_s$ and $\hat z_f$ axes however is different in both cases.
-The same can be generalized for every $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\mathbf{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\mathbf{B} \in \{110\}$ type resonance.
+The same can be generalized for every $\vec{B} \in \{110\}$ and $\vec{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\vec{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\vec{B} \in \{110\}$ type resonance.
 
 Because $\bar G$ is greater for $\psi=109.5 \ ^\circ$ than it is for $\psi=70.5 \ ^\circ$, then following eq.
-\ref{eq.
-eta bar r bar g bar}, we should expect a faster depolarization rate when $\mathbf{B} \in \{110\}$ than when $\mathbf{B} \in \{100\}$.
-Fig.
-\ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\mathbf{B} \in \{110\}$ scenario and four in the $\mathbf{B} \in \{100\}$ scenario.
-We can clearly see a slower relaxation rate for all four $\mathbf{B} \in \{100\}$ measurements compared to all four $\mathbf{B} \in \{110\}$, which agrees with the predictions of the model.
+(\ref{eq.
+eta bar r bar g bar}), we should expect a faster depolarization rate when $\vec{B} \in \{110\}$ than when $\vec{B} \in \{100\}$.
+Fig.
+\ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\vec{B} \in \{110\}$ scenario and four in the $\vec{B} \in \{100\}$ scenario.
+We can clearly see a slower relaxation rate for all four $\vec{B} \in \{100\}$ measurements compared to all four $\vec{B} \in \{110\}$, which agrees with the predictions of the model.
 
@@
 no detuning between the central frequencies of the different resonant classes), we can assume that $\bar R$ is a constant ($\bar R \equiv \bar R^0 \approx \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}$ according to eq.
-[REF]).
+(\ref{eq 1/T avec T2*})).
 We then have:
@@
 \bar \eta(\rm{1\ class})&=\bar G(\psi=0) \bar R_0 \\[1em]
-\frac{\bar \eta(\mathbf{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
-\frac{\bar \eta(\mathbf{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
-\frac{\bar \eta(\mathbf{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
-\frac{\bar \eta(\mathbf{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
+\frac{\bar \eta(\vec{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\vec{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\vec{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
+\frac{\bar \eta(\vec{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
 \end{align}
 
-Finally, the values in Table [REF] are computed thanks to the relation: \begin{equation}
+Finally, the values in Table \ref{T1 champ mag} are computed thanks to the relation: \begin{equation}
 \frac{\Gamma_1}{\Gamma_0}= \left(\frac{\bar \eta}{\bar \eta(\rm{1\ class})} \right)^2.
@@
 
-Because of the symmetry breaking of the $(xy)$ plane, we now have to consider the relative orientation of $x_1$ and $x_2$ on top of the previously mentioned angles $\theta$, $\phi$ and $\psi$.
-We define $\alpha_1$ and $\alpha_2$ as the relative angle formed by the $x_1$ and $x_2$ axes with an arbitrary axis in their respective $(xy)$ plane (the arbitrary axis can be defined for instance in the lab frame of reference).
-
-Because we assume that the physical orientation of the $x_1$ and $x_2$ axes in zero external magnetic field is determined by a random local electric field, we will sample than angles $\alpha_1$ and $\alpha_2$ on a uniform [0,$2\pi$] distribution independently for each spin.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\textwidth]{chapter 3/Figures/appendix_alpha_s_et_f}
+\caption{Illustration of $\alpha_s$ and $\alpha_f$ defined as the angles formed by $\hat x_s$ and $\hat x_f$ with the reference axes $\hat x_s^0$ and $\hat x_f^0$.
+Blue arrows symbolize the local transverse electric field acting on the NV and the fluctuator respectively.}
+\label{alpha s et f}
+\end{figure}
+
+Because of the symmetry breaking of the $(xy)$ plane, we now have to consider the relative orientation of $x_s$ and $x_f$ on top of the previously mentioned angles $\theta$, $\phi$ and $\psi$.
+We define $\alpha_s$ and $\alpha_f$ in Fig.
+\ref{alpha s et f} as the relative angle formed by the $x_s$ and $x_f$ axes with the arbitrary axes $\hat x_s^0$ and $\hat x_f^0$ in their respective $(xy)$ plane (the arbitrary axis can be defined for instance in the lab frame of reference).
+
+Because we assume that the physical orientation of the $x_s$ and $x_f$ axes in zero external magnetic field is determined by a random local electric field, we will sample than angles $\alpha_s$ and $\alpha_f$ on a uniform [0,$2\pi$] distribution independently for each spin.
 
@@
 \begin{equation}
-\bar G= \frac{1}{\sqrt{3}} \iiiint \abs{3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi} \frac{d \alpha_1}{2\pi} \frac{d \alpha_2}{2\pi}.
+\bar G= \frac{1}{\sqrt{3}} \iiiint \abs{3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f)} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi} \frac{d \alpha_s}{2\pi} \frac{d \alpha_f}{2\pi}.
 \end{equation}
@@
 
-Finally the value $\Gamma_1^{\rm th}= 51.4 \Gamma_0^{\rm th}$ shown in Fig. [REF] was computed from:
+Finally the value $\Gamma_1^{\rm th}= 51.4 \Gamma_0^{\rm th}$ shown in Fig. \ref{summary_theory} was computed from:
 \begin{equation}
@@
 These values are used to compute the double-flip term in Fig.
-[REF]
+\ref{summary_theory}
 
@@
 We will focus here on the case detailed in Fig.
-[REF]: a single class of NV centers whose Hamiltonian is dominated by the transverse magnetic field.
+\ref{champ tranverse exp}: a single class of NV centers whose Hamiltonian is dominated by the transverse magnetic field.
 Unlike the case presented in Sec.

### 2022-10-22 Manuscrit Thèse/These.tex
FINITO LESPAGNOLO

@@
 
-While the NV physics can explain cross-relaxations between non-equivalent NV centers, it cannot explain cross-relaxation between equivalent NV centers.
+While the single NV physics can explain cross-relaxations between non-equivalent NV centers, it cannot explain cross-relaxation between equivalent NV centers.
 Nevertheless, as we will see in this section, CR between both equivalent and non-equivalent NV centers is observed for samples with high NV density.
@@
 \ref{non-equivalent NV-NV}-a) illustrates such a case: we label here the four possible NV axis orientation (or NV classes) 1 to 4 and we assume that the magnetic field is aligned with class 1.
-In this conditions, an NV center belong to class 1 will always be non-equivalent with NV centers from classes 2-4, as the projection of the magnetic field on their respective axis will never be the same.
+In this conditions, an NV center belonging to class 1 will always be non-equivalent with NV centers from classes 2-4, as the projection of the magnetic field on their respective axis will never be the same.
 On the other hand, NV centers from the classes 2-4 are all equivalents, even though they do not necessary belong to the same class.
@@
 \includegraphics[width=\textwidth]{chapter 3/Figures/NV-NV_equivalent}
-\caption{CR between equivalent NV centers for sample ADM-15-1.
+\caption{CR between equivalent NV centers for sample ADM-15-1 (see sample details in appendix \ref{Appendix samples}).
 (a) Representation of equivalent NV centers: either two NVs from the same class, or two classes with the same projected magnetic field.
@@
 b) Average population of the two NV centers in the $\ket{0}$ state in the steady state as a function of $\Gamma_{\rm dd}$ and $\Gamma_1^*$.
-The steady state was computed by solving a classical rate equation where we fixed $\Gamma_{\rm las}=1000\ \rm Hz$ and $\Gamma_1=300\ \rm Hz$.}
+The steady state was computed by solving for a classical rate equation where we fixed $\Gamma_{\rm las}=1000\ \rm Hz$ and $\Gamma_1=300\ \rm Hz$.}
 \label{inhomogene}
@@
 Indeed, if we assume that, for instance, the relaxation rate $\Gamma_1=1/T_1$ is not strictly the same for each NV centers, but instead follows a certain distribution $\rho(\Gamma_1)$, then CR between NV centers with different $\Gamma_1$ can explain the observations in Fig.
-\ref{equivalent NV-NV}.
+\ref{equivalent NV-NV}-d).
 
@@
 Our model here consists of two (resonantly coupled) NV centers with two different relaxation rates $\Gamma_1$ and $\Gamma_1'$.
-To complete the model, we add the rate $\Gamma_{\rm las}$ which is the optical pumping rate from the $\ket{-1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins \footnote{We consider here a two level system instead of the full three levels of the spin 1.
+To complete the model, we add the rate $\Gamma_{\rm las}$ which is the optical pumping rate from the $\ket{-1}$ to the $\ket{0}$ state of each spin, and $\Gamma_{\rm dd}$ the flip-flop rate between the two spins \footnote{We only consider two levels for each spin instead of the full three levels of the spin 1.
 The choice of $\ket{-1}$ instead of $\ket{+1}$ is arbitrary.}.
@@
 We can also notice that the influence of the flip-flop rate on $\rho_{00}$ increases when the difference in decay rates between NV$_1$ and NV$_2$ increases.
-As a result, we can suppose that a greater inhomogeneity between the NV centers would result in a stronger NV-NV cross-relaxation effect.
+As a result, we conclude that a greater inhomogeneity between the NV centers results in a stronger NV-NV cross-relaxation effect.
 
@@
 Choi et al.
-in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime outside of dipole-dipole interaction, and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism leading to a new depolarization rate $\gamma_f$.
+in \citep{choi2017depolarization} take this approach a step further by separating the NV centers into two groups: ``normal" NV centers with a phonon-limited lifetime (not taking into account the dipole-dipole interaction), and ``fluctuators" which are NV centers with an intrinsic, extremely fast relaxation mechanism leading to a new depolarization rate $\gamma_f$.
 Normal and fluctuator NV are represented in Fig.
@@
 \end{equation}
-where the expression of $J_0, g, h$ and $q$ is given in appendix \ref{Appendix eta}.
+where the expression of $J_0, g, h$ and $q$ are given in appendix \ref{Appendix eta}.
 $g, h$ and $q$ are dimensionless factors that are function of the relative radial positions and orientation of the two dipoles.
@@
 \end{equation}
-where $\gamma_f$ is the fluctuator decay rate for each possible channel between $\ket{0}, \ket{-1}$ and $\ket{+1}$.
+where $\gamma_f$ is the fluctuator decay rate for each possible channel between $\ket{0}, \ket{-1}$ and $\ket{+1}$ (cf Fig.
+\ref{NV vs fluct}).
 This number $\eta$ encapsulates both the angular dependency (through $g$ and $h$) and the resonance condition between the NV center and the fluctuator.
@@
 We consider the situation pictured in Fig.
-\ref{NV + bain fluct}: the NV center is at the center of the frame and the position of each fluctuator is labeled $\vec{r}_i$.
+\ref{NV + bain fluct}: the NV center is at the center of the frame of reference and the position of each fluctuator is labeled $\vec{r}_i$.
 We consider that each fluctuator is an independent depolarization source, therefore the depolarization on a single NV center reads:
@@
 
-We will start by separating the factor $\eta$ given in eq. (\ref{eq eta}) into a geometric and a spectral part:
+We will start by rewriting the factor $\eta$ given in eq. (\ref{eq eta}) as the product of a geometric and spectral part:
 \begin{align}
@@
 
-The computation of $\bar G$ will be further discussed in sec. \ref{sec quantitative T1} and in appendix \ref{delta gamma 1}. We will focus here on $\bar R$.
+The computation of $\bar G$ will be further discussed in sec. \ref{sec quantitative T1} and in appendix \ref{Appendix eta}. We will focus here on $\bar R$.
 
@@
 
-In the case where $\omega^0_{NV}=\omega^0_{f}$, meaning that either the considered NVs and fluctuators are from the same class, or from two classes perfectly resonant, by substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f \gg 1/T_2^*$) in eq.
-(\ref{eq 1/T}), we find that:
+By substituting $\overline{R ^2}$ for $(\bar{R})^2$ (which again is only approximately valid for $2 \gamma_f \gg 1/T_2^*$) in eq. (\ref{eq 1/T}), we find that:
 \begin{equation}
@@
 To further enhance the cross-relaxation rate, the experiment was performed in zero external magnetic field so that every class of NV centers would be resonant.
-We can observe a relaxation profile that deviates significantly from tthe exponential profile, and which is nicely fitted by a stretched exponential.
+We can observe a relaxation profile that deviates significantly from the exponential profile, and which is nicely fitted by a stretched exponential.
 The $T_1$ value obtained from the fit is also an order of magnetite lower than the phonon-limited value, which is coherent with the fact that dipole-dipole interactions dominate the spin relaxation in this case.
@@
 For many samples, the $T_1$ profile is neither fully exponential nor fully stretched exponential, because the relaxation time associated with CR, which we will call $T_1^{\rm dd}$, is of the same order of magnitude as the phonon limited lifetime $T_1^{\rm ph}$.
-Sometimes the same sample can go from mostly exponential to mostly stretched exponential depending on the number of NV-NV co-resonances, meaning that we have to include both aspects in the fitting formula.
+Some samples have both predominantly exponential and predominantly stretched exponential profile depending on the number of NV-NV co-resonances.
+We therefore have to include both aspects in the fitting formula.
 
@@
 
-The fits used here are either purely exponential or purely stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$.
+The fit functions investigated here are: pure exponential, stretched exponential, or a combination of both where the exponential lifetime $T_1^{\rm ph}$ was either left as a free parameter or fixed at a value $T_1^{\rm ph}=5\ \rm ms$.
 The values of the different fitting parameters used is reported in Table \ref{fitting table}.
 
-We can see that neither the purely exponential nor stretched exponential fits can be satisfying for both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is poorly fitted by the stretched exponential.
+We can see that neither the pure exponential nor stretched exponential can satisfyingly fit both measurements: the $B=0$ curve is poorly fitted by the exponential fit and the $B \neq 0$ curve is poorly fitted by the stretched exponential.
 The protocols that include both exponential and stretched lifetimes correctly fit both curves.
@@
 Indeed, in the case of Fig.
-\ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms with satisfying fits, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
+\ref{various_fit_formulae}, the value of $T_1^{\rm ph}$ can be somewhat arbitrarily fixed between 3 and 6 ms, and the resulting values $T_1^{\rm dd}$ can vary by almost an order of magnitude in the case of $B\neq0$.
 The values of $T_1^{\rm dd}$ obtained with this method have therefore to be compared to values from the same method and the same fixed value of $T_1^{\rm ph}$.
@@
 Even though this method also yields satisfying fits, it employs more free parameters ($\beta$ and $T_1$) than the method of \ref{various_fit_formulae}-d).
-We therefore decided to go with the later.
+We therefore decided to work with the later.
 
@@
 
-\subsubsection{The fluctuators spectral response}
 \begin{figure}[h]
@@
 b) Measurement of $T_1^{\rm dd}$ as a function of the splitting $\Delta \nu$ between the two central classes, fitted with a Lorentzian of half-width $\sigma=8.8 \pm 0.2\ \rm MHz$.
-The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed line correspond to the same spectrm scaled up by a factor of 2.}
+The green dashed line correspond to the ODMR spectrum of a single class of NV centers, and the red dashed line correspond to the same spectrum scaled up by a factor of 2.}
 \label{fluct linewidth}
@@
 We now move to the second major prediction of the fluctuator model which is the spectral broadening of the fluctuators caused by the their relaxation rate $\gamma_f$.
-Indeed, following the arguments presented in \citep{choi2017depolarization}, the fluctuators homogeneous spectral response is a Lorentzian of half-width $2\gamma_f$.
-The inhomogeneous spectral response of the ensemble of fluctuators is then further broadened by the same electric, magnetic or strain noises than the NV centers, which is characterized by $1/T_2^*$.
+Indeed, following the arguments presented in \citep{choi2017depolarization}, the inhomogeneous spectral response of the fluctuators is broadened by a Lorentzian of half-width $2\gamma_f$ on top of the ``normal" $T_2^*$ broadening resulting from magnetic, electric or strain noise.
 
@@
 
-We can clearly see that the $\Gamma^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
+We can clearly see that the $\Gamma_1^{\rm dd}$ profile is much broader than the ODMR line of the NV centers, whereas in a standard CR process, we expect the decay rate to be proportional to the spectral overlap of the two classes \citep{hall2016detection}.
 If we assume that both NV classes spectral response are Lorentzian of same width $\sigma$, then the spectral overlap between the two classes is also a Lorentzian of width $2 \sigma$ \footnote{For other line profile, the overlap would be smaller than $2 \sigma$.
@@
 
-\subsubsection{The fluctuators lifetime}
-\label{sec fluctuators lifetime}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\textwidth]{chapter 3/Figures/deconvolution}
-\caption{Blue curve $1/T_1^{\rm dd}$ data presented in Fig.
-\ref{fluct linewidth}.
-Red curve: deconvolution of the blue curve by the ODMR line shown on the same figure.
-Green curve: deconvolution of the red curve by the ODMR line.
-All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters).}
-\label{deconvolution}
-\end{figure}
-
-We now want to extract the fluctuators lifetime $T_1^f=1/\gamma_f$ to get further information on their potential nature.
-
-To do this we first assume that the dipole induced decay rate is proportional to the overlap of the spectral responses of the NV centers and fluctuators $S_{NV}$ and $S_f$.
-We will take $S_{NV}( \omega)$ and $S_{f}( \omega)$ as functions centered in $\omega=0$, and consider that the spectral responses are shifted by $\omega^0_{NV}$ and $\omega^0_{f}$ respectively, the central frequencies of each group.
-We then have:
-\begin{align*}
-\frac{1}{T_1^{\rm dd}} &\propto \int S_{NV} (\omega-\omega_{NV}^0) S_f (\omega-\omega_{f}^0) d\omega \\
-&= \int S_{NV} (\omega-\Delta \omega) S_f (\omega)
- d\omega \\
- &= (S_{NV}*S_{f})(\Delta \omega),
-\end{align*}
-
-where $\Delta \omega=\omega_{NV}^0-\omega_{f}^0$.
-
-Using the approximation $\bar {R ^2} \approx (\bar{R})^2$ detailed in sec.
-\ref{sec computation eta}, we can then decompose the fluctuator spectral response as a convolution of the lifetime contribution and the inhomogeneous broadening:
-\begin{equation}
-S_{f}(\omega)=\left(S_f^*(\omega)\right)*\left(\frac{1}{1+\left(\frac{\omega}{2 \gamma_f}\right)^2} \right),
-\end{equation}
-
-where $S_f^*$ is the spectral broadening caused by the inhomogeneous distribution of the fluctuators, which we will suppose to be equal to $S_{NV}$.
-With these assumptions, we are now ready to extract $\gamma_f$ from the data of Fig.
-\ref{fluct linewidth}.
-
-The first way to do this, as was done for NV-P1 CR in \citep{hall2016detection}, is to deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm.
-The result of the deconvolution is shown in Fig.
-\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve).
-Fitting with Lorentzians, we find a final value $\Gamma_2 =2\gamma_f= (2\pi) 6.4 \pm 0.4\ \rm MHz$ which corresponds to a fluctuator lifetime $T_1^f=\frac{1}{\gamma_f}=50 \pm 3\ \rm ns$\footnote{We define here the fluctuators lifetime $T_1^f$ as $\frac{1}{\gamma_f}$ for simplicity, but it should be noted that the actual lifetime of the fluctuators states is $\frac{1}{2\gamma_f}$ due to the competition between the two possible decay channels for each state (cf Fig.
-\ref{NV vs fluct}).}.
-
-Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\frac{\sqrt{\ln (2)}}{\pi T_2^*}$ and a Lorentzian of HWHM $f_L=\frac{2 \gamma_f}{2 \pi}$:
-\begin{align*}
-\Gamma_1^{\rm dd}&\propto \mathcal{G}(f_G)*(\mathcal{L}(f_L)*\mathcal{G}(f_G)) \\
-&\propto \mathcal{L}(f_L)*\mathcal{G}(\sqrt{2} f_G),
-\end{align*}
-where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
-We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{ida2000extended}:
-\begin{equation}
-f = [f_G^5 + 2.69269 f_G^4 f_L + 2.42843 f_G^3 f_L^2 + 4.47163 f_G^2 f_L^3 + 0.07842 f_G f_L^4 + f_L^5]^{1/5}.
-\end{equation}
-
-With $f=8.8 \pm 0.2\ \rm MHz$ and $\sqrt{2}f_G=4.3 \pm 0.2\ \rm MHz$, we find $f_L=6.5 \pm 0.3\ \rm MHz$ for a final lifetime value $T_1^f = 49 \pm 2\ \rm ns$, a value consistent with the one found with the deconvolution method.
-
-Doing similar experiments and calculations, \citep{choi2017depolarization} found $\gamma_f=1/T_1^f=(2\pi) 3.3\ \rm MHz$ which gives them a fluctuator lifetime value $T_1^f = 48\ \rm ns$, a value very close to the ones we obtained above, despite using a completely different sample.
+We proceed to extract the value of $\gamma_f$ from the data presented in Fig.
+\ref{fluct linewidth} by two methods: one involves the deconvolution of the $\Gamma_1^{\rm dd}$ profile by the ODMR profile, and the other involves approximations with pseudo-Voigt profiles.
+Both methods are detailed in appendix \ref{Appendix fluct T1} and result in $\gamma_f$ values of $(2\pi) 3.3 \pm 0.2\ \rm MHz$ and $(2\pi) 3.2 \pm 0.2\ \rm MHz$ respectively, which corresponds to a fluctuator spin lifetime $T_1^f=\frac{1}{3 \gamma_f}\approx 17\ \rm ns$\footnote{The expression of $T_1^f$ as a function of $\gamma_f$ is found by solving for the time dependent solution of the rate equations associated with the fluctuator model shown in Fig.
+\ref{NV vs fluct}.}.
+
+It is worth noting that that \citep{choi2017depolarization} found on their sample $\gamma_f=(2\pi) 3.3\ \rm MHz$ by performing similar experiments, a value very close to the ones we obtained above.
 
@@
 The main tool to probe NV-NV CR is to use the co-resonances between the NV classes.
-While it is not possible to completely shut down the CR, because the NV centers from the same class are always resonant with each other, we can tune the density of resonant NV centers (and resonant fluctuators) by a factor of 4 by playing with the inter-class co-resonances.
+While it is not possible to completely shut down the CR, because NV centers and fluctuators from the same class are always resonant with each other, we can tune the density of resonant NV centers and fluctuators by a factor of 4 by playing with the inter-class co-resonances.
 In this section, we will discuss the various scenarios where NV classes overlap, and quantify theoretically and experimentally the dipole-induced relaxation rate in each case.
@@
 Fig.
-\ref{Carte} shows a map where we monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet on a motorized goniometer.
+\ref{Carte} shows a map where we monitored the PL of sample CVD-pink while scanning the magnetic field angle around the [001] axis, thanks to a permanent magnet mounted on a motorized goniometer.
 We have also reported the locus of $\vec{B}$ being in the planes [100], [010], [110] and [$1\bar 1 0$].
@@
 
-Table \ref{T1 champ mag} shows these predicted values, along with experimental ones on sample ADM-15-3.
+Table \ref{T1 champ mag} shows these predicted values, along with experimental ones measured on sample ADM-15-3 (see sample details in appendix \ref{Appendix samples}).
 The experimental values were obtained with the protocol described in Fig.
@@
 It also correctly predicts that the fluctuators have a broader spectral response than the other NV centers.
-Nevertheless, the model does not describe perfectly every aspects of NV-NV relaxation, and there still are many aspects of this problem, both theoretical and experimental, to explore.
+Nevertheless, the model does not describe perfectly every aspects of NV-NV relaxation.
 
@@
 
-We mentioned previously that the discrepancy between the theoretical and experimental $T_1$ values were in contradiction with the current fluctuator model, although the difference could come in part from the $T_1$ fitting procedure.
-Another, more glaring problem is the fact that some $T_1$ profile are not stretched exponential.
-
-Fig.
-\ref{anormal T1} shows $T_1$ measurement on sample ADM-15-4, either on a single isolated class ,where the dipole-dipole interactions are at their weakest, or on all four resonant classes at $\vec{B}=0$, where the interactions are at their strongest.
-
-Although the spin lifetime still gets $\sim$ 10 times shorter in the $\vec{B}=0$ case, indicating that dipole-dipole interaction still plays an important role in the spin relaxation, this time the lifetime profile does not converge to a purely stretched exponential.
+We mentioned previously that the discrepancy between the theoretical and experimental $T_1$ values was a limitation of the fluctuator model, although the difference could come in part from the $T_1$ fitting procedure.
+Another, more glaring problem is the fact that $T_1$ profiles are not always stretched exponential.
+
+Fig.
+\ref{anormal T1} shows two $T_1$ measurement on sample ADM-15-4: one on a single isolated class (weak dipole-dipole relaxation) and one on all four resonant classes at $\vec{B}=0$ (strong dipole-dipole relaxation).
+
+Although the spin lifetime is $\sim$ 10 times shorter in the $\vec{B}=0$ case, indicating that dipole-dipole interaction still plays an important role in the spin relaxation, on this example the lifetime profile does not converge to a purely stretched exponential.
 In fact it seems to be the opposite: the lifetime profile is closer to a pure exponential for the shorter $T_1$ (especially for the very short times), and closer to a stretched exponential for the longer $T_1$, which is in complete contradiction with the fluctuator model and the previous observations in Fig.
@@
 In the last chapter we have identified a dipole-dipole relaxation mechanism within dense NV centers ensemble, and we have studied its influence in non-zero magnetic field.
-In this chapter we will cover the low to zero magnetic field region and see how the depolarization at low magnetic field can be used for magnetometry.
-
-This chapter is constructed as follows: We will first describe in sec. \ref{sec 4.1} the physics of NV centers under low or transverse magnetic field. In sec. \ref{sec causes zero field}
+In this chapter we will cover the low to zero magnetic field regime and see how the depolarization at low magnetic field can be used for magnetometry.
+
+This chapter is constructed as follows: We will first describe in sec. \ref{sec 4.1} the physics of single NV centers under low or transverse magnetic field. In sec. \ref{sec causes zero field}
 we will present the different zero field depolarization causes that we have identified as well as their theoretical impact on the NV relaxation rates according to the fluctuator model.
@@
 We will start this chapter by studying the single NV physics under low and transverse magnetic field.
-On of the main difference with the previously studied regime is that the eigenstates of the Hamiltonian are no longer $\{ \ket{0}, \ket{+1}, \ket{-1} \}$, which will then in turn modify the dipole-dipole interaction between the NV centers.
-
-We will first describe the spin Hamiltonian of the NV center under low magnetic field and determine the dominant factors in our samples, then do the same for transverse magnetic field, and finally investigate the modification of the ODMR lineshapes in those regimes.
+One of the main difference with the previously studied regime is that the eigenstates of the Hamiltonian are no longer $\{ \ket{0}, \ket{+1}, \ket{-1} \}$, which will then in turn modify the dipole-dipole interaction between the NV centers.
+
+We will first describe the spin Hamiltonian of the NV center under low magnetic field and determine the dominant terms in our samples, then do the same for transverse magnetic field, and finally investigate the modification of the ODMR lineshapes in those regimes.
 
@@
 c) Electric noise: each of the $\vec{E}$ components is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$.
-d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2.}
+d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2 (see sample details in appendix \ref{Appendix samples}).}
 \label{simus ESR}
@@
 \end{equation}
-where $\tan(\phi_E)=E_y/E_x$.
+and where $\tan(\phi_E)=E_y/E_x$.
 
@@
 We now turn to the case of a magnetic field that is orthogonal to the NV axis, i.e.
-$\vec{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$.
+$\vec{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} \ll D$.
 In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.
@@
 the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis, as we established in the previous part.
-Further details on the low magnetic field regime are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.
+Further details on the NV low magnetic field regime are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.
 
@@
 Fig.
-\ref{summary_theory} summarizes these four mechanisms along with the theoretical increase in the depolarization rate caused by each of them.
-We will detail below the physical origin of these four mechanisms and the calculations that lead to the values presented in Fig.
-\ref{summary_theory}.
+\ref{summary_theory} summarizes these four mechanisms along with the theoretical increase in the depolarization rate computed from the fluctuator model for each of them.
+We will detail below the physical origin of these four mechanisms.
+The computation of the theoretical rates will be mostly detailed in appendix \ref{Appendix eta}.
 
@@
 
-Thank to the two previous experiments, we were able to discriminate the contribution of the 4-classes degeneracy from the three other mechanisms that are which are the eigenbasis change, the double-flips and the $T_2^* increase$.
+Thank to the two previous experiments, we were able to discriminate the contribution of the 4-classes degeneracy from the three other mechanisms that are which are the eigenbasis change, the double-flips and the $T_2^*$ increase.
 We now seek to discriminate the contribution of each of these three mechanisms.
 
-We unfortunately were not able to discriminate the change of eigenbasis from the $T_2^* increase$, which would likely require a statistical study over different samples with different $T_2^*$ values.
+We unfortunately were not able to discriminate the change of eigenbasis from the $T_2^*$ increase, which would likely require a statistical study over different samples with different $T_2^*$ values.
 We were however able to discriminate the double flip contribution from the two other ones by scanning the magnetic field in a direction orthogonal to one NV class ($\vec{B} \perp [111]$).
@@
 \ref{champ tranverse exp}-a) shows an ODMR spectrum for $B=20\ \rm G$ with the two central lines corresponding to the $\ket{0} \to \ket{+}$ and $\ket{0} \to \ket{-}$ transitions of the class orthogonal to the magnetic field.
-We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\vec{B}$ was far enough ($> 20 MHz\ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
+We decided to start the scan of the magnetic field at $20\ \rm G$ because the class orthogonal to $\vec{B}$ was far enough ($> 20 \ \rm MHz$) from the three other ones that flip-flop between them could be neglected\footnote{This is confirmed by the small plateau at the beginning of Fig.
+\ref{champ tranverse exp}-c)}.
+
+%\footnote{The small plateau at the beginning of the $1/T_1^{\rm dd}$ curve in Fig.
 \ref{champ tranverse exp}-c) confirms that flip-flop between the class orthogonal to $\vec{B}$ and the three other classes are indeed negligible compared to the other depolarization factors.}.
@@
 \end{itemize}
-We have labeled here $\Gamma_0$ the relaxation rate coming from the flip-flop inside of a single class in the $\ket{\pm 1}$ basis (same definition as in table \ref{T1 champ mag} (table des valeurs chap 3).
+We have labeled here $\Gamma_0$ the relaxation rate coming from the flip-flop inside of a single class in the $\ket{\pm 1}$ basis (same definition as in table \ref{T1 champ mag} and Fig.
+\ref{summary_theory}).
 Experimentally, $\Gamma_0$ corresponds to the baseline shown in Fig.
@@
 There are two classes of NV magnetometer: single NV or ensemble.
-Single NVs offer a spatial resolution limited only by how far the magnetic source can be from the NV center (down to $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
-
-Ensemble NV centers center on the other hand offer sensitivities as low as $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of sizes up to several mm.
+Single NVs offer a spatial resolution limited only by how far the magnetic source can be from the NV center (down to $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically down to $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).
+
+Ensemble NV centers on the other hand offer sensitivities as low as $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of sizes up to several mm.
 
 We will only focus in this section on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR.
-We will not try to give a complete overview of this field, but only of the elements close to the technique we present.
+Furthermore, we will only focus on certain magnetometry aspects relevant to our study.
 For a general overview on NV magnetometry, we recommend the excellent reviews \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.
@@
 \subsection{AC and DC magnetometry}
-The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted into spin population.
+The most sensitive NV magnetometers use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted into spin population and read out.
 These protocols are generally separated between DC-broadband and AC-narrow band.
 
-For DC-broadband magnetometry, the general technique consists of a Ramsey interferometry experiment, presented in chapter 1, the sensitivity of which is limited by the spin inhomogeneous coherence time $T_2^*$ (which can reach up to $\ \rm{\mu s}$ for NV ensembles).
+For DC-broadband magnetometry, the general technique consists of a Ramsey interferometry experiment, presented in chapter 1, the sensitivity of which is limited by the spin inhomogeneous coherence time $T_2^*$ (which can reach up to $\sim 1\ \rm{\mu s}$ for the best NV ensembles).
 The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$.
 Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry.
-The state of the art DC ensemble magnetometr operates with a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
-
-For AC magnetometry, the basic protocol consists of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}.
+The state of the art DC ensemble magnetometer operates with a sensitivity of $\sim 20\ \rm pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.
+
+For AC magnetometry, the basic protocol consists of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, and which can be extended even further to $\sim \rm ms$ times with more sophisticated dynamical decoupling \citep{pham2012enhanced}.
 Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field.
@@
 The technology of flux concentrator could be applied to any NV magnetometry protocol.
-We will here only compare experimental results obtained without a flux concentrator.
+We will here only compare experimental results obtained without them.
 
@@
 \ref{sec NV-13c-NV} which because of the magnetic field offset caused by the $^{13}\rm{C}$ nucleus are first order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}.
-With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers.
+With natural abundance $^{13}\rm{C}$ however, this means that only $\sim 3\%$ of the NV centers are available, and increasing the $^{13}\rm{C}$ concentration can lower the coherence of the NV centers.
 
@@
 b) Histogram of the DC measurement over 50 s.
-The data is fitted with Gaussians of width $1.6\ \rm{\mu T}$.}
+The data is fitted with Gaussians of width $1.5\ \rm{\mu T}$.}
 \label{sensei alternee}
@@
 
-We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\sqrt{\langle \delta V^2 \rangle} \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\sqrt{\langle \delta V^2 \rangle} < 1\ \rm mV$).
-The remaining noise could also come from laser intensity fluctuations.
+The same LIA voltage noise ($\sqrt{\langle \delta V^2 \rangle} \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\sqrt{\langle \delta V^2 \rangle} < 1\ \rm mV$).
+The sensitivity in this case is either shot-noise limited or limited by the laser intensity fluctuations.
 
@@
 We will now study the dependence of the sensitivity with respect to the magnetic field angle.
-In particular, we will compare the sensitivity for two scenarios: the case $\vec{B}\parallel [100]$and the four classes remains degenerate when $\vec{B}$ increases, and the case $\vec{B}\nparallel [100]$ where there is a lift of degeneracy between the four classes.
+In particular, we will compare the sensitivity for two scenarios: the case $\vec{B}\parallel [100]$ for which the four classes remains degenerate when $\vec{B}$ increases, and the case $\vec{B}\nparallel [100]$ for which there is a lift of degeneracy between the four classes.
 We previously saw in Fig.
-\ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\vec{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.
+\ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast were much smaller when $\vec{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.
 
@@
 \ref{comparaison 0B MW} shows a comparison between LFDM and microwave-based magnetometry using our setup.
-A fixed microwave tone (here at 2600 MHz) is sent to the diamond and a magnetic field $B_0$ close to the [111] axis is scanned along an oscillating field $\delta B$ to perform a lock-in detection.
+A fixed microwave tone (here at 2600 MHz) is sent to the diamond and a magnetic field $B_0$ close to the [111] axis is scanned along with an oscillating field $\tilde B_{\rm AC}$ to perform a lock-in detection.
 
@@
 \subsection{Comparison with microwave-less protocols}
-The most studied NV-based microwave-less protocols are the GSLAC protocols that we described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
-Unfortunately the sensing volume in this studies is not known, meaning that we can not directly compare the volume normalized sensitivity of both approaches.
+The most studied NV-based microwave-less protocol is the GSLAC protocol that we described previously \citep{wickenbrock2016microwave, zheng2020microwave}.
+Unfortunately the sensing volume in these studies is not known, meaning that we can not directly compare the volume normalized sensitivity of both approaches.
 Both GSLAC protocols and LFDM however work under the same basis, which is to exploit the dependency of the NV PL with the magnetic field.
-In order to compare both approach, we can simply focus on the characteristics (width and depth) of the GSLAC PL feature and the zero-field PL feature.
+In order to compare both approaches, we can simply focus on the characteristics (width and depth) of the GSLAC PL feature and the zero-field PL feature.
 
 In Fig.
-\ref{setup magneto}-c) we can observe a PL dip with a contrast $C=4.7\pm 1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
+\ref{setup magneto}-c) we can observe a PL dip with a contrast $C=4.7\pm 0.1 \% $ for a full width at half maximum $\Gamma=20 \pm 1\ \rm G$.
 In comparison, in \citep{wickenbrock2016microwave} the authors observe a GSLAC PL dip with a contrast $C=4.5 \%$ for a full width of $12 \ \rm G$ with a type Ib irradiated diamond, similar to the ones we use.
@@
 
-The zero field technique is therefore comparable to the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear what kind of material engineering would improve its sensitivity.
+The zero field technique is therefore comparable to the GSLAC technique for samples with high NV and impurity concentration, but unlike the GSLAC magnetometry, it is still unclear how to improve its sensitivity with material engineering.
 
@@
 We have explored in this chapter the reasons behind the depolarization of dense NV ensemble at low magnetic field.
-We have identified four mechanisms by which the depolarization is increased in zero-field: the degeneracy of the classes, the double flips, the change in the NV eigenstates and the modification of the spins $T_2^*$.We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
+We have identified four mechanisms increasing the depolarization in zero-field: the degeneracy of the classes, the double flips, the change in the NV eigenstates and the modification of the spins $T_2^*$.
+We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.
 
@@
 
-We should note that we observed large discrepancy in the zero field depolarization among the samples that we studied, even from samples coming from the same batch.
+We should note that we observed large discrepancies in the zero field depolarization among the samples that we studied, even from samples coming from the same batch.
 
@@
 The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}.
-While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biologic samples \citep{scholten2021widefield}.
+While the diamond surface can be polished within a few nm, the same is not always true for the sample, especially for paleomagnetism or biologic samples \citep{scholten2021widefield}.
 
@@
 Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations ,does not require a homogeneous magnetic field or microwave field along the sample, and can be used with polycrystalline samples which are easier to produce in large size.
-A recent paper demonstrates an energy-efficient and cost-effective NV magnetometer which uses sunlight as the optical excitation source, and exploits the transverse field dependence of the NV PL to measure the magnetic field\citep{zhu2022sunlight}.
+A recent paper demonstrated an energy-efficient and cost-effective NV magnetometer which uses sunlight as the optical excitation source, and exploits the transverse field dependence of the NV PL to measure the magnetic field\citep{zhu2022sunlight}.
 The use of LFDM instead of transverse field could improve the sensitivity of this magnetometer by an order of magnitude.
@@
 
-Dipole-dipole interaction within dense ensemble of NV centers ([NV] > ppm) can be responsible for an increase of the depolarization rate by more than a factor of 10.
+Dipole-dipole interaction within dense ensemble of NV centers ([NV] $>$ ppm) can be responsible for an increase of the depolarization rate by more than a factor of 10.
 This changes considerably the properties of NV ensemble compared to single NV.
@@
 \subsection{CVD-B and CVD-C}
-The growth process and characterization of sample CVD pink are detailed in \citep{ngambou2022improving}.
+The growth process and characterization of samples CVD-B and CVD-C are detailed in \citep{ngambou2022improving}.
 These two samples were grown with a N$_2$O concentration of 100 ppm in the gas phase, and were irradiated similarly to sample CVD-pink.
@@
 \section{HPHT samples}
-\subsection{Adamas fluorescent microdimaonds}
+\subsection{Adamas fluorescent microdimaonds (ADM-15-$n$ and ADM-150-$n$)}
 \begin{table}[htbp]
@@
 The two samples were further annealed at 800 $^\circ$C for two hours.
-The NV and N concentration were not measured on this sample.
+The NV and N concentration were not measured on these samples.
 \subsubsection{SBST-B and SBST-C}
@@
 The relative position between the two spins is noted with the vector $\vec{r}=r\hat{u}$.
-The vector's spherical coordinate are given in the Cartesian basis of the NV center ($\hat x_s,\hat y_s,\hat z_s$).
+The vector's spherical coordinate are given in the Cartesian basis of the NV center ($\hat x_s,\hat y_s,\hat z_s$) with the usual definitions for the polar and azimuthal angles.
 
@@
 The arbitrary choice of the $\hat x$ direction is justified by the symmetry of the problem in the ($\hat x, \hat y$) plane.
-We chose here to neglect the symmetry-breaking part of the transverse magnetic field which is justified by the fact that we do not take into consideration the mixing of the eigenstates caused by the transverse field.
+We chose here to neglect the symmetry-breaking part of the transverse magnetic field which is justified by the assumption that $\gamma B_z \gg \frac{(\gamma B_\perp)^2}{D}$ for both spins.
 
@@
 We consider here an NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis.
-Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\vec{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.
+Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\vec{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and $\hat z_f=[\bar 1 11]$, and in in the right cases $\hat z_s=[\bar 1 \bar 1 \bar 1]$ and $\hat z_f=[\bar 1 11]$.
 
@@
 
-In this part we will study the averaged flip-flop and double flip rates in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis which corresponds to the various mechanisms discribed in chapter 4.
+In this part we will study the averaged flip-flop and double flip rates in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis which is relevant for the low and transverse field regime as described in chapter 4.
 
@@
 \label{sec flip-flop +/-}
-Rewriting the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-}\} \times \{ \ket{0}, \ket{+}, \ket{-}\}$ changes the value of the flip-flop and double flip prefactors.
+Rewriting the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-}\} \times \{ \ket{0}, \ket{+}, \ket{-}\}$ basis changes the value of the flip-flop and double flip prefactors.
 For instance, the prefactor of the $\ket{+,0}\bra{0,+}$ term is: \begin{equation}
@@
 
-Because we assume that the physical orientation of the $x_s$ and $x_f$ axes in zero external magnetic field is determined by a random local electric field, we will sample than angles $\alpha_s$ and $\alpha_f$ on a uniform [0,$2\pi$] distribution independently for each spin.
+Because we assume that the physical orientation of the $x_s$ and $x_f$ axes in zero external magnetic field is determined by a random local electric field, we will sample the angles $\alpha_s$ and $\alpha_f$ on a uniform [0,$2\pi$] distribution independently for each spin.
 
@@
 \subsection{Double-flips in the $\{ \ket{0}, \ket{+}, \ket{-}\}$ basis}
-Including the double flips, there are two possible relaxation path for each transition of the regular NV center.
+Including the double flips, there are two possible relaxation path for each transition of the ``normal" NV center.
 For instance the transition $\ket{0} \to \ket{-}$ can involve either the flip flop $\ket{0,-} \bra{-,0}$ or the double-flip $\ket{0,+} \bra{-,0}$.
@@
 Eq.
-(\ref{eq eta2 avec DQ}) can no longer be factorized as a product of a spectral and spatial function.
-To compute the value of $\bar eta$ in zero magnetic field, we will proceed to the following assumptions: $\Delta_{\omega_{NV}}^{(0,\pm)}=\Delta_{\omega_{f}}^{(0,\pm)}$, and $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$\footnote{This second assumption is justified by experimental measurements.}.
+(\ref{eq eta2 avec DQ}) can no longer be factorized as a product of a spectral and spatial functions.
+To compute the value of $\bar \eta$ in zero magnetic field, we will proceed to the following assumptions: $\Delta_{\omega_{NV}}^{(0,\pm)}=\Delta_{\omega_{f}}^{(0,\pm)}$, and $|\Delta_{\omega_{NV}}^{(0,\pm)}-\Delta_{\omega_{f}}^{(0,\mp)}|=2 \gamma_f$\footnote{This second assumption is justified by experimental measurements.}.
 Eq.
@@
 
-The expression of $\bar G$ then becomes:
+The expression of $\bar G$ when considering only the flip-flops then becomes:
 \begin{equation}
@@
 \end{equation}
+
+\chapter{Extraction of the fluctuator depolarization rate $\gamma_f$}
+
+\label{Appendix fluct T1}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.6\textwidth]{chapter 3/Figures/deconvolution}
+\caption{Blue curve $1/T_1^{\rm dd}$ data presented in Fig.
+\ref{fluct linewidth}.
+Red curve: deconvolution of the blue curve by the ODMR line shown on the same figure.
+Green curve: deconvolution of the red curve by the ODMR line.
+All three curves are fitted with Lorentzian of respective half widths $8.8 \pm 0.2$, $7.5 \pm 0.3$ and $6.4 \pm 0.4$ MHz (the precision is determined by the variance of the optimal fit parameters).}
+\label{deconvolution}
+\end{figure}
+
+We will detail here the process used to extract the fluctuator depolarization rate $\gamma_f$ from the data presented in Fig. \ref{fluct linewidth}.
+
+To do this we first assume that the dipole induced decay rate is proportional to the overlap of the spectral responses of the NV centers and fluctuators $S_{NV}$ and $S_f$.
+We will take $S_{NV}( \omega)$ and $S_{f}( \omega)$ as functions centered in $\omega=0$, and consider that the spectral responses are shifted by $\omega^0_{NV}$ and $\omega^0_{f}$ respectively, the central frequencies of each group.
+We then have:
+\begin{align*}
+\frac{1}{T_1^{\rm dd}} &\propto \int S_{NV} (\omega-\omega_{NV}^0) S_f (\omega-\omega_{f}^0) d\omega \\
+&= \int S_{NV} (\omega-\Delta \omega) S_f (\omega)
+ d\omega \\
+ &= (S_{NV}*S_{f})(\Delta \omega),
+\end{align*}
+
+where $\Delta \omega=\omega_{NV}^0-\omega_{f}^0$.
+
+Using the approximation $\bar {R ^2} \approx (\bar{R})^2$ detailed in sec.
+\ref{sec computation eta}, we can then decompose the fluctuator spectral response as a convolution of the lifetime contribution and the inhomogeneous broadening:
+\begin{equation}
+S_{f}(\omega)=\left(S_f^*(\omega)\right)*\left(\frac{1}{1+\left(\frac{\omega}{2 \gamma_f}\right)^2} \right),
+\end{equation}
+
+where $S_f^*$ is the spectral broadening caused by the inhomogeneous distribution of the fluctuators, which we will suppose to be equal to $S_{NV}$.
+With these assumptions, we are now ready to extract $\gamma_f$ from the data of Fig.
+\ref{fluct linewidth}.
+
+The first way to do this, as was done for NV-P1 CR in \citep{hall2016detection}, is to deconvolve the $1/T_1^{\rm dd}$ data by the ODMR spectrum using a Wiener deconvolution algorithm.
+The result of the deconvolution is shown in Fig.
+\ref{deconvolution}: the data is deconvolved once to get the spectral response of the fluctuator (red curve), and a second time to get only the lifetime contribution of the fluctuator (green curve).
+Fitting with Lorentzians, we find a final value $\Gamma_2 =2\gamma_f= (2\pi) 6.4 \pm 0.4\ \rm MHz$.
+
+Another, simpler alternative is to consider that the $1/T_1^{\rm dd}$ profile is a Voigt profile given by the convolution of two Gaussians of half width at half maximum (HWHM) $f_G=\frac{\sqrt{\ln (2)}}{\pi T_2^*}$ and a Lorentzian of HWHM $f_L=\frac{2 \gamma_f}{2 \pi}$:
+\begin{align*}
+\Gamma_1^{\rm dd}&\propto \mathcal{G}(f_G)*(\mathcal{L}(f_L)*\mathcal{G}(f_G)) \\
+&\propto \mathcal{L}(f_L)*\mathcal{G}(\sqrt{2} f_G),
+\end{align*}
+where $\mathcal{G}(f_G)$ and $\mathcal{L}(f_L)$ denote Gaussian and Lorentzian functions of HWHM $f_G$ and $f_L$ respectively.
+We can then use the formula of a pseudo-Voigt profile which gives an approximation of the Voigt profile's HWHM $f$ as a function of $f_G$ and $f_L$ \citep{ida2000extended}:
+\begin{equation}
+f = [f_G^5 + 2.69269 f_G^4 f_L + 2.42843 f_G^3 f_L^2 + 4.47163 f_G^2 f_L^3 + 0.07842 f_G f_L^4 + f_L^5]^{1/5}.
+\end{equation}
+
+With $f=8.8 \pm 0.2\ \rm MHz$ and $\sqrt{2}f_G=4.3 \pm 0.2\ \rm MHz$, we find $f_L= 2 \gamma_f =6.5 \pm 0.3\ \rm MHz$.
 
@@
 We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$.
-We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.
+We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects associated with the fluctuators.
 

### 2022-11-04 Talks/Talk Basel 2022/presentation.tex

--- 2022-10-17 Talks/Talk Stuttgart/presentation.tex
@@
 %\section{Physics of the NV center}
-\begin{frame}{Preamble : the NV center}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide  NV properties}
-\end{frame}
+%\begin{frame}{Preamble : the NV center}
+%\centering
+%\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide  NV properties}
+%\end{frame}
 
@@
 
-\begin{frame}{Subject of this presentation}
+\begin{frame}{Context: Levitation of micro-diamonds in ion traps}
+\centering
+\includegraphics[width=0.9\textwidth,height=0.9\textheight,keepaspectratio]{Shema_meca}
+\end{frame}
+
+\begin{frame}{Context: Levitation of micro-diamonds in ion traps}
+\centering
+\includegraphics[width=0.9\textwidth,height=0.9\textheight,keepaspectratio]{Shema_meca_barre}
+\end{frame}
+
+\begin{frame}{Depolarization of dense NV ensemble at low magnetic field}
 \centering

### 2022-11-04 python_clément/analyse.py
tata

@@ -195,7 +195,6 @@ def gauss_derivative_fit(x,y,amp=None,x0=None,sigma=None,ss=0):
 	popt, pcov = curve_fit(f, x, y, p0)
 	return(popt,f(x,*popt))
 
-
 def lor_fit(x,y,amp=None,x0=None,sigma=None,ss=None,err=False) :
 	#sigma=HWHM
 	if not ss :
@@ -233,7 +232,6 @@ def lor_fit_fixed(x,y,amp=None,x0=0,sigma=1,ss=None):
 	popt, pcov = curve_fit(f, x, y, p0)
 	return(popt,f(x,*popt))
 
-
 def cos_fit(x,y,amp=None,omega=None,phi=0,ss=None):
 	if not amp :
 		amp=max(y)-min(y)
@@ -256,7 +254,6 @@ def invert_fit(x,y,amp=None):
 	p0=[amp]
 	popt, pcov = curve_fit(f, x, y, p0)
 	return(popt,f(x,*popt))
-
 
 def exp_fit(x,y,amp=None,ss=None,tau=None,err=False) :
 	if not amp :
@@ -294,7 +291,6 @@ def exp_fit_zero(x,y,amp=None,tau=None,norm=False,err=False) :
 		return(popt,f(x,*popt),np.sqrt(np.diag(pcov)))
 	else :
 		return(popt,f(x,*popt))
-
 
 def stretch_exp_fit(x,y,amp=None,ss=None,tau=None,err=False) :
 	if not amp :

### 2022-11-04 python_clément/lab.py

@@ -714,8 +714,6 @@ class fitButton():
 		self.addFitButton.addToBox(box)
 		self.removeFitButton.addToBox(box)
 
-
-
 class device(): #Pour l'instant ça sert juste à les regrouper pour les fermer
 	def __init__(self):
 		self.toBeClosed=True

### 2022-11-22 Talks/Soutenance 2022/presentation.tex
la tete a toto

--- 2022-10-17 Talks/Talk Stuttgart/presentation.tex
@@
 \usepackage{tcolorbox} % toujours plus de boites
+\usepackage{tabularx}
 %\usepackage[backend=biber]{biblatex}
@@
 
-\title{Low field depolarization of dense NV centers ensemble: application to magnetometry}
-\author{Clément Pellet-Mary, Maxime Perdriat, Paul Huillery,\\ Gabriel Hétet}
-\date{Nano-optics group}
+\title{Cross-relaxation in dense ensembles of NV centers and
+application to magnetometry}
+\author{Clément Pellet-Mary}
+\date{PhD Defense}
 
@@
 \maketitle
-\begin{center}
-\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{logos}
-\end{center}
-\end{frame}
-%\begin{frame}{Outline}
-%\tableofcontents
-%\end{frame}
-%\section{Physics of the NV center}
-\begin{frame}{Preamble : the NV center}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide  NV properties}
-\end{frame}
-
-\begin{frame}{Preamble : the 4 classes of NV centers}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide ODMR 4 classes}
-\end{frame}
-
-\begin{frame}{Subject of this presentation}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_presentation_sujet}
+%\begin{center}
+%\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{logos}
+%\end{center}
 \end{frame}
@@
 
-\section{Cross-relaxation with NV centers}
+\section{Sensing with quantum mechanics}
+\begin{frame}{Outline}
+\tableofcontents[currentsection]
+\end{frame}
+
+\begin{frame}{Quantum sensing and metrology}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_quantum_metrology}
+\end{frame}
+
+\begin{frame}{Magnetometry applications}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_example_magnetometry}
+\end{frame}
+
+\begin{frame}{Quantum magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_quantum_magnetometers}
+\end{frame}
+
+\section{Magnetometry with NV centers}
+\begin{frame}{Outline}
+\tableofcontents[currentsection]
+\end{frame}
+
+\begin{frame}{Colored centers in diamond}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide diamant}
+\end{frame}
+
+\begin{frame}{Synthetic diamond and NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_fab_sample}
+\end{frame}
+
+\begin{frame}{The NV center energy levels}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_NV_levels_1}
+\end{frame}
+
+\begin{frame}{The NV center energy levels}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_NV_levels}
+\end{frame}
+
+\begin{frame}{Optically detected magnetic resonance (ODMR)}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_ODMR_0}
+\end{frame}
+
+\begin{frame}{Experimental setup}
+\centering
+\includegraphics[width=0.9\textwidth,height=0.9\textheight,keepaspectratio]{Slide_setup}
+\end{frame}
+
+\begin{frame}{ODMR with NV ensemble}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_ODMR_8_classes}
+\end{frame}
+
+\begin{frame}{Sensitivity limit for NV ensemble magnetometry}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_NV_ensemble_sensi}
+\end{frame}
+
+\begin{frame}{Dipole-dipole interaction between two electronic spins}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_dipole_dipole}
+\end{frame}
+
+\begin{frame}{Interaction limit to NV ensemble magnetometry}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_interaction_limit}
+\end{frame}
+
+\section{Low field depolarization magnetometry (LFDM)}
+\begin{frame}{Outline}
+\tableofcontents[currentsection]
+\end{frame}
+
+\begin{frame}{Depolarization of dense NV ensemble at low magnetic field}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_presentation_sujet}
+\end{frame}
+
+\begin{frame}{Pesentation of low field depolarization magnetometry (LFDM)}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_principle_LFDM}
+\end{frame}
+
+\begin{frame}{Sensitivity of LFDM}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_sensi_LFDM}
+\end{frame}
+
+\begin{frame}{Comparison with the state of the art}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_litterature}
+\end{frame}
+
+\begin{frame}{Comparison with CW ODMR}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_microwave}
+\end{frame}
+
+\begin{frame}{Application: wide-field magnetometry on irregular surfaces}
+\centering
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_applications_wide_field}
+\end{frame}
+
+\section{Depolarization mechanisms in dense NV ensemble}
 \begin{frame}{Outline}
@@
 \end{frame}
-\section{The NV-fluctuator model and experimental verification}
-\begin{frame}{Outline}
-\tableofcontents[currentsection]
-\end{frame}
+
 \begin{frame}{Presentation of the fluctuator model}
@@
 \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro}
-\end{frame}
-
-\begin{frame}{Predictions of the fluctuator model}
-\begin{itemize}
-\item $\Gamma_1$ increases when classes overlap spectrally (increase in the resonant fluctuator density).
-\item The dipole induced depolarization has a stretched exponential profile:
-$$ \rho_{00}(t) \propto \exp(-\sqrt{\frac{t}{T_1}})$$
-\item The Fluctuators spectral response is broadened by their decay rate $\gamma_f$ (lifetime limit).
-\end{itemize}
-\end{frame}
-
-\begin{frame}{$T_1$ measurement protocol}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_T1_protocol}
 \end{frame}
@@
 
-\begin{frame}{Competition between stretched and exponential decay}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_exp_et_stretch}
-\end{frame}
-
-\begin{frame}{Spectral response of the fluctuators}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_fluct_linewidth}
-\end{frame}
-
-\begin{frame}{Geometry conditions for class resonances}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_geométries}
-\end{frame}
-
-\begin{frame}{PL mapping of the crystalline planes}
-\centering
-\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_carte}
-\end{frame}
-
-\begin{frame}{Limitations of the fluctuator model}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_limit_fluct}
-\end{frame}
-
-\section{Dipole-dipole interaction under low magnetic field}
-\begin{frame}{Outline}
-\tableofcontents[currentsection]
-\end{frame}
 \begin{frame}{Zero field depolarization sources (theory)}
@@
 
-\section{Low field depolarization magnetometry (LFDM)}
-\begin{frame}{Outline}
-\tableofcontents[currentsection]
-\end{frame}
-\begin{frame}{Principle of LFDM}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_principle_LFDM}
-\end{frame}
-
-\begin{frame}{Sensitivity of LFDM}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_sensi_LFDM}
-\end{frame}
-
-\begin{frame}{Comparison with the state of the art}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_litterature}
-\end{frame}
-
-\begin{frame}{Comparison with CW ODMR}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_microwave}
-\end{frame}
-
 \begin{frame}{Angular sensitivity of LFDM}
@@
 
-\begin{frame}{Application: wide-field magnetometry on irregular surfaces}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_applications_wide_field}
+\begin{frame}{Conclusion}
+
 \end{frame}
 
-\begin{frame}{Other applications}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_application_autre}
-\end{frame}
-
-\begin{frame}{Take home messages}
-\begin{itemize}
-\item The spin depolarization in dense NV ensemble is dominated by dipole-dipole interaction with fast decaying centers (fluctuators).
-\medskip
-\item The leading cause of depolarization in zero field is the spectral overlap of the NV classes, followed by the double flips.
-\medskip
-\item Low field depolarization magnetometry achieves a sensitivity $\sim 6\ \mu\rm T \cdot \mu \rm m^{3/2} / Hz^{1/2}$ with commercial samples.
-\medskip
-\item LFDM can be used with powder or polycristalline samples.
-\end{itemize}
+\begin{frame}{Acknowledgments}
 

### 2022-11-22 data/20211015/Le bon adamas/traitement.py

@@ -4,7 +4,7 @@ sys.path.append('C:\\Users\\cleme\\OneDrive\\Documents\\these\\python_clément')
 sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
-plt.figure(num=1,figsize=(4.5,3),dpi=80)
+plt.figure(num=1,figsize=(4,3),dpi=80)
 plt.xticks(fontsize=16)
 plt.yticks(fontsize=16)
 plt.locator_params(axis='x', nbins=5)
@@ -30,8 +30,13 @@ plt.locator_params(axis='y', nbins=5)
 # print(popt)
 # plt.show()
 
+# fname='ESR 0B zoom +10 dBm avec un T et 50 ohms'
 fname='Vrai ESR 1x1x1x1'
 x,y=extract_data(fname)
+# plt.ylim(0.96,1.0015)
+# cs=[2624,3141]
+# popt,yfit=ESR_n_pics(x,y,cs)
+# plt.plot(x,yfit)
 plt.plot(x,y)
 
 plt.show()

### 2022-11-29 Talks/Soutenance 2022/presentation.tex
manger

@@
 
-\title{Cross-relaxation in dense ensembles of NV centers and
+\title[PhD Defense]{Cross-relaxation in dense ensembles of NV centers and
 application to magnetometry}
@@
 
-\mode<presentation> {\usetheme{Rochester}}
+\usetheme{Malmoe}
+\usecolortheme{seahorse}
 
@@
 
+\subsection{Quantum sensing and magnetometry}
 \begin{frame}{Quantum sensing and metrology}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_quantum_metrology}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_metrology}
 \end{frame}
 
-\begin{frame}{Magnetometry applications}
+\begin{frame}{Magnetometry key parameters}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_example_magnetometry}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_magnetometer_size}
 \end{frame}
 
-\begin{frame}{Quantum magnetometers}
+\begin{frame}{Sate of the art magnetometers}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_quantum_magnetometers}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_bare}
 \end{frame}
 
-\section{Magnetometry with NV centers}
+\begin{frame}{Sate of the art magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_OPM}
+\end{frame}
+
+\begin{frame}{Sate of the art magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_SQUID}
+\end{frame}
+
+\begin{frame}{Sate of the art magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV}
+\end{frame}
+
+\subsection{NV center magnetometry}
+\begin{frame}{NV center overview}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_rapido}
+\end{frame}
+
+\begin{frame}{NV center magnetometry sensitivity}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_limite_sensi}
+\end{frame}
+
+\begin{frame}{NV center magnetometry: the interaction limit}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_interaction_limit}
+\end{frame}
+
+\section{Diamond and NV centers in practice}
+\subsection{Diamond properties}
 \begin{frame}{Outline}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide diamant}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide diamant}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_fab_sample}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_fab_sample}
+\end{frame}
+
+\subsection{NV center energy levels}
+\begin{frame}{The NV center energy levels}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_levels_1}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_NV_levels_1}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_levels}
 \end{frame}
 
-\begin{frame}{The NV center energy levels}
+\subsection{Basic experiment with NV centers}
+\begin{frame}{Experimental setup}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_NV_levels}
+\includegraphics[width=0.9\textwidth,height=0.85\textheight,keepaspectratio]{Slide_setup}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_ODMR_0}
-\end{frame}
-
-\begin{frame}{Experimental setup}
-\centering
-\includegraphics[width=0.9\textwidth,height=0.9\textheight,keepaspectratio]{Slide_setup}
+\includegraphics[width=\textwidth,height=0.80\textheight,keepaspectratio]{Slide_ODMR_0}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_ODMR_8_classes}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_ODMR_8_classes}
 \end{frame}
 
-\begin{frame}{Sensitivity limit for NV ensemble magnetometry}
+\begin{frame}{Transverse magnetic field effect}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_NV_ensemble_sensi}
-\end{frame}
-
-\begin{frame}{Dipole-dipole interaction between two electronic spins}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_dipole_dipole}
-\end{frame}
-
-\begin{frame}{Interaction limit to NV ensemble magnetometry}
-\centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_interaction_limit}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_champs_transverse}
 \end{frame}
@@
 \section{Low field depolarization magnetometry (LFDM)}
+\subsection{Principle}
 \begin{frame}{Outline}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_presentation_sujet}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_presentation_sujet}
 \end{frame}
 
-\begin{frame}{Pesentation of low field depolarization magnetometry (LFDM)}
+\begin{frame}{LFDM experimental setup}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_principle_LFDM}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_principle_LFDM}
 \end{frame}
 
+\subsection{Characterization}
 \begin{frame}{Sensitivity of LFDM}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_sensi_LFDM}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_sensi_LFDM}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_litterature}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_comparison_microwave}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_microwave}
 \end{frame}
 
+\subsection{Applications}
 \begin{frame}{Application: wide-field magnetometry on irregular surfaces}
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_applications_wide_field}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_applications_wide_field}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_CR_VH}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_VH}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_CR_adamas}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_adamas}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_exp_stretch}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_exp_stretch}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_0B_theorie}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_T1_PL_100}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{slide_champs_transverse}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_champs_transverse}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{shema_summary_exp}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{shema_summary_exp}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{Slide_angular_sensitivity}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_angular_sensitivity}
 \end{frame}

### 2022-11-29 Talks/Soutenance 2022/Plot_T2*_P1.py

@@ -4,26 +4,58 @@ sys.path.append('C:\\Users\\cleme\\OneDrive\\Documents\\these\\python_clément')
 sys.path.append('/home/pellet-mary/these/python_clément')
 from analyse import *
 
-plt.figure(num=1,figsize=(4,3),dpi=80)
-plt.xticks(fontsize=16)
-plt.yticks(fontsize=16)
-plt.locator_params(axis='x', nbins=5)
-plt.locator_params(axis='y', nbins=5)
 
-x=np.logspace(-3,3,300)
-DeltaNu=160+16*x
 
-N=1e-12*1.75e23*1e-6*1e-1*x
 
-hbar=1.05e-34
-g=2
-muB=9.27e-24
-C=1e-3
+def plot_sensi_P1():
+	plt.figure(num=1,figsize=(4,3),dpi=80)
+	plt.xticks(fontsize=16)
+	plt.yticks(fontsize=16)
+	plt.locator_params(axis='x', nbins=5)
+	plt.locator_params(axis='y', nbins=5)
+	x=np.logspace(-3,3,300)
+	DeltaNu=160+16*x
 
-eta=hbar/(g*muB*C)*sqrt(DeltaNu/N)
-eta=eta*1e9
+	N=1e-12*1.75e23*1e-6*1e-1*x
 
-plt.plot(x,eta)
-plt.yscale('log')
-plt.xscale('log')
-plt.show()
+	hbar=1.05e-34
+	g=2
+	muB=9.27e-24
+	C=1e-3
+
+	eta=hbar/(g*muB*C)*sqrt(DeltaNu/N)
+	eta=eta*1e9
+
+	plt.plot(x,eta)
+	plt.yscale('log')
+	plt.xscale('log')
+	plt.show()
+
+def max_sensi_single_spin():
+	DeltaNu=100
+
+	N=1
+
+	hbar=1.05e-34
+	g=2
+	muB=9.27e-24
+	C=1e-2
+	eta=hbar/(g*muB*C)*sqrt(DeltaNu/N)
+	eta=eta*1e9
+	print(eta)
+
+max_sensi_single_spin()
+
+def plot_PL_CR_simu():
+	plt.figure(num=1,figsize=(4,3),dpi=80)
+	plt.xticks(fontsize=16)
+	plt.yticks(fontsize=16)
+	# plt.locator_params(axis='x', nbins=5)
+	# plt.locator_params(axis='y', nbins=5)
[…]

### 2022-12-02 Talks/Soutenance 2022/presentation.tex
fatiguééé
tata

@@
 \usecolortheme{seahorse}
+\setcounter{tocdepth}{1}
 
@@
 \maketitle
-%\begin{center}
-%\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{logos}
-%\end{center}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{logos}
+\end{center}
+\end{frame}
+
+\begin{frame}{Context of my PhD}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_contexte_f-4}
+\end{frame}
+
+\begin{frame}{Context of my PhD}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_contexte_f-3}
+\end{frame}
+
+\begin{frame}{Context of my PhD}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_contexte_f-2}
+\end{frame}
+
+\begin{frame}{Context of my PhD}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_contexte_f-1}
+\end{frame}
+
+\begin{frame}{Context of my PhD}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_contexte_f}
 \end{frame}
@@
 
-\subsection{Quantum sensing and magnetometry}
+\subsection{Quantum sensing}
 \begin{frame}{Quantum sensing and metrology}
@@
 
-\begin{frame}{Magnetometry key parameters}
+\subsection{Magnetometry}
+\begin{frame}{Key properties of magnetometers}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_magnetometer_size}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_magnetometer_size_f-2}
+\end{frame}
+
+\begin{frame}{Key properties of magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_magnetometer_size_f-1}
+\end{frame}
+
+\begin{frame}{Key properties of magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_magnetometer_size_f}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_bare}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_bare_1}
 \end{frame}
@@
 
-\begin{frame}{Sate of the art magnetometers}
+\begin{frame}{State of the art magnetometers}
 \centering
@@
 
-\begin{frame}{Sate of the art magnetometers}
+\begin{frame}{State of the art magnetometers}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV_n-3}
 \end{frame}
 
-\subsection{NV center magnetometry}
-\begin{frame}{NV center overview}
+\begin{frame}{State of the art magnetometers}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_rapido}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV_n-2}
 \end{frame}
 
-\begin{frame}{NV center magnetometry sensitivity}
+\begin{frame}{State of the art magnetometers}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_limite_sensi}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV_n-1}
 \end{frame}
 
-\begin{frame}{NV center magnetometry: the interaction limit}
+\begin{frame}{State of the art magnetometers}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_interaction_limit}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV_n}
 \end{frame}
 
-\section{Diamond and NV centers in practice}
+\begin{frame}{State of the art magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV_n+1}
+\end{frame}
+
+\begin{frame}{State of the art magnetometers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_quantum_magnetometers_NV_n+2}
+\end{frame}
+
+%\begin{frame}{NV center overview}
+%\centering
+%\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_rapido}
+%\end{frame}
+
+\section{NV centers and diamonds in practice}
 \subsection{Diamond properties}
@@
 \tableofcontents[currentsection]
+\end{frame}
+
+\begin{frame}{Colored centers in diamond}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide diamant_0}
 \end{frame}
@@
 \centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_fab_sample-3}
+\end{frame}
+
+\begin{frame}{Synthetic diamond and NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_fab_sample-2}
+\end{frame}
+
+\begin{frame}{Synthetic diamond and NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_fab_sample-1}
+\end{frame}
+
+\begin{frame}{Synthetic diamond and NV centers}
+\centering
 \includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_fab_sample}
@@
 \subsection{NV center energy levels}
+\begin{frame}{The NV center energy levels}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_levels_3}
+\end{frame}
+
+%\begin{frame}{The NV center energy levels}
+%\centering
+%\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_levels_2}
+%\end{frame}
+
 \begin{frame}{The NV center energy levels}
@@
 \centering
+\includegraphics[width=0.9\textwidth,height=0.85\textheight,keepaspectratio]{Slide_setup_-4}
+\end{frame}
+
+\begin{frame}{Experimental setup}
+\centering
+\includegraphics[width=0.9\textwidth,height=0.85\textheight,keepaspectratio]{Slide_setup_-3}
+\end{frame}
+
+\begin{frame}{Experimental setup}
+\centering
+\includegraphics[width=0.9\textwidth,height=0.85\textheight,keepaspectratio]{Slide_setup_-2}
+\end{frame}
+
+\begin{frame}{Experimental setup}
+\centering
+\includegraphics[width=0.9\textwidth,height=0.85\textheight,keepaspectratio]{Slide_setup_-1}
+\end{frame}
+
+\begin{frame}{Experimental setup}
+\centering
 \includegraphics[width=0.9\textwidth,height=0.85\textheight,keepaspectratio]{Slide_setup}
+\end{frame}
+
+\begin{frame}{Optically detected magnetic resonance (ODMR)}
+\centering
+\includegraphics[width=\textwidth,height=0.80\textheight,keepaspectratio]{Slide_ODMR_0_-3}
+\end{frame}
+
+\begin{frame}{Optically detected magnetic resonance (ODMR)}
+\centering
+\includegraphics[width=\textwidth,height=0.80\textheight,keepaspectratio]{Slide_ODMR_0_-2}
+\end{frame}
+
+\begin{frame}{Optically detected magnetic resonance (ODMR)}
+\centering
+\includegraphics[width=\textwidth,height=0.80\textheight,keepaspectratio]{Slide_ODMR_0_-1}
 \end{frame}
@@
 
-\begin{frame}{ODMR with NV ensemble}
+\begin{frame}{ODMR with NV ensemble: the 4 classes}
 \centering
@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_champs_transverse}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_champs_transverse_allege}
 \end{frame}
@@
 \section{Depolarization mechanisms in dense NV ensemble}
+\subsection{Cross-relaxations}
 \begin{frame}{Outline}
@@
 
+\subsection{The fluctuator model}
 \begin{frame}{Presentation of the fluctuator model}
@@
 
+\subsection{Low field depolarization mechanisms}
 \begin{frame}{Zero field depolarization sources (theory)}
@@
 \end{frame}
+
+\begin{frame}{Summary of the experimental observations}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{shema_summary_exp}
+\end{frame}
+
+\begin{frame}{Angular sensitivity of LFDM}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_angular_sensitivity}
+\end{frame}
+\subsection{Conclusion}
+\begin{frame}{Conclusion}
+
+\begin{itemize}
+\item The use of NV center ensemble for magnetometry is currently limited by the interactions between the spin defects.
+\bigskip
+\item Dipole-dipole interaction within dense ensemble of NV centers result in spin depolarization which is exacerbated at low magnetic field.
+\bigskip
+\item This depolarization can be used to perform microwave-free and orientation-free magnetometry at low magneitc field.
+\bigskip
+\item the LFDM sensitivity seem to be determined by the double-flips.
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Acknowledgments}
+
+\end{frame}
+
+\section*{Bonus slides}
 
@@
 
-\begin{frame}{Summary of the experimental observations}
+\begin{frame}{NV center magnetometry sensitivity}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{shema_summary_exp}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_NV_limite_sensi}
 \end{frame}
+\end{document}
 
-\begin{frame}{Angular sensitivity of LFDM}
+\begin{frame}{NV center magnetometry: the interaction limit}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_angular_sensitivity}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_interaction_limit}
 \end{frame}
-
-\begin{frame}{Conclusion}
-
-\end{frame}
-
-\begin{frame}{Acknowledgments}
-
-\end{frame}
-
-\end{document}

### 2022-12-03 Talks/Soutenance 2022/presentation.tex
tata

@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_presentation_sujet}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_low_field_depolarization_f_-3}
+\end{frame}
+
+\begin{frame}{Depolarization of dense NV ensemble at low magnetic field}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_low_field_depolarization_f_-2}
+\end{frame}
+
+\begin{frame}{Depolarization of dense NV ensemble at low magnetic field}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_low_field_depolarization_f_-1}
+\end{frame}
+
+\begin{frame}{Depolarization of dense NV ensemble at low magnetic field}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_low_field_depolarization_f}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_principle_LFDM}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_principle_LFDM_f-2}
+\end{frame}
+
+\begin{frame}{LFDM experimental setup}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_principle_LFDM_f-1}
+\end{frame}
+
+\begin{frame}{LFDM experimental setup}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_principle_LFDM_f}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature_f-6}
 \end{frame}
 
-\begin{frame}{Comparison with CW ODMR}
+\begin{frame}{Comparison with the state of the art}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_microwave}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature_f-5}
+\end{frame}
+
+\begin{frame}{Comparison with the state of the art}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature_f-4}
+\end{frame}
+
+\begin{frame}{Comparison with the state of the art}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature_f-3}
+\end{frame}
+
+\begin{frame}{Comparison with the state of the art}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature_f-2}
+\end{frame}
+
+\begin{frame}{Comparison with the state of the art}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature_f-1}
+\end{frame}
+
+\begin{frame}{Comparison with the state of the art}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_litterature_f}
 \end{frame}
@@
 \subsection{Applications}
-\begin{frame}{Application: wide-field magnetometry on irregular surfaces}
+\begin{frame}{Application: wide-field magnetometry}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_applications_wide_field}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_applications_wide_field_f-3}
+\end{frame}
+
+\subsection{Applications}
+\begin{frame}{Application: wide-field magnetometry}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_applications_wide_field_f-2}
+\end{frame}
+
+\subsection{Applications}
+\begin{frame}{Application: wide-field magnetometry}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_applications_wide_field_f-1}
+\end{frame}
+
+\subsection{Applications}
+\begin{frame}{Application: wide-field magnetometry}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_applications_wide_field_f}
 \end{frame}
@@
 \end{frame}
+
+\begin{frame}{LFDM: shot-noise limit}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_bonus_shot_noise_limit}
+\end{frame}
+
+\begin{frame}{Comparison with CW ODMR}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_microwave}
+\end{frame}

### 2022-12-04 Talks/Soutenance 2022/presentation.tex
aled

@@
 
+%\beamertemplatenavigationsymbolsempty
+%\setbeamerfont{page number in head/foot}{size=\large}
+%\setbeamertemplate{footline}[frame number]
+
+\expandafter\def\expandafter\insertshorttitle\expandafter{%
+  \insertshorttitle\hfill%
+  \insertframenumber\,}
+
 \begin{document}
+%\begin{frame}
+%\maketitle
+%\begin{center}
+%\includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{logos}
+%\end{center}
+%\end{frame}
+
 \begin{frame}
-\maketitle
-\begin{center}
-\includegraphics[width=\textwidth,height=0.3\textheight,keepaspectratio]{logos}
-\end{center}
+\centering
+\begin{tcolorbox}[colback=blue!20!white]
+\centering
+\Large
+Cross-relaxation in dense ensembles of NV centers and
+application to magnetometry
+\end{tcolorbox}
+\bigskip
+Clément Pellet-Mary PhD Defense
+\bigskip
+\includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{logos}
 \end{frame}
@@
 
-\section{NV centers and diamonds in practice}
+\section{NV center spin properties}
 \subsection{Diamond properties}
@@
 \centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_sensi_LFDM-1}
+\end{frame}
+
+\begin{frame}{Sensitivity of LFDM}
+\centering
 \includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_sensi_LFDM}
@@
 
-\subsection{Applications}
 \begin{frame}{Application: wide-field magnetometry}
@@
 \end{frame}
+
 \begin{frame}{Principle of cross-relaxation with NV centers}
 \centering
-\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation}
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation_f-5}
+\end{frame}
+
+\begin{frame}{Principle of cross-relaxation with NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation_f-4}
+\end{frame}
+
+\begin{frame}{Principle of cross-relaxation with NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation_f-3}
+\end{frame}
+
+\begin{frame}{Principle of cross-relaxation with NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation_f-2}
+\end{frame}
+
+\begin{frame}{Principle of cross-relaxation with NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation_f-1}
+\end{frame}
+
+\begin{frame}{Principle of cross-relaxation with NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_CR_presentation_f}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_VH}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_VH_0}
+\end{frame}
+
+\begin{frame}{Example: Cross-relaxation between NV centers and VH$^-$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_VH_1}
+\end{frame}
+
+\begin{frame}{Example: Cross-relaxation between NV centers and VH$^-$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_VH_2}
+\end{frame}
+
+\begin{frame}{Example: Cross-relaxation between NV centers and VH$^-$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_VH_3}
+\end{frame}
+
+\begin{frame}{Example: Cross-relaxation between NV centers and VH$^-$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_VH_4}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_adamas}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_adamas_f-3}
+\end{frame}
+
+\begin{frame}{Cross-relaxation between NV centers and NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_adamas_f-2}
+\end{frame}
+
+\begin{frame}{Cross-relaxation between NV centers and NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_adamas_f-1}
+\end{frame}
+
+\begin{frame}{Cross-relaxation between NV centers and NV centers}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CR_adamas_f}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro}
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro_f-4}
+\end{frame}
+
+\begin{frame}{Presentation of the fluctuator model}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro_f-3}
+\end{frame}
+
+\begin{frame}{Presentation of the fluctuator model}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro_f-2}
+\end{frame}
+
+\begin{frame}{Presentation of the fluctuator model}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro_f-1}
+\end{frame}
+
+\begin{frame}{Presentation of the fluctuator model}
+\centering
+\includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{Slide_fluct_intro_f}
 \end{frame}
@@
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_exp_stretch}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_exp_stretch_f-1}
+\end{frame}
+
+\begin{frame}{Stretched exponential decay profile}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_exp_stretch_f}
 \end{frame}
@@
 \subsection{Low field depolarization mechanisms}
-\begin{frame}{Zero field depolarization sources (theory)}
+\begin{frame}{Zero field depolarization mechanisms (theory)}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie_f-5}
 \end{frame}
 
-\begin{frame}{Summary of the experimental observations}
+\begin{frame}{Zero field depolarization mechanisms (theory)}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{shema_summary_exp}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie_f-4}
 \end{frame}
 
-\begin{frame}{Angular sensitivity of LFDM}
+\begin{frame}{Zero field depolarization mechanisms (theory)}
 \centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_angular_sensitivity}
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie_f-3}
 \end{frame}
+
+\begin{frame}{Zero field depolarization mechanisms (theory)}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie_f-2}
+\end{frame}
+
+\begin{frame}{Zero field depolarization mechanisms (theory)}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie_f-1}
+\end{frame}
+
+\begin{frame}{Zero field depolarization mechanisms (theory)}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_0B_theorie_f}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B$ in arbitrary direction}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1_f-5}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B$ in arbitrary direction}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1_f-4}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B$ in arbitrary direction}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1_f-3}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B$ in arbitrary direction}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1_f-2}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B$ in arbitrary direction}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1_f-1}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B$ in arbitrary direction}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_1x1x1x1_f}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100_f-6}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100_f-5}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100_f-4}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100_f-3}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100_f-2}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100_f-1}
+\end{frame}
+
+\begin{frame}{Experiment: $\vec B \parallel [100]$}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_PL_100_f}
+\end{frame}
+
 \subsection{Conclusion}
 \begin{frame}{Conclusion}
-
 \begin{itemize}
 \item The use of NV center ensemble for magnetometry is currently limited by the interactions between the spin defects.
+\pause
 \bigskip
-\item Dipole-dipole interaction within dense ensemble of NV centers result in spin depolarization which is exacerbated at low magnetic field.
+\item Dipole-dipole interaction within dense ensemble of NV centers result in spin depolarization. This effect is stronger in zero-field.
+\pause
 \bigskip
-\item This depolarization can be used to perform microwave-free and orientation-free magnetometry at low magneitc field.
+\item This depolarization can be used to perform microwave-free and orientation-free magnetometry at low magnetic field.
+\pause
 \bigskip
-\item the LFDM sensitivity seem to be determined by the double-flips.
+\item The physics of low-field depolarization is complex and need to be further studied to improve the performances of LFDM.
 \end{itemize}
@@
 \begin{frame}{Acknowledgments}
-
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Remerciements}
 \end{frame}
@@
 \end{frame}
-\end{document}
 
@@
 \end{frame}
+
+\begin{frame}{Summary of the experimental observations}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{shema_summary_exp}
+\end{frame}
+
+\begin{frame}{Angular sensitivity of LFDM}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_angular_sensitivity}
+\end{frame}
+
+\end{document}

### 2022-12-06 Talks/Soutenance 2022/presentation.tex
vf?

@@
 \subsection{Conclusion}
+
 \begin{frame}{Conclusion}
-\begin{itemize}
-\item The use of NV center ensemble for magnetometry is currently limited by the interactions between the spin defects.
-\pause
-\bigskip
-\item Dipole-dipole interaction within dense ensemble of NV centers result in spin depolarization. This effect is stronger in zero-field.
-\pause
-\bigskip
-\item This depolarization can be used to perform microwave-free and orientation-free magnetometry at low magnetic field.
-\pause
-\bigskip
-\item The physics of low-field depolarization is complex and need to be further studied to improve the performances of LFDM.
-\end{itemize}
+\end{frame}
 
+\begin{frame}{Conclusion}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CCL_f-4}
+\end{frame}
+
+\begin{frame}{Conclusion}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CCL_f-3}
+\end{frame}
+
+\begin{frame}{Conclusion}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CCL_f-2}
+\end{frame}
+
+\begin{frame}{Conclusion}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CCL_f-1}
+\end{frame}
+
+\begin{frame}{Conclusion}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_CCL_f}
 \end{frame}
@@
 
-\section*{Bonus slides}
+%\section*{Bonus slides}
 
@@
 
+\begin{frame}{Summary of the experimental observations}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{shema_summary_exp}
+\end{frame}
+
+\begin{frame}{LFDM: shot-noise limit}
+\centering
+\includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{Slide_bonus_shot_noise_limit}
+\end{frame}
+
+\begin{frame}{Comparison with CW ODMR}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_microwave}
+\end{frame}
+
+\begin{frame}{Angular sensitivity of LFDM}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_angular_sensitivity}
+\end{frame}
+
+\begin{frame}{Protocol T1}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_protocol}
+\end{frame}
+
+\begin{frame}{Protocol T1}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_T1_exp_et_stretch}
+\end{frame}
+
+\begin{frame}{Fluctuator linewidth}
+\centering
+\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{slide_fluct_linewidth}
+\end{frame}
+
 \begin{frame}{NV center magnetometry sensitivity}
@@
 
-\begin{frame}{LFDM: shot-noise limit}
-\centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_bonus_shot_noise_limit}
-\end{frame}
-
-\begin{frame}{Comparison with CW ODMR}
-\centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_comparison_microwave}
-\end{frame}
-
-\begin{frame}{Summary of the experimental observations}
-\centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{shema_summary_exp}
-\end{frame}
-
-\begin{frame}{Angular sensitivity of LFDM}
-\centering
-\includegraphics[width=\textwidth,height=0.85\textheight,keepaspectratio]{Slide_angular_sensitivity}
-\end{frame}
-
 \end{document}
